### 2021-12-16 Gandhi_Project_Report/draft1.tex
first commit. Line 250

\documentclass[12 pt]{article}
\usepackage[utf8]{inputenc}
\usepackage{filecontents}
\usepackage[many]{tcolorbox}
\usepackage{graphicx}
\usepackage{harpoon}
%\usepackage{apacite} 
\usepackage{xcolor}
\usepackage{latexsym}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{mdframed}
%\usepackage[authoryear]{natbib}
\usepackage{hyperref, url}
\usepackage[show]{ed}
\usepackage{subfigure}
\usepackage[ntheorem]{empheq} 
\usepackage{amsthm, amssymb, amsfonts, latexsym}
\usepackage{pst-all}

\usepackage{enumitem}

\setlength{\parskip}{1.7em}
\setlength{\parindent}{0em}

%\setlength{\baselineskip}{20pt}

\setlength{\textwidth}{429.75499pt}
\setlength{\textheight}{643.20255pt}
\setlength{\oddsidemargin}{5 mm}
\setlength{\evensidemargin}{5 mm}
\setlength{\topmargin}{0 mm}
\setlength{\headsep}{0 mm}
\setlength{\headheight}{0 mm}

\newcommand{\citealtt}[1]{\citeauthor{#1},\citeyear{#1}}
\newcommand{\myycite}[1]{\citep{#1}}

\mathchardef\mhyphen="2D

%\newcommand{\cev}[1]{\reflectbox{\ensuremath{\vec{\reflectbox{\ensuremath{#1}}}}}}

\usepackage{pst-node,graphicx,pst-blur}
%\uspackage{auto-pst-pdf}
%\usepackage{tikz-cd}

\makeatletter
\DeclareRobustCommand{\cev}[1]{%
  \mathpalette\do@cev{#1}%
}
\newcommand{\do@cev}[2]{%
  \fix@cev{#1}{+}%
  \reflectbox{$\m@th#1\vec{\reflectbox{$\fix@cev{#1}{-}\m@th#1#2\fix@cev{#1}{+}$}}$}%
  \fix@cev{#1}{-}%
}
\newcommand{\fix@cev}[2]{%
  \ifx#1\displaystyle
    \mkern#20mu
  \else
    \ifx#1\textstyle
      \mkern#20mu
    \else
      \ifx#1\scriptstyle
        \mkern#26mu
      \else
        \mkern#26mu
      \fi
    \fi
  \fi
}

\makeatother

\newcommand{\xcm}{\epsfxsize=3.1cm}
\newcommand{\fig}[1]{\epsfbox}
% \newcommand{\bp}{\begin{minipage}{3.1cm}}
% \newcommand{\ep}{\end{minipage}}

\newtheorem{Definition}{Definition}[]
%\newmdtheoremenv{Theorem}{Theorem}[]
\newtheorem{Theorem}{Theorem}[]
\newtheorem{Lemma}{Lemma}[]
\newtheorem{Proposition}{Proposition}[]
\newtheorem{Corollary}{Corollary}[]
\newtheorem{Remark}{Remark}[]

\makeatletter 
\renewcommand{\thefigure}{\@arabic\c@figure}
\makeatother
\usepackage{xr}

%%%% PNAS GRAPHICS SETUP- FROM THE PNAS STYLE LATEX FILE
\RequirePackage{graphicx,xcolor}
\RequirePackage{colortbl}
\RequirePackage{booktabs}
\RequirePackage{algorithm}
\RequirePackage[noend]{algpseudocode}
\RequirePackage{changepage}
\RequirePackage[twoside,%
				letterpaper,includeheadfoot,%
				layoutsize={8.125in,10.875in},%
                layouthoffset=0.1875in,%
                layoutvoffset=0.0625in,%
                left=38.5pt,%
                right=43pt,%
                top=43pt,% 10pt provided by headsep
                bottom=32pt,%
                headheight=0pt,% No Header
                headsep=10pt,%
                footskip=25pt,
                marginparwidth=38pt]{geometry}
\RequirePackage[labelfont={bf,sf},%
                labelsep=period,%
                figurename=Fig. ]{caption}
\setlength{\columnsep}{13.5pt} % Distance between the two columns of text
\setlength{\parindent}{12pt} % Paragraph indent

%% Figure caption style
\DeclareCaptionFormat{pnasformat}{\normalfont\sffamily\fontsize{7}{9}\selectfont#1#2#3}
\captionsetup*{format=pnasformat}

%%% GREYBOX AROUND FIG
\definecolor{lightgray}{gray}{0.95}
\newcommand\greybox[1]{%
  \vskip\baselineskip%
  \par\noindent\colorbox{lightgray}{%
    \begin{minipage}{\textwidth}#1\end{minipage}%
  }%
  \vskip\baselineskip%
}

\newtcolorbox{paperbox}[1][]{
    enhanced,
    colback=white,
    boxrule=0pt,
    boxsep=0pt,
    arc=0mm,
    width=0.8\linewidth,
    fuzzy shadow={0mm}{-4pt}{-4pt}{1mm}{black!30!white},
    #1
}

\externaldocument{Supp_Koopman-0808}
\begin{document}

\title{Project Thesis/Report}

\author{} 
\date{}
\date{}
\maketitle

\section{Introduction}

Experimenting on biological, physical, and artificial systems in order to generate a more informative dynamical model is a well-established practice in modern science. Traditional methods for modelling physical systems are based on laws of physics that are based either on empirical relationships or on intuition. For systems that evolve with time, physical laws yield mathematical equations that govern how the quantities evolve with time. However our world around is, for the most part, much more complex than that which can be distilled into elegant equations. We do not fully understand many of the more complex systems, nor do they even provide us with a good physical intuition of the underlying principles governing the dynamics.

To complicate this - the underlying systems we observe often display sensitive dependence on initial conditions; despite having highly similar initial conditions, differing orbits diverge quickly and to such an extent that it becomes seemingly impossible to retrace their steps back to the original conditions. Even the presence of usually 'negligible' computational noise or measurement error renders long-term point-wise prediction infeasible. 
There are many difficulties that one encounters while modelling such systems:
\vspace{-8mm}
\begin{enumerate}[noitemsep, label=\roman*.]
  \item One may not have access to the complete states of the systems
  \item The actual system may be described by functions that behave wildly in the sense their graphs have a wild oscillatory behaviour, i.e., they have a large functional complexity.    
\end{enumerate}

Models derived primarily from data can be classified into three categories: 
\vspace{-8mm}
\begin{enumerate}[noitemsep, label=\roman*.]
  \item  Interpretable models (i.e., they establish relationships between internal physical mechanisms), 
  \item Partially interpretable models capturing some modes of the dynamics, 
  \item Non-Interpretable models, defined as such mainly due to the fact that they are defined on a different phase space that is usually high-dimensional. 
\end{enumerate}

Examples in the literature attempting to forecast data from such systems have tried a number of different approaches with differing degrees of success. One could learn a system conjugate to the underlying system by applying the Takens delay embedding; this learnt system could then be used for forecasting the original system. \cite{takens1981detecting}. An ordinary differential equation can be obtained from data (Citations to Sindy). Alternatively, one could approximate the vector field by a library of functions (Citations to Sindy) to obtain interpretable models. Takens delay methods suffer from lack of robustness to noise and may not yield a learnable map that has lower functional complexity (See Section 3). For the methods that build interpretable models (Citations to Sindy) the complete system states are needed which is rarely available. Pure machine learning methodology that processes temporal information (like the echo state networks (Citations to echo state networks)) is an approach that maps data onto a higher dimensional space for further processing. Although they perform  well on forecasting some dynamical systems, they fail completely on others (Citations here) as there is often guarantee that the right function was learnt during training.

Recent partially interpretable models available in the literature have been based on the Koopman operator (e.g.,\cite{koopman1932dynamical,budivsic2012applied}) to  employ observables mapping the data onto a higher-dimensional space. This then makes the dynamics in the higher-dimensional space more amenable for approximation by a linear transformation. Such methods, obviously, do not guarantee an exact reconstructions for nonlinear models, and in practice provide poor long-term consistency for a large class of systems. (Add citation)

The non-interpretable models include the delay embedding and the machine learning algorithms. Under some generic conditions, the Takens delay embedding theorem \cite{takens1981detecting} and its various generalisations (e.g., \cite{sauer1991embedology, stark1999delay, gutman2018embedding, gutman2018embedding}) establishes the learnability of a system constructed by concatenating sufficiently large previous time-series observations of a dynamical system into a vector (called delay coordinates). This then establishes the existence of a map on the space of delay coordinates equivalent (or topologically conjugate) to the underlying map from which the observed time-series was first obtained. Although topological conjugacy guarantees an alternate representation of the underlying system, the quality of thsi representationstill depends on numerous parameters, making the comprehension of the dynamics rather unreliable (e.g., \cite{principe1992prediction}). One reason for this fragility is that the embedded attractor in the reconstruction space not always an attractor itself, despite  unquestionably being an invariant set. Since the embedded object is not \emph{(definitely/always?)} an attractor (as explained in Chapter 3) it can cause predictions to fail.

Practically, the application of Takens embedding involves learning a map through some technique and consequently one wishes that these would have low functional complexity\cite{manjunath2021universal}, i.e., functions with fewer oscillatory graphs.

This thesis deals with the implementation and analysis of non-interpretable models that can guarantee \emph{exact/accurate} reconstruction. The project work concerns the study and implementation of a method (Citation to paper with Adriaan) that incorporates learning a function by mapping the data on to a higher dimensional space. With a clear understanding of how the data is mapped onto the higher dimensional space, the method then permits the learning of a dynamical system topologically conjugate to that of the underlying system. Instead of linear regression, deep learning methods are employed to learn the correct function. With slight modifications to the implementation in the paper (with Adriaan), we show that one can construct accurate non-interpretable models with the ability to reconstruct attractors from more hard-to-forecast systems like the double pendulum. (The forecasting of the time-series from a double pendulum has not been reported before.) Moreover, we also demonstrate that long-term statistical consistency is preserved.

By solidifying the mathematical underpinnings of our theory, we hope to guarantee the ability to construct models with predictive power ranging from molecular biology to neuroscience. in the near future. (We can modify this sentence when the report is completed.)

The report is organised as follows: 
\newline In Section 2, we recall the definition of a discrete-time dynamical system, how a discrete-system arises from a flow of an ODE and then proceed to define the inverse limit space and topological conjugacy of autonomous systems. 
\newline In Section 3, we introduce the problem of forecasting dynamical systems, state the Takens delay embedding theorem and discuss various issues faced while forecasting. 
\newline In Section 4, a driven dynamical system is defined and discuss the properties of a specific class of driven dynamical systems that we make use of in this theses/report/paper. 
\newline Finally in Section 5 we show the implementation of these forecasting methods, and conclusions are provided in Section 6.

\section{Discrete-time Dynamical Systems}

In this chapter, we provide a brief description of what a discrete-time dynamical system is and what it means for it to exhibit chaos. We refer to \cite{devaney2018introduction, de2013elements} for more details.

At its most elementary level, a dynamical system is just \emph{something} that evolves deterministically through time. In the context of this thesis, deterministic refers to the fact that a system evolves according to specified rules rather than based on random events. 
Dynamical systems arise in a variety of situations. A continuous dynamical system, and here specifically the motion of a pendulum, is one in which quantities such as the angular position and angular momentum are known at all times. 
The equations of the dynamical system take the form of one or more ordinary differential equation that determine the relevant quantities at any future time if we know the initial location and momentum. 
In ecology, discrete dynamical systems are widely used to model population growth. The model in this case is a function calculating the following generation's population given the population of the previous generation. If we know the starting population, we may once again calculate the population at any time in the future.

Formally, a function $T: U \to U$, where $U$ is some set is  a \emph{discrete-time dynamical system} and its iterates $\{u,Tu,T^2u,\ldots\}$, where $T^n$ denotes the $n$-fold composition of $T$ with itself, describe the evolution of an initial condition $u\in U$ (Note that we drop the brackets and denote $T(u)$ by $Tu$ for simplicity).

Continuous dynamical systems modelled using ordinary differential equations can give rise to discrete-time dynamical systems. To see this, consider differential equation $\dot{x} = f(x)$, $f: \mathbb{R}^n \to \mathbb{R}^n$, $n\in\mathbb{N}$ given to have a unique solution passing through 
each point $x\in\mathbb{R}^{n}$, the \emph{flow of the equation} is defined to be a mapping $\varphi: \mathbb{R}^n \times \mathbb{R} \to \mathbb{R}$ that satisfies the following properties:
\vspace{-8mm}
\begin{enumerate}[noitemsep, label=\roman*.]
  \item $x(t)$ is a solution of the differential equation,
  \item $x(0)=x_0$, and
  \item $\varphi(x_0,t) = x(t)$.
\end{enumerate}
By fixing $t=K$, we can define the \emph{time-$K$ map} as  $T(x):= \varphi(x_0,K)$, and it is easily verified that $T\circ T(x_0) = \varphi(x_0,2K)$. In general the $m^{th}$ iterate of $x_0$ under $T$ would be the value of the solution of the ODE evaluated at time $mK$ with the initial condition $x_0$. 
Thus ordinary differential equations give rise to a discrete-time dynamical systems by sampling the value of $x$ at time intervals $K$ units apart.

Of course, discrete-time dynamical systems need not always arise through a differential equation, and once again we may consider the field of ecology where discrete dynamical systems are often directly derived or assumed. 
\emph{There is a school of thought that advocates discrete-time dynamical systems to be more natural for modelling real-world observations than differential equations (see \cite{saber2010introduction}).  <- too philosphical, or actually good?}

A numerical discretization of a differential equation can also give rise to a discrete-time dynamical system. For instance, Euler's method approximates $\dot{x}(t)$ by $(x(t+h)-x(t))/h$; if $h$ is fixed throughout, the solution of a differential equation $\dot{x}=f(x)$ 
can be approximated at the time instant $t+(m+1)h$ by iterating the equation 
$$x(t+(m+1)h) = x(t+mh) + h f(x(t+mh)).$$ 
Adopting more succinct notation by replacing $x(t+mh)$ with $u_m$, we rewrite the above equation as
$$u_{m+1} = u_m + hf(u_m)$$
or in even simpler terms as the discrete-time dynamical system with map $T(u) = u + hf(u)$, where $T$, $u$ and $f$ are understood to be as above.

Given a discrete-time dynamical system $T: U \to U$, we call a sequence $\bar{u} = \{u_n\}_{n\in \mathbb{Z}}$ that obeys the update equation, $u_{n+1}=Tu_n$, $n \in \mathbb{Z}$ the \emph{orbit of $T$}. A core concept in the study of dynamical systems is that of invariance. A subset $A \subset U$ is said to be an \emph{invariant set} if $T(A) =A$. 
\newline Two examples of invariant sets include a fixed point where $Tu=u$ for  $u\in U$, and a periodic orbit, i.e., a set of iterates $\{u,Tu, T^2u,\ldots,T^pu\}$, where $T^{p+1}u=u$ for some $p$.  The entire space $U$ could be also be invariant. Consider for example the space $U=[0,1]$, where $Tu=4u(1-u)$ and then $U$ is invariant.

%%%RIGHT HERE

The kind of invariant sets a discrete-time dynamical system has can tell a great deal about the iterates of a dynamical system. For example, if $U=[0,1]$ has map $Tu= u/2$, then the only invariant set is $\{0\}$, and all orbits approach this invariant set as time flows in the forward direction. One may ask if every orbit approaches an invariant set? In general the answer is no, since for the dynamical system $Tu=2u$, an orbit not intersecting the invariant set $\{0\}$ would not approach any invariant set. However,  when the space $U$ is compact, all orbits approach an invariant set. In fact the $\omega$-limit set $\omega(u;T)$ of a point $u$ defined to the collection of limit points of the sequence $\{x,Tu,T^2u,\ldots\}$ is nonempty, and $\omega(u;T)$ is invariant.  Now, invariant sets can be attracting or repelling depending on how orbits in its vicinity behave as in the examples of $Tu =u/2$ and $Tu=2u$. We are interested in attractive invariant sets, and in particular what are called attractors whose definition we recall next. 
\begin{Definition} \rm Let $T: U \to U$, where $U$  is a metric space with metric $d$. A compact subset $A \subset U$ is said to be an attractor if it satisfies the three conditions: 
\begin{enumerate}
	\item $A$ is invariant. 
	\item $A$ is asymptotically stable, i.e., for every $\epsilon > 0$ and for all $u$ so that $d(u,A) < \epsilon$, $d(T^nu,A) \to 0$ as $n\to \infty$. 
	\item $A$ has Lyapunov stability, i.e., for every $\epsilon > 0$  there exists a $\delta(\epsilon) > 0$ so that $d(u,A) < \delta$ implies $d(T^nu,A) < \epsilon$ for all $n\ge 0$.  
\end{enumerate}
\end{Definition}

For instance, in the system  $Tu=u/2$ the set $\{0\}$ is an attractor. For the system,  $Tu=4u(1-u)$ on $[0,1]$, it can be verified that the only attractor is the entire space $[0,1]$. A dynamical system can have several attractors, and an attractor can be contained in another attractor as well.

If the dynamics on the attractor is somewhat complicated in the sense that the attractor cannot be decomposed further, and if the attractor is infinite then there is a possibility of complicated dynamics, and a particular well-studied phenomenon of such complexity gives rise to what is called a chaotic behaviour.

In the 1960s, a number of mathematicians and mathematically interested scientists independently discovered chaos in the mathematical sense. Edward Lorenz, a meteorologist, may have been the first to explain this phenomenon in his 1963 paper \cite{lorenz1963deterministic}. We note that the notions of invariance, attractivity and chaos can be described for continuous systems as well, and Lorenz's system comprised of a system of differential equations. The narrative of Lorenz's discovery of chaos, as well as the history of other forerunners in this subject, are fascinating. For those who are interested, we highly recommend James Gleick's book Chaos: The Making of a New Science \cite{gleick2008chaos}. It clearly illustrates these experiences, explains why chaos was such a startling and crucial mathematical and scientific discovery, and describes the underlying mathematical notions for non-specialists.

There are several intricately different notions of chaos each of them indicating some complexity. In practice, it is not possible to verify which notion is satisfied when data is observed from a dynamical system. However, for the sake of understanding some nuances, we recall the notion of the definition of chaos in the sense of Devaney \cite{devaney2018introduction,de2013elements}. Devaney's definitio has three ingredients, and the first one is the notion of sensitive dependence on initial conditions.

\begin{Definition} \rm 
A dynamical system $T: U \to U$ is said to have sensitive dependence on initial conditions (SDIC) if there exists a $\delta > 0$ such that for every $u \in U$ and in every neighborhood of $u \in U$ there exists an $v$ and an integer $N$ such that $d(T^Nu,T^Nv)>\delta$. 	
\end{Definition}

This definition has a lot of quantifiers in it to analyse, so we should take our time to acquire a sense of what it says. The common misconception is to interpret SDIC as two close points ($u$ and $v$) eventually become separated by a distance $\delta$ under iteration by $T$. But this is not true. To completely comprehend all of the subtleties of the concept, we need to talk about it more thoroughly. The simplest way to achieve this is to go through each sentence in the definition slowly and attentively, then examine how they fit together to convey the concept of sensitive dependence.

First, the $\delta>0$ in the definition of SDIC should be independent of $u$. Second, in every neighborhood of $u$, we may not necessarily find all points $v$ in the neighborhood distinct from $u$ that would separate from the forward iterates of $u$. Next, $N$ depends upon $u$ and $v$ chosen, and their iterates may not separate for ever (i.e., for all $n>N$) and we allow their iterates to get arbitrarily close in the future.

\ednote{M: It is probably best to see a few numerical simulations to understand SDIC. (Please use the logistic map and perhaps the trajectories of the double pendulum data). }

We next recall the next ingredient of Devaney's definition of chaos.

\begin{Definition}
	A dynamical system $F: U \to U$ is topologically transive if for any pair of nonempty open sets $E_1$ and $E_2$ there exists a $n>0$ such that $T^n(E_1) \cap E_2 \not= \emptyset$. 
\end{Definition}

Topologically transitivity implies that the iterates of an open set of initial conditions gets mixed up with other open sets. On a compact metric space, it can be shown that topological transitivity also implies the existence of a point whose forward iterates are dense. In fact, it is more topologically likely that one picks a point whose iterates are dense in the sense that the complement of the set of points whose forward iterates are not dense in $U$  are topologically insignificant -- such an insignificant set can be written as a countable union of sets with empty interior.

Now we recall the definition of chaos in the sense of Devaney \cite{devaney2018introduction}.
\begin{Definition}
	A dynamical system $T: U \to U$ is said to exhibit chaos in the sense of Devaney if it satisfies the three properties:
	\begin{enumerate}
		\item $T$ has SDIC
		\item $T$ has topologically transitive
		\item the set of periodic points of $T$ are dense in $U$. 
	\end{enumerate}
\end{Definition}

An example of a dynamical system that satisfies these three properties is the standard tent map $T(u)=1-|2u-1|$ defined on $[0,1]$. We now reason as why this map exhibits chaos in the sense of Devaney.   The graph of the map $T$ is piecewise linear with two straight lines, one that connects the points $(0,0)$ and $(1/2,1)$ and the other that connects $(1/2,1)$ to $(1,0)$ and they form a tent with a base centered at $u=1/2$. The graph of the map $T^2$ comprises two symmetric tents with their base centered at $1/4$ and $3/4$.  So given any point $u$ and an open neighborhoow $W  \subset [0,1]$ of $u$, we can find $k$ large enough so that $T^k(W) = [0,1]$ since since we can accommodate a tent whose base is contained in $W$. This implies that there are two points $w_1$ and $w_2$ in $W$ so that $d(T^kw_1,T^kw_2)= 1$, and by triangle inequality  at least one of the inequalities $d(T^ku,T^kw_1)> \delta$ or $d(T^ku,T^kw_2)> \delta$ hold when $\delta\in (0,1/2)$.  So $T$ has sensitive dependence on initial conditions.  Next, let $W_1$ and $W_2$ be two nonempty open sets. Since given any open set $W_1$, we can find a $k$ large enough so that $T^k(W_1) = [0,1]$ since we can accommodate a tent whose base is contained in $W$. So $T^k(W_1) \cap W_2 \not=\emptyset$, and thus $T$ has topological transitivity. Next, since for every open interval $(a,b)$,  there exists a $k$ such that $T^k(a,b) = [0,1]$, the graph of $T^k$ intersects the graph of the identity map on $[0,1]$, and if intersection has coordinates $(p,p)$,  the point $p$ is a fixed point of $T^k$, which is a periodic point of $T$. Thus the set of periodic points are dense.

We next talk on another concept of conjugacy. 
To show two metric or topological spaces are topologically the same one needs to show that they are homeomorphic, i.e., to establish a homeomorphism between the spaces. In the study of dynamical systems defined on two spaces, showing two spaces to be homeomorphic does not mean that the two dynamical systems are somehow related. One needs to establish some kind of dynamical equivalence. The easiest way to illustrate dynamical equivalence between two systems $T:U\to U$ and $S:V\to V$  is through a commutativity diagram:

% Giving errors
\begin{equation}  \label{eqn_conjugacy}
%    \[ 
    \psset{arrows=->, arrowinset=0.25, linewidth=0.6pt, nodesep=3pt, labelsep=2pt, rowsep=0.7cm, colsep = 1.1cm, shortput =tablr}
 \everypsbox{\scriptstyle}
%  \begin{psmatrix}
% U & U\\%
% V & V.
%  %%%
%  \ncline{1,1}{1,2}^{T} 
%  \ncline{1,1}{2,1} <{\phi}
%   \ncline{2,1}{2,2}^{S}
%  \ncline{1,2}{2,2} > {\phi}
%  \end{psmatrix}
% \]
\end{equation} 
To understand the diagram we note that if we travel ``right and then down," the diagram instructs us to use $T$ (top arrow right) first, followed by $\phi$. (right arrow downwards). As a result, the pathway amounts to finding $\phi(T(u))$. If we proceed ``down and then right," the diagram instructs us to apply $\phi$ (left arrow downwards) first, and then apply $\phi$ (right arrow downwards) and then $S$ (bottom arrow right). As a result, the pathway amounts to finding  $S(\phi(u))$. When the diagram holds, we say 
$\phi(T(u))= S(\phi(u))$ and write it more formally as
$\phi \circ T=S\circ \phi$.
When the above commutativity holds in addition that $\phi$ is a homeomorphism, then we say we say that  $S$ is conjugate to $T$.  
If we relax the criterion on $\phi$ to merely require $\phi$ to be continuous and surjective rather than being a homeomorphism, then $\phi$ is called a semi-conjugacy, and declare that $S$ is semi-conjugate to $T$. When $S$ is conjugate to $T$, it indicates the dynamics between the two systems are in one-to-one correspondence, however when $S$ is semi-conjugate to $T$ with $\phi$ a many-to-one mapping, the dynamics on $V$ provides a coarse-grained description of the dynamics on $V$. It is common to remark that $S$ is a factor of $T$ or that $T$ is an extension of $S$ when $S$ is semi-conjugate to $T$. In essence, an extension (e.g., \cite{de2013elements}) is a larger system that captures all of the important dynamics of its factor.

It is a very hard or nearly an impossible task to establish the existence of a conjugacy or a semi-conjugacy $\phi$ between two systems. However, one can verify if given a function $\phi$ if it satisfies the commutativity.  For example $\phi(x)=\sin(\pi x/2)^2$ is a conjugacy between two systems $Tu=1-|2u-1|$ and $Sv=4v(1-v)$  defined on $[0,1]$.  The importance of conjugacy is that when we know it exists we can prefer to work with  the dynamical systems to gain information of the other, and this is useful for us while forecasting dynamical systems.

\vspace{-1cm}
\bibliographystyle{pnas-new}
\bibliography{pnas}

\end{document}

### 2021-12-17 Gandhi_Project_Report/draft1.tex
up to SDIC

@@
 $$u_{m+1} = u_m + hf(u_m)$$
-or in even simpler terms as the discrete-time dynamical system with map $T(u) = u + hf(u)$, where $T$, $u$ and $f$ are understood to be as above.
+or, in even simpler terms, as the discrete-time dynamical system with map $T(u) = u + hf(u)$, where $T$, $u$ and $f$ are understood to be as above.
 
-Given a discrete-time dynamical system $T: U \to U$, we call a sequence $\bar{u} = \{u_n\}_{n\in \mathbb{Z}}$ that obeys the update equation, $u_{n+1}=Tu_n$, $n \in \mathbb{Z}$ the \emph{orbit of $T$}.
-A core concept in the study of dynamical systems is that of invariance.
-A subset $A \subset U$ is said to be an \emph{invariant set} if $T(A) =A$.
-\newline Two examples of invariant sets include a fixed point where $Tu=u$ for  $u\in U$, and a periodic orbit, i.e., a set of iterates $\{u,Tu, T^2u,\ldots,T^pu\}$, where $T^{p+1}u=u$ for some $p$.
+\subsection{Invariant Sets}
+
+A core concept in the study of dynamical systems is that of invariance. Given a discrete-time dynamical system $T: U \to U$, a subset $A \subset U$ is said to be an \emph{invariant set} if $T(A) =A$. 
+We also define the \emph{orbit of T} to be a sequence $\bar{u} = \{u_n\}_{n\in \mathbb{Z}}$ obeying the update equation, $u_{n+1}=Tu_n$, $n \in \mathbb{Z}$.
+
+Two examples of invariant sets include a fixed point where $Tu=u$ for  $u\in U$, and a periodic orbit, i.e., a set of iterates $\{u,Tu, T^2u,\ldots,T^pu\}$, where $T^{p+1}u=u$ for some $p$.
 The entire space $U$ could be also be invariant.
-Consider for example the space $U=[0,1]$, where $Tu=4u(1-u)$ and then $U$ is invariant.
+Consider for example the space $U=[0,1]$, where $Tu=4u(1-u)$ and then $U$ is invariant. \emph{(as every $u\in{U}$ can be written as $u = 4x(1-x)$ for some $x\in{U}$)}
 
-%%%RIGHT HERE
+We may learn a great deal about the iterates of a dynamical system by considering the types of invariant sets of a discrete-time dynamical system.
+For example, if $U=[0,1]$ has map $Tu= u/2$, then the only invariant set is $\{0\}$, and all orbits approach this invariant set as time flows in the forward direction.
+Indeed, if a some non-zero invariant set (call it $B$) exists, then there is  some $r\in(0,1]\cap{B}$.
+But $r\notin{T(B)}$ since any orbit with initial value $r$ will be a decreasing sequence.
+Moreover, every orbit of T will be decreasing and therefore approach the value $0$ as $n\rightarrow\infty$
 
-The kind of invariant sets a discrete-time dynamical system has can tell a great deal about the iterates of a dynamical system.
-For example, if $U=[0,1]$ has map $Tu= u/2$, then the only invariant set is $\{0\}$, and all orbits approach this invariant set as time flows in the forward direction.
 One may ask if every orbit approaches an invariant set?
-In general the answer is no, since for the dynamical system $Tu=2u$, an orbit not intersecting the invariant set $\{0\}$ would not approach any invariant set.
+In general the answer is no, since for the dynamical system $Tu=2u$, any orbit that does not intersect the invariant set $\{0\}$, will not approach any invariant set.
+
 However,  when the space $U$ is compact, all orbits approach an invariant set.
-In fact the $\omega$-limit set $\omega(u;T)$ of a point $u$ defined to the collection of limit points of the sequence $\{x,Tu,T^2u,\ldots\}$ is nonempty, and $\omega(u;T)$ is invariant.
-Now, invariant sets can be attracting or repelling depending on how orbits in its vicinity behave as in the examples of $Tu =u/2$ and $Tu=2u$.
-We are interested in attractive invariant sets, and in particular what are called attractors whose definition we recall next.
+\textbf{For let $\bar{u} = \{u_n\}_{n\in \mathbb{Z}}$ be an orbit in the compact space U, then $\bar{u}$ has a convergent subsequence }
+
+\textbf{In fact the $\omega$-limit set $\omega(u;T)$ of a point $u$ defined to the collection of limit points of the sequence $\{x,Tu,T^2u,\ldots\}$ is nonempty, and $\omega(u;T)$ is invariant.
+} (not relevant other than being an additional example perhaps?)
+
+We discuss various properties of invariant sets: Invariant sets can be attracting or repelling depending on how orbits in its vicinity behave (recall the examples $Tu =u/2$ and $Tu=2u$).
+We are interested in attractive invariant sets, and in particular so-called attractors which we define below.
 \begin{Definition} \rm Let $T: U \to U$, where $U$  is a metric space with metric $d$. A compact subset $A \subset U$ is said to be an attractor if it satisfies the three conditions: 
+  \vspace{-8mm}
   \begin{enumerate}
 	\item $A$ is invariant. 
-	\item $A$ is asymptotically stable, i.e., for every $\epsilon > 0$ and for all $u$ so that $d(u,A) < \epsilon$, $d(T^nu,A) \to 0$ as $n\to \infty$. 
+	\item $A$ is asymptotically stable, i.e., for every $\epsilon > 0$ and for all $u$ so that $d(u,A) < \epsilon$, we have $d(T^nu,A) \to 0$ as $n\to \infty$. 
 	\item $A$ has Lyapunov stability, i.e., for every $\epsilon > 0$  there exists a $\delta(\epsilon) > 0$ so that $d(u,A) < \delta$ implies $d(T^nu,A) < \epsilon$ for all $n\ge 0$.  
@@
 
-For instance, in the system  $Tu=u/2$ the set $\{0\}$ is an attractor.
-For the system,  $Tu=4u(1-u)$ on $[0,1]$, it can be verified that the only attractor is the entire space $[0,1]$.
-A dynamical system can have several attractors, and an attractor can be contained in another attractor as well.
+In our previous examples, the system  $Tu=u/2$ had the singleton set $\{0\}$ as an attractor.
+For the system,  $Tu=4u(1-u)$ on $[0,1]$, one may easily verify that the only attractor is the entire space $[0,1]$.
+\textbf{For let $B$ be an attractor of U.
+From above, we know that B is itself invariant, is asymptotically stable and has Lyapunov stability.
+Now suppose there is some attractor $C\subsetneq{U}=B$, so there is a $b\in{U}$ which is not in $C$.
+........
+}
 
-If the dynamics on the attractor is somewhat complicated in the sense that the attractor cannot be decomposed further, and if the attractor is infinite then there is a possibility of complicated dynamics, and a particular well-studied phenomenon of such complexity gives rise to what is called a chaotic behaviour.
+A dynamical system can have several attractors, and may also be contained in another attractor.
 
-In the 1960s, a number of mathematicians and mathematically interested scientists independently discovered chaos in the mathematical sense.
-Edward Lorenz, a meteorologist, may have been the first to explain this phenomenon in his 1963 paper \cite{lorenz1963deterministic}.
-We note that the notions of invariance, attractivity and chaos can be described for continuous systems as well, and Lorenz's system comprised of a system of differential equations.
+% If the dynamics on the attractor are somewhat complicated in the sense that the attractor cannot be decomposed further, and if the attractor is infinite then there is a possibility of complicated dynamics, and a particular well-studied phenomenon of such complexity gives rise to what is called a chaotic behaviour.
+
+If the dynamics on an attractor cannot be decomposed further and the attractor is itself also infinite, then there is a possibility of complicated dynamics, and a particular well-studied phenomenon of such complexity gives rise to what is called chaotic behaviour.
+
+\subsection{Chaos}
+
+\emph{$\downarrow$ - should one add paragraphs like these?}
+\newline In the 1960s, a number of mathematicians and mathematically interested scientists independently discovered chaos in the mathematical sense.
+The meteorologist Edward Lorenz may have been the first to explain this phenomenon in his 1963 paper \cite{lorenz1963deterministic}.
+The notions of invariance, attractivity and chaos may also bedescribed for continuous systems, and Lorenz's system comprised of a system of differential equations.
 The narrative of Lorenz's discovery of chaos, as well as the history of other forerunners in this subject, are fascinating.
@@
 
-There are several intricately different notions of chaos each of them indicating some complexity.
-In practice, it is not possible to verify which notion is satisfied when data is observed from a dynamical system.
+There are several intricately different notions of chaos, with each of them indicating some complexity.
+\emph{ $rightarrow$ what are we saying here?}In practice, it is not possible to verify which notion is satisfied when data is observed from a dynamical system.
 However, for the sake of understanding some nuances, we recall the notion of the definition of chaos in the sense of Devaney \cite{devaney2018introduction,de2013elements}.
-Devaney's definitio has three ingredients, and the first one is the notion of sensitive dependence on initial conditions.
+Devaney's definition has three ingredients, and the first one is the notion of sensitive dependence on initial conditions.
 
@@
  \everypsbox{\scriptstyle}
-%  \begin{psmatrix}
-% U & U\\%
-% V & V.
-%  %%%
+ \begin{psmatrix}
+U & U\\%
+V & V.
+ %%%
 %  \ncline{1,1}{1,2}^{T} 
@@
 %  \ncline{1,2}{2,2} > {\phi}
-%  \end{psmatrix}
+ \end{psmatrix}
 % \]

### 2021-12-19 Gandhi_Project_Report/draft1.tex
Halfway through 2.3-Conjugacy

@@
 \newtheorem{Remark}{Remark}[]
+\newtheorem{Example}{Example}[]
 
@@
 There are several intricately different notions of chaos, with each of them indicating some complexity.
-\emph{ $rightarrow$ what are we saying here?}In practice, it is not possible to verify which notion is satisfied when data is observed from a dynamical system.
-However, for the sake of understanding some nuances, we recall the notion of the definition of chaos in the sense of Devaney \cite{devaney2018introduction,de2013elements}.
-Devaney's definition has three ingredients, and the first one is the notion of sensitive dependence on initial conditions.
+\emph{ $\leftarrow  $ what are we saying here?} In practice, it is not possible to verify which notion is satisfied when data is observed from a dynamical system.
+We do, however, specifically recall the definition of chaos in the sense of Devaney \cite{devaney2018introduction,de2013elements} so as to understand some nuances.
+Devaney's definition consists of three parts, with the first being the notion  of sensitive dependence on initial conditions.
 
 \begin{Definition} \rm 
-A dynamical system $T: U \to U$ is said to have sensitive dependence on initial conditions (SDIC) if there exists a $\delta > 0$ such that for every $u \in U$ and in every neighborhood of $u \in U$ there exists an $v$ and an integer $N$ such that $d(T^Nu,T^Nv)>\delta$.
+A dynamical system $T: U \to U$ is said to have sensitive dependence on initial conditions (SDIC) if there exists a $\delta > 0$ such that for every $u \in U$ and in every neighborhood of $u \in U$ there exists a $v\in{U}$ and an integer $N:=N{(u,v)}\in\mathbb{Z}$ such that $d(T^Nu,T^Nv)>\delta$.
 \end{Definition}
 
-This definition has a lot of quantifiers in it to analyse, so we should take our time to acquire a sense of what it says.
-The common misconception is to interpret SDIC as two close points ($u$ and $v$) eventually become separated by a distance $\delta$ under iteration by $T$.
+It is important to acquire a sense of this definition as it can easily be misunderstood.
+It is common misconception to interpret SDIC as two close points ($u$ and $v$) that eventually become separated by a distance $\delta$ under iteration by $T$.
 But this is not true.
-To completely comprehend all of the subtleties of the concept, we need to talk about it more thoroughly.
-The simplest way to achieve this is to go through each sentence in the definition slowly and attentively, then examine how they fit together to convey the concept of sensitive dependence.
+To fully comprehend all the subtleties of the concept, one needs to discuss it more thoroughly by considering each sentence with care and attention, whereafter one may examine how they fit together to convey the concept of sensitive dependence.
+\emph{ $\leftarrow$ is this talking too simply perhaps?}.
 
-First, the $\delta>0$ in the definition of SDIC should be independent of $u$.
-Second, in every neighborhood of $u$, we may not necessarily find all points $v$ in the neighborhood distinct from $u$ that would separate from the forward iterates of $u$.
-Next, $N$ depends upon $u$ and $v$ chosen, and their iterates may not separate for ever (i.e., for all $n>N$) and we allow their iterates to get arbitrarily close in the future.
+To this end we make 3 remarks:
+\vspace{-5mm}
+\begin{enumerate}
+  \item First, the $\delta>0$ in the definition of SDIC is independent of $u$. 
+  \item Second, in every neighborhood of $u$, we may not necessarily find all points $v$ in the neighborhood distinct from $u$ that would separate from the forward iterates of $u$. 
+  \item  Finally, $N$ depends upon $u$ and $v$ chosen, and their iterates may not separate for ever (i.e., for all $n>N$) and we allow their iterates to get arbitrarily close in the future. 
+\end{enumerate}
 
@@
 
-We next recall the next ingredient of Devaney's definition of chaos.
+The second part in Devaney's definition of chaos concerns topological transitivity.
 
 \begin{Definition}
-	A dynamical system $F: U \to U$ is topologically transive if for any pair of nonempty open sets $E_1$ and $E_2$ there exists a $n>0$ such that $T^n(E_1) \cap E_2 \not= \emptyset$. 
+	A dynamical system $F: U \to U$ is topologically transitive if for any pair of nonempty open sets $E_1$ and $E_2$ there exists a $n\in\mathbb{N}$ such that $T^n(E_1) \cap E_2 \not= \emptyset$. 
 \end{Definition}
@@
 Topologically transitivity implies that the iterates of an open set of initial conditions gets mixed up with other open sets.
-On a compact metric space, it can be shown that topological transitivity also implies the existence of a point whose forward iterates are dense.
-In fact, it is more topologically likely that one picks a point whose iterates are dense in the sense that the complement of the set of points whose forward iterates are not dense in $U$  are topologically insignificant -- such an insignificant set can be written as a countable union of sets with empty interior.
+On a compact metric space, it can be shown that topological transitivity also implies the existence of a point whose forward iterates are dense; or in other words, the orbit going through this point will be dense in the compact metric space.
 
-Now we recall the definition of chaos in the sense of Devaney \cite{devaney2018introduction}.
+In fact, it is topologically more likely that the choice of an arbitrary point will be one  whose iterates are almost dense.
+% I inserted a definition here: it makes more sense to me. 
+\begin{Definition}
+A compact metric space $U$ is said to be almost dense if the complement of the set of points which are dense in $U$ may be written as a countable union of sets of with empty interior.
+In other words, the set on points not dense on $U$ is topologically insignificant.
+\end{Definition}
+
+We may now define the notion of Chaos as formulated by Devaney\cite{devaney2018introduction}.
 \begin{Definition}
 	A dynamical system $T: U \to U$ is said to exhibit chaos in the sense of Devaney if it satisfies the three properties:
+	\vspace{-5mm}
   \begin{enumerate}
-		\item $T$ has SDIC
-		\item $T$ has topologically transitive
-		\item the set of periodic points of $T$ are dense in $U$. 
+		\item $T$ has SDIC.
+		\item $T$ is topologically transitive.
+		\item The set of periodic points of $T$ are dense in $U$. 
 	\end{enumerate}
@@
 
-An example of a dynamical system that satisfies these three properties is the standard tent map $T(u)=1-|2u-1|$ defined on $[0,1]$.
-We now reason as why this map exhibits chaos in the sense of Devaney.
-The graph of the map $T$ is piecewise linear with two straight lines, one that connects the points $(0,0)$ and $(1/2,1)$ and the other that connects $(1/2,1)$ to $(1,0)$ and they form a tent with a base centered at $u=1/2$.
+\begin{figure}
+  \centering
+  \begin{minipage}{.5\textwidth}
+    \centering
+    \includegraphics[scale=0.18]{chaos1.png}
+    \captionof{figure}{Starting to diverge}
+    \label{fig:chaos1}
+  \end{minipage}%
+  \begin{minipage}{.5\textwidth}
+    \centering
+    \includegraphics[scale=0.18]{chaos2.png}
+    \captionof{figure}{Completely chaotic}
+    \label{fig:chaos2}
+  \end{minipage}
+  \end{figure}
+
+\begin{Example}
+  The standard tent map $T(u)=1-|2u-1|$ defined on $[0,1]$ is a common example of a dynamical system satisfying these three properties. 
+  \begin{figure}
+    \includegraphics[scale=0.6]{T2.png}
+    \centering
+    \label{fig:T2tentmap}
+  \end{figure}
+
+  Let us now consider why this map exhibit's Devaneys's sense of chaos.
+The graph of the map $T$ is piecewise linear with two straight lines, one connecting the points $(0,0)$ and $(1/2,1)$ and the other connecting $(1/2,1)$ with $(1,0)$.
+They form a so-called tent with base centered at $u=1/2$.
 The graph of the map $T^2$ comprises two symmetric tents with their base centered at $1/4$ and $3/4$.
-So given any point $u$ and an open neighborhoow $W  \subset [0,1]$ of $u$, we can find $k$ large enough so that $T^k(W) = [0,1]$ since since we can accommodate a tent whose base is contained in $W$.
-This implies that there are two points $w_1$ and $w_2$ in $W$ so that $d(T^kw_1,T^kw_2)= 1$, and by triangle inequality  at least one of the inequalities $d(T^ku,T^kw_1)> \delta$ or $d(T^ku,T^kw_2)> \delta$ hold when $\delta\in (0,1/2)$.
-So $T$ has sensitive dependence on initial conditions.
+  % We now reason as why this map exhibits chaos in the sense of Devaney.
+
+  Given any point $u$ and an open neighborhood $W \subset[0,1]$ of $u$, we can find $k\in\mathbb{Z}$ large enough so that $T^k(W) = [0,1]$ because we can accommodate a tent whose base is contained in $W$.
+\emph{This last phrase doesn't make sense to me.} This implies the existence of two points $w_1$ and $w_2$ in $W$ so that $d(T^kw_1,T^kw_2)= 1$, and by the triangle inequality  at least one of the inequalities $d(T^ku,T^kw_1)> \delta$ or $d(T^ku,T^kw_2)> \delta$ hold when $\delta\in (0,1/2)$.
+\newline So $T$ has sensitive dependence on initial conditions.
+
   Next, let $W_1$ and $W_2$ be two nonempty open sets.
-Since given any open set $W_1$, we can find a $k$ large enough so that $T^k(W_1) = [0,1]$ since we can accommodate a tent whose base is contained in $W$.
+Given any open set $W_1$, we can find a $k\in\mathbb{N}$ large enough so that $T^k(W_1) = [0,1]$ because we can accommodate a tent with base contained in $W$.
 So $T^k(W_1) \cap W_2 \not=\emptyset$, and thus $T$ has topological transitivity.
-Next, since for every open interval $(a,b)$,  there exists a $k$ such that $T^k(a,b) = [0,1]$, the graph of $T^k$ intersects the graph of the identity map on $[0,1]$, and if intersection has coordinates $(p,p)$,  the point $p$ is a fixed point of $T^k$, which is a periodic point of $T$.
-Thus the set of periodic points are dense.
 
-We next talk on another concept of conjugacy. 
-To show two metric or topological spaces are topologically the same one needs to show that they are homeomorphic, i.e., to establish a homeomorphism between the spaces.
-In the study of dynamical systems defined on two spaces, showing two spaces to be homeomorphic does not mean that the two dynamical systems are somehow related.
-One needs to establish some kind of dynamical equivalence.
-The easiest way to illustrate dynamical equivalence between two systems $T:U\to U$ and $S:V\to V$  is through a commutativity diagram:
+  Finally, for every open interval $(a,b)$, the graph of $T^k$ intersects the graph of the identity map on $[0,1]$.
+This follows from the fact established above that there exists a $k\in\mathbb{N}$ such that $T^k(a,b) = [0,1]$.If the intersection point has coordinates of the form $(p,p)$, the point p is then a fixed point of $T^k$ and therefore also a periodic point of T.
+It follows that the set of periodic points are dense in T.
+  \textbf{Expand further on reason for density.}
+\end{Example}
 
+\subsection{Conjugacy}
+
+We now turn our attention to the subject of conjugacy.
+
+To show topological similarity or sameness between two metric or topological spaces, one needs to establish a homeomorphism between the two spaces. 
+However, in the study of dynamical systems defined on two spaces, establishing a homeomorphism does not indicate that the systems are (physically) related in any way.
+\textbf{"related" - in what sense am I saying it here?
+Physically?}  For this one rather needs to establish a kind of dynamical equivalence which we illustrate in a simple commutativity diagram below.
 % Giving errors

### 2021-12-19 graphing.py

@@ -1,0 +1,26 @@
+def T(u):
+    return 1 - abs(2*u-1)
+
+def Tn(n, U):
+    for i in range(n):
+        U = T(U)
+    return U
+
+import numpy as np
+X1 = np.linspace(0,1,1200)
+Y= T(T(X1))
+
+import matplotlib.pyplot as plt
+
+# plt.plot(X1, Y, '-g')
+# plt.title(r'Plot of the map $T^2$, $T=1-|2u-1|$')
+# plt.xlabel(r'$x_n$')
+# plt.ylabel(r'$x_{n+2}$')
+# plt.savefig("T2.png")
+# plt.show()
+
+plt.plot(X1, Tn(54, X1))  #52=close, 53=line
+plt.show()
+
+# plt.plot(X3, T(X3), '-k')
+# plt.show()

### 2021-12-20 Gandhi_Project_Report/draft1.tex
Started with Section 3

@@
 \end{equation}
+
+%Composite functions
 To understand the diagram we note that if we travel ``right and then down," the diagram instructs us to use $T$ (top arrow right) first, followed by $\phi$.
 (right arrow downwards).
-As a result, the pathway amounts to finding $\phi(T(u))$.
-If we proceed ``down and then right," the diagram instructs us to apply $\phi$ (left arrow downwards) first, and then apply $\phi$ (right arrow downwards) and then $S$ (bottom arrow right).
-As a result, the pathway amounts to finding  $S(\phi(u))$.
-When the diagram holds, we say
-$\phi(T(u))= S(\phi(u))$ and write it more formally as
-$\phi \circ T=S\circ \phi$.
-When the above commutativity holds in addition that $\phi$ is a homeomorphism, then we say we say that  $S$ is conjugate to $T$.  
-If we relax the criterion on $\phi$ to merely require $\phi$ to be continuous and surjective rather than being a homeomorphism, then $\phi$ is called a semi-conjugacy, and declare that $S$ is semi-conjugate to $T$.
-When $S$ is conjugate to $T$, it indicates the dynamics between the two systems are in one-to-one correspondence, however when $S$ is semi-conjugate to $T$ with $\phi$ a many-to-one mapping, the dynamics on $V$ provides a coarse-grained description of the dynamics on $V$.
-It is common to remark that $S$ is a factor of $T$ or that $T$ is an extension of $S$ when $S$ is semi-conjugate to $T$.
-In essence, an extension (e.g., \cite{de2013elements}) is a larger system that captures all of the important dynamics of its factor.
+Consequently the pathway amounts to finding $\phi(T(u))$.
+If we proceed "down and then right", the diagram instructs us to apply $\phi$ (left arrow downwards) first, and then apply $S$ (bottom arrow right).
+This then amounts to finding  $S(\phi(u))$.
+When the relationships in this diagram hold, we say $\phi(T(u))= S(\phi(u))$ and formally denote it as $\phi \circ T=S\circ \phi$.
+
+\begin{Definition}
+  Consider the dynamical systems $T:U\to{U}$, $S:V\to{V}$ and suppose the relationship $\phi \circ T=S\circ \phi$ holds.
+If $\phi$ is a homeomorphism, then S is said to be conjugate to T and $\phi$ is called a conugacy.
+If we relax the criterion on $\phi$ and merely require $\phi$ to be continuous and surjective, $\phi$ is then a semi-conjugacy between $T$ and $S$; S is said to be semi-conjugate to T.
+\end{Definition}
+
+When $S$ is conjugate to $T$, the dynamics of the two systems are in some way ''dynamically equivalent'' and specifically they are in one-to-one correspondence with one another.
+However when $S$ is semi-conjugate to $T$ with $\phi$ a many-to-one mapping, the dynamics on $V$ provides a coarse-grained \textbf{cite?} description of the dynamics on $U$ .
+When $S$ is semi-conjugate to $T$, it is also common to call $S$ a factor of $T$, or conversely that $T$ is an extension of $S$.
+In essence, an extension (e.g., \cite{de2013elements}) is a larger system capturing all of the important dynamics of its factor.
 
 It is a very hard or nearly an impossible task to establish the existence of a conjugacy or a semi-conjugacy $\phi$ between two systems.
-However, one can verify if given a function $\phi$ if it satisfies the commutativity.
-For example $\phi(x)=\sin(\pi x/2)^2$ is a conjugacy between two systems $Tu=1-|2u-1|$ and $Sv=4v(1-v)$  defined on $[0,1]$.
-The importance of conjugacy is that when we know it exists we can prefer to work with  the dynamical systems to gain information of the other, and this is useful for us while forecasting dynamical systems.
+\textbf{cite?} However, one can verify that a function $\phi$ satisfies the commutativity diagram \ref{eqn_conjugacy}.
+For example $\phi(x)=\sin(\pi x/2)^2$ is a conjugacy between two systems $Tu=1-|2u-1|$ and $Sv=4v(1-v)$  defined on $U=[0,1]$.
+
+In establishing that two systems are conjugate (semi-conjugate) to one another, we have also shown that one may choose to work with one systems as opposed to the second and still be guaranteed to obtain information on the latter.
+This is incredibly useful while forecasting the future evolution of dynamical systems.
+
+\section{The Learning Problem}
+
+\emph{Introduction...}
+
+Consider a relatively simple learning problem:
+
+Given $(u_0, u_1, \ldots, u_m)$ for $m\in\mathbb{Z}$, a finite segment of an orbit of the map $T$ where $T$ is defined by the update equation $u_{n+1} = Tu_n$, forecast the values $u_{m+1}, u_{m+1}$ where the map $T$ is unknown, given that $u_m$.
+
+A practical example of this would be - given the sequential coordinates of an object moving in space, predict the future positions of that object.
+However we are very rarely, if at all, presented with a problem where the entire state information is available to us in the form $u_n$ at some time-step $n\in\mathbb{N}$.
+Consider then a more involved learning problem:
+
+Suppose we only have the observations $\theta(u_0), \theta(u_1), \ldots, \theta(u_m)$ of the true system states $u$ in an unknown dynamical system $(U,T)$ and we wish to predict the values $\theta(u_{m+1}), \theta(u_{m+2})$.
+
+Here we consider a continuous-time dynamical system $(U,T)$ with true state $u$ in the input space $U$ and then a measured signal $\theta$ (where $\theta(u_n)=\theta(u(t_n))$ for $n\in\mathbb{N}$.
+The sequence $\{x_n\}$ represents a scalar time-series and intuitively $\theta$ may be thought to represent a probe inserted into a larger system and itself measuring/extracting only a small part of the greater system state at time $t=n$.
+Consider for example a thermometer erected to measure the ambient temperature in a local village.
+This measurement function, the thermometer is capturing a single aspect, the temperature, of a grander dynamical system entailing the present weather of the surrounding area, but even more than that, it is measuring an incredibly small part of the global weather system.
+
+From here we construct a multidimensional observable using the method of delay-coordinate stacking. 
+\textbf{move to later} Given the sequence $\{\theta_1, \theta_2, \ldots\}$, we define the variable $y_n:=[\theta_n, \theta_{n-\tau}, \ldots, \theta_{n-(2d-2)\tau}, \theta_{n-(2d-1)\tau}]^{T}\in\mathbb{R}_{d}\times{1}$ where $\tau$ represents the lag and $d$ the dimension of the attractor.
+(We shall return to these terms shortly and their discussion may be put on hold for the time being).
+
+However, the true state $u$ of a system is seldom, if ever fully known.
+In almost all cases we can merely insert probes into a system to obtain partial information by measurements taken.
+Moreover, the process of taking a measurement itself introduces 2 additional difficulties:
+\vspace{-8mm}
+\begin{enumerate}[noitemsep, label=\roman*.]
+  \item A series of measurements over a specific time-interval is inherently a discretisation
+  procedure of the underlying continuous-time dynamical system. (Hence why we restricted our attention to discrete-time systems in the preceding section)
+  \item A probe will never be fully accurate and so the act of measurement introduces a certain
+measure of numerical noise/inaccuracy.
+\end{enumerate}
+
+More formally we may say that in the process of measuring a system, we obtain a 1-dimensional time-series $\{\theta(u(t))\}$, where $\theta:\mathbb{R}^n\to\mathbb{R}$ is a smooth observation or measurement function which is probably also introducing noise
 

### 2021-12-20 Gandhi_Project_Report/draft1.tex
nearly done with first draft of chapter 3. Definitely still need introduction and conclusion as well as adding graphs and their explanations

@@
 \newtheorem{Example}{Example}[]
+
+%Shortcut symbols
+\newcommand{\Ftheta}{\ensuremath{F_\theta}}
 
@@
 To see this, consider differential equation $\dot{x} = f(x)$, $f: \mathbb{R}^n \to \mathbb{R}^n$, $n\in\mathbb{N}$ given to have a unique solution passing through
-each point $x\in\mathbb{R}^{n}$, the \emph{flow of the equation} is defined to be a mapping $\varphi: \mathbb{R}^n \times \mathbb{R} \to \mathbb{R}$ that satisfies the following properties:
+each point $x\in\mathbb{R}^{n}$, the \emph{flow of the equation}\label{defn_flow} is defined to be a mapping $\varphi: \mathbb{R}^n \times \mathbb{R} \to \mathbb{R}$ that satisfies the following properties:
 \vspace{-8mm}
@@
 
-\begin{figure}
+\begin{figure}[ht]
   \centering
@@
   The standard tent map $T(u)=1-|2u-1|$ defined on $[0,1]$ is a common example of a dynamical system satisfying these three properties. 
-  \begin{figure}
+  \begin{figure}[ht]
     \includegraphics[scale=0.6]{T2.png}
@@
 From here we construct a multidimensional observable using the method of delay-coordinate stacking. 
-\textbf{move to later} Given the sequence $\{\theta_1, \theta_2, \ldots\}$, we define the variable $y_n:=[\theta_n, \theta_{n-\tau}, \ldots, \theta_{n-(2d-2)\tau}, \theta_{n-(2d-1)\tau}]^{T}\in\mathbb{R}_{d}\times{1}$ where $\tau$ represents the lag and $d$ the dimension of the attractor.
+\textbf{move to later} Given the sequence $\{\theta_1, \theta_2, \ldots\}$, we define the variable \newline $y_n:=[\theta_n, \theta_{n-\tau}, \ldots, \theta_{n-(2d-2)\tau}, \theta_{n-(2d-1)\tau}]^{T}\in\mathbb{R}_{d}\times{1}$ where $\tau$ represents the lag and $d$ the dimension of the attractor.
 (We shall return to these terms shortly and their discussion may be put on hold for the time being).
@@
 
-More formally we may say that in the process of measuring a system, we obtain a 1-dimensional time-series $\{\theta(u(t))\}$, where $\theta:\mathbb{R}^n\to\mathbb{R}$ is a smooth observation or measurement function which is probably also introducing noise
+More formally we may say that in the process of measuring a system, we obtain a 1-dimensional time-series $\{\theta(u(t))\}$, where $\theta:\mathbb{R}^n\to\mathbb{R}$ is a smooth observation or measurement function.
+This series represents a possibly noise-augmented dataset containing information of certain parts of the original system.
+
+We may then ask ourselves the question: Is it in any way possible to retain information about the system state $x(t)$ in this temporal data-series $\theta(x(t))$?
+The answer is easily yes if T is known.
+However, neither this, nor even the exact function $\theta$ is available.
+
+\subsection{Taken Embedding Theorem}
+
+We define first the concepts of homeomorphism and embedding:
+\begin{Definition}
+  A homeomorphism is a function $f:Z\rightarrow Y$ between two topological spaces $Z$ and $Y$ that is continuous, bijective and has a continuous inverse. 
+\end{Definition}
+
+\begin{Definition}
+  Consider a homeomorphism $f:Z\rightarrow Y$ for $Y\in X$. $Z$ is said to be embedded in $X$ by $f$.
+\end{Definition}
+
+Takens Theorem states a result that establishes a relationship between the observed and underlying dynamical systems by showing that the concetantion of a sufficiently large number of previous observations into a vector will, under certain conditions, generate a map between the vectors from the respective systems.
+\textbf{check wording} We formultate the theorem from \cite{takens1981detecting}.
+
+\begin{Theorem} 
+	[\bf Takens Embedding Theorem (adopted from \cite{takens1981detecting}] \label{Thm_Takens}
+            Let $W$ be a compact manifold of dimension $m$, and $d\ge m$ so that $2d\in\mathbb{Z}$. It is a 
+            generic property for the pair $(T, \theta)$,  where $T:W \to W$ is
+            a smooth diffeomorphism, and $\theta:W \to \mathbb{R}$ a smooth function, that the map $\Phi_{2d,\theta}:W \to \mathbb{R}^{2d+1}$ defined on $W$ by 
+            $\Phi_{2d,\theta}(w) := (\theta(T^{-2d}w)\ldots,\theta(T^{-1}w),\theta(w))$
+            is a diffeomorphic embedding; by `smooth' we mean at least $C^2$. Consequently, there exists a map $\Ftheta: \Phi_{2d,\theta}(W) \to \Phi_{2d,\theta}(W)$ defined by 
+            \begin{equation}  \label{eqn_takens}
+              \Ftheta: (\theta(T^{-2d}w),\ldots,\theta(T^{-1}w),\theta(w)) \mapsto  (\theta(T^{-2d+1}w),\ldots,\theta(T^{-1}w),\theta(Tw))
+            \end{equation}
+           so that $(W,T)$ is topologically conjugate to 
+            $(\Phi_{2d,\theta}(W), \Ftheta)$. 
+\end{Theorem}
+
+(In our scenario, the input space $U$ is considered the attractor.)
+
+Some explanation is beneficial.
+Takens establishes a delay-coordinate map $\Ftheta$ defined in \ref{eqn_takens} for $T$ the flow (\ref{defn_flow}) defined by the update equation  $Tu_n=u_{n+1}$ and $theta$ our measurement function.
+When confining a dynamical system to the manifold U (in which the underlying system is contained), Takens showed that if certain smoothness conditions are satisfied on $T$ and $\theta$, then the delay-coordinate map $\Ftheta$ embeds $U$ in the reconstruction space $R^{2d+1}$ \textbf{we haven't used the term reconstruction space yet, so this is falling out of the blue}, for almost every choice of measurement function $\theta$, the observable.
+By almost every choice here, we mean that we exclude some observations which do not give any information.
+To see a more precise discussion, see \textbf{xxx} for we do not concern ourselves with a further elaboration on this thesis/report.
+
+Alternatively we may also make the statement: Takens' Embedding Theorem guarantees that almost all dynamical systems can be reconstructed from just one noiseless observation sequence \textbf{this isn't obvious from the theorem's formulation, is it?} i.e.
+for a great number of possible observation functions $\theta$, $\Ftheta$ preserves the topology of U.
+\textbf{'preserving the topology' is not something we've touched on before.
+Perhaps I should add to previous discussion so logic flows seamlessly here}.
+
+\begin{figure}[ht]
+  \includegraphics[scale=0.3]{takensmap.png}
+  \centering
+  \label{fig:takensmap}
+\end{figure}
+
+Once again we pause and consider the graph to consolidate our understanding of the conjugacy (\textbf{are we being too simplistic?}).
+If we are in state $u$ and then evolve towards state $Tu$ before embedding into $\mathbb{R}^{2m-1}$, then this is equivalent to first embedding into $\mathbb{R}^{2m-1}$ and then evolving through the map $\Ftheta$, i.e.
+$\Phi_{2d,\theta}\circ{F}_{\theta} = T\circ\Phi_{2d,\theta}$.
+The map $F_theta$ is a homeomorphism as well.
+
+Thus information about U can be retained in the time series' (or observation's) output.
+By preserving the topology on the manifold $U$ in the reconstruction space $X$ (\textbf{not discussed before}), we are also guaranteed the preservation of topological invariants of the manifold, of which dimensionality is one such invariant.
+(We note here that dimensionality will again be considered later in this project/report).
+
+\subsection{Obstacles to Takens}
+Despite the fact that the Takens' Embedding Theorem is a powerful result and provides compelling reason to believe that it is possible to accurately ()\emph{right word?}) reconstruct the underlying space, it does present some serious limitations:
+\vspace{-5mm}
+\begin{enumerate}
+\item Even if we can indeed find \Ftheta{}  our \emph{approximation} of \Ftheta{}  is a map from a larger set $\mathbb{R}^{2m+1}$ containing the embedded attractor.
+There are, however, no theoretical guarantees that \Ftheta{}  will keep this attractor $U$  the same.(\textbf{We haven't properly said that U is an attractor, or have we.})
+\item Takens only holds for noiseless observations.
+Suppose the case where \Ftheta{} was learnt, but due to measurement error or other noise we obtained the set $V$ and not the original attractor.
+Due to the chaotic nature of the underlying system (i.e.
+the fact that it has SIDC), this new system $V$ will then move in a different direction.
+This problem will later be eliminated by some properties of a new function $g$, namely the Unique Solution Property.
+\item It is an unfortunate case that the functional complexity (i.e., the relative number of oscillatory graphs \cite{manjunath2021universal}) as well as stability of the embedding \Ftheta{} (\textbf{cite?})depends on the choice of observable $\theta$.
+This means that we may even learn a map that complicates our learning task and in so doing adds obstacles to the road to learning a conjugate map with predictive power.
+(\textbf{Am I talking nonsense here?})
+\end{enumerate}
+
+Another limitation owes itself to the requirement that T be a surjective map. This is not always the case in practical examples and we may consider the case of a damped double pendulum.
+
+\textbf{Continue writing here}
+
+\begin{figure}[ht]
+  \includegraphics[scale=0.3]{learningfailure.png}
+  \centering
+  \label{fig:supp_learning_failure}
+  \caption{Illustration of a system having failed to learn the map}
+\end{figure}
+
+\textbf{Explanation continued here}
+Traalala lorem ipsum et dolore faria maria didi klapogaani koro di tolore mi strem srem krem lem lemmy bemmy bem te dem dem
+
+\begin{figure}[h]
+  \includegraphics[scale=0.4]{temporary_damp_fig.png}
+  \centering
+  \label{fig:damped_pendulum}
+\caption{A damped pendulum}
+\end{figure}
+
+\textbf{Continue explaining here}
 

### 2021-12-21 Gandhi_Project_Report/draft1.tex
Section 4 - 3/4's done but need to add the flesh in between

@@
 One reason for this fragility is that the embedded attractor in the reconstruction space not always an attractor itself, despite  unquestionably being an invariant set.
-Since the embedded object is not \emph{(definitely/always?)} an attractor (as explained in Chapter 3) it can cause predictions to fail.
+Since the embedded object is not \textbf{(definitely/always?)} an attractor (as explained in Chapter 3) it can cause predictions to fail.
 
@@
 
-This thesis deals with the implementation and analysis of non-interpretable models that can guarantee \emph{exact/accurate} reconstruction.
+This thesis deals with the implementation and analysis of non-interpretable models that can guarantee \textbf{exact/accurate} reconstruction.
 The project work concerns the study and implementation of a method (Citation to paper with Adriaan) that incorporates learning a function by mapping the data on to a higher dimensional space.
@@
 \newline In Section 3, we introduce the problem of forecasting dynamical systems, state the Takens delay embedding theorem and discuss various issues faced while forecasting. 
-\newline In Section 4, a driven dynamical system is defined and discuss the properties of a specific class of driven dynamical systems that we make use of in this theses/report/paper. 
+\newline In Section 4, a driven dynamical system is defined and discuss the properties of a specific class of driven dynamical systems that we make use of in this thesis/report/paper. 
 \newline Finally in Section 5 we show the implementation of these forecasting methods, and conclusions are provided in Section 6.
@@
 
-At its most elementary level, a dynamical system is just \emph{something} that evolves deterministically through time.
+At its most elementary level, a dynamical system is just something that evolves deterministically through time.
 In the context of this thesis, deterministic refers to the fact that a system evolves according to specified rules rather than based on random events.
@@
 To see this, consider differential equation $\dot{x} = f(x)$, $f: \mathbb{R}^n \to \mathbb{R}^n$, $n\in\mathbb{N}$ given to have a unique solution passing through
-each point $x\in\mathbb{R}^{n}$, the \emph{flow of the equation}\label{defn_flow} is defined to be a mapping $\varphi: \mathbb{R}^n \times \mathbb{R} \to \mathbb{R}$ that satisfies the following properties:
+each point $x\in\mathbb{R}^{n}$, the flow of the equation \label{defn_flow} is defined to be a mapping $\varphi: \mathbb{R}^n \times \mathbb{R} \to \mathbb{R}$ that satisfies the following properties:
 \vspace{-8mm}
@@
 Of course, discrete-time dynamical systems need not always arise through a differential equation, and once again we may consider the field of ecology where discrete dynamical systems are often directly derived or assumed.
-\emph{There is a school of thought that advocates discrete-time dynamical systems to be more natural for modelling real-world observations than differential equations (see \cite{saber2010introduction}).
-<- too philosphical, or actually good?}
+\emph{There is a school of thought that advocates discrete-time dynamical systems to be more natural for modelling real-world observations than differential equations (see \cite{saber2010introduction})}.
+\textbf{too philosphical, or actually good?}
 
@@
 
-\textbf{In fact the $\omega$-limit set $\omega(u;T)$ of a point $u$ defined to the collection of limit points of the sequence $\{x,Tu,T^2u,\ldots\}$ is nonempty, and $\omega(u;T)$ is invariant.
-} (not relevant other than being an additional example perhaps?)
+In fact the $\omega$-limit set $\omega(u;T)$ of a point $u$ defined to the collection of limit points of the sequence $\{x,Tu,T^2u,\ldots\}$ is nonempty, and $\omega(u;T)$ is invariant.
+\textbf{(not relevant other than being an additional example perhaps?)}
 
@@
 
-\emph{$\downarrow$ - should one add paragraphs like these?}
+\textbf{$\downarrow$ - should one add paragraphs like these?}
 \newline In the 1960s, a number of mathematicians and mathematically interested scientists independently discovered chaos in the mathematical sense.
@@
 There are several intricately different notions of chaos, with each of them indicating some complexity.
-\emph{ $\leftarrow  $ what are we saying here?} In practice, it is not possible to verify which notion is satisfied when data is observed from a dynamical system.
+\textbf{ $\leftarrow  $ what are we saying here?} In practice, it is not possible to verify which notion is satisfied when data is observed from a dynamical system.
 We do, however, specifically recall the definition of chaos in the sense of Devaney \cite{devaney2018introduction,de2013elements} so as to understand some nuances.
@@
 To fully comprehend all the subtleties of the concept, one needs to discuss it more thoroughly by considering each sentence with care and attention, whereafter one may examine how they fit together to convey the concept of sensitive dependence.
-\emph{ $\leftarrow$ is this talking too simply perhaps?}.
+\textbf{ $\leftarrow$ is this talking too simply perhaps?}.
 
@@
 Takens Theorem states a result that establishes a relationship between the observed and underlying dynamical systems by showing that the concetantion of a sufficiently large number of previous observations into a vector will, under certain conditions, generate a map between the vectors from the respective systems.
-\textbf{check wording} We formultate the theorem from \cite{takens1981detecting}.
+\textbf{Check wording} We formultate the theorem from \cite{takens1981detecting}.
 
@@
 \textbf{Explanation continued here}
-Traalala lorem ipsum et dolore faria maria didi klapogaani koro di tolore mi strem srem krem lem lemmy bemmy bem te dem dem
 
@@
 
+\section{Obtaining X through Driven Dynamical Systems}
+
+In this chapter, we recall results on mapping temporal data acquired from a discrete-dynamical system onto a different space through a driven dynamical system.
+We recall the conditions the driven system should have not to add distortion to its state space representation, so that the single-delay dynamics (SDD) gets conjugate or semi-conjugate to the system.
+The SDD can be then use to forecast and reconstruct the underlying attractor
+
+\subsection{Nonautonomous and Driven Dynamical Systems}
+
+A nonautonomous dynamical system is simply a dynamical system where the input $u$ from the input space $U$, a  topological space is time-dependent.
+\textbf{This means it can no longer be generated by a flow?}
+
+We now recall the figure first defined in \ref{eqn_conjugacy} below and the greater goal to construct a system - such as $(V,S)$  - conjugate to $(U,T)$,
+
+%#####################################
+\begin{center}
+\psset{arrows=->, arrowinset=0.25, linewidth=0.6pt, nodesep=3pt, labelsep=2pt, rowsep=0.7cm, colsep = 1.1cm, shortput =tablr}
+\everypsbox{\scriptstyle}
+\begin{psmatrix}
+U & U\\%
+V & V.
+%%%
+%  \ncline{1,1}{1,2}^{T} 
+%  \ncline{1,1}{2,1} <{\phi}
+%   \ncline{2,1}{2,2}^{S}
+%  \ncline{1,2}{2,2} > {\phi}
+\end{psmatrix}
+%#####################################
+\end{center}
+
+To do this, we need to consider a driven dynamical system taking input from both an input space $U$, and an underlying state space $X$.
+This is done to account for influence of both the input and the actual state of the system at a specific timestep.
+
+\begin{Definition}
+A driven dynamical system comprises two topological spaces $U$, $X$ and a continuous function  $g:U\times{X}\to{X}$ where $g(u_n, x_n)=x_{n+1}$.
+The dynamics on X are generated by the update equation $x_{n+1}=g(u_n, x_n)$ where $n\in\mathbb{Z}$, $u_n$ from the metric input space $U$ and state $x_n$ belonging to the state space $X$, a compact metric space.
+If the input space $U$ is compact, we refer to the system as compactly driven.
+Abbreviated, we shall refer only to the "driven system $g$", with all other entities being understood implicitly.
+\end{Definition}
+
+In particular, a nonautonomous dynamical system may be generated from $U$; any input $\overline{u}$, a bi-infinite sequence from $U$ gives rise the sequence of self-maps $\{g(u_n, \ldotp)\}_{n\in\mathbb{Z}}$ contained in $X$.
+\textbf{explain better why}
+Physically, one may think of this bi-infinite as referring to a system that has been running for an incredibly long period at the time of the first measurement/observation taken from the system.
+
+\begin{Definition}
+  A sequence $\{x_n\}_{n\in\mathbb{Z}}\subset{U}$ is called an entire solution(or simply a solution) to the driven system  $g$ with input $\overline{u}$ when it satisfies 
+  $$g(u_{n-1}, x_{n-1})=x_n$$ for all $n$ in $\mathbb{Z}$
+\end{Definition}
+
+It is important to emphasise that a sequence satisfying the update equation above can only be a solution if $x_n\in{U}$ for all $n\in\mathbb{Z}$. Consider the example below.
+
+\begin{Example}
+  Consider the driven system $g(u_n,x_n)=x^n$ for $X=[0,1]$.
+This system has an uncountable number of solutions, because for every $x\in{X}$, there exists a solution passing through the point $x$ and $lim_{n\to\infty}x_n=0$, $lim_{n\to{-}\infty}x_n=1$.
+\end{Example}
+
+\begin{Definition}
+Let $X_n{\overline{u}}:= \cap_{m<n}{\phi_{\overline{u}}(n,m,X)}$ be the set of all reachable states at time $n$ for input $\overline{u}$.
+  A function $g:U\times{X}\to{X}$ is a topological contraction if for all $n\in\mathbb{Z}$ and all $\overline{u}\subset{U}$, $X_n(\overline{u})$ is a singleton subset of X.  
+\end{Definition}
+
+Note that $x\in{X_n{\overline{u}}}$ if and only g has a solution $\{x_k\}$ s.t.
+$x_n=x$ and the input is $\overline{u}$.
+\textbf{cite paper}.
+This will lead to a result established later, but which is worth taking note of now: $g$ being a topological contraction is equivalent to the existence of a unique entire solution for every $n\in\mathbb{Z}$.
+\textbf{Adding worth or not?} \label{here_UAP_firstestablished}
+
+\begin{Example}\label{ex_halfux}
+  If $X=[0,1]$, $U=[0,1]$, then the only possible solution $\{x_n\}$ to the driven system $g(u,x)=\frac{ux}{2}$ is the zero solution $x_n\equiv0$.
+
+  To see this, we show that for every input $\{u_n\}$, $g(u_k,\dot)$ is a contraction map on $X$ for every $k\in\mathbb{Z}$. \textbf{Show}
+  Consequently, we easily conclude that $\{x_n\}$ where $x_n=0$ for all $n\in\mathbb{Z}$ is the only solution. \textbf{Show}
+\end{Example}
+
+One may ask if a driven system always has a solution, and if so, whether it has certain properties such as uniqueness.
+Should the system be compact, existence follows immediately as can be seen in the following result.
+
+%%%%%%%%%%%%%%%%%%%%%ISSUE HERE - fix before sending
+\begin{Theorem}
+  If $X$ is compact then for each input $\overline{u}$, there exists at least one solution to the driven system $g(\ldotp, x)$
+\end{Theorem}
+\begin{proof}
+  Consider ${u_n}_{n\in\mathbb{Z}}\subset{U}$ and $g:U\times{X}\to{X}$  generating a sequence ${x_n}_{n\in\mathbb{Z}}$ in the compact space.
+(If U is a metric space then $x_n$ has a convergent subseqeunce)
+\end{proof}
+
+\begin{Example}
+  The driven system $g(u,x)=x$ has only the constant solution $x$ and so for $U=[-1,1]$, the system has no solution if $|X|>1$
+\end{Example}
+
+The examples above present systems with trivial sets of solutions. To refine the scenario, we define systems with unique solutions.
+
+\begin{Definition}
+  A driven system $g$ is said to have the Unique Solution Property if for each input $\overline{u}$ there exists exactly one solution.
+Alternatively we may formulate the USP as follows: $g$ has the Unique Solution Property if there exists a well-defined map $\Psi:{U}\to{X}$ where $\Psi({\overline{u}})$ denotes the unique solution
+\end{Definition}
+
+\emph{(More discussion on USP … - not yet filled in ...)}
+
+We next identify a subspace $X_U$ of $X$ that contains all possible solutions. To realize such a subspace of a driven system $g$, we define 
+the \textit{reachable set} of the driven system $g$ to be the union of all the elements of all the solutions, i.e., 
+$$X_U :=\Big \{x \in X:  x = x_k \mbox{ where  $\{x_n\}$  is a solution for some  $\bar{u}$} \Big \}.$$
+
+We denote $\cev{u}^{n}:=(\ldots,u_{n-2} ,u_{n-1})$ and $\overleftarrow{U}$ denote all the left-infinite sequences in $U$. 
+Symbolically we let $\cev{u}^{n}v:=(\ldots,u_{n-2} ,u_{n-1}, v)$ denote the input up to time $n$ with $v \in U$ being the input value at time $n$.
+This introduction of a new input at time $n$ can be described by a mapping $\sigma_v:
+\cev{u}^{n} \mapsto \cev{u}^{n}v$.   \textbf{Is this in the right place?}
+
+The question now becomes whether we may establish a conjugacy as presented below for g driven
+
+\begin{equation}  \label{Scomm_h}
+  %    \[ 
+      \psset{arrows=->, arrowinset=0.25, linewidth=0.6pt, nodesep=3pt, labelsep=2pt, rowsep=0.7cm, colsep = 1.1cm, shortput =tablr}
+   \everypsbox{\scriptstyle}
+   \begin{psmatrix}
+   \overleftarrow{U} & \overleftarrow{U}\\%
+   X_U & X_U.
+   %%%
+  %  \ncline{1,1}{1,2}^{\sigma_v} \ncline{1,1}{2,1} <{h}
+  %  \ncline{1,2}{2,2} > {h}
+  %  \ncline{2,1}{2,2}^{g(v,\cdot)}
+   \end{psmatrix}
+  % \]
+  \end{equation}
+
+  \begin{Definition} \rm \label{Def_UC}
+    Given a driven system $g$, we  call a continuous and surjective map $h : \overleftarrow{U} \to X_U$ a universal semi-conjugacy if  diagram \ref{Scomm_h} commutes for all $v \in U$.
+  \end{Definition}
+
+If the universal semi-conjugacy $h$ exists (i.e.
+the diagram  \ref{Scomm_h} commutes) then the solution $\Psi$ will be at least a coarse-grained representation of the input $u$; we may now consider the question of the existence of such a function $h$ for the driven system defined above.
+
+If $g$ has the USP and $\Psi(u)=\{x_n\}$ then $h$, defined by  $h(\overleftarrow{u}_n)=x_n=g(u_n,x_{n-1})$, will satisfy the semi-conjugacy in the graph above \ref{Scomm_h}.
+
+In general the mapping h is not guaranteed to exist. We thus state the following result.
+
+\begin{Theorem}
+  For a compactly driven system, a causal mapping $H$ exists if and only if $g$ has the USP. 
+\end{Theorem}
+\begin{proof}
+  \emph{See proof from p.15 of \cite{manjunath2021universal}}
+\end{proof}
+
+Note that even when $h$ does exist, we are not guaranteed its injectivity.
+Considering again example 3* \ref{ex_halfux}, we conclude that even if $h$ were to exist, it could not be injective as $X_U=\{0\}$.
+
+An embedding of the space $\overleftarrow{U}$   would allow one to establish topological conjugacy, which in turn provides stronger results than merely obtaining a coarse-grained representation via a causal mapping.
+
+Re-sketching the graph \ref{Scomm_h} above by fixing $v$ in $g$ and replacing $X_U$ by its left-infinite sequence space $\overleftarrow{U}$, we obtain the graph below.
+In this case, the function $H:\overleftarrow{U}\to\overleftarrow{X}_U$, a map that is both continuous and surjective, is called a \emph{causal mapping}.
+
+\begin{equation}  \label{SCausal_H}
+    %    \[ 
+        \psset{arrows=->, arrowinset=0.25, linewidth=0.6pt, nodesep=3pt, labelsep=2pt, rowsep=0.7cm, colsep = 1.1cm, shortput =tablr}
+     \everypsbox{\scriptstyle}
+     \begin{psmatrix}
+     \overleftarrow{U} & \overleftarrow{U}\\%
+     \overleftarrow{X}_U & \overleftarrow{X}_U.
+     %%%
+    %  \ncline{1,1}{1,2}^{\sigma_v} \ncline{1,1}{2,1} <{h}
+    %  \ncline{1,2}{2,2} > {h}
+    %  \ncline{2,1}{2,2}^{\tilde{g}_v}
+     \end{psmatrix}
+    % \]
+    \end{equation}
+
+  Note that $\tilde{g}_v$ maps $(\ldots, u_{-2}, u_{-1})$ to $(\ldots, u_{-2}, u_{-1}, g(v, u_{-1}))$.
+  The driven system $g$ can induce an embedding of $\overleftarrow{U}$ in $\overleftarrow{X}$ as follows: 
+  If $H$ is also injective (and hence bijective), it becomes the embedding of $\overleftarrow{U}$ in $\overleftarrow{X}$ induced by the driven system $g$ and we refer to $H$ as a \emph{causal embedding}.
+We formalise this in the following definition.
+
+\begin{Definition}
+  A driven system $g$ is said to causally embed the dynamical system $(U,T)$ if 
+  \vspace{-8mm}
+\begin{enumerate}[noitemsep, label=\roman*.]
+  \item The diagram below commutes (i.e. $g$ is a universal semi-conjugacy),
+  \item $H_2(\overline{u}):=(h(r\overline{u}, h(\overline{u})))$ embeds the inverse-limit space $(\hat{U}, \hat{T})$ of $(U,T)$ in $X\times{X}$.
+\end{enumerate}
+\end{Definition}
+
+\begin{Theorem}
+ If $g(\ldotp,x)$ is invertible and has the USP then $H$ is a causal embedding. 
+\end{Theorem}
+
+When g has the USP, the diagram below illustrates the operation of the the mappings h and H the mapping $h:\overleftarrow{U}\to{X_U}$ is an observable as discussed (above).
+
+\begin{figure}[ht]
+  \includegraphics[scale=0.3]{actionofh_H.png}
+  \centering
+  \label{fig:actionh_H}
+\end{figure}
+
+\begin{Theorem}
+  The following statements are equivalent:
+  \vspace{-8mm}
+  \begin{enumerate}[noitemsep, label=\roman*.]
+    \item $g$ has the USP 
+    \item $g$ is topological contraction 
+    \item $g$ has the Unique Attraction Property (UAP). 
+  \end{enumerate}
+\end{Theorem}
+\begin{proof}
+  (we’ll show this)  [11] eqn.6 
+\end{proof}
+
+We have not as of yet treated the UAP, and establish the necessary result here (without proof - left to literature or later) so as not to break up the completeness of theorem
+
+It is important to note here that one must be careful to avoid a choice of g which adds additional complexity to the obtained solution.  
+When the causal embedding exists for the driven system $g$, one may map an arbitrary input $u$ onto the solution space $X$ without additional distortion or information-loss.  
+Establishing an embedding removes this problem by guaranteeing that, since the systems are conjugate (semi-conjugate), $g$ does not add any (only “some”) complexity to the system.  
+It is, however, a balancing act as one also does not want to choose a function g that quenches the temporal structure in $u$ by contracting to such a degree that one loses the ability to recover information from the original system completely.
+To obtain this we want the reachable set of a driven system to be large enough to relate to the input.
+
+In the example above \ref{ex_halfux}, we cannot relate the input's temporal variation and the reachable set because $X_U$ consists of a single element.
+
+The reachable set $Y_T$ of a driven system must therefore be such that we can embed the inverse-limit space of $U_T$ can be embedded in some finite self-product of the reachable set of $T$.
+To this end, consider the notion of State-Input (SI) Invertibility.
+
+\begin{Definition}
+  A map $g$ is said to be SI-Invertible if $g$ is invertible for all $x\in{X}$.
+Alternatively, if, given $x_n$ and $x_{n-1}$, $u_n$ can be uniquely determined from $x_n=g(u_n,x_{n-1})$, then $g$ is said to be SI-invertible.
+\end{Definition}
+
+We make special note of a specific driven system.
+The function $$g(u,x)=(1-a)x+atanh(Au+\alpha{a}Bx)$$  is both SI-invertible and possesses the USP.
+This function $g$ is used in our implementation and is more completely discussed in Section 5 \ref{sect5}
+The proof of these two facts is rather involved and the proof, instead of being replicated here, may be found in (cite relevant article)
+
+\textbf{Expand: UAP + Embedding -> perturbation isn’t an issue + we can drop left-infinite sequence.  }
+
+Despite the ease that one may work with left-infinite sequences in the realm of theory, it is impossible to obtain such a sequence in any real-life application.
+One does not in practice, however, need the entire left-infinite history of the input due to the Uniform Attraction Property (UAP).
+The next definition is stated in an informal manner as the more rigourous definition makes use of processes, a concept which would take some amount of paper to establish and explain - something which would detract from the overarching thrust of this paper/project/thesis.
+\textbf{is it VERY obsious to the reader at this point what the general 'thrust' of the paper IS?} \textbf{Cite articles}
+
+\begin{Definition}
+  A driven system $g$ has the Uniform Attraction Property (UAP) if, regardless of starting position, all trajectories converge to a single trajectory as time flows forward. 
+  This trajectory is also the unique solution sequence $x$ to the input sequence $u$ as mentioned above \ref{here_UAP_firstestablished}.
+\end{Definition}
+
+Incredibly, one may then initialise a driven system with a completely arbitrary initial value $y_m\in{X}$ where $m\in\mathbb{Z}$ and the UAP then guarantees that the sequence $\{y_{m+1}, y_{m+2},\ldots\}$ satisfying $y_n=g(u_n, y_{n-1})$  with $k\geq{m}$  will uniformly approach the elements $\{x_n\}$ of the actual solution.
+(see [41, Theorem 1] or [11, Eqn.
+(18)])
+
+One may now appreciate an even more astounding result.
+$g$ having the USP is equivalent to the UAP and so one need not establish any additional results to ensure that the sequence uniformly approach the unique solution $\Psi$.
+This vastly simplifies the work needed to be done in setting up a problem to guarantee that the underlying system will be accurately 'represented' by the conjugate system.
+\textbf{Check this statement again to make sure it's not nonsensical.}
+
+\begin{figure}[h]
+  \includegraphics[scale=0.4]{memloss_conttime.png}
+  \centering
+  \label{fig:memoryLosscont_time}
+%\caption{Approaching the true solution if g has the UAP}
+\end{figure}
+
+This also solves the problem of perturbations or noise introduced by the observable or measurement function. \textbf{(rephrase)}
+
+Next we define the relation $Y_T$ induced by $(U,T)$ on $X_U\times{X_U}$ for a driven system $g$ possessing SI-invertibility.
+To describe the  single-delay lag dynamics formally, we consider a dynamical system $T: U \to U$ and we define a relation on the reachable set $X_U$, i.e., a subset defined on $X_U \times X_U$ by
+$$Y_T:=\{(x_{n-1},x_n): \{x_k\}_{k\in \mathbb{Z}} \mbox{ is a solution for some orbit of } T \mbox{ and } n \in \mathbb{Z}\}.$$
+
+The following theorem establishes the existence of a well-defined map GT describing the single-delay dynamics of the system above.
+
+\begin{Theorem}
+If we let $G_T:Y_T\to{Y_T}$ be a map defined by the relation $(x_{n-1},x_n)\mapsto(x_n,x_{n-1})$, then $G_T$ is well-defined (and this results holds even in the absence of $g$ possessing over the USP)
+\end{Theorem}
+
+We're not getting quite close to where we want to be and our results carry more and more weight. Recall that $g$ is only being given inputs from the orbits of $T$.
+
+\begin{Theorem}
+  Graph 1 is exactly the inverse-limit system $(\hat{U}, \hat{T})$.    
+\end{Theorem}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\section{Implementation} \label{sect5}
+
 \vspace{-1cm}

### 2021-12-22 Gandhi_Project_Report/draft1.tex
Nearly at the end of Section 4

@@
 \begin{Definition}
-  A driven system $g$ is said to have the Unique Solution Property if for each input $\overline{u}$ there exists exactly one solution.
+  A driven system $g$ is said to have the Unique Solution Property(USP) if for each input $\overline{u}$ there exists exactly one solution.
 Alternatively we may formulate the USP as follows: $g$ has the Unique Solution Property if there exists a well-defined map $\Psi:{U}\to{X}$ where $\Psi({\overline{u}})$ denotes the unique solution
@@
   If $H$ is also injective (and hence bijective), it becomes the embedding of $\overleftarrow{U}$ in $\overleftarrow{X}$ induced by the driven system $g$ and we refer to $H$ as a \emph{causal embedding}.
-We formalise this in the following definition.
+
+  Before formalising this, we make mention of the concept of the \emph{inverse-limit system} of a system $(U,T)$.
+In broad sweeps, the inverse-limit system of a dynamical system is a self-map on a subset of an infinite-dimensional space \textbf{(expand what exactly this is?)} where each point in the inverse-limit space corresponds to a backward orbit of the map $T$.
+In literature, the inverse-limit space is denoted by $\hat{U}_T\subset\overleftarrow{U}$.
+\textbf{cite from supp.koopman}.
+  The map $T$, then, induces a self-map $\hat{T}:\widehat{T}: (\ldots,u_{-2},u_{-1}) \mapsto  (\ldots,u_{-2},u_{-1},T(u_{-1}))$ on $\widehat{U}_T$
 
@@
 
-We have not as of yet treated the UAP, and establish the necessary result here (without proof - left to literature or later) so as not to break up the completeness of theorem
+We have not as of yet treated the UAP, and establish the necessary result here (without proof - left to literature. Or later) so as not to break up the completeness of theorem
 
@@
 
-We're not getting quite close to where we want to be and our results carry more and more weight. Recall that $g$ is only being given inputs from the orbits of $T$.
+We're now getting quite close to where we want to be and our results carry more and more weight. Recall that $g$ is only being given inputs from the orbits of $T$.
+
+\begin{Definition}
+  The inverse-limit space $\widehat{U}_T$, a subspace of $\overleftarrow{U}$, is defined by $$\widehat{U}_T:=\{(\ldots, u_{-2}, u_{-1}):T(u_n)=u_{n+1}\}$$ This space is well-defined since $T:U\to{U}$ is surjective by assumption.
+\textbf{Expand why}
+\end{Definition}
 
 \begin{Theorem}
-  Graph 1 is exactly the inverse-limit system $(\hat{U}, \hat{T})$.    
+  Graph \ref{Scomm_h} is exactly the inverse-limit system $(\hat{U}, \hat{T})$.    
 \end{Theorem}
+
+Note that $H_2$ maps an entire left-infinite solution sequence from $\Psi$ to an element in $X\times{X}$.
+
+We now have the following (compare with \ref{SCausal_H} above):
+
+\emph{graph here causes issues currently}
+\begin{center}
+%    \[ 
+      \psset{arrows=->, arrowinset=0.25, linewidth=0.6pt, nodesep=3pt, labelsep=2pt, rowsep=0.7cm, colsep = 1.1cm, shortput =tablr}
+      \everypsbox{\scriptstyle}
+      \begin{psmatrix}
+      \widehat{U}_T  && \widehat{U}_T \\
+      Y_T && Y_T \\
+      %%%
+     %  \ncline{1,1}{1,2}^{\widehat{T}} \ncline{1,1}{2,1} <{h}
+     %  \ncline{1,2}{2,2} > {H_2}
+     %  \ncline{2,1}{2,2}^{G_T}
+      \end{psmatrix}
+     % \]
+    \end{center}
+
+\textbf{Show that $(Y_T, G_T)$ is semi-conjugate to $(\widehat{U}, \widehat{T})$. }
+
+\textbf{Summarising the discussing thus far:}
+It is easy to lose onself in the symbols, so we take a moment to review our progress up until this point.
+
+\vspace{-8mm}
+\begin{enumerate}
+\item We are interested in a dynamical system $(U,T)$ with unknown dynamics for whichever reason.
+\item To determine information about this system $(U,T)$, we shift our focus by rather aiming to determining the dynamics of the inverse-limit system $(\widehat{U}, \widehat{T})$ which, given certain assumptions, we can guarantee to be at least semi-conjugate to $(U,T)$.
+\item If the driven system $g$ (and $\{u_n\}$ in $U$  is an orbit of $T$), we have that $G_T$ exists. 
+\item If, furthermore, $g$ may be shown to have the USP, we showed that $(Y_T, G_T)$ is semi-conjugate to $(\widehat{U}, \widehat{T})$.
+\item If $T$ is also (and at this stage this is very much our hope) a homeomorphism, then $(Y_T, G_T)$ topologically conjugate to $(\widehat{U}, \widehat{T})$ in which $(U,T)$ has been embedded!
+
+One can therefore learn the single-delay lag dynamics of the driven system states via $G_T$ with enough data thanks to the USP.
+This enables us to do at least 2 things: 1) Forecast future values of $x_n$ via iterates of $G_T$ (as $G_T$ can be determined.
+2) Forecast future values of $u_n$.
+
+\end{enumerate}
 

### 2021-12-29 Gandhi_Project_Report/draft1.tex
Section 4: third of the way in with refining

@@
 
-\section{Introduction}
+\section{Introduction}\label{ch1}
 
-Experimenting on biological, physical, and artificial systems in order to generate a more informative dynamical model is a well-established practice in modern science.
+% Experimenting on biological, physical, and artificial systems in order to generate a more informative dynamical model is a well-established practice in modern science.
 Traditional methods for modelling physical systems are based on laws of physics that are based either on empirical relationships or on intuition.
@@
 
-To complicate this - the underlying systems we observe often display sensitive dependence on initial conditions; despite having highly similar initial conditions, differing orbits diverge quickly and to such an extent that it becomes seemingly impossible to retrace their steps back to the original conditions.
+% To complicate this - the underlying systems we observe often display sensitive dependence on initial conditions; despite having highly similar initial conditions, differing orbits diverge quickly and to such an extent that it becomes seemingly impossible to retrace their steps back to the original conditions.
 Even the presence of usually 'negligible' computational noise or measurement error renders long-term point-wise prediction infeasible.
-There are many difficulties that one encounters while modelling such systems:
-\vspace{-8mm}
-\begin{enumerate}[noitemsep, label=\roman*.]
-  \item One may not have access to the complete states of the systems
-  \item The actual system may be described by functions that behave wildly in the sense their graphs have a wild oscillatory behaviour, i.e., they have a large functional complexity.    
-\end{enumerate}
+% There are many difficulties that one encounters while modelling such systems:
+% \vspace{-8mm}
+% \begin{enumerate}[noitemsep, label=\roman*.]
+%   \item One may not have access to the complete states of the systems
+%   \item The actual system may be described by functions that behave wildly in the sense their graphs have a wild oscillatory behaviour, i.e., they have a large functional complexity.    
+% \end{enumerate}
 
-Models derived primarily from data can be classified into three categories: 
-\vspace{-8mm}
-\begin{enumerate}[noitemsep, label=\roman*.]
-  \item  Interpretable models (i.e., they establish relationships between internal physical mechanisms), 
-  \item Partially interpretable models capturing some modes of the dynamics, 
-  \item Non-Interpretable models, defined as such mainly due to the fact that they are defined on a different phase space that is usually high-dimensional. 
-\end{enumerate}
+% Models derived primarily from data can be classified into three categories: 
+% \vspace{-8mm}
+% \begin{enumerate}[noitemsep, label=\roman*.]
+%   \item  Interpretable models (i.e., they establish relationships between internal physical mechanisms), 
+%   \item Partially interpretable models capturing some modes of the dynamics, 
+%   \item Non-Interpretable models, defined as such mainly due to the fact that they are defined on a different phase space that is usually high-dimensional. 
+% \end{enumerate}
 
-Examples in the literature attempting to forecast data from such systems have tried a number of different approaches with differing degrees of success.
+% Examples in the literature attempting to forecast data from such systems have tried a number of different approaches with differing degrees of success.
 One could learn a system conjugate to the underlying system by applying the Takens delay embedding; this learnt system could then be used for forecasting the original system.
@@
 Alternatively, one could approximate the vector field by a library of functions (Citations to Sindy) to obtain interpretable models.
-Takens delay methods suffer from lack of robustness to noise and may not yield a learnable map that has lower functional complexity (See Section 3).
+Takens delay methods suffer from lack of robustness to noise and may not yield a learnable map that has lower functional complexity (See Chapter 3).
 For the methods that build interpretable models (Citations to Sindy) the complete system states are needed which is rarely available.
@@
 
-Recent partially interpretable models available in the literature have been based on the Koopman operator (e.g.,\cite{koopman1932dynamical,budivsic2012applied}) to  employ observables mapping the data onto a higher-dimensional space.
+% Recent partially interpretable models available in the literature have been based on the Koopman operator (e.g.,\cite{koopman1932dynamical,budivsic2012applied}) to  employ observables mapping the data onto a higher-dimensional space.
 This then makes the dynamics in the higher-dimensional space more amenable for approximation by a linear transformation.
@@
 
-The non-interpretable models include the delay embedding and the machine learning algorithms.
+% The non-interpretable models include the delay embedding and the machine learning algorithms.
 Under some generic conditions, the Takens delay embedding theorem \cite{takens1981detecting} and its various generalisations (e.g., \cite{sauer1991embedology, stark1999delay, gutman2018embedding, gutman2018embedding}) establishes the learnability of a system constructed by concatenating sufficiently large previous time-series observations of a dynamical system into a vector (called delay coordinates).
@@
 
-Practically, the application of Takens embedding involves learning a map through some technique and consequently one wishes that these would have low functional complexity\cite{manjunath2021universal}, i.e., functions with fewer oscillatory graphs.
+% Practically, the application of Takens embedding involves learning a map through some technique and consequently one wishes that these would have low functional complexity\cite{manjunath2021universal}, i.e., functions with fewer oscillatory graphs.
 
-This thesis deals with the implementation and analysis of non-interpretable models that can guarantee \textbf{exact/accurate} reconstruction.
+% This thesis deals with the implementation and analysis of non-interpretable models that can guarantee \textbf{exact/accurate} reconstruction.
 The project work concerns the study and implementation of a method (Citation to paper with Adriaan) that incorporates learning a function by mapping the data on to a higher dimensional space.
@@
 
-By solidifying the mathematical underpinnings of our theory, we hope to guarantee the ability to construct models with predictive power ranging from molecular biology to neuroscience.
+% By solidifying the mathematical underpinnings of our theory, we hope to guarantee the ability to construct models with predictive power ranging from molecular biology to neuroscience.
 in the near future.
@@
 
-The report is organised as follows: 
-\newline In Section 2, we recall the definition of a discrete-time dynamical system, how a discrete-system arises from a flow of an ODE and then proceed to define the inverse limit space and topological conjugacy of autonomous systems.
-\newline In Section 3, we introduce the problem of forecasting dynamical systems, state the Takens delay embedding theorem and discuss various issues faced while forecasting. 
-\newline In Section 4, a driven dynamical system is defined and discuss the properties of a specific class of driven dynamical systems that we make use of in this thesis/report/paper. 
-\newline Finally in Section 5 we show the implementation of these forecasting methods, and conclusions are provided in Section 6.
+% The report is organised as follows: 
+% \newline In Chapter 2, we recall the definition of a discrete-time dynamical system, how a discrete-system arises from a flow of an ODE and then proceed to define the inverse limit space and topological conjugacy of autonomous systems.
+% \newline In Chapter 3, we introduce the problem of forecasting dynamical systems, state the Takens delay embedding theorem and discuss various issues faced while forecasting. 
+% \newline In Chapter 4, a driven dynamical system is defined and discuss the properties of a specific class of driven dynamical systems that we make use of in this thesis/report/paper. 
+% \newline Finally in Chapter 5 we show the implementation of these forecasting methods, and conclusions are provided in Chapter 6.
 
-\section{Discrete-time Dynamical Systems}
+\section{Discrete-time Dynamical Systems}\label{ch2}
 
-In this chapter, we provide a brief description of what a discrete-time dynamical system is and what it means for it to exhibit chaos.
+% In this chapter, we provide a brief description of what a discrete-time dynamical system is and what it means for it to exhibit chaos.
 We refer to \cite{devaney2018introduction, de2013elements} for more details.
 
-At its most elementary level, a dynamical system is just something that evolves deterministically through time.
+% At its most elementary level, a dynamical system is just something that evolves deterministically through time.
 In the context of this thesis, deterministic refers to the fact that a system evolves according to specified rules rather than based on random events.
-Dynamical systems arise in a variety of situations.
+% Dynamical systems arise in a variety of situations.
 A continuous dynamical system, and here specifically the motion of a pendulum, is one in which quantities such as the angular position and angular momentum are known at all times.
-The equations of the dynamical system take the form of one or more ordinary differential equation that determine the relevant quantities at any future time if we know the initial location and momentum.
-In ecology, discrete dynamical systems are widely used to model population growth.
+% The equations of the dynamical system take the form of one or more ordinary differential equation that determine the relevant quantities at any future time if we know the initial location and momentum.
+% In ecology, discrete dynamical systems are widely used to model population growth.
 The model in this case is a function calculating the following generation's population given the population of the previous generation.
@@
 
-Formally, a function $T: U \to U$, where $U$ is some set is  a \emph{discrete-time dynamical system} and its iterates $\{u,Tu,T^2u,\ldots\}$, where $T^n$ denotes the $n$-fold composition of $T$ with itself, describe the evolution of an initial condition $u\in U$ (Note that we drop the brackets and denote $T(u)$ by $Tu$ for simplicity).
+% Formally, a function $T: U \to U$, where $U$ is some set is  a \emph{discrete-time dynamical system} and its iterates $\{u,Tu,T^2u,\ldots\}$, where $T^n$ denotes the $n$-fold composition of $T$ with itself, describe the evolution of an initial condition $u\in U$ (Note that we drop the brackets and denote $T(u)$ by $Tu$ for simplicity).
 
-Continuous dynamical systems modelled using ordinary differential equations can give rise to discrete-time dynamical systems.
+% Continuous dynamical systems modelled using ordinary differential equations can give rise to discrete-time dynamical systems.
 To see this, consider differential equation $\dot{x} = f(x)$, $f: \mathbb{R}^n \to \mathbb{R}^n$, $n\in\mathbb{N}$ given to have a unique solution passing through
-each point $x\in\mathbb{R}^{n}$, the flow of the equation \label{defn_flow} is defined to be a mapping $\varphi: \mathbb{R}^n \times \mathbb{R} \to \mathbb{R}$ that satisfies the following properties:
-\vspace{-8mm}
-\begin{enumerate}[noitemsep, label=\roman*.]
-  \item $x(t)$ is a solution of the differential equation,
-  \item $x(0)=x_0$, and
-  \item $\varphi(x_0,t) = x(t)$.
-\end{enumerate}
-By fixing $t=K$, we can define the \emph{time-$K$ map} as  $T(x):= \varphi(x_0,K)$, and it is easily verified that $T\circ T(x_0) = \varphi(x_0,2K)$.
+% each point $x\in\mathbb{R}^{n}$, the flow of the equation \label{defn_flow} is defined to be a mapping $\varphi: \mathbb{R}^n \times \mathbb{R} \to \mathbb{R}$ that satisfies the following properties:
+% \vspace{-8mm}
+% \begin{enumerate}[noitemsep, label=\roman*.]
+%   \item $x(t)$ is a solution of the differential equation,
+%   \item $x(0)=x_0$, and
+%   \item $\varphi(x_0,t) = x(t)$.
+% \end{enumerate}
+% By fixing $t=K$, we can define the \emph{time-$K$ map} as  $T(x):= \varphi(x_0,K)$, and it is easily verified that $T\circ T(x_0) = \varphi(x_0,2K)$.
 In general the $m^{th}$ iterate of $x_0$ under $T$ would be the value of the solution of the ODE evaluated at time $mK$ with the initial condition $x_0$.
-Thus ordinary differential equations give rise to a discrete-time dynamical systems by sampling the value of $x$ at time intervals $K$ units apart.
+% Thus ordinary differential equations give rise to a discrete-time dynamical systems by sampling the value of $x$ at time intervals $K$ units apart.
 
-Of course, discrete-time dynamical systems need not always arise through a differential equation, and once again we may consider the field of ecology where discrete dynamical systems are often directly derived or assumed.
-\emph{There is a school of thought that advocates discrete-time dynamical systems to be more natural for modelling real-world observations than differential equations (see \cite{saber2010introduction})}.
+% Of course, discrete-time dynamical systems need not always arise through a differential equation, and once again we may consider the field of ecology where discrete dynamical systems are often directly derived or assumed.
+% \emph{There is a school of thought that advocates discrete-time dynamical systems to be more natural for modelling real-world observations than differential equations (see \cite{saber2010introduction})}.
 \textbf{too philosphical, or actually good?}
 
-A numerical discretization of a differential equation can also give rise to a discrete-time dynamical system.
+% A numerical discretization of a differential equation can also give rise to a discrete-time dynamical system.
 For instance, Euler's method approximates $\dot{x}(t)$ by $(x(t+h)-x(t))/h$; if $h$ is fixed throughout, the solution of a differential equation $\dot{x}=f(x)$
-can be approximated at the time instant $t+(m+1)h$ by iterating the equation 
-$$x(t+(m+1)h) = x(t+mh) + h f(x(t+mh)).$$ 
-Adopting more succinct notation by replacing $x(t+mh)$ with $u_m$, we rewrite the above equation as
-$$u_{m+1} = u_m + hf(u_m)$$
-or, in even simpler terms, as the discrete-time dynamical system with map $T(u) = u + hf(u)$, where $T$, $u$ and $f$ are understood to be as above.
+% can be approximated at the time instant $t+(m+1)h$ by iterating the equation 
+% $$x(t+(m+1)h) = x(t+mh) + h f(x(t+mh)).$$ 
+% Adopting more succinct notation by replacing $x(t+mh)$ with $u_m$, we rewrite the above equation as
+% $$u_{m+1} = u_m + hf(u_m)$$
+% or, in even simpler terms, as the discrete-time dynamical system with map $T(u) = u + hf(u)$, where $T$, $u$ and $f$ are understood to be as above.
 
-\subsection{Invariant Sets}
+% \subsection{Invariant Sets}
 
-A core concept in the study of dynamical systems is that of invariance. Given a discrete-time dynamical system $T: U \to U$, a subset $A \subset U$ is said to be an \emph{invariant set} if $T(A) =A$. 
-We also define the \emph{orbit of T} to be a sequence $\bar{u} = \{u_n\}_{n\in \mathbb{Z}}$ obeying the update equation, $u_{n+1}=Tu_n$, $n \in \mathbb{Z}$.
+% A core concept in the study of dynamical systems is that of invariance.
+% \begin{Definition}
+%   Given a discrete-time dynamical system $T: U \to U$, a subset $A \subseteq U$ is said to be an invariant set if $T(A) =A$. 
+% \end{Definition}
+% We also define the orbit of T to be a sequence $\bar{u} = \{u_n\}_{n\in \mathbb{Z}}$ obeying the update equation, $u_{n+1}=Tu_n$, $n \in \mathbb{Z}$.
 
-Two examples of invariant sets include a fixed point where $Tu=u$ for  $u\in U$, and a periodic orbit, i.e., a set of iterates $\{u,Tu, T^2u,\ldots,T^pu\}$, where $T^{p+1}u=u$ for some $p$.
+% Two examples of invariant sets include a fixed point where $Tu=u$ for  $u\in U$, and a periodic orbit, i.e., a set of iterates $\{u,Tu, T^2u,\ldots,T^pu\}$, where $T^{p+1}u=u$ for some $p$.
 The entire space $U$ could be also be invariant.
-Consider for example the space $U=[0,1]$, where $Tu=4u(1-u)$ and then $U$ is invariant. \emph{(as every $u\in{U}$ can be written as $u = 4x(1-x)$ for some $x\in{U}$)}
+% Consider for example the space $U=[0,1]$, where $Tu=4u(1-u)$ and then $U$ is invariant. \emph{(as every $u\in{U}$ can be written as $u = 4x(1-x)$ for some $x\in{U}$)}
 
-We may learn a great deal about the iterates of a dynamical system by considering the types of invariant sets of a discrete-time dynamical system.
+% We may learn a great deal about the iterates of a dynamical system by considering the types of invariant sets of a discrete-time dynamical system.
 For example, if $U=[0,1]$ has map $Tu= u/2$, then the only invariant set is $\{0\}$, and all orbits approach this invariant set as time flows in the forward direction.
-Indeed, if a some non-zero invariant set (call it $B$) exists, then there is  some $r\in(0,1]\cap{B}$.
+% Indeed, if a some non-zero invariant set (call it $B$) exists, then there is  some $r\in(0,1]\cap{B}$.
 But $r\notin{T(B)}$ since any orbit with initial value $r$ will be a decreasing sequence.
@@
 
-One may ask if every orbit approaches an invariant set?
+% One may ask if every orbit approaches an invariant set?
 In general the answer is no, since for the dynamical system $Tu=2u$, any orbit that does not intersect the invariant set $\{0\}$, will not approach any invariant set.
 
-However,  when the space $U$ is compact, all orbits approach an invariant set.
+% However,  when the space $U$ is compact, all orbits approach an invariant set.
 \textbf{For let $\bar{u} = \{u_n\}_{n\in \mathbb{Z}}$ be an orbit in the compact space U, then $\bar{u}$ has a convergent subsequence }
 
-In fact the $\omega$-limit set $\omega(u;T)$ of a point $u$ defined to the collection of limit points of the sequence $\{x,Tu,T^2u,\ldots\}$ is nonempty, and $\omega(u;T)$ is invariant.
+% In fact the $\omega$-limit set $\omega(u;T)$ of a point $u$ defined to the collection of limit points of the sequence $\{x,Tu,T^2u,\ldots\}$ is nonempty, and $\omega(u;T)$ is invariant.
 \textbf{(not relevant other than being an additional example perhaps?)}
 
-We discuss various properties of invariant sets: Invariant sets can be attracting or repelling depending on how orbits in its vicinity behave (recall the examples $Tu =u/2$ and $Tu=2u$).
+% We discuss various properties of invariant sets: Invariant sets can be attracting or repelling depending on how orbits in its vicinity behave (recall the examples $Tu =u/2$ and $Tu=2u$).
 We are interested in attractive invariant sets, and in particular so-called attractors which we define below.
-\begin{Definition} \rm Let $T: U \to U$, where $U$  is a metric space with metric $d$. A compact subset $A \subset U$ is said to be an attractor if it satisfies the three conditions: 
-  \vspace{-8mm}
-  \begin{enumerate}
-	\item $A$ is invariant. 
-	\item $A$ is asymptotically stable, i.e., for every $\epsilon > 0$ and for all $u$ so that $d(u,A) < \epsilon$, we have $d(T^nu,A) \to 0$ as $n\to \infty$. 
-	\item $A$ has Lyapunov stability, i.e., for every $\epsilon > 0$  there exists a $\delta(\epsilon) > 0$ so that $d(u,A) < \delta$ implies $d(T^nu,A) < \epsilon$ for all $n\ge 0$.  
-\end{enumerate}
-\end{Definition}
+% \begin{Definition} \rm Let $T: U \to U$, where $U$  is a metric space with metric $d$. A compact subset $A \subseteq U$ is said to be an attractor if it satisfies the three conditions: 
+%   \vspace{-8mm}
+%   \begin{enumerate}
+% 	\item $A$ is invariant. 
+% 	\item $A$ is asymptotically stable, i.e., for every $\epsilon > 0$ and for all $u$ so that $d(u,A) < \epsilon$, we have $d(T^nu,A) \to 0$ as $n\to \infty$. 
+% 	\item $A$ has Lyapunov stability, i.e., for every $\epsilon > 0$  there exists a $\delta(\epsilon) > 0$ so that $d(u,A) < \delta$ implies $d(T^nu,A) < \epsilon$ for all $n\ge 0$.  
+% \end{enumerate}
+% \end{Definition}
 
-In our previous examples, the system  $Tu=u/2$ had the singleton set $\{0\}$ as an attractor.
+% In our previous examples, the system  $Tu=u/2$ had the singleton set $\{0\}$ as an attractor.
 For the system,  $Tu=4u(1-u)$ on $[0,1]$, one may easily verify that the only attractor is the entire space $[0,1]$.
@@
 From above, we know that B is itself invariant, is asymptotically stable and has Lyapunov stability.
-Now suppose there is some attractor $C\subsetneq{U}=B$, so there is a $b\in{U}$ which is not in $C$.
+Now suppose there is some attractor $C\subseteqneq{U}=B$, so there is a $b\in{U}$ which is not in $C$.
 ........
@@
 
-A dynamical system can have several attractors, and may also be contained in another attractor.
+% A dynamical system can have several attractors, and may also be contained in another attractor.
 
-% If the dynamics on the attractor are somewhat complicated in the sense that the attractor cannot be decomposed further, and if the attractor is infinite then there is a possibility of complicated dynamics, and a particular well-studied phenomenon of such complexity gives rise to what is called a chaotic behaviour.
+% % If the dynamics on the attractor are somewhat complicated in the sense that the attractor cannot be decomposed further, and if the attractor is infinite then there is a possibility of complicated dynamics, and a particular well-studied phenomenon of such complexity gives rise to what is called a chaotic behaviour.
 
-If the dynamics on an attractor cannot be decomposed further and the attractor is itself also infinite, then there is a possibility of complicated dynamics, and a particular well-studied phenomenon of such complexity gives rise to what is called chaotic behaviour.
+% If the dynamics on an attractor cannot be decomposed further and the attractor is itself also infinite, then there is a possibility of complicated dynamics, and a particular well-studied phenomenon of such complexity gives rise to what is called chaotic behaviour.
 
-\subsection{Chaos}
+% \subsection{Chaos}
 
-\textbf{$\downarrow$ - should one add paragraphs like these?}
-\newline In the 1960s, a number of mathematicians and mathematically interested scientists independently discovered chaos in the mathematical sense.
+% \textbf{$\downarrow$ - should one add paragraphs like these?}
+% \newline In the 1960s, a number of mathematicians and mathematically interested scientists independently discovered chaos in the mathematical sense.
 The meteorologist Edward Lorenz may have been the first to explain this phenomenon in his 1963 paper \cite{lorenz1963deterministic}.
@@
 
-There are several intricately different notions of chaos, with each of them indicating some complexity.
+% There are several intricately different notions of chaos, with each of them indicating some complexity.
 \textbf{ $\leftarrow  $ what are we saying here?} In practice, it is not possible to verify which notion is satisfied when data is observed from a dynamical system.
@@
 
-\begin{Definition} \rm 
-A dynamical system $T: U \to U$ is said to have sensitive dependence on initial conditions (SDIC) if there exists a $\delta > 0$ such that for every $u \in U$ and in every neighborhood of $u \in U$ there exists a $v\in{U}$ and an integer $N:=N{(u,v)}\in\mathbb{Z}$ such that $d(T^Nu,T^Nv)>\delta$.
-\end{Definition}
+% \begin{Definition} \rm 
+% A dynamical system $T: U \to U$ is said to have sensitive dependence on initial conditions (SDIC) if there exists a $\delta > 0$ such that for every $u \in U$ and in every neighborhood of $u \in U$ there exists a $v\in{U}$ and an integer $N:=N{(u,v)}\in\mathbb{Z}$ such that $d(T^Nu,T^Nv)>\delta$.
+% \end{Definition}
 
-It is important to acquire a sense of this definition as it can easily be misunderstood.
+% It is important to acquire a sense of this definition as it can easily be misunderstood.
 It is common misconception to interpret SDIC as two close points ($u$ and $v$) that eventually become separated by a distance $\delta$ under iteration by $T$.
@@
 To fully comprehend all the subtleties of the concept, one needs to discuss it more thoroughly by considering each sentence with care and attention, whereafter one may examine how they fit together to convey the concept of sensitive dependence.
-\textbf{ $\leftarrow$ is this talking too simply perhaps?}.
+% \textbf{ $\leftarrow$ is this talking too simply perhaps?}.
 
-To this end we make 3 remarks:
-\vspace{-5mm}
-\begin{enumerate}
-  \item First, the $\delta>0$ in the definition of SDIC is independent of $u$. 
-  \item Second, in every neighborhood of $u$, we may not necessarily find all points $v$ in the neighborhood distinct from $u$ that would separate from the forward iterates of $u$. 
-  \item  Finally, $N$ depends upon $u$ and $v$ chosen, and their iterates may not separate for ever (i.e., for all $n>N$) and we allow their iterates to get arbitrarily close in the future. 
-\end{enumerate}
+% To this end we make 3 remarks:
+% \vspace{-5mm}
+% \begin{enumerate}
+%   \item First, the $\delta>0$ in the definition of SDIC is independent of $u$. 
+%   \item Second, in every neighborhood of $u$, we may not necessarily find all points $v$ in the neighborhood distinct from $u$ that would separate from the forward iterates of $u$. 
+%   \item  Finally, $N$ depends upon $u$ and $v$ chosen, and their iterates may not separate for ever (i.e., for all $n>N$) and we allow their iterates to get arbitrarily close in the future. 
+% \end{enumerate}
 
-\ednote{M: It is probably best to see a few numerical simulations to understand SDIC. (Please use the logistic map and perhaps the trajectories of the double pendulum data). }
+% \ednote{M: It is probably best to see a few numerical simulations to understand SDIC. (Please use the logistic map and perhaps the trajectories of the double pendulum data). }
 
-The second part in Devaney's definition of chaos concerns topological transitivity.
+% The second part in Devaney's definition of chaos concerns topological transitivity.
 
-\begin{Definition}
-	A dynamical system $F: U \to U$ is topologically transitive if for any pair of nonempty open sets $E_1$ and $E_2$ there exists a $n\in\mathbb{N}$ such that $T^n(E_1) \cap E_2 \not= \emptyset$. 
-\end{Definition}
+% \begin{Definition}
+% 	A dynamical system $F: U \to U$ is topologically transitive if for any pair of nonempty open sets $E_1$ and $E_2$ there exists a $n\in\mathbb{N}$ such that $T^n(E_1) \cap E_2 \not= \emptyset$. 
+% \end{Definition}
 
-Topologically transitivity implies that the iterates of an open set of initial conditions gets mixed up with other open sets.
+% Topologically transitivity implies that the iterates of an open set of initial conditions gets mixed up with other open sets.
 On a compact metric space, it can be shown that topological transitivity also implies the existence of a point whose forward iterates are dense; or in other words, the orbit going through this point will be dense in the compact metric space.
 
-In fact, it is topologically more likely that the choice of an arbitrary point will be one  whose iterates are almost dense.
-% I inserted a definition here: it makes more sense to me. 
-\begin{Definition}
-A compact metric space $U$ is said to be almost dense if the complement of the set of points which are dense in $U$ may be written as a countable union of sets of with empty interior.
+% In fact, it is topologically more likely that the choice of an arbitrary point will be one  whose iterates are almost dense.
+% % I inserted a definition here: it makes more sense to me. 
+% \begin{Definition}
+% A compact metric space $U$ is said to be almost dense if the complement of the set of points which are dense in $U$ may be written as a countable union of sets of with empty interior.
 In other words, the set on points not dense on $U$ is topologically insignificant.
-\end{Definition}
+% \end{Definition}
 
-We may now define the notion of Chaos as formulated by Devaney\cite{devaney2018introduction}.
-\begin{Definition}
-	A dynamical system $T: U \to U$ is said to exhibit chaos in the sense of Devaney if it satisfies the three properties:
-	\vspace{-5mm}
-  \begin{enumerate}
-		\item $T$ has SDIC.
-		\item $T$ is topologically transitive.
-		\item The set of periodic points of $T$ are dense in $U$. 
-	\end{enumerate}
-\end{Definition}
+% We may now define the notion of Chaos as formulated by Devaney\cite{devaney2018introduction}.
+% \begin{Definition}
+% 	A dynamical system $T: U \to U$ is said to exhibit chaos in the sense of Devaney if it satisfies the three properties:
+% 	\vspace{-5mm}
+%   \begin{enumerate}
+% 		\item $T$ has SDIC.
+% 		\item $T$ is topologically transitive.
+% 		\item The set of periodic points of $T$ are dense in $U$. 
+% 	\end{enumerate}
+% \end{Definition}
 
-\begin{figure}[ht]
-  \centering
-  \begin{minipage}{.5\textwidth}
-    \centering
-    \includegraphics[scale=0.18]{chaos1.png}
-    \captionof{figure}{Starting to diverge}
-    \label{fig:chaos1}
-  \end{minipage}%
-  \begin{minipage}{.5\textwidth}
-    \centering
-    \includegraphics[scale=0.18]{chaos2.png}
-    \captionof{figure}{Completely chaotic}
-    \label{fig:chaos2}
-  \end{minipage}
-  \end{figure}
+% \begin{figure}[ht]
+%   \centering
+%   \begin{minipage}{.5\textwidth}
+%     \centering
+%     \includegraphics[scale=0.18]{chaos1.png}
+%     \captionof{figure}{Starting to diverge}
+%     \label{fig:chaos1}
+%   \end{minipage}%
+%   \begin{minipage}{.5\textwidth}
+%     \centering
+%     \includegraphics[scale=0.18]{chaos2.png}
+%     \captionof{figure}{Completely chaotic}
+%     \label{fig:chaos2}
+%   \end{minipage}
+%   \end{figure}
 
-\begin{Example}
-  The standard tent map $T(u)=1-|2u-1|$ defined on $[0,1]$ is a common example of a dynamical system satisfying these three properties. 
-  \begin{figure}[ht]
-    \includegraphics[scale=0.6]{T2.png}
-    \centering
-    \label{fig:T2tentmap}
-  \end{figure}
+% \begin{Example}
+%   The standard tent map $T(u)=1-|2u-1|$ defined on $[0,1]$ is a common example of a dynamical system satisfying these three properties. 
+%   \begin{figure}[ht]
+%     \includegraphics[scale=0.6]{T2.png}
+%     \centering
+%     \label{fig:T2tentmap}
+%   \end{figure}
 
-  Let us now consider why this map exhibit's Devaneys's sense of chaos.
+%   Let us now consider why this map exhibit's Devaneys's sense of chaos.
 The graph of the map $T$ is piecewise linear with two straight lines, one connecting the points $(0,0)$ and $(1/2,1)$ and the other connecting $(1/2,1)$ with $(1,0)$.
@@
 The graph of the map $T^2$ comprises two symmetric tents with their base centered at $1/4$ and $3/4$.
-  % We now reason as why this map exhibits chaos in the sense of Devaney.
+%   % We now reason as why this map exhibits chaos in the sense of Devaney.
 
-  Given any point $u$ and an open neighborhood $W \subset[0,1]$ of $u$, we can find $k\in\mathbb{Z}$ large enough so that $T^k(W) = [0,1]$ because we can accommodate a tent whose base is contained in $W$.
+%   Given any point $u$ and an open neighborhood $W \subseteq[0,1]$ of $u$, we can find $k\in\mathbb{Z}$ large enough so that $T^k(W) = [0,1]$ because we can accommodate a tent whose base is contained in $W$.
 \emph{This last phrase doesn't make sense to me.} This implies the existence of two points $w_1$ and $w_2$ in $W$ so that $d(T^kw_1,T^kw_2)= 1$, and by the triangle inequality  at least one of the inequalities $d(T^ku,T^kw_1)> \delta$ or $d(T^ku,T^kw_2)> \delta$ hold when $\delta\in (0,1/2)$.
@@
 
-  Next, let $W_1$ and $W_2$ be two nonempty open sets.
+%   Next, let $W_1$ and $W_2$ be two nonempty open sets.
 Given any open set $W_1$, we can find a $k\in\mathbb{N}$ large enough so that $T^k(W_1) = [0,1]$ because we can accommodate a tent with base contained in $W$.
@@
 
-  Finally, for every open interval $(a,b)$, the graph of $T^k$ intersects the graph of the identity map on $[0,1]$.
+%   Finally, for every open interval $(a,b)$, the graph of $T^k$ intersects the graph of the identity map on $[0,1]$.
 This follows from the fact established above that there exists a $k\in\mathbb{N}$ such that $T^k(a,b) = [0,1]$.If the intersection point has coordinates of the form $(p,p)$, the point p is then a fixed point of $T^k$ and therefore also a periodic point of T.
 It follows that the set of periodic points are dense in T.
-  \textbf{Expand further on reason for density.}
-\end{Example}
+%   \textbf{Expand further on reason for density.}
+% \end{Example}
 
-\subsection{Conjugacy}
+% \subsection{Conjugacy}
 
-We now turn our attention to the subject of conjugacy.
+% We now turn our attention to the subject of conjugacy.
 
-To show topological similarity or sameness between two metric or topological spaces, one needs to establish a homeomorphism between the two spaces. 
-However, in the study of dynamical systems defined on two spaces, establishing a homeomorphism does not indicate that the systems are (physically) related in any way.
+% To show topological similarity or sameness between two metric or topological spaces, one needs to establish a homeomorphism between the two spaces. 
+% However, in the study of dynamical systems defined on two spaces, establishing a homeomorphism does not indicate that the systems are (physically) related in any way.
 \textbf{"related" - in what sense am I saying it here?
 Physically?}  For this one rather needs to establish a kind of dynamical equivalence which we illustrate in a simple commutativity diagram below.
-% Giving errors
-\begin{equation}  \label{eqn_conjugacy}
-%    \[ 
-    \psset{arrows=->, arrowinset=0.25, linewidth=0.6pt, nodesep=3pt, labelsep=2pt, rowsep=0.7cm, colsep = 1.1cm, shortput =tablr}
- \everypsbox{\scriptstyle}
- \begin{psmatrix}
-U & U\\%
-V & V.
- %%%
-%  \ncline{1,1}{1,2}^{T} 
-%  \ncline{1,1}{2,1} <{\phi}
-%   \ncline{2,1}{2,2}^{S}
-%  \ncline{1,2}{2,2} > {\phi}
- \end{psmatrix}
-% \]
-\end{equation}
+% % Giving errors
+% \begin{equation}  \label{eqn_conjugacy}
+% %    \[ 
+%     \psset{arrows=->, arrowinset=0.25, linewidth=0.6pt, nodesep=3pt, labelsep=2pt, rowsep=0.7cm, colsep = 1.1cm, shortput =tablr}
+%  \everypsbox{\scriptstyle}
+%  \begin{psmatrix}
+% U & U\\%
+% V & V.
+%  %%%
+% %  \ncline{1,1}{1,2}^{T} 
+% %  \ncline{1,1}{2,1} <{\phi}
+% %   \ncline{2,1}{2,2}^{S}
+% %  \ncline{1,2}{2,2} > {\phi}
+%  \end{psmatrix}
+% % \]
+% \end{equation}
 
-%Composite functions
-To understand the diagram we note that if we travel ``right and then down," the diagram instructs us to use $T$ (top arrow right) first, followed by $\phi$.
+% %Composite functions
+% To understand the diagram we note that if we travel ``right and then down," the diagram instructs us to use $T$ (top arrow right) first, followed by $\phi$.
 (right arrow downwards).
@@
 
-\begin{Definition}
-  Consider the dynamical systems $T:U\to{U}$, $S:V\to{V}$ and suppose the relationship $\phi \circ T=S\circ \phi$ holds.
+% \begin{Definition}
+%   Consider the dynamical systems $T:U\to{U}$, $S:V\to{V}$ and suppose the relationship $\phi \circ T=S\circ \phi$ holds.
 If $\phi$ is a homeomorphism, then S is said to be conjugate to T and $\phi$ is called a conugacy.
 If we relax the criterion on $\phi$ and merely require $\phi$ to be continuous and surjective, $\phi$ is then a semi-conjugacy between $T$ and $S$; S is said to be semi-conjugate to T.
-\end{Definition}
+% \end{Definition}
 
-When $S$ is conjugate to $T$, the dynamics of the two systems are in some way ''dynamically equivalent'' and specifically they are in one-to-one correspondence with one another.
+% When $S$ is conjugate to $T$, the dynamics of the two systems are in some way ''dynamically equivalent'' and specifically they are in one-to-one correspondence with one another.
 However when $S$ is semi-conjugate to $T$ with $\phi$ a many-to-one mapping, the dynamics on $V$ provides a coarse-grained \textbf{cite?} description of the dynamics on $U$ .
@@
 
-It is a very hard or nearly an impossible task to establish the existence of a conjugacy or a semi-conjugacy $\phi$ between two systems.
+% It is a very hard or nearly an impossible task to establish the existence of a conjugacy or a semi-conjugacy $\phi$ between two systems.
 \textbf{cite?} However, one can verify that a function $\phi$ satisfies the commutativity diagram \ref{eqn_conjugacy}.
@@
 
-In establishing that two systems are conjugate (semi-conjugate) to one another, we have also shown that one may choose to work with one systems as opposed to the second and still be guaranteed to obtain information on the latter.
+% In establishing that two systems are conjugate (semi-conjugate) to one another, we have also shown that one may choose to work with one systems as opposed to the second and still be guaranteed to obtain information on the latter.
 This is incredibly useful while forecasting the future evolution of dynamical systems.
 
-\section{The Learning Problem}
+\section{The Learning Problem}\label{ch3}
 
-\emph{Introduction...}
+% \emph{Introduction...}
 
-Consider a relatively simple learning problem:
+% Consider a relatively simple learning problem:
 
-Given $(u_0, u_1, \ldots, u_m)$ for $m\in\mathbb{Z}$, a finite segment of an orbit of the map $T$ where $T$ is defined by the update equation $u_{n+1} = Tu_n$, forecast the values $u_{m+1}, u_{m+1}$ where the map $T$ is unknown, given that $u_m$.
+% Given $(u_0, u_1, \ldots, u_m)$ for $m\in\mathbb{Z}$, a finite segment of an orbit of the map $T$ where $T$ is defined by the update equation $u_{n+1} = Tu_n$, forecast the values $u_{m+1}, u_{m+1}$ where the map $T$ is unknown, given that $u_m$.
 
-A practical example of this would be - given the sequential coordinates of an object moving in space, predict the future positions of that object.
+% A practical example of this would be - given the sequential coordinates of an object moving in space, predict the future positions of that object.
 However we are very rarely, if at all, presented with a problem where the entire state information is available to us in the form $u_n$ at some time-step $n\in\mathbb{N}$.
@@
 
-Suppose we only have the observations $\theta(u_0), \theta(u_1), \ldots, \theta(u_m)$ of the true system states $u$ in an unknown dynamical system $(U,T)$ and we wish to predict the values $\theta(u_{m+1}), \theta(u_{m+2})$.
+% Suppose we only have the observations $\theta(u_0), \theta(u_1), \ldots, \theta(u_m)$ of the true system states $u$ in an unknown dynamical system $(U,T)$ and we wish to predict the values $\theta(u_{m+1}), \theta(u_{m+2})$.
 
-Here we consider a continuous-time dynamical system $(U,T)$ with true state $u$ in the input space $U$ and then a measured signal $\theta$ (where $\theta(u_n)=\theta(u(t_n))$ for $n\in\mathbb{N}$.
+% Here we consider a continuous-time dynamical system $(U,T)$ with true state $u$ in the input space $U$ and then a measured signal $\theta$ (where $\theta(u_n)=\theta(u(t_n))$ for $n\in\mathbb{N}$.
 The sequence $\{x_n\}$ represents a scalar time-series and intuitively $\theta$ may be thought to represent a probe inserted into a larger system and itself measuring/extracting only a small part of the greater system state at time $t=n$.
@@
 
-From here we construct a multidimensional observable using the method of delay-coordinate stacking. 
-\textbf{move to later} Given the sequence $\{\theta_1, \theta_2, \ldots\}$, we define the variable \newline $y_n:=[\theta_n, \theta_{n-\tau}, \ldots, \theta_{n-(2d-2)\tau}, \theta_{n-(2d-1)\tau}]^{T}\in\mathbb{R}_{d}\times{1}$ where $\tau$ represents the lag and $d$ the dimension of the attractor.
+% From here we construct a multidimensional observable using the method of delay-coordinate stacking. 
+% \textbf{move to later} Given the sequence $\{\theta_1, \theta_2, \ldots\}$, we define the variable \newline $y_n:=[\theta_n, \theta_{n-\tau}, \ldots, \theta_{n-(2d-2)\tau}, \theta_{n-(2d-1)\tau}]^{T}\in\mathbb{R}_{d}\times{1}$ where $\tau$ represents the lag and $d$ the dimension of the attractor.
 (We shall return to these terms shortly and their discussion may be put on hold for the time being).
 
-However, the true state $u$ of a system is seldom, if ever fully known.
+% However, the true state $u$ of a system is seldom, if ever fully known.
 In almost all cases we can merely insert probes into a system to obtain partial information by measurements taken.
 Moreover, the process of taking a measurement itself introduces 2 additional difficulties:
-\vspace{-8mm}
-\begin{enumerate}[noitemsep, label=\roman*.]
-  \item A series of measurements over a specific time-interval is inherently a discretisation
-  procedure of the underlying continuous-time dynamical system. (Hence why we restricted our attention to discrete-time systems in the preceding section)
-  \item A probe will never be fully accurate and so the act of measurement introduces a certain
-measure of numerical noise/inaccuracy.
-\end{enumerate}
+% \vspace{-8mm}
+% \begin{enumerate}[noitemsep, label=\roman*.]
+%   \item A series of measurements over a specific time-interval is inherently a discretisation
+%   procedure of the underlying continuous-time dynamical system. (Hence why we restricted our attention to discrete-time systems in the preceding Chapter)
+%   \item A probe will never be fully accurate and so the act of measurement introduces a certain
+% measure of numerical noise/inaccuracy.
+% \end{enumerate}
 
-More formally we may say that in the process of measuring a system, we obtain a 1-dimensional time-series $\{\theta(u(t))\}$, where $\theta:\mathbb{R}^n\to\mathbb{R}$ is a smooth observation or measurement function.
+% More formally we may say that in the process of measuring a system, we obtain a 1-dimensional time-series $\{\theta(u(t))\}$, where $\theta:\mathbb{R}^n\to\mathbb{R}$ is a smooth observation or measurement function.
 This series represents a possibly noise-augmented dataset containing information of certain parts of the original system.
 
-We may then ask ourselves the question: Is it in any way possible to retain information about the system state $x(t)$ in this temporal data-series $\theta(x(t))$?
+% We may then ask ourselves the question: Is it in any way possible to retain information about the system state $x(t)$ in this temporal data-series $\theta(x(t))$?
 The answer is easily yes if T is known.
@@
 
-\subsection{Taken Embedding Theorem}
+% \subsection{Taken Embedding Theorem}
 
-We define first the concepts of homeomorphism and embedding:
-\begin{Definition}
-  A homeomorphism is a function $f:Z\rightarrow Y$ between two topological spaces $Z$ and $Y$ that is continuous, bijective and has a continuous inverse. 
-\end{Definition}
+% We define first the concepts of homeomorphism and embedding:
+% \begin{Definition}
+% [\bf Homeomorphism] \label{Dfn_Homeo}
+%   A homeomorphism is a function $f:Z\rightarrow Y$ between two topological spaces $Z$ and $Y$ that is continuous, bijective and has a continuous inverse. 
+% \end{Definition}
 
-\begin{Definition}
-  Consider a homeomorphism $f:Z\rightarrow Y$ for $Y\in X$. $Z$ is said to be embedded in $X$ by $f$.
-\end{Definition}
+% \begin{Definition}
+% [\bf Embedding] \label{Dfn_Homeo}
+%   Consider a homeomorphism $f:Z\rightarrow Y$ for $Y\in X$. $f$ is said to be an embedding and $Z$ to be embedded in $X$ by $f$.
+% \end{Definition}
 
-Takens Theorem states a result that establishes a relationship between the observed and underlying dynamical systems by showing that the concetantion of a sufficiently large number of previous observations into a vector will, under certain conditions, generate a map between the vectors from the respective systems.
+% Takens Theorem states a result that establishes a relationship between the observed and underlying dynamical systems by showing that the concetantion of a sufficiently large number of previous observations into a vector will, under certain conditions, generate a map between the vectors from the respective systems.
 \textbf{Check wording} We formultate the theorem from \cite{takens1981detecting}.
 
-\begin{Theorem} 
-	[\bf Takens Embedding Theorem (adopted from \cite{takens1981detecting}] \label{Thm_Takens}
-            Let $W$ be a compact manifold of dimension $m$, and $d\ge m$ so that $2d\in\mathbb{Z}$. It is a 
-            generic property for the pair $(T, \theta)$,  where $T:W \to W$ is
-            a smooth diffeomorphism, and $\theta:W \to \mathbb{R}$ a smooth function, that the map $\Phi_{2d,\theta}:W \to \mathbb{R}^{2d+1}$ defined on $W$ by 
-            $\Phi_{2d,\theta}(w) := (\theta(T^{-2d}w)\ldots,\theta(T^{-1}w),\theta(w))$
-            is a diffeomorphic embedding; by `smooth' we mean at least $C^2$. Consequently, there exists a map $\Ftheta: \Phi_{2d,\theta}(W) \to \Phi_{2d,\theta}(W)$ defined by 
-            \begin{equation}  \label{eqn_takens}
-              \Ftheta: (\theta(T^{-2d}w),\ldots,\theta(T^{-1}w),\theta(w)) \mapsto  (\theta(T^{-2d+1}w),\ldots,\theta(T^{-1}w),\theta(Tw))
-            \end{equation}
-           so that $(W,T)$ is topologically conjugate to 
-            $(\Phi_{2d,\theta}(W), \Ftheta)$.   
-\end{Theorem}
+% \begin{Theorem} 
+% 	[\bf Takens Embedding Theorem (adopted from \cite{takens1981detecting}] \label{Thm_Takens}
+%             Let $W$ be a compact manifold of dimension $m$, and $d\ge m$ so that $2d\in\mathbb{Z}$. It is a 
+%             generic property for the pair $(T, \theta)$,  where $T:W \to W$ is
+%             a smooth diffeomorphism, and $\theta:W \to \mathbb{R}$ a smooth function, that the map $\Phi_{2d,\theta}:W \to \mathbb{R}^{2d+1}$ defined on $W$ by 
+%             $\Phi_{2d,\theta}(w) := (\theta(T^{-2d}w)\ldots,\theta(T^{-1}w),\theta(w))$
+%             is a diffeomorphic embedding; by `smooth' we mean at least $C^2$. Consequently, there exists a map $\Ftheta: \Phi_{2d,\theta}(W) \to \Phi_{2d,\theta}(W)$ defined by 
+%             \begin{equation}  \label{eqn_takens}
+%               \Ftheta: (\theta(T^{-2d}w),\ldots,\theta(T^{-1}w),\theta(w)) \mapsto  (\theta(T^{-2d+1}w),\ldots,\theta(T^{-1}w),\theta(Tw))
+%             \end{equation}
+%            so that $(W,T)$ is topologically conjugate to 
+%             $(\Phi_{2d,\theta}(W), \Ftheta)$.   
+% \end{Theorem}
 
-(In our scenario, the input space $U$ is considered the attractor.)
+% (In our scenario, the input space $U$ is considered the attractor.)
 
-Some explanation is beneficial.
+% Some explanation is beneficial.
 Takens establishes a delay-coordinate map $\Ftheta$ defined in \ref{eqn_takens} for $T$ the flow (\ref{defn_flow}) defined by the update equation  $Tu_n=u_{n+1}$ and $theta$ our measurement function.
@@
 
-Alternatively we may also make the statement: Takens' Embedding Theorem guarantees that almost all dynamical systems can be reconstructed from just one noiseless observation sequence \textbf{this isn't obvious from the theorem's formulation, is it?} i.e.
+% Alternatively we may also make the statement: Takens' Embedding Theorem guarantees that almost all dynamical systems can be reconstructed from just one noiseless observation sequence \textbf{this isn't obvious from the theorem's formulation, is it?} i.e.
 for a great number of possible observation functions $\theta$, $\Ftheta$ preserves the topology of U.
@@
 
-\begin{figure}[ht]
-  \includegraphics[scale=0.3]{takensmap.png}
-  \centering
-  \label{fig:takensmap}
-\end{figure}
+% \begin{figure}[ht]
+%   \includegraphics[scale=0.3]{takensmap.png}
+%   \centering
+%   \label{fig:takensmap}
+% \end{figure}
 
-Once again we pause and consider the graph to consolidate our understanding of the conjugacy (\textbf{are we being too simplistic?}).
+% Once again we pause and consider the graph to consolidate our understanding of the conjugacy (\textbf{are we being too simplistic?}).
 If we are in state $u$ and then evolve towards state $Tu$ before embedding into $\mathbb{R}^{2m-1}$, then this is equivalent to first embedding into $\mathbb{R}^{2m-1}$ and then evolving through the map $\Ftheta$, i.e.
@@
 
-Thus information about U can be retained in the time series' (or observation's) output.
+% Thus information about U can be retained in the time series' (or observation's) output.
 By preserving the topology on the manifold $U$ in the reconstruction space $X$ (\textbf{not discussed before}), we are also guaranteed the preservation of topological invariants of the manifold, of which dimensionality is one such invariant.
@@
 
-\subsection{Obstacles to Takens}
-Despite the fact that the Takens' Embedding Theorem is a powerful result and provides compelling reason to believe that it is possible to accurately ()\emph{right word?}) reconstruct the underlying space, it does present some serious limitations:
-\vspace{-5mm}
-\begin{enumerate}
-\item Even if we can indeed find \Ftheta{}  our \emph{approximation} of \Ftheta{}  is a map from a larger set $\mathbb{R}^{2m+1}$ containing the embedded attractor.
+% \subsection{Obstacles to Takens}
+% Despite the fact that the Takens' Embedding Theorem is a powerful result and provides compelling reason to believe that it is possible to accurately ()\emph{right word?}) reconstruct the underlying space, it does present some serious limitations:
+% \vspace{-5mm}
+% \begin{enumerate}
+% \item Even if we can indeed find \Ftheta{}  our \emph{approximation} of \Ftheta{}  is a map from a larger set $\mathbb{R}^{2m+1}$ containing the embedded attractor.
 There are, however, no theoretical guarantees that \Ftheta{}  will keep this attractor $U$  the same.(\textbf{We haven't properly said that U is an attractor, or have we.})
-\item Takens only holds for noiseless observations.
+% \item Takens only holds for noiseless observations.
 Suppose the case where \Ftheta{} was learnt, but due to measurement error or other noise we obtained the set $V$ and not the original attractor.
@@
 This problem will later be eliminated by some properties of a new function $g$, namely the Unique Solution Property.
-\item It is an unfortunate case that the functional complexity (i.e., the relative number of oscillatory graphs \cite{manjunath2021universal}) as well as stability of the embedding \Ftheta{} (\textbf{cite?})depends on the choice of observable $\theta$.
+% \item It is an unfortunate case that the functional complexity (i.e., the relative number of oscillatory graphs \cite{manjunath2021universal}) as well as stability of the embedding \Ftheta{} (\textbf{cite?})depends on the choice of observable $\theta$.
 This means that we may even learn a map that complicates our learning task and in so doing adds obstacles to the road to learning a conjugate map with predictive power.
 (\textbf{Am I talking nonsense here?})
-\end{enumerate}
+% \end{enumerate}
 
-Another limitation owes itself to the requirement that T be a surjective map. This is not always the case in practical examples and we may consider the case of a damped double pendulum.
+% Another limitation owes itself to the requirement that T be a surjective map. This is not always the case in practical examples and we may consider the case of a damped double pendulum.
 
-\textbf{Continue writing here}
+% \textbf{Continue writing here}
 
-\begin{figure}[ht]
-  \includegraphics[scale=0.3]{learningfailure.png}
-  \centering
-  \label{fig:supp_learning_failure}
-  \caption{Illustration of a system having failed to learn the map}
-\end{figure}
+% \begin{figure}[ht]
+%   \includegraphics[scale=0.3]{learningfailure.png}
+%   \centering
+%   \label{fig:supp_learning_failure}
+%   \caption{Illustration of a system having failed to learn the map}
+% \end{figure}
 
-\textbf{Explanation continued here}
+% \textbf{Explanation continued here}
 
-\begin{figure}[h]
-  \includegraphics[scale=0.4]{temporary_damp_fig.png}
-  \centering
-  \label{fig:damped_pendulum}
-\caption{A damped pendulum}
-\end{figure}
+% \begin{figure}[h]
+%   \includegraphics[scale=0.4]{temporary_damp_fig.png}
+%   \centering
+%   \label{fig:damped_pendulum}
+% \caption{A damped pendulum}
+% \end{figure}
 
-\textbf{Continue explaining here}
+% \textbf{Continue explaining here}
+\newpage
 
-\section{Obtaining X through Driven Dynamical Systems}
+\section{Driven Dynamical Systems to Forecast Problems} \label{ch4}
 
-In this chapter, we recall results on mapping temporal data acquired from a discrete-dynamical system onto a different space through a driven dynamical system.
-We recall the conditions the driven system should have not to add distortion to its state space representation, so that the single-delay dynamics (SDD) gets conjugate or semi-conjugate to the system.
-The SDD can be then use to forecast and reconstruct the underlying attractor
+In this chapter, we discuss results pertaining to the mapping of temporal data obtained from a discrete-dynamical system onto a different space through the notion of a driven dynamical system. 
+We also consider the conditions a driven system should have to avoid adding distortion to its state space representation, as well as then ensuring that the single-delay dynamics (SDD) are conjugate (or at least semi-conjugate to the system).
+The SDD can then be used to forecast and reconstruct the underlying attractor. \textbf{Is this attractor assumed to exist?}
 
@@
 
-A nonautonomous dynamical system is simply a dynamical system where the input $u$ from the input space $U$, a  topological space is time-dependent.
+A nonautonomous dynamical system is simply a dynamical system (as defined before in \textbf{link to above}) where the input $u$ from the input space $U$, a  topological space, is time-dependent.
 \textbf{This means it can no longer be generated by a flow?}
 
-We now recall the figure first defined in \ref{eqn_conjugacy} below and the greater goal to construct a system - such as $(V,S)$  - conjugate to $(U,T)$,
+We immediately recall the figure first defined in \ref{eqn_conjugacy} and remind ourselves of our ultimate goal to forecast the evolution of an unknown nonautonomous dynamical system $(U,T)$ by constructing a conjugate system such as $(V,S)$ below.
 
@@
 
-To do this, we need to consider a driven dynamical system taking input from both an input space $U$, and an underlying state space $X$.
-This is done to account for influence of both the input and the actual state of the system at a specific timestep.
+To do this, the need arises to consider a driven dynamical system where input is taken from both an input space $U$, and an underlying state space $X$. 
+This is done to account, in a manner more similar to the true scenario, for the influence of both the input and the actual state of the system at a specific timestep.
 
 \begin{Definition}
-A driven dynamical system comprises two topological spaces $U$, $X$ and a continuous function  $g:U\times{X}\to{X}$ where $g(u_n, x_n)=x_{n+1}$.
-The dynamics on X are generated by the update equation $x_{n+1}=g(u_n, x_n)$ where $n\in\mathbb{Z}$, $u_n$ from the metric input space $U$ and state $x_n$ belonging to the state space $X$, a compact metric space.
+  [\bf Driven and Compactly Driven Dynamical System] \label{Dfn_DDS}
+A driven dynamical system comprises two topological metric spaces $U$, $X$ and a continuous function  $g:U\times{X}\to{X}$ where $g(u_n, x_n)=x_{n+1}$. 
+The dynamics on X are generated by the update equation $x_{n+1}=g(u_n, x_n)$ where $n\in\mathbb{Z}$, input $u_n$ from $U$ and state $x_n$ belonging to $X$, where $X$ is compact. 
 If the input space $U$ is compact, we refer to the system as compactly driven. 
-Abbreviated, we shall refer only to the "driven system $g$", with all other entities being understood implicitly.
+Abbreviated, we shall refer only to the \emph{driven system $g$}, with all other entities being understood implicitly.
 \end{Definition}
 
-In particular, a nonautonomous dynamical system may be generated from $U$; any input $\overline{u}$, a bi-infinite sequence from $U$ gives rise the sequence of self-maps $\{g(u_n, \ldotp)\}_{n\in\mathbb{Z}}$ contained in $X$.
-\textbf{explain better why}
-Physically, one may think of this bi-infinite as referring to a system that has been running for an incredibly long period at the time of the first measurement/observation taken from the system.
+In particular, a nonautonomous dynamical system may be generated from $U$; any input $\overline{u}$, a bi-infinite sequence from $U$, gives rise to the sequence of self-maps $\{g(u_n, \cdot)\}_{n\in\mathbb{Z}}$ contained in $X$.
+Physically, one may think of this bi-infinite sequence as referring to a system that has been running for an incredibly long period at the time of the first measurement taken from the system (alternatively, the first time a probe is inserted into the system to take an observation).
 
 \begin{Definition}
-  A sequence $\{x_n\}_{n\in\mathbb{Z}}\subset{U}$ is called an entire solution(or simply a solution) to the driven system  $g$ with input $\overline{u}$ when it satisfies 
-  $$g(u_{n-1}, x_{n-1})=x_n$$ for all $n$ in $\mathbb{Z}$
+  [\bf Entire Solution] \label{Dfn_Soln}
+  A sequence $\{x_n\}_{n\in\mathbb{Z}}\subseteq{U}$ is called an entire solution(or simply a solution) to the driven system  $g$ with input $\overline{u}$ when it satisfies 
+  \[g(u_{n-1}, x_{n-1})=x_n\] for all $n\in\mathbb{Z}$
 \end{Definition}
@@
 
-\begin{Example}
-  Consider the driven system $g(u_n,x_n)=x^n$ for $X=[0,1]$.
-This system has an uncountable number of solutions, because for every $x\in{X}$, there exists a solution passing through the point $x$ and $lim_{n\to\infty}x_n=0$, $lim_{n\to{-}\infty}x_n=1$.
+\begin{Example}\label{ex_halfux}
+  The only solution  $\{x_n\}_{n\in\mathbb{Z}}$ to the driven system  $g(u,x)=\frac{ux}{2}$, where $X=[0,1]$, $U=[0,1]$,  is the zero solution $x_n\equiv0$.
+  To see this, consider any $x_n=a\in[0,1]$ where $a\neq{0}$.  Let $\overline{u}\in{U}$ be an non-zero constant sequence , say $u_n=0.5$. 
+  The driven system may be rewritten as $x_{n-1}=\frac{2x_n}{u_{n-1}}=4x_n$ and the  iterates of $x_n$ in backward time will increase by a factor of 4 at each timestep. 
+  Thus for some $m\leq{n}$,  $1<x_m$ i.e. $x_m\notin{X}$. So ${\{x_n\}}_{n\in\mathbb{Z}}$ is not a solution and it follows that the only possible solution is the zero solution.
 \end{Example}
 
+A system may also have multiple solutions as is evidenced in the example below.
+
+\begin{Example} \label{Ex_exp}
+ Consider the driven system $g(u_n,x_n)=x^n$ for $X=[0,1]$, $U=\mathbb{R}$.
+The system has an uncountable number of solutions, as there exists a solution for every $x\in{X}$ which also passes through the point $x$ and $lim_{n\to\infty}x_n=0$, $lim_{n\to{-}\infty}x_n=1$.
+ The proof is deferred to immediately after the next paragraph.
+
+\end{Example}
+
+As the solutions to a driven system are often considered, we next identify a subspace $X_U$ of $X$ that contains all possible solutions.
+To realize such a subspace of a driven system $g$, the concept of a reachable set is defined.
+
 \begin{Definition}
-Let $X_n{\overline{u}}:= \cap_{m<n}{\phi_{\overline{u}}(n,m,X)}$ be the set of all reachable states at time $n$ for input $\overline{u}$.
-  A function $g:U\times{X}\to{X}$ is a topological contraction if for all $n\in\mathbb{Z}$ and all $\overline{u}\subset{U}$, $X_n(\overline{u})$ is a singleton subset of X.  
+  [\bf Reachable Set] \label{Dfn_ReachableSet}
+The reachable set of a driven system $g$ is exactly the union of all the elements of all the solutions, i.e., 
+\[X_U :=\Big \{x \in X:  x = x_k \mbox{ where  $\{x_n\}$  is a solution for some  $\bar{u}$} \Big \}.\]
+The set of all reachable states at a specific time $n$ for input $\overline{u}$ is denoted by $X_n(\overline{u})$
 \end{Definition}
 
-Note that $x\in{X_n{\overline{u}}}$ if and only g has a solution $\{x_k\}$ s.t.
-$x_n=x$ and the input is $\overline{u}$.
-\textbf{cite paper}.
-This will lead to a result established later, but which is worth taking note of now: $g$ being a topological contraction is equivalent to the existence of a unique entire solution for every $n\in\mathbb{Z}$.
-\textbf{Adding worth or not?} \label{here_UAP_firstestablished}
+We note that $x\in{X_n(\overline{u})}$ if and only $g$ has a solution $\{x_k\}$ for $x_n=x$ and input $\overline{u}$. \textbf{Cite}.
+This will lead to a result established later, but which is worth taking note of now: \textit{$g$ being a topological contraction is equivalent to the existence of a unique entire solution.} To this end we define the notion of a topological contraction.
 
-\begin{Example}\label{ex_halfux}
-  If $X=[0,1]$, $U=[0,1]$, then the only possible solution $\{x_n\}$ to the driven system $g(u,x)=\frac{ux}{2}$ is the zero solution $x_n\equiv0$.
+\begin{Definition}
+  [\bf Topological Contraction] \label{Dfn_TopContr}
+  A function $g:U\times{X}\to{X}$ is a topological contraction if for all $n\in\mathbb{Z}$ and all $\overline{u}\subseteq{U}$, $X_n(\overline{u})$ is a singleton subset of X.  
+\end{Definition}
 
-  To see this, we show that for every input $\{u_n\}$, $g(u_k,\dot)$ is a contraction map on $X$ for every $k\in\mathbb{Z}$. \textbf{Show}
-  Consequently, we easily conclude that $\{x_n\}$ where $x_n=0$ for all $n\in\mathbb{Z}$ is the only solution. \textbf{Show}
-\end{Example}
+\begin{proof}
+  [\bf Proof of Example \ref{Ex_exp}] 
+  To see this, we show that for every input $\overline{u}=\{u_n\}_{n\in\mathbb{Z}}$, $g(u_n,\cdot)$ is a contraction map on $X$ for every $k\in\mathbb{Z}$. 
+  Indeed $g(u_n,x_n)=x^n$ depends only on the state at time $n$ and so for each input $\overline{u}$, $X_n(\overline{u})=\{x^n\}$ is a singleton subset.
+  We then easily conclude that $\{x_n\}\equiv0$ is the only solution. 
+\end{proof}
 
-One may ask if a driven system always has a solution, and if so, whether it has certain properties such as uniqueness.
-Should the system be compact, existence follows immediately as can be seen in the following result.
+Thus far it has been demonstrated that a system may have one or more solutions; one may ask if a driven system always has a solution and, if so, whether it satisfies certain properties such as uniqueness.
+Should the driven system be compact, existence follows immediately as shown in the following result.
 
 %%%%%%%%%%%%%%%%%%%%%ISSUE HERE - fix before sending
-\begin{Theorem}
+\begin{Theorem} \label{Thm_CompactExistence}
   If $X$ is compact then for each input $\overline{u}$, there exists at least one solution to the driven system $g(\ldotp, x)$
@@
 \begin{proof}
-  Consider ${u_n}_{n\in\mathbb{Z}}\subset{U}$ and $g:U\times{X}\to{X}$  generating a sequence ${x_n}_{n\in\mathbb{Z}}$ in the compact space.
-(If U is a metric space then $x_n$ has a convergent subseqeunce)
+  Consider an input $\overline{u}={\{u_n\}}_{n\in\mathbb{Z}}\subseteq{U}$ and driven system $g:U\times{X}\to{X}$  generating a sequence ${\{x_n\}}_{n\in\mathbb{Z}}$ in the compact space $X$. 
+  Since $X$ is a metric space, it follows immediately from a well-known result in Analysis (\textbf{Cite}) that $\{x_n\}$ has a convergent subsequence which then is a solution. 
 \end{proof}
 
-\begin{Example}
-  The driven system $g(u,x)=x$ has only the constant solution $x$ and so for $U=[-1,1]$, the system has no solution if $|X|>1$
-\end{Example}
-
-The examples above present systems with trivial sets of solutions. To refine the scenario, we define systems with unique solutions.
+We may easily construct many systems with trivial solution-sets, such as $g(u,x)=x$ which has only the constant solution $x$ and so for $U=[-1,1]$, the system would have no solution if $|x|>1$.
+To refine the scenario, we consider only systems with unique solutions.
 
 \begin{Definition}
-  A driven system $g$ is said to have the Unique Solution Property(USP) if for each input $\overline{u}$ there exists exactly one solution.
-Alternatively we may formulate the USP as follows: $g$ has the Unique Solution Property if there exists a well-defined map $\Psi:{U}\to{X}$ where $\Psi({\overline{u}})$ denotes the unique solution
+  [\bf Unique Solution Property)] \label{Dfn_usp}
+  A driven system $g$ is said to have the Unique Solution Property (USP) if for each input $\overline{u}$ there exists exactly one solution. 
+  Alternatively we may formulate the USP as follows: $g$ has the Unique Solution Property if there exists a well-defined map $\Psi:{U}\to{X}$ with $\Psi({\overline{u}})$ denoting the unique solution.
 \end{Definition}
 
-\emph{(More discussion on USP … - not yet filled in ...)}
+\emph{(More discussion on USP?)}
 
-We next identify a subspace $X_U$ of $X$ that contains all possible solutions. To realize such a subspace of a driven system $g$, we define 
-the \textit{reachable set} of the driven system $g$ to be the union of all the elements of all the solutions, i.e., 
-$$X_U :=\Big \{x \in X:  x = x_k \mbox{ where  $\{x_n\}$  is a solution for some  $\bar{u}$} \Big \}.$$
+Having already defined the reachable set $X_U$, we pause for a moment to fix additional notation.
+Letting $\cev{u}^{n}:=(\ldots,u_{n-2} ,u_{n-1})$ be the left-infinite subsequence of an input, $\overleftarrow{U}$ then denotes all these left-infinite sequences in $U$. 
+Moreover, $\cev{u}^{n}v:=(\ldots,u_{n-2} ,u_{n-1}, v)$ is to symbolise the input up to time $n$ with $v \in U$ being the specific input value at time $n$. 
+The introduction of a new input at time $n$ can be described by a mapping $\sigma_v:   \cev{u}^{n} \mapsto \cev{u}^{n}v$.
 
-We denote $\cev{u}^{n}:=(\ldots,u_{n-2} ,u_{n-1})$ and $\overleftarrow{U}$ denote all the left-infinite sequences in $U$. 
-Symbolically we let $\cev{u}^{n}v:=(\ldots,u_{n-2} ,u_{n-1}, v)$ denote the input up to time $n$ with $v \in U$ being the input value at time $n$.
-This introduction of a new input at time $n$ can be described by a mapping $\sigma_v:
-\cev{u}^{n} \mapsto \cev{u}^{n}v$.   \textbf{Is this in the right place?}
-
-The question now becomes whether we may establish a conjugacy as presented below for g driven
+The question now becomes whether we may establish a conjugacy as presented below for $g$ driven
 
@@
 
-  \begin{Definition} \rm \label{Def_UC}
+  Restricting our attention more and more, we now consider a specific subclass of conjugacies.
+
+  \begin{Definition} 
+    [\bf Universal Semi-Conjugacy)] \label{Def_UnivSemiConj}
     Given a driven system $g$, we  call a continuous and surjective map $h : \overleftarrow{U} \to X_U$ a universal semi-conjugacy if  diagram \ref{Scomm_h} commutes for all $v \in U$.
@@
 
-If the universal semi-conjugacy $h$ exists (i.e.
-the diagram  \ref{Scomm_h} commutes) then the solution $\Psi$ will be at least a coarse-grained representation of the input $u$; we may now consider the question of the existence of such a function $h$ for the driven system defined above.
+If the universal semi-conjugacy $h$ exists (i.e. the diagram in \ref{Scomm_h} commutes) then the solution $\Psi$ will be at least a coarse-grained representation of the input $u$.
 
-If $g$ has the USP and $\Psi(u)=\{x_n\}$ then $h$, defined by  $h(\overleftarrow{u}_n)=x_n=g(u_n,x_{n-1})$, will satisfy the semi-conjugacy in the graph above \ref{Scomm_h}.
+Does such a function $h$ for the driven system defined above exist?
+If $g$ has the USP and $\Psi(u)=\{x_n\}_{n\in\mathbb{Z}}$ then $h$, defined by  $h(\cev{u}_n):=g(u_n,x_{n-1})=x_n$, will satisfy the semi-conjugacy in the graph above \ref{Scomm_h}.
+Regrettably, the mapping $h$ is not guaranteed to exist in general.
 
-In general the mapping h is not guaranteed to exist. We thus state the following result.
-
-\begin{Theorem}
-  For a compactly driven system, a causal mapping $H$ exists if and only if $g$ has the USP. 
-\end{Theorem}
-\begin{proof}
-  \emph{See proof from p.15 of \cite{manjunath2021universal}}
-\end{proof}
-
-Note that even when $h$ does exist, we are not guaranteed its injectivity.
-Considering again example 3* \ref{ex_halfux}, we conclude that even if $h$ were to exist, it could not be injective as $X_U=\{0\}$.
-
-An embedding of the space $\overleftarrow{U}$   would allow one to establish topological conjugacy, which in turn provides stronger results than merely obtaining a coarse-grained representation via a causal mapping.
-
-Re-sketching the graph \ref{Scomm_h} above by fixing $v$ in $g$ and replacing $X_U$ by its left-infinite sequence space $\overleftarrow{U}$, we obtain the graph below.
+Re-sketching the graph \ref{Scomm_h} above by fixing the input $v$ in $g$ and replacing $X_U$ by its left-infinite sequence space $\overleftarrow{U}$, we obtain the graph below.
 In this case, the function $H:\overleftarrow{U}\to\overleftarrow{X}_U$, a map that is both continuous and surjective, is called a \emph{causal mapping}.
@@
   The driven system $g$ can induce an embedding of $\overleftarrow{U}$ in $\overleftarrow{X}$ as follows:
-  If $H$ is also injective (and hence bijective), it becomes the embedding of $\overleftarrow{U}$ in $\overleftarrow{X}$ induced by the driven system $g$ and we refer to $H$ as a \emph{causal embedding}.
+
+\begin{Definition}
+  [\bf Causal Embedding]\label{Def_CausEmbed}
+  If $H:\overleftarrow{U}\to\overleftarrow{X}_U$ is also injective (in addition to being surjective and hence bijective), it becomes the embedding of $\overleftarrow{U}$ in $\overleftarrow{X}$ induced by the driven system $g$ and we refer to $H$ as a \emph{causal embedding}.
+\end{Definition}
+
+\begin{Theorem}
+  For a compactly driven system, a causal mapping $H$ exists if and only if $g$ has the USP. 
+\end{Theorem}
+\begin{proof}
+  \emph{See proof from p.15 of \cite{manjunath2021universal}}
+\end{proof}
+
+Note that even when $h$ does exist, we are not guaranteed its injectivity.
+Considering again example \ref{ex_halfux}, we conclude that even if $h$ were to exist, it could not be injective as $X_U=\{0\}$.
+
+An embedding of the space $\overleftarrow{U}$   would allow one to establish topological conjugacy, which in turn provides stronger results than merely obtaining a coarse-grained representation via a causal mapping.
 
@@
 In broad sweeps, the inverse-limit system of a dynamical system is a self-map on a subset of an infinite-dimensional space \textbf{(expand what exactly this is?)} where each point in the inverse-limit space corresponds to a backward orbit of the map $T$.
-In literature, the inverse-limit space is denoted by $\hat{U}_T\subset\overleftarrow{U}$.
+In literature, the inverse-limit space is denoted by $\hat{U}_T\subseteq\overleftarrow{U}$.
 \textbf{cite from supp.koopman}.
@@
 The function $$g(u,x)=(1-a)x+atanh(Au+\alpha{a}Bx)$$  is both SI-invertible and possesses the USP.
-This function $g$ is used in our implementation and is more completely discussed in Section 5 \ref{sect5}
+This function $g$ is used in our implementation and is more completely discussed in chapter 5 \ref{sect5}
 The proof of these two facts is rather involved and the proof, instead of being replicated here, may be found in (cite relevant article)
@@
 \emph{graph here causes issues currently}
-\begin{center}
-%    \[ 
-      \psset{arrows=->, arrowinset=0.25, linewidth=0.6pt, nodesep=3pt, labelsep=2pt, rowsep=0.7cm, colsep = 1.1cm, shortput =tablr}
-      \everypsbox{\scriptstyle}
-      \begin{psmatrix}
-      \widehat{U}_T  && \widehat{U}_T \\
-      Y_T && Y_T \\
-      %%%
-     %  \ncline{1,1}{1,2}^{\widehat{T}} \ncline{1,1}{2,1} <{h}
-     %  \ncline{1,2}{2,2} > {H_2}
-     %  \ncline{2,1}{2,2}^{G_T}
-      \end{psmatrix}
-     % \]
-    \end{center}
+% \begin{center}
+% %    \[ 
+%       \psset{arrows=->, arrowinset=0.25, linewidth=0.6pt, nodesep=3pt, labelsep=2pt, rowsep=0.7cm, colsep = 1.1cm, shortput =tablr}
+%       \everypsbox{\scriptstyle}
+%       \begin{psmatrix}
+%       \widehat{U}_T  && \widehat{U}_T \\
+%       Y_T && Y_T \\
+%       %%%
+%      %  \ncline{1,1}{1,2}^{\widehat{T}} \ncline{1,1}{2,1} <{h}
+%      %  \ncline{1,2}{2,2} > {H_2}
+%      %  \ncline{2,1}{2,2}^{G_T}
+%       \end{psmatrix}
+%      % \]
+%     \end{center}
 
@@
 \item If $T$ is also (and at this stage this is very much our hope) a homeomorphism, then $(Y_T, G_T)$ topologically conjugate to $(\widehat{U}, \widehat{T})$ in which $(U,T)$ has been embedded! 
-
+\end{enumerate} 
 One can therefore learn the single-delay lag dynamics of the driven system states via $G_T$ with enough data thanks to the USP.
@@
 
-\end{enumerate}
+\subsection{A discussion of $G_T$ }
+
+By establishing the existence of the map $G_T$, we’ve essentially embedded the attractor $U$ \textbf{So far we haven't really spoken about an attractor} into the higher dimensional space $X\times{X}$.
+In layman’s terms, this ensures that there is more “dimensional room” for the underlying system’s underlying dynamics to “move”.
+As the dynamics aren’t as “squashed”, we might therefore hope that the dynamics of $G_T$ are in some sense simpler than that of $T$.
+(Taken note of the fact that $G_T$ is a homeomorphism even when $T$ is just continuous)
+
+In \emph{[Koopman]} it is illustrated in an empirical fashion in that the map $G_T$ describes dynamics which are less functionally complex than that of $T$ or of the map $\Phi_{2d,\theta}$ (as discussed above in the section on Takens).
+This is done by implementing a Recurrent Neural Network (RNN), but not discussed as of yet.
+\emph{(Add in that we'll be using Pearson coefficient?)}
+
+We opt to learn $G_T$ in an indirect manner by defining a new map $\Gamma$. It follows immediately from $G_T$’s existence that $\Gamma$ also exists.
+
+The reasons for taking this roundabout approach remain to be discussed, but a pat answer may be immediately given: When we have learnt $\Gamma$, we can drive the system autonomously and then we know $G_T$ anyway.
+
+\begin{Theorem}
+  When $x_n$ and $x_{n-1}$ are successive points on a solution obtained for input-orbit of $T$ $\{u_n\}$, then the map $\Gamma: X\times{X}\to{U}$ defined by $(x_{n-1},x_n)\mapsto{u_n}$ exists whenever $G_T$ exists [Koopman, Theorem 3c]
+\end{Theorem}
+
+\begin{Definition}
+ The projection mapping $\pi_i$ is defined as follows: $$\pi_1:(a_1,a_2)\to{a_1}$$ and $$\pi_2:(a_1,a_2)\to{a_2}$$ 
+\end{Definition}
+
+We now have the graph below:
+
+\begin{figure}[ht]
+  \includegraphics[scale=0.3]{gammatemp.png}
+  \centering
+  \label{fig:conjugacy_with_Gamma}
+\end{figure}
+
+And a final set of equations - equations that have been constructed from data!
+\begin{eqnarray}
+	u_{k+1} &=& \pi_1 \circ (\Gamma, \pi_2) \circ (\pi_2,g) (u_k,x_k) \label{Seqn_u}\\
+	x_{k+1} &=& \pi_2 \circ (\Gamma, \pi_2) \circ (\pi_2,g) (u_k,x_k). \label{Seqn_x}
+\end{eqnarray}
+
+The problem thus simplifies to the issue of learning the map $\Gamma$ and combining this with the projection mapping $\pi_2$ and the function $g$, which will be known.
+
+\subsection{Advantages of learning $\Gamma$}
+
+Why are we going such a lengthy roundabout route, one may immediately ask.
+Why not just learn the map $G_T$ from the get-go?
+On the surface, this seems to be an arbitrarily chosen path with no real reason for the choice made, so we take a pause again and discuss our reasoning for learning Gamma.
+There are a number of distinct advantages.
+
+Learning $\Gamma$ saves computational resources.
+This follows from the fact that the input $u_n$ at time $n$ lies in a low-dimensional subspace of the high-dimensional subspace $X\times{X}$ due to the fact that it is zero-padded.
+
+Input- and parameter-related stability is achieved if $g$ has the USP
+\newline \emph{Define the two types of stabilities}
+\newline \emph{Sentence on $u$, $v$, tails, USP and product topology}
+\newline If we opt to work with $Y-T$, errors can occur. \emph{Expand}
+\newline $\Gamma$ is more stable. \emph{Explain why}
+
+Global dissipativity prevents large numerical errors due to input data.
+$\Gamma$ is globally dissipative (\emph{Define}) whereas $G_T$ is not guaranteed to be.
+IF we were to rather learn $G_T$, we would risk having exactly the same problems arise that could arise also for Takens.
+\textbf{\emph{ADD THIS TO ERRORS IN TAKENS: NOT GUARANTEED GLOBALLY DISSIPATIVE}}
+\newline \emph{Expand here ...}
+
+\emph{Where to add:If $g$ has USP, then any solution to $g$ is a non-autonomous uniform attractor}
 

### 2021-12-29 Gandhi_Project_Report/draft2.tex
Section 4: third of the way in with refining

%\documentclass[12 pt]{article}

\documentclass[a4paper,12pt,twoside]{book}
\pagestyle{headings}
\usepackage[utf8]{inputenc}
\usepackage{filecontents}
\usepackage[many]{tcolorbox}
\usepackage{graphicx}
\usepackage{harpoon}
%\usepackage{apacite} 
\usepackage{xcolor}
\usepackage{latexsym}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{mdframed}
%\usepackage[authoryear]{natbib}
\usepackage{hyperref, url}
\usepackage[show]{ed}
\usepackage{subfigure}
\usepackage[ntheorem]{empheq} 
\usepackage{amsthm, amssymb, amsfonts, latexsym}
%\usepackage{pst-all}

\usepackage{enumitem}

\setlength{\parskip}{1.7em}
\setlength{\parindent}{0em}

%\setlength{\baselineskip}{20pt}

\setlength{\textwidth}{429.75499pt}
\setlength{\textheight}{643.20255pt}
\setlength{\oddsidemargin}{5 mm}
\setlength{\evensidemargin}{5 mm}
\setlength{\topmargin}{0 mm}
\setlength{\headsep}{0 mm}
\setlength{\headheight}{0 mm}

\usepackage{pst-node}

\newcommand{\citealtt}[1]{\citeauthor{#1},\citeyear{#1}}
\newcommand{\myycite}[1]{\citep{#1}}

\mathchardef\mhyphen="2D

%\newcommand{\cev}[1]{\reflectbox{\ensuremath{\vec{\reflectbox{\ensuremath{#1}}}}}}

\usepackage{pst-node,graphicx,pst-blur}
%\uspackage{auto-pst-pdf}
%\usepackage{tikz-cd}

\makeatletter
\DeclareRobustCommand{\cev}[1]{%
  \mathpalette\do@cev{#1}%
}
\newcommand{\do@cev}[2]{%
  \fix@cev{#1}{+}%
  \reflectbox{$\m@th#1\vec{\reflectbox{$\fix@cev{#1}{-}\m@th#1#2\fix@cev{#1}{+}$}}$}%
  \fix@cev{#1}{-}%
}
\newcommand{\fix@cev}[2]{%
  \ifx#1\displaystyle
    \mkern#20mu
  \else
    \ifx#1\textstyle
      \mkern#20mu
    \else
      \ifx#1\scriptstyle
        \mkern#26mu
      \else
        \mkern#26mu
      \fi
    \fi
  \fi
}

\makeatother

\newcommand{\xcm}{\epsfxsize=3.1cm}
\newcommand{\fig}[1]{\epsfbox}
% \newcommand{\bp}{\begin{minipage}{3.1cm}}
% \newcommand{\ep}{\end{minipage}}

\newtheorem{Definition}{Definition}[]
%\newmdtheoremenv{Theorem}{Theorem}[]
\newtheorem{Theorem}{Theorem}[]
\newtheorem{Lemma}{Lemma}[]
\newtheorem{Proposition}{Proposition}[]
\newtheorem{Corollary}{Corollary}[]
\newtheorem{Remark}{Remark}[]
\newtheorem{Example}{Example}[]

%Shortcut symbols
\newcommand{\Ftheta}{\ensuremath{F_\theta}}

\makeatletter 
\renewcommand{\thefigure}{\@arabic\c@figure}
\makeatother
\usepackage{xr}

%%%% PNAS GRAPHICS SETUP- FROM THE PNAS STYLE LATEX FILE
\RequirePackage{graphicx,xcolor}
\RequirePackage{colortbl}
\RequirePackage{booktabs}
\RequirePackage{algorithm}
\RequirePackage[noend]{algpseudocode}
\RequirePackage{changepage}
\RequirePackage[twoside,%
				letterpaper,includeheadfoot,%
				layoutsize={8.125in,10.875in},%
                layouthoffset=0.1875in,%
                layoutvoffset=0.0625in,%
                left=38.5pt,%
                right=43pt,%
                top=43pt,% 10pt provided by headsep
                bottom=32pt,%
                headheight=0pt,% No Header
                headsep=10pt,%
                footskip=25pt,
                marginparwidth=38pt]{geometry}
\RequirePackage[labelfont={bf,sf},%
                labelsep=period,%
                figurename=Fig. ]{caption}
\setlength{\columnsep}{13.5pt} % Distance between the two columns of text
\setlength{\parindent}{12pt} % Paragraph indent

%% Figure caption style
\DeclareCaptionFormat{pnasformat}{\normalfont\sffamily\fontsize{7}{9}\selectfont#1#2#3}
\captionsetup*{format=pnasformat}

%%% GREYBOX AROUND FIG
\definecolor{lightgray}{gray}{0.95}
\newcommand\greybox[1]{%
  \vskip\baselineskip%
  \par\noindent\colorbox{lightgray}{%
    \begin{minipage}{\textwidth}#1\end{minipage}%
  }%
  \vskip\baselineskip%
}

\newtcolorbox{paperbox}[1][]{
    enhanced,
    colback=white,
    boxrule=0pt,
    boxsep=0pt,
    arc=0mm,
    width=0.8\linewidth,
    fuzzy shadow={0mm}{-4pt}{-4pt}{1mm}{black!30!white},
    #1
}

%\externaldocument{Supp_Koopman-0808}
\begin{document}

\title{Project Thesis/Report}

\author{} 
\date{}
\date{}
\maketitle

\tableofcontents

\chapter{Introduction}

Experimenting on biological, physical, and artificial systems in order to generate a more informative dynamical model is a well-established practice in modern science. Traditional methods for modelling physical systems are based on laws of physics that are based either on empirical relationships or on intuition. For systems that evolve with time, physical laws yield mathematical equations that govern how the quantities evolve with time. However our world around is, for the most part, much more complex than that which can be distilled into elegant equations. We do not fully understand many of the more complex systems, nor do they even provide us with a good physical intuition of the underlying principles governing the dynamics. To complicate this - the underlying systems we observe often display sensitive dependence on initial conditions; despite having highly similar initial conditions, differing orbits diverge quickly and to such an extent that it becomes seemingly impossible to retrace their steps back to the original conditions. Even the presence of usually 'negligible' computational noise or measurement error renders long-term point-wise prediction infeasible. 
There are many difficulties that one encounters while modelling such systems:
\vspace{-8mm}
\begin{enumerate}[noitemsep, label=\roman*.]
  \item One may not have access to the complete states of the systems
  \item The actual system may be described by functions that behave wildly in the sense their graphs have a wild oscillatory behaviour, i.e., they have a large functional complexity.    
\end{enumerate}

Models derived primarily from data can be classified into three categories: 
\vspace{-8mm}
\begin{enumerate}[noitemsep, label=\roman*.]
  \item  Interpretable models (i.e., they establish relationships between internal physical mechanisms), 
  \item Partially interpretable models capturing some modes of the dynamics, 
  \item Non-Interpretable models, defined as such mainly due to the fact that they are defined on a different phase space that is usually high-dimensional. 
\end{enumerate}

Examples in the literature attempting to forecast data from such systems have tried a number of different approaches with differing degrees of success. 
When we have complete access to states of the system, an ordinary differential equation can be obtained from data (Citations to Sindy and \cite{small2002modeling,xu2006modeling}) wherein one could approximate the vector field by a library of functions to obtain interpretable models.

Recent partially interpretable models available in the literature have been based on the Koopman operator (e.g.,\cite{koopman1932dynamical,budivsic2012applied}) to  employ observables mapping the data onto a higher-dimensional space. This then makes the dynamics in the higher-dimensional space more amenable for approximation by a linear transformation. Such methods, obviously, do not guarantee an exact reconstructions for nonlinear models, and in practice provide poor long-term consistency for a large class of systems. (Add citation)

The non-interpretable models include the delay embedding and the machine learning algorithms. One could learn a system conjugate to the underlying system by applying the Takens delay embedding \cite{takens1981detecting} when one has ``good" observations from the system; this learnt system could then be used for forecast the observed data.  Takens delay embedding theorem \cite{takens1981detecting} and its various generalisations (e.g., \cite{sauer1991embedology, stark1999delay, gutman2018embedding, gutman2018embedding}) establishes the learnability of a system constructed by concatenating sufficiently large previous time-series observations of a dynamical system into a vector (called delay coordinates). This then establishes the existence of a map on the space of delay coordinates equivalent (or topologically conjugate) to the underlying map from which the observed time-series was first obtained. Although topological conjugacy guarantees an alternate representation of the underlying system, the quality of this representation still depends on numerous parameters, making the comprehension of the dynamics rather unreliable (e.g., \cite{principe1992prediction}). One reason for this fragility is that the embedded attractor in the reconstruction space not always an attractor of the map that is learnt in the reconstruction space, despite  unquestionably being an invariant set. When the embedded object is not \textbf{(definitely/always?)} an attractor (as explained in Chapter 3) it can cause predictions to fail.

Practically, the application of Takens embedding involves learning a map through some technique and consequently one wishes that these would have low functional complexity\cite{manjunath2021universal}, i.e., functions with fewer oscillatory graphs. Pure machine learning methodology that processes temporal information (like the echo state networks (Citations to echo state networks)) by mapping data onto a higher dimensional space for further processing. Although they perform  well on forecasting some dynamical systems, they fail completely on others (Citations here) as there is often no guarantee that the right function was learnt during training.

This thesis deals with the implementation and analysis of non-interpretable models that can guarantee \textbf{exact/accurate} reconstruction. The project work concerns the study and implementation of a method (Citation to paper with Adriaan) that incorporates learning a function by mapping the data on to a higher dimensional space using what is called a driven dynamical system (See Chapter 4). With a clear understanding of how the data is mapped onto the higher dimensional space, the method then permits the learning of a dynamical system topologically conjugate to that of the underlying system. Instead of linear regression as in the training of echo state networks, deep learning methods are employed to learn the correct function. With slight modifications to the implementation in the paper (with Adriaan), we show that one can construct accurate non-interpretable models with the ability to reconstruct attractors from more hard-to-forecast systems like the double pendulum. (The forecasting of the time-series from a double pendulum has not been reported before.) Moreover, we also demonstrate that long-term statistical consistency is preserved.

By solidifying the mathematical underpinnings of our theory, we hope to guarantee the ability to construct models with predictive power ranging from molecular biology to neuroscience. in the near future. (We can modify this sentence when the report is completed.)

The report is organised as follows: 
\newline In \ednote{M: Please label Sections as Chapters as this is apt for a thesis; for the paper that we will write soon, it would be in Sections. Please note the labelling in the tex file, and you may like it to be cross references like this}Chapter~\ref{Chapter_Ds}, we recall the definition of a discrete-time dynamical system, how a discrete-system arises from a flow of an ODE and then proceed to define the inverse limit space and topological conjugacy of autonomous systems. 
\newline In Section 3, we introduce the problem of forecasting dynamical systems, state the Takens delay embedding theorem and discuss various issues faced while forecasting. 
\newline In Section 4, a driven dynamical system is defined and discuss the properties of a specific class of driven dynamical systems that we make use of in this thesis/report/paper. 
\newline Finally in Section 5 we show the implementation of these forecasting methods, and conclusions are provided in Section 6.

\chapter{Discrete-time Dynamical Systems} \label{Chapter_Ds}

In this chapter, we provide a brief description of what a discrete-time dynamical system is and what it means for it to exhibit chaos. We refer to \cite{devaney2018introduction, de2013elements} for more details.

At its most elementary level, a dynamical system is just something that evolves deterministically through time. In the context of this thesis, deterministic refers to the fact that a system evolves according to specified rules rather than based on random events. 
Dynamical systems arise in a variety of situations. A continuous-time dynamical system describes the states for all values of the time. Specifically, if the motion of a pendulum in which the quantities such as the angular position and angular momentum are known at all times, then it is a continuous-time dynamical system. The equations of the dynamical system can take the form of one or more ordinary differential equations that determine the relevant quantities at any future time if we know the initial location and momentum. 
In ecology, discrete-time dynamical systems are widely used to model population growth. The model in this case is a function calculating the following generation's population given the population of the previous generation. If we know the starting population, we may once again calculate the population at any time in the future.

Formally, a function $T: U \to U$, where $U$ is some set is  a \emph{discrete-time dynamical system} and its iterates $\{u,Tu,T^2u,\ldots\}$, where $T^n$ denotes the $n$-fold composition of $T$ with itself, describe the evolution of an initial condition $u\in U$ (Note that we drop the brackets and denote $T(u)$ by $Tu$ frequently for simplicity in notation).

Continuous-time dynamical systems modelled using ordinary differential equations can give rise to discrete-time dynamical systems. To see this, consider a differential equation $\dot{x} = f(x)$, $f: \mathbb{R}^n \to \mathbb{R}^n$, $n\in\mathbb{N}$ given to have a unique solution passing through 
each point $x\in\mathbb{R}^{n}$, the flow of the equation \label{defn_flow} is defined to be a mapping $\varphi: \mathbb{R}^n \times \mathbb{R} \to \mathbb{R}$ that satisfies the following properties \ednote{M: Please change this or revert back to how it was defined earlier. We are describing $\varphi$ and in the first two points, it does not concern $\varphi$}
\vspace{-8mm}
\begin{enumerate}[noitemsep, label=\roman*.]
  \item $x(t)$ is a solution of the differential equation,
  \item $x(0)=x_0$, and
  \item $\varphi(x_0,t) = x(t)$.
\end{enumerate}
By fixing $t=K \in (0,\infty)$, we can define the \emph{time-$K$ map} as  $T(x):= \varphi(x_0,K)$, and it is easily verified that $T\circ T(x_0) = \varphi(x_0,2K)$. In general the $m^{\mbox{th}}$ iterate of $x_0$ under $T$ would be the value of the solution of the ODE evaluated at time $mK$ with the initial condition $x_0$. 
Thus ordinary differential equations give rise to a discrete-time dynamical system by sampling the value of $x$ at time intervals $K$ units apart.

A numerical discretization of a differential equation can also give rise to a discrete-time dynamical system. For instance, Euler's method approximates $\dot{x}(t)$ by $(x(t+h)-x(t))/h$; if $h$ is fixed throughout, the solution of a differential equation $\dot{x}=f(x)$ 
can be approximated at the time instant $t+(m+1)h$ by iterating the equation 
$$x(t+(m+1)h) = x(t+mh) + h f(x(t+mh)).$$ 
Adopting more succinct notation by replacing $x(t+mh)$ with $u_m$, we rewrite the above equation as
$$u_{m+1} = u_m + hf(u_m)$$
or, in even simpler terms, as the discrete-time dynamical system with map $T(u) = u + hf(u)$, where $T$, $u$ and $f$ are understood to be as above.

Of course, discrete-time dynamical systems need not always arise through a differential equation, and once again we may consider the field of ecology where discrete dynamical systems are often directly derived or assumed. 
There is a school of thought that advocates discrete-time dynamical systems to be more natural for modelling real-world observations than differential equations and we refer the interested reader to \cite{saber2010introduction}. \textbf{too philosphical, or actually good?} \ednote{M: It is good! I have moved this paragraph}

\subsection{Invariant Sets}

A core concept in the study of dynamical systems is that of invariance. Given a discrete-time dynamical system $T: U \to U$, a subset $A \subset U$ is said to be an \emph{invariant set} if $T(A) =A$. 
We also define the \emph{orbit of $T$} to be a sequence $\bar{u} = \{u_n\}_{n\in \mathbb{Z}}$ obeying the update equation, $u_{n+1}=Tu_n$, $n \in \mathbb{Z}$.

Two examples of invariant sets include a fixed point where $Tu=u$ for  $u\in U$, and a periodic orbit, i.e., a set of iterates $\{u,Tu, T^2u,\ldots,T^pu\}$, where $T^{p+1}u=u$ for some $p$.  The entire space $U$ could be also be invariant.
Consider for example the space $U=[0,1]$, where $Tu=4u(1-u)$ and then $U$ is invariant. \emph{(as every $u\in{U}$ can be written as $u = 4x(1-x)$ for some $x\in{U}$)}

We may learn a great deal about the iterates of a dynamical system by considering the types of invariant sets of a discrete-time dynamical system.  For example, if $U=[0,1]$ has map $Tu= u/2$, then the only invariant set is $\{0\}$, and all orbits approach this invariant set as time flows in the forward direction. 
Indeed, if a some non-zero invariant set (call it $B$) exists, then there is  some $r\in(0,1]\cap{B}$. But $r\notin{T(B)}$ since any orbit with initial value $r$ will be a decreasing sequence. Moreover, every orbit of T will be decreasing and therefore approach the value $0$ as $n\rightarrow\infty$.

One may ask if every orbit approaches an invariant set? In general the answer is no, since for the dynamical system $T: \mathbb{R} \to \mathbb{R}$ defined by $Tu=2u$, any orbit that does not intersect the invariant set $\{0\}$, will not approach any invariant set.

However,  when the space $U$ is compact, all orbits approach an invariant set. This is since the set of limit points of the orbit can be shown to be invariant \cite{de2013elements}. \ednote{M: I have edited here} So, when $U$ is compact,  the $\omega$-limit set $\omega(u;T)$ of a point $u$ defined to the collection of limit points of the sequence $\{x,Tu,T^2u,\ldots\}$ is nonempty, and $\omega(u;T)$ is invariant. For instance, for the map $Tu=u^2$ defined on $[0,1]$ all orbits lie in the invariant set $\{1\}$ or else would would approach the invariant set $\{0\}$. 
\ednote{B: not relevant other than being an additional example perhaps? M: Have added another simple example}.

We discuss various properties of invariant sets: Invariant sets can be attracting or repelling depending on how orbits in its vicinity behave. Recall the examples $Tu =u/2$ and $Tu=2u$ defined on $\mathbb{R}$ ($\{0\}$ ``attracts" orbits) or the example, $Tu=u^2$ defined on $[0,1]$ ($\{1\}$ ``repels" orbits). We are interested in attractive invariant sets since they capture the long-term dynamics as time increases, and in particular the invariant sets that are called attractors whose definition we recall next. \begin{Definition} \rm Let $T: U \to U$, where $U$  is a metric space with metric $d$. A compact subset $A \subset U$ is said to be an attractor if it satisfies the three conditions: 
  \vspace{-8mm}
  \begin{enumerate}
	\item $A$ is invariant. 
	\item $A$ is asymptotically stable, i.e., for every $\epsilon > 0$ and for all $u$ so that $d(u,A) < \epsilon$, we have $d(T^nu,A) \to 0$ as $n\to \infty$. 
	\item $A$ has Lyapunov stability, i.e., for every $\epsilon > 0$  there exists a $\delta(\epsilon) > 0$ so that $d(u,A) < \delta$ implies $d(T^nu,A) < \epsilon$ for all $n\ge 0$.  
\end{enumerate}
\end{Definition}

Indeed in our previous examples, it can be verified that the system  $Tu=u/2$ defined on  $\mathbb{R}$ and $Tu=u^2$ defined on $[0,1]$  the singleton set $\{0\}$ is an attractor.  For the system,  $Tu=1-|2u-1|$ on $[0,1]$, one may easily verify that the only attractor is the entire space $[0,1]$. This is since between any two points $u< v$ in $[0,1]$, and for any $a,b$ so that $u\le a < b \le v$, we can find an $n$ so that $T^n(a,b)=[0,1]$ (as would be explained later in this chapter).

\ednote{For let $B$ be an attractor of U. From above, we know that B is itself invariant, is asymptotically stable and has Lyapunov stability. Now suppose there is some attractor $C\subsetneq{U}=B$, so there is a $b\in{U}$ which is not in $C$. ........  M: Have changed the example}

A dynamical system can have several attractors, and may also be contained in another attractor. For the example, $Tu =u^2$ on $[0,1]$, both $\{0\}$ and $[0,1]$ are attractors. It is a known result that for dynamical systems on a compact space there always exists an attractor.

The dynamics restricted to an invariant set can be complicated. For instance an invariant set could be just a single point or it could have an infinite set.    If the invariant set is infinite then there is a possibility of complicated dynamics, and a particular well-studied phenomenon of such complexity gives rise to what is called  chaotic behaviour, and has been studied in detail over the last fifty years.

% If the dynamics on the attractor are somewhat complicated in the sense that the attractor cannot be decomposed further, and if the attractor is infinite then there is a possibility of complicated dynamics, and a particular well-studied phenomenon of such complexity gives rise to what is called a chaotic behaviour.

%If the dynamics on an attractor cannot be decomposed further and the attractor is itself also infinite, then there is a possibility of complicated dynamics, and a particular well-studied phenomenon of such complexity gives rise to what is called chaotic behaviour.

\subsection{Chaos}

\ednote{ should one add paragraphs like these? M: this is a thesis, and not a paper; the instructions sent to the external examiner regarding the thesis evaluation says that the student suggest strongly that it is good to write}
\newline In the 1960s, a number of mathematicians and mathematically interested scientists independently discovered chaos in the mathematical sense. The meteorologist Edward Lorenz may have been the first to explain this phenomenon in his 1963 paper \cite{lorenz1963deterministic}. The notions of invariance, attractivity and chaos may also bedescribed for continuous systems, and Lorenz's system comprised of a system of differential equations. The narrative of Lorenz's discovery of chaos, as well as the history of other forerunners in this subject, are fascinating. For those who are interested, we highly recommend James Gleick's book Chaos: The Making of a New Science \cite{gleick2008chaos}. It clearly illustrates these experiences, explains why chaos was such a startling and crucial mathematical and scientific discovery, and describes the underlying mathematical notions for non-specialists.

There are several intricately different definitions of chaos available in the literature, with each of them indicating some aspect of complexity \ednote{B:  what are we saying here? M: Devaney's definition is not the only one, there are about half a dozen definitions as well} In practice, it is not possible to verify which definition or notion does the underlying dynamical system satisfies when only data is observed from it. We do, however, specifically recall the definition of chaos in the sense of Devaney \cite{devaney2018introduction,de2013elements} so as to understand some nuances behind the complexity. Devaney's definition of chaos has three requirements, with the first being the notion  of sensitive dependence on initial conditions.

\begin{Definition} \rm 
A dynamical system $T: U \to U$ is said to have sensitive dependence on initial conditions (SDIC) if there exists a $\delta > 0$ such that for every $u \in U$ and in every neighborhood of $u \in U$ there exists a $v\in{U}$ and an integer $N:=N{(u,v)}\in\mathbb{Z}$ such that $d(T^Nu,T^Nv)>\delta$. 	
\end{Definition}

It is important to acquire a sense of this definition as it can easily be misunderstood. It is common misconception to interpret SDIC as two close points ($u$ and $v$) that eventually become separated by a distance $\delta$ under iteration by $T$. But this is not true. To fully comprehend all the subtleties of the concept, one needs to discuss it more thoroughly by considering each sentence with care and attention, whereafter one may examine how they fit together to convey the concept of sensitive dependence. 
\ednote{ $\leftarrow$ is this talking too simply perhaps? M: No! it is not so}.

To this end, we make 3 remarks:
\vspace{-5mm}
\begin{enumerate}
  \item First, the $\delta>0$ in the definition of SDIC is independent of $u$. 
  \item Second, in every neighborhood of $u$, we may not necessarily find all points $v$ in the neighborhood distinct from $u$ that would separate from the forward iterates of $u$. 
  \item  Finally, $N$ depends upon $u$ and $v$ chosen, and their iterates may not separate for ever (i.e., for all $n>N$) and we allow their iterates to get arbitrarily close in the future. 
\end{enumerate}

\ednote{M: It is probably best to see a few numerical simulations to understand SDIC. (Please use the logistic map and perhaps the trajectories of the double pendulum data). }

The second part in Devaney's definition of chaos concerns topological transitivity.

\begin{Definition}
	A dynamical system $F: U \to U$ is topologically transitive if for any pair of nonempty open sets $E_1$ and $E_2$ there exists a $n\in\mathbb{N}$ such that $T^n(E_1) \cap E_2 \not= \emptyset$. 
\end{Definition}

Topologically transitivity implies that the iterates of an open set of initial conditions gets mixed up with other open sets. On a compact metric space, it can be shown that topological transitivity also implies the existence of a point whose forward iterates are dense \cite{de2013elements}; or in other words, the orbit going through this point will be dense in the compact metric space. 
In fact, it is topologically more likely that the choice of an arbitrary point will be one  whose iterates are almost dense. 
% I inserted a definition here: it makes more sense to me. 
\begin{Definition} \rm
A subset of a topological space $U$ is said to be a meager set if it cab be written as a countable union of sets of with empty interior. The complement of the a meager set is set to be a residual set.
\end{Definition}

To be more precise (see \cite{de2013elements}), for a discrete-time dynamical system on a compact space, the set of points with dense iterates are residual, and they are typical or likely to be observed in practice. In this thesis, we assume that when data is observed from a topologically transitive system, we assume that the data arises from a dense orbit.

We may now define the notion of Chaos as formulated by Devaney\cite{devaney2018introduction}.
\begin{Definition} \rm
	A dynamical system $T: U \to U$ is said to exhibit chaos in the sense of Devaney if it satisfies the three properties:
	\vspace{-5mm}
  \begin{enumerate}
		\item $T$ has SDIC.
		\item $T$ is topologically transitive.
		\item The set of periodic points of $T$ are dense in $U$. 
	\end{enumerate}
\end{Definition}

\begin{figure}[ht]
  \centering
  \begin{minipage}{.5\textwidth}
    \centering
\includegraphics[scale=0.18]{chaos1.png}
    \captionof{figure}{Starting to diverge}
    \label{fig:chaos1}
  \end{minipage}%
  \begin{minipage}{.5\textwidth}
    \centering
    \includegraphics[scale=0.18]{chaos2.png}
    \captionof{figure}{After more time-steps}
    \label{fig:chaos2}
  \end{minipage}
  \end{figure}

\ednote{M: Please put this picture in perspective and refer after the definition of SDIC. You can also say that you pick two positions that are slightly different but with the same angular velocity. Also in the text, say that you describe the DP in a later chapter.}
\ednote{M: I realize that need figures in the eps format to have the commutativity diagrams compile as well. }

\begin{Example} \rm
  The standard tent map $Tu=1-|2u-1|$ defined on $[0,1]$ is a well known example of a dynamical system satisfying these three properties. 
\end{Example}

  \begin{figure}[ht]
\includegraphics[scale=0.6]{T2.png}
    \centering
    \label{fig:T2tentmap}
  \end{figure}

Let us now consider why this map in the above example exhibits chaos in the sense of Devaneys. The graph of the map $T$ is piecewise linear with two straight lines, one connecting the points $(0,0)$ and $(1/2,1)$ and the other connecting $(1/2,1)$ with $(1,0)$. They form a so-called tent with base centered at $u=1/2$. The graph of the map $T^2$ comprises two symmetric tents with their base centered at $1/4$ and $3/4$.
  % We now reason as why this map exhibits chaos in the sense of Devaney.

  In general, the graph of  $T^k$ contains $2^{k-1}$ tents.
Given any point $u$ and an open neighborhood $W \subset[0,1]$ of $u$, we can find $k\in\mathbb{Z}$ large enough so that $T^k(W) = [0,1]$ because we can accommodate a tent whose base is contained in $W$. \ednote{B: This last phrase doesn't make sense to me. M: The number of tents grow with $k$ and hence this is obvious; I explained in one of our meetings. Does it make sense?} This implies the existence of two points $w_1$ and $w_2$ in $W$ so that $d(T^kw_1,T^kw_2)= 1$, and by the triangle inequality  at least one of the inequalities $d(T^ku,T^kw_1)> \delta$ or $d(T^ku,T^kw_2)> \delta$ hold when $\delta\in (0,1/2)$.  So $T$ has sensitive dependence on initial conditions.  \ednote{M: It is advised not to introduced newlines in a thesis. For the presentation, yes}

  Next, let $W_1$ and $W_2$ be two nonempty open sets. Given any open set $W_1$, we can find a $k\in\mathbb{N}$ large enough so that $T^k(W_1) = [0,1]$ because we can accommodate a tent with base contained in $W$. So $T^k(W_1) \cap W_2 \not=\emptyset$, and thus $T$ has topological transitivity.

  Finally, for every open interval $(a,b)$, the graph of $T^k$ intersects the graph of the identity map on $[0,1]$. This follows from the fact established above that there exists a $k\in\mathbb{N}$ such that $T^k(a,b) = [0,1]$.If the intersection point has coordinates of the form $(p,p)$, the point $p$ is then a fixed point of $T^k$ and therefore also a periodic point of $T$. Since we have found a periodic point in the interval $W$,  the set of periodic points are dense in $T$.
  \ednote{Expand further on reason for density. M: fixed point of $T^k$ is a periodic point, and we have found in an arbitrary interval.}

\section{Conjugacy}

We now turn our attention to the subject of conjugacy that describes when two dynamical systems are equivalent dynamically.

To show topological similarity or sameness between two metric or topological spaces, one needs to establish a homeomorphism between the two spaces. 
However, in the study of dynamical systems defined on two spaces, establishing a homeomorphism does not indicate that the systems are (physically) related in any way.  For instance, the maps $Tu=u^2$ and $Tu=1-|2u-1|$ defined on $[0,1]$ have totally different behaviour. \ednote{"related" - in what sense am I saying it here? Physically? M: Dynamically}  To find systems that are dynamically similar, one rather needs to establish a kind of dynamical equivalence which we illustrate in a simple commutativity diagram below.
% Giving errors
\begin{equation}  \label{eqn_conjugacy}
%    \[ 
    \psset{arrows=->, arrowinset=0.25, linewidth=0.6pt, nodesep=3pt, labelsep=2pt, rowsep=0.7cm, colsep = 1.1cm, shortput =tablr}
 \everypsbox{\scriptstyle}
 \begin{psmatrix}
U & U\\%
V & V.
 %%%
%  \ncline{1,1}{1,2}^{T} 
%  \ncline{1,1}{2,1} <{\phi}
%   \ncline{2,1}{2,2}^{S}
%  \ncline{1,2}{2,2} > {\phi}
 \end{psmatrix}
% \]
\end{equation}

%Composite functions
To understand the diagram we note that if we travel ``right and then down," the diagram instructs us to use $T$ (top arrow right) first, followed by $\phi$. (right arrow downwards). Consequently the pathway amounts to finding $\phi(T(u))$. If we proceed "down and then right", the diagram instructs us to apply $\phi$ (left arrow downwards) first, and then apply $S$ (bottom arrow right). This then amounts to finding  $S(\phi(u))$. When the relationships in this diagram hold, we say $\phi(T(u))= S(\phi(u))$, and we formally denote it as $\phi \circ T=S\circ \phi$.

\begin{Definition} \rm 
  Consider the dynamical systems $T:U\to{U}$, $S:V\to{V}$ and suppose the relationship $\phi \circ T=S\circ \phi$ holds. If $\phi$ is a homeomorphism, then S is said to be conjugate to T and $\phi$ is called a conugacy. If we relax the criterion on $\phi$ and merely require $\phi$ to be continuous and surjective, $\phi$ is then a semi-conjugacy between $T$ and $S$; S is said to be semi-conjugate to T. 
\end{Definition}

\ednote{M: Saying $\phi$ is then a semi-conjugacy between $T$ and $S$ could be ambiguous as it does not specify which system is on the top in the commutativity. So please edit accodingly}

When $S$ is conjugate to $T$, the dynamics of the two systems are in some way ''dynamically equivalent'' and specifically they are in one-to-one correspondence with one another. However when $S$ is semi-conjugate to $T$ with $\phi$ a many-to-one mapping, the dynamics on $V$ provides a coarse-grained description of the dynamics on $U$ \cite{de2013elements}. When $S$ is semi-conjugate to $T$, it is also common to call $S$ a factor of $T$, or conversely that $T$ is an extension of $S$. In essence, an extension (e.g., \cite{de2013elements}) is a larger system capturing all of the important dynamics of its factor.

It is a very hard or nearly an impossible task to establish the existence of a conjugacy or a semi-conjugacy $\phi$ between two systems \cite{devaney2018introduction}. However, one can verify that a function $\phi$ satisfies the commutativity diagram \ref{eqn_conjugacy}.  For example $\phi(x)=\sin(\pi x/2)^2$ is a conjugacy between two systems $Tu=1-|2u-1|$ and $Sv=4v(1-v)$  defined on $U=[0,1]$.

In establishing that two systems are conjugate (semi-conjugate) to one another, we have also shown that one may choose to work with one systems as opposed to the second and still be guaranteed to obtain information on the latter. This is incredibly useful while forecasting the future evolution of dynamical systems.

\chapter{The Learning Problem}

\emph{Introduction...}

Consider a relatively simple learning problem:

Given $(u_0, u_1, \ldots, u_m)$ for $m\in\mathbb{Z}$, a finite segment of an orbit of the map $T$ where $T$ is defined by the update equation $u_{n+1} = Tu_n$, forecast the values $u_{m+1}, u_{m+1}$ where the map $T$ is unknown, given that $u_m$.

A practical example of this would be - given the sequential coordinates of an object moving in space, predict the future positions of that object. However we are very rarely, if at all, presented with a problem where the entire state information is available to us in the form $u_n$ at some time-step $n\in\mathbb{N}$. Consider then a more involved learning problem:

Suppose we only have the observations $\theta(u_0), \theta(u_1), \ldots, \theta(u_m)$ of the true system states $u$ in an unknown dynamical system $(U,T)$ and we wish to predict the values $\theta(u_{m+1}), \theta(u_{m+2})$.

Here we consider a continuous-time dynamical system $(U,T)$ with true state $u$ in the input space $U$ and then a measured signal $\theta$ (where $\theta(u_n)=\theta(u(t_n))$ for $n\in\mathbb{N}$. The sequence $\{x_n\}$ represents a scalar time-series and intuitively $\theta$ may be thought to represent a probe inserted into a larger system and itself measuring/extracting only a small part of the greater system state at time $t=n$. Consider for example a thermometer erected to measure the ambient temperature in a local village. This measurement function, the thermometer is capturing a single aspect, the temperature, of a grander dynamical system entailing the present weather of the surrounding area, but even more than that, it is measuring an incredibly small part of the global weather system.

From here we construct a multidimensional observable using the method of delay-coordinate stacking. 
\textbf{move to later} Given the sequence $\{\theta_1, \theta_2, \ldots\}$, we define the variable \newline $y_n:=[\theta_n, \theta_{n-\tau}, \ldots, \theta_{n-(2d-2)\tau}, \theta_{n-(2d-1)\tau}]^{T}\in\mathbb{R}_{d}\times{1}$ where $\tau$ represents the lag and $d$ the dimension of the attractor. (We shall return to these terms shortly and their discussion may be put on hold for the time being).

However, the true state $u$ of a system is seldom, if ever fully known. In almost all cases we can merely insert probes into a system to obtain partial information by measurements taken. Moreover, the process of taking a measurement itself introduces 2 additional difficulties:
\vspace{-8mm}
\begin{enumerate}[noitemsep, label=\roman*.]
  \item A series of measurements over a specific time-interval is inherently a discretisation
  procedure of the underlying continuous-time dynamical system. (Hence why we restricted our attention to discrete-time systems in the preceding section)
  \item A probe will never be fully accurate and so the act of measurement introduces a certain
measure of numerical noise/inaccuracy.
\end{enumerate}

More formally we may say that in the process of measuring a system, we obtain a 1-dimensional time-series $\{\theta(u(t))\}$, where $\theta:\mathbb{R}^n\to\mathbb{R}$ is a smooth observation or measurement function. This series represents a possibly noise-augmented dataset containing information of certain parts of the original system.

We may then ask ourselves the question: Is it in any way possible to retain information about the system state $x(t)$ in this temporal data-series $\theta(x(t))$? The answer is easily yes if T is known. However, neither this, nor even the exact function $\theta$ is available.

\subsection{Taken Embedding Theorem}

We define first the concepts of homeomorphism and embedding:
\begin{Definition}
  A homeomorphism is a function $f:Z\rightarrow Y$ between two topological spaces $Z$ and $Y$ that is continuous, bijective and has a continuous inverse. 
\end{Definition}

\begin{Definition}
  Consider a homeomorphism $f:Z\rightarrow Y$ for $Y\in X$. $Z$ is said to be embedded in $X$ by $f$.
\end{Definition}

Takens Theorem states a result that establishes a relationship between the observed and underlying dynamical systems by showing that the concetantion of a sufficiently large number of previous observations into a vector will, under certain conditions, generate a map between the vectors from the respective systems. \textbf{Check wording} We formultate the theorem from \cite{takens1981detecting}.

\begin{Theorem} 
	[\bf Takens Embedding Theorem (adopted from \cite{takens1981detecting}] \label{Thm_Takens}
            Let $W$ be a compact manifold of dimension $m$, and $d\ge m$ so that $2d\in\mathbb{Z}$. It is a 
            generic property for the pair $(T, \theta)$,  where $T:W \to W$ is
            a smooth diffeomorphism, and $\theta:W \to \mathbb{R}$ a smooth function, that the map $\Phi_{2d,\theta}:W \to \mathbb{R}^{2d+1}$ defined on $W$ by 
            $\Phi_{2d,\theta}(w) := (\theta(T^{-2d}w)\ldots,\theta(T^{-1}w),\theta(w))$
            is a diffeomorphic embedding; by `smooth' we mean at least $C^2$. Consequently, there exists a map $\Ftheta: \Phi_{2d,\theta}(W) \to \Phi_{2d,\theta}(W)$ defined by 
            \begin{equation}  \label{eqn_takens}
              \Ftheta: (\theta(T^{-2d}w),\ldots,\theta(T^{-1}w),\theta(w)) \mapsto  (\theta(T^{-2d+1}w),\ldots,\theta(T^{-1}w),\theta(Tw))
            \end{equation}
           so that $(W,T)$ is topologically conjugate to 
            $(\Phi_{2d,\theta}(W), \Ftheta)$. 
\end{Theorem}

(In our scenario, the input space $U$ is considered the attractor.)

Some explanation is beneficial. Takens establishes a delay-coordinate map $\Ftheta$ defined in \ref{eqn_takens} for $T$ the flow (\ref{defn_flow}) defined by the update equation  $Tu_n=u_{n+1}$ and $theta$ our measurement function. When confining a dynamical system to the manifold U (in which the underlying system is contained), Takens showed that if certain smoothness conditions are satisfied on $T$ and $\theta$, then the delay-coordinate map $\Ftheta$ embeds $U$ in the reconstruction space $R^{2d+1}$ \textbf{we haven't used the term reconstruction space yet, so this is falling out of the blue}, for almost every choice of measurement function $\theta$, the observable. By almost every choice here, we mean that we exclude some observations which do not give any information. To see a more precise discussion, see \textbf{xxx} for we do not concern ourselves with a further elaboration on this thesis/report.

Alternatively we may also make the statement: Takens' Embedding Theorem guarantees that almost all dynamical systems can be reconstructed from just one noiseless observation sequence \textbf{this isn't obvious from the theorem's formulation, is it?} i.e. for a great number of possible observation functions $\theta$, $\Ftheta$ preserves the topology of U. \textbf{'preserving the topology' is not something we've touched on before. Perhaps I should add to previous discussion so logic flows seamlessly here}.

\begin{figure}[ht]
  \includegraphics[scale=0.3]{takensmap.png}
  \centering
  \label{fig:takensmap}
\end{figure}

Once again we pause and consider the graph to consolidate our understanding of the conjugacy (\textbf{are we being too simplistic?}). If we are in state $u$ and then evolve towards state $Tu$ before embedding into $\mathbb{R}^{2m-1}$, then this is equivalent to first embedding into $\mathbb{R}^{2m-1}$ and then evolving through the map $\Ftheta$, i.e. $\Phi_{2d,\theta}\circ{F}_{\theta} = T\circ\Phi_{2d,\theta}$. The map $F_theta$ is a homeomorphism as well.

Thus information about U can be retained in the time series' (or observation's) output. By preserving the topology on the manifold $U$ in the reconstruction space $X$ (\textbf{not discussed before}), we are also guaranteed the preservation of topological invariants of the manifold, of which dimensionality is one such invariant. (We note here that dimensionality will again be considered later in this project/report).

\subsection{Obstacles to Takens}
Despite the fact that the Takens' Embedding Theorem is a powerful result and provides compelling reason to believe that it is possible to accurately ()\emph{right word?}) reconstruct the underlying space, it does present some serious limitations:
\vspace{-5mm}
\begin{enumerate}
\item Even if we can indeed find \Ftheta{}  our \emph{approximation} of \Ftheta{}  is a map from a larger set $\mathbb{R}^{2m+1}$ containing the embedded attractor. There are, however, no theoretical guarantees that \Ftheta{}  will keep this attractor $U$  the same.(\textbf{We haven't properly said that U is an attractor, or have we.})
\item Takens only holds for noiseless observations. Suppose the case where \Ftheta{} was learnt, but due to measurement error or other noise we obtained the set $V$ and not the original attractor. Due to the chaotic nature of the underlying system (i.e. the fact that it has SIDC), this new system $V$ will then move in a different direction. This problem will later be eliminated by some properties of a new function $g$, namely the Unique Solution Property. 
\item It is an unfortunate case that the functional complexity (i.e., the relative number of oscillatory graphs \cite{manjunath2021universal}) as well as stability of the embedding \Ftheta{} (\textbf{cite?})depends on the choice of observable $\theta$. This means that we may even learn a map that complicates our learning task and in so doing adds obstacles to the road to learning a conjugate map with predictive power. (\textbf{Am I talking nonsense here?})
\end{enumerate}

Another limitation owes itself to the requirement that T be a surjective map. This is not always the case in practical examples and we may consider the case of a damped double pendulum.

\textbf{Continue writing here}

\begin{figure}[ht]
  \includegraphics[scale=0.3]{learningfailure.png}
  \centering
  \label{fig:supp_learning_failure}
  \caption{Illustration of a system having failed to learn the map}
\end{figure}

\textbf{Explanation continued here}
Traalala lorem ipsum et dolore faria maria didi klapogaani koro di tolore mi strem srem krem lem lemmy bemmy bem te dem dem

\begin{figure}[h]
  \includegraphics[scale=0.4]{temporary_damp_fig.png}
  \centering
  \label{fig:damped_pendulum}
\caption{A damped pendulum}
\end{figure}

\textbf{Continue explaining here}

\vspace{-1cm}
\bibliographystyle{pnas-new}
\bibliography{pnas}

\end{document}

### 2021-12-31 Gandhi_Project_Report/draft1.tex
Section 4 - just before memloss diagram. Need to split Th.4

@@
 
-% \subsection{Taken Embedding Theorem}
+\subsection{Taken Embedding Theorem} \label{subs_TakensEmbedding}
 
@@
 
-\begin{Example} \label{Ex_exp}
+\begin{Example}\label{Ex_exp}
  Consider the driven system $g(u_n,x_n)=x^n$ for $X=[0,1]$, $U=\mathbb{R}$.
@@
 \begin{Definition}
-  [\bf Reachable Set] \label{Dfn_ReachableSet}
+  [\bf Reachable Set]\label{Dfn_ReachableSet}
 The reachable set of a driven system $g$ is exactly the union of all the elements of all the solutions, i.e., 
@@
 \begin{Definition}
-  [\bf Topological Contraction] \label{Dfn_TopContr}
+  [\bf Topological Contraction]\label{Dfn_TopContr}
   A function $g:U\times{X}\to{X}$ is a topological contraction if for all $n\in\mathbb{Z}$ and all $\overline{u}\subseteq{U}$, $X_n(\overline{u})$ is a singleton subset of X.  
@@
 \begin{proof}
-  [\bf Proof of Example \ref{Ex_exp}] 
+  [\bf Proof of Example\ref{Ex_exp}] 
   To see this, we show that for every input $\overline{u}=\{u_n\}_{n\in\mathbb{Z}}$, $g(u_n,\cdot)$ is a contraction map on $X$ for every $k\in\mathbb{Z}$. 
@@
 
-%%%%%%%%%%%%%%%%%%%%%ISSUE HERE - fix before sending
-\begin{Theorem} \label{Thm_CompactExistence}
+\begin{Theorem}\label{Thm_CompactExistence}
   If $X$ is compact then for each input $\overline{u}$, there exists at least one solution to the driven system $g(\ldotp, x)$
@@
 
-\begin{Definition}
-  [\bf Unique Solution Property)] \label{Dfn_usp}
+\subsection {Unique Solution Property}
+
+\begin{Definition}
+  [\bf Unique Solution Property] \label{Dfn_usp}
   A driven system $g$ is said to have the Unique Solution Property (USP) if for each input $\overline{u}$ there exists exactly one solution. 
@@
   \begin{Definition} 
-    [\bf Universal Semi-Conjugacy)] \label{Def_UnivSemiConj}
+    [\bf Universal Semi-Conjugacy]\label{Def_UnivSemiConj}
     Given a driven system $g$, we  call a continuous and surjective map $h : \overleftarrow{U} \to X_U$ a universal semi-conjugacy if  diagram \ref{Scomm_h} commutes for all $v \in U$.
@@
 
-Re-sketching the graph \ref{Scomm_h} above by fixing the input $v$ in $g$ and replacing $X_U$ by its left-infinite sequence space $\overleftarrow{U}$, we obtain the graph below.
+Re-sketching the graph\ref{Scomm_h} above by fixing the input $v$ in $g$ and replacing $X_U$ by its left-infinite sequence space $\overleftarrow{U}$, we obtain the graph below.
 In this case, the function $H:\overleftarrow{U}\to\overleftarrow{X}_U$, a map that is both continuous and surjective, is called a \emph{causal mapping}.
+
+\begin{Definition}
+  [\bf Causal Mapping]\label{Def_CausMap}
+  A continuous, surjective map $H:\overleftarrow{U}\to\overleftarrow{X}_U$ such that \[H\circ\tilde{g}_v=\sigma\circ{H}\] where $\tilde{g}_v$ maps $(\ldots, u_{-2}, u_{-1})$ to $(\ldots, u_{-2}, u_{-1}, g(v, u_{-1}))$.
+\end{Definition}
 
@@
 
-  Note that $\tilde{g}_v$ maps $(\ldots, u_{-2}, u_{-1})$ to $(\ldots, u_{-2}, u_{-1}, g(v, u_{-1}))$.
-  The driven system $g$ can induce an embedding of $\overleftarrow{U}$ in $\overleftarrow{X}$ as follows:
-
-\begin{Definition}
-  [\bf Causal Embedding]\label{Def_CausEmbed}
-  If $H:\overleftarrow{U}\to\overleftarrow{X}_U$ is also injective (in addition to being surjective and hence bijective), it becomes the embedding of $\overleftarrow{U}$ in $\overleftarrow{X}$ induced by the driven system $g$ and we refer to $H$ as a \emph{causal embedding}.
-\end{Definition}
-
 \begin{Theorem}
@@
   \emph{See proof from p.15 of \cite{manjunath2021universal}}
+  \textbf{Expand.}
 \end{proof}
 
-Note that even when $h$ does exist, we are not guaranteed its injectivity.
-Considering again example \ref{ex_halfux}, we conclude that even if $h$ were to exist, it could not be injective as $X_U=\{0\}$.
+Note that even when $h$ does exist, we are not guaranteed its injectivity. Considering again example \ref{ex_halfux}, we see that even if $h$ were to exist, it could not be injective as $X_U=\{0\}$.
 
@@
 
-  Before formalising this, we make mention of the concept of the \emph{inverse-limit system} of a system $(U,T)$.
-In broad sweeps, the inverse-limit system of a dynamical system is a self-map on a subset of an infinite-dimensional space \textbf{(expand what exactly this is?)} where each point in the inverse-limit space corresponds to a backward orbit of the map $T$.
-In literature, the inverse-limit space is denoted by $\hat{U}_T\subseteq\overleftarrow{U}$.
-\textbf{cite from supp.koopman}.
+  Before formalising this, we make mention of the concept of the inverse-limit system of a dynamical system $(U,T)$. 
+  In broad sweeps, the inverse-limit system of a dynamical system is a self-map on a subset of an infinite-dimensional space \textbf{(Expand)} where each point in the inverse-limit space corresponds to a backward orbit of the map $T$.
+In literature, the inverse-limit space is denoted by $\hat{U}_T\subseteq\overleftarrow{U}$ \cite{manjunath2021universal}.
   The map $T$, then, induces a self-map $\hat{T}:\widehat{T}: (\ldots,u_{-2},u_{-1}) \mapsto  (\ldots,u_{-2},u_{-1},T(u_{-1}))$ on $\widehat{U}_T$
@@
 \begin{Definition}
+  [\bf Causal Embedding]\label{Dfn_CausalEmbed}
   A driven system $g$ is said to causally embed the dynamical system $(U,T)$ if 
@@
 \begin{enumerate}[noitemsep, label=\roman*.]
-  \item The diagram below commutes (i.e. $g$ is a universal semi-conjugacy),
+  \item The diagram in \ref{SCausal_H} commutes (i.e. $g$ is a universal semi-conjugacy),
   \item $H_2(\overline{u}):=(h(r\overline{u}, h(\overline{u})))$ embeds the inverse-limit space $(\hat{U}, \hat{T})$ of $(U,T)$ in $X\times{X}$.
@@
 \end{Definition}
+
+The driven system $g$ can induce an embedding of $\overleftarrow{U}$ in $\overleftarrow{X}$ as follows: If $H:\overleftarrow{U}\to\overleftarrow{X}_U$ is also injective (in addition to being surjective, and hence bijective), it becomes the embedding of $\overleftarrow{U}$ in $\overleftarrow{X}$ induced by the driven system $g$ and we refer to $H$ as a \emph{causal embedding}.
 
@@
 \end{Theorem}
-
-When g has the USP, the diagram below illustrates the operation of the the mappings h and H the mapping $h:\overleftarrow{U}\to{X_U}$ is an observable as discussed (above).
+\begin{proof}
+  \emph{See proof from \cite{manjunath2021universal}}  \textbf{Expand.}
+\end{proof}
+
+When g has the USP, the diagram below illustrates the operation of the the mappings $h$ and $H$.
+The mapping $h:\overleftarrow{U}\to{X_U}$ is an observable as discussed in an above section (\ref{TakensEmbedding}).
 
@@
 
+The next result is of relative import but to fully comprehend its influence, some discussion still remains.
+Before stating it, we define a concept which will directly be shown to be equivalent to the USP.
+
+\begin{Definition}
+  [\bf Uniform Attraction Property]\label{Dfn_UAP}
+  A driven system $g$ has the Uniform Attraction Property (UAP) if, regardless of starting position, all trajectories converge to a single trajectory as time flows forward. 
+  This trajectory is also the unique solution sequence $x$ to the input sequence $u$ as mentioned above \ref{here_UAP_firstestablished}.
+\end{Definition}
+
 \begin{Theorem}
@@
     \item $g$ has the USP 
-    \item $g$ is topological contraction 
+    \item $g$ is a topological contraction 
     \item $g$ has the Unique Attraction Property (UAP). 
@@
 
-We have not as of yet treated the UAP, and establish the necessary result here (without proof - left to literature. Or later) so as not to break up the completeness of theorem
-
-It is important to note here that one must be careful to avoid a choice of g which adds additional complexity to the obtained solution.  
-When the causal embedding exists for the driven system $g$, one may map an arbitrary input $u$ onto the solution space $X$ without additional distortion or information-loss.  
-Establishing an embedding removes this problem by guaranteeing that, since the systems are conjugate (semi-conjugate), $g$ does not add any (only “some”) complexity to the system.  
-It is, however, a balancing act as one also does not want to choose a function g that quenches the temporal structure in $u$ by contracting to such a degree that one loses the ability to recover information from the original system completely.
-To obtain this we want the reachable set of a driven system to be large enough to relate to the input.
-
-In the example above \ref{ex_halfux}, we cannot relate the input's temporal variation and the reachable set because $X_U$ consists of a single element.
-
-The reachable set $Y_T$ of a driven system must therefore be such that we can embed the inverse-limit space of $U_T$ can be embedded in some finite self-product of the reachable set of $T$.
+We have not as of yet treated the UAP properly, and establish the necessary result here (without proof) so as not to break up the completeness of theorem.
+
+\textbf{The proof of the result for the UAP is left to literature. Why? Cite.}
+
+\subsection {Choosing the driven system $g$}
+
+One must be careful to avoid a choice of $g$ which would  add complexity to the obtained solution.
+
+When a causal embedding $H$ exists for the driven system $g$, one can map an arbitrary input ${u}$ onto the solution space $X$ without additional distortion or information-loss.  \textbf{(Cite.)} 
+When an embedding is established, the question of possible additional complexity in the solution is removed by guaranteeing that, since the systems are conjugate (semi-conjugate, (refer)), $g$ does not add any (some) complexity to the system.
+It is, however, a balancing act as it also undesirable to choose a function $g$ that quenches the temporal structure in $u$ by contracting to such a degree that the ability to recover information from the original system is lost completely.
+To obtain a suitably complex function $g$, it is desired that the the reachable set of a driven system  be large enough to relate to the input.  \textbf{(Expand here. It stops a bit abruptly.)}
+
+In the example above (\ref{ex_halfux}), the input's temporal variation cannot be related to the reachable set as $X_U$ consists of a single element and so little, if not no, information is encoded.
+Rather, the reachable set of a driven system must therefore be such that the inverse-limit space of $U_T$ can be embedded in some finite self-product of the reachable set of $T$ \textbf{(refer or cite)}.
 To this end, consider the notion of State-Input (SI) Invertibility.
@@
 \begin{Definition}
+  [\bf SI-Invertibility]\label{Dfn_SIinv}
   A map $g$ is said to be SI-Invertible if $g$ is invertible for all $x\in{X}$.
-Alternatively, if, given $x_n$ and $x_{n-1}$, $u_n$ can be uniquely determined from $x_n=g(u_n,x_{n-1})$, then $g$ is said to be SI-invertible.
-\end{Definition}
-
-We make special note of a specific driven system.
-The function $$g(u,x)=(1-a)x+atanh(Au+\alpha{a}Bx)$$  is both SI-invertible and possesses the USP.
-This function $g$ is used in our implementation and is more completely discussed in chapter 5 \ref{sect5}
-The proof of these two facts is rather involved and the proof, instead of being replicated here, may be found in (cite relevant article)
-
-\textbf{Expand: UAP + Embedding -> perturbation isn’t an issue + we can drop left-infinite sequence.  }
-
-Despite the ease that one may work with left-infinite sequences in the realm of theory, it is impossible to obtain such a sequence in any real-life application.
-One does not in practice, however, need the entire left-infinite history of the input due to the Uniform Attraction Property (UAP).
-The next definition is stated in an informal manner as the more rigourous definition makes use of processes, a concept which would take some amount of paper to establish and explain - something which would detract from the overarching thrust of this paper/project/thesis.
-\textbf{is it VERY obsious to the reader at this point what the general 'thrust' of the paper IS?} \textbf{Cite articles}
-
-\begin{Definition}
-  A driven system $g$ has the Uniform Attraction Property (UAP) if, regardless of starting position, all trajectories converge to a single trajectory as time flows forward. 
-  This trajectory is also the unique solution sequence $x$ to the input sequence $u$ as mentioned above \ref{here_UAP_firstestablished}.
-\end{Definition}
-
-Incredibly, one may then initialise a driven system with a completely arbitrary initial value $y_m\in{X}$ where $m\in\mathbb{Z}$ and the UAP then guarantees that the sequence $\{y_{m+1}, y_{m+2},\ldots\}$ satisfying $y_n=g(u_n, y_{n-1})$  with $k\geq{m}$  will uniformly approach the elements $\{x_n\}$ of the actual solution.
+Alternatively it may be said that if, given $x_n$ and $x_{n-1}$, $u_n$ can be uniquely determined from $x_n=g(u_n,x_{n-1})$, then $g$ is said to be SI-invertible.
+\end{Definition}
+
+\textbf{Explain more here.}
+
+We make special note of a specific driven system. The function
+
+\begin{equation}  \label{eqn_driveng}
+%    \[ 
+g(u,x)=(1-a)x+a{\cdot}tanh(Au+\alpha{a}Bx)
+% \]
+\end{equation}
+
+is both SI-invertible and possesses the USP. The proof of these two facts is rather involved and  hence the proof, instead of being replicated here, may be found in (cite relevant article) 
+This specific driving function $g$ is used in our implementation and is more completely discussed in chapter \ref{sect5}
+
+\textbf{Expand: UAP + Embedding $\rightarrow$  perturbation isn’t an issue + we can drop left-infinite sequence.  }
+
+Despite the ease that one may work with a left-infinite history in the realm of theory, it is impossible to obtain such a sequence in any real-life application.  
+One does not in practice, fortunately, need the entire left-infinite history of the input thanks to the Uniform Attraction Property (UAP). 
+The next definition is stated in an informal manner as the more rigourous definition makes use of processes, a concept which would take some amount of paper to establish and prove - an exercise which would detract from the principal thrust of this paper/project/thesis.
+
+\textbf{Is it VERY obvious to the reader at this point what the general 'thrust' of the paper IS?} \textbf{Cite articles}
+
+Recall that the Unique Attraction Property guarantees that all trajectories \textit{(which ones?)} will converge to the same trajectory in forward time. 
+Incredibly, this permits one to initialise a driven system $g$ with an altogether arbitrary initial value $y_m\in{X}$ where $m\in\mathbb{Z}$ and the UAP guarantees that the sequence $\{y_{m+1}, y_{m+2},\ldots\}$ satisfying the relation $y_n=g(u_n, y_{n-1})$  where $k\geq{m}$  will uniformly approach the elements $\{x_n\}$ of the actual solution.
 (see [41, Theorem 1] or [11, Eqn.
@@
 
-One may now appreciate an even more astounding result.
-$g$ having the USP is equivalent to the UAP and so one need not establish any additional results to ensure that the sequence uniformly approach the unique solution $\Psi$.
-This vastly simplifies the work needed to be done in setting up a problem to guarantee that the underlying system will be accurately 'represented' by the conjugate system.
-\textbf{Check this statement again to make sure it's not nonsensical.}
-
-\begin{figure}[h]
+%Change the theorem above so that th.4 is split up into two parts.
+
+\begin{figure}[ht]
   \includegraphics[scale=0.4]{memloss_conttime.png}
@@
 
-This also solves the problem of perturbations or noise introduced by the observable or measurement function. \textbf{(rephrase)}
+One may now appreciate an even more astounding result.
+$g$ having the USP is equivalent to the UAP and so one need not establish any additional results to ensure that the sequence uniformly approach the unique solution $\Psi$.
+This vastly simplifies the work needed to be done in setting up a problem to guarantee that the underlying system will be accurately 'represented' by the conjugate system.
+\textbf{Check this statement again to make sure it's not nonsensical.}
+
+This also solves the problem of perturbations or noise introduced by the observable or measurement function. \textbf{(rephrase, expand and explain)}
+
+\subsection{The next step in Dynamics}
 

### 2022-01-01 Gandhi_Project_Report/draft1.tex
Done 4.4

@@
 
-The next result is of relative import but to fully comprehend its influence, some discussion still remains.
-Before stating it, we define a concept which will directly be shown to be equivalent to the USP.
-
-\begin{Definition}
-  [\bf Uniform Attraction Property]\label{Dfn_UAP}
-  A driven system $g$ has the Uniform Attraction Property (UAP) if, regardless of starting position, all trajectories converge to a single trajectory as time flows forward. 
-  This trajectory is also the unique solution sequence $x$ to the input sequence $u$ as mentioned above \ref{here_UAP_firstestablished}.
-\end{Definition}
+The next result is of relative import but to fully comprehend its influence, some discussion still remains. We state it in part and postpone the proof.
+% Before stating it, we define a concept which will directly be shown to be equivalent to the USP.
 
@@
   \begin{enumerate}[noitemsep, label=\roman*.]
-    \item $g$ has the USP 
-    \item $g$ is a topological contraction 
-    \item $g$ has the Unique Attraction Property (UAP). 
+    \item $g$ has the USP. 
+    \item $g$ is a topological contraction.
   \end{enumerate}
@@
 \begin{proof}
-  (we’ll show this)  [11] eqn.6 
+  (postponed)  [11] eqn.6 
 \end{proof}
 
-We have not as of yet treated the UAP properly, and establish the necessary result here (without proof) so as not to break up the completeness of theorem.
-
-\textbf{The proof of the result for the UAP is left to literature. Why? Cite.}
+% We have not as of yet treated the UAP properly, and establish the necessary result here (without proof) so as not to break up the completeness of theorem.
+
+% \textbf{The proof of the result for the UAP is left to literature. Why? Cite.}
 
@@
 We make special note of a specific driven system. The function 
-
-\begin{equation}  \label{eqn_driveng}
-%    \[ 
+\begin{equation}  \label{eqn_driving}
 g(u,x)=(1-a)x+a{\cdot}tanh(Au+\alpha{a}Bx)
-% \]
 \end{equation} 
-
-is both SI-invertible and possesses the USP. The proof of these two facts is rather involved and  hence the proof, instead of being replicated here, may be found in (cite relevant article) 
+is both SI-invertible and possesses the USP. The proof of these two facts is rather involved and  hence the proof, instead of being replicated here, may be found in \textbf{(cite relevant article)}
 This specific driving function $g$ is used in our implementation and is more completely discussed in chapter \ref{sect5}
-
-\textbf{Expand: UAP + Embedding $\rightarrow$  perturbation isn’t an issue + we can drop left-infinite sequence.  }
 
@@
 One does not in practice, fortunately, need the entire left-infinite history of the input thanks to the Uniform Attraction Property (UAP).
-The next definition is stated in an informal manner as the more rigourous definition makes use of processes, a concept which would take some amount of paper to establish and prove - an exercise which would detract from the principal thrust of this paper/project/thesis.
+
+\begin{Definition}
+  [\bf Uniform Attraction Property]\label{Dfn_UAP}
+  A driven system $g$ has the Uniform Attraction Property (UAP) if, regardless of starting position, all trajectories \textbf{(of what?)} converge to a single trajectory as time flows forward. 
+  This trajectory is also the unique solution sequence $x$ to the input sequence $u$ as mentioned above \ref{here_UAP_firstestablished}.
+\end{Definition}
+
+This definition is stated in a not completely rigourous manner as the more formal definition makes use of processes, a concept which would take some time to establish and detracts from the principal thrust of this paper/project/thesis.
 
@@
 
-Recall that the Unique Attraction Property guarantees that all trajectories \textit{(which ones?)} will converge to the same trajectory in forward time. 
-Incredibly, this permits one to initialise a driven system $g$ with an altogether arbitrary initial value $y_m\in{X}$ where $m\in\mathbb{Z}$ and the UAP guarantees that the sequence $\{y_{m+1}, y_{m+2},\ldots\}$ satisfying the relation $y_n=g(u_n, y_{n-1})$  where $k\geq{m}$  will uniformly approach the elements $\{x_n\}$ of the actual solution.
+As the Unique Attraction Property guarantees that all trajectories \textit{(which ones?)} will converge to the same trajectory as time moves forward. 
+Incredibly, this permits one to initialise a driven system $g$ with an altogether arbitrary initial value $y_m\in{X}$ where $m\in\mathbb{Z}$ and the UAP then guarantees that the sequence $\{y_{m+1}, y_{m+2},\ldots\}$ which satisfies the relation $y_n=g(u_n, y_{n-1})$  for  $k\geq{m}$  will uniformly approach the elements $\{x_n\}$ of the actual solution.
 (see [41, Theorem 1] or [11, Eqn.
@@
   \label{fig:memoryLosscont_time}
-%\caption{Approaching the true solution if g has the UAP}
+\caption{Approaching the true solution if g has the UAP}
 \end{figure}
 
-One may now appreciate an even more astounding result.
-$g$ having the USP is equivalent to the UAP and so one need not establish any additional results to ensure that the sequence uniformly approach the unique solution $\Psi$.
-This vastly simplifies the work needed to be done in setting up a problem to guarantee that the underlying system will be accurately 'represented' by the conjugate system.
-\textbf{Check this statement again to make sure it's not nonsensical.}
-
-This also solves the problem of perturbations or noise introduced by the observable or measurement function. \textbf{(rephrase, expand and explain)}
+One may now appreciate an even more astounding result: $g$ having the USP is equivalent to the UAP. We restate the above theorem.
+
+\begin{Theorem}
+  The following statements are equivalent:
+  \vspace{-8mm}
+  \begin{enumerate}[noitemsep, label=\roman*.]
+    \item $g$ has the USP.
+    \item $g$ is a topological contraction.
+    \item $g$ has the UAP.
+  \end{enumerate}
+\end{Theorem}
+\begin{proof}
+  (We'll show this)  [11] eqn.6 
+\end{proof}
+
+One need therefore not establish any additional results to ensure that the sequence uniformly approaches the unique solution $\Psi$.
+This vastly simplifies the effort necessary to set up a problem in order to guarantee that the underlying system will be accurately represented by the conjugate system.
+
+This also solves the problem of perturbation/noise introduced by the observable or measurement function.
+Recall from before that an observable is inherently a map that discretises the underlying continuous-time system.
+Moreover, measurement and device error introduce mistakes.
+Since we are working with a chaotic system displaying sensitive dependence on initial conditions, these small errors could potentially send the trajectory into a completely different attractor.
+In the presence of the UAP, however, small measurement errors introduced into the system do not pose the same danger as before.
+
+\textbf{(Rephrase and/or expand)}
 
@@
 
-The following theorem establishes the existence of a well-defined map GT describing the single-delay dynamics of the system above.
+The following theorem establishes the existence of a well-defined map $G_T$ describing the single-delay dynamics of the system above.
 
@@
 Note that $H_2$ maps an entire left-infinite solution sequence from $\Psi$ to an element in $X\times{X}$.
-
 We now have the following (compare with \ref{SCausal_H} above):
 
-\emph{graph here causes issues currently}
-% \begin{center}
-% %    \[ 
-%       \psset{arrows=->, arrowinset=0.25, linewidth=0.6pt, nodesep=3pt, labelsep=2pt, rowsep=0.7cm, colsep = 1.1cm, shortput =tablr}
-%       \everypsbox{\scriptstyle}
-%       \begin{psmatrix}
-%       \widehat{U}_T  && \widehat{U}_T \\
-%       Y_T && Y_T \\
-%       %%%
-%      %  \ncline{1,1}{1,2}^{\widehat{T}} \ncline{1,1}{2,1} <{h}
-%      %  \ncline{1,2}{2,2} > {H_2}
-%      %  \ncline{2,1}{2,2}^{G_T}
-%       \end{psmatrix}
-%      % \]
-%     \end{center}
-
-\textbf{Show that $(Y_T, G_T)$ is semi-conjugate to $(\widehat{U}, \widehat{T})$. }
-
-\textbf{Summarising the discussing thus far:}
-It is easy to lose onself in the symbols, so we take a moment to review our progress up until this point.
+\begin{equation}
+  %  \[ 
+      \psset{arrows=->, arrowinset=0.25, linewidth=0.6pt, nodesep=3pt, labelsep=2pt, rowsep=0.7cm, colsep = 1.1cm, shortput =tablr}
+      \everypsbox{\scriptstyle}
+      \begin{psmatrix}
+      \widehat{U}_T  && \widehat{U}_T \\
+      Y_T && Y_T \\
+      %%%
+     %  \ncline{1,1}{1,2}^{\widehat{T}} \ncline{1,1}{2,1} <{h}
+     %  \ncline{1,2}{2,2} > {H_2}
+     %  \ncline{2,1}{2,2}^{G_T}
+      \end{psmatrix}
+    %  \]
+    \end{equation}
+
+\begin{Theorem}
+    $(Y_T, G_T)$ is semi-conjugate to $(\widehat{U}, \widehat{T})$.
+\end{Theorem}
+\begin{proof}
+  \textbf{Show}
+\end{proof}
+
+\subsubsection*{Summarising the discussing thus far:}
+
+It is easy to lose the birds-eye view, so we take a moment to review our progress up until this point.
 
@@
 \begin{enumerate}
-\item We are interested in a dynamical system $(U,T)$ with unknown dynamics for whichever reason.
-\item To determine information about this system $(U,T)$, we shift our focus by rather aiming to determining the dynamics of the inverse-limit system $(\widehat{U}, \widehat{T})$ which, given certain assumptions, we can guarantee to be at least semi-conjugate to $(U,T)$.
-\item If the driven system $g$ (and $\{u_n\}$ in $U$  is an orbit of $T$), we have that $G_T$ exists. 
-\item If, furthermore, $g$ may be shown to have the USP, we showed that $(Y_T, G_T)$ is semi-conjugate to $(\widehat{U}, \widehat{T})$.
-\item If $T$ is also (and at this stage this is very much our hope) a homeomorphism, then $(Y_T, G_T)$ topologically conjugate to $(\widehat{U}, \widehat{T})$ in which $(U,T)$ has been embedded! 
+\item We are interested in a some dynamical system $(U,T)$ with unknown dynamics.
+\item To determine properties about this system $(U,T)$ and predict its future evolution, we determine the dynamics of the inverse-limit system $(\widehat{U}, \widehat{T})$.
+Given certain assumptions, we can guarantee that $(\widehat{U}, \widehat{T})$ is at least semi-conjugate to $(U,T)$.
+\item If the driven system $g$ is SI-invertible (and $\{u_n\}\in{U}$  is an orbit of $T$), the map $G_T$ exists. 
+\item If, furthermore, $g$ has the USP, $(Y_T, G_T)$ is semi-conjugate to $(\widehat{U}, \widehat{T})$.
+\item If we can assume that $T$ is a homeomorphism, $(Y_T, G_T)$ is topologically conjugate to $(\widehat{U}, \widehat{T})$, an extension space of $(U,T)$
 \end{enumerate}
-One can therefore learn the single-delay lag dynamics of the driven system states via $G_T$ with enough data thanks to the USP.
-This enables us to do at least 2 things: 1) Forecast future values of $x_n$ via iterates of $G_T$ (as $G_T$ can be determined.
-2) Forecast future values of $u_n$.
+
+One can therefore learn the single-delay lag dynamics of the driven system states via $G_T$ with enough data thanks to the USP/UAP. This enables us to do at least 2 things: 
+\vspace{-8mm}
+\begin{itemize}
+\item Forecast future values of $x_n$ via iterates of $G_T$ (as $G_T$ can be determined.
+\item Forecast future values of $u_n$. 
+\end{itemize}
 
@@
 
-By establishing the existence of the map $G_T$, we’ve essentially embedded the attractor $U$ \textbf{So far we haven't really spoken about an attractor} into the higher dimensional space $X\times{X}$.
+By establishing the existence of the map $G_T$, we’ve essentially embedded the attractor $U$ \textbf{(So far we haven't really spoken about an attractor)} into the higher dimensional space $X\times{X}$.
+\textbf{(Why is it true that it's higher-dimensional?)}
 In layman’s terms, this ensures that there is more “dimensional room” for the underlying system’s underlying dynamics to “move”.

### 2022-01-03 Gandhi_Project_Report/draft1.tex
cleaning git

@@
-\documentclass[12 pt]{article}
+% \documentclass[12 pt]{article}
+\documentclass[a4paper,12pt,twoside]{book}
+\pagestyle{headings}
 \usepackage[utf8]{inputenc}
@@
 \usepackage{amsthm, amssymb, amsfonts, latexsym}
-\usepackage{pst-all}
+% \usepackage{pst-all}
 
@@
 \setlength{\headheight}{0 mm}
+
+\usepackage{pst-node}
 
@@
                 headheight=0pt,% No Header
-                headsep=10pt,%
+                headsep=30pt,%
                 footskip=25pt,
@@
 
-\externaldocument{Supp_Koopman-0808}
+% \externaldocument{Supp_Koopman-0808}
 \begin{document}
@@
 
-\section{Introduction}\label{ch1}
+\tableofcontents
+
+% \chapter{Introduction}\label{ch1}
 
@@
 
-\section{Discrete-time Dynamical Systems}\label{ch2}
+% \chapter{Discrete-time Dynamical Systems}\label{ch2}
 
@@
 We refer to \cite{devaney2018introduction, de2013elements} for more details.
+\label{Dfn_DS}
 
@@
 
-% \subsection{Invariant Sets}
+% \section{Invariant Sets}
 
@@
 
-% \subsection{Chaos}
+% \section{Chaos}
 
@@
 
-% \subsection{Conjugacy}
+% \section{Conjugacy}
 
@@
 
-\section{The Learning Problem}\label{ch3}
+% \chapter{The Learning Problem}\label{ch3}
 
@@
 
-\subsection{Taken Embedding Theorem} \label{subs_TakensEmbedding}
+% \section{Taken Embedding Theorem} \label{sect_TakensEmbedding}
 
@@
 
-% \subsection{Obstacles to Takens}
+% \section{Obstacles to Takens}
 % Despite the fact that the Takens' Embedding Theorem is a powerful result and provides compelling reason to believe that it is possible to accurately ()\emph{right word?}) reconstruct the underlying space, it does present some serious limitations:
@@
 
+% \emph{ADD THIS TO ERRORS IN TAKENS: NOT GUARANTEED GLOBALLY DISSIPATIVE}
 % \textbf{Explanation continued here}
@@
 % \textbf{Continue explaining here}
-\newpage
-
-\section{Driven Dynamical Systems to Forecast Problems} \label{ch4}
+% \newpage
+
+\chapter{Driven Dynamical Systems to Forecast Problems} \label{ch4}
 
@@
 We also consider the conditions a driven system should have to avoid adding distortion to its state space representation, as well as then ensuring that the single-delay dynamics (SDD) are conjugate (or at least semi-conjugate to the system).
-The SDD can then be used to forecast and reconstruct the underlying attractor. \textbf{Is this attractor assumed to exist?}
-
-\subsection{Nonautonomous and Driven Dynamical Systems}
-
-A nonautonomous dynamical system is simply a dynamical system (as defined before in \textbf{link to above}) where the input $u$ from the input space $U$, a  topological space, is time-dependent.
-\textbf{This means it can no longer be generated by a flow?}
+The SDD can then be used to forecast and reconstruct the underlying attractor. \ednote{Is this attractor assumed to exist?}
+
+\section{Nonautonomous and Driven Dynamical Systems}
+
+A nonautonomous dynamical system is simply a dynamical system (as defined before in \ref{Dfn_DS}) where the input $u$ from the input space $U$, a  topological space, is time-dependent.
 
@@
 
-\subsection {Unique Solution Property}
+\section {Unique Solution Property}
 
@@
 
-\emph{(More discussion on USP?)}
-
+One of the first result obtained upon defining the USP, is the fact that every solution will attract any different initial conditions towards the component parts of the solution, i.e.
+that the solution $\Psi$ is a non-autonomous uniform attractor.
+\ednote{Define Uniform Attractor before stating theorem.}
+
+\begin{Theorem} \label{Thm_AttractorSoln}
+  If $g$ has USP, then any solution to $g$ is a non-autonomous uniform attractor 
+\end{Theorem}
+\begin{proof}
+  \cite{Manju_Nonlinearity}
+\end{proof}
+
+\ednote{Additional discussion on USP?} 
 Having already defined the reachable set $X_U$, we pause for a moment to fix additional notation.
@@
 In this case, the function $H:\overleftarrow{U}\to\overleftarrow{X}_U$, a map that is both continuous and surjective, is called a \emph{causal mapping}.
-
+make special note of a 
 \begin{Definition}
@@
 \begin{proof}
-  (postponed)  [11] eqn.6 
+  (postponed) 
 \end{proof}
@@
 
-\subsection {Choosing the driven system $g$}
+\section {Choosing the driven system $g$}
 
@@
 
-When a causal embedding $H$ exists for the driven system $g$, one can map an arbitrary input ${u}$ onto the solution space $X$ without additional distortion or information-loss.  \textbf{(Cite.)} 
+When a causal embedding $H$ exists for the driven system $g$, one can map an arbitrary input ${u}$ onto the solution space $X$ without additional distortion or information-loss.  \ednote{(Cite.)} 
 When an embedding is established, the question of possible additional complexity in the solution is removed by guaranteeing that, since the systems are conjugate (semi-conjugate, (refer)), $g$ does not add any (some) complexity to the system.
 It is, however, a balancing act as it also undesirable to choose a function $g$ that quenches the temporal structure in $u$ by contracting to such a degree that the ability to recover information from the original system is lost completely.
-To obtain a suitably complex function $g$, it is desired that the the reachable set of a driven system  be large enough to relate to the input.  \textbf{(Expand here. It stops a bit abruptly.)}
+To obtain a suitably complex function $g$, it is desired that the the reachable set of a driven system  be large enough to relate to the input.  \ednote{(Expand here. It stops a bit abruptly.)}
 
 In the example above (\ref{ex_halfux}), the input's temporal variation cannot be related to the reachable set as $X_U$ consists of a single element and so little, if not no, information is encoded.
-Rather, the reachable set of a driven system must therefore be such that the inverse-limit space of $U_T$ can be embedded in some finite self-product of the reachable set of $T$ \textbf{(refer or cite)}.
+Rather, the reachable set of a driven system must therefore be such that the inverse-limit space of $U_T$ can be embedded in some finite self-product of the reachable set of $T$ \ednote{(refer or cite)}.
 To this end, consider the notion of State-Input (SI) Invertibility.
@@
 
-\textbf{Explain more here.}
-
-We make special note of a specific driven system. The function 
+SI-invertibility promises that 'enough' information is retained when $g$ is chosen without introducing additional unwanted complexity. \ednote{Expand why this is true.}
+
+It is worth taking note of a specific driven system. The function 
 \begin{equation}  \label{eqn_driving}
@@
 This definition is stated in a not completely rigourous manner as the more formal definition makes use of processes, a concept which would take some time to establish and detracts from the principal thrust of this paper/project/thesis.
-
-\textbf{Is it VERY obvious to the reader at this point what the general 'thrust' of the paper IS?} \textbf{Cite articles}
+\ednote{Is it VERY obvious to the reader at this point what the general thrust of the paper is?}
 
@@
 In the presence of the UAP, however, small measurement errors introduced into the system do not pose the same danger as before.
-
-\textbf{(Rephrase and/or expand)}
-
-\subsection{The next step in Dynamics}
+\ednote{(Rephrase and/or expand)}
+
+\section{The next step in Dynamics}
 
@@
 
-The following theorem establishes the existence of a well-defined map $G_T$ describing the single-delay dynamics of the system above.
+The following theorem establishes the existence of a well-defined map $G_T$ describing the single-delay dynamics(SDD) of the system above.
 
@@
 \begin{Definition}
-  The inverse-limit space $\widehat{U}_T$, a subspace of $\overleftarrow{U}$, is defined by $$\widehat{U}_T:=\{(\ldots, u_{-2}, u_{-1}):T(u_n)=u_{n+1}\}$$ This space is well-defined since $T:U\to{U}$ is surjective by assumption.
-\textbf{Expand why}
+  [\bf Inverse-Limit Space]\label{Dfn_InvLimSpace}
+  The inverse-limit space $\widehat{U}_T$, a subspace of $\overleftarrow{U}$, is defined by $$\widehat{U}_T:=\{(\ldots, u_{-2}, u_{-1}):T(u_n)=u_{n+1}\}$$ 
 \end{Definition}
+
+The inverse-limit space is well-defined since $T:U\to{U}$ is surjective by assumption. \ednote{Should I explain this in more details or am  I then over-simplifiying?}
 
@@
 
-\begin{equation}
+\begin{equation} \label{Fig_inverse_limsystem}
   %  \[ 
@@
 
-\subsubsection*{Summarising the discussing thus far:}
+\section*{Summarising the discussing thus far:}
 
@@
 
-One can therefore learn the single-delay lag dynamics of the driven system states via $G_T$ with enough data thanks to the USP/UAP. This enables us to do at least 2 things: 
+One can therefore learn the SDD of the driven system states via $G_T$ with enough data thanks to the USP/UAP. This enables us to do at least 2 things: 
 \vspace{-8mm}
-\begin{itemize}
+\begin{itemize}[noitemsep]
 \item Forecast future values of $x_n$ via iterates of $G_T$ (as $G_T$ can be determined.
@@
 
-\subsection{A discussion of $G_T$ }
-
-By establishing the existence of the map $G_T$, we’ve essentially embedded the attractor $U$ \textbf{(So far we haven't really spoken about an attractor)} into the higher dimensional space $X\times{X}$.
-\textbf{(Why is it true that it's higher-dimensional?)}
+\section{A discussion of $G_T$ }
+
+In the above sections, we established the map $G_T$ describing the SDD of a driven system.
+By establishing the existence of this map, we’ve essentially embedded the attractor $U$ \ednote{(So far we haven't really spoken about an attractor)} into the higher dimensional space $X\times{X}$.
+\ednote{(Why is it true that it's higher-dimensional?)}
 In layman’s terms, this ensures that there is more “dimensional room” for the underlying system’s underlying dynamics to “move”.
@@
 
-In \emph{[Koopman]} it is illustrated in an empirical fashion in that the map $G_T$ describes dynamics which are less functionally complex than that of $T$ or of the map $\Phi_{2d,\theta}$ (as discussed above in the section on Takens).
+In \cite{manjunath2021universal} it is illustrated in an empirical fashion in that the map $G_T$ describes dynamics which are less functionally complex than that of $T$ or of the map $\Phi_{2d,\theta}$.
 This is done by implementing a Recurrent Neural Network (RNN), but not discussed as of yet.
-\emph{(Add in that we'll be using Pearson coefficient?)}
-
-We opt to learn $G_T$ in an indirect manner by defining a new map $\Gamma$. It follows immediately from $G_T$’s existence that $\Gamma$ also exists.
-
-The reasons for taking this roundabout approach remain to be discussed, but a pat answer may be immediately given: When we have learnt $\Gamma$, we can drive the system autonomously and then we know $G_T$ anyway.
+\ednote{(Add in that we'll be using Pearson coefficient?)}
+
+We opt to learn $G_T$ in an indirect manner by defining a new map $\Gamma:(x_{n-1},x_n)\mapsto{u_n}$. It will follow immediately from $G_T$’s existence that $\Gamma$ also exists.
+
+The reasons for taking this roundabout approach remain to be discussed in \ref{subs_LearnGamma}, but a pat answer may immediately be given: When $\Gamma$ has been learnt, the system can be driven autonomously and then $G_T$ is known anyway.
+We formalise this with a theorem.
 
 \begin{Theorem}
-  When $x_n$ and $x_{n-1}$ are successive points on a solution obtained for input-orbit of $T$ $\{u_n\}$, then the map $\Gamma: X\times{X}\to{U}$ defined by $(x_{n-1},x_n)\mapsto{u_n}$ exists whenever $G_T$ exists [Koopman, Theorem 3c]
+  When $x_n$ and $x_{n-1}$ are successive points on a solution obtained for input-orbit of $T$ $\{u_n\}$, then the map $\Gamma: X\times{X}\to{U}$ defined by $(x_{n-1},x_n)\mapsto{u_n}$ exists whenever $G_T$ exists
 \end{Theorem}
-
-\begin{Definition}
- The projection mapping $\pi_i$ is defined as follows: $$\pi_1:(a_1,a_2)\to{a_1}$$ and $$\pi_2:(a_1,a_2)\to{a_2}$$ 
-\end{Definition}
-
-We now have the graph below:
-
-\begin{figure}[ht]
+\begin{proof}
+  From [Koopman, Theorem 3c] 
+\end{proof}
+
+Projection mappings $\pi_i$ are defined in the traditional meaning where a $k$-dimensionalvector is projected to it's $i^{th}$ component such that $$\pi_i:(a_1,a_2, \ldotp, a_{k-1}, a_k)\to{a_i}$$
+
+% \begin{Definition}
+%   [\bf Projection Mapping]\label{Dfn_ProjMap}
+%  The projection mapping $\pi_i$ is defined as follows: $$\pi_1:(a_1,a_2)\to{a_1}$$ and $$\pi_2:(a_1,a_2)\to{a_2}$$ 
+% \end{Definition}
+
+The graph in equation \ref{Fig_inverse_limsystem} is then extended as below:
+
+\begin{figure}[ht] \label{Fig_Gamma}
   \includegraphics[scale=0.3]{gammatemp.png}
@@
 
-And a final set of equations - equations that have been constructed from data!
+The problem finally simplifies to the issue of learning the map $\Gamma$ and combining this with the projection mapping $\pi_2$ and the function $g$, which will be known.
+A final set of equations is obtained - equations that have been entirely constructed from data.
 \begin{eqnarray}
@@
 
-The problem thus simplifies to the issue of learning the map $\Gamma$ and combining this with the projection mapping $\pi_2$ and the function $g$, which will be known.
-
-\subsection{Advantages of learning $\Gamma$}
-
-Why are we going such a lengthy roundabout route, one may immediately ask.
-Why not just learn the map $G_T$ from the get-go?
-On the surface, this seems to be an arbitrarily chosen path with no real reason for the choice made, so we take a pause again and discuss our reasoning for learning Gamma.
+\section{Advantages of learning $\Gamma$} \label{subs_LearnGamma}
+
+One may immediately ask why we opt to take such a roundabout route; why not just learn the map $G_T$ from the get-go?
+On the surface, this seems to be an arbitrary decision path with no real reasoning, so we take a pause again and discuss the motivation for learning Gamma.
 There are a number of distinct advantages.
@@
 Learning $\Gamma$ saves computational resources.
-This follows from the fact that the input $u_n$ at time $n$ lies in a low-dimensional subspace of the high-dimensional subspace $X\times{X}$ due to the fact that it is zero-padded.
+This is due to the fact that the input $u_n$ lies in a lower-dimensional subspace of the high-dimensional space $X\times{X}$.
+In practice, if the input is of a  lower dimension, one may easily embed it into the space $X\times{X}$ by padding the vector $u_n$ with zeroes.
+
+\textbf{Can we discuss how to put this across well? I would like to add it in as I think it beneficial to the discussion.}
+
+Moreover, two types of stability are achieved when $g$ has the USP.
 
 Input- and parameter-related stability is achieved if $g$ has the USP
-\newline \emph{Define the two types of stabilities}
+\newline \emph{Define/Explain the two types of stabilities}
 \newline \emph{Sentence on $u$, $v$, tails, USP and product topology}
-\newline If we opt to work with $Y-T$, errors can occur. \emph{Expand}
-\newline $\Gamma$ is more stable. \emph{Explain why}
-
-Global dissipativity prevents large numerical errors due to input data.
-$\Gamma$ is globally dissipative (\emph{Define}) whereas $G_T$ is not guaranteed to be.
-IF we were to rather learn $G_T$, we would risk having exactly the same problems arise that could arise also for Takens.
-\textbf{\emph{ADD THIS TO ERRORS IN TAKENS: NOT GUARANTEED GLOBALLY DISSIPATIVE}}
-\newline \emph{Expand here ...}
-
-\emph{Where to add:If $g$ has USP, then any solution to $g$ is a non-autonomous uniform attractor}
+\newline If we opt to work with $Y_T$, errors can occur. Expand
+\newline $\Gamma$ is more stable. Explain why.
+
+Thirdly, $\Gamma$ is globally dissipative whereas $G_T$ is not guaranteed to be(\emph{Cite}). \ednote{Add linking sentence.} Global dissipativity prevents large numerical errors due to input data.
+
+The $\omega$-limit set $\omega(u;T)$ of a point $u$ is defined to be the collection of limit points of the sequence $\{x,Tu,T^2u,\ldots\}$.
+This set is nonempty, and $\omega(u;T)$ is invariant.
+\textbf{(Change)}
+
+\begin{Definition}
+  [\bf Global Dissipativity]\label{Dfn_GlobalDiss}
+	We say a dynamical system $(Z,f)$ is globally dissipative if there exists a nonempty proper closed subset $B$ of $Z$  so that for all $x\in Z\setminus B$, (i).
+$\omega(x;f) \subset B$  and (ii).
+$B$ is positively invariant that is $f(B) \subset B$.
+We call any such closed subset $B$ a trapping set of $f$.
+\end{Definition}
+
+If a system is not globally dissipative, neglibly small errors result in major errors which thus induce predictions to fail utterly.
+If one learns $G_T$ directly, we are in danger of replicating the problems arising for Takens.
 
@@
 
-\section{Implementation} \label{sect5}
+\chapter{Implementation} \label{sect5}
 

### 2022-01-03 Gandhi_Project_Report/draft2.tex
cleaning git

@@
 
-\ednote{M: Saying $\phi$ is then a semi-conjugacy between $T$ and $S$ could be ambiguous as it does not specify which system is on the top in the commutativity. So please edit accodingly}
+\ednote{M: Saying $\phi$ is then a semi-conjugacy between $T$ and $S$ could be ambiguous as it does not specify which system is on the top in the commutativity. So please edit accordingly}
 
@@
 
-Suppose we only have the observations $\theta(u_0), \theta(u_1), \ldots, \theta(u_m)$ of the true system states $u$ in an unknown dynamical system $(U,T)$ and we wish to predict the values $\theta(u_{m+1}), \theta(u_{m+2})$.
+Suppose we only have the observations $\theta(u_0), \theta(u_1), \ldots, \theta(u_m)$ of the true system states $u$ in an unknown dynamical system $(U,T)$ and we wish to predict the values $\theta(u_{m+1}), \theta(u_{m+2})$ and so on.
+First we recall the method of Takens delay embedding.
 
-Here we consider a continuous-time dynamical system $(U,T)$ with true state $u$ in the input space $U$ and then a measured signal $\theta$ (where $\theta(u_n)=\theta(u(t_n))$ for $n\in\mathbb{N}$.
-The sequence $\{x_n\}$ represents a scalar time-series and intuitively $\theta$ may be thought to represent a probe inserted into a larger system and itself measuring/extracting only a small part of the greater system state at time $t=n$.
+In this method we consider a discrete-time dynamical system  
+defined on a nice space (a smooth manifold) that can be obtained as time-$K$ map of a continuous time-dynamical system \ednote{M: edited here}.
+We make observations from such a system, i.e., we observe the evolution of $\theta(w_0)$, i.e., we observe  a finite set of values of $u_n :=  \theta(w_n) = \theta(Tw_{n-1})$.
+The sequence $\{u_n\}$ represents a scalar time-series and intuitively $\theta$ may be thought to represent a probe inserted into a larger system and itself measuring/extracting only a small part of the greater system state at time $t=n$.
 Consider for example a thermometer erected to measure the ambient temperature in a local village.
@@
 
-From here we construct a multidimensional observable using the method of delay-coordinate stacking. 
-\textbf{move to later} Given the sequence $\{\theta_1, \theta_2, \ldots\}$, we define the variable \newline $y_n:=[\theta_n, \theta_{n-\tau}, \ldots, \theta_{n-(2d-2)\tau}, \theta_{n-(2d-1)\tau}]^{T}\in\mathbb{R}_{d}\times{1}$ where $\tau$ represents the lag and $d$ the dimension of the attractor.
+%Given the sequence $\{u_1, u_2, \ldots, u_n}$, we define the variable \newline $y_n:=[\theta_n, \theta_{n-\tau}, \ldots, \theta_{n-(2d-2)\tau}, \theta_{n-(2d-1)\tau}]^{T}\in\mathbb{R}_{d}\times{1}$ where $\tau$ represents the lag and $d$ the dimension of the attractor.
 (We shall return to these terms shortly and their discussion may be put on hold for the time being).
@@
 In almost all cases we can merely insert probes into a system to obtain partial information by measurements taken.
-Moreover, the process of taking a measurement itself introduces 2 additional difficulties:
+Moreover, the process of taking a measurement itself introduces 2 additional difficulties: \ednote{M: Why is discretization a difficulty?
+Maybe you replace difficulty by saying that we need consider two aspects}
 \vspace{-8mm}
@@
 
-More formally we may say that in the process of measuring a system, we obtain a 1-dimensional time-series $\{\theta(u(t))\}$, where $\theta:\mathbb{R}^n\to\mathbb{R}$ is a smooth observation or measurement function.
-This series represents a possibly noise-augmented dataset containing information of certain parts of the original system.
+From here we construct a multidimensional observable using the method of stacking previous observations, i.e., we create delay-coordinate map defined by
+$\Phi_{k,\theta}(w) := (\theta(T^{-k}w)\ldots,\theta(T^{-1}w),\theta(w))$.
+The essence of Takens theorem is that When $k$ is sufficiently large, we can define a dynamical system on the $\mathbb{R}^{k+1}$ that whose states are $\Phi_{k,\theta}(w)$, $\Phi_{k,\theta}(Tw)$, $\Phi_{k,\theta}(T^2w)$ and this dynamical system is topologically conjugate to the unknown underlying system $(W,T)$.
+We recall the Takens delay embedding theorem next.
 
-We may then ask ourselves the question: Is it in any way possible to retain information about the system state $x(t)$ in this temporal data-series $\theta(x(t))$?
+\ednote{M: I have omitted many statements here}
+
+%More formally we may say that in the process of measuring a system, we obtain a 1-dimensional time-series $\{\theta(u(t))\}$, where $\theta:\mathbb{R}^n\to\mathbb{R}$ is a smooth observation or measurement function.
+This series represents a possibly noise-augmented \ednote{M: Noise-augmented is not the right word} dataset containing information of certain parts of the original system.
+
+%We may then ask ourselves the question: Is it in any way possible to retain information about the system state $x(t)$ in this temporal data-series $\theta(x(t))$?
 The answer is easily yes if T is known.
@@
 
-Takens Theorem states a result that establishes a relationship between the observed and underlying dynamical systems by showing that the concetantion of a sufficiently large number of previous observations into a vector will, under certain conditions, generate a map between the vectors from the respective systems.
-\textbf{Check wording} We formultate the theorem from \cite{takens1981detecting}.
+Takens Theorem states a result that establishes a relationship between the observed and underlying dynamical systems by showing that the concatenation of a sufficiently large number of previous observations into a vector will, under certain conditions, generate a map between the vectors from the respective systems.
+\ednote{B: Check wording.
+M: Edited the statement of the theorem} We formulate the theorem from \cite{takens1981detecting}.
 
@@
 	[\bf Takens Embedding Theorem (adopted from \cite{takens1981detecting}] \label{Thm_Takens}
-            Let $W$ be a compact manifold of dimension $m$, and $d\ge m$ so that $2d\in\mathbb{Z}$. It is a 
+         Let $W$ be a compact manifold of dimension $m$, and $d\ge m$ so that $2d$ is an integer. It is a 
             generic property for the pair $(T, \theta)$,  where $T:W \to W$ is
-            a smooth diffeomorphism, and $\theta:W \to \mathbb{R}$ a smooth function, that the map $\Phi_{2d,\theta}:W \to \mathbb{R}^{2d+1}$ defined on $W$ by 
+            a smooth diffeomorphism, and $\theta:W \to \mathbb{R}$ a smooth function, the map $\Phi_{2d,\theta}:W \to \mathbb{R}^{2d+1}$ defined on $W$ by 
             $\Phi_{2d,\theta}(w) := (\theta(T^{-2d}w)\ldots,\theta(T^{-1}w),\theta(w))$
-            is a diffeomorphic embedding; by `smooth' we mean at least $C^2$. Consequently, there exists a map $\Ftheta: \Phi_{2d,\theta}(W) \to \Phi_{2d,\theta}(W)$ defined by 
-            \begin{equation}  \label{eqn_takens}
-              \Ftheta: (\theta(T^{-2d}w),\ldots,\theta(T^{-1}w),\theta(w)) \mapsto  (\theta(T^{-2d+1}w),\ldots,\theta(T^{-1}w),\theta(Tw))
-            \end{equation}
+            is a diffeomorphic embedding; by `smooth' we mean at least $C^2$.
+Consequently, there exists a map $F_\theta: \Phi_{2d,\theta}(W) \to \Phi_{2d,\theta}(W)$ defined by $$F_\theta: (\theta(T^{-2d}w),\ldots,\theta(T^{-1}w),\theta(w)) \mapsto
+            (\theta(T^{-2d+1}w),\ldots,\theta(w),\theta(Tw))$$
            so that $(W,T)$ is topologically conjugate to 
-            $(\Phi_{2d,\theta}(W), \Ftheta)$. 
+            $(\Phi_{2d,\theta}(W), F_\theta)$.    
 \end{Theorem}
 
+By generic we mean a residual set on a certain topology on an appropriate spaces of functions (that we do not describe here).
+By $C^2$, we mean a function that is twice differentiable, and the second derivative is continuous.
 (In our scenario, the input space $U$ is considered the attractor.)
 
-Some explanation is beneficial.
+%Some explanation is beneficial.
 Takens establishes a delay-coordinate map $\Ftheta$ defined in \ref{eqn_takens} for $T$ the flow (\ref{defn_flow}) defined by the update equation  $Tu_n=u_{n+1}$ and $theta$ our measurement function.
@@
 
-Alternatively we may also make the statement: Takens' Embedding Theorem guarantees that almost all dynamical systems can be reconstructed from just one noiseless observation sequence \textbf{this isn't obvious from the theorem's formulation, is it?} i.e.
+%Alternatively we may also make the statement: Takens' Embedding Theorem guarantees that almost all dynamical systems can be reconstructed from just one noiseless observation sequence \textbf{this isn't obvious from the theorem's formulation, is it?} i.e.
 for a great number of possible observation functions $\theta$, $\Ftheta$ preserves the topology of U.
@@
 \end{figure}
+\ednote{M: Please replace this figure with the one with in the Supplement in Adriaan's paper where $w$ appears instead of $u$ and this would be consistent with the statement in the theorem.
+Here, and throughout if you reproduce a figure, it must be stated where it was reproduced from.}
 
-Once again we pause and consider the graph to consolidate our understanding of the conjugacy (\textbf{are we being too simplistic?}).
+If we can learn the map $F_\theta$, from sufficient number of data points of $\{\phi_{2d,\theta}(w_n)\}$, then we also know how to forecast how $\theta(w_n)$.
+
+%Once again we pause and consider the graph to consolidate our understanding of the conjugacy (\textbf{are we being too simplistic?}).
 If we are in state $u$ and then evolve towards state $Tu$ before embedding into $\mathbb{R}^{2m-1}$, then this is equivalent to first embedding into $\mathbb{R}^{2m-1}$ and then evolving through the map $\Ftheta$, i.e.
@@
 
-Thus information about U can be retained in the time series' (or observation's) output.
+%Thus information about U can be retained in the time series' (or observation's) output.
 By preserving the topology on the manifold $U$ in the reconstruction space $X$ (\textbf{not discussed before}), we are also guaranteed the preservation of topological invariants of the manifold, of which dimensionality is one such invariant.
@@
 
-\subsection{Obstacles to Takens}
-Despite the fact that the Takens' Embedding Theorem is a powerful result and provides compelling reason to believe that it is possible to accurately ()\emph{right word?}) reconstruct the underlying space, it does present some serious limitations:
+\subsection{Practical Issues of Takens theorem}
+Despite the fact that the Takens' Embedding Theorem is a powerful result and provides compelling reason to believe that it is possible to accurately \emph{B: right word?
+M: Yes, I have edited the sentence} reconstruct a system that is conjugate to the underlying system, it does present some serious practical limitations:
 \vspace{-5mm}
 \begin{enumerate}
-\item Even if we can indeed find \Ftheta{}  our \emph{approximation} of \Ftheta{}  is a map from a larger set $\mathbb{R}^{2m+1}$ containing the embedded attractor.
-There are, however, no theoretical guarantees that \Ftheta{}  will keep this attractor $U$  the same.(\textbf{We haven't properly said that U is an attractor, or have we.})
-\item Takens only holds for noiseless observations.
-Suppose the case where \Ftheta{} was learnt, but due to measurement error or other noise we obtained the set $V$ and not the original attractor.
-Due to the chaotic nature of the underlying system (i.e.
-the fact that it has SIDC), this new system $V$ will then move in a different direction.
-This problem will later be eliminated by some properties of a new function $g$, namely the Unique Solution Property.
+\item Even if we can indeed find $F_\theta$  our \emph{approximation} of  $F_\theta$ is a map from a larger set $\mathbb{R}^{2d+1}$ containing the embedded attractor.
+There are, however, no theoretical guarantees that
+$F_\theta$ will retain  $V=\Phi_{2d,\theta}(W)$  as an attractor although $W$ could be an attractor. \ednote{B: We haven't properly said that U is an attractor, or have we. M: I have rephrased }
+\item Takens theorem is stated for noiseless observations.
+Due to noise $\epsilon_n$, the delay vector $\Phi_{2d,\theta}(w_n) + \epsilon_n$,  may lie outside $V$, and further, due to the chaotic nature of the underlying system (i.e.
+the fact that it has SIDC), the evolution of $\Phi_{2d,\theta}(w_n) + \epsilon_n$ under the map $F_\theta$ may move completely out of $V$.
+This problem can be overcome  by using a driven dynamical system with some properties, and we discuss this in Chapter *.
  \item It is an unfortunate case that the functional complexity (i.e., the relative number of oscillatory graphs \cite{manjunath2021universal}) as well as stability of the embedding \Ftheta{} (\textbf{cite?})depends on the choice of observable $\theta$.
 This means that we may even learn a map that complicates our learning task and in so doing adds obstacles to the road to learning a conjugate map with predictive power.
-(\textbf{Am I talking nonsense here?})
+\ednote{Am I talking nonsense here?
+M: Please avoid this point and delete it}
 \end{enumerate}
 
-Another limitation owes itself to the requirement that T be a surjective map. This is not always the case in practical examples and we may consider the case of a damped double pendulum.
+Another limitation owes itself to the requirement that T be a surjective map.
+This is not always the case in practical examples and we may consider the case of a damped double pendulum \ednote{M: Please remove the paragraph as well.
+We do not overcome this problem.
+You can add this in the next Chapter where you need T to be surjective or else you could include this at the begining of this chapter where you explain the forecasting problem.}
 

### 2022-01-03 Supp_Koopman-0808.tex
cleaning git

\documentclass[12 pt]{article}
\usepackage{graphicx}
\usepackage{harpoon}
%\usepackage{apacite} 
\usepackage{color} 
\usepackage{latexsym}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{mdframed}
%\usepackage[authoryear]{natbib}
\usepackage{hyperref, url}
\usepackage[show]{ed}
%\usepackage{subfigure}
\usepackage[ntheorem]{empheq} 
\usepackage{amsthm, amssymb, amsfonts, latexsym}
\usepackage{multirow}
\usepackage{multicol}
\usepackage[utf8]{inputenc}

\usepackage{subcaption}

\setlength{\parskip}{1.7em}
\setlength{\parindent}{0em}
\usepackage[ntheorem]{empheq}
%\setlength{\baselineskip}{20pt}

\setlength{\textwidth}{429.75499pt}
\setlength{\textheight}{643.20255pt}
\setlength{\oddsidemargin}{5 mm}
\setlength{\evensidemargin}{5 mm}
\setlength{\topmargin}{0 mm}
\setlength{\headsep}{0 mm}
\setlength{\headheight}{0 mm}

\newcommand{\citealtt}[1]{\citeauthor{#1},\citeyear{#1}}
\newcommand{\myycite}[1]{\citep{#1}}

\mathchardef\mhyphen="2D

%\newcommand{\cev}[1]{\reflectbox{\ensuremath{\vec{\reflectbox{\ensuremath{#1}}}}}}

\renewcommand{\thefigure}{S\@arabic\c@figure}

\usepackage{pst-node}
\usepackage{auto-pst-pdf}
\usepackage{tikz-cd}

\makeatletter
\DeclareRobustCommand{\cev}[1]{%
  \mathpalette\do@cev{#1}%
}
\newcommand{\do@cev}[2]{%
  \fix@cev{#1}{+}%
  \reflectbox{$\m@th#1\vec{\reflectbox{$\fix@cev{#1}{-}\m@th#1#2\fix@cev{#1}{+}$}}$}%
  \fix@cev{#1}{-}%
}
\newcommand{\fix@cev}[2]{%
  \ifx#1\displaystyle
    \mkern#20mu
  \else
    \ifx#1\textstyle
      \mkern#20mu
    \else
      \ifx#1\scriptstyle
        \mkern#26mu
      \else
        \mkern#26mu
      \fi
    \fi
  \fi
}

\makeatother

\newcommand{\xcm}{\epsfxsize=3.1cm}
\newcommand{\fig}[1]{\epsfbox}
% \newcommand{\bp}{\begin{minipage}{3.1cm}}
% \newcommand{\ep}{\end{minipage}}

\newtheorem{Definition}{Definition}[]
%\newmdtheoremenv{Theorem}{Theorem}[]
\newtheorem{Theorem}{Theorem}[]
\newtheorem{Lemma}{Lemma}[]
\newtheorem{Proposition}{Proposition}[]
\newtheorem{Corollary}{Corollary}[]
\newtheorem{Remark}{Remark}[]
\newtheorem{Conjecture}{Conjecture}[]

\makeatletter 
\renewcommand{\thefigure}{\@arabic\c@figure}
\makeatother
\usepackage{xr}

\externaldocument{Draft_penultimate}

\begin{document}

\title{Supporting Information for: Universal set of Observables for the Koopman Operator through Causal Embedding}
\author{G Manjunath and A de Clercq\\
{\small \it Department of Mathematics \& Applied Mathematics, University of Pretoria, Pretoria 0028} \\ {\small Email: manjunath.gandhi@up.ac.za \& decle029@umn.edu} \\}
\date{}
\date{}
\maketitle

This supporting information document is meant to serve three purposes: (i). to provide a detailed motivation and include an additional explanation of some ideas in the main article (ii). mathematical proofs of results claimed in the main article (iii). more details of the methods of the numerical results in the main article.

\section{Detailed Motivation}
Performing experimental measurements 
on  biological, physical, and artificial systems to obtain 
a more informative dynamical model has been well established in modern-day science. The practical purpose of obtaining high fidelity models is not only for minimizing the point-wise prediction error. Although better prediction indeed helps both manage better service interruptions and resource management, models that have long-term consistency  in their time-averaged characteristics
are useful for understanding the response of statistical properties of the perturbations of models that have potentially distinguished applications, for instance in climate science (e.g., \cite{majda2010high,ghil2020physics} and references therein). 
Although attempts made to build model equations from complex data such as sunspot cycles date back to the 1920s in \cite{yule1927vii}, a major breakthrough came through the Takens embedding theorem \cite{takens1981detecting} that
made state-space reconstruction a valuable tool for any kind of analysis.  Often experimental measurements from such systems are not directly the system's states, but a univariate time-series whose span is smaller than the underlying system's dynamics.  Takens embedding theorem, under some generic conditions,  establishes the learnability of a system that is created out of concatenating sufficiently large previous observations of a dynamical system into a vector (called a delay coordinate map). The system determined by such vectors is equivalent (topologically conjugate) to the system from which the observed time-series was derived.  While this topological result ensures an alternate representation of the underlying system, the quality of the representation remains sensitive to various parameters, and importantly stability. The main reason why stability affects performance is that there is no guarantee that the embedding in the reconstruction space does  not lead to globally dissipative dynamics although it is locally conservative. In the absence of global dissipativity, small errors lead to major errors before predictions can totally fail (schematic in Fig.~\ref{Fig_Takens}). 
Theoretical conditions under which the geometry of the attractor can be preserved (e.g.,\cite{eftekhari2018stabilizing}) are not adequate for attractor reconstruction in practice.
Moreover, global approximation techniques that find a single map to fit the data often work well only when the data can be fit by functions with low functional complexity, i.e., functions that have relatively fewer oscillatory graphs. An example, where neural networks often failing to learn a representation of the underlying attractor   is illustrated in   \cite{principe1992prediction} and in this article in Fig.~\ref{Fig_learnt_takens}.

Although not theoretically assuring accurate reconstruction as in Takens delay embedding, the more recent approaches in machine learning that employ the reservoir computing techniques \cite{jaeger2004harnessing, lu2018attractor} and data-driven algorithms based on Koopman's theory \cite{koopman1932dynamical,budivsic2012applied}  have shown greater quantitative accuracy in forecasting dynamical systems than what has been achieved through delay-embedding alone. The prominent idea in reservoir computing approaches is to transform the available temporal data into another higher dimensional space through the dynamics of a driven dynamical system (e.g., \cite{Manju_ESP,manjunath2020stability}).  Often, the driven dynamical system is an arbitrary network with recurrent connections, and its state is post-processed to fit the prediction task by adopting very few connections in the network rather than rigging the entire connectivity in the network from data.  Although there is no theoretical framework that guarantees the existence of a post-processing function, it is approximated typically by linear regression. These methods often give good short-term predictions for some specific systems \cite{jaeger2004harnessing, lu2018attractor}, albeit with adaptations. Unfortunately, the well-cited literature barring perhaps \cite{herteux2020breaking,Manju_IEEE,wu2018statistical, faranda2019boosting} in this field does not consider a broad class of chaotic dynamical systems where they fail.

The idea behind forecasting using 
Koopman's theory is to transform the observed data generated from an unknown dynamical system into a new space or a new coordinate system so that the induced dynamics in the new coordinate system can be approximated by a more simple set of equations, for instance by a linear system of equations. The well-known algorithms like Dynamic Mode Decomposition and Extended-Dynamic Mode Decomposition \cite{schmid2010dynamic,williams2015data} involve a priori determination of a Koopman invariant subspace obtained by a handpicked choice of observables. If the observables and the Koopman invariant subspace they determine are not adequate enough to capture some essential dynamics, the resultant approximation to the Koopman operator is poor, and hence the qualitative behavior like the phase portrait of the forecasted data veers down to a different one in a very few time-steps.  In any case, linear systems in finite dimensions cannot be topologically equivalent to a chaotic map, and hence such methods are limited in performance while modeling complex/chaotic data.

Balancing model complexity with accuracy has recently led to the idea of sparse representations where the vector field of a differential equation that describes the dynamics is approximated by an optimal linear combination of a very few functions (determined through sparse regression) that are `possibly nonlinear'  from a chosen library. Such representations have yielded not only accurate prediction results for some classes of chaotic systems but can be used to build simpler model equations from data e.g., the popular SINDy algorithms in \cite{brunton2016discovering, champion2019data}. However, if the library is not big enough to span the vector field, then long-term consistent modeling is not possible for complex data. 
 Besides requiring data obtained at high sampling rates since derivatives are approximated another crucial issue is that the method requires full knowledge of the underlying state space variables that govern the underlying system which is rare in many real-world scenarios.

Library free approaches such as combining delay-embedding and Koopman theory allow to build intermittently forced linear models \cite{brunton2017chaos,
 champion2019discovery} combining delay embedding and Koopman theory.  However, unlike Takens embedding, the existing reservoir computing or data-driven methods do not ensure finding an equivalent (conjugate) model of the data. In the case of Koopman analysis,  by capturing a portion of the spectrum like the eigenvalues \cite{budivsic2012applied,korda2020data} of the operator when they exist, only a topological semi-conjugacy can be obtained (see Section~\ref{Sec_Prelim}). Although there is  talk about finding  faithful representations in the Koopman framework leading to topological conjugacy \cite{mezic2020koopman}
there is currently no framework that guarantees existence or a method to find them.

On the positive side, through  generalization of Takens embedding  (e.g., \cite{sauer1991embedology,stark1999delay,robinson2010dimensions,gutman2018embedding}, and notably \cite[Theorem 1]{gutman2018embedding}, generic continuous-time-lagged observations of attractors of homeomorphisms on compact spaces have been shown  sufficient for attractor reconstruction. On preserving the geometry of the reconstructed attractor, some recent theoretical and empirical studies suggest that it may be possible to construct delay embeddings that are not only homeomorphic or diffeomorphic to the original system but nearly isometric \cite{sauer1991embedology, baraniuk2009random}.  Isometric embeddings for instance obtained by the Nash embedding theorem help preserving local neighborhoods of points on the attractor, a geometric feature desired during forecasting and reconstructing attractors. More recent studies are suggestive that with large enough time-lagged observations
\cite{eftekhari2018stabilizing, yair2017reconstruction}, isometric embeddings can be obtained.

With the objective of getting close to an isometric embedding through large, perhaps infinite, time-lagged observations we employ a driven (dynamical) system with the ability to causally embed a dynamical system so as to transform the entire left-infinite history of the temporal data onto just a pair of points (in practice, a pair of finite-dimensional vectors) in a different space. 
Remarkably, we find that such infinite-delay maps induced by such driven systems determine the observables for the Koopman operator of the inverse-limit system (e.g., \cite{ingram2011inverse}) of the dynamical system generating the data.  
This, in conclusion, gives a topological conjugacy to the dynamical systems arising 
from the discretization of ordinary differential equations, or more generally homeomorphisms. The conjugacy determines a finite faithful representation raised in \cite{mezic2020koopman}. Remarkably, the observables in the faithful representation are universal, in the sense that they remain unchanged and all the driven system does is to restrict the observables to a new subspace of the inverse-limit system once the underlying dynamical system is changed.  This not only solves a highly pursued problem in finding a set of observables for the Koopman operator needed for theoretically precise forecasting dynamical systems but also prevents expert human intuition or machine optimization needed for choosing observables or a library of functions specific to a dynamical system.

Furthermore, we obtain \emph{exact} equations from the data of a dynamical system, the data could be the actual states or obtained as observations in Takens delay-embedding. The equations are determined after learning a map from the data and have several auxiliary variables. Since the driven system that appears in these equations has mathematically proven stability properties \cite{manjunath2013echo, grigoryeva2018echo}, the forecasting results show greater stability to noise and parameters although it introduces auxiliary variables. Empirical evidence shows that a linearity measure like a Pearson correlation coefficient points at reduced functional complexity while learning the dynamics in the state space of the driven system.  
We demonstrate long-term topological 
consistency (attractor is learnt)
and statistical consistency (density of the orbits) of the models obtained through data from standard benchmark chaotic systems and also on systems that show intermittency -- Type I intermittency is a feature of transition to turbulence and convection in confined spaces \cite{pomeau1980intermittent}, seismic data \cite{bottiglieri2007off} and anomalous diffusion in biology \cite{klages2013weak} and are extremely difficult to model from data.  In summary, we learn the action of the Koopman operator on observables of the inverse-limit system of the underlying dynamical system, and these observables are determined by a driven system and data. 
This in turn helps obtain exact equations from data and thus robust high-fidelity models.

%that has several auxiliary variables but has the advantage of long-term topological and statistical consistency while forecasting dynamical systems.

\section{Preliminaries} \label{Sec_Prelim}

%We map the input $\{u_n\}$ onto a different space for predicting.  If the other space is higher-dimensional data often unfolds.  After mapping $\{u_n\}$ to $\{x_n\}$, we want to know that $\{x_n\}$ has all the information that can generate $\{u_n\}$, but nothing more -- to illustrate why we are interested in this for forecasting, and contextualize our methodology we recall Takens delay embedding and the idea of learning the Koopman's operator.

A dynamical system in this work is a tuple $(U,T)$ where $U$ is a compact metric space and $T: U \to U$ is a function (that is not necessarily continuous). Henceforth, we consider only surjective dynamical systems, i.e., systems for which $T$ is surjective. In fact, if $A$ is any invariant set, i.e., if $T(A):=\cup_{u\in A} Tu = A$,  then $T$ is surjective on $A$, so if $T$ is not surjective to start with we can restrict the non-transient dynamics to an invariant set.  We call any sequence $\bar{u} = \{u_n\}_{n\in \mathbb{Z}}$ that obeys the update equation, $u_{n+1}=Tu_n$ where $n \in \mathbb{Z}$ as an orbit of $T$. Subsets of $\mathbb{R}^d$ are endowed with the standard topology induced from 
$\mathbb{R}^d$.  The $n$-fold composition of a self-map $f$ on any space is denoted by $f^{(n)}$. The $\omega$-limit set $\omega(x;f)$ of a point $x$ is the collection of limit points of the sequence $\{x,f(x),f^{2}(x),\ldots\}$; note that when $f$ is defined on a compact space, $\omega(x;f)$ is always non-empty. Throughout, if $F: X \to Y$ is any function, and $Z$ is a subset of $X$, for brevity of notation we denote the restriction of $F$ on $Z$, $F_{|Z}$ by just $F$ when the context is clear.

 A core concept in dynamical systems theory is the notion of equivalent dynamical systems.
Finding an equivalent dynamical system to $(U,T)$ means finding another dynamical system $(V,S)$ so that there exists a homeomorphism $\phi:U \to V$ with the property that  $\phi \circ T=S\circ \phi$. Such a map $\phi$ is called a  conjugacy and we say that $(V,S)$ is conjugate to $(U,T)$. If we relax the condition on $\phi$ where, instead of having a homeomorphism, we only require $\phi$ to be continuous, then we call $\phi$ a semi-conjugacy and say the systems are semi-conjugate. A system being conjugate to the original means that there is a one-to-one correspondence between the two systems, whereas a semi-conjugacy when it is many-to-one mapping provides a coarse-grained description of the original system. When  $(V,S)$ is semi-conjugate to $(U,T)$ then it is customary to say that $(V,S)$ is a factor of $(U,T)$ or that $(U,T)$ is an  extension of $(V,S)$. When the underlying spaces are clear, we just say $S$ is a factor of $T$ or $T$ is an extension of $S$.

\section{Takens Delay Embedding}
The Takens Delay Embedding Theorem establishes that concatenating a sufficiently large number of previous observations of a dynamical system into a vector (schematic in Fig. ~\ref{Fig_Takens}) can generate a map between the vectors under some conditions. We recall the theorem from \cite{takens1981detecting} below.

\begin{Theorem} 
	[\bf Takens Embedding Theorem (adopted from \cite{takens1981detecting}] \label{Thm_Takens}
            Let $W$ be a compact manifold of dimension $m$, and $d\ge m$ so that $2d$ is an integer. It is a 
            generic property for the pair $(T, \theta)$,  where $T:W \to W$ is
            a smooth diffeomorphism, and $\theta:W \to \mathbb{R}$ a smooth function, the map $\Phi_{2d,\theta}:W \to \mathbb{R}^{2d+1}$ defined on $W$ by 
            $\Phi_{2d,\theta}(w) := (\theta(T^{-2d}w)\ldots,\theta(T^{-1}w),\theta(w))$
            is a diffeomorphic embedding; by `smooth' we mean at least $C^2$. Consequently, there exists a map $F_\theta: \Phi_{2d,\theta}(W) \to \Phi_{2d,\theta}(W)$ defined by $$F_\theta: (\theta(T^{-2d}w),\ldots,\theta(T^{-1}w),\theta(w)) \mapsto 
            (\theta(T^{-2d+1}w),\ldots,\theta(T^{-1}w),\theta(Tw))$$
           so that $(W,T)$ is topologically conjugate to 
            $(\Phi_{2d,\theta}(W), F_\theta)$. 
\end{Theorem}

The map $\Phi_{2d,\theta}$ is called the delay-coordinate map. 
There are now various relaxations on the hypotheses required for such a map $F_\theta$ to exist in recent works using the central idea of using a delay-coordinate map 
$\Phi_{2d,\theta}$ to embed an attractor. 
The map $F_\theta$ can be learnt through a finite segment of an orbit of $F_{\theta}$, and any such learnt map $\widetilde{F}_{\theta}$ would also accept arguments that are outside $\Phi_{2d,\theta}(W)$ as well. 
The above theorem does not guarantee that the set $\Phi_{2d,\theta}(W)$ is an attractor of the system defined by the map $\widetilde{F}_{\theta}$ (See Fig.~\ref{Fig_Takens}). 
When $\Phi_{2d,\theta}(W)$ is not an attractor, as schematically illustrated in Fig.~\ref{Fig_Takens}),  a small amount of error  could leave an iterate of $\widetilde{F}_{\theta}$  outside $\Phi_{2d,\theta}(W)$. This could then result in the future iterates under $\widetilde{F}_{\theta}$ moving far away from $\Phi_{2d,\theta}(W)$ resulting in erroneous prediction.

To demonstrate such fragility in forecasting, we consider the evolution of the image of the time-series obtained by an observation of the Lorenz system under the delay-coordinate map with a delay equal to $10$. 
Specifically, if $(w_n)$ denote the samples of the Lorenz System (as in \eqref{Seq_Lor}), with $(w_{n,x}),(w_{n,y}),$ and $(w_{n,z})$ denoting the $x,y,$ and $z$ coordinates of $(w_n),$
then we consider the observation 
$\theta_n=\theta(w_n)=\frac{1}{10}(\sin(0.1w_{n,x})+\sin(0.1w_{n,y})+\sin(0.1w_{n,z})).$
The map $F_\theta$ is thus the mapping $(\theta_{n-10},\cdots,\theta_{n-1}) \mapsto (\theta_{n-9},\cdots,\theta_{n})$, which we approximate using a feedforward neural network.
To illustrate the failure in forecasting after having learnt $\widetilde{F}_{\theta}$, we plot the three main principal component of the iterates $\widetilde{F}_{\theta}$ (in blue), and the 
three main principal component of the delay coordinates $\Phi_{2d,\theta}(T^{(k)}w)$ (in red) in  Fig.~\ref{Fig_learnt_takens}).

\title{Figure was here}
% \begin{center}
% \begin{figure}[t]
% \centering
% \includegraphics[width=16cm]{Fig_Takens.png}
% \caption{Schematic of employing the Takens embedding and the consequence of how iterates of $\widetilde{F}_\theta$ could slip out of $\Phi_{\theta,2d}(W)$.}
% \label{Fig_Takens}
%  \end{figure}
% \end{center}

In practice, data-driven algorithms and other machine learning algorithms outperform Takens embedding often.

\title{Figure was here}
% \begin{center}
% \begin{figure}[t]
% \centering
% \includegraphics[width=16cm]{Fig_F_smart.png}
% \caption{Illustration of the failure to learn $\widetilde{F}_{\theta}$ accurately. The map $\widetilde{F}_{\theta}$ was constructed 
% by actually learning $(\theta_{n-2d-1}, \cdots \theta_{n-1}) \mapsto \theta_n$ through a Vanilla feedforward neural network;  plotted are the principal component of trajectories of $\widetilde{F}_{\theta}$ of randomly initialised points (in blue), and the 
% three principal component of the delay coordinates $\Phi_{2d,\theta}(T^{(k)}w)$ (in red). The data and observation $\theta$ is identical to that in the \cite[Fig.~\ref{Fig_1}C]{Main_article}, and a delay of 10 was used. }
% \label{Fig_learnt_takens}
%  \end{figure}
% \end{center}

%\ednote{(I strictly did not approximate $F_\theta$ using NN's. I learnt a map $F: (x_{n-d}, \cdots x_{n-1}) \mapsto x_n$ then constructed $F_\theta$ using $(\pi_2,\pi_3\, \cdots, F)(x_{n-d}, \cdots x_{n-1})=(x_{n-d+1}, \cdots x_{n})$. Please let me know if I should explain this thoroughly. The vanilla approximation using NNs is Fig\_F\_stupid.png) M: Please see the caption now.}
%Stability is desirable while modeling because rarely experimental measurements determine the exact values, so a physically meaningful feature of a mathematical model should exhibit persistence under small perturbations. Incidentally, the USP always satisfies this robustness to noise, that is small perturbations of inputs lead to small changes in the solution $\{x_n\}$ (see Theorem **). Another important criterion is that when a finite segment of the temporal input $(u_{n-m}\ldots,u_{n-2},u_{n-1})$ is fed into the system instead of the entire-left input $\cev{x_n}$, the system at time $n$ should be closer to the state $x_n$, and this closeness should have some uniformity for all inputs and data lengths $m$. Remarkably, USP is equivalent to the uniform attraction property (UAP) that ensures that not only do all states approach $X_U$, but they approach specific trajectories (solutions) within $X_U$, and they approach at a rate which does not fall below a certain threshold (see Theorem ***).

%This work is about forecasting observed trajectories from dynamical systems without recourse to the underlying equations of motion.

%\title{A Theory for Precise Forecasting: Koopman's Operators and Causal Embedding}

%\author{G. Manjunath \\{\small \it Department of Mathematics \& Applied Mathematics, University of Pretoria, Pretoria 0028} \\ {\small Email: manjunath.gandhi@up.ac.za } \\}
\date{}
\date{}
\maketitle

\section{Koopman Operator and Forecasting} \label{Sec_Koopman}

Given a dynamical system $(U,T)$, and a vector space $V$ of observables $f$ whose domain is $U$, the operator $\mathcal{K}: V \to V$ so that 
$\mathcal{K} f = f \circ T$ holds is called the Koopman operator.  If one knows the action of the Kooopman operator on an observable $f$, then since 
$\mathcal{K} f(u) = f(T(u))$ and likewise $\mathcal{K} f(Tu) = f(T^{(2)}u)$, one can forecast the observed values $(f(Tx), f(T^{(2)}u), f(T^{(3)}u),\ldots)$. Often 
for an analysis or approximation of the Koopman operator, the space of observables $V$ is a Banach space or a Hilbert space over the field of complex numbers. In this case a complex number $\lambda \in \mathbb{C}$ and an associated $\phi \in V$ is called an eigenvalue and eigenfunction respectively if $\mathcal{K} \phi = \lambda \phi$. Thus, when $\phi$ is an eigenfunction associated with $\lambda$, then it follows from \cite{schmid2010dynamic,tu2013dynamic,williams2015data} that the following diagram commutes:

\begin{equation}  \label{comm_phi}
%    \[ 
    \psset{arrows=->, arrowinset=0.25, linewidth=0.6pt, nodesep=3pt, labelsep=2pt, rowsep=0.7cm, colsep = 1.1cm, shortput =tablr}
 \everypsbox{\scriptstyle}
 \begin{psmatrix}
U & U\\%
 \phi(U) & \phi(U)
 %%%
 \ncline{1,1}{1,2}^{T} \ncline{1,1}{2,1} <{\phi}
 \ncline{1,2}{2,2} > {\phi}
 \ncline{2,1}{2,2}^{u\mapsto\lambda u}
 \end{psmatrix}
% \]
\end{equation} 
and hence the dynamical system $(\phi(U), F_\lambda: u \mapsto \lambda u)$ is topologically semi-conjugate to $(U,T)$. So every eigenvalue captures a coarse-grain description of $(U,T)$, and the representation gains physical meaning when the eigenfunctions are non-constants. Finding non-constant eigenfunctions is difficult even while the map $T$ is known. However, they can be determined from data (e.g., \cite{schmid2010dynamic,williams2015data}). One of the central ideas in using Koopman's theory for forecasting dynamical systems is to obtain a collection of observables $\mathbf{f}=\{f_1,\ldots,f_N\}$ so that the resultant dynamics from $\mathbf{f}(u) \mapsto \mathbf{f}(Tu)$ can be approximated by a map, say $F$  with a lower functional complexity exploiting the fact that $\mathcal{K}$ is linear in $V$. The choice of the $L^p$ space considered determines the spectrum \cite{eisner2015operator} of the Koopman operator $\mathcal{K}$, and when it is a Hilbert space, and the span of the observables is invariant under $\mathcal{K}$, one can capture a portion of the spectrum of $\mathcal{K}$ through $F$. More specifically eigenvalues of $F$ would belong to the spectrum of  $\mathcal{K}$ (e.g., \cite{williams2015data,korda2020data}). The interesting feature of such approximation is that $F$ (in fact its matrix representation) can be derived from the observed data $\{f_i(T^{(k)}u)\}_{0\le k \le m, 1\le i\le N}$. However, there are many potential issues: (i).  the span of arbitrarily chosen observables is not necessarily invariant under the Koopman operator (ii) one does not know how to expand the set of observables to capture more eigenvalues and concomitantly retaining invariance under the Koopman operator (iii). Above all, there is no guarantee that the Koopman operator has relevant eigenvalues for approximating the salient feature of the dynamics of the underlying system.  For example, non-isolated eigenvalues and/or continuous spectra of the Koopman operator are a feature of many complex systems especially those which exhibit chaos (e.g., \cite{korda2020data}).

%In any case, since $S$ may not be invariant under $\mathcal{K}$, it is an open question on how to choose observables that gives an invariant subspace that good approximation to such systems.

Sparse identification of nonlinear dynamical systems (SINDy)
\cite{brunton2016discovering,champion2019data} overcomes the difficulty with linear approximations considerably -- rather than aiming at capturing eigenvalues, an optimal linear combination of elements in a pre-determined library is determined from the observed data to approximate the vector field of the unknown dynamical systems. The main disadvantage of this method is that the method requires full knowledge of the underlying state space variables that govern the underlying system which is rare in many real-world scenarios. Further, when the library is not big enough to span the vector field, then long-term consistent modeling is not possible for complex data.  For instance, if the library comprises polynomials and rational functions good approximations are obtainable only if the underlying system has polynomial and rational nonlinearities, and can fail when the underlying system has a nonrational and nonpolynomial term \cite{pan2018long}. Also since time-derivatives are found, one needs data at a very high sampling rate with lower noise.

Library free approaches such as combining delay-embedding and Koopman theory allows to build intermittently forced linear models \cite{brunton2017chaos,champion2019discovery}.  However, unlike Takens embedding, the existing reservoir computing or data-driven methods do not guarantee an equivalent (conjugate) model of the data.  Recently, Igor Mezi\'{c} proposed the idea \cite{mezic2020koopman} of learning the action of the Koopman operator on a finite set of observables $\mathbf{f}= (f_1,f_2,\ldots,f_N)$ so that we can find a map $G$ so that $\mathbf{f}$ is a topological conjugacy between $T$ and $G$, i.e., we determine $(\mathbf{f},G)$ so that the following diagram commutes: 
\begin{equation}  \label{comm_h2}
%    \[ 
    \psset{arrows=->, arrowinset=0.25, linewidth=0.6pt, nodesep=3pt, labelsep=2pt, rowsep=0.7cm, colsep = 1.1cm, shortput =tablr}
 \everypsbox{\scriptstyle}
 \begin{psmatrix}
U & U\\%
 \mathbf{f}(U) & \mathbf{f}(U).
 %%%
 \ncline{1,1}{1,2}^{T} \ncline{1,1}{2,1} <{\mathbf{f}}
  \ncline{2,1}{2,2}^{G}
 \ncline{1,2}{2,2} > { \mathbf{f}}
 \end{psmatrix}
% \]
\end{equation} 
When such commutativity holds,  the author in \cite{mezic2020koopman} calls $(\mathbf{f},G)$ a finite faithful representation of the dynamical system. Finding a finite faithful representation would avoid approximating the Koopman operator but would instead enable to learn the action of the Koopman operator on the set of observables determined by $\mathbf{f}$. In essence, we can abandon the whole idea of approximating the spectrum of the Koopman operator while forecasting.
In this work we approach to solve the problem of finding a finite faithful representation of the inverse-limit system of a dynamical system \cite{ingram2011inverse} which is an extension of the original system $(U,T)$; extensions always contain the spectrum of their factors \cite{eisner2015operator}. Also, we observe that employing observables from inverse-limit systems would help us forecast a system that is greatly sensitive to the distant past much more than the immediate past.

\section{Driven Systems: Some Basics}

A driven system comprises an input metric space $(U,d_U)$, a compact metric (state) space $(X,d)$ and a {\bf continuous function} $g: U \times X \to X$. For brevity, we refer to $g$ as a driven system with all underlying entities quietly understood.

A bi-infinite sequence $\bar{u}=\{u_n\}_{n \in \mathbb{Z}} \subset U$ which we call an input, induces a sequence of self-maps $\{g(u_n,\cdot)\}_{n\in \mathbb{Z}}$ defined on $X$, and  the dynamics on $X$ generated through this sequence os self-maps is given by the  update equation $x_{n+1}=g(u_n,x_n)$.

Given a driven system $g$ and an input $\bar{u}$, we call a sequence $\{x_n\}$ a solution if it satisfies  $x_{n+1}=g(u_n,x_n)$ for all $n\in \mathbb{Z}$. We denote a solution obtained by $\bar{u}$ as $\{x_n(\bar{u})\}$.

We next identify a subspace $X_U$ of $X$ that contains all possible solutions. To realize such a subspace of a driven system $g$, we define 
the \textit{reachable set} of the driven system $g$ to be the union of all the elements of all the solutions, i.e., 
$$X_U :=\Big \{x \in X:  x = x_k \mbox{ where  $\{x_n\}$  is a solution for some  $\bar{u}$} \Big \}.$$

For example, when $U=[0,1]$ and $X=[0,1]$, for the driven system 
$g(u,x) := \frac{ux}{2}$ regardless of any input sequence in $U$, $x_n\equiv 0$ for $n \in \mathbb{Z}$ is the only solution of $g$, and hence the reachable set $X_U=\{0\}$. Also for example, when  $U=[0,1]$ and $X=[0,1]$, $g(u,x)=x$ regardless of any input sequence in $U$, all constant sequences contained in $X$ are solutions of $g$, and hence the reachable set $X_U=X$.

In both these examples, we have hit somewhat extreme cases that are not useful, and we would need the reachable set to be relatable to the temporal input; we will make precise of what we mean by ``relatable" later in Definition~\ref{Def_UC}. To restrict to our attention to more useful cases, we  say $g$ is SI-invertible if $g(\cdot,x):U \to X$ is invertible for all $x$, i.e., if the current state $x_n$ and the future state $x_{n+1}$ are given, then the current input $u_n$ can be uniquely determined, since  the inverse of   $g(\cdot,x_n)$ exists. For instance, consider a recurrent neural network (RNN) with  $N$ artifical neurons and $U \subset \mathbb{R}^N$
and $X=[-1,1]^N$ (the cartesian product of $N$ copies of $[-1,1]$) given by
\begin{equation} \label{Seq_RNN}
	g(u,x) = (1-a)x + a\overline{\tanh}(Au + \alpha Bx),
\end{equation}
where $A$ is a $N \times N$ matrix with input connections called the input matrix. The matrix $B$ is also of dimension $N \times N$ representing the strength of the interconnections called a reservoir matrix, and $a$ and $\alpha$ are real-valued parameter that are normally called  leak rate and scaling of the reservoir 
 $B$ respectively and $\overline{\tanh}(*)$ is (the nonlinear activation) $\tanh$ performed component-wise on $*$.  The RNN accepts inputs as vectors with $N$ elements, and if an input $v_n$ is of dimension $K<N$, it can be embedded into $\mathbb{R}^N$, for instance one can pad $N-K$ zeroes to obtain an input of dimension $N$ i.e., $v_n \mapsto (v^1_n,v^2_n,\ldots,v^K_n,0,0,\ldots 0) = u_n$. Since  given $x_{n+1}$ and $x_n$, we can recover $u_n$ by
\begin{equation} \label{eqn_SI_RNN}
u_n := A^{-1}\bigg(\overline{\tanh}^{^{-1}}\frac{1}{a}\Big(x_{n+1}-(1-a)x_n\Big) \bigg) - \alpha B x_n
\end{equation}
$g$ in \eqref{Seq_RNN} is SI-invertible.

We now describe the set of all solutions of $g$ for a given input $\bar{u}$ in the nonautonomous dynamical systems setting (e.g., \cite{kloeden2011nonautonomous, Manju_ESP}). Suppose a driven system $g$ has been fed input values $u_{m},u_{m+1},\ldots,u_{n-1}$ starting at time $m$. Then the map $g$ transports a state-value $x\in X$ at time $m$ to give a state-value $g_{u_{n-1}}\circ \cdots \circ g_{u_{m}}(x)$ at time $n$.

Formally, for every choice of  $\bar{u} = (\ldots,u_{-1},u_0,u_1,\ldots)$, we define for all pair of integers $m\le n$, the function that `transports' a system state at $x$ at time $m$ through the inputs $u_{m}, u_{m+1}, \dots u_{n-1}$ 
to the state at time $n$ given by a composition-operator called a process by several authors (e.g.,\cite{kloeden2011nonautonomous}) by the map $\phi_{\bar{u}}: \mathbb{Z}^2_{\ge} \times X \to X$, where $\mathbb{Z}^2_{\ge} := \{ (n,m) : n \ge m, n,m \in \mathbb{Z} \}$ and
\begin{align} \label{eq_Process}
  \phi_{\bar{u}}(n,m,x) := \begin{cases}
        x \quad \quad \quad \quad \quad \quad \quad \text\quad \quad \quad \quad \quad \quad \:\text{ if $n=m$,}
        \\
        g_{u_{n-1}}\circ \cdots \circ g_{u_{m+1}} \circ g_{u_{m}}(x)  \quad \: \: \: \: \;\;\text{ if }  m< n.
        \end{cases}
\end{align}

Since $g(u,x):X \to X$,  it easily follows that $\phi_{\bar{u}}(n,m-1,X) \subset \phi_{\bar{u}}(n,m,X)$ (see \cite{Manju_ESP} \cite{manjunath2020stability}), and since $\phi$ is continuous in the variable $x$, these sets are all closed. Further, since $X$ is compact the nested intersection  \begin{equation} \label{SSeqn_xn}
X_n(\bar{u}) := \bigcap_{m<n} \phi_{\bar{u}}(n,m,X)
\end{equation}
is nonempty when $X$ is nonempty.

The set $X_n(\bar{u})$ denotes the set of all reachable states at time $n$ if the input is $\bar{u}$. It is a well known result in nonautonomous dynamical systems literature (e.g., \cite{kloeden2011nonautonomous,Manju_ESP,manjunath2014dynamics}) that  $x\in X_n(\bar{u})$ if and only if there is a solution $\{x_k\}$ of $g$ with input $\bar{u}$ so that $x_n = x$. Thus in the special case where we have a topological contraction in \eqref{SSeqn_xn} where $X_n(\bar{u})$ is a singleton subset of $X$ for each $n$ and $\bar{u}$, then we have exactly one entire-solution for each $n$.

\section{Universal Semi-Conjugacy and the Causal Embedding Theorem}
Since the state $x_{n+1}$ of a driven system $g$ at time $n+1$ depends on both the input value $u_n$ and the state value $x_n$ at time $n$, we consider a question on whether the ``complexity" in a solution is exclusively contributed by the input or if the sequence of maps $\{g_{u_{n}}(\cdot)\}_{n\in \mathbb{Z}}$ also contributes to the complexity in the solution. In the case of autonomous systems on $X$, i.e., for a self-map $f:X \to X$, the general feature of complex dynamics is that the sequence $\{f^{(n)} \}$ of self-compositions is not equicontinuous. So the question in the case of non-autonomous system $\{g_{u_{n}}(\cdot)\}_{n\in \mathbb{Z}}$ is not just about equi-continuity, but something more general since the input is involved.

To illustrate the idea behind the question, consider an example of a numerical simulation of a solution of a RNN (as in \eqref{Seq_RNN}) with two different parameter sets (see Fig.~\ref{Fig_Complexity}).  A coordinate of a solution is plotted in red and blue for the two parameters and the input sequence is shown in black. As it may be observed, the coordinate of the solution shown in red seems to just follow the input in its temporal variation, while that in blue has wild behavior with an oscillatory envelope. We say that the driven system has introduced new additional complexity to the solution indicated in blue that was not there in the input. In order to mathematically describe the scenario of  $\{g_{u_{n}}(\cdot)\}$ not adding on to the complexity to the solution we consider  a definition in Definition~\ref{Def_UC}. 
We denote $\cev{u}^{n}:=(\ldots,u_{n-2} ,u_{n-1})$ and $\overleftarrow{U}$ denote all the left-infinite sequences in $U$. 
Symbolically we let $\cev{u}^{n}v:=(\ldots,u_{n-2} ,u_{n-1}, v)$ denote the input up to time $n$ with $v \in U$ being the input value at time $n$. This introduction of a new input at time $n$ can be described by a mapping $\sigma_v:   
\cev{u}^{n} \mapsto \cev{u}^{n}v$. We would like to talk of continuity of functions defined  on the space of left-infinite sequences contained in $U$ hence we adopt the notation: if $Y$ is a metric space then we denote the product space
$\overleftarrow{Y} := \prod_{i=-\infty}^{-1} Z_i$
 where $Z_i \equiv Y$ and equip this space with the product topology, and consider the definition of the universal semi-conjugacy.

 \title{Figure was here}
% \begin{center}
% \begin{figure}[t]
% \centering
% \includegraphics[width=10cm]{Fig_Complexity_label.png}
% \caption{A coordinate of a simulated solution $(x_0,x_1,\ldots,x_{5000})$ of a RNN in \eqref{Seq_RNN}   plotted in red (with parameter-set $a=1$, $\alpha=0.99$) and blue ($a=0.99$ and $\alpha=1.05$) while the choice of matrices $A$ and $B$ are the same.}
% \label{Fig_Complexity}
%  \end{figure}
% \end{center}

\begin{Definition} \rm \label{Def_UC}
 Given a driven system $g$, we  call a continuous and surjective map $h : \overleftarrow{U} \to X_U$ a universal semi-conjugacy if the following diagram commutes for all $v \in U$:
	\begin{equation}  \label{Scomm_h}
%    \[ 
    \psset{arrows=->, arrowinset=0.25, linewidth=0.6pt, nodesep=3pt, labelsep=2pt, rowsep=0.7cm, colsep = 1.1cm, shortput =tablr}
 \everypsbox{\scriptstyle}
 \begin{psmatrix}
 \overleftarrow{U} & \overleftarrow{U}\\%
 X_U & X_U.
 %%%
 \ncline{1,1}{1,2}^{\sigma_v} \ncline{1,1}{2,1} <{h}
 \ncline{1,2}{2,2} > {h}
 \ncline{2,1}{2,2}^{g(v,\cdot)}
 \end{psmatrix}
% \]
\end{equation} 	
	\end{Definition}

We next say a driven system $g$ has the unique solution property (USP) if for each input $\bar{u}$ there exists exactly one solution. In other words, $g$ has the USP if there exists a well-defined solution-map $\Psi$ so that $\Psi(\bar{u})$ denotes the unique solution obtained from the input $\bar{u}$.  In our context, the USP is equivalent to saying $g$ is a topological contraction, i.e., each $X_n(\bar{u})$ is a singleton subset of $X$ for all $n \in \mathbb{Z}$ and all $\bar{u}$. The USP notion is independent of SI-invertibility.

The left-finite sequence $\cev{u}:= (\ldots,u_{-2},u_{-1})$ belonging to $\overleftarrow{U}$
can also be used to equivalently define USP. Consider $\Pi_{n=-\infty}^{n=+\infty} U_i$, where $U_i \equiv U$. 
Since  any left-infinite portion of an element of  
$\Pi_{n=-\infty}^{n=+\infty} U_i$
is an element of $\overleftarrow{U}$,
we can express the definition of the USP more succinctly by  denoting $\mathcal{E}(\cev{u}) := X_0(\bar{u})$, and then say that $g$ has the USP if $\mathcal{E}(\cev{u})$ is a singleton subset for all $\cev{u} \in \overleftarrow{U}$. We call $\mathcal{E}$ the encoding function.

We denote the
collection of all nonempty closed subsets of $X$ by $\mathsf{H}_X$. On $\mathsf{H}_X$ we employ the Hausdorff metric defined by
$d_H(A,B) := \max (dist(A,B) , dist(B,A)) = \inf\{ \epsilon : A
\subset B_\epsilon(B) \: \: \& \: \:B\subset B_\epsilon(A) \}$, where
$B_\epsilon(A): = \{ x \in X : d(x,A) < \epsilon\}$ is the open
$\epsilon$-neighborhood of $A$.
 We could treat the encoding function $\mathcal{E}(\cdot)$  and describe its continuity by treating it either as a multivalued function of $\cev{u}$ or as a regular function taking values in the space of nonempty compact subsets $\mathsf{H}_X$.  In particular when $X$ is a compact metric space the continuity notions become equivalent (e.g., \cite[Theorem 1, p. 126]{berge1997topological}).   Henceforth, we consider the continuity of $\mathcal{E}(\cdot)$ a $\mathsf{H}_X$-valued function $f$, with of course $\mathsf{H}_X$ being equipped with the Hausdorff metric. When $\mathcal{E}(\cdot)$ is a singleton subset of $X$ then it is always continuous as a set-valued function
  (e.g., \cite{manjunath2020stability,Manju_Nonlinearity}).  We define the subspace of $\mathsf{H}_X$ that contains the singleton subsets of $X$ by $\mathsf{S}_X$.

 We borrow the following facts from \cite[Section 3]{manjunath2020stability}: {\bf (F1).} If $\mathcal{E}(\cev{v})$ is a singleton subset of  $X$ then $\mathcal{E}(\cdot)$ is continuous at $\cev{v}$.  To state our theorem, we define the subspace of $\mathsf{H}_X$ that contains the singleton subsets of $X$ by $\mathsf{S}_X$, and define the mapping $i: (X,d) \to (\mathsf{S}_X,d_H)$  by $i(a) = \{a\}$. 
Clearly $i$ is invertible. Note that $i$ is an isometry since \begin{eqnarray*}d_H(i(a),i(b)) &=& \max \bigg( \sup_{a \in \{a\}} d(a,b), \sup_{b \in \{b\}} d(b,a) \bigg) \\& =& \max(d(a,b),d(b,a)) = d(a,b).\end{eqnarray*}

We recall results from \cite{Manju_IEEE, Manju_Nonlinearity} and state them in a way to suit the context here.

%Hencewhen $n=0$, we denote $\mathcal{E}(\cev{u}) = X_n(\bar{u})$ and call it the encoding of the left-infinite sequence $\cev{u}$.

\begin{Theorem}{{\bf (Universal Semi-Conjugacy Theorem.)}}
 \label{Thm_UCT}
            Let $g$ be a driven system.  Then $g$ induces a universal semi-conjugacy $h: \overleftarrow{U} \to X_U$ if  $g$ has the unique solution property (USP), i.e., for every input $\{u_n\}$ there exists exactly one entire-solution. Moreover, $h(\dots,u_{k-2},u_{k-1})=x_k$, where $\{x_n\}$ is the solution of $g$. 
            \end{Theorem}

{\bf Proof. } Let $h(\dots,u_{k-2},u_{k-1}):=x_k$, where $\{x_n\}$ is the solution for any input whose left-infinite sequence is
            $(\dots,u_{k-2},u_{k-1})$. Hence $h(\cev{u}^k) = \mathit{i}^{-1}(\mathcal{E}(\cev{u}^k))$. By definition of $\mathcal{E}$, we find (the required commutativity in the diagram in \eqref{Scomm_h}) through the deduction: 
\begin{eqnarray*}
g_v\circ \mathit{i}^{-1} \circ \mathcal{E}(\cev{u}) & = & \mathit{i}^{-1} \circ \mathcal{E}(\cev{u}v), \\
& =& \mathit{i}^{-1} \circ \mathcal{E}(\sigma_v(\cev{u})).
\end{eqnarray*}

It remains to be shown that $h$ is surjective and continuous. 
By definition of $X_U$,  it follows that $h(\overleftarrow{U})=X_U$ and hence $h$ is surjective. Also since
 $h = \mathit{i}^{-1} \circ \mathcal{E}$, and $\mathcal{E}$ is continuous and $\mathit{i}^{-1}$ is continuous as it is an isometry, the function $h$ is continuous.  $\blacksquare$

Interestingly, the converse of the above result is also true \cite[Lemma 5]{Manju_Nonlinearity}.  The map $g(v, \cdot)$ in \eqref{Scomm_h} actually depends on $v$. In general, it is not possible to find a map on $X_U$ that is independent of $v$ so that the diagram in \eqref{Scomm_h} commutes. However, when the inputs originate from a dynamical 
system, we can restrict $h$ to a subspace of $\overleftarrow{U}$ and we then can establish a $v$-independent map when the \textit{single-delay lag dynamics} is considered on a subset of $X_U \times X_U$. To describe the  single-delay lag dynamics formally, we consider a dynamical system $T: U \to U$ and we define a relation on the reachable set $X_U$, i.e., a subset defined on $X_U \times X_U$ by 
$$Y_T:=\{(x_{n-1},x_n): \{x_k\}_{k\in \mathbb{Z}} \mbox{ is a solution for some orbit of } T \mbox{ and } n \in \mathbb{Z}\}.$$

The following result shows that $T$ induces a self-map $G_T$ on $Y_T$ and its iterates describes the single-delay lag dynamics.

\begin{Theorem} \label{Thm_GT_Exists}
Let $g$ be a driven system that is SI-invertible, and $(U,T)$ be a dynamical system. Consider the subspace $Y_T$ of $X_U \times X_U$ to be the tuple arising from two successive points of any solution of $g$, i.e.,
$$Y_T:=\{(x_{n-1},x_n): \{x_k\}_{k\in \mathbb{Z}} \mbox{ is a solution for some orbit of } T \mbox{ and } n \in \mathbb{Z}\}.$$ Then we have a well-defined map $G_T: Y_T \to Y_T$ defined by $G_T : (x_{n-1},x_n) \mapsto (x_n,x_{n+1})$. Consequently, the mapping $(x_{n-1},x_n) \mapsto u_{n}$ is well-defined when $x_{n-1}$ and $x_n$ are successive points on a solution obtained for an input 
$\{u_n\}$ that is an orbit of $T$.
\end{Theorem}

{\bf Proof.} Since $x_n=g(u_{n-1},x_{n-1})$ and $g(\cdot,x):U \to X$ is invertible, we have 
\begin{equation}
u_{n-1} = g_{*,x_{n-1}}^{-1}(x_n),
\end{equation}
where  $g_{*,x}^{-1}$ is the inverse of the map $g(\cdot,x):U \to X$. Using this in $x_{n+1}=g(u_n,x_n)=g(Tu_{n-1},x_n)$, we have $x_{n+1}=g(Tg_{*,x_{n-1}}^{-1}(x_n),x_n)$. Therefore, there exists a function $\theta_T: (x_{n-1},x_n) \mapsto x_{n+1}$. As a consequence we can define a map $G_T : (x_{n-1},x_n) \mapsto 
(x_n,x_{n+1})$ by 
\begin{equation} \label{eqn_GT}
	G_T: (x_{n-1},x_n) \mapsto 
(x_n,\theta_T(x_{n-1},x_n)). 
\end{equation}
Lastly, since $u_{n-1} = g_{*,x_{n-1}}^{-1}(x_n)$ and $u_{n} = Tu_{n-1}$, there exists a mapping $(x_{n-1},x_{n}) \mapsto u_n$.
$\blacksquare$

\begin{Definition} \rm
	If $(U,T)$ is a dynamical system then we call $(Y_T,G_T)$ the dynamical system induced by $g$. 
\end{Definition}

The natural question that arises is if we can obtain a relationship like a topological conjugacy or a semi-conjugacy between the single-delay lag dynamics described by $(Y_T, G_T)$ and $(U,T)$. The answer is affirmative when $g$ has the USP but indirect. We can establish such a conjugacy/semi-conjugacy with what we call the inverse-limit system of $(U,T)$ that we explain next.

Roughly, the inverse-limit system of $(U,T)$ comprises a self-map on a subset of an infinite dimensional
space (e.g.,  the Hilbert cube) where each point in the inverse-limit space corresponds to a backward orbit. 
Formally, any dynamical system $(U,T)$  determines a nonempty subspace $\widehat{U}_T$ of $\overleftarrow{U}$ given by \begin{equation}
\label{eqn_Inv}
 \widehat{U}_T: = \{(\ldots,u_{-2},u_{-1}) : Tu_{n} = u_{n+1} \},
\end{equation} 
and $\widehat{U}_T$ is equivalent to the inverse-limit space of $(U,T)$ or the natural extension of $(U,T)$ in the  literature (e.g., \cite{ingram2011inverse}). Note that the inverse-limit space is well-defined since we have assumed that $T: U \to U$ is surjective. It is customary to write the inverse-limit space as a space comprising right-infinite sequences in the literature instead of left-infinite sequences that we have considered in \eqref{eqn_Inv}.

The map $T$ also induces a self-map $\widehat{T}$ on 
$\widehat{U}_T$ defined by  $\widehat{T}: (\ldots,u_{-2},u_{-1}) \mapsto 
(\ldots,u_{-2},u_{-1},T(u_{-1}))$.  
We refer to the dynamical system $(\widehat{U}_T,\widehat{T})$ as the inverse-limit system of $(U,T)$.
The inverse-limit system is an extension of $(U,T)$, since
\begin{equation}  \label{comm_inv}
%    \[ 
    \psset{arrows=->, arrowinset=0.25, linewidth=0.6pt, nodesep=3pt, labelsep=2pt, rowsep=0.7cm, colsep = 2.1cm, shortput =tablr}
 \everypsbox{\scriptstyle}
 \begin{psmatrix}
 \widehat{U}_T & \widehat{U}_T\\%
U &  U,
 %%%
 \ncline{1,1}{1,2}^{ \widehat{T}} \ncline{1,1}{2,1} <{\pi_{-1}}
 \ncline{1,2}{2,2} > {\pi_{-1}}
 \ncline{2,1}{2,2}^{T}
 \end{psmatrix}
% \]
\end{equation}
holds where $\pi_{-1}(\ldots,u_{-2},u_{-1}) = u_{-1}$. Further, the inverse-limit systems do not  introduce any new complexity into the dynamics in the sense the topological entropies of $(U,T)$ and $(\widehat{U}_T,\widehat{T})$ are identical \cite{bowen1970topological}. Also, $(U,T)$  satisfies several other topological properties \cite{liang1993dynamical} if and only if $(\widehat{U}_T,\widehat{T})$ does.

%When the input to the driven system $g$ is an orbit of $(U,T)$, and we consider the value of the solution $x_n$ at time $n$, the value of $x_n$ would have been influenced by a left-infinite orbit $(\ldots,u_{n-2},u_{n-1})$ which belongs to $\widehat{U}_T$.

For our analysis with inputs being restricted to be orbits of $T$, and particularly while dealing with left-infinite spaces as in \eqref{Scomm_h}, it 
 is sufficient to restrict  the system on the top in \eqref{Scomm_h} to the inverse-limit system 
$(\widehat{U}_T, \widehat{T})$. This is since the new input value at any time would be the image of the previous value under $T$, i.e., if $u_{-1}$ is the current input value, then $Tu_{-1}$ is the next input value. Hence $\sigma_v(\cev{u})$ in \eqref{Scomm_h} is
$\sigma_{Tu_{-1}}(\cev{u})$ and this is equal to  $\widehat{T}(\cev{u})$. Note that if $\{x_k(\bar{u})\}$ is a solution, then $x_n$ would have been influenced only by the left-infinite sequence $(\ldots,u_{n-2},u_{n-1})$ which belongs to $\widehat{U}_T$. 
In this setting of inputs originating from a dynamical system $(U,T)$, we establish a relationship between the inverse-limit system of $(U,T)$ and its induced dynamical system $(Y_T,G_T)$:

\begin{Theorem}{{\bf (Causal Embedding Theorem.)}}
 \label{Thm_CET}
	Let $g$ be a driven system that is SI-invertible and has the USP. Let $h$ denote the universal semi-conjugacy and $H_2(\overleftarrow{u}) := (h(r\overleftarrow{u}),h(\overleftarrow{u}))$, where $r$ is the right-shift $r: (\cdots, u_{-2},u_{-1}) \mapsto 
  (\cdots, u_{-3},u_{-2})$.  Let $(\widehat{U}_T, \widehat{T})$  be the inverse-limit system of any dynamical system $(U,T)$. 
 The function $H_2$ restricted to $\widehat{U}_T$ is a topological semi-conjugacy between the inverse-limit system 
  $(\widehat{U}_T, \widehat{T})$  and the induced dynamical system  
  $(Y_T,G_T)$, i.e., the following diagram commutes
\begin{equation} \label{comm_H_CET}
\psset{arrows=->, arrowinset=0.25, linewidth=0.6pt, nodesep=3pt, labelsep=2pt, rowsep=0.7cm, colsep = 1.1cm, shortput =tablr}
 \everypsbox{\scriptstyle}
 \begin{psmatrix}
 \widehat{U}_T & \widehat{U}_T\\%
 Y_T &  Y_T.
 %%%
 \ncline{1,1}{1,2}^{\widehat{T}} \ncline{1,1}{2,1} <{H_2}
 \ncline{1,2}{2,2} > {H_2}
 \ncline{2,1}{2,2}^{G_T}
 \end{psmatrix}
 \end{equation}
or in other words, $(Y_T, G_T)$ is a factor of  $(\widehat{U}_T, \widehat{T})$. Further, if $T:U \to U$ is a homeomorphism, then $H_2$ embeds $\widehat{U}_T$ in $X_U \times X_U$, and hence
$(Y_T, G_T)$ is conjugate to $(\widehat{U}_T, \widehat{T})$.
\end{Theorem}

{\bf Proof. } From Theorem~\ref{Thm_UCT}, we know that $h$ is continuous and the following diagram commutes:  \begin{equation}  \label{comm_h_repeat}
%    \[ 
    \psset{arrows=->, arrowinset=0.25, linewidth=0.6pt, nodesep=3pt, labelsep=2pt, rowsep=0.7cm, colsep = 1.1cm, shortput =tablr}
 \everypsbox{\scriptstyle}
 \begin{psmatrix}
 \overleftarrow{U} & \overleftarrow{U}\\%
 X_U & X_U.
 %%%
 \ncline{1,1}{1,2}^{\sigma_v} \ncline{1,1}{2,1} <{h}
 \ncline{1,2}{2,2} > {h}
 \ncline{2,1}{2,2}^{g(v,\cdot)}
 \end{psmatrix}
% \]
\end{equation}

Consider the inverse-limit space $\widehat{U}_T \subset \overleftarrow{U}$, and a bi-infinite orbit $\bar{u}$ of $T$. The left-infinite segment $\cev{u} = (\ldots,u_{-2},u_{-1})$ then belongs to $\widehat{U}_T$. If we restrict the map $\sigma_v$ to $\widehat{U}_T$, and select $v=Tu_{-1}$, then we have 
$\sigma_v(\ldots,u_{-2},u_{-1})=\sigma_{Tu_{-1}}(\ldots,u_{-2},u_{-1})=\widehat{T}(\ldots,u_{-2},u_{-1})$.
Let $(x_{n-1},x_n) \in Y_T$. Without loss of generality, let $n=0$, i.e., $(x_{-1},x_0) \in Y_T$. By definition of $Y_T$, there exists an orbit $\{u_k\}$ of $T$  for which $\{x_n\} = \Psi(\{u_k\})$ 
so that $(x_{-1},x_0) \in Y_T$ (recall $\Psi$ is the solution map). Also by Theorem~\ref{Thm_UCT}, and since $T$ is surjective, $h(\cev{u}) = x_0$ and $h(r\cev{u}) = x_{-1}$ where $\cev{u} = (\ldots,u_{-2},u_{-1})$ is the left-infinite segment of the orbit $\{u_k\}$. 
Hence $H_2: \widehat{U} \to Y_T$ is surjective. 
 Since $h$ and $r$ are continuous, it follows that $H_2$ is also continuous. 
Also, since $h\circ \widehat{T}(\cev{u}) = x_1$, we have $H_2 \circ \widehat{T}(\cev{u}) =  (x_{0},x_{1}) = G_T(x_{-1},x_{0})$. Using all of these in \eqref{comm_h_repeat},  the diagram in \eqref{comm_H_CET} commutes, and $(Y_T,G_T)$ is a factor of $(\widehat{U}_T, \widehat{T})$.

It remains to be shown whenever $T$ is a homeomorphism  then $H_2$ does not map two distinct points in $\widehat{U}_T$ to the same point. Suppose there exists two orbits $\{u_k\}$ and $\{v_k\}$ so that $\cev{u} \not= \cev{v}$ and
$H_2(\cev{u}) = H_2(\cev{v})$. This means that the solutions $\{x_k\} = \Psi(\{u_k\})$ and $\{y_k\}= \Psi(\{v_k\})$ are such that $(x_{-1},x_0) = 
(y_{-1},y_0)$. Since $g$ is SI-invertible, this implies $u_{-1}=v_{-1}$. Since $T$ is a homeomorphism, $u_{-1}$ and $v_{-1}$ have the same left-infinite history which implies $\cev{u} = \cev{v}$. Hence when $T$ is a homeomorphism, $H_2$ embeds $\widehat{U}_T$ in $X_U \times X_U$. By definition of $Y_T$, $H_2( \widehat{U}_T) = Y_T$ and hence $H_2: \widehat{U}_T \to Y_T$ is a homeomorphism. Thus $(Y_T,G_T)$ is topologically conjugate to $(\widehat{U}_T, \widehat{T})$. $\blacksquare$

\begin{Corollary} \label{Coroll}
Let $X\subset \mathbb{R}^N$, and $[H_2]$ denote the collection of all component functions of $H_2:\widehat{U}_T \to Y_T$ each of which that maps $X$ into $\mathbb{R}^2$. 
        	If $\mathcal{K}$ denotes the Koopman operator of the dynamical system $(\widehat{U}_T, \widehat{T})$ then $([H_2], G_T)$ is a finite-dimensional faithful representation and, for each $f_i \in [H_2]$,   $\mathcal{K} f_i = \pi_i \circ G_T \circ H_2$, where  $\pi_i$ picks the $i^{\mathit{th}}$ component function of $H_2 :\widehat{U}_T \to X_U \times X_U$.
\end{Corollary}

{\bf Proof. } If $\pi_i$ picks the $i^{\mathit{th}}$ component function of $H_2$ then $f_i = \pi_i \circ H_2$. When the diagram in \eqref{comm_h_repeat}  commutes, we have $H_2 \circ \widehat{T} =G_T \circ H_2$. By composing $\pi_i$ on both sides, we have implies $f_i \circ \widehat{T} = \pi_i \circ G_T \circ H_2$. But by definition of $\mathcal{K}$, $f_i \circ \widehat{T} = \mathcal{K} f_i$. Hence, $\mathcal{K} f_i  = \pi_i \circ G_T \circ H_2$ when $\mathcal{K}$ is the Koopman operator of $\widehat{T}$.  $\blacksquare$

The reader may note that when $g$ is a RNN as in \eqref{Seq_RNN} and if  $\{f_1,\ldots,f_N\}$ denotes the $N$ component functions of $H_2(\cev{u}^n)$ then each $f_i(\cev{u}^n)$ pairs the $i^{\mathit{th}}$ coordinate of $x_{n-1}$, i.e., the state of the $i^{\mathit{th}}$ node at time $n-1$ with the state of $i^{\mathit{th}}$ node at time $n$ ($i^{\mathit{th}}$ coordinate of $x_n$) where $x_n = h(\cev{u}^n)$, and thus $f_i$'s define a set of $\mathbb{R}^2$-valued observables  of the inverse-limit space $(\widehat{U}, \widehat{T})$. According to the definition of a faithful representation \cite{mezic2020koopman} made in Section~\ref{Sec_Koopman}, $\mathbf{f}$ the collection of component functions of $H_2$ and the map $G_T$ together form a finite faithful representation of $(\widehat{U}, \widehat{T})$.

\begin{Definition} \rm
	We say a driven system $g$ \emph{causally embeds} a dynamical system $(U,T)$ if it satisfies the two properties: (i) a universal semi-conjugacy exists, i.e.,  the diagram in \eqref{Scomm_h} commutes (and thus, \eqref{Scomm_imp} also commutes) (ii)  $H_2(\cev{u}) := (h(r\cev{u}),h(\cev{u}))$ embeds the inverse-limit space $\widehat{U}_T$ in $X \times X$. \end{Definition}

Since when $g$ is SI-invertible and has the USP then $H_2: \overleftarrow{U}_T \to Y_T$  is a homeomorphism, $H_2$ can embed the inverse-limit space $(\widehat{U}, \widehat{T})$ of any dynamical system in $X \times X$, and thus causally embed any dynamical system as long as its inverse-limit space is contained in $\overleftarrow{U}$. 
Synoptically,  restricting $H_2$ to inverse-limit spaces of different dynamical systems contained in $\overleftarrow{U}$ establishes semi-conjugacies/conjugacies  between  $G_T$ and those inverse-limit spaces. 
 Thus one does not need to change $g$ to learn $G_T$ when the dynamical system that drives it changes.  Bringing out this fact is a theoretical advancement that drives us to obtain a universal set of observables for learning the Koopman operator \cite{koopman1932dynamical} of any $(\widehat{U}_T, \widehat{T})$ as long as $\widehat{U}_T$ is contained in $\overleftarrow{U}$.

%\eqref{comm_h_repeat} in  Theorem~\ref{Thm_UCT} finite-faithful representation in \eqref{comm_h2}, the following diagram commutes:\begin{equation*}
%    \[   \psset{arrows=->, arrowinset=0.25, linewidth=0.6pt, nodesep=3pt, labelsep=2pt, rowsep=1.3cm, colsep = 1.1cm, shortput =tablr}\everypsbox{\scriptstyle} \begin{psmatrix} \widehat{U}_T & \widehat{U}_T\\% H_2(\widehat{U}_T) & H_2(\widehat{U}_T)\\
 %%% \ncline{1,1}{1,2}^{\widehat{T}} \ncline{1,1}{2,1} <{[H_2]} \ncline{1,2}{2,2} > {[H_2]}\ncline{2,1}{2,2}^{G_T}\end{psmatrix}
% \]\end{equation*}  $\blacksquare$

\section{Equations from Data and Forecasting}

Since two successive points $(x_{n-1},x_n)$ of a solution of $g$ with an input $\bar{u}$ lie in $Y_T$, one can learn the single-lag dynamics of the driven states through the map $G_T: (x_{n-1},x_{n}) \mapsto (x_n,x_{n+1})$ from a sufficiently large finite set of data points $(x_0,x_1)$ , $(x_1,x_2), \ldots, (x_{m-1},x_m)$. Once having a learnt version of $G_T$, one can iterate  to forecast two successive points on a solution corresponding to $\bar{u}$.  Forthwith, one can also predict the input value $u_n$, since two successive states $x_n$ and $x_{n+1}$ determine $u_n$ since $g$ is SI-invertible \eqref{eqn_g*}. We recall that $g$ is SI-invertible,  any two successive values of the solution $(x_n,x_{n+1})$  determine $u_n$ uniquely since \begin{equation} \label{eqn_g*}
u_n=g_{*,x_{n}}^{-1}(x_{n+1})
\end{equation}
where  $g_{*,x}^{-1}$ is the inverse of the map $g(\cdot,x):U \to X$. 
In summary, if $G_T$ is learnt without errors, prediction would be exact whenever $H_2$ causally embeds $(U,T)$, or else the predicted value of $u_n$ would be an approximation to it obtained from the system $(Y_T,G_T)$ that is semi-conjugate to $(\widehat{U}_T,\widehat{T})$.

We empirically illustrate that the single-lag dynamics described by $G_T$ is less functionally complex than learning $T$. We consider an RNN as in \eqref{Seq_RNN} with the USP, and by varying the dimensions of $X$ (i.e., the neurons) in the RNN we measure the functional complexity of the resultant single-delay dynamics. 
Empirically (see Table~\ref{Table_FC}), increasing the dimension of $X$ is found to increase the linear relationship (or intuitively reduce the functional complexity of $G_T$) that is measured as a generalization of the Pearson correlation coefficient to random vectors (e.g., \cite{puccetti2019measuring}) between 
$(x_{n-1},x_n)$ and $G_T(x_{n-1},x_n)$. 
If $\Sigma_{a}$ and $\Sigma_{b}$ denotes the covariance matrices of the vectors $(x_{n-1},x_n)$ and $G_T(x_{n-1},x_n)$ respectively, and if $\Sigma_{ab}$ denotes the covariance matrix between the vectors $(x_{n-1},x_n)$ and $G_T(x_{n-1},x_n)$, then the multidimensional correlation coefficient is computed by using the traces of these matrices:
$$
    \rho= \frac{\text{tr}({\Sigma_{ab}})}{\text{tr}({\sqrt{\Sigma_a\Sigma_b}})}.
$$
This multidimensional correlation coefficient satisfies most of the well-known properties of the one-dimensional Pearson coefficient. In particular, $\rho=\pm 1$ if and only if $Y\overset{d}{=}AX+b$ for some invertible matrix $A$ and vector $b$. Hence $\rho$, and in practice, the estimator $\hat{\rho}$, can be used 
to measure the linear relationship between two random vectors of the same dimension and thus serves as an indicator of functional complexity.

We remark that the estimation of such linear relationship in (see Table~\ref{Table_FC}) obtained by sample correlations alone is justified whenever $\{u_n\}$ is a realization of an ergodic process since then when $g$ has the USP, any solution $\{x_n\}$ is also a realization of an ergodic process \cite{manjunath2014dynamics}. In the ergodic input case the generalized Pearson correlation coefficient  (between $(x_{n-1},x_n)$ and $G_T(x_{n-1},x_n)$) is independent of $n$.

\begin{center}
\begin{table}
    \scalebox{0.75}{\begin{tabular}{|c|c| c c |} 
        \hline
        Input & Dimension of $X$ & $u_n$ vs $u_{n+1}$ 
        & $\begin{bmatrix}
            x_{n-1}\\
            x_n
        \end{bmatrix}$ vs $\begin{bmatrix}
            x_n\\
            x_{n+1}
        \end{bmatrix}$ \\
        \hline
        \hline
        $(w_n)$ is the Lorenz states, sampled every 0.1 timestamps. & & &\\
        \multirow{3}{*}{$u_n=\frac{1}{100}w_n$} & 10 &0.9311  & 0.9731\\
            & 100 &  & 0.9934\\
            & 1000 & & 0.9930\\

        \multirow{3}{*}{$u_n=\frac{1}{10}(\sin(0.1w_{n,x})+\sin(0.1w_{n,y})+\sin(0.1w_{n,z}))$} 
            & 10 &0.8401  & 0.9392\\
            & 100 & & 0.9616\\
            & 1000 & & 0.9737\\
        \hline
        \hline 
        $(w_n)$ comes from the H\'enon map: & & &  \\
         {$w_{n+1}  = \begin{bmatrix} 1 - 1.4w_{n,x}^2+w_{n,y} \\
                0.3 w_{n,y}\end{bmatrix}$} & & & \\
        \multirow{3}{*}{$u_n = w_n - \overline{w}$} 
            & 10 &-0.2996  & 0.4278\\
            & 100 &  & 0.5733\\
            & 1000 & & 0.4953\\
        \hline
        \hline 
        $(w_n)$ comes from a Pomeau-Manneville map: & & & \\
        {$ w_{n+1}  = \begin{cases} 
            w_n(1+2^{0.6}w_n^{0.6}) & \text{ if } w_n\leq 0.5\\
            2w_n-1 & \text{ if } 0.5 < w_n
        \end{cases}$} & & & \\
        \multirow{3}{*}{$u_n = w_n - \overline{w}$} 
            & 10 &0.6943  & 0.9260\\
            & 100 & & 0.9542\\
            & 1000 & &  0.9669\\
        \hline
        \hline 
        $(w_n)$ comes from the Logistic map: & & & \\
        {$ w_{n+1}  = 4w_n(1-w_n)$} & & & \\
        \multirow{3}{*}{$u_n = w_n - 0.5$} 
            & 10 &-0.0314  & 0.5577\\
            & 100 &  & 0.7150\\
            & 1000 & & 0.7826\\
        \hline
        % \hline 
        % $(w_n)$ comes from an intermittent H\'enon map: & & &  \\
        %  {$w_{n+1}  = \begin{bmatrix} w_{n,y} \\
        %         -0.015w_{n,x}+1.76+w_{n,y}^2\end{bmatrix}$} & & & \\
        % \multirow{3}{*}{$u_n = \frac{1}{10}(w_n - \overline{w})$} 
        %     & 10 &-0.5135 & 0.5121\\
        %     & 100 & & 0.5098\\
        %     & 1000 & & 0.4466\\
        % \hline
    \end{tabular}}
    \caption{Multidimensional Correlation Coefficients $\rho$ to indicate a general reduction in the functional complexity of the map $G_T$ that arises from a RNN. Each row corresponds to a different dynamical system used in the experiments in \cite{Main_article}. In the second column corresponds to the dimension of $X$ (number of artificial neurons) in the RNN used. 
    The numerical estimate of $\rho$ for the relevant vectors is plotted in the last two columns. The reader can compare the linear relationship between $u_n$ vs $T(u_{n})$ and the linear relationship $\begin{bmatrix}
            x_{n-1}\\
            x_n
        \end{bmatrix}$ vs $G_T \left( \begin{bmatrix}
            x_{n-1}\\
            x_n
        \end{bmatrix}\right)$. The dimension of the RNN from \eqref{Seq_RNN} is also given, showcasing that increasing the dimension of the driven system typically yields a map $G_T$ of lower functional complexity.} \label{Table_FC}
    \end{table}
\end{center}

Having observed that learning the single-delay dynamics entails the learning of a less functional complex map, we also note that Theorem~\ref{Thm_GT_Exists} there exists a well-defined map $(x_{n-1},x_n) \mapsto u_{n}$ when $(x_{n-1},x_n) \in Y_T$. We denote this map by $\Gamma$.

%In principle,  feeding an orbit of $(U,T)$ to a driven system that is SI-invertible and has USP enables learning the map $G_T$  from the simulated solution so that the diagram printed in black in \eqref{Scomm_imp} commutes.

Hence,  one can learn $G_T$ indirectly by first learning the map $\Gamma: (x_{n-1},x_{n}) \mapsto u_n$. Then clearly $(\Gamma, \pi_2)(x_{k-1},x_{k}) =(u_k,x_k)$ where $\pi_2 : (a,b) \mapsto b$ is the coordinate projection. Next, $(\pi_2,g)(u_k,x_k) = (x_{k},x_{k+1})$. Therefore, we can rewrite the commutativity in \eqref{Scomm_h} to include the map $\Gamma$ as:

 \begin{equation}  \label{Scomm_imp}
%    \[ 
    \psset{arrows=->, arrowinset=0.25, linewidth=0.6pt, nodesep=3pt, labelsep=2pt, rowsep=1.3cm, colsep = 1.1cm, shortput =tablr}
 \everypsbox{\scriptstyle}
 \begin{psmatrix}
 \widehat{U}_T & \widehat{U}_T\\%
 H_2(\widehat{U}_T) & H_2(\widehat{U}_T)\\
 & \textcolor{red}{U \times X}
 %%%
 \ncline{1,1}{1,2}^{\widehat{T}} \ncline{1,1}{2,1} <{H_2}
 \ncline{1,2}{2,2} > {H_2}
 \ncline{2,1}{2,2}^{G_T}
 \psset{linecolor=red}
 \textcolor{red}{\ncline{2,1}{3,2}<{(\Gamma,\pi_2)}}
 \ncline{3,2}{2,2}>{\textcolor{red}{(\pi_2,g)}}.
 \end{psmatrix}
% \]
\end{equation}

Since, $(\pi_2,g)(u_k,x_k) = (x_{k},x_{k+1})$  (see the vertical line in red in \eqref{Scomm_imp} or that in Fig.~\ref{Fig_1}A) -- we can realize  the mapping $(u_k,x_k) \mapsto (x_{k},x_{k+1})$ by feeding $u_k$ back to the driven system (a schematic in Fig.~\ref{Fig_RCN}). Thus, we obtain  iterative roots of $G_T$,  i.e., $G_T = (\pi_2,g) \circ (\Gamma,\pi_2)$.  This equivalent representation of $G_T$ entails another map (see the vertical line in red and the diagonal line in red in that order in Fig.~\ref{Fig_1}A),   $S: (u_k,x_k) \mapsto (u_{k+1},x_{k+1})$ defined by 
 \begin{eqnarray}
	u_{k+1} &=& \pi_1 \circ (\Gamma, \pi_2) \circ (\pi_2,g) (u_k,x_k) \label{Seqn_u}\\
	x_{k+1} &=& \pi_2 \circ (\Gamma, \pi_2) \circ (\pi_2,g) (u_k,x_k). \label{Seqn_x}
\end{eqnarray}
The expressions \eqref{Seqn_u}-\eqref{Seqn_x} are equations constructed from data! We digress to note that for a system with the USP, we can replace $h(\cev{u}^k)$ by $x_k$ in \eqref{Seqn_u} to obtain
$$
u_{k+1} = \pi_1 \circ (\Gamma, \pi_2) \circ (\pi_2,g) (u_k,h(\cev{u}^k)),
$$ 
which is a nonlinear difference equation that only refers to the variable $U$. 
Thus the variable $x_k$ acts as an auxiliary variable in the system defined by $S$.

There are advantages of employing \eqref{Seqn_u}-\eqref{Seqn_x} instead of learning $G_T$. First,  we can save resources by learning $\Gamma$ instead of learning the map $G_T$ on a subset of a very large dimensional space $X \times X$.
This is since in practice $u_n$ is an isometrically embedded image (due to the zero-padding)  of the actual input $v_n$, and since $v_n$  lies in a space with a much lower dimension than $X$ to learn $\Gamma$, it is sufficient to learn the map $(x_{n-1},x_{n}) \mapsto v_n$.

Secondly, if $\cev{u}$ and say its noisy version $\cev{v}$ are nearby then their tails are permitted to be significantly different in the product topology. In that case, $h(\cev{u})$ and $h(\cev{v})$ are nearby owing to the continuity of $h$ due to the USP \cite{manjunath2020stability}, and thereby provides robustness to input noise. The continuity in the product topology can be interpreted as follows: when the noise in the input that occurs in the distant past is mitigated by the driven system as time flows. Such robustness to the input noise in the driven system is called input-related stability in \cite{manjunath2020stability}. More precisely, when there is input-related stability $h(\cev{u})$ is affected to an arbitrarily small extent by that left-infinite segment of the past of the input beyond some arbitrarily large but finite time  \cite[Theorem 3.1]{manjunath2020stability}. Further, when a parameter $\lambda$ is such that $\lambda \mapsto g_\lambda$ is continuous, the universal semi-conjugacy  $h_\lambda$ is also continuous with respect to the parameter $\lambda$. This is referred to as parameter-related stability \cite[Theorem 3.2]{manjunath2020stability}, a consequence of the USP.

 Lastly, learning $G_T$ directly would lead to the same risk as learning a map after employing delay coordinates in Takens embedding where global dissipativity is not guaranteed (see Fig.~\ref{Fig_Takens} and Fig.~\ref{Fig_GTplus}). This is especially crucial when the underlying dynamical system $T$ exhibits chaos, in which case $G_T$ also would exhibit chaos. This is since $\widehat{T}$ inherits chaos since it is an extension of $T$, and when  $G_T$ is conjugate to $\widehat{T}$, $G_T$ also exhibits chaos. Global approximation of maps that exhibit chaos is numerically stable when it is locally conservative on an invariant set and globally dissipative lest would often lead to errors. An empirical illustration would follow in Fig.~\ref{Fig_Glob_Diss}). First, we formally define  global dissipativity and  elaborate with a detailed analysis.

%Thirdly, If $T$ is chaotic, then $\widehat{T}$ is also chaotic and the map $S$ on a domain $D$ contained in $U \times X$ is also chaotic. This entails a small open ball of initial conditions contained in $D$ to grow quickly in time. Separately,  due to either noise in the input or due to approximation of $\Gamma$ by its learnt version, the noise could expand the domain of the map $S$ beyond $D$. We note that $g$ can be designed so that the solutions of $g$ lie always in the compact space $X$ regardless of the expansion of the input space $U$ (for e.g., \eqref{eq_RNN}). Hence the dynamics that spills far beyond $D$ can be quenched by a saturation function like in the RNN in \eqref{eq_RNN}. This restricts the growth of the error while forecasting \ednote{M: Will include a figure in the supp}. With an RNN in \eqref{eq_RNN}, the set $H_2(\widehat{U}_T)$ can be made to expand by setting $\alpha$ large enough while retaining the USP \cite{Supp}) so that the quenching gets close to the boundary of the domain of $S$.

%We make an analysis to show that the expansion $(\pi_2,g^+) \circ (\Gamma_{exp},\pi_2)$ is globally dissipative, thereby obtaining some degree of stability when small computational errors are made, a feature was lacking with Takens delay embedding.

\begin{Definition} \rm
	We say a dynamical system $(Z,f)$ is globally dissipative if there exists a nonempty proper closed subset $B$ of $Z$  so that for all $x\in Z\setminus B$, (i). $\omega(x;f) \subset B$  and (ii).  $B$ is positively invariant that is $f(B) \subset B$. We call any such closed subset $B$ a trapping set of $f$. 
\end{Definition}

It is obvious when $(Z,f)$ is globally dissipative, every invariant set or attractor of $f$ would be contained in a trapping set $B$. In fact one can find that $\bigcap_{n=1}^\infty f^{(n)}B$ is an
attractor.

An expansion of a set $A$ contained is a fattening of $A$ that is also closed. Formally, an expansion of a set $A \subset \mathbb{R}^n$ ($A \subset \mathbb{R}^n \times \mathbb{R}^n$) is any closed subset $A^+ \subset \mathbb{R}^n$ ($A^+ \subset \mathbb{R}^n \times \mathbb{R}^n$) that  contains $A$. 
Given a function $f$ defined on $A$, we denote a function $f^+$ to be some continuous function that preserves $f$ on $A$ and also leaves $A^+$ positively-invariant. Formally, suppose $f$ is a continuous function defined on $A$, then $f^+$ denotes a choice of a continuous function defined on $A^+$ so that $f^+(a)=f(a)$ for all $a \in A$, and  $f^+(A^+) \subset A^+$. Now if $(A^+,f^+)$ is globally dissipative with a trapping set contained in $A$, then the forward asymptotic behavior of the orbits under $f^+$ and $f$ would be identical. Our aim is to show that a system $(Y_T^+, (\pi_2,g^+) \circ (\Gamma_{exp},\pi_2))$ that is a learnt version of $(Y_T,(\pi_2,g) \circ (\Gamma,\pi_2))$ can have global dissipativity, and hence their forward asymptotic dynamics are identical; $\Gamma_{exp}$ would be defined in the discussion later. Note that we have assumed that the dynamics of the learnt version has the same dynamics of $(\pi_2,g) \circ (\Gamma,\pi_2)$ on $Y_T$, and the purpose of this discussion is to analyse what can happen if the iterates of $G_T$ slip on to  $Y_T^+$ due to transient or noise at isolated moments in the input.

First we would consider the map $G_T: Y_T \to Y_T$. 
Suppose successive iterates of  of $(x_{n-1},x_{n})$ under $G_T$ is employed to forecast successive values of a solution of $g$ contained in $Y_T$, then these iterates would remain in $Y_T$. However, in practice, due to noise at some moment, an iterate could slip beyond $Y_T$ and into an 
expansion $Y_T^+$; for instance $Y_T^+$ in this case could be 
$$Y_T^+ : = \mbox{Closure}\bigg(Y_T \cup \cup_{n > 0 } (y_{n-1},y_{n})\bigg),$$  where $y_i$ is an approximation of $x_i$.  
In practice, a function like a feedforward neural network that is made to learn $G_T$ would also accept a value outside $Y_T$ as an argument.  So in practice, we have a map $G_T^+$ on an expansion $Y_T^+$. In general, $Y_T$ need not be an attractor of $G_T^+: Y_T^+ \to Y_T^+$, or more generally there is no guarantee that $(Y_T^+,G_T^+)$ is globally dissipative. Especially for chaotic maps, a map $G_T^+$ could continue to show sensitive dependence on initial conditions on $Y_T^+\setminus Y_T$ as well, which could take the iterates of $G_T^+$ further away from $Y_T$, and hence forecasting leads to large errors (see Fig.~\ref{Fig_GTplus} for a schematic, and an empirical illustration in the figure on the left in Fig.~\ref{Fig_Glob_Diss}). If we learn $\Gamma$ instead of $G_T$, then the system $(Y_T^+, (\pi_2,g^+) \circ (\Gamma_{exp},\pi_2))$ which is what we realise in practice while learning $(\pi_2,g) \circ (\Gamma,\pi_2)$ in \eqref{Scomm_imp} turns out to be globally dissipative (Theorem~\ref{Thm_GD}) under some conditions on $g$; here $\Gamma_{exp}$ denotes some function defined on $Y_T^+$ that agrees with $\Gamma$ on $Y_T$. In summary, by feeding the predicted values of the input into the driven system $g$ during forecasting as in Fig.~\ref{Fig_RCN} or more formally, as indicated in  \cite[Fig.~\ref{Fig_1}A]{Main_article}, the iterates under   
$(Y_T^+, (\pi_2,g^+) \circ (\Gamma_{exp},\pi_2))$ would not wander far away from the iterates of $(\pi_2,g) \circ (\Gamma,\pi_2)$ in practice. (for a schematic see Fig.~\ref{Fig_Takens2} and for 
an empirical illustration see the figure on the right in Fig.~\ref{Fig_Glob_Diss}).

\title{Figure was here}
% \begin{center}
% \begin{figure}[t]
% \centering
% \includegraphics[width=16cm]{Fig_GT_plus}
% \caption{Schematic to explain the errors that could occur if $G_T$ is learnt to forecast values in $Y_T$.}
% \label{Fig_GTplus}
%  \end{figure}
% \end{center}

Before proving  Theorem~\ref{Thm_GD}, we deal with how $\Gamma$ can be learnt reliably from successive points on a simulated solution. Note that for obtaining an actual solution,  the entire left-infinite input has to be fed to the driven system. When $g$ has the USP, the solutions of $g$ can be recognized as non-autonomous uniform attractors \cite{Manju_Nonlinearity}, that is each solution attracts all other initial conditions towards the components of the solution, and moreover, this attractivity is uniform for all solutions -- uniform attraction property \cite{Manju_Nonlinearity}, which we state next.  Recall that when $A$ and $B$ belong to  $\mathsf{H}_X$, the quantity 
$dist(A,B):= \inf\{ \epsilon : A
\subset B_\epsilon(B)\}$ is the Hausdorff semi-distance between $A$ and $B$.

\begin{Definition} \rm \label{Def_UAP}
Let $g$ be a driven system. Then $g$ is said to have the uniform attraction property (UAP) if for each $\bar{u} \in \overleftarrow{U}$ the process $\phi_{\bar{u}}$	is such that 
there exists a sequence of singleton subsets  $\{A_k(\bar{u})\}$ of $X$ so that $\phi_{\bar{u}}(k,k-j,A_k(\bar{u})) = A_{k+1}(\bar{u})$ and 
\begin{equation} \label{eq_UAP}
	\lim_{j\to \infty} \sup_k \sup_{\bar{u}} dist\Big( \phi_{\bar{u}}(k,k-j,X), A_k(\bar{u})\Big) = 0.
\end{equation}
\end{Definition}
It is a result in \cite[Theorem 1]{Manju_Nonlinearity} that the UAP is equivalent to the USP. Now, clearly, when $g$ has the USP, the sequence of subsets that satisfy Definition~\ref{Def_UAP} are subsets containing the elements of the solution $\Psi(\bar{u})$.  Hence from  \eqref{eq_UAP}, we have 
$$ 
	\lim_{j\to \infty}  \sup_k \sup_{\bar{u}} d\Big( \phi_{\bar{u}}(k,k-j,y), x_k(\bar{u})\Big) = 0,
$$
for all $y \in X$, and where  $\{x_k(\bar{u})\} = \Psi(\bar{u})$.  By replacing $k$ by $k+j$, we obtain 
\begin{equation} \label{Seqn_uniform}
	\lim_{j\to \infty}  \sup_k \sup_{\bar{u}} d\Big( \phi_{\bar{u}}(k+j,k,y), x_{k+j}(\bar{u})\Big) = 0,
\end{equation}
for all $y \in X$, and where  $\{x_k(\bar{u})\} = \Psi(\bar{u})$. This in turn implies for any $\bar{u}$, and $k \in \mathbb{Z}$ and any pair $y_1,y_2 \in X$
\begin{equation} \label{Seqn_uniform2}
	\lim_{j\to \infty} 
	\max\Bigg(d\Big(\phi_{\bar{u}}(k+j-1,k,y_1), x_{k+j-1}(\bar{u})\Big),
	d\Big(\phi_{\bar{u}}(k+j,k,y_2), x_{k+j}(\bar{u})\Big)\Bigg) = 0,
\end{equation}
where  $\{x_k(\bar{u})\} = \Psi(\bar{u})$.

The limit in \eqref{Seqn_uniform2} ensures that if we initialize the driven system
with an arbitrary initial value $y_m \in X$, then the sequence $y_{m+1}, y_{m+2}, y_{m+3},...$ satisfying $y_{k+1}= g(u_k,y_k)$ for $k \geq m$ approximates the actual solution $\{x_n\}$ \textit{uniformly} (see \eqref{eq_UAP}). To be precise, given $\epsilon>0$ (independent of $y_m$) there is an integer $n$ so that the distance between  $x_{n+i}$ and $y_{n+i}$ is less than $\epsilon$ for all $i\ge 0$, where $\{x_m\} = \Psi(\bar{u})$  and $y_{n+i}$ is generated by $y_{k+1} = g(u_k,y_k)$ for $k\ge m$. Thus by leaving out a few values of $y_i$ (termed as washing out initial conditions in the RC literature \cite{jaeger2004harnessing}), for practical purposes the subsequent values of $y_i$ are indistinct from the actual solution values. In contrast, the Takens delay embedding theorem in \cite{takens1981detecting} does not guarantee embedding a subset of the manifold as an attractor (recall Fig.~\ref{Fig_Takens}).

We next elaborate on the conditions on $g^+$ so that $(Y_T^+, (\pi_2,g^+) \circ (\Gamma_{exp},\pi_2))$ is globally dissipative.   Now suppose $g^+$ is such that  $g^+(U^+\times X^+) \subset X$. An example of this is the RNN in \eqref{Seq_RNN}. In this case all solutions of $g^+$ would lie within $X=[-1,1]^N$ thanks to the $\tanh$ function. Hence $Y_T^+$, an expansion of $Y_T$ would be always contained in $X \times X$. Next, let $U^+ = \Gamma_{exp}(Y_T^+)$ be an expansion of $U$, where $\Gamma_{exp}$ is the learnt version of $\Gamma$ with domain $Y_T^+$.  Clearly, $U^+$ is compact since $X$ is compact and $\Gamma_{exp}$ is continuous.  In this scenario, if $g^+$ has the USP, then it turns out that the product of the reachable set $X_{U^{+}} \times X_{U^{+}}$ is a trapping set of the map $(\pi_2,g^+) \circ (\Gamma_{exp},\pi_2)$ as indicated in Fig.~\ref{Fig_GD}. As promised a formal proof is in Theorem~\ref{Thm_GD}.

%Then $(\pi_2,g) \circ (\Gamma_{exp},\pi_2)(x_1,x_2) \subset X \times X$. Hence $Y_T^+$ would be always contained in $X$. Next, let $U^+$ denote $\Gamma_{exp}(Y_T^+)$, where $\Gamma_{exp}(x_1,x_2) = \Gamma(x_1,x_2)$ and $\Gamma_{exp}$ is the map obtained after learning $\Gamma$, and let $U^+ = \Gamma_{exp}(Y_T^+)$.  Hence $(Y_T^+,(\pi_2,g^+) \circ (\Gamma_{exp},\pi_2))$ is well-defined. To discuss the  global dissipativity of $(Y_T^+,(\pi_2,g^+) \circ (\Gamma_{exp},\pi_2))$, we first adopt a definition from \cite{Manju_Nonlinearity}.

%The next result shows that $X_{U^{+}} \times X_{U^{+}}$ is a trapping set for the system $(Y_T^+,(\pi_2,g^+) \circ (\Gamma_{exp},\pi_2))$ making it globally dissipative. Since $X_{U^{+}} \times X_{U^{+}}$ would contain $Y_T$, we have some degree of stability while we learn since $(Y_T^+,(\pi_2,g^+) \circ (\Gamma_{exp},\pi_2))$  would map an initial condition outside $X_{U^{+}} \times X_{U^{+}}$ towards it (see schematic in Fig.~\ref{Fig_GD}).

\title{Figure was here}
% \begin{center}
% \begin{figure}[t]
% \centering
% \includegraphics[width=12cm]{Fig_GD}
% \caption{Schematic to illustrate the effect of global dissipativity.}
% \label{Fig_GD}
%  \end{figure}
% \end{center}

%Thus $\lim_{j\to \infty}  \sup_k d\Big( \phi_{\bar{u}}(k,k-j,y), x_k(\bar{u})\Big) = 0$ for all $\bar{u}$ and $y \in X$.  By replacing $k$ by $k+j$, we get  $\lim_{j\to \infty}  \sup_k d\Big( \phi_{\bar{u}}(k+j,k,y), x_{k+j}(\bar{u})\Big) = 0$for all $\bar{u}$ and $y \in X$.Thus $\lim_{j\to \infty} d\Big( \phi_{\bar{u}}(k+j,k,y), X_U\Big) = 0$ for all $\bar{u}$  $y \in X$ and $k \in \mathbb{Z}$, since the reachable set contains all the components of a solution. This in turn implies\begin{equation} \label{eqn_product}$\lim_{j\to \infty}  \max( d\Big( \phi_{\bar{u}}(k+j-1,k,y_1), X_U\Big),d\Big( \phi_{\bar{u}}(k+j,k,y_2), X_U\Big) ) = 0$ for all $\bar{u}$  $y_1,y_2 \in X$ and $k \in \mathbb{Z}$.\end{equation}

\begin{Theorem} \label{Thm_GD}
If $g^+$ is such that $g^+ : U^+ \times X^+ \to X$ and has the USP then $(\pi_2,g^+) \circ (\Gamma_{exp},\pi_2)$ is globally dissipative with $X_{U^{+}} \times X_{U^{+}}$ being a trapping set. 
\end{Theorem}

{\bf Proof. } By definition $(\pi_2,g^+) \circ (\Gamma_{exp},\pi_2)$ is a self-map on $Y_T^+$. For brevity of notation, set $f := (\pi_2,g^+) \circ (\Gamma_{exp},\pi_2)$. To prove the theorem, we  show that  $\omega((y_1,y_2);f) \subset X_{U^{+}} \times X_{U^{+}}$.

Fix $(y_{-1},y_0) \in Y_T^+$. Let $u_0 := \Gamma_{exp}(y_{-1},y_{0})$. In general denote $u_i := \Gamma_{exp}(f^{(i)}(y_{-1},y_0))$. 
 So under the repeated iterations of $f$ we also find an input sequence $\{u_0,u_1,u_2,\ldots \subset  U^+\}$. Let $\bar{u}$ be the bi-finite sequence contained in $U^+$ so that its right infinite sequence matches with $\{u_1u_2,\ldots\} \subset  U^+$. Let $\phi_{\bar{u}}^+$ be the composition operator of $g^+$ with input ${\bar{u}}$. Then 
 by definition of $\bar{u}$, we have $f(y_{-1},y_{0}) =  (y_0,\phi_{\bar{u}}^+(1,0,y_{0})) = (\phi_{\bar{u}}^+(0,0,y_{0}),\phi_{\bar{u}}^+(1,0,y_{0}))$.
 In general, for all $j>0$,  $$f^{(j)}(y_{-1},y_{0}) =   
 \Big(\phi_{\bar{u}}^+(j-1,0,y_0), \phi_{\bar{u}}^+(j,0,y_0)\Big).$$

Let $\{x_k\} :=  \Psi(\bar{u})$. By definition of the reachable set $X_{U^{+}}$,   $(x_{k},x_{k+1}) \in X_{U^{+}} \times X_{U^{+}}$ for all $k \in \mathbb{Z}$. The product  $X_{U^{+}} \times X_{U^{+}}$ is compact since by definition $X_{U^{+}}$ is compact. Hence to prove that 
 $\omega((y_1,y_2); f)\subset X_{U^{+}} \times X_{U^{+}}$ it is sufficient if we show that 
$f^{(j)}(y_{-1},y_0) \to (x_{j-1},x_{j})$ as $j \to \infty$. Recalling \eqref{Seqn_uniform2}, we have  for all $k \in \mathbb{Z}$, and $(z_1,z_2)$,  $$\lim_{j\to \infty}\max\Bigg(d\Big(\phi_{\bar{u}}^+(k+j-1,k,z_1), x_{k+j-1}(\bar{u})\Big),
	d\Big(\phi_{\bar{u}}^+(k+j,k,z_2), x_{k+j}(\bar{u})\Big)\Bigg) = 0.$$  
 In particular by setting $k=0$, $z_1=z_2=y_0$, we have 
 $$\lim_{j\to \infty}\max\Bigg(d\Big(\phi_{\bar{u}}^+(j-1,0,y_0), x_{j-1}(\bar{u})\Big),
	d\Big(\phi_{\bar{u}}^+(j,0,y_0), x_{j}(\bar{u})\Big)\Bigg) = 0$$ which implies  $f^{(j)}(y_{-1},y_{0}) \to (x_{j-1},x_{j})$ as $j \to \infty$. 
$\blacksquare$

Figure~\ref{Fig_Glob_Diss} illustrates the importance of Theorem~\ref{Thm_GD} where we plot the three principal components of different evolution of the states after learning $G_T$. The principal components of $(G_T^{(n)}(y_0,y_1))$ for different randomly initialized values of $(y_0,y_1)$ is plotted on the left. On the right is plotted the vectors $(x_{n-1},x_n)$ where $(x_n)$ is the solution to the actual data, i.e. it satisfies the update equation $x_{n+1}=g(x_n,u_n)$. 
The Lorenz data from the main article (\cite[Fig.~\ref{Fig_1}B]{Main_article}) was used in the update equation and learn $G_T$.

% and this was used to construct $G_T := (\pi_2,g) \circ (\Gamma,\pi_2)$. Plotted are the first three principal components of the trajectories in space $X_{U^+}\times X_{U^+}$. Randomly initialized points were driven using $G_T^+$, and this is compared with the trajectory of the system as it is being driven by the actual input data. It can be clearly seen that the randomly initialized trajectories tend toward the same attractor as the driven system, illustrating global dissipativity. This is in stark contrast to Figure \ref{Fig_Delay_Lorenz} where the Takens map $F_\theta$ was learnt and which completely fails to learn the attractor. \ednote{I added this paragraph. Please see.}

%In view of the theorem, the reader may compare the result in Theorem~\ref{Thm_GD}, we can summarize the causal embedding property and the advantage of learning $\Gamma$ in Fig.~\ref{Fig_Takens2} (the reader may compare this with Fig.~\ref{Fig_Takens}).

Intuitively from Fig.~\ref{Fig_GD}, if the  diameter of the set $Y_T$ gets larger, the diameter of $X_{U^{+}} \times X_{U^{+}}$ gets larger as well. 
A larger diameter of $X_{U^{+}} \times X_{U^{+}}$ would ensure that the subset indicated in yellow in Fig.~\ref{Fig_GD} would shrink in size since the set $X_{U^{+}} \times X_{U^{+}}$ is always contained in the compact set $X \times X$ when $g^+:(U^+ \times X^+) \to X$. This prevents large errors during forecasting while trajectories slip out of it. To get the diameter of $Y_T$ larger we must make $g(u,\cdot)$  not ``too contractive" -- recall the examples that $g(u,x) = ux/2$ would have $Y_T$  to be a set with a single point $\{0\} \times \{0\}$.  In the case of a RNN as in \eqref{Seq_RNN}, a large diameter of $Y_T$ can be realized by setting $\alpha$ that is large enough so that it also simultaneously satisfies the USP (see \cite{manjunath2020stability} for a detailed study).  In summary, when we employ learning $\Gamma$ instead of learning $G_T$, the commutativity as indicated in  Fig.~\ref{Fig_Takens2} holds. The reader may compare this with Fig.~\ref{Fig_Takens}. In the discussion above, we have not considered the effect of errors made while learning $\Gamma$. This would entirely depend on the functional complexity of the map and the sophisticated method to find it.

\title{Figure was here}
% \begin{center}
% \begin{figure}[t]
% \centering
% \includegraphics[width=12cm]{Fig_GD_Commutativity.png}
% \caption{Causal embedding and global dissipativity together. Compare this with Fig.~\ref{Fig_Takens}}
% \label{Fig_Takens2}
%  \end{figure}
% \end{center}

%Consider a driven system $g$ defined on $U\subset \mathbb{R}$ and $X=[-1,1]$$$g(u,x) = \alpha \overline{\tanh}(u+x),$$where $\alpha \in (0,1)$. Suppose we consider $U= \mathbb{R}$, since $\alpha \in (0,1)$, it is possible to find a $y \in X$ so that there exists no $u \in \mathbb{R}$ and $x \in X$ so that $g(u,x) = y$. Hence $g$ is not surjective and there exists a nonempty open subset of $X$ that does not intersect with the image of $g$ under $(\mathbb{R},X)$. We want to capture driven systems that show this phenomenon through a formal definition.  We say $\tilde{U}$ to be an enlargement of $U \mathbb{R}^N$ if $U \subset \tilde{U} \subset \mathbb{R}^N$ and $\tilde{g}$, the extension of $g$ is such that  $(\tilde{U},X)$, and $\tilde{g}: (\tilde{U},X) \to X$.  We say $g$ is not completely reachable with respect to some enlargement $\tilde{U}$ of $U \in \mathbb{R}^N$, if the extension $\tilde{g}$ is such that $X \setminus \tilde{g}(\tilde{U},X)$ contains a nonempty open-subset of $X$.

%Now suppose, while learning the map $\Gamma$, we make an error and this results in learning an extension $\tilde{Gamma}$ on an enlargement $\tilde{Y_T}$. Suppose $\tilde{U} = \tilde{\Gamma}(\tilde{Y_T}$ is an enlargement of $U$.

%\begin{Theorem} Suppose the driven system $\tilde{g}: \tilde{U} \times X \to X$ has the USP and also not completely reachable, then the extension $\tilde{R}_T$ is globally dissipative. \end{Theorem} {\bf Proof. } We need to find a compact set $A \subset \tilde{Y}_T$ so that for all $z\in \tilde{Y}_T\setminus A$ we have $\omega(z; \tilde{R}_T) \subset A$. {\bf Claim:} The set $A=X_{\tilde{U}}$ is non-empty and compact.{\bf Claim:} For all $z \in X \setminus X_{\tilde{U}}$, $\omega(z; \tilde{R}_T) \subset A$.$\blacksquare$

%Finally, when a parameter $\lambda$ is such that $\lambda \mapsto g_\lambda$ is continuous the universal semi-conjugacy  $h_\lambda$ is also continuous with respect to the parameter $\lambda$. This is referred to as parameter-related stability \cite[Theorem 3.2]{manjunath2020stability}, a consequence of the USP.

%The solutions of $g$ are non-autonomous uniform attractors, that is each solution attracts all other initial conditions towards the components of the solution, and this attractivity is uniform for all solutions -- uniform attraction property \cite{Manju_Nonlinearity}.  This also ensures that if we initialize the driven systemwith an arbitrary initial value $y_m \in X$, then the sequence $y_{m+1}, y_{m+2}, y_{m+3},...$ satisfying $y_{k+1}= g(u_k,y_k)$ for $k \geq m$ approximates the actual solution $\{x_n\}$ \textit{uniformly}. To be precise, given $\epsilon>0$ (independent of $y_m$) there is an integer $n$ so that the distance between  $x_{n+i}$ and $y_{n+i}$ is less than $\epsilon$ for all $i\ge 0$, where $\{x_m\} = \Psi(\bar{u})$  and $y_{n+i}$ is generated by $y_{k+1} = g(u_k,y_k)$ for $k\ge m$. This is a property of $g$ and hence the same convergence to the actual solution holds after learning $\Gamma$ in \eqref{Seqn_u}-\eqref{Seqn_x}. In contrast, the Takens delay embedding theorem in \cite{takens1981detecting} does not guarantee a stable embedding, and in particular the attractor embedded is not necessarily an attractor in the reconstructed  space. Theoretical conditions under which a kind of stable embedding can be obtained depends actually on the observables used in the delay-coordinate mapping \cite{eftekhari2018stabilizing}.

\title{Figure was here}
% \begin{center}
% \begin{figure}[t]
% \centering
% \includegraphics[width=0.7\linewidth]{Fig_Lorenz_Aut.png}
% \caption{Empirical illustration of why global dissipativity as in Theorem \ref{Thm_GD} is important. Plotted are the three principal components of the trajectories of randomly initialised points (blue) obtain through iterating a learnt version of $G_T$ as against the trajectory $(x_{n-1},x_n)$ (red) of the driven states that is obtained when the Lorenz data is fed into the system; data used for learning $G_T$ was that used in the simulation in \cite[Fig.~\ref{Fig_1}B]{Main_article} }
% \label{Fig_Glob_Diss}
%  \end{figure}
% \end{center}

\section{Delay Coordinates as Inputs} \label{Sec_DC}

We next consider a more general problem of learning a dynamical system when only observations of an orbit are available. Explicitly, if $(W,T)$ is a dynamical system with dynamics generated by $w_{n+1}=Tw_n$, and if $\theta:W \to \mathbb{R}$ is an observable, then the task is to learn a system that is topologically conjugate to $(W,T)$ and predict $\theta(w_{m+1}),\theta(w_{m+2}),...$ using the data $\theta(w_{0}),\theta(w_{1}),\ldots,\theta(w_{m})$.  
Suppose the input from the delay-coordinate map $\Phi_{\theta,2d}(\theta(w_{n})) := (\theta(w_{n-2d}),\ldots,\theta(w_{n-1}),\theta(w_{n}))$ is fed into the driven system as $u_n$ and Takens delay embedding theorem (see Theorem~\ref{Thm_Takens}) holds, in which case, there exists a homeomorphism  $F_\theta: \Phi_{\theta,2d}(\theta(w_{n})) \mapsto \Phi_{\theta,2d}(\theta(w_{n+1}))$. Hence if the input values $u_n := \Phi_{\theta,2d}(\theta(w_{n}))$ are fed to a driven system $g$ that is SI-invertible and has USP, the induced dynamical system $(Y_F,G_F)$ would be topologically conjugate to the inverse-limit space of
$(\Phi_{\theta,2d}(W), F_\theta)$ due to Theorem~\ref{Thm_CET}, and one could forecast $u_n,u_{n+1},\ldots$, and hence the values $\theta(w_n), \theta(w_{n+1}),\ldots$. The advantage of feeding delay-coordinates to a driven system is that the embedding is stable in the sense of global dissipativity that we have described in Theorem~\ref{Thm_GD}. From the Koopman operator standpoint, when $X \subset \mathbb{R}^N$, the component functions of $H_2$ are the set of observables 
of the inverse-limit system of $(\Phi_{\theta,2d}(W), F)$ on which the action of the Koopman operator can be learnt from data.

One of the perplexities of employing the delay coordinate map is that the required delay to embed the attractor is not known.  In the case when the required delay is not known, we demonstrated appealing numerical results (\cite[Fig.~\ref{Fig_1}D]{Main_article}) where the theory is based on a conjecture below.

\begin{Conjecture} \label{Conject} Let $g$ be a driven system that is SI-invertible and has the USP. Let $(W,T)$ be a dynamical system and $\theta: W \to \mathbb{R}$ 
be an observable so that Takens delay embedding theorem holds and the 
the component functions of $H_2$  can embed $$\Theta(\widehat{W}_T) := 
\{(\ldots, \theta(w_{-2}),\theta(w_{-1})) : w_{n+1} = T w_n\}$$  in $X_U \times X_U,$.  Then there exists a homeomorphism on $ \widetilde{\mathcal{F}}_{\theta}$ on  $\Theta(\widehat{W}_T)$ and a 
self-map $G_{\Theta,T}$ on   $H_2(\Theta(\widehat{W}_T))$ so that the  following diagram commutes:
\begin{equation}  \label{comm_Takens}
%    \[ 
    \psset{arrows=->, arrowinset=0.25, linewidth=0.6pt, nodesep=3pt, labelsep=2pt, rowsep=0.7cm, colsep = 1.1cm, shortput =tablr}
 \everypsbox{\scriptstyle}
 \begin{psmatrix}
 \Theta(\widehat{W}_T) &  \Theta(\widehat{W}_T)\\%
  H_2(\Theta(\widehat{W}_T)) &  H_2(\Theta(\widehat{W}_T))
 %%%
 \ncline{1,1}{1,2}^{\widetilde{\mathcal{F}}_{\theta}} \ncline{1,1}{2,1} <{H_2}
 \ncline{1,2}{2,2} > {H_2}
 \ncline{2,1}{2,2}^{G_{\Theta,T}}.
 \end{psmatrix}
% \]
\end{equation}

\end{Conjecture}

If the conjecture is true then one  can learn the single-delay lag dynamics $G_{\Theta,T}$ 
 by feeding  $\theta(w_{0}),\theta(w_{1}),\ldots,\theta(w_{m})$ into the driven system $g$ as before to forecast $\theta(w_{m}),\theta(w_{m+11}),\ldots,$.  The premise as to why the conjecture can be true can explained as follows. If Takens embedding theorem for  the system $(W,T)$ and the observable $\theta$, then it easily follows that there is a homeomorphism $$\widetilde{\mathcal{F}}_{\theta} : = (\ldots, \theta(w_{k-2}),\theta(w_{k-1})) \mapsto (\ldots, \theta(w_{k-1}),\theta(w_{k})).$$ Next, it can be shown that the component functions of $H_2$ are also sufficiently smooth \cite[Theorem III.1]{grigoryeva2020chaos} when $g$ has the USP and is sufficiently smooth.  If $U$ is a smooth manifold of dimension $M$, and there are at least $2M+1$ component functions of $H_2$ that are generic in the sense of Whitney's embedding theorem --  then  $H_2$ would embed $\Theta(\widehat{W}_T)$ in $X_U \times X_U$. In this case, $G_{\Theta,T} : (x_{n-1}, x_n) \mapsto (x_{n}, x_{n+1})$ exists and is given by $G_{\Theta,T} = H_2 \circ \widetilde{\mathcal{F}}_{\theta} \circ H_2^{-1}$. However, it is an open question whether there are enough independent component functions in $H_2$ that are generic in the sense of Whitney so that $H_2$ embeds $\Theta(\widehat{W}_T)$.  The only result the authors are currently aware of is that there are enough independent functions in  the universal semi-conjugacy $h$ (actually $h$ restricted to $\Theta(\widehat{W}_T)$)  that makes $h$ a limit point of embeddings in the sense of Whitney \cite[Corollary 2.3.2]{hart2019embedding}; $h$ true turns out to be the so-called echo state map in \cite{hart2019embedding}.

\section{Methods and Forecasting Results} \label{Sec_Methods}

We demonstrate numerical results after learning the map $\Gamma$ and using equations \eqref{Seqn_u}--\eqref{Seqn_x} for forecasting. 
We realize $g$ through a RNN of the type \eqref{Seq_RNN} with the causal embedding property. Regardless of how the map $\Gamma$ is learnt and implemented, we call a system with a RNN that implements $g$ in \eqref{Seqn_u}--\eqref{Seqn_x} as a \emph{recurrent conjugate network} (RCN) in view of $G_T$ being conjugate/semi-conjugate to the inverse-limit system as in \eqref{Scomm_imp}. A schematic of a RCN is shown in Fig.~\ref{Fig_RCN}. Throughout, the matrices $A$ and $B$ in the RNN are randomly initialized, and in particular, we set the spectral radius of the matrix $\alpha B$ to be $\alpha$ by using a matrix $B$ with spectral radius  $1$. For $\alpha$ sufficiently small, usually in $(0,1)$, RNNs have the USP and this can also be verified empirically (using the parameter-stability plot in \cite{manjunath2020stability}) when needed. Unless mentioned otherwise, we retain the same parameter values  $a=0.5$ and $\alpha=0.99$ as in \cite{Main_article} in the experiments. We describe the systems employed in \cite{Main_article} for the experiments  in detail and also supplement them with more numerical results.

\title{Figure was here}
% \begin{center}
% \begin{figure}[t]
% \centering
% \includegraphics[width=16cm]{Fig_RCN.png}
% \caption{Schematic of a RCN with  three nodes in a RNN: (the print in red indicates the connections made to run in the autonomous mode)}
% \label{Fig_RCN}
%  \end{figure}
% \end{center}

%By observing driven systems states with a finite segment of an input, one gets an approximation to a solution fragment as well. The approximation to the solution gets accurate as the length increases

%By collecting finite set of data points $(x_0,x_1)$ , $(x_1,x_2), \ldots, (x_{m-1},x_m)$ belonging to $Y_T$,

In our RCN implementation, we learn $\Gamma: (x_n,x_{n-1}) \mapsto u_n$ by first obtaining the principal components of the states $x_n$ and then 
learning a feedforward network (NN) on these principal components with the target $u_n$. The principal components are used only for an efficient state representation possibly to reduce the errors while learning and is not for a lossy approximation since all principal components are used.  To be explicit, we denote the matrix with the first $N$ states of the network data as row vectors by $X_{1:N}$. If $X_{1:N}=U\Sigma P^T$ denotes the singular value decomposition of $X_{1:N}$, then the principal component matrix is given by $P$ and the principal components are given by 
 $Z_{1:N}=X_{1:N}P.$
These principal components are used to train a feedforward neural network. If we denote the row vectors of $Z_{1:N}$ by 
$z_i^T, i=1,2,\cdots,N$ and the neural network by $NN$, then we learn an approximation of the map
$NN: (z_{n-1},z_n) \mapsto  u_n.
$
We use the learnt approximation of $NN$ to approximate $\Gamma$ since 
$$
NN\left( \begin{bmatrix} 
P^Tx_{n-1} \\
P^Tx_n
\end{bmatrix}
\right) = NN \circ 
\begin{bmatrix}
P^T & 0 \\
0 & P^T 
\end{bmatrix}\begin{bmatrix}
x_{n-1}\\
x_n
\end{bmatrix} = u_n = \Gamma(x_{n-1},x_n).
$$

The feedforward neural network is implemented in {\it Python} using the {\it Keras} \cite{chollet2015keras} library built on {\it Tensorflow}. 
Throughout our experiments, the network is constructed with 12 hidden layers with a layer dimension equal to 128. The activation function on the hidden layers is the {\it ReLU } function built into {\it Keras}, whereas the output layer has a {\it tanh} output. Note that the output of the {\it tanh} activation function lies within $(-1,1)$, so before training, the sequence $(u_n)$ needs to be re-scaled as to fit inside $(-1,1)$. We typically accomplish this by first subtracting the mean $\overline{u}$, and then multiplying the input by a scalar so that the input range is inside $[-0.5,0.5]$ so as to keep the range of $(u_n)$ well within the bounds of the {\it tanh} function.

Training is accomplished using the {\it Adam optimizer}, optimizing the {\it Mean Squared Error} loss function. We train the network three times, using three different learning rates for the {\it Adam optimizer} equal to 0.001,0.0001, and 0.00001. Each time we learn $\Gamma$, we use 150 training epochs and a batch size of 128.

 Chaotic systems form an abstraction of complex behaviors in large-dimensional systems. Hence, chaotic systems provide an amiable set of toy examples on which to test our forecasting theory behind causal embedding. We now describe the chaotic system used in the numerical simulations in the main article \cite{Main_article}. The well-known system of ordinary differential equations developed by Edward Lorenz:
\begin{align} 
\label{Seq_Lor}
\begin{split}
    \dot{x} &= \sigma(x-y) \\
    \dot{y} &= x(\rho -z) -y \\
    \dot{z} &= xy - \beta z,    
\end{split}
\end{align}
with parameters $\sigma=10, \beta=8/3$ and $\rho=28$ exhibits chaos.  We sample the flow of this Lorenz system by using numerical integration techniques (implemented using the {\it SciPy} library's ODE solver \cite{2020SciPy-NMeth}), and in our experiment we use a step size of $0.1$ between successive samples.   Next, we consider the discrete time dynamical system driven by the (full) logistic map, given by: 
\begin{align}
    \begin{split}
        \label{Seq_log}
         v_{n+1} &= 4v_n(1-v_n).
    \end{split}
\end{align}
Note that the logistic map is not invertible, which means the theory (Theorem~\ref{Thm_CET}) only guarantees a semi-conjugacy.
For chaotic systems, even negligible errors in estimating $\Gamma$ would make it hard to minimize the long-term point-wise prediction error. However long-term topological consistency can be achieved: the orbit of the prediction lies on the attractor.

The behavior of deterministic dynamical systems can alternatively be described macroscopically (e.g., \cite{lasota2013chaos}). This study involves describing how the density of an ensemble of initial conditions evolves upon iterating the map on such an ensemble.  A useful tool is the Perron Frobenius operator that determines the evolution of an initial density \cite{lasota2013chaos}.  An important concept in studying stochastic dynamics is the notion of an invariant density $p$ that is a fixed point of the Perron Frobenius operator and when it exists it can determine the visitation frequency of typical orbits to any (measurable) set.   We refer the reader to \cite{lasota2013chaos} for more details.
The accuracy of the reconstructed or forecasted data relies on how the statistical properties like the invariant density  are retained by the forecasted data.  In the main article, we exhibited invariant densities of the logistic map \cite[Fig.~\ref{Fig_1}]{Main_article}. The invariant density of the full logistic map  on $[0,1]$ is found to be $\frac{1}{\pi\sqrt{x(1-x)}}$ (e.g., \cite{lasota2013chaos}.). There exist arbitrary small perturbations of the map $T$ under which the invariant density is much different than that of the invariant density of the logistic map, i.e., there is a non-smooth change in the invariant density. In more technical terms the map $T$ lacks the linear response \cite{baladi2014linear}. Despite this statistical instability, our forecasting results from the RCNs show greater numerical accuracy while the invariant density was simulated (see Fig.~\ref{Fig_1}E). Wherever we cannot graphically illustrate the reconstructed invariant density of a higher dimensional system like the Lorenz system, we tabulate the Wasserstein distance (Table~\ref{Tbl_Wass}) 
between the numerically found densities of the actual and predicted. In this case, we use the 1-Wasserstein distance \cite{dobrushin1970prescribing} which is a metric between one-dimensional probability distributions, implemented programmatically through the {\it SciPy } library \cite{2020SciPy-NMeth}. In addition, we exhibit the cumulative distribution functions (Figure \ref{Fig_CDF}) of the various coordinates of the numerical experiment in  \cite[Fig.~\ref{Fig_1}B]{Main_article}.

\begin{center}
\begin{table}
    \scalebox{0.75}{\begin{tabular}{|c | c c c c c c|} 
        \hline
        Time-steps &1000 &2000 &5000 & 10000&20000 &50000 \\
        \hline
        \hline
        $(w_n)$ is the Lorenz states, sampled every 0.1 timestamps. & & & & & &\\
        {$u_n=\frac{1}{100}w_n$} & & & & & & \\
        $x$-coordinate & 0.0088& 0.0144& 0.0211& 0.0329&0.0452 &0.0686\\
        $y$-coordinate & 0.0088& 0.0144& 0.0211& 0.0329& 0.0452&  0.0686\\
        $z$-coordinate & 0.0022& 0.0065& 0.0136& 0.0204& 0.0285&0.0715\\

        \hline
    \end{tabular} }
    \caption{Wasserstein distances between each coordinate of the forecasted and the actual data. The table illustrates long-term consistency - the data point distribution remains close even after 50,000 prediction steps. Furthermore, it showcases the learnt attractor being close to the actual one.} \label{Tbl_Wass}
    \end{table}
\end{center}

\title{Figure was here}
% \begin{figure}
%     \centering
%     \minipage{0.33\linewidth}
%         \centering
%         \includegraphics[width=\linewidth]{Fig_cdf-x.png}
%         \includegraphics[width=\linewidth]{Fig_hist-x.png}
%         \caption*{$x$-coordinate}
%     \endminipage\hfill
%     \minipage{0.33\linewidth}
%         \centering
%         \includegraphics[width=\linewidth]{Fig_cdf-y.png}
%         \includegraphics[width=\linewidth]{Fig_hist-y.png}
%         \caption*{$y$-coordinate}
%     \endminipage\hfill
%     \minipage{0.33\linewidth}
%         \centering
%         \includegraphics[width=\linewidth]{Fig_cdf-z.png}
%         \includegraphics[width=\linewidth]{Fig_hist-z.png}
%         \caption*{$z$-coordinate}
%     \endminipage
%     \caption{Distributions of the coordinates of the learnt system (blue) and the actual Lorenz system (red). The top row shows the empirical cumulative distribution function (CDF) of each coordinate, and the bottom row their respective histograms. These graphs were constructed by predicting the Lorenz system 50,000 steps into the future, illustrating the long-term consistency and accuracy of the learnt system.}
%     \label{Fig_CDF}
% \end{figure}

We next illustrate the forecasting results obtained by feeding the delay-coordinates of an observation  of \eqref{Seq_Lor} in Figure \ref{Fig_Delay_Lorenz}.
We consider observations $\theta(w_{0}),\theta(w_{1}),\ldots,\theta(w_{m})$ from a  dynamical system determined  by sampling the flow of the Lorenz system $(W,T)$, and where $\theta:W \to \mathbb{R}$ is an observation -- observation we use is the $x$-coordinate of \eqref{Seq_Lor}. The delay-coordinates $\Phi_{\theta,2d}(\theta(w_{n})) := (\theta(w_{n-2d}),\ldots,\theta(w_{n-1}),\theta(w_{n}))$ is fed into the driven system as $u_n$ and the attractor is learnt by learning a map $(x_{n-1},x_{n}) \mapsto w_n$. The original and reconstructed attractor for the Lorenz system is shown in Fig.~\ref{Fig_Delay_Lorenz}. To obtain this result, data $(w_n)$ is generated from the Lorenz system, sampled every 0.1 time-steps. The input into the network is the 10-delayed first coordinate of $(w_n)$, scaled down to fit inside $[-0.5,0.5]$. Specifically $u_n = \frac{1}{100}  (w^{(n-9)}_x,w^{(n-8)}_x,\cdots ,w^{(n)}_x )$. Training of $\Gamma$ was accomplished using 2000 data points, after discarding the first 500 to allow the RNN to forget its initial state.

\title{Figure was here}
% \begin{center}
% \begin{figure}[t]
% \centering
% \includegraphics[width=0.9\linewidth]{Fig_Lorenz_Delay.png}
% \caption{RCN learnt Lorenz attractor in the three dimensional-Euclidean space obtained by feeding delay-coordinates as the input.}
% \label{Fig_Delay_Lorenz}
%  \end{figure}
% \end{center}

One of the great challenges in learning dynamics comes from those systems where the present is greatly sensitive to the distant past much more than the immediate past, that is the past does not pass away.   Examples of systems that show intermittency occurs in dynamical systems whenever the system appears to switch back and forth between two qualitatively different behaviors. 
We consider a type of intermittency (type-I intermittency) that has been found difficult to forecast, one in which the orbits slowly drift from a fixed/periodic point and then move to a chaotic regime but then return to the slowly drifting neighborhood of the fixed point. The time spent by the orbit in a neighborhood of the fixed/periodic point  depends rather sensitively on how closely the system entered its vicinity and the length of this slow drifting phase is unpredictable. Such intermittency is a feature of transition to turbulence and convection in confined spaces \cite{pomeau1980intermittent}, earthquake occurence \cite{bottiglieri2007off},  and anamolous diffusion in biology \cite{klages2013weak}. We consider an invertible map of type-I intermittency in the constant-J sub-families of the H\`enon maps: \cite{kaplan1992return}
\begin{align}
    \begin{split}
        \label{Seq_Hen}
        w_{n+1}  = \begin{bmatrix} w_{n,y} \\
                -Jw_{n,x}+\mu+w_{n,y}^2\end{bmatrix}.
    \end{split}
\end{align}
The parameters $J$ and $\mu$ can be adjusted to alter the  intermittent behavior, and in our case, we used $J=0.015$ and $\mu=1.76$ in our experiment in \cite[Fig.~\ref{Fig_2}A,B,C]{Main_article}.
We next consider the non-invertible Pomeau-Manneville family of maps \cite{pomeau1980intermittent} in which the intermittency is more pronounced:
\begin{align}
    \begin{split}
        \label{Seq_PM}
w_{n+1}= \begin{cases} 
            w_n(1+2^{\gamma}w_n^{\gamma}) & \text{ if } w_n\leq 0.5\\
            2w_n-1 & \text{ if } 0.5 < w_n,
        \end{cases}
    \end{split}
\end{align}
where $0<\gamma<1$. A larger value of $\gamma$ makes the on-off intermittent behaviour more pronounced, and in the main article we had employed $\gamma=0.6$.

The change in qualitative behavior as the parameter $\gamma$ in the Pomeau-Manneville map is tuned is illustrated in Fig.~ \ref{Fig_Interm}. In the main article, we opted to keep the hyperparameters for our choice of RNN in \eqref{Seq_RNN} constant.  We note that we are able to obtain results for maps with stronger intermittency by increasing the parameter $\alpha$ in the RNN slightly greater than $1$ in \eqref{Seq_RNN}. On an intuitive level, this makes the size (diameter) of $Y_T^+$ larger.   With $\alpha=1.2$ instead of $\alpha=0.99$ in \cite{Main_article}, we can obtain both topological and statistical consistency while forecasting for even more pronounced intermittency.

Authors in \cite{gaspard1988sporadicity} have called the behavior exhibited by such intermittent maps as sporadicity that ``fills in a gap between multi-periodic and chaotic dynamical behaviors or equivalently between predictable and random patterns".  Approximations of the Koopman operators of such dynamical systems or feedforward neural network-based approaches are amnesiacs regarding their distant past. We have demonstrated that RCNs can reconstruct the attractor of a map in the Pomeau-Manneville family \cite[Fig. ~\ref{Fig_2}D]{Main_article}. This can be attributed to the fact that RCNs can remember nostalgically the distant past as we consider observables of the inverse-limit systems.

\title{Figure was here}
% \begin{figure}
%     \centering
%     \minipage{0.49\linewidth}
%         \centering
%         \includegraphics[width=\linewidth]{Fig_PM6_phase.png}
%         \includegraphics[width=\linewidth]{Fig_PM6_hist.png}
%         \includegraphics[width=\linewidth]{Fig_PM6.png}
%         \caption*{$\gamma=0.6$}
%     \endminipage\hfill
%     \minipage{0.49\linewidth}
%         \centering
%         \includegraphics[width=\linewidth]{Fig_PM7_phase.png}
%         \includegraphics[width=\linewidth]{Fig_PM7_hist.png}
%         \includegraphics[width=\linewidth]{Fig_PM7.png}
%         \caption*{$\gamma=0.7$}
%     \endminipage\hfill
%     \minipage{0.49\linewidth}
%         \centering
%         \includegraphics[width=\linewidth]{Fig_PM8_phase.png}
%         \includegraphics[width=\linewidth]{Fig_PM8_hist.png}
%         \includegraphics[width=\linewidth]{Fig_PM8.png}
%         \caption*{$\gamma=0.8$}
%     \endminipage
%     \caption{Forecasting results of the Pomeau-Manneville maps.
%      In each of these cases, training was accomplished using the standard RCN from \eqref{Seq_RNN}, differing from the main article \cite{Main_article} only in setting $\alpha=1.2$ instead of $\alpha=0.99$. Also, we use 5000 data points for training, after discarding the first 1000 data points.}
%     \label{Fig_Interm}
% \end{figure}

We also show the effectiveness of our methods to forecast physical data of temperatures across South Africa that shows variations on multiple time scales. Data considered for forecasting is the real-world weather data observed at weather stations and then interpolated using physical models to allow for finer detail. In particular, the data is the {\it ERA5 hourly data on single levels collected from 1979}  available from {\it the Copernicus Climate Change Service (C3S) Climate Data Store (CDS)} \cite{era5data}, and we use the temperature 2 meters above the earth's surface and mean sea level pressure data sampled every 6 hours. The data takes the form of a grid of $65 \times 105$ data points, where grid points refer to a location in South Africa. Due to computational constraints, we scale down the resolution by only considering the 5th data point in each dimension, leaving a grid of size $13 \times 21$ (see top-panel Fig.~\ref{Fig_Clim}). Note that associated with each grid point, there are two numbers: the temperature (in {\it K}), and the Pressure (in {\it Pa}).  For each time-step, we unroll the grid into two vectors of length 173 $(= 13\times 21)$ each. After normalizing each vector by subtracting the mean and scaling down, we stack these vectors atop each other. This final vector of length 546 is fed as input to the network.

As training, we use 26,000 data points, the first 1000 of which are discarded to allow the RCN to forget its initial state values. At a 6-hour sample rate, this translates to training on around 12 year's worth of climate data from late 2005 to 2017. We then use the trained network to predict 5,000 time-steps into the future,  which is roughly 3 years, ending in April 2021. 
Fig.~\ref{Fig_Clim} illustrates the result of the prediction of the temperatures (which is of interest to us). As can be seen, even though the short-term prediction quickly fails (bottom panel of Fig.~\ref{Fig_Clim}), long-term characteristics of the data set have been modeled with the seasonal variation in temperature of the predicted and actual data clearly visible (middle panel of Fig.~\ref{Fig_Clim}). This is despite using only one in five samples from the dataset in our experiment due to computational constraints.

\title{Figure was here}
% \begin{figure}[h!]
%     \centering
%     \minipage{0.49\linewidth}
%         \centering
%         \includegraphics[width=0.9\linewidth]{Fig_Fine_Temp.png}
%         \caption*{A temperature data point}
%     \endminipage\hfill
%     \minipage{0.49\linewidth}
%         \centering
%         \includegraphics[width=0.9\linewidth]{Fig_Coarse_Temp.png}
%         \caption*{Low-resolution data point that is derived}
%     \endminipage

%     \includegraphics[width=0.9\linewidth]{Fig_Ave_Temp_Long.png}
%     \includegraphics[width=0.9\linewidth]{Fig_Ave_Temp_Short.png}
%     \caption{Forecasting average temperature (in Kelvin) across South Africa. 
%     Top-panel: data point from the data set, and derived low resolution data point to minimize computational time. 
%      Middle and Bottom Panels: Mean of the predicted temperature vector on each time-step for 5000 steps and for the first 300 time-steps}
%     \label{Fig_Clim}
% \end{figure}

\section{Conclusions}
 Finding a set of observables of data that determine a less functionally complex learnable map so that its dynamics in the observed space gets closer to that of the action of the Koopman operator on the observables has been pursued by many researchers for the last decade. In practice, however, the Koopman operators of complex and chaotic systems tend to have significantly more complicated spectral properties (e.g., non-isolated eigenvalues and/or continuous spectra) hindering the performance of data-driven approximation techniques that capture only a portion of the spectrum (e.g., \cite{budivsic2012applied,korda2020data}). Also, in the temporal domain, approximations of the Koopman operators of the underlying dynamical systems are amnesiacs regarding their distant past. 
  Here, we show a driven dynamical system that can causally embed dynamical systems is capable of determining a set of observables of the inverse-limit system of the underlying dynamical system and learning the Koopman operator rather than learning an approximation of it. 
 In particular, we produce a topological conjugacy (semi-conjugacy) between the single-delay lag dynamics of the driven system's data and the underlying homeomorphism (non-invertible map that can be discontinuous) dynamical system.

 This methodology induces equations from data. In particular, we have shown that for data arising from a self-map on a compact metric space,  a first-order system of difference equations with auxiliary variables or a single difference equation of infinite order can be obtained.  The states of the driven system are affected to an arbitrarily small extent by that left-infinite segment of the past of the input beyond some arbitrarily large but finite time thanks to the continuity of the universal semi-conjugacy of the driven system. As a consequence finite length of input data is adequate for forecasting. Empirically it is found that the map with the single-delay lag dynamics has a stronger linear relationship and intuitively less functional complexity than the map that describes the data or the map the delay coordinates induce.  As a consequence, through the use of recurrent conjugate networks (RCNs),  we obtain exceptional forecasting results with long-term topological and statistical consistency which is illustrated on chaotic maps and data showing intermittency. 
The robustness of forecasting against external noise is also observed.

%  Mapping temporal data into a new phase space through a driven (dynamical) system that can causally embed a dynamical system lies at the heart of obtaining the equations, and the single-lag dynamics in the new phase space is topologically conjugate or semi-conjugate to the system generating the input depending on whether the input is generated by a homeomorphism or a non-invertible map. Thus we achieve learning the action of the Koopman operator itself rather than approximating it.

The advantage over other known data-driven approaches is not only long-term consistency, but it does not need expert human intuition or physical insights to decide on observables and library functions that are commonly employed in data-driven approaches. This is very useful in high-dimensional forecasting tasks where such insights are rare. 
 The performance of such data-driven approaches is affected in the absence of such physical precognition, and when one overcomes it through combining delay-embedding techniques still exact reconstruction is lacking.
Hence, whenever high-fidelity models are needed, the RCNs out-perform such data-driven methods considering that the step size of a discretization can be even up to {\bf 100 times} than that have been used in some algorithms like SINDy and other echo state network methods. Lastly, and remarkably, we do not have to resort to finding new observables or guessing the library of governing equations for new temporal datasets as the observables induced by the driven systems are universal.

From the perspective of the reservoir computing approaches, we show the existence of a learnable map in a RCN if single-delay lag dynamics is chosen. Although learning a function that the single-delay dynamics entails is more computationally expensive than performing a simple linear regression as in a customary echo state network training, we overcome extensive ad hoc adaptations, like introducing feedback conditions and other parameters, which make the art of training in reservoir computing methods a craft rather than a science.  
The driven system in the RCN  has proven properties of robustness to perturbations of the input and parameters in the system, making these networks attractive for hardware implementations. The theory of causally embedding a dynamical system is general and hence does not depend on a particular choice of the driven system. A future exploration into designing more sophisticated driven dynamical systems to reduce the functional complexity of the map that the single-delay dynamics entails could help in furthering the accuracy that we have obtained with a RCN.

{\bf Data availability.}  All the data used is either computer-generatable or obtained from publicly accessible sources.

{\bf Code availability.} The Python/Matlab code employed for all the simulations in the supplement is also submitted.

\bibliographystyle{IEEEtran}
\bibliography{pnas_supp}

\end{document}

% DISCUSSION: EdNote1, discuss reference to figure 1: small errors lead to major errors
% Talk about how to present code... Too much to fit into a few lines
% \omega not added?

### 2022-01-03 writeup.tex
cleaning git

\documentclass[12pt, oneside]{article}
\usepackage[T1]{fontenc}
\usepackage{lmodern}

\usepackage{amsmath, amssymb, amsthm, latexsym, bm,enumerate, fancyhdr, tikz,afterpage, float}
\usepackage{graphics,graphicx,geometry}
\usepackage{hyperref,color}
%beamer uses these
\usepackage{etex}
\usepackage{epsfig, subfigure, psfrag}
%\usepackage[]{hyperref}
\usepackage[english]{babel}
\usepackage{multimedia}
\usepackage[all]{xypic}
\usepackage{alltt}
\usepackage[numbers,square,sort&compress]{natbib}
\usetikzlibrary{shapes.geometric}
\usetikzlibrary{arrows}
\usetikzlibrary{decorations.markings}
\usepackage{xspace}
\usepackage{comment}
\usepackage{mathtools}

\usepackage[framemethod=tikz]{mdframed}
%color definitions

%berné added
\usepackage{pst-node}
%\newcommand*\circled[1]{\tikz[baseline=(char.base)]{\node[shape=circle,draw,inner sep=2pt] (char) {#1};}}

\colorlet{darkgreen}{green!50!black}
\definecolor{darkblue}{rgb}{0, 0, .6}
\definecolor{midnightblue}{rgb}{.2, .2, .7}
\definecolor{grey}{rgb}{.7, .7, .7}
\definecolor{Red}{rgb}{.9,0,0}
\definecolor{Blue}{rgb}{0,0,.9}
\definecolor{Green}{rgb}{0,.9,0}

\newenvironment{greenblock}[3]{%
  \setbeamercolor{block body}{bg=structure!10,fg=black}
  \setbeamercolor{block title}{bg=green!20,fg=black}
  \begin{block}{#1}}{\end{block}}

\newenvironment{redblock}[3]{%
  \setbeamercolor{block body}{bg=red!10,fg=black}
  \setbeamercolor{block title}{bg=red!20,fg=black}
  \begin{block}{#1}}{\end{block}}

%%% more color definitions

\colorlet{lightred}{red!30!white}
\colorlet{lightblue}{blue!30!white}
\colorlet{lightyellow}{yellow!30!white}
\colorlet{keylime}{green!10!white}
\colorlet{darkgreen}{green!50!black}

\def\docolor#1#2{\textcolor{#1}{#2}}
\def\highlight#1{\docolor{blue}{#1}}
\def\underscore#1{\underline{#1}}
\def\boldify#1{{\bfseries #1}}

\restylefloat{figure}

\newtheorem{thm}{Theorem}[section]
\newtheorem{lem}[thm]{Lemma}
\newtheorem{prop}[thm]{Proposition}
\newtheorem{cor}[thm]{Corollary}
\newtheorem{nt}[thm]{Note}
\newtheorem{exs}[thm]{Examples}
\theoremstyle{definition}
\newtheorem{dfn}[thm]{Definition}
%\theoremstyle{remark}
\newtheorem*{ex}{Example}
\newtheorem{rmk}[thm]{Remark}
\newtheorem*{keg}{Key example}
\newtheorem*{keyeg}{Key example}
\newtheorem*{term}{Terminology}
\newtheorem*{notn}{Notation}
\newcommand{\mt}{}
\newcommand{\del}{\ensuremath{\backslash}}
\newcommand{\la}{\langle}
\newcommand{\ra}{\rangle}
% \newcommand{\beq}{\begin{eqnarray}}
% \newcommand{\eeq}{\end{eqnarray}}
% \newcommand{\beqn}{\begin{eqnarray*}}
% \newcommand{\eeqn}{\end{eqnarray*}}
\newcommand\ix[1]{#1\index{#1}}
\newcommand{\ls}{\leqslant}
\newcommand{\gs}{\geqslant}
\newcommand\lset[2]{\left\{\left.#1\ \right|\ #2\right\}}
\newcommand\rset[2]{\left\{#1\ \left|\ #2\right.\right\}}
\newcommand{\ol}{\overline}
\newcommand{\mbb}[1]{\ensuremath{\mathbb{#1}}}
\newcommand{\mbf}[1]{\ensuremath{\mathbf{#1}}}
\newcommand{\bbfq}{\mathbb{F}_q}
\newcommand\bbp{\mathbb{P}}
\newcommand{\bbc}{\mathbb{C}}
\newcommand{\bbf}{\mathbb{F}}
\newcommand{\bbq}{\mathbb{Q}}
\newcommand{\bbr}{\mathbb{R}}
\newcommand{\bba}{\mathbb{A}}
\newcommand{\mcal}[1]{\ensuremath{\mathcal{#1}}}
\newcommand{\CC}{\mcal{C}}
%\newcommand{\CD}{\mbox{$\cal D$}}
\newcommand{\DD}{\mathcal{D}}
\newcommand{\CE}{\mcal{E}}
\newcommand{\CF}{\mcal{F}}
\newcommand{\CG}{\mcal{G}}
\newcommand{\CO}{\mcal{O}}
%\newcommand\weight{\operatorname{weight}}
\newcommand{\wt}{\operatorname{weight}}
\newcommand{\rank}{\operatorname{rank}}
\newcommand{\columnrank}{\operatorname{column~rank}}
\newcommand{\rowrank}{\operatorname{row~rank}}
\newcommand{\im}{\operatorname{Im}}
\newcommand{\ifff}{\Longleftrightarrow}
\newcommand{\ham}{\operatorname{Ham}}
\newcommand{\tran}{^\textup{T}}
\newcommand{\li}{linearly independent}
\newcommand{\ld}{linearly dependent}
\newcommand{\sgn}[1]{\ensuremath{\mathrm{sgn}(#1)}}
\newcommand{\krn}[1]{\ensuremath{\mathrm{Ker}(#1)}}
\newcommand{\ch}[1]{\ensuremath{\mathrm{char}(#1)}}
 \newcommand{\jmod}[2]{\mbox{ $\equiv #1$}\mbox{ \rm (mod $#2$)}}
\newcommand{\group}[2]{\ensuremath{(#1,\, #2)}}
\newcommand{\ring}[3]{\ensuremath{\langle #1,\, #2,\, #3 \rangle}}
\newcommand{\gl}[2]{\ensuremath{\mathrm{GL}(#1,\, #2)}}
%\newcommand\ex[1]{\emph{#1}\index{#1|btemp}}
\newenvironment{pfenum}{\begin{proof}\indent\begin{enumerate}\vspace{-\topsep}}{\qedhere\end{enumerate}\end{proof}}
\makeatletter
\def\imod#1{\allowbreak\mkern10mu({\operator@font mod}\,\,#1)}
\makeatother

\pagestyle{fancy}
\fancyhead[L]{Draft 1}
% \fancyhead[C]{Algebra}
\fancyhead[R]{\leftmark}

% \fancypagestyle{plain}{
% \fancyhead[L]{WTW 381}
% \fancyhead[C]{Algebra}
% \fancyhead[R]{\leftmark}}

\renewcommand{\sectionmark}[1]{\markboth{\thesection.\ #1}{}}

\renewcommand{\headrulewidth}{0.4pt}

\addtolength{\headheight}{\baselineskip}

\sloppy

\begin{document}

\title{Title}

\author{B. Nortier\\
% {\small \it Department of Mathematics \& Applied Mathematics, University of Pretoria, Pretoria 0028} \\ {\small Email: u17091820@tuks.co.za \& } \\
}
\date{2021}
\date{}
\maketitle

\begin{abstract}
Abstract goes here.
\end{abstract}

% {\bf Keywords} Data-driven dynamical systems, Koopman Operator, Causal Embedding, Nonautonomous Dynamical Systems.

% \newpage

\section{Introduction}
Traditional methods for modelling physical systems involves deriving governing equations by
considering the world and discover some fundamental relationships between encountered
quantities and phenomena. These fundamental relationships are then expressed mathematically and
used as the foundations to derive more complicated relationships. Consider the example of
Newton’s law F=ma
However our world around is, for the most part, much more complex than that which can be
distilled into elegant equations. Many of the more complex systems we don’t fully understand, nor
even have good physical intuition of.
To complicate this – many of the physical phenomena we observe in nature are chaotic.
Mathematically, this would imply what we call sensitive dependence on initial conditions. Despite
having highly similar initial conditions, differing orbits will diverge quickly and by such a large degree
that it becomes seemingly impossible to retrace them steps back to the origin conditions.The
presence of the smallest of computational noise or measurement error makes long-term pointwise
prediction very difficult to minimise.
We may however exploit the environment of data-overload we currently find ourselves in. This is
one where an incredible amount of information is being gathered and generated on a daily basis.
Alongside this we have also witnessed the advent of greater computing power and storage.
This project work focusees on constructing governing models for real-world systems in some sense
‘blindly’, by making use of the abundance of information we have at our disposal. An advantage of
this approach is the fact that we need only make a very few starting assumptions about the system
we choose to investigate, allowing the methodology it to be greatly generalised. In establishing 
topological conjugacy between underlying systems we can then, amongst others, establish long-
term accurate results.

By solidifying the mathematical underpinnings of our theory, we may guarantee having the ability to
construct models with predictive power ranging from molecular biology to neuroscience.
The goal of the methodology described here is to be able to predict the dynamics into the future by
constructing a model that will accurately mimic the dynamics of a system and hence also provide a
framework for future evolution of a dynamical system.\\

\emph{Additional to Intro?}\\
Examples in literature attempting to forecast data from such systems abound: One could model the
unobserved states as stochastic quantities. Or one could map the data to a higher-dimensional space
by means of delayed-time map of observations and then hope to see some of the the unobserved
characteristics of the data revealed.

\newpage

\section{Possible addition or Post-introduction:}

\emph{Expand on non-autonomous systems, their unique challenges and chaotic dynamics?}

Long-term consistency while modeling \textbf{chaotic} dynamical systems is an issue. When systems have 
sensitive dependence on initial conditions, negligible numerical errors multiply within a very short 
time gap to the point that it becomes unfeasible to track specific orbits.
We’ll focus on case where attractor is present.

\section{Setup and Learning Problem}

Consider a relatively simple learning problem:

Given ($u_{0}$, $u_{1}$, …, $u_{m}$), forecast um+1, um+2, …, given that $u_{n}$ is defined by relation  $u_{n+1} = Tu_{n}$.

However, T is unknown.

More involved learning problem: We have $\theta$(u0), $\theta$(u1), …, $\theta$(um) in an unknown dynamical system (U,T). 
Forecast $\theta$(um+1), $\theta$(um+2)

\textbf{Now consider a more refined scenario:}
\begin{enumerate}
  \item  a continuous time system with true state u in U (input space)
  \item Measured signal = $x_{n}$ (where: $x_{n}$ homeo x(u(tn)) + $\eta_{n}$ <-- noise 
\end{enumerate}

We have a scalar time-series.

We construct multi-dimensional observable

How? 
\begin{enumerate}
  \item Delay-coordinates: Given {x1,x2,$\ldots$}: define $x_{n}$ = [$x_{n}$,$x_{n}$-$\tau$,$\ldots$, $x_{n}$-(d-2)$\tau$,$x_{n}$-(d-1)$\tau$].T $\in\mathbb{R}^{d}$ 
  \item $\tau$=lag
  \item d=embedding dimension (park those two for now)
\end{enumerate}

Given: $x' = f(x)$ where $x(t)\in\mathbb{R}$  (x(t) are the states, f is assumed smooth)

Define flow: $\Phi$(x(t0),M)=x(t0+M), $\Phi$:U$\times$R $\rightarrow$ U

we can relate the two:

d/dt$\Phi$ (x(t0),t) = d/dt x(t0+M) = f(x(t0+M)) =f($\Phi$(x(t0),t)

Now define time-map $\Phi$M : x(t0+M)= $\Phi$T(x(t0))

i.e. x(t0+kM) = $\Phi$k

M(x(t0))= $\Phi$M$\circ$ $\Phi$M $\circ$$\ldots$$\circ$ $\Phi$M(x(t0) = k-fold composition

BUT: we only get to see a 1D time series $\theta$(x(t)), where $\theta$: R

N$\rightarrow$R is a smooth observation/measurement function.

Can information about the system state x(t) be retained in this time series data $\theta$(x(t))?

If we know f, then easily. But we don’t.

\textbf{However: Takens’ theorem gives some hope by saying yes! }

\subsection{Some side-questions}
\emph{Let’s talk about the fact that the manifold U is an attractor etc? Takens only holds for attractors, so
that is then why we immediately “restrict” U by only considering the subset containing the attractor
right? (ie. Only consider the space occupied by attractor. This is then why we demand that U is
compact and T surjective?
What about density of attractors? (1)}

$\ldots$

\section{Takens Theorem}
\begin{enumerate}
  \item What's that?
  \begin{enumerate}
        \item Takens defines delay-coord-map F$\theta$
        \begin{enumerate}
              \item F$\theta$ : [$\theta$(T-2du), $\ldots$ , $\theta$(T-1u), $\theta$(u)] $\rightarrow$ [$\theta$(T-2d+1u), $\ldots$ , $\theta$(T-1u), $\theta$(Tu)]
              \item Where T is flow T$u_{n-1}$ = $u_{n}$ (previously our flow was $\Phi$)
              \item $\theta$ is measurement function,
        \end{enumerate}  
        \item When confining DS to manifold U (in which system is contained), Takens showed
        that, if certain smoothness conditions are satisfied on T and $\theta$ then delay-coord
        map F$\theta$ is embedding of U onto the reconstruction space for \emph{almost every
        choice(2)} of measurement function $\theta$ (the observable)
        \item Or alternatively: Takens' embedding theorem guarantees that almost all
        dynamical systems can be reconstructed from just one noiseless observation
        sequence.
        \item i.e. for great number of possible observation functions $\theta$, F$\theta$ preserves the
        topology of U
        \item \emph {See Graph 1 }
        \begin{enumerate}
          \item If we are in a state u, and then evolve towards state Tu before  embedding into R(2m+1), then it is the same as embedding into R(2m+1) and then evolving through the map F$\theta$
        \end{enumerate}
        \item Thus: information about U can be retained in the time series’ (observation)
        output. By preserving the topology of the manifold U in the reconstruction space
        X, topological invariants of manifold are preserved, of which dimensionality is
        one (again considered later important later. (3)
  \end{enumerate}
  \item Problems with Takens
  \begin{enumerate}
    \item Even if we can find F$\theta$, our approximation of F$\theta$ is a map from a larger set R2m+1
    containing the embedded attractor. No theoretical guarantees that F$\theta$ will keep
    this attractor the same. (should have defined attractor earlier?)
    \item Takens only holds for noiseless observations.
        \begin{enumerate}
          \item Assume F was learnt but due to measurement error/noise we get V, not an attractor, then will tend somewhere else. This problem will later be removed by some properties of g (USP/UAP)
        \end{enumerate}
    \item Takens only holds if T is invertible. 
    \item  Functional complexity and stability of embedding F$\theta$ depend on choice of  observable $\theta$
  \end{enumerate}
\end{enumerate}

Now: Although we cannot necessarily find a map T:U $\rightarrow$ U, there does exist a map from
$\Phi$: $\Phi$2d,f(um) $\rightarrow$ $\Phi$2d,f(um+1) whenever U is a manifold of dimension less than d (should we say more
about d?)

\emph{Graph 2}
(In sketch: W==U )

\textbf{We’re one step closer to recovering information about the system states}

\section{Problem}

We're trying to approximate a map that might not even exist.

\emph{Goal:} 
predict dynamics into future and construct model from data which mimics dynamics.

\section{So how do we guarantee existence?}

\begin{enumerate}
  \item Assumptions
  \begin{enumerate}
      \item assume a compact, discrete time, dynamical system (U,T) (implicitly T cts)
      \begin{enumerate}
          \item T must be surjective. (This is important, we return to it later)  
      \end{enumerate}
      \item  A driven dynamical system g, who takes input in U. (implicitly g is cts)
      \begin{enumerate}
         \item g will be randomly initiated neural net <--  wil explain why later
        \item  g($u_{n}$-1, $x_{n}$-1) = $x_{n}$
      \end{enumerate}
      \item Bi-infinite sequence u <-- physically: running for long time
  \end{enumerate}

  \item We can easily show existence of the function g:
  \begin{enumerate}
    \item   g = identity on X 
    \newline But then: g ignores the input u and any constant bi-infinite sequence would be a solution to input u.
    \item  However this gives trivial soln, can we find something better?
    \item  Define: reachable soln space XU, entire solutions, YT
    \begin{enumerate}
      \item Give own example of entire sol'n
      \item \emph{Graph 3}
    \end{enumerate}
  \end{enumerate}
  \item Choosing g
  \begin{enumerate}
      \item Choosing g st “not quench temporal structure in u”
      \item(4)
      \item  g must be cts, invertible (for SI-invertibility)
      \item Specific w.r.t tanh:
       \newline \emph{Describe set of all solutions of g for given input u}
  \end{enumerate}
\end{enumerate}

\section{Can we show uniqueness of an entire solution?}

\emph{Graph 4}

\begin{enumerate}
  \item UAP: regardless of starting position, all trajectories converge to a single trajectory. This
  trajectory is the unique solution sequence x to the input sequence u from the previous
  slide.
  \item  USP: \emph{insert more info here}
  \begin{enumerate}
      \item The USP - we can start the system anywhere and in time it will forget
      \item  No new complexity
      \item USP $\equiv$ g is topological contraction $\equiv$ UAP (we’ll show this) 
      \begin{enumerate}
            \item 1. Define topological contraction      
      \end{enumerate}
  \end{enumerate}
  \item but infinite sequences are difficult to manage
  \begin{enumerate}
      \item   i. If both SI-invertibility and USP hold for g, then our causal mapping becomes much simpler: we can drop the infinite sequence to focus only on the most recent two terms. (Call it H2)
      \newline \textbf{AND} H2 becomes embedding    
  \end{enumerate}
\end{enumerate}

\section{Establishing Embedding and Conjugacy}

\begin{enumerate}
        \item Define: Embedding, Causal Embedding, Topological Conjugacy
        \begin{enumerate}
            \item Explain conjugacy more perhaps?  
        \end{enumerate}
        \item \emph{Graph 5}
        \item SI-invertibility, guaranteeing us the function $G_{T}$ on X$\times$X. <--- theoretical basis to try approximate!
        \item If g is SI-invertible and {$u_{n}$} in U is an orbit of T then:
        \begin{enumerate}
          \item   $G_{T}$ exists
          \item if g also has USP then (YT, $G_{T}$) top.semi-conjug to ($\hat{U}$, $\hat{T}$)
          \item if T:U $\rightarrow$ U is homeo then (YT, $G_{T}$) is top.conjug to ($\hat{U}$, $\hat{T}$) and top.conjug to (U,T) so g is causal embedding
          \item WE can learn single-delay dynamics
          \begin{enumerate}
            \item forecast future values of $G_{T}$
            \item forecast future values of $u_{n}$
          \end{enumerate}
        \end{enumerate}
        \item Question: why can one learn single-delay lag dynamics of the driven states through $G_{T}$
        with \textbf{enough} data?
\end{enumerate}

\section{Why $G_{T}$?}

Why do we do this?

\begin{enumerate}
  \item Essentially embed attractor into higher dimensional space X$\times$X.
  \item  Hence: more `dimensional room' for the dynamics to move, and we might hope
  that the dynamics of $G_{T}$ is in some sense simpler than that of T.
  \newline \textbf{OR} 
  \newline  Embed into state space with much higher dimension s.t. unobserved features of sequential iput data ($u_{n}$) can emerge in ($x_{n}$)
\end{enumerate}

\section{Progression to $\Gamma$ma instead of $G_{T}$}

\begin{enumerate}
  \item Define $\Gamma$:($x_{n}$-1, $x_{n}$) $\rightarrow$ $u_{n}$
  \begin{enumerate}
      \item Existence? $\Gamma$ exists when G exists  
  \end{enumerate}
  \item New conjugacy  
  \newline \emph{Graph 6}
  \item Why not $G_{T}$?
  \begin{enumerate}
    \item dimension on which $G_{T}$ is defined can be high (wasn’t it the goal to increase dimension?)
    \item  If we know $\Gamma$ then we can drive the system autonomously. 
    \newline  And then we know $G_{T}$ anyway
    \item Also: numerical errors could accumulate  when working
    \newline \emph{Graph 7}
  \end{enumerate}
\end{enumerate}

\section{Programmatically} 
\begin{enumerate}
  \item Implementation
  \begin{enumerate}
    \item Explain the choice of g
    \item Explain DC, SDD, MDD, AMDD
    \item Sparse matrices
    \item Explain why RNN is chosen (above, for example, FNN)
  \end{enumerate}
  \item Some Examples
  \begin{enumerate}
    \item Double Pendulum successes
    \item Data dying down (ie. Not surjective and make more remarks here on surjectivity as promised in the beginning)
    \item Attractors from previous article such as Lorenz, Henon
    \item Other attractors I’ve experimented with.
  \end{enumerate}
\end{enumerate}

\section{Comparison with/Mention of previous work}

\begin{enumerate}
  \item Mention previous work, such as SyndiPi, and that it doesn’t work
  \item  Data-driven approaches alongside machine-learning algorithms have outperformed the classic delay-coordinate embedding as formulated by Takens in the prediction of chaotic dynamical systems’ evolution.
\end{enumerate}

\section{Further Work}

MD

\section{More things to Add}
\emph{Unsure here whether it's important}

\begin{enumerate}
  \item Defining attractors
  \item Instability due to perturbation and how prevented
  \item Defining a process?
  \item \emph{Graph 8}
  \item State contractions and their importance
  \item  More actual theorems
  \item Worry:
  \begin{enumerate}
    \item  So far I have very little “mathematical” theorems or proofs. Is this a problem
    \newline This will, however, probably be added when expanding on attractors and USP
  \end{enumerate}
  \item Should I show/state that USP == UAP (equivalent)?

\end{enumerate}

% \section{Definitions}

% \begin{dfn}{Discrete-time autonomous system}
%   A discrete-time autonomous system on a state space $X$ is given by the map $g:X\mapsto{X}$, where the dynamics are generated by compositions of $g$ with itself.
% \end{dfn}

% The system evolves according to the equation $x_{n} = g(x_{n-1})$ for $n\geq1$ and generates the sequence
% $\ldots, x_{-1}, x_{0}, x_{1},\ldots$.

% \begin{dfn}{Discrete-time nonautonomous systems}
%     A discrete-time nonautonomous system on a state space $X$ is then a collection of continuous maps ${g_n}$ such that $g_n:X\mapsto{X}$ satisfying the equation
%     $x_{n} = g_{n-1}(x_{n-1})$ at each timestep $n$.
% \end{dfn}

% Now consider a subset of this discretised version of nonautonomous systems in which external factors are taken into account.

% \begin{dfn}{Input-driven system}
%   An input-driven system is understood to be a nonautonomous, discretised dynamical system on the state space $X$ consisting of an input, or driving, sequence, 
%   ${u_{n}}\subset{U}$ and a continuous function $g:U\times{X}\mapsto{X}$ such that at time $n$ the state $x_{n}$ is determined by $x_{n}=g(u_{n-1}, x_{n-1})$
% \end{dfn}

% \begin{rmk}
%   $g^{n}$ denotes the self-composition of the function $g:U\times{X}\mapsto{X}$ n times.  
% \end{rmk}

% \begin{dfn}
% Discrete-time state space models of the form 
% \begin{equation}
%     x_{n+1} = g(u_{n}, x_{n})
% \end{equation}
% \begin{equation}
%     y_{n+1} = f(x_{n+1}),
% \end{equation}
% where $n \in \mathbb{Z}, u_{n}$ belongs to an input space $U$ and state $x_{n}$ belonging to a state space $X, g: U \times X \rightarrow X$, and a measurement $f:X \rightarrow Y$
% \end{dfn}

% \begin{dfn}
% The dynamics on $X$ are generated by the update equation $x_{n+1} = g(u_{n},x_{n}$
% \end{dfn}

% \begin{dfn} {Entire solution} \\
% Given a driven system $g$ and an input $\overline{u}$, we call a sequence $\{x_{n}\}_{n\in\mathbb{Z}})$ an entire-solution, or just a solution, if it satisfies the update equation $x_{n+1} = g(u_{n},x_{n}$ for all $n\in\mathbb{Z}$
% \end{dfn}

% \begin{dfn} {Entire solution of process a $\phi$} \\
%   An entire solution of a process $\phi$ on the state space $X$ is a sequence $\{x_{n}\}_{n\in\mathbb{Z}}$ where $x_{n}\in{X}$ for all $n\in\mathbb{Z}$ and 
%   $\phi(n,m,x_{n})=x_{n}$ whenever $n\leq{m}$.
% \end{dfn}

% \begin{dfn} {$\phi$-Invariant Set}
%   The set $\mathbb{A} = \{A_{n} : A_{n} \subset X\}$ where $\mathbb{A}_{m}\subset X$ for all $m$ and $\phi(n,m,A_{m})=A_{n}$ for $n\geq{m}$.
% \end{dfn}

% \begin{rmk}
%   Note: when we speak of g, then we refer implicitly to the entire underlying dynamical system.
% \end{rmk}

% The infinite product space $\overline{Y}$ is defined as follows: $\overline{Y}:=\prod_{i=-\infty}^{\infty}{Y_{i}}$ such 
% that $Y_{i} \equiv$.

% \begin{dfn} {Unique Solution Property} \\
% A driven system $g$ has the unique solution property (USP) if for each input $\overline{u}$ there exists exactly one solution. \\
% Expressed more rigourously, we may state that $g$ has the USP if there exists a well-defined map $\Psi:\overline{U}\mapsto\overline{X}$.
% $\Psi(\overline{u})$ then represents the unique entire solution when the input is $\overline{u}$. 
% % Alternatively, g has the USP if there exist a well-defined map $\Psi$ between the infinite product space $\overline{U}$ and $\overline{X}$, where $\overline{Y}:=\prod_{i=-\infty}^{\infty} Y_{i}$ and $Y_{i}\equivY$ so that $\Psi(\overline{u})$ denotes the unique solution obtained from the input $\overline{u}$.

% \end{dfn}

% \begin{dfn}
% Echo State Property or Topological Contraction
% \end{dfn}

% \begin{thm}{Contraction Mapping Theorem}
%   $f:M\rightarrow M$ where $(M,d)$ is a metric space then $\exists k\in[0,1]$ such that (s.t.) $\forall{x,y}\in\mathbb{R}$ we have
%   $d(f(x),f(y)) \leq d(x,y)$
% \end{thm}

% \begin{dfn}
% Dissipative System
% \end{dfn}

% \begin{dfn} {Basin of Attraction }
% Basin of Attraction for a stable fixed point
% \end{dfn}

% \begin{dfn}
% Local Irreducible Attractor
% \end{dfn}

% \begin{dfn}
% Pullback Attractor
% \end{dfn}

% \begin{dfn}
% Forward Attractor
% \end{dfn}

% \begin{dfn}
% Uniform Attraction Property:
% $\forall \epsilon \exists N \in \mathbb{N} s.t. \forall y_{0} \in X \forall \overline{u}\in\overline{U}, d(x_{n}, y_{n})<\epsilon holds for n>\mathbb{N}.$ Here $y_{n+1}:=g(u_{n},y_{n})$ for $n\geq0$ and $\{x_{n}\}$ is a solution for input $\overline{u}$
% \end{dfn}

% \begin{dfn}
%     $\hat{U}_{T} = $ the space of all left-infinite T-orbits = $\{(\ldots, u_{-2}, u_{-1} : Tu_{n} = u_{n+1})\}$
% \end{dfn}

% \begin{dfn}
% Left-Infinite Sequence Space $\overleftarrow{U}$ where

% $\overline{Y}:=\prod_{i=-\infty}^{\infty} Y_{i}$ and
% $Y_{i} \equiv Y$ equipped with the product topology.

% \end{dfn}

% \begin{dfn}
% The left-infinite fragment of any bi-infinite input sequence $\overline{u}$ is denoted by $ \overleftarrow {u}^{n}:=($\ldots$,u_{n-2}, u_{n-1})$
% \end{dfn}

% \begin{dfn}
% The reachable set of a driven system ${g}$ is the union of all elements of all possible solutions. \\
% $X_{U} := \{ x \in X : \exists$ a solution $\{x_{n}\}$ for some $\overline{u}$ and $x=x_{k}$ for some $k \in \mathbb{Z}$
% \end{dfn}

% \begin{rmk}
% An input value is denoted as $v$.
% \end{rmk}

% \begin{dfn}
%   Orbit of T
% \end{dfn}

% \begin{dfn}
%   \begin{equation}
%     Y_{T} = \{ (x_{n-1}, x_{n}) : \{x_{k}\}  \textrm{is a solution of some orbit of} T \}  
%   \end{equation}
% \end{dfn}

% \begin{rmk}
%   The following notation is used:\\
%   $\overleftarrow{u} = (\ldots, u_{-2}, u_{-1})$ \newline
%   $\overleftarrow{u}^{n} = (\ldots, u_{n-2}, u_{n-1})$
% \end{rmk}

% % \begin{def}
% % The update of the input is defined as $u_n$. 
% % \end{def}

% \begin{dfn}
% The input up to time $n$ with an input $v$ at time $n$ is denoted 
% $\overleftarrow{u}^{n}v:=(\dots,u_{n-2}, u_{n-1}, v)$.
% \end{dfn}

% \begin{dfn}
% H is a causal mapping if $H:\overleftarrow{U}\rightarrow\overleftarrow{X}_{U}$ is a continuous and surjective map such that 
% \end{dfn}

% \begin{dfn}{Topological Conjugacy}
% A homeomorphism between two dynamical systems (U,T) and (V,S) is a a function $\phi:U\rightarrow V$ such that $\phi\circ\ T = S\circ\phi$.\\ We then say that 
% (V,S) is \emph{topologically conjugate} to (U,T).
% \end{dfn}

% When $\phi$ is merely continuous and not a homeomorphism, we call the map a topological semi-conjugacy. This many-to-one relationship guarantees at least a 
% coarse-grained depiction of the original system.

% \begin{dfn}
%   Universal Semi-Conjugacy: A map $h:\overleftarrow{X}\rightarrow X_{U}$ which is continuous and surjective s.t. $\forall v\in U$ the diagram commutes: \\
%   \textbf{XXX}
%   \end{dfn}

% \begin{dfn}{Embedding}
% A homeomorphism $f:Z\rightarrow Y$ for $Y\in X$. \\
% We then say that $Z$ is embedded in $X$ by $f$
% \end{dfn}

% \begin{dfn}{Observable}
% A one-dimensional time-series $y(t)=\phi(x(t))$, where $\phi:\mathbb{R}^{n}\rightarrow\mathbb{R}$ is a smooth so-called observation or measurement function.
% \end{dfn}

% \begin{dfn} {Process}\\
%   Consider the subspace $\mathbb{Z}_{\leq}^{2} := {(n,m) : n,m \in \mathbb{Z}, n\geq{m}}$. \\
%   The process $\phi$ is a mapping $\phi:\mathbb{Z}_{\leq}^{2}\times{X}\mapsto{X}$ satisfying the following:
%   \begin{enumerate}
%     \item $\phi(m,m,x)=x $ \quad for all $ m\in\mathbb{Z}, x\in{X}$
%     \item  $\phi(n,m,x)= \phi(n,k,\phi(k,m,x)) \quad $ for all $ m,k,n \in\mathbb{Z}$ where $m\leq{k}\leq{n}$, $x\in{X}$
%     \item $\phi(n,m,\dot)$ is continuous on $X \forall{n,m}$ 
%   \end{enumerate}
% \end{dfn}

% % Process: Two-parameter semigroup formalism (10,15) defined for nonautonomous systems

% \begin{thm}
%   A nonautonomous system $g_{n}$ on $X$ exists if and only if a process $\phi$ on $X$ exists.
% \end{thm}

% The proof is almost trivial:
% \begin{proof}
%   A process $\phi$ on $X$ generates a nonautonomous system $\{g_{n}\}$ if we define $g_{n}(x):=\phi(n+1,n,x)$. In so doing we obtain the 
%   system state $x_{n}$ at time $n$ by $\phi(n,m,x)$ where $x_{0}$ was the initial state at time $m$.

%   Conversely, the nonautonomous system  with $\{g_{n}\}$ generates a process $\phi$ on $X$ if  we define as follows: 
%   $\phi(m,m,x):=x$ and $\phi(n,m,x):= g_{n-1}\circ{\ldots}\circ{g_{m}(x)} = x_{n}$.   
% \end{proof}

% % the nonautonomous system $(X, g)$

% \begin{dfn} {Flow} \\
%   The unique global solution $\Phi(t,x_{0})$  of $x'=f(x)$ passing through some $x_{0}$ at $t=0$ where
%    $\Phi$ satisfies the following properties:
%    \begin{enumerate}
%     \item $\phi(0,x_0)=x_{0}$
%     \item $\phi(t,x_{0})$ is continuous on its domain
%     \item $\Phi(t+s,x_{0})=\Phi(t,\Phi(s,x_{0}))$
%    \end{enumerate}
% \end{dfn}

% It follows that $\Phi(t,x_{0})$ attains the value $x_t$ at time $t$, and so we may define the function 
% $T(x_0):=\Phi(M,x_{0})$ , which can be referred to as the \emph{Time-T Map}. 
% \newline

% Now consider the following: \\
% \begin{equation}
%   T\circ{T(x_0)} = T\circ{\Phi(M,x_{0})} = \Phi(M,\Phi(M,x_{0})) = \Phi(2M,x_{0}) = T^{2}(x_0)
% \end{equation}

% In general then: {  $T^{n}(x_{0}) = $ the value of the solution after  nxM  time-units} and any solution $(x_k)$in U is of 
% the form ${x, Tx, T^{2}x, \ldots}$, i.e. it satisfies the difference equation 
% \begin{equation}
%   x_{k+1} = T(x_k)
% \end{equation}

% \begin{dfn}{Dynamical System} \\
% A tuple $(U,T)$ for an attractor/state space U which is a compact metric space and a function T that is not necessarily continuous.
% \end{dfn}

% It may easily be verified that if T is obtained from a flow, then T is a homeomorphism. \emph{CITATION}\\

% One can also assume, without loss of generality, that T is surjective. To see this, consider 
% $\ldots$  \emph{CITATION}

% The evolution of a dynamical system may happen smoothly over time or over discrete time-units. Though evolution in nature occurs
% in a continuous fashion, we may only observe this change in discrete-time due to the fact that we can only gather data through
% observations at specific intervals of time. It is thus necessary to discretise the system. \emph{(which automatically arise per dis-
% cretization in simulations on digital computers).}

% \bigskip
% Assume that the system has state $u$ in the input space $X$. We cannot directly observe u, (for if we did then this project would 
% be superfluous). Rather, we have some observed signal $(x_{n})$ of the form
% \begin{equation}
% x_{n} = x(u)(t_n) + \nu_{n}
% \end{equation}
% for measurement noise $\nu{n}$.

% Measurement noise can be a great hindrance to accurate reconstruction of the system and could even become indistinguishable from a random time series. 
% \emph{The case of low-D data from high-D system}

% \subsection{Functions}

% \begin{equation}
%     \Gamma: Y_{T} \mapsto {U} \textrm{ s.t. } \Gamma:(x_{n-1}, x_{n}) \mapsto u_{n}
% \end{equation}

% \begin{equation}
%   \hat{T}(\overleftarrow{u_n}) = \hat{T}(\ldots, u_{-2}, u_{-1})
%   % & = (\ldots, u_{-2}, u_{-1}, Tu_{-1}) 
%   % & = (\ldots, u_{-2}, u_{-1}, u_{0})
% \end{equation}

% \bigskip
% \newpage

% \bigskip

\end{document}

### 2022-01-06 Gandhi_Project_Report/draft2.tex
Done working through all edits Ch1-Ch3

@@
 
-\chapter{Introduction}
+\chapter{Introduction}\label{ch1}
 
@@
 The report is organised as follows: 
-\newline In \ednote{M: Please label Sections as Chapters as this is apt for a thesis; for the paper that we will write soon, it would be in Sections.
-Please note the labelling in the tex file, and you may like it to be cross references like this}Chapter~\ref{Chapter_Ds}, we recall the definition of a discrete-time dynamical system, how a discrete-system arises from a flow of an ODE and then proceed to define the inverse limit space and topological conjugacy of autonomous systems.
-\newline In Section 3, we introduce the problem of forecasting dynamical systems, state the Takens delay embedding theorem and discuss various issues faced while forecasting. 
+\newline In Chapter~\ref{ch2}, we recall the definition of a discrete-time dynamical system, how a discrete-system arises from a flow of an ODE and then proceed to define the inverse limit space and topological conjugacy of autonomous systems.
+\newline In Chapter~\ref{ch3}, we introduce the problem of forecasting dynamical systems, state the Takens delay embedding theorem and discuss various issues faced while forecasting. 
 \newline In Section 4, a driven dynamical system is defined and discuss the properties of a specific class of driven dynamical systems that we make use of in this thesis/report/paper. 
@@
 
-\chapter{Discrete-time Dynamical Systems} \label{Chapter_Ds}
+\chapter{Discrete-time Dynamical Systems} \label{ch2}
 
@@
 
-Formally, a function $T: U \to U$, where $U$ is some set is  a \emph{discrete-time dynamical system} and its iterates $\{u,Tu,T^2u,\ldots\}$, where $T^n$ denotes the $n$-fold composition of $T$ with itself, describe the evolution of an initial condition $u\in U$ (Note that we drop the brackets and denote $T(u)$ by $Tu$ frequently for simplicity in notation).
+Formally, a function $T: U \to U$, where $U$ is some set is  a \emph{discrete-time dynamical system} and its iterates $\{u,Tu,T^2u,\ldots\}$, where $T^n$ denotes the $n$-fold composition of $T$ with itself, describe the evolution of an initial condition $u\in U$ (Note that we frequently drop the brackets and denote $T(u)$ by $Tu$ so as to simplify notations).
 
@@
 To see this, consider a differential equation $\dot{x} = f(x)$, $f: \mathbb{R}^n \to \mathbb{R}^n$, $n\in\mathbb{N}$ given to have a unique solution passing through
-each point $x\in\mathbb{R}^{n}$, the flow of the equation \label{defn_flow} is defined to be a mapping $\varphi: \mathbb{R}^n \times \mathbb{R} \to \mathbb{R}$ that satisfies the following properties \ednote{M: Please change this or revert back to how it was defined earlier.
-We are describing $\varphi$ and in the first two points, it does not concern $\varphi$}
-\vspace{-8mm}
-\begin{enumerate}[noitemsep, label=\roman*.]
-  \item $x(t)$ is a solution of the differential equation,
-  \item $x(0)=x_0$, and
-  \item $\varphi(x_0,t) = x(t)$.
-\end{enumerate}
+each point $x\in\mathbb{R}^{n}$, call it $x(t)$ where $x(0)=x_0$.
+
+\begin{Definition}
+  [\bf Flow of an Equation] \label{Dfn_Flow}
+  The flow of the equation  $\dot{x} = f(x)$  is defined to be a mapping $\varphi: \mathbb{R}^n \times \mathbb{R} \to \mathbb{R}$ where $\varphi(x_0,t)= x(t)$ for the solution $x(t)$ with $x(0)=x_0$ 
+\end{Definition}
+
 By fixing $t=K \in (0,\infty)$, we can define the \emph{time-$K$ map} as  $T(x):= \varphi(x_0,K)$, and it is easily verified that $T\circ T(x_0) = \varphi(x_0,2K)$.
-In general the $m^{\mbox{th}}$ iterate of $x_0$ under $T$ would be the value of the solution of the ODE evaluated at time $mK$ with the initial condition $x_0$.
-Thus ordinary differential equations give rise to a discrete-time dynamical system by sampling the value of $x$ at time intervals $K$ units apart.
+In general the $m^{\mbox{th}}$ iterate of $x_0$ under $T$ would be the value of the solution of the ODE evaluated at time $mK$ with the initial condition $x_0$, i.e.
+$\varphi(x_0, mK)$.
+Thus ordinary differential equations give rise to a discrete-time dynamical system by sampling the value of the solution $x(t)$ at time intervals $K$ units apart.
 
@@
 can be approximated at the time instant $t+(m+1)h$ by iterating the equation 
-$$x(t+(m+1)h) = x(t+mh) + h f(x(t+mh)).$$ 
+\begin{equation}
+  x(t+(m+1)h) = x(t+mh) + h f(x(t+mh))
+\end{equation}
+
 Adopting more succinct notation by replacing $x(t+mh)$ with $u_m$, we rewrite the above equation as
-$$u_{m+1} = u_m + hf(u_m)$$
-or, in even simpler terms, as the discrete-time dynamical system with map $T(u) = u + hf(u)$, where $T$, $u$ and $f$ are understood to be as above.
+\begin{equation}
+u_{m+1} = u_m + hf(u_m)
+\end{equation}
+or in even simpler terms as the discrete-time dynamical system with map $T(u) = u + hf(u)$, where $T$, $u$ and $f$ are understood to be as above.
 
 Of course, discrete-time dynamical systems need not always arise through a differential equation, and once again we may consider the field of ecology where discrete dynamical systems are often directly derived or assumed.
+\textbf{Cite.}
 There is a school of thought that advocates discrete-time dynamical systems to be more natural for modelling real-world observations than differential equations and we refer the interested reader to \cite{saber2010introduction}.
-\textbf{too philosphical, or actually good?} \ednote{M: It is good!
-I have moved this paragraph}
 
-\subsection{Invariant Sets}
+\section{Invariant Sets}
 
-A core concept in the study of dynamical systems is that of invariance. Given a discrete-time dynamical system $T: U \to U$, a subset $A \subset U$ is said to be an \emph{invariant set} if $T(A) =A$. 
+A core concept in the study of dynamical systems is that of invariance. 
+\begin{Definition}
+  [\bf Invariant Set] \label{Dfn_Flow}
+  Given a discrete-time dynamical system $T: U \to U$, a subset $A \subset U$ is said to be an \emph{invariant set} if $T(A) =A$. 
+\end{Definition}
+
 We also define the \emph{orbit of $T$} to be a sequence $\bar{u} = \{u_n\}_{n\in \mathbb{Z}}$ obeying the update equation, $u_{n+1}=Tu_n$, $n \in \mathbb{Z}$.
 
-Two examples of invariant sets include a fixed point where $Tu=u$ for  $u\in U$, and a periodic orbit, i.e., a set of iterates $\{u,Tu, T^2u,\ldots,T^pu\}$, where $T^{p+1}u=u$ for some $p$.
+Two examples of invariant sets include a fixed point where $Tu=u$ for  $u\in U$, and a periodic orbit, i.e., a set of iterates $\{u,Tu, T^2u,\ldots,T^pu\}$, where $T^{p+1}u=u$ for some $p\in\mathbb{Z}$.
 The entire space $U$ could be also be invariant.
-Consider for example the space $U=[0,1]$, where $Tu=4u(1-u)$ and then $U$ is invariant. \emph{(as every $u\in{U}$ can be written as $u = 4x(1-x)$ for some $x\in{U}$)}
+Consider for example the space $U=[0,1]$, where $Tu=4u(1-u)$ and then $U$ is invariant (as every $u\in{U}$ can be written as $u = 4x(1-x)$ for some $x\in{U}$).
 
@@
 This is since the set of limit points of the orbit can be shown to be invariant \cite{de2013elements}.
-\ednote{M: I have edited here} So, when $U$ is compact,  the $\omega$-limit set $\omega(u;T)$ of a point $u$ defined to the collection of limit points of the sequence $\{x,Tu,T^2u,\ldots\}$ is nonempty, and $\omega(u;T)$ is invariant.
-For instance, for the map $Tu=u^2$ defined on $[0,1]$ all orbits lie in the invariant set $\{1\}$ or else would would approach the invariant set $\{0\}$.
-\ednote{B: not relevant other than being an additional example perhaps? M: Have added another simple example}.
+So, when $U$ is compact,  the $\omega$-limit set $\omega(u;T)$ of a point $u$ defined to the collection of limit points of the sequence $\{x,Tu,T^2u,\ldots\}$ is nonempty, and $\omega(u;T)$ is invariant.
 
-We discuss various properties of invariant sets: Invariant sets can be attracting or repelling depending on how orbits in its vicinity behave.
-Recall the examples $Tu =u/2$ and $Tu=2u$ defined on $\mathbb{R}$ ($\{0\}$ ``attracts" orbits) or the example, $Tu=u^2$ defined on $[0,1]$ ($\{1\}$ ``repels" orbits).
-We are interested in attractive invariant sets since they capture the long-term dynamics as time increases, and in particular the invariant sets that are called attractors whose definition we recall next.
-\begin{Definition} \rm Let $T: U \to U$, where $U$  is a metric space with metric $d$.
-A compact subset $A \subset U$ is said to be an attractor if it satisfies the three conditions:
+\begin{Example}
+  For the map $Tu=u^2$ defined on $[0,1]$ all orbits lie in the invariant set $\{1\}$ or else would would approach the invariant set $\{0\}$. 
+  % \ednote{B: not relevant other than being an additional example perhaps? M: Have added another simple example}.
+\end{Example}
+
+Invariant sets have various properties. Invariant sets can be attracting or repelling depending on how orbits in its vicinity behave. 
+Recall the examples $Tu =u/2$ and $Tu=2u$ defined on $\mathbb{R}$ (where $\{0\}$ ``attracts" orbits) or the example, $Tu=u^2$ defined on $[0,1]$ (where $\{1\}$ ``repels" orbits).
+We are interested in attractive invariant sets since they capture the long-term dynamics as time increases.
+In particular we are interested in those invariant sets named attractors.
+
+\begin{Definition} 
+  [\bf Attractor] \label{Dfn_Attractor}
+  Let $T: U \to U$, where $U$  is a metric space with metric $d$. A compact subset $A \subset U$ is said to be an attractor if it satisfies the three conditions: 
   \vspace{-8mm}
@@
 
-\subsection{Chaos}
+\section{Chaos}
 
-\ednote{ should one add paragraphs like these?
+% \ednote{ should one add paragraphs like these?
 M: this is a thesis, and not a paper; the instructions sent to the external examiner regarding the thesis evaluation says that the student suggest strongly that it is good to write}
-\newline In the 1960s, a number of mathematicians and mathematically interested scientists independently discovered chaos in the mathematical sense.
+In the 1960s, a number of mathematicians and mathematically interested scientists independently discovered chaos in the mathematical sense.
 The meteorologist Edward Lorenz may have been the first to explain this phenomenon in his 1963 paper \cite{lorenz1963deterministic}.
-The notions of invariance, attractivity and chaos may also bedescribed for continuous systems, and Lorenz's system comprised of a system of differential equations.
+The notions of invariance, attractivity and chaos may also be described for continuous systems, and Lorenz's system comprised of a system of differential equations.
 The narrative of Lorenz's discovery of chaos, as well as the history of other forerunners in this subject, are fascinating.
@@
 
-There are several intricately different definitions of chaos available in the literature, with each of them indicating some aspect of complexity \ednote{B:  what are we saying here?
-M: Devaney's definition is not the only one, there are about half a dozen definitions as well} In practice, it is not possible to verify which definition or notion does the underlying dynamical system satisfies when only data is observed from it.
+There are several intricately different definitions of chaos available in the literature \textbf{(Cite)}, with each of them indicating some aspect of complexity.
+\ednote%{B:  what are we saying here?
+M: Devaney's definition is not the only one, there are about half a dozen definitions as well}
+In practice, it is not possible to verify which definition or notion does the underlying dynamical system satisfies when only data is observed from it.
 We do, however, specifically recall the definition of chaos in the sense of Devaney \cite{devaney2018introduction,de2013elements} so as to understand some nuances behind the complexity.
@@
 
-\begin{Definition} \rm 
+\begin{Definition}
+  [\bf {Sensitive Dependence on Initial Conditions}] \label{Dfn_SDIC}
 A dynamical system $T: U \to U$ is said to have sensitive dependence on initial conditions (SDIC) if there exists a $\delta > 0$ such that for every $u \in U$ and in every neighborhood of $u \in U$ there exists a $v\in{U}$ and an integer $N:=N{(u,v)}\in\mathbb{Z}$ such that $d(T^Nu,T^Nv)>\delta$.
@@
 To fully comprehend all the subtleties of the concept, one needs to discuss it more thoroughly by considering each sentence with care and attention, whereafter one may examine how they fit together to convey the concept of sensitive dependence.
-\ednote{ $\leftarrow$ is this talking too simply perhaps? M: No! it is not so}.
 
@@
 
-\ednote{M: It is probably best to see a few numerical simulations to understand SDIC. (Please use the logistic map and perhaps the trajectories of the double pendulum data). }
+To illustrate the concept of SDIC, we present two examples - one from Mathematics and the other from the field of Physics.
+
+\begin{Example}
+  The logistic map(LM), a recursion relation of the form $x_{n+1}=rx_n(1-x_n)$ where $x_n\in[0,1]$ and $r<4$ is of particular interest to us as it is classically used to illustrate the influence of chaos on a system \textbf{(Cite)}.
+Below are plotted the first 75 iterates of the LM for $r=3.7$ and $x_0=0.2$ with an adjustment of 0.001 in $x_0$ and then $r$.
+
+  \begin{figure}[ht]
+    \includegraphics[scale=0.45]{log_sdic.eps}
+        \centering
+        \label{fig:log_sdic}
+        \captionof{figure}{Logistic Map for $r=3.7$, $x_0=0.2$}
+      \end{figure}
+
+  For an intuitive sense of the fact that the LM exhibits SDIC, we draw two green neighbourhoods of diameters $\delta_1=0.1, \delta_2=0.02$ around the points $u_1=0.7, u_2=0.4$ respectively.
+It is easy to see that every neighbourhood around these two points will contain some $v_1, v_2\in[0,1]$ and natural number $N_1, N_2 \in\mathbb{N}$ respectively s.t.
+the distance $d(T^Nu_1,T^Nv)>\delta_1=0.02$ and $d(T^Nu_2,T^Nv)>\delta_2=0.1$.
+
+  In a more rigourous fashion ....
+  \ednote{B:Should I be more rigourous here? }
+
+\end{Example}
+
+\begin{Example}
+  A second example pertains to a simple physical system called the double pendulum - a pendulum with another pendulum attached to its head given an initial position and angular velociy.
+
+  Below are two figures denoting the trajectories of a double pendulum's second head after some elapsed time.
+Depicted are three pendulum heads with equal angular velocity, but differing ever so slightly in initial position .
+As can be seen below, the trajectories diverge very quickly to become completely different.
+  The double pendulum will be discussed in much greater detail in a later chapter (\ref{ch5}), but we mention the example here to provide a numerical example of a practical system exhibiting SDIC.
+
+\begin{figure}[ht]
+  \centering
+\includegraphics[scale=0.8]{dp_sdic.eps}
+    \captionof{figure}{Three double pendulum heads with initial angle differing by 0,025 radians}\label{fig:dp_sdic}
+  \end{figure}
+
+\end{Example}
 
@@
 \begin{Definition}
+  [\bf {Topologically Transitive}] \label{Dfn_TopTrans}
 	A dynamical system $F: U \to U$ is topologically transitive if for any pair of nonempty open sets $E_1$ and $E_2$ there exists a $n\in\mathbb{N}$ such that $T^n(E_1) \cap E_2 \not= \emptyset$. 
@@
 In fact, it is topologically more likely that the choice of an arbitrary point will be one  whose iterates are almost dense.
-% I inserted a definition here: it makes more sense to me. 
-\begin{Definition} \rm
-A subset of a topological space $U$ is said to be a meager set if it cab be written as a countable union of sets of with empty interior. The complement of the a meager set is set to be a residual set.
+
+\begin{Definition} 
+  [\bf {Meager Set}] \label{Dfn_Meager Set}
+A subset of a topological space $U$ is said to be a meager set if it can be written as a countable union of sets of with empty interior. The complement of the a meager set is set to be a residual set.
 \end{Definition}
@@
 We may now define the notion of Chaos as formulated by Devaney\cite{devaney2018introduction}.
-\begin{Definition} \rm
+\begin{Definition}
+  [\bf {Devaney's Chaos}]\label{Dfn_ChaosDec}
 	A dynamical system $T: U \to U$ is said to exhibit chaos in the sense of Devaney if it satisfies the three properties:
@@
 
-\begin{figure}[ht]
-  \centering
-  \begin{minipage}{.5\textwidth}
-    \centering
-\includegraphics[scale=0.18]{chaos1.png}
-    \captionof{figure}{Starting to diverge}
-    \label{fig:chaos1}
-  \end{minipage}%
-  \begin{minipage}{.5\textwidth}
-    \centering
-    \includegraphics[scale=0.18]{chaos2.png}
-    \captionof{figure}{After more time-steps}
-    \label{fig:chaos2}
-  \end{minipage}
-  \end{figure}
+\begin{Example} 
+  The standard tent map $Tu=1-|2u-1|$ defined on $[0,1]$ is a well-known example of a dynamical system satisfying these three properties.
 
-\ednote{M: Please put this picture in perspective and refer after the definition of SDIC.
-You can also say that you pick two positions that are slightly different but with the same angular velocity.
-Also in the text, say that you describe the DP in a later chapter.}
-\ednote{M: I realize that need figures in the eps format to have the commutativity diagrams compile as well. }
-
-\begin{Example} \rm
-  The standard tent map $Tu=1-|2u-1|$ defined on $[0,1]$ is a well known example of a dynamical system satisfying these three properties. 
-\end{Example}
+  We now reason as to why this is true.
+The graph of the map $T$ is piecewise linear with two straight lines, one connecting the points $(0,0)$ and $(\frac{1}{2},1)$ and the other connecting $(\frac{1}{2},1)$ with $(1,0)$.
+They form a so-called tent with base centered at $u=1/2$.
+The graph of the map $T^2$ comprises two symmetric tents with their base centered at $1/4$ and $3/4$.
+See Figure \ref{fig:T2tentmap}
 
@@
         \centering
-    \label{fig:T2tentmap}
+        \captionof{figure}{Three double pendulum heads with initial angle differing by 0,025 radians}\label{fig:T2tentmap}
   \end{figure}
-
-Let us now consider why this map in the above example exhibits chaos in the sense of Devaneys.
-The graph of the map $T$ is piecewise linear with two straight lines, one connecting the points $(0,0)$ and $(1/2,1)$ and the other connecting $(1/2,1)$ with $(1,0)$.
-They form a so-called tent with base centered at $u=1/2$.
-The graph of the map $T^2$ comprises two symmetric tents with their base centered at $1/4$ and $3/4$.
-  % We now reason as why this map exhibits chaos in the sense of Devaney.
 
@@
   Given any point $u$ and an open neighborhood $W \subset[0,1]$ of $u$, we can find $k\in\mathbb{Z}$ large enough so that $T^k(W) = [0,1]$ because we can accommodate a tent whose base is contained in $W$.
-\ednote{B: This last phrase doesn't make sense to me.
+% \ednote{B: This last phrase doesn't make sense to me.
 M: The number of tents grow with $k$ and hence this is obvious; I explained in one of our meetings.
-Does it make sense?} This implies the existence of two points $w_1$ and $w_2$ in $W$ so that $d(T^kw_1,T^kw_2)= 1$, and by the triangle inequality  at least one of the inequalities $d(T^ku,T^kw_1)> \delta$ or $d(T^ku,T^kw_2)> \delta$ hold when $\delta\in (0,1/2)$.
+Does it make sense?}
+  This implies the existence of two points $w_1$ and $w_2$ in $W$ so that $d(T^kw_1,T^kw_2)= 1$, and by the triangle inequality  at least one of the inequalities $d(T^ku,T^kw_1)> \delta$ or $d(T^ku,T^kw_2)> \delta$ hold when $\delta\in (0,\frac{1}{2})$.
 So $T$ has sensitive dependence on initial conditions.
-\ednote{M: It is advised not to introduced newlines in a thesis.
-For the presentation, yes}
 
@@
 Since we have found a periodic point in the interval $W$,  the set of periodic points are dense in $T$.
-  \ednote{Expand further on reason for density. M: fixed point of $T^k$ is a periodic point, and we have found in an arbitrary interval.}
+  This follows because any fixed point of $T^k$ is also a periodic point, and we have found this in an arbitrary interval.% \ednote{Expand further on reason for density.
+M: fixed point of $T^k$ is a periodic point, and we have found in an arbitrary interval.}
+
+\end{Example}
 
@@
 To show topological similarity or sameness between two metric or topological spaces, one needs to establish a homeomorphism between the two spaces. 
-However, in the study of dynamical systems defined on two spaces, establishing a homeomorphism does not indicate that the systems are (physically) related in any way.
+However, in the study of dynamical systems defined on two spaces, establishing a homeomorphism does not indicate that the systems are dynamically related in any way.
 For instance, the maps $Tu=u^2$ and $Tu=1-|2u-1|$ defined on $[0,1]$ have totally different behaviour.
-\ednote{"related" - in what sense am I saying it here?
-Physically?
-M: Dynamically}  To find systems that are dynamically similar, one rather needs to establish a kind of dynamical equivalence which we illustrate in a simple commutativity diagram below.
+To find systems that are dynamically similar, one rather needs to establish a kind of dynamical equivalence which we illustrate in a simple commutativity diagram below.
+
 % Giving errors
@@
 
-\begin{Definition} \rm 
+\begin{Definition}  
+  [\bf {Conjugacy \& Semi-Conjugacy}]\label{Dfn_Conjugate}
   Consider the dynamical systems $T:U\to{U}$, $S:V\to{V}$ and suppose the relationship $\phi \circ T=S\circ \phi$ holds.
-If $\phi$ is a homeomorphism, then S is said to be conjugate to T and $\phi$ is called a conugacy.
-If we relax the criterion on $\phi$ and merely require $\phi$ to be continuous and surjective, $\phi$ is then a semi-conjugacy between $T$ and $S$; S is said to be semi-conjugate to T.
+If $\phi:U\to{V}$ is a homeomorphism, then $S$ is said to be conjugate to $T$ and $\phi$ is called a conjugacy.
+If we relax the criterion on $\phi$ and merely require $\phi$ to be continuous and surjective where $\phi:U\to{V}$, then $\phi$ is a semi-conjugacy between $T$ and $S$ where T has domain $U$ and S has domain $V$; S is said to be semi-conjugate to T.
 \end{Definition}
 
-\ednote{M: Saying $\phi$ is then a semi-conjugacy between $T$ and $S$ could be ambiguous as it does not specify which system is on the top in the commutativity. So please edit accordingly}
-
-When $S$ is conjugate to $T$, the dynamics of the two systems are in some way ''dynamically equivalent'' and specifically they are in one-to-one correspondence with one another.
-However when $S$ is semi-conjugate to $T$ with $\phi$ a many-to-one mapping, the dynamics on $V$ provides a coarse-grained description of the dynamics on $U$ \cite{de2013elements}.
-When $S$ is semi-conjugate to $T$, it is also common to call $S$ a factor of $T$, or conversely that $T$ is an extension of $S$.
+\ednote{M: Saying $\phi$ is then a semi-conjugacy between $T$ and $S$ could be ambiguous as it does not specify which system is on the top in the commutativity.
+So please edit accordingly.
+B: Does defining explicitly the domain and codomain of $phi$ make it clear enough?
+}
+When $S$ is conjugate to $T$, the dynamics of the two systems are in some way ''dynamically equivalent''.
+Specifically, they are in one-to-one correspondence with one another.
+However when $S$ is semi-conjugate to $T$ with $\phi:U\to{V}$ a many-to-one mapping, the dynamics on $V$ provide merely a coarse-grained description of the dynamics on $U$ \cite{de2013elements}.
+When $S$ is semi-conjugate to $T$, it is also common to call $S$ a \emph{factor of $T$}, or conversely that $T$ is an \emph{extension of $S$}.
 In essence, an extension (e.g., \cite{de2013elements}) is a larger system capturing all of the important dynamics of its factor.
@@
 However, one can verify that a function $\phi$ satisfies the commutativity diagram \ref{eqn_conjugacy}.
-For example $\phi(x)=\sin(\pi x/2)^2$ is a conjugacy between two systems $Tu=1-|2u-1|$ and $Sv=4v(1-v)$  defined on $U=[0,1]$.
 
-In establishing that two systems are conjugate (semi-conjugate) to one another, we have also shown that one may choose to work with one systems as opposed to the second and still be guaranteed to obtain information on the latter.
-This is incredibly useful while forecasting the future evolution of dynamical systems.
+\textbf{NB NB FINISH}
+\begin{Example}
+  For example $\phi(x)=\sin(\frac{\pi}{2}x)^2$ is a conjugacy between two systems $Tu=1-|2u-1|$ and $Sv=4v(1-v)$  defined on $U=[0,1]$.
 
-\chapter{The Learning Problem}
+  To see this, we consider $\phi\circ{T} = \sin(\frac{\pi}{2}(1-|2u-1|))^2$
 
-\emph{Introduction...}
+  and $S\circ\phi = 4\sin(\frac{\pi}{2}x)^{2}(1-\sin(\frac{\pi}{2}x)^2)$
+\end{Example}
+
+In establishing that two systems are conjugate (semi-conjugate) to one another, we have also shown that one may choose to work with one system as opposed to the second and still be guaranteed to obtain information on the latter.
+This is especially useful while forecasting the future evolution of dynamical systems.
+
+\chapter{The Learning Problem}\label{ch3}
+
+In this chapter we acquaint ourselves with the question of forecasting dynamical systems with unknown underlying structures, state and discuss the shortcomings of the Takens delay embedding theorem and discuss various issues faced while forecasting.
 
 Consider a relatively simple learning problem: 
+Given the sequence $(u_0, u_1, \ldots, u_m)$ for $m\in\mathbb{Z}$, a finite segment of an orbit of the map $T$ where $T$ is defined by the update equation $u_{n+1} = Tu_n$, forecast the values $u_{m+1}, u_{m+1}$ where the map $T$ is unknown, given that $u_m$.
 
-Given $(u_0, u_1, \ldots, u_m)$ for $m\in\mathbb{Z}$, a finite segment of an orbit of the map $T$ where $T$ is defined by the update equation $u_{n+1} = Tu_n$, forecast the values $u_{m+1}, u_{m+1}$ where the map $T$ is unknown, given that $u_m$.
-
-A practical example of this would be - given the sequential coordinates of an object moving in space, predict the future positions of that object.
+A practical example of this would be the subsequent scenario: given the time-sequential coordinates of an object moving in space, predict the future positions of that object.
 However we are very rarely, if at all, presented with a problem where the entire state information is available to us in the form $u_n$ at some time-step $n\in\mathbb{N}$.
@@
 
-Suppose we only have the observations $\theta(u_0), \theta(u_1), \ldots, \theta(u_m)$ of the true system states $u$ in an unknown dynamical system $(U,T)$ and we wish to predict the values $\theta(u_{m+1}), \theta(u_{m+2})$ and so on.
-First we recall the method of Takens delay embedding.
+Suppose we only have the observations $\theta(u_0), \theta(u_1), \ldots, \theta(u_m)$ of the true system states $u$ in an unknown dynamical system $(U,T)$ and we wish to predict the values $\theta(u_{m+1}), \theta(u_{m+2})$ and so forth.
 
-In this method we consider a discrete-time dynamical system  
-defined on a nice space (a smooth manifold) that can be obtained as time-$K$ map of a continuous time-dynamical system \ednote{M: edited here}.
-We make observations from such a system, i.e., we observe the evolution of $\theta(w_0)$, i.e., we observe  a finite set of values of $u_n :=  \theta(w_n) = \theta(Tw_{n-1})$.
-The sequence $\{u_n\}$ represents a scalar time-series and intuitively $\theta$ may be thought to represent a probe inserted into a larger system and itself measuring/extracting only a small part of the greater system state at time $t=n$.
+First we establish the method of Takens delay embedding. 
+In this method, one considers a discrete-time dynamical system defined on a 'nice' space (a smooth manifold)\ednote{Should this be further explained?} that can be obtained as time-$K$ map of a continuous time-dynamical system \ednote{B: Should one add here: "These notions will be made formal below"?}.
+We make observations from such a system, i.e., we observe the evolution of $\theta(w_0)$ (where $w_0$ represents the initial state) i.e., we observe  a finite set of values of $u_n :=  \theta(w_n) = \theta(Tw_{n-1})$.
+The sequence $\{u_n\}$ represents a scalar time-series and intuitively $\theta$ may be thought to represent a probe inserted into a larger system which is itself only measuring/extracting a small part of the greater system state at time $t=n$.
 Consider for example a thermometer erected to measure the ambient temperature in a local village.
-This measurement function, the thermometer is capturing a single aspect, the temperature, of a grander dynamical system entailing the present weather of the surrounding area, but even more than that, it is measuring an incredibly small part of the global weather system.
+This measurement function, the thermometer, is capturing only a single aspect, the temperature, of a much grander dynamical system entailing the present weather of the surrounding area.
+Even more than that though- it is measuring a miniscule part of the global weather system.
 
@@
 
-However, the true state $u$ of a system is seldom, if ever fully known.
-In almost all cases we can merely insert probes into a system to obtain partial information by measurements taken.
-Moreover, the process of taking a measurement itself introduces 2 additional difficulties: \ednote{M: Why is discretization a difficulty?
-Maybe you replace difficulty by saying that we need consider two aspects}
+However, the true state $u$ of a system is seldom, if ever, fully known.
+In almost all cases we can at most insert probes into a system to obtain partial information by means of the measurements taken.
+Moreover, the process of taking a measurement itself introduces 2 additional aspects which complicate the problem: \ednote{M: Maybe replace difficulty by saying that we need to consider two aspects.
+B: Perhaps this edit makes more sense?}
 \vspace{-8mm}
 \begin{enumerate}[noitemsep, label=\roman*.]
-  \item A series of measurements over a specific time-interval is inherently a discretisation
-  procedure of the underlying continuous-time dynamical system. (Hence why we restricted our attention to discrete-time systems in the preceding section)
-  \item A probe will never be fully accurate and so the act of measurement introduces a certain
-measure of numerical noise/inaccuracy.
+  \item A series of measurements over a specific time-interval is inherently a discretisation procedure of the underlying continuous-time dynamical system.
+(Hence why we restricted our attention to discrete-time systems in the preceding section)
+  \item A probe will never be fully accurate, and so the act of measurement introduces a certain measure of numerical noise/inaccuracy.
 \end{enumerate}
@@
 $\Phi_{k,\theta}(w) := (\theta(T^{-k}w)\ldots,\theta(T^{-1}w),\theta(w))$.
-The essence of Takens theorem is that When $k$ is sufficiently large, we can define a dynamical system on the $\mathbb{R}^{k+1}$ that whose states are $\Phi_{k,\theta}(w)$, $\Phi_{k,\theta}(Tw)$, $\Phi_{k,\theta}(T^2w)$ and this dynamical system is topologically conjugate to the unknown underlying system $(W,T)$.
+The essence of Takens theorem relates to the fact that when $k$ is sufficiently large, we can define a dynamical system on the space $\mathbb{R}^{k+1}$ whose states are $\Phi_{k,\theta}(w)$, $\Phi_{k,\theta}(Tw)$, $\Phi_{k,\theta}(T^2w)$, etc.
+and this dynamical system is topologically conjugate to the unknown underlying system $(W,T)$.
 We recall the Takens delay embedding theorem next.
-
-\ednote{M: I have omitted many statements here}
 
@@
 
-\subsection{Taken Embedding Theorem}
+\section{Taken Embedding Theorem}
 
@@
 \begin{Definition}
+  [\bf {Homeomorphism}]\label{Dfn_homeo}
   A homeomorphism is a function $f:Z\rightarrow Y$ between two topological spaces $Z$ and $Y$ that is continuous, bijective and has a continuous inverse. 
@@
 \begin{Definition}
-  Consider a homeomorphism $f:Z\rightarrow Y$ for $Y\in X$. $Z$ is said to be embedded in $X$ by $f$.
+  [\bf {Embedding}]\label{Dfn_embed}
+  Consider a homeomorphism $f:Z\rightarrow Y$ for $Y\subset X$. $Z$ is said to be embedded in $X$ by $f$.
 \end{Definition}
 
-Takens Theorem states a result that establishes a relationship between the observed and underlying dynamical systems by showing that the concatenation of a sufficiently large number of previous observations into a vector will, under certain conditions, generate a map between the vectors from the respective systems.
-\ednote{B: Check wording.
-M: Edited the statement of the theorem} We formulate the theorem from \cite{takens1981detecting}.
+Takens Theorem states a result establishing a relationship between the observed and underlying dynamical systems by showing that the concatenation of a sufficiently large number of previous observations into a vector will, under certain conditions, generate a map between the vectors from the respective systems.
+We formulate the theorem from \cite{takens1981detecting}.
 
@@
 
-By generic we mean a residual set on a certain topology on an appropriate spaces of functions (that we do not describe here).
-By $C^2$, we mean a function that is twice differentiable, and the second derivative is continuous.
-(In our scenario, the input space $U$ is considered the attractor.)
+By generic we mean a residual set on a certain topology on appropriate spaces of functions (that we do not describe here) \textbf{(Cite)}.
+%\ednote{How vague may one be in general by leaving things out?}
+By $C^2$, we make reference to a twice-differentiable function  with a continuous second derivative.
+In our scenario (and this remains important throughout) the input space $U$ is considered the attractor.
+
+Below we reproduce Figure 1 from the supplementary material of \cite{manjunath2021universal}
 
@@
 \begin{figure}[ht]
-  \includegraphics[scale=0.3]{takensmap.png}
+  \includegraphics[scale=0.25]{takensmap.png}
   \centering
+  \captionof{figure}{Schematic of Takens Embedding Theorem's usage and consequence of how iterates \ldots }
   \label{fig:takensmap}
 \end{figure}
-\ednote{M: Please replace this figure with the one with in the Supplement in Adriaan's paper where $w$ appears instead of $u$ and this would be consistent with the statement in the theorem.
-Here, and throughout if you reproduce a figure, it must be stated where it was reproduced from.}
+\ednote{M: Here, and throughout if you reproduce a figure, it must be stated where it was reproduced from.
+B: How does one write that a figure is reproduced?
+Also, should $\tilde{F}$ be left out for now?}
+\ednote{B: Could I perhaps have the original figure? This is only a screenshot used as a temporary placeholder.}
 
@@
 
-\subsection{Practical Issues of Takens theorem}
-Despite the fact that the Takens' Embedding Theorem is a powerful result and provides compelling reason to believe that it is possible to accurately \emph{B: right word?
-M: Yes, I have edited the sentence} reconstruct a system that is conjugate to the underlying system, it does present some serious practical limitations:
+\section{Practical Issues of Takens theorem}
+
+It is a fact that Takens' Embedding Theorem is both a powerful result and that it provides compelling reason to believe that one could conceivably reconstruct accurately a system conjugate to the underlying system.
+Nonetheless, it does present some serious practical limitations:
 \vspace{-5mm}
 \begin{enumerate}
-\item Even if we can indeed find $F_\theta$  our \emph{approximation} of  $F_\theta$ is a map from a larger set $\mathbb{R}^{2d+1}$ containing the embedded attractor.
-There are, however, no theoretical guarantees that
-$F_\theta$ will retain  $V=\Phi_{2d,\theta}(W)$  as an attractor although $W$ could be an attractor. \ednote{B: We haven't properly said that U is an attractor, or have we. M: I have rephrased }
+\item Even supposing that we could indeed find $F_\theta$ , our \emph{approximation} of  $F_\theta$ is a map from a larger set $\mathbb{R}^{2d+1}$ containing the embedded attractor.
+There are, however, no theoretical guarantees that $F_\theta$ will retain  $V=\Phi_{2d,\theta}(W)$  as an attractor although $W$ could itself be an attractor.
 \item Takens theorem is stated for noiseless observations.
-Due to noise $\epsilon_n$, the delay vector $\Phi_{2d,\theta}(w_n) + \epsilon_n$,  may lie outside $V$, and further, due to the chaotic nature of the underlying system (i.e.
-the fact that it has SIDC), the evolution of $\Phi_{2d,\theta}(w_n) + \epsilon_n$ under the map $F_\theta$ may move completely out of $V$.
+Due to noise $\epsilon_n$ the delay vector $\Phi_{2d,\theta}(w_n) + \epsilon_n$  may lie outside $V$.Furthermore, due to the chaotic nature of the underlying system (i.e.
+the fact that it has SIDC), the evolution of $\Phi_{2d,\theta}(w_n) + \epsilon_n$ under the map $F_\theta$ could move out of $V$ completely.
 This problem can be overcome  by using a driven dynamical system with some properties, and we discuss this in Chapter *.
- \item It is an unfortunate case that the functional complexity (i.e., the relative number of oscillatory graphs \cite{manjunath2021universal}) as well as stability of the embedding \Ftheta{} (\textbf{cite?})depends on the choice of observable $\theta$.
+%  \item It is an unfortunate case that the functional complexity (i.e., the relative number of oscillatory graphs \cite{manjunath2021universal}) as well as stability of the embedding \Ftheta{} (\textbf{cite?})depends on the choice of observable $\theta$.
 This means that we may even learn a map that complicates our learning task and in so doing adds obstacles to the road to learning a conjugate map with predictive power.
@@
 
-Another limitation owes itself to the requirement that T be a surjective map.
+% Another limitation owes itself to the requirement that T be a surjective map.
 This is not always the case in practical examples and we may consider the case of a damped double pendulum \ednote{M: Please remove the paragraph as well.
@@
 
+\chapter{Driven Dynamical Systems to Forecast Problems}\label{ch4}
+
+In this chapter, we discuss results pertaining to the mapping of temporal data obtained from a discrete-dynamical system onto a different space through the notion of a driven dynamical system. 
+We also consider the conditions a driven system should have to avoid adding distortion to its state space representation, as well as then ensuring that the single-delay dynamics (SDD) are conjugate (or at least semi-conjugate to the system).
+The SDD can then be used to forecast and reconstruct the underlying attractor. \ednote{Is this attractor assumed to exist?}
+
+\section{Nonautonomous and Driven Dynamical Systems}
+
+A nonautonomous dynamical system is simply a dynamical system (as defined before in \ref{Dfn_DS}) where the input $u$ from the input space $U$, a  topological space, is time-dependent.
+
+We immediately recall the figure first defined in \ref{eqn_conjugacy} and remind ourselves of our ultimate goal to forecast the evolution of an unknown nonautonomous dynamical system $(U,T)$ by constructing a conjugate system such as $(V,S)$ below.
+
+%#####################################
+\begin{center}
+\psset{arrows=->, arrowinset=0.25, linewidth=0.6pt, nodesep=3pt, labelsep=2pt, rowsep=0.7cm, colsep = 1.1cm, shortput =tablr}
+\everypsbox{\scriptstyle}
+\begin{psmatrix}
+U & U\\%
+V & V.
+%%%
+%  \ncline{1,1}{1,2}^{T} 
+%  \ncline{1,1}{2,1} <{\phi}
+%   \ncline{2,1}{2,2}^{S}
+%  \ncline{1,2}{2,2} > {\phi}
+\end{psmatrix}
+%#####################################
+\end{center}
+
+To do this, the need arises to consider a driven dynamical system where input is taken from both an input space $U$, and an underlying state space $X$. 
+This is done to account, in a manner more similar to the true scenario, for the influence of both the input and the actual state of the system at a specific timestep.
+
+\begin{Definition}
+  [\bf Driven and Compactly Driven Dynamical System] \label{Dfn_DDS}
+A driven dynamical system comprises two topological metric spaces $U$, $X$ and a continuous function  $g:U\times{X}\to{X}$ where $g(u_n, x_n)=x_{n+1}$. 
+The dynamics on X are generated by the update equation $x_{n+1}=g(u_n, x_n)$ where $n\in\mathbb{Z}$, input $u_n$ from $U$ and state $x_n$ belonging to $X$, where $X$ is compact. 
+If the input space $U$ is compact, we refer to the system as compactly driven. 
+Abbreviated, we shall refer only to the \emph{driven system $g$}, with all other entities being understood implicitly.
+\end{Definition}
+
+In particular, a nonautonomous dynamical system may be generated from $U$; any input $\overline{u}$, a bi-infinite sequence from $U$, gives rise to the sequence of self-maps $\{g(u_n, \cdot)\}_{n\in\mathbb{Z}}$ contained in $X$.
+Physically, one may think of this bi-infinite sequence as referring to a system that has been running for an incredibly long period at the time of the first measurement taken from the system (alternatively, the first time a probe is inserted into the system to take an observation).
+
+\begin{Definition}
+  [\bf Entire Solution] \label{Dfn_Soln}
+  A sequence $\{x_n\}_{n\in\mathbb{Z}}\subseteq{U}$ is called an entire solution(or simply a solution) to the driven system  $g$ with input $\overline{u}$ when it satisfies 
+  \[g(u_{n-1}, x_{n-1})=x_n\] for all $n\in\mathbb{Z}$
+\end{Definition}
+
+It is important to emphasise that a sequence satisfying the update equation above can only be a solution if $x_n\in{U}$ for all $n\in\mathbb{Z}$. Consider the example below.
+
+\begin{Example}\label{ex_halfux}
+  The only solution  $\{x_n\}_{n\in\mathbb{Z}}$ to the driven system  $g(u,x)=\frac{ux}{2}$, where $X=[0,1]$, $U=[0,1]$,  is the zero solution $x_n\equiv0$.
+  To see this, consider any $x_n=a\in[0,1]$ where $a\neq{0}$.  Let $\overline{u}\in{U}$ be an non-zero constant sequence , say $u_n=0.5$. 
+  The driven system may be rewritten as $x_{n-1}=\frac{2x_n}{u_{n-1}}=4x_n$ and the  iterates of $x_n$ in backward time will increase by a factor of 4 at each timestep. 
+  Thus for some $m\leq{n}$,  $1<x_m$ i.e. $x_m\notin{X}$. So ${\{x_n\}}_{n\in\mathbb{Z}}$ is not a solution and it follows that the only possible solution is the zero solution.
+\end{Example}
+
+A system may also have multiple solutions as is evidenced in the example below.
+
+\begin{Example}\label{Ex_exp}
+ Consider the driven system $g(u_n,x_n)=x^n$ for $X=[0,1]$, $U=\mathbb{R}$.
+The system has an uncountable number of solutions, as there exists a solution for every $x\in{X}$ which also passes through the point $x$ and $lim_{n\to\infty}x_n=0$, $lim_{n\to{-}\infty}x_n=1$.
+ The proof is deferred to immediately after the next paragraph.
+
+\end{Example}
+
+As the solutions to a driven system are often considered, we next identify a subspace $X_U$ of $X$ that contains all possible solutions.
+To realize such a subspace of a driven system $g$, the concept of a reachable set is defined.
+
+\begin{Definition}
+  [\bf Reachable Set]\label{Dfn_ReachableSet}
+The reachable set of a driven system $g$ is exactly the union of all the elements of all the solutions, i.e., 
+\[X_U :=\Big \{x \in X:  x = x_k \mbox{ where  $\{x_n\}$  is a solution for some  $\bar{u}$} \Big \}.\]
+The set of all reachable states at a specific time $n$ for input $\overline{u}$ is denoted by $X_n(\overline{u})$
+\end{Definition}
+
+We note that $x\in{X_n(\overline{u})}$ if and only $g$ has a solution $\{x_k\}$ for $x_n=x$ and input $\overline{u}$. \textbf{Cite}.
+This will lead to a result established later, but which is worth taking note of now: \textit{$g$ being a topological contraction is equivalent to the existence of a unique entire solution.} To this end we define the notion of a topological contraction.
+
+\begin{Definition}
+  [\bf Topological Contraction]\label{Dfn_TopContr}
+  A function $g:U\times{X}\to{X}$ is a topological contraction if for all $n\in\mathbb{Z}$ and all $\overline{u}\subseteq{U}$, $X_n(\overline{u})$ is a singleton subset of X.  
+\end{Definition}
+
+\begin{proof}
+  [\bf Proof of Example\ref{Ex_exp}] 
+  To see this, we show that for every input $\overline{u}=\{u_n\}_{n\in\mathbb{Z}}$, $g(u_n,\cdot)$ is a contraction map on $X$ for every $k\in\mathbb{Z}$. 
+  Indeed $g(u_n,x_n)=x^n$ depends only on the state at time $n$ and so for each input $\overline{u}$, $X_n(\overline{u})=\{x^n\}$ is a singleton subset.
+  We then easily conclude that $\{x_n\}\equiv0$ is the only solution. 
+\end{proof}
+
+Thus far it has been demonstrated that a system may have one or more solutions; one may ask if a driven system always has a solution and, if so, whether it satisfies certain properties such as uniqueness.
+Should the driven system be compact, existence follows immediately as shown in the following result.
+
+\begin{Theorem}\label{Thm_CompactExistence}
+  If $X$ is compact then for each input $\overline{u}$, there exists at least one solution to the driven system $g(\ldotp, x)$
+\end{Theorem}
+\begin{proof}
+  Consider an input $\overline{u}={\{u_n\}}_{n\in\mathbb{Z}}\subseteq{U}$ and driven system $g:U\times{X}\to{X}$  generating a sequence ${\{x_n\}}_{n\in\mathbb{Z}}$ in the compact space $X$. 
+  Since $X$ is a metric space, it follows immediately from a well-known result in Analysis (\textbf{Cite}) that $\{x_n\}$ has a convergent subsequence which then is a solution. 
+\end{proof}
+
+We may easily construct many systems with trivial solution-sets, such as $g(u,x)=x$ which has only the constant solution $x$ and so for $U=[-1,1]$, the system would have no solution if $|x|>1$.
+To refine the scenario, we consider only systems with unique solutions.
+
+\section {Unique Solution Property}
+
+\begin{Definition}
+  [\bf Unique Solution Property] \label{Dfn_usp}
+  A driven system $g$ is said to have the Unique Solution Property (USP) if for each input $\overline{u}$ there exists exactly one solution. 
+  Alternatively we may formulate the USP as follows: $g$ has the Unique Solution Property if there exists a well-defined map $\Psi:{U}\to{X}$ with $\Psi({\overline{u}})$ denoting the unique solution.
+\end{Definition}
+
+One of the first result obtained upon defining the USP, is the fact that every solution will attract any different initial conditions towards the component parts of the solution, i.e.
+that the solution $\Psi$ is a non-autonomous uniform attractor.
+\ednote{Define Uniform Attractor before stating theorem.}
+
+\begin{Theorem} \label{Thm_AttractorSoln}
+  If $g$ has USP, then any solution to $g$ is a non-autonomous uniform attractor 
+\end{Theorem}
+\begin{proof}
+  \cite{Manju_Nonlinearity}
+\end{proof}
+
+\ednote{Additional discussion on USP?} 
+Having already defined the reachable set $X_U$, we pause for a moment to fix additional notation.
+Letting $\cev{u}^{n}:=(\ldots,u_{n-2} ,u_{n-1})$ be the left-infinite subsequence of an input, $\overleftarrow{U}$ then denotes all these left-infinite sequences in $U$. 
+Moreover, $\cev{u}^{n}v:=(\ldots,u_{n-2} ,u_{n-1}, v)$ is to symbolise the input up to time $n$ with $v \in U$ being the specific input value at time $n$. 
+The introduction of a new input at time $n$ can be described by a mapping $\sigma_v:   \cev{u}^{n} \mapsto \cev{u}^{n}v$.
+
+The question now becomes whether we may establish a conjugacy as presented below for $g$ driven
+
+\begin{equation}  \label{Scomm_h}
+  %    \[ 
+      \psset{arrows=->, arrowinset=0.25, linewidth=0.6pt, nodesep=3pt, labelsep=2pt, rowsep=0.7cm, colsep = 1.1cm, shortput =tablr}
+   \everypsbox{\scriptstyle}
+   \begin{psmatrix}
+   \overleftarrow{U} & \overleftarrow{U}\\%
+   X_U & X_U.
+   %%%
+  %  \ncline{1,1}{1,2}^{\sigma_v} \ncline{1,1}{2,1} <{h}
+  %  \ncline{1,2}{2,2} > {h}
+  %  \ncline{2,1}{2,2}^{g(v,\cdot)}
+   \end{psmatrix}
+  % \]
+  \end{equation}
+
+  Restricting our attention more and more, we now consider a specific subclass of conjugacies.
+
+  \begin{Definition} 
+    [\bf Universal Semi-Conjugacy]\label{Def_UnivSemiConj}
+    Given a driven system $g$, we  call a continuous and surjective map $h : \overleftarrow{U} \to X_U$ a universal semi-conjugacy if  diagram \ref{Scomm_h} commutes for all $v \in U$.
+  \end{Definition}
+
+If the universal semi-conjugacy $h$ exists (i.e. the diagram in \ref{Scomm_h} commutes) then the solution $\Psi$ will be at least a coarse-grained representation of the input $u$.
+
+Does such a function $h$ for the driven system defined above exist?
+If $g$ has the USP and $\Psi(u)=\{x_n\}_{n\in\mathbb{Z}}$ then $h$, defined by  $h(\cev{u}_n):=g(u_n,x_{n-1})=x_n$, will satisfy the semi-conjugacy in the graph above \ref{Scomm_h}.
+Regrettably, the mapping $h$ is not guaranteed to exist in general.
+
+Re-sketching the graph\ref{Scomm_h} above by fixing the input $v$ in $g$ and replacing $X_U$ by its left-infinite sequence space $\overleftarrow{U}$, we obtain the graph below.
+In this case, the function $H:\overleftarrow{U}\to\overleftarrow{X}_U$, a map that is both continuous and surjective, is called a \emph{causal mapping}.
+make special note of a 
+\begin{Definition}
+  [\bf Causal Mapping]\label{Def_CausMap}
+  A continuous, surjective map $H:\overleftarrow{U}\to\overleftarrow{X}_U$ such that \[H\circ\tilde{g}_v=\sigma\circ{H}\] where $\tilde{g}_v$ maps $(\ldots, u_{-2}, u_{-1})$ to $(\ldots, u_{-2}, u_{-1}, g(v, u_{-1}))$.
+\end{Definition}
+
+\begin{equation} \label{SCausal_H}
+    %    \[ 
+        \psset{arrows=->, arrowinset=0.25, linewidth=0.6pt, nodesep=3pt, labelsep=2pt, rowsep=0.7cm, colsep = 1.1cm, shortput =tablr}
+     \everypsbox{\scriptstyle}
+     \begin{psmatrix}
+     \overleftarrow{U} & \overleftarrow{U}\\%
+     \overleftarrow{X}_U & \overleftarrow{X}_U.
+     %%%
+    %  \ncline{1,1}{1,2}^{\sigma_v} \ncline{1,1}{2,1} <{h}
+    %  \ncline{1,2}{2,2} > {h}
+    %  \ncline{2,1}{2,2}^{\tilde{g}_v}
+     \end{psmatrix}
+    % \]
+    \end{equation}
+
+\begin{Theorem}
+  For a compactly driven system, a causal mapping $H$ exists if and only if $g$ has the USP. 
+\end{Theorem}
+\begin{proof}
+  \emph{See proof from p.15 of \cite{manjunath2021universal}}
+  \textbf{Expand.}
+\end{proof}
+
+Note that even when $h$ does exist, we are not guaranteed its injectivity. Considering again example \ref{ex_halfux}, we see that even if $h$ were to exist, it could not be injective as $X_U=\{0\}$.
+
+An embedding of the space $\overleftarrow{U}$   would allow one to establish topological conjugacy, which in turn provides stronger results than merely obtaining a coarse-grained representation via a causal mapping.
+
+  Before formalising this, we make mention of the concept of the inverse-limit system of a dynamical system $(U,T)$. 
+  In broad sweeps, the inverse-limit system of a dynamical system is a self-map on a subset of an infinite-dimensional space \textbf{(Expand)} where each point in the inverse-limit space corresponds to a backward orbit of the map $T$.
+In literature, the inverse-limit space is denoted by $\hat{U}_T\subseteq\overleftarrow{U}$ \cite{manjunath2021universal}.
+  The map $T$, then, induces a self-map $\hat{T}:\widehat{T}: (\ldots,u_{-2},u_{-1}) \mapsto  (\ldots,u_{-2},u_{-1},T(u_{-1}))$ on $\widehat{U}_T$
+
+\begin{Definition}
+  [\bf Causal Embedding]\label{Dfn_CausalEmbed}
+  A driven system $g$ is said to causally embed the dynamical system $(U,T)$ if 
+  \vspace{-8mm}
+\begin{enumerate}[noitemsep, label=\roman*.]
+  \item The diagram in \ref{SCausal_H} commutes (i.e. $g$ is a universal semi-conjugacy),
+  \item $H_2(\overline{u}):=(h(r\overline{u}, h(\overline{u})))$ embeds the inverse-limit space $(\hat{U}, \hat{T})$ of $(U,T)$ in $X\times{X}$.
+\end{enumerate}
+\end{Definition}
+
+The driven system $g$ can induce an embedding of $\overleftarrow{U}$ in $\overleftarrow{X}$ as follows: If $H:\overleftarrow{U}\to\overleftarrow{X}_U$ is also injective (in addition to being surjective, and hence bijective), it becomes the embedding of $\overleftarrow{U}$ in $\overleftarrow{X}$ induced by the driven system $g$ and we refer to $H$ as a \emph{causal embedding}.
+
+\begin{Theorem}
+ If $g(\ldotp,x)$ is invertible and has the USP then $H$ is a causal embedding. 
+\end{Theorem}
+\begin{proof}
+  \emph{See proof from \cite{manjunath2021universal}}  \textbf{Expand.}
+\end{proof}
+
+When g has the USP, the diagram below illustrates the operation of the the mappings $h$ and $H$.
+The mapping $h:\overleftarrow{U}\to{X_U}$ is an observable as discussed in an above section (\ref{TakensEmbedding}).
+
+\begin{figure}[ht]
+  \includegraphics[scale=0.3]{actionofh_H.png}
+  \centering
+  \label{fig:actionh_H}
+\end{figure}
+
+The next result is of relative import but to fully comprehend its influence, some discussion still remains. We state it in part and postpone the proof.
+% Before stating it, we define a concept which will directly be shown to be equivalent to the USP.
+
+\begin{Theorem}
+  The following statements are equivalent:
+  \vspace{-8mm}
+  \begin{enumerate}[noitemsep, label=\roman*.]
+    \item $g$ has the USP. 
+    \item $g$ is a topological contraction.
+  \end{enumerate}
+\end{Theorem}
+\begin{proof}
+  (postponed) 
+\end{proof}
+
+% We have not as of yet treated the UAP properly, and establish the necessary result here (without proof) so as not to break up the completeness of theorem.
+
+% \textbf{The proof of the result for the UAP is left to literature. Why? Cite.}
+
+\section {Choosing the driven system $g$}
+
+One must be careful to avoid a choice of $g$ which would  add complexity to the obtained solution.
+
+When a causal embedding $H$ exists for the driven system $g$, one can map an arbitrary input ${u}$ onto the solution space $X$ without additional distortion or information-loss.  \ednote{(Cite.)} 
+When an embedding is established, the question of possible additional complexity in the solution is removed by guaranteeing that, since the systems are conjugate (semi-conjugate, (refer)), $g$ does not add any (some) complexity to the system.
+It is, however, a balancing act as it also undesirable to choose a function $g$ that quenches the temporal structure in $u$ by contracting to such a degree that the ability to recover information from the original system is lost completely.
+To obtain a suitably complex function $g$, it is desired that the the reachable set of a driven system  be large enough to relate to the input.  \ednote{(Expand here. It stops a bit abruptly.)}
+
+In the example above (\ref{ex_halfux}), the input's temporal variation cannot be related to the reachable set as $X_U$ consists of a single element and so little, if not no, information is encoded.
+Rather, the reachable set of a driven system must therefore be such that the inverse-limit space of $U_T$ can be embedded in some finite self-product of the reachable set of $T$ \ednote{(refer or cite)}.
+To this end, consider the notion of State-Input (SI) Invertibility.
+
+\begin{Definition}
+  [\bf SI-Invertibility]\label{Dfn_SIinv}
+  A map $g$ is said to be SI-Invertible if $g$ is invertible for all $x\in{X}$.
+Alternatively it may be said that if, given $x_n$ and $x_{n-1}$, $u_n$ can be uniquely determined from $x_n=g(u_n,x_{n-1})$, then $g$ is said to be SI-invertible.
+\end{Definition}
+
+SI-invertibility promises that 'enough' information is retained when $g$ is chosen without introducing additional unwanted complexity. \ednote{Expand why this is true.}
+
+It is worth taking note of a specific driven system. The function 
+\begin{equation}  \label{eqn_driving}
+g(u,x)=(1-a)x+a{\cdot}tanh(Au+\alpha{a}Bx)
+\end{equation} 
+is both SI-invertible and possesses the USP. The proof of these two facts is rather involved and  hence the proof, instead of being replicated here, may be found in \textbf{(cite relevant article)}
+This specific driving function $g$ is used in our implementation and is more completely discussed in chapter \ref{sect5}
+
+Despite the ease that one may work with a left-infinite history in the realm of theory, it is impossible to obtain such a sequence in any real-life application.  
+One does not in practice, fortunately, need the entire left-infinite history of the input thanks to the Uniform Attraction Property (UAP).
+
+\begin{Definition}
+  [\bf Uniform Attraction Property]\label{Dfn_UAP}
+  A driven system $g$ has the Uniform Attraction Property (UAP) if, regardless of starting position, all trajectories \textbf{(of what?)} converge to a single trajectory as time flows forward. 
+  This trajectory is also the unique solution sequence $x$ to the input sequence $u$ as mentioned above \ref{here_UAP_firstestablished}.
+\end{Definition}
+
+This definition is stated in a not completely rigourous manner as the more formal definition makes use of processes, a concept which would take some time to establish and detracts from the principal thrust of this paper/project/thesis.
+\ednote{Is it VERY obvious to the reader at this point what the general thrust of the paper is?}
+
+As the Unique Attraction Property guarantees that all trajectories \textit{(which ones?)} will converge to the same trajectory as time moves forward. 
+Incredibly, this permits one to initialise a driven system $g$ with an altogether arbitrary initial value $y_m\in{X}$ where $m\in\mathbb{Z}$ and the UAP then guarantees that the sequence $\{y_{m+1}, y_{m+2},\ldots\}$ which satisfies the relation $y_n=g(u_n, y_{n-1})$  for  $k\geq{m}$  will uniformly approach the elements $\{x_n\}$ of the actual solution.
+(see [41, Theorem 1] or [11, Eqn.
+(18)])
+
+%Change the theorem above so that th.4 is split up into two parts.
+
+\begin{figure}[ht]
+  \includegraphics[scale=0.4]{memloss_conttime.png}
+  \centering
+  \label{fig:memoryLosscont_time}
+\caption{Approaching the true solution if g has the UAP}
+\end{figure}
+
+One may now appreciate an even more astounding result: $g$ having the USP is equivalent to the UAP. We restate the above theorem.
+
+\begin{Theorem}
+  The following statements are equivalent:
+  \vspace{-8mm}
+  \begin{enumerate}[noitemsep, label=\roman*.]
+    \item $g$ has the USP.
+    \item $g$ is a topological contraction.
+    \item $g$ has the UAP.
+  \end{enumerate}
+\end{Theorem}
+\begin{proof}
+  (We'll show this)  [11] eqn.6 
+\end{proof}
+
+One need therefore not establish any additional results to ensure that the sequence uniformly approaches the unique solution $\Psi$.
+This vastly simplifies the effort necessary to set up a problem in order to guarantee that the underlying system will be accurately represented by the conjugate system.
+
+This also solves the problem of perturbation/noise introduced by the observable or measurement function.
+Recall from before that an observable is inherently a map that discretises the underlying continuous-time system.
+Moreover, measurement and device error introduce mistakes.
+Since we are working with a chaotic system displaying sensitive dependence on initial conditions, these small errors could potentially send the trajectory into a completely different attractor.
+In the presence of the UAP, however, small measurement errors introduced into the system do not pose the same danger as before.
+\ednote{(Rephrase and/or expand)}
+
+\section{The next step in Dynamics}
+
+Next we define the relation $Y_T$ induced by $(U,T)$ on $X_U\times{X_U}$ for a driven system $g$ possessing SI-invertibility.
+To describe the  single-delay lag dynamics formally, we consider a dynamical system $T: U \to U$ and we define a relation on the reachable set $X_U$, i.e., a subset defined on $X_U \times X_U$ by
+$$Y_T:=\{(x_{n-1},x_n): \{x_k\}_{k\in \mathbb{Z}} \mbox{ is a solution for some orbit of } T \mbox{ and } n \in \mathbb{Z}\}.$$
+
+The following theorem establishes the existence of a well-defined map $G_T$ describing the single-delay dynamics(SDD) of the system above.
+
+\begin{Theorem}
+If we let $G_T:Y_T\to{Y_T}$ be a map defined by the relation $(x_{n-1},x_n)\mapsto(x_n,x_{n-1})$, then $G_T$ is well-defined (and this results holds even in the absence of $g$ possessing over the USP)
+\end{Theorem}
+
+We're now getting quite close to where we want to be and our results carry more and more weight. Recall that $g$ is only being given inputs from the orbits of $T$.
+
+\begin{Definition}
+  [\bf Inverse-Limit Space]\label{Dfn_InvLimSpace}
+  The inverse-limit space $\widehat{U}_T$, a subspace of $\overleftarrow{U}$, is defined by $$\widehat{U}_T:=\{(\ldots, u_{-2}, u_{-1}):T(u_n)=u_{n+1}\}$$ 
+\end{Definition}
+
+The inverse-limit space is well-defined since $T:U\to{U}$ is surjective by assumption. \ednote{Should I explain this in more details or am  I then over-simplifiying?}
+
+\begin{Theorem}
+  Graph \ref{Scomm_h} is exactly the inverse-limit system $(\hat{U}, \hat{T})$.    
+\end{Theorem}
+
+Note that $H_2$ maps an entire left-infinite solution sequence from $\Psi$ to an element in $X\times{X}$.
+We now have the following (compare with \ref{SCausal_H} above):
+
+\begin{equation} \label{Fig_inverse_limsystem}
+  %  \[ 
+      \psset{arrows=->, arrowinset=0.25, linewidth=0.6pt, nodesep=3pt, labelsep=2pt, rowsep=0.7cm, colsep = 1.1cm, shortput =tablr}
+      \everypsbox{\scriptstyle}
+      \begin{psmatrix}
+      \widehat{U}_T  && \widehat{U}_T \\
+      Y_T && Y_T \\
+      %%%
+     %  \ncline{1,1}{1,2}^{\widehat{T}} \ncline{1,1}{2,1} <{h}
+     %  \ncline{1,2}{2,2} > {H_2}
+     %  \ncline{2,1}{2,2}^{G_T}
+      \end{psmatrix}
+    %  \]
+  \end{equation}
+
+\begin{Theorem}
+    $(Y_T, G_T)$ is semi-conjugate to $(\widehat{U}, \widehat{T})$.
+\end{Theorem}
+\begin{proof}
+  \textbf{Show}
+\end{proof}
+
+\section*{Summarising the discussing thus far:}
+
+It is easy to lose the birds-eye view, so we take a moment to review our progress up until this point.
+
+\vspace{-8mm}
+\begin{enumerate}
+\item We are interested in a some dynamical system $(U,T)$ with unknown dynamics.
+\item To determine properties about this system $(U,T)$ and predict its future evolution, we determine the dynamics of the inverse-limit system $(\widehat{U}, \widehat{T})$.
+Given certain assumptions, we can guarantee that $(\widehat{U}, \widehat{T})$ is at least semi-conjugate to $(U,T)$.
+\item If the driven system $g$ is SI-invertible (and $\{u_n\}\in{U}$  is an orbit of $T$), the map $G_T$ exists. 
+\item If, furthermore, $g$ has the USP, $(Y_T, G_T)$ is semi-conjugate to $(\widehat{U}, \widehat{T})$.
+\item If we can assume that $T$ is a homeomorphism, $(Y_T, G_T)$ is topologically conjugate to $(\widehat{U}, \widehat{T})$, an extension space of $(U,T)$
+\end{enumerate}
+
+One can therefore learn the SDD of the driven system states via $G_T$ with enough data thanks to the USP/UAP. This enables us to do at least 2 things: 
+\vspace{-8mm}
+\begin{itemize}[noitemsep]
+\item Forecast future values of $x_n$ via iterates of $G_T$ (as $G_T$ can be determined.
+\item Forecast future values of $u_n$. 
+\end{itemize}
+
+\section{A discussion of $G_T$ }
+
+In the above sections, we established the map $G_T$ describing the SDD of a driven system.
+By establishing the existence of this map, we’ve essentially embedded the attractor $U$ \ednote{(So far we haven't really spoken about an attractor)} into the higher dimensional space $X\times{X}$.
+\ednote{(Why is it true that it's higher-dimensional?)}
+In layman’s terms, this ensures that there is more “dimensional room” for the underlying system’s underlying dynamics to “move”.
+As the dynamics aren’t as “squashed”, we might therefore hope that the dynamics of $G_T$ are in some sense simpler than that of $T$.
+(Taken note of the fact that $G_T$ is a homeomorphism even when $T$ is just continuous)
+
+In \cite{manjunath2021universal} it is illustrated in an empirical fashion in that the map $G_T$ describes dynamics which are less functionally complex than that of $T$ or of the map $\Phi_{2d,\theta}$.
+This is done by implementing a Recurrent Neural Network (RNN), but not discussed as of yet.
+\ednote{(Add in that we'll be using Pearson coefficient?)}
+
+We opt to learn $G_T$ in an indirect manner by defining a new map $\Gamma:(x_{n-1},x_n)\mapsto{u_n}$. It will follow immediately from $G_T$’s existence that $\Gamma$ also exists.
+
+The reasons for taking this roundabout approach remain to be discussed in \ref{subs_LearnGamma}, but a pat answer may immediately be given: When $\Gamma$ has been learnt, the system can be driven autonomously and then $G_T$ is known anyway.
+We formalise this with a theorem.
+
+\begin{Theorem}
+  When $x_n$ and $x_{n-1}$ are successive points on a solution obtained for input-orbit of $T$ $\{u_n\}$, then the map $\Gamma: X\times{X}\to{U}$ defined by $(x_{n-1},x_n)\mapsto{u_n}$ exists whenever $G_T$ exists
+\end{Theorem}
+\begin{proof}
+  From [Koopman, Theorem 3c] 
+\end{proof}
+
+Projection mappings $\pi_i$ are defined in the traditional meaning where a $k$-dimensionalvector is projected to it's $i^{th}$ component such that $$\pi_i:(a_1,a_2, \ldotp, a_{k-1}, a_k)\to{a_i}$$
+
+% \begin{Definition}
+%   [\bf Projection Mapping]\label{Dfn_ProjMap}
+%  The projection mapping $\pi_i$ is defined as follows: $$\pi_1:(a_1,a_2)\to{a_1}$$ and $$\pi_2:(a_1,a_2)\to{a_2}$$ 
+% \end{Definition}
+
+The graph in equation \ref{Fig_inverse_limsystem} is then extended as below:
+
+\begin{figure}[ht] \label{Fig_Gamma}
+  \includegraphics[scale=0.3]{gammatemp.png}
+  \centering
+  \label{fig:conjugacy_with_Gamma}
+\end{figure}
+
+The problem finally simplifies to the issue of learning the map $\Gamma$ and combining this with the projection mapping $\pi_2$ and the function $g$, which will be known.
+A final set of equations is obtained - equations that have been entirely constructed from data.
+\begin{eqnarray}
+	u_{k+1} &=& \pi_1 \circ (\Gamma, \pi_2) \circ (\pi_2,g) (u_k,x_k) \label{Seqn_u}\\
+	x_{k+1} &=& \pi_2 \circ (\Gamma, \pi_2) \circ (\pi_2,g) (u_k,x_k). \label{Seqn_x}
+\end{eqnarray}
+
+\section{Advantages of learning $\Gamma$} \label{subs_LearnGamma}
+
+One may immediately ask why we opt to take such a roundabout route; why not just learn the map $G_T$ from the get-go?
+On the surface, this seems to be an arbitrary decision path with no real reasoning, so we take a pause again and discuss the motivation for learning Gamma.
+There are a number of distinct advantages.
+
+Learning $\Gamma$ saves computational resources.
+This is due to the fact that the input $u_n$ lies in a lower-dimensional subspace of the high-dimensional space $X\times{X}$.
+In practice, if the input is of a  lower dimension, one may easily embed it into the space $X\times{X}$ by padding the vector $u_n$ with zeroes.
+
+\textbf{Can we discuss how to put this across well? I would like to add it in as I think it beneficial to the discussion.}
+
+Moreover, two types of stability are achieved when $g$ has the USP.
+
+Input- and parameter-related stability is achieved if $g$ has the USP
+\newline \emph{Define/Explain the two types of stabilities}
+\newline \emph{Sentence on $u$, $v$, tails, USP and product topology}
+\newline If we opt to work with $Y_T$, errors can occur. Expand
+\newline $\Gamma$ is more stable. Explain why.
+
+Thirdly, $\Gamma$ is globally dissipative whereas $G_T$ is not guaranteed to be(\emph{Cite}). \ednote{Add linking sentence.} Global dissipativity prevents large numerical errors due to input data.
+
+The $\omega$-limit set $\omega(u;T)$ of a point $u$ is defined to be the collection of limit points of the sequence $\{x,Tu,T^2u,\ldots\}$.
+This set is nonempty, and $\omega(u;T)$ is invariant.
+\textbf{(Change)}
+
+\begin{Definition}
+  [\bf Global Dissipativity]\label{Dfn_GlobalDiss}
+	We say a dynamical system $(Z,f)$ is globally dissipative if there exists a nonempty proper closed subset $B$ of $Z$  so that for all $x\in Z\setminus B$, (i).
+$\omega(x;f) \subset B$  and (ii).
+$B$ is positively invariant that is $f(B) \subset B$.
+We call any such closed subset $B$ a trapping set of $f$.
+\end{Definition}
+
+If a system is not globally dissipative, neglibly small errors result in major errors which thus induce predictions to fail utterly.
+If one learns $G_T$ directly, we are in danger of replicating the problems arising for Takens.
+
 \end{document}

### 2022-01-06 Gandhi_Project_Report/draft3.tex
Done working through all edits Ch1-Ch3

--- 2022-01-03 Gandhi_Project_Report/draft1.tex
@@
 In this chapter, we discuss results pertaining to the mapping of temporal data obtained from a discrete-dynamical system onto a different space through the notion of a driven dynamical system. 
-We also consider the conditions a driven system should have to avoid adding distortion to its state space representation, as well as then ensuring that the single-delay dynamics (SDD) are conjugate (or at least semi-conjugate to the system).
-The SDD can then be used to forecast and reconstruct the underlying attractor. \ednote{Is this attractor assumed to exist?}
+We also consider the conditions a driven system should have to avoid adding distortion to its state space representation.
+We then describe its \emph{single-delay dynamics} (SDD) and then recall conditions on the driven system so that the SDD is conjugate (or at least semi-conjugate) to the underlying system.
+The SDD can then be used to forecast and reconstruct the underlying system.
+\ednote{Is this attractor assumed to exist?
+M:  I replaced  it with the system.
+Attractor is also a commonly used terminology since we observe data from an attractor in almost all cases.}
 
@@
 
-A nonautonomous dynamical system is simply a dynamical system (as defined before in \ref{Dfn_DS}) where the input $u$ from the input space $U$, a  topological space, is time-dependent.
-
-We immediately recall the figure first defined in \ref{eqn_conjugacy} and remind ourselves of our ultimate goal to forecast the evolution of an unknown nonautonomous dynamical system $(U,T)$ by constructing a conjugate system such as $(V,S)$ below.
+A nonautonomous dynamical system is simply a dynamical system (as defined before in \ref{Dfn_DS}) \ednote{M:I am not sure where it was defined before.
+If not, please say, A NDS comprises a family of map $\{g_n\}_{n \in \mathbb{Z}}$ and say  each $g_n$ could arise as a consequence of an input $u_n$.} from the input space $U$, a  topological space, is time-dependent.
+
+We immediately recall the figure first defined in \ref{eqn_conjugacy}  \ednote{Please use eqref instead of ref while cross-citing equations} and remind ourselves of our ultimate goal to forecast the evolution of an unknown nonautonomous dynamical system $(U,T)$ \ednote{We are not forecasting a nonautonomous, but an autonomous system} by constructing a conjugate system such as $(V,S)$ below.
+\ednote{M: Maybe something like this sounds better: We learn a system that is conjugate to an underlying or unknown dynamical system by using the measurements from  the unknown system.}
 
@@
 
-To do this, the need arises to consider a driven dynamical system where input is taken from both an input space $U$, and an underlying state space $X$. 
-This is done to account, in a manner more similar to the true scenario, for the influence of both the input and the actual state of the system at a specific timestep.
+To find a $\phi$ \ednote{M: Have edited here}, the need arises to consider a driven dynamical system where input is taken from an input space $U$, and the state of a space $X$ is updated. 
+This is done to account, in a manner more similar to the true scenario, for the influence of both the input and the actual state of the system at a specific time-step.
+\ednote{M : I understand this sentence, but am not sure if a general audience would.
+So please elaborate the sentence or omit}
 
 \begin{Definition}
-  [\bf Driven and Compactly Driven Dynamical System] \label{Dfn_DDS}
+  [\bf Driven and Compactly Driven Dynamical System] \label{Dfn_DDS} \rm
 A driven dynamical system comprises two topological metric spaces $U$, $X$ and a continuous function  $g:U\times{X}\to{X}$ where $g(u_n, x_n)=x_{n+1}$. 
 The dynamics on X are generated by the update equation $x_{n+1}=g(u_n, x_n)$ where $n\in\mathbb{Z}$, input $u_n$ from $U$ and state $x_n$ belonging to $X$, where $X$ is compact.
+\ednote{M : The sentence ``The dynamics on X are generated by ..." does not define anything, so mathematicians do not include it as part of a definition, and can be moved after the definition.}
 If the input space $U$ is compact, we refer to the system as compactly driven. 
@@
 In particular, a nonautonomous dynamical system may be generated from $U$; any input $\overline{u}$, a bi-infinite sequence from $U$, gives rise to the sequence of self-maps $\{g(u_n, \cdot)\}_{n\in\mathbb{Z}}$ contained in $X$.
-Physically, one may think of this bi-infinite sequence as referring to a system that has been running for an incredibly long period at the time of the first measurement taken from the system (alternatively, the first time a probe is inserted into the system to take an observation).
+Physically, one may think of this bi-infinite sequence as referring to a system that has been running for an incredibly long period at the time of the first measurement taken from the system (alternatively, the first time a probe is inserted into the system to take an observation) \ednote{Time is a label we put.
+We can label the first sample as $0$th or $-10000$th.
+So the word incredibly is not necessary, all we need is good amount of data.}
 
@@
   [\bf Entire Solution] \label{Dfn_Soln}
-  A sequence $\{x_n\}_{n\in\mathbb{Z}}\subseteq{U}$ is called an entire solution(or simply a solution) to the driven system  $g$ with input $\overline{u}$ when it satisfies 
+  A sequence $\{x_n\}_{n\in\mathbb{Z}}\subset X$ \ednote{Please avoid subseteq as it is ambiguous} is called an entire solution(or simply a solution) of the driven system  $g$ with input $\overline{u}$ when it satisfies
   \[g(u_{n-1}, x_{n-1})=x_n\] for all $n\in\mathbb{Z}$
@@
 
-It is important to emphasise that a sequence satisfying the update equation above can only be a solution if $x_n\in{U}$ for all $n\in\mathbb{Z}$. Consider the example below.
-
-\begin{Example}\label{ex_halfux}
+It is important to emphasise that a sequence $\{x_n\}$ satisfying the update equation above can only be a solution if $x_n\in{U}$ holds for all $n\in\mathbb{Z}$ and just for $n>0$.
+Consider the example below.
+
+\ednote{M: Only theorems, lemmas and propositions are in italics. For Definitions, remarks and examples, please add slash-rm command to keep it in normal mode}
+
+\begin{Example} \rm \label{ex_halfux}
   The only solution  $\{x_n\}_{n\in\mathbb{Z}}$ to the driven system  $g(u,x)=\frac{ux}{2}$, where $X=[0,1]$, $U=[0,1]$,  is the zero solution $x_n\equiv0$.
@@
 \end{Example}
+\ednote{Nicely written!}
 
@@
 
-\begin{Example}\label{Ex_exp}
+\begin{Example}\label{Ex_exp} \rm 
  Consider the driven system $g(u_n,x_n)=x^n$ for $X=[0,1]$, $U=\mathbb{R}$.
 The system has an uncountable number of solutions, as there exists a solution for every $x\in{X}$ which also passes through the point $x$ and $lim_{n\to\infty}x_n=0$, $lim_{n\to{-}\infty}x_n=1$.
- The proof is deferred to immediately after the next paragraph.
+ The proof is deferred to immediately after the next paragraph. \ednote{Make the proof a remark and please cite the remark}
 
@@
 
-As the solutions to a driven system are often considered, we next identify a subspace $X_U$ of $X$ that contains all possible solutions.
+As the solutions to a driven system are often considered to be an important entity \cite{M: edited the sentence here}, we next identify a subspace $X_U$ of $X$ that contains all possible solutions.
 To realize such a subspace of a driven system $g$, the concept of a reachable set is defined.
@@
 This will lead to a result established later, but which is worth taking note of now: \textit{$g$ being a topological contraction is equivalent to the existence of a unique entire solution.} To this end we define the notion of a topological contraction.
+\ednote{I do not think you need the definition of a topological contraction unless it is defined through a process.
+Please note that you have defined $X_n(\bar{u})$ differently than what was in the earlier works.
+Now it is just the collection of solution values at time $n$ with the input $\bar{u}$ and topological contraction is just paraphrasing the USP definition.}
 
@@
   [\bf Topological Contraction]\label{Dfn_TopContr}
-  A function $g:U\times{X}\to{X}$ is a topological contraction if for all $n\in\mathbb{Z}$ and all $\overline{u}\subseteq{U}$, $X_n(\overline{u})$ is a singleton subset of X.  
+  A function $g:U\times{X}\to{X}$ is a topological contraction if for all $n\in\mathbb{Z}$ and all $\overline{u}\subseteq{U}$, $X_n(\overline{u})$ is a singleton subset of $X$.  
 \end{Definition}
@@
 \begin{Theorem}\label{Thm_CompactExistence}
-  If $X$ is compact then for each input $\overline{u}$, there exists at least one solution to the driven system $g(\ldotp, x)$
+ Let $g$ be a driven system.  If $X$ is compact then for each input $\overline{u}$, there exists at least one solution to the driven system $g(\ldotp, x)$
 \end{Theorem}
 \begin{proof}
-  Consider an input $\overline{u}={\{u_n\}}_{n\in\mathbb{Z}}\subseteq{U}$ and driven system $g:U\times{X}\to{X}$  generating a sequence ${\{x_n\}}_{n\in\mathbb{Z}}$ in the compact space $X$. 
+  Consider an input $\overline{u}={\{u_n\}}_{n\in\mathbb{Z}}\subseteq{U}$ and driven system $g:U\times{X}\to{X}$  generating a sequence ${\{x_n\}}_{n\in\mathbb{Z}}$ in the compact space $X$ \ednote{This statement is not correct.}.
   Since $X$ is a metric space, it follows immediately from a well-known result in Analysis (\textbf{Cite}) that $\{x_n\}$ has a convergent subsequence which then is a solution.
+\ednote{This is not a valid proof.
+Please refer to the proof existence of a solution in \cite{Manju_ESP}.}
 \end{proof}
@@
 that the solution $\Psi$ is a non-autonomous uniform attractor.
-\ednote{Define Uniform Attractor before stating theorem.}
+\ednote{Define Uniform Attractor before stating theorem or else omit the statement.}
 
@@
 \begin{proof}
-  \cite{Manju_Nonlinearity}
+\ednote{M: You can omit the theorem, and just say it is a uniform attractor in the nonautonomous dynamical systems literature \cite{Manju_Nonlinearity}, and say the discussion on nonautonomous attractors is beyond the scope of the thesis.}
+
 \end{proof}
 
-\ednote{Additional discussion on USP?} 
+\ednote{Additional discussion on USP? M: Please do so by stating that it is equivalent to the echos state property. See the paper with Adriaan and paraphrase} 
 Having already defined the reachable set $X_U$, we pause for a moment to fix additional notation.
@@
 
-The question now becomes whether we may establish a conjugacy as presented below for $g$ driven
+The question now becomes whether we may establish a semi-conjugacy \ednote{M: I have changed it to a semi-conjugacy since conjugacy amounts to asking a infinite-dimensional space $\overleftarrow{U}$ to be embedded in $X_U$ which is not possible.
+} as presented below for the driven system $g$.
 
@@
 If $g$ has the USP and $\Psi(u)=\{x_n\}_{n\in\mathbb{Z}}$ then $h$, defined by  $h(\cev{u}_n):=g(u_n,x_{n-1})=x_n$, will satisfy the semi-conjugacy in the graph above \ref{Scomm_h}.
-Regrettably, the mapping $h$ is not guaranteed to exist in general.
-
-Re-sketching the graph\ref{Scomm_h} above by fixing the input $v$ in $g$ and replacing $X_U$ by its left-infinite sequence space $\overleftarrow{U}$, we obtain the graph below.
+Regrettably, such a continuous mapping $h$ is not guaranteed to exist \cite{M: Have edited here} when $g$ does not have the USP \cite[Lemma 5]{Manju_Nonlinearity}.
+
+Re-sketching the commutativity diagram \ref{Scomm_h} above by fixing the input $v$ in $g$ and replacing $X_U$ by its left-infinite sequence space $\overleftarrow{U}$, we obtain the diagram below.
 In this case, the function $H:\overleftarrow{U}\to\overleftarrow{X}_U$, a map that is both continuous and surjective, is called a \emph{causal mapping}.
-make special note of a 
+make special note of a \ednote{something you want to fill?} 
 \begin{Definition}
@@
 
-Note that even when $h$ does exist, we are not guaranteed its injectivity. Considering again example \ref{ex_halfux}, we see that even if $h$ were to exist, it could not be injective as $X_U=\{0\}$.
-
-An embedding of the space $\overleftarrow{U}$   would allow one to establish topological conjugacy, which in turn provides stronger results than merely obtaining a coarse-grained representation via a causal mapping.
+Note that even when $h$ does exist, we are not guaranteed its injectivity.
+Considering again example \ref{ex_halfux}, we see that even if $h$ were to exist, it could not be injective as $X_U=\{0\}$ \ednote{M: I feel these two sentences should be made after the definition of $h$ rather than here}.
+
+An embedding of the space $\overleftarrow{U}$   \ednote{M: in which space?
+is it the inver-limit space instead?} would allow one to establish topological conjugacy, which in turn provides stronger results than merely obtaining a coarse-grained representation via a causal mapping \cite{M: I feel this is very confusing to the reader as the co-domain of causal mapping is the left-infinite space and , so please omit if possible.
+You can just say we want to restrict inputs to the left-infinite orbits of a dynamical system and then explore conjugacy instead of a mere semi-conjugacy}.
 
@@
   In broad sweeps, the inverse-limit system of a dynamical system is a self-map on a subset of an infinite-dimensional space \textbf{(Expand)} where each point in the inverse-limit space corresponds to a backward orbit of the map $T$.
-In literature, the inverse-limit space is denoted by $\hat{U}_T\subseteq\overleftarrow{U}$ \cite{manjunath2021universal}.
-  The map $T$, then, induces a self-map $\hat{T}:\widehat{T}: (\ldots,u_{-2},u_{-1}) \mapsto  (\ldots,u_{-2},u_{-1},T(u_{-1}))$ on $\widehat{U}_T$
+Following \cite{manjunath2021universal}, the inverse-limit space is denoted by $\widehat{U}_T\subset \overleftarrow{U}$, and defined as $\widehat{U}_T\subset: = \{ (\ldots,u_{-2},u_{-1}): Tu_{i} = u_{i+1}$.
+  The map $T$, then, induces a self-map $\widehat{T}: (\ldots,u_{-2},u_{-1}) \mapsto  (\ldots,u_{-2},u_{-1},T(u_{-1}))$ on $\widehat{U}_T$.
 
@@
 \begin{enumerate}[noitemsep, label=\roman*.]
-  \item The diagram in \ref{SCausal_H} commutes (i.e. $g$ is a universal semi-conjugacy),
-  \item $H_2(\overline{u}):=(h(r\overline{u}, h(\overline{u})))$ embeds the inverse-limit space $(\hat{U}, \hat{T})$ of $(U,T)$ in $X\times{X}$.
+  \item The diagram in \ref{SCausal_H} \ednote{M: It is actually Eqn. 1.1 and not 1.2}commutes (i.e. $h$ is a universal semi-conjugacy) ,
+  \item $H_2(\overline{u}):=(h(r\overline{u}, h(\overline{u})))$ embeds the inverse-limit space $(\widehat{U}, \hat{T})$ of $(U,T)$ in $X\times{X}$ \ednote{M: Please replace $\widehat{U}$ by $\widehat{U}_T$ here and throughout}.
 \end{enumerate}
@@
 
-The driven system $g$ can induce an embedding of $\overleftarrow{U}$ in $\overleftarrow{X}$ as follows: If $H:\overleftarrow{U}\to\overleftarrow{X}_U$ is also injective (in addition to being surjective, and hence bijective), it becomes the embedding of $\overleftarrow{U}$ in $\overleftarrow{X}$ induced by the driven system $g$ and we refer to $H$ as a \emph{causal embedding}.
+The driven system $g$ can induce an embedding of $\overleftarrow{U}$ in $\overleftarrow{X}$ as follows: If $H:\overleftarrow{U}\to\overleftarrow{X}_U$ is also injective (in addition to being surjective, and hence bijective), it becomes the embedding of $\overleftarrow{U}$ in $\overleftarrow{X}$ induced by the driven system $g$ and we refer to $H$ as a \emph{causal embedding} \ednote{M : Causal embedding a dynamical system is different from the definition of causal embedding, so please edit/delete this paragraph and the next theorem and also in the remaining part till Section 1.2}.
 
@@
 
-When g has the USP, the diagram below illustrates the operation of the the mappings $h$ and $H$.
+When $g$ has the USP, the diagram below illustrates the operation of the the mappings $h$ and $H$.
 The mapping $h:\overleftarrow{U}\to{X_U}$ is an observable as discussed in an above section (\ref{TakensEmbedding}).
@@
 \begin{figure}[ht]
-  \includegraphics[scale=0.3]{actionofh_H.png}
+  %\includegraphics[scale=0.3]{actionofh_H.png}
   \centering
@@
 \begin{figure}[ht]
-  \includegraphics[scale=0.4]{memloss_conttime.png}
+  %\includegraphics[scale=0.4]{memloss_conttime.png}
   \centering
@@
 \begin{figure}[ht] \label{Fig_Gamma}
-  \includegraphics[scale=0.3]{gammatemp.png}
+  % \includegraphics[scale=0.3]{gammatemp.png}
   \centering

### 2022-01-07 Gandhi_Project_Report/draft2.tex
Sortof done withCh.4 edits. Need to do a lot of work on 4.6

@@
 By $C^2$, we make reference to a twice-differentiable function  with a continuous second derivative.
-In our scenario (and this remains important throughout) the input space $U$ is considered the attractor.
-
-Below we reproduce Figure 1 from the supplementary material of \cite{manjunath2021universal}
+In our scenario (and this remains important throughout) the input space $U$ is considered the attractor.~\label{attractor_U}
+
+Below we reproduce Figure 1 from \cite{Supp}.
 
@@
 \vspace{-1cm}
-\bibliographystyle{pnas-new}
-\bibliography{pnas}
+% \bibliographystyle{pnas-new}
+% \bibliography{pnas}
 

### 2022-01-07 Gandhi_Project_Report/draft3.tex
Sortof done withCh.4 edits. Need to do a lot of work on 4.6

@@
 
-% \chapter{Introduction}\label{ch1}
-
-% Experimenting on biological, physical, and artificial systems in order to generate a more informative dynamical model is a well-established practice in modern science.
+\chapter{Introduction}\label{ch1}
+
+Experimenting on biological, physical, and artificial systems in order to generate a more informative dynamical model is a well-established practice in modern science.
 Traditional methods for modelling physical systems are based on laws of physics that are based either on empirical relationships or on intuition.
@@
 We do not fully understand many of the more complex systems, nor do they even provide us with a good physical intuition of the underlying principles governing the dynamics.
-
-% To complicate this - the underlying systems we observe often display sensitive dependence on initial conditions; despite having highly similar initial conditions, differing orbits diverge quickly and to such an extent that it becomes seemingly impossible to retrace their steps back to the original conditions.
+To complicate this - the underlying systems we observe often display sensitive dependence on initial conditions; despite having highly similar initial conditions, differing orbits diverge quickly and to such an extent that it becomes seemingly impossible to retrace their steps back to the original conditions.
 Even the presence of usually 'negligible' computational noise or measurement error renders long-term point-wise prediction infeasible.
-% There are many difficulties that one encounters while modelling such systems:
-% \vspace{-8mm}
-% \begin{enumerate}[noitemsep, label=\roman*.]
-%   \item One may not have access to the complete states of the systems
-%   \item The actual system may be described by functions that behave wildly in the sense their graphs have a wild oscillatory behaviour, i.e., they have a large functional complexity.    
-% \end{enumerate}
-
-% Models derived primarily from data can be classified into three categories: 
-% \vspace{-8mm}
-% \begin{enumerate}[noitemsep, label=\roman*.]
-%   \item  Interpretable models (i.e., they establish relationships between internal physical mechanisms), 
-%   \item Partially interpretable models capturing some modes of the dynamics, 
-%   \item Non-Interpretable models, defined as such mainly due to the fact that they are defined on a different phase space that is usually high-dimensional. 
-% \end{enumerate}
-
-% Examples in the literature attempting to forecast data from such systems have tried a number of different approaches with differing degrees of success.
-One could learn a system conjugate to the underlying system by applying the Takens delay embedding; this learnt system could then be used for forecasting the original system.
-\cite{takens1981detecting}.
-An ordinary differential equation can be obtained from data (Citations to Sindy).
-Alternatively, one could approximate the vector field by a library of functions (Citations to Sindy) to obtain interpretable models.
-Takens delay methods suffer from lack of robustness to noise and may not yield a learnable map that has lower functional complexity (See Chapter 3).
-For the methods that build interpretable models (Citations to Sindy) the complete system states are needed which is rarely available.
-Pure machine learning methodology that processes temporal information (like the echo state networks (Citations to echo state networks)) is an approach that maps data onto a higher dimensional space for further processing.
-Although they perform  well on forecasting some dynamical systems, they fail completely on others (Citations here) as there is often guarantee that the right function was learnt during training.
-
-% Recent partially interpretable models available in the literature have been based on the Koopman operator (e.g.,\cite{koopman1932dynamical,budivsic2012applied}) to  employ observables mapping the data onto a higher-dimensional space.
+There are many difficulties that one encounters while modelling such systems:
+\vspace{-8mm}
+\begin{enumerate}[noitemsep, label=\roman*.]
+  \item One may not have access to the complete states of the systems
+  \item The actual system may be described by functions that behave wildly in the sense their graphs have a wild oscillatory behaviour, i.e., they have a large functional complexity.    
+\end{enumerate}
+
+Models derived primarily from data can be classified into three categories: 
+\vspace{-8mm}
+\begin{enumerate}[noitemsep, label=\roman*.]
+  \item  Interpretable models (i.e., they establish relationships between internal physical mechanisms), 
+  \item Partially interpretable models capturing some modes of the dynamics, 
+  \item Non-Interpretable models, defined as such mainly due to the fact that they are defined on a different phase space that is usually high-dimensional. 
+\end{enumerate}
+
+Examples in the literature attempting to forecast data from such systems have tried a number of different approaches with differing degrees of success. 
+When we have complete access to states of the system, an ordinary differential equation can be obtained from data (Citations to Sindy and \cite{small2002modeling,xu2006modeling}) wherein one could approximate the vector field by a library of functions to obtain interpretable models.
+
+Recent partially interpretable models available in the literature have been based on the Koopman operator (e.g.,\cite{koopman1932dynamical,budivsic2012applied}) to  employ observables mapping the data onto a higher-dimensional space.
 This then makes the dynamics in the higher-dimensional space more amenable for approximation by a linear transformation.
@@
 
-% The non-interpretable models include the delay embedding and the machine learning algorithms.
-Under some generic conditions, the Takens delay embedding theorem \cite{takens1981detecting} and its various generalisations (e.g., \cite{sauer1991embedology, stark1999delay, gutman2018embedding, gutman2018embedding}) establishes the learnability of a system constructed by concatenating sufficiently large previous time-series observations of a dynamical system into a vector (called delay coordinates).
+The non-interpretable models include the delay embedding and the machine learning algorithms.
+One could learn a system conjugate to the underlying system by applying the Takens delay embedding \cite{takens1981detecting} when one has ``good" observations from the system; this learnt system could then be used for forecast the observed data.
+Takens delay embedding theorem \cite{takens1981detecting} and its various generalisations (e.g., \cite{sauer1991embedology, stark1999delay, gutman2018embedding, gutman2018embedding}) establishes the learnability of a system constructed by concatenating sufficiently large previous time-series observations of a dynamical system into a vector (called delay coordinates).
 This then establishes the existence of a map on the space of delay coordinates equivalent (or topologically conjugate) to the underlying map from which the observed time-series was first obtained.
-Although topological conjugacy guarantees an alternate representation of the underlying system, the quality of thsi representationstill depends on numerous parameters, making the comprehension of the dynamics rather unreliable (e.g., \cite{principe1992prediction}).
-One reason for this fragility is that the embedded attractor in the reconstruction space not always an attractor itself, despite  unquestionably being an invariant set.
-Since the embedded object is not \textbf{(definitely/always?)} an attractor (as explained in Chapter 3) it can cause predictions to fail.
-
-% Practically, the application of Takens embedding involves learning a map through some technique and consequently one wishes that these would have low functional complexity\cite{manjunath2021universal}, i.e., functions with fewer oscillatory graphs.
-
-% This thesis deals with the implementation and analysis of non-interpretable models that can guarantee \textbf{exact/accurate} reconstruction.
-The project work concerns the study and implementation of a method (Citation to paper with Adriaan) that incorporates learning a function by mapping the data on to a higher dimensional space.
+Although topological conjugacy guarantees an alternate representation of the underlying system, the quality of this representation still depends on numerous parameters, making the comprehension of the dynamics rather unreliable (e.g., \cite{principe1992prediction}).
+One reason for this fragility is that the embedded attractor in the reconstruction space not always an attractor of the map that is learnt in the reconstruction space, despite  unquestionably being an invariant set.
+When the embedded object is not \textbf{(definitely/always?)} an attractor (as explained in Chapter 3) it can cause predictions to fail.
+
+Practically, the application of Takens embedding involves learning a map through some technique and consequently one wishes that these would have low functional complexity\cite{manjunath2021universal}, i.e., functions with fewer oscillatory graphs.
+Pure machine learning methodology that processes temporal information (like the echo state networks (Citations to echo state networks)) by mapping data onto a higher dimensional space for further processing.
+Although they perform  well on forecasting some dynamical systems, they fail completely on others (Citations here) as there is often no guarantee that the right function was learnt during training.
+
+This thesis deals with the implementation and analysis of non-interpretable models that can guarantee \textbf{exact/accurate} reconstruction.
+The project work concerns the study and implementation of a method (Citation to paper with Adriaan) that incorporates learning a function by mapping the data on to a higher dimensional space using what is called a driven dynamical system (See Chapter 4).
 With a clear understanding of how the data is mapped onto the higher dimensional space, the method then permits the learning of a dynamical system topologically conjugate to that of the underlying system.
-Instead of linear regression, deep learning methods are employed to learn the correct function.
+Instead of linear regression as in the training of echo state networks, deep learning methods are employed to learn the correct function.
 With slight modifications to the implementation in the paper (with Adriaan), we show that one can construct accurate non-interpretable models with the ability to reconstruct attractors from more hard-to-forecast systems like the double pendulum.
@@
 
-% By solidifying the mathematical underpinnings of our theory, we hope to guarantee the ability to construct models with predictive power ranging from molecular biology to neuroscience.
+By solidifying the mathematical underpinnings of our theory, we hope to guarantee the ability to construct models with predictive power ranging from molecular biology to neuroscience.
 in the near future.
@@
 
-% The report is organised as follows: 
-% \newline In Chapter 2, we recall the definition of a discrete-time dynamical system, how a discrete-system arises from a flow of an ODE and then proceed to define the inverse limit space and topological conjugacy of autonomous systems.
-% \newline In Chapter 3, we introduce the problem of forecasting dynamical systems, state the Takens delay embedding theorem and discuss various issues faced while forecasting. 
-% \newline In Chapter 4, a driven dynamical system is defined and discuss the properties of a specific class of driven dynamical systems that we make use of in this thesis/report/paper. 
-% \newline Finally in Chapter 5 we show the implementation of these forecasting methods, and conclusions are provided in Chapter 6.
-
-% \chapter{Discrete-time Dynamical Systems}\label{ch2}
-
-% In this chapter, we provide a brief description of what a discrete-time dynamical system is and what it means for it to exhibit chaos.
+The report is organised as follows: 
+\newline In Chapter~\ref{ch2}, we recall the definition of a discrete-time dynamical system, how a discrete-system arises from a flow of an ODE and then proceed to define the inverse limit space and topological conjugacy of autonomous systems.
+\newline In Chapter~\ref{ch3}, we introduce the problem of forecasting dynamical systems, state the Takens delay embedding theorem and discuss various issues faced while forecasting. 
+\newline In Section 4, a driven dynamical system is defined and discuss the properties of a specific class of driven dynamical systems that we make use of in this thesis/report/paper. 
+\newline Finally in Section 5 we show the implementation of these forecasting methods, and conclusions are provided in Section 6.
+
+\chapter{Discrete-time Dynamical Systems} \label{ch2}
+
+In this chapter, we provide a brief description of what a discrete-time dynamical system is and what it means for it to exhibit chaos.
 We refer to \cite{devaney2018introduction, de2013elements} for more details.
-\label{Dfn_DS}
-
-% At its most elementary level, a dynamical system is just something that evolves deterministically through time.
+
+At its most elementary level, a dynamical system is just something that evolves deterministically through time.
 In the context of this thesis, deterministic refers to the fact that a system evolves according to specified rules rather than based on random events.
-% Dynamical systems arise in a variety of situations.
-A continuous dynamical system, and here specifically the motion of a pendulum, is one in which quantities such as the angular position and angular momentum are known at all times.
-% The equations of the dynamical system take the form of one or more ordinary differential equation that determine the relevant quantities at any future time if we know the initial location and momentum.
-% In ecology, discrete dynamical systems are widely used to model population growth.
+Dynamical systems arise in a variety of situations.
+A continuous-time dynamical system describes the states for all values of the time.
+Specifically, if the motion of a pendulum in which the quantities such as the angular position and angular momentum are known at all times, then it is a continuous-time dynamical system.
+The equations of the dynamical system can take the form of one or more ordinary differential equations that determine the relevant quantities at any future time if we know the initial location and momentum.
+In ecology, discrete-time dynamical systems are widely used to model population growth.
 The model in this case is a function calculating the following generation's population given the population of the previous generation.
@@
 
-% Formally, a function $T: U \to U$, where $U$ is some set is  a \emph{discrete-time dynamical system} and its iterates $\{u,Tu,T^2u,\ldots\}$, where $T^n$ denotes the $n$-fold composition of $T$ with itself, describe the evolution of an initial condition $u\in U$ (Note that we drop the brackets and denote $T(u)$ by $Tu$ for simplicity).
-
-% Continuous dynamical systems modelled using ordinary differential equations can give rise to discrete-time dynamical systems.
-To see this, consider differential equation $\dot{x} = f(x)$, $f: \mathbb{R}^n \to \mathbb{R}^n$, $n\in\mathbb{N}$ given to have a unique solution passing through
-% each point $x\in\mathbb{R}^{n}$, the flow of the equation \label{defn_flow} is defined to be a mapping $\varphi: \mathbb{R}^n \times \mathbb{R} \to \mathbb{R}$ that satisfies the following properties:
-% \vspace{-8mm}
-% \begin{enumerate}[noitemsep, label=\roman*.]
-%   \item $x(t)$ is a solution of the differential equation,
-%   \item $x(0)=x_0$, and
-%   \item $\varphi(x_0,t) = x(t)$.
-% \end{enumerate}
-% By fixing $t=K$, we can define the \emph{time-$K$ map} as  $T(x):= \varphi(x_0,K)$, and it is easily verified that $T\circ T(x_0) = \varphi(x_0,2K)$.
-In general the $m^{th}$ iterate of $x_0$ under $T$ would be the value of the solution of the ODE evaluated at time $mK$ with the initial condition $x_0$.
-% Thus ordinary differential equations give rise to a discrete-time dynamical systems by sampling the value of $x$ at time intervals $K$ units apart.
-
-% Of course, discrete-time dynamical systems need not always arise through a differential equation, and once again we may consider the field of ecology where discrete dynamical systems are often directly derived or assumed.
-% \emph{There is a school of thought that advocates discrete-time dynamical systems to be more natural for modelling real-world observations than differential equations (see \cite{saber2010introduction})}.
-\textbf{too philosphical, or actually good?}
-
-% A numerical discretization of a differential equation can also give rise to a discrete-time dynamical system.
+Formally, a function $T: U \to U$, where $U$ is some set is  a \emph{discrete-time dynamical system} and its iterates $\{u,Tu,T^2u,\ldots\}$, where $T^n$ denotes the $n$-fold composition of $T$ with itself, describe the evolution of an initial condition $u\in U$ (Note that we frequently drop the brackets and denote $T(u)$ by $Tu$ so as to simplify notations).
+
+Continuous-time dynamical systems modelled using ordinary differential equations can give rise to discrete-time dynamical systems.
+To see this, consider a differential equation $\dot{x} = f(x)$, $f: \mathbb{R}^n \to \mathbb{R}^n$, $n\in\mathbb{N}$ given to have a unique solution passing through
+each point $x\in\mathbb{R}^{n}$, call it $x(t)$ where $x(0)=x_0$.
+
+\begin{Definition}
+  [\bf Flow of an Equation] \label{Dfn_Flow}
+  The flow of the equation  $\dot{x} = f(x)$  is defined to be a mapping $\varphi: \mathbb{R}^n \times \mathbb{R} \to \mathbb{R}$ where $\varphi(x_0,t)= x(t)$ for the solution $x(t)$ with $x(0)=x_0$ 
+\end{Definition}
+
+By fixing $t=K \in (0,\infty)$, we can define the \emph{time-$K$ map} as  $T(x):= \varphi(x_0,K)$, and it is easily verified that $T\circ T(x_0) = \varphi(x_0,2K)$.
+In general the $m^{\mbox{th}}$ iterate of $x_0$ under $T$ would be the value of the solution of the ODE evaluated at time $mK$ with the initial condition $x_0$, i.e.
+$\varphi(x_0, mK)$.
+Thus ordinary differential equations give rise to a discrete-time dynamical system by sampling the value of the solution $x(t)$ at time intervals $K$ units apart.
+
+A numerical discretization of a differential equation can also give rise to a discrete-time dynamical system.
 For instance, Euler's method approximates $\dot{x}(t)$ by $(x(t+h)-x(t))/h$; if $h$ is fixed throughout, the solution of a differential equation $\dot{x}=f(x)$
-% can be approximated at the time instant $t+(m+1)h$ by iterating the equation 
-% $$x(t+(m+1)h) = x(t+mh) + h f(x(t+mh)).$$ 
-% Adopting more succinct notation by replacing $x(t+mh)$ with $u_m$, we rewrite the above equation as
-% $$u_{m+1} = u_m + hf(u_m)$$
-% or, in even simpler terms, as the discrete-time dynamical system with map $T(u) = u + hf(u)$, where $T$, $u$ and $f$ are understood to be as above.
-
-% \section{Invariant Sets}
-
-% A core concept in the study of dynamical systems is that of invariance.
-% \begin{Definition}
-%   Given a discrete-time dynamical system $T: U \to U$, a subset $A \subseteq U$ is said to be an invariant set if $T(A) =A$. 
-% \end{Definition}
-% We also define the orbit of T to be a sequence $\bar{u} = \{u_n\}_{n\in \mathbb{Z}}$ obeying the update equation, $u_{n+1}=Tu_n$, $n \in \mathbb{Z}$.
-
-% Two examples of invariant sets include a fixed point where $Tu=u$ for  $u\in U$, and a periodic orbit, i.e., a set of iterates $\{u,Tu, T^2u,\ldots,T^pu\}$, where $T^{p+1}u=u$ for some $p$.
+can be approximated at the time instant $t+(m+1)h$ by iterating the equation 
+\begin{equation}
+  x(t+(m+1)h) = x(t+mh) + h f(x(t+mh))
+\end{equation}
+
+Adopting more succinct notation by replacing $x(t+mh)$ with $u_m$, we rewrite the above equation as
+\begin{equation}
+u_{m+1} = u_m + hf(u_m)
+\end{equation}
+or in even simpler terms as the discrete-time dynamical system with map $T(u) = u + hf(u)$, where $T$, $u$ and $f$ are understood to be as above.
+
+Of course, discrete-time dynamical systems need not always arise through a differential equation, and once again we may consider the field of ecology where discrete dynamical systems are often directly derived or assumed.
+\textbf{Cite.}
+There is a school of thought that advocates discrete-time dynamical systems to be more natural for modelling real-world observations than differential equations and we refer the interested reader to \cite{saber2010introduction}.
+
+\section{Invariant Sets}
+
+A core concept in the study of dynamical systems is that of invariance. 
+\begin{Definition}
+  [\bf Invariant Set] \label{Dfn_Flow}
+  Given a discrete-time dynamical system $T: U \to U$, a subset $A \subset U$ is said to be an \emph{invariant set} if $T(A) =A$. 
+\end{Definition}
+
+We also define the \emph{orbit of $T$} to be a sequence $\bar{u} = \{u_n\}_{n\in \mathbb{Z}}$ obeying the update equation, $u_{n+1}=Tu_n$, $n \in \mathbb{Z}$.
+
+Two examples of invariant sets include a fixed point where $Tu=u$ for  $u\in U$, and a periodic orbit, i.e., a set of iterates $\{u,Tu, T^2u,\ldots,T^pu\}$, where $T^{p+1}u=u$ for some $p\in\mathbb{Z}$.
 The entire space $U$ could be also be invariant.
-% Consider for example the space $U=[0,1]$, where $Tu=4u(1-u)$ and then $U$ is invariant. \emph{(as every $u\in{U}$ can be written as $u = 4x(1-x)$ for some $x\in{U}$)}
-
-% We may learn a great deal about the iterates of a dynamical system by considering the types of invariant sets of a discrete-time dynamical system.
+Consider for example the space $U=[0,1]$, where $Tu=4u(1-u)$ and then $U$ is invariant (as every $u\in{U}$ can be written as $u = 4x(1-x)$ for some $x\in{U}$).
+
+We may learn a great deal about the iterates of a dynamical system by considering the types of invariant sets of a discrete-time dynamical system.
 For example, if $U=[0,1]$ has map $Tu= u/2$, then the only invariant set is $\{0\}$, and all orbits approach this invariant set as time flows in the forward direction.
-% Indeed, if a some non-zero invariant set (call it $B$) exists, then there is  some $r\in(0,1]\cap{B}$.
+Indeed, if a some non-zero invariant set (call it $B$) exists, then there is  some $r\in(0,1]\cap{B}$.
 But $r\notin{T(B)}$ since any orbit with initial value $r$ will be a decreasing sequence.
-Moreover, every orbit of T will be decreasing and therefore approach the value $0$ as $n\rightarrow\infty$
-
-% One may ask if every orbit approaches an invariant set?
-In general the answer is no, since for the dynamical system $Tu=2u$, any orbit that does not intersect the invariant set $\{0\}$, will not approach any invariant set.
-
-% However,  when the space $U$ is compact, all orbits approach an invariant set.
-\textbf{For let $\bar{u} = \{u_n\}_{n\in \mathbb{Z}}$ be an orbit in the compact space U, then $\bar{u}$ has a convergent subsequence }
-
-% In fact the $\omega$-limit set $\omega(u;T)$ of a point $u$ defined to the collection of limit points of the sequence $\{x,Tu,T^2u,\ldots\}$ is nonempty, and $\omega(u;T)$ is invariant.
-\textbf{(not relevant other than being an additional example perhaps?)}
-
-% We discuss various properties of invariant sets: Invariant sets can be attracting or repelling depending on how orbits in its vicinity behave (recall the examples $Tu =u/2$ and $Tu=2u$).
-We are interested in attractive invariant sets, and in particular so-called attractors which we define below.
-% \begin{Definition} \rm Let $T: U \to U$, where $U$  is a metric space with metric $d$. A compact subset $A \subseteq U$ is said to be an attractor if it satisfies the three conditions: 
-%   \vspace{-8mm}
-%   \begin{enumerate}
-% 	\item $A$ is invariant. 
-% 	\item $A$ is asymptotically stable, i.e., for every $\epsilon > 0$ and for all $u$ so that $d(u,A) < \epsilon$, we have $d(T^nu,A) \to 0$ as $n\to \infty$. 
-% 	\item $A$ has Lyapunov stability, i.e., for every $\epsilon > 0$  there exists a $\delta(\epsilon) > 0$ so that $d(u,A) < \delta$ implies $d(T^nu,A) < \epsilon$ for all $n\ge 0$.  
-% \end{enumerate}
-% \end{Definition}
-
-% In our previous examples, the system  $Tu=u/2$ had the singleton set $\{0\}$ as an attractor.
-For the system,  $Tu=4u(1-u)$ on $[0,1]$, one may easily verify that the only attractor is the entire space $[0,1]$.
-\textbf{For let $B$ be an attractor of U.
+Moreover, every orbit of T will be decreasing and therefore approach the value $0$ as $n\rightarrow\infty$.
+
+One may ask if every orbit approaches an invariant set?
+In general the answer is no, since for the dynamical system $T: \mathbb{R} \to \mathbb{R}$ defined by $Tu=2u$, any orbit that does not intersect the invariant set $\{0\}$, will not approach any invariant set.
+
+However,  when the space $U$ is compact, all orbits approach an invariant set.
+This is since the set of limit points of the orbit can be shown to be invariant \cite{de2013elements}.
+So, when $U$ is compact,  the $\omega$-limit set $\omega(u;T)$ of a point $u$ defined to the collection of limit points of the sequence $\{x,Tu,T^2u,\ldots\}$ is nonempty, and $\omega(u;T)$ is invariant.
+
+\begin{Example}
+  For the map $Tu=u^2$ defined on $[0,1]$ all orbits lie in the invariant set $\{1\}$ or else would would approach the invariant set $\{0\}$. 
+  % \ednote{B: not relevant other than being an additional example perhaps? M: Have added another simple example}.
+\end{Example}
+
+Invariant sets have various properties. Invariant sets can be attracting or repelling depending on how orbits in its vicinity behave. 
+Recall the examples $Tu =u/2$ and $Tu=2u$ defined on $\mathbb{R}$ (where $\{0\}$ ``attracts" orbits) or the example, $Tu=u^2$ defined on $[0,1]$ (where $\{1\}$ ``repels" orbits).
+We are interested in attractive invariant sets since they capture the long-term dynamics as time increases.
+In particular we are interested in those invariant sets named attractors.
+
+\begin{Definition} 
+  [\bf Attractor] \label{Dfn_Attractor}
+  Let $T: U \to U$, where $U$  is a metric space with metric $d$. A compact subset $A \subset U$ is said to be an attractor if it satisfies the three conditions: 
+  \vspace{-8mm}
+  \begin{enumerate}
+	\item $A$ is invariant. 
+	\item $A$ is asymptotically stable, i.e., for every $\epsilon > 0$ and for all $u$ so that $d(u,A) < \epsilon$, we have $d(T^nu,A) \to 0$ as $n\to \infty$. 
+	\item $A$ has Lyapunov stability, i.e., for every $\epsilon > 0$  there exists a $\delta(\epsilon) > 0$ so that $d(u,A) < \delta$ implies $d(T^nu,A) < \epsilon$ for all $n\ge 0$.  
+\end{enumerate}
+\end{Definition}
+
+Indeed in our previous examples, it can be verified that the system  $Tu=u/2$ defined on  $\mathbb{R}$ and $Tu=u^2$ defined on $[0,1]$  the singleton set $\{0\}$ is an attractor.
+For the system,  $Tu=1-|2u-1|$ on $[0,1]$, one may easily verify that the only attractor is the entire space $[0,1]$.
+This is since between any two points $u< v$ in $[0,1]$, and for any $a,b$ so that $u\le a < b \le v$, we can find an $n$ so that $T^n(a,b)=[0,1]$ (as would be explained later in this chapter).
+
+\ednote{For let $B$ be an attractor of U.
 From above, we know that B is itself invariant, is asymptotically stable and has Lyapunov stability.
-Now suppose there is some attractor $C\subseteqneq{U}=B$, so there is a $b\in{U}$ which is not in $C$.
+Now suppose there is some attractor $C\subsetneq{U}=B$, so there is a $b\in{U}$ which is not in $C$.
 ........
-}
-
-% A dynamical system can have several attractors, and may also be contained in another attractor.
-
-% % If the dynamics on the attractor are somewhat complicated in the sense that the attractor cannot be decomposed further, and if the attractor is infinite then there is a possibility of complicated dynamics, and a particular well-studied phenomenon of such complexity gives rise to what is called a chaotic behaviour.
-
-% If the dynamics on an attractor cannot be decomposed further and the attractor is itself also infinite, then there is a possibility of complicated dynamics, and a particular well-studied phenomenon of such complexity gives rise to what is called chaotic behaviour.
-
-% \section{Chaos}
-
-% \textbf{$\downarrow$ - should one add paragraphs like these?}
-% \newline In the 1960s, a number of mathematicians and mathematically interested scientists independently discovered chaos in the mathematical sense.
+M: Have changed the example}
+
+A dynamical system can have several attractors, and may also be contained in another attractor.
+For the example, $Tu =u^2$ on $[0,1]$, both $\{0\}$ and $[0,1]$ are attractors.
+It is a known result that for dynamical systems on a compact space there always exists an attractor.
+
+The dynamics restricted to an invariant set can be complicated.
+For instance an invariant set could be just a single point or it could have an infinite set.
+If the invariant set is infinite then there is a possibility of complicated dynamics, and a particular well-studied phenomenon of such complexity gives rise to what is called  chaotic behaviour, and has been studied in detail over the last fifty years.
+
+% If the dynamics on the attractor are somewhat complicated in the sense that the attractor cannot be decomposed further, and if the attractor is infinite then there is a possibility of complicated dynamics, and a particular well-studied phenomenon of such complexity gives rise to what is called a chaotic behaviour.
+
+%If the dynamics on an attractor cannot be decomposed further and the attractor is itself also infinite, then there is a possibility of complicated dynamics, and a particular well-studied phenomenon of such complexity gives rise to what is called chaotic behaviour.
+
+\section{Chaos}
+
+% \ednote{ should one add paragraphs like these?
+M: this is a thesis, and not a paper; the instructions sent to the external examiner regarding the thesis evaluation says that the student suggest strongly that it is good to write}
+In the 1960s, a number of mathematicians and mathematically interested scientists independently discovered chaos in the mathematical sense.
 The meteorologist Edward Lorenz may have been the first to explain this phenomenon in his 1963 paper \cite{lorenz1963deterministic}.
-The notions of invariance, attractivity and chaos may also bedescribed for continuous systems, and Lorenz's system comprised of a system of differential equations.
+The notions of invariance, attractivity and chaos may also be described for continuous systems, and Lorenz's system comprised of a system of differential equations.
 The narrative of Lorenz's discovery of chaos, as well as the history of other forerunners in this subject, are fascinating.
@@
 
-% There are several intricately different notions of chaos, with each of them indicating some complexity.
-\textbf{ $\leftarrow  $ what are we saying here?} In practice, it is not possible to verify which notion is satisfied when data is observed from a dynamical system.
-We do, however, specifically recall the definition of chaos in the sense of Devaney \cite{devaney2018introduction,de2013elements} so as to understand some nuances.
-Devaney's definition consists of three parts, with the first being the notion  of sensitive dependence on initial conditions.
-
-% \begin{Definition} \rm 
-% A dynamical system $T: U \to U$ is said to have sensitive dependence on initial conditions (SDIC) if there exists a $\delta > 0$ such that for every $u \in U$ and in every neighborhood of $u \in U$ there exists a $v\in{U}$ and an integer $N:=N{(u,v)}\in\mathbb{Z}$ such that $d(T^Nu,T^Nv)>\delta$.
-% \end{Definition}
-
-% It is important to acquire a sense of this definition as it can easily be misunderstood.
+There are several intricately different definitions of chaos available in the literature \textbf{(Cite)}, with each of them indicating some aspect of complexity.
+\ednote%{B:  what are we saying here?
+M: Devaney's definition is not the only one, there are about half a dozen definitions as well}
+In practice, it is not possible to verify which definition or notion does the underlying dynamical system satisfies when only data is observed from it.
+We do, however, specifically recall the definition of chaos in the sense of Devaney \cite{devaney2018introduction,de2013elements} so as to understand some nuances behind the complexity.
+Devaney's definition of chaos has three requirements, with the first being the notion  of sensitive dependence on initial conditions.
+
+\begin{Definition}
+  [\bf {Sensitive Dependence on Initial Conditions}] \label{Dfn_SDIC}
+A dynamical system $T: U \to U$ is said to have sensitive dependence on initial conditions (SDIC) if there exists a $\delta > 0$ such that for every $u \in U$ and in every neighborhood of $u \in U$ there exists a $v\in{U}$ and an integer $N:=N{(u,v)}\in\mathbb{Z}$ such that $d(T^Nu,T^Nv)>\delta$.
+\end{Definition}
+
+It is important to acquire a sense of this definition as it can easily be misunderstood.
 It is common misconception to interpret SDIC as two close points ($u$ and $v$) that eventually become separated by a distance $\delta$ under iteration by $T$.
@@
 To fully comprehend all the subtleties of the concept, one needs to discuss it more thoroughly by considering each sentence with care and attention, whereafter one may examine how they fit together to convey the concept of sensitive dependence.
-% \textbf{ $\leftarrow$ is this talking too simply perhaps?}.
-
-% To this end we make 3 remarks:
-% \vspace{-5mm}
-% \begin{enumerate}
-%   \item First, the $\delta>0$ in the definition of SDIC is independent of $u$. 
-%   \item Second, in every neighborhood of $u$, we may not necessarily find all points $v$ in the neighborhood distinct from $u$ that would separate from the forward iterates of $u$. 
-%   \item  Finally, $N$ depends upon $u$ and $v$ chosen, and their iterates may not separate for ever (i.e., for all $n>N$) and we allow their iterates to get arbitrarily close in the future. 
-% \end{enumerate}
-
-% \ednote{M: It is probably best to see a few numerical simulations to understand SDIC. (Please use the logistic map and perhaps the trajectories of the double pendulum data). }
-
-% The second part in Devaney's definition of chaos concerns topological transitivity.
-
-% \begin{Definition}
-% 	A dynamical system $F: U \to U$ is topologically transitive if for any pair of nonempty open sets $E_1$ and $E_2$ there exists a $n\in\mathbb{N}$ such that $T^n(E_1) \cap E_2 \not= \emptyset$. 
-% \end{Definition}
-
-% Topologically transitivity implies that the iterates of an open set of initial conditions gets mixed up with other open sets.
-On a compact metric space, it can be shown that topological transitivity also implies the existence of a point whose forward iterates are dense; or in other words, the orbit going through this point will be dense in the compact metric space.
-
-% In fact, it is topologically more likely that the choice of an arbitrary point will be one  whose iterates are almost dense.
-% % I inserted a definition here: it makes more sense to me. 
-% \begin{Definition}
-% A compact metric space $U$ is said to be almost dense if the complement of the set of points which are dense in $U$ may be written as a countable union of sets of with empty interior.
-In other words, the set on points not dense on $U$ is topologically insignificant.
-% \end{Definition}
-
-% We may now define the notion of Chaos as formulated by Devaney\cite{devaney2018introduction}.
-% \begin{Definition}
-% 	A dynamical system $T: U \to U$ is said to exhibit chaos in the sense of Devaney if it satisfies the three properties:
-% 	\vspace{-5mm}
-%   \begin{enumerate}
-% 		\item $T$ has SDIC.
-% 		\item $T$ is topologically transitive.
-% 		\item The set of periodic points of $T$ are dense in $U$. 
-% 	\end{enumerate}
-% \end{Definition}
-
-% \begin{figure}[ht]
-%   \centering
-%   \begin{minipage}{.5\textwidth}
-%     \centering
-%     \includegraphics[scale=0.18]{chaos1.png}
-%     \captionof{figure}{Starting to diverge}
-%     \label{fig:chaos1}
-%   \end{minipage}%
-%   \begin{minipage}{.5\textwidth}
-%     \centering
-%     \includegraphics[scale=0.18]{chaos2.png}
-%     \captionof{figure}{Completely chaotic}
-%     \label{fig:chaos2}
-%   \end{minipage}
-%   \end{figure}
-
-% \begin{Example}
-%   The standard tent map $T(u)=1-|2u-1|$ defined on $[0,1]$ is a common example of a dynamical system satisfying these three properties. 
-%   \begin{figure}[ht]
-%     \includegraphics[scale=0.6]{T2.png}
-%     \centering
-%     \label{fig:T2tentmap}
-%   \end{figure}
-
-%   Let us now consider why this map exhibit's Devaneys's sense of chaos.
-The graph of the map $T$ is piecewise linear with two straight lines, one connecting the points $(0,0)$ and $(1/2,1)$ and the other connecting $(1/2,1)$ with $(1,0)$.
+
+To this end, we make 3 remarks:
+\vspace{-5mm}
+\begin{enumerate}
+  \item First, the $\delta>0$ in the definition of SDIC is independent of $u$. 
+  \item Second, in every neighborhood of $u$, we may not necessarily find all points $v$ in the neighborhood distinct from $u$ that would separate from the forward iterates of $u$. 
+  \item  Finally, $N$ depends upon $u$ and $v$ chosen, and their iterates may not separate for ever (i.e., for all $n>N$) and we allow their iterates to get arbitrarily close in the future. 
+\end{enumerate}
+
+To illustrate the concept of SDIC, we present two examples - one from Mathematics and the other from the field of Physics.
+
+\begin{Example}
+  The logistic map(LM), a recursion relation of the form $x_{n+1}=rx_n(1-x_n)$ where $x_n\in[0,1]$ and $r<4$ is of particular interest to us as it is classically used to illustrate the influence of chaos on a system \textbf{(Cite)}.
+Below are plotted the first 75 iterates of the LM for $r=3.7$ and $x_0=0.2$ with an adjustment of 0.001 in $x_0$ and then $r$.
+
+  \begin{figure}[ht]
+    \includegraphics[scale=0.45]{log_sdic.eps}
+        \centering
+        \label{fig:log_sdic}
+        \captionof{figure}{Logistic Map for $r=3.7$, $x_0=0.2$}
+      \end{figure}
+
+  For an intuitive sense of the fact that the LM exhibits SDIC, we draw two green neighbourhoods of diameters $\delta_1=0.1, \delta_2=0.02$ around the points $u_1=0.7, u_2=0.4$ respectively.
+It is easy to see that every neighbourhood around these two points will contain some $v_1, v_2\in[0,1]$ and natural number $N_1, N_2 \in\mathbb{N}$ respectively s.t.
+the distance $d(T^Nu_1,T^Nv)>\delta_1=0.02$ and $d(T^Nu_2,T^Nv)>\delta_2=0.1$.
+
+  In a more rigourous fashion ....
+  \ednote{B:Should I be more rigourous here? }
+
+\end{Example}
+
+\begin{Example}
+  A second example pertains to a simple physical system called the double pendulum - a pendulum with another pendulum attached to its head given an initial position and angular velociy.
+
+  Below are two figures denoting the trajectories of a double pendulum's second head after some elapsed time.
+Depicted are three pendulum heads with equal angular velocity, but differing ever so slightly in initial position .
+As can be seen below, the trajectories diverge very quickly to become completely different.
+  The double pendulum will be discussed in much greater detail in a later chapter (\ref{ch5}), but we mention the example here to provide a numerical example of a practical system exhibiting SDIC.
+
+\begin{figure}[ht]
+  \centering
+\includegraphics[scale=0.8]{dp_sdic.eps}
+    \captionof{figure}{Three double pendulum heads with initial angle differing by 0,025 radians}\label{fig:dp_sdic}
+  \end{figure}
+
+\end{Example}
+
+The second part in Devaney's definition of chaos concerns topological transitivity.
+
+\begin{Definition}
+  [\bf {Topologically Transitive}] \label{Dfn_TopTrans}
+	A dynamical system $F: U \to U$ is topologically transitive if for any pair of nonempty open sets $E_1$ and $E_2$ there exists a $n\in\mathbb{N}$ such that $T^n(E_1) \cap E_2 \not= \emptyset$. 
+\end{Definition}
+
+Topologically transitivity implies that the iterates of an open set of initial conditions gets mixed up with other open sets.
+On a compact metric space, it can be shown that topological transitivity also implies the existence of a point whose forward iterates are dense \cite{de2013elements}; or in other words, the orbit going through this point will be dense in the compact metric space.
+In fact, it is topologically more likely that the choice of an arbitrary point will be one  whose iterates are almost dense.
+
+\begin{Definition} 
+  [\bf {Meager Set}] \label{Dfn_Meager Set}
+A subset of a topological space $U$ is said to be a meager set if it can be written as a countable union of sets of with empty interior. The complement of the a meager set is set to be a residual set.
+\end{Definition}
+
+To be more precise (see \cite{de2013elements}), for a discrete-time dynamical system on a compact space, the set of points with dense iterates are residual, and they are typical or likely to be observed in practice.
+In this thesis, we assume that when data is observed from a topologically transitive system, we assume that the data arises from a dense orbit.
+
+We may now define the notion of Chaos as formulated by Devaney\cite{devaney2018introduction}.
+\begin{Definition}
+  [\bf {Devaney's Chaos}]\label{Dfn_ChaosDec}
+	A dynamical system $T: U \to U$ is said to exhibit chaos in the sense of Devaney if it satisfies the three properties:
+	\vspace{-5mm}
+  \begin{enumerate}
+		\item $T$ has SDIC.
+		\item $T$ is topologically transitive.
+		\item The set of periodic points of $T$ are dense in $U$. 
+	\end{enumerate}
+\end{Definition}
+
+\begin{Example} 
+  The standard tent map $Tu=1-|2u-1|$ defined on $[0,1]$ is a well-known example of a dynamical system satisfying these three properties.
+
+  We now reason as to why this is true.
+The graph of the map $T$ is piecewise linear with two straight lines, one connecting the points $(0,0)$ and $(\frac{1}{2},1)$ and the other connecting $(\frac{1}{2},1)$ with $(1,0)$.
 They form a so-called tent with base centered at $u=1/2$.
 The graph of the map $T^2$ comprises two symmetric tents with their base centered at $1/4$ and $3/4$.
-%   % We now reason as why this map exhibits chaos in the sense of Devaney.
-
-%   Given any point $u$ and an open neighborhood $W \subseteq[0,1]$ of $u$, we can find $k\in\mathbb{Z}$ large enough so that $T^k(W) = [0,1]$ because we can accommodate a tent whose base is contained in $W$.
-\emph{This last phrase doesn't make sense to me.} This implies the existence of two points $w_1$ and $w_2$ in $W$ so that $d(T^kw_1,T^kw_2)= 1$, and by the triangle inequality  at least one of the inequalities $d(T^ku,T^kw_1)> \delta$ or $d(T^ku,T^kw_2)> \delta$ hold when $\delta\in (0,1/2)$.
-\newline So $T$ has sensitive dependence on initial conditions.
-
-  % Next, let $W_1$ and $W_2$ be two nonempty open sets.
+See Figure \ref{fig:T2tentmap}
+
+  \begin{figure}[ht]
+    \includegraphics[scale=0.6]{T2.png}
+        \centering
+        \captionof{figure}{Three double pendulum heads with initial angle differing by 0,025 radians}\label{fig:T2tentmap}
+  \end{figure}
+
+  In general, the graph of  $T^k$ contains $2^{k-1}$ tents.
+  Given any point $u$ and an open neighborhood $W \subset[0,1]$ of $u$, we can find $k\in\mathbb{Z}$ large enough so that $T^k(W) = [0,1]$ because we can accommodate a tent whose base is contained in $W$.
+% \ednote{B: This last phrase doesn't make sense to me.
+M: The number of tents grow with $k$ and hence this is obvious; I explained in one of our meetings.
+Does it make sense?}
+  This implies the existence of two points $w_1$ and $w_2$ in $W$ so that $d(T^kw_1,T^kw_2)= 1$, and by the triangle inequality  at least one of the inequalities $d(T^ku,T^kw_1)> \delta$ or $d(T^ku,T^kw_2)> \delta$ hold when $\delta\in (0,\frac{1}{2})$.
+So $T$ has sensitive dependence on initial conditions.
+
+  Next, let $W_1$ and $W_2$ be two nonempty open sets.
 Given any open set $W_1$, we can find a $k\in\mathbb{N}$ large enough so that $T^k(W_1) = [0,1]$ because we can accommodate a tent with base contained in $W$.
@@
 
-%   Finally, for every open interval $(a,b)$, the graph of $T^k$ intersects the graph of the identity map on $[0,1]$.
-This follows from the fact established above that there exists a $k\in\mathbb{N}$ such that $T^k(a,b) = [0,1]$.If the intersection point has coordinates of the form $(p,p)$, the point p is then a fixed point of $T^k$ and therefore also a periodic point of T.
-It follows that the set of periodic points are dense in T.
-%   \textbf{Expand further on reason for density.}
-% \end{Example}
-
-% \section{Conjugacy}
-
-% We now turn our attention to the subject of conjugacy.
-
-% To show topological similarity or sameness between two metric or topological spaces, one needs to establish a homeomorphism between the two spaces. 
-% However, in the study of dynamical systems defined on two spaces, establishing a homeomorphism does not indicate that the systems are (physically) related in any way.
-\textbf{"related" - in what sense am I saying it here?
-Physically?}  For this one rather needs to establish a kind of dynamical equivalence which we illustrate in a simple commutativity diagram below.
-% % Giving errors
-% \begin{equation}  \label{eqn_conjugacy}
-% %    \[ 
-%     \psset{arrows=->, arrowinset=0.25, linewidth=0.6pt, nodesep=3pt, labelsep=2pt, rowsep=0.7cm, colsep = 1.1cm, shortput =tablr}
-%  \everypsbox{\scriptstyle}
-%  \begin{psmatrix}
-% U & U\\%
-% V & V.
-%  %%%
-% %  \ncline{1,1}{1,2}^{T} 
-% %  \ncline{1,1}{2,1} <{\phi}
-% %   \ncline{2,1}{2,2}^{S}
-% %  \ncline{1,2}{2,2} > {\phi}
-%  \end{psmatrix}
-% % \]
-% \end{equation}
-
-% %Composite functions
-% To understand the diagram we note that if we travel ``right and then down," the diagram instructs us to use $T$ (top arrow right) first, followed by $\phi$.
+  Finally, for every open interval $(a,b)$, the graph of $T^k$ intersects the graph of the identity map on $[0,1]$.
+This follows from the fact established above that there exists a $k\in\mathbb{N}$ such that $T^k(a,b) = [0,1]$.If the intersection point has coordinates of the form $(p,p)$, the point $p$ is then a fixed point of $T^k$ and therefore also a periodic point of $T$.
+Since we have found a periodic point in the interval $W$,  the set of periodic points are dense in $T$.
+  This follows because any fixed point of $T^k$ is also a periodic point, and we have found this in an arbitrary interval.% \ednote{Expand further on reason for density.
+M: fixed point of $T^k$ is a periodic point, and we have found in an arbitrary interval.}
+
+\end{Example}
+
+\section{Conjugacy}
+
+We now turn our attention to the subject of conjugacy that describes when two dynamical systems are equivalent dynamically.
+
+To show topological similarity or sameness between two metric or topological spaces, one needs to establish a homeomorphism between the two spaces. 
+However, in the study of dynamical systems defined on two spaces, establishing a homeomorphism does not indicate that the systems are dynamically related in any way.
+For instance, the maps $Tu=u^2$ and $Tu=1-|2u-1|$ defined on $[0,1]$ have totally different behaviour.
+To find systems that are dynamically similar, one rather needs to establish a kind of dynamical equivalence which we illustrate in a simple commutativity diagram below.
+
+% Giving errors
+\begin{equation}  \label{eqn_conjugacy}
+%    \[ 
+    \psset{arrows=->, arrowinset=0.25, linewidth=0.6pt, nodesep=3pt, labelsep=2pt, rowsep=0.7cm, colsep = 1.1cm, shortput =tablr}
+ \everypsbox{\scriptstyle}
+ \begin{psmatrix}
+U & U\\%
+V & V.
+ %%%
+%  \ncline{1,1}{1,2}^{T} 
+%  \ncline{1,1}{2,1} <{\phi}
+%   \ncline{2,1}{2,2}^{S}
+%  \ncline{1,2}{2,2} > {\phi}
+ \end{psmatrix}
+% \]
+\end{equation}
+
+%Composite functions
+To understand the diagram we note that if we travel ``right and then down," the diagram instructs us to use $T$ (top arrow right) first, followed by $\phi$.
 (right arrow downwards).
@@
 This then amounts to finding  $S(\phi(u))$.
-When the relationships in this diagram hold, we say $\phi(T(u))= S(\phi(u))$ and formally denote it as $\phi \circ T=S\circ \phi$.
-
-% \begin{Definition}
-%   Consider the dynamical systems $T:U\to{U}$, $S:V\to{V}$ and suppose the relationship $\phi \circ T=S\circ \phi$ holds.
-If $\phi$ is a homeomorphism, then S is said to be conjugate to T and $\phi$ is called a conugacy.
-If we relax the criterion on $\phi$ and merely require $\phi$ to be continuous and surjective, $\phi$ is then a semi-conjugacy between $T$ and $S$; S is said to be semi-conjugate to T.
-% \end{Definition}
-
-% When $S$ is conjugate to $T$, the dynamics of the two systems are in some way ''dynamically equivalent'' and specifically they are in one-to-one correspondence with one another.
-However when $S$ is semi-conjugate to $T$ with $\phi$ a many-to-one mapping, the dynamics on $V$ provides a coarse-grained \textbf{cite?} description of the dynamics on $U$ .
-When $S$ is semi-conjugate to $T$, it is also common to call $S$ a factor of $T$, or conversely that $T$ is an extension of $S$.
+When the relationships in this diagram hold, we say $\phi(T(u))= S(\phi(u))$, and we formally denote it as $\phi \circ T=S\circ \phi$.
+
+\begin{Definition}  
+  [\bf {Conjugacy \& Semi-Conjugacy}]\label{Dfn_Conjugate}
+  Consider the dynamical systems $T:U\to{U}$, $S:V\to{V}$ and suppose the relationship $\phi \circ T=S\circ \phi$ holds.
+If $\phi:U\to{V}$ is a homeomorphism, then $S$ is said to be conjugate to $T$ and $\phi$ is called a conjugacy.
+If we relax the criterion on $\phi$ and merely require $\phi$ to be continuous and surjective where $\phi:U\to{V}$, then $\phi$ is a semi-conjugacy between $T$ and $S$ where T has domain $U$ and S has domain $V$; S is said to be semi-conjugate to T.
+\end{Definition}
+
+\ednote{M: Saying $\phi$ is then a semi-conjugacy between $T$ and $S$ could be ambiguous as it does not specify which system is on the top in the commutativity.
+So please edit accordingly.
+B: Does defining explicitly the domain and codomain of $phi$ make it clear enough?
+}
+When $S$ is conjugate to $T$, the dynamics of the two systems are in some way ''dynamically equivalent''.
+Specifically, they are in one-to-one correspondence with one another.
+However when $S$ is semi-conjugate to $T$ with $\phi:U\to{V}$ a many-to-one mapping, the dynamics on $V$ provide merely a coarse-grained description of the dynamics on $U$ \cite{de2013elements}.
+When $S$ is semi-conjugate to $T$, it is also common to call $S$ a \emph{factor of $T$}, or conversely that $T$ is an \emph{extension of $S$}.
 In essence, an extension (e.g., \cite{de2013elements}) is a larger system capturing all of the important dynamics of its factor.
 
-% It is a very hard or nearly an impossible task to establish the existence of a conjugacy or a semi-conjugacy $\phi$ between two systems.
-\textbf{cite?} However, one can verify that a function $\phi$ satisfies the commutativity diagram \ref{eqn_conjugacy}.
-For example $\phi(x)=\sin(\pi x/2)^2$ is a conjugacy between two systems $Tu=1-|2u-1|$ and $Sv=4v(1-v)$  defined on $U=[0,1]$.
-
-% In establishing that two systems are conjugate (semi-conjugate) to one another, we have also shown that one may choose to work with one systems as opposed to the second and still be guaranteed to obtain information on the latter.
-This is incredibly useful while forecasting the future evolution of dynamical systems.
-
-% \chapter{The Learning Problem}\label{ch3}
-
-% \emph{Introduction...}
-
-% Consider a relatively simple learning problem:
-
-% Given $(u_0, u_1, \ldots, u_m)$ for $m\in\mathbb{Z}$, a finite segment of an orbit of the map $T$ where $T$ is defined by the update equation $u_{n+1} = Tu_n$, forecast the values $u_{m+1}, u_{m+1}$ where the map $T$ is unknown, given that $u_m$.
-
-% A practical example of this would be - given the sequential coordinates of an object moving in space, predict the future positions of that object.
+It is a very hard or nearly an impossible task to establish the existence of a conjugacy or a semi-conjugacy $\phi$ between two systems \cite{devaney2018introduction}.
+However, one can verify that a function $\phi$ satisfies the commutativity diagram \ref{eqn_conjugacy}.
+
+\textbf{NB NB FINISH}
+\begin{Example}
+  For example $\phi(x)=\sin(\frac{\pi}{2}x)^2$ is a conjugacy between two systems $Tu=1-|2u-1|$ and $Sv=4v(1-v)$  defined on $U=[0,1]$.
+
+  To see this, we consider $\phi\circ{T} = \sin(\frac{\pi}{2}(1-|2u-1|))^2$
+
+  and $S\circ\phi = 4\sin(\frac{\pi}{2}x)^{2}(1-\sin(\frac{\pi}{2}x)^2)$
+\end{Example}
+
+In establishing that two systems are conjugate (semi-conjugate) to one another, we have also shown that one may choose to work with one system as opposed to the second and still be guaranteed to obtain information on the latter.
+This is especially useful while forecasting the future evolution of dynamical systems.
+
+\chapter{The Learning Problem}\label{ch3}
+
+In this chapter we acquaint ourselves with the question of forecasting dynamical systems with unknown underlying structures, state and discuss the shortcomings of the Takens delay embedding theorem and discuss various issues faced while forecasting.
+
+Consider a relatively simple learning problem: 
+Given the sequence $(u_0, u_1, \ldots, u_m)$ for $m\in\mathbb{Z}$, a finite segment of an orbit of the map $T$ where $T$ is defined by the update equation $u_{n+1} = Tu_n$, forecast the values $u_{m+1}, u_{m+1}$ where the map $T$ is unknown, given that $u_m$.
+
+A practical example of this would be the subsequent scenario: given the time-sequential coordinates of an object moving in space, predict the future positions of that object.
 However we are very rarely, if at all, presented with a problem where the entire state information is available to us in the form $u_n$ at some time-step $n\in\mathbb{N}$.
@@
 
-% Suppose we only have the observations $\theta(u_0), \theta(u_1), \ldots, \theta(u_m)$ of the true system states $u$ in an unknown dynamical system $(U,T)$ and we wish to predict the values $\theta(u_{m+1}), \theta(u_{m+2})$.
-
-% Here we consider a continuous-time dynamical system $(U,T)$ with true state $u$ in the input space $U$ and then a measured signal $\theta$ (where $\theta(u_n)=\theta(u(t_n))$ for $n\in\mathbb{N}$.
-The sequence $\{x_n\}$ represents a scalar time-series and intuitively $\theta$ may be thought to represent a probe inserted into a larger system and itself measuring/extracting only a small part of the greater system state at time $t=n$.
+Suppose we only have the observations $\theta(u_0), \theta(u_1), \ldots, \theta(u_m)$ of the true system states $u$ in an unknown dynamical system $(U,T)$ and we wish to predict the values $\theta(u_{m+1}), \theta(u_{m+2})$ and so forth.
+
+First we establish the method of Takens delay embedding. 
+In this method, one considers a discrete-time dynamical system defined on a 'nice' space (a smooth manifold)\ednote{Should this be further explained?} that can be obtained as time-$K$ map of a continuous time-dynamical system \ednote{B: Should one add here: "These notions will be made formal below"?}.
+We make observations from such a system, i.e., we observe the evolution of $\theta(w_0)$ (where $w_0$ represents the initial state) i.e., we observe  a finite set of values of $u_n :=  \theta(w_n) = \theta(Tw_{n-1})$.
+The sequence $\{u_n\}$ represents a scalar time-series and intuitively $\theta$ may be thought to represent a probe inserted into a larger system which is itself only measuring/extracting a small part of the greater system state at time $t=n$.
 Consider for example a thermometer erected to measure the ambient temperature in a local village.
-This measurement function, the thermometer is capturing a single aspect, the temperature, of a grander dynamical system entailing the present weather of the surrounding area, but even more than that, it is measuring an incredibly small part of the global weather system.
-
-% From here we construct a multidimensional observable using the method of delay-coordinate stacking. 
-% \textbf{move to later} Given the sequence $\{\theta_1, \theta_2, \ldots\}$, we define the variable \newline $y_n:=[\theta_n, \theta_{n-\tau}, \ldots, \theta_{n-(2d-2)\tau}, \theta_{n-(2d-1)\tau}]^{T}\in\mathbb{R}_{d}\times{1}$ where $\tau$ represents the lag and $d$ the dimension of the attractor.
+This measurement function, the thermometer, is capturing only a single aspect, the temperature, of a much grander dynamical system entailing the present weather of the surrounding area.
+Even more than that though- it is measuring a miniscule part of the global weather system.
+
+%Given the sequence $\{u_1, u_2, \ldots, u_n}$, we define the variable \newline $y_n:=[\theta_n, \theta_{n-\tau}, \ldots, \theta_{n-(2d-2)\tau}, \theta_{n-(2d-1)\tau}]^{T}\in\mathbb{R}_{d}\times{1}$ where $\tau$ represents the lag and $d$ the dimension of the attractor.
 (We shall return to these terms shortly and their discussion may be put on hold for the time being).
 
-% However, the true state $u$ of a system is seldom, if ever fully known.
-In almost all cases we can merely insert probes into a system to obtain partial information by measurements taken.
-Moreover, the process of taking a measurement itself introduces 2 additional difficulties:
-% \vspace{-8mm}
-% \begin{enumerate}[noitemsep, label=\roman*.]
-%   \item A series of measurements over a specific time-interval is inherently a discretisation
-%   procedure of the underlying continuous-time dynamical system. (Hence why we restricted our attention to discrete-time systems in the preceding Chapter)
-%   \item A probe will never be fully accurate and so the act of measurement introduces a certain
-% measure of numerical noise/inaccuracy.
-% \end{enumerate}
-
-% More formally we may say that in the process of measuring a system, we obtain a 1-dimensional time-series $\{\theta(u(t))\}$, where $\theta:\mathbb{R}^n\to\mathbb{R}$ is a smooth observation or measurement function.
-This series represents a possibly noise-augmented dataset containing information of certain parts of the original system.
-
-% We may then ask ourselves the question: Is it in any way possible to retain information about the system state $x(t)$ in this temporal data-series $\theta(x(t))$?
+However, the true state $u$ of a system is seldom, if ever, fully known.
+In almost all cases we can at most insert probes into a system to obtain partial information by means of the measurements taken.
+Moreover, the process of taking a measurement itself introduces 2 additional aspects which complicate the problem: \ednote{M: Maybe replace difficulty by saying that we need to consider two aspects.
+B: Perhaps this edit makes more sense?}
+\vspace{-8mm}
+\begin{enumerate}[noitemsep, label=\roman*.]
+  \item A series of measurements over a specific time-interval is inherently a discretisation procedure of the underlying continuous-time dynamical system.
+(Hence why we restricted our attention to discrete-time systems in the preceding section)
+  \item A probe will never be fully accurate, and so the act of measurement introduces a certain measure of numerical noise/inaccuracy.
+\end{enumerate}
+
+From here we construct a multidimensional observable using the method of stacking previous observations, i.e., we create delay-coordinate map defined by
+$\Phi_{k,\theta}(w) := (\theta(T^{-k}w)\ldots,\theta(T^{-1}w),\theta(w))$.
+The essence of Takens theorem relates to the fact that when $k$ is sufficiently large, we can define a dynamical system on the space $\mathbb{R}^{k+1}$ whose states are $\Phi_{k,\theta}(w)$, $\Phi_{k,\theta}(Tw)$, $\Phi_{k,\theta}(T^2w)$, etc.
+and this dynamical system is topologically conjugate to the unknown underlying system $(W,T)$.
+We recall the Takens delay embedding theorem next.
+
+%More formally we may say that in the process of measuring a system, we obtain a 1-dimensional time-series $\{\theta(u(t))\}$, where $\theta:\mathbb{R}^n\to\mathbb{R}$ is a smooth observation or measurement function.
+This series represents a possibly noise-augmented \ednote{M: Noise-augmented is not the right word} dataset containing information of certain parts of the original system.
+
+%We may then ask ourselves the question: Is it in any way possible to retain information about the system state $x(t)$ in this temporal data-series $\theta(x(t))$?
 The answer is easily yes if T is known.
@@
 
-% \section{Taken Embedding Theorem} \label{sect_TakensEmbedding}
-
-% We define first the concepts of homeomorphism and embedding:
-% \begin{Definition}
-% [\bf Homeomorphism] \label{Dfn_Homeo}
-%   A homeomorphism is a function $f:Z\rightarrow Y$ between two topological spaces $Z$ and $Y$ that is continuous, bijective and has a continuous inverse. 
-% \end{Definition}
-
-% \begin{Definition}
-% [\bf Embedding] \label{Dfn_Homeo}
-%   Consider a homeomorphism $f:Z\rightarrow Y$ for $Y\in X$. $f$ is said to be an embedding and $Z$ to be embedded in $X$ by $f$.
-% \end{Definition}
-
-% Takens Theorem states a result that establishes a relationship between the observed and underlying dynamical systems by showing that the concetantion of a sufficiently large number of previous observations into a vector will, under certain conditions, generate a map between the vectors from the respective systems.
-\textbf{Check wording} We formultate the theorem from \cite{takens1981detecting}.
-
-% \begin{Theorem} 
-% 	[\bf Takens Embedding Theorem (adopted from \cite{takens1981detecting}] \label{Thm_Takens}
-%             Let $W$ be a compact manifold of dimension $m$, and $d\ge m$ so that $2d\in\mathbb{Z}$. It is a 
-%             generic property for the pair $(T, \theta)$,  where $T:W \to W$ is
-%             a smooth diffeomorphism, and $\theta:W \to \mathbb{R}$ a smooth function, that the map $\Phi_{2d,\theta}:W \to \mathbb{R}^{2d+1}$ defined on $W$ by 
-%             $\Phi_{2d,\theta}(w) := (\theta(T^{-2d}w)\ldots,\theta(T^{-1}w),\theta(w))$
-%             is a diffeomorphic embedding; by `smooth' we mean at least $C^2$. Consequently, there exists a map $\Ftheta: \Phi_{2d,\theta}(W) \to \Phi_{2d,\theta}(W)$ defined by 
-%             \begin{equation}  \label{eqn_takens}
-%               \Ftheta: (\theta(T^{-2d}w),\ldots,\theta(T^{-1}w),\theta(w)) \mapsto  (\theta(T^{-2d+1}w),\ldots,\theta(T^{-1}w),\theta(Tw))
-%             \end{equation}
-%            so that $(W,T)$ is topologically conjugate to 
-%             $(\Phi_{2d,\theta}(W), \Ftheta)$.   
-% \end{Theorem}
-
-% (In our scenario, the input space $U$ is considered the attractor.)
-
-% Some explanation is beneficial.
+\section{Taken Embedding Theorem}
+
+We define first the concepts of homeomorphism and embedding:
+\begin{Definition}
+  [\bf {Homeomorphism}]\label{Dfn_homeo}
+  A homeomorphism is a function $f:Z\rightarrow Y$ between two topological spaces $Z$ and $Y$ that is continuous, bijective and has a continuous inverse. 
+\end{Definition}
+
+\begin{Definition}
+  [\bf {Embedding}]\label{Dfn_embed}
+  Consider a homeomorphism $f:Z\rightarrow Y$ for $Y\subset X$. $Z$ is said to be embedded in $X$ by $f$.
+\end{Definition}
+
+Takens Theorem states a result establishing a relationship between the observed and underlying dynamical systems by showing that the concatenation of a sufficiently large number of previous observations into a vector will, under certain conditions, generate a map between the vectors from the respective systems.
+We formulate the theorem from \cite{takens1981detecting}.
+
+\begin{Theorem} 
+	[\bf Takens Embedding Theorem (adopted from \cite{takens1981detecting}] \label{Thm_Takens}
+         Let $W$ be a compact manifold of dimension $m$, and $d\ge m$ so that $2d$ is an integer. It is a 
+            generic property for the pair $(T, \theta)$,  where $T:W \to W$ is
+            a smooth diffeomorphism, and $\theta:W \to \mathbb{R}$ a smooth function, the map $\Phi_{2d,\theta}:W \to \mathbb{R}^{2d+1}$ defined on $W$ by 
+            $\Phi_{2d,\theta}(w) := (\theta(T^{-2d}w)\ldots,\theta(T^{-1}w),\theta(w))$
+            is a diffeomorphic embedding; by `smooth' we mean at least $C^2$.
+Consequently, there exists a map $F_\theta: \Phi_{2d,\theta}(W) \to \Phi_{2d,\theta}(W)$ defined by $$F_\theta: (\theta(T^{-2d}w),\ldots,\theta(T^{-1}w),\theta(w)) \mapsto
+            (\theta(T^{-2d+1}w),\ldots,\theta(w),\theta(Tw))$$
+           so that $(W,T)$ is topologically conjugate to 
+            $(\Phi_{2d,\theta}(W), F_\theta)$.    
+\end{Theorem}
+
+By generic we mean a residual set on a certain topology on appropriate spaces of functions (that we do not describe here) \textbf{(Cite)}.
+%\ednote{How vague may one be in general by leaving things out?}
+By $C^2$, we make reference to a twice-differentiable function  with a continuous second derivative.
+In our scenario (and this remains important throughout) the input space $U$ is considered the attractor.~\label{attractor_U}
+
+Below we reproduce Figure 1 from \cite{Supp}.
+
+%Some explanation is beneficial.
 Takens establishes a delay-coordinate map $\Ftheta$ defined in \ref{eqn_takens} for $T$ the flow (\ref{defn_flow}) defined by the update equation  $Tu_n=u_{n+1}$ and $theta$ our measurement function.
@@
 
-% Alternatively we may also make the statement: Takens' Embedding Theorem guarantees that almost all dynamical systems can be reconstructed from just one noiseless observation sequence \textbf{this isn't obvious from the theorem's formulation, is it?} i.e.
+%Alternatively we may also make the statement: Takens' Embedding Theorem guarantees that almost all dynamical systems can be reconstructed from just one noiseless observation sequence \textbf{this isn't obvious from the theorem's formulation, is it?} i.e.
 for a great number of possible observation functions $\theta$, $\Ftheta$ preserves the topology of U.
@@
 
-% \begin{figure}[ht]
-%   \includegraphics[scale=0.3]{takensmap.png}
-%   \centering
-%   \label{fig:takensmap}
-% \end{figure}
-
-% Once again we pause and consider the graph to consolidate our understanding of the conjugacy (\textbf{are we being too simplistic?}).
+\begin{figure}[ht]
+  \includegraphics[scale=0.25]{takensmap.png}
+  \centering
+  \captionof{figure}{Schematic of Takens Embedding Theorem's usage and consequence of how iterates \ldots }
+  \label{fig:takensmap}
+\end{figure}
+\ednote{M: Here, and throughout if you reproduce a figure, it must be stated where it was reproduced from.
+B: How does one write that a figure is reproduced?
+Also, should $\tilde{F}$ be left out for now?}
+\ednote{B: Could I perhaps have the original figure? This is only a screenshot used as a temporary placeholder.}
+
+If we can learn the map $F_\theta$, from sufficient number of data points of $\{\phi_{2d,\theta}(w_n)\}$, then we also know how to forecast how $\theta(w_n)$.
+
+%Once again we pause and consider the graph to consolidate our understanding of the conjugacy (\textbf{are we being too simplistic?}).
 If we are in state $u$ and then evolve towards state $Tu$ before embedding into $\mathbb{R}^{2m-1}$, then this is equivalent to first embedding into $\mathbb{R}^{2m-1}$ and then evolving through the map $\Ftheta$, i.e.
@@
 
-% Thus information about U can be retained in the time series' (or observation's) output.
+%Thus information about U can be retained in the time series' (or observation's) output.
 By preserving the topology on the manifold $U$ in the reconstruction space $X$ (\textbf{not discussed before}), we are also guaranteed the preservation of topological invariants of the manifold, of which dimensionality is one such invariant.
@@
 
-% \section{Obstacles to Takens}
-% Despite the fact that the Takens' Embedding Theorem is a powerful result and provides compelling reason to believe that it is possible to accurately ()\emph{right word?}) reconstruct the underlying space, it does present some serious limitations:
-% \vspace{-5mm}
-% \begin{enumerate}
-% \item Even if we can indeed find \Ftheta{}  our \emph{approximation} of \Ftheta{}  is a map from a larger set $\mathbb{R}^{2m+1}$ containing the embedded attractor.
-There are, however, no theoretical guarantees that \Ftheta{}  will keep this attractor $U$  the same.(\textbf{We haven't properly said that U is an attractor, or have we.})
-% \item Takens only holds for noiseless observations.
-Suppose the case where \Ftheta{} was learnt, but due to measurement error or other noise we obtained the set $V$ and not the original attractor.
-Due to the chaotic nature of the underlying system (i.e.
-the fact that it has SIDC), this new system $V$ will then move in a different direction.
-This problem will later be eliminated by some properties of a new function $g$, namely the Unique Solution Property.
+\section{Practical Issues of Takens theorem}
+
+It is a fact that Takens' Embedding Theorem is both a powerful result and that it provides compelling reason to believe that one could conceivably reconstruct accurately a system conjugate to the underlying system.
+Nonetheless, it does present some serious practical limitations:
+\vspace{-5mm}
+\begin{enumerate}
+\item Even supposing that we could indeed find $F_\theta$ , our \emph{approximation} of  $F_\theta$ is a map from a larger set $\mathbb{R}^{2d+1}$ containing the embedded attractor.
+There are, however, no theoretical guarantees that $F_\theta$ will retain  $V=\Phi_{2d,\theta}(W)$  as an attractor although $W$ could itself be an attractor.
+\item Takens theorem is stated for noiseless observations.
+Due to noise $\epsilon_n$ the delay vector $\Phi_{2d,\theta}(w_n) + \epsilon_n$  may lie outside $V$.Furthermore, due to the chaotic nature of the underlying system (i.e.
+the fact that it has SIDC), the evolution of $\Phi_{2d,\theta}(w_n) + \epsilon_n$ under the map $F_\theta$ could move out of $V$ completely.
+This problem can be overcome  by using a driven dynamical system with some properties, and we discuss this in Chapter *.
 %  \item It is an unfortunate case that the functional complexity (i.e., the relative number of oscillatory graphs \cite{manjunath2021universal}) as well as stability of the embedding \Ftheta{} (\textbf{cite?})depends on the choice of observable $\theta$.
 This means that we may even learn a map that complicates our learning task and in so doing adds obstacles to the road to learning a conjugate map with predictive power.
-(\textbf{Am I talking nonsense here?})
-% \end{enumerate}
-
-% Another limitation owes itself to the requirement that T be a surjective map. This is not always the case in practical examples and we may consider the case of a damped double pendulum.
-
-% \textbf{Continue writing here}
-
-% \begin{figure}[ht]
-%   \includegraphics[scale=0.3]{learningfailure.png}
-%   \centering
-%   \label{fig:supp_learning_failure}
-%   \caption{Illustration of a system having failed to learn the map}
-% \end{figure}
-
-% \emph{ADD THIS TO ERRORS IN TAKENS: NOT GUARANTEED GLOBALLY DISSIPATIVE}
-% \textbf{Explanation continued here}
-
-% \begin{figure}[h]
-%   \includegraphics[scale=0.4]{temporary_damp_fig.png}
-%   \centering
-%   \label{fig:damped_pendulum}
-% \caption{A damped pendulum}
-% \end{figure}
-
-% \textbf{Continue explaining here}
-% \newpage
-
-\chapter{Driven Dynamical Systems to Forecast Problems} \label{ch4}
+\ednote{Am I talking nonsense here?
+M: Please avoid this point and delete it}
+\end{enumerate}
+
+% Another limitation owes itself to the requirement that T be a surjective map.
+This is not always the case in practical examples and we may consider the case of a damped double pendulum \ednote{M: Please remove the paragraph as well.
+We do not overcome this problem.
+You can add this in the next Chapter where you need T to be surjective or else you could include this at the begining of this chapter where you explain the forecasting problem.}
+
+\textbf{Continue writing here}
+
+\begin{figure}[ht]
+  \includegraphics[scale=0.3]{learningfailure.png}
+  \centering
+  \label{fig:supp_learning_failure}
+  \caption{Illustration of a system having failed to learn the map}
+\end{figure}
+
+\textbf{Explanation continued here}
+
+\begin{figure}[h]
+  \includegraphics[scale=0.4]{temporary_damp_fig.png}
+  \centering
+  \label{fig:damped_pendulum}
+\caption{A damped pendulum}
+\end{figure}
+
+\textbf{Continue explaining here}
+
+\chapter{Driven Dynamical Systems to Forecast Problems}\label{ch4}
 
 In this chapter, we discuss results pertaining to the mapping of temporal data obtained from a discrete-dynamical system onto a different space through the notion of a driven dynamical system. 
-We also consider the conditions a driven system should have to avoid adding distortion to its state space representation.
-We then describe its \emph{single-delay dynamics} (SDD) and then recall conditions on the driven system so that the SDD is conjugate (or at least semi-conjugate) to the underlying system.
+We also consider the conditions a driven system should possess to avoid adding distortion to its state-space representation.
+We then describe the \emph{single-delay dynamics} (SDD) of the system and and finally recall conditions on the driven system so that the SDD are conjugate (or at least semi-conjugate) to the underlying system.
 The SDD can then be used to forecast and reconstruct the underlying system. 
-\ednote{Is this attractor assumed to exist?
-M:  I replaced  it with the system.
-Attractor is also a commonly used terminology since we observe data from an attractor in almost all cases.}
+% \ednote{Is this attractor assumed to exist? M:  I replaced  it with the system. Attractor is also a commonly used term since we observe data from an attractor in almost all cases.}
 
@@
 
-A nonautonomous dynamical system is simply a dynamical system (as defined before in \ref{Dfn_DS}) \ednote{M:I am not sure where it was defined before.
-If not, please say, A NDS comprises a family of map $\{g_n\}_{n \in \mathbb{Z}}$ and say  each $g_n$ could arise as a consequence of an input $u_n$.} from the input space $U$, a  topological space, is time-dependent.
-
-We immediately recall the figure first defined in \ref{eqn_conjugacy}  \ednote{Please use eqref instead of ref while cross-citing equations} and remind ourselves of our ultimate goal to forecast the evolution of an unknown nonautonomous dynamical system $(U,T)$ \ednote{We are not forecasting a nonautonomous, but an autonomous system} by constructing a conjugate system such as $(V,S)$ below.
-\ednote{M: Maybe something like this sounds better: We learn a system that is conjugate to an underlying or unknown dynamical system by using the measurements from  the unknown system.}
+\begin{Definition}
+  [\bf Nonautonomous Dynamical System (adopted from~\cite{Manju_ESP}))] \label{Dfn_Soln}\rm
+  A nonautonomous dynamical system (NDS) on a space $X$ is simply a dynamical system comprising a family of maps $\{g_n\}_{n \in \mathbb{Z}}$ where $g_n(\ldotp):=g(u_n,\ldotp):X\to{X}$ is continuous.
+Each $g_n$ arises as a consequence of an input $u_n$ from the input space $U$, a topological space.
+\ednote{Does the reader immediately remember/know what is meant by a dynamical system?}
+\end{Definition}
+
+We immediately recall the figure first defined in \eqref{eqn_conjugacy}  and remind ourselves of our ultimate goal to learn a system that is conjugate to an underlying or unknown dynamical system $(U,T)$ by making use of the measurements obtained from  the unknown system.
 
@@
 
-To find a $\phi$ \ednote{M: Have edited here}, the need arises to consider a driven dynamical system where input is taken from an input space $U$, and the state of a space $X$ is updated. 
-This is done to account, in a manner more similar to the true scenario, for the influence of both the input and the actual state of the system at a specific time-step.
+To find a function $\phi$, the need now arises to consider a so-called driven dynamical system, a special case of a nonautonomous dynamical system, where two spaces are considered: input is taken from an input space $U$ and the state of a space $X$ is updated.
+(Previously, only the state was considered.)
+We do this to mimic the true conditions.
+By accounting for both external and internal factors which influence the  system-state at a specific time-step, our model resembles nature much more realistically.
 \ednote{M : I understand this sentence, but am not sure if a general audience would.
-So please elaborate the sentence or omit}
-
-\begin{Definition}
-  [\bf Driven and Compactly Driven Dynamical System] \label{Dfn_DDS} \rm
+So please elaborate or omit.
+B: Is this satisfactory?}
+
+\begin{Definition}
+  [\bf Driven, Compactly Driven Dynamical System] \label{Dfn_DDS} \rm
 A driven dynamical system comprises two topological metric spaces $U$, $X$ and a continuous function  $g:U\times{X}\to{X}$ where $g(u_n, x_n)=x_{n+1}$.
+If the input space $U$ is compact, we refer to the system as compactly driven. 
+\end{Definition}
+
 The dynamics on X are generated by the update equation $x_{n+1}=g(u_n, x_n)$ where $n\in\mathbb{Z}$, input $u_n$ from $U$ and state $x_n$ belonging to $X$, where $X$ is compact.
-\ednote{M : The sentence ``The dynamics on X are generated by ..." does not define anything, so mathematicians do not include it as part of a definition, and can be moved after the definition.}
-If the input space $U$ is compact, we refer to the system as compactly driven. 
 Abbreviated, we shall refer only to the \emph{driven system $g$}, with all other entities being understood implicitly.
-\end{Definition}
-
-In particular, a nonautonomous dynamical system may be generated from $U$; any input $\overline{u}$, a bi-infinite sequence from $U$, gives rise to the sequence of self-maps $\{g(u_n, \cdot)\}_{n\in\mathbb{Z}}$ contained in $X$.
-Physically, one may think of this bi-infinite sequence as referring to a system that has been running for an incredibly long period at the time of the first measurement taken from the system (alternatively, the first time a probe is inserted into the system to take an observation) \ednote{Time is a label we put.
-We can label the first sample as $0$th or $-10000$th.
-So the word incredibly is not necessary, all we need is good amount of data.}
-
-\begin{Definition}
-  [\bf Entire Solution] \label{Dfn_Soln}
-  A sequence $\{x_n\}_{n\in\mathbb{Z}}\subset X$ \ednote{Please avoid subseteq as it is ambiguous} is called an entire solution(or simply a solution) of the driven system  $g$ with input $\overline{u}$ when it satisfies
+
+Notably, a nonautonomous dynamical system may be generated from $U$; any input $\overline{u}$, a bi-infinite sequence from $U$, gives rise to the sequence of self-maps $\{g(u_n, \cdot)\}_{n\in\mathbb{Z}}$ contained in $X$.
+Physically, one may think of this bi-infinite sequence as referring to a system that has been running for a long period at the time of the first measurement taken from the system (alternatively, the first time a probe is inserted into the system to take an observation)
+
+\begin{Definition}
+  [\bf Entire Solution] \label{Dfn_Soln} \rm
+  A sequence $\{x_n\}_{n\in\mathbb{Z}}\subset X$ is called an entire solution(or simply a solution) of the driven system  $g$ with input $\overline{u}$ when it satisfies 
   \[g(u_{n-1}, x_{n-1})=x_n\] for all $n\in\mathbb{Z}$
@@
 Consider the example below.
-
-\ednote{M: Only theorems, lemmas and propositions are in italics. For Definitions, remarks and examples, please add slash-rm command to keep it in normal mode}
 
@@
 \end{Example}
-\ednote{Nicely written!}
+% \ednote{Nicely written!}
 
@@
 The system has an uncountable number of solutions, as there exists a solution for every $x\in{X}$ which also passes through the point $x$ and $lim_{n\to\infty}x_n=0$, $lim_{n\to{-}\infty}x_n=1$.
- The proof is deferred to immediately after the next paragraph. \ednote{Make the proof a remark and please cite the remark}
+ The proof is deferred to immediately after the next paragraph (see~\ref{rem_proofEx}).
 
@@
 \begin{Definition}
-  [\bf Reachable Set]\label{Dfn_ReachableSet}
+  [\bf Reachable Set]\label{Dfn_ReachableSet}\rm
 The reachable set of a driven system $g$ is exactly the union of all the elements of all the solutions, i.e., 
@@
 
-We note that $x\in{X_n(\overline{u})}$ if and only $g$ has a solution $\{x_k\}$ for $x_n=x$ and input $\overline{u}$. \textbf{Cite}.
+We note that $x\in{X_n(\overline{u})}$ if and only $g$ has a solution $\{x_k\}$ for $x_n=x$ and input $\overline{u}$. \textbf{(Cite)}.
+
+$X_n$ is exactly the set of points $x$ through which some entire solution $\Psi$ passes at time $n$.
+
 This will lead to a result established later, but which is worth taking note of now: \textit{$g$ being a topological contraction is equivalent to the existence of a unique entire solution.} To this end we define the notion of a topological contraction.
-\ednote{I do not think you need the definition of a topological contraction unless it is defined through a process.
+\ednote{M: I do not think you need the definition of a topological contraction unless it is defined through a process.
 Please note that you have defined $X_n(\bar{u})$ differently than what was in the earlier works.
-Now it is just the collection of solution values at time $n$ with the input $\bar{u}$ and topological contraction is just paraphrasing the USP definition.}
-
-\begin{Definition}
-  [\bf Topological Contraction]\label{Dfn_TopContr}
-  A function $g:U\times{X}\to{X}$ is a topological contraction if for all $n\in\mathbb{Z}$ and all $\overline{u}\subseteq{U}$, $X_n(\overline{u})$ is a singleton subset of $X$.  
-\end{Definition}
-
-\begin{proof}
-  [\bf Proof of Example\ref{Ex_exp}] 
+Now it is just the collection of solution values at time $n$ with the input $\bar{u}$ and topological contraction is just paraphrasing the USP definition.
+B: Is it better to then just leave out the entire sentence?}
+
+\begin{Definition}
+  [\bf Topological Contraction]\label{Dfn_TopContr}\rm
+  A function $g:U\times{X}\to{X}$ is a topological contraction if for all $n\in\mathbb{Z}$ and all $\overline{u}\subset{U}$, $X_n(\overline{u})$ is a singleton subset of $X$.  
+\end{Definition}
+
+\begin{Remark}
+  [\bf Proof of Example~\ref{Ex_exp}] \label{rem_proofEx} \rm
   To see this, we show that for every input $\overline{u}=\{u_n\}_{n\in\mathbb{Z}}$, $g(u_n,\cdot)$ is a contraction map on $X$ for every $k\in\mathbb{Z}$. 
@@
   We then easily conclude that $\{x_n\}\equiv0$ is the only solution. 
-\end{proof}
+\end{Remark}
 
@@
 \begin{proof}
-  Consider an input $\overline{u}={\{u_n\}}_{n\in\mathbb{Z}}\subseteq{U}$ and driven system $g:U\times{X}\to{X}$  generating a sequence ${\{x_n\}}_{n\in\mathbb{Z}}$ in the compact space $X$ \ednote{This statement is not correct.}.
-  Since $X$ is a metric space, it follows immediately from a well-known result in Analysis (\textbf{Cite}) that $\{x_n\}$ has a convergent subsequence which then is a solution.
+  \textbf{I've muddled myself here.}
+  % Consider an input $\overline{u}={\{u_n\}}_{n\in\mathbb{Z}}\subset{U}$ and driven system $g:U\times{X}\to{X}$  generating a sequence ${\{x_n\}}_{n\in\mathbb{Z}}$ in the compact space $X$ \ednote{This statement is not correct.}.
+  % Since $X$ is a metric space, it follows immediately from a well-known result in Analysis (\textbf{Cite}) that $\{x_n\}$ has a convergent subsequence which then is a solution.
 \ednote{This is not a valid proof.
@@
 
-We may easily construct many systems with trivial solution-sets, such as $g(u,x)=x$ which has only the constant solution $x$ and so for $U=[-1,1]$, the system would have no solution if $|x|>1$.
+One may easily construct many systems with trivial solution-sets, such as $g(u,x)=x$ which has only the constant solution $x$ and so for $U=[-1,1]$, the system would have no solution if $|x|>1$.
 To refine the scenario, we consider only systems with unique solutions.
@@
 \begin{Definition}
-  [\bf Unique Solution Property] \label{Dfn_usp}
+  [\bf Unique Solution Property] \label{Dfn_usp}\rm
   A driven system $g$ is said to have the Unique Solution Property (USP) if for each input $\overline{u}$ there exists exactly one solution. 
@@
 
-One of the first result obtained upon defining the USP, is the fact that every solution will attract any different initial conditions towards the component parts of the solution, i.e.
+One of the first result obtained upon defining the USP, is the fact that every solution will attract any different initial conditions towards the component parts of the solution.
+%, i.e.
 that the solution $\Psi$ is a non-autonomous uniform attractor.
 \ednote{Define Uniform Attractor before stating theorem or else omit the statement.}
-
-\begin{Theorem} \label{Thm_AttractorSoln}
-  If $g$ has USP, then any solution to $g$ is a non-autonomous uniform attractor 
-\end{Theorem}
-\begin{proof}
-\ednote{M: You can omit the theorem, and just say it is a uniform attractor in the nonautonomous dynamical systems literature \cite{Manju_Nonlinearity}, and say the discussion on nonautonomous attractors is beyond the scope of the thesis.}
-
-\end{proof}
-
-\ednote{Additional discussion on USP? M: Please do so by stating that it is equivalent to the echos state property. See the paper with Adriaan and paraphrase} 
+If $g$ has the USP, then any solution to $g$ is a uniform attractor in the nonautonomous dynamical systems literature \cite{Manju_Nonlinearity}.
+The discussion on nonautonomous attractors is beyond the scope of this thesis and we refer the reader to \cite{Manju_ESP, esann2012ids} for further reading.
+
+% \ednote{M: You can omit the theorem, and just say it is a uniform attractor in the nonautonomous dynamical systems literature \cite{Manju_Nonlinearity}, and say the discussion on nonautonomous attractors is beyond the scope of the thesis.}
+\textbf{NEED TO PARAPHRASE}
+\newline In our context, $g$ having the USP is equivalent to $g$ being a topological contraction \cite{manjunath2021universal}, or alternatively that $g$ exhibits the Echo State Property(ESP). 
+In \cite{Manju_ESP} the echo state property is a key for the design and training of recurrent neural networks within the paradigm of reservoir computing.
+Available definitions and theorems concerning the echo state property, however, are of little practical use because they do not relate the network response to temporal or statistical properties of the driving input.
+
+A greater discussion is presented in \cite{jaeger2001echo}
+a particular stability concept for which a number of equivalent definitions are available \cite{jaeger2001echo}.
+
+xxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxx
+
 Having already defined the reachable set $X_U$, we pause for a moment to fix additional notation.
@@
 
-The question now becomes whether we may establish a semi-conjugacy \ednote{M: I have changed it to a semi-conjugacy since conjugacy amounts to asking a infinite-dimensional space $\overleftarrow{U}$ to be embedded in $X_U$ which is not possible.
-} as presented below for the driven system $g$.
+The question now becomes whether we may establish a semi-conjugacy %\ednote{M: I have changed it to a semi-conjugacy since conjugacy amounts to asking a infinite-dimensional space $\overleftarrow{U}$ to be embedded in $X_U$ which is not possible.}
+as presented below for the driven system $g$.
 
@@
 Regrettably, such a continuous mapping $h$ is not guaranteed to exist \cite{M: Have edited here} when $g$ does not have the USP \cite[Lemma 5]{Manju_Nonlinearity}.
+Note that even when $h$ does exist, we are not guaranteed its injectivity. Considering again example \ref{ex_halfux}, we see that even if $h$ were to exist, it could not be injective as $X_U=\{0\}$.
 
@@
 In this case, the function $H:\overleftarrow{U}\to\overleftarrow{X}_U$, a map that is both continuous and surjective, is called a \emph{causal mapping}.
-make special note of a \ednote{something you want to fill?} 
+
+\textbf{make special note of a} %\ednote{something you want to fill?} }
 \begin{Definition}
@@
 
-Note that even when $h$ does exist, we are not guaranteed its injectivity.
-Considering again example \ref{ex_halfux}, we see that even if $h$ were to exist, it could not be injective as $X_U=\{0\}$ \ednote{M: I feel these two sentences should be made after the definition of $h$ rather than here}.
-
-An embedding of the space $\overleftarrow{U}$   \ednote{M: in which space?
-is it the inver-limit space instead?} would allow one to establish topological conjugacy, which in turn provides stronger results than merely obtaining a coarse-grained representation via a causal mapping \cite{M: I feel this is very confusing to the reader as the co-domain of causal mapping is the left-infinite space and , so please omit if possible.
+\textbf{When considering the map $H$, we once again restrict restrict our focus of $H$ to the function $H_2(\overline{u}):=(h(r\overline{u}, h(\overline{u})))$. }
+
+An embedding of the space $\overleftarrow{U}$  in $\overleftarrow{U}$ would allow one to establish topological conjugacy, which in turn provides stronger results than merely obtaining a coarse-grained representation via a causal mapping \cite{M: I feel this is very confusing to the reader as the co-domain of causal mapping is the left-infinite space and , so please omit if possible.
 You can just say we want to restrict inputs to the left-infinite orbits of a dynamical system and then explore conjugacy instead of a mere semi-conjugacy}.
@@
 \begin{Definition}
-  [\bf Causal Embedding]\label{Dfn_CausalEmbed}
+  [\bf Causal Embedding]\label{Dfn_CausalEmbed}\rm
   A driven system $g$ is said to causally embed the dynamical system $(U,T)$ if 
@@
 \begin{enumerate}[noitemsep, label=\roman*.]
-  \item The diagram in \ref{SCausal_H} \ednote{M: It is actually Eqn. 1.1 and not 1.2}commutes (i.e. $h$ is a universal semi-conjugacy) ,
-  \item $H_2(\overline{u}):=(h(r\overline{u}, h(\overline{u})))$ embeds the inverse-limit space $(\widehat{U}, \hat{T})$ of $(U,T)$ in $X\times{X}$ \ednote{M: Please replace $\widehat{U}$ by $\widehat{U}_T$ here and throughout}.
+  \item The diagram in \ref{Scomm_h} commutes (i.e. $h$ is a universal semi-conjugacy),
+  \item $H_2(\overline{u}):=(h(r\overline{u}, h(\overline{u})))$ embeds the inverse-limit space $(\widehat{U}_T, \hat{T})$ of $(U,T)$ in $X\times{X}$.
 \end{enumerate}
@@
 
-When a causal embedding $H$ exists for the driven system $g$, one can map an arbitrary input ${u}$ onto the solution space $X$ without additional distortion or information-loss.  \ednote{(Cite.)} 
-When an embedding is established, the question of possible additional complexity in the solution is removed by guaranteeing that, since the systems are conjugate (semi-conjugate, (refer)), $g$ does not add any (some) complexity to the system.
+When a causal embedding $H$ exists for the driven system $g$, one can map an arbitrary input ${u}$ onto the solution space $X$ without additional distortion or information-loss \textbf{(Cite.')}.
+When an embedding is established, the question of possible additional complexity in the solution is removed by guaranteeing that, since the systems are conjugate (semi-conjugate, \textbf{(refer)}), $g$ does not add any (some) complexity to the system.
+
 It is, however, a balancing act as it also undesirable to choose a function $g$ that quenches the temporal structure in $u$ by contracting to such a degree that the ability to recover information from the original system is lost completely.
-To obtain a suitably complex function $g$, it is desired that the the reachable set of a driven system  be large enough to relate to the input.  \ednote{(Expand here. It stops a bit abruptly.)}
-
-In the example above (\ref{ex_halfux}), the input's temporal variation cannot be related to the reachable set as $X_U$ consists of a single element and so little, if not no, information is encoded.
-Rather, the reachable set of a driven system must therefore be such that the inverse-limit space of $U_T$ can be embedded in some finite self-product of the reachable set of $T$ \ednote{(refer or cite)}.
+In the above example (\ref{ex_halfux}), the input's temporal variation cannot be related to the reachable set since $X_U$ consists of a single element; little, if not no, information is encoded.
+To obtain a suitably complex function $g$, it is thus desired that the reachable set of a driven system  be large enough to relate to the input. 
+ We  reguire that the reachable set of a driven system to be such that the inverse-limit space of $U_T$ can be embedded in some finite self-product of the reachable set of $T$ \textbf{(refer or cite)}.
 To this end, consider the notion of State-Input (SI) Invertibility.
-
-\begin{Definition}
-  [\bf SI-Invertibility]\label{Dfn_SIinv}
+ \ednote{(M: Expand here. It stops a bit abruptly. B: Does the above reorganisation read better?)}
+
+\begin{Definition}
+  [\bf SI-Invertibility]\label{Dfn_SIinv}\rm
   A map $g$ is said to be SI-Invertible if $g$ is invertible for all $x\in{X}$.
@@
 
-SI-invertibility promises that 'enough' information is retained when $g$ is chosen without introducing additional unwanted complexity. \ednote{Expand why this is true.}
+SI-invertibility promises that 'enough' information is retained when $g$ is chosen without introducing additional unwanted complexity. \ednote{(M: Expand why this is true. B: Is this satisfactory?)} 
+'Enough' here refers to the fact that we may always recover the previous input value $u_n$ (if we know successive states $x_{n-1}, x_n$).
+'Without additional complexity' is guaranteed by the invertibility of the function $g$ that guarantees a one-to-one mapping between $u_n, x_{n-1}$ and $x_n$. 
+In obtaining an SI-invertible function, therefore, we are guaranteed a system which will not lose so much information about its previous states in the forward-flow of time that one may not make any credible claims as to the original system's behaviour and properties.
+Simultaneously, it does guarantee us that the important information encoded in previous inputs is indeed preserved.
 
@@
 \begin{Definition}
-  [\bf Uniform Attraction Property]\label{Dfn_UAP}
+  [\bf Uniform Attraction Property]\label{Dfn_UAP}\rm
   A driven system $g$ has the Uniform Attraction Property (UAP) if, regardless of starting position, all trajectories \textbf{(of what?)} converge to a single trajectory as time flows forward. 
@@
 
-This definition is stated in a not completely rigourous manner as the more formal definition makes use of processes, a concept which would take some time to establish and detracts from the principal thrust of this paper/project/thesis.
-\ednote{Is it VERY obvious to the reader at this point what the general thrust of the paper is?}
-
-As the Unique Attraction Property guarantees that all trajectories \textit{(which ones?)} will converge to the same trajectory as time moves forward. 
-Incredibly, this permits one to initialise a driven system $g$ with an altogether arbitrary initial value $y_m\in{X}$ where $m\in\mathbb{Z}$ and the UAP then guarantees that the sequence $\{y_{m+1}, y_{m+2},\ldots\}$ which satisfies the relation $y_n=g(u_n, y_{n-1})$  for  $k\geq{m}$  will uniformly approach the elements $\{x_n\}$ of the actual solution.
-(see [41, Theorem 1] or [11, Eqn.
-(18)])
+This definition is stated in a not completely rigourous manner as the more formal definition makes use of processes, a concept which would take some time to establish and detracts from the principal thrust of this project.
+\ednote{B: Is it very obvious to the reader at this point what the general thrust of the paper is?}
+
+As the Unique Attraction Property guarantees that all trajectories \ednote{(B: which ones?)} will converge to the same trajectory as time moves forward. 
+Incredibly, this permits one to initialise a driven system $g$ with an altogether arbitrary initial value $y_m\in{X}$ where $m\in\mathbb{Z}$ and the UAP then guarantees that the sequence $\{y_{m+1}, y_{m+2},\ldots\}$ which satisfies the relation $y_n=g(u_n, y_{n-1})$  for  $k\geq{m}$  will uniformly approach the elements $\{x_n\}$ of the actual solution (See \cite[Th.11,]{Manju_Nonlinearity} or \cite[Eqn.18]{Supp}).
 
@@
   \centering
-  \label{fig:memoryLosscont_time}
-\caption{Approaching the true solution if g has the UAP}
+\captionof{figure}{Approaching the true solution if g has the UAP}
+\label{fig:memloss_conttime.png} 
 \end{figure}
@@
 Since we are working with a chaotic system displaying sensitive dependence on initial conditions, these small errors could potentially send the trajectory into a completely different attractor.
-In the presence of the UAP, however, small measurement errors introduced into the system do not pose the same danger as before.
-\ednote{(Rephrase and/or expand)}
+In the presence of the UAP, however, small measurement errors introduced into the system do not pose the same danger as before since the trajectories will eventually tend towards the actual solution.
+\ednote{(B: Complete enough?)}
 
@@
 \begin{Definition}
-  [\bf Inverse-Limit Space]\label{Dfn_InvLimSpace}
+  [\bf Inverse-Limit Space]\label{Dfn_InvLimSpace}\rm
   The inverse-limit space $\widehat{U}_T$, a subspace of $\overleftarrow{U}$, is defined by $$\widehat{U}_T:=\{(\ldots, u_{-2}, u_{-1}):T(u_n)=u_{n+1}\}$$ 
@@
 
-The inverse-limit space is well-defined since $T:U\to{U}$ is surjective by assumption. \ednote{Should I explain this in more details or am  I then over-simplifiying?}
+The inverse-limit space is well-defined since $T:U\to{U}$ is surjective by assumption.
 
@@
 In the above sections, we established the map $G_T$ describing the SDD of a driven system.
-By establishing the existence of this map, we’ve essentially embedded the attractor $U$ \ednote{(So far we haven't really spoken about an attractor)} into the higher dimensional space $X\times{X}$.
-\ednote{(Why is it true that it's higher-dimensional?)}
+By establishing the existence of this map, we’ve essentially embedded the attractor $U$ (recall from \ref{attractor_U}) into the higher dimensional space $X\times{X}$.
 In layman’s terms, this ensures that there is more “dimensional room” for the underlying system’s underlying dynamics to “move”.
-As the dynamics aren’t as “squashed”, we might therefore hope that the dynamics of $G_T$ are in some sense simpler than that of $T$.
+Since the dynamics aren’t as “squashed”, we might therefore hope that the dynamics of $G_T$ are in some sense simpler than that of $T$.
 (Taken note of the fact that $G_T$ is a homeomorphism even when $T$ is just continuous)
@@
 This is done by implementing a Recurrent Neural Network (RNN), but not discussed as of yet.
-\ednote{(Add in that we'll be using Pearson coefficient?)}
+\ednote{(B: Add in that we'll be using Pearson coefficient?)}
 
@@
 \begin{proof}
-  From [Koopman, Theorem 3c] 
+  The proof may be found in~\cite[Th. 3c]{manjunath2021universal}.
 \end{proof}
 
-Projection mappings $\pi_i$ are defined in the traditional meaning where a $k$-dimensionalvector is projected to it's $i^{th}$ component such that $$\pi_i:(a_1,a_2, \ldotp, a_{k-1}, a_k)\to{a_i}$$
+Projection mappings $\pi_i$ are defined in the traditional meaning where a $k$-dimensional vector is projected to it's $i^{th}$ component such that $$\pi_i:(a_1,a_2, \ldotp, a_{k-1}, a_k)\to{a_i}$$
 
 % \begin{Definition}
-%   [\bf Projection Mapping]\label{Dfn_ProjMap}
+%   [\bf Projection Mapping]\label{Dfn_ProjMap}\rm
 %  The projection mapping $\pi_i$ is defined as follows: $$\pi_1:(a_1,a_2)\to{a_1}$$ and $$\pi_2:(a_1,a_2)\to{a_2}$$ 
@@
 \begin{figure}[ht] \label{Fig_Gamma}
-  % \includegraphics[scale=0.3]{gammatemp.png}
+  \includegraphics[scale=0.3]{gammatemp.png}
   \centering
@@
 
-Thirdly, $\Gamma$ is globally dissipative whereas $G_T$ is not guaranteed to be(\emph{Cite}). \ednote{Add linking sentence.} Global dissipativity prevents large numerical errors due to input data.
+Thirdly, $\Gamma$ is globally dissipative whereas $G_T$ is not guaranteed to be(\emph{Cite}). \ednote{B: Add linking sentence.} Global dissipativity prevents large numerical errors due to input data.
 
@@
 \begin{Definition}
-  [\bf Global Dissipativity]\label{Dfn_GlobalDiss}
+  [\bf Global Dissipativity]\label{Dfn_GlobalDiss}\rm
 	We say a dynamical system $(Z,f)$ is globally dissipative if there exists a nonempty proper closed subset $B$ of $Z$  so that for all $x\in Z\setminus B$, (i).
@@
 
-\chapter{Implementation} \label{sect5}
+% \chapter{Implementation} \label{sect5}
 

### 2022-01-08 Gandhi_Project_Report/draft3.tex
making a mess with all my comments

@@
-% \documentclass[12 pt]{article}
-\documentclass[a4paper,12pt,twoside]{book}
+% \documentclass[a4paper,12pt,twoside]{book}
+\documentclass[a4paper,12pt,twoside]{report}
 \pagestyle{headings}
@@
 \maketitle
-
 \tableofcontents
@@
   \begin{figure}[ht]
-    \includegraphics[scale=0.45]{log_sdic.eps}
+    \includegraphics[scale=0.45]{_log_sdic.eps}
         \centering
@@
   \centering
-\includegraphics[scale=0.8]{dp_sdic.eps}
+\includegraphics[scale=0.8]{_dp_sdic.eps}
     \captionof{figure}{Three double pendulum heads with initial angle differing by 0,025 radians}\label{fig:dp_sdic}
@@
   \begin{figure}[ht]
-    \includegraphics[scale=0.6]{T2.png}
+    \includegraphics[scale=0.6]{_tentmap_2.eps}
         \centering
@@
 \begin{figure}[ht]
-  \includegraphics[scale=0.25]{takensmap.png}
+  \includegraphics[scale=0.25]{_takensmap.png}
   \centering
@@
 \begin{figure}[ht]
-  \includegraphics[scale=0.3]{learningfailure.png}
+  \includegraphics[scale=0.3]{_learningfailure.png}
   \centering
@@
 
-\begin{figure}[h]
-  \includegraphics[scale=0.4]{temporary_damp_fig.png}
+\begin{figure}[ht]
+  \includegraphics[scale=0.4]{_dp_dying.eps}
   \centering
@@
 
+In most of the reservoir computing (RC) literature, the idea of memory-loss or fading memory popularly called
+the echo state property (ESP) is treated as a stability property of the system valid
+for all inputs. In such a case, the ESP guarantees that the entire past history of the
+input determines the state of the system precisely, i.e., there is no possibility of two
+or more reachable states at any given point in time [5] if the entire past values of the
+input have influenced the dynamics. The formal definition linking ESP intrinsically
+to an input is available in [4].
+
+In the context of information processing, when a driven system g has a causal em-
+bedding, it helps carry over any input ū onto the solution space without distortion or
+loss of information. This is a very important feature, and the current discrete-time
+state space models used in reservoir computing do not exploit the causal embedding.
+We intend to demonstrate the utility of these theoretical results in a future work
+directed at applications.
+
+For natural systems, say, for instance, when the human brain is viewed as a driven
+system where a time-varying external stimulus (for instance, a visual stimulus) is
+an input, the input can be causally embedded into the nervous system without the
+need for a particular structure that is custom-made for each stimulus as long as
+the system has the USP. From the perspective of perturbation of an autonomous
+dynamical system, if the resulting nonautonomous system has the USP, then the
+temporal structure of noise relates to the dynamics on a nonautonomous uniform
+attractor through a topological semi-conjugacy or a conjugacy.
+
+The continuity of E reflects on whether small changes in u ~ would result in a ‘propor-
+tionate’ response in the solution space – a notion of input related stability as defined in \cite{Manju_ESP}
+
+To understand
+how solutions become sensitive to a perturbation of the input for systems without
+USP, we define a stability notion that describes how the collection of all solutions
+obtained from an input ū are mapped onto the solution space.
+
+Sufficient to consider continuity of encoding map.
+
+owing to the USP of the driven system $g$ one gains input-related, parameter-related stability \cite{manjunath2020stability}
+
+------
 xxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxx
@@
 \begin{equation}  \label{eqn_driving}
-g(u,x)=(1-a)x+a{\cdot}tanh(Au+\alpha{a}Bx)
+  g(u,x) = (1-a)x + a\overline{\tanh}(Au + \alpha Bx)
 \end{equation} 
-is both SI-invertible and possesses the USP. The proof of these two facts is rather involved and  hence the proof, instead of being replicated here, may be found in \textbf{(cite relevant article)}
+is both SI-invertible and possesses the USP. 
+One may easily show that $g$ is SI-invertible by recovering $u_n$ by 
+\begin{equation} \label{eqn_SI_RNN}
+  u_n := A^{-1}\bigg(\overline{\tanh}^{^{-1}}\frac{1}{a}\Big(x_{n+1}-(1-a)x_n\Big) \bigg) - \alpha B x_n
+  \end{equation}
+  for $x_{n-1}, x_n$ known.
+
+The proof of these $g$ possessing the USP is rather involved and hence the proof, instead of being replicated here, may be found in \textbf{(cite relevant article)}
 This specific driving function $g$ is used in our implementation and is more completely discussed in chapter \ref{sect5}
@@
 \ednote{(B: Complete enough?)}
+\textbf{From the perspective of perturbation of an autonomous
+dynamical system, if the resulting nonautonomous system has the USP, then the
+temporal structure of noise relates to the dynamics on a nonautonomous uniform
+attractor through a topological semi-conjugacy or a conjugacy.}
 
@@
 We now have the following (compare with \ref{SCausal_H} above):
+
+\textbf{NB NB NB - Causal Embedding Theorem! - Th.4}
 
@@
 \begin{figure}[ht] \label{Fig_Gamma}
-  \includegraphics[scale=0.3]{gammatemp.png}
+  \includegraphics[scale=0.3]{_gammatemp.png}
   \centering
@@
 A final set of equations is obtained - equations that have been entirely constructed from data.
-\begin{eqnarray}
+\begin{eqnarray}\label{eqns_from_data}
 	u_{k+1} &=& \pi_1 \circ (\Gamma, \pi_2) \circ (\pi_2,g) (u_k,x_k) \label{Seqn_u}\\
@@
 
-% \chapter{Implementation} \label{sect5}
-
-\vspace{-1cm}
-\bibliographystyle{pnas-new}
-\bibliography{pnas}
+\chapter{Implementation} \label{ch5}
+
+In this section we turn to the implementation of the function $G_T$ by discussing a method to learn the function $\Gamma$.
+We then indicate where this implementation differs from previous work done by (\cite{manjunath2021universal}) and show that our results do hold water by considering the Pearson correlation coefficient.
+\ednote{(this sentence needs to be changed) Are you talking of MDD?
+Or the sparse matrix?.}
+Finally, we present and discuss some numerical results for a focus system of this project: the double pendulum. We also consider other chaotic attractors.
+
+\section{Implementing $G_T$}
+The map $G_T$ is implemented by applying the functions $\Gamma$, $g$ and the projection function $\pi_2$ in the relevant order. To reach this point, one must first learn $\Gamma$.
+After 1.normalising the data to ensure it lies within the interval $|x| < 1$ and has zero mean value, we 2.drive the system with external input through the function $g$.
+
+From before (\eqref{eqn_driving}): $ g(u,x)=(1-a)x+a{\cdot}\tanh(Au+\alpha{a}Bx)$. Remember that $g$ is SI-invertible. $A$ is the so-called “input matrix” and $B$ the “reservoir matrix”.
+
+We make use of a Recurrent Neural Network (RNN) in our implementation.
+RNN's are widely used in the literature for discrete-time processing and are considered to be discrete-time version of continuous dynamical systems modelling a network of neurons.
+(\textbf{Cite.})
+A complete discussion of the RNN is outside the scope of this thesis, so we only present the details necessary to the current discussion. For information, (\textbf{Cite, Cite, Cite}) may be consulted.
+
+If the RNN we will be using in the next step has $N$ neurons, then $A$ and $B$ are each $N\times{N}$ matrices.
+Inputs $u_n$ of dimension $K < N$ are padded by zeroes to obtain an input of dimension $N$.
+
+No need for entire left-infinite history of input: An arbitrary initialisation of the system with initial values and the driving of the system with the function $g$ means that the generated values will come to approximate the actual solution’s elements in a uniform manner thanks to the UAP.
+
+We then 3.set up/prepare the RNN.
+Utilising a network $N$ neurons (the same $N$ as in the dimensions of the matrices $A,B$).
+Training is accomplished using the Adam Optimiser where the MSE loss function is optimised.
+The system is trained several times(in our case 4) with incrementally smaller learning rates on each run (0.001 and thereafter divided by 10 at each run).
+\ednote{Please present the number of layers in the network, number of neurons in each layer and the length of the training validation and testing for learning Gamma.}
+
+We then reach the prediction step. Making use of $g$'s SI-invertibility and $g$ as defined in \eqref{eqn_driving}, we have the following expression for $u_n$ replicated from~\eqref{eqn_SI_RNN}.
+\begin{equation*}
+  u_n := A^{-1}\bigg(\overline{\tanh}^{^{-1}}\frac{1}{a}\Big(x_{n+1}-(1-a)x_n\Big) \bigg) - \alpha B x_n
+\end{equation*}
+
+The two model equations for learning from data (\eqref{eqns_from_data}) also allows us to determine what the state $x_{n+1}$ at time $n+1$ will be. 
+We specifically remind ourselves of \[x_{k+1}=\pi_2 \circ (\Gamma, \pi_2) \circ (\pi_2,g) (u_k,x_k)\]
+
+\section{Changes in this thesis/project/report to previous work of~\cite{manjunath2021universal}}
+
+We made several changes which gave improved results:
+\vspace{-8mm}
+\begin{enumerate}
+\item Using Delay Coordinates (DC) when driving the system vs.
+using Single/Multi Delay Dynamics (SDD/MDD) after driving (and before training).
+Explain/show why these two methods are equivalent.
+\ednote{Would MDD  be not computationally intensive ?
+So may be you just explain if you can’t show it.}
+\item Sparse matrix implementation. Slight improvement in some scenarios, but not fantastic. \ednote{Did it not help you a great deal with the DP forecasting?}
+
+(Possible paragraph here/later on implementing sparsity of A/B: Generate sparse matrix that is also well-conditioned and correctly scaled matrix.)
+\item Density of matrices A/B (separate from sparse matrix implementation) \ednote{Is it the probability distributions of the entries in A and B?}
+\end{enumerate}
+
+\section{Numerical simulations}
+\subsection{Double Pendulum}
+\begin{itemize}
+  \item What it is. 
+  \item Why it's chaotic
+  \item How data was generated for this. (noiseless)
+  \item Why data dying down doesn't work (ie. Not surjective)
+  \item Other possibilities
+\end{itemize}
+
+The Double Pendulum consists of two pendulums fastened to one another such that the system moves as a whole.
+The first (or higher pendulum) is attached to a fixed point while the second is attached to the endpoint of the first.
+(See below)
+
+\begin{figure}[ht] \label{Fig_Gamma}
+  \includegraphics[scale=0.25]{_dbl_Pend.png}
+  \centering
+  \label{fig:conjugacy_with_Gamma}
+\end{figure}
+
+An ideal system is assumed where the rods are massless and no friction in the pivots are considered.
+One is easily reminded of the Lorenz system, another three-equation deterministic model which exhibits chaotic behaviour.
+The scalar equations of motion are presented below and were derived with the use of Newtonian physics and can be found in \textbf{(Cite.)}
+
+\begin{eqnarray}
+\theta_{1}^{''} = \frac{-g(2m_1+m_2)\sin{\theta_1} - m_{2}g\sin(\theta_1 - 2\theta_2) -2\sin(\theta_1-\theta_2)m_2({\theta_{2}'}^{2}L_2 + \theta_{1}^{2}L_{1}\cos(\theta_1 - \theta_2))}{L_1(2m_1+m_2-m_2\cos({2\theta_1 -2\theta_2}))}
+\\
+\theta_{2}^{''} = \frac{ 2\sin(\theta_1-\theta_2)(\theta_{1}'^2 L_1 (m_1+m_2) + g(m_1+m_2)\cos{\theta_1} + \theta_{2}'^2 L_2 m_2 \cos(\theta_1 - \theta_2))}{L_2(2m_1+m_2-m_2\cos({2\theta_1 -2\theta_2}))}
+\end{eqnarray}
+
+This second-order system was then converted into a first-order system by defining the first derivatives as separate values.`'
+
+\subsection{Other attractors I’ve experimented with: Clifford \& Thomas}
+We follow the methodology presented in \cite{manjunath2021universal} but opt to consider different attractors this time.
+
+\begin{itemize}
+\item Thomas:
+\begin{eqnarray}\label{eqns_thomas}
+  u_x = \sin(y) - \beta{x} \\
+  u_y = \sin(z) - \beta{y} \\
+  u_z = \sin(x) - \beta{z}
+\end{eqnarray}
+
+\item Clifford:
+\begin{equation}\label{eqn_clifford}
+  \frac{\sin(ay) + c\cos(ax)}{\sin(bx)+d\cos(by)}
+\end{equation}
+\end{itemize}
+
+\section{Generalized Pearson correlation coefficient for the DP and other attractors}
+\textbf{Insert table: D.Pendulum, Lorenz, Others (Clifford \& Thomas)}
+
+\chapter{Conclusion}\label{ch6}
+
+The driven states $x_{n-1},x_n)$ holds the information of the left-infinite history of the input.
+Furthermore, there is some evidence that the inverse-limit space (not system) seems to distinguish two dynamical systems that are not topologically conjugate for a class of systems in the sense their inverse-limit spaces are not homeomorphic to each other.
+So the relation $Y_T$, which is homeomorphic to the inverse-limit system, may be unique to a class of systems.
+(The reference is a paper on Ingram’s Conjecture by Marcy Barge, Henk Bruin and Sonja Stimac.
+) Further investigation is needed.
 

### 2022-01-08 graphing.py

@@ -19,8 +19,10 @@ import matplotlib.pyplot as plt
 # plt.savefig("T2.png")
 # plt.show()
 
-plt.plot(X1, Tn(54, X1))  #52=close, 53=line
+plt.plot(X1, Tn(2, X1), 'forestgreen')  #52=close, 53=line
+plt.savefig('tentmap_2.svg', format='svg', dpi=1000)
 plt.show()
+
 
 # plt.plot(X3, T(X3), '-k')
 # plt.show()

### 2022-01-10 Gandhi_Project_Report/draft3.tex
Done with S 4.6

@@
 \usepackage{enumitem}
+\usepackage{multirow}
 
@@
 \begin{Definition}
-  [\bf Invariant Set] \label{Dfn_Flow}
+  [\bf Invariant Set] \label{Dfn_InvariantSet}
   Given a discrete-time dynamical system $T: U \to U$, a subset $A \subset U$ is said to be an \emph{invariant set} if $T(A) =A$. 
@@
 
-We also define the \emph{orbit of $T$} to be a sequence $\bar{u} = \{u_n\}_{n\in \mathbb{Z}}$ obeying the update equation, $u_{n+1}=Tu_n$, $n \in \mathbb{Z}$.
+We also define the orbit of a function $T$.
+\begin{Definition}
+  [\bf Orbit of $T$] \label{Dfn_Orbit}
+  The orbit of $T$ is to be defined the sequence $\bar{u} = \{u_n\}_{n\in \mathbb{Z}}$ obeying the update equation, $u_{n+1}=Tu_n$, $n \in \mathbb{Z}$. 
+\end{Definition}
 
@@
 There are several intricately different definitions of chaos available in the literature \textbf{(Cite)}, with each of them indicating some aspect of complexity.
-\ednote%{B:  what are we saying here?
+%\ednote{B:  what are we saying here?
 M: Devaney's definition is not the only one, there are about half a dozen definitions as well}
@@
 \begin{Definition}
-  [\bf Nonautonomous Dynamical System (adopted from~\cite{Manju_ESP}))] \label{Dfn_Soln}\rm
+  [\bf Nonautonomous Dynamical System (adopted from~\cite{Manju_ESP}))] \label{Dfn_NDS}\rm
   A nonautonomous dynamical system (NDS) on a space $X$ is simply a dynamical system comprising a family of maps $\{g_n\}_{n \in \mathbb{Z}}$ where $g_n(\ldotp):=g(u_n,\ldotp):X\to{X}$ is continuous.
 Each $g_n$ arises as a consequence of an input $u_n$ from the input space $U$, a topological space.
-\ednote{Does the reader immediately remember/know what is meant by a dynamical system?}
 \end{Definition}
@@
  The proof is deferred to immediately after the next paragraph (see~\ref{rem_proofEx}).
-
 \end{Example}
@@
   [\bf Reachable Set]\label{Dfn_ReachableSet}\rm
-The reachable set of a driven system $g$ is exactly the union of all the elements of all the solutions, i.e., 
+The reachable set of a driven system $g$ is exactly the union of all the elements of all the (entire) solutions, i.e., 
 \[X_U :=\Big \{x \in X:  x = x_k \mbox{ where  $\{x_n\}$  is a solution for some  $\bar{u}$} \Big \}.\]
@@
 
-We note that $x\in{X_n(\overline{u})}$ if and only $g$ has a solution $\{x_k\}$ for $x_n=x$ and input $\overline{u}$. \textbf{(Cite)}.
+The reachable set itself can defined independently from the fact that $g$ does or does not have the ESP/USP.
+We note that $x\in{X_n(\overline{u})}$ if and only $g$ has a solution $\{x_k\}$ for $x_n=x$ and input $\overline{u}$.
+\textbf{(Cite)}.
 
@@
 
-This will lead to a result established later, but which is worth taking note of now: \textit{$g$ being a topological contraction is equivalent to the existence of a unique entire solution.} To this end we define the notion of a topological contraction.
-\ednote{M: I do not think you need the definition of a topological contraction unless it is defined through a process.
-Please note that you have defined $X_n(\bar{u})$ differently than what was in the earlier works.
-Now it is just the collection of solution values at time $n$ with the input $\bar{u}$ and topological contraction is just paraphrasing the USP definition.
-B: Is it better to then just leave out the entire sentence?}
-
-\begin{Definition}
-  [\bf Topological Contraction]\label{Dfn_TopContr}\rm
-  A function $g:U\times{X}\to{X}$ is a topological contraction if for all $n\in\mathbb{Z}$ and all $\overline{u}\subset{U}$, $X_n(\overline{u})$ is a singleton subset of $X$.  
-\end{Definition}
+This will lead to a result established later, but which is worth taking note of now: \textit{$g$ being a topological contraction is equivalent to the existence of a unique entire solution.}
+
+% \begin{Definition}
+%   [\bf Topological Contraction]\label{Dfn_TopContr}\rm
+%   A function $g:U\times{X}\to{X}$ is a topological contraction if for all $n\in\mathbb{Z}$ and all $\overline{u}\subset{U}$, $X_n(\overline{u})$ is a singleton subset of $X$.  
+% \end{Definition}
 
@@
   We then easily conclude that $\{x_n\}\equiv0$ is the only solution. 
-\end{Remark}
+\end{Remark} \ednote{B: Is this sufficient?}
 
@@
 \begin{proof}
-  \textbf{I've muddled myself here.}
-  % Consider an input $\overline{u}={\{u_n\}}_{n\in\mathbb{Z}}\subset{U}$ and driven system $g:U\times{X}\to{X}$  generating a sequence ${\{x_n\}}_{n\in\mathbb{Z}}$ in the compact space $X$ \ednote{This statement is not correct.}.
-  % Since $X$ is a metric space, it follows immediately from a well-known result in Analysis (\textbf{Cite}) that $\{x_n\}$ has a convergent subsequence which then is a solution.
-\ednote{This is not a valid proof.
-Please refer to the proof existence of a solution in \cite{Manju_ESP}.}
+  May be found in \cite{kloeden2011nonautonomous, manjunath2014dynamics, manjunath2013echo}
 \end{proof}
@@
 
-To understand
-how solutions become sensitive to a perturbation of the input for systems without
-USP, we define a stability notion that describes how the collection of all solutions
-obtained from an input ū are mapped onto the solution space.
+To understand how solutions become sensitive to a perturbation of the input for systems without USP, we define a stability notion that describes how the collection of all solutions obtained from an input ū are mapped onto the solution space.
 
@@
 
-------
+Historically, data have been transformed into a different space for processing to gain some advantages.
+For example, according to Vapnik–Chervonenkis theory (e.g., [16]) mapping data into a higher dimension often provides much greater classification power.
+Mapping temporal data through a delay-coordinate mapping in Takens delay embedding uncovers hidden information [17]. In data-driven modeling employing Koopman’s theory (e.g., [18]–[20]), 
+approximating the dynamics of a dynamical system through a linear dynamicalsystem in a higher dimensional space is often possible by mapping the data samplewise into the higher dimensional space
+
+we first note that when T is not invertible, the dynamics of (U, T ) is only topologically semi-conjugate to $(\hat{U}, \hat{T})$ and hence the dynamics in $(\hat{U}, \hat{T})$ could have additional complexity.
+
 xxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxx
@@
 \begin{proof}
-  \emph{See proof from p.15 of \cite{manjunath2021universal}}
-  \textbf{Expand.}
+  See \cite[Th.3]{manjunath2013echo}.
 \end{proof}
 
-\textbf{When considering the map $H$, we once again restrict restrict our focus of $H$ to the function $H_2(\overline{u}):=(h(r\overline{u}, h(\overline{u})))$. }
+\textbf{When considering the map $H$, we once again restrict our focus of $H$ to the function $H_2(\overline{u}):=(h(r\overline{u}, h(\overline{u})))$. }
 
@@
 Before formalising this, we make mention of the concept of the inverse-limit system of a dynamical system $(U,T)$. 
-  In broad sweeps, the inverse-limit system of a dynamical system is a self-map on a subset of an infinite-dimensional space \textbf{(Expand)} where each point in the inverse-limit space corresponds to a backward orbit of the map $T$.
-Following \cite{manjunath2021universal}, the inverse-limit space is denoted by $\widehat{U}_T\subset \overleftarrow{U}$, and defined as $\widehat{U}_T\subset: = \{ (\ldots,u_{-2},u_{-1}): Tu_{i} = u_{i+1}$.
-  The map $T$, then, induces a self-map $\widehat{T}: (\ldots,u_{-2},u_{-1}) \mapsto  (\ldots,u_{-2},u_{-1},T(u_{-1}))$ on $\widehat{U}_T$.
+These spaces are noteworthy as their topology can help one ascertain much about the original system \cite{Manju_IEEE}.
+The topological structure for inverse-limit systems of chaotic dynamical systems is complicated, so we relegate further discussion to \cite{kennedy2008inverse_limit, ingram2011inverse}.
+In broad sweeps, the inverse-limit space of a dynamical system is a self-map on a subset of an infinite-dimensional space where each point in the inverse-limit space corresponds to a backward orbit of the map $T$.
+Following \cite{manjunath2021universal, ingram2011inverse}:
+
+\begin{Definition}
+  [\bf Inverse-Limit Space, Inverse-Limit System]\label{Dfn_InverseL}\rm
+ The subset $\widehat{U}_T\subset \overleftarrow{U}$ defined by \[\widehat{U}_T: = \{ (\ldots,u_{-2},u_{-1}): Tu_{i} = u_{i+1}\]is called the inverse-limit space of $(U,T)$.
+ The self-map $\widehat{T}$  induced by $T$ on $\widehat{U}_T$  is defined by \[\widehat{T}: (\ldots,u_{-2},u_{-1}) \mapsto  (\ldots,u_{-2},u_{-1},T(u_{-1}))\] and the resulting dynamical system $(\widehat{U_T}, \widehat{T})$ is called the inverse-limit system of $(U,T)$.
+\end{Definition}
 
@@
 \begin{proof}
-  \emph{See proof from \cite{manjunath2021universal}}  \textbf{Expand.}
+  See \cite{manjunath2021universal}.
 \end{proof}
@@
 
-The next result is of relative import but to fully comprehend its influence, some discussion still remains. We state it in part and postpone the proof.
+The next result is of relative import but to fully comprehend its influence, some discussion still remains. We state it in part here.
 % Before stating it, we define a concept which will directly be shown to be equivalent to the USP.
@@
 \end{Theorem}
-\begin{proof}
-  (postponed) 
-\end{proof}
+% \begin{proof}
+%   (postponed) 
+% \end{proof}
 
 % We have not as of yet treated the UAP properly, and establish the necessary result here (without proof) so as not to break up the completeness of theorem.
-
-% \textbf{The proof of the result for the UAP is left to literature. Why? Cite.}
 
@@
 
-It is worth taking note of a specific driven system. The function 
+It is worth at this stage taking note of a specific driven system in the form of a discrete state-space model that has acquired some adherence in applications \cite{Manju_IEEE} and especially those pertaining to Echo State Networks and ESP.
+The function
+
 \begin{equation}  \label{eqn_driving}
@@
 \end{equation} 
-is both SI-invertible and possesses the USP. 
+is both SI-invertible and, if $\alpha\beta<1$, also possesses the USP. 
 One may easily show that $g$ is SI-invertible by recovering $u_n$ by 
@@
 
-The proof of these $g$ possessing the USP is rather involved and hence the proof, instead of being replicated here, may be found in \textbf{(cite relevant article)}
-This specific driving function $g$ is used in our implementation and is more completely discussed in chapter \ref{sect5}
+The proof that the map $g$ possesses the USP is rather involved and hence, instead of being replicated here, may be found in \cite[Th.2]{manjunath2013echo }
+This specific driving function $g$ is used in our implementation and is discussed more completely in chapter \ref{sect5}
 
@@
 
-As the Unique Attraction Property guarantees that all trajectories \ednote{(B: which ones?)} will converge to the same trajectory as time moves forward. 
+As the Unique Attraction Property guarantees that all trajectories will converge to the same trajectory as time moves forward. 
 Incredibly, this permits one to initialise a driven system $g$ with an altogether arbitrary initial value $y_m\in{X}$ where $m\in\mathbb{Z}$ and the UAP then guarantees that the sequence $\{y_{m+1}, y_{m+2},\ldots\}$ which satisfies the relation $y_n=g(u_n, y_{n-1})$  for  $k\geq{m}$  will uniformly approach the elements $\{x_n\}$ of the actual solution (See \cite[Th.11,]{Manju_Nonlinearity} or \cite[Eqn.18]{Supp}).
@@
 
-\begin{Theorem}
+\begin{Theorem}\label{thm_usp-contr-uap}
   The following statements are equivalent:
@@
 \begin{proof}
-  (We'll show this)  [11] eqn.6 
+  May be found in \cite[Th.1]{Manju_Nonlinearity}
 \end{proof}
@@
 \ednote{(B: Complete enough?)}
-\textbf{From the perspective of perturbation of an autonomous
-dynamical system, if the resulting nonautonomous system has the USP, then the
-temporal structure of noise relates to the dynamics on a nonautonomous uniform
-attractor through a topological semi-conjugacy or a conjugacy.}
+% \textbf{From the perspective of perturbation of an autonomous dynamical system, if the resulting nonautonomous system has the USP, then the temporal structure of noise relates to the dynamics on a nonautonomous uniform attractor through a topological semi-conjugacy or a conjugacy.}
 
@@
 \begin{proof}
-  \textbf{Show}
+  See \cite[Th.3, Th.4]{manunath2021universal}
 \end{proof}
@@
 
+% Below is reproduced a pictorial representation of the relationship between different entities and properties. \textbf{(Cite SIG_25May)}
+% Below is reproduced a pictorial representation of the relationship between the different entities and their properties.
+
+\begin{figure}[ht]
+  \includegraphics[scale=0.3]{_summarypictorial.png}
+  \centering
+  \label{fig:fig_pictorialSummary}
+\end{figure}
+
 \section{A discussion of $G_T$ }
@@
 In \cite{manjunath2021universal} it is illustrated in an empirical fashion in that the map $G_T$ describes dynamics which are less functionally complex than that of $T$ or of the map $\Phi_{2d,\theta}$.
-This is done by implementing a Recurrent Neural Network (RNN), but not discussed as of yet.
-\ednote{(B: Add in that we'll be using Pearson coefficient?)}
+This is done by implementing a Recurrent Neural Network (RNN), which is discussed in~\ref{ch5}.
+\ednote{(B: Should I add in that we'll be using Pearson coefficient?)}
 
@@
 \begin{proof}
-  The proof may be found in~\cite[Th. 3c]{manjunath2021universal}.
+  See \cite[Th. 3c]{manjunath2021universal}.
 \end{proof}
@@
 
-\begin{figure}[ht] \label{Fig_Gamma}
+\begin{figure}[ht]
   \includegraphics[scale=0.3]{_gammatemp.png}
   \centering
-  \label{fig:conjugacy_with_Gamma}
+  \label{fig:temp_Gamma}
 \end{figure}
@@
 
-Learning $\Gamma$ saves computational resources.
-This is due to the fact that the input $u_n$ lies in a lower-dimensional subspace of the high-dimensional space $X\times{X}$.
+In the first place, Learning $\Gamma$ saves computational resources.
+This is due to the fact that the input $u_n$ lies in a lower-dimensional subspace of the high-dimensional space $X\times{X}$ and we can learn $\Gamma$ on a small subset of the space.
 In practice, if the input is of a  lower dimension, one may easily embed it into the space $X\times{X}$ by padding the vector $u_n$ with zeroes.
 
-\textbf{Can we discuss how to put this across well? I would like to add it in as I think it beneficial to the discussion.}
-
-Moreover, two types of stability are achieved when $g$ has the USP.
-
-Input- and parameter-related stability is achieved if $g$ has the USP
-\newline \emph{Define/Explain the two types of stabilities}
-\newline \emph{Sentence on $u$, $v$, tails, USP and product topology}
-\newline If we opt to work with $Y_T$, errors can occur. Expand
-\newline $\Gamma$ is more stable. Explain why.
-
-Thirdly, $\Gamma$ is globally dissipative whereas $G_T$ is not guaranteed to be(\emph{Cite}). \ednote{B: Add linking sentence.} Global dissipativity prevents large numerical errors due to input data.
-
-The $\omega$-limit set $\omega(u;T)$ of a point $u$ is defined to be the collection of limit points of the sequence $\{x,Tu,T^2u,\ldots\}$.
+Secondly, the function $\Gamma$ is known to be stable.
+Learning Gamma makes use of equations \ref{eqns_from_data} which in turn employs $g$, a function possessing the USP, and offers distinct advantages with regards to stability in the presence of perturbation.
+It is desirable that inputs that differ only slightly would have only ‘slight’ effects on the output of the system.
+Moreover, one would hope to prevent numerical errors originating from input and measurement noise (see~\cite[Th.
+5]{manjunath2021universal}).
+As the data fed into the system is a function of both the input given and the parameters of the system itself (i.e.
+the value a, $\alpha$ in $\tanh$), we investigate both input-related and parameter-related stability.
+Both these notions are defined by way of the continuity of an encoding map of an input and design parameter respectively in \cite{manjunath2020stability}.\
+
+Informally, we may consider input-related stability as concerned with the question of whether or not small variations in input would result in small responses.   
+If two inputs $u$ and $u_noisy$ are close in the product topology, then their tails could differ greatly due to the metric's insensitivity to the differences in sequence's tails.
+ However if the system has the USP then the function $h$ is continuous and so $h(u), h(u_noisy)$ remain near – maintaining a measure of robustness to input and measurement noise.\cite{manjunath2021universal}.
+In a similar fashion, parameter-related stability is obtained thanks to a result \cite[Lemma 3.2]{manjunath2020stability} relating this form of stability with the ESP (which is here equivalent to the USP).
+
+% Secondly,  due to $g$ possessing the USP, both input- and parameter-related stability~\cite{manjunath2020stability} is obtained, which in turn prevents numerical errors that caused by input and measurement noise (see~\cite[Th.
+5]{manjunath2021universal}).
+
+\begin{figure}[ht]
+  \includegraphics[scale=0.4]{_AutovsDriven.eps}
+  \centering
+  \label{fig:learningFailure}
+  \captionof{figure}{Illustrating empirically the importance of global dissipativity.
+Plotted below the three principal components of randomly initiated trajectories of a learnt version of $G_T$ versus the trajectory $(x_{n-1}, x_n)$(grey) of driven stated obtained from the Thomas Attractor's data (see below).
+}
+\end{figure}
+
+\ednote{B: Would you please help me with the writing here? Otherwise I shall leave this out.}
+\emph{Thirdly, $\Gamma$ is globally dissipative whereas $G_T$ is not guaranteed to be~\cite{manjunath2021universal}). 
+Global dissipativity prevents large numerical errors due to input data.If a system is not globally dissipative, neglibly small errors result in major errors which thus induce predictions to fail utterly.
+If one learns $G_T$ directly, we are in danger of replicating the problems arising for Takens.
+}
+
+% The $\omega$-limit set $\omega(u;T)$ of a point $u$ is defined to be the collection of limit points of the sequence $\{x,Tu,T^2u,\ldots\}$.
 This set is nonempty, and $\omega(u;T)$ is invariant.
 \textbf{(Change)}
-
-\begin{Definition}
-  [\bf Global Dissipativity]\label{Dfn_GlobalDiss}\rm
-	We say a dynamical system $(Z,f)$ is globally dissipative if there exists a nonempty proper closed subset $B$ of $Z$  so that for all $x\in Z\setminus B$, (i).
+% \begin{Definition}
+%   [\bf Global Dissipativity]\label{Dfn_GlobalDiss}\rm
+% 	We say a dynamical system $(Z,f)$ is globally dissipative if there exists a nonempty proper closed subset $B$ of $Z$  so that for all $x\in Z\setminus B$, (i).
 $\omega(x;f) \subset B$  and (ii).
@@
 We call any such closed subset $B$ a trapping set of $f$.
-\end{Definition}
-
-If a system is not globally dissipative, neglibly small errors result in major errors which thus induce predictions to fail utterly.
-If one learns $G_T$ directly, we are in danger of replicating the problems arising for Takens.
+% \end{Definition}
 
@@
 In this section we turn to the implementation of the function $G_T$ by discussing a method to learn the function $\Gamma$.
-We then indicate where this implementation differs from previous work done by (\cite{manjunath2021universal}) and show that our results do hold water by considering the Pearson correlation coefficient.
-\ednote{(this sentence needs to be changed) Are you talking of MDD?
-Or the sparse matrix?.}
+We then indicate where this implementation differs from previous work done by (\cite{manjunath2021universal}) and empirically show that the SDD of $G_T$ are less functionally complex than $\Phi_{k, \theta}$ or $T$ by considering the Pearson correlation coefficient.
 Finally, we present and discuss some numerical results for a focus system of this project: the double pendulum. We also consider other chaotic attractors.
@@
 \section{Implementing $G_T$}
-The map $G_T$ is implemented by applying the functions $\Gamma$, $g$ and the projection function $\pi_2$ in the relevant order. To reach this point, one must first learn $\Gamma$.
-After 1.normalising the data to ensure it lies within the interval $|x| < 1$ and has zero mean value, we 2.drive the system with external input through the function $g$.
-
-From before (\eqref{eqn_driving}): $ g(u,x)=(1-a)x+a{\cdot}\tanh(Au+\alpha{a}Bx)$. Remember that $g$ is SI-invertible. $A$ is the so-called “input matrix” and $B$ the “reservoir matrix”.
+The map $G_T$ is implemented by applying the functions $\Gamma$, $g$ and the projection function $\pi_2$ in the relevant order as described by the equations~\ref{eqns_from_data}.
+To reach this point, one must first learn $\Gamma$.
+After first normalising the data to ensure it lies within the interval $|x| < 1$ and has zero mean value, we continue by driving the system with external input through the function $g$.
+
+From \eqref{eqn_driving} we know that $ g(u,x)=(1-a)x+a{\cdot}\tanh(Au+\alpha{a}Bx)$.
+Recall that $g$ is SI-invertible.
+%Adopting the terminology of \cite{manjunath2021universal}, $A$ is called the “input matrix” and $B$ the “reservoir matrix”.
 
@@
 (\textbf{Cite.})
-A complete discussion of the RNN is outside the scope of this thesis, so we only present the details necessary to the current discussion. For information, (\textbf{Cite, Cite, Cite}) may be consulted.
+A complete discussion of RNN's is outside the scope of this thesis, so we present only the details necessary to the current discussion.
+For information, (\textbf{Cite, Cite, Cite}) may be consulted.
+We implement our RNN in the \emph{Python} programming language and making use of the \emph{Scipy} and \emph{Tensorflow} libraries.
 
@@
 Inputs $u_n$ of dimension $K < N$ are padded by zeroes to obtain an input of dimension $N$.
-
-No need for entire left-infinite history of input: An arbitrary initialisation of the system with initial values and the driving of the system with the function $g$ means that the generated values will come to approximate the actual solution’s elements in a uniform manner thanks to the UAP.
-
-We then 3.set up/prepare the RNN.
-Utilising a network $N$ neurons (the same $N$ as in the dimensions of the matrices $A,B$).
-Training is accomplished using the Adam Optimiser where the MSE loss function is optimised.
+As we do not need the entire left-infinite history of input, an arbitrary initialisation of the system with initial values followed by the driving of the system with the function $g$ means that the generated values will approach the actual solution’s elements in a uniform manner thanks to the UAP.
+\textbf{(Cite.)}
+
+Thirdly we prepare the network.
+Utilising a network $N$ neurons (the same value $N$ as in the dimensions of the matrices $A,B$).
+Training is done using the Adam Optimiser where the MSE loss function is optimised.
 The system is trained several times(in our case 4) with incrementally smaller learning rates on each run (0.001 and thereafter divided by 10 at each run).
-\ednote{Please present the number of layers in the network, number of neurons in each layer and the length of the training validation and testing for learning Gamma.}
+Additional parameters such as the number of network-layers, neurons per layer and training-length are presented case-by-case below.
 
@@
 
-The two model equations for learning from data (\eqref{eqns_from_data}) also allows us to determine what the state $x_{n+1}$ at time $n+1$ will be. 
+The two model equations for learning from data (\eqref{eqns_from_data}) allows us to determine what the state $x_{n+1}$ at time $n+1$ will be. 
 We specifically remind ourselves of \[x_{k+1}=\pi_2 \circ (\Gamma, \pi_2) \circ (\pi_2,g) (u_k,x_k)\]
@@
 \subsection{Double Pendulum}
-\begin{itemize}
-  \item What it is. 
-  \item Why it's chaotic
-  \item How data was generated for this. (noiseless)
-  \item Why data dying down doesn't work (ie. Not surjective)
-  \item Other possibilities
-\end{itemize}
 
@@
 
-An ideal system is assumed where the rods are massless and no friction in the pivots are considered.
-One is easily reminded of the Lorenz system, another three-equation deterministic model which exhibits chaotic behaviour.
-The scalar equations of motion are presented below and were derived with the use of Newtonian physics and can be found in \textbf{(Cite.)}
-
-\begin{eqnarray}
-\theta_{1}^{''} = \frac{-g(2m_1+m_2)\sin{\theta_1} - m_{2}g\sin(\theta_1 - 2\theta_2) -2\sin(\theta_1-\theta_2)m_2({\theta_{2}'}^{2}L_2 + \theta_{1}^{2}L_{1}\cos(\theta_1 - \theta_2))}{L_1(2m_1+m_2-m_2\cos({2\theta_1 -2\theta_2}))}
-\\
-\theta_{2}^{''} = \frac{ 2\sin(\theta_1-\theta_2)(\theta_{1}'^2 L_1 (m_1+m_2) + g(m_1+m_2)\cos{\theta_1} + \theta_{2}'^2 L_2 m_2 \cos(\theta_1 - \theta_2))}{L_2(2m_1+m_2-m_2\cos({2\theta_1 -2\theta_2}))}
-\end{eqnarray}
-
-This second-order system was then converted into a first-order system by defining the first derivatives as separate values.`'
+This system has four variables: angle and angular velocity of each pendulum rod. An ideal system is assumed where the rods are massless and no friction in the pivots are considered.
+The scalar equations of motion are derived using Newtonian physics and can be found in \cite{DPFormulas}.
+The second-order system is converted to a first-order ODE and solved using the Runge-Kutta numerical method.
+
+This system is a classical example of a system exhibiting chaos. \textbf{Cite.}
+
+% \begin{eqnarray}
+% \theta_{1}^{''} = \frac{-g(2m_1+m_2)\sin{\theta_1} - m_{2}g\sin(\theta_1 - 2\theta_2) -2\sin(\theta_1-\theta_2)m_2({\theta_{2}'}^{2}L_2 + \theta_{1}^{2}L_{1}\cos(\theta_1 - \theta_2))}{L_1(2m_1+m_2-m_2\cos({2\theta_1 -2\theta_2}))}
+% \\
+% \theta_{2}^{''} = \frac{ 2\sin(\theta_1-\theta_2)(\theta_{1}'^2 L_1 (m_1+m_2) + g(m_1+m_2)\cos{\theta_1} + \theta_{2}'^2 L_2 m_2 \cos(\theta_1 - \theta_2))}{L_2(2m_1+m_2-m_2\cos({2\theta_1 -2\theta_2}))}
+% \end{eqnarray}
+
+\begin{itemize}
+  \item Why data dying down doesn't work (ie. Not surjective)
+  \item Parameter Values \& Success
+\end{itemize}
 
@@
 
+One is easily reminded of the Lorenz system, another three-equation deterministic model which exhibits chaotic behaviour.
+
 \section{Generalized Pearson correlation coefficient for the DP and other attractors}
+
 \textbf{Insert table: D.Pendulum, Lorenz, Others (Clifford \& Thomas)}
 
+\begin{center}
+  \begin{table}
+      \scalebox{0.75}
+      {\begin{tabular}{|c|c| c c |} 
+          \hline
+          Input & Dimension of $X$ & $u_n$ vs $u_{n+1}$ 
+          & $\begin{bmatrix}
+              x_{n-1}\\
+              x_n
+          \end{bmatrix}$ vs $\begin{bmatrix}
+              x_n\\
+              x_{n+1}
+          \end{bmatrix}$ \\
+          \hline
+
+          \hline
+          $(w_n)$ is the Lorenz states, sampled every 0.1 timestamps. & & &\\
+          \multirow{3}{*}{$u_n=\frac{1}{100}w_n$} & 10 &0.9311  & 0.9731\\
+              & 100 &  & 0.9934\\
+              & 1000 & & 0.9930\\
+
+          \multirow{3}{*}{$u_n=\frac{1}{10}(\sin(0.1w_{n,x})+\sin(0.1w_{n,y})+\sin(0.1w_{n,z}))$} 
+              & 10 &0.8401  & 0.9392\\
+              & 100 & & 0.9616\\
+              & 1000 & & 0.9737\\
+          \hline
+
+          \hline 
+          $(w_n)$ comes from the Thomas Attractor (for $\beta=0.1056$): & & & \\
+          {$w_{n+1}= \begin{bmatrix} \sin(w_{n,y}) - \beta{w_{n,x}} \\
+              u_y = \sin(w_{n,z}) - \beta{w_{n,y}} \\
+          u_z = \sin(w_{n,x}) - \beta{w_{n,z}} \end{bmatrix}$} & & & \\
+          \multirow{3}{*}{$u_n = w_n - \overline{w}$}
+              & 10 & 0.99989 & 0.999886 \\
+              & 100 & &  0.999913 \\
+              & 1000 & & 0.999901 \\
+          \hline
+
+          \hline
+          $(w_n)$ comes from Noise-Added Thomas Attractor (for $\beta=0.1056$): & & & \\
+          {$w_{n+1}= \begin{bmatrix} \sin(w_{n,y}) - \beta{w_{n,x}} \\
+              u_y = \sin(w_{n,z}) - \beta{w_{n,y}} \\
+          u_z = \sin(w_{n,x}) - \beta{w_{n,z}} \end{bmatrix}$} & & & \\
+          \multirow{3}{*}{$u_n = w_n - \overline{w}$}
+              & 10 & 0.9994 & 0.99985 \\
+              & 100 & &  0.99991 \\
+              & 1000 & & 0.999895\\
+          \hline
+
+          \hline
+          $(w_n)$ comes from the Thomas Attractor (for $\beta=0.208186$): & & & \\
+          {$w_{n+1}= \begin{bmatrix} \sin(w_{n,y}) - \beta{w_{n,x}} \\
+              u_y = \sin(w_{n,z}) - \beta{w_{n,y}} \\
+          u_z = \sin(w_{n,x}) - \beta{w_{n,z}} \end{bmatrix}$} & & & \\
+          \multirow{3}{*}{$u_n = w_n - \overline{w}$}
+              & 10 & 0.99982 & 0.99981\\
+              & 100 & &  0.999851 \\
+              & 1000 & & 0.99984 \\
+          \hline
+
+          \hline
+          $(w_n)$ comes from the Clifford Attractor \\ (for $a = -1.7; b = 1.8; c = -1.9; d = -0.4$): & & & \\
+        {$w_{n+1}= \begin{bmatrix} \sin(aw_{n,y}) + c\cos(aw_{n,x}) \\ 
+                                                                    \sin(bw_{n,x})+d\cos(bw_{n,y}) \end{bmatrix}$} & & & \\
+          \multirow{3}{*}{$u_n = w_n-\overline{w}$}
+              & 10 & -0.21713 & 0.78719 \\
+              & 100 & &  0.8334 \\
+              & 1000 & & 0.84854 \\
+          \hline
+
+          \hline 
+          $(w_n)$ comes from the Logistic map: & & & \\
+          {$ w_{n+1}  = 4w_n(1-w_n)$} & & & \\
+          \multirow{3}{*}{$u_n = w_n - 0.5$} 
+              & 10 &-0.0314  & 0.5577\\
+              & 100 &  & 0.7150\\
+              & 1000 & & 0.7826\\
+          \hline
+
+          \hline
+          $(w_n)$ comes from the Svensson Attractor \\ (for $a = -1.7; b = 1.8; c = -1.9; d = -0.4$): & & & \\
+          {$w_{n+1}= \begin{bmatrix} d\sin(aw_{n,x}) - \sin(bw_{n,y}) \\ 
+                                                                      c\cos(aw_{n,x})+\cos(bw_{n,y}) \end{bmatrix}$} & & & \\
+            \multirow{3}{*}{$u_n = w_n-\overline{w}$}
+                & 10 & -0.2243 & 0.49698 \\
+                & 100 & &  0.63310 \\
+                & 1000 & & 0.71151 \\
+          \hline
+
+      \end{tabular}} %\ednote{http://paulbourke.net/fractals/peterdejong/}
+
+      \caption{Multidimensional Correlation Coefficients $\rho$ to indicate a general reduction in the functional complexity of the map $G_T$ that arises from a RNN.
+Each row corresponds to a different dynamical system used in the experiments in \cite{Main_article}.
+In the second column corresponds to the dimension of $X$ (number of artificial neurons) in the RNN used.
+      The numerical estimate of $\rho$ for the relevant vectors is plotted in the last two columns.
+The reader can compare the linear relationship between $u_n$ vs $T(u_{n})$ and the linear relationship $\begin{bmatrix}
+              x_{n-1}\\
+              x_n
+          \end{bmatrix}$ vs $G_T \left( \begin{bmatrix}
+              x_{n-1}\\
+              x_n
+          \end{bmatrix}\right)$.
+The dimension of the RNN from \eqref{Seq_RNN} is also given, showcasing that increasing the dimension of the driven system typically yields a map $G_T$ of lower functional complexity.} \label{Table_FC}
+      \end{table}
+  \end{center}
+
 \chapter{Conclusion}\label{ch6}
 
-The driven states $x_{n-1},x_n)$ holds the information of the left-infinite history of the input.
+The driven states $(x_{n-1},x_n)$ holds the information of the left-infinite history of the input.
 Furthermore, there is some evidence that the inverse-limit space (not system) seems to distinguish two dynamical systems that are not topologically conjugate for a class of systems in the sense their inverse-limit spaces are not homeomorphic to each other.

### 2022-01-10 Gandhi_Project_Report/draft3.tex
Clifford, Thomas mostly done. USP discussion done. Working on introducing H2

@@
 \usepackage{multirow}
+\usepackage{longtable}
+\usepackage[export]{adjustbox}
 
@@
 \begin{Definition}
-  [\bf Flow of an Equation] \label{Dfn_Flow}
+  [\bf Flow of an Equation] \label{Dfn_Flow}\rm
   The flow of the equation  $\dot{x} = f(x)$  is defined to be a mapping $\varphi: \mathbb{R}^n \times \mathbb{R} \to \mathbb{R}$ where $\varphi(x_0,t)= x(t)$ for the solution $x(t)$ with $x(0)=x_0$ 
@@
 \begin{Definition}
-  [\bf Invariant Set] \label{Dfn_InvariantSet}
+  [\bf Invariant Set]\label{Dfn_InvariantSet}\rm
   Given a discrete-time dynamical system $T: U \to U$, a subset $A \subset U$ is said to be an \emph{invariant set} if $T(A) =A$. 
@@
 \begin{Definition}
-  [\bf Orbit of $T$] \label{Dfn_Orbit}
+  [\bf Orbit of $T$]\label{Dfn_Orbit}\rm
   The orbit of $T$ is to be defined the sequence $\bar{u} = \{u_n\}_{n\in \mathbb{Z}}$ obeying the update equation, $u_{n+1}=Tu_n$, $n \in \mathbb{Z}$. 
@@
 \begin{Definition}
-  [\bf Attractor] \label{Dfn_Attractor}
+  [\bf Attractor]\label{Dfn_Attractor}\rm
   Let $T: U \to U$, where $U$  is a metric space with metric $d$. A compact subset $A \subset U$ is said to be an attractor if it satisfies the three conditions: 
@@
 
-\begin{Definition}
-  [\bf {Sensitive Dependence on Initial Conditions}] \label{Dfn_SDIC}
+\begin{Definition}\rm
+  [\bf {Sensitive Dependence on Initial Conditions}]\label{Dfn_SDIC}\rm
 A dynamical system $T: U \to U$ is said to have sensitive dependence on initial conditions (SDIC) if there exists a $\delta > 0$ such that for every $u \in U$ and in every neighborhood of $u \in U$ there exists a $v\in{U}$ and an integer $N:=N{(u,v)}\in\mathbb{Z}$ such that $d(T^Nu,T^Nv)>\delta$.
@@
 \begin{Definition}
-  [\bf {Topologically Transitive}] \label{Dfn_TopTrans}
+  [\bf {Topologically Transitive}]\label{Dfn_TopolTrans}\rm
 	A dynamical system $F: U \to U$ is topologically transitive if for any pair of nonempty open sets $E_1$ and $E_2$ there exists a $n\in\mathbb{N}$ such that $T^n(E_1) \cap E_2 \not= \emptyset$. 
@@
 \begin{Definition}
-  [\bf {Meager Set}] \label{Dfn_Meager Set}
+  [\bf {Meager Set}]\label{Dfn_Meager Set}\rm
 A subset of a topological space $U$ is said to be a meager set if it can be written as a countable union of sets of with empty interior. The complement of the a meager set is set to be a residual set.
@@
 \begin{Definition}
-  [\bf {Devaney's Chaos}]\label{Dfn_ChaosDec}
+  [\bf {Devaney's Chaos}]\label{Dfn_ChaosDec}\rm
 	A dynamical system $T: U \to U$ is said to exhibit chaos in the sense of Devaney if it satisfies the three properties:
@@
 
-\begin{Definition}  
-  [\bf {Conjugacy \& Semi-Conjugacy}]\label{Dfn_Conjugate}
+\begin{Definition}\rm  
+  [\bf {Conjugacy \& Semi-Conjugacy}]\label{Dfn_Conjugate}\rm
   Consider the dynamical systems $T:U\to{U}$, $S:V\to{V}$ and suppose the relationship $\phi \circ T=S\circ \phi$ holds.
@@
 We define first the concepts of homeomorphism and embedding:
-\begin{Definition}
-  [\bf {Homeomorphism}]\label{Dfn_homeo}
+\begin{Definition}\rm
+  [\bf {Homeomorphism}]\label{Dfn_homeo}\rm
   A homeomorphism is a function $f:Z\rightarrow Y$ between two topological spaces $Z$ and $Y$ that is continuous, bijective and has a continuous inverse. 
@@
 \begin{Definition}
-  [\bf {Embedding}]\label{Dfn_embed}
+  [\bf {Embedding}]\label{Dfn_embed}\rm
   Consider a homeomorphism $f:Z\rightarrow Y$ for $Y\subset X$. $Z$ is said to be embedded in $X$ by $f$.
@@
 % \ednote{M: You can omit the theorem, and just say it is a uniform attractor in the nonautonomous dynamical systems literature \cite{Manju_Nonlinearity}, and say the discussion on nonautonomous attractors is beyond the scope of the thesis.}
-\textbf{NEED TO PARAPHRASE}
-\newline In our context, $g$ having the USP is equivalent to $g$ being a topological contraction \cite{manjunath2021universal}, or alternatively that $g$ exhibits the Echo State Property(ESP). 
-In \cite{Manju_ESP} the echo state property is a key for the design and training of recurrent neural networks within the paradigm of reservoir computing.
-Available definitions and theorems concerning the echo state property, however, are of little practical use because they do not relate the network response to temporal or statistical properties of the driving input.
-
-A greater discussion is presented in \cite{jaeger2001echo}
-a particular stability concept for which a number of equivalent definitions are available \cite{jaeger2001echo}.
-
-In most of the reservoir computing (RC) literature, the idea of memory-loss or fading memory popularly called
-the echo state property (ESP) is treated as a stability property of the system valid
-for all inputs. In such a case, the ESP guarantees that the entire past history of the
-input determines the state of the system precisely, i.e., there is no possibility of two
-or more reachable states at any given point in time [5] if the entire past values of the
-input have influenced the dynamics. The formal definition linking ESP intrinsically
-to an input is available in [4].
-
-In the context of information processing, when a driven system g has a causal em-
-bedding, it helps carry over any input ū onto the solution space without distortion or
-loss of information. This is a very important feature, and the current discrete-time
-state space models used in reservoir computing do not exploit the causal embedding.
-We intend to demonstrate the utility of these theoretical results in a future work
-directed at applications.
-
-For natural systems, say, for instance, when the human brain is viewed as a driven
-system where a time-varying external stimulus (for instance, a visual stimulus) is
-an input, the input can be causally embedded into the nervous system without the
-need for a particular structure that is custom-made for each stimulus as long as
-the system has the USP. From the perspective of perturbation of an autonomous
-dynamical system, if the resulting nonautonomous system has the USP, then the
-temporal structure of noise relates to the dynamics on a nonautonomous uniform
-attractor through a topological semi-conjugacy or a conjugacy.
-
-The continuity of E reflects on whether small changes in u ~ would result in a ‘propor-
-tionate’ response in the solution space – a notion of input related stability as defined in \cite{Manju_ESP}
-
-To understand how solutions become sensitive to a perturbation of the input for systems without USP, we define a stability notion that describes how the collection of all solutions obtained from an input ū are mapped onto the solution space.
-
-Sufficient to consider continuity of encoding map.
-
-owing to the USP of the driven system $g$ one gains input-related, parameter-related stability \cite{manjunath2020stability}
-
-Historically, data have been transformed into a different space for processing to gain some advantages.
-For example, according to Vapnik–Chervonenkis theory (e.g., [16]) mapping data into a higher dimension often provides much greater classification power.
-Mapping temporal data through a delay-coordinate mapping in Takens delay embedding uncovers hidden information [17]. In data-driven modeling employing Koopman’s theory (e.g., [18]–[20]), 
-approximating the dynamics of a dynamical system through a linear dynamicalsystem in a higher dimensional space is often possible by mapping the data samplewise into the higher dimensional space
-
-we first note that when T is not invertible, the dynamics of (U, T ) is only topologically semi-conjugate to $(\hat{U}, \hat{T})$ and hence the dynamics in $(\hat{U}, \hat{T})$ could have additional complexity.
-
-xxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxx
-
-Having already defined the reachable set $X_U$, we pause for a moment to fix additional notation.
+% Paraphrased.
+ In our context, $g$ having the USP is equivalent to $g$ being a topological contraction \cite{manjunath2021universal}, or alternatively that $g$ exhibits the Echo State Property (ESP).
+
+This property is used to describe the concept of fading memory or memory-loss in the majority of literature within the field reservoir computing (RC) \textbf{(Citation)}. 
+$g$ having the ESP guarantees that the entire left-infinite history of an input will determine exactly the current system state; there is only ever the possibility of a single reachable state at a given instance of time \cite{jaeger2001echo}
+The ESP is also often discussed in terms of a stability property that holds for all inputs of a system~\cite{manjunath2020stability} and, furthermore, plays a key role in the design and training of recurrent neural networks within the field of RC \cite{Manju_ESP}.
+
+Defined formally, the ESP is intrinsically linked to an input (see \cite{Manju_ESP}) Unfortunately, definitions and results concerning the ESP do not have much practical use since a relationship of the network's response to the temporo-statistical properties of the input is not directly described.
+A more detailed presentation may be found in~\cite{jaeger2001echo}.
+
+\section{Next section? \ednote{I feel a break is necessary here, but would hear from you what your thoughts are.}}
+
+Having already defined the reachable set $X_U$, we now pause for a moment to fix additional notation.
 Letting $\cev{u}^{n}:=(\ldots,u_{n-2} ,u_{n-1})$ be the left-infinite subsequence of an input, $\overleftarrow{U}$ then denotes all these left-infinite sequences in $U$. 
@@
 
-  Restricting our attention more and more, we now consider a specific subclass of conjugacies.
+  Continuing to refine our attention, we proceed to consider a specific subclass of conjugacies.
 
@@
 
-\textbf{make special note of a} %\ednote{something you want to fill?} }
 \begin{Definition}
@@
 
-\textbf{When considering the map $H$, we once again restrict our focus of $H$ to the function $H_2(\overline{u}):=(h(r\overline{u}, h(\overline{u})))$. }
+\textbf{We direct our attention to the function $H_2(\overline{u}):=(h(r\overline{u}, h(\overline{u})))$. }
 
@@
 \section{Numerical simulations}
+In this section we showcase numerical results pertaining to 3 attractors that were simulated using the methodology described in the previous chapters. We first consider a simple physical system.
 \subsection{Double Pendulum}
@@
 
-This system has four variables: angle and angular velocity of each pendulum rod. An ideal system is assumed where the rods are massless and no friction in the pivots are considered.
+The system has four variables: angle and angular velocity of each pendulum rod. An ideal system is assumed where the rods are massless and no friction in the pivots are considered.
 The scalar equations of motion are derived using Newtonian physics and can be found in \cite{DPFormulas}.
@@
 
-\begin{itemize}
-\item Thomas:
+\subsubsection{Thomas' Cyclically Symmetric Strange Attractor}
+Thomas' Cyclically Symmetric Strange Attractor, a 3D attractor proposed in 1999 by Ren\'e Thomas in \cite{ThomasAttractor}, is described by a set of three equations which is distinctly reminiscent of the Lorenz system, another deterministic three-equation model exhibiting chaotic behaviour.It has a single parameter $\beta$ and has been shown to transition to chaotic behavious when $\beta<0.208186$~\cite{Thomas_BetaParameter}.
 \begin{eqnarray}\label{eqns_thomas}
-  u_x = \sin(y) - \beta{x} \\
-  u_y = \sin(z) - \beta{y} \\
-  u_z = \sin(x) - \beta{z}
+  x_{n+1} = \sin(y_n) - \beta{x_n} \\
+  y_{n+1} = \sin(z_n) - \beta{y_n} \\
+  z_{n+1} = \sin(x_n) - \beta{z_n}
 \end{eqnarray}
 
-\item Clifford:
-\begin{equation}\label{eqn_clifford}
-  \frac{\sin(ay) + c\cos(ax)}{\sin(bx)+d\cos(by)}
-\end{equation}
-\end{itemize}
-
-One is easily reminded of the Lorenz system, another three-equation deterministic model which exhibits chaotic behaviour.
-
-\section{Generalized Pearson correlation coefficient for the DP and other attractors}
-
-\textbf{Insert table: D.Pendulum, Lorenz, Others (Clifford \& Thomas)}
+The behaviour of the system with a $\beta$-parameter value of 0.1056 is simulated by scaling it to fit inside the interval $[-1,1]$ and both perturbed with noise as well as leaving it intact.
+Specifically the noise is obtained from a normally distribution with mean 0 and standard deviation 0.05, which translates approximately to *dB.
+ A sequence of 3000 observations in the clear and noisy data-set respectively is fed into the system with 300 entries discarded before predicting the next 9000 entries.
+
+\begin{figure}[ht]
+  \centering
+  \minipage{0.5\linewidth}
+    \centering
+    \includegraphics[width=\linewidth]{Thomas_nonoise.eps}
+    \caption*{True and predicted trajectories in the absence of noise.}
+  \endminipage\hfill
+  \minipage{0.5\linewidth}
+    \centering
+    \includegraphics[width=\linewidth]{ThomasNoisy.eps}
+    \caption*{True and predicted trajectories for noise added.}
+  \endminipage
+\end{figure}
+
+The RNN was initialised as follows: The dimension of the network is set equal to 500 neurons and 12 hidden layers of dimension 64 each were initialised.
+Training is realised through 150 epochs of batch sizes of 128 each; the map is learnt via the Adam Optimiser using finer and finer learning rates (0.001 and then divided by 10 at every iteration).
+
+\begin{figure}[ht]
+  \centering
+  \includegraphics[scale=0.35]{Thomas_1.eps}\caption*{Predicted trajectories for the attractor demonstrate empirically the ability to predict the evolution of the trajectory for the next ~100 timesteps}
+  \includegraphics[scale=0.5]{Thomas_3.eps}\caption*{Represented here the learnt (blue) and actual (red) densities of the first coordinate of the dataset.}
+\end{figure}
+
+\subsubsection{Fractal Dream Attractor}
+
+Another 2-equation discrete-time map named the Fractal Dream Attractor, or more commonly known as the Pickover Map (first discovered by Clifford A.
+Pickover and discussed in his fascinating book "Chaos in Wonderland"\cite{PickoverChaos}), was also considered.
+\begin{eqnarray}\label{eqns_clifford}
+  {x_{n+1}=\sin(ay_n) + c\cos(ax_n)} \\
+  {y_{n+1}=\sin(bx_n)+d\cos(by_n)}
+\end{eqnarray}
+
+\begin{figure}[ht]
+  \centering
+  \includegraphics[width=1.0\textwidth,left]{Clifford_1.eps}
+  \caption*{These graphs were constructed by predicting the Clifford system 15000 steps into the future and in so doing illustrating the long-term consistency and accuracy of the learnt system.
+As perceived here, we are able to lock on to the trajectory of the Clifford map almost exactly for the first 25 steps.}
+  \minipage{0.5\textwidth}
+      \centering
+      \includegraphics[width=\linewidth]{Clifford_2.eps}
+  \endminipage\hfill
+  \minipage{0.5\textwidth}
+    \centering
+    \includegraphics[width=\linewidth]{Clifford_3.eps}
+  \endminipage
+  \caption{The bottom row, left, shows a plot of the true attractor(red) and the predicted values(blue).
+Bottom row, right, shows the distribution of the first coordinate of the learnt system (blue) and the actual Clifford system (red).
+}
+  \label{Fig_Clifford}
+\end{figure}
+
+\section{Generalized Pearson Correlation}
+\ednote{B: Need to change heading here.}
+
+\textbf{Pearson Introduction}
 
@@
           \hline  
-
           \hline 
@@
           \hline
-
           \hline
@@
           \hline
-
+          \hline
+          $(w_n)$ comes from the DP \\ (for ): & & & \\
+          \multirow{3}{*}{$u_n = w_n-\overline{w}$}
+                & 10 & 0.99141 & 0.98916\\
+                & 100 & &  0.99238 \\
+                & 1000 & & 0.99278 \\
+          \hline        
       \end{tabular}} %\ednote{http://paulbourke.net/fractals/peterdejong/}
@@
 The dimension of the RNN from \eqref{Seq_RNN} is also given, showcasing that increasing the dimension of the driven system typically yields a map $G_T$ of lower functional complexity.} \label{Table_FC}
-      \end{table}
+      \end{table}\label{tbl_attractorsPearson}
 \end{center}

### 2022-01-11 Gandhi_Project_Report/draft3.tex
Submitted to Gandhi for review. Still lots to be done

@@
+\documentclass[a4paper,12pt,twoside]{report}
 % \documentclass[a4paper,12pt,twoside]{book}
-\documentclass[a4paper,12pt,twoside]{report}
 \pagestyle{headings}
@@
 
-\title{Project Thesis/Report}
+\title{Project Report}
 
@@
 Examples in the literature attempting to forecast data from such systems have tried a number of different approaches with differing degrees of success. 
-When we have complete access to states of the system, an ordinary differential equation can be obtained from data (Citations to Sindy and \cite{small2002modeling,xu2006modeling}) wherein one could approximate the vector field by a library of functions to obtain interpretable models.
+When we have complete access to states of the system, an ordinary differential equation can be obtained from data (\cite{brunton2016discovering, champion2019data} and \cite{small2002modeling,xu2006modeling}) wherein one could approximate the vector field by a library of functions to obtain interpretable models.
 
@@
 Such methods, obviously, do not guarantee an exact reconstructions for nonlinear models, and in practice provide poor long-term consistency for a large class of systems.
-(Add citation)
+\textbf{(Citation)}
 
@@
 One could learn a system conjugate to the underlying system by applying the Takens delay embedding \cite{takens1981detecting} when one has ``good" observations from the system; this learnt system could then be used for forecast the observed data.
-Takens delay embedding theorem \cite{takens1981detecting} and its various generalisations (e.g., \cite{sauer1991embedology, stark1999delay, gutman2018embedding, gutman2018embedding}) establishes the learnability of a system constructed by concatenating sufficiently large previous time-series observations of a dynamical system into a vector (called delay coordinates).
+Takens delay embedding theorem \cite{takens1981detecting} and its various generalisations (e.g., \cite{sauer1991embedology, stark1999delay, gutman2018embedding}) establishes the learnability of a system constructed by concatenating sufficiently large previous time-series observations of a dynamical system into a vector (called delay coordinates).
 This then establishes the existence of a map on the space of delay coordinates equivalent (or topologically conjugate) to the underlying map from which the observed time-series was first obtained.
@@
 One reason for this fragility is that the embedded attractor in the reconstruction space not always an attractor of the map that is learnt in the reconstruction space, despite  unquestionably being an invariant set.
-When the embedded object is not \textbf{(definitely/always?)} an attractor (as explained in Chapter 3) it can cause predictions to fail.
+When the embedded object is not explicitly known to be an attractor (as explained in Chapter \ref{ch3}) it can cause predictions to fail.
 
@@
 
-This thesis deals with the implementation and analysis of non-interpretable models that can guarantee \textbf{exact/accurate} reconstruction.
-The project work concerns the study and implementation of a method (Citation to paper with Adriaan) that incorporates learning a function by mapping the data on to a higher dimensional space using what is called a driven dynamical system (See Chapter 4).
+This thesis deals with the implementation and analysis of non-interpretable models that can guarantee exact reconstruction.
+The project work concerns the study and implementation of a method (\cite{manjunath2021universal}) that incorporates learning a function by mapping the data on to a higher dimensional space using what is called a driven dynamical system (See Chapter 4).
 With a clear understanding of how the data is mapped onto the higher dimensional space, the method then permits the learning of a dynamical system topologically conjugate to that of the underlying system.
@@
 
-By solidifying the mathematical underpinnings of our theory, we hope to guarantee the ability to construct models with predictive power ranging from molecular biology to neuroscience.
+\emph{By solidifying the mathematical underpinnings of our theory, we hope to guarantee the ability to construct models with predictive power ranging from molecular biology to neuroscience.
 in the near future.
-(We can modify this sentence when the report is completed.)
-
-The report is organised as follows: 
+(We can modify this sentence when the report is completed.)}
+
+The more intricate details of proofs are referred to where relevant throughout. This report is organised as follows: 
 \newline In Chapter~\ref{ch2}, we recall the definition of a discrete-time dynamical system, how a discrete-system arises from a flow of an ODE and then proceed to define the inverse limit space and topological conjugacy of autonomous systems.
 \newline In Chapter~\ref{ch3}, we introduce the problem of forecasting dynamical systems, state the Takens delay embedding theorem and discuss various issues faced while forecasting. 
-\newline In Section 4, a driven dynamical system is defined and discuss the properties of a specific class of driven dynamical systems that we make use of in this thesis/report/paper. 
-\newline Finally in Section 5 we show the implementation of these forecasting methods, and conclusions are provided in Section 6.
-
-\chapter{Discrete-time Dynamical Systems} \label{ch2}
+\newline In Chapter~\ref{ch4}, a driven dynamical system is defined and discuss the properties of a specific class of driven dynamical systems that we make use of in this thesis/report/paper. 
+\newline Finally in Chapter~\ref{ch5} we show the implementation of these forecasting methods, and conclusions are provided in Chapter~\ref{ch6}.
+
+\chapter{Discrete-time Dynamical Systems}\label{ch2}
 
@@
 Of course, discrete-time dynamical systems need not always arise through a differential equation, and once again we may consider the field of ecology where discrete dynamical systems are often directly derived or assumed.
-\textbf{Cite.}
+\textbf{Citation}
 There is a school of thought that advocates discrete-time dynamical systems to be more natural for modelling real-world observations than differential equations and we refer the interested reader to \cite{saber2010introduction}.
@@
 
-\ednote{For let $B$ be an attractor of U.
+% \ednote{For let $B$ be an attractor of U.
 From above, we know that B is itself invariant, is asymptotically stable and has Lyapunov stability.
@@
 
-  In a more rigourous fashion ....
-  \ednote{B:Should I be more rigourous here? }
+    \ednote{B: Is it desirable to have a more rigourous explanation here? }
 
@@
 So please edit accordingly.
-B: Does defining explicitly the domain and codomain of $phi$ make it clear enough?
+B: Does defining explicitly the domain and codomain of $\phi$ make it clear enough?
 }
@@
 First we establish the method of Takens delay embedding. 
-In this method, one considers a discrete-time dynamical system defined on a 'nice' space (a smooth manifold)\ednote{Should this be further explained?} that can be obtained as time-$K$ map of a continuous time-dynamical system \ednote{B: Should one add here: "These notions will be made formal below"?}.
+In this method, one considers a discrete-time dynamical system defined on a 'nice' space (a smooth manifold)\ednote{Should this be further explained?} that can be obtained as time-$K$ map of a continuous time-dynamical system, concepts which will be formalised below.
 We make observations from such a system, i.e., we observe the evolution of $\theta(w_0)$ (where $w_0$ represents the initial state) i.e., we observe  a finite set of values of $u_n :=  \theta(w_n) = \theta(Tw_{n-1})$.
@@
 By almost every choice here, we mean that we exclude some observations which do not give any information.
-To see a more precise discussion, see \textbf{xxx} for we do not concern ourselves with a further elaboration on this thesis/report.
+To see a more precise discussion, see \textbf{xxx} for we do not concern ourselves with a further elaboration on this report.
 
@@
 \end{figure}
-\ednote{M: Here, and throughout if you reproduce a figure, it must be stated where it was reproduced from.
-B: How does one write that a figure is reproduced?
-Also, should $\tilde{F}$ be left out for now?}
-\ednote{B: Could I perhaps have the original figure? This is only a screenshot used as a temporary placeholder.}
+\ednote{M: Here, and throughout if you reproduce a figure, it must be stated where it was reproduced from. B: Is this satisfactory? Also, should $\tilde{F}$ be left out for now?}
+\ednote{B: This is only a screenshot used as a temporary placeholder. If sufficient, I shall convert it to .eps. If not, may I perhaps have the original picture?}
 
@@
 By preserving the topology on the manifold $U$ in the reconstruction space $X$ (\textbf{not discussed before}), we are also guaranteed the preservation of topological invariants of the manifold, of which dimensionality is one such invariant.
-(We note here that dimensionality will again be considered later in this project/report).
+(We note here that dimensionality will again be considered later in this project).
 
@@
 There are, however, no theoretical guarantees that $F_\theta$ will retain  $V=\Phi_{2d,\theta}(W)$  as an attractor although $W$ could itself be an attractor.
-\item Takens theorem is stated for noiseless observations.
+\item Takens theorem is stated only for noiseless observations.
 Due to noise $\epsilon_n$ the delay vector $\Phi_{2d,\theta}(w_n) + \epsilon_n$  may lie outside $V$.Furthermore, due to the chaotic nature of the underlying system (i.e.
 the fact that it has SIDC), the evolution of $\Phi_{2d,\theta}(w_n) + \epsilon_n$ under the map $F_\theta$ could move out of $V$ completely.
-This problem can be overcome  by using a driven dynamical system with some properties, and we discuss this in Chapter *.
-%  \item It is an unfortunate case that the functional complexity (i.e., the relative number of oscillatory graphs \cite{manjunath2021universal}) as well as stability of the embedding \Ftheta{} (\textbf{cite?})depends on the choice of observable $\theta$.
-This means that we may even learn a map that complicates our learning task and in so doing adds obstacles to the road to learning a conjugate map with predictive power.
-\ednote{Am I talking nonsense here?
-M: Please avoid this point and delete it}
+This problem can be overcome  by using a driven dynamical system with some properties, and we discuss this in Chapter\ref{ch3}.
 \end{enumerate}
@@
 
-\textbf{Continue writing here}
-
-\begin{figure}[ht]
-  \includegraphics[scale=0.3]{_learningfailure.png}
-  \centering
-  \label{fig:supp_learning_failure}
-  \caption{Illustration of a system having failed to learn the map}
-\end{figure}
-
-\textbf{Explanation continued here}
-
-\begin{figure}[ht]
-  \includegraphics[scale=0.4]{_dp_dying.eps}
-  \centering
-  \label{fig:damped_pendulum}
-\caption{A damped pendulum}
-\end{figure}
-
-\textbf{Continue explaining here}
+The Takens embedding theorem does not guarantee global dissipativity.
+\ednote{B:Links with~\ref{subs_LearnGamma} in discussion of dissipativity of $\Gamma$.
+Sir, I'm not sure how to write this here - I think it a beneficial point to make so as to set the stage for the later discussion in ~\ref{subs_LearnGamma}, but I am unsure there as well.}
+
+% \begin{figure}[ht]
+%   \includegraphics[scale=0.3]{_learningfailure.png}
+%   \centering
+%   \caption{Illustration of a system having failed to learn the map}
+%   \label{fig:supp_learning_failure}
+% \end{figure}
 
@@
 
-\section{Next section? \ednote{I feel a break is necessary here, but would hear from you what your thoughts are.}}
+\section{Next section?}
+\ednote{I feel a break is necessary here, but would hear from you what your thoughts are.}
 
@@
 \begin{proof}
-  See \cite[Th.3]{manjunath2013echo}.
+  See~\cite[Th.3]{manjunath2013echo}.
 \end{proof}
 
-\textbf{We direct our attention to the function $H_2(\overline{u}):=(h(r\overline{u}, h(\overline{u})))$. }
+We direct our attention to the function \[H_2{\overline{}} := (h(r\overline{u}, h(\overline{u}))) \]
 
@@
 
-The driven system $g$ can induce an embedding of $\overleftarrow{U}$ in $\overleftarrow{X}$ as follows: If $H:\overleftarrow{U}\to\overleftarrow{X}_U$ is also injective (in addition to being surjective, and hence bijective), it becomes the embedding of $\overleftarrow{U}$ in $\overleftarrow{X}$ induced by the driven system $g$ and we refer to $H$ as a \emph{causal embedding} \ednote{M : Causal embedding a dynamical system is different from the definition of causal embedding, so please edit/delete this paragraph and the next theorem and also in the remaining part till Section 1.2}.
+The driven system $g$ can induce an embedding of $\overleftarrow{U}$ in $\overleftarrow{X}$ as follows: 
+If $H:\overleftarrow{U}{\to}{\overleftarrow{X}_U}$ is also injective (in addition to being surjective, and hence bijective), 
+it becomes the embedding of $\overleftarrow{U}$ in $\overleftarrow{X}$ induced by the driven system $g$ and we refer to $H$ as a \emph{causal embedding} 
+\ednote{M : Causal embedding a dynamical system is different from the definition of causal embedding, so please edit/delete this paragraph and the next theorem and also in the remaining part till Section 1.2}.
 
@@
 
-When $g$ has the USP, the diagram below illustrates the operation of the the mappings $h$ and $H$.
+When $g$ has the USP, the diagram below illustrates the operation of the mappings $h$ and $H$.
 The mapping $h:\overleftarrow{U}\to{X_U}$ is an observable as discussed in an above section (\ref{TakensEmbedding}).
@@
 
-% We have not as of yet treated the UAP properly, and establish the necessary result here (without proof) so as not to break up the completeness of theorem.
-
-\section {Choosing the driven system $g$}
+\section{Choosing the driven system $g$}
 
@@
 
-When a causal embedding $H$ exists for the driven system $g$, one can map an arbitrary input ${u}$ onto the solution space $X$ without additional distortion or information-loss \textbf{(Cite.')}.
+When a causal embedding $H$ exists for the driven system $g$, one can map an arbitrary input ${u}$ onto the solution space $X$ without additional distortion or information-loss \textbf{(Cite.)}.
 When an embedding is established, the question of possible additional complexity in the solution is removed by guaranteeing that, since the systems are conjugate (semi-conjugate, \textbf{(refer)}), $g$ does not add any (some) complexity to the system.
@@
   [\bf Uniform Attraction Property]\label{Dfn_UAP}\rm
-  A driven system $g$ has the Uniform Attraction Property (UAP) if, regardless of starting position, all trajectories \textbf{(of what?)} converge to a single trajectory as time flows forward. 
-  This trajectory is also the unique solution sequence $x$ to the input sequence $u$ as mentioned above \ref{here_UAP_firstestablished}.
+  A driven system $g$ has the Uniform Attraction Property (UAP) if, regardless of starting position, all trajectories converge to a single trajectory as time flows forward. 
+  This trajectory is also the unique solution sequence $x$ to the input sequence $u$ as mentioned above.
 \end{Definition}
@@
 
-%Change the theorem above so that th.4 is split up into two parts.
-
 \begin{figure}[ht]
-  %\includegraphics[scale=0.4]{memloss_conttime.png}
-  \centering
-\captionof{figure}{Approaching the true solution if g has the UAP}
+  \includegraphics[scale=0.4]{_memloss_conttime.png}
+  \centering
+\captionof{figure}{Reproduced from \cite{manunath2021universal}. Illustration of all trajectories approaching the true solution if $g$ has the UAP}
 \label{fig:memloss_conttime.png} 
@@
 In the presence of the UAP, however, small measurement errors introduced into the system do not pose the same danger as before since the trajectories will eventually tend towards the actual solution.
-\ednote{(B: Complete enough?)}
+\ednote{(B: Is this Complete enough?)}
 % \textbf{From the perspective of perturbation of an autonomous dynamical system, if the resulting nonautonomous system has the USP, then the temporal structure of noise relates to the dynamics on a nonautonomous uniform attractor through a topological semi-conjugacy or a conjugacy.}
@@
 
-\textbf{NB NB NB - Causal Embedding Theorem! - Th.4}
+% \ednote{Add Causal Embedding Theorem}
+% \textbf{NB NB NB - Causal Embedding Theorem! - Th.4}
 
@@
 
-\begin{figure}[ht]
-  \includegraphics[scale=0.3]{_gammatemp.png}
-  \centering
-  \label{fig:temp_Gamma}
-\end{figure}
+% \begin{figure}[ht]
+%   \includegraphics[scale=0.3]{_gammatemp.png}
+%   \centering
+%   \label{fig:temp_Gamma}
+% \end{figure}
+
+\begin{equation}  \label{Scomm_imp}
+  %    \[ 
+      \psset{arrows=->, arrowinset=0.25, linewidth=0.6pt, nodesep=3pt, labelsep=2pt, rowsep=1.3cm, colsep = 1.1cm, shortput =tablr}
+   \everypsbox{\scriptstyle}
+   \begin{psmatrix}
+   \widehat{U}_T & \widehat{U}_T\\%
+   Y_T & Y_T\\
+   & \textcolor{red}{U \times X}
+   %%%
+  %  \ncline{1,1}{1,2}^{\widehat{T}} \ncline{1,1}{2,1} <{H_2}
+  %  \ncline{1,2}{2,2} > {H_2}
+  %  \ncline{2,1}{2,2}^{G_T}
+  %  \psset{linecolor=red}
+  %  \textcolor{red}{\ncline{2,1}{3,2}<{(\Gamma,\pi_2)}}
+  %  \ncline{3,2}{2,2}>{\textcolor{red}{(\pi_2,g)}}.
+   \end{psmatrix}
+  % \]
+  \end{equation}
 
@@
 Informally, we may consider input-related stability as concerned with the question of whether or not small variations in input would result in small responses.   
-If two inputs $u$ and $u_noisy$ are close in the product topology, then their tails could differ greatly due to the metric's insensitivity to the differences in sequence's tails.
- However if the system has the USP then the function $h$ is continuous and so $h(u), h(u_noisy)$ remain near – maintaining a measure of robustness to input and measurement noise.\cite{manjunath2021universal}.
+If two inputs $u$ and $u_{noisy}$ are close in the product topology, then their tails could differ greatly due to the metric's insensitivity to the differences in sequence's tails.
+ However if the system has the USP then the function $h$ is continuous and so $h(u), h(u_{noisy})$ remain near – maintaining a measure of robustness to input and measurement noise.\cite{manjunath2021universal}.
 In a similar fashion, parameter-related stability is obtained thanks to a result \cite[Lemma 3.2]{manjunath2020stability} relating this form of stability with the ESP (which is here equivalent to the USP).
@@
 \begin{figure}[ht]
-  \includegraphics[scale=0.4]{_AutovsDriven.eps}
+  \includegraphics[scale=0.5]{_AutovsDriven.eps}
   \centering
@@
 RNN's are widely used in the literature for discrete-time processing and are considered to be discrete-time version of continuous dynamical systems modelling a network of neurons.
-(\textbf{Cite.})
+(\textbf{Citation.})
 A complete discussion of RNN's is outside the scope of this thesis, so we present only the details necessary to the current discussion.
@@
 As we do not need the entire left-infinite history of input, an arbitrary initialisation of the system with initial values followed by the driving of the system with the function $g$ means that the generated values will approach the actual solution’s elements in a uniform manner thanks to the UAP.
-\textbf{(Cite.)}
+\textbf{(Citation)}
 
@@
 
-\section{Changes in this thesis/project/report to previous work of~\cite{manjunath2021universal}}
+\section{Changes in this project to previous work of~\cite{manjunath2021universal}}
 
@@
 So may be you just explain if you can’t show it.}
-\item Sparse matrix implementation. Slight improvement in some scenarios, but not fantastic. \ednote{Did it not help you a great deal with the DP forecasting?}
+\item Sparse matrix implementation.
+Slight improvement in some scenarios, but not fantastic.
+\ednote{M: Did it not help you a great deal with the DP forecasting?
+B: Sparse matrices tended to be finicky.
+In some cases they helped whereas in others they actually worsened the solution somewhat.
+There's some tuning to be done in terms of condition number, scaling and density}
 
 (Possible paragraph here/later on implementing sparsity of A/B: Generate sparse matrix that is also well-conditioned and correctly scaled matrix.)
-\item Density of matrices A/B (separate from sparse matrix implementation) \ednote{Is it the probability distributions of the entries in A and B?}
+\item Density of matrices A/B(i.e. number of non-zero entries in each matrix) (separate from sparse matrix implementation)
 \end{enumerate}
@@
 (See below)
-
-\begin{figure}[ht] \label{Fig_Gamma}
-  \includegraphics[scale=0.25]{_dbl_Pend.png}
-  \centering
-  \label{fig:conjugacy_with_Gamma}
+This system is a classical example of a system exhibiting chaos. \textbf{(Citation.)} Consider again~\ref{fig:dp_sdic}.
+The system has four variables: angle and angular velocity of each pendulum rod. An ideal system is assumed where the rods are massless and no friction in the pivots are considered.
+
+\begin{figure}[ht]
+  \includegraphics[scale=0.35]{_dp_setup.png}
+  \centering
+  \captionof{figure}{Pictorial representation of the setup of a DP. Image taken from~\cite{DPSetup}}
+  \label{fig:dp_setup}
 \end{figure}
 
-The system has four variables: angle and angular velocity of each pendulum rod. An ideal system is assumed where the rods are massless and no friction in the pivots are considered.
-The scalar equations of motion are derived using Newtonian physics and can be found in \cite{DPFormulas}.
+The scalar equations of motion are derived using Newtonian physics and can be found in~\cite{DPFormulas}.
 The second-order system is converted to a first-order ODE and solved using the Runge-Kutta numerical method.
-
-This system is a classical example of a system exhibiting chaos. \textbf{Cite.}
-
-% \begin{eqnarray}
-% \theta_{1}^{''} = \frac{-g(2m_1+m_2)\sin{\theta_1} - m_{2}g\sin(\theta_1 - 2\theta_2) -2\sin(\theta_1-\theta_2)m_2({\theta_{2}'}^{2}L_2 + \theta_{1}^{2}L_{1}\cos(\theta_1 - \theta_2))}{L_1(2m_1+m_2-m_2\cos({2\theta_1 -2\theta_2}))}
-% \\
-% \theta_{2}^{''} = \frac{ 2\sin(\theta_1-\theta_2)(\theta_{1}'^2 L_1 (m_1+m_2) + g(m_1+m_2)\cos{\theta_1} + \theta_{2}'^2 L_2 m_2 \cos(\theta_1 - \theta_2))}{L_2(2m_1+m_2-m_2\cos({2\theta_1 -2\theta_2}))}
-% \end{eqnarray}
-
-\begin{itemize}
-  \item Why data dying down doesn't work (ie. Not surjective)
-  \item Parameter Values \& Success
-\end{itemize}
+Noise-free, undamped DP data-sets were generated and utilised in this project due to the requirement that the map $T$ be surjective.
+It may easily be verified that a DP undergoing damping (as seen in \ref{conjugacy_with_Gamma}) is not surjective and hence the required assumptions to apply our theory do not hold.
+\ednote{B: Change sentence.}
+
+\begin{figure}[ht]
+  \includegraphics[scale=0.4]{_dp_dying.eps}
+  \centering
+\caption{A damped pendulum is not surjective.}
+\label{fig:damped_pendulum}
+\end{figure}
+
+\textbf{Parameter Values \& Successes}
+
+Train length: 15000. Prediction Length: 1000. Discard to wait for network to forget: 500. Dimension of network: 128 neurons.
+$\alpha=0.99$. $a=0.5$.
+
+Hidden Layers: 16. Layer dimension: 64. Epochs:256. Batch Size: 128. Delay(for MDD): 8. Density of Sparse Matrices: 0.2. $d=1$
+
+\ednote{B: Write this paragraph.}
+
+\begin{figure}[ht]
+  \centering
+  \includegraphics[width=\linewidth]{_DP_Fig-2F2023.eps} 
+  \captionof{figure}{Predicted(red) and actual(blue) trajectories of the $x$- and $y$-coordinates of each of the pendulum heads.} 
+\end{figure}
+\begin{figure}[ht]
+  \centering
+  \includegraphics[width=\linewidth]{_DP_Fig-2B2023.eps}
+  \captionof{figure}{Densities of predicted and actual values of x-coordinate of first pendulum head. }
+  \end{figure}
+  \begin{figure}[ht]
+    \centering 
+  \includegraphics[width=\linewidth]{_DP_Fig-2D_5.eps}
+    \captionof{figure}{Takens Delay Map with delay=5 to illustrate behaviour of predicted and actual values of double pendulum data.}
+\end{figure}
 
 \subsection{Other attractors I’ve experimented with: Clifford \& Thomas}
-We follow the methodology presented in \cite{manjunath2021universal} but opt to consider different attractors this time.
+We follow the methodology presented in \cite{manjunath2021universal} but opt to consider attractors different from the Lorenz system and H\'enon map in this project.
 
@@
 
-\section{Generalized Pearson Correlation}
-\ednote{B: Need to change heading here.}
-
-\textbf{Pearson Introduction}
+\ednote{B: Expand more.}
+
+\section{Generalized Pearson Correlation}\ednote{B: Alter heading.}
+
+\ednote{B: Write Pearson Introduction. Could you give me some help here, Sir?}
 
@@
 
-          \hline
-          $(w_n)$ is the Lorenz states, sampled every 0.1 timestamps. & & &\\
-          \multirow{3}{*}{$u_n=\frac{1}{100}w_n$} & 10 &0.9311  & 0.9731\\
-              & 100 &  & 0.9934\\
-              & 1000 & & 0.9930\\
-
-          \multirow{3}{*}{$u_n=\frac{1}{10}(\sin(0.1w_{n,x})+\sin(0.1w_{n,y})+\sin(0.1w_{n,z}))$} 
-              & 10 &0.8401  & 0.9392\\
-              & 100 & & 0.9616\\
-              & 1000 & & 0.9737\\
-          \hline
+          % \hline
+          % $(w_n)$ is the Lorenz states, sampled every 0.1 timestamps. & & &\\
+          % \multirow{3}{*}{$u_n=\frac{1}{100}w_n$} & 10 &0.9311  & 0.9731\\
+          %     & 100 &  & 0.9934\\
+          %     & 1000 & & 0.9930\\
+
+          % \multirow{3}{*}{$u_n=\frac{1}{10}(\sin(0.1w_{n,x})+\sin(0.1w_{n,y})+\sin(0.1w_{n,z}))$} 
+          %     & 10 &0.8401  & 0.9392\\
+          %     & 100 & & 0.9616\\
+          %     & 1000 & & 0.9737\\
+          % \hline
 
@@
           \hline  
-          \hline 
-          $(w_n)$ comes from the Logistic map: & & & \\
-          {$ w_{n+1}  = 4w_n(1-w_n)$} & & & \\
-          \multirow{3}{*}{$u_n = w_n - 0.5$} 
-              & 10 &-0.0314  & 0.5577\\
-              & 100 &  & 0.7150\\
-              & 1000 & & 0.7826\\
-          \hline
-          \hline
-          $(w_n)$ comes from the Svensson Attractor \\ (for $a = -1.7; b = 1.8; c = -1.9; d = -0.4$): & & & \\
-          {$w_{n+1}= \begin{bmatrix} d\sin(aw_{n,x}) - \sin(bw_{n,y}) \\ 
-                                                                      c\cos(aw_{n,x})+\cos(bw_{n,y}) \end{bmatrix}$} & & & \\
-            \multirow{3}{*}{$u_n = w_n-\overline{w}$}
-                & 10 & -0.2243 & 0.49698 \\
-                & 100 & &  0.63310 \\
-                & 1000 & & 0.71151 \\
-          \hline
+          % \hline 
+          % $(w_n)$ comes from the Logistic map: & & & \\
+          % {$ w_{n+1}  = 4w_n(1-w_n)$} & & & \\
+          % \multirow{3}{*}{$u_n = w_n - 0.5$} 
+          %     & 10 &-0.0314  & 0.5577\\
+          %     & 100 &  & 0.7150\\
+          %     & 1000 & & 0.7826\\
+          % \hline
+          % \hline
+          % $(w_n)$ comes from the Svensson Attractor \\ (for $a = -1.7; b = 1.8; c = -1.9; d = -0.4$): & & & \\
+          % {$w_{n+1}= \begin{bmatrix} d\sin(aw_{n,x}) - \sin(bw_{n,y}) \\ 
+          %                                                             c\cos(aw_{n,x})+\cos(bw_{n,y}) \end{bmatrix}$} & & & \\
+          %   \multirow{3}{*}{$u_n = w_n-\overline{w}$}
+          %       & 10 & -0.2243 & 0.49698 \\
+          %       & 100 & &  0.63310 \\
+          %       & 1000 & & 0.71151 \\
+          % \hline
           \hline
@@
 
-      \caption{Multidimensional Correlation Coefficients $\rho$ to indicate a general reduction in the functional complexity of the map $G_T$ that arises from a RNN.
-Each row corresponds to a different dynamical system used in the experiments in \cite{Main_article}.
-In the second column corresponds to the dimension of $X$ (number of artificial neurons) in the RNN used.
-      The numerical estimate of $\rho$ for the relevant vectors is plotted in the last two columns.
-The reader can compare the linear relationship between $u_n$ vs $T(u_{n})$ and the linear relationship $\begin{bmatrix}
+      \caption{Following \cite{manjunath2021universal}, we determine the multidimensional correlation coefficients $\rho$ to evidence a general reduction in the functional complexity of the map $G_T$ emerging due to a RNN.
+Rows correspond to distinct dynamical systems considered in the experiments
+      The second column corresponds to the amount of artificial neurons (the dimension of $X$) in the implemented RNN.
+      The last two columns comprise numerical estimates of $\rho$ for the relevant vectors. The linear association between $u_n$ vs $T(u_{n})$, as well as the linear relationship $\begin{bmatrix}
               x_{n-1}\\
@@
               x_n
-          \end{bmatrix}\right)$.
-The dimension of the RNN from \eqref{Seq_RNN} is also given, showcasing that increasing the dimension of the driven system typically yields a map $G_T$ of lower functional complexity.} \label{Table_FC}
+          \end{bmatrix}\right)$ may be contrasted by the reader. 
+          Finally, the RNN's dimension (from \eqref{Seq_RNN}) is provided to exhibiting that an increase in dimension of the driven system will 
+          typically result in a map $G_T$ with less functional complexity.} 
+          \label{Table_FC}
       \end{table}\label{tbl_attractorsPearson}
 \end{center}
+\ednote{B: Functional Complexity will be explained before this}
 
@@
 So the relation $Y_T$, which is homeomorphic to the inverse-limit system, may be unique to a class of systems.
-(The reference is a paper on Ingram’s Conjecture by Marcy Barge, Henk Bruin and Sonja Stimac.
-) Further investigation is needed.
+(The reference is a paper on Ingram’s Conjecture by Marcy Barge, Henk Bruin and Sonja Stimac.) Further investigation is needed.
+
+\ednote{B: Expand}
 

### 2022-01-11 Gandhi_Project_Report/draft4.tex
Submitted to Gandhi for review. Still lots to be done

--- 2022-01-11 Gandhi_Project_Report/draft3.tex
@@
 \date{}
-\maketitle
-\tableofcontents
+% \maketitle
+% \tableofcontents
 
@@
 Instead of linear regression as in the training of echo state networks, deep learning methods are employed to learn the correct function.
-With slight modifications to the implementation in the paper (with Adriaan), we show that one can construct accurate non-interpretable models with the ability to reconstruct attractors from more hard-to-forecast systems like the double pendulum.
+With slight modifications to the implementation in the paper (\cite{manjunath2021universal}), we show that one can construct accurate non-interpretable models with the ability to reconstruct attractors from more hard-to-forecast systems like the double pendulum.
 (The forecasting of the time-series from a double pendulum has not been reported before.) Moreover, we also demonstrate that long-term statistical consistency is preserved.
@@
 Of course, discrete-time dynamical systems need not always arise through a differential equation, and once again we may consider the field of ecology where discrete dynamical systems are often directly derived or assumed.
-\textbf{Citation}
+\textbf{(Citation)}
 There is a school of thought that advocates discrete-time dynamical systems to be more natural for modelling real-world observations than differential equations and we refer the interested reader to \cite{saber2010introduction}.
@@
 
-A core concept in the study of dynamical systems is that of invariance. 
-\begin{Definition}
-  [\bf Invariant Set]\label{Dfn_InvariantSet}\rm
-  Given a discrete-time dynamical system $T: U \to U$, a subset $A \subset U$ is said to be an \emph{invariant set} if $T(A) =A$. 
-\end{Definition}
-
-We also define the orbit of a function $T$.
-\begin{Definition}
-  [\bf Orbit of $T$]\label{Dfn_Orbit}\rm
-  The orbit of $T$ is to be defined the sequence $\bar{u} = \{u_n\}_{n\in \mathbb{Z}}$ obeying the update equation, $u_{n+1}=Tu_n$, $n \in \mathbb{Z}$. 
-\end{Definition}
-
-Two examples of invariant sets include a fixed point where $Tu=u$ for  $u\in U$, and a periodic orbit, i.e., a set of iterates $\{u,Tu, T^2u,\ldots,T^pu\}$, where $T^{p+1}u=u$ for some $p\in\mathbb{Z}$.
-The entire space $U$ could be also be invariant.
-Consider for example the space $U=[0,1]$, where $Tu=4u(1-u)$ and then $U$ is invariant (as every $u\in{U}$ can be written as $u = 4x(1-x)$ for some $x\in{U}$).
-
-We may learn a great deal about the iterates of a dynamical system by considering the types of invariant sets of a discrete-time dynamical system.
-For example, if $U=[0,1]$ has map $Tu= u/2$, then the only invariant set is $\{0\}$, and all orbits approach this invariant set as time flows in the forward direction.
-Indeed, if a some non-zero invariant set (call it $B$) exists, then there is  some $r\in(0,1]\cap{B}$.
-But $r\notin{T(B)}$ since any orbit with initial value $r$ will be a decreasing sequence.
-Moreover, every orbit of T will be decreasing and therefore approach the value $0$ as $n\rightarrow\infty$.
-
-One may ask if every orbit approaches an invariant set?
-In general the answer is no, since for the dynamical system $T: \mathbb{R} \to \mathbb{R}$ defined by $Tu=2u$, any orbit that does not intersect the invariant set $\{0\}$, will not approach any invariant set.
-
-However,  when the space $U$ is compact, all orbits approach an invariant set.
-This is since the set of limit points of the orbit can be shown to be invariant \cite{de2013elements}.
-So, when $U$ is compact,  the $\omega$-limit set $\omega(u;T)$ of a point $u$ defined to the collection of limit points of the sequence $\{x,Tu,T^2u,\ldots\}$ is nonempty, and $\omega(u;T)$ is invariant.
-
-\begin{Example}
-  For the map $Tu=u^2$ defined on $[0,1]$ all orbits lie in the invariant set $\{1\}$ or else would would approach the invariant set $\{0\}$. 
-  % \ednote{B: not relevant other than being an additional example perhaps? M: Have added another simple example}.
-\end{Example}
-
-Invariant sets have various properties. Invariant sets can be attracting or repelling depending on how orbits in its vicinity behave. 
-Recall the examples $Tu =u/2$ and $Tu=2u$ defined on $\mathbb{R}$ (where $\{0\}$ ``attracts" orbits) or the example, $Tu=u^2$ defined on $[0,1]$ (where $\{1\}$ ``repels" orbits).
-We are interested in attractive invariant sets since they capture the long-term dynamics as time increases.
-In particular we are interested in those invariant sets named attractors.
-
-\begin{Definition}
-  [\bf Attractor]\label{Dfn_Attractor}\rm
-  Let $T: U \to U$, where $U$  is a metric space with metric $d$. A compact subset $A \subset U$ is said to be an attractor if it satisfies the three conditions: 
-  \vspace{-8mm}
-  \begin{enumerate}
-	\item $A$ is invariant. 
-	\item $A$ is asymptotically stable, i.e., for every $\epsilon > 0$ and for all $u$ so that $d(u,A) < \epsilon$, we have $d(T^nu,A) \to 0$ as $n\to \infty$. 
-	\item $A$ has Lyapunov stability, i.e., for every $\epsilon > 0$  there exists a $\delta(\epsilon) > 0$ so that $d(u,A) < \delta$ implies $d(T^nu,A) < \epsilon$ for all $n\ge 0$.  
-\end{enumerate}
-\end{Definition}
-
-Indeed in our previous examples, it can be verified that the system  $Tu=u/2$ defined on  $\mathbb{R}$ and $Tu=u^2$ defined on $[0,1]$  the singleton set $\{0\}$ is an attractor.
-For the system,  $Tu=1-|2u-1|$ on $[0,1]$, one may easily verify that the only attractor is the entire space $[0,1]$.
-This is since between any two points $u< v$ in $[0,1]$, and for any $a,b$ so that $u\le a < b \le v$, we can find an $n$ so that $T^n(a,b)=[0,1]$ (as would be explained later in this chapter).
-
-% \ednote{For let $B$ be an attractor of U.
-From above, we know that B is itself invariant, is asymptotically stable and has Lyapunov stability.
-Now suppose there is some attractor $C\subsetneq{U}=B$, so there is a $b\in{U}$ which is not in $C$.
-........
-M: Have changed the example}
-
-A dynamical system can have several attractors, and may also be contained in another attractor.
-For the example, $Tu =u^2$ on $[0,1]$, both $\{0\}$ and $[0,1]$ are attractors.
-It is a known result that for dynamical systems on a compact space there always exists an attractor.
-
-The dynamics restricted to an invariant set can be complicated.
-For instance an invariant set could be just a single point or it could have an infinite set.
-If the invariant set is infinite then there is a possibility of complicated dynamics, and a particular well-studied phenomenon of such complexity gives rise to what is called  chaotic behaviour, and has been studied in detail over the last fifty years.
-
-% If the dynamics on the attractor are somewhat complicated in the sense that the attractor cannot be decomposed further, and if the attractor is infinite then there is a possibility of complicated dynamics, and a particular well-studied phenomenon of such complexity gives rise to what is called a chaotic behaviour.
-
-%If the dynamics on an attractor cannot be decomposed further and the attractor is itself also infinite, then there is a possibility of complicated dynamics, and a particular well-studied phenomenon of such complexity gives rise to what is called chaotic behaviour.
-
 \section{Chaos}
 
-% \ednote{ should one add paragraphs like these?
-M: this is a thesis, and not a paper; the instructions sent to the external examiner regarding the thesis evaluation says that the student suggest strongly that it is good to write}
-In the 1960s, a number of mathematicians and mathematically interested scientists independently discovered chaos in the mathematical sense.
-The meteorologist Edward Lorenz may have been the first to explain this phenomenon in his 1963 paper \cite{lorenz1963deterministic}.
-The notions of invariance, attractivity and chaos may also be described for continuous systems, and Lorenz's system comprised of a system of differential equations.
-The narrative of Lorenz's discovery of chaos, as well as the history of other forerunners in this subject, are fascinating.
-For those who are interested, we highly recommend James Gleick's book Chaos: The Making of a New Science \cite{gleick2008chaos}.
-It clearly illustrates these experiences, explains why chaos was such a startling and crucial mathematical and scientific discovery, and describes the underlying mathematical notions for non-specialists.
-
-There are several intricately different definitions of chaos available in the literature \textbf{(Cite)}, with each of them indicating some aspect of complexity.
-%\ednote{B:  what are we saying here?
-M: Devaney's definition is not the only one, there are about half a dozen definitions as well}
-In practice, it is not possible to verify which definition or notion does the underlying dynamical system satisfies when only data is observed from it.
-We do, however, specifically recall the definition of chaos in the sense of Devaney \cite{devaney2018introduction,de2013elements} so as to understand some nuances behind the complexity.
-Devaney's definition of chaos has three requirements, with the first being the notion  of sensitive dependence on initial conditions.
-
-\begin{Definition}\rm
-  [\bf {Sensitive Dependence on Initial Conditions}]\label{Dfn_SDIC}\rm
-A dynamical system $T: U \to U$ is said to have sensitive dependence on initial conditions (SDIC) if there exists a $\delta > 0$ such that for every $u \in U$ and in every neighborhood of $u \in U$ there exists a $v\in{U}$ and an integer $N:=N{(u,v)}\in\mathbb{Z}$ such that $d(T^Nu,T^Nv)>\delta$.
-\end{Definition}
-
-It is important to acquire a sense of this definition as it can easily be misunderstood.
-It is common misconception to interpret SDIC as two close points ($u$ and $v$) that eventually become separated by a distance $\delta$ under iteration by $T$.
-But this is not true.
-To fully comprehend all the subtleties of the concept, one needs to discuss it more thoroughly by considering each sentence with care and attention, whereafter one may examine how they fit together to convey the concept of sensitive dependence.
-
-To this end, we make 3 remarks:
-\vspace{-5mm}
-\begin{enumerate}
-  \item First, the $\delta>0$ in the definition of SDIC is independent of $u$. 
-  \item Second, in every neighborhood of $u$, we may not necessarily find all points $v$ in the neighborhood distinct from $u$ that would separate from the forward iterates of $u$. 
-  \item  Finally, $N$ depends upon $u$ and $v$ chosen, and their iterates may not separate for ever (i.e., for all $n>N$) and we allow their iterates to get arbitrarily close in the future. 
-\end{enumerate}
-
-To illustrate the concept of SDIC, we present two examples - one from Mathematics and the other from the field of Physics.
-
-\begin{Example}
-  The logistic map(LM), a recursion relation of the form $x_{n+1}=rx_n(1-x_n)$ where $x_n\in[0,1]$ and $r<4$ is of particular interest to us as it is classically used to illustrate the influence of chaos on a system \textbf{(Cite)}.
-Below are plotted the first 75 iterates of the LM for $r=3.7$ and $x_0=0.2$ with an adjustment of 0.001 in $x_0$ and then $r$.
-
-  \begin{figure}[ht]
-    \includegraphics[scale=0.45]{_log_sdic.eps}
-        \centering
-        \label{fig:log_sdic}
-        \captionof{figure}{Logistic Map for $r=3.7$, $x_0=0.2$}
-      \end{figure}
-
-  For an intuitive sense of the fact that the LM exhibits SDIC, we draw two green neighbourhoods of diameters $\delta_1=0.1, \delta_2=0.02$ around the points $u_1=0.7, u_2=0.4$ respectively.
-It is easy to see that every neighbourhood around these two points will contain some $v_1, v_2\in[0,1]$ and natural number $N_1, N_2 \in\mathbb{N}$ respectively s.t.
-the distance $d(T^Nu_1,T^Nv)>\delta_1=0.02$ and $d(T^Nu_2,T^Nv)>\delta_2=0.1$.
-
-    \ednote{B: Is it desirable to have a more rigourous explanation here? }
-
-\end{Example}
-
-\begin{Example}
-  A second example pertains to a simple physical system called the double pendulum - a pendulum with another pendulum attached to its head given an initial position and angular velociy.
-
-  Below are two figures denoting the trajectories of a double pendulum's second head after some elapsed time.
-Depicted are three pendulum heads with equal angular velocity, but differing ever so slightly in initial position .
-As can be seen below, the trajectories diverge very quickly to become completely different.
-  The double pendulum will be discussed in much greater detail in a later chapter (\ref{ch5}), but we mention the example here to provide a numerical example of a practical system exhibiting SDIC.
-
-\begin{figure}[ht]
-  \centering
-\includegraphics[scale=0.8]{_dp_sdic.eps}
-    \captionof{figure}{Three double pendulum heads with initial angle differing by 0,025 radians}\label{fig:dp_sdic}
-  \end{figure}
-
-\end{Example}
-
-The second part in Devaney's definition of chaos concerns topological transitivity.
-
-\begin{Definition}
-  [\bf {Topologically Transitive}]\label{Dfn_TopolTrans}\rm
-	A dynamical system $F: U \to U$ is topologically transitive if for any pair of nonempty open sets $E_1$ and $E_2$ there exists a $n\in\mathbb{N}$ such that $T^n(E_1) \cap E_2 \not= \emptyset$. 
-\end{Definition}
-
-Topologically transitivity implies that the iterates of an open set of initial conditions gets mixed up with other open sets.
-On a compact metric space, it can be shown that topological transitivity also implies the existence of a point whose forward iterates are dense \cite{de2013elements}; or in other words, the orbit going through this point will be dense in the compact metric space.
-In fact, it is topologically more likely that the choice of an arbitrary point will be one  whose iterates are almost dense.
-
-\begin{Definition}
-  [\bf {Meager Set}]\label{Dfn_Meager Set}\rm
-A subset of a topological space $U$ is said to be a meager set if it can be written as a countable union of sets of with empty interior. The complement of the a meager set is set to be a residual set.
-\end{Definition}
-
-To be more precise (see \cite{de2013elements}), for a discrete-time dynamical system on a compact space, the set of points with dense iterates are residual, and they are typical or likely to be observed in practice.
-In this thesis, we assume that when data is observed from a topologically transitive system, we assume that the data arises from a dense orbit.
-
-We may now define the notion of Chaos as formulated by Devaney\cite{devaney2018introduction}.
-\begin{Definition}
-  [\bf {Devaney's Chaos}]\label{Dfn_ChaosDec}\rm
-	A dynamical system $T: U \to U$ is said to exhibit chaos in the sense of Devaney if it satisfies the three properties:
-	\vspace{-5mm}
-  \begin{enumerate}
-		\item $T$ has SDIC.
-		\item $T$ is topologically transitive.
-		\item The set of periodic points of $T$ are dense in $U$. 
-	\end{enumerate}
-\end{Definition}
-
-\begin{Example} 
-  The standard tent map $Tu=1-|2u-1|$ defined on $[0,1]$ is a well-known example of a dynamical system satisfying these three properties.
-
-  We now reason as to why this is true.
-The graph of the map $T$ is piecewise linear with two straight lines, one connecting the points $(0,0)$ and $(\frac{1}{2},1)$ and the other connecting $(\frac{1}{2},1)$ with $(1,0)$.
-They form a so-called tent with base centered at $u=1/2$.
-The graph of the map $T^2$ comprises two symmetric tents with their base centered at $1/4$ and $3/4$.
-See Figure \ref{fig:T2tentmap}
-
-  \begin{figure}[ht]
-    \includegraphics[scale=0.6]{_tentmap_2.eps}
-        \centering
-        \captionof{figure}{Three double pendulum heads with initial angle differing by 0,025 radians}\label{fig:T2tentmap}
-  \end{figure}
-
-  In general, the graph of  $T^k$ contains $2^{k-1}$ tents.
-  Given any point $u$ and an open neighborhood $W \subset[0,1]$ of $u$, we can find $k\in\mathbb{Z}$ large enough so that $T^k(W) = [0,1]$ because we can accommodate a tent whose base is contained in $W$.
-% \ednote{B: This last phrase doesn't make sense to me.
-M: The number of tents grow with $k$ and hence this is obvious; I explained in one of our meetings.
-Does it make sense?}
-  This implies the existence of two points $w_1$ and $w_2$ in $W$ so that $d(T^kw_1,T^kw_2)= 1$, and by the triangle inequality  at least one of the inequalities $d(T^ku,T^kw_1)> \delta$ or $d(T^ku,T^kw_2)> \delta$ hold when $\delta\in (0,\frac{1}{2})$.
-So $T$ has sensitive dependence on initial conditions.
-
-  Next, let $W_1$ and $W_2$ be two nonempty open sets.
-Given any open set $W_1$, we can find a $k\in\mathbb{N}$ large enough so that $T^k(W_1) = [0,1]$ because we can accommodate a tent with base contained in $W$.
-So $T^k(W_1) \cap W_2 \not=\emptyset$, and thus $T$ has topological transitivity.
-
-  Finally, for every open interval $(a,b)$, the graph of $T^k$ intersects the graph of the identity map on $[0,1]$.
-This follows from the fact established above that there exists a $k\in\mathbb{N}$ such that $T^k(a,b) = [0,1]$.If the intersection point has coordinates of the form $(p,p)$, the point $p$ is then a fixed point of $T^k$ and therefore also a periodic point of $T$.
-Since we have found a periodic point in the interval $W$,  the set of periodic points are dense in $T$.
-  This follows because any fixed point of $T^k$ is also a periodic point, and we have found this in an arbitrary interval.% \ednote{Expand further on reason for density.
-M: fixed point of $T^k$ is a periodic point, and we have found in an arbitrary interval.}
-
-\end{Example}
-
 \section{Conjugacy}
-
-We now turn our attention to the subject of conjugacy that describes when two dynamical systems are equivalent dynamically.
-
-To show topological similarity or sameness between two metric or topological spaces, one needs to establish a homeomorphism between the two spaces. 
-However, in the study of dynamical systems defined on two spaces, establishing a homeomorphism does not indicate that the systems are dynamically related in any way.
-For instance, the maps $Tu=u^2$ and $Tu=1-|2u-1|$ defined on $[0,1]$ have totally different behaviour.
-To find systems that are dynamically similar, one rather needs to establish a kind of dynamical equivalence which we illustrate in a simple commutativity diagram below.
-
-% Giving errors
-\begin{equation}  \label{eqn_conjugacy}
-%    \[ 
-    \psset{arrows=->, arrowinset=0.25, linewidth=0.6pt, nodesep=3pt, labelsep=2pt, rowsep=0.7cm, colsep = 1.1cm, shortput =tablr}
- \everypsbox{\scriptstyle}
- \begin{psmatrix}
-U & U\\%
-V & V.
- %%%
-%  \ncline{1,1}{1,2}^{T} 
-%  \ncline{1,1}{2,1} <{\phi}
-%   \ncline{2,1}{2,2}^{S}
-%  \ncline{1,2}{2,2} > {\phi}
- \end{psmatrix}
-% \]
-\end{equation}
-
-%Composite functions
-To understand the diagram we note that if we travel ``right and then down," the diagram instructs us to use $T$ (top arrow right) first, followed by $\phi$.
-(right arrow downwards).
-Consequently the pathway amounts to finding $\phi(T(u))$.
-If we proceed "down and then right", the diagram instructs us to apply $\phi$ (left arrow downwards) first, and then apply $S$ (bottom arrow right).
-This then amounts to finding  $S(\phi(u))$.
-When the relationships in this diagram hold, we say $\phi(T(u))= S(\phi(u))$, and we formally denote it as $\phi \circ T=S\circ \phi$.
-
-\begin{Definition}\rm  
-  [\bf {Conjugacy \& Semi-Conjugacy}]\label{Dfn_Conjugate}\rm
-  Consider the dynamical systems $T:U\to{U}$, $S:V\to{V}$ and suppose the relationship $\phi \circ T=S\circ \phi$ holds.
-If $\phi:U\to{V}$ is a homeomorphism, then $S$ is said to be conjugate to $T$ and $\phi$ is called a conjugacy.
-If we relax the criterion on $\phi$ and merely require $\phi$ to be continuous and surjective where $\phi:U\to{V}$, then $\phi$ is a semi-conjugacy between $T$ and $S$ where T has domain $U$ and S has domain $V$; S is said to be semi-conjugate to T.
-\end{Definition}
-
-\ednote{M: Saying $\phi$ is then a semi-conjugacy between $T$ and $S$ could be ambiguous as it does not specify which system is on the top in the commutativity.
-So please edit accordingly.
-B: Does defining explicitly the domain and codomain of $\phi$ make it clear enough?
-}
-When $S$ is conjugate to $T$, the dynamics of the two systems are in some way ''dynamically equivalent''.
-Specifically, they are in one-to-one correspondence with one another.
-However when $S$ is semi-conjugate to $T$ with $\phi:U\to{V}$ a many-to-one mapping, the dynamics on $V$ provide merely a coarse-grained description of the dynamics on $U$ \cite{de2013elements}.
-When $S$ is semi-conjugate to $T$, it is also common to call $S$ a \emph{factor of $T$}, or conversely that $T$ is an \emph{extension of $S$}.
-In essence, an extension (e.g., \cite{de2013elements}) is a larger system capturing all of the important dynamics of its factor.
-
-It is a very hard or nearly an impossible task to establish the existence of a conjugacy or a semi-conjugacy $\phi$ between two systems \cite{devaney2018introduction}.
-However, one can verify that a function $\phi$ satisfies the commutativity diagram \ref{eqn_conjugacy}.
-
-\textbf{NB NB FINISH}
-\begin{Example}
-  For example $\phi(x)=\sin(\frac{\pi}{2}x)^2$ is a conjugacy between two systems $Tu=1-|2u-1|$ and $Sv=4v(1-v)$  defined on $U=[0,1]$.
-
-  To see this, we consider $\phi\circ{T} = \sin(\frac{\pi}{2}(1-|2u-1|))^2$
-
-  and $S\circ\phi = 4\sin(\frac{\pi}{2}x)^{2}(1-\sin(\frac{\pi}{2}x)^2)$
-\end{Example}
-
-In establishing that two systems are conjugate (semi-conjugate) to one another, we have also shown that one may choose to work with one system as opposed to the second and still be guaranteed to obtain information on the latter.
-This is especially useful while forecasting the future evolution of dynamical systems.
 
@@
 First we establish the method of Takens delay embedding. 
-In this method, one considers a discrete-time dynamical system defined on a 'nice' space (a smooth manifold)\ednote{Should this be further explained?} that can be obtained as time-$K$ map of a continuous time-dynamical system, concepts which will be formalised below.
+In this method, one considers a discrete-time dynamical system defined on a 'nice' space (a smooth manifold) that can be obtained as time-$K$ map of a continuous time-dynamical system, concepts which will be formalised below.
 We make observations from such a system, i.e., we observe the evolution of $\theta(w_0)$ (where $w_0$ represents the initial state) i.e., we observe  a finite set of values of $u_n :=  \theta(w_n) = \theta(Tw_{n-1})$.
@@
 In almost all cases we can at most insert probes into a system to obtain partial information by means of the measurements taken.
-Moreover, the process of taking a measurement itself introduces 2 additional aspects which complicate the problem: \ednote{M: Maybe replace difficulty by saying that we need to consider two aspects.
-B: Perhaps this edit makes more sense?}
+Moreover, the process of taking a measurement itself introduces 2 additional aspects which complicate the problem:
 \vspace{-8mm}
@@
 
-\section{Taken Embedding Theorem}
+\section{Takens Embedding Theorem}
 
-We define first the concepts of homeomorphism and embedding:
-\begin{Definition}\rm
-  [\bf {Homeomorphism}]\label{Dfn_homeo}\rm
-  A homeomorphism is a function $f:Z\rightarrow Y$ between two topological spaces $Z$ and $Y$ that is continuous, bijective and has a continuous inverse. 
-\end{Definition}
+% \section{Practical Issues of Takens theorem}
 
-\begin{Definition}
-  [\bf {Embedding}]\label{Dfn_embed}\rm
-  Consider a homeomorphism $f:Z\rightarrow Y$ for $Y\subset X$. $Z$ is said to be embedded in $X$ by $f$.
-\end{Definition}
+% \chapter{Driven Dynamical Systems to Forecast Problems}\label{ch4}
 
-Takens Theorem states a result establishing a relationship between the observed and underlying dynamical systems by showing that the concatenation of a sufficiently large number of previous observations into a vector will, under certain conditions, generate a map between the vectors from the respective systems.
-We formulate the theorem from \cite{takens1981detecting}.
+% In this chapter, we discuss results pertaining to the mapping of temporal data obtained from a discrete-dynamical system onto a different space through the notion of a driven dynamical system. 
+% We also consider the conditions a driven system should possess to avoid adding distortion to its state-space representation.
+We then describe the \emph{single-delay dynamics} (SDD) of the system and and finally recall conditions on the driven system so that the SDD are conjugate (or at least semi-conjugate) to the underlying system.
+% The SDD can then be used to forecast and reconstruct the underlying system. 
+% % \ednote{Is this attractor assumed to exist? M:  I replaced  it with the system. Attractor is also a commonly used term since we observe data from an attractor in almost all cases.}
 
-\begin{Theorem} 
-	[\bf Takens Embedding Theorem (adopted from \cite{takens1981detecting}] \label{Thm_Takens}
-         Let $W$ be a compact manifold of dimension $m$, and $d\ge m$ so that $2d$ is an integer. It is a 
-            generic property for the pair $(T, \theta)$,  where $T:W \to W$ is
-            a smooth diffeomorphism, and $\theta:W \to \mathbb{R}$ a smooth function, the map $\Phi_{2d,\theta}:W \to \mathbb{R}^{2d+1}$ defined on $W$ by 
-            $\Phi_{2d,\theta}(w) := (\theta(T^{-2d}w)\ldots,\theta(T^{-1}w),\theta(w))$
-            is a diffeomorphic embedding; by `smooth' we mean at least $C^2$.
-Consequently, there exists a map $F_\theta: \Phi_{2d,\theta}(W) \to \Phi_{2d,\theta}(W)$ defined by $$F_\theta: (\theta(T^{-2d}w),\ldots,\theta(T^{-1}w),\theta(w)) \mapsto
-            (\theta(T^{-2d+1}w),\ldots,\theta(w),\theta(Tw))$$
-           so that $(W,T)$ is topologically conjugate to 
-            $(\Phi_{2d,\theta}(W), F_\theta)$.    
-\end{Theorem}
+% \section{Nonautonomous and Driven Dynamical Systems}
 
-By generic we mean a residual set on a certain topology on appropriate spaces of functions (that we do not describe here) \textbf{(Cite)}.
-%\ednote{How vague may one be in general by leaving things out?}
-By $C^2$, we make reference to a twice-differentiable function  with a continuous second derivative.
-In our scenario (and this remains important throughout) the input space $U$ is considered the attractor.~\label{attractor_U}
+% \section {Unique Solution Property}
 
-Below we reproduce Figure 1 from \cite{Supp}.
+% % \section{Next section?}
 
-%Some explanation is beneficial.
-Takens establishes a delay-coordinate map $\Ftheta$ defined in \ref{eqn_takens} for $T$ the flow (\ref{defn_flow}) defined by the update equation  $Tu_n=u_{n+1}$ and $theta$ our measurement function.
-When confining a dynamical system to the manifold U (in which the underlying system is contained), Takens showed that if certain smoothness conditions are satisfied on $T$ and $\theta$, then the delay-coordinate map $\Ftheta$ embeds $U$ in the reconstruction space $R^{2d+1}$ \textbf{we haven't used the term reconstruction space yet, so this is falling out of the blue}, for almost every choice of measurement function $\theta$, the observable.
-By almost every choice here, we mean that we exclude some observations which do not give any information.
-To see a more precise discussion, see \textbf{xxx} for we do not concern ourselves with a further elaboration on this report.
+% \section{Choosing the driven system $g$}
 
-%Alternatively we may also make the statement: Takens' Embedding Theorem guarantees that almost all dynamical systems can be reconstructed from just one noiseless observation sequence \textbf{this isn't obvious from the theorem's formulation, is it?} i.e.
-for a great number of possible observation functions $\theta$, $\Ftheta$ preserves the topology of U.
-\textbf{'preserving the topology' is not something we've touched on before.
-Perhaps I should add to previous discussion so logic flows seamlessly here}.
+% \section{The next step in Dynamics}
 
-\begin{figure}[ht]
-  \includegraphics[scale=0.25]{_takensmap.png}
-  \centering
-  \captionof{figure}{Schematic of Takens Embedding Theorem's usage and consequence of how iterates \ldots }
-  \label{fig:takensmap}
-\end{figure}
-\ednote{M: Here, and throughout if you reproduce a figure, it must be stated where it was reproduced from. B: Is this satisfactory? Also, should $\tilde{F}$ be left out for now?}
-\ednote{B: This is only a screenshot used as a temporary placeholder. If sufficient, I shall convert it to .eps. If not, may I perhaps have the original picture?}
+% % \section*{Summarising the discussing thus far:}
 
-If we can learn the map $F_\theta$, from sufficient number of data points of $\{\phi_{2d,\theta}(w_n)\}$, then we also know how to forecast how $\theta(w_n)$.
+% \section{A discussion of $G_T$ }
 
-%Once again we pause and consider the graph to consolidate our understanding of the conjugacy (\textbf{are we being too simplistic?}).
-If we are in state $u$ and then evolve towards state $Tu$ before embedding into $\mathbb{R}^{2m-1}$, then this is equivalent to first embedding into $\mathbb{R}^{2m-1}$ and then evolving through the map $\Ftheta$, i.e.
-$\Phi_{2d,\theta}\circ{F}_{\theta} = T\circ\Phi_{2d,\theta}$.
-The map $F_theta$ is a homeomorphism as well.
+% \section{Advantages of learning $\Gamma$} \label{sect_LearnGamma}
 
-%Thus information about U can be retained in the time series' (or observation's) output.
-By preserving the topology on the manifold $U$ in the reconstruction space $X$ (\textbf{not discussed before}), we are also guaranteed the preservation of topological invariants of the manifold, of which dimensionality is one such invariant.
-(We note here that dimensionality will again be considered later in this project).
+% \chapter{Implementation} \label{ch5}
 
-\section{Practical Issues of Takens theorem}
+% In this chapter we turn to the implementation of the function $G_T$ by discussing a method to learn the function $\Gamma$.
+We then indicate where this implementation differs from previous work done by (\cite{manjunath2021universal}) and empirically show that the SDD of $G_T$ are less functionally complex than $\Phi_{k, \theta}$ or $T$ by considering the Pearson correlation coefficient.
+% Finally, we present and discuss some numerical results for a focus system of this project: the double pendulum. We also consider other chaotic attractors.
 
-It is a fact that Takens' Embedding Theorem is both a powerful result and that it provides compelling reason to believe that one could conceivably reconstruct accurately a system conjugate to the underlying system.
-Nonetheless, it does present some serious practical limitations:
-\vspace{-5mm}
-\begin{enumerate}
-\item Even supposing that we could indeed find $F_\theta$ , our \emph{approximation} of  $F_\theta$ is a map from a larger set $\mathbb{R}^{2d+1}$ containing the embedded attractor.
-There are, however, no theoretical guarantees that $F_\theta$ will retain  $V=\Phi_{2d,\theta}(W)$  as an attractor although $W$ could itself be an attractor.
-\item Takens theorem is stated only for noiseless observations.
-Due to noise $\epsilon_n$ the delay vector $\Phi_{2d,\theta}(w_n) + \epsilon_n$  may lie outside $V$.Furthermore, due to the chaotic nature of the underlying system (i.e.
-the fact that it has SIDC), the evolution of $\Phi_{2d,\theta}(w_n) + \epsilon_n$ under the map $F_\theta$ could move out of $V$ completely.
-This problem can be overcome  by using a driven dynamical system with some properties, and we discuss this in Chapter\ref{ch3}.
-\end{enumerate}
+% \section{Implementing $G_T$}
 
-% Another limitation owes itself to the requirement that T be a surjective map.
-This is not always the case in practical examples and we may consider the case of a damped double pendulum \ednote{M: Please remove the paragraph as well.
-We do not overcome this problem.
-You can add this in the next Chapter where you need T to be surjective or else you could include this at the begining of this chapter where you explain the forecasting problem.}
+% \section{Changes in this project to previous work of~\cite{manjunath2021universal}}
 
-The Takens embedding theorem does not guarantee global dissipativity.
-\ednote{B:Links with~\ref{subs_LearnGamma} in discussion of dissipativity of $\Gamma$.
-Sir, I'm not sure how to write this here - I think it a beneficial point to make so as to set the stage for the later discussion in ~\ref{subs_LearnGamma}, but I am unsure there as well.}
+% \section{Numerical simulations}
+% % In this section we showcase numerical results pertaining to 3 attractors that were simulated using the methodology described in the previous chapters. We first consider a simple physical system.
+% \subsection{Double Pendulum}
+% \subsection{Alternative Attractors}
+% \subsubsection{Thomas' Cyclically Symmetric Strange Attractor}
+% \subsubsection{Fractal Dream Attractor}
+% \section{Generalized Pearson Correlation} %\ednote{B: Alter heading.}
 
-% \begin{figure}[ht]
-%   \includegraphics[scale=0.3]{_learningfailure.png}
-%   \centering
-%   \caption{Illustration of a system having failed to learn the map}
-%   \label{fig:supp_learning_failure}
-% \end{figure}
+% \chapter{Conclusion}\label{ch6}
 
-\chapter{Driven Dynamical Systems to Forecast Problems}\label{ch4}
-
-In this chapter, we discuss results pertaining to the mapping of temporal data obtained from a discrete-dynamical system onto a different space through the notion of a driven dynamical system. 
-We also consider the conditions a driven system should possess to avoid adding distortion to its state-space representation.
-We then describe the \emph{single-delay dynamics} (SDD) of the system and and finally recall conditions on the driven system so that the SDD are conjugate (or at least semi-conjugate) to the underlying system.
-The SDD can then be used to forecast and reconstruct the underlying system. 
-% \ednote{Is this attractor assumed to exist? M:  I replaced  it with the system. Attractor is also a commonly used term since we observe data from an attractor in almost all cases.}
-
-\section{Nonautonomous and Driven Dynamical Systems}
-
-\begin{Definition}
-  [\bf Nonautonomous Dynamical System (adopted from~\cite{Manju_ESP}))] \label{Dfn_NDS}\rm
-  A nonautonomous dynamical system (NDS) on a space $X$ is simply a dynamical system comprising a family of maps $\{g_n\}_{n \in \mathbb{Z}}$ where $g_n(\ldotp):=g(u_n,\ldotp):X\to{X}$ is continuous.
-Each $g_n$ arises as a consequence of an input $u_n$ from the input space $U$, a topological space.
-\end{Definition}
-
-We immediately recall the figure first defined in \eqref{eqn_conjugacy}  and remind ourselves of our ultimate goal to learn a system that is conjugate to an underlying or unknown dynamical system $(U,T)$ by making use of the measurements obtained from  the unknown system.
-
-%#####################################
-\begin{center}
-\psset{arrows=->, arrowinset=0.25, linewidth=0.6pt, nodesep=3pt, labelsep=2pt, rowsep=0.7cm, colsep = 1.1cm, shortput =tablr}
-\everypsbox{\scriptstyle}
-\begin{psmatrix}
-U & U\\%
-V & V.
-%%%
-%  \ncline{1,1}{1,2}^{T} 
-%  \ncline{1,1}{2,1} <{\phi}
-%   \ncline{2,1}{2,2}^{S}
-%  \ncline{1,2}{2,2} > {\phi}
-\end{psmatrix}
-%#####################################
-\end{center}
-
-To find a function $\phi$, the need now arises to consider a so-called driven dynamical system, a special case of a nonautonomous dynamical system, where two spaces are considered: input is taken from an input space $U$ and the state of a space $X$ is updated.
-(Previously, only the state was considered.)
-We do this to mimic the true conditions.
-By accounting for both external and internal factors which influence the  system-state at a specific time-step, our model resembles nature much more realistically.
-\ednote{M : I understand this sentence, but am not sure if a general audience would.
-So please elaborate or omit.
-B: Is this satisfactory?}
-
-\begin{Definition}
-  [\bf Driven, Compactly Driven Dynamical System] \label{Dfn_DDS} \rm
-A driven dynamical system comprises two topological metric spaces $U$, $X$ and a continuous function  $g:U\times{X}\to{X}$ where $g(u_n, x_n)=x_{n+1}$.
-If the input space $U$ is compact, we refer to the system as compactly driven. 
-\end{Definition}
-
-The dynamics on X are generated by the update equation $x_{n+1}=g(u_n, x_n)$ where $n\in\mathbb{Z}$, input $u_n$ from $U$ and state $x_n$ belonging to $X$, where $X$ is compact.
-Abbreviated, we shall refer only to the \emph{driven system $g$}, with all other entities being understood implicitly.
-
-Notably, a nonautonomous dynamical system may be generated from $U$; any input $\overline{u}$, a bi-infinite sequence from $U$, gives rise to the sequence of self-maps $\{g(u_n, \cdot)\}_{n\in\mathbb{Z}}$ contained in $X$.
-Physically, one may think of this bi-infinite sequence as referring to a system that has been running for a long period at the time of the first measurement taken from the system (alternatively, the first time a probe is inserted into the system to take an observation)
-
-\begin{Definition}
-  [\bf Entire Solution] \label{Dfn_Soln} \rm
-  A sequence $\{x_n\}_{n\in\mathbb{Z}}\subset X$ is called an entire solution(or simply a solution) of the driven system  $g$ with input $\overline{u}$ when it satisfies 
-  \[g(u_{n-1}, x_{n-1})=x_n\] for all $n\in\mathbb{Z}$
-\end{Definition}
-
-It is important to emphasise that a sequence $\{x_n\}$ satisfying the update equation above can only be a solution if $x_n\in{U}$ holds for all $n\in\mathbb{Z}$ and just for $n>0$.
-Consider the example below.
-
-\begin{Example} \rm \label{ex_halfux}
-  The only solution  $\{x_n\}_{n\in\mathbb{Z}}$ to the driven system  $g(u,x)=\frac{ux}{2}$, where $X=[0,1]$, $U=[0,1]$,  is the zero solution $x_n\equiv0$.
-  To see this, consider any $x_n=a\in[0,1]$ where $a\neq{0}$.  Let $\overline{u}\in{U}$ be an non-zero constant sequence , say $u_n=0.5$. 
-  The driven system may be rewritten as $x_{n-1}=\frac{2x_n}{u_{n-1}}=4x_n$ and the  iterates of $x_n$ in backward time will increase by a factor of 4 at each timestep. 
-  Thus for some $m\leq{n}$,  $1<x_m$ i.e. $x_m\notin{X}$. So ${\{x_n\}}_{n\in\mathbb{Z}}$ is not a solution and it follows that the only possible solution is the zero solution.
-\end{Example}
-% \ednote{Nicely written!}
-
-A system may also have multiple solutions as is evidenced in the example below.
-
-\begin{Example}\label{Ex_exp} \rm 
- Consider the driven system $g(u_n,x_n)=x^n$ for $X=[0,1]$, $U=\mathbb{R}$.
-The system has an uncountable number of solutions, as there exists a solution for every $x\in{X}$ which also passes through the point $x$ and $lim_{n\to\infty}x_n=0$, $lim_{n\to{-}\infty}x_n=1$.
- The proof is deferred to immediately after the next paragraph (see~\ref{rem_proofEx}).
-\end{Example}
-
-As the solutions to a driven system are often considered to be an important entity \cite{M: edited the sentence here}, we next identify a subspace $X_U$ of $X$ that contains all possible solutions.
-To realize such a subspace of a driven system $g$, the concept of a reachable set is defined.
-
-\begin{Definition}
-  [\bf Reachable Set]\label{Dfn_ReachableSet}\rm
-The reachable set of a driven system $g$ is exactly the union of all the elements of all the (entire) solutions, i.e., 
-\[X_U :=\Big \{x \in X:  x = x_k \mbox{ where  $\{x_n\}$  is a solution for some  $\bar{u}$} \Big \}.\]
-The set of all reachable states at a specific time $n$ for input $\overline{u}$ is denoted by $X_n(\overline{u})$
-\end{Definition}
-
-The reachable set itself can defined independently from the fact that $g$ does or does not have the ESP/USP.
-We note that $x\in{X_n(\overline{u})}$ if and only $g$ has a solution $\{x_k\}$ for $x_n=x$ and input $\overline{u}$.
-\textbf{(Cite)}.
-
-$X_n$ is exactly the set of points $x$ through which some entire solution $\Psi$ passes at time $n$.
-
-This will lead to a result established later, but which is worth taking note of now: \textit{$g$ being a topological contraction is equivalent to the existence of a unique entire solution.}
-
-% \begin{Definition}
-%   [\bf Topological Contraction]\label{Dfn_TopContr}\rm
-%   A function $g:U\times{X}\to{X}$ is a topological contraction if for all $n\in\mathbb{Z}$ and all $\overline{u}\subset{U}$, $X_n(\overline{u})$ is a singleton subset of $X$.  
-% \end{Definition}
-
-\begin{Remark}
-  [\bf Proof of Example~\ref{Ex_exp}] \label{rem_proofEx} \rm
-  To see this, we show that for every input $\overline{u}=\{u_n\}_{n\in\mathbb{Z}}$, $g(u_n,\cdot)$ is a contraction map on $X$ for every $k\in\mathbb{Z}$. 
-  Indeed $g(u_n,x_n)=x^n$ depends only on the state at time $n$ and so for each input $\overline{u}$, $X_n(\overline{u})=\{x^n\}$ is a singleton subset.
-  We then easily conclude that $\{x_n\}\equiv0$ is the only solution. 
-\end{Remark} \ednote{B: Is this sufficient?}
-
-Thus far it has been demonstrated that a system may have one or more solutions; one may ask if a driven system always has a solution and, if so, whether it satisfies certain properties such as uniqueness.
-Should the driven system be compact, existence follows immediately as shown in the following result.
-
-\begin{Theorem}\label{Thm_CompactExistence}
- Let $g$ be a driven system.  If $X$ is compact then for each input $\overline{u}$, there exists at least one solution to the driven system $g(\ldotp, x)$
-\end{Theorem}
-\begin{proof}
-  May be found in \cite{kloeden2011nonautonomous, manjunath2014dynamics, manjunath2013echo}
-\end{proof}
-
-One may easily construct many systems with trivial solution-sets, such as $g(u,x)=x$ which has only the constant solution $x$ and so for $U=[-1,1]$, the system would have no solution if $|x|>1$.
-To refine the scenario, we consider only systems with unique solutions.
-
-\section {Unique Solution Property}
-
-\begin{Definition}
-  [\bf Unique Solution Property] \label{Dfn_usp}\rm
-  A driven system $g$ is said to have the Unique Solution Property (USP) if for each input $\overline{u}$ there exists exactly one solution. 
-  Alternatively we may formulate the USP as follows: $g$ has the Unique Solution Property if there exists a well-defined map $\Psi:{U}\to{X}$ with $\Psi({\overline{u}})$ denoting the unique solution.
-\end{Definition}
-
-One of the first result obtained upon defining the USP, is the fact that every solution will attract any different initial conditions towards the component parts of the solution.
-%, i.e.
-that the solution $\Psi$ is a non-autonomous uniform attractor.
-\ednote{Define Uniform Attractor before stating theorem or else omit the statement.}
-If $g$ has the USP, then any solution to $g$ is a uniform attractor in the nonautonomous dynamical systems literature \cite{Manju_Nonlinearity}.
-The discussion on nonautonomous attractors is beyond the scope of this thesis and we refer the reader to \cite{Manju_ESP, esann2012ids} for further reading.
-
-% \ednote{M: You can omit the theorem, and just say it is a uniform attractor in the nonautonomous dynamical systems literature \cite{Manju_Nonlinearity}, and say the discussion on nonautonomous attractors is beyond the scope of the thesis.}
-% Paraphrased.
- In our context, $g$ having the USP is equivalent to $g$ being a topological contraction \cite{manjunath2021universal}, or alternatively that $g$ exhibits the Echo State Property (ESP).
-
-This property is used to describe the concept of fading memory or memory-loss in the majority of literature within the field reservoir computing (RC) \textbf{(Citation)}. 
-$g$ having the ESP guarantees that the entire left-infinite history of an input will determine exactly the current system state; there is only ever the possibility of a single reachable state at a given instance of time \cite{jaeger2001echo}
-The ESP is also often discussed in terms of a stability property that holds for all inputs of a system~\cite{manjunath2020stability} and, furthermore, plays a key role in the design and training of recurrent neural networks within the field of RC \cite{Manju_ESP}.
-
-Defined formally, the ESP is intrinsically linked to an input (see \cite{Manju_ESP}) Unfortunately, definitions and results concerning the ESP do not have much practical use since a relationship of the network's response to the temporo-statistical properties of the input is not directly described.
-A more detailed presentation may be found in~\cite{jaeger2001echo}.
-
-\section{Next section?}
-\ednote{I feel a break is necessary here, but would hear from you what your thoughts are.}
-
-Having already defined the reachable set $X_U$, we now pause for a moment to fix additional notation.
-Letting $\cev{u}^{n}:=(\ldots,u_{n-2} ,u_{n-1})$ be the left-infinite subsequence of an input, $\overleftarrow{U}$ then denotes all these left-infinite sequences in $U$. 
-Moreover, $\cev{u}^{n}v:=(\ldots,u_{n-2} ,u_{n-1}, v)$ is to symbolise the input up to time $n$ with $v \in U$ being the specific input value at time $n$. 
-The introduction of a new input at time $n$ can be described by a mapping $\sigma_v:   \cev{u}^{n} \mapsto \cev{u}^{n}v$.
-
-The question now becomes whether we may establish a semi-conjugacy %\ednote{M: I have changed it to a semi-conjugacy since conjugacy amounts to asking a infinite-dimensional space $\overleftarrow{U}$ to be embedded in $X_U$ which is not possible.}
-as presented below for the driven system $g$.
-
-\begin{equation}  \label{Scomm_h}
-  %    \[ 
-      \psset{arrows=->, arrowinset=0.25, linewidth=0.6pt, nodesep=3pt, labelsep=2pt, rowsep=0.7cm, colsep = 1.1cm, shortput =tablr}
-   \everypsbox{\scriptstyle}
-   \begin{psmatrix}
-   \overleftarrow{U} & \overleftarrow{U}\\%
-   X_U & X_U.
-   %%%
-  %  \ncline{1,1}{1,2}^{\sigma_v} \ncline{1,1}{2,1} <{h}
-  %  \ncline{1,2}{2,2} > {h}
-  %  \ncline{2,1}{2,2}^{g(v,\cdot)}
-   \end{psmatrix}
-  % \]
-  \end{equation}
-
-Continuing to refine our attention, we proceed to consider a specific subclass of conjugacies.
-
-  \begin{Definition} 
-    [\bf Universal Semi-Conjugacy]\label{Def_UnivSemiConj}
-    Given a driven system $g$, we  call a continuous and surjective map $h : \overleftarrow{U} \to X_U$ a universal semi-conjugacy if  diagram \ref{Scomm_h} commutes for all $v \in U$.
-  \end{Definition}
-
-If the universal semi-conjugacy $h$ exists (i.e. the diagram in \ref{Scomm_h} commutes) then the solution $\Psi$ will be at least a coarse-grained representation of the input $u$.
-
-Does such a function $h$ for the driven system defined above exist?
-If $g$ has the USP and $\Psi(u)=\{x_n\}_{n\in\mathbb{Z}}$ then $h$, defined by  $h(\cev{u}_n):=g(u_n,x_{n-1})=x_n$, will satisfy the semi-conjugacy in the graph above \ref{Scomm_h}.
-Regrettably, such a continuous mapping $h$ is not guaranteed to exist \cite{M: Have edited here} when $g$ does not have the USP \cite[Lemma 5]{Manju_Nonlinearity}.
-Note that even when $h$ does exist, we are not guaranteed its injectivity. Considering again example \ref{ex_halfux}, we see that even if $h$ were to exist, it could not be injective as $X_U=\{0\}$.
-
-Re-sketching the commutativity diagram \ref{Scomm_h} above by fixing the input $v$ in $g$ and replacing $X_U$ by its left-infinite sequence space $\overleftarrow{U}$, we obtain the diagram below.
-In this case, the function $H:\overleftarrow{U}\to\overleftarrow{X}_U$, a map that is both continuous and surjective, is called a \emph{causal mapping}.
-
-\begin{Definition}
-  [\bf Causal Mapping]\label{Def_CausMap}
-  A continuous, surjective map $H:\overleftarrow{U}\to\overleftarrow{X}_U$ such that \[H\circ\tilde{g}_v=\sigma\circ{H}\] where $\tilde{g}_v$ maps $(\ldots, u_{-2}, u_{-1})$ to $(\ldots, u_{-2}, u_{-1}, g(v, u_{-1}))$.
-\end{Definition}
-
-\begin{equation} \label{SCausal_H}
-    %    \[ 
-        \psset{arrows=->, arrowinset=0.25, linewidth=0.6pt, nodesep=3pt, labelsep=2pt, rowsep=0.7cm, colsep = 1.1cm, shortput =tablr}
-     \everypsbox{\scriptstyle}
-     \begin{psmatrix}
-     \overleftarrow{U} & \overleftarrow{U}\\%
-     \overleftarrow{X}_U & \overleftarrow{X}_U.
-     %%%
-    %  \ncline{1,1}{1,2}^{\sigma_v} \ncline{1,1}{2,1} <{h}
-    %  \ncline{1,2}{2,2} > {h}
-    %  \ncline{2,1}{2,2}^{\tilde{g}_v}
-     \end{psmatrix}
-    % \]
-  \end{equation}
-
- \begin{Theorem}
-  For a compactly driven system, a causal mapping $H$ exists if and only if $g$ has the USP. 
-\end{Theorem}
-\begin{proof}
-  See~\cite[Th.3]{manjunath2013echo}.
-\end{proof}
-
-We direct our attention to the function \[H_2{\overline{}} := (h(r\overline{u}, h(\overline{u}))) \]
-
-An embedding of the space $\overleftarrow{U}$  in $\overleftarrow{U}$ would allow one to establish topological conjugacy, which in turn provides stronger results than merely obtaining a coarse-grained representation via a causal mapping \cite{M: I feel this is very confusing to the reader as the co-domain of causal mapping is the left-infinite space and , so please omit if possible.
-You can just say we want to restrict inputs to the left-infinite orbits of a dynamical system and then explore conjugacy instead of a mere semi-conjugacy}.
-
-Before formalising this, we make mention of the concept of the inverse-limit system of a dynamical system $(U,T)$. 
-These spaces are noteworthy as their topology can help one ascertain much about the original system \cite{Manju_IEEE}.
-The topological structure for inverse-limit systems of chaotic dynamical systems is complicated, so we relegate further discussion to \cite{kennedy2008inverse_limit, ingram2011inverse}.
-In broad sweeps, the inverse-limit space of a dynamical system is a self-map on a subset of an infinite-dimensional space where each point in the inverse-limit space corresponds to a backward orbit of the map $T$.
-Following \cite{manjunath2021universal, ingram2011inverse}:
-
-\begin{Definition}
-  [\bf Inverse-Limit Space, Inverse-Limit System]\label{Dfn_InverseL}\rm
- The subset $\widehat{U}_T\subset \overleftarrow{U}$ defined by \[\widehat{U}_T: = \{ (\ldots,u_{-2},u_{-1}): Tu_{i} = u_{i+1}\]is called the inverse-limit space of $(U,T)$.
- The self-map $\widehat{T}$  induced by $T$ on $\widehat{U}_T$  is defined by \[\widehat{T}: (\ldots,u_{-2},u_{-1}) \mapsto  (\ldots,u_{-2},u_{-1},T(u_{-1}))\] and the resulting dynamical system $(\widehat{U_T}, \widehat{T})$ is called the inverse-limit system of $(U,T)$.
-\end{Definition}
-
-\begin{Definition}
-  [\bf Causal Embedding]\label{Dfn_CausalEmbed}\rm
-  A driven system $g$ is said to causally embed the dynamical system $(U,T)$ if 
-  \vspace{-8mm}
-\begin{enumerate}[noitemsep, label=\roman*.]
-  \item The diagram in \ref{Scomm_h} commutes (i.e. $h$ is a universal semi-conjugacy),
-  \item $H_2(\overline{u}):=(h(r\overline{u}, h(\overline{u})))$ embeds the inverse-limit space $(\widehat{U}_T, \hat{T})$ of $(U,T)$ in $X\times{X}$.
-\end{enumerate}
-\end{Definition}
-
-The driven system $g$ can induce an embedding of $\overleftarrow{U}$ in $\overleftarrow{X}$ as follows: 
-If $H:\overleftarrow{U}{\to}{\overleftarrow{X}_U}$ is also injective (in addition to being surjective, and hence bijective), 
-it becomes the embedding of $\overleftarrow{U}$ in $\overleftarrow{X}$ induced by the driven system $g$ and we refer to $H$ as a \emph{causal embedding} 
-\ednote{M : Causal embedding a dynamical system is different from the definition of causal embedding, so please edit/delete this paragraph and the next theorem and also in the remaining part till Section 1.2}.
-
-\begin{Theorem}
- If $g(\ldotp,x)$ is invertible and has the USP then $H$ is a causal embedding. 
-\end{Theorem}
-\begin{proof}
-  See \cite{manjunath2021universal}.
-\end{proof}
-
-When $g$ has the USP, the diagram below illustrates the operation of the mappings $h$ and $H$.
-The mapping $h:\overleftarrow{U}\to{X_U}$ is an observable as discussed in an above section (\ref{TakensEmbedding}).
-
-\begin{figure}[ht]
-  %\includegraphics[scale=0.3]{actionofh_H.png}
-  \centering
-  \label{fig:actionh_H}
-\end{figure}
-
-The next result is of relative import but to fully comprehend its influence, some discussion still remains. We state it in part here.
-% Before stating it, we define a concept which will directly be shown to be equivalent to the USP.
-
-\begin{Theorem}
-  The following statements are equivalent:
-  \vspace{-8mm}
-  \begin{enumerate}[noitemsep, label=\roman*.]
-    \item $g$ has the USP. 
-    \item $g$ is a topological contraction.
-  \end{enumerate}
-\end{Theorem}
-% \begin{proof}
-%   (postponed) 
-% \end{proof}
-
-\section{Choosing the driven system $g$}
-
-One must be careful to avoid a choice of $g$ which would  add complexity to the obtained solution.
-
-When a causal embedding $H$ exists for the driven system $g$, one can map an arbitrary input ${u}$ onto the solution space $X$ without additional distortion or information-loss \textbf{(Cite.)}.
-When an embedding is established, the question of possible additional complexity in the solution is removed by guaranteeing that, since the systems are conjugate (semi-conjugate, \textbf{(refer)}), $g$ does not add any (some) complexity to the system.
-
-It is, however, a balancing act as it also undesirable to choose a function $g$ that quenches the temporal structure in $u$ by contracting to such a degree that the ability to recover information from the original system is lost completely.
-In the above example (\ref{ex_halfux}), the input's temporal variation cannot be related to the reachable set since $X_U$ consists of a single element; little, if not no, information is encoded.
-To obtain a suitably complex function $g$, it is thus desired that the reachable set of a driven system  be large enough to relate to the input. 
- We  reguire that the reachable set of a driven system to be such that the inverse-limit space of $U_T$ can be embedded in some finite self-product of the reachable set of $T$ \textbf{(refer or cite)}.
-To this end, consider the notion of State-Input (SI) Invertibility.
- \ednote{(M: Expand here. It stops a bit abruptly. B: Does the above reorganisation read better?)}
-
-\begin{Definition}
-  [\bf SI-Invertibility]\label{Dfn_SIinv}\rm
-  A map $g$ is said to be SI-Invertible if $g$ is invertible for all $x\in{X}$.
-Alternatively it may be said that if, given $x_n$ and $x_{n-1}$, $u_n$ can be uniquely determined from $x_n=g(u_n,x_{n-1})$, then $g$ is said to be SI-invertible.
-\end{Definition}
-
-SI-invertibility promises that 'enough' information is retained when $g$ is chosen without introducing additional unwanted complexity. \ednote{(M: Expand why this is true. B: Is this satisfactory?)} 
-'Enough' here refers to the fact that we may always recover the previous input value $u_n$ (if we know successive states $x_{n-1}, x_n$).
-'Without additional complexity' is guaranteed by the invertibility of the function $g$ that guarantees a one-to-one mapping between $u_n, x_{n-1}$ and $x_n$. 
-In obtaining an SI-invertible function, therefore, we are guaranteed a system which will not lose so much information about its previous states in the forward-flow of time that one may not make any credible claims as to the original system's behaviour and properties.
-Simultaneously, it does guarantee us that the important information encoded in previous inputs is indeed preserved.
-
-It is worth at this stage taking note of a specific driven system in the form of a discrete state-space model that has acquired some adherence in applications \cite{Manju_IEEE} and especially those pertaining to Echo State Networks and ESP.
-The function
-
-\begin{equation}  \label{eqn_driving}
-  g(u,x) = (1-a)x + a\overline{\tanh}(Au + \alpha Bx)
-\end{equation} 
-is both SI-invertible and, if $\alpha\beta<1$, also possesses the USP. 
-One may easily show that $g$ is SI-invertible by recovering $u_n$ by 
-\begin{equation} \label{eqn_SI_RNN}
-  u_n := A^{-1}\bigg(\overline{\tanh}^{^{-1}}\frac{1}{a}\Big(x_{n+1}-(1-a)x_n\Big) \bigg) - \alpha B x_n
-  \end{equation}
-  for $x_{n-1}, x_n$ known.
-
-The proof that the map $g$ possesses the USP is rather involved and hence, instead of being replicated here, may be found in \cite[Th.2]{manjunath2013echo }
-This specific driving function $g$ is used in our implementation and is discussed more completely in chapter \ref{sect5}
-
-Despite the ease that one may work with a left-infinite history in the realm of theory, it is impossible to obtain such a sequence in any real-life application.  
-One does not in practice, fortunately, need the entire left-infinite history of the input thanks to the Uniform Attraction Property (UAP).
-
-\begin{Definition}
-  [\bf Uniform Attraction Property]\label{Dfn_UAP}\rm
-  A driven system $g$ has the Uniform Attraction Property (UAP) if, regardless of starting position, all trajectories converge to a single trajectory as time flows forward. 
-  This trajectory is also the unique solution sequence $x$ to the input sequence $u$ as mentioned above.
-\end{Definition}
-
-This definition is stated in a not completely rigourous manner as the more formal definition makes use of processes, a concept which would take some time to establish and detracts from the principal thrust of this project.
-\ednote{B: Is it very obvious to the reader at this point what the general thrust of the paper is?}
-
-As the Unique Attraction Property guarantees that all trajectories will converge to the same trajectory as time moves forward. 
-Incredibly, this permits one to initialise a driven system $g$ with an altogether arbitrary initial value $y_m\in{X}$ where $m\in\mathbb{Z}$ and the UAP then guarantees that the sequence $\{y_{m+1}, y_{m+2},\ldots\}$ which satisfies the relation $y_n=g(u_n, y_{n-1})$  for  $k\geq{m}$  will uniformly approach the elements $\{x_n\}$ of the actual solution (See \cite[Th.11,]{Manju_Nonlinearity} or \cite[Eqn.18]{Supp}).
-
-\begin{figure}[ht]
-  \includegraphics[scale=0.4]{_memloss_conttime.png}
-  \centering
-\captionof{figure}{Reproduced from \cite{manunath2021universal}. Illustration of all trajectories approaching the true solution if $g$ has the UAP}
-\label{fig:memloss_conttime.png} 
-\end{figure}
-
-One may now appreciate an even more astounding result: $g$ having the USP is equivalent to the UAP. We restate the above theorem.
-
-\begin{Theorem}\label{thm_usp-contr-uap}
-  The following statements are equivalent:
-  \vspace{-8mm}
-  \begin{enumerate}[noitemsep, label=\roman*.]
-    \item $g$ has the USP.
-    \item $g$ is a topological contraction.
-    \item $g$ has the UAP.
-  \end{enumerate}
-\end{Theorem}
-\begin{proof}
-  May be found in \cite[Th.1]{Manju_Nonlinearity}
-\end{proof}
-
-One need therefore not establish any additional results to ensure that the sequence uniformly approaches the unique solution $\Psi$.
-This vastly simplifies the effort necessary to set up a problem in order to guarantee that the underlying system will be accurately represented by the conjugate system.
-
-This also solves the problem of perturbation/noise introduced by the observable or measurement function.
-Recall from before that an observable is inherently a map that discretises the underlying continuous-time system.
-Moreover, measurement and device error introduce mistakes.
-Since we are working with a chaotic system displaying sensitive dependence on initial conditions, these small errors could potentially send the trajectory into a completely different attractor.
-In the presence of the UAP, however, small measurement errors introduced into the system do not pose the same danger as before since the trajectories will eventually tend towards the actual solution.
-\ednote{(B: Is this Complete enough?)}
-% \textbf{From the perspective of perturbation of an autonomous dynamical system, if the resulting nonautonomous system has the USP, then the temporal structure of noise relates to the dynamics on a nonautonomous uniform attractor through a topological semi-conjugacy or a conjugacy.}
-
-\section{The next step in Dynamics}
-
-Next we define the relation $Y_T$ induced by $(U,T)$ on $X_U\times{X_U}$ for a driven system $g$ possessing SI-invertibility.
-To describe the  single-delay lag dynamics formally, we consider a dynamical system $T: U \to U$ and we define a relation on the reachable set $X_U$, i.e., a subset defined on $X_U \times X_U$ by
-$$Y_T:=\{(x_{n-1},x_n): \{x_k\}_{k\in \mathbb{Z}} \mbox{ is a solution for some orbit of } T \mbox{ and } n \in \mathbb{Z}\}.$$
-
-The following theorem establishes the existence of a well-defined map $G_T$ describing the single-delay dynamics(SDD) of the system above.
-
-\begin{Theorem}
-If we let $G_T:Y_T\to{Y_T}$ be a map defined by the relation $(x_{n-1},x_n)\mapsto(x_n,x_{n-1})$, then $G_T$ is well-defined (and this results holds even in the absence of $g$ possessing over the USP)
-\end{Theorem}
-
-We're now getting quite close to where we want to be and our results carry more and more weight. Recall that $g$ is only being given inputs from the orbits of $T$.
-
-\begin{Definition}
-  [\bf Inverse-Limit Space]\label{Dfn_InvLimSpace}\rm
-  The inverse-limit space $\widehat{U}_T$, a subspace of $\overleftarrow{U}$, is defined by $$\widehat{U}_T:=\{(\ldots, u_{-2}, u_{-1}):T(u_n)=u_{n+1}\}$$ 
-\end{Definition}
-
-The inverse-limit space is well-defined since $T:U\to{U}$ is surjective by assumption.
-
-\begin{Theorem}
-  Graph \ref{Scomm_h} is exactly the inverse-limit system $(\hat{U}, \hat{T})$.    
-\end{Theorem}
-
-Note that $H_2$ maps an entire left-infinite solution sequence from $\Psi$ to an element in $X\times{X}$.
-We now have the following (compare with \ref{SCausal_H} above):
-
-% \ednote{Add Causal Embedding Theorem}
-% \textbf{NB NB NB - Causal Embedding Theorem! - Th.4}
-
-\begin{equation} \label{Fig_inverse_limsystem}
-  %  \[ 
-      \psset{arrows=->, arrowinset=0.25, linewidth=0.6pt, nodesep=3pt, labelsep=2pt, rowsep=0.7cm, colsep = 1.1cm, shortput =tablr}
-      \everypsbox{\scriptstyle}
-      \begin{psmatrix}
-      \widehat{U}_T  && \widehat{U}_T \\
-      Y_T && Y_T \\
-      %%%
-     %  \ncline{1,1}{1,2}^{\widehat{T}} \ncline{1,1}{2,1} <{h}
-     %  \ncline{1,2}{2,2} > {H_2}
-     %  \ncline{2,1}{2,2}^{G_T}
-      \end{psmatrix}
-    %  \]
-  \end{equation}
-
-\begin{Theorem}
-    $(Y_T, G_T)$ is semi-conjugate to $(\widehat{U}, \widehat{T})$.
-\end{Theorem}
-\begin{proof}
-  See \cite[Th.3, Th.4]{manunath2021universal}
-\end{proof}
-
-\section*{Summarising the discussing thus far:}
-
-It is easy to lose the birds-eye view, so we take a moment to review our progress up until this point.
-
-\vspace{-8mm}
-\begin{enumerate}
-\item We are interested in a some dynamical system $(U,T)$ with unknown dynamics.
-\item To determine properties about this system $(U,T)$ and predict its future evolution, we determine the dynamics of the inverse-limit system $(\widehat{U}, \widehat{T})$.
-Given certain assumptions, we can guarantee that $(\widehat{U}, \widehat{T})$ is at least semi-conjugate to $(U,T)$.
-\item If the driven system $g$ is SI-invertible (and $\{u_n\}\in{U}$  is an orbit of $T$), the map $G_T$ exists. 
-\item If, furthermore, $g$ has the USP, $(Y_T, G_T)$ is semi-conjugate to $(\widehat{U}, \widehat{T})$.
-\item If we can assume that $T$ is a homeomorphism, $(Y_T, G_T)$ is topologically conjugate to $(\widehat{U}, \widehat{T})$, an extension space of $(U,T)$
-\end{enumerate}
-
-One can therefore learn the SDD of the driven system states via $G_T$ with enough data thanks to the USP/UAP. This enables us to do at least 2 things: 
-\vspace{-8mm}
-\begin{itemize}[noitemsep]
-\item Forecast future values of $x_n$ via iterates of $G_T$ (as $G_T$ can be determined.
-\item Forecast future values of $u_n$. 
-\end{itemize}
-
-% Below is reproduced a pictorial representation of the relationship between different entities and properties. \textbf{(Cite SIG_25May)}
-% Below is reproduced a pictorial representation of the relationship between the different entities and their properties.
-
-\begin{figure}[ht]
-  \includegraphics[scale=0.3]{_summarypictorial.png}
-  \centering
-  \label{fig:fig_pictorialSummary}
-\end{figure}
-
-\section{A discussion of $G_T$ }
-
-In the above sections, we established the map $G_T$ describing the SDD of a driven system.
-By establishing the existence of this map, we’ve essentially embedded the attractor $U$ (recall from \ref{attractor_U}) into the higher dimensional space $X\times{X}$.
-In layman’s terms, this ensures that there is more “dimensional room” for the underlying system’s underlying dynamics to “move”.
-Since the dynamics aren’t as “squashed”, we might therefore hope that the dynamics of $G_T$ are in some sense simpler than that of $T$.
-(Taken note of the fact that $G_T$ is a homeomorphism even when $T$ is just continuous)
-
-In \cite{manjunath2021universal} it is illustrated in an empirical fashion in that the map $G_T$ describes dynamics which are less functionally complex than that of $T$ or of the map $\Phi_{2d,\theta}$.
-This is done by implementing a Recurrent Neural Network (RNN), which is discussed in~\ref{ch5}.
-\ednote{(B: Should I add in that we'll be using Pearson coefficient?)}
-
-We opt to learn $G_T$ in an indirect manner by defining a new map $\Gamma:(x_{n-1},x_n)\mapsto{u_n}$. It will follow immediately from $G_T$’s existence that $\Gamma$ also exists.
-
-The reasons for taking this roundabout approach remain to be discussed in \ref{subs_LearnGamma}, but a pat answer may immediately be given: When $\Gamma$ has been learnt, the system can be driven autonomously and then $G_T$ is known anyway.
-We formalise this with a theorem.
-
-\begin{Theorem}
-  When $x_n$ and $x_{n-1}$ are successive points on a solution obtained for input-orbit of $T$ $\{u_n\}$, then the map $\Gamma: X\times{X}\to{U}$ defined by $(x_{n-1},x_n)\mapsto{u_n}$ exists whenever $G_T$ exists
-\end{Theorem}
-\begin{proof}
-  See \cite[Th. 3c]{manjunath2021universal}.
-\end{proof}
-
-Projection mappings $\pi_i$ are defined in the traditional meaning where a $k$-dimensional vector is projected to it's $i^{th}$ component such that $$\pi_i:(a_1,a_2, \ldotp, a_{k-1}, a_k)\to{a_i}$$
-
-% \begin{Definition}
-%   [\bf Projection Mapping]\label{Dfn_ProjMap}\rm
-%  The projection mapping $\pi_i$ is defined as follows: $$\pi_1:(a_1,a_2)\to{a_1}$$ and $$\pi_2:(a_1,a_2)\to{a_2}$$ 
-% \end{Definition}
-
-The graph in equation \ref{Fig_inverse_limsystem} is then extended as below:
-
-% \begin{figure}[ht]
-%   \includegraphics[scale=0.3]{_gammatemp.png}
-%   \centering
-%   \label{fig:temp_Gamma}
-% \end{figure}
-
-\begin{equation}  \label{Scomm_imp}
-  %    \[ 
-      \psset{arrows=->, arrowinset=0.25, linewidth=0.6pt, nodesep=3pt, labelsep=2pt, rowsep=1.3cm, colsep = 1.1cm, shortput =tablr}
-   \everypsbox{\scriptstyle}
-   \begin{psmatrix}
-   \widehat{U}_T & \widehat{U}_T\\%
-   Y_T & Y_T\\
-   & \textcolor{red}{U \times X}
-   %%%
-  %  \ncline{1,1}{1,2}^{\widehat{T}} \ncline{1,1}{2,1} <{H_2}
-  %  \ncline{1,2}{2,2} > {H_2}
-  %  \ncline{2,1}{2,2}^{G_T}
-  %  \psset{linecolor=red}
-  %  \textcolor{red}{\ncline{2,1}{3,2}<{(\Gamma,\pi_2)}}
-  %  \ncline{3,2}{2,2}>{\textcolor{red}{(\pi_2,g)}}.
-   \end{psmatrix}
-  % \]
-  \end{equation}
-
-The problem finally simplifies to the issue of learning the map $\Gamma$ and combining this with the projection mapping $\pi_2$ and the function $g$, which will be known.
-A final set of equations is obtained - equations that have been entirely constructed from data.
-\begin{eqnarray}\label{eqns_from_data}
-	u_{k+1} &=& \pi_1 \circ (\Gamma, \pi_2) \circ (\pi_2,g) (u_k,x_k) \label{Seqn_u}\\
-	x_{k+1} &=& \pi_2 \circ (\Gamma, \pi_2) \circ (\pi_2,g) (u_k,x_k). \label{Seqn_x}
-\end{eqnarray}
-
-\section{Advantages of learning $\Gamma$} \label{subs_LearnGamma}
-
-One may immediately ask why we opt to take such a roundabout route; why not just learn the map $G_T$ from the get-go?
-On the surface, this seems to be an arbitrary decision path with no real reasoning, so we take a pause again and discuss the motivation for learning Gamma.
-There are a number of distinct advantages.
-
-In the first place, Learning $\Gamma$ saves computational resources.
-This is due to the fact that the input $u_n$ lies in a lower-dimensional subspace of the high-dimensional space $X\times{X}$ and we can learn $\Gamma$ on a small subset of the space.
-In practice, if the input is of a  lower dimension, one may easily embed it into the space $X\times{X}$ by padding the vector $u_n$ with zeroes.
-
-Secondly, the function $\Gamma$ is known to be stable.
-Learning Gamma makes use of equations \ref{eqns_from_data} which in turn employs $g$, a function possessing the USP, and offers distinct advantages with regards to stability in the presence of perturbation.
-It is desirable that inputs that differ only slightly would have only ‘slight’ effects on the output of the system.
-Moreover, one would hope to prevent numerical errors originating from input and measurement noise (see~\cite[Th.
-5]{manjunath2021universal}).
-As the data fed into the system is a function of both the input given and the parameters of the system itself (i.e.
-the value a, $\alpha$ in $\tanh$), we investigate both input-related and parameter-related stability.
-Both these notions are defined by way of the continuity of an encoding map of an input and design parameter respectively in \cite{manjunath2020stability}.\
-
-Informally, we may consider input-related stability as concerned with the question of whether or not small variations in input would result in small responses.   
-If two inputs $u$ and $u_{noisy}$ are close in the product topology, then their tails could differ greatly due to the metric's insensitivity to the differences in sequence's tails.
- However if the system has the USP then the function $h$ is continuous and so $h(u), h(u_{noisy})$ remain near – maintaining a measure of robustness to input and measurement noise.\cite{manjunath2021universal}.
-In a similar fashion, parameter-related stability is obtained thanks to a result \cite[Lemma 3.2]{manjunath2020stability} relating this form of stability with the ESP (which is here equivalent to the USP).
-
-% Secondly,  due to $g$ possessing the USP, both input- and parameter-related stability~\cite{manjunath2020stability} is obtained, which in turn prevents numerical errors that caused by input and measurement noise (see~\cite[Th.
-5]{manjunath2021universal}).
-
-\begin{figure}[ht]
-  \includegraphics[scale=0.5]{_AutovsDriven.eps}
-  \centering
-  \label{fig:learningFailure}
-  \captionof{figure}{Illustrating empirically the importance of global dissipativity.
-Plotted below the three principal components of randomly initiated trajectories of a learnt version of $G_T$ versus the trajectory $(x_{n-1}, x_n)$(grey) of driven stated obtained from the Thomas Attractor's data (see below).
-}
-\end{figure}
-
-\ednote{B: Would you please help me with the writing here? Otherwise I shall leave this out.}
-\emph{Thirdly, $\Gamma$ is globally dissipative whereas $G_T$ is not guaranteed to be~\cite{manjunath2021universal}). 
-Global dissipativity prevents large numerical errors due to input data.If a system is not globally dissipative, neglibly small errors result in major errors which thus induce predictions to fail utterly.
-If one learns $G_T$ directly, we are in danger of replicating the problems arising for Takens.
-}
-
-% The $\omega$-limit set $\omega(u;T)$ of a point $u$ is defined to be the collection of limit points of the sequence $\{x,Tu,T^2u,\ldots\}$.
-This set is nonempty, and $\omega(u;T)$ is invariant.
-\textbf{(Change)}
-% \begin{Definition}
-%   [\bf Global Dissipativity]\label{Dfn_GlobalDiss}\rm
-% 	We say a dynamical system $(Z,f)$ is globally dissipative if there exists a nonempty proper closed subset $B$ of $Z$  so that for all $x\in Z\setminus B$, (i).
-$\omega(x;f) \subset B$  and (ii).
-$B$ is positively invariant that is $f(B) \subset B$.
-We call any such closed subset $B$ a trapping set of $f$.
-% \end{Definition}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\chapter{Implementation} \label{ch5}
-
-In this section we turn to the implementation of the function $G_T$ by discussing a method to learn the function $\Gamma$.
-We then indicate where this implementation differs from previous work done by (\cite{manjunath2021universal}) and empirically show that the SDD of $G_T$ are less functionally complex than $\Phi_{k, \theta}$ or $T$ by considering the Pearson correlation coefficient.
-Finally, we present and discuss some numerical results for a focus system of this project: the double pendulum. We also consider other chaotic attractors.
-
-\section{Implementing $G_T$}
-The map $G_T$ is implemented by applying the functions $\Gamma$, $g$ and the projection function $\pi_2$ in the relevant order as described by the equations~\ref{eqns_from_data}.
-To reach this point, one must first learn $\Gamma$.
-After first normalising the data to ensure it lies within the interval $|x| < 1$ and has zero mean value, we continue by driving the system with external input through the function $g$.
-
-From \eqref{eqn_driving} we know that $ g(u,x)=(1-a)x+a{\cdot}\tanh(Au+\alpha{a}Bx)$.
-Recall that $g$ is SI-invertible.
-%Adopting the terminology of \cite{manjunath2021universal}, $A$ is called the “input matrix” and $B$ the “reservoir matrix”.
-
-We make use of a Recurrent Neural Network (RNN) in our implementation.
-RNN's are widely used in the literature for discrete-time processing and are considered to be discrete-time version of continuous dynamical systems modelling a network of neurons.
-(\textbf{Citation.})
-A complete discussion of RNN's is outside the scope of this thesis, so we present only the details necessary to the current discussion.
-For information, (\textbf{Cite, Cite, Cite}) may be consulted.
-We implement our RNN in the \emph{Python} programming language and making use of the \emph{Scipy} and \emph{Tensorflow} libraries.
-
-If the RNN we will be using in the next step has $N$ neurons, then $A$ and $B$ are each $N\times{N}$ matrices.
-Inputs $u_n$ of dimension $K < N$ are padded by zeroes to obtain an input of dimension $N$.
-As we do not need the entire left-infinite history of input, an arbitrary initialisation of the system with initial values followed by the driving of the system with the function $g$ means that the generated values will approach the actual solution’s elements in a uniform manner thanks to the UAP.
-\textbf{(Citation)}
-
-Thirdly we prepare the network.
-Utilising a network $N$ neurons (the same value $N$ as in the dimensions of the matrices $A,B$).
-Training is done using the Adam Optimiser where the MSE loss function is optimised.
-The system is trained several times(in our case 4) with incrementally smaller learning rates on each run (0.001 and thereafter divided by 10 at each run).
-Additional parameters such as the number of network-layers, neurons per layer and training-length are presented case-by-case below.
-
-We then reach the prediction step. Making use of $g$'s SI-invertibility and $g$ as defined in \eqref{eqn_driving}, we have the following expression for $u_n$ replicated from~\eqref{eqn_SI_RNN}.
-\begin{equation*}
-  u_n := A^{-1}\bigg(\overline{\tanh}^{^{-1}}\frac{1}{a}\Big(x_{n+1}-(1-a)x_n\Big) \bigg) - \alpha B x_n
-\end{equation*}
-
-The two model equations for learning from data (\eqref{eqns_from_data}) allows us to determine what the state $x_{n+1}$ at time $n+1$ will be. 
-We specifically remind ourselves of \[x_{k+1}=\pi_2 \circ (\Gamma, \pi_2) \circ (\pi_2,g) (u_k,x_k)\]
-
-\section{Changes in this project to previous work of~\cite{manjunath2021universal}}
-
-We made several changes which gave improved results:
-\vspace{-8mm}
-\begin{enumerate}
-\item Using Delay Coordinates (DC) when driving the system vs.
-using Single/Multi Delay Dynamics (SDD/MDD) after driving (and before training).
-Explain/show why these two methods are equivalent.
-\ednote{Would MDD  be not computationally intensive ?
-So may be you just explain if you can’t show it.}
-\item Sparse matrix implementation.
-Slight improvement in some scenarios, but not fantastic.
-\ednote{M: Did it not help you a great deal with the DP forecasting?
-B: Sparse matrices tended to be finicky.
-In some cases they helped whereas in others they actually worsened the solution somewhat.
-There's some tuning to be done in terms of condition number, scaling and density}
-
-(Possible paragraph here/later on implementing sparsity of A/B: Generate sparse matrix that is also well-conditioned and correctly scaled matrix.)
-\item Density of matrices A/B(i.e. number of non-zero entries in each matrix) (separate from sparse matrix implementation)
-\end{enumerate}
-
-\section{Numerical simulations}
-In this section we showcase numerical results pertaining to 3 attractors that were simulated using the methodology described in the previous chapters. We first consider a simple physical system.
-\subsection{Double Pendulum}
-
-The Double Pendulum consists of two pendulums fastened to one another such that the system moves as a whole.
-The first (or higher pendulum) is attached to a fixed point while the second is attached to the endpoint of the first.
-(See below)
-This system is a classical example of a system exhibiting chaos. \textbf{(Citation.)} Consider again~\ref{fig:dp_sdic}.
-The system has four variables: angle and angular velocity of each pendulum rod. An ideal system is assumed where the rods are massless and no friction in the pivots are considered.
-
-\begin{figure}[ht]
-  \includegraphics[scale=0.35]{_dp_setup.png}
-  \centering
-  \captionof{figure}{Pictorial representation of the setup of a DP. Image taken from~\cite{DPSetup}}
-  \label{fig:dp_setup}
-\end{figure}
-
-The scalar equations of motion are derived using Newtonian physics and can be found in~\cite{DPFormulas}.
-The second-order system is converted to a first-order ODE and solved using the Runge-Kutta numerical method.
-Noise-free, undamped DP data-sets were generated and utilised in this project due to the requirement that the map $T$ be surjective.
-It may easily be verified that a DP undergoing damping (as seen in \ref{conjugacy_with_Gamma}) is not surjective and hence the required assumptions to apply our theory do not hold.
-\ednote{B: Change sentence.}
-
-\begin{figure}[ht]
-  \includegraphics[scale=0.4]{_dp_dying.eps}
-  \centering
-\caption{A damped pendulum is not surjective.}
-\label{fig:damped_pendulum}
-\end{figure}
-
-\textbf{Parameter Values \& Successes}
-
-Train length: 15000. Prediction Length: 1000. Discard to wait for network to forget: 500. Dimension of network: 128 neurons.
-$\alpha=0.99$. $a=0.5$.
-
-Hidden Layers: 16. Layer dimension: 64. Epochs:256. Batch Size: 128. Delay(for MDD): 8. Density of Sparse Matrices: 0.2. $d=1$
-
-\ednote{B: Write this paragraph.}
-
-\begin{figure}[ht]
-  \centering
-  \includegraphics[width=\linewidth]{_DP_Fig-2F2023.eps} 
-  \captionof{figure}{Predicted(red) and actual(blue) trajectories of the $x$- and $y$-coordinates of each of the pendulum heads.} 
-\end{figure}
-\begin{figure}[ht]
-  \centering
-  \includegraphics[width=\linewidth]{_DP_Fig-2B2023.eps}
-  \captionof{figure}{Densities of predicted and actual values of x-coordinate of first pendulum head. }
-  \end{figure}
-  \begin{figure}[ht]
-    \centering 
-  \includegraphics[width=\linewidth]{_DP_Fig-2D_5.eps}
-    \captionof{figure}{Takens Delay Map with delay=5 to illustrate behaviour of predicted and actual values of double pendulum data.}
-\end{figure}
-
-\subsection{Other attractors I’ve experimented with: Clifford \& Thomas}
-We follow the methodology presented in \cite{manjunath2021universal} but opt to consider attractors different from the Lorenz system and H\'enon map in this project.
-
-\subsubsection{Thomas' Cyclically Symmetric Strange Attractor}
-Thomas' Cyclically Symmetric Strange Attractor, a 3D attractor proposed in 1999 by Ren\'e Thomas in \cite{ThomasAttractor}, is described by a set of three equations which is distinctly reminiscent of the Lorenz system, another deterministic three-equation model exhibiting chaotic behaviour.It has a single parameter $\beta$ and has been shown to transition to chaotic behavious when $\beta<0.208186$~\cite{Thomas_BetaParameter}.
-\begin{eqnarray}\label{eqns_thomas}
-  x_{n+1} = \sin(y_n) - \beta{x_n} \\
-  y_{n+1} = \sin(z_n) - \beta{y_n} \\
-  z_{n+1} = \sin(x_n) - \beta{z_n}
-\end{eqnarray}
-
-The behaviour of the system with a $\beta$-parameter value of 0.1056 is simulated by scaling it to fit inside the interval $[-1,1]$ and both perturbed with noise as well as leaving it intact.
-Specifically the noise is obtained from a normally distribution with mean 0 and standard deviation 0.05, which translates approximately to *dB.
- A sequence of 3000 observations in the clear and noisy data-set respectively is fed into the system with 300 entries discarded before predicting the next 9000 entries.
-
-\begin{figure}[ht]
-  \centering
-  \minipage{0.5\linewidth}
-    \centering
-    \includegraphics[width=\linewidth]{Thomas_nonoise.eps}
-    \caption*{True and predicted trajectories in the absence of noise.}
-  \endminipage\hfill
-  \minipage{0.5\linewidth}
-    \centering
-    \includegraphics[width=\linewidth]{ThomasNoisy.eps}
-    \caption*{True and predicted trajectories for noise added.}
-  \endminipage
-\end{figure}
-
-The RNN was initialised as follows: The dimension of the network is set equal to 500 neurons and 12 hidden layers of dimension 64 each were initialised.
-Training is realised through 150 epochs of batch sizes of 128 each; the map is learnt via the Adam Optimiser using finer and finer learning rates (0.001 and then divided by 10 at every iteration).
-
-\begin{figure}[ht]
-  \centering
-  \includegraphics[scale=0.35]{Thomas_1.eps}\caption*{Predicted trajectories for the attractor demonstrate empirically the ability to predict the evolution of the trajectory for the next ~100 timesteps}
-  \includegraphics[scale=0.5]{Thomas_3.eps}\caption*{Represented here the learnt (blue) and actual (red) densities of the first coordinate of the dataset.}
-\end{figure}
-
-\subsubsection{Fractal Dream Attractor}
-
-Another 2-equation discrete-time map named the Fractal Dream Attractor, or more commonly known as the Pickover Map (first discovered by Clifford A.
-Pickover and discussed in his fascinating book "Chaos in Wonderland"\cite{PickoverChaos}), was also considered.
-\begin{eqnarray}\label{eqns_clifford}
-  {x_{n+1}=\sin(ay_n) + c\cos(ax_n)} \\
-  {y_{n+1}=\sin(bx_n)+d\cos(by_n)}
-\end{eqnarray}
-
-\begin{figure}[ht]
-  \centering
-  \includegraphics[width=1.0\textwidth,left]{Clifford_1.eps}
-  \caption*{These graphs were constructed by predicting the Clifford system 15000 steps into the future and in so doing illustrating the long-term consistency and accuracy of the learnt system.
-As perceived here, we are able to lock on to the trajectory of the Clifford map almost exactly for the first 25 steps.}
-  \minipage{0.5\textwidth}
-      \centering
-      \includegraphics[width=\linewidth]{Clifford_2.eps}
-  \endminipage\hfill
-  \minipage{0.5\textwidth}
-    \centering
-    \includegraphics[width=\linewidth]{Clifford_3.eps}
-  \endminipage
-  \caption{The bottom row, left, shows a plot of the true attractor(red) and the predicted values(blue).
-Bottom row, right, shows the distribution of the first coordinate of the learnt system (blue) and the actual Clifford system (red).
-}
-  \label{Fig_Clifford}
-\end{figure}
-
-\ednote{B: Expand more.}
-
-\section{Generalized Pearson Correlation}\ednote{B: Alter heading.}
-
-\ednote{B: Write Pearson Introduction. Could you give me some help here, Sir?}
-
-\begin{center}
-  \begin{table} 
-      \scalebox{0.75}
-      {\begin{tabular}{|c|c| c c |} 
-          \hline
-          Input & Dimension of $X$ & $u_n$ vs $u_{n+1}$ 
-          & $\begin{bmatrix}
-              x_{n-1}\\
-              x_n
-          \end{bmatrix}$ vs $\begin{bmatrix}
-              x_n\\
-              x_{n+1}
-          \end{bmatrix}$ \\
-          \hline
-
-          % \hline
-          % $(w_n)$ is the Lorenz states, sampled every 0.1 timestamps. & & &\\
-          % \multirow{3}{*}{$u_n=\frac{1}{100}w_n$} & 10 &0.9311  & 0.9731\\
-          %     & 100 &  & 0.9934\\
-          %     & 1000 & & 0.9930\\
-
-          % \multirow{3}{*}{$u_n=\frac{1}{10}(\sin(0.1w_{n,x})+\sin(0.1w_{n,y})+\sin(0.1w_{n,z}))$} 
-          %     & 10 &0.8401  & 0.9392\\
-          %     & 100 & & 0.9616\\
-          %     & 1000 & & 0.9737\\
-          % \hline
-
-          \hline
-          $(w_n)$ comes from the Thomas Attractor (for $\beta=0.1056$): & & & \\
-          {$w_{n+1}= \begin{bmatrix} \sin(w_{n,y}) - \beta{w_{n,x}} \\
-              u_y = \sin(w_{n,z}) - \beta{w_{n,y}} \\
-          u_z = \sin(w_{n,x}) - \beta{w_{n,z}} \end{bmatrix}$} & & & \\
-          \multirow{3}{*}{$u_n = w_n - \overline{w}$}
-              & 10 & 0.99989 & 0.999886 \\
-              & 100 & &  0.999913 \\
-              & 1000 & & 0.999901 \\
-          \hline
-
-          \hline 
-          $(w_n)$ comes from Noise-Added Thomas Attractor (for $\beta=0.1056$): & & & \\
-          {$w_{n+1}= \begin{bmatrix} \sin(w_{n,y}) - \beta{w_{n,x}} \\
-              u_y = \sin(w_{n,z}) - \beta{w_{n,y}} \\
-          u_z = \sin(w_{n,x}) - \beta{w_{n,z}} \end{bmatrix}$} & & & \\
-          \multirow{3}{*}{$u_n = w_n - \overline{w}$}
-              & 10 & 0.9994 & 0.99985 \\
-              & 100 & &  0.99991 \\
-              & 1000 & & 0.999895\\
-          \hline
-
-          \hline
-          $(w_n)$ comes from the Thomas Attractor (for $\beta=0.208186$): & & & \\
-          {$w_{n+1}= \begin{bmatrix} \sin(w_{n,y}) - \beta{w_{n,x}} \\
-              u_y = \sin(w_{n,z}) - \beta{w_{n,y}} \\
-          u_z = \sin(w_{n,x}) - \beta{w_{n,z}} \end{bmatrix}$} & & & \\
-          \multirow{3}{*}{$u_n = w_n - \overline{w}$}
-              & 10 & 0.99982 & 0.99981\\
-              & 100 & &  0.999851 \\
-              & 1000 & & 0.99984 \\
-          \hline
-
-          \hline
-          $(w_n)$ comes from the Clifford Attractor \\ (for $a = -1.7; b = 1.8; c = -1.9; d = -0.4$): & & & \\
-        {$w_{n+1}= \begin{bmatrix} \sin(aw_{n,y}) + c\cos(aw_{n,x}) \\ 
-                                                                    \sin(bw_{n,x})+d\cos(bw_{n,y}) \end{bmatrix}$} & & & \\
-          \multirow{3}{*}{$u_n = w_n-\overline{w}$}
-              & 10 & -0.21713 & 0.78719 \\
-              & 100 & &  0.8334 \\
-              & 1000 & & 0.84854 \\
-          \hline  
-          % \hline 
-          % $(w_n)$ comes from the Logistic map: & & & \\
-          % {$ w_{n+1}  = 4w_n(1-w_n)$} & & & \\
-          % \multirow{3}{*}{$u_n = w_n - 0.5$} 
-          %     & 10 &-0.0314  & 0.5577\\
-          %     & 100 &  & 0.7150\\
-          %     & 1000 & & 0.7826\\
-          % \hline
-          % \hline
-          % $(w_n)$ comes from the Svensson Attractor \\ (for $a = -1.7; b = 1.8; c = -1.9; d = -0.4$): & & & \\
-          % {$w_{n+1}= \begin{bmatrix} d\sin(aw_{n,x}) - \sin(bw_{n,y}) \\ 
-          %                                                             c\cos(aw_{n,x})+\cos(bw_{n,y}) \end{bmatrix}$} & & & \\
-          %   \multirow{3}{*}{$u_n = w_n-\overline{w}$}
-          %       & 10 & -0.2243 & 0.49698 \\
-          %       & 100 & &  0.63310 \\
-          %       & 1000 & & 0.71151 \\
-          % \hline
-          \hline
-          $(w_n)$ comes from the DP \\ (for ): & & & \\
-          \multirow{3}{*}{$u_n = w_n-\overline{w}$}
-                & 10 & 0.99141 & 0.98916\\
-                & 100 & &  0.99238 \\
-                & 1000 & & 0.99278 \\
-          \hline        
-      \end{tabular}} %\ednote{http://paulbourke.net/fractals/peterdejong/}
-
-      \caption{Following \cite{manjunath2021universal}, we determine the multidimensional correlation coefficients $\rho$ to evidence a general reduction in the functional complexity of the map $G_T$ emerging due to a RNN.
-Rows correspond to distinct dynamical systems considered in the experiments
-      The second column corresponds to the amount of artificial neurons (the dimension of $X$) in the implemented RNN.
-      The last two columns comprise numerical estimates of $\rho$ for the relevant vectors. The linear association between $u_n$ vs $T(u_{n})$, as well as the linear relationship $\begin{bmatrix}
-              x_{n-1}\\
-              x_n
-          \end{bmatrix}$ vs $G_T \left( \begin{bmatrix}
-              x_{n-1}\\
-              x_n
-          \end{bmatrix}\right)$ may be contrasted by the reader. 
-          Finally, the RNN's dimension (from \eqref{Seq_RNN}) is provided to exhibiting that an increase in dimension of the driven system will 
-          typically result in a map $G_T$ with less functional complexity.} 
-          \label{Table_FC}
-      \end{table}\label{tbl_attractorsPearson}
-\end{center}
-\ednote{B: Functional Complexity will be explained before this}
-
-\chapter{Conclusion}\label{ch6}
-
-The driven states $(x_{n-1},x_n)$ holds the information of the left-infinite history of the input.
+% The driven states $(x_{n-1},x_n)$ holds the information of the left-infinite history of the input.
 Furthermore, there is some evidence that the inverse-limit space (not system) seems to distinguish two dynamical systems that are not topologically conjugate for a class of systems in the sense their inverse-limit spaces are not homeomorphic to each other.
 So the relation $Y_T$, which is homeomorphic to the inverse-limit system, may be unique to a class of systems.
-(The reference is a paper on Ingram’s Conjecture by Marcy Barge, Henk Bruin and Sonja Stimac.) Further investigation is needed.
-
-\ednote{B: Expand}
+(The reference is a paper on Ingram’s Conjecture by Marcy Barge, Henk Bruin and Sonja Stimac.
+) Further investigation is needed.
 

### 2022-01-12 Gandhi_Project_Report/draft3.tex
fiddling with bib loading

@@
 
-\textbf{NB NB FINISH}
-\begin{Example}
-  For example $\phi(x)=\sin(\frac{\pi}{2}x)^2$ is a conjugacy between two systems $Tu=1-|2u-1|$ and $Sv=4v(1-v)$  defined on $U=[0,1]$.
-
-  To see this, we consider $\phi\circ{T} = \sin(\frac{\pi}{2}(1-|2u-1|))^2$
-
-  and $S\circ\phi = 4\sin(\frac{\pi}{2}x)^{2}(1-\sin(\frac{\pi}{2}x)^2)$
+\begin{Example}\rm
+  For example $\phi(x)=\sin\big(\frac{\pi}{2}x\big)^2$ is a conjugacy between two systems $Tu=1-|2u-1|$ and $Sv=4v(1-v)$  defined on $U=[0,1]$.
+
+  Indeed, $\phi:[0,1]\to[0,1]$ is a homeomorphism and by employing simple trigonometric identities, it is easily proved that $\phi$ satisfies~\ref{eqn_conjugacy}
+  \[
+    \begin{aligned}[t]
+      (\phi\circ{T})u 
+                &=\sin^{2}\Big(\frac{\pi}{2} - \frac{\pi}{2}|2u-1|\Big) = \sin^{2}(\pi{u})=\Big(2\sin(\frac{\pi}{2})\cos(\frac{\pi}{2})\Big)^{2}=4\sin^{2}(\frac{\pi}{2}u)\Big(1-\sin^{2}\frac{\pi}{2}u\Big) \\
+                                    &=({S}\circ\phi)u.
+    \end{aligned}
+\]
+And it follows that $\phi$ is a conjugacy between the two systems.
 \end{Example}
@@
 $\Phi_{k,\theta}(w) := (\theta(T^{-k}w)\ldots,\theta(T^{-1}w),\theta(w))$.  
-The essence of Takens theorem relates to the fact that when $k$ is sufficiently large, we can define a dynamical system on the space $\mathbb{R}^{k+1}$ whose states are $\Phi_{k,\theta}(w)$, $\Phi_{k,\theta}(Tw)$, $\Phi_{k,\theta}(T^2w)$, etc.
+The concept of an observable is understood in the sense of an observable as originally introduced by \cite{takens1981detecting, genericObservableAeyels} to refer to a probe or measurement function inserted into the system.
+The essence of Takens' theorem relates to the fact that when $k$ is sufficiently large, we can define a dynamical system on the space $\mathbb{R}^{k+1}$ whose states are $\Phi_{k,\theta}(w)$, $\Phi_{k,\theta}(Tw)$, $\Phi_{k,\theta}(T^2w)$, etc.
 and this dynamical system is topologically conjugate to the unknown underlying system $(W,T)$.
@@
 \begin{figure}[ht]
-  \includegraphics[scale=0.25]{_takensmap.png}
+  \includegraphics[scale=0.25]{_takensmap.eps}
   \centering
@@
 % \begin{figure}[ht]
-%   \includegraphics[scale=0.3]{_learningfailure.png}
+%   \includegraphics[scale=0.3]{_learningfailure.eps}
 %   \centering
@@
 If $H:\overleftarrow{U}{\to}{\overleftarrow{X}_U}$ is also injective (in addition to being surjective, and hence bijective), 
-it becomes the embedding of $\overleftarrow{U}$ in $\overleftarrow{X}$ induced by the driven system $g$ and we refer to $H$ as a \emph{causal embedding} 
-\ednote{M : Causal embedding a dynamical system is different from the definition of causal embedding, so please edit/delete this paragraph and the next theorem and also in the remaining part till Section 1.2}.
-
-\begin{Theorem}
- If $g(\ldotp,x)$ is invertible and has the USP then $H$ is a causal embedding. 
-\end{Theorem}
-\begin{proof}
-  See \cite{manjunath2021universal}.
-\end{proof}
-
-When $g$ has the USP, the diagram below illustrates the operation of the mappings $h$ and $H$.
-The mapping $h:\overleftarrow{U}\to{X_U}$ is an observable as discussed in an above section (\ref{TakensEmbedding}).
+it becomes the embedding of $\overleftarrow{U}$ in $\overleftarrow{X}$ induced by the driven system $g$. %and we refer to $H$ as a \emph{causal embedding} 
+% \ednote{M : Causal embedding a dynamical system is different from the definition of causal embedding, so please edit/delete this paragraph and the next theorem and also in the remaining part till Section 1.2}.
+
+% \begin{Theorem}
+%  If $g(\ldotp,x)$ is invertible and has the USP then $H$ is a causal embedding. 
+% \end{Theorem}
+% \begin{proof}
+%   See \cite{manjunath2021universal}.
+% \end{proof}
+
+When $g$ has the USP, the diagram below (adopted from~\cite{Manju_Nonlinearity}) illustrates the operation of the mappings $h$ and $H$.
+The mapping $h:\overleftarrow{U}\to{X_U}$ is also considered an observable as mentioned in the introduction of Chapter \ref{ch3}).
 
 \begin{figure}[ht]
-  %\includegraphics[scale=0.3]{actionofh_H.png}
-  \centering
+  \includegraphics[scale=0.3]{_actionofh_H.png}
+  \centering
+  \captionof{figure}{Progressive actions of $h$ and $H$ to illustrate the causal mapping. }
   \label{fig:actionh_H}
@@
 \begin{figure}[ht]
-  \includegraphics[scale=0.4]{_memloss_conttime.png}
+  \includegraphics[scale=0.4]{_memloss_conttime.eps}
   \centering
 \captionof{figure}{Reproduced from \cite{manunath2021universal}. Illustration of all trajectories approaching the true solution if $g$ has the UAP}
-\label{fig:memloss_conttime.png} 
+\label{fig:memloss_conttime} 
 \end{figure}
@@
 
-One can therefore learn the SDD of the driven system states via $G_T$ with enough data thanks to the USP/UAP. This enables us to do at least 2 things: 
+One can therefore learn the SDD of the driven system states via $G_T$ with enough data thanks to the USP/UAP. This enables us to do at least two things: 
 \vspace{-8mm}
-\begin{itemize}[noitemsep]
+\begin{itemize}
 \item Forecast future values of $x_n$ via iterates of $G_T$ (as $G_T$ can be determined.
@@
 \begin{figure}[ht]
-  \includegraphics[scale=0.3]{_summarypictorial.png}
+  \includegraphics[scale=0.3]{_summarypictorial.eps}
   \centering
@@
 % \begin{figure}[ht]
-%   \includegraphics[scale=0.3]{_gammatemp.png}
+%   \includegraphics[scale=0.3]{_gammatemp.eps}
 %   \centering
@@
 \begin{figure}[ht]
-  \includegraphics[scale=0.35]{_dp_setup.png}
+  \includegraphics[scale=0.35]{_dp_setup.eps}
   \centering

### 2022-01-13 Gandhi_Project_Report/draft3.tex
Nearly done with changes

@@
 
-\title{Project Report}
-
-\author{} 
-\date{}
+\title{Data-Driven Modelling \& \\ Forecasting of Chaotic Systems}
+
+\author{B.L. Nortier} 
+\date{January, 2022}
 \date{}
@@
 \tableofcontents
+
+% \begin{abstract}
+%   Abstract goes here.
+%   \end{abstract}
 
@@
 The more intricate details of proofs are referred to where relevant throughout. This report is organised as follows: 
-\newline In Chapter~\ref{ch2}, we recall the definition of a discrete-time dynamical system, how a discrete-system arises from a flow of an ODE and then proceed to define the inverse limit space and topological conjugacy of autonomous systems.
+\newline In Chapter~\ref{ch2}, we recall the definition of a discrete-time dynamical system, how a discrete-system arises from a flow of an ODE and then proceed to define the inverse-limit space and topological conjugacy of autonomous systems.
 \newline In Chapter~\ref{ch3}, we introduce the problem of forecasting dynamical systems, state the Takens delay embedding theorem and discuss various issues faced while forecasting. 
@@
 
-\section{Taken Embedding Theorem}
+\section{Takens Embedding Theorem}\label{sect_Takens}
 
@@
 % \end{figure}
+
+Though not a uniquely related to the Takens embedding theorem, we do opt to make mention of this requirement for our work. 
+The methods described here are applicable for data originating from a surjective map. 
+This follows from the fact that when $T$ maps the space $U$ into a proper subset $B$ of $U$, we me conclude that $T$ is not surjective as there exists some point in $U-A$ which does not have a preimage.
+An example of this would be the case when a system displays energy-loss.
+Consider, for instance, consider the oscillations dying down for a damped pendulum.
+(A more in-depth account will be given in \ref{ch5})
+
+This requirement is not overly restrictive, as a great number of chaotic systems do have a surjective map.
+(Cite) \ednote{I rememeber you saying this in one of our sessions, but don't have a source for it.}
 
@@
   [\bf Entire Solution] \label{Dfn_Soln} \rm
-  A sequence $\{x_n\}_{n\in\mathbb{Z}}\subset X$ is called an entire solution(or simply a solution) of the driven system  $g$ with input $\overline{u}$ when it satisfies 
+  A sequence $\{x_n\}_{n\in\mathbb{Z}}\subset X$ is called an entire solution (or simply a solution) of the driven system  $g$ with input $\overline{u}$ when it satisfies 
   \[g(u_{n-1}, x_{n-1})=x_n\] for all $n\in\mathbb{Z}$
@@
   A driven system $g$ is said to have the Unique Solution Property (USP) if for each input $\overline{u}$ there exists exactly one solution. 
-  Alternatively we may formulate the USP as follows: $g$ has the Unique Solution Property if there exists a well-defined map $\Psi:{U}\to{X}$ with $\Psi({\overline{u}})$ denoting the unique solution.
-\end{Definition}
-
-One of the first result obtained upon defining the USP, is the fact that every solution will attract any different initial conditions towards the component parts of the solution.
+  Alternatively we may formulate the USP as follows: $g$ has the Unique Solution Property if there exists a well-defined map $\Psi:{U}\to{X}$, with $\Psi({\overline{u}})$ denoting the unique solution.
+\end{Definition}
+
+One of the first result obtained after defining the USP is the fact that every solution will attract different initial conditions towards the component parts of the solution.
 %, i.e.
@@
 \ednote{Define Uniform Attractor before stating theorem or else omit the statement.}
-If $g$ has the USP, then any solution to $g$ is a uniform attractor in the nonautonomous dynamical systems literature \cite{Manju_Nonlinearity}.
-The discussion on nonautonomous attractors is beyond the scope of this thesis and we refer the reader to \cite{Manju_ESP, esann2012ids} for further reading.
+If $g$ has the USP, then any solution to $g$ is also a uniform attractor in nonautonomous dynamical systems literature \cite{Manju_Nonlinearity}.
+The discussion on nonautonomous attractors is beyond the scope of this project and we refer the reader to \cite{Manju_ESP, esann2012ids} for further reading.
 
@@
 % Paraphrased.
- In our context, $g$ having the USP is equivalent to $g$ being a topological contraction \cite{manjunath2021universal}, or alternatively that $g$ exhibits the Echo State Property (ESP).
-
-This property is used to describe the concept of fading memory or memory-loss in the majority of literature within the field reservoir computing (RC) \textbf{(Citation)}. 
-$g$ having the ESP guarantees that the entire left-infinite history of an input will determine exactly the current system state; there is only ever the possibility of a single reachable state at a given instance of time \cite{jaeger2001echo}
-The ESP is also often discussed in terms of a stability property that holds for all inputs of a system~\cite{manjunath2020stability} and, furthermore, plays a key role in the design and training of recurrent neural networks within the field of RC \cite{Manju_ESP}.
-
-Defined formally, the ESP is intrinsically linked to an input (see \cite{Manju_ESP}) Unfortunately, definitions and results concerning the ESP do not have much practical use since a relationship of the network's response to the temporo-statistical properties of the input is not directly described.
+Additionally, $g$ having the USP is in our context also equivalent to $g$ being a topological contraction \cite{manjunath2021universal}, or alternatively also that $g$ exhibits the Echo State Property (ESP).
+
+In the majority of Reservoir Computing(RC) literature (a machine learning algorithm approach)\ednote{How should I explain/qualify RC better here? It's mentioned without really explaining.}, 
+the ESP is used to describe the concept of fading memory or memory-loss  \cite{jaeger2004harnessing, lu2018attractor}. 
+If $g$ possesses the ESP (equivalent in our context to the USP), the we are guaranteed that the whole of an input's left-infinite history will exactly determine the current system state; i.e.
+there is only ever the possibility of a single reachable state at a given instance of time \cite{jaeger2001echo}
+
+Defined formally, the ESP is intrinsically linked to an input (see \cite{Manju_ESP}) Unfortunately, definitions and results concerning the ESP do not have much practical benefit since a relationship between the network's response and the temporo-statistical properties of the input are not directly described.
 A more detailed presentation may be found in~\cite{jaeger2001echo}.
+
+We do however consider it important to mention that the ESP is also often discussed in terms of a stability property that holds for all inputs of a system~\cite{manjunath2020stability} and, furthermore, plays a key role in the design and training of recurrent neural networks(RNN) within the field of RC \cite{Manju_ESP}.
 
@@
 
-Having already defined the reachable set $X_U$, we now pause for a moment to fix additional notation.
-Letting $\cev{u}^{n}:=(\ldots,u_{n-2} ,u_{n-1})$ be the left-infinite subsequence of an input, $\overleftarrow{U}$ then denotes all these left-infinite sequences in $U$. 
-Moreover, $\cev{u}^{n}v:=(\ldots,u_{n-2} ,u_{n-1}, v)$ is to symbolise the input up to time $n$ with $v \in U$ being the specific input value at time $n$. 
-The introduction of a new input at time $n$ can be described by a mapping $\sigma_v:   \cev{u}^{n} \mapsto \cev{u}^{n}v$.
-
-The question now becomes whether we may establish a semi-conjugacy %\ednote{M: I have changed it to a semi-conjugacy since conjugacy amounts to asking a infinite-dimensional space $\overleftarrow{U}$ to be embedded in $X_U$ which is not possible.}
-as presented below for the driven system $g$.
+Having already defined the reachable set $X_U$ as the set of all elements of all entire solutions, we pause briefly in order to fix additional notation.
+Defining $\cev{u}^{n}:=(\ldots,u_{n-2} ,u_{n-1})$ as the left-infinite subsequence of an input up until time $n$, $\overleftarrow{U}$ is then the notation for all these left-infinite sequences in $U$.
+Moreover, $\cev{u}^{n}v:=(\ldots,u_{n-2} ,u_{n-1}, v)$ will symbolise the input up to time $n$ with $v \in U$ being the specific input value at time $n$. 
+The introduction of a new input at time $n$ can be described by the mapping $\sigma_v:   \cev{u}^{n} \mapsto \cev{u}^{n}v$.
+
+The goal now assumes the form of the question on our ability to establish a semi-conjugacy as presented below for the driven system $g$. 
+%\ednote{M: I have changed it to a semi-conjugacy since conjugacy amounts to asking a infinite-dimensional space $\overleftarrow{U}$ to be embedded in $X_U$ which is not possible.}
 
@@
 
-Continuing to refine our attention, we proceed to consider a specific subclass of conjugacies.
-
-  \begin{Definition} 
-    [\bf Universal Semi-Conjugacy]\label{Def_UnivSemiConj}
+We proceed to consider a specific subclass of conjugacies.
+
+  \begin{Definition}
+    [\bf Universal Semi-Conjugacy]\label{Def_UnivSemiConj} \rm
     Given a driven system $g$, we  call a continuous and surjective map $h : \overleftarrow{U} \to X_U$ a universal semi-conjugacy if  diagram \ref{Scomm_h} commutes for all $v \in U$.
@@
 
-If the universal semi-conjugacy $h$ exists (i.e. the diagram in \ref{Scomm_h} commutes) then the solution $\Psi$ will be at least a coarse-grained representation of the input $u$.
-
-Does such a function $h$ for the driven system defined above exist?
-If $g$ has the USP and $\Psi(u)=\{x_n\}_{n\in\mathbb{Z}}$ then $h$, defined by  $h(\cev{u}_n):=g(u_n,x_{n-1})=x_n$, will satisfy the semi-conjugacy in the graph above \ref{Scomm_h}.
+If the universal semi-conjugacy $h$ exists (i.e.
+the diagram in \ref{Scomm_h} commutes) then the solution $\Psi$ will be at least a coarse-grained representation of the input $u$.
+\ednote{B: Expand/Cite}
+
+But does such a function $h$ for the above driven system $g$ above exist?
+Whenever $g$ has the USP and $\Psi(u)=\{x_n\}_{n\in\mathbb{Z}}$ it follows that $h$, defined by  $h(\cev{u}_n):=g(u_n,x_{n-1})=x_n$, will satisfy the semi-conjugacy in the graph above \ref{Scomm_h}.
 Regrettably, such a continuous mapping $h$ is not guaranteed to exist \cite{M: Have edited here} when $g$ does not have the USP \cite[Lemma 5]{Manju_Nonlinearity}.
@@
   [\bf Causal Mapping]\label{Def_CausMap}
-  A continuous, surjective map $H:\overleftarrow{U}\to\overleftarrow{X}_U$ such that \[H\circ\tilde{g}_v=\sigma\circ{H}\] where $\tilde{g}_v$ maps $(\ldots, u_{-2}, u_{-1})$ to $(\ldots, u_{-2}, u_{-1}, g(v, u_{-1}))$.
+  A continuous, surjective map $H:\overleftarrow{U}\to\overleftarrow{X}_U$ such that \[H\circ\tilde{g}_v=\sigma\circ{H}\] where $\tilde{g}_v$ maps $(\ldots, u_{-2}, u_{-1})$ to $(\ldots, u_{-2}, u_{-1}, g(v, u_{-1}))$ is called a causal mapping.
 \end{Definition}
@@
 \end{proof}
-
-We direct our attention to the function \[H_2{\overline{}} := (h(r\overline{u}, h(\overline{u}))) \]
-
-An embedding of the space $\overleftarrow{U}$  in $\overleftarrow{U}$ would allow one to establish topological conjugacy, which in turn provides stronger results than merely obtaining a coarse-grained representation via a causal mapping \cite{M: I feel this is very confusing to the reader as the co-domain of causal mapping is the left-infinite space and , so please omit if possible.
+To establish whether $\widehat{T}$ and $G_T$ are related, we consider the the function $H_2{\overline{}} := (h(r\overline{u}, h(\overline{u})))$.
+$H_2$ send the entirety of a left-infinite sequence to some element of $X\times{X}$.
+
+% An embedding of the space $\overleftarrow{U}$  in $\overleftarrow{U}$ would allow one to establish topological conjugacy, which in turn provides stronger results than merely obtaining a coarse-grained representation via a causal mapping \ednote{M: I feel this is very confusing to the reader as the co-domain of causal mapping is the left-infinite space and , so please omit if possible.
 You can just say we want to restrict inputs to the left-infinite orbits of a dynamical system and then explore conjugacy instead of a mere semi-conjugacy}.
 
-Before formalising this, we make mention of the concept of the inverse-limit system of a dynamical system $(U,T)$. 
+% Before formalising this, we make mention of the concept of the inverse-limit system of a dynamical system $(U,T)$. 
+We would like to restrict inputs to the left-infinite orbits of a dynamical system and accordingly explore the possibility of establishing a conjugacy instead of merely obtaining a semi-conjugacy.
 These spaces are noteworthy as their topology can help one ascertain much about the original system \cite{Manju_IEEE}.
@@
 
-One must be careful to avoid a choice of $g$ which would  add complexity to the obtained solution.
+The driving function $g$ cannot be chosen in a frivolous manner as it could possibly complicate our work.
+Specifically, one must be careful to avoid a choice of $g$ which would  add complexity to the obtained solution.
 
@@
   A map $g$ is said to be SI-Invertible if $g$ is invertible for all $x\in{X}$.
-Alternatively it may be said that if, given $x_n$ and $x_{n-1}$, $u_n$ can be uniquely determined from $x_n=g(u_n,x_{n-1})$, then $g$ is said to be SI-invertible.
+Alternatively it may be said that if, given $x_n$ and $x_{n-1}$, $u_{n-1}$ can be uniquely determined from $x_n=g(u_n,x_{n-1})$, then $g$ is said to be SI-invertible.
 \end{Definition}
@@
 \begin{equation} \label{eqn_SI_RNN}
-  u_n := A^{-1}\bigg(\overline{\tanh}^{^{-1}}\frac{1}{a}\Big(x_{n+1}-(1-a)x_n\Big) \bigg) - \alpha B x_n
+  u_{n-1} := A^{-1}\bigg(\overline{\tanh}^{^{-1}}\frac{1}{a}\Big(x_{n+1}-(1-a)x_n\Big) \bigg) - \alpha B x_n
   \end{equation}
@@
 
-\begin{equation} \label{Fig_inverse_limsystem}
+\begin{equation} \label{fig:inverse_limsystem}
   %  \[ 
@@
 
-The graph in equation \ref{Fig_inverse_limsystem} is then extended as below:
+The graph in equation \ref{fig:inverse_limsystem} is then extended as below:
 
@@
 Recall that $g$ is SI-invertible.
-%Adopting the terminology of \cite{manjunath2021universal}, $A$ is called the “input matrix” and $B$ the “reservoir matrix”.
+Adopting the terminology of \cite{manjunath2021universal}, $A$ is called the “input matrix” and $B$ the “reservoir matrix”.
 
@@
 For information, (\textbf{Cite, Cite, Cite}) may be consulted.
-We implement our RNN in the \emph{Python} programming language and making use of the \emph{Scipy} and \emph{Tensorflow} libraries.
+We implement our RNN in the \emph{Python} programming language and making use of the \emph{Scipy} library, alongside the \emph{Keras} package in the \emph{Tensorflow} library.
 
@@
 
-We made several changes which gave improved results:
+To understand the first and most drastic\ednote{change word} change, we recall the the conjugacy in~\ref{fig:inverse_limsystem}:
+
+\begin{equation*}
+  %  \[ 
+      \psset{arrows=->, arrowinset=0.25, linewidth=0.6pt, nodesep=3pt, labelsep=2pt, rowsep=0.7cm, colsep = 1.1cm, shortput =tablr}
+      \everypsbox{\scriptstyle}
+      \begin{psmatrix}
+      \widehat{U}_T  && \widehat{U}_T \\
+      Y_T && Y_T \\
+      %%%
+     %  \ncline{1,1}{1,2}^{\widehat{T}} \ncline{1,1}{2,1} <{h}
+     %  \ncline{1,2}{2,2} > {H_2}
+     %  \ncline{2,1}{2,2}^{G_T}
+      \end{psmatrix}
+    %  \]
+  \end{equation*}
+
+  We recall that $G_T$ is topologically conjugate to $\widehat{T}$, a result which we have already made use of.
+
+  The authors of~\cite{manjunath2021universal} did not, however, consider the relationship between the inverse-limit space $\widehat{U}_T$ and the input $u_n=\theta(w_n)$ (where $\theta$ is our observable and $w_n$ the complete input), a concept we now explore.
+  We may construct a delay-coordinate $v_n=[\theta(w_n), \theta(w_{n-1}), \ldots, \theta(w_{n-d})]^{T} = [u_n, u_{n-1}, \ldots, u_{n-d}]^T$
+    to accordingly obtain a new dynamical system $(V,F)$ where $F$ is the map $F:v_n\mapsto{v_{n+1}}$.
+If $d$ is sufficiently large, we may apply Takens Embedding Theorem~\ref{Thm_Takens} to establish the existence of a topological conjugacy between the new system $(V,F)$ and the original system $(U,T)$.
+
+  Using the same notation as before, we may construct a new conjugacy which now relates delay coordinates to one another.
+
+  \begin{equation}\label{eqn_VFdelay}
+    %  \[ 
+        \psset{arrows=->, arrowinset=0.25, linewidth=0.6pt, nodesep=3pt, labelsep=2pt, rowsep=0.7cm, colsep = 1.1cm, shortput =tablr}
+        \everypsbox{\scriptstyle}
+        \begin{psmatrix}
+        \widehat{V}_F  && \widehat{V}_F \\
+        Y_F && Y_F \\
+        %%%
+       %  \ncline{1,1}{1,2}^{\widehat{F}} \ncline{1,1}{2,1} <{h}
+       %  \ncline{1,2}{2,2} > {*}
+       %  \ncline{2,1}{2,2}^{G_F}
+        \end{psmatrix}
+      %  \]
+    \end{equation}
+
+The existence of $\widehat{F}$ is guaranteed, but the map $v_{n}\mapsto{v_{n+1}}$ itself is unknown.
+We may, however, make use of the SI-invertibility of the function $g(u_{n-1}, x_{n-1})=x_n$ to conclude that
+$(x_{n-1}, x_{n})$ determines $u_{n-1}$ uniquely.
+
+Accordingly $(x_{n-2}, x_{n-1}, x_{n})$ determines $u_{n-2}$ and $u_{n-1}$ uniquely and generally $(x_{n-d}, \ldots , x_{n-1}, x_{n})$ determines $u_{n-d}, \ldots u_{n-2}$ and $u_{n-1}$ uniquely.
+
+We therefore need not know what $\widehat{F}$ is, as we can determine $v_{n+1}={[u_n, u_{n-1}, \ldots, u_{n-d}]}^T$ indirectly by determining the value of $(x_{n-d}, \ldots , x_{n-1}, x_{n})$, a much more feasible task.
+
+Accordingly we define the following abbreviations:
+\begin{itemize}
+  \item Delay Coordinates (DC): A vector composed by appending a fixed amount of previous values to the final value.
+  \item Single-Delay Dynamics(SDD): Dynamics generated by 2-element delay-coordinate vectors of the form $(x_{n-1}, x_n)$ where only one previous value is used.
+  \item Multi-Delay Dynamics(MDD): Dynamics generated by arbitrary-length delay-coordinate vectors of the form $(x_{n-d}, \ldots , x_{n-1}, x_{n})$.
+($d$ must be sufficiently large, but is unknown.
+We vary this parameter in experimentation.)
+\end{itemize}
+
+Within this framework, the PCA-implementation of~\cite{manjunath2021universal} is then considered to fall underneath the of auspices of what we term \emph{Approximated Multi-Delay Dynamics(AMDD)}.
+
+We have now seen that one may comfortably switch between the use of SDD or MDD when $g$ has SI-invertibility. /ednote{This sentence needs to be improved.}
+
+Naturally, one would immediately suppose that the MDD approach would increased computation time and resource use as the internal matrices and datasets generated during learning would grow with every increase in the delay-parameter $d$.
+
+In fact, preliminary results indicate the opposite due the observation that the implementation of MDD is inversely proportional to the the dimension of the network (i.e.
+the number of neurons in the RNN) and possibly also the number of epochs/batch size necessary to produce reliable predictions.
+This, in turn, actually speeds up the computation-time and provides results of similar accuracy.
+\ednote{I did not have enough time to explore this properly.
+I can only hazard guesses and promise future research work in this project.
+How should I best convey this?
+}
+
+\textbf{Secondly}
+
+Sparse Matrices:
 \vspace{-8mm}
-\begin{enumerate}
-\item Using Delay Coordinates (DC) when driving the system vs.
-using Single/Multi Delay Dynamics (SDD/MDD) after driving (and before training).
-Explain/show why these two methods are equivalent.
-\ednote{Would MDD  be not computationally intensive ?
-So may be you just explain if you can’t show it.}
-\item Sparse matrix implementation.
+\begin{itemize}[noitemsep]
+\item Well-conditioned: condition number close to 1
+\item Scaled
+\item Sparsity/density ratio
+\end{itemize}
+
+% Sparse matrix implementation.
 Slight improvement in some scenarios, but not fantastic.
@@
 
-(Possible paragraph here/later on implementing sparsity of A/B: Generate sparse matrix that is also well-conditioned and correctly scaled matrix.)
-\item Density of matrices A/B(i.e. number of non-zero entries in each matrix) (separate from sparse matrix implementation)
-\end{enumerate}
+\textbf{Third}
+
+Density of matrices A/B(i.e. number of non-zero entries in each matrix) (separate from sparse matrix implementation)
 
@@
 Noise-free, undamped DP data-sets were generated and utilised in this project due to the requirement that the map $T$ be surjective.
-It may easily be verified that a DP undergoing damping (as seen in \ref{conjugacy_with_Gamma}) is not surjective and hence the required assumptions to apply our theory do not hold.
-\ednote{B: Change sentence.}
+We discuss the case of a DP exhibiting energy-loss (i.e.
+where the map $T$ is not surjective) below.
+
+\textbf{Data originating from a non-surjective map.}
+
+The importance of the requirement that data originates from a surjective map is observed learning failure of a map describing a damped DP.
+Due to the assumption, the program discounted any possibility of energy-loss and we consider the example below.
 
@@
   \centering
-\caption{A damped pendulum is not surjective.}
+\caption{The map generating data for a damped pendulum is not surjective.}
 \label{fig:damped_pendulum}
@@
 
-\textbf{Parameter Values \& Successes}
-
-Train length: 15000. Prediction Length: 1000. Discard to wait for network to forget: 500. Dimension of network: 128 neurons.
-$\alpha=0.99$. $a=0.5$.
-
-Hidden Layers: 16. Layer dimension: 64. Epochs:256. Batch Size: 128. Delay(for MDD): 8. Density of Sparse Matrices: 0.2. $d=1$
-
-\ednote{B: Write this paragraph.}
+Despite feeding data that originated from a DP with dying occilations into the system, a map was instead learnt which continued to oscillate in a chaotic, but conservative manner.
 
 \begin{figure}[ht]
-  \centering
-  \includegraphics[width=\linewidth]{_DP_Fig-2F2023.eps} 
-  \captionof{figure}{Predicted(red) and actual(blue) trajectories of the $x$- and $y$-coordinates of each of the pendulum heads.} 
+  \includegraphics[scale=0.4]{_dpfailure_nonsurj.eps}
+  \centering
+\caption{Failure to learn a map which is not surjective.}
+\label{fig:dp_notsurjective}
 \end{figure}
+
+Having spent ample time to discuss the challenges, we now change tone so as to satisfy the optimist.
+We present successful run for DP-data to repdict the future movement of the pendulum.
+
+The system was driven for 15500 steps, of which the first 500 were immediately discarded to simulate the network 'forgetting'.
+The driving function $g$ had input and reservoir matrices were implemented as sparse matrices filled to 20\%.
+Parameter-values of $\alpha=0.99$ and $a=0.5$ were used.
+The remaining 15000 datapoints were used to train a network with 128 neurons. 
+The network was initialised to contain 16 hidden layers each with a layer dimension of 64, activated with the tanh activation function.
+Training was accomplished using 256 epochs with batch sizes 128 and a delay of 1.
+
+% Trainlen:15000; Predlen: 1000; Discard: 500; Dimension:128 neurons; $\alpha=0.99$; $a=0.5$.
+Hidden Layers: 16.
+Layer dimension: 64.
+Epochs:256.
+Batch Size: 128.
+Delay(for MDD): 8.
+Density of Sparse Matrices: 0.2.
+$d=1$
+
 \begin{figure}[ht]
   \centering
-  \includegraphics[width=\linewidth]{_DP_Fig-2B2023.eps}
-  \captionof{figure}{Densities of predicted and actual values of x-coordinate of first pendulum head. }
+  \includegraphics[width=0.8\linewidth]{_DP_Fig-2F2023.eps} 
+  \captionof*{figure}{Predicted(red) and actual(blue) trajectories of the $x$- and $y$-coordinates of each of the pendulum heads.} 
+% \end{figure}
+% \begin{figure}[ht]
+%   \centering
+  \includegraphics[width=0.8\linewidth]{_DP_Fig-2B2023.eps}
+  \captionof*{figure}{Densities of predicted and actual values of x-coordinate of first pendulum head. }
+  % \end{figure}
+  % \begin{figure}[ht]
+  %   \centering 
+  \includegraphics[width=0.8\linewidth]{_DP_Fig-2D_5.eps}
+    \captionof*{figure}{Takens Delay Map with delay=5 to illustrate behaviour of predicted and actual values of double pendulum data.}
+    \label{fig:dp_predictions}
 \end{figure}
-  \begin{figure}[ht]
-    \centering 
-  \includegraphics[width=\linewidth]{_DP_Fig-2D_5.eps}
-    \captionof{figure}{Takens Delay Map with delay=5 to illustrate behaviour of predicted and actual values of double pendulum data.}
-\end{figure}
-
-\subsection{Other attractors I’ve experimented with: Clifford \& Thomas}
+
+\subsection{Additional Attractors: Clifford \& Thomas}
 We follow the methodology presented in \cite{manjunath2021universal} but opt to consider attractors different from the Lorenz system and H\'enon map in this project.
@@
 
-The behaviour of the system with a $\beta$-parameter value of 0.1056 is simulated by scaling it to fit inside the interval $[-1,1]$ and both perturbed with noise as well as leaving it intact.
-Specifically the noise is obtained from a normally distribution with mean 0 and standard deviation 0.05, which translates approximately to *dB.
- A sequence of 3000 observations in the clear and noisy data-set respectively is fed into the system with 300 entries discarded before predicting the next 9000 entries.
+The behaviour of the system with a $\beta$-parameter value of 0.1056 was simulated by scaling it to fit inside the interval $[-1,1]$.
+Two cases were considered: noise-free and perturbed.
+Noise was generated from a normal distribution with mean 0 and standard deviation 0.05, which translates approximately to *dB.
+ A sequence of 3000 observations in both the clear and noisy data-sets respectively were fed into the system with 300 entries discarded to simulate the network's loss of memory.
+The next 9000 steps were predicted and compared with true data.
 
@@
 
-The RNN was initialised as follows: The dimension of the network is set equal to 500 neurons and 12 hidden layers of dimension 64 each were initialised.
+The RNN was initialised as follows: the dimension of the network is set equal to 500 neurons and 12 hidden layers of dimension 64 each were initialised.
 Training is realised through 150 epochs of batch sizes of 128 each; the map is learnt via the Adam Optimiser using finer and finer learning rates (0.001 and then divided by 10 at every iteration).
@@
 }
-  \label{Fig_Clifford}
+  \label{fig:Clifford}
 \end{figure}
@@
 \end{center}
-\ednote{B: Functional Complexity will be explained before this}
 
@@
 
-\ednote{B: Expand}
-
 \end{document}

### 2022-01-13 Gandhi_Project_Report/ch4.tex
Changed 4.3 and H2. Conclusion writtend and added

--- 2022-01-13 Gandhi_Project_Report/draft3.tex
@@
 \usepackage[export]{adjustbox}
+% $f\arrowvert_A$ To restrict functions
 
@@
 \maketitle
-\tableofcontents
+% \tableofcontents
 
@@
 %   \end{abstract}
-
-\chapter{Introduction}\label{ch1}
-
-Experimenting on biological, physical, and artificial systems in order to generate a more informative dynamical model is a well-established practice in modern science.
-Traditional methods for modelling physical systems are based on laws of physics that are based either on empirical relationships or on intuition.
-For systems that evolve with time, physical laws yield mathematical equations that govern how the quantities evolve with time.
-However our world around is, for the most part, much more complex than that which can be distilled into elegant equations.
-We do not fully understand many of the more complex systems, nor do they even provide us with a good physical intuition of the underlying principles governing the dynamics.
-To complicate this - the underlying systems we observe often display sensitive dependence on initial conditions; despite having highly similar initial conditions, differing orbits diverge quickly and to such an extent that it becomes seemingly impossible to retrace their steps back to the original conditions.
-Even the presence of usually 'negligible' computational noise or measurement error renders long-term point-wise prediction infeasible.
-There are many difficulties that one encounters while modelling such systems:
-\vspace{-8mm}
-\begin{enumerate}[noitemsep, label=\roman*.]
-  \item One may not have access to the complete states of the systems
-  \item The actual system may be described by functions that behave wildly in the sense their graphs have a wild oscillatory behaviour, i.e., they have a large functional complexity.    
-\end{enumerate}
-
-Models derived primarily from data can be classified into three categories: 
-\vspace{-8mm}
-\begin{enumerate}[noitemsep, label=\roman*.]
-  \item  Interpretable models (i.e., they establish relationships between internal physical mechanisms), 
-  \item Partially interpretable models capturing some modes of the dynamics, 
-  \item Non-Interpretable models, defined as such mainly due to the fact that they are defined on a different phase space that is usually high-dimensional. 
-\end{enumerate}
-
-Examples in the literature attempting to forecast data from such systems have tried a number of different approaches with differing degrees of success. 
-When we have complete access to states of the system, an ordinary differential equation can be obtained from data (\cite{brunton2016discovering, champion2019data} and \cite{small2002modeling,xu2006modeling}) wherein one could approximate the vector field by a library of functions to obtain interpretable models.
-
-Recent partially interpretable models available in the literature have been based on the Koopman operator (e.g.,\cite{koopman1932dynamical,budivsic2012applied}) to  employ observables mapping the data onto a higher-dimensional space.
-This then makes the dynamics in the higher-dimensional space more amenable for approximation by a linear transformation.
-Such methods, obviously, do not guarantee an exact reconstructions for nonlinear models, and in practice provide poor long-term consistency for a large class of systems.
-\textbf{(Citation)}
-
-The non-interpretable models include the delay embedding and the machine learning algorithms.
-One could learn a system conjugate to the underlying system by applying the Takens delay embedding \cite{takens1981detecting} when one has ``good" observations from the system; this learnt system could then be used for forecast the observed data.
-Takens delay embedding theorem \cite{takens1981detecting} and its various generalisations (e.g., \cite{sauer1991embedology, stark1999delay, gutman2018embedding}) establishes the learnability of a system constructed by concatenating sufficiently large previous time-series observations of a dynamical system into a vector (called delay coordinates).
-This then establishes the existence of a map on the space of delay coordinates equivalent (or topologically conjugate) to the underlying map from which the observed time-series was first obtained.
-Although topological conjugacy guarantees an alternate representation of the underlying system, the quality of this representation still depends on numerous parameters, making the comprehension of the dynamics rather unreliable (e.g., \cite{principe1992prediction}).
-One reason for this fragility is that the embedded attractor in the reconstruction space not always an attractor of the map that is learnt in the reconstruction space, despite  unquestionably being an invariant set.
-When the embedded object is not explicitly known to be an attractor (as explained in Chapter \ref{ch3}) it can cause predictions to fail.
-
-Practically, the application of Takens embedding involves learning a map through some technique and consequently one wishes that these would have low functional complexity\cite{manjunath2021universal}, i.e., functions with fewer oscillatory graphs.
-Pure machine learning methodology that processes temporal information (like the echo state networks (Citations to echo state networks)) by mapping data onto a higher dimensional space for further processing.
-Although they perform  well on forecasting some dynamical systems, they fail completely on others (Citations here) as there is often no guarantee that the right function was learnt during training.
-
-This thesis deals with the implementation and analysis of non-interpretable models that can guarantee exact reconstruction.
-The project work concerns the study and implementation of a method (\cite{manjunath2021universal}) that incorporates learning a function by mapping the data on to a higher dimensional space using what is called a driven dynamical system (See Chapter 4).
-With a clear understanding of how the data is mapped onto the higher dimensional space, the method then permits the learning of a dynamical system topologically conjugate to that of the underlying system.
-Instead of linear regression as in the training of echo state networks, deep learning methods are employed to learn the correct function.
-With slight modifications to the implementation in the paper (with Adriaan), we show that one can construct accurate non-interpretable models with the ability to reconstruct attractors from more hard-to-forecast systems like the double pendulum.
-(The forecasting of the time-series from a double pendulum has not been reported before.) Moreover, we also demonstrate that long-term statistical consistency is preserved.
-
-\emph{By solidifying the mathematical underpinnings of our theory, we hope to guarantee the ability to construct models with predictive power ranging from molecular biology to neuroscience.
-in the near future.
-(We can modify this sentence when the report is completed.)}
-
-The more intricate details of proofs are referred to where relevant throughout. This report is organised as follows: 
-\newline In Chapter~\ref{ch2}, we recall the definition of a discrete-time dynamical system, how a discrete-system arises from a flow of an ODE and then proceed to define the inverse-limit space and topological conjugacy of autonomous systems.
-\newline In Chapter~\ref{ch3}, we introduce the problem of forecasting dynamical systems, state the Takens delay embedding theorem and discuss various issues faced while forecasting. 
-\newline In Chapter~\ref{ch4}, a driven dynamical system is defined and discuss the properties of a specific class of driven dynamical systems that we make use of in this thesis/report/paper. 
-\newline Finally in Chapter~\ref{ch5} we show the implementation of these forecasting methods, and conclusions are provided in Chapter~\ref{ch6}.
-
-\chapter{Discrete-time Dynamical Systems}\label{ch2}
-
-In this chapter, we provide a brief description of what a discrete-time dynamical system is and what it means for it to exhibit chaos.
-We refer to \cite{devaney2018introduction, de2013elements} for more details.
-
-At its most elementary level, a dynamical system is just something that evolves deterministically through time.
-In the context of this thesis, deterministic refers to the fact that a system evolves according to specified rules rather than based on random events.
-Dynamical systems arise in a variety of situations.
-A continuous-time dynamical system describes the states for all values of the time.
-Specifically, if the motion of a pendulum in which the quantities such as the angular position and angular momentum are known at all times, then it is a continuous-time dynamical system.
-The equations of the dynamical system can take the form of one or more ordinary differential equations that determine the relevant quantities at any future time if we know the initial location and momentum.
-In ecology, discrete-time dynamical systems are widely used to model population growth.
-The model in this case is a function calculating the following generation's population given the population of the previous generation.
-If we know the starting population, we may once again calculate the population at any time in the future.
-
-Formally, a function $T: U \to U$, where $U$ is some set is  a \emph{discrete-time dynamical system} and its iterates $\{u,Tu,T^2u,\ldots\}$, where $T^n$ denotes the $n$-fold composition of $T$ with itself, describe the evolution of an initial condition $u\in U$ (Note that we frequently drop the brackets and denote $T(u)$ by $Tu$ so as to simplify notations).
-
-Continuous-time dynamical systems modelled using ordinary differential equations can give rise to discrete-time dynamical systems.
-To see this, consider a differential equation $\dot{x} = f(x)$, $f: \mathbb{R}^n \to \mathbb{R}^n$, $n\in\mathbb{N}$ given to have a unique solution passing through
-each point $x\in\mathbb{R}^{n}$, call it $x(t)$ where $x(0)=x_0$.
-
-\begin{Definition}
-  [\bf Flow of an Equation] \label{Dfn_Flow}\rm
-  The flow of the equation  $\dot{x} = f(x)$  is defined to be a mapping $\varphi: \mathbb{R}^n \times \mathbb{R} \to \mathbb{R}$ where $\varphi(x_0,t)= x(t)$ for the solution $x(t)$ with $x(0)=x_0$ 
-\end{Definition}
-
-By fixing $t=K \in (0,\infty)$, we can define the \emph{time-$K$ map} as  $T(x):= \varphi(x_0,K)$, and it is easily verified that $T\circ T(x_0) = \varphi(x_0,2K)$.
-In general the $m^{\mbox{th}}$ iterate of $x_0$ under $T$ would be the value of the solution of the ODE evaluated at time $mK$ with the initial condition $x_0$, i.e.
-$\varphi(x_0, mK)$.
-Thus ordinary differential equations give rise to a discrete-time dynamical system by sampling the value of the solution $x(t)$ at time intervals $K$ units apart.
-
-A numerical discretization of a differential equation can also give rise to a discrete-time dynamical system.
-For instance, Euler's method approximates $\dot{x}(t)$ by $(x(t+h)-x(t))/h$; if $h$ is fixed throughout, the solution of a differential equation $\dot{x}=f(x)$
-can be approximated at the time instant $t+(m+1)h$ by iterating the equation 
-\begin{equation}
-  x(t+(m+1)h) = x(t+mh) + h f(x(t+mh))
-\end{equation}
-
-Adopting more succinct notation by replacing $x(t+mh)$ with $u_m$, we rewrite the above equation as
-\begin{equation}
-u_{m+1} = u_m + hf(u_m)
-\end{equation}
-or in even simpler terms as the discrete-time dynamical system with map $T(u) = u + hf(u)$, where $T$, $u$ and $f$ are understood to be as above.
-
-Of course, discrete-time dynamical systems need not always arise through a differential equation, and once again we may consider the field of ecology where discrete dynamical systems are often directly derived or assumed.
-\textbf{Citation}
-There is a school of thought that advocates discrete-time dynamical systems to be more natural for modelling real-world observations than differential equations and we refer the interested reader to \cite{saber2010introduction}.
-
-\section{Invariant Sets}
-
-A core concept in the study of dynamical systems is that of invariance. 
-\begin{Definition}
-  [\bf Invariant Set]\label{Dfn_InvariantSet}\rm
-  Given a discrete-time dynamical system $T: U \to U$, a subset $A \subset U$ is said to be an \emph{invariant set} if $T(A) =A$. 
-\end{Definition}
-
-We also define the orbit of a function $T$.
-\begin{Definition}
-  [\bf Orbit of $T$]\label{Dfn_Orbit}\rm
-  The orbit of $T$ is to be defined the sequence $\bar{u} = \{u_n\}_{n\in \mathbb{Z}}$ obeying the update equation, $u_{n+1}=Tu_n$, $n \in \mathbb{Z}$. 
-\end{Definition}
-
-Two examples of invariant sets include a fixed point where $Tu=u$ for  $u\in U$, and a periodic orbit, i.e., a set of iterates $\{u,Tu, T^2u,\ldots,T^pu\}$, where $T^{p+1}u=u$ for some $p\in\mathbb{Z}$.
-The entire space $U$ could be also be invariant.
-Consider for example the space $U=[0,1]$, where $Tu=4u(1-u)$ and then $U$ is invariant (as every $u\in{U}$ can be written as $u = 4x(1-x)$ for some $x\in{U}$).
-
-We may learn a great deal about the iterates of a dynamical system by considering the types of invariant sets of a discrete-time dynamical system.
-For example, if $U=[0,1]$ has map $Tu= u/2$, then the only invariant set is $\{0\}$, and all orbits approach this invariant set as time flows in the forward direction.
-Indeed, if a some non-zero invariant set (call it $B$) exists, then there is  some $r\in(0,1]\cap{B}$.
-But $r\notin{T(B)}$ since any orbit with initial value $r$ will be a decreasing sequence.
-Moreover, every orbit of T will be decreasing and therefore approach the value $0$ as $n\rightarrow\infty$.
-
-One may ask if every orbit approaches an invariant set?
-In general the answer is no, since for the dynamical system $T: \mathbb{R} \to \mathbb{R}$ defined by $Tu=2u$, any orbit that does not intersect the invariant set $\{0\}$, will not approach any invariant set.
-
-However,  when the space $U$ is compact, all orbits approach an invariant set.
-This is since the set of limit points of the orbit can be shown to be invariant \cite{de2013elements}.
-So, when $U$ is compact,  the $\omega$-limit set $\omega(u;T)$ of a point $u$ defined to the collection of limit points of the sequence $\{x,Tu,T^2u,\ldots\}$ is nonempty, and $\omega(u;T)$ is invariant.
-
-\begin{Example}
-  For the map $Tu=u^2$ defined on $[0,1]$ all orbits lie in the invariant set $\{1\}$ or else would would approach the invariant set $\{0\}$. 
-  % \ednote{B: not relevant other than being an additional example perhaps? M: Have added another simple example}.
-\end{Example}
-
-Invariant sets have various properties. Invariant sets can be attracting or repelling depending on how orbits in its vicinity behave. 
-Recall the examples $Tu =u/2$ and $Tu=2u$ defined on $\mathbb{R}$ (where $\{0\}$ ``attracts" orbits) or the example, $Tu=u^2$ defined on $[0,1]$ (where $\{1\}$ ``repels" orbits).
-We are interested in attractive invariant sets since they capture the long-term dynamics as time increases.
-In particular we are interested in those invariant sets named attractors.
-
-\begin{Definition}
-  [\bf Attractor]\label{Dfn_Attractor}\rm
-  Let $T: U \to U$, where $U$  is a metric space with metric $d$. A compact subset $A \subset U$ is said to be an attractor if it satisfies the three conditions: 
-  \vspace{-8mm}
-  \begin{enumerate}
-	\item $A$ is invariant. 
-	\item $A$ is asymptotically stable, i.e., for every $\epsilon > 0$ and for all $u$ so that $d(u,A) < \epsilon$, we have $d(T^nu,A) \to 0$ as $n\to \infty$. 
-	\item $A$ has Lyapunov stability, i.e., for every $\epsilon > 0$  there exists a $\delta(\epsilon) > 0$ so that $d(u,A) < \delta$ implies $d(T^nu,A) < \epsilon$ for all $n\ge 0$.  
-\end{enumerate}
-\end{Definition}
-
-Indeed in our previous examples, it can be verified that the system  $Tu=u/2$ defined on  $\mathbb{R}$ and $Tu=u^2$ defined on $[0,1]$  the singleton set $\{0\}$ is an attractor.
-For the system,  $Tu=1-|2u-1|$ on $[0,1]$, one may easily verify that the only attractor is the entire space $[0,1]$.
-This is since between any two points $u< v$ in $[0,1]$, and for any $a,b$ so that $u\le a < b \le v$, we can find an $n$ so that $T^n(a,b)=[0,1]$ (as would be explained later in this chapter).
-
-% \ednote{For let $B$ be an attractor of U.
-From above, we know that B is itself invariant, is asymptotically stable and has Lyapunov stability.
-Now suppose there is some attractor $C\subsetneq{U}=B$, so there is a $b\in{U}$ which is not in $C$.
-........
-M: Have changed the example}
-
-A dynamical system can have several attractors, and may also be contained in another attractor.
-For the example, $Tu =u^2$ on $[0,1]$, both $\{0\}$ and $[0,1]$ are attractors.
-It is a known result that for dynamical systems on a compact space there always exists an attractor.
-
-The dynamics restricted to an invariant set can be complicated.
-For instance an invariant set could be just a single point or it could have an infinite set.
-If the invariant set is infinite then there is a possibility of complicated dynamics, and a particular well-studied phenomenon of such complexity gives rise to what is called  chaotic behaviour, and has been studied in detail over the last fifty years.
-
-% If the dynamics on the attractor are somewhat complicated in the sense that the attractor cannot be decomposed further, and if the attractor is infinite then there is a possibility of complicated dynamics, and a particular well-studied phenomenon of such complexity gives rise to what is called a chaotic behaviour.
-
-%If the dynamics on an attractor cannot be decomposed further and the attractor is itself also infinite, then there is a possibility of complicated dynamics, and a particular well-studied phenomenon of such complexity gives rise to what is called chaotic behaviour.
-
-\section{Chaos}
-
-% \ednote{ should one add paragraphs like these?
-M: this is a thesis, and not a paper; the instructions sent to the external examiner regarding the thesis evaluation says that the student suggest strongly that it is good to write}
-In the 1960s, a number of mathematicians and mathematically interested scientists independently discovered chaos in the mathematical sense.
-The meteorologist Edward Lorenz may have been the first to explain this phenomenon in his 1963 paper \cite{lorenz1963deterministic}.
-The notions of invariance, attractivity and chaos may also be described for continuous systems, and Lorenz's system comprised of a system of differential equations.
-The narrative of Lorenz's discovery of chaos, as well as the history of other forerunners in this subject, are fascinating.
-For those who are interested, we highly recommend James Gleick's book Chaos: The Making of a New Science \cite{gleick2008chaos}.
-It clearly illustrates these experiences, explains why chaos was such a startling and crucial mathematical and scientific discovery, and describes the underlying mathematical notions for non-specialists.
-
-There are several intricately different definitions of chaos available in the literature \textbf{(Cite)}, with each of them indicating some aspect of complexity.
-%\ednote{B:  what are we saying here?
-M: Devaney's definition is not the only one, there are about half a dozen definitions as well}
-In practice, it is not possible to verify which definition or notion does the underlying dynamical system satisfies when only data is observed from it.
-We do, however, specifically recall the definition of chaos in the sense of Devaney \cite{devaney2018introduction,de2013elements} so as to understand some nuances behind the complexity.
-Devaney's definition of chaos has three requirements, with the first being the notion  of sensitive dependence on initial conditions.
-
-\begin{Definition}\rm
-  [\bf {Sensitive Dependence on Initial Conditions}]\label{Dfn_SDIC}\rm
-A dynamical system $T: U \to U$ is said to have sensitive dependence on initial conditions (SDIC) if there exists a $\delta > 0$ such that for every $u \in U$ and in every neighborhood of $u \in U$ there exists a $v\in{U}$ and an integer $N:=N{(u,v)}\in\mathbb{Z}$ such that $d(T^Nu,T^Nv)>\delta$.
-\end{Definition}
-
-It is important to acquire a sense of this definition as it can easily be misunderstood.
-It is common misconception to interpret SDIC as two close points ($u$ and $v$) that eventually become separated by a distance $\delta$ under iteration by $T$.
-But this is not true.
-To fully comprehend all the subtleties of the concept, one needs to discuss it more thoroughly by considering each sentence with care and attention, whereafter one may examine how they fit together to convey the concept of sensitive dependence.
-
-To this end, we make 3 remarks:
-\vspace{-5mm}
-\begin{enumerate}
-  \item First, the $\delta>0$ in the definition of SDIC is independent of $u$. 
-  \item Second, in every neighborhood of $u$, we may not necessarily find all points $v$ in the neighborhood distinct from $u$ that would separate from the forward iterates of $u$. 
-  \item  Finally, $N$ depends upon $u$ and $v$ chosen, and their iterates may not separate for ever (i.e., for all $n>N$) and we allow their iterates to get arbitrarily close in the future. 
-\end{enumerate}
-
-To illustrate the concept of SDIC, we present two examples - one from Mathematics and the other from the field of Physics.
-
-\begin{Example}
-  The logistic map(LM), a recursion relation of the form $x_{n+1}=rx_n(1-x_n)$ where $x_n\in[0,1]$ and $r<4$ is of particular interest to us as it is classically used to illustrate the influence of chaos on a system \textbf{(Cite)}.
-Below are plotted the first 75 iterates of the LM for $r=3.7$ and $x_0=0.2$ with an adjustment of 0.001 in $x_0$ and then $r$.
-
-  \begin{figure}[ht]
-    \includegraphics[scale=0.45]{_log_sdic.eps}
-        \centering
-        \label{fig:log_sdic}
-        \captionof{figure}{Logistic Map for $r=3.7$, $x_0=0.2$}
-      \end{figure}
-
-  For an intuitive sense of the fact that the LM exhibits SDIC, we draw two green neighbourhoods of diameters $\delta_1=0.1, \delta_2=0.02$ around the points $u_1=0.7, u_2=0.4$ respectively.
-It is easy to see that every neighbourhood around these two points will contain some $v_1, v_2\in[0,1]$ and natural number $N_1, N_2 \in\mathbb{N}$ respectively s.t.
-the distance $d(T^Nu_1,T^Nv)>\delta_1=0.02$ and $d(T^Nu_2,T^Nv)>\delta_2=0.1$.
-
-    \ednote{B: Is it desirable to have a more rigourous explanation here? }
-
-\end{Example}
-
-\begin{Example}
-  A second example pertains to a simple physical system called the double pendulum - a pendulum with another pendulum attached to its head given an initial position and angular velociy.
-
-  Below are two figures denoting the trajectories of a double pendulum's second head after some elapsed time.
-Depicted are three pendulum heads with equal angular velocity, but differing ever so slightly in initial position .
-As can be seen below, the trajectories diverge very quickly to become completely different.
-  The double pendulum will be discussed in much greater detail in a later chapter (\ref{ch5}), but we mention the example here to provide a numerical example of a practical system exhibiting SDIC.
-
-\begin{figure}[ht]
-  \centering
-\includegraphics[scale=0.8]{_dp_sdic.eps}
-    \captionof{figure}{Three double pendulum heads with initial angle differing by 0,025 radians}\label{fig:dp_sdic}
-  \end{figure}
-
-\end{Example}
-
-The second part in Devaney's definition of chaos concerns topological transitivity.
-
-\begin{Definition}
-  [\bf {Topologically Transitive}]\label{Dfn_TopolTrans}\rm
-	A dynamical system $F: U \to U$ is topologically transitive if for any pair of nonempty open sets $E_1$ and $E_2$ there exists a $n\in\mathbb{N}$ such that $T^n(E_1) \cap E_2 \not= \emptyset$. 
-\end{Definition}
-
-Topologically transitivity implies that the iterates of an open set of initial conditions gets mixed up with other open sets.
-On a compact metric space, it can be shown that topological transitivity also implies the existence of a point whose forward iterates are dense \cite{de2013elements}; or in other words, the orbit going through this point will be dense in the compact metric space.
-In fact, it is topologically more likely that the choice of an arbitrary point will be one  whose iterates are almost dense.
-
-\begin{Definition}
-  [\bf {Meager Set}]\label{Dfn_Meager Set}\rm
-A subset of a topological space $U$ is said to be a meager set if it can be written as a countable union of sets of with empty interior. The complement of the a meager set is set to be a residual set.
-\end{Definition}
-
-To be more precise (see \cite{de2013elements}), for a discrete-time dynamical system on a compact space, the set of points with dense iterates are residual, and they are typical or likely to be observed in practice.
-In this thesis, we assume that when data is observed from a topologically transitive system, we assume that the data arises from a dense orbit.
-
-We may now define the notion of Chaos as formulated by Devaney\cite{devaney2018introduction}.
-\begin{Definition}
-  [\bf {Devaney's Chaos}]\label{Dfn_ChaosDec}\rm
-	A dynamical system $T: U \to U$ is said to exhibit chaos in the sense of Devaney if it satisfies the three properties:
-	\vspace{-5mm}
-  \begin{enumerate}
-		\item $T$ has SDIC.
-		\item $T$ is topologically transitive.
-		\item The set of periodic points of $T$ are dense in $U$. 
-	\end{enumerate}
-\end{Definition}
-
-\begin{Example} 
-  The standard tent map $Tu=1-|2u-1|$ defined on $[0,1]$ is a well-known example of a dynamical system satisfying these three properties.
-
-  We now reason as to why this is true.
-The graph of the map $T$ is piecewise linear with two straight lines, one connecting the points $(0,0)$ and $(\frac{1}{2},1)$ and the other connecting $(\frac{1}{2},1)$ with $(1,0)$.
-They form a so-called tent with base centered at $u=1/2$.
-The graph of the map $T^2$ comprises two symmetric tents with their base centered at $1/4$ and $3/4$.
-See Figure \ref{fig:T2tentmap}
-
-  \begin{figure}[ht]
-    \includegraphics[scale=0.6]{_tentmap_2.eps}
-        \centering
-        \captionof{figure}{Three double pendulum heads with initial angle differing by 0,025 radians}\label{fig:T2tentmap}
-  \end{figure}
-
-  In general, the graph of  $T^k$ contains $2^{k-1}$ tents.
-  Given any point $u$ and an open neighborhood $W \subset[0,1]$ of $u$, we can find $k\in\mathbb{Z}$ large enough so that $T^k(W) = [0,1]$ because we can accommodate a tent whose base is contained in $W$.
-% \ednote{B: This last phrase doesn't make sense to me.
-M: The number of tents grow with $k$ and hence this is obvious; I explained in one of our meetings.
-Does it make sense?}
-  This implies the existence of two points $w_1$ and $w_2$ in $W$ so that $d(T^kw_1,T^kw_2)= 1$, and by the triangle inequality  at least one of the inequalities $d(T^ku,T^kw_1)> \delta$ or $d(T^ku,T^kw_2)> \delta$ hold when $\delta\in (0,\frac{1}{2})$.
-So $T$ has sensitive dependence on initial conditions.
-
-  Next, let $W_1$ and $W_2$ be two nonempty open sets.
-Given any open set $W_1$, we can find a $k\in\mathbb{N}$ large enough so that $T^k(W_1) = [0,1]$ because we can accommodate a tent with base contained in $W$.
-So $T^k(W_1) \cap W_2 \not=\emptyset$, and thus $T$ has topological transitivity.
-
-  Finally, for every open interval $(a,b)$, the graph of $T^k$ intersects the graph of the identity map on $[0,1]$.
-This follows from the fact established above that there exists a $k\in\mathbb{N}$ such that $T^k(a,b) = [0,1]$.If the intersection point has coordinates of the form $(p,p)$, the point $p$ is then a fixed point of $T^k$ and therefore also a periodic point of $T$.
-Since we have found a periodic point in the interval $W$,  the set of periodic points are dense in $T$.
-  This follows because any fixed point of $T^k$ is also a periodic point, and we have found this in an arbitrary interval.% \ednote{Expand further on reason for density.
-M: fixed point of $T^k$ is a periodic point, and we have found in an arbitrary interval.}
-
-\end{Example}
-
-\section{Conjugacy}
-
-We now turn our attention to the subject of conjugacy that describes when two dynamical systems are equivalent dynamically.
-
-To show topological similarity or sameness between two metric or topological spaces, one needs to establish a homeomorphism between the two spaces. 
-However, in the study of dynamical systems defined on two spaces, establishing a homeomorphism does not indicate that the systems are dynamically related in any way.
-For instance, the maps $Tu=u^2$ and $Tu=1-|2u-1|$ defined on $[0,1]$ have totally different behaviour.
-To find systems that are dynamically similar, one rather needs to establish a kind of dynamical equivalence which we illustrate in a simple commutativity diagram below.
-
-% Giving errors
-\begin{equation}  \label{eqn_conjugacy}
-%    \[ 
-    \psset{arrows=->, arrowinset=0.25, linewidth=0.6pt, nodesep=3pt, labelsep=2pt, rowsep=0.7cm, colsep = 1.1cm, shortput =tablr}
- \everypsbox{\scriptstyle}
- \begin{psmatrix}
-U & U\\%
-V & V.
- %%%
-%  \ncline{1,1}{1,2}^{T} 
-%  \ncline{1,1}{2,1} <{\phi}
-%   \ncline{2,1}{2,2}^{S}
-%  \ncline{1,2}{2,2} > {\phi}
- \end{psmatrix}
-% \]
-\end{equation}
-
-%Composite functions
-To understand the diagram we note that if we travel ``right and then down," the diagram instructs us to use $T$ (top arrow right) first, followed by $\phi$.
-(right arrow downwards).
-Consequently the pathway amounts to finding $\phi(T(u))$.
-If we proceed "down and then right", the diagram instructs us to apply $\phi$ (left arrow downwards) first, and then apply $S$ (bottom arrow right).
-This then amounts to finding  $S(\phi(u))$.
-When the relationships in this diagram hold, we say $\phi(T(u))= S(\phi(u))$, and we formally denote it as $\phi \circ T=S\circ \phi$.
-
-\begin{Definition}\rm  
-[\bf {Conjugacy \& Semi-Conjugacy}]\label{Dfn_Conjugate}\rm
-   Consider the dynamical systems $T:U\to{U}$, $S:V\to{V}$ and suppose the relationship $\phi \circ T=S\circ \phi$ holds.
-If $\phi:U\to{V}$ is a homeomorphism, then $S$ is said to be conjugate to $T$ and $\phi$ is called a conjugacy.
-If we relax the criterion on $\phi$ and merely require $\phi$ to be continuous and surjective where $\phi:U\to{V}$, then $\phi$ is a semi-conjugacy between $T$ and $S$ where T has domain $U$ and S has domain $V$; S is said to be semi-conjugate to T.
-\end{Definition}
-
-\ednote{M: Saying $\phi$ is then a semi-conjugacy between $T$ and $S$ could be ambiguous as it does not specify which system is on the top in the commutativity.
-So please edit accordingly.
-B: Does defining explicitly the domain and codomain of $\phi$ make it clear enough?
-}
-When $S$ is conjugate to $T$, the dynamics of the two systems are in some way ''dynamically equivalent''.
-Specifically, they are in one-to-one correspondence with one another.
-However when $S$ is semi-conjugate to $T$ with $\phi:U\to{V}$ a many-to-one mapping, the dynamics on $V$ provide merely a coarse-grained description of the dynamics on $U$ \cite{de2013elements}.
-When $S$ is semi-conjugate to $T$, it is also common to call $S$ a \emph{factor of $T$}, or conversely that $T$ is an \emph{extension of $S$}.
-In essence, an extension (e.g., \cite{de2013elements}) is a larger system capturing all of the important dynamics of its factor.
-
-It is a very hard or nearly an impossible task to establish the existence of a conjugacy or a semi-conjugacy $\phi$ between two systems \cite{devaney2018introduction}.
-However, one can verify that a function $\phi$ satisfies the commutativity diagram \ref{eqn_conjugacy}.
-
-\begin{Example}\rm
-  For example $\phi(x)=\sin\big(\frac{\pi}{2}x\big)^2$ is a conjugacy between two systems $Tu=1-|2u-1|$ and $Sv=4v(1-v)$  defined on $U=[0,1]$.
-
-  Indeed, $\phi:[0,1]\to[0,1]$ is a homeomorphism and by employing simple trigonometric identities, it is easily proved that $\phi$ satisfies~\ref{eqn_conjugacy}
-  \[
-    \begin{aligned}[t]
-      (\phi\circ{T})u 
-                &=\sin^{2}\Big(\frac{\pi}{2} - \frac{\pi}{2}|2u-1|\Big) = \sin^{2}(\pi{u})=\Big(2\sin(\frac{\pi}{2})\cos(\frac{\pi}{2})\Big)^{2}=4\sin^{2}(\frac{\pi}{2}u)\Big(1-\sin^{2}\frac{\pi}{2}u\Big) \\
-                                    &=({S}\circ\phi)u.
-    \end{aligned}
-\]
-And it follows that $\phi$ is a conjugacy between the two systems.
-\end{Example}
-
-In establishing that two systems are conjugate (semi-conjugate) to one another, we have also shown that one may choose to work with one system as opposed to the second and still be guaranteed to obtain information on the latter.
-This is especially useful while forecasting the future evolution of dynamical systems.
-
-\chapter{The Learning Problem}\label{ch3}
-
-In this chapter we acquaint ourselves with the question of forecasting dynamical systems with unknown underlying structures, state and discuss the shortcomings of the Takens delay embedding theorem and discuss various issues faced while forecasting.
-
-Consider a relatively simple learning problem: 
-Given the sequence $(u_0, u_1, \ldots, u_m)$ for $m\in\mathbb{Z}$, a finite segment of an orbit of the map $T$ where $T$ is defined by the update equation $u_{n+1} = Tu_n$, forecast the values $u_{m+1}, u_{m+1}$ where the map $T$ is unknown, given that $u_m$.
-
-A practical example of this would be the subsequent scenario: given the time-sequential coordinates of an object moving in space, predict the future positions of that object.
-However we are very rarely, if at all, presented with a problem where the entire state information is available to us in the form $u_n$ at some time-step $n\in\mathbb{N}$.
-Consider then a more involved learning problem:
-
-Suppose we only have the observations $\theta(u_0), \theta(u_1), \ldots, \theta(u_m)$ of the true system states $u$ in an unknown dynamical system $(U,T)$ and we wish to predict the values $\theta(u_{m+1}), \theta(u_{m+2})$ and so forth.
-
-First we establish the method of Takens delay embedding. 
-In this method, one considers a discrete-time dynamical system defined on a 'nice' space (a smooth manifold)\ednote{Should this be further explained?} that can be obtained as time-$K$ map of a continuous time-dynamical system, concepts which will be formalised below.
-We make observations from such a system, i.e., we observe the evolution of $\theta(w_0)$ (where $w_0$ represents the initial state) i.e., we observe  a finite set of values of $u_n :=  \theta(w_n) = \theta(Tw_{n-1})$.
-The sequence $\{u_n\}$ represents a scalar time-series and intuitively $\theta$ may be thought to represent a probe inserted into a larger system which is itself only measuring/extracting a small part of the greater system state at time $t=n$.
-Consider for example a thermometer erected to measure the ambient temperature in a local village.
-This measurement function, the thermometer, is capturing only a single aspect, the temperature, of a much grander dynamical system entailing the present weather of the surrounding area.
-Even more than that though- it is measuring a miniscule part of the global weather system.
-
-%Given the sequence $\{u_1, u_2, \ldots, u_n}$, we define the variable \newline $y_n:=[\theta_n, \theta_{n-\tau}, \ldots, \theta_{n-(2d-2)\tau}, \theta_{n-(2d-1)\tau}]^{T}\in\mathbb{R}_{d}\times{1}$ where $\tau$ represents the lag and $d$ the dimension of the attractor.
-(We shall return to these terms shortly and their discussion may be put on hold for the time being).
-
-However, the true state $u$ of a system is seldom, if ever, fully known.
-In almost all cases we can at most insert probes into a system to obtain partial information by means of the measurements taken.
-Moreover, the process of taking a measurement itself introduces 2 additional aspects which complicate the problem: \ednote{M: Maybe replace difficulty by saying that we need to consider two aspects.
-B: Perhaps this edit makes more sense?}
-\vspace{-8mm}
-\begin{enumerate}[noitemsep, label=\roman*.]
-  \item A series of measurements over a specific time-interval is inherently a discretisation procedure of the underlying continuous-time dynamical system.
-(Hence why we restricted our attention to discrete-time systems in the preceding section)
-  \item A probe will never be fully accurate, and so the act of measurement introduces a certain measure of numerical noise/inaccuracy.
-\end{enumerate}
-
-From here we construct a multidimensional observable using the method of stacking previous observations, i.e., we create delay-coordinate map defined by
-$\Phi_{k,\theta}(w) := (\theta(T^{-k}w)\ldots,\theta(T^{-1}w),\theta(w))$.  
-The concept of an observable is understood in the sense of an observable as originally introduced by \cite{takens1981detecting, genericObservableAeyels} to refer to a probe or measurement function inserted into the system.
-The essence of Takens' theorem relates to the fact that when $k$ is sufficiently large, we can define a dynamical system on the space $\mathbb{R}^{k+1}$ whose states are $\Phi_{k,\theta}(w)$, $\Phi_{k,\theta}(Tw)$, $\Phi_{k,\theta}(T^2w)$, etc.
-and this dynamical system is topologically conjugate to the unknown underlying system $(W,T)$.
-We recall the Takens delay embedding theorem next.
-
-%More formally we may say that in the process of measuring a system, we obtain a 1-dimensional time-series $\{\theta(u(t))\}$, where $\theta:\mathbb{R}^n\to\mathbb{R}$ is a smooth observation or measurement function.
-This series represents a possibly noise-augmented \ednote{M: Noise-augmented is not the right word} dataset containing information of certain parts of the original system.
-
-%We may then ask ourselves the question: Is it in any way possible to retain information about the system state $x(t)$ in this temporal data-series $\theta(x(t))$?
-The answer is easily yes if T is known.
-However, neither this, nor even the exact function $\theta$ is available.
-
-\section{Takens Embedding Theorem}\label{sect_Takens}
-
-We define first the concepts of homeomorphism and embedding:
-\begin{Definition}\rm
-  [\bf {Homeomorphism}]\label{Dfn_homeo}\rm
-  A homeomorphism is a function $f:Z\rightarrow Y$ between two topological spaces $Z$ and $Y$ that is continuous, bijective and has a continuous inverse. 
-\end{Definition}
-
-\begin{Definition}
-  [\bf {Embedding}]\label{Dfn_embed}\rm
-  Consider a homeomorphism $f:Z\rightarrow Y$ for $Y\subset X$. $Z$ is said to be embedded in $X$ by $f$.
-\end{Definition}
-
-Takens Theorem states a result establishing a relationship between the observed and underlying dynamical systems by showing that the concatenation of a sufficiently large number of previous observations into a vector will, under certain conditions, generate a map between the vectors from the respective systems.
-We formulate the theorem from \cite{takens1981detecting}.
-
-\begin{Theorem} 
-	[\bf Takens Embedding Theorem (adopted from \cite{takens1981detecting}] \label{Thm_Takens}
-         Let $W$ be a compact manifold of dimension $m$, and $d\ge m$ so that $2d$ is an integer. It is a 
-            generic property for the pair $(T, \theta)$,  where $T:W \to W$ is
-            a smooth diffeomorphism, and $\theta:W \to \mathbb{R}$ a smooth function, the map $\Phi_{2d,\theta}:W \to \mathbb{R}^{2d+1}$ defined on $W$ by 
-            $\Phi_{2d,\theta}(w) := (\theta(T^{-2d}w)\ldots,\theta(T^{-1}w),\theta(w))$
-            is a diffeomorphic embedding; by `smooth' we mean at least $C^2$.
-Consequently, there exists a map $F_\theta: \Phi_{2d,\theta}(W) \to \Phi_{2d,\theta}(W)$ defined by $$F_\theta: (\theta(T^{-2d}w),\ldots,\theta(T^{-1}w),\theta(w)) \mapsto
-            (\theta(T^{-2d+1}w),\ldots,\theta(w),\theta(Tw))$$
-           so that $(W,T)$ is topologically conjugate to 
-            $(\Phi_{2d,\theta}(W), F_\theta)$.    
-\end{Theorem}
-
-By generic we mean a residual set on a certain topology on appropriate spaces of functions (that we do not describe here) \textbf{(Cite)}.
-%\ednote{How vague may one be in general by leaving things out?}
-By $C^2$, we make reference to a twice-differentiable function  with a continuous second derivative.
-In our scenario (and this remains important throughout) the input space $U$ is considered the attractor.~\label{attractor_U}
-
-Below we reproduce Figure 1 from \cite{Supp}.
-
-%Some explanation is beneficial.
-Takens establishes a delay-coordinate map $\Ftheta$ defined in \ref{eqn_takens} for $T$ the flow (\ref{defn_flow}) defined by the update equation  $Tu_n=u_{n+1}$ and $theta$ our measurement function.
-When confining a dynamical system to the manifold U (in which the underlying system is contained), Takens showed that if certain smoothness conditions are satisfied on $T$ and $\theta$, then the delay-coordinate map $\Ftheta$ embeds $U$ in the reconstruction space $R^{2d+1}$ \textbf{we haven't used the term reconstruction space yet, so this is falling out of the blue}, for almost every choice of measurement function $\theta$, the observable.
-By almost every choice here, we mean that we exclude some observations which do not give any information.
-To see a more precise discussion, see \textbf{xxx} for we do not concern ourselves with a further elaboration on this report.
-
-%Alternatively we may also make the statement: Takens' Embedding Theorem guarantees that almost all dynamical systems can be reconstructed from just one noiseless observation sequence \textbf{this isn't obvious from the theorem's formulation, is it?} i.e.
-for a great number of possible observation functions $\theta$, $\Ftheta$ preserves the topology of U.
-\textbf{'preserving the topology' is not something we've touched on before.
-Perhaps I should add to previous discussion so logic flows seamlessly here}.
-
-\begin{figure}[ht]
-  \includegraphics[scale=0.25]{_takensmap.eps}
-  \centering
-  \captionof{figure}{Schematic of Takens Embedding Theorem's usage and consequence of how iterates \ldots }
-  \label{fig:takensmap}
-\end{figure}
-\ednote{M: Here, and throughout if you reproduce a figure, it must be stated where it was reproduced from. B: Is this satisfactory? Also, should $\tilde{F}$ be left out for now?}
-\ednote{B: This is only a screenshot used as a temporary placeholder. If sufficient, I shall convert it to .eps. If not, may I perhaps have the original picture?}
-
-If we can learn the map $F_\theta$, from sufficient number of data points of $\{\phi_{2d,\theta}(w_n)\}$, then we also know how to forecast how $\theta(w_n)$.
-
-%Once again we pause and consider the graph to consolidate our understanding of the conjugacy (\textbf{are we being too simplistic?}).
-If we are in state $u$ and then evolve towards state $Tu$ before embedding into $\mathbb{R}^{2m-1}$, then this is equivalent to first embedding into $\mathbb{R}^{2m-1}$ and then evolving through the map $\Ftheta$, i.e.
-$\Phi_{2d,\theta}\circ{F}_{\theta} = T\circ\Phi_{2d,\theta}$.
-The map $F_theta$ is a homeomorphism as well.
-
-%Thus information about U can be retained in the time series' (or observation's) output.
-By preserving the topology on the manifold $U$ in the reconstruction space $X$ (\textbf{not discussed before}), we are also guaranteed the preservation of topological invariants of the manifold, of which dimensionality is one such invariant.
-(We note here that dimensionality will again be considered later in this project).
-
-\section{Practical Issues of Takens theorem}
-
-It is a fact that Takens' Embedding Theorem is both a powerful result and that it provides compelling reason to believe that one could conceivably reconstruct accurately a system conjugate to the underlying system.
-Nonetheless, it does present some serious practical limitations:
-\vspace{-5mm}
-\begin{enumerate}
-\item Even supposing that we could indeed find $F_\theta$ , our \emph{approximation} of  $F_\theta$ is a map from a larger set $\mathbb{R}^{2d+1}$ containing the embedded attractor.
-There are, however, no theoretical guarantees that $F_\theta$ will retain  $V=\Phi_{2d,\theta}(W)$  as an attractor although $W$ could itself be an attractor.
-\item Takens theorem is stated only for noiseless observations.
-Due to noise $\epsilon_n$ the delay vector $\Phi_{2d,\theta}(w_n) + \epsilon_n$  may lie outside $V$.Furthermore, due to the chaotic nature of the underlying system (i.e.
-the fact that it has SIDC), the evolution of $\Phi_{2d,\theta}(w_n) + \epsilon_n$ under the map $F_\theta$ could move out of $V$ completely.
-This problem can be overcome  by using a driven dynamical system with some properties, and we discuss this in Chapter\ref{ch3}.
-\end{enumerate}
-
-% Another limitation owes itself to the requirement that T be a surjective map.
-This is not always the case in practical examples and we may consider the case of a damped double pendulum \ednote{M: Please remove the paragraph as well.
-We do not overcome this problem.
-You can add this in the next Chapter where you need T to be surjective or else you could include this at the begining of this chapter where you explain the forecasting problem.}
-
-The Takens embedding theorem does not guarantee global dissipativity.
-\ednote{B:Links with~\ref{subs_LearnGamma} in discussion of dissipativity of $\Gamma$.
-Sir, I'm not sure how to write this here - I think it a beneficial point to make so as to set the stage for the later discussion in ~\ref{subs_LearnGamma}, but I am unsure there as well.}
-
-% \begin{figure}[ht]
-%   \includegraphics[scale=0.3]{_learningfailure.eps}
-%   \centering
-%   \caption{Illustration of a system having failed to learn the map}
-%   \label{fig:supp_learning_failure}
-% \end{figure}
-
-Though not a uniquely related to the Takens embedding theorem, we do opt to make mention of this requirement for our work. 
-The methods described here are applicable for data originating from a surjective map. 
-This follows from the fact that when $T$ maps the space $U$ into a proper subset $B$ of $U$, we me conclude that $T$ is not surjective as there exists some point in $U-A$ which does not have a preimage.
-An example of this would be the case when a system displays energy-loss.
-Consider, for instance, consider the oscillations dying down for a damped pendulum.
-(A more in-depth account will be given in \ref{ch5})
-
-This requirement is not overly restrictive, as a great number of chaotic systems do have a surjective map.
-(Cite) \ednote{I rememeber you saying this in one of our sessions, but don't have a source for it.}
 
@@
 
-\section {Unique Solution Property}
+\section{Unique Solution Property}
 
@@
 The introduction of a new input at time $n$ can be described by the mapping $\sigma_v:   \cev{u}^{n} \mapsto \cev{u}^{n}v$. 
+The right-shift map, $r\overleftarrow{U}\to\overleftarrow{U}$, of an input sequence is defined $r: (\cdots, u_{-2},u_{-1}) \mapsto(\cdots, u_{-3},u_{-2})$
 
@@
 \end{proof}
-To establish whether $\widehat{T}$ and $G_T$ are related, we consider the the function $H_2{\overline{}} := (h(r\overline{u}, h(\overline{u})))$.
-$H_2$ send the entirety of a left-infinite sequence to some element of $X\times{X}$.
-
-% An embedding of the space $\overleftarrow{U}$  in $\overleftarrow{U}$ would allow one to establish topological conjugacy, which in turn provides stronger results than merely obtaining a coarse-grained representation via a causal mapping \ednote{M: I feel this is very confusing to the reader as the co-domain of causal mapping is the left-infinite space and , so please omit if possible.
-You can just say we want to restrict inputs to the left-infinite orbits of a dynamical system and then explore conjugacy instead of a mere semi-conjugacy}.
-
-% Before formalising this, we make mention of the concept of the inverse-limit system of a dynamical system $(U,T)$. 
-We would like to restrict inputs to the left-infinite orbits of a dynamical system and accordingly explore the possibility of establishing a conjugacy instead of merely obtaining a semi-conjugacy.
-These spaces are noteworthy as their topology can help one ascertain much about the original system \cite{Manju_IEEE}.
-The topological structure for inverse-limit systems of chaotic dynamical systems is complicated, so we relegate further discussion to \cite{kennedy2008inverse_limit, ingram2011inverse}.
-In broad sweeps, the inverse-limit space of a dynamical system is a self-map on a subset of an infinite-dimensional space where each point in the inverse-limit space corresponds to a backward orbit of the map $T$.
-Following \cite{manjunath2021universal, ingram2011inverse}:
-
-\begin{Definition}
-  [\bf Inverse-Limit Space, Inverse-Limit System]\label{Dfn_InverseL}\rm
- The subset $\widehat{U}_T\subset \overleftarrow{U}$ defined by \[\widehat{U}_T: = \{ (\ldots,u_{-2},u_{-1}): Tu_{i} = u_{i+1}\]is called the inverse-limit space of $(U,T)$.
- The self-map $\widehat{T}$  induced by $T$ on $\widehat{U}_T$  is defined by \[\widehat{T}: (\ldots,u_{-2},u_{-1}) \mapsto  (\ldots,u_{-2},u_{-1},T(u_{-1}))\] and the resulting dynamical system $(\widehat{U_T}, \widehat{T})$ is called the inverse-limit system of $(U,T)$.
-\end{Definition}
-
-\begin{Definition}
-  [\bf Causal Embedding]\label{Dfn_CausalEmbed}\rm
-  A driven system $g$ is said to causally embed the dynamical system $(U,T)$ if 
-  \vspace{-8mm}
-\begin{enumerate}[noitemsep, label=\roman*.]
-  \item The diagram in \ref{Scomm_h} commutes (i.e. $h$ is a universal semi-conjugacy),
-  \item $H_2(\overline{u}):=(h(r\overline{u}, h(\overline{u})))$ embeds the inverse-limit space $(\widehat{U}_T, \hat{T})$ of $(U,T)$ in $X\times{X}$.
-\end{enumerate}
-\end{Definition}
 
 The driven system $g$ can induce an embedding of $\overleftarrow{U}$ in $\overleftarrow{X}$ as follows: 
-If $H:\overleftarrow{U}{\to}{\overleftarrow{X}_U}$ is also injective (in addition to being surjective, and hence bijective), 
-it becomes the embedding of $\overleftarrow{U}$ in $\overleftarrow{X}$ induced by the driven system $g$. %and we refer to $H$ as a \emph{causal embedding} 
+If the causal mapping $H:\overleftarrow{U}{\to}{\overleftarrow{X}_U}$ is injective (in addition to being surjective), it becomes the embedding of $\overleftarrow{U}$ in $\overleftarrow{X}$ induced by the driven system $g$.
+\ednote{Explain why $H_{-1}$ is continuous} and we refer to $H$ as a \emph{causal embedding}
 % \ednote{M : Causal embedding a dynamical system is different from the definition of causal embedding, so please edit/delete this paragraph and the next theorem and also in the remaining part till Section 1.2}.
-
 % \begin{Theorem}
@@
 % \end{proof}
-
 When $g$ has the USP, the diagram below (adopted from~\cite{Manju_Nonlinearity}) illustrates the operation of the mappings $h$ and $H$.
@@
 \begin{figure}[ht]
-  \includegraphics[scale=0.3]{_actionofh_H.png}
+  \includegraphics[scale=0.3]{_actionofh_H.eps}
   \centering
-  \captionof{figure}{Progressive actions of $h$ and $H$ to illustrate the causal mapping. }
+  \captionof{figure}{Ladder-like behaviour of $h$ and $H$ to illustrate the causal mapping. }
   \label{fig:actionh_H}
@@
 
-The next result is of relative import but to fully comprehend its influence, some discussion still remains. We state it in part here.
-% Before stating it, we define a concept which will directly be shown to be equivalent to the USP.
-
-\begin{Theorem}
-  The following statements are equivalent:
-  \vspace{-8mm}
-  \begin{enumerate}[noitemsep, label=\roman*.]
-    \item $g$ has the USP. 
-    \item $g$ is a topological contraction.
-  \end{enumerate}
-\end{Theorem}
-% \begin{proof}
-%   (postponed) 
-% \end{proof}
+We wish to explore the possibility of establishing a conjugacy rather than a mere semi-conjugacy and therefore are interested in the restriction of inputs to the left-infinite orbits of a dynamical system.
+The left-infinite orbit spaces are significant because a great amount of information about the original system can be gleaned from their topological structure~\cite{Manju_IEEE}. 
+The formal topological structure for the so-called inverse-limit space of chaotic dynamical systems is complicated, and further discussion is relegated to~\cite{kennedy2008inverse_limit, ingram2011inverse}.
+Broadly, we may conceive of the inverse-limit space of a dynamical system as a self-map on a subset of an infinite-dimensional space where each point in the inverse-limit space corresponds to a backward orbit of the map $T$.
+Following~\cite{manjunath2021universal, ingram2011inverse} we provide a definition:
+
+\begin{Definition}
+  [\bf Inverse-Limit Space, Inverse-Limit System]\label{Dfn_InverseL}\rm
+ The subset $\widehat{U}_T\subset \overleftarrow{U}$ defined by \[\widehat{U}_T: = \{ (\ldots,u_{-2},u_{-1}): Tu_{i} = u_{i+1}\]is called the inverse-limit space of $(U,T)$.
+ The self-map $\widehat{T}$  induced by $T$ on $\widehat{U}_T$  is defined by \[\widehat{T}: (\ldots,u_{-2},u_{-1}) \mapsto  (\ldots,u_{-2},u_{-1},T(u_{-1}))\] and the resulting dynamical system $(\widehat{U_T}, \widehat{T})$ is called the inverse-limit system of $(U,T)$.
+\end{Definition}
+
+The inverse-limit space is well-defined since T : U →U is surjective by assumption.
 
@@
 When a causal embedding $H$ exists for the driven system $g$, one can map an arbitrary input ${u}$ onto the solution space $X$ without additional distortion or information-loss \textbf{(Cite.)}.
-When an embedding is established, the question of possible additional complexity in the solution is removed by guaranteeing that, since the systems are conjugate (semi-conjugate, \textbf{(refer)}), $g$ does not add any (some) complexity to the system.
+Established such an embedding eliminates the question of possible additional complexity in the solution by guaranteeing that, since the systems are conjugate (semi-conjugate, \textbf{(refer)}), $g$ does not add any (some) complexity to the system.
 
@@
 
-We're now getting quite close to where we want to be and our results carry more and more weight. Recall that $g$ is only being given inputs from the orbits of $T$.
-
-\begin{Definition}
-  [\bf Inverse-Limit Space]\label{Dfn_InvLimSpace}\rm
-  The inverse-limit space $\widehat{U}_T$, a subspace of $\overleftarrow{U}$, is defined by $$\widehat{U}_T:=\{(\ldots, u_{-2}, u_{-1}):T(u_n)=u_{n+1}\}$$ 
-\end{Definition}
-
-The inverse-limit space is well-defined since $T:U\to{U}$ is surjective by assumption.
+% Recall that $g$ is only being given inputs from the orbits of $T$.
+
+It is of interest to ascertain whether $\widehat{T}$ and $G_T$ are related\ednote{B: Should I expand more here?}, and we consequently introduce  the function $H_2{\overline{}} := (h(r\overline{u}, h(\overline{u})))$ mapping the entirety of a left-infinite sequence to some element in $X\times{X}$.
+Indeed, it may be shown that $g$ having SI-invertibility and the USP immediately guarantees the existence of at least a semi-conjugacy between the systems $(Y_T, G_T)$ and $(\widehat{U}_T, \widehat{T})$.
+This is formalised in Causal Embedding Theorem, which we state next:
+
+\begin{Theorem}
+  [\bf Causal Embedding Theorem (adopted from \cite{Supp})]
+  % {\bf (Causal Embedding Theorem.)}
+ \label{Thm_CET}
+	Let $g$ be a driven system with SI-invertibility and the USP.
+Let $h$ denote the universal semi-conjugacy and $H_2(\overleftarrow{u}) := (h(r\overleftarrow{u}),h(\overleftarrow{u}))$, where $r$ is the right-shift map.
+ Let $(\widehat{U}_T, \widehat{T})$  be the inverse-limit system of a dynamical system $(U,T)$. 
+ Then the restriction of $H_2$ to $\widehat{U}_T$ is a topological semi-conjugacy between the inverse-limit system $(\widehat{U}_T, \widehat{T})$ 
+and the induced dynamical system  $(Y_T,G_T)$, i.e., the following diagram commutes
+\begin{equation} \label{comm_H_CET}
+\psset{arrows=->, arrowinset=0.25, linewidth=0.6pt, nodesep=3pt, labelsep=2pt, rowsep=0.7cm, colsep = 1.1cm, shortput =tablr}
+ \everypsbox{\scriptstyle}
+ \begin{psmatrix}
+ \widehat{U}_T & \widehat{U}_T\\%
+ Y_T &  Y_T.
+ %%%
+%  \ncline{1,1}{1,2}^{\widehat{T}} \ncline{1,1}{2,1} <{H_2}
+%  \ncline{1,2}{2,2} > {H_2}
+%  \ncline{2,1}{2,2}^{G_T}
+ \end{psmatrix}
+ \end{equation}
+or in other words, $(Y_T, G_T)$ is a factor of  $(\widehat{U}_T, \widehat{T})$. Further, if $T:U \to U$ is a homeomorphism, 
+then $H_2$ embeds $\widehat{U}_T$ in $X_U \times X_U$, and hence $(Y_T, G_T)$ is conjugate to $(\widehat{U}_T, \widehat{T})$.
+\end{Theorem}
+\begin{proof}
+  May be found as the proof of~\cite[Th.4]{Supp}
+\end{proof}
+
+We are drawing nearer and nearer to our principal target and our results carry more and more weight. The next result is remarkable.
 
@@
 
-Note that $H_2$ maps an entire left-infinite solution sequence from $\Psi$ to an element in $X\times{X}$.
-We now have the following (compare with \ref{SCausal_H} above):
-
-% \ednote{Add Causal Embedding Theorem}
-% \textbf{NB NB NB - Causal Embedding Theorem! - Th.4}
+Recall that $H_2$ maps an entire left-infinite solution sequence from $\Psi$ to an element in $X\times{X}$.
+We now have the following (which may be compared with \ref{SCausal_H} above):
 
@@
 
-\section{A discussion of $G_T$ }
-
-In the above sections, we established the map $G_T$ describing the SDD of a driven system.
-By establishing the existence of this map, we’ve essentially embedded the attractor $U$ (recall from \ref{attractor_U}) into the higher dimensional space $X\times{X}$.
-In layman’s terms, this ensures that there is more “dimensional room” for the underlying system’s underlying dynamics to “move”.
-Since the dynamics aren’t as “squashed”, we might therefore hope that the dynamics of $G_T$ are in some sense simpler than that of $T$.
-(Taken note of the fact that $G_T$ is a homeomorphism even when $T$ is just continuous)
-
-In \cite{manjunath2021universal} it is illustrated in an empirical fashion in that the map $G_T$ describes dynamics which are less functionally complex than that of $T$ or of the map $\Phi_{2d,\theta}$.
-This is done by implementing a Recurrent Neural Network (RNN), which is discussed in~\ref{ch5}.
-\ednote{(B: Should I add in that we'll be using Pearson coefficient?)}
-
-We opt to learn $G_T$ in an indirect manner by defining a new map $\Gamma:(x_{n-1},x_n)\mapsto{u_n}$. It will follow immediately from $G_T$’s existence that $\Gamma$ also exists.
-
-The reasons for taking this roundabout approach remain to be discussed in \ref{subs_LearnGamma}, but a pat answer may immediately be given: When $\Gamma$ has been learnt, the system can be driven autonomously and then $G_T$ is known anyway.
-We formalise this with a theorem.
-
-\begin{Theorem}
-  When $x_n$ and $x_{n-1}$ are successive points on a solution obtained for input-orbit of $T$ $\{u_n\}$, then the map $\Gamma: X\times{X}\to{U}$ defined by $(x_{n-1},x_n)\mapsto{u_n}$ exists whenever $G_T$ exists
-\end{Theorem}
-\begin{proof}
-  See \cite[Th. 3c]{manjunath2021universal}.
-\end{proof}
-
-Projection mappings $\pi_i$ are defined in the traditional meaning where a $k$-dimensional vector is projected to it's $i^{th}$ component such that $$\pi_i:(a_1,a_2, \ldotp, a_{k-1}, a_k)\to{a_i}$$
-
-% \begin{Definition}
-%   [\bf Projection Mapping]\label{Dfn_ProjMap}\rm
-%  The projection mapping $\pi_i$ is defined as follows: $$\pi_1:(a_1,a_2)\to{a_1}$$ and $$\pi_2:(a_1,a_2)\to{a_2}$$ 
-% \end{Definition}
-
-The graph in equation \ref{fig:inverse_limsystem} is then extended as below:
-
-% \begin{figure}[ht]
-%   \includegraphics[scale=0.3]{_gammatemp.eps}
-%   \centering
-%   \label{fig:temp_Gamma}
-% \end{figure}
-
-\begin{equation}  \label{Scomm_imp}
-  %    \[ 
-      \psset{arrows=->, arrowinset=0.25, linewidth=0.6pt, nodesep=3pt, labelsep=2pt, rowsep=1.3cm, colsep = 1.1cm, shortput =tablr}
-   \everypsbox{\scriptstyle}
-   \begin{psmatrix}
-   \widehat{U}_T & \widehat{U}_T\\%
-   Y_T & Y_T\\
-   & \textcolor{red}{U \times X}
-   %%%
-  %  \ncline{1,1}{1,2}^{\widehat{T}} \ncline{1,1}{2,1} <{H_2}
-  %  \ncline{1,2}{2,2} > {H_2}
-  %  \ncline{2,1}{2,2}^{G_T}
-  %  \psset{linecolor=red}
-  %  \textcolor{red}{\ncline{2,1}{3,2}<{(\Gamma,\pi_2)}}
-  %  \ncline{3,2}{2,2}>{\textcolor{red}{(\pi_2,g)}}.
-   \end{psmatrix}
-  % \]
-  \end{equation}
-
-The problem finally simplifies to the issue of learning the map $\Gamma$ and combining this with the projection mapping $\pi_2$ and the function $g$, which will be known.
-A final set of equations is obtained - equations that have been entirely constructed from data.
-\begin{eqnarray}\label{eqns_from_data}
-	u_{k+1} &=& \pi_1 \circ (\Gamma, \pi_2) \circ (\pi_2,g) (u_k,x_k) \label{Seqn_u}\\
-	x_{k+1} &=& \pi_2 \circ (\Gamma, \pi_2) \circ (\pi_2,g) (u_k,x_k). \label{Seqn_x}
-\end{eqnarray}
-
-\section{Advantages of learning $\Gamma$} \label{subs_LearnGamma}
-
-One may immediately ask why we opt to take such a roundabout route; why not just learn the map $G_T$ from the get-go?
-On the surface, this seems to be an arbitrary decision path with no real reasoning, so we take a pause again and discuss the motivation for learning Gamma.
-There are a number of distinct advantages.
-
-In the first place, Learning $\Gamma$ saves computational resources.
-This is due to the fact that the input $u_n$ lies in a lower-dimensional subspace of the high-dimensional space $X\times{X}$ and we can learn $\Gamma$ on a small subset of the space.
-In practice, if the input is of a  lower dimension, one may easily embed it into the space $X\times{X}$ by padding the vector $u_n$ with zeroes.
-
-Secondly, the function $\Gamma$ is known to be stable.
-Learning Gamma makes use of equations \ref{eqns_from_data} which in turn employs $g$, a function possessing the USP, and offers distinct advantages with regards to stability in the presence of perturbation.
-It is desirable that inputs that differ only slightly would have only ‘slight’ effects on the output of the system.
-Moreover, one would hope to prevent numerical errors originating from input and measurement noise (see~\cite[Th.
-5]{manjunath2021universal}).
-As the data fed into the system is a function of both the input given and the parameters of the system itself (i.e.
-the value a, $\alpha$ in $\tanh$), we investigate both input-related and parameter-related stability.
-Both these notions are defined by way of the continuity of an encoding map of an input and design parameter respectively in \cite{manjunath2020stability}.\
-
-Informally, we may consider input-related stability as concerned with the question of whether or not small variations in input would result in small responses.   
-If two inputs $u$ and $u_{noisy}$ are close in the product topology, then their tails could differ greatly due to the metric's insensitivity to the differences in sequence's tails.
- However if the system has the USP then the function $h$ is continuous and so $h(u), h(u_{noisy})$ remain near – maintaining a measure of robustness to input and measurement noise.\cite{manjunath2021universal}.
-In a similar fashion, parameter-related stability is obtained thanks to a result \cite[Lemma 3.2]{manjunath2020stability} relating this form of stability with the ESP (which is here equivalent to the USP).
-
-% Secondly,  due to $g$ possessing the USP, both input- and parameter-related stability~\cite{manjunath2020stability} is obtained, which in turn prevents numerical errors that caused by input and measurement noise (see~\cite[Th.
-5]{manjunath2021universal}).
-
-\begin{figure}[ht]
-  \includegraphics[scale=0.5]{_AutovsDriven.eps}
-  \centering
-  \label{fig:learningFailure}
-  \captionof{figure}{Illustrating empirically the importance of global dissipativity.
-Plotted below the three principal components of randomly initiated trajectories of a learnt version of $G_T$ versus the trajectory $(x_{n-1}, x_n)$(grey) of driven stated obtained from the Thomas Attractor's data (see below).
-}
-\end{figure}
-
-\ednote{B: Would you please help me with the writing here? Otherwise I shall leave this out.}
-\emph{Thirdly, $\Gamma$ is globally dissipative whereas $G_T$ is not guaranteed to be~\cite{manjunath2021universal}). 
-Global dissipativity prevents large numerical errors due to input data.If a system is not globally dissipative, neglibly small errors result in major errors which thus induce predictions to fail utterly.
-If one learns $G_T$ directly, we are in danger of replicating the problems arising for Takens.
-}
-
-% The $\omega$-limit set $\omega(u;T)$ of a point $u$ is defined to be the collection of limit points of the sequence $\{x,Tu,T^2u,\ldots\}$.
-This set is nonempty, and $\omega(u;T)$ is invariant.
-\textbf{(Change)}
-% \begin{Definition}
-%   [\bf Global Dissipativity]\label{Dfn_GlobalDiss}\rm
-% 	We say a dynamical system $(Z,f)$ is globally dissipative if there exists a nonempty proper closed subset $B$ of $Z$  so that for all $x\in Z\setminus B$, (i).
-$\omega(x;f) \subset B$  and (ii).
-$B$ is positively invariant that is $f(B) \subset B$.
-We call any such closed subset $B$ a trapping set of $f$.
-% \end{Definition}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\chapter{Implementation} \label{ch5}
-
-In this section we turn to the implementation of the function $G_T$ by discussing a method to learn the function $\Gamma$.
-We then indicate where this implementation differs from previous work done by (\cite{manjunath2021universal}) and empirically show that the SDD of $G_T$ are less functionally complex than $\Phi_{k, \theta}$ or $T$ by considering the Pearson correlation coefficient.
-Finally, we present and discuss some numerical results for a focus system of this project: the double pendulum. We also consider other chaotic attractors.
-
-\section{Implementing $G_T$}
-The map $G_T$ is implemented by applying the functions $\Gamma$, $g$ and the projection function $\pi_2$ in the relevant order as described by the equations~\ref{eqns_from_data}.
-To reach this point, one must first learn $\Gamma$.
-After first normalising the data to ensure it lies within the interval $|x| < 1$ and has zero mean value, we continue by driving the system with external input through the function $g$.
-
-From \eqref{eqn_driving} we know that $ g(u,x)=(1-a)x+a{\cdot}\tanh(Au+\alpha{a}Bx)$.
-Recall that $g$ is SI-invertible.
-Adopting the terminology of \cite{manjunath2021universal}, $A$ is called the “input matrix” and $B$ the “reservoir matrix”.
-
-We make use of a Recurrent Neural Network (RNN) in our implementation.
-RNN's are widely used in the literature for discrete-time processing and are considered to be discrete-time version of continuous dynamical systems modelling a network of neurons.
-(\textbf{Citation.})
-A complete discussion of RNN's is outside the scope of this thesis, so we present only the details necessary to the current discussion.
-For information, (\textbf{Cite, Cite, Cite}) may be consulted.
-We implement our RNN in the \emph{Python} programming language and making use of the \emph{Scipy} library, alongside the \emph{Keras} package in the \emph{Tensorflow} library.
-
-If the RNN we will be using in the next step has $N$ neurons, then $A$ and $B$ are each $N\times{N}$ matrices.
-Inputs $u_n$ of dimension $K < N$ are padded by zeroes to obtain an input of dimension $N$.
-As we do not need the entire left-infinite history of input, an arbitrary initialisation of the system with initial values followed by the driving of the system with the function $g$ means that the generated values will approach the actual solution’s elements in a uniform manner thanks to the UAP.
-\textbf{(Citation)}
-
-Thirdly we prepare the network.
-Utilising a network $N$ neurons (the same value $N$ as in the dimensions of the matrices $A,B$).
-Training is done using the Adam Optimiser where the MSE loss function is optimised.
-The system is trained several times(in our case 4) with incrementally smaller learning rates on each run (0.001 and thereafter divided by 10 at each run).
-Additional parameters such as the number of network-layers, neurons per layer and training-length are presented case-by-case below.
-
-We then reach the prediction step. Making use of $g$'s SI-invertibility and $g$ as defined in \eqref{eqn_driving}, we have the following expression for $u_n$ replicated from~\eqref{eqn_SI_RNN}.
-\begin{equation*}
-  u_n := A^{-1}\bigg(\overline{\tanh}^{^{-1}}\frac{1}{a}\Big(x_{n+1}-(1-a)x_n\Big) \bigg) - \alpha B x_n
-\end{equation*}
-
-The two model equations for learning from data (\eqref{eqns_from_data}) allows us to determine what the state $x_{n+1}$ at time $n+1$ will be. 
-We specifically remind ourselves of \[x_{k+1}=\pi_2 \circ (\Gamma, \pi_2) \circ (\pi_2,g) (u_k,x_k)\]
-
-\section{Changes in this project to previous work of~\cite{manjunath2021universal}}
-
-To understand the first and most drastic\ednote{change word} change, we recall the the conjugacy in~\ref{fig:inverse_limsystem}:
-
-\begin{equation*}
-  %  \[ 
-      \psset{arrows=->, arrowinset=0.25, linewidth=0.6pt, nodesep=3pt, labelsep=2pt, rowsep=0.7cm, colsep = 1.1cm, shortput =tablr}
-      \everypsbox{\scriptstyle}
-      \begin{psmatrix}
-      \widehat{U}_T  && \widehat{U}_T \\
-      Y_T && Y_T \\
-      %%%
-     %  \ncline{1,1}{1,2}^{\widehat{T}} \ncline{1,1}{2,1} <{h}
-     %  \ncline{1,2}{2,2} > {H_2}
-     %  \ncline{2,1}{2,2}^{G_T}
-      \end{psmatrix}
-    %  \]
-  \end{equation*}
-
-  We recall that $G_T$ is topologically conjugate to $\widehat{T}$, a result which we have already made use of.
-
-  The authors of~\cite{manjunath2021universal} did not, however, consider the relationship between the inverse-limit space $\widehat{U}_T$ and the input $u_n=\theta(w_n)$ (where $\theta$ is our observable and $w_n$ the complete input), a concept we now explore.
-  We may construct a delay-coordinate $v_n=[\theta(w_n), \theta(w_{n-1}), \ldots, \theta(w_{n-d})]^{T} = [u_n, u_{n-1}, \ldots, u_{n-d}]^T$
-    to accordingly obtain a new dynamical system $(V,F)$ where $F$ is the map $F:v_n\mapsto{v_{n+1}}$.
-If $d$ is sufficiently large, we may apply Takens Embedding Theorem~\ref{Thm_Takens} to establish the existence of a topological conjugacy between the new system $(V,F)$ and the original system $(U,T)$.
-
-  Using the same notation as before, we may construct a new conjugacy which now relates delay coordinates to one another.
-
-  \begin{equation}\label{eqn_VFdelay}
-    %  \[ 
-        \psset{arrows=->, arrowinset=0.25, linewidth=0.6pt, nodesep=3pt, labelsep=2pt, rowsep=0.7cm, colsep = 1.1cm, shortput =tablr}
-        \everypsbox{\scriptstyle}
-        \begin{psmatrix}
-        \widehat{V}_F  && \widehat{V}_F \\
-        Y_F && Y_F \\
-        %%%
-       %  \ncline{1,1}{1,2}^{\widehat{F}} \ncline{1,1}{2,1} <{h}
-       %  \ncline{1,2}{2,2} > {*}
-       %  \ncline{2,1}{2,2}^{G_F}
-        \end{psmatrix}
-      %  \]
-    \end{equation}
-
-The existence of $\widehat{F}$ is guaranteed, but the map $v_{n}\mapsto{v_{n+1}}$ itself is unknown.
-We may, however, make use of the SI-invertibility of the function $g(u_{n-1}, x_{n-1})=x_n$ to conclude that
-$(x_{n-1}, x_{n})$ determines $u_{n-1}$ uniquely.
-
-Accordingly $(x_{n-2}, x_{n-1}, x_{n})$ determines $u_{n-2}$ and $u_{n-1}$ uniquely and generally $(x_{n-d}, \ldots , x_{n-1}, x_{n})$ determines $u_{n-d}, \ldots u_{n-2}$ and $u_{n-1}$ uniquely.
-
-We therefore need not know what $\widehat{F}$ is, as we can determine $v_{n+1}={[u_n, u_{n-1}, \ldots, u_{n-d}]}^T$ indirectly by determining the value of $(x_{n-d}, \ldots , x_{n-1}, x_{n})$, a much more feasible task.
-
-Accordingly we define the following abbreviations:
-\begin{itemize}
-  \item Delay Coordinates (DC): A vector composed by appending a fixed amount of previous values to the final value.
-  \item Single-Delay Dynamics(SDD): Dynamics generated by 2-element delay-coordinate vectors of the form $(x_{n-1}, x_n)$ where only one previous value is used.
-  \item Multi-Delay Dynamics(MDD): Dynamics generated by arbitrary-length delay-coordinate vectors of the form $(x_{n-d}, \ldots , x_{n-1}, x_{n})$.
-($d$ must be sufficiently large, but is unknown.
-We vary this parameter in experimentation.)
-\end{itemize}
-
-Within this framework, the PCA-implementation of~\cite{manjunath2021universal} is then considered to fall underneath the of auspices of what we term \emph{Approximated Multi-Delay Dynamics(AMDD)}.
-
-We have now seen that one may comfortably switch between the use of SDD or MDD when $g$ has SI-invertibility. /ednote{This sentence needs to be improved.}
-
-Naturally, one would immediately suppose that the MDD approach would increased computation time and resource use as the internal matrices and datasets generated during learning would grow with every increase in the delay-parameter $d$.
-
-In fact, preliminary results indicate the opposite due the observation that the implementation of MDD is inversely proportional to the the dimension of the network (i.e.
-the number of neurons in the RNN) and possibly also the number of epochs/batch size necessary to produce reliable predictions.
-This, in turn, actually speeds up the computation-time and provides results of similar accuracy.
-\ednote{I did not have enough time to explore this properly.
-I can only hazard guesses and promise future research work in this project.
-How should I best convey this?
-}
-
-\textbf{Secondly}
-
-Sparse Matrices:
-\vspace{-8mm}
-\begin{itemize}[noitemsep]
-\item Well-conditioned: condition number close to 1
-\item Scaled
-\item Sparsity/density ratio
-\end{itemize}
-
-% Sparse matrix implementation.
-Slight improvement in some scenarios, but not fantastic.
-\ednote{M: Did it not help you a great deal with the DP forecasting?
-B: Sparse matrices tended to be finicky.
-In some cases they helped whereas in others they actually worsened the solution somewhat.
-There's some tuning to be done in terms of condition number, scaling and density}
-
-\textbf{Third}
-
-Density of matrices A/B(i.e. number of non-zero entries in each matrix) (separate from sparse matrix implementation)
-
-\section{Numerical simulations}
-In this section we showcase numerical results pertaining to 3 attractors that were simulated using the methodology described in the previous chapters. We first consider a simple physical system.
-\subsection{Double Pendulum}
-
-The Double Pendulum consists of two pendulums fastened to one another such that the system moves as a whole.
-The first (or higher pendulum) is attached to a fixed point while the second is attached to the endpoint of the first.
-(See below)
-This system is a classical example of a system exhibiting chaos. \textbf{(Citation.)} Consider again~\ref{fig:dp_sdic}.
-The system has four variables: angle and angular velocity of each pendulum rod. An ideal system is assumed where the rods are massless and no friction in the pivots are considered.
-
-\begin{figure}[ht]
-  \includegraphics[scale=0.35]{_dp_setup.eps}
-  \centering
-  \captionof{figure}{Pictorial representation of the setup of a DP. Image taken from~\cite{DPSetup}}
-  \label{fig:dp_setup}
-\end{figure}
-
-The scalar equations of motion are derived using Newtonian physics and can be found in~\cite{DPFormulas}.
-The second-order system is converted to a first-order ODE and solved using the Runge-Kutta numerical method.
-Noise-free, undamped DP data-sets were generated and utilised in this project due to the requirement that the map $T$ be surjective.
-We discuss the case of a DP exhibiting energy-loss (i.e.
-where the map $T$ is not surjective) below.
-
-\textbf{Data originating from a non-surjective map.}
-
-The importance of the requirement that data originates from a surjective map is observed learning failure of a map describing a damped DP.
-Due to the assumption, the program discounted any possibility of energy-loss and we consider the example below.
-
-\begin{figure}[ht]
-  \includegraphics[scale=0.4]{_dp_dying.eps}
-  \centering
-\caption{The map generating data for a damped pendulum is not surjective.}
-\label{fig:damped_pendulum}
-\end{figure}
-
-Despite feeding data that originated from a DP with dying occilations into the system, a map was instead learnt which continued to oscillate in a chaotic, but conservative manner.
-
-\begin{figure}[ht]
-  \includegraphics[scale=0.4]{_dpfailure_nonsurj.eps}
-  \centering
-\caption{Failure to learn a map which is not surjective.}
-\label{fig:dp_notsurjective}
-\end{figure}
-
-Having spent ample time to discuss the challenges, we now change tone so as to satisfy the optimist.
-We present successful run for DP-data to repdict the future movement of the pendulum.
-
-The system was driven for 15500 steps, of which the first 500 were immediately discarded to simulate the network 'forgetting'.
-The driving function $g$ had input and reservoir matrices were implemented as sparse matrices filled to 20\%.
-Parameter-values of $\alpha=0.99$ and $a=0.5$ were used.
-The remaining 15000 datapoints were used to train a network with 128 neurons. 
-The network was initialised to contain 16 hidden layers each with a layer dimension of 64, activated with the tanh activation function.
-Training was accomplished using 256 epochs with batch sizes 128 and a delay of 1.
-
-% Trainlen:15000; Predlen: 1000; Discard: 500; Dimension:128 neurons; $\alpha=0.99$; $a=0.5$.
-Hidden Layers: 16.
-Layer dimension: 64.
-Epochs:256.
-Batch Size: 128.
-Delay(for MDD): 8.
-Density of Sparse Matrices: 0.2.
-$d=1$
-
-\begin{figure}[ht]
-  \centering
-  \includegraphics[width=0.8\linewidth]{_DP_Fig-2F2023.eps} 
-  \captionof*{figure}{Predicted(red) and actual(blue) trajectories of the $x$- and $y$-coordinates of each of the pendulum heads.} 
-% \end{figure}
-% \begin{figure}[ht]
-%   \centering
-  \includegraphics[width=0.8\linewidth]{_DP_Fig-2B2023.eps}
-  \captionof*{figure}{Densities of predicted and actual values of x-coordinate of first pendulum head. }
-  % \end{figure}
-  % \begin{figure}[ht]
-  %   \centering 
-  \includegraphics[width=0.8\linewidth]{_DP_Fig-2D_5.eps}
-    \captionof*{figure}{Takens Delay Map with delay=5 to illustrate behaviour of predicted and actual values of double pendulum data.}
-    \label{fig:dp_predictions}
-\end{figure}
-
-\subsection{Additional Attractors: Clifford \& Thomas}
-We follow the methodology presented in \cite{manjunath2021universal} but opt to consider attractors different from the Lorenz system and H\'enon map in this project.
-
-\subsubsection{Thomas' Cyclically Symmetric Strange Attractor}
-Thomas' Cyclically Symmetric Strange Attractor, a 3D attractor proposed in 1999 by Ren\'e Thomas in \cite{ThomasAttractor}, is described by a set of three equations which is distinctly reminiscent of the Lorenz system, another deterministic three-equation model exhibiting chaotic behaviour.It has a single parameter $\beta$ and has been shown to transition to chaotic behavious when $\beta<0.208186$~\cite{Thomas_BetaParameter}.
-\begin{eqnarray}\label{eqns_thomas}
-  x_{n+1} = \sin(y_n) - \beta{x_n} \\
-  y_{n+1} = \sin(z_n) - \beta{y_n} \\
-  z_{n+1} = \sin(x_n) - \beta{z_n}
-\end{eqnarray}
-
-The behaviour of the system with a $\beta$-parameter value of 0.1056 was simulated by scaling it to fit inside the interval $[-1,1]$.
-Two cases were considered: noise-free and perturbed.
-Noise was generated from a normal distribution with mean 0 and standard deviation 0.05, which translates approximately to *dB.
- A sequence of 3000 observations in both the clear and noisy data-sets respectively were fed into the system with 300 entries discarded to simulate the network's loss of memory.
-The next 9000 steps were predicted and compared with true data.
-
-\begin{figure}[ht]
-  \centering
-  \minipage{0.5\linewidth}
-    \centering
-    \includegraphics[width=\linewidth]{Thomas_nonoise.eps}
-    \caption*{True and predicted trajectories in the absence of noise.}
-  \endminipage\hfill
-  \minipage{0.5\linewidth}
-    \centering
-    \includegraphics[width=\linewidth]{ThomasNoisy.eps}
-    \caption*{True and predicted trajectories for noise added.}
-  \endminipage
-\end{figure}
-
-The RNN was initialised as follows: the dimension of the network is set equal to 500 neurons and 12 hidden layers of dimension 64 each were initialised.
-Training is realised through 150 epochs of batch sizes of 128 each; the map is learnt via the Adam Optimiser using finer and finer learning rates (0.001 and then divided by 10 at every iteration).
-
-\begin{figure}[ht]
-  \centering
-  \includegraphics[scale=0.35]{Thomas_1.eps}\caption*{Predicted trajectories for the attractor demonstrate empirically the ability to predict the evolution of the trajectory for the next ~100 timesteps}
-  \includegraphics[scale=0.5]{Thomas_3.eps}\caption*{Represented here the learnt (blue) and actual (red) densities of the first coordinate of the dataset.}
-\end{figure}
-
-\subsubsection{Fractal Dream Attractor}
-
-Another 2-equation discrete-time map named the Fractal Dream Attractor, or more commonly known as the Pickover Map (first discovered by Clifford A.
-Pickover and discussed in his fascinating book "Chaos in Wonderland"\cite{PickoverChaos}), was also considered.
-\begin{eqnarray}\label{eqns_clifford}
-  {x_{n+1}=\sin(ay_n) + c\cos(ax_n)} \\
-  {y_{n+1}=\sin(bx_n)+d\cos(by_n)}
-\end{eqnarray}
-
-\begin{figure}[ht]
-  \centering
-  \includegraphics[width=1.0\textwidth,left]{Clifford_1.eps}
-  \caption*{These graphs were constructed by predicting the Clifford system 15000 steps into the future and in so doing illustrating the long-term consistency and accuracy of the learnt system.
-As perceived here, we are able to lock on to the trajectory of the Clifford map almost exactly for the first 25 steps.}
-  \minipage{0.5\textwidth}
-      \centering
-      \includegraphics[width=\linewidth]{Clifford_2.eps}
-  \endminipage\hfill
-  \minipage{0.5\textwidth}
-    \centering
-    \includegraphics[width=\linewidth]{Clifford_3.eps}
-  \endminipage
-  \caption{The bottom row, left, shows a plot of the true attractor(red) and the predicted values(blue).
-Bottom row, right, shows the distribution of the first coordinate of the learnt system (blue) and the actual Clifford system (red).
-}
-  \label{fig:Clifford}
-\end{figure}
-
-\ednote{B: Expand more.}
-
-\section{Generalized Pearson Correlation}\ednote{B: Alter heading.}
-
-\ednote{B: Write Pearson Introduction. Could you give me some help here, Sir?}
-
-\begin{center}
-  \begin{table} 
-      \scalebox{0.75}
-      {\begin{tabular}{|c|c| c c |} 
-          \hline
-          Input & Dimension of $X$ & $u_n$ vs $u_{n+1}$ 
-          & $\begin{bmatrix}
-              x_{n-1}\\
-              x_n
-          \end{bmatrix}$ vs $\begin{bmatrix}
-              x_n\\
-              x_{n+1}
-          \end{bmatrix}$ \\
-          \hline
-
-          % \hline
-          % $(w_n)$ is the Lorenz states, sampled every 0.1 timestamps. & & &\\
-          % \multirow{3}{*}{$u_n=\frac{1}{100}w_n$} & 10 &0.9311  & 0.9731\\
-          %     & 100 &  & 0.9934\\
-          %     & 1000 & & 0.9930\\
-
-          % \multirow{3}{*}{$u_n=\frac{1}{10}(\sin(0.1w_{n,x})+\sin(0.1w_{n,y})+\sin(0.1w_{n,z}))$} 
-          %     & 10 &0.8401  & 0.9392\\
-          %     & 100 & & 0.9616\\
-          %     & 1000 & & 0.9737\\
-          % \hline
-
-          \hline
-          $(w_n)$ comes from the Thomas Attractor (for $\beta=0.1056$): & & & \\
-          {$w_{n+1}= \begin{bmatrix} \sin(w_{n,y}) - \beta{w_{n,x}} \\
-              u_y = \sin(w_{n,z}) - \beta{w_{n,y}} \\
-          u_z = \sin(w_{n,x}) - \beta{w_{n,z}} \end{bmatrix}$} & & & \\
-          \multirow{3}{*}{$u_n = w_n - \overline{w}$}
-              & 10 & 0.99989 & 0.999886 \\
-              & 100 & &  0.999913 \\
-              & 1000 & & 0.999901 \\
-          \hline
-
-          \hline 
-          $(w_n)$ comes from Noise-Added Thomas Attractor (for $\beta=0.1056$): & & & \\
-          {$w_{n+1}= \begin{bmatrix} \sin(w_{n,y}) - \beta{w_{n,x}} \\
-              u_y = \sin(w_{n,z}) - \beta{w_{n,y}} \\
-          u_z = \sin(w_{n,x}) - \beta{w_{n,z}} \end{bmatrix}$} & & & \\
-          \multirow{3}{*}{$u_n = w_n - \overline{w}$}
-              & 10 & 0.9994 & 0.99985 \\
-              & 100 & &  0.99991 \\
-              & 1000 & & 0.999895\\
-          \hline
-
-          \hline
-          $(w_n)$ comes from the Thomas Attractor (for $\beta=0.208186$): & & & \\
-          {$w_{n+1}= \begin{bmatrix} \sin(w_{n,y}) - \beta{w_{n,x}} \\
-              u_y = \sin(w_{n,z}) - \beta{w_{n,y}} \\
-          u_z = \sin(w_{n,x}) - \beta{w_{n,z}} \end{bmatrix}$} & & & \\
-          \multirow{3}{*}{$u_n = w_n - \overline{w}$}
-              & 10 & 0.99982 & 0.99981\\
-              & 100 & &  0.999851 \\
-              & 1000 & & 0.99984 \\
-          \hline
-
-          \hline
-          $(w_n)$ comes from the Clifford Attractor \\ (for $a = -1.7; b = 1.8; c = -1.9; d = -0.4$): & & & \\
-        {$w_{n+1}= \begin{bmatrix} \sin(aw_{n,y}) + c\cos(aw_{n,x}) \\ 
-                                                                    \sin(bw_{n,x})+d\cos(bw_{n,y}) \end{bmatrix}$} & & & \\
-          \multirow{3}{*}{$u_n = w_n-\overline{w}$}
-              & 10 & -0.21713 & 0.78719 \\
-              & 100 & &  0.8334 \\
-              & 1000 & & 0.84854 \\
-          \hline  
-          % \hline 
-          % $(w_n)$ comes from the Logistic map: & & & \\
-          % {$ w_{n+1}  = 4w_n(1-w_n)$} & & & \\
-          % \multirow{3}{*}{$u_n = w_n - 0.5$} 
-          %     & 10 &-0.0314  & 0.5577\\
-          %     & 100 &  & 0.7150\\
-          %     & 1000 & & 0.7826\\
-          % \hline
-          % \hline
-          % $(w_n)$ comes from the Svensson Attractor \\ (for $a = -1.7; b = 1.8; c = -1.9; d = -0.4$): & & & \\
-          % {$w_{n+1}= \begin{bmatrix} d\sin(aw_{n,x}) - \sin(bw_{n,y}) \\ 
-          %                                                             c\cos(aw_{n,x})+\cos(bw_{n,y}) \end{bmatrix}$} & & & \\
-          %   \multirow{3}{*}{$u_n = w_n-\overline{w}$}
-          %       & 10 & -0.2243 & 0.49698 \\
-          %       & 100 & &  0.63310 \\
-          %       & 1000 & & 0.71151 \\
-          % \hline
-          \hline
-          $(w_n)$ comes from the DP \\ (for ): & & & \\
-          \multirow{3}{*}{$u_n = w_n-\overline{w}$}
-                & 10 & 0.99141 & 0.98916\\
-                & 100 & &  0.99238 \\
-                & 1000 & & 0.99278 \\
-          \hline        
-      \end{tabular}} %\ednote{http://paulbourke.net/fractals/peterdejong/}
-
-      \caption{Following \cite{manjunath2021universal}, we determine the multidimensional correlation coefficients $\rho$ to evidence a general reduction in the functional complexity of the map $G_T$ emerging due to a RNN.
-Rows correspond to distinct dynamical systems considered in the experiments
-      The second column corresponds to the amount of artificial neurons (the dimension of $X$) in the implemented RNN.
-      The last two columns comprise numerical estimates of $\rho$ for the relevant vectors. The linear association between $u_n$ vs $T(u_{n})$, as well as the linear relationship $\begin{bmatrix}
-              x_{n-1}\\
-              x_n
-          \end{bmatrix}$ vs $G_T \left( \begin{bmatrix}
-              x_{n-1}\\
-              x_n
-          \end{bmatrix}\right)$ may be contrasted by the reader. 
-          Finally, the RNN's dimension (from \eqref{Seq_RNN}) is provided to exhibiting that an increase in dimension of the driven system will 
-          typically result in a map $G_T$ with less functional complexity.} 
-          \label{Table_FC}
-      \end{table}\label{tbl_attractorsPearson}
-\end{center}
-
-\chapter{Conclusion}\label{ch6}
-
-The driven states $(x_{n-1},x_n)$ holds the information of the left-infinite history of the input.
-Furthermore, there is some evidence that the inverse-limit space (not system) seems to distinguish two dynamical systems that are not topologically conjugate for a class of systems in the sense their inverse-limit spaces are not homeomorphic to each other.
-So the relation $Y_T$, which is homeomorphic to the inverse-limit system, may be unique to a class of systems.
-(The reference is a paper on Ingram’s Conjecture by Marcy Barge, Henk Bruin and Sonja Stimac.) Further investigation is needed.
-
 \end{document}

### 2022-01-13 Gandhi_Project_Report/draft3.tex
Changed 4.3 and H2. Conclusion writtend and added

@@
 
-\section{Practical Issues of Takens theorem}
+\section{Practical Limitations to Takens' Theorem}
 
@@
 
+\ednote{Include note on learning failure because of sensitivity as in graph commented out?}
 % \begin{figure}[ht]
@@
 \begin{figure}[ht]
-  \includegraphics[scale=0.3]{_actionofh_H.png}
-  \centering
-  \captionof{figure}{Progressive actions of $h$ and $H$ to illustrate the causal mapping. }
+  \includegraphics[scale=0.3]{_actionofh_H.eps}
+  \centering
+  \captionof{figure}{Ladder-like behaviour of $h$ and $H$ to illustrate the causal mapping. }
   \label{fig:actionh_H}
@@
 
-The driven states $(x_{n-1},x_n)$ holds the information of the left-infinite history of the input.
+% The driven states $(x_{n-1},x_n)$ hold the information of the left-infinite history of the input.
 Furthermore, there is some evidence that the inverse-limit space (not system) seems to distinguish two dynamical systems that are not topologically conjugate for a class of systems in the sense their inverse-limit spaces are not homeomorphic to each other.
@@
 (The reference is a paper on Ingram’s Conjecture by Marcy Barge, Henk Bruin and Sonja Stimac.) Further investigation is needed.
+\cite{budivsic2012applied, korda2020data}
+
+As technology has advanced over the past century, humanity’s ability to record and manipulate rich, complex data in order to produce forecasts has profoundly increased. 
+Collecting measurements from underlying artificial, natural, and physical systems through observables to realise a dynamical model with greater descriptive power is well-established in present-day science.
+
+A key to the success of these techniques is the fact that dynamics discerned from the observed space can often be characterised by a mapping of much lower functional complexity when compared to the map describing the unknown underlying system.
+Takens theorem and the numerous generalizations (e.g., [5, 6, 7, 8, 8]) thereof is one such method establishing (under some generic conditions) the learnability of a system produced by creating a delay-coordinate vector from a sufficiently large number of previous observations of a dynamical system.
+This system is topologically conjugate to the system from which the observed time-series was obtained.
+
+A critical concern of this method is the requirement of full knowledge of the variables of the underlying state space that govern the complex dynamical system one is interested in.
+In real-world conditions this intelligence is seldom available and one must depend heavily on human expertise and insight.
+
+Each discrete-time state space model has at its core a driven dynamical system.
+In this project, it has been shown that an input originating from a compact space entailing exactly one solution (i.e., if the driven system has the USP) is equivalent to representing information without distortion, a concept which was formalised by establishing a causal mapping’s existence.
+The established results are general, implying that any driven dynamical system with specific properties can give rise to a causal mapping, and in so doing  providing an answer as to the success of arbitrary choices of driven systems in information processing applications in [4, 5].
+
+In particular, we have a topological conjugacy (semi-conjugacy) between the data comprising the 11single-delay lag dynamics of a driven system’s and the underlying homeomorphic dynamical system.
+Finite-length input data is sufficient for forecasting as driven system states are influenced only in an almost negligibly small manner by the left-infinite segment of the past of the input beyond some arbitrarily large but finite time (a result which holds due to the universal semi-conjugacy of a driven system’s continuity)
+
+Viewed through the lens reservoir computing approaches, we have also demonstrated that when single-delay dynamics are utilised, a learnable map in a reservoir computing network certainly exists.
+Moreover, such a driven system within a reservoir computing network enjoys certain proven qualities pertaining to the robustness w.r.t input- and parameter-perturbation.
+Accordingly, we are able to preserve the quality of a representation that would otherwise remain sensitive to various fluctuations, rendering the learning of the system’s dynamics wholly unpredictable.
+In addition, we obtain the robustness of the system against external noise, a phenomenon which is observed for the Fractal Dream Attractor.
+
+Finally, we consider some physical systems to observe the methodology in practice and verify that we can indeed predict the future evolution of chaotic systems such as the double pendulum.
+Empirical evidence shows that a linearity measure like a Pearson correlation coefficient points at reduced functional complexity while learning the dynamics in the state space of the driven system
+
+Considered with a philosophical-scientific lens, this project has aimed to establish a juncture between rationalism (the theory of nonautonomous dynamical systems) and empiricism (machine learning) to yield results profitable to the discipline of Applied Mathematics.
+This was done by showing that dynamics in an observed space can be established as topologically conjugate to the underlying system.
+As a result of the topological conjugacy, deep learning techniques are then subsequently applied to physically such equations from the data.
+In so doing, we acquire models of high fidelity that provide exceptional forecasting results on chaotic data that exhibit both statistical and topological consistency in the long run.
 

### 2022-01-13 Gandhi_Project_Report/draft4.tex
Changed 4.3 and H2. Conclusion writtend and added

--- 2022-01-11 Gandhi_Project_Report/draft3.tex
@@
 
+\ednote{Include note on learning failure because of sensitivity as in graph commented out?}
 % \begin{figure}[ht]
-%   \includegraphics[scale=0.3]{_learningfailure.png}
+%   \includegraphics[scale=0.3]{_learningfailure.eps}
 %   \centering
@@
 
-Having already defined the reachable set $X_U$, we now pause for a moment to fix additional notation.
-Letting $\cev{u}^{n}:=(\ldots,u_{n-2} ,u_{n-1})$ be the left-infinite subsequence of an input, $\overleftarrow{U}$ then denotes all these left-infinite sequences in $U$. 
-Moreover, $\cev{u}^{n}v:=(\ldots,u_{n-2} ,u_{n-1}, v)$ is to symbolise the input up to time $n$ with $v \in U$ being the specific input value at time $n$. 
-The introduction of a new input at time $n$ can be described by a mapping $\sigma_v:   \cev{u}^{n} \mapsto \cev{u}^{n}v$.
-
-The question now becomes whether we may establish a semi-conjugacy %\ednote{M: I have changed it to a semi-conjugacy since conjugacy amounts to asking a infinite-dimensional space $\overleftarrow{U}$ to be embedded in $X_U$ which is not possible.}
-as presented below for the driven system $g$.
+Having already defined the reachable set $X_U$ as the set of all elements of all entire solutions, we pause briefly in order to fix additional notation.
+Defining $\cev{u}^{n}:=(\ldots,u_{n-2} ,u_{n-1})$ as the left-infinite subsequence of an input up until time $n$, $\overleftarrow{U}$ is then the notation for all these left-infinite sequences in $U$.
+Moreover, $\cev{u}^{n}v:=(\ldots,u_{n-2} ,u_{n-1}, v)$ will symbolise the input up to time $n$ with $v \in U$ being the specific input value at time $n$. 
+The introduction of a new input at time $n$ can be described by the mapping $\sigma_v:   \cev{u}^{n} \mapsto \cev{u}^{n}v$. 
+The right-shift map, $r\overleftarrow{U}\to\overleftarrow{U}$, of an input sequence is defined $r: (\cdots, u_{-2},u_{-1}) \mapsto(\cdots, u_{-3},u_{-2})$
+
+The goal now assumes the form of the question on our ability to establish a semi-conjugacy as presented below for the driven system $g$. 
+%\ednote{M: I have changed it to a semi-conjugacy since conjugacy amounts to asking a infinite-dimensional space $\overleftarrow{U}$ to be embedded in $X_U$ which is not possible.}
 
@@
 
-Continuing to refine our attention, we proceed to consider a specific subclass of conjugacies.
+We proceed to consider a specific subclass of conjugacies.
 
   \begin{Definition}
-    [\bf Universal Semi-Conjugacy]\label{Def_UnivSemiConj}
+    [\bf Universal Semi-Conjugacy]\label{Def_UnivSemiConj} \rm
     Given a driven system $g$, we  call a continuous and surjective map $h : \overleftarrow{U} \to X_U$ a universal semi-conjugacy if  diagram \ref{Scomm_h} commutes for all $v \in U$.
@@
 
-We direct our attention to the function \[H_2{\overline{}} := (h(r\overline{u}, h(\overline{u}))) \]
-
-An embedding of the space $\overleftarrow{U}$  in $\overleftarrow{U}$ would allow one to establish topological conjugacy, which in turn provides stronger results than merely obtaining a coarse-grained representation via a causal mapping \cite{M: I feel this is very confusing to the reader as the co-domain of causal mapping is the left-infinite space and , so please omit if possible.
-You can just say we want to restrict inputs to the left-infinite orbits of a dynamical system and then explore conjugacy instead of a mere semi-conjugacy}.
-
-Before formalising this, we make mention of the concept of the inverse-limit system of a dynamical system $(U,T)$. 
-These spaces are noteworthy as their topology can help one ascertain much about the original system \cite{Manju_IEEE}.
-The topological structure for inverse-limit systems of chaotic dynamical systems is complicated, so we relegate further discussion to \cite{kennedy2008inverse_limit, ingram2011inverse}.
-In broad sweeps, the inverse-limit space of a dynamical system is a self-map on a subset of an infinite-dimensional space where each point in the inverse-limit space corresponds to a backward orbit of the map $T$.
-Following \cite{manjunath2021universal, ingram2011inverse}:
+The driven system $g$ can induce an embedding of $\overleftarrow{U}$ in $\overleftarrow{X}$ as follows: 
+If the causal mapping $H:\overleftarrow{U}{\to}{\overleftarrow{X}_U}$ is injective (in addition to being surjective), it becomes the embedding of $\overleftarrow{U}$ in $\overleftarrow{X}$ induced by the driven system $g$.
+\ednote{Explain why $H_{-1}$ is continuous} and we refer to $H$ as a \emph{causal embedding}
+% \ednote{M : Causal embedding a dynamical system is different from the definition of causal embedding, so please edit/delete this paragraph and the next theorem and also in the remaining part till Section 1.2}.
+% \begin{Theorem}
+%  If $g(\ldotp,x)$ is invertible and has the USP then $H$ is a causal embedding. 
+% \end{Theorem}
+% \begin{proof}
+%   See \cite{manjunath2021universal}.
+% \end{proof}
+When $g$ has the USP, the diagram below (adopted from~\cite{Manju_Nonlinearity}) illustrates the operation of the mappings $h$ and $H$.
+The mapping $h:\overleftarrow{U}\to{X_U}$ is also considered an observable as mentioned in the introduction of Chapter \ref{ch3}).
+
+\begin{figure}[ht]
+  \includegraphics[scale=0.3]{_actionofh_H.eps}
+  \centering
+  \captionof{figure}{Ladder-like behaviour of $h$ and $H$ to illustrate the causal mapping. }
+  \label{fig:actionh_H}
+\end{figure}
+
+We wish to explore the possibility of establishing a conjugacy rather than a mere semi-conjugacy and therefore are interested in the restriction of inputs to the left-infinite orbits of a dynamical system.
+The left-infinite orbit spaces are significant because a great amount of information about the original system can be gleaned from their topological structure~\cite{Manju_IEEE}. 
+The formal topological structure for the so-called inverse-limit space of chaotic dynamical systems is complicated, and further discussion is relegated to~\cite{kennedy2008inverse_limit, ingram2011inverse}.
+Broadly, we may conceive of the inverse-limit space of a dynamical system as a self-map on a subset of an infinite-dimensional space where each point in the inverse-limit space corresponds to a backward orbit of the map $T$.
+Following~\cite{manjunath2021universal, ingram2011inverse} we provide a definition:
 
@@
 
-\begin{Definition}
-  [\bf Causal Embedding]\label{Dfn_CausalEmbed}\rm
-  A driven system $g$ is said to causally embed the dynamical system $(U,T)$ if 
-  \vspace{-8mm}
-\begin{enumerate}[noitemsep, label=\roman*.]
-  \item The diagram in \ref{Scomm_h} commutes (i.e. $h$ is a universal semi-conjugacy),
-  \item $H_2(\overline{u}):=(h(r\overline{u}, h(\overline{u})))$ embeds the inverse-limit space $(\widehat{U}_T, \hat{T})$ of $(U,T)$ in $X\times{X}$.
-\end{enumerate}
-\end{Definition}
-
-The driven system $g$ can induce an embedding of $\overleftarrow{U}$ in $\overleftarrow{X}$ as follows: 
-If $H:\overleftarrow{U}{\to}{\overleftarrow{X}_U}$ is also injective (in addition to being surjective, and hence bijective), 
-it becomes the embedding of $\overleftarrow{U}$ in $\overleftarrow{X}$ induced by the driven system $g$ and we refer to $H$ as a \emph{causal embedding} 
-\ednote{M : Causal embedding a dynamical system is different from the definition of causal embedding, so please edit/delete this paragraph and the next theorem and also in the remaining part till Section 1.2}.
-
-\begin{Theorem}
- If $g(\ldotp,x)$ is invertible and has the USP then $H$ is a causal embedding. 
-\end{Theorem}
-\begin{proof}
-  See \cite{manjunath2021universal}.
-\end{proof}
-
-When $g$ has the USP, the diagram below illustrates the operation of the mappings $h$ and $H$.
-The mapping $h:\overleftarrow{U}\to{X_U}$ is an observable as discussed in an above section (\ref{TakensEmbedding}).
-
-\begin{figure}[ht]
-  %\includegraphics[scale=0.3]{actionofh_H.png}
-  \centering
-  \label{fig:actionh_H}
-\end{figure}
-
-The next result is of relative import but to fully comprehend its influence, some discussion still remains. We state it in part here.
-% Before stating it, we define a concept which will directly be shown to be equivalent to the USP.
-
-\begin{Theorem}
-  The following statements are equivalent:
-  \vspace{-8mm}
-  \begin{enumerate}[noitemsep, label=\roman*.]
-    \item $g$ has the USP. 
-    \item $g$ is a topological contraction.
-  \end{enumerate}
-\end{Theorem}
-% \begin{proof}
-%   (postponed) 
-% \end{proof}
+The inverse-limit space is well-defined since T : U →U is surjective by assumption.
 
@@
 
-One must be careful to avoid a choice of $g$ which would  add complexity to the obtained solution.
+The driving function $g$ cannot be chosen in a frivolous manner as it could possibly complicate our work.
+Specifically, one must be careful to avoid a choice of $g$ which would  add complexity to the obtained solution.
 
 When a causal embedding $H$ exists for the driven system $g$, one can map an arbitrary input ${u}$ onto the solution space $X$ without additional distortion or information-loss \textbf{(Cite.)}.
-When an embedding is established, the question of possible additional complexity in the solution is removed by guaranteeing that, since the systems are conjugate (semi-conjugate, \textbf{(refer)}), $g$ does not add any (some) complexity to the system.
+Established such an embedding eliminates the question of possible additional complexity in the solution by guaranteeing that, since the systems are conjugate (semi-conjugate, \textbf{(refer)}), $g$ does not add any (some) complexity to the system.
 
@@
 
-Note that $H_2$ maps an entire left-infinite solution sequence from $\Psi$ to an element in $X\times{X}$.
-We now have the following (compare with \ref{SCausal_H} above):
-
-% \ednote{Add Causal Embedding Theorem}
-% \textbf{NB NB NB - Causal Embedding Theorem! - Th.4}
-
-\begin{equation} \label{Fig_inverse_limsystem}
+Recall that $H_2$ maps an entire left-infinite solution sequence from $\Psi$ to an element in $X\times{X}$.
+We now have the following (which may be compared with \ref{SCausal_H} above):
+
+\begin{equation} \label{fig:inverse_limsystem}
   %  \[ 
@@
 
-The graph in equation \ref{Fig_inverse_limsystem} is then extended as below:
+The graph in equation \ref{fig:inverse_limsystem} is then extended as below:
 
 % \begin{figure}[ht]
-%   \includegraphics[scale=0.3]{_gammatemp.png}
+%   \includegraphics[scale=0.3]{_gammatemp.eps}
 %   \centering
@@
     \centering
-    \includegraphics[width=\linewidth]{Thomas_nonoise.eps}
+    \includegraphics[width=\linewidth]{_Thomas_nonoise.eps}
     \caption*{True and predicted trajectories in the absence of noise.}
@@
     \centering
-    \includegraphics[width=\linewidth]{ThomasNoisy.eps}
+    \includegraphics[width=\linewidth]{_ThomasNoisy.eps}
     \caption*{True and predicted trajectories for noise added.}
@@
   \centering
-  \includegraphics[scale=0.35]{Thomas_1.eps}\caption*{Predicted trajectories for the attractor demonstrate empirically the ability to predict the evolution of the trajectory for the next ~100 timesteps}
-  \includegraphics[scale=0.5]{Thomas_3.eps}\caption*{Represented here the learnt (blue) and actual (red) densities of the first coordinate of the dataset.}
+  \includegraphics[scale=0.35]{_Thomas_1.eps}\caption*{Predicted trajectories for the attractor demonstrate empirically the ability to predict the evolution of the trajectory for the next ~100 timesteps}
+  \includegraphics[scale=0.5]{_Thomas_3.eps}\caption*{Represented here the learnt (blue) and actual (red) densities of the first coordinate of the dataset.}
 \end{figure}
@@
 (The reference is a paper on Ingram’s Conjecture by Marcy Barge, Henk Bruin and Sonja Stimac.) Further investigation is needed.
-
-\ednote{B: Expand}
+\cite{budivsic2012applied, korda2020data}
+
+As technology has advanced over the past century, humanity’s ability to record and manipulate rich, complex data in order to produce forecasts has profoundly increased. 
+Collecting measurements from underlying artificial, natural, and physical systems through observables to realise a dynamical model with greater descriptive power is well-established in present-day science.
+
+A key to the success of these techniques is the fact that dynamics discerned from the observed space can often be characterised by a mapping of much lower functional complexity when compared to the map describing the unknown underlying system.
+Takens theorem and the numerous generalizations (e.g., [5, 6, 7, 8, 8]) thereof is one such method establishing (under some generic conditions) the learnability of a system produced by creating a delay-coordinate vector from a sufficiently large number of previous observations of a dynamical system.
+This system is topologically conjugate to the system from which the observed time-series was obtained.
+
+A critical concern of this method is the requirement of full knowledge of the variables of the underlying state space that govern the complex dynamical system one is interested in.
+In real-world conditions this intelligence is seldom available and one must depend heavily on human expertise and insight.
+
+Each discrete-time state space model has at its core a driven dynamical system.
+In this project, it has been shown that an input originating from a compact space entailing exactly one solution (i.e., if the driven system has the USP) is equivalent to representing information without distortion, a concept which was formalised by establishing a causal mapping’s existence.
+The established results are general, implying that any driven dynamical system with specific properties can give rise to a causal mapping, and in so doing  providing an answer as to the success of arbitrary choices of driven systems in information processing applications in [4, 5].
+
+In particular, we have a topological conjugacy (semi-conjugacy) between the data comprising the 11single-delay lag dynamics of a driven system’s and the underlying homeomorphic dynamical system.
+Finite-length input data is sufficient for forecasting as driven system states are influenced only in an almost negligibly small manner by the left-infinite segment of the past of the input beyond some arbitrarily large but finite time (a result which holds due to the universal semi-conjugacy of a driven system’s continuity)
+
+Viewed through the lens reservoir computing approaches, we have also demonstrated that when single-delay dynamics are utilised, a learnable map in a reservoir computing network certainly exists.
+Moreover, such a driven system within a reservoir computing network enjoys certain proven qualities pertaining to the robustness w.r.t input- and parameter-perturbation.
+Accordingly, we are able to preserve the quality of a representation that would otherwise remain sensitive to various fluctuations, rendering the learning of the system’s dynamics wholly unpredictable.
+In addition, we obtain the robustness of the system against external noise, a phenomenon which is observed for the Fractal Dream Attractor.
+
+Finally, we consider some physical systems to observe the methodology in practice and verify that we can indeed predict the future evolution of chaotic systems such as the double pendulum.
+Empirical evidence shows that a linearity measure like a Pearson correlation coefficient points at reduced functional complexity while learning the dynamics in the state space of the driven system
+
+Considered with a philosophical-scientific lens, this project has aimed to establish a juncture between rationalism (the theory of nonautonomous dynamical systems) and empiricism (machine learning) to yield results profitable to the discipline of Applied Mathematics.
+This was done by showing that dynamics in an observed space can be established as topologically conjugate to the underlying system.
+As a result of the topological conjugacy, deep learning techniques are then subsequently applied to physically such equations from the data.
+In so doing, we acquire models of high fidelity that provide exceptional forecasting results on chaotic data that exhibit both statistical and topological consistency in the long run.
 

### 2022-01-13 Gandhi_Project_Report/draft5.tex
Changed 4.3 and H2. Conclusion writtend and added

--- 2022-01-11 Gandhi_Project_Report/draft4.tex
@@
 
-\title{Project Report}
+\title{Data-Driven Modelling \& \\ Forecasting of Chaotic Systems}
 
-\author{} 
+\author{B.L. Nortier} 
+\date{January, 2022}
 \date{}
-\date{}
-% \maketitle
+\maketitle
 % \tableofcontents
@@
 % \subsubsection{Fractal Dream Attractor}
-% \section{Generalized Pearson Correlation} %\ednote{B: Alter heading.}
+% \section{Generalised Pearson Correlation} %\ednote{B: Alter heading.}
 
-% \chapter{Conclusion}\label{ch6}
+\chapter{Conclusion}\label{ch6}
 
-% The driven states $(x_{n-1},x_n)$ holds the information of the left-infinite history of the input.
+% The driven states $(x_{n-1},x_n)$ hold the information of the left-infinite history of the input.
 Furthermore, there is some evidence that the inverse-limit space (not system) seems to distinguish two dynamical systems that are not topologically conjugate for a class of systems in the sense their inverse-limit spaces are not homeomorphic to each other.
 So the relation $Y_T$, which is homeomorphic to the inverse-limit system, may be unique to a class of systems.
-(The reference is a paper on Ingram’s Conjecture by Marcy Barge, Henk Bruin and Sonja Stimac.
-) Further investigation is needed.
+(The reference is a paper on Ingram’s Conjecture by Marcy Barge, Henk Bruin and Sonja Stimac.) Further investigation is needed.
+\cite{budivsic2012applied, korda2020data}
+
+As technology has advanced over the past century, humanity’s ability to record and manipulate rich, complex data in order to produce forecasts has profoundly increased. 
+Collecting measurements from underlying artificial, natural, and physical systems through observables to realise a dynamical model with greater descriptive power is well-established in present-day science.
+
+A key to the success of these techniques is the fact that dynamics discerned from the observed space can often be characterised by a mapping of much lower functional complexity when compared to the map describing the unknown underlying system.
+Takens theorem and the numerous generalisations (e.g., [5, 6, 7, 8, 8]) thereof is one such method establishing (under some generic conditions) the learnability of a system produced by creating a delay-coordinate vector from a sufficiently large number of previous observations of a dynamical system.
+This system is topologically conjugate to the system from which the observed time-series was obtained.
+
+A critical concern of this method is the requirement of full knowledge of the variables of the underlying state space that govern the complex dynamical system one is interested in.
+In real-world conditions this intelligence is seldom available and one must depend heavily on human expertise and insight.
+
+Each discrete-time state space model has at its core a driven dynamical system.
+In this project, it has been shown that an input originating from a compact space entailing exactly one solution (i.e., if the driven system has the USP) is equivalent to representing information without distortion, a concept which was formalised by establishing a causal mapping’s existence.
+The established results are general, implying that any driven dynamical system with specific properties can give rise to a causal mapping, and in so doing  providing an answer as to the success of arbitrary choices of driven systems in information processing applications in [4, 5].
+
+In particular, we have a topological conjugacy (semi-conjugacy) between the data comprising the 11single-delay lag dynamics of a driven system’s and the underlying homeomorphic dynamical system.
+Finite-length input data is sufficient for forecasting as driven system states are influenced only in an almost negligibly small manner by the left-infinite segment of the past of the input beyond some arbitrarily large but finite time (a result which holds due to the universal semi-conjugacy of a driven system’s continuity)
+
+Viewed through the lens reservoir computing approaches, we have also demonstrated that when single-delay dynamics are utilised, a learnable map in a reservoir computing network certainly exists.
+Moreover, such a driven system within a reservoir computing network enjoys certain proven qualities pertaining to the robustness w.r.t input- and parameter-perturbation.
+Accordingly, we are able to preserve the quality of a representation that would otherwise remain sensitive to various fluctuations, rendering the learning of the system’s dynamics wholly unpredictable.
+In addition, we obtain the robustness of the system against external noise, a phenomenon which is observed for the Fractal Dream Attractor.
+
+Finally, we consider some physical systems to observe the methodology in practice and verify that we can indeed predict the future evolution of chaotic systems such as the double pendulum.
+Empirical evidence shows that a linearity measure like a Pearson correlation coefficient points at reduced functional complexity while learning the dynamics in the state space of the driven system
+
+Considered with a philosophical-scientific lens, this project has aimed to establish a juncture between rationalism (the theory of nonautonomous dynamical systems) and empiricism (machine learning) to yield results profitable to the discipline of Applied Mathematics.
+This was done by showing that dynamics in an observed space can be established as topologically conjugate to the underlying system.
+As a result of the topological conjugacy, deep learning techniques are then subsequently applied to physically such equations from the data.
+In so doing, we acquire models of high fidelity that provide exceptional forecasting results on chaotic data that exhibit both statistical and topological consistency in the long run.
 

### 2022-01-15 Gandhi_Project_Report/draft4.tex
Gandhi's edits of Ch.4

@@
 \maketitle
+
+\begin{abstract}
+  Collecting measurements from underlying artificial, natural, and physical systems through observables to realise a dynamical model with greater descriptive power is a well-established practice in present-day science.
+However, we often do not fully understand the many facets of the more complicated systems we interact with on a daily basis.
+
+  In this project we formulate the concept of discrete-time dynamical systems, particularly those which are chaotic.
+The learning problem is formulated and a well-known approach (first proposed by Takens) to solving the problem alongside its limitations is considered.
+  We then introduce the notion of driven dynamical systems and establish the theoretical existence of dynamical systems possessing certain properties that guarantee them to be topologically conjugate (or dynamically equivalent) to the dynamics of the underlying system from which measurements were taken.
+The methodology is practically implemented by making use of a recurrent neural network, a form of deep learning within the reservoir computing framework and we show that we may obtain long-term consistent predictions for simple physical systems such as the double pendulum as well as attractors exhibiting chaotic behaviour.
+
+  \end{abstract}
+
   \tableofcontents
-
-% \begin{abstract}
-%   Abstract goes here.
-%   \end{abstract}
 
@@
   For the map $Tu=u^2$ defined on $[0,1]$ all orbits lie in the invariant set $\{1\}$ or else would would approach the invariant set $\{0\}$. 
-  % \ednote{B: not relevant other than being an additional example perhaps? M: Have added another simple example}.
 \end{Example}
@@
 
-% \ednote{For let $B$ be an attractor of U.
-From above, we know that B is itself invariant, is asymptotically stable and has Lyapunov stability.
-Now suppose there is some attractor $C\subsetneq{U}=B$, so there is a $b\in{U}$ which is not in $C$.
-........
-M: Have changed the example}
-
 A dynamical system can have several attractors, and may also be contained in another attractor.
@@
 
-% \ednote{ should one add paragraphs like these?
-M: this is a thesis, and not a paper; the instructions sent to the external examiner regarding the thesis evaluation says that the student suggest strongly that it is good to write}
 In the 1960s, a number of mathematicians and mathematically interested scientists independently discovered chaos in the mathematical sense.
@@
 There are several intricately different definitions of chaos available in the literature \textbf{(Cite)}, with each of them indicating some aspect of complexity.
-%\ednote{B:  what are we saying here?
-M: Devaney's definition is not the only one, there are about half a dozen definitions as well}
 In practice, it is not possible to verify which definition or notion does the underlying dynamical system satisfies when only data is observed from it.
@@
   The logistic map(LM), a recursion relation of the form $x_{n+1}=rx_n(1-x_n)$ where $x_n\in[0,1]$ and $r<4$ is of particular interest to us as it is classically used to illustrate the influence of chaos on a system \textbf{(Cite)}.
-Below are plotted the first 75 iterates of the LM for $r=3.7$ and $x_0=0.2$ with an adjustment of 0.001 in $x_0$ and then $r$.
+Below are plotted the first 75 iterates of the LM for $r=3.7$ and $x_0=0.2$ with an adjustment of 0.001 in $x_0$ and $r$ respectively.
 
@@
         \label{fig:log_sdic}
-        \captionof{figure}{Logistic Map for $r=3.7$, $x_0=0.2$}
+        \captionof{figure}{Figure generated for Logistic Map for $r=3.7$, $x_0=0.2$ where each parameter is varied separately by 0.001.}
       \end{figure}
@@
 \includegraphics[scale=0.8]{_dp_sdic.eps}
-    \captionof{figure}{Three double pendulum heads with initial angle differing by 0,025 radians}\label{fig:dp_sdic}
+    \captionof{figure}{Three double pendulum heads with equal angular velocity and initial angles differing by 0,025 radians}\label{fig:dp_sdic}
   \end{figure}
@@
         \centering
-        \captionof{figure}{Three double pendulum heads with initial angle differing by 0,025 radians}\label{fig:T2tentmap}
+        \captionof{figure}{Graph of the $T^2$ map with bases at $x=0$, $x=\frac{1}{2}$ and $x=1$ respectively.}\label{fig:T2tentmap}
   \end{figure}
@@
 
-%More formally we may say that in the process of measuring a system, we obtain a 1-dimensional time-series $\{\theta(u(t))\}$, where $\theta:\mathbb{R}^n\to\mathbb{R}$ is a smooth observation or measurement function.
-This series represents a possibly noise-augmented \ednote{M: Noise-augmented is not the right word} dataset containing information of certain parts of the original system.
-
 %We may then ask ourselves the question: Is it in any way possible to retain information about the system state $x(t)$ in this temporal data-series $\theta(x(t))$?
@@
 By generic we mean a residual set on a certain topology on appropriate spaces of functions (that we do not describe here) \textbf{(Cite)}. 
-%\ednote{How vague may one be in general by leaving things out?}
 By $C^2$, we make reference to a twice-differentiable function  with a continuous second derivative.
@@
   \centering
-  \captionof{figure}{Schematic of Takens Embedding Theorem's usage and consequence of how iterates \ldots }
+  \captionof{figure}{Schematic of Takens Embedding Theorem's usage and impact of iterates slipping from the attractor when the approximation $\tilde{F}_\theta$ of $F_\theta$ is learnt.}
   \label{fig:takensmap}
 \end{figure}
-\ednote{M: Here, and throughout if you reproduce a figure, it must be stated where it was reproduced from. B: Is this satisfactory? Also, should $\tilde{F}$ be left out for now?}
-\ednote{B: This is only a screenshot used as a temporary placeholder. If sufficient, I shall convert it to .eps. If not, may I perhaps have the original picture?}
 
@@
 
-% Another limitation owes itself to the requirement that T be a surjective map.
-This is not always the case in practical examples and we may consider the case of a damped double pendulum \ednote{M: Please remove the paragraph as well.
-We do not overcome this problem.
-You can add this in the next Chapter where you need T to be surjective or else you could include this at the begining of this chapter where you explain the forecasting problem.}
-
 The Takens embedding theorem does not guarantee global dissipativity.
@@
 This requirement is not overly restrictive, as a great number of chaotic systems do have a surjective map.
-(Cite) \ednote{I rememeber you saying this in one of our sessions, but don't have a source for it.}
+(Cite) \ednote{B: I remember you saying this in one of our sessions, but don't have a source for it.}
 
@@
 The SDD can then be used to forecast and reconstruct the underlying system.
-% \ednote{Is this attractor assumed to exist? M:  I replaced  it with the system. Attractor is also a commonly used term since we observe data from an attractor in almost all cases.}
 
@@
 
-To find a function $\phi$, the need now arises to consider a so-called driven dynamical system, a special case of a nonautonomous dynamical system, where two spaces are considered: input is taken from an input space $U$ and the state of a space $X$ is updated.
+To find a function $\phi$, the need now arises to consider a so-called driven dynamical system, one that can generate a  nonautonomous dynamical system, where two spaces are considered: input is taken from an input space $U$ and the state of a space $X$ is updated.
 (Previously, only the state was considered.)
 We do this to mimic the true conditions.
-By accounting for both external and internal factors which influence the  system-state at a specific time-step, our model resembles nature much more realistically.
-\ednote{M : I understand this sentence, but am not sure if a general audience would.
-So please elaborate or omit.
-B: Is this satisfactory?}
+By accounting for an extrogenous input sequence in $U$ which can also influence the  system-state at a specific time-step, it produces more general models than an autonomous system.
 
@@
 \end{Example}
-% \ednote{Nicely written!}
+% \ednote{Nicely written! Thank you}
 
@@
 \begin{Example}\label{Ex_exp} \rm 
- Consider the driven system $g(u_n,x_n)=x^n$ for $X=[0,1]$, $U=\mathbb{R}$.
+  Consider the driven system $g(u,x)=x^2$ for $X=[0,1]$, $U=\mathbb{R}$.
 The system has an uncountable number of solutions, as there exists a solution for every $x\in{X}$ which also passes through the point $x$ and $lim_{n\to\infty}x_n=0$, $lim_{n\to{-}\infty}x_n=1$.
@@
   [\bf Proof of Example~\ref{Ex_exp}] \label{rem_proofEx} \rm
-  To see this, we show that for every input $\overline{u}=\{u_n\}_{n\in\mathbb{Z}}$, $g(u_n,\cdot)$ is a contraction map on $X$ for every $k\in\mathbb{Z}$. 
-  Indeed $g(u_n,x_n)=x^n$ depends only on the state at time $n$ and so for each input $\overline{u}$, $X_n(\overline{u})=\{x^n\}$ is a singleton subset.
-  We then easily conclude that $\{x_n\}\equiv0$ is the only solution. 
-\end{Remark} \ednote{B: Is this sufficient?}
+  For the driven system,  
+  $g(u,x)=x^2$ the evolution does not depend on the input at all.
+Hence $g(u,x)$ can be written as a single map $f(x)=x^2$ which is a homeomorphism on $[0,1]$.
+A left-infinite orbit is defined and it converges to $1$ while the right-infinite orbit converges to $0$.
+Or in other words the iterates of $f^{-1}$ converges to $1$ while the iterates of $f$ converge to $0$
+\end{Remark}
 
@@
 
-One may easily construct many systems with trivial solution-sets, such as $g(u,x)=x$ which has only the constant solution $x$ and so for $U=[-1,1]$, the system would have no solution if $|x|>1$.
+% One may easily construct many systems with trivial solution-sets, such as $g(u,x)=x$ which has only the constant solution $x$ and so for $U=[-1,1]$, the system would have no solution if $|x|>1$.
 To refine the scenario, we consider only systems with unique solutions.
@@
 
-% \ednote{M: You can omit the theorem, and just say it is a uniform attractor in the nonautonomous dynamical systems literature \cite{Manju_Nonlinearity}, and say the discussion on nonautonomous attractors is beyond the scope of the thesis.}
 % Paraphrased.
@@
 
-In the majority of Reservoir Computing(RC) literature (a machine learning algorithm approach)\ednote{How should I explain/qualify RC better here? It's mentioned without really explaining.}, 
-the ESP is used to describe the concept of fading memory or memory-loss  \cite{jaeger2004harnessing, lu2018attractor}. 
+In the majority of Reservoir Computing(RC) literature (a machine learning approach) where the input is mapped onto a different but higher dimensional through a system called the reservoir and a readout that measures the state of the reservoir is trained, a notion of forgetting the states of the reservoir asymptotically is often used.
+
+% \ednote{How should I explain/qualify RC better here?
+It's mentioned without really explaining.
+M: Have put it in one sentence; just to let you know people use optical devices to map signal into a ``reservoir".
+B: Thanks!},
+Concepts like the ESP,  fading memory \cite{boyd1985fading} are some of these notions.  
 If $g$ possesses the ESP (equivalent in our context to the USP), the we are guaranteed that the whole of an input's left-infinite history will exactly determine the current system state; i.e.
-there is only ever the possibility of a single reachable state at a given instance of time \cite{jaeger2001echo}
-
-Defined formally, the ESP is intrinsically linked to an input (see \cite{Manju_ESP}) Unfortunately, definitions and results concerning the ESP do not have much practical benefit since a relationship between the network's response and the temporo-statistical properties of the input are not directly described.
+there is only ever the possibility of a single reachable state at a given instance of time \cite{jaeger2001echo,Manju_2020}.
+% Defined formally, the ESP is intrinsically linked to an input (see \cite{Manju_ESP}) Unfortunately, definitions and results concerning the ESP do not have much practical benefit since a relationship between the network's response and the temporo-statistical properties of the input are not directly described.
 A more detailed presentation may be found in~\cite{jaeger2001echo}.
 
-We do however consider it important to mention that the ESP is also often discussed in terms of a stability property that holds for all inputs of a system~\cite{manjunath2020stability} and, furthermore, plays a key role in the design and training of recurrent neural networks(RNN) within the field of RC \cite{Manju_ESP}.
-
-\section{Next section?}
-\ednote{I feel a break is necessary here, but would hear from you what your thoughts are.}
-
-Having already defined the reachable set $X_U$ as the set of all elements of all entire solutions, we pause briefly in order to fix additional notation.
+% We do however consider it important to mention that the ESP is also often discussed in terms of a stability property that holds for all inputs of a system~\cite{manjunath2020stability} and, furthermore, plays a key role in the design and training of recurrent neural networks(RNN) within the field of RC \cite{Manju_ESP}.
+Beyond forgetting the past states the ESP is also often discussed in terms of a stability property that is nearby inputs create close by responses (solutions), a concept called input-related stability~\cite{manjunath2020stability} and, furthermore, plays a key role in the robustness of recurrent neural networks(RNN).
+
+\section{Conjugacies}
+
+Having already defined the reachable set $X_U$ as the collection of all elements of all entire solutions, we pause briefly in order to fix additional notation.
 Defining $\cev{u}^{n}:=(\ldots,u_{n-2} ,u_{n-1})$ as the left-infinite subsequence of an input up until time $n$, $\overleftarrow{U}$ is then the notation for all these left-infinite sequences in $U$.
@@
 The goal now assumes the form of the question on our ability to establish a semi-conjugacy as presented below for the driven system $g$.
-%\ednote{M: I have changed it to a semi-conjugacy since conjugacy amounts to asking a infinite-dimensional space $\overleftarrow{U}$ to be embedded in $X_U$ which is not possible.}
 
@@
 
-If the universal semi-conjugacy $h$ exists (i.e.
-the diagram in \ref{Scomm_h} commutes) then the solution $\Psi$ will be at least a coarse-grained representation of the input $u$.
-\ednote{B: Expand/Cite}
+  If the universal semi-conjugacy $h$ exists (i.e. the diagram in \ref{Scomm_h} commutes) then the solution $\Psi(\bar{u})$ will intuitively have no more ``complexity' than the input $\bar{u}$.
 
@@
 
-Re-sketching the commutativity diagram \ref{Scomm_h} above by fixing the input $v$ in $g$ and replacing $X_U$ by its left-infinite sequence space $\overleftarrow{U}$, we obtain the diagram below.
-In this case, the function $H:\overleftarrow{U}\to\overleftarrow{X}_U$, a map that is both continuous and surjective, is called a \emph{causal mapping}.
+Re-sketching the commutativity diagram \ref{Scomm_h} above by replacing $X_U$ by its left-infinite sequence space $\overleftarrow{X}_U$, we obtain the diagram below.
+In this case, the function $H:\overleftarrow{U}\to\overleftarrow{X}_U$, a map that is both continuous and surjective, is called a \emph{causal mapping} that is defined next.
 
@@
   [\bf Causal Mapping]\label{Def_CausMap}
-  A continuous, surjective map $H:\overleftarrow{U}\to\overleftarrow{X}_U$ such that \[H\circ\tilde{g}_v=\sigma\circ{H}\] where $\tilde{g}_v$ maps $(\ldots, u_{-2}, u_{-1})$ to $(\ldots, u_{-2}, u_{-1}, g(v, u_{-1}))$ is called a causal mapping.
+  A continuous, surjective map $H:\overleftarrow{U}\to\overleftarrow{X}_U$ such that \[H\circ\tilde{g}_v=\sigma\circ{H}\] holds for all $v \in U$ where $\tilde{g}_v$ maps $(\ldots, u_{-2}, u_{-1})$ to $(\ldots, u_{-2}, u_{-1}, g(v, u_{-1}))$ is called a causal mapping.
 \end{Definition}
@@
 \end{proof}
+\ednote{M: this sentence is freely hanging} To establish whether $\widehat{T}$ and $G_T$ are related, we consider the the function $H_2{\overline{}} := (h(r\overline{u}, h(\overline{u})))$ where $H_2$ is a mapping from the entirety of a left-infinite sequence to some element of $X\times{X}$.
 
@@
 \ednote{Explain why $H_{-1}$ is continuous} and we refer to $H$ as a \emph{causal embedding}
-% \ednote{M : Causal embedding a dynamical system is different from the definition of causal embedding, so please edit/delete this paragraph and the next theorem and also in the remaining part till Section 1.2}.
-% \begin{Theorem}
-%  If $g(\ldotp,x)$ is invertible and has the USP then $H$ is a causal embedding. 
-% \end{Theorem}
-% \begin{proof}
-%   See \cite{manjunath2021universal}.
-% \end{proof}
+
 When $g$ has the USP, the diagram below (adopted from~\cite{Manju_Nonlinearity}) illustrates the operation of the mappings $h$ and $H$.
@@
 \begin{figure}[ht]
-  \includegraphics[scale=0.3]{_actionofh_H.eps}
-  \centering
-  \captionof{figure}{Ladder-like behaviour of $h$ and $H$ to illustrate the causal mapping. }
+  \includegraphics[scale=0.4]{_actionofh_H.eps}
+  \centering
+  \captionof{figure}{Ladder-like behaviour of $h$ and $H$ to illustrate the causal mapping. (Figure reproduced from~\cite{Manju_Nonlinearity}) }
   \label{fig:actionh_H}
@@
 When a causal embedding $H$ exists for the driven system $g$, one can map an arbitrary input ${u}$ onto the solution space $X$ without additional distortion or information-loss \textbf{(Cite.)}.
-Established such an embedding eliminates the question of possible additional complexity in the solution by guaranteeing that, since the systems are conjugate (semi-conjugate, \textbf{(refer)}), $g$ does not add any (some) complexity to the system.
+When an embedding is established, the question of possible additional complexity in the solution is removed by guaranteeing that, since the systems are conjugate (semi-conjugate, \textbf{(refer)}), $g$ does not add any (some) complexity to the system.
 
@@
   \centering
-\captionof{figure}{Reproduced from \cite{manunath2021universal}. Illustration of all trajectories approaching the true solution if $g$ has the UAP}
+\captionof{figure}{Illustration of trajectories that all approach the true solution if $g$ has the UAP. Reproduced from \cite{manunath2021universal}.}
 \label{fig:memloss_conttime} 
@@
 
-It is of interest to ascertain whether $\widehat{T}$ and $G_T$ are related\ednote{B: Should I expand more here?}, and we consequently introduce  the function $H_2{\overline{}} := (h(r\overline{u}, h(\overline{u})))$ mapping the entirety of a left-infinite sequence to some element in $X\times{X}$.
+It is of interest to ascertain whether $\widehat{T}$ and $G_T$ are related \ednote{B: Should I expand more here?}, and we consequently introduce  the function $H_2{\overline{}} := (h(r\overline{u}, h(\overline{u})))$ mapping the entirety of a left-infinite sequence to some element in $X\times{X}$.
 Indeed, it may be shown that $g$ having SI-invertibility and the USP immediately guarantees the existence of at least a semi-conjugacy between the systems $(Y_T, G_T)$ and $(\widehat{U}_T, \widehat{T})$.
@@
 \begin{figure}[ht]
-  \includegraphics[scale=0.5]{_AutovsDriven.eps}
+  \includegraphics[scale=0.3]{_autovsdriven.eps}
   \centering
@@
   \captionof{figure}{Illustrating empirically the importance of global dissipativity.
-Plotted below the three principal components of randomly initiated trajectories of a learnt version of $G_T$ versus the trajectory $(x_{n-1}, x_n)$(grey) of driven stated obtained from the Thomas Attractor's data (see below).
+Plotted below the three principal components of randomly initiated points(blue) of a learnt version of $G_T$ versus the trajectory $(x_{n-1}, x_n)$(red) of driven states obtained from the Fractal Dream Attractor's data~\ref{Thomas_Attractor}).
+Figure generated by altering the code used to produce Figure 7 in \cite{Supp}.
 }
@@
 }
-
-% The $\omega$-limit set $\omega(u;T)$ of a point $u$ is defined to be the collection of limit points of the sequence $\{x,Tu,T^2u,\ldots\}$.
-This set is nonempty, and $\omega(u;T)$ is invariant.
-\textbf{(Change)}
-% \begin{Definition}
-%   [\bf Global Dissipativity]\label{Dfn_GlobalDiss}\rm
-% 	We say a dynamical system $(Z,f)$ is globally dissipative if there exists a nonempty proper closed subset $B$ of $Z$  so that for all $x\in Z\setminus B$, (i).
-$\omega(x;f) \subset B$  and (ii).
-$B$ is positively invariant that is $f(B) \subset B$.
-We call any such closed subset $B$ a trapping set of $f$.
-% \end{Definition}
 
@@
 
-We have now seen that one may comfortably switch between the use of SDD or MDD when $g$ has SI-invertibility. /ednote{This sentence needs to be improved.}
+% We have now seen that one may comfortably switch between the use of SDD or MDD when $g$ has SI-invertibility. /ednote{This sentence needs to be improved.}
 
@@
   \centering
-\caption{The map generating data for a damped pendulum is not surjective.}
+\caption{The map generating data for a damped pendulum is not surjective. Data obtained from chaotic double pendulum dataset generated by~\cite{asseman2018learning}}
 \label{fig:damped_pendulum}
@@
 \begin{figure}[ht]
-  \includegraphics[scale=0.4]{_dpfailure_nonsurj.eps}
+  \includegraphics[scale=0.4]{_dpfail_nonsurj.eps}
   \centering
@@
 
-\subsubsection{Thomas' Cyclically Symmetric Strange Attractor}
+\subsubsection{Thomas' Cyclically Symmetric Strange Attractor}\label{Thomas_Attractor}
 Thomas' Cyclically Symmetric Strange Attractor, a 3D attractor proposed in 1999 by Ren\'e Thomas in \cite{ThomasAttractor}, is described by a set of three equations which is distinctly reminiscent of the Lorenz system, another deterministic three-equation model exhibiting chaotic behaviour.It has a single parameter $\beta$ and has been shown to transition to chaotic behavious when $\beta<0.208186$~\cite{Thomas_BetaParameter}.
@@
 \section{Generalized Pearson Correlation}\ednote{B: Alter heading.}
+
+% Besides being able to learn the action of the Koopman operator exactly, our setup has other advantages over selecting or optimizing the choice of observables as in the Extended Dynamic Mode Decomposition (EDMD) algorithm (e.g., [21, 22]).
+We can alter the functional complexity of G T by changing the number of observables or equivalently by altering the dimension of X – in the case of a RNN implementation of g, by simply changing the number of neurons in the network.
+Empirically, increasing the dimension of X is found to increase the linear relationship (or intuitively reduces the functional complexity of G T ) that is measured as a generalization of the Pearson correlation coefficient to random vectors (e.g., [44]) between (x_{n−1} , x_{n}) and G_T (x_{n−1} , x_n), and such numerical evidence is tabulated in [11, Table 1].
+In contrast, expanding the set of observables while using EDMD to learn a map with a lower functional complexity is a highly involved task, not least because so much that one often does not know how to guess a new observable, as well as the fact that adding observables does not necessarily retain the invariance of the span of the observables.
+
+% Empirical evidence shows that a linearity measure like a Pearson correlation coefficient points at reduced functional complexity while learning the dynamics in the state space of the driven system.
+
+% Empirically (see Table ??), increasing the dimension of X is found to increase the linear relationship
+% (or intuitively reduce the functional complexity of G T ) that is measured as a generalization of the Pearson correlation coefficient to random vectors (e.g., [?]) between (x n−1 , x n ) and G T (x n−1 , x n ).
+If Σ a and Σ b denotes the covariance matrices of the vectors (x n−1 , x n ) and G T (x n-1 , x n ) respectively, and if Σ ab denotes the covariance matrix between the vectors (x n-1 , x_n ) and G_T (x n-1 , x_n ), then the multidimensional 18correlation coefficient is computed by using the traces of these matrices:
+% \[\pho = \frac{tr(\sum_{ab})}{tr(\sqrt{\sum_{a}\sum_{b}})}\]
+
+% This multidimensional correlation coefficient satisfies most of the well-known properties of the one-dimensional Pearson coefficient.
+In particular, ρ = ±1 if and only Y = AX + b for some invertible matrix A and vector b.
+Hence ρ, and in practice, the estimator ρ̂, can be used to measure the linear relationship between two random vectors of the same dimension and thus serves as an indicator of functional complexity.
+
+% We remark that the estimation of such linear relationship in (see Table ??) obtained by sample correlations alone is justified whenever {u n } is a realization of an ergodic process since then when g has the USP, any solution {x n } is also a realization of an ergodic process [?].
+In the ergodic input case the generalized Pearson correlation coefficient (between (x n−1 , x n ) and G T (x n−1 , x n )) is independent of n.
 
@@
 
-A key to the success of these techniques is the fact that dynamics discerned from the observed space can often be characterised by a mapping of much lower functional complexity when compared to the map describing the unknown underlying system.
-Takens theorem and the numerous generalizations (e.g., [5, 6, 7, 8, 8]) thereof is one such method establishing (under some generic conditions) the learnability of a system produced by creating a delay-coordinate vector from a sufficiently large number of previous observations of a dynamical system.
+Core to the success of these techniques is the fact that dynamics discerned from the observed space can often be characterised by a mapping of much lower functional complexity when compared to the map describing the unknown underlying system.
+Takens theorem and the numerous generalizations thereof is one such method establishing (under some generic conditions) the learnability of a system produced by creating a delay-coordinate vector from a sufficiently large number of previous observations of a dynamical system.
 This system is topologically conjugate to the system from which the observed time-series was obtained.

### 2022-01-15 Gandhi_Project_Report/draft5.tex
Gandhi's edits of Ch.4

--- 2022-01-13 Gandhi_Project_Report/draft4.tex
@@
 \maketitle
+
+\begin{abstract}
+  Collecting measurements from underlying artificial, natural, and physical systems through observables to realise a dynamical model with greater descriptive power is a well-established practice.
+Physical systems can often be modelled by instance with differential equations derived from the laws of physics, if we knew the laws.
+However, we often do not fully understand the many complexities of the more intricate systems we interact with on a daily basis.
+  In this project we start with the hypothesis that an discrete-time dynamical systems, particularly those which are chaotic generates the data.
+The general learning problem is then formulated and a well-known approach to the problem, alongside its practical limitations are considered.
+  We then introduce the notion of driven dynamical systems and establish the theoretical existence of dynamical systems possessing certain properties that guarantee them to be topologically conjugate ('dynamically equivalent') to the dynamics of the underlying system from which measurements were taken.
+The methodology is practically implemented by making use of a recurrent neural network and using deep learning methods to train on the recurrent neural network states to show that we may obtain long-term consistent predictions for simple physical systems such as the double pendulum as well as a few other attractors exhibiting chaotic behaviour.
+The machine learning algorithms for modelling the double pendulum thus far have not been successful when it exhibits chaos.
+
+  \end{abstract}
+
   \tableofcontents
-
-% \begin{abstract}
-%   Abstract goes here.
-%   \end{abstract}
 
@@
   Given any point $u$ and an open neighborhood $W \subset[0,1]$ of $u$, we can find $k\in\mathbb{Z}$ large enough so that $T^k(W) = [0,1]$ because we can accommodate a tent whose base is contained in $W$.
-% \ednote{B: This last phrase doesn't make sense to me.
-M: The number of tents grow with $k$ and hence this is obvious; I explained in one of our meetings.
-Does it make sense?}
   This implies the existence of two points $w_1$ and $w_2$ in $W$ so that $d(T^kw_1,T^kw_2)= 1$, and by the triangle inequality  at least one of the inequalities $d(T^ku,T^kw_1)> \delta$ or $d(T^ku,T^kw_2)> \delta$ hold when $\delta\in (0,\frac{1}{2})$.
@@
 Since we have found a periodic point in the interval $W$,  the set of periodic points are dense in $T$.
-  This follows because any fixed point of $T^k$ is also a periodic point, and we have found this in an arbitrary interval.% \ednote{Expand further on reason for density.
-M: fixed point of $T^k$ is a periodic point, and we have found in an arbitrary interval.}
+  This follows because any fixed point of $T^k$ is also a periodic point, and we have found this in an arbitrary interval.
 
@@
 First we establish the method of Takens delay embedding. 
-In this method, one considers a discrete-time dynamical system defined on a 'nice' space (a smooth manifold)\ednote{Should this be further explained?} that can be obtained as time-$K$ map of a continuous time-dynamical system, concepts which will be formalised below.
+In this method, one considers a discrete-time dynamical system defined on a 'nice' space (a smooth manifold)\ednote{Should this manifold be further explained?} that can be obtained as time-$K$ map of a continuous time-dynamical system, concepts which will be formalised below.
 We make observations from such a system, i.e., we observe the evolution of $\theta(w_0)$ (where $w_0$ represents the initial state) i.e., we observe  a finite set of values of $u_n :=  \theta(w_n) = \theta(Tw_{n-1})$.
@@
 We recall the Takens delay embedding theorem next.
-
-%More formally we may say that in the process of measuring a system, we obtain a 1-dimensional time-series $\{\theta(u(t))\}$, where $\theta:\mathbb{R}^n\to\mathbb{R}$ is a smooth observation or measurement function.
-This series represents a possibly noise-augmented \ednote{M: Noise-augmented is not the right word} dataset containing information of certain parts of the original system.
 
@@
 \begin{Definition}
-  [\bf Nonautonomous Dynamical System (adopted from~\cite{Manju_ESP}))] \label{Dfn_NDS}\rm
+  [\bf Nonautonomous Dynamical System (adopted from~\cite{Manju_ESP})]\label{Dfn_NDS}\rm
   A nonautonomous dynamical system (NDS) on a space $X$ is simply a dynamical system comprising a family of maps $\{g_n\}_{n \in \mathbb{Z}}$ where $g_n(\ldotp):=g(u_n,\ldotp):X\to{X}$ is continuous.
@@
 \end{Example}
-% \ednote{Nicely written!}
 
@@
 
-The reachable set itself can defined independently from the fact that $g$ does or does not have the ESP/USP.
+The reachable set itself can defined independently from the fact that $g$ does or does not have the USP.
 We note that $x\in{X_n(\overline{u})}$ if and only $g$ has a solution $\{x_k\}$ for $x_n=x$ and input $\overline{u}$.
 \textbf{(Cite)}.
-
-$X_n$ is exactly the set of points $x$ through which some entire solution $\Psi$ passes at time $n$.
-
-This will lead to a result established later, but which is worth taking note of now: \textit{$g$ being a topological contraction is equivalent to the existence of a unique entire solution.}
-
-% \begin{Definition}
-%   [\bf Topological Contraction]\label{Dfn_TopContr}\rm
-%   A function $g:U\times{X}\to{X}$ is a topological contraction if for all $n\in\mathbb{Z}$ and all $\overline{u}\subset{U}$, $X_n(\overline{u})$ is a singleton subset of $X$.  
-% \end{Definition}
+The set $X_n(\overline{u})$ is exactly the set of all points $x$ that have some entire solution $\Psi$ passes through it at time $n$.
+\ednote{B: Changed sentence for better readability - still understandable?}
+
+We can give an alternate formulation of the set $X_n(\overline{u})$:
+suppose, for some fixed input 
+$\overline{u}$, we define $g_i = g(u_i,\cdot)$ for all $i\in \mathbb{Z}$, then the set of states
+\begin{equation} \label{eqn_association}
+X_{n,i}(\overline{u}) := g_{n-1} \circ \cdots g_{i+1} \circ g_i(X).
+\end{equation}
+
+As the image of a finite composition of continuous maps, the set $X_{n,i}(\overline{u})$ is compact whenever $X$ is compact and is also nonempty.
+Further, $X_{n,i}(\overline{u}) \supset X_{n,i-1}(\overline{u})$. Hence $\bigcap_{i<n}
+X_{n,i}(\overline{u})$ is a nested intersection of closed nonempty subsets, and
+whenever $X$ is compact, the intersection is nonempty.  \ednote{B: Should I cite here?}
+
+\begin{Definition}
+  [\bf Topological Contraction]\label{Dfn_TopContr}\rm
+  We say $g$ is a topological contraction if $\bigcap_{i<n}X_{n,i}(\overline{u})$ is a singleton subset of $X$ for each $\overline{u}$. 
+\end{Definition}
+
+It turns out that $\bigcap_{i<n}X_{n,i}(\overline{u})$ is identically equal to $X_n(\overline{u})$ that we had defined earlier \textbf{(Cite.)}.
+This will lead to a result (for.
+e.g., \cite{manjunath2013echo}) that  \textit{$g$ being a topological contraction is equivalent to the existence of a exactly one entire solution.}
 
@@
   [\bf Proof of Example~\ref{Ex_exp}] \label{rem_proofEx} \rm
-  To see this, we show that for every input $\overline{u}=\{u_n\}_{n\in\mathbb{Z}}$, $g(u_n,\cdot)$ is a contraction map on $X$ for every $k\in\mathbb{Z}$. 
-  Indeed $g(u_n,x_n)=x^n$ depends only on the state at time $n$ and so for each input $\overline{u}$, $X_n(\overline{u})=\{x^n\}$ is a singleton subset.
-  We then easily conclude that $\{x_n\}\equiv0$ is the only solution. 
-\end{Remark} \ednote{B: Is this sufficient?}
+  For the driven system,  
+  $g(u,x)=x^2$ the evolution does not depend on the input at all.
+Hence $g(u,x)$ can be written as a single map $f(x)=x^2$ which is a homeomorphism on $[0,1]$.
+A left-infinite orbit is defined and it converges to $1$ while the right-infinite orbit converges to $0$.
+Or in other words the iterates of $f^{-1}$ converges to $1$ while the iterates of $f$ converge to $0$
+\end{Remark}
 
 Thus far it has been demonstrated that a system may have one or more solutions; one may ask if a driven system always has a solution and, if so, whether it satisfies certain properties such as uniqueness.
-Should the driven system be compact, existence follows immediately as shown in the following result.
+Should the driven system be compact, existence follows immediately.
 
 \begin{Theorem}\label{Thm_CompactExistence}
- Let $g$ be a driven system.  If $X$ is compact then for each input $\overline{u}$, there exists at least one solution to the driven system $g(\ldotp, x)$
+ Let $g$ be a driven system.  If $X$ is compact, then for each input $\overline{u}$, there exists at least one solution to the driven system $g(\ldotp, x)$
 \end{Theorem}
@@
 
-% \ednote{M: You can omit the theorem, and just say it is a uniform attractor in the nonautonomous dynamical systems literature \cite{Manju_Nonlinearity}, and say the discussion on nonautonomous attractors is beyond the scope of the thesis.}
 % Paraphrased. 
-Additionally, $g$ having the USP is in our context also equivalent to $g$ being a topological contraction \cite{manjunath2021universal}, or alternatively also that $g$ exhibits the Echo State Property (ESP).
-
-In the majority of Reservoir Computing(RC) literature (a machine learning algorithm approach)\ednote{How should I explain/qualify RC better here? It's mentioned without really explaining.}, 
-the ESP is used to describe the concept of fading memory or memory-loss  \cite{jaeger2004harnessing, lu2018attractor}. 
-If $g$ possesses the ESP (equivalent in our context to the USP), the we are guaranteed that the whole of an input's left-infinite history will exactly determine the current system state; i.e.
-there is only ever the possibility of a single reachable state at a given instance of time \cite{jaeger2001echo}
-
-Defined formally, the ESP is intrinsically linked to an input (see \cite{Manju_ESP}) Unfortunately, definitions and results concerning the ESP do not have much practical benefit since a relationship between the network's response and the temporo-statistical properties of the input are not directly described.
-A more detailed presentation may be found in~\cite{jaeger2001echo}.
-
-We do however consider it important to mention that the ESP is also often discussed in terms of a stability property that holds for all inputs of a system~\cite{manjunath2020stability} and, furthermore, plays a key role in the design and training of recurrent neural networks(RNN) within the field of RC \cite{Manju_ESP}.
-
-\section{Next section?}
-\ednote{I feel a break is necessary here, but would hear from you what your thoughts are.}
-
-Having already defined the reachable set $X_U$ as the set of all elements of all entire solutions, we pause briefly in order to fix additional notation.
+%Additionally, $g$ having the USP is in our context also equivalent to $g$ being a topological contraction \cite{manjunath2021universal}, or alternatively also that $g$ exhibits the Echo State Property (ESP).
+
+\ednote{RC introduced a bit abruptly} Reservoir Computing(RC) refers to a machine learning approach in which input is mapped onto a different, but higher dimensional, space through a system called the reservoir after which a readout measuring the reservoir state is trained.
+In the majority of Reservoir Computing(RC) literature \textbf{(Cite)}, a notion of forgetting the states of the reservoir asymptotically is often used.  
+Concepts like the echo state property(ESP) and  fading memory \cite{boyd1985fading} are two of these.  
+If $g$ possesses the ESP (equivalent in our context to the USP \textbf{(Cite)}), we are guaranteed that the whole of an input's left-infinite history will exactly determine the current system state; i.e.
+there is only ever the possibility of a single reachable state at a given instance of time \cite{jaeger2001echo,Manju_2020}.
+% \cite{Manju_ESP}.
+Beyond forgetting the past states, the ESP is also often discussed in terms of a stability property that plays a key role in the robustness of recurrent neural networks(RNN).
+The concept, termed input-related stability~\cite{manjunath2020stability}, refers the phenomenon of nearby inputs creating responses (solutions) close to one another.
+
+\section{Establishing Conjugacies}
+
+Having already defined the reachable set $X_U$ as the collection of all elements of all entire solutions, we pause briefly in order to fix additional notation.
 Defining $\cev{u}^{n}:=(\ldots,u_{n-2} ,u_{n-1})$ as the left-infinite subsequence of an input up until time $n$, $\overleftarrow{U}$ is then the notation for all these left-infinite sequences in $U$.
@@
 If the causal mapping $H:\overleftarrow{U}{\to}{\overleftarrow{X}_U}$ is injective (in addition to being surjective), it becomes the embedding of $\overleftarrow{U}$ in $\overleftarrow{X}$ induced by the driven system $g$.
-\ednote{Explain why $H_{-1}$ is continuous} and we refer to $H$ as a \emph{causal embedding}
-% \ednote{M : Causal embedding a dynamical system is different from the definition of causal embedding, so please edit/delete this paragraph and the next theorem and also in the remaining part till Section 1.2}.
-% \begin{Theorem}
-%  If $g(\ldotp,x)$ is invertible and has the USP then $H$ is a causal embedding. 
-% \end{Theorem}
-% \begin{proof}
-%   See \cite{manjunath2021universal}.
-% \end{proof}
+The continuity of $H^{-1}$ follows from the fact that $H$ is itself a continuous and surjective mapping of a compact space $\overleftarrow{U}$ in a Hausdorff space.
+and we refer to the map $H$ as a \emph{causal embedding}.
+
 When $g$ has the USP, the diagram below (adopted from~\cite{Manju_Nonlinearity}) illustrates the operation of the mappings $h$ and $H$.
@@
 
-We wish to explore the possibility of establishing a conjugacy rather than a mere semi-conjugacy and therefore are interested in the restriction of inputs to the left-infinite orbits of a dynamical system.
-The left-infinite orbit spaces are significant because a great amount of information about the original system can be gleaned from their topological structure~\cite{Manju_IEEE}. 
-The formal topological structure for the so-called inverse-limit space of chaotic dynamical systems is complicated, and further discussion is relegated to~\cite{kennedy2008inverse_limit, ingram2011inverse}.
-Broadly, we may conceive of the inverse-limit space of a dynamical system as a self-map on a subset of an infinite-dimensional space where each point in the inverse-limit space corresponds to a backward orbit of the map $T$.
+The function $H$ contains a countably infinite number of coordinate (or factor maps) since its codomain is $\overleftarrow{X}_U$.  
+As it is not practically feasible to consider a left-infinite sequence in $X_U$ (in other words an element of  $\overleftarrow{X}_U$) for learning, we wish to restrict the inputs we consider to a subspace of $\overleftarrow{U}$.
+Particularly, we restrict our attention to the  so-called inverse-limit space of a dynamical system and hope to embed it within some finite product of $X$.
+
+%We wish to explore the possibility of establishing a conjugacy rather than a mere semi-conjugacy and therefore are interested in the restriction of inputs to the left-infinite orbits of a dynamical system.
+
+The left-infinite orbit space is significant because a great amount of information about the original system can be gleaned from its topological structure~\cite{ingram2011inverse,Manju_IEEE}. 
+The formal topological structure for the inverse-limit space of a chaotic dynamical system is complicated, and further discussion is relegated to~\cite{kennedy2008inverse_limit, ingram2011inverse}.
+Broadly, we may conceive of the inverse-limit system of a dynamical system as a subspace of an infinite-dimensional space where each point in the inverse-limit space corresponds to a backward orbit of the map $T$.
 Following~\cite{manjunath2021universal, ingram2011inverse} we provide a definition:
@@
   [\bf Inverse-Limit Space, Inverse-Limit System]\label{Dfn_InverseL}\rm
- The subset $\widehat{U}_T\subset \overleftarrow{U}$ defined by \[\widehat{U}_T: = \{ (\ldots,u_{-2},u_{-1}): Tu_{i} = u_{i+1}\]is called the inverse-limit space of $(U,T)$.
+ The subset $\widehat{U}_T\subset \overleftarrow{U}$ defined by \[\widehat{U}_T: = \{ (\ldots,u_{-2},u_{-1}): Tu_{i} = u_{i+1}\} \] is called the inverse-limit space of $(U,T)$.
  The self-map $\widehat{T}$  induced by $T$ on $\widehat{U}_T$  is defined by \[\widehat{T}: (\ldots,u_{-2},u_{-1}) \mapsto  (\ldots,u_{-2},u_{-1},T(u_{-1}))\] and the resulting dynamical system $(\widehat{U_T}, \widehat{T})$ is called the inverse-limit system of $(U,T)$.
@@
 
-The inverse-limit space is well-defined since T : U →U is surjective by assumption.
+The inverse-limit space is well-defined since $T : U \to{U}$ is surjective by assumption.We now get to the concept of embedding the inverse limit-space of the dynamical system in $X \times X$.
+
+\begin{Definition} \rm
+	We say a driven system $g$ \emph{causally embeds} a dynamical system $(U,T)$ if it satisfies the two properties: (i) a universal semi-conjugacy exists and  (ii)  $H_2(\cev{u}) := (h(r\cev{u}),h(\cev{u}))$ embeds the inverse-limit space $\widehat{U}_T$ in $X \times X$.
+\end{Definition}
+
+It is imperative to take note of a subtlety here: we refer to the \emph{map} $H$ as a (causal) embedding, whereas the driven \emph{system} $g$ (causally) embeds another \emph{system} $(U,T)$.
 
@@
 
-The driving function $g$ cannot be chosen in a frivolous manner as it could possibly complicate our work.
+So far it is not clear if the USP alone ensures the ability to causally embedding a dynamical system and we mention that a driving function $g$ cannot just be chosen in a frivolous manner.
+This could possibly complicate our work.
+\ednote{Dié lees vreemd}
+
+%The driving function $g$ cannot be chosen in a frivolous manner as it could possibly complicate our work.
 Specifically, one must be careful to avoid a choice of $g$ which would  add complexity to the obtained solution.
 
-When a causal embedding $H$ exists for the driven system $g$, one can map an arbitrary input ${u}$ onto the solution space $X$ without additional distortion or information-loss \textbf{(Cite.)}.
-Established such an embedding eliminates the question of possible additional complexity in the solution by guaranteeing that, since the systems are conjugate (semi-conjugate, \textbf{(refer)}), $g$ does not add any (some) complexity to the system.
-
-It is, however, a balancing act as it also undesirable to choose a function $g$ that quenches the temporal structure in $u$ by contracting to such a degree that the ability to recover information from the original system is lost completely.
-In the above example (\ref{ex_halfux}), the input's temporal variation cannot be related to the reachable set since $X_U$ consists of a single element; little, if not no, information is encoded.
+%When a causal embedding $H$ exists for the driven system $g$, one can map an arbitrary input ${u}$ onto the solution space $X$ without additional distortion or information-loss \textbf{(Cite.)}.
+
+%When an embedding is established, the question of possible additional complexity in the solution is removed by guaranteeing that, since the systems are conjugate (semi-conjugate, \textbf{(refer)}), $g$ does not add any (some) complexity to the system.
+
+%To achieve a causal embedding embedding not just a causal mapping, it is undesirable to choose a function $g$ that quenches the temporal structure in $u$ by contracting to such a degree that the ability to recover information from the original system is lost completely.
+In the above example(\ref{ex_halfux}), the input's temporal variation could not be related to the reachable set as $X_U$ only consisted of a single element; little, if not no, information is encoded.
 To obtain a suitably complex function $g$, it is thus desired that the reachable set of a driven system  be large enough to relate to the input. 
- We  reguire that the reachable set of a driven system to be such that the inverse-limit space of $U_T$ can be embedded in some finite self-product of the reachable set of $T$ \textbf{(refer or cite)}.
-To this end, consider the notion of State-Input (SI) Invertibility.
- \ednote{(M: Expand here. It stops a bit abruptly. B: Does the above reorganisation read better?)}
+This is true even for embedding  the inverse-limit space of a dynamical system. \textbf{(B: Expand or refer?)}
+
+ To this end, we recall the notion of State-Input (SI) Invertibility.
 
@@
   [\bf SI-Invertibility]\label{Dfn_SIinv}\rm
-  A map $g$ is said to be SI-Invertible if $g$ is invertible for all $x\in{X}$.
+  A driven system $g$ is said to be SI-Invertible if $g(*,x): U \to X$ is invertible for all $x\in X$.
 Alternatively it may be said that if, given $x_n$ and $x_{n-1}$, $u_{n-1}$ can be uniquely determined from $x_n=g(u_n,x_{n-1})$, then $g$ is said to be SI-invertible.
@@
 
-SI-invertibility promises that 'enough' information is retained when $g$ is chosen without introducing additional unwanted complexity. \ednote{(M: Expand why this is true. B: Is this satisfactory?)} 
+SI-invertibility means that one's choice of $g$ will still ensure 'enough' information is retained without introducing unwanted complexity. 
 'Enough' here refers to the fact that we may always recover the previous input value $u_n$ (if we know successive states $x_{n-1}, x_n$).
-'Without additional complexity' is guaranteed by the invertibility of the function $g$ that guarantees a one-to-one mapping between $u_n, x_{n-1}$ and $x_n$. 
+
+Subsequently we define the relation $Y_T$ induced by $(U,T)$ on $X_U\times{X_U}$ for a driven system $g$ possessing SI-invertibility.  
+To describe the  SDD formally, we consider a dynamical system $T: U \to U$ and define a relation on the reachable set $X_U$, i.e. a subset on $X_U \times X_U$  defined by 
+$$Y_T:=\{(x_{n-1},x_n): \{x_k\}_{k\in \mathbb{Z}} \mbox{ is a solution for some orbit of } T \mbox{ and } n \in \mathbb{Z}\}.$$
+
+The following theorem establishes the existence of a well-defined map $G_T$ describing the SDD of the system above.
+
+\begin{Theorem}
+If we let $G_T:Y_T\to{Y_T}$ be a map defined by the relation $(x_{n-1},x_n)\mapsto(x_n,x_{n-1})$, then $G_T$ is well-defined (and this results holds even in the absence of $g$ possessing over the USP)
+\end{Theorem}
+\begin{proof}
+  \textbf{See .....................}
+\end{proof}
+
+%'Without additional complexity' is guaranteed by the invertibility of the function $g$ that guarantees a one-to-one mapping between $u_n, x_{n-1}$ and $x_n$.
 In obtaining an SI-invertible function, therefore, we are guaranteed a system which will not lose so much information about its previous states in the forward-flow of time that one may not make any credible claims as to the original system's behaviour and properties.
@@
 
-It is worth at this stage taking note of a specific driven system in the form of a discrete state-space model that has acquired some adherence in applications \cite{Manju_IEEE} and especially those pertaining to Echo State Networks and ESP.
+At this stage it is worth taking note of a specific driven system in the form of a discrete state-space model which has acquired some adherence in applications \cite{Manju_IEEE}- especially those pertaining to Echo State Networks and the ESP.
 The function
-
 \begin{equation}  \label{eqn_driving}
@@
 \end{equation} 
-is both SI-invertible and, if $\alpha\beta<1$, also possesses the USP. 
-One may easily show that $g$ is SI-invertible by recovering $u_n$ by 
+is SI-invertible and, if $\alpha B$ has a spectral norm $<1$, also possesses the USP \cite[Th.2]{manjunath2013echo }. 
+It is easy to show that $g$ is SI-invertible by recovering $u_n$ in 
 \begin{equation} \label{eqn_SI_RNN}
@@
   \end{equation}
-  for $x_{n-1}, x_n$ known.
-
-The proof that the map $g$ possesses the USP is rather involved and hence, instead of being replicated here, may be found in \cite[Th.2]{manjunath2013echo }
-This specific driving function $g$ is used in our implementation and is discussed more completely in chapter \ref{sect5}
-
-Despite the ease that one may work with a left-infinite history in the realm of theory, it is impossible to obtain such a sequence in any real-life application.  
-One does not in practice, fortunately, need the entire left-infinite history of the input thanks to the Uniform Attraction Property (UAP).
+  when $x_{n-1}$ and $x_n$ are known.
+
+This specific driving function $g$ is used in our implementation and is discussed more completely in chapter~\ref{ch5}
+
+Despite the ease that with which one manipulates a left-infinite history in the realm of theory, it is impossible to obtain or use such a sequence in any real-life application.  
+Fortunately, one does not need the entire left-infinite history of an input in practice thanks to the Uniform Attraction Property(UAP). 
+We use an alternate version of the definition of UAP -- since the notion of nonautonomous attractors in this thesis would take some time to establish and detracts from the principal thrust of this project, we refer to \cite{Manju_Nonlinearity}.
 
@@
   [\bf Uniform Attraction Property]\label{Dfn_UAP}\rm
-  A driven system $g$ has the Uniform Attraction Property (UAP) if, regardless of starting position, all trajectories converge to a single trajectory as time flows forward. 
-  This trajectory is also the unique solution sequence $x$ to the input sequence $u$ as mentioned above.
-\end{Definition}
-
-This definition is stated in a not completely rigourous manner as the more formal definition makes use of processes, a concept which would take some time to establish and detracts from the principal thrust of this project.
-\ednote{B: Is it very obvious to the reader at this point what the general thrust of the paper is?}
-
-As the Unique Attraction Property guarantees that all trajectories will converge to the same trajectory as time moves forward. 
-Incredibly, this permits one to initialise a driven system $g$ with an altogether arbitrary initial value $y_m\in{X}$ where $m\in\mathbb{Z}$ and the UAP then guarantees that the sequence $\{y_{m+1}, y_{m+2},\ldots\}$ which satisfies the relation $y_n=g(u_n, y_{n-1})$  for  $k\geq{m}$  will uniformly approach the elements $\{x_n\}$ of the actual solution (See \cite[Th.11,]{Manju_Nonlinearity} or \cite[Eqn.18]{Supp}).
+  A driven system $g$ has the uniform attraction property (UAP) if we initialize the driven system
+with an arbitrary initial value $y_m \in X$, and the sequence $y_{m+1}, y_{m+2}, y_{m+3},...$ satisfying $y_{k+1}= g(u_k,y_k)$ for $k \geq m$ then approximates an actual solution $\{x_n\}$ uniformly in the sense that given $\epsilon>0$ (independent of $y_m$) there is an integer $n$ so that $d(x_{n+i}, y_{n+i})<\epsilon$ for all $i\ge 0$, where $\{x_m\}$ is a solution.
+\end{Definition}
+
+The UAP guarantees that all trajectories will converge to the same trajectory as time moves forward.
+An illustration of this is shown in Fig.~\ref{fig:memloss_conttime} where the trajectory in red locks on to the supposed actual solution when the system has the USP.
 
 \begin{figure}[ht]
-  \includegraphics[scale=0.4]{_memloss_conttime.eps}
-  \centering
-\captionof{figure}{Reproduced from \cite{manunath2021universal}. Illustration of all trajectories approaching the true solution if $g$ has the UAP}
+  \includegraphics[scale=0.35]{_Nonlinearity_Fig_response.eps}
+  \centering
+\captionof{figure}{A coordinate of a simulated solution $(x_0,x_1,\ldots,x_{5000})$ of a RNN  plotted in red (with parameter $\alpha=0.99$ (where $g$ has the USP) and blue ($a=1$ and $\alpha=1.05$ where $g$ does not have the USP) while the matrices $A$ and $B$ are randomly generated, and $B$ has unit spectral radius.
+Reproduced from \cite{Manju_Nonlinearity}.}
 \label{fig:memloss_conttime} 
-\end{figure}
-
-One may now appreciate an even more astounding result: $g$ having the USP is equivalent to the UAP. We restate the above theorem.
-
-\begin{Theorem}\label{thm_usp-contr-uap}
-  The following statements are equivalent:
-  \vspace{-8mm}
-  \begin{enumerate}[noitemsep, label=\roman*.]
-    \item $g$ has the USP.
+\end{figure} \ednote{This graph with the black line does not immediately make sense}
+
+One may now appreciate an even more astounding result: $g$ having the USP is equivalent to the UAP as is proved in~\cite{Manju_Nonlinearity}.
+
+%We restate the above theorem.
+
+%\begin{Theorem}\label{thm_usp-contr-uap}  The following statements are equivalent:  \vspace{-8mm}  \begin{enumerate}[noitemsep, label=\roman*.]    \item $g$ has the USP.
 \item $g$ is a topological contraction.
 \item $g$ has the UAP.
-  \end{enumerate}
-\end{Theorem}
-\begin{proof}
-  May be found in \cite[Th.1]{Manju_Nonlinearity}
-\end{proof}
-
-One need therefore not establish any additional results to ensure that the sequence uniformly approaches the unique solution $\Psi$.
+\end{enumerate\end{Theorem}\begin{proof}  May be found in \cite[Th.1]{Manju_Nonlinearity}\end{proof}
+
+% One need therefore not establish any additional results to ensure that the sequence uniformly approaches the unique solution $\Psi$.
 This vastly simplifies the effort necessary to set up a problem in order to guarantee that the underlying system will be accurately represented by the conjugate system.
 
-This also solves the problem of perturbation/noise introduced by the observable or measurement function.
+%This also solves the problem of perturbation/noise introduced by the observable or measurement function.
 Recall from before that an observable is inherently a map that discretises the underlying continuous-time system.
@@
 Since we are working with a chaotic system displaying sensitive dependence on initial conditions, these small errors could potentially send the trajectory into a completely different attractor.
-In the presence of the UAP, however, small measurement errors introduced into the system do not pose the same danger as before since the trajectories will eventually tend towards the actual solution.
-\ednote{(B: Is this Complete enough?)}
+%In the presence of the UAP, however, small measurement errors introduced into the system do not pose the same danger as before since the trajectories will eventually tend towards the actual solution.
+% \ednote{(B: Is this Complete enough?)}
 % \textbf{From the perspective of perturbation of an autonomous dynamical system, if the resulting nonautonomous system has the USP, then the temporal structure of noise relates to the dynamics on a nonautonomous uniform attractor through a topological semi-conjugacy or a conjugacy.}
@@
 
-Next we define the relation $Y_T$ induced by $(U,T)$ on $X_U\times{X_U}$ for a driven system $g$ possessing SI-invertibility.
-To describe the  single-delay lag dynamics formally, we consider a dynamical system $T: U \to U$ and we define a relation on the reachable set $X_U$, i.e., a subset defined on $X_U \times X_U$ by
-$$Y_T:=\{(x_{n-1},x_n): \{x_k\}_{k\in \mathbb{Z}} \mbox{ is a solution for some orbit of } T \mbox{ and } n \in \mathbb{Z}\}.$$
-
-The following theorem establishes the existence of a well-defined map $G_T$ describing the single-delay dynamics(SDD) of the system above.
-
-\begin{Theorem}
-If we let $G_T:Y_T\to{Y_T}$ be a map defined by the relation $(x_{n-1},x_n)\mapsto(x_n,x_{n-1})$, then $G_T$ is well-defined (and this results holds even in the absence of $g$ possessing over the USP)
-\end{Theorem}
-
 % Recall that $g$ is only being given inputs from the orbits of $T$.
 
-It is of interest to ascertain whether $\widehat{T}$ and $G_T$ are related\ednote{B: Should I expand more here?}, and we consequently introduce  the function $H_2{\overline{}} := (h(r\overline{u}, h(\overline{u})))$ mapping the entirety of a left-infinite sequence to some element in $X\times{X}$.
+%Incredibly, this permits one to initialise a driven system $g$ with an altogether arbitrary initial value $y_m\in{X}$ where $m\in\mathbb{Z}$ and the UAP then guarantees that the sequence $\{y_{m+1}, y_{m+2},\ldots\}$ which satisfies the relation $y_n=g(u_n, y_{n-1})$  for  $k\geq{m}$  will approach the elements $\{x_n\}$ of the actual solution.
+This vastly simplifies the effort necessary to set up a problem in order to guarantee that the underlying system will be accurately represented by the conjugate system.
+
+We are interested in determining whether $\widehat{T}$ and $G_T$ are related.
+For this, we introduce  the function $H_2{\overline{}} := (h(r\overline{u}, h(\overline{u})))$ mapping the entirety of a left-infinite sequence to some element in $X\times{X}$.
 Indeed, it may be shown that $g$ having SI-invertibility and the USP immediately guarantees the existence of at least a semi-conjugacy between the systems $(Y_T, G_T)$ and $(\widehat{U}_T, \widehat{T})$.
-This is formalised in Causal Embedding Theorem, which we state next:
+This is formalised in the Causal Embedding Theorem, which we state next:
 
@@
 
-We are drawing nearer and nearer to our principal target and our results carry more and more weight. The next result is remarkable.
-
-\begin{Theorem}
-  Graph \ref{Scomm_h} is exactly the inverse-limit system $(\hat{U}, \hat{T})$.    
-\end{Theorem}
-
-Recall that $H_2$ maps an entire left-infinite solution sequence from $\Psi$ to an element in $X\times{X}$.
-We now have the following (which may be compared with \ref{SCausal_H} above):
-
-\begin{equation} \label{fig:inverse_limsystem}
-  %  \[ 
-      \psset{arrows=->, arrowinset=0.25, linewidth=0.6pt, nodesep=3pt, labelsep=2pt, rowsep=0.7cm, colsep = 1.1cm, shortput =tablr}
-      \everypsbox{\scriptstyle}
-      \begin{psmatrix}
-      \widehat{U}_T  && \widehat{U}_T \\
-      Y_T && Y_T \\
+Recall that $H_2$ maps an entire left-infinite solution sequence from $\Psi$ \ednote{Check}to an element in $X\times{X}$. 
+\ednote{Rephrase}We are drawing nearer and nearer to our principal target and our results carry more and more weight.
+Since if learn $G_T$ we would have learnt $\widehat{T}$ and since $\widehat{T}$ is an extension of $T$, we would have learnt $T$ as well.
+
+\ednote{Why is theorem on inverse-limit system taken out?}
+%\begin{Theorem} Graph \ref{Scomm_h} is exactly the inverse-limit system $(\hat{U}, \hat{T})$.    \end{Theorem}
+
+%We now have the following (which may be compared with \ref{SCausal_H} above):
+
+%\begin{equation} \label{fig:inverse_limsystem}
+  %  \[       \psset{arrows=->, arrowinset=0.25, linewidth=0.6pt, nodesep=3pt, labelsep=2pt, rowsep=0.7cm, colsep = 1.1cm, shortput =tablr}      \everypsbox{\scriptstyle}
+%      \begin{psmatrix}      \widehat{U}_T  && \widehat{U}_T \\     Y_T && Y_T \\
       %%%
@@
      %  \ncline{2,1}{2,2}^{G_T}
-      \end{psmatrix}
+%      \end{psmatrix}
     %  \]
-  \end{equation}
-
-\begin{Theorem}
-    $(Y_T, G_T)$ is semi-conjugate to $(\widehat{U}, \widehat{T})$.
-\end{Theorem}
-\begin{proof}
-  See \cite[Th.3, Th.4]{manunath2021universal}
-\end{proof}
+%  \end{equation}
+
+%\begin{Theorem}
+ %   $(Y_T, G_T)$ is semi-conjugate to $(\widehat{U}, \widehat{T})$.
+%\end{Theorem}
+%\begin{proof}
+%  See \cite[Th.3, Th.4]{manunath2021universal}
+%\end{proof}
 
@@
 \item To determine properties about this system $(U,T)$ and predict its future evolution, we determine the dynamics of the inverse-limit system $(\widehat{U}, \widehat{T})$.
-Given certain assumptions, we can guarantee that $(\widehat{U}, \widehat{T})$ is at least semi-conjugate to $(U,T)$.
-\item If the driven system $g$ is SI-invertible (and $\{u_n\}\in{U}$  is an orbit of $T$), the map $G_T$ exists. 
+% Given certain assumptions, we can guarantee that $(\widehat{U}, \widehat{T})$ is at least semi-conjugate to $(U,T)$.
+\item If the driven system $g$ is SI-invertible (and $\{u_n\}\in{U}$  is an orbit of $T$), the map $G_T$ exists that describes the single delay dynamics. 
 \item If, furthermore, $g$ has the USP, $(Y_T, G_T)$ is semi-conjugate to $(\widehat{U}, \widehat{T})$.
@@
 
-One can therefore learn the SDD of the driven system states via $G_T$ with enough data thanks to the USP/UAP. This enables us to do at least two things: 
+One can therefore in practice learn the SDD of the driven system states via $G_T$ with enough data thanks to the USP/UAP. This enables us to do at least two things: 
 \vspace{-8mm}
 \begin{itemize}
-\item Forecast future values of $x_n$ via iterates of $G_T$ (as $G_T$ can be determined.
-\item Forecast future values of $u_n$. 
+\item Forecast  $x_{n+1},x_{n+2}, \ldots$ via iterates of $G_T$ (if $G_T$ can be learnt).
+\item Forecast future values of $u_n$ since $x_n$ and $x_{n+1}$ determine $u_n$ since $g$ is SI-invertible. 
 \end{itemize}
+
+Finally, we note that although $G_T$ exists with SI-invertibility, we need USP  otherwise the driven states has to be running for all time since the driven states would not have forgotten its past states.
 
@@
 \end{figure}
+\ednote{M: Please put a caption on this figure and cite the figure in the main text}
 
@@
 
-% \begin{Definition}
-%   [\bf Projection Mapping]\label{Dfn_ProjMap}\rm
-%  The projection mapping $\pi_i$ is defined as follows: $$\pi_1:(a_1,a_2)\to{a_1}$$ and $$\pi_2:(a_1,a_2)\to{a_2}$$ 
-% \end{Definition}
-
-The graph in equation \ref{fig:inverse_limsystem} is then extended as below:
-
-% \begin{figure}[ht]
-%   \includegraphics[scale=0.3]{_gammatemp.eps}
-%   \centering
-%   \label{fig:temp_Gamma}
-% \end{figure}
+The graph in equation \ref{fig:comm_H_CET} is then extended as below:
 
@@
 
-In the first place, Learning $\Gamma$ saves computational resources.
-This is due to the fact that the input $u_n$ lies in a lower-dimensional subspace of the high-dimensional space $X\times{X}$ and we can learn $\Gamma$ on a small subset of the space.
-In practice, if the input is of a  lower dimension, one may easily embed it into the space $X\times{X}$ by padding the vector $u_n$ with zeroes.
+In the first place, learning $\Gamma$ saves computational resources.
+This is due to the fact that the input $u_n$ lies in a lower-dimensional subspace in comparison to the high-dimensional space $X\times{X}$.
+
+%In practice, if the input is of a  lower dimension, one may easily embed it into the space $X\times{X}$ by padding the vector $u_n$ with zeroes.
 
 Secondly, the function $\Gamma$ is known to be stable.
-Learning Gamma makes use of equations \ref{eqns_from_data} which in turn employs $g$, a function possessing the USP, and offers distinct advantages with regards to stability in the presence of perturbation.
+Learning $\Gamma$ makes use of equations \ref{eqns_from_data} which in turn employs the driven system $g$ possessing the USP, and offers distinct advantages with regards to stability in the presence of perturbation.
 It is desirable that inputs that differ only slightly would have only ‘slight’ effects on the output of the system.
@@
 As the data fed into the system is a function of both the input given and the parameters of the system itself (i.e.
-the value a, $\alpha$ in $\tanh$), we investigate both input-related and parameter-related stability.
+the value a, $\alpha$ in $\tanh$), we exploit both input-related and parameter-related stability.
 Both these notions are defined by way of the continuity of an encoding map of an input and design parameter respectively in \cite{manjunath2020stability}.\
@@
 Informally, we may consider input-related stability as concerned with the question of whether or not small variations in input would result in small responses.   
-If two inputs $u$ and $u_{noisy}$ are close in the product topology, then their tails could differ greatly due to the metric's insensitivity to the differences in sequence's tails.
- However if the system has the USP then the function $h$ is continuous and so $h(u), h(u_{noisy})$ remain near – maintaining a measure of robustness to input and measurement noise.\cite{manjunath2021universal}.
+If two inputs $\bar{u}$ and $\bar{u}_{noisy}$ are close in the product topology, then their tails could differ greatly due to the metric that generates the product topology has insensitivity to the differences in sequence's tails.
+ However if the system has the USP then the function $h$ is continuous and so $h(\bar{u}), h(\bar{u}_{noisy})$ remain close-by for small levels of noise maintaining a measure of robustness to input and measurement noise.\cite{manjunath2021universal}.
 In a similar fashion, parameter-related stability is obtained thanks to a result \cite[Lemma 3.2]{manjunath2020stability} relating this form of stability with the ESP (which is here equivalent to the USP).
@@
 \begin{figure}[ht]
-  \includegraphics[scale=0.5]{_AutovsDriven.eps}
+  \includegraphics[scale=0.25]{_autovsdriven.eps}
   \centering
@@
   \captionof{figure}{Illustrating empirically the importance of global dissipativity.
-Plotted below the three principal components of randomly initiated trajectories of a learnt version of $G_T$ versus the trajectory $(x_{n-1}, x_n)$(grey) of driven stated obtained from the Thomas Attractor's data (see below).
+Plotted below are the three principal components of randomly initiated points(blue) of a learnt version of $G_T$ versus the trajectory $(x_{n-1}, x_n)$(red) of driven states obtained from the Fractal Dream Attractor's data~\ref{Thomas_Attractor}).
 }
 \end{figure}
-
-\ednote{B: Would you please help me with the writing here? Otherwise I shall leave this out.}
-\emph{Thirdly, $\Gamma$ is globally dissipative whereas $G_T$ is not guaranteed to be~\cite{manjunath2021universal}). 
-Global dissipativity prevents large numerical errors due to input data.If a system is not globally dissipative, neglibly small errors result in major errors which thus induce predictions to fail utterly.
+\ednote{nice figure: Please rescale the figure and also you do not have to state that code was borrowed.
+So I have omitted that in the caption.
+Please put a caption to the figure and then cite in the text.
+}
+
+Thirdly, whenever we learn a chaotic map on a set (here $Y_T$), as we had remarked in the disadvantages of Takens delay embedding a little noise would mean that the learnt system would accept arguments out of the set, so we will have a map $G_T^+$ acting on $Y_T^+$, and since $Y_T$ may not be an attractor of $G_T$ this could lead to errors.
+To circumvent this if the driven system with a saturation function is used, then we can ensure that there exists an attractor $A$ of the map that implements $G_T$ indirectly via $g$ containing $Y_T^+$.
+This prevents from large errors occuring since the dynamics is confined to $A$.
+We refer to \cite{Supp} for details where this is referred to as the global dissipativity of $g$.
+
+In essence, global dissipativity prevents large numerical errors due to the noise in input data.
+If a system is not globally dissipative, neglibly small errors result could lead to major errors which thus induce predictions to fail utterly.
 If one learns $G_T$ directly, we are in danger of replicating the problems arising for Takens.
-}
-
-% The $\omega$-limit set $\omega(u;T)$ of a point $u$ is defined to be the collection of limit points of the sequence $\{x,Tu,T^2u,\ldots\}$.
-This set is nonempty, and $\omega(u;T)$ is invariant.
-\textbf{(Change)}
-% \begin{Definition}
-%   [\bf Global Dissipativity]\label{Dfn_GlobalDiss}\rm
-% 	We say a dynamical system $(Z,f)$ is globally dissipative if there exists a nonempty proper closed subset $B$ of $Z$  so that for all $x\in Z\setminus B$, (i).
-$\omega(x;f) \subset B$  and (ii).
-$B$ is positively invariant that is $f(B) \subset B$.
-We call any such closed subset $B$ a trapping set of $f$.
-% \end{Definition}
 

### 2022-01-16 Gandhi_Project_Report/draft5.tex
Many graphs added and comments added. Edits of Ch.5. Starting to prep to add Grammarly back in

@@
 
-\section{Practical Limitations to Takens' Theorem}
+\section{Practical Limitations to Takens' Theorem}\label{sect_takenslimits}
 
@@
 
-It turns out that $\bigcap_{i<n}X_{n,i}(\overline{u})$ is identically equal to $X_n(\overline{u})$ that we had defined earlier \textbf{(Cite.)}.
+It turns out that $\bigcap_{i<n}X_{n,i}(\overline{u})$ is identically equal to $X_n(\overline{u})$ that we had defined earlier \ednote{(Insert citations.)}.
 This will lead to a result (for.
@@
 \end{Theorem}
-\begin{proof}
-  May be found in \cite{kloeden2011nonautonomous, manjunath2014dynamics, manjunath2013echo}
-\end{proof}
+{\bf Proof.}  May be found in \cite{kloeden2011nonautonomous, manjunath2014dynamics, manjunath2013echo}
 
@@
 
-\ednote{RC introduced a bit abruptly} Reservoir Computing(RC) refers to a machine learning approach in which input is mapped onto a different, but higher dimensional, space through a system called the reservoir after which a readout measuring the reservoir state is trained.
+\ednote{B: RC introduced a bit abruptly.
+Adjust.} Reservoir Computing(RC) refers to a machine learning approach in which input is mapped onto a different, but higher dimensional, space through a system called the reservoir after which a readout measuring the reservoir state is trained.
 In the majority of Reservoir Computing(RC) literature \textbf{(Cite)}, a notion of forgetting the states of the reservoir asymptotically is often used.  
@@
 \end{Theorem}
-\begin{proof}
-  See~\cite[Th.3]{manjunath2013echo}.
-\end{proof}
+{\bf Proof.}  See~\cite[Th.3]{manjunath2013echo}.
 
@@
 This could possibly complicate our work.
-\ednote{Dié lees vreemd}
+%\ednote{B: Grammarly}
 
@@
 To obtain a suitably complex function $g$, it is thus desired that the reachable set of a driven system  be large enough to relate to the input. 
-This is true even for embedding  the inverse-limit space of a dynamical system. \textbf{(B: Expand or refer?)}
+This is true even for embedding  the inverse-limit space of a dynamical system. \ednote{B: Expand or refer?}
 
@@
 
-\begin{Theorem}
-If we let $G_T:Y_T\to{Y_T}$ be a map defined by the relation $(x_{n-1},x_n)\mapsto(x_n,x_{n-1})$, then $G_T$ is well-defined (and this results holds even in the absence of $g$ possessing over the USP)
+\begin{Theorem}\label{Thm_GT_Exists}
+  For an SI-invertible driven system $g$ and a dynamical system $(U,T)$, the map $G_T: Y_T \to Y_T$ defined by the relation $(x_{n-1},x_n) \mapsto (x_n,x_{n+1})$ is well-defined. 
+  (This results holds even in the absence of $g$ possessing over the USP) 
+  % Moreover, the mapping $(x_{n-1},x_n) \mapsto u_{n}$ is well-defined when $x_{n-1}$ and $x_n$ are successive points on a solution obtained for an input  $\{u_n\}$ that is an orbit of $T$.
   \end{Theorem}
-\begin{proof}
-  \textbf{See .....................}
-\end{proof}
+  {\bf Proof.} See~\cite[Th.3]{Supp}
 
@@
 \captionof{figure}{A coordinate of a simulated solution $(x_0,x_1,\ldots,x_{5000})$ of a RNN  plotted in red (with parameter $\alpha=0.99$ (where $g$ has the USP) and blue ($a=1$ and $\alpha=1.05$ where $g$ does not have the USP) while the matrices $A$ and $B$ are randomly generated, and $B$ has unit spectral radius.
-Reproduced from \cite{Manju_Nonlinearity}.}
+Reproduced from \cite{Manju_Nonlinearity}. }
 \label{fig:memloss_conttime} 
-\end{figure} \ednote{This graph with the black line does not immediately make sense}
+\end{figure} 
+\ednote{B: How can I explain the black line? I don't think this makes sense to the reader immediately.}
 
@@
 \item $g$ has the UAP.
-\end{enumerate\end{Theorem}\begin{proof}  May be found in \cite[Th.1]{Manju_Nonlinearity}\end{proof}
+\end{enumerate\end{Theorem} {\bf Proof.}  May be found in \cite[Th.1]{Manju_Nonlinearity}
 
@@
 \end{Theorem}
-\begin{proof}
-  May be found as the proof of~\cite[Th.4]{Supp}
-\end{proof}
-
-Recall that $H_2$ maps an entire left-infinite solution sequence from $\Psi$ \ednote{Check}to an element in $X\times{X}$. 
-\ednote{Rephrase}We are drawing nearer and nearer to our principal target and our results carry more and more weight.
-Since if learn $G_T$ we would have learnt $\widehat{T}$ and since $\widehat{T}$ is an extension of $T$, we would have learnt $T$ as well.
-
-\ednote{Why is theorem on inverse-limit system taken out?}
+{\bf Proof.}  May be found as the proof of~\cite[Th.4]{Supp}
+
+Recall that $H_2$ maps an entire left-infinite solution sequence from $\Psi$ to an element in $X\times{X}$. 
+We are drawing nearer and nearer to our principal target and our results carry more and more weight: If we are able to learn $G_T$, we will also have learnt $\widehat{T}$; and since $\widehat{T}$ is an extension of $T$, we will have obtained $T$ as well.
+%\ednote{B: Grammarly check}
+
 %\begin{Theorem} Graph \ref{Scomm_h} is exactly the inverse-limit system $(\hat{U}, \hat{T})$.    \end{Theorem}
@@
 %\end{Theorem}
-%\begin{proof}
-%  See \cite[Th.3, Th.4]{manunath2021universal}
-%\end{proof}
+% {\bf Proof.} See \cite[Th.3, Th.4]{manunath2021universal}
 
@@
 
-Finally, we note that although $G_T$ exists with SI-invertibility, we need USP  otherwise the driven states has to be running for all time since the driven states would not have forgotten its past states.
-
-% Below is reproduced a pictorial representation of the relationship between different entities and properties. \textbf{(Cite SIG_25May)}
-% Below is reproduced a pictorial representation of the relationship between the different entities and their properties.
+Finally, we note that although $G_T$ exists with SI-invertibility, we need the USP as well(see \ref{fig_pictorialSummary} below for a visual illustration).
+If not, the driven states would have to be running for all time since they would not have forgotten their past states.
 
@@
   \centering
+  \captionof{figure}{Pictorial summary of results obtained up to now and how they overlap. Reproduced from \textbf{(Cite SIG-25May)} }
   \label{fig:fig_pictorialSummary}
 \end{figure}
-\ednote{M: Please put a caption on this figure and cite the figure in the main text}
 
@@
 This is done by implementing a Recurrent Neural Network (RNN), which is discussed in~\ref{ch5}.
-\ednote{(B: Should I add in that we'll be using Pearson coefficient?)}
+\ednote{B: Should I add in that we'll be using Pearson coefficient?}
 
@@
 \end{Theorem}
-\begin{proof}
-  See \cite[Th. 3c]{manjunath2021universal}.
-\end{proof}
+{\bf Proof.}  See \cite[Th. 3c]{manjunath2021universal}.
 
@@
 the value a, $\alpha$ in $\tanh$), we exploit both input-related and parameter-related stability.
-Both these notions are defined by way of the continuity of an encoding map of an input and design parameter respectively in \cite{manjunath2020stability}.\
+Both these notions are defined by way of the continuity of an encoding map of an input and design parameter respectively in \cite{manjunath2020stability}.\ednote{B: Is this satisfactory?}
 
@@
 
+Thirdly, learning a chaotic map on a set(here $Y_T$) with even a small measure of noise  means that the learnt system accepts arguments out of the set (Problems similar to Takens, cf.~\ref{sect_takenslimits}).
+We thus have a map $G_T^+$ acting on $Y_T^+$ and since $Y_T$ is not guaranteed to be an attractor of $G_T$ this could lead to errors.
+If one learns $G_T$ directly, we are in danger of replicating the problems arising for Takens.  To see this, consider Fig.~\ref{fig:YtGtFailure}.
+
+\begin{figure}[ht]
+  \includegraphics[scale=0.25]{_YTerrors.eps}
+  \centering
+  \captionof{figure}{Diagram of potential errors developing when $G_T$ is learnt.}
+  \label{fig:YtGtFailure}
+\end{figure}
+
+To circumvent this: if a driven system with a saturation function\ednote{B: What is a saturation function?} is used, we can indirectly (via $g$) ensure the existence of an attractor $A$ of the map which implements $G_T$, and containing $Y_T^+$.
+\ednote{B: I'm not sure this sentence makes complete sense.}
+By defining the dynamics to $A$, this prevents large errors from occuring.  
+More details on the global dissipativity of $g$, as it is termed there, may be found in~\cite{Supp}.
+
 \begin{figure}[ht]
@@
   \centering
-  \label{fig:learningFailure}
   \captionof{figure}{Illustrating empirically the importance of global dissipativity.
@@
 }
+  \label{fig:learningFailure}
 \end{figure}
-\ednote{nice figure: Please rescale the figure and also you do not have to state that code was borrowed.
-So I have omitted that in the caption.
-Please put a caption to the figure and then cite in the text.
-}
-
-Thirdly, whenever we learn a chaotic map on a set (here $Y_T$), as we had remarked in the disadvantages of Takens delay embedding a little noise would mean that the learnt system would accept arguments out of the set, so we will have a map $G_T^+$ acting on $Y_T^+$, and since $Y_T$ may not be an attractor of $G_T$ this could lead to errors.
-To circumvent this if the driven system with a saturation function is used, then we can ensure that there exists an attractor $A$ of the map that implements $G_T$ indirectly via $g$ containing $Y_T^+$.
-This prevents from large errors occuring since the dynamics is confined to $A$.
-We refer to \cite{Supp} for details where this is referred to as the global dissipativity of $g$.
-
-In essence, global dissipativity prevents large numerical errors due to the noise in input data.
-If a system is not globally dissipative, neglibly small errors result could lead to major errors which thus induce predictions to fail utterly.
-If one learns $G_T$ directly, we are in danger of replicating the problems arising for Takens.
+
+In effect, global dissipativity prevents large numerical errors due to the noise in input data from arising (see Fig.~\ref{fig:learningFailure}). 
+If a system is not globally dissipative, neglibly small errors could lead to major inaccuracy which thus induce predictions to utterly fail.
 

### 2022-01-16 Gandhi_Project_Report/draft6.tex
Many graphs added and comments added. Edits of Ch.5. Starting to prep to add Grammarly back in

--- 2022-01-16 Gandhi_Project_Report/draft5.tex
@@
   The authors of~\cite{manjunath2021universal} did not, however, consider the relationship between the inverse-limit space $\widehat{U}_T$ and the input $u_n=\theta(w_n)$ (where $\theta$ is our observable and $w_n$ the complete input), a concept we now explore.
-  We may construct a delay-coordinate $v_n=[\theta(w_n), \theta(w_{n-1}), \ldots, \theta(w_{n-d})]^{T} = [u_n, u_{n-1}, \ldots, u_{n-d}]^T$
-    to accordingly obtain a new dynamical system $(V,F)$ where $F$ is the map $F:v_n\mapsto{v_{n+1}}$.
-If $d$ is sufficiently large, we may apply Takens Embedding Theorem~\ref{Thm_Takens} to establish the existence of a topological conjugacy between the new system $(V,F)$ and the original system $(U,T)$.
-
-  Using the same notation as before, we may construct a new conjugacy which now relates delay coordinates to one another.
+
+Let $(V,T)$ be a dynamical system with dynamics generated by the update equation $v_{n+1}=Tv_n$
+Consider the case where the only available information is observations $\theta$ from an orbit of the map $T$ where $\theta:V\to\mathbb{R}$ is an observable. 
+Our task is still to find some system topologically conjugate to $(V,T)$ so that we may use the datapoints $\theta(v_{0}),\theta(v_{1}),\ldots,\theta(v_{m})$ to predict the subsequent observations $\theta(v_{m+1}),\theta(v_{m+2}),\ldots$.
+
+Applying the Takens embedding theorem, we may conclude that when an input from the delay-coordinate mapping $\Phi_{\theta,2d}(\theta(v_{n})) := (\theta(v_{n-2d}),\ldots,\theta(v_{n-1}),\theta(v_{n}))$ is fed to the driven system $g$ as the input $u_n$,
+there exists a homeomorphism $F_\theta: \Phi_{\theta,2d}(\theta(v_{n})) \mapsto \Phi_{\theta,2d}(\theta(v_{n+1}))$.
+
+This may be stated alternatively as follows: If the inpute $u_n$ is of the form $u_n := \Phi_{\theta,2d}(\theta(v_{n}))$ and is fed to a driven system $g$ possessing both the USP and SI-invertibility,
+it follows from the Causal Embedding Theorem~\ref{Thm_CET} that the newly generated space $(Y_F,G_F)$ topologically conjugate to the inverse-limit space of $(\Phi_{\theta,2d}(V), F_\theta)$.
+Moreover, we can forecast the values $u_n,u_{n+1},\ldots$ and $\theta(v_n), \theta(v_{n+1}),\ldots$ as well as having the guarantee that this embedding is stable when considered in terms of global dissipativity as discussed before.
 
@@
     \begin{psmatrix}
-        \widehat{V}_F  && \widehat{V}_F \\
+    \widehat{\Phi}_{2d,\theta}  && \widehat{\Phi}_{2d,\theta} \\
     Y_F && Y_F \\
     %%%
-       %  \ncline{1,1}{1,2}^{\widehat{F}} \ncline{1,1}{2,1} <{h}
+    %  \ncline{1,1}{1,2}^{\widehat{F}_\theta} \ncline{1,1}{2,1} <{h}
     %  \ncline{1,2}{2,2} > {*}
@@
 \end{equation}
+
+%   We may construct a delay-coordinate $v_n=[\theta(w_n), \theta(w_{n-1}), \ldots, \theta(w_{n-d})]^{T} = [u_n, u_{n-1}, \ldots, u_{n-d}]^T$
+%     to accordingly obtain a new dynamical system $(V,F)$ where $F$ is the map $F:v_n\mapsto{v_{n+1}}$.
+If $d$ is sufficiently large, we may apply Takens Embedding Theorem~\ref{Thm_Takens} to establish the existence of a topological conjugacy between the new system $(V,F)$ and the original system $(U,T)$.
+
+%   Using the same notation as before, we may construct a new conjugacy which now relates delay coordinates to one another.
+
+%   \begin{equation}\label{eqn_VFdelay}
+%     %  \[ 
+%         \psset{arrows=->, arrowinset=0.25, linewidth=0.6pt, nodesep=3pt, labelsep=2pt, rowsep=0.7cm, colsep = 1.1cm, shortput =tablr}
+%         \everypsbox{\scriptstyle}
+%         \begin{psmatrix}
+%         \widehat{V}_F  && \widehat{V}_F \\
+%         Y_F && Y_F \\
+%         %%%
+%        %  \ncline{1,1}{1,2}^{\widehat{F}} \ncline{1,1}{2,1} <{h}
+%        %  \ncline{1,2}{2,2} > {*}
+%        %  \ncline{2,1}{2,2}^{G_F}
+%         \end{psmatrix}
+%       %  \]
+%     \end{equation}
 
@@
 
+\bibliographystyle{pnas-new}
+\bibliography{pnas}
+\printbibliography
 \end{document}

### 2022-01-16 draft_16Jan/draft5.tex
Many graphs added and comments added. Edits of Ch.5. Starting to prep to add Grammarly back in

--- 2022-01-16 Gandhi_Project_Report/draft5.tex
@@
 The machine learning algorithms for modelling the double pendulum thus far have not been successful when it exhibits chaos.
-
   \end{abstract}
@@
   Given any point $u$ and an open neighborhood $W \subset[0,1]$ of $u$, we can find $k\in\mathbb{Z}$ large enough so that $T^k(W) = [0,1]$ because we can accommodate a tent whose base is contained in $W$.
+% \ednote{B: This last phrase doesn't make sense to me.
+M: The number of tents grow with $k$ and hence this is obvious; I explained in one of our meetings.
+Does it make sense?}
   This implies the existence of two points $w_1$ and $w_2$ in $W$ so that $d(T^kw_1,T^kw_2)= 1$, and by the triangle inequality  at least one of the inequalities $d(T^ku,T^kw_1)> \delta$ or $d(T^ku,T^kw_2)> \delta$ hold when $\delta\in (0,\frac{1}{2})$.
@@
 
-\ednote{M: Saying $\phi$ is then a semi-conjugacy between $T$ and $S$ could be ambiguous as it does not specify which system is on the top in the commutativity.
-So please edit accordingly.
-B: Does defining explicitly the domain and codomain of $\phi$ make it clear enough?
-}
+\ednote{B: Does defining explicitly the domain and codomain of $\phi$ make it clear enough?
+M: Saying $\phi$ is then a semi-conjugacy between $T$ and $S$ could be ambiguous as it does not specify which system is on the top in the commutativity.
+So please edit accordingly.}
 When $S$ is conjugate to $T$, the dynamics of the two systems are in some way ''dynamically equivalent''.
@@
 First we establish the method of Takens delay embedding. 
-In this method, one considers a discrete-time dynamical system defined on a 'nice' space (a smooth manifold)\ednote{Should this manifold be further explained?} that can be obtained as time-$K$ map of a continuous time-dynamical system, concepts which will be formalised below.
+In this method, one considers a discrete-time dynamical system defined on a 'nice' space (a smooth manifold)\ednote{B: Should this be further explained?} that can be obtained as time-$K$ map of a continuous time-dynamical system, concepts which will be formalised below.
 We make observations from such a system, i.e., we observe the evolution of $\theta(w_0)$ (where $w_0$ represents the initial state) i.e., we observe  a finite set of values of $u_n :=  \theta(w_n) = \theta(Tw_{n-1})$.
@@
 In almost all cases we can at most insert probes into a system to obtain partial information by means of the measurements taken.
-Moreover, the process of taking a measurement itself introduces 2 additional aspects which complicate the problem: \ednote{M: Maybe replace difficulty by saying that we need to consider two aspects.
-B: Perhaps this edit makes more sense?}
+Moreover, the process of taking a measurement itself introduces 2 additional aspects which complicate the problem: \ednote{B: Perhaps this edit makes more sense?
+M: Maybe replace difficulty by saying that we need to consider two aspects.
+}
 \vspace{-8mm}
@@
 
-\section{Practical Limitations to Takens' Theorem}\label{sect_takenslimits}
+\section{Practical Limitations to Takens' Theorem}
 
@@
 
-The Takens embedding theorem does not guarantee global dissipativity.
-\ednote{B:Links with~\ref{subs_LearnGamma} in discussion of dissipativity of $\Gamma$.
+% The Takens embedding theorem does not guarantee global dissipativity.
+% \ednote{B:Links with~\ref{subs_LearnGamma} in discussion of dissipativity of $\Gamma$.
 Sir, I'm not sure how to write this here - I think it a beneficial point to make so as to set the stage for the later discussion in ~\ref{subs_LearnGamma}, but I am unsure there as well.}
-
-\ednote{Include note on learning failure because of sensitivity as in graph commented out?}
-% \begin{figure}[ht]
-%   \includegraphics[scale=0.3]{_learningfailure.eps}
-%   \centering
-%   \caption{Illustration of a system having failed to learn the map}
-%   \label{fig:supp_learning_failure}
-% \end{figure}
 
@@
 \begin{Definition}
-  [\bf Nonautonomous Dynamical System (adopted from~\cite{Manju_ESP})]\label{Dfn_NDS}\rm
+  [\bf Nonautonomous Dynamical System (adopted from~\cite{Manju_ESP}))] \label{Dfn_NDS}\rm
   A nonautonomous dynamical system (NDS) on a space $X$ is simply a dynamical system comprising a family of maps $\{g_n\}_{n \in \mathbb{Z}}$ where $g_n(\ldotp):=g(u_n,\ldotp):X\to{X}$ is continuous.
@@
 \end{Example}
+% \ednote{Nicely written! Thank you}
 
@@
 
-As the solutions to a driven system are often considered to be an important entity \cite{M: edited the sentence here}, we next identify a subspace $X_U$ of $X$ that contains all possible solutions.
+As the solutions to a driven system are often considered to be an important entity, we next identify a subspace $X_U$ of $X$ that contains all possible solutions.
 To realize such a subspace of a driven system $g$, the concept of a reachable set is defined.
@@
 
-The reachable set itself can defined independently from the fact that $g$ does or does not have the USP.
+The reachable set itself can defined independently from the fact that $g$ does or does not have the ESP/USP.
 We note that $x\in{X_n(\overline{u})}$ if and only $g$ has a solution $\{x_k\}$ for $x_n=x$ and input $\overline{u}$.
 \textbf{(Cite)}.
-The set $X_n(\overline{u})$ is exactly the set of all points $x$ that have some entire solution $\Psi$ passes through it at time $n$.
-\ednote{B: Changed sentence for better readability - still understandable?}
-
-We can give an alternate formulation of the set $X_n(\overline{u})$:
-suppose, for some fixed input 
+
+The set $X_n(\overline{u})$ is exactly the set of points $x$ through which some entire solution $\Psi$ passes at time $n$. We can give an alternate formulation of the set $X_n(\overline{u})$. 
+Now suppose, we for some fixed input 
 $\overline{u}$, we define $g_i = g(u_i,\cdot)$ for all $i\in \mathbb{Z}$, then the set of states
+
 \begin{equation} \label{eqn_association}
@@
 
-As the image of a finite composition of continuous maps, the set $X_{n,i}(\overline{u})$ is compact whenever $X$ is compact and is also nonempty.
-Further, $X_{n,i}(\overline{u}) \supset X_{n,i-1}(\overline{u})$. Hence $\bigcap_{i<n}
+The set $X_{n,i}(\overline{u})$ being the image of a finite composition of
+continuous maps is compact whenever $X$ is compact. Also $X_{n,i}(\overline{u})$ is
+nonempty.  Further, $X_{n,i}(\overline{u}) \supset X_{n,i-1}$. Hence $\bigcap_{i<n}
 X_{n,i}(\overline{u})$ is a nested intersection of closed nonempty subsets, and
-whenever $X$ is compact, the intersection is nonempty.  \ednote{B: Should I cite here?}
-
-\begin{Definition}
-  [\bf Topological Contraction]\label{Dfn_TopContr}\rm
-  We say $g$ is a topological contraction if $\bigcap_{i<n}X_{n,i}(\overline{u})$ is a singleton subset of $X$ for each $\overline{u}$. 
-\end{Definition}
-
-It turns out that $\bigcap_{i<n}X_{n,i}(\overline{u})$ is identically equal to $X_n(\overline{u})$ that we had defined earlier \ednote{(Insert citations.)}.
+whenever $X$ is compact, the intersection is nonempty. We say $g$ is a topological contraction if $\bigcap_{i<n}X_{n,i}(\overline{u})$ is a singleton subset of $X$ for each $\overline{u}$.
+
+It turns out that $\bigcap_{i<n}X_{n,i}(\overline{u})$ is identically equal to $X_n((\overline{u})$ that we had defined earlier.
 This will lead to a result (for.
 e.g., \cite{manjunath2013echo}) that  \textit{$g$ being a topological contraction is equivalent to the existence of a exactly one entire solution.}
+
+% \begin{Definition}
+%   [\bf Topological Contraction]\label{Dfn_TopContr}\rm
+%   A function $g:U\times{X}\to{X}$ is a topological contraction if for all $n\in\mathbb{Z}$ and all $\overline{u}\subset{U}$, $X_n(\overline{u})$ is a singleton subset of $X$.  
+% \end{Definition}
 
@@
 Thus far it has been demonstrated that a system may have one or more solutions; one may ask if a driven system always has a solution and, if so, whether it satisfies certain properties such as uniqueness.
-Should the driven system be compact, existence follows immediately.
+Should the driven system be compact, existence follows immediately as shown in the following result.
 
 \begin{Theorem}\label{Thm_CompactExistence}
- Let $g$ be a driven system.  If $X$ is compact, then for each input $\overline{u}$, there exists at least one solution to the driven system $g(\ldotp, x)$
+ Let $g$ be a driven system.  If $X$ is compact then for each input $\overline{u}$, there exists at least one solution to the driven system $g(\ldotp, x)$
 \end{Theorem}
-{\bf Proof.}  May be found in \cite{kloeden2011nonautonomous, manjunath2014dynamics, manjunath2013echo}
+\begin{proof}
+  May be found in \cite{kloeden2011nonautonomous, manjunath2014dynamics, manjunath2013echo}
+\end{proof}
 
@@
 
-\ednote{B: RC introduced a bit abruptly.
-Adjust.} Reservoir Computing(RC) refers to a machine learning approach in which input is mapped onto a different, but higher dimensional, space through a system called the reservoir after which a readout measuring the reservoir state is trained.
-In the majority of Reservoir Computing(RC) literature \textbf{(Cite)}, a notion of forgetting the states of the reservoir asymptotically is often used.  
-Concepts like the echo state property(ESP) and  fading memory \cite{boyd1985fading} are two of these.  
-If $g$ possesses the ESP (equivalent in our context to the USP \textbf{(Cite)}), we are guaranteed that the whole of an input's left-infinite history will exactly determine the current system state; i.e.
+In the majority of Reservoir Computing(RC) literature (a machine learning approach) where the input is mapped onto a different but higher dimensional through a system called the reservoir and a readout that measures the state of the reservoir is trained, a notion of forgetting the states of the reservoir asymptotically is often used.
+
+% \ednote{How should I explain/qualify RC better here?
+It's mentioned without really explaining.
+M: Have put it in one sentence; just to let you know people use optical devices to map signal into a ``reservoir".
+B: Thanks!},
+Concepts like the ESP,  fading memory \cite{boyd1985fading} are some of these notions.  
+If $g$ possesses the ESP (equivalent in our context to the USP), the we are guaranteed that the whole of an input's left-infinite history will exactly determine the current system state; i.e.
 there is only ever the possibility of a single reachable state at a given instance of time \cite{jaeger2001echo,Manju_2020}.
-% \cite{Manju_ESP}.
-Beyond forgetting the past states, the ESP is also often discussed in terms of a stability property that plays a key role in the robustness of recurrent neural networks(RNN).
-The concept, termed input-related stability~\cite{manjunath2020stability}, refers the phenomenon of nearby inputs creating responses (solutions) close to one another.
-
-\section{Establishing Conjugacies}
+% Defined formally, the ESP is intrinsically linked to an input (see \cite{Manju_ESP}) Unfortunately, definitions and results concerning the ESP do not have much practical benefit since a relationship between the network's response and the temporo-statistical properties of the input are not directly described.
+A more detailed presentation may be found in~\cite{jaeger2001echo}.
+
+% We do however consider it important to mention that the ESP is also often discussed in terms of a stability property that holds for all inputs of a system~\cite{manjunath2020stability} and, furthermore, plays a key role in the design and training of recurrent neural networks(RNN) within the field of RC \cite{Manju_ESP}.
+Beyond forgetting the past states the ESP is also often discussed in terms of a stability property that is nearby inputs create close by responses (solutions), a concept called input-related stability~\cite{manjunath2020stability} and, furthermore, plays a key role in the robustness of recurrent neural networks(RNN), a component of the implementation discussed in~\ref{ch5}.
+
+\section{Conjugacies}
 
@@
 Whenever $g$ has the USP and $\Psi(u)=\{x_n\}_{n\in\mathbb{Z}}$ it follows that $h$, defined by  $h(\cev{u}_n):=g(u_n,x_{n-1})=x_n$, will satisfy the semi-conjugacy in the graph above \ref{Scomm_h}.
-Regrettably, such a continuous mapping $h$ is not guaranteed to exist \cite{M: Have edited here} when $g$ does not have the USP \cite[Lemma 5]{Manju_Nonlinearity}.
+Regrettably, such a continuous mapping $h$ is not guaranteed to exist when $g$ does not have the USP \cite[Lemma 5]{Manju_Nonlinearity}.
 Note that even when $h$ does exist, we are not guaranteed its injectivity. Considering again example \ref{ex_halfux}, we see that even if $h$ were to exist, it could not be injective as $X_U=\{0\}$.
@@
 and we refer to the map $H$ as a \emph{causal embedding}. 
+\ednote{B: edited the sentence here}
 
@@
 
-The inverse-limit space is well-defined since $T : U \to{U}$ is surjective by assumption.We now get to the concept of embedding the inverse limit-space of the dynamical system in $X \times X$.
-
+The inverse-limit space is well-defined since $T : U \to{U}$ is surjective by assumption.
+We now get to the concept of embedding the inverse limit-space of the dynamical system in $X \times X$. 
 \begin{Definition} \rm
@@
   It is imperative to take note of a subtlety here: we refer to the \emph{map} $H$ as a (causal) embedding, whereas the driven \emph{system} $g$ (causally) embeds another \emph{system} $(U,T)$.
+\ednote{B: inserted line}
 
@@
 
-The following theorem establishes the existence of a well-defined map $G_T$ describing the SDD of the system above.
+The following theorem establishes the existence of a well-defined map $G_T$ describing the single-delay dynamics(SDD) of the system above.
 
@@
   [\bf Uniform Attraction Property]\label{Dfn_UAP}\rm
-  A driven system $g$ has the uniform attraction property (UAP) if we initialize the driven system
+  A driven system $g$ has the Uniform Attraction Property (UAP) if we initialize the driven system
 with an arbitrary initial value $y_m \in X$, and the sequence $y_{m+1}, y_{m+2}, y_{m+3},...$ satisfying $y_{k+1}= g(u_k,y_k)$ for $k \geq m$ then approximates an actual solution $\{x_n\}$ uniformly in the sense that given $\epsilon>0$ (independent of $y_m$) there is an integer $n$ so that $d(x_{n+i}, y_{n+i})<\epsilon$ for all $i\ge 0$, where $\{x_m\}$ is a solution.
@@
 \item $g$ has the UAP.
-\end{enumerate\end{Theorem} {\bf Proof.}  May be found in \cite[Th.1]{Manju_Nonlinearity}
+\end{enumerate\end{Theorem}{\bf Proof.}  May be found in \cite[Th.1]{Manju_Nonlinearity}
 
@@
 
-In this section we turn to the implementation of the function $G_T$ by discussing a method to learn the function $\Gamma$.
-We then indicate where this implementation differs from previous work done by (\cite{manjunath2021universal}) and empirically show that the SDD of $G_T$ are less functionally complex than $\Phi_{k, \theta}$ or $T$ by considering the Pearson correlation coefficient.
-Finally, we present and discuss some numerical results for a focus system of this project: the double pendulum. We also consider other chaotic attractors.
+In this chapter we turn to the implementation of the function $G_T$ by discussing a method to learn the function $\Gamma$.
+We apply our methodology to some chaotic attractors.
+Most notably, we consider the reconstruction of the chaotic attractor of the double pendulum - an  attractor that has not been successfully reconstructed because of its intricate structure.
+In fact the previous data-driven approaches also have not been successful in building a model from the double pendulum.
+
+%We then indicate where this implementation differs from previous work done by (\cite{manjunath2021universal}) and empirically show that the SDD of $G_T$ are less functionally complex than $\Phi_{k, \theta}$ or $T$ by considering the Pearson correlation coefficient.
+Finally, we present and discuss some numerical results for a focus system of this project: the double pendulum.
+We also consider other chaotic attractors.
 
@@
 To reach this point, one must first learn $\Gamma$.
-After first normalising the data to ensure it lies within the interval $|x| < 1$ and has zero mean value, we continue by driving the system with external input through the function $g$.
-
-From \eqref{eqn_driving} we know that $ g(u,x)=(1-a)x+a{\cdot}\tanh(Au+\alpha{a}Bx)$.
-Recall that $g$ is SI-invertible.
-Adopting the terminology of \cite{manjunath2021universal}, $A$ is called the “input matrix” and $B$ the “reservoir matrix”.
-
-We make use of a Recurrent Neural Network (RNN) in our implementation.
-RNN's are widely used in the literature for discrete-time processing and are considered to be discrete-time version of continuous dynamical systems modelling a network of neurons.
-(\textbf{Citation.})
-A complete discussion of RNN's is outside the scope of this thesis, so we present only the details necessary to the current discussion.
-For information, (\textbf{Cite, Cite, Cite}) may be consulted.
-We implement our RNN in the \emph{Python} programming language and making use of the \emph{Scipy} library, alongside the \emph{Keras} package in the \emph{Tensorflow} library.
+After first normalising the data to ensure that it has zero mean value and lies within the range $[-1,1]$ and required zero-padding \ednote{M: this zero padding is something that we need to tell the examiner}, we continue by driving the system with external input through the function $g$.
+
+From \eqref{eqn_driving} we know that $g(u,x) = (1-a) x + a \tanh(Au + \alpha B x)$. Recall that $g$ is SI-invertible. We refer to $A$ as the input matrix and $B$ the reservoir matrix.
+
+We make use of a Recurrent Neural Network (RNN) in our implementation to project the data on to a higher dimensional space, but we do not train the RNN in the conventional manner.
+\ednote{B: Should we say what the conventional manner is?} A few words about the RNNs: RNNs are widely used in the literature for temporal processing of data as a type of deep learning technique.
+They take the form of equations considered to be discrete-time version of continuous dynamical systems modelling a neural network of neurons \cite{jaeger2001echo}.
+
+We implement a feedforward network, (a network with no memory, unlike the RNN)\ednote{B: Added so that FFN is in more context} required to learn $\Gamma$ in the \emph{Python} programming language and making use of the \emph{Scipy} library, alongside the \emph{Keras} package in the \emph{Tensorflow} library \ednote{B: Is this something you want me to write in?
+M: Note!
+We do not implement the RNN, but the FN that learns Gamma through the Tensorflow.
+We called the RNN + FN set up a RCN in Adriaan's paper.
+FYI: FN's work only if there is a definite relationship between the $u_n$ and $u_{n+1}$ which is rare in temporal data.
+RNNs with ESP work when there is a definite relationship between $(\ldots,u_{n-1},u_n)$ and $u_{n+1}$ and is suited for temporal processing.}.
 
 If the RNN we will be using in the next step has $N$ neurons, then $A$ and $B$ are each $N\times{N}$ matrices.
-Inputs $u_n$ of dimension $K < N$ are padded by zeroes to obtain an input of dimension $N$.
-As we do not need the entire left-infinite history of input, an arbitrary initialisation of the system with initial values followed by the driving of the system with the function $g$ means that the generated values will approach the actual solution’s elements in a uniform manner thanks to the UAP.
-\textbf{(Citation)}
+Inputs $u_n$ of dimension $K < N$ are padded by zeroes (as remarked earlier) to embed the input in a space of dimension $N$.
+As we do not need the entire left-infinite history of input, an arbitrary initialisation of the system with initial values followed by the driving of the system with the function $g$ means that the generated values will approach the actual solution elements in a uniform manner thanks to the UAP that we had discussed in the preceding chapter.
 
@@
 Utilising a network $N$ neurons (the same value $N$ as in the dimensions of the matrices $A,B$).
-Training is done using the Adam Optimiser where the MSE loss function is optimised.
-The system is trained several times(in our case 4) with incrementally smaller learning rates on each run (0.001 and thereafter divided by 10 at each run).
+Learning of $\Gamma$ is done using the Adam Optimiser where the means square error (MSE) loss function is optimised.
+The system is trained several times (in our case 4) with incrementally smaller learning rates on each run (0.001 and thereafter divided by 10 at each run).
 Additional parameters such as the number of network-layers, neurons per layer and training-length are presented case-by-case below.
@@
 
-\section{Changes in this project to previous work of~\cite{manjunath2021universal}}
-
-To understand the first and most drastic\ednote{change word} change, we recall the the conjugacy in~\ref{fig:inverse_limsystem}:
-
-\begin{equation*}
-  %  \[ 
-      \psset{arrows=->, arrowinset=0.25, linewidth=0.6pt, nodesep=3pt, labelsep=2pt, rowsep=0.7cm, colsep = 1.1cm, shortput =tablr}
-      \everypsbox{\scriptstyle}
-      \begin{psmatrix}
-      \widehat{U}_T  && \widehat{U}_T \\
-      Y_T && Y_T \\
-      %%%
-     %  \ncline{1,1}{1,2}^{\widehat{T}} \ncline{1,1}{2,1} <{h}
-     %  \ncline{1,2}{2,2} > {H_2}
-     %  \ncline{2,1}{2,2}^{G_T}
-      \end{psmatrix}
-    %  \]
-  \end{equation*}
-
-  We recall that $G_T$ is topologically conjugate to $\widehat{T}$, a result which we have already made use of.
-
-  The authors of~\cite{manjunath2021universal} did not, however, consider the relationship between the inverse-limit space $\widehat{U}_T$ and the input $u_n=\theta(w_n)$ (where $\theta$ is our observable and $w_n$ the complete input), a concept we now explore.
-  We may construct a delay-coordinate $v_n=[\theta(w_n), \theta(w_{n-1}), \ldots, \theta(w_{n-d})]^{T} = [u_n, u_{n-1}, \ldots, u_{n-d}]^T$
-    to accordingly obtain a new dynamical system $(V,F)$ where $F$ is the map $F:v_n\mapsto{v_{n+1}}$.
-If $d$ is sufficiently large, we may apply Takens Embedding Theorem~\ref{Thm_Takens} to establish the existence of a topological conjugacy between the new system $(V,F)$ and the original system $(U,T)$.
-
-  Using the same notation as before, we may construct a new conjugacy which now relates delay coordinates to one another.
-
-  \begin{equation}\label{eqn_VFdelay}
-    %  \[ 
-        \psset{arrows=->, arrowinset=0.25, linewidth=0.6pt, nodesep=3pt, labelsep=2pt, rowsep=0.7cm, colsep = 1.1cm, shortput =tablr}
-        \everypsbox{\scriptstyle}
-        \begin{psmatrix}
-        \widehat{V}_F  && \widehat{V}_F \\
-        Y_F && Y_F \\
-        %%%
-       %  \ncline{1,1}{1,2}^{\widehat{F}} \ncline{1,1}{2,1} <{h}
-       %  \ncline{1,2}{2,2} > {*}
-       %  \ncline{2,1}{2,2}^{G_F}
-        \end{psmatrix}
-      %  \]
-    \end{equation}
-
-The existence of $\widehat{F}$ is guaranteed, but the map $v_{n}\mapsto{v_{n+1}}$ itself is unknown.
-We may, however, make use of the SI-invertibility of the function $g(u_{n-1}, x_{n-1})=x_n$ to conclude that
-$(x_{n-1}, x_{n})$ determines $u_{n-1}$ uniquely.
-
-Accordingly $(x_{n-2}, x_{n-1}, x_{n})$ determines $u_{n-2}$ and $u_{n-1}$ uniquely and generally $(x_{n-d}, \ldots , x_{n-1}, x_{n})$ determines $u_{n-d}, \ldots u_{n-2}$ and $u_{n-1}$ uniquely.
-
-We therefore need not know what $\widehat{F}$ is, as we can determine $v_{n+1}={[u_n, u_{n-1}, \ldots, u_{n-d}]}^T$ indirectly by determining the value of $(x_{n-d}, \ldots , x_{n-1}, x_{n})$, a much more feasible task.
-
-Accordingly we define the following abbreviations:
-\begin{itemize}
-  \item Delay Coordinates (DC): A vector composed by appending a fixed amount of previous values to the final value.
-  \item Single-Delay Dynamics(SDD): Dynamics generated by 2-element delay-coordinate vectors of the form $(x_{n-1}, x_n)$ where only one previous value is used.
-  \item Multi-Delay Dynamics(MDD): Dynamics generated by arbitrary-length delay-coordinate vectors of the form $(x_{n-d}, \ldots , x_{n-1}, x_{n})$.
-($d$ must be sufficiently large, but is unknown.
-We vary this parameter in experimentation.)
-\end{itemize}
-
-Within this framework, the PCA-implementation of~\cite{manjunath2021universal} is then considered to fall underneath the of auspices of what we term \emph{Approximated Multi-Delay Dynamics(AMDD)}.
-
-% We have now seen that one may comfortably switch between the use of SDD or MDD when $g$ has SI-invertibility. /ednote{This sentence needs to be improved.}
-
-Naturally, one would immediately suppose that the MDD approach would increased computation time and resource use as the internal matrices and datasets generated during learning would grow with every increase in the delay-parameter $d$.
-
-In fact, preliminary results indicate the opposite due the observation that the implementation of MDD is inversely proportional to the the dimension of the network (i.e.
-the number of neurons in the RNN) and possibly also the number of epochs/batch size necessary to produce reliable predictions.
-This, in turn, actually speeds up the computation-time and provides results of similar accuracy.
-\ednote{I did not have enough time to explore this properly.
-I can only hazard guesses and promise future research work in this project.
-How should I best convey this?
-}
-
-\textbf{Secondly}
-
-Sparse Matrices:
-\vspace{-8mm}
-\begin{itemize}[noitemsep]
-\item Well-conditioned: condition number close to 1
-\item Scaled
-\item Sparsity/density ratio
-\end{itemize}
-
-% Sparse matrix implementation.
-Slight improvement in some scenarios, but not fantastic.
-\ednote{M: Did it not help you a great deal with the DP forecasting?
-B: Sparse matrices tended to be finicky.
-In some cases they helped whereas in others they actually worsened the solution somewhat.
-There's some tuning to be done in terms of condition number, scaling and density}
-
-\textbf{Third}
-
-Density of matrices A/B(i.e. number of non-zero entries in each matrix) (separate from sparse matrix implementation)
-
-\section{Numerical simulations}
-In this section we showcase numerical results pertaining to 3 attractors that were simulated using the methodology described in the previous chapters. We first consider a simple physical system.
+\section{Delay Coordinates}
+
+We also consider a more general problem of learning an attractor of dynamical system when only observations of an orbit are given.
+Explicitly, if $(W,T)$ is a dynamical system with dynamics generated by $w_{n+1}=Tw_n$, and if $\theta:W \to \mathbb{R}$ is an observable, then the task is to learn a dynamical system that is topologically conjugate to $(W,T)$ and predict $\theta(w_{m+1}),\theta(w_{m+2}),...$ using the data $\theta(w_{0}),\theta(w_{1}),\ldots,\theta(w_{m})$.
+So how do we do it?
+Suppose the input generated from the delay-coordinate map $\Phi_{\theta,2d}(\theta(w_{n})) := (\theta(w_{n-2d}),\ldots,\theta(w_{n-1}),\theta(w_{n}))$ is such that Takens delay embedding theorem (see Theorem~\ref{Thm_Takens}) holds, in which case, there exists a homeomorphism  $F_\theta: \Phi_{\theta,2d}(\theta(w_{n})) \mapsto \Phi_{\theta,2d}(\theta(w_{n+1}))$.
+Then if we feed  the input values $u_n := \Phi_{\theta,2d}(\theta(w_{n}))$ to a driven system $g$ that is SI-invertible and has USP, the induced system $(Y_F,G_F)$ would be topologically conjugate to the inverse-limit space of
+$(\Phi_{\theta,2d}(W), F_\theta)$ as in Theorem~\ref{Thm_CET} \ednote{M: please see if this refers to the causal embedding theorem in the thesis}, and hence one could forecast $u_n,u_{n+1},\ldots$.
+In other words, one could forecast  the values $\theta(w_n), \theta(w_{n+1}),\ldots$ by feeding these observations as delay coordinates.
+The advantage of feeding delay-coordinates to a driven system is that the embedding is stable in the sense of global dissipativity that we have described in the previous chapter, i.e., $Y_F$ is made to contained in an attractor of a mapping induced on $Y_F$ that is induced by $g$.
+
+We note that the twin equations together~\ref{eqns_from_data} generate a model from data.
+For a chaotic system, a small amount of noise due either to the input-noise or computational error would cause the solutions from two accurate models with the same initial conditions to  diverge as they are not identical.
+\ednote{B: edited this sentence} However, statistically, they would be the same since due to ergodic theory all typical orbits have the same visiting frequency to any region of the phase space of the dynamical system.
+\ednote{B: I'll need to recap this sentence with you before presenting as it's new to me.} To illustrate that we plot the densities of the trajectories of the actual and predicted solutions.
+
+Our implementation of the reservoir(A) and input(B) matrices in the RNN makes use of sparse matrices.
+We discuss three specifics relating to this concept.
+Firstly, the matrices are assigned a variable density in terms of the amount of non-zero entries and are scaled to have values within the range $[-1,1]$.
+For this, we make use of \textit{Python's SciPy} library and specifically the \textit{Sparse, LinAlg} modules.
+In our experiments, we generally opted for a density ratio of 10\%-25\%.
+Secondly, we opted to enforce the requirement that both matrices be well-conditioned, i.e.
+that they have a condition number of \~1.
+The condition number of a matrix $M$ (denoted $cond(M)$) is the product of the matrix norms of the matrix $M$ and its inverse $M^{-1}$ and is a measure of the sensitivity of output of a matrix multiplication when an input is changed.
+\textbf{(Cite.)} By ensuring a condition number close to 1, we also ensure that our driven function $g$, depending on the matrices $A$ and $B$, will not be sensitive to adjustments in the input.
+In so doing we then also guarantee a measure of stability.
+\ednote{B: Not satisfactory.
+Edit again.}
+
+Lastly, we ensure the matrix $B$ has unit spectral radius.
+This is necessary due to the Causal Embedding Theorem~\ref{Thm_GT_Exists} which can then guarantee the existence of semi-conjugacy between the systems $(Y_T, G_T)$ and $(\widehat{U}_T, \widehat{T})$.
+In fact, the theorem actually requires that $\alpha{B}$ have spectral radius $<1$, so we impose the restriction $\alpha<1$ and consider only the matrix $B$'s spectral radius from thereon.
+
+Note that to learn $\Gamma: (x_{n-1},x_{n}) \mapsto u_n$ throughout feedforward network, one could learn from the principal components of $(x_{n-1},x_{n})$ as well. 
+We employ this since empirically it gives slightly better results and not essential.
+The principal components are used only for an efficient state representation possibly to reduce the errors while learning and is not for a lossy approximation since all principal components are used.
+  To be explicit, we denote $X_{1:N}$ to be the matrix with the first $N$ states of the network data as row vectors.
+So, if $X_{1:N}=U\Sigma P^T$ represents the singular value decomposition of $X_{1:N}$, then the principal component matrix is denoted by $P$, and the principal components are given by
+ $Z_{1:N}=X_{1:N}P.$
+These principal components are then used to train a feedforward neural network. If we denote the row vectors of $Z_{1:N}$ by 
+$z_i^T, i=1,2,\cdots,N$ and the neural network by $NN$, then the network is trained to learn an approximation of the map
+$NN: (z_{n-1},z_n) \mapsto  u_n.
+$
+The learnt approximation of $NN$ can be used to approximate $\Gamma$ since 
+$$
+NN\left( \begin{bmatrix} 
+P^Tx_{n-1} \\
+P^Tx_n
+\end{bmatrix}
+\right) = NN \circ 
+\begin{bmatrix}
+P^T & 0 \\
+0 & P^T 
+\end{bmatrix}\begin{bmatrix}
+x_{n-1}\\
+x_n
+\end{bmatrix} = u_n = \Gamma(x_{n-1},x_n).
+$$
+
+\section{Experimental Results}
+In this section we showcase numerical results pertaining to 3 attractors that were simulated using the methodology described.
+We first consider a simple physical system that exhibits complicated dynamics.
 \subsection{Double Pendulum}
@@
 The system has four variables: angle and angular velocity of each pendulum rod. An ideal system is assumed where the rods are massless and no friction in the pivots are considered.
-
 \begin{figure}[ht]
@@
 
-The scalar equations of motion are derived using Newtonian physics and can be found in~\cite{DPFormulas}.
-The second-order system is converted to a first-order ODE and solved using the Runge-Kutta numerical method.
-Noise-free, undamped DP data-sets were generated and utilised in this project due to the requirement that the map $T$ be surjective.
-We discuss the case of a DP exhibiting energy-loss (i.e.
-where the map $T$ is not surjective) below.
-
-\textbf{Data originating from a non-surjective map.}
-
-The importance of the requirement that data originates from a surjective map is observed learning failure of a map describing a damped DP.
-Due to the assumption, the program discounted any possibility of energy-loss and we consider the example below.
+The scalar equations of motion are derived using Newtonian physics and are reproduced from~\cite{DPFormulas} in~\ref{eqns_dp}.
+Each pendulum component has an associated mass of the head, $m$, and length of the rod, $L_i$.
+Acceleration due to gravity is denoted by $g$ and initial velocities by $\dot{\theta}$
+\begin{eqnarray}\label{eqns_dp}
+  \ddot{\theta_{1}}  = \frac{-g(2m_1+m_2)\sin{\theta_1} - m_2g\sin(\theta_1-2\theta_2) - 2\sin(\theta_1-\theta_2)m_2({\dot{\theta_{2}}}^{2}L_2 + {\theta_{1}'}L_1\cos(\theta_1-\theta_2))} {L_1(2m_1 + m_2 -m_2\cos(2\theta_1 - 2\theta_2))}
+  \\
+  \ddot{\theta_{2}} = \frac{2\sin(\theta_1-\theta_2)(\theta_{1}'^{2}L_1(m_1+m_2) + g(m_1+m_2)\cos\theta_1 + \dot{\theta_{2}}^{2}L_2m_2\cos(\theta_1-\theta_2))}{L_2(2m_1 + m_2 -m_2\cos(2\theta_1 - 2\theta_2))}
+\end{eqnarray}
+
+The second-order system is converted to a first-order ODE system and solved using the Runge-Kutta numerical method.  
+Here we chose the values $m_1=2.0$, $m_2=2.0$, $L_1=1.5$ and $m_2=1.5$ with initial velocities $v_1=1.5, v_2=-2$ for the respective pendulum heads.
+Undamped DP data-sets were generated and used in this project due to the requirement that the map $T$ be surjective that we must not forget. 
+An example of this would be the case of a DP exhibiting energy-loss (i.e.
+where the map $T$ is not surjective).
+In such a case, the attractor is a single point in the phase space when the double pendulum comes to rest.
+It must be mentioned once again that if we use data outside the attractor, learning is not reliable. To see this, consder the graphs ~\ref{fig:damped_pendulum} and ~\ref{fig:dp_notsurjective}.
 
@@
   \centering
-\caption{The map generating data for a damped pendulum is not surjective. Data obtained from chaotic double pendulum dataset generated by~\cite{asseman2018learning}}
+\caption{$x$- and $y$-coordinates of the two pendulum heads plotted for a dataset that exhibits dying oscillations.
+The map generating the (four-dimensional data) for a damped pendulum is not surjective.
+Data obtained from chaotic double pendulum dataset generated from videos takes of experimental pendulums~\cite{asseman2018learning}}
 \label{fig:damped_pendulum}
@@
 
-Despite feeding data that originated from a DP with dying occilations into the system, a map was instead learnt which continued to oscillate in a chaotic, but conservative manner.
-
-\begin{figure}[ht]
-  \includegraphics[scale=0.4]{_dpfail_nonsurj.eps}
-  \centering
-\caption{Failure to learn a map which is not surjective.}
+\begin{figure}[ht]
+  \includegraphics[scale=0.3]{_dpfail_nonsurj.eps}
+  \centering
+\caption{Failure to forecast or learn a map which is not surjective: Evolution of the $y$-coordinate of the bottommost pendulum plotted here for 15000 timestep.
+Plotted here the predicted value(red) versus the actual value(blue).
+The plots overlap for the first 10000 steps, the duration of training to learn the map $\Gamma$.
+Afterwards, the trajectories start diverging and it the predicted value no longer accurately represents the complexity or energy-loss of the actual data.
+}
 \label{fig:dp_notsurjective}
@@
 
-Having spent ample time to discuss the challenges, we now change tone so as to satisfy the optimist.
-We present successful run for DP-data to repdict the future movement of the pendulum.
+Having spent ample time to discuss the challenges, we now change tone so as to satisfy the optimist and present a successful run of the DP-data to predict the future movement of the pendulum.
 
@@
 Training was accomplished using 256 epochs with batch sizes 128 and a delay of 1.
-
-% Trainlen:15000; Predlen: 1000; Discard: 500; Dimension:128 neurons; $\alpha=0.99$; $a=0.5$.
-Hidden Layers: 16.
-Layer dimension: 64.
-Epochs:256.
-Batch Size: 128.
-Delay(for MDD): 8.
-Density of Sparse Matrices: 0.2.
-$d=1$
-
-\begin{figure}[ht]
-  \centering
-  \includegraphics[width=0.8\linewidth]{_DP_Fig-2F2023.eps} 
-  \captionof*{figure}{Predicted(red) and actual(blue) trajectories of the $x$- and $y$-coordinates of each of the pendulum heads.} 
-% \end{figure}
-% \begin{figure}[ht]
-%   \centering
-  \includegraphics[width=0.8\linewidth]{_DP_Fig-2B2023.eps}
-  \captionof*{figure}{Densities of predicted and actual values of x-coordinate of first pendulum head. }
-  % \end{figure}
-  % \begin{figure}[ht]
-  %   \centering 
-  \includegraphics[width=0.8\linewidth]{_DP_Fig-2D_5.eps}
-    \captionof*{figure}{Takens Delay Map with delay=5 to illustrate behaviour of predicted and actual values of double pendulum data.}
-    \label{fig:dp_predictions}
+See~\ref{fig:dp_success_traj} for results pertaining to the trajectory taken by the DP's components and then~\ref{fig:dp_success_density} to witness associated densities of predicted vs.
+actual values.
+
+\begin{figure}[ht]
+  \centering
+  \includegraphics[width=0.95\linewidth]{_dp_success_4coords_traj.eps} 
+  \captionof{figure}{Predicted(red) and actual(blue) trajectories of the $x$- and $y$-coordinates of each of the pendulum heads.
+From top to bottom are as follows: $x$-coordinate of topmost pendulum head, $y$-coordinate of topmost pendulum head, $x$-coordinate of lower pendulum head, $y$-coordinate of lower pendulum head.
+  These graphs were constructed by predicting the DP 1000 timesteps into the future and in so doing illustrating the long-term consistency and accuracy of the learnt system.
+Here we lock onto the near-exact trajectory for about 300 timesteps and stay close up until about 600 seconds.
+}
+  \label{fig:dp_success_traj}
+\end{figure}
+\begin{figure}[ht]
+  \centering
+  \includegraphics[width=\linewidth]{_dp_success_4coords_hist.eps}
+  \captionof{figure}{Densities of predicted and actual values of $x$- and $y$-coordinate of pendulum heads.}
+  \label{fig:dp_success_density}
+ \end{figure}
+
+We now consider an experiment where noise is added to the system and find it to be incredibly robust.
+We employ the same parameters as before and add noise from a normal-distribution with mean zero and standard deviation equal to 0.1.
+Once again consider the actual and predicted trajectories for the noisy DP dataset in Figure~\ref{fig:noisydp_success_traj}.
+Next we consider the densities of the predicted and actual values in Figure~\ref{fig:noisydp_success_density} and observe that they display highly similar distributions.
+
+\begin{figure}[ht]
+  \centering
+  \includegraphics[width=0.95\linewidth]{_dp_noise_FigCD_2030.eps} 
+  \captionof{figure}{Predicted(red) and actual(blue) trajectories of the $x$- and $y$-coordinates of each of the pendulum heads with noise added from a normal distribution with mean zero and standard deviation 0.1.
+  These graphs were constructed by predicting the DP 1000 timesteps into the future and in so doing illustrating the long-term consistency and accuracy of the learnt system.
+Here we lock onto the near-exact trajectory for about 250 timesteps and stay close up until about 400 seconds.
+}
+  \label{fig:noisydp_success_traj}
+\end{figure}
+\begin{figure}[ht]
+  \centering
+  \includegraphics[width=\linewidth]{_dp_noise_2B_2030.eps}
+  \captionof{figure}{Densities of predicted and actual values of x-coordinate of first pendulum head.}
+  \label{fig:noisydp_success_density}
  \end{figure}
@@
 The behaviour of the system with a $\beta$-parameter value of 0.1056 was simulated by scaling it to fit inside the interval $[-1,1]$.
-Two cases were considered: noise-free and perturbed.
-Noise was generated from a normal distribution with mean 0 and standard deviation 0.05, which translates approximately to *dB.
+Two cases were considered: noise-free and the one perturbed by noise.
+Noise was generated from a normal distribution with mean 0 and standard deviation 0.05, which translates approximately to * dB \ednote{Please calculate dB as signal to noise ratio }
  A sequence of 3000 observations in both the clear and noisy data-sets respectively were fed into the system with 300 entries discarded to simulate the network's loss of memory.
@@
     \centering
-    \includegraphics[width=\linewidth]{_Thomas_nonoise.eps}
-    \caption*{True and predicted trajectories in the absence of noise.}
+    \includegraphics[width=0.8\linewidth]{_thomas_noisy_0.1056.eps}
   \endminipage\hfill
   \minipage{0.5\linewidth}
-    \centering
-    \includegraphics[width=\linewidth]{_ThomasNoisy.eps}
-    \caption*{True and predicted trajectories for noise added.}
+    \centering    \includegraphics[width=0.8\linewidth]{_thomas_clean_0.1056.eps}
   \endminipage
+  \captionof{figure}{True(red) and predicted(blue) trajectories for the Thomas attractor with parameter value $\beta=0.1056$.
+To the left: Noise added from normal distribution with mean zero and standard deviation 0.05.
+Right: No noise added to dataset.}
 \end{figure}
@@
   \centering
-  \includegraphics[scale=0.35]{_Thomas_1.eps}\caption*{Predicted trajectories for the attractor demonstrate empirically the ability to predict the evolution of the trajectory for the next ~100 timesteps}
-  \includegraphics[scale=0.5]{_Thomas_3.eps}\caption*{Represented here the learnt (blue) and actual (red) densities of the first coordinate of the dataset.}
+  \includegraphics[scale=0.35]{_Thomas_1.eps}\caption*{Predicted trajectories of the $x$- and $y$-coordinates for the Thomas attractor with parameter-value $\beta=0.1056$ demonstrate empirically the ability to predict the evolution of the trajectory for the next an estimated 100 timesteps into the future near-exactly.
+Here no noise was added.}
+  \includegraphics[scale=0.5]{_Thomas_3.eps}\caption*{Represented here the learnt (blue) and actual (red) densities of the first coordinate ($x$-coordinate) of the dataset for the Cyclically Symmetric Attractor with parameter-value $\beta=0.1056$.}
 \end{figure}
@@
   \centering
-  \includegraphics[width=1.0\textwidth,left]{Clifford_1.eps}
-  \caption*{These graphs were constructed by predicting the Clifford system 15000 steps into the future and in so doing illustrating the long-term consistency and accuracy of the learnt system.
+  \includegraphics[width=1.0\textwidth,left]{_Clifford_1_nonoise.eps}
+  \caption*{These graphs were constructed by predicting the Clifford system 1000 steps into the future and in so doing illustrating the long-term consistency and accuracy of the learnt system.
 As perceived here, we are able to lock on to the trajectory of the Clifford map almost exactly for the first 25 steps.}
@@
       \centering
-      \includegraphics[width=\linewidth]{Clifford_2.eps}
+      \includegraphics[width=\linewidth]{_Clifford_2_nonoise.eps}
   \endminipage\hfill
@@
     \centering
-    \includegraphics[width=\linewidth]{Clifford_3.eps}
+    \includegraphics[width=\linewidth]{_Clifford_3_nonoise.eps}
   \endminipage
@@
 
-\ednote{B: Expand more.}
-
-\section{Generalized Pearson Correlation}\ednote{B: Alter heading.}
+The behaviour of the system with a parameter values of $x_0=0$, $y_0=0$, $a = -1.7$,  $b = 1.8$, $c = -1.9$ and $d = -0.4$ was examined.
+Data was scaled to fit inside the interval $[-1,1]$. Only the noise-free case was investigated. 
+ A sequence of 25000 observationswas fed into the system with 15000 entries discarded to simulate the network's loss of memory. The next 1000 steps were predicted and compared with true data.
+The RNN was initialised exactly as for Thomas' Cyclically Symmetric Strange attractor: the dimension of the network is set equal to 500 neurons and 12 hidden layers of dimension 64 each were initialised.
+Training is realised through 150 epochs of batch sizes of 128 each; the map is learnt via the Adam Optimiser using finer and finer learning rates (0.001 and then divided by 10 at every iteration).
+
+\section{Functional Complexity Reduction in Noninvertible maps}
 
@@
 
-\ednote{B: Write Pearson Introduction. Could you give me some help here, Sir?}
+In the case of non-invertible maps, we note that the map $G_T$ is a homemorphism although $T$ is not.
+Further, empirically it is found that $G_T$ is much more linear than $T$ as measured by what is called the multidimensional correlation coefficient $\rho$.
+This coefficient is a generalization of the Pearson correlation coefficient for real-valued random variables to random vectors (e.g., \cite{puccetti2019measuring}).
+We state teh details of computing this coefficient between
+$(x_{n-1},x_n)$ and $G_T(x_{n-1},x_n)$. 
+If $\Sigma_{a}$ and $\Sigma_{b}$ denotes the covariance matrices of the vectors $(x_{n-1},x_n)$ and $G_T(x_{n-1},x_n)$ respectively, and if $\Sigma_{ab}$ denotes the covariance matrix between the vectors $(x_{n-1},x_n)$ and $G_T(x_{n-1},x_n)$, then the multidimensional correlation coefficient is computed by using the traces of these matrices:
+\[
+    \rho= \frac{\text{tr}({\Sigma_{ab}})}{\text{tr}({\sqrt{\Sigma_a\Sigma_b}})}.
+\]
+This multidimensional correlation coefficient satisfies one of the desired and well-known properties of the one-dimensional Pearson coefficient.
+That is, $\rho=\pm 1$ if and only if $Y\overset{d}{=}AX+b$ for some invertible matrix $A$ and vector $b$.
+Hence $\rho$, and in practice can be used  to measure the linear relationship between two random vectors of the same dimension and thus serves as an indicator of functional complexity.
+For the non-invertible system defined by the Clifford attractor, we indicate the difference by which $\rho$ changes for $G_T$ in Table~\ref{tbl_attractorsPearson}.
+Following \cite{manjunath2021universal}, we say that this multidimensional correlation coefficients $\rho$ to evidence a general reduction in the functional complexity of the map $G_T$ emerging due to a RNN.
+In the
+Table ~\ref{tbl_attractorsPearson}, the second column corresponds to the amount of artificial neurons (the dimension of $X$) in the implemented RNN.
 
@@
           \hline      
-
           % \hline
@@
           %     & 1000 & & 0.9930\\              
-
           % \multirow{3}{*}{$u_n=\frac{1}{10}(\sin(0.1w_{n,x})+\sin(0.1w_{n,y})+\sin(0.1w_{n,z}))$} 
@@
           % \hline
-
-          \hline
-          $(w_n)$ comes from the Thomas Attractor (for $\beta=0.1056$): & & & \\
-          {$w_{n+1}= \begin{bmatrix} \sin(w_{n,y}) - \beta{w_{n,x}} \\
-              u_y = \sin(w_{n,z}) - \beta{w_{n,y}} \\
-          u_z = \sin(w_{n,x}) - \beta{w_{n,z}} \end{bmatrix}$} & & & \\
-          \multirow{3}{*}{$u_n = w_n - \overline{w}$}
-              & 10 & 0.99989 & 0.999886 \\
-              & 100 & &  0.999913 \\
-              & 1000 & & 0.999901 \\
-          \hline
-
-          \hline 
-          $(w_n)$ comes from Noise-Added Thomas Attractor (for $\beta=0.1056$): & & & \\
-          {$w_{n+1}= \begin{bmatrix} \sin(w_{n,y}) - \beta{w_{n,x}} \\
-              u_y = \sin(w_{n,z}) - \beta{w_{n,y}} \\
-          u_z = \sin(w_{n,x}) - \beta{w_{n,z}} \end{bmatrix}$} & & & \\
-          \multirow{3}{*}{$u_n = w_n - \overline{w}$}
-              & 10 & 0.9994 & 0.99985 \\
-              & 100 & &  0.99991 \\
-              & 1000 & & 0.999895\\
-          \hline
-
-          \hline
-          $(w_n)$ comes from the Thomas Attractor (for $\beta=0.208186$): & & & \\
-          {$w_{n+1}= \begin{bmatrix} \sin(w_{n,y}) - \beta{w_{n,x}} \\
-              u_y = \sin(w_{n,z}) - \beta{w_{n,y}} \\
-          u_z = \sin(w_{n,x}) - \beta{w_{n,z}} \end{bmatrix}$} & & & \\
-          \multirow{3}{*}{$u_n = w_n - \overline{w}$}
-              & 10 & 0.99982 & 0.99981\\
-              & 100 & &  0.999851 \\
-              & 1000 & & 0.99984 \\
-          \hline
 
@@
               & 1000 & & 0.84854 \\
-          \hline  
           % \hline 
@@
           %       & 1000 & & 0.71151 \\
-          % \hline
-          \hline
-          $(w_n)$ comes from the DP \\ (for ): & & & \\
-          \multirow{3}{*}{$u_n = w_n-\overline{w}$}
-                & 10 & 0.99141 & 0.98916\\
-                & 100 & &  0.99238 \\
-                & 1000 & & 0.99278 \\
            \hline
                 \end{tabular}} %\ednote{http://paulbourke.net/fractals/peterdejong/}
-
+ \label{Table_FC}
  \caption{Following \cite{manjunath2021universal}, we determine the multidimensional correlation coefficients $\rho$ to evidence a general reduction in the functional complexity of the map $G_T$ emerging due to a RNN.
@@
           Finally, the RNN's dimension (from \eqref{Seq_RNN}) is provided to exhibiting that an increase in dimension of the driven system will 
-          typically result in a map $G_T$ with less functional complexity.} 
-          \label{Table_FC}
+          typically result in a map $G_T$ with less functional complexity.}.
       \end{table}\label{tbl_attractorsPearson}
@@
 
-\chapter{Conclusion}\label{ch6}
+\chapter{Conclusions} \label{ch6}
 
@@
 In this project, it has been shown that an input originating from a compact space entailing exactly one solution (i.e., if the driven system has the USP) is equivalent to representing information without distortion, a concept which was formalised by establishing a causal mapping’s existence.
-The established results are general, implying that any driven dynamical system with specific properties can give rise to a causal mapping, and in so doing  providing an answer as to the success of arbitrary choices of driven systems in information processing applications in [4, 5].
-
-In particular, we have a topological conjugacy (semi-conjugacy) between the data comprising the 11single-delay lag dynamics of a driven system’s and the underlying homeomorphic dynamical system.
+The established results are general, implying that any driven dynamical system with specific properties can give rise to a causal mapping, and in so doing  providing an answer as to the success of arbitrary choices of driven systems in information processing applications in \cite{jaeger2004harnessing, mass2002realtime}.
+
+In particular, we have a topological conjugacy (semi-conjugacy) between the data comprising the single-delay lag dynamics of a driven system~\cite{manjunath2013echo} and the underlying homeomorphic dynamical system.
 Finite-length input data is sufficient for forecasting as driven system states are influenced only in an almost negligibly small manner by the left-infinite segment of the past of the input beyond some arbitrarily large but finite time (a result which holds due to the universal semi-conjugacy of a driven system’s continuity)
@@
 
+\bibliographystyle{pnas-new}
+\bibliography{pnas}
+
 \end{document}

### 2022-01-16 forcomparison.tex
Many graphs added and comments added. Edits of Ch.5. Starting to prep to add Grammarly back in

--- 2022-01-16 draft_16Jan/draft5.tex
@@
 
-Experimenting on biological, physical, and artificial systems in order to generate a more informative dynamical model is a well-established practice in modern science.
-Traditional methods for modelling physical systems are based on laws of physics that are based either on empirical relationships or on intuition.
+Experimenting with biological, physical, and artificial systems to generate a more informative dynamical model is a well-established practice in modern science.
+Traditional methods for modelling physical systems are based on laws of physics that are based either on empirical relationships or intuition.
 For systems that evolve with time, physical laws yield mathematical equations that govern how the quantities evolve with time.
-However our world around is, for the most part, much more complex than that which can be distilled into elegant equations.
-We do not fully understand many of the more complex systems, nor do they even provide us with a good physical intuition of the underlying principles governing the dynamics.
-To complicate this - the underlying systems we observe often display sensitive dependence on initial conditions; despite having highly similar initial conditions, differing orbits diverge quickly and to such an extent that it becomes seemingly impossible to retrace their steps back to the original conditions.
-Even the presence of usually 'negligible' computational noise or measurement error renders long-term point-wise prediction infeasible.
+However, our world is much more complex than that which can be distilled into elegant equations.
+We do not fully understand many of the more complex systems, nor do they provide us with a good physical intuition of the underlying principles governing system dynamics.
+To complicate this, the underlying systems we observe often display sensitive dependence on initial conditions despite having highly similar initial conditions.
+This is because differing orbits diverge quickly and to such an extent that it becomes seemingly impossible to retrace their steps back to the original conditions.
+Even the presence of usually' negligible' computational noise or measurement error renders long-term pointwise prediction infeasible.
 There are many difficulties that one encounters while modelling such systems:
+
 \vspace{-8mm}
@@
   \item One may not have access to the complete states of the systems
-  \item The actual system may be described by functions that behave wildly in the sense their graphs have a wild oscillatory behaviour, i.e., they have a large functional complexity.    
+  \item The system may be described by functions that behave wildly because their graphs have a wild oscillatory behaviour, i.e., a large functional complexity.
 \end{enumerate}
@@
 
+Examples in the literature attempting to forecast data from such systems have tried many different approaches with differing degrees of success.
+When we have complete access to states of the system, an ordinary differential equation can be obtained from data ([?, ?] and [?, ?]) wherein one could approximate the vector field by a library of functions to get interpretable models.
 Examples in the literature attempting to forecast data from such systems have tried a number of different approaches with differing degrees of success. 
@@
 
-Recent partially interpretable models available in the literature have been based on the Koopman operator (e.g.,\cite{koopman1932dynamical,budivsic2012applied}) to  employ observables mapping the data onto a higher-dimensional space.
-This then makes the dynamics in the higher-dimensional space more amenable for approximation by a linear transformation.
-Such methods, obviously, do not guarantee an exact reconstructions for nonlinear models, and in practice provide poor long-term consistency for a large class of systems.
-\textbf{(Citation)}
+Recent partially interpretable models available in the literature have been based on the Koopman operator (e.g.,[?, ?]) to employ observables mapping the data onto a higher-dimensional space.
+This makes the dynamics in the higher-dimensional space more amenable for approximation by a linear transformation.
+Such methods do not guarantee exact reconstructions for nonlinear models, and in practice, provide poor long-term consistency for a large class of systems.
+(Citation)
 
 The non-interpretable models include the delay embedding and the machine learning algorithms.
-One could learn a system conjugate to the underlying system by applying the Takens delay embedding \cite{takens1981detecting} when one has ``good" observations from the system; this learnt system could then be used for forecast the observed data.
-Takens delay embedding theorem \cite{takens1981detecting} and its various generalisations (e.g., \cite{sauer1991embedology, stark1999delay, gutman2018embedding}) establishes the learnability of a system constructed by concatenating sufficiently large previous time-series observations of a dynamical system into a vector (called delay coordinates).
-This then establishes the existence of a map on the space of delay coordinates equivalent (or topologically conjugate) to the underlying map from which the observed time-series was first obtained.
-Although topological conjugacy guarantees an alternate representation of the underlying system, the quality of this representation still depends on numerous parameters, making the comprehension of the dynamics rather unreliable (e.g., \cite{principe1992prediction}).
-One reason for this fragility is that the embedded attractor in the reconstruction space not always an attractor of the map that is learnt in the reconstruction space, despite  unquestionably being an invariant set.
-When the embedded object is not explicitly known to be an attractor (as explained in Chapter \ref{ch3}) it can cause predictions to fail.
-
-Practically, the application of Takens embedding involves learning a map through some technique and consequently one wishes that these would have low functional complexity\cite{manjunath2021universal}, i.e., functions with fewer oscillatory graphs.
-Pure machine learning methodology that processes temporal information (like the echo state networks (Citations to echo state networks)) by mapping data onto a higher dimensional space for further processing.
-Although they perform  well on forecasting some dynamical systems, they fail completely on others (Citations here) as there is often no guarantee that the right function was learnt during training.
-
-This thesis deals with the implementation and analysis of non-interpretable models that can guarantee exact reconstruction.
-The project work concerns the study and implementation of a method (\cite{manjunath2021universal}) that incorporates learning a function by mapping the data on to a higher dimensional space using what is called a driven dynamical system (See Chapter 4).
+For example, one could learn a system conjugate to the underlying system by applying the Takens delay embedding [?] when one has "good" observations from the system; this learnt system could then be used to forecast the observed data.
+Takens delay embedding theorem [?] and its various generalisations (e.g., [?, ?, ?]) establishes the learnability of a system constructed by concatenating sufficiently large previous time-series observations of a dynamical system into a vector (called delay coordinates).
+This then confirms the existence of a map on the space of delay coordinates equivalent (or topologically conjugate) to the underlying map from which the observed time-series was first obtained.
+Although topological conjugacy guarantees an alternate representation of the underlying system, the quality still depends on numerous parameters, making the comprehension of the dynamics unreliable (e.g., [?]).
+One reason for this fragility is that the embedded attractor in the reconstruction space is not always an attractor of the map learnt in the reconstruction space, despite unquestionably being an invariant set.
+When the embedded object is not explicitly known to be an attractor (as explained in Chapter 3), it can cause predictions to fail.
+
+Practically, the application of Takens embedding involves learning a map through some technique, and consequently, one wishes that these would have low functional complexity[?], i.e., functions with fewer oscillatory graphs.
+On the other hand, pure machine learning methodology processes temporal information (like the echo state networks (Citations to echo state networks)) by mapping data onto a higher dimensional space for further processing.
+Although they perform well on forecasting some dynamical systems, they fail on others (Citations here) as there is often no guarantee that the correct function was learnt during training.
+
+This thesis involves implementing and analysing non-interpretable models that can guarantee exact reconstruction.
+The project work concerns the study and implementation of a method ([?]) that incorporates learning a function by mapping the data onto a higher dimensional space using what is called a driven dynamical system (See Chapter 4).
 With a clear understanding of how the data is mapped onto the higher dimensional space, the method then permits the learning of a dynamical system topologically conjugate to that of the underlying system.
@@
 The more intricate details of proofs are referred to where relevant throughout. This report is organised as follows:
-\newline In Chapter~\ref{ch2}, we recall the definition of a discrete-time dynamical system, how a discrete-system arises from a flow of an ODE and then proceed to define the inverse-limit space and topological conjugacy of autonomous systems.
-\newline In Chapter~\ref{ch3}, we introduce the problem of forecasting dynamical systems, state the Takens delay embedding theorem and discuss various issues faced while forecasting. 
-\newline In Chapter~\ref{ch4}, a driven dynamical system is defined and discuss the properties of a specific class of driven dynamical systems that we make use of in this thesis/report/paper. 
-\newline Finally in Chapter~\ref{ch5} we show the implementation of these forecasting methods, and conclusions are provided in Chapter~\ref{ch6}.
+\newline In Chapter 2, we recall the definition of a discrete-time dynamical system, how a discrete-system arises from a flow of an ODE, and then proceed to define the inverse-limit space and topological conjugacy of autonomous systems.
+\newline Chapter 3 introduces the problem of forecasting dynamical systems, states the Takens delay embedding theorem, and discusses various issues faced while forecasting.
+\newline In Chapter 4, a driven dynamical system is defined and discusses the properties of a specific class of driven dynamical systems that we use in this thesis/report/paper.
+\newline Finally, Chapter 5 shows the implementation of these forecasting methods, and conclusions are provided in Chapter 6.
 
@@
 In ecology, discrete-time dynamical systems are widely used to model population growth.
-The model in this case is a function calculating the following generation's population given the population of the previous generation.
+The model, in this case, is a function calculating the following generation's population given the population of the previous generation.
 If we know the starting population, we may once again calculate the population at any time in the future.
@@
 
-\ednote{B: Does defining explicitly the domain and codomain of $\phi$ make it clear enough?
-M: Saying $\phi$ is then a semi-conjugacy between $T$ and $S$ could be ambiguous as it does not specify which system is on the top in the commutativity.
-So please edit accordingly.}
+\ednote{M: Saying $\phi$ is then a semi-conjugacy between $T$ and $S$ could be ambiguous as it does not specify which system is on the top in the commutativity.
+So please edit accordingly.
+B: Does defining explicitly the domain and codomain of $\phi$ make it clear enough?
+}
 When $S$ is conjugate to $T$, the dynamics of the two systems are in some way ''dynamically equivalent''.
@@
 In almost all cases we can at most insert probes into a system to obtain partial information by means of the measurements taken.
-Moreover, the process of taking a measurement itself introduces 2 additional aspects which complicate the problem: \ednote{B: Perhaps this edit makes more sense?
-M: Maybe replace difficulty by saying that we need to consider two aspects.
-}
+Moreover, the process of taking a measurement itself introduces 2 additional aspects which complicate the problem: \ednote{M: Maybe replace difficulty by saying that we need to consider two aspects.
+B: Perhaps this edit makes more sense?}
 \vspace{-8mm}
@@
 
-% The Takens embedding theorem does not guarantee global dissipativity.
-% \ednote{B:Links with~\ref{subs_LearnGamma} in discussion of dissipativity of $\Gamma$.
+The Takens embedding theorem does not guarantee global dissipativity.
+\ednote{B:Links with~\ref{subs_LearnGamma} in discussion of dissipativity of $\Gamma$.
 Sir, I'm not sure how to write this here - I think it a beneficial point to make so as to set the stage for the later discussion in ~\ref{subs_LearnGamma}, but I am unsure there as well.}
@@
 and we refer to the map $H$ as a \emph{causal embedding}. 
-\ednote{B: edited the sentence here}
+\ednote{B: edited here}
 
@@
 
-We implement a feedforward network, (a network with no memory, unlike the RNN)\ednote{B: Added so that FFN is in more context} required to learn $\Gamma$ in the \emph{Python} programming language and making use of the \emph{Scipy} library, alongside the \emph{Keras} package in the \emph{Tensorflow} library \ednote{B: Is this something you want me to write in?
+We implement a feedforward network, (a network with no memory, unlike the RNN)\ednote{B: So that FFN is in more context} required to learn $\Gamma$ in the \emph{Python} programming language and making use of the \emph{Scipy} library, alongside the \emph{Keras} package in the \emph{Tensorflow} library \ednote{B: Is this something you want me to write in?
 M: Note!
@@
 In so doing we then also guarantee a measure of stability.
-\ednote{B: Not satisfactory.
-Edit again.}
+/ednote{B: Not satisfactory.
+Edit.}
 

### 2022-01-17 draft_16Jan/forcomparison.tex
Grammarly changes copied in

--- 2022-01-16 forcomparison.tex
@@
 There are many difficulties that one encounters while modelling such systems:
-
 \vspace{-8mm}
@@
 Examples in the literature attempting to forecast data from such systems have tried many different approaches with differing degrees of success. 
-When we have complete access to states of the system, an ordinary differential equation can be obtained from data ([?, ?] and [?, ?]) wherein one could approximate the vector field by a library of functions to get interpretable models.
-Examples in the literature attempting to forecast data from such systems have tried a number of different approaches with differing degrees of success. 
 When we have complete access to states of the system, an ordinary differential equation can be obtained from data (\cite{brunton2016discovering, champion2019data} and \cite{small2002modeling,xu2006modeling}) wherein one could approximate the vector field by a library of functions to obtain interpretable models.
@@
 
-In this chapter, we provide a brief description of what a discrete-time dynamical system is and what it means for it to exhibit chaos.
-We refer to \cite{devaney2018introduction, de2013elements} for more details.
+This chapter briefly describes a discrete-time dynamical system and what it means to exhibit chaos. We refer to \cite{devaney2018introduction, de2013elements} for more details.
 
@@
 
-Formally, a function $T: U \to U$, where $U$ is some set is  a \emph{discrete-time dynamical system} and its iterates $\{u,Tu,T^2u,\ldots\}$, where $T^n$ denotes the $n$-fold composition of $T$ with itself, describe the evolution of an initial condition $u\in U$ (Note that we frequently drop the brackets and denote $T(u)$ by $Tu$ so as to simplify notations).
+Formally, a function $T : U \to U$, where U is some set is a discrete-time dynamical system and its iterates ${u, Tu, T^{2}{u},\ldots}$, where Tn denotes the n-fold composition of T with itself, describe the evolution of an initial condition $u \in U$ (Note that we frequently drop the brackets and represent $T(u)$ by $Tu$ to simplify notations).
 
@@
 
-Of course, discrete-time dynamical systems need not always arise through a differential equation, and once again we may consider the field of ecology where discrete dynamical systems are often directly derived or assumed.
-\textbf{Citation}
-There is a school of thought that advocates discrete-time dynamical systems to be more natural for modelling real-world observations than differential equations and we refer the interested reader to \cite{saber2010introduction}.
+Of course, discrete-time dynamical systems need not always arise through a differential equation.
+Once again, we may consider the field of ecology, where discrete dynamical systems are often directly derived or assumed.
+Citation
+There is a school of thought that advocates discrete-time dynamical systems to be more natural for modelling real-world observations than differential equations.
+We refer the interested reader to \cite{saber2010introduction}.
 
@@
 Two examples of invariant sets include a fixed point where $Tu=u$ for  $u\in U$, and a periodic orbit, i.e., a set of iterates $\{u,Tu, T^2u,\ldots,T^pu\}$, where $T^{p+1}u=u$ for some $p\in\mathbb{Z}$.
-The entire space $U$ could be also be invariant.
+The entire space $U$ could also be invariant.
 Consider for example the space $U=[0,1]$, where $Tu=4u(1-u)$ and then $U$ is invariant (as every $u\in{U}$ can be written as $u = 4x(1-x)$ for some $x\in{U}$).
@@
 
-However,  when the space $U$ is compact, all orbits approach an invariant set.
-This is since the set of limit points of the orbit can be shown to be invariant \cite{de2013elements}.
+However, when the space U is compact, all orbits approach an invariant set.
+It follows because one can show the set of limit points of the orbit to be invariant~\cite{de2013elements}.
 So, when $U$ is compact,  the $\omega$-limit set $\omega(u;T)$ of a point $u$ defined to the collection of limit points of the sequence $\{x,Tu,T^2u,\ldots\}$ is nonempty, and $\omega(u;T)$ is invariant.
@@
 
-Invariant sets have various properties. Invariant sets can be attracting or repelling depending on how orbits in its vicinity behave. 
-Recall the examples $Tu =u/2$ and $Tu=2u$ defined on $\mathbb{R}$ (where $\{0\}$ ``attracts" orbits) or the example, $Tu=u^2$ defined on $[0,1]$ (where $\{1\}$ ``repels" orbits).
+Invariant sets have various properties. Invariant sets can be attracting or repelling depending on how orbits in their vicinity behave. 
+Recall the examples Tu = u/2 and Tu = 2u defined on R (where {0} “attracts” orbits) or the example, Tu = u2 defined on [0,1] (where {1} “repels” orbits). 
 We are interested in attractive invariant sets since they capture the long-term dynamics as time increases. 
-In particular we are interested in those invariant sets named attractors.
+In particular, we are interested in those invariant sets named attractors.
 
@@
 For the system,  $Tu=1-|2u-1|$ on $[0,1]$, one may easily verify that the only attractor is the entire space $[0,1]$.
-This is since between any two points $u< v$ in $[0,1]$, and for any $a,b$ so that $u\le a < b \le v$, we can find an $n$ so that $T^n(a,b)=[0,1]$ (as would be explained later in this chapter).
-
-A dynamical system can have several attractors, and may also be contained in another attractor.
-For the example, $Tu =u^2$ on $[0,1]$, both $\{0\}$ and $[0,1]$ are attractors.
-It is a known result that for dynamical systems on a compact space there always exists an attractor.
+This follows from the fact that between any two points $u< v$ in $[0,1]$ and for any $a,b$ so that $u\le a < b \le v$, we can find an $n$ so that $T^n(a,b)=[0,1]$ (as would be explained later in this chapter).
+
+A dynamical system can have several attractors and may also be contained in another attractor.
+For the example, $Tu = ^u2$ on $[0,1]$, both $\{0\}$ and [0,1] are attractors.
+It is known that an attractor for a dynamical system always exists in a compact space.
+\textbf{(Cite).}
 
 The dynamics restricted to an invariant set can be complicated.
-For instance an invariant set could be just a single point or it could have an infinite set.
-If the invariant set is infinite then there is a possibility of complicated dynamics, and a particular well-studied phenomenon of such complexity gives rise to what is called  chaotic behaviour, and has been studied in detail over the last fifty years.
+For instance, an invariant set could be just a single point, or it could have an infinite set.
+If the invariant set is infinite, then complicated dynamics are possible.
+A particular well-studied phenomenon of such complexity gives rise to so-called chaotic behaviour, studied in detail over the past fifty years.
 
@@
 
-In the 1960s, a number of mathematicians and mathematically interested scientists independently discovered chaos in the mathematical sense.
-The meteorologist Edward Lorenz may have been the first to explain this phenomenon in his 1963 paper \cite{lorenz1963deterministic}.
-The notions of invariance, attractivity and chaos may also be described for continuous systems, and Lorenz's system comprised of a system of differential equations.
-The narrative of Lorenz's discovery of chaos, as well as the history of other forerunners in this subject, are fascinating.
-For those who are interested, we highly recommend James Gleick's book Chaos: The Making of a New Science \cite{gleick2008chaos}.
-It clearly illustrates these experiences, explains why chaos was such a startling and crucial mathematical and scientific discovery, and describes the underlying mathematical notions for non-specialists.
-
-There are several intricately different definitions of chaos available in the literature \textbf{(Cite)}, with each of them indicating some aspect of complexity.
-In practice, it is not possible to verify which definition or notion does the underlying dynamical system satisfies when only data is observed from it.
-We do, however, specifically recall the definition of chaos in the sense of Devaney \cite{devaney2018introduction,de2013elements} so as to understand some nuances behind the complexity.
+In the 1960s, several mathematicians and mathematically interested scientists independently discovered chaos in the mathematical sense.
+The meteorologist Edward Lorenz may have been the first to explain this phenomenon in his 1963 paper [?].
+The notions of invariance, attractivity, and chaos may also be described for continuous systems, and Lorenz's system comprised of a system of differential equations.
+The narrative of Lorenz's discovery of chaos and the history of other forerunners in this subject is fascinating.
+We highly recommend James Gleick's book Chaos: The Making of a New Science [?] for those interested.
+It clearly illustrates these experiences, explains why chaos was such a startling and crucial mathematical and scientific discovery and describes the underlying mathematical notions for non-specialists.
+
+Several intricately different definitions of chaos are available in the literature (Cite), each indicating some aspect of complexity.
+In practice, it is not possible to verify which definition or notion the underlying dynamical system satisfies when only data is observed from it.
+However, we specifically recall the definition of chaos in the sense of Devaney [?, ?] to understand some nuances behind the complexity.
 Devaney's definition of chaos has three requirements, with the first being the notion of sensitive dependence on initial conditions.
@@
 
-It is important to acquire a sense of this definition as it can easily be misunderstood.
-It is common misconception to interpret SDIC as two close points ($u$ and $v$) that eventually become separated by a distance $\delta$ under iteration by $T$.
-But this is not true.
-To fully comprehend all the subtleties of the concept, one needs to discuss it more thoroughly by considering each sentence with care and attention, whereafter one may examine how they fit together to convey the concept of sensitive dependence.
+It is vital to acquire a sense of this definition as it can easily be misunderstood.
+It is a common misconception to interpret SDIC as two close points (u and v) that eventually become separated by a distance $\delta$ under iteration by T, but this is not true.
+To fully comprehend all the subtleties of the concept, one needs to discuss it more thoroughly by considering each sentence with care and attention; whereafter, one may examine how they fit together to convey the idea of sensitive dependence.
 
@@
   \item Second, in every neighborhood of $u$, we may not necessarily find all points $v$ in the neighborhood distinct from $u$ that would separate from the forward iterates of $u$. 
-  \item  Finally, $N$ depends upon $u$ and $v$ chosen, and their iterates may not separate for ever (i.e., for all $n>N$) and we allow their iterates to get arbitrarily close in the future. 
+  \item  Finally, $N$ depends upon $u$ and $v$ chosen, and their iterates may not separate forever (i.e., for all $n>N$) and we allow their iterates to get arbitrarily close in the future. 
 \end{enumerate}
@@
 
-\begin{Example}
-  The logistic map(LM), a recursion relation of the form $x_{n+1}=rx_n(1-x_n)$ where $x_n\in[0,1]$ and $r<4$ is of particular interest to us as it is classically used to illustrate the influence of chaos on a system \textbf{(Cite)}.
-Below are plotted the first 75 iterates of the LM for $r=3.7$ and $x_0=0.2$ with an adjustment of 0.001 in $x_0$ and $r$ respectively.
-
-  \begin{figure}[ht]
-    \includegraphics[scale=0.45]{_log_sdic.eps}
-        \centering
+\begin{Example} \rm
+  The logistic map(LM), a recursion relation of the form $x_{n+1}=rx_n(1-x_n)$ where $x_n\in[0,1]$ and $r\leq{4}$ is of particular interest to us as it is classically used to illustrate the influence of chaos on a system \textbf{(Cite)}.
+  In the graph below~\ref{fig:log_sdic} we plot the first 50 iterates of the LM for $r=4$ and two slightly different initial values $x_0$.
+
+  \begin{figure}[ht]
+    \includegraphics[scale=0.74]{_logistic_sdic.eps}
+        \centering
+        \captionof{figure}{Figure generated for Logistic Map $x_{n+1}=rx_n(1-x_n)$ with $r=4$ to exhibit the presence of SDIC.
+Plotted are the values $x_n$ against time $n$ for the the first 50 step with initial values $x_0=0.2$ (blue) and $x_0=0.20001$ (red).
+Initially the two trajectories overlap, but they diverge completely at $n=12$ whereafter they follow distinct paths.}
         \label{fig:log_sdic}
-        \captionof{figure}{Figure generated for Logistic Map for $r=3.7$, $x_0=0.2$ where each parameter is varied separately by 0.001.}
-      \end{figure}
-
-  For an intuitive sense of the fact that the LM exhibits SDIC, we draw two green neighbourhoods of diameters $\delta_1=0.1, \delta_2=0.02$ around the points $u_1=0.7, u_2=0.4$ respectively.
-It is easy to see that every neighbourhood around these two points will contain some $v_1, v_2\in[0,1]$ and natural number $N_1, N_2 \in\mathbb{N}$ respectively s.t.
-the distance $d(T^Nu_1,T^Nv)>\delta_1=0.02$ and $d(T^Nu_2,T^Nv)>\delta_2=0.1$.
-
-    \ednote{B: Is it desirable to have a more rigourous explanation here? }
-
+      \end{figure}
+  \ednote{B: Caption and figure adjusted}
+  % For an intuitive sense of the fact that the LM exhibits SDIC, consider again the definition and substitute specific values.
+Let $\delta=0.05$ and for purposes of illustration consider the value $u=0.4$.
+The green band in~\ref{fig:log_sdic} shades the region $[u-\delta, u+\delta]$ and we can see that regardless the value $v$  and in every neighborhood of $u \in U$ there exists a $v\in{U}$ and an integer $N:=N{(u,v)}\in\mathbb{Z}$ such that $d(T^Nu,T^Nv)>\delta$
+  % specific point, say $u=0.4$.
+Draw a green neighbourhood $[0.35,0.45]$ of diameters $\delta=0.02$ around the point.
+It is easy to see that every neighbourhood around $u=0.4$ will contain some $v_1, v_2\in[0,1]$ and natural number $N_1, N_2 \in\mathbb{N}$ respectively s.t.
+the distance $d(T^Nu_1,T^Nv)>\delta=0.05$ .
 \end{Example}
 
-\begin{Example}
-  A second example pertains to a simple physical system called the double pendulum - a pendulum with another pendulum attached to its head given an initial position and angular velociy.
-
-  Below are two figures denoting the trajectories of a double pendulum's second head after some elapsed time.
-Depicted are three pendulum heads with equal angular velocity, but differing ever so slightly in initial position .
+\begin{Example} \rm
+  A second example pertains to a simple physical system called the double pendulum - a pendulum with another pendulum attached to its head given an initial position and angular velocity.
+
+  Below are two figures denoting trajectories of a double pendulum's second head after some elapsed time.
+Depicted are three pendulum heads with equal angular velocity but differing slightly in initial position.
 As can be seen below, the trajectories diverge very quickly to become completely different.
-  The double pendulum will be discussed in much greater detail in a later chapter (\ref{ch5}), but we mention the example here to provide a numerical example of a practical system exhibiting SDIC.
+We discuss the double pendulum in greater detail in a later chapter (\ref{ch5}), but we mention the example here to provide a numerical example of a practical system exhibiting SDIC.
 
@@
 
-Topologically transitivity implies that the iterates of an open set of initial conditions gets mixed up with other open sets.
-On a compact metric space, it can be shown that topological transitivity also implies the existence of a point whose forward iterates are dense \cite{de2013elements}; or in other words, the orbit going through this point will be dense in the compact metric space.
+Topological transitivity implies that iterates of an open set of initial conditions get mixed up with other open sets.
+On a compact metric space, one may show that topological transitivity also implies the existence of a point whose forward iterates are dense [?]; or in other words, the orbit going through this point will be dense in the compact metric space.
 In fact, it is topologically more likely that the choice of an arbitrary point will be one whose iterates are almost dense.
@@
 To be more precise (see \cite{de2013elements}), for a discrete-time dynamical system on a compact space, the set of points with dense iterates are residual, and they are typical or likely to be observed in practice.
-In this thesis, we assume that when data is observed from a topologically transitive system, we assume that the data arises from a dense orbit.
+In this thesis, when data is observed from a topologically transitive system, we assume it arises from a dense orbit.
 
@@
 
-\begin{Example} 
+\begin{Example} \rm
   The standard tent map $Tu=1-|2u-1|$ defined on $[0,1]$ is a well-known example of a dynamical system satisfying these three properties.
@@
 
-To show topological similarity or sameness between two metric or topological spaces, one needs to establish a homeomorphism between the two spaces. 
+To show topological similarity (or sameness) between two metric or topological spaces, one needs to establish a homeomorphism between the two spaces. 
 However, in the study of dynamical systems defined on two spaces, establishing a homeomorphism does not indicate that the systems are dynamically related in any way. 
 For instance, the maps $Tu=u^2$ and $Tu=1-|2u-1|$ defined on $[0,1]$ have totally different behaviour. 
-To find systems that are dynamically similar, one rather needs to establish a kind of dynamical equivalence which we illustrate in a simple commutativity diagram below.
+Therefore, to find dynamically similar systems, one must establish a dynamical equivalence that we illustrate in a simple commutativity diagram below~\ref{eqn_conjugacy}.
 
@@
 %Composite functions
-To understand the diagram we note that if we travel ``right and then down," the diagram instructs us to use $T$ (top arrow right) first, followed by $\phi$.
+To understand the diagram, we note that if we travel "right and then down," the diagram instructs us to use T (top arrow right) first, followed by $\phi$.
 (right arrow downwards).
-Consequently the pathway amounts to finding $\phi(T(u))$.
+Consequently, the pathway amounts to finding $\phi(T(u))$.
 If we proceed "down and then right", the diagram instructs us to apply $\phi$ (left arrow downwards) first, and then apply $S$ (bottom arrow right).
@@
 
-\ednote{M: Saying $\phi$ is then a semi-conjugacy between $T$ and $S$ could be ambiguous as it does not specify which system is on the top in the commutativity.
-So please edit accordingly.
-B: Does defining explicitly the domain and codomain of $\phi$ make it clear enough?
-}
+\ednote{B: Does defining explicitly the domain and codomain of $\phi$ make it clear enough?
+M: Saying $\phi$ is then a semi-conjugacy between $T$ and $S$ could be ambiguous as it does not specify which system is on the top in the commutativity.
+So please edit accordingly.}
 When $S$ is conjugate to $T$, the dynamics of the two systems are in some way ''dynamically equivalent''.
 Specifically, they are in one-to-one correspondence with one another.
-However when $S$ is semi-conjugate to $T$ with $\phi:U\to{V}$ a many-to-one mapping, the dynamics on $V$ provide merely a coarse-grained description of the dynamics on $U$ \cite{de2013elements}.
+However when $S$ is semi-conjugate to $T$ with $\phi:U\to{V}$ a many-to-one mapping, the dynamics on $V$ provide merely a coarse-grained description of the dynamics on $U$~\cite{de2013elements}.
 When $S$ is semi-conjugate to $T$, it is also common to call $S$ a \emph{factor of $T$}, or conversely that $T$ is an \emph{extension of $S$}.
-In essence, an extension (e.g., \cite{de2013elements}) is a larger system capturing all of the important dynamics of its factor.
+In essence, an extension (e.g., \cite{de2013elements}) is a more extensive system capturing all of the essential dynamics of its factor.
 
@@
 
-In establishing that two systems are conjugate (semi-conjugate) to one another, we have also shown that one may choose to work with one system as opposed to the second and still be guaranteed to obtain information on the latter.
-This is especially useful while forecasting the future evolution of dynamical systems.
+By establishing that two systems are conjugate (semi-conjugate) to one another, we have also shown that one may choose to work with one system instead of the second and still be guaranteed to obtain information on the latter.
+It is instrumental during the process of forecasting the future evolution of dynamical systems.
 
@@
 
-In this chapter we acquaint ourselves with the question of forecasting dynamical systems with unknown underlying structures, state and discuss the shortcomings of the Takens delay embedding theorem and discuss various issues faced while forecasting.
+In this chapter, we acquaint ourselves with the question of forecasting dynamical systems with unknown underlying structures, state and discuss the shortcomings of the Takens delay embedding theorem and discuss various issues faced while forecasting.
 
@@
 A practical example of this would be the subsequent scenario: given the time-sequential coordinates of an object moving in space, predict the future positions of that object.
-However we are very rarely, if at all, presented with a problem where the entire state information is available to us in the form $u_n$ at some time-step $n\in\mathbb{N}$.
-Consider then a more involved learning problem:
+However, we are very rarely, if at all, presented with a problem where the entire state information is available to us in the form un at some timestep $n\in\mathbb{N}$.
+Consider then a more complicated learning problem:
 
@@
 First we establish the method of Takens delay embedding. 
-In this method, one considers a discrete-time dynamical system defined on a 'nice' space (a smooth manifold)\ednote{B: Should this be further explained?} that can be obtained as time-$K$ map of a continuous time-dynamical system, concepts which will be formalised below.
-We make observations from such a system, i.e., we observe the evolution of $\theta(w_0)$ (where $w_0$ represents the initial state) i.e., we observe  a finite set of values of $u_n :=  \theta(w_n) = \theta(Tw_{n-1})$.
-The sequence $\{u_n\}$ represents a scalar time-series and intuitively $\theta$ may be thought to represent a probe inserted into a larger system which is itself only measuring/extracting a small part of the greater system state at time $t=n$.
-Consider for example a thermometer erected to measure the ambient temperature in a local village.
-This measurement function, the thermometer, is capturing only a single aspect, the temperature, of a much grander dynamical system entailing the present weather of the surrounding area.
-Even more than that though- it is measuring a miniscule part of the global weather system.
+In this method, one considers a discrete-time dynamical system defined on a 'nice' space (a smooth manifold)\ednote{B: Should this be further explained?} that can be obtained as time-$K$ map of a continuous time-dynamical system, concepts formalised below.
+We make observations from such a system, i.e., we observe the evolution of $\theta(w_0)$ (where $w_0$ represents the initial state) i.e.\ednote{two i.e.'s}, we examine  a finite set of values of $u_n :=  \theta(w_n) = \theta(Tw_{n-1})$.
+The sequence $\{u_n\}$ represents a scalar time-series and intuitively $\theta$ represents a probe inserted into a bigger system which is itself only measuring/extracting a small part of the greater system state at time $t=n$.
+Consider, for example, a thermometer erected to measure the ambient temperature in a local village.
+This measurement function, the thermometer, is capturing only a single aspect, the temperature of a much grander dynamical system entailing the current weather of the surrounding area.
+Even more than that, though- it is measuring a miniscule part of the global weather system.
 
@@
 
-However, the true state $u$ of a system is seldom, if ever, fully known.
-In almost all cases we can at most insert probes into a system to obtain partial information by means of the measurements taken.
-Moreover, the process of taking a measurement itself introduces 2 additional aspects which complicate the problem: \ednote{M: Maybe replace difficulty by saying that we need to consider two aspects.
+However, the actual state $u$ of a system is seldom, if ever, fully known.
+In almost all cases, the most one can do is to insert probes into a system to obtain partial information through the measurements taken.
+Moreover, the process of taking a measurement itself introduces two additional aspects which complicate the problem: \ednote{M: Maybe replace difficulty by saying that we need to consider two aspects.
 B: Perhaps this edit makes more sense?}
@@
 \begin{enumerate}[noitemsep, label=\roman*.]
-  \item A series of measurements over a specific time-interval is inherently a discretisation procedure of the underlying continuous-time dynamical system.
+  \item A series of measurements over a specific time interval is inherently a discretisation procedure of the underlying continuous-time dynamical system.
 (Hence why we restricted our attention to discrete-time systems in the preceding section)
-  \item A probe will never be fully accurate, and so the act of measurement introduces a certain measure of numerical noise/inaccuracy.
+  \item A probe will never be entirely accurate, and so the act of measurement introduces a certain measure of numerical noise/inaccuracy.
 \end{enumerate}
@@
 $\Phi_{k,\theta}(w) := (\theta(T^{-k}w)\ldots,\theta(T^{-1}w),\theta(w))$.  
-The concept of an observable is understood in the sense of an observable as originally introduced by \cite{takens1981detecting, genericObservableAeyels} to refer to a probe or measurement function inserted into the system.
-The essence of Takens' theorem relates to the fact that when $k$ is sufficiently large, we can define a dynamical system on the space $\mathbb{R}^{k+1}$ whose states are $\Phi_{k,\theta}(w)$, $\Phi_{k,\theta}(Tw)$, $\Phi_{k,\theta}(T^2w)$, etc.
-and this dynamical system is topologically conjugate to the unknown underlying system $(W,T)$.
+The concept of an observable is understood in the sense of an observable as originally introduced by~\cite{takens1981detecting, genericObservableAeyels} to refer to a probe or measurement function inserted into the system.
+Takens' theorem, in essence,  refers to the fact that for a sufficiently large $k$, we can define a dynamical system on the space $\mathbb{R}^{k+1}$ with states $\Phi_{k,\theta}(w)$, $\Phi_{k,\theta}(Tw)$, $\Phi_{k,\theta}(T^2w)$, etc.
+This dynamical system is topologically conjugate to the unknown underlying system $(W,T)$.
 We recall the Takens delay embedding theorem next.
@@
 
-It is a fact that Takens' Embedding Theorem is both a powerful result and that it provides compelling reason to believe that one could conceivably reconstruct accurately a system conjugate to the underlying system.
-Nonetheless, it does present some serious practical limitations:
+Takens' Embedding Theorem is a powerful result, and it provides compelling reason to believe that one could conceivably reconstruct a system conjugate accurately to the underlying system.
+Nonetheless, it does present some severe practical limitations:
 \vspace{-5mm}
@@
 
-The Takens embedding theorem does not guarantee global dissipativity.
-\ednote{B:Links with~\ref{subs_LearnGamma} in discussion of dissipativity of $\Gamma$.
+% The Takens embedding theorem does not guarantee global dissipativity.
+% \ednote{B:Links with~\ref{subs_LearnGamma} in discussion of dissipativity of $\Gamma$.
 Sir, I'm not sure how to write this here - I think it a beneficial point to make so as to set the stage for the later discussion in ~\ref{subs_LearnGamma}, but I am unsure there as well.}
 
-Though not a uniquely related to the Takens embedding theorem, we do opt to make mention of this requirement for our work. 
+Though not a uniquely related to the Takens embedding theorem, we do opt to make mention of another requirement for our work. 
 The methods described here are applicable for data originating from a surjective map. 
-This follows from the fact that when $T$ maps the space $U$ into a proper subset $B$ of $U$, we me conclude that $T$ is not surjective as there exists some point in $U-A$ which does not have a preimage.
-An example of this would be the case when a system displays energy-loss.
-Consider, for instance, consider the oscillations dying down for a damped pendulum.
-(A more in-depth account will be given in \ref{ch5})
+This follows from the fact that when T maps the space U into a proper subset B of U, there exists some point in $U - A$ that does not have a preimage and we conclude that T is not surjective. 
+An example of this would be the case when a system displays energy-loss. Consider, for instance, the dying oscillations for a damped pendulum. (A more in-depth account will be provided in 5)
 
@@
 
-In this chapter, we discuss results pertaining to the mapping of temporal data obtained from a discrete-dynamical system onto a different space through the notion of a driven dynamical system. 
+In this chapter, we discuss results on the mapping of temporal data obtained from a discrete-dynamical system onto a different space through the notion of a driven dynamical system. 
 We also consider the conditions a driven system should possess to avoid adding distortion to its state-space representation.
-We then describe the \emph{single-delay dynamics} (SDD) of the system and and finally recall conditions on the driven system so that the SDD are conjugate (or at least semi-conjugate) to the underlying system.
+We then describe the single-delay Dynamics (SDD) of the system and finally recall conditions on the driven system so that the SDD are conjugate (or at least semi-conjugate) to the underlying system.
 The SDD can then be used to forecast and reconstruct the underlying system.
@@
 
-To find a function $\phi$, the need now arises to consider a so-called driven dynamical system, one that can generate a  nonautonomous dynamical system, where two spaces are considered: input is taken from an input space $U$ and the state of a space $X$ is updated.
+To find a function $\phi$, the need  now arises to consider a so-called driven dynamical system, a special case of a nonautonomous dynamical system.
+Two spaces are considered: input is taken from an input space U and the state of a space X is updated.
 (Previously, only the state was considered.)
@@
 
-The dynamics on X are generated by the update equation $x_{n+1}=g(u_n, x_n)$ where $n\in\mathbb{Z}$, input $u_n$ from $U$ and state $x_n$ belonging to $X$, where $X$ is compact.
-Abbreviated, we shall refer only to the \emph{driven system $g$}, with all other entities being understood implicitly.
-
-Notably, a nonautonomous dynamical system may be generated from $U$; any input $\overline{u}$, a bi-infinite sequence from $U$, gives rise to the sequence of self-maps $\{g(u_n, \cdot)\}_{n\in\mathbb{Z}}$ contained in $X$.
-Physically, one may think of this bi-infinite sequence as referring to a system that has been running for a long period at the time of the first measurement taken from the system (alternatively, the first time a probe is inserted into the system to take an observation)
+The update equation $x_{n+1} = g(u_n,x_n)$ for $n \in\mathbb{Z}$, input $u_n$ from $U$ and state $x_n$ belonging to $X$ (a compact space) generate the dynamics on X. 
+Abbreviated, we refer only to the \textit{driven system g} where all other entities are understood implicitly.
+
+Notably, a nonautonomous dynamical system can be generated from $U$; any input $\overline{u}$, a bi-infinite sequence from $U$, gives rise to the sequence of self-maps $\{g(u_n, \cdot)\}_{n\in\mathbb{Z}}$ contained in $X$.
+Physically, one may think of this bi-infinite sequence as referring to a system that has been running for an extended period at the time of the first measurement taken from the system (alternatively, the first time a probe is inserted into the system to take an observation)
 
@@
 
-A system may also have multiple solutions as is evidenced in the example below.
+A system may also have multiple solutions, as evidenced in the example below.
 
@@
 
-As the solutions to a driven system are often considered to be an important entity, we next identify a subspace $X_U$ of $X$ that contains all possible solutions.
-To realize such a subspace of a driven system $g$, the concept of a reachable set is defined.
+As the solutions to a driven system are often considered an important entity, we next identify a subspace $X_U$ of $X$ that contains all possible solutions.
+To realise such a subspace of a driven system $g$, the concept of a reachable set is defined.
 
@@
 
-The reachable set itself can defined independently from the fact that $g$ does or does not have the ESP/USP.
+The reachable set itself can defined independently from $g$ possessing the ESP/USP.
 We note that $x\in{X_n(\overline{u})}$ if and only $g$ has a solution $\{x_k\}$ for $x_n=x$ and input $\overline{u}$.
@@
 
-The set $X_n(\overline{u})$ is exactly the set of points $x$ through which some entire solution $\Psi$ passes at time $n$. We can give an alternate formulation of the set $X_n(\overline{u})$. 
+The set $X_n(\overline{u})$ is precisely the set of points $x$ through which some entire solution $\Psi$ passes at time $n$. We can give an alternate formulation of the set $X_n(\overline{u})$. 
 Now suppose, we for some fixed input 
@@
 
-Thus far it has been demonstrated that a system may have one or more solutions; one may ask if a driven system always has a solution and, if so, whether it satisfies certain properties such as uniqueness.
-Should the driven system be compact, existence follows immediately as shown in the following result.
+Thus far, it has been demonstrated that a system may have one or more solutions; one may ask if a driven system always has a solution and, if so, whether it satisfies specific properties such as uniqueness.
+Should the driven system be compact, existence follows immediately, as shown in the following result.
 
@@
 
-One of the first result obtained after defining the USP is the fact that every solution will attract different initial conditions towards the component parts of the solution.
-%, i.e.
-that the solution $\Psi$ is a non-autonomous uniform attractor.
-\ednote{Define Uniform Attractor before stating theorem or else omit the statement.}
+One of the first results obtained after defining the USP is that every solution will attract different initial conditions towards the component parts of the solution. 
 If $g$ has the USP, then any solution to $g$ is also a uniform attractor in nonautonomous dynamical systems literature \cite{Manju_Nonlinearity}.
-The discussion on nonautonomous attractors is beyond the scope of this project and we refer the reader to \cite{Manju_ESP, esann2012ids} for further reading.
+The discussion on nonautonomous attractors is beyond the scope of this project, and we refer the reader to \cite{Manju_ESP, esann2012ids} for further reading.
 
@@
 
-The goal now assumes the form of the question on our ability to establish a semi-conjugacy as presented below for the driven system $g$.
+The question now assumes the form: Can we establish a semi-conjugacy for the driven system $g$ as presented below.
 
@@
 
-SI-invertibility means that one's choice of $g$ will still ensure 'enough' information is retained without introducing unwanted complexity. 
+SI-invertibility promises that 'enough' information is retained when $g$ is chosen without introducing one's choice of $g$ without introducing unwanted complexity. 
 'Enough' here refers to the fact that we may always recover the previous input value $u_n$ (if we know successive states $x_{n-1}, x_n$).
@@
 /ednote{B: Not satisfactory.
-Edit.}
+Edit again.}
 

### 2022-01-17 draft_16Jan/hiscurrent.tex
Grammarly changes copied in

--- 2022-01-17 draft_16Jan/forcomparison.tex
@@
 (Citation)
+\textbf{(Citation)}
 
 The non-interpretable models include the delay embedding and the machine learning algorithms.
-For example, one could learn a system conjugate to the underlying system by applying the Takens delay embedding [?] when one has "good" observations from the system; this learnt system could then be used to forecast the observed data.
-Takens delay embedding theorem [?] and its various generalisations (e.g., [?, ?, ?]) establishes the learnability of a system constructed by concatenating sufficiently large previous time-series observations of a dynamical system into a vector (called delay coordinates).
+For example, one could learn a system conjugate to the underlying system by applying the Takens delay embedding~\cite{takens1981detecting} when one has "good" observations from the system; this learnt system could then be used to forecast the observed data.
+Takens delay embedding theorem ~\cite{takens1981detecting} and its various generalisations (e.g., \cite{sauer1991embedology, stark1999delay, gutman2018embedding}) establishes the learnability of a system constructed by concatenating sufficiently large previous time-series observations of a dynamical system into a vector (called delay coordinates).
 This then confirms the existence of a map on the space of delay coordinates equivalent (or topologically conjugate) to the underlying map from which the observed time-series was first obtained.
@@
 One reason for this fragility is that the embedded attractor in the reconstruction space is not always an attractor of the map learnt in the reconstruction space, despite unquestionably being an invariant set.
-When the embedded object is not explicitly known to be an attractor (as explained in Chapter 3), it can cause predictions to fail.
-
-Practically, the application of Takens embedding involves learning a map through some technique, and consequently, one wishes that these would have low functional complexity[?], i.e., functions with fewer oscillatory graphs.
+When the embedded object is not explicitly known to be an attractor (as explained in Chapter~ \ref{ch3}), it can cause predictions to fail.
+
+Practically, the application of Takens embedding involves learning a map through some technique, and consequently one wishes that these would have low functional complexity\cite{manjunath2021universal}, i.e., functions with fewer oscillatory graphs.
 On the other hand, pure machine learning methodology processes temporal information (like the echo state networks (Citations to echo state networks)) by mapping data onto a higher dimensional space for further processing.
-Although they perform well on forecasting some dynamical systems, they fail on others (Citations here) as there is often no guarantee that the correct function was learnt during training.
+Although they perform  well on forecasting some dynamical systems, they fail completely on others (Citations here) as there is often no guarantee that the right function was learnt during training.
 
 This thesis involves implementing and analysing non-interpretable models that can guarantee exact reconstruction. 
-The project work concerns the study and implementation of a method ([?]) that incorporates learning a function by mapping the data onto a higher dimensional space using what is called a driven dynamical system (See Chapter 4).
+The project work concerns the study and implementation of a method (\cite{manjunath2021universal}) that incorporates learning a function by mapping the data on to a higher dimensional space using what is called a driven dynamical system (See Chapter~ref{ch.4}).
 With a clear understanding of how the data is mapped onto the higher dimensional space, the method then permits the learning of a dynamical system topologically conjugate to that of the underlying system.
@@
 The more intricate details of proofs are referred to where relevant throughout. This report is organised as follows: 
-\newline In Chapter 2, we recall the definition of a discrete-time dynamical system, how a discrete-system arises from a flow of an ODE, and then proceed to define the inverse-limit space and topological conjugacy of autonomous systems.
-\newline Chapter 3 introduces the problem of forecasting dynamical systems, states the Takens delay embedding theorem, and discusses various issues faced while forecasting.
-\newline In Chapter 4, a driven dynamical system is defined and discusses the properties of a specific class of driven dynamical systems that we use in this thesis/report/paper.
-\newline Finally, Chapter 5 shows the implementation of these forecasting methods, and conclusions are provided in Chapter 6.
+\newline In Chapter~\ref{ch2}, we recall the definition of a discrete-time dynamical system, how a discrete-system arises from a flow of an ODE and then proceed to define the inverse-limit space and topological conjugacy of autonomous systems.
+\newline Chapter~\ref{ch3} introduces the problem of forecasting dynamical systems, states the Takens delay embedding theorem, and discusses various issues faced while forecasting. 
+\newline In Chapter~\ref{ch4}, a driven dynamical system is defined and discuss the properties of a specific class of driven dynamical systems that we make use of in this thesis/report/paper. 
+\newline Finally, in Chapter~\ref{ch5} we show the implementation of these forecasting methods, and conclusions are provided in Chapter~\ref{ch6}.
 
@@
 In ecology, discrete-time dynamical systems are widely used to model population growth.
-The model, in this case, is a function calculating the following generation's population given the population of the previous generation.
+The model in this case is a function calculating the following generation's population given the population of the previous generation.
 If we know the starting population, we may once again calculate the population at any time in the future.
 
-Formally, a function $T : U \to U$, where U is some set is a discrete-time dynamical system and its iterates ${u, Tu, T^{2}{u},\ldots}$, where Tn denotes the n-fold composition of T with itself, describe the evolution of an initial condition $u \in U$ (Note that we frequently drop the brackets and represent $T(u)$ by $Tu$ to simplify notations).
+Formally, a function $T: U \to U$, where $U$ is some set is  a \emph{discrete-time dynamical system} and its iterates $\{u,Tu,T^2u,\ldots\}$, where $T^n$ denotes the $n$-fold composition of $T$ with itself, describe the evolution of an initial condition $u\in U$ (Note that we frequently drop the brackets and denote $T(u)$ by $Tu$ so as to simplify notations).
 
@@
 Once again, we may consider the field of ecology, where discrete dynamical systems are often directly derived or assumed.
-Citation
+\textbf{Citation}
 There is a school of thought that advocates discrete-time dynamical systems to be more natural for modelling real-world observations than differential equations.
@@
 Two examples of invariant sets include a fixed point where $Tu=u$ for  $u\in U$, and a periodic orbit, i.e., a set of iterates $\{u,Tu, T^2u,\ldots,T^pu\}$, where $T^{p+1}u=u$ for some $p\in\mathbb{Z}$.
-The entire space $U$ could also be invariant.
+The entire space $U$ could be also be invariant.
 Consider for example the space $U=[0,1]$, where $Tu=4u(1-u)$ and then $U$ is invariant (as every $u\in{U}$ can be written as $u = 4x(1-x)$ for some $x\in{U}$).
@@
 
-However, when the space U is compact, all orbits approach an invariant set.
-It follows because one can show the set of limit points of the orbit to be invariant~\cite{de2013elements}.
+However,  when the space $U$ is compact, all orbits approach an invariant set.
+This is since the set of limit points of the orbit can be shown to be invariant \cite{de2013elements}.
 So, when $U$ is compact,  the $\omega$-limit set $\omega(u;T)$ of a point $u$ defined to the collection of limit points of the sequence $\{x,Tu,T^2u,\ldots\}$ is nonempty, and $\omega(u;T)$ is invariant.
@@
 Invariant sets have various properties. Invariant sets can be attracting or repelling depending on how orbits in their vicinity behave. 
-Recall the examples Tu = u/2 and Tu = 2u defined on R (where {0} “attracts” orbits) or the example, Tu = u2 defined on [0,1] (where {1} “repels” orbits). 
+Recall the examples $Tu =u/2$ and $Tu=2u$ defined on $\mathbb{R}$ (where $\{0\}$ “attracts" orbits) or the example, $Tu=u^2$ defined on $[0,1]$ (where $\{1\}$ “repels" orbits). 
 We are interested in attractive invariant sets since they capture the long-term dynamics as time increases. 
@@
 For the system,  $Tu=1-|2u-1|$ on $[0,1]$, one may easily verify that the only attractor is the entire space $[0,1]$.
-This follows from the fact that between any two points $u< v$ in $[0,1]$ and for any $a,b$ so that $u\le a < b \le v$, we can find an $n$ so that $T^n(a,b)=[0,1]$ (as would be explained later in this chapter).
+This follows from the fact that between any two points $u< v$ in $[0,1]$, and for any $a,b$ so that $u\le a < b \le v$, we can find an $n$ so that $T^n(a,b)=[0,1]$ (as would be explained later in this chapter).
 
 A dynamical system can have several attractors and may also be contained in another attractor.
-For the example, $Tu = ^u2$ on $[0,1]$, both $\{0\}$ and [0,1] are attractors.
+For the example, $Tu =u^2$ on $[0,1]$, both $\{0\}$ and $[0,1]$ are attractors.
 It is known that an attractor for a dynamical system always exists in a compact space.
@@
 The dynamics restricted to an invariant set can be complicated.
-For instance, an invariant set could be just a single point, or it could have an infinite set.
-If the invariant set is infinite, then complicated dynamics are possible.
-A particular well-studied phenomenon of such complexity gives rise to so-called chaotic behaviour, studied in detail over the past fifty years.
+For instance an invariant set could be just a single point or it could have an infinite set.
+If the invariant set is infinite, then complicated dynamics are possible.A particular well-studied phenomenon of such complexity gives rise to so-called  chaotic behaviour, studied in detail over the past fifty years.
 
@@
 In the 1960s, several mathematicians and mathematically interested scientists independently discovered chaos in the mathematical sense.
-The meteorologist Edward Lorenz may have been the first to explain this phenomenon in his 1963 paper [?].
+The meteorologist Edward Lorenz may have been the first to explain this phenomenon in his 1963 paper \cite{lorenz1963deterministic}.
 The notions of invariance, attractivity, and chaos may also be described for continuous systems, and Lorenz's system comprised of a system of differential equations.
 The narrative of Lorenz's discovery of chaos and the history of other forerunners in this subject is fascinating.
-We highly recommend James Gleick's book Chaos: The Making of a New Science [?] for those interested.
+We highly recommend James Gleick's book Chaos: The Making of a New Science~\cite{gleick2008chaos} for those interested.
 It clearly illustrates these experiences, explains why chaos was such a startling and crucial mathematical and scientific discovery and describes the underlying mathematical notions for non-specialists.
 
-Several intricately different definitions of chaos are available in the literature (Cite), each indicating some aspect of complexity.
-In practice, it is not possible to verify which definition or notion the underlying dynamical system satisfies when only data is observed from it.
-However, we specifically recall the definition of chaos in the sense of Devaney [?, ?] to understand some nuances behind the complexity.
+There are several intricately different definitions of chaos available in the literature \textbf{(Cite)}, with each of them indicating some aspect of complexity.
+In practice, it is not possible to verify which definition or notion does the underlying dynamical system satisfies when only data is observed from it.
+We do, however, specifically recall the definition of chaos in the sense of Devaney \cite{devaney2018introduction,de2013elements} so as to understand some nuances behind the complexity.
 Devaney's definition of chaos has three requirements, with the first being the notion  of sensitive dependence on initial conditions.
@@
 It is vital to acquire a sense of this definition as it can easily be misunderstood.
-It is a common misconception to interpret SDIC as two close points (u and v) that eventually become separated by a distance $\delta$ under iteration by T, but this is not true.
-To fully comprehend all the subtleties of the concept, one needs to discuss it more thoroughly by considering each sentence with care and attention; whereafter, one may examine how they fit together to convey the idea of sensitive dependence.
+It is a common misconception to interpret SDIC as two close points ($u$ and $v$) that eventually become separated by a distance $\delta$ under iteration by $T$.
+But this is not true.
+To fully comprehend all the subtleties of the concept, one needs to discuss it more thoroughly by considering each sentence with care and attention, whereafter one may examine how they fit together to convey the idea of sensitive dependence.
 
@@
 Topological transitivity implies that iterates of an open set of initial conditions get mixed up with other open sets.
-On a compact metric space, one may show that topological transitivity also implies the existence of a point whose forward iterates are dense [?]; or in other words, the orbit going through this point will be dense in the compact metric space.
+On a compact metric space, one may show that topological transitivity also implies the existence of a point whose forward iterates are dense \cite{de2013elements}; or in other words, the orbit going through this point will be dense in the compact metric space.
 In fact, it is topologically more likely that the choice of an arbitrary point will be one whose iterates are almost dense.
@@
 To be more precise (see \cite{de2013elements}), for a discrete-time dynamical system on a compact space, the set of points with dense iterates are residual, and they are typical or likely to be observed in practice.
-In this thesis, when data is observed from a topologically transitive system, we assume it arises from a dense orbit.
+In this project, when data is observed from a topologically transitive system, we assume it arises from a dense orbit.
 
@@
 %Composite functions
-To understand the diagram, we note that if we travel "right and then down," the diagram instructs us to use T (top arrow right) first, followed by $\phi$.
+To understand the diagram, we note that if we travel "right and then down," the diagram instructs us to use $T$ (top arrow right) first, followed by $\phi$.
 (right arrow downwards).
 Consequently, the pathway amounts to finding $\phi(T(u))$.
-If we proceed "down and then right", the diagram instructs us to apply $\phi$ (left arrow downwards) first, and then apply $S$ (bottom arrow right).
+If we proceed "down and then right," the diagram instructs us to apply $\phi$ (left arrow downwards) first, and then apply $S$ (bottom arrow right).
 This then amounts to finding  $S(\phi(u))$.
@@
 A practical example of this would be the subsequent scenario: given the time-sequential coordinates of an object moving in space, predict the future positions of that object.
-However, we are very rarely, if at all, presented with a problem where the entire state information is available to us in the form un at some timestep $n\in\mathbb{N}$.
+We are very rarely, if at all, presented with a problem where the entire state information is available to us in the form $u_n$ at some timestep $n\in\mathbb{N}$.
 Consider then a more complicated learning problem:
@@
 Consider, for example, a thermometer erected to measure the ambient temperature in a local village.
-This measurement function, the thermometer, is capturing only a single aspect, the temperature of a much grander dynamical system entailing the current weather of the surrounding area.
-Even more than that, though- it is measuring a miniscule part of the global weather system.
+This measurement function, the thermometer, is capturing only a single aspect, the temperature, of a much grander dynamical system entailing the current weather of the surrounding area.
+Even more than that though- it is measuring a miniscule part of the global weather system.
 
@@
 However, the actual state $u$ of a system is seldom, if ever, fully known.
-In almost all cases, the most one can do is to insert probes into a system to obtain partial information through the measurements taken.
-Moreover, the process of taking a measurement itself introduces two additional aspects which complicate the problem: \ednote{M: Maybe replace difficulty by saying that we need to consider two aspects.
-B: Perhaps this edit makes more sense?}
+the most one can do is to insert probes into a system to obtain partial information through the measurements taken.
+Moreover, the process of taking a measurement itself introduces two additional aspects which complicate the problem: \ednote{B: Perhaps this edit makes more sense?
+M: Maybe replace difficulty by saying that we need to consider two aspects.
+}
 \vspace{-8mm}
@@
 \section{Practical Limitations to Takens' Theorem}
-
 Takens' Embedding Theorem is a powerful result, and it provides compelling reason to believe that one could conceivably reconstruct a system conjugate accurately to the underlying system. 
@@
 The methods described here are applicable for data originating from a surjective map. 
-This follows from the fact that when T maps the space U into a proper subset B of U, there exists some point in $U - A$ that does not have a preimage and we conclude that T is not surjective. 
-An example of this would be the case when a system displays energy-loss. Consider, for instance, the dying oscillations for a damped pendulum. (A more in-depth account will be provided in 5)
+This follows from the fact that when $T$ maps the space $U$ into a proper subset $B$ of$ $U, there exists some point in $U - A$ that does not have a preimage and we conclude that $T$ is not surjective.
+An example of this would be the case when a system displays energy-loss.
+Consider, for instance, consider the dying oscillations for a damped pendulum.
+(An account of greater depth  is provided in~\ref{ch5})
 
@@
 We also consider the conditions a driven system should possess to avoid adding distortion to its state-space representation.
-We then describe the single-delay Dynamics (SDD) of the system and finally recall conditions on the driven system so that the SDD are conjugate (or at least semi-conjugate) to the underlying system.
+We then describe the \emph{single-delay dynamics} (SDD) of the system and finally recall conditions on the driven system so that the SDD are conjugate (or at least semi-conjugate) to the underlying system.
 The SDD can then be used to forecast and reconstruct the underlying system.
@@
 To find a function $\phi$, the need now arises to consider a so-called driven dynamical system, a special case of a nonautonomous dynamical system.
-Two spaces are considered: input is taken from an input space U and the state of a space X is updated.
+Two spaces are considered: input is taken from an input space $U$ and the state of a space $X$ is updated.
 (Previously, only the state was considered.)
@@
 The update equation $x_{n+1} = g(u_n,x_n)$ for $n \in\mathbb{Z}$, input $u_n$ from $U$ and state $x_n$ belonging to $X$ (a compact space) generate the dynamics on X. 
-Abbreviated, we refer only to the \textit{driven system g} where all other entities are understood implicitly.
+Abbreviated, we refer only to the \textit{driven system $g$}, where all other entities are understood implicitly.
 
@@
 
-As the solutions to a driven system are often considered an important entity, we next identify a subspace $X_U$ of $X$ that contains all possible solutions.
-To realise such a subspace of a driven system $g$, the concept of a reachable set is defined.
+As the solutions to a driven system are considered an important entity, we next identify a subspace $X_U$ of $X$ that contains all possible solutions.
+To realize such a subspace of a driven system $g$, the concept of a reachable set is defined.
 
@@
 One of the first results obtained after defining the USP is that every solution will attract different initial conditions towards the component parts of the solution.
-If $g$ has the USP, then any solution to $g$ is also a uniform attractor in nonautonomous dynamical systems literature \cite{Manju_Nonlinearity}.
+If $g$ has the USP, then any solution to $g$ is also a uniform attractor in nonautonomous dynamical systems literature~\cite{Manju_Nonlinearity}.
 The discussion on nonautonomous attractors is beyond the scope of this project, and we refer the reader to \cite{Manju_ESP, esann2012ids} for further reading.
@@
 and we refer to the map $H$ as a \emph{causal embedding}. 
-\ednote{B: edited here}
+\ednote{B: edited the sentence here}
 
@@
 
-SI-invertibility promises that 'enough' information is retained when $g$ is chosen without introducing one's choice of $g$ without introducing unwanted complexity. 
+SI-invertibility promises that 'enough' information is retained when $g$ is chosen without introducing one's choice of $g$ will still ensure 'enough' information is retained without introducing unwanted complexity.
 'Enough' here refers to the fact that we may always recover the previous input value $u_n$ (if we know successive states $x_{n-1}, x_n$).
@@
 
-We implement a feedforward network, (a network with no memory, unlike the RNN)\ednote{B: So that FFN is in more context} required to learn $\Gamma$ in the \emph{Python} programming language and making use of the \emph{Scipy} library, alongside the \emph{Keras} package in the \emph{Tensorflow} library \ednote{B: Is this something you want me to write in?
+We implement a feedforward network, (a network with no memory, unlike the RNN)\ednote{B: Added so that FFN is in more context} required to learn $\Gamma$ in the \emph{Python} programming language and making use of the \emph{Scipy} library, alongside the \emph{Keras} package in the \emph{Tensorflow} library \ednote{B: Is this something you want me to write in?
 M: Note!
@@
 In so doing we then also guarantee a measure of stability.
-/ednote{B: Not satisfactory.
+\ednote{B: Not satisfactory.
 Edit again.}

### 2022-01-17 graphing.py

@@ -1,28 +1,56 @@
-def T(u):
-    return 1 - abs(2*u-1)
+# def T(u):
+#     return 1 - abs(2*u-1)
 
-def Tn(n, U):
-    for i in range(n):
-        U = T(U)
-    return U
+# def Tn(n, U):
+#     for i in range(n):
+#         U = T(U)
+#     return U
 
-import numpy as np
-X1 = np.linspace(0,1,1200)
-Y= T(T(X1))
+# import numpy as np
+# X1 = np.linspace(0,1,1200)
+# Y= T(T(X1))
+
+# import matplotlib.pyplot as plt
+
+# # plt.plot(X1, Y, '-g')
+# # plt.title(r'Plot of the map $T^2$, $T=1-|2u-1|$')
+# # plt.xlabel(r'$x_n$')
+# # plt.ylabel(r'$x_{n+2}$')
+# # plt.savefig("T2.png")
+# # plt.show()
+
+# plt.plot(X1, Tn(2, X1), 'forestgreen')  #52=close, 53=line
+# plt.savefig('tentmap_2.svg', format='svg', dpi=1000)
+# plt.show()
+
+# # plt.plot(X3, T(X3), '-k')
+# # plt.show()
+
+
+def logistic(r, xn):
+    return r*xn*(1-xn)
+
+x0=0.2
+x0_p=0.20001
+r0=4 #3.7
+N=50
+
+Y1 = [x0]
+Y2= [x0_p]
+X = [i for i in range(N)]
+for i in range(1, N):
+    Y1.append(logistic(r0,Y1[i-1]))
+    Y2.append(logistic(r0,Y2[i-1]))
 
 import matplotlib.pyplot as plt
+plt.rcParams['figure.figsize'] = [10, 4]
+fig, ax = plt.figure(), plt.axes()
+plt.plot(X, Y1, 'darkblue', lw=0.75)
+plt.plot(X, Y2, 'orangered', lw=0.75)
 
-# plt.plot(X1, Y, '-g')
-# plt.title(r'Plot of the map $T^2$, $T=1-|2u-1|$')
-# plt.xlabel(r'$x_n$')
-# plt.ylabel(r'$x_{n+2}$')
-# plt.savefig("T2.png")
-# plt.show()
+ax.axhspan(0.4-0.05, 0.4+0.05, alpha=0.6, color='lightgreen')
+ax.set_xlabel(r"n", fontsize=12)
[…]

### 2022-01-17 draft_16Jan/draft7.tex
Changed abstract and conclusion

--- 2022-01-17 draft_16Jan/hiscurrent.tex
@@
 \begin{abstract}
-  Collecting measurements from underlying artificial, natural, and physical systems through observables to realise a dynamical model with greater descriptive power is a well-established practice.
-Physical systems can often be modelled by instance with differential equations derived from the laws of physics, if we knew the laws.
-However, we often do not fully understand the many complexities of the more intricate systems we interact with on a daily basis.
-  In this project we start with the hypothesis that an discrete-time dynamical systems, particularly those which are chaotic generates the data.
-The general learning problem is then formulated and a well-known approach to the problem, alongside its practical limitations are considered.
-  We then introduce the notion of driven dynamical systems and establish the theoretical existence of dynamical systems possessing certain properties that guarantee them to be topologically conjugate ('dynamically equivalent') to the dynamics of the underlying system from which measurements were taken.
-The methodology is practically implemented by making use of a recurrent neural network and using deep learning methods to train on the recurrent neural network states to show that we may obtain long-term consistent predictions for simple physical systems such as the double pendulum as well as a few other attractors exhibiting chaotic behaviour.
-The machine learning algorithms for modelling the double pendulum thus far have not been successful when it exhibits chaos.
+  It is a well-established practice to collect measurements from underlying artificial, natural, and physical systems through observables in order to realise a dynamical model with greater descriptive power.
+  Many data-driven approaches to forecast the future evolution of systems exist.
+However, many of these methods rely on an initial understanding of the system and we often do not fully understand the many complexities of more intricate systems that we interact with on a daily basis.
+
+  In this project, the base assumption is that data is originates from  some discrete-time dynamical system, particularly one of a chaotic nature but no assumptions are made pertaining to a specific dynamical system.
+  A general learning problem is formulated and a well-known approach to the problem, alongside its practical limitations, is then considered.
+  The notion of driven dynamical systems is introduced, and we subsequently establish the theoretical existence of a dynamical system possessing certain properties which guarantees it to be topologically conjugate (`dynamically equivalent`) to the dynamics of the underlying system from which measurements were originally taken.
+  Our methodology is implemented practically by making use of recurrent neural networks, a form of deep learning, to learn a map topologically equivalent to the underlying system that forecasts future states.
+  In so doing we present the workability of an approach to obtain long-term topological and statistically consistent predictions for simple physical systems as well as a some chaotic attractors.
   \end{abstract}
@@
 Practically, the application of Takens embedding involves learning a map through some technique, and consequently one wishes that these would have low functional complexity\cite{manjunath2021universal}, i.e., functions with fewer oscillatory graphs.
-On the other hand, pure machine learning methodology processes temporal information (like the echo state networks (Citations to echo state networks)) by mapping data onto a higher dimensional space for further processing.
+On the other hand, pure machine learning methodology processes temporal information (like the echo state networks \cite{jaeger2001echo, jaeger2004harnessing} by mapping data onto a higher dimensional space for further processing.
 Although they perform  well on forecasting some dynamical systems, they fail completely on others (Citations here) as there is often no guarantee that the right function was learnt during training.
@@
 With a clear understanding of how the data is mapped onto the higher dimensional space, the method then permits the learning of a dynamical system topologically conjugate to that of the underlying system.
-Instead of linear regression as in the training of echo state networks, deep learning methods are employed to learn the correct function.
+Instead of linear regression as in the training of echo state networks \cite{jaeger2001echo}, deep learning methods are employed to learn the correct function.
 With slight modifications to the implementation in the paper (with Adriaan), we show that one can construct accurate non-interpretable models with the ability to reconstruct attractors from more hard-to-forecast systems like the double pendulum.
@@
 Formally, a function $T: U \to U$, where $U$ is some set is  a \emph{discrete-time dynamical system} and its iterates $\{u,Tu,T^2u,\ldots\}$, where $T^n$ denotes the $n$-fold composition of $T$ with itself, describe the evolution of an initial condition $u\in U$ (Note that we frequently drop the brackets and denote $T(u)$ by $Tu$ so as to simplify notations).
+Also, we assume that $U$ is endowed with a metric $d$.
 
@@
 \begin{Example} \rm
-  The logistic map(LM), a recursion relation of the form $x_{n+1}=rx_n(1-x_n)$ where $x_n\in[0,1]$ and $r\leq{4}$ is of particular interest to us as it is classically used to illustrate the influence of chaos on a system \textbf{(Cite)}.
+  The logistic map (LM), a recursion relation of the form $x_{n+1}=rx_n(1-x_n)$ where $x_n\in[0,1]$ and $0<r\leq{4}$.
+It is of particular interest to us as it is classically used to illustrate chaos \cite{devaney2018introduction} for certain parameters.
   In the graph below~\ref{fig:log_sdic} we plot the first 50 iterates of the LM for $r=4$ and two slightly different initial values $x_0$.
@@
       \end{figure}
-  \ednote{B: Caption and figure adjusted}
+
   % For an intuitive sense of the fact that the LM exhibits SDIC, consider again the definition and substitute specific values.
@@
    Consider the dynamical systems $T:U\to{U}$, $S:V\to{V}$ and suppose the relationship $\phi \circ T=S\circ \phi$ holds.
-If $\phi:U\to{V}$ is a homeomorphism, then $S$ is said to be conjugate to $T$ and $\phi$ is called a conjugacy.
-If we relax the criterion on $\phi$ and merely require $\phi$ to be continuous and surjective where $\phi:U\to{V}$, then $\phi$ is a semi-conjugacy between $T$ and $S$ where T has domain $U$ and S has domain $V$; S is said to be semi-conjugate to T.
+If $\phi:U\to{V}$ is a homeomorphism, then $S$ is said to be conjugate to $T$ and $\phi$ is called a conjugacy with the underlying spaces.
+If we relax the criterion on $\phi$ and merely require $\phi$ to be continuous and surjective where $\phi:U\to{V}$, then $\phi$ is a semi-conjugacy between $T$ and $S$.
 \end{Definition}
@@
 M: Saying $\phi$ is then a semi-conjugacy between $T$ and $S$ could be ambiguous as it does not specify which system is on the top in the commutativity.
-So please edit accordingly.}
+So please edit accordingly.
+M: Very often they are implicitly understood...
+so no worries, and I have edited.}
 When $S$ is conjugate to $T$, the dynamics of the two systems are in some way ''dynamically equivalent''.
@@
 First we establish the method of Takens delay embedding. 
-In this method, one considers a discrete-time dynamical system defined on a 'nice' space (a smooth manifold)\ednote{B: Should this be further explained?} that can be obtained as time-$K$ map of a continuous time-dynamical system, concepts formalised below.
-We make observations from such a system, i.e., we observe the evolution of $\theta(w_0)$ (where $w_0$ represents the initial state) i.e.\ednote{two i.e.'s}, we examine  a finite set of values of $u_n :=  \theta(w_n) = \theta(Tw_{n-1})$.
+In this method, one considers a discrete-time dynamical system defined on a 'nice' space (a smooth manifold\footnote{A manifold is a special topological space that is locally homeomorphic at each point to an open set of $\mathbb{R}^d$} that can be obtained as time-$K$ map of a continuous time-dynamical system, concepts formalised below.
+We make observations from such a system, i.e., we observe the evolution of $\theta(w_0)$ (where $w_0$ represents the initial state) by examining  a finite set of values of $u_n :=  \theta(w_n) = \theta(Tw_{n-1})$.
 The sequence $\{u_n\}$ represents a scalar time-series and intuitively $\theta$ represents a probe inserted into a bigger system which is itself only measuring/extracting a small part of the greater system state at time $t=n$.
@@
 
-However, the actual state $u$ of a system is seldom, if ever, fully known.
+Hence, when the actual state $u$ of a system is seldom, if ever, fully known.
 the most one can do is to insert probes into a system to obtain partial information through the measurements taken.
-Moreover, the process of taking a measurement itself introduces two additional aspects which complicate the problem: \ednote{B: Perhaps this edit makes more sense?
-M: Maybe replace difficulty by saying that we need to consider two aspects.
-}
+Moreover, the process of taking a measurement itself introduces two additional aspects which complicate the problem:
 \vspace{-8mm}
 \begin{enumerate}[noitemsep, label=\roman*.]
-  \item A series of measurements over a specific time interval is inherently a discretisation procedure of the underlying continuous-time dynamical system.
+  \item A series of measurements over a specific time interval is inherently an orbit if discretisation  of the underlying continuous-time dynamical system.
 (Hence why we restricted our attention to discrete-time systems in the preceding section)
-  \item A probe will never be entirely accurate, and so the act of measurement introduces a certain measure of numerical noise/inaccuracy.
+  \item A probe will never be entirely accurate, and so the act of measurement introduces a certain measure of numerical noise/inaccuracy (and so we could consider some noise to be involved).
 \end{enumerate}
 
-From here we construct a multidimensional observable using the method of stacking previous observations, i.e., we create delay-coordinate map defined by
+Takens embedding theorem does not take into account the noise, and for describing it, we construct a multidimensional observable using the method of stacking previous observations, i.e., we create delay-coordinate map defined by
 $\Phi_{k,\theta}(w) := (\theta(T^{-k}w)\ldots,\theta(T^{-1}w),\theta(w))$.  
 The concept of an observable is understood in the sense of an observable as originally introduced by~\cite{takens1981detecting, genericObservableAeyels} to refer to a probe or measurement function inserted into the system.
-Takens' theorem, in essence, refers to the fact that for a sufficiently large $k$, we can define a dynamical system on the space $\mathbb{R}^{k+1}$ with states $\Phi_{k,\theta}(w)$, $\Phi_{k,\theta}(Tw)$, $\Phi_{k,\theta}(T^2w)$, etc.
+Takens' theorem, in essence, refers to the fact that for a sufficiently large $k$, we can define a dynamical system on the space $\mathbb{R}^{k+1}$ with states evolving as $\Phi_{k,\theta}(w)$, $\Phi_{k,\theta}(Tw)$, $\Phi_{k,\theta}(T^2w)$, etc.
 This dynamical system is topologically conjugate to the unknown underlying system $(W,T)$.
@@
 
-This requirement is not overly restrictive, as a great number of chaotic systems do have a surjective map.
-(Cite) \ednote{B: I remember you saying this in one of our sessions, but don't have a source for it.}
+This requirement is not overly restrictive, since if a map is topologically transitive on a subspace it has to be surjective on it, since if it were to be not surjective there would be an open set $V$ that does not intersect with any future iterates of an open set  $U$ disjoint from it.
+\ednote{B: I remember you saying this in one of our sessions, but don't have a source for it.
+Do you have a specific reference?
+M: Follows from the transitivity definition.
+It is enough}
 
@@
 
+The USP is equivalent called  the echo state property  \cite{Manju_ESP}, and systems that satisfy this property have other important and useful properties \cite{grigoryeva2018echo, grigoryeva2019differentiable, manjunath2020stability} for information processing.
 One of the first results obtained after defining the USP is that every solution will attract different initial conditions towards the component parts of the solution.
@@
   [\bf Causal Mapping]\label{Def_CausMap}
-  A continuous, surjective map $H:\overleftarrow{U}\to\overleftarrow{X}_U$ such that \[H\circ\tilde{g}_v=\sigma\circ{H}\] holds for all $v \in U$ where $\tilde{g}_v$ maps $(\ldots, u_{-2}, u_{-1})$ to $(\ldots, u_{-2}, u_{-1}, g(v, u_{-1}))$ is called a causal mapping.
+  A continuous, surjective map $H:\overleftarrow{U}\to\overleftarrow{X}_U$ such that \[H\circ\tilde{g}_v=\sigma_v \circ{H}\] holds for all $v \in U$ where $\tilde{g}_v$ maps $(\ldots, u_{-2}, u_{-1})$ to $(\ldots, u_{-2}, u_{-1}, g(v, u_{-1}))$ is called a causal mapping.
 \end{Definition}
@@
 If the causal mapping $H:\overleftarrow{U}{\to}{\overleftarrow{X}_U}$ is injective (in addition to being surjective), it becomes the embedding of $\overleftarrow{U}$ in $\overleftarrow{X}$ induced by the driven system $g$.
-The continuity of $H^{-1}$ follows from the fact that $H$ is itself a continuous and surjective mapping of a compact space $\overleftarrow{U}$ in a Hausdorff space.
+The map $H^{-1}$ is continuous since $H$ is a continuous and surjective mapping of a compact space $\overleftarrow{U}$ to a Hausdorff space.
 and we refer to the map $H$ as a \emph{causal embedding}.
-\ednote{B: edited the sentence here}
 
@@
 
-  It is imperative to take note of a subtlety here: we refer to the \emph{map} $H$ as a (causal) embedding, whereas the driven \emph{system} $g$ (causally) embeds another \emph{system} $(U,T)$.
+  It is imperative to take note of a subtlety here: we refer to the \emph{map} $H$ as a (causal) embedding, whereas when the driven \emph{system} $g$ (causally) embeds another \emph{system} $(U,T)$ we are concerned with $H_2$.
 \ednote{B: inserted line}
@@
 To obtain a suitably complex function $g$, it is thus desired that the reachable set of a driven system  be large enough to relate to the input. 
-This is true even for embedding  the inverse-limit space of a dynamical system. \ednote{B: Expand or refer?}
+This is true even for embedding  the inverse-limit space of a dynamical system \cite{manjunath2021universal}.
 
@@
 \captionof{figure}{A coordinate of a simulated solution $(x_0,x_1,\ldots,x_{5000})$ of a RNN  plotted in red (with parameter $\alpha=0.99$ (where $g$ has the USP) and blue ($a=1$ and $\alpha=1.05$ where $g$ does not have the USP) while the matrices $A$ and $B$ are randomly generated, and $B$ has unit spectral radius.
+The input is shown in black.
 Reproduced from \cite{Manju_Nonlinearity}. }
@@
 \end{figure} 
-\ednote{B: How can I explain the black line? I don't think this makes sense to the reader immediately.}
+\ednote{B: How can I explain the black line? I don't think this makes sense to the reader immediately. M: Oh it was the input in black that I have added.}
 
@@
 In \cite{manjunath2021universal} it is illustrated in an empirical fashion in that the map $G_T$ describes dynamics which are less functionally complex than that of $T$ or of the map $\Phi_{2d,\theta}$.
-This is done by implementing a Recurrent Neural Network (RNN), which is discussed in~\ref{ch5}.
-\ednote{B: Should I add in that we'll be using Pearson coefficient?}
+This is done by implementing a Recurrent Neural Network (RNN), which is later mentiond in Chapter~\ref{ch5}.
+\ednote{B: Should I add in that we'll be using Pearson coefficient?
+Not necessary.
+Please check all the cross citation.
+You need to explicitly mention that it is a Chapter or else only a number would appear}
 
@@
 
-The reasons for taking this roundabout approach remain to be discussed in \ref{subs_LearnGamma}, but a pat answer may immediately be given: When $\Gamma$ has been learnt, the system can be driven autonomously and then $G_T$ is known anyway.
+The reasons for taking this roundabout approach remain to be discussed in Section~\ref{subs_LearnGamma}, but a pat answer may immediately be given: When $\Gamma$ has been learnt, the system can be driven autonomously and then $G_T$ is known anyway.
 We formalise this with a theorem.
@@
 the value a, $\alpha$ in $\tanh$), we exploit both input-related and parameter-related stability.
-Both these notions are defined by way of the continuity of an encoding map of an input and design parameter respectively in \cite{manjunath2020stability}.\ednote{B: Is this satisfactory?}
+Both these notions are defined by way of the continuity of an encoding map of an input and design parameter respectively in \cite{manjunath2020stability}.
 
@@
 
-Thirdly, learning a chaotic map on a set(here $Y_T$) with even a small measure of noise  means that the learnt system accepts arguments out of the set (Problems similar to Takens, cf.~\ref{sect_takenslimits}).
-We thus have a map $G_T^+$ acting on $Y_T^+$ and since $Y_T$ is not guaranteed to be an attractor of $G_T$ this could lead to errors.
-If one learns $G_T$ directly, we are in danger of replicating the problems arising for Takens.  To see this, consider Fig.~\ref{fig:YtGtFailure}.
-
-\begin{figure}[ht]
-  \includegraphics[scale=0.25]{_YTerrors.eps}
-  \centering
-  \captionof{figure}{Diagram of potential errors developing when $G_T$ is learnt.}
+Thirdly, learning a chaotic map on a set(here $Y_T$) with even a small measure of noise  means that the learnt system accepts arguments out of the set (Problems similar to Takens, cf.
+Section~\ref{sect_takenslimits}) \ednote{Please insert the right label}.
+Suppose iterates of  $(x_{n-1},x_{n})$ under the map $G_T$ is used to forecast two successive values of a solution of $g$ contained in $Y_T$, then these iterates would be contained in $Y_T$.
+However, in practice we are confronted with noise.
+Due to noise at some moment, an iterate could slip beyond $Y_T$ and into an
+expansion $Y_T^+$, for example a closed set that contains all the iterates of $(x_{n-1},x_{n})$ under the map $G_T^+$, where $G_T^+$ acts on $Y_T^+$.
+
+Now since $Y_T$ is not guaranteed to be an attractor of $G_T^+$ this could lead to errors. To see this, consider Fig.~\ref{fig:YtGtFailure}.
+Thus if one learns $G_T$ directly, we are in danger of replicating the problems arising for Takens.
+
+\begin{figure}[ht]\includegraphics[scale=0.25]{_YTerrors.eps}
+  \centering
+  \captionof{figure}{Diagram of potential errors developing when $G_T$ is learnt. (Figure from \cite{Supp}}
   \label{fig:YtGtFailure}
@@
 
-To circumvent this: if a driven system with a saturation function\ednote{B: What is a saturation function?} is used, we can indirectly (via $g$) ensure the existence of an attractor $A$ of the map which implements $G_T$, and containing $Y_T^+$.
-\ednote{B: I'm not sure this sentence makes complete sense.}
-By defining the dynamics to $A$, this prevents large errors from occuring.  
-More details on the global dissipativity of $g$, as it is termed there, may be found in~\cite{Supp}.
+To circumvent this: if a driven system with a saturation function \footnote{a function $f(t)$ that is monotonic and approaches a constant as $|t|\to \infty$, for e.g., $\tanh(t)$ on $\mathbb{R}$} is used, we can ensure the existence of an attractor $A$ of the map that implements
+$(\pi_2,g) \circ (\Gamma,\pi_2)$ (printed in the red in ...\ednote{Please cite the commutativity diagram on the previous page}) on 
+ which in practice turns out to be defined on $Y_T^+$ has an attractor $A$  containing $Y_T$.
+\ednote{B: A little vague.
+Are we saying that an attractor of a map implementing $G_T$ is also the one that contains $Y_T^+$?
+M: Please see now..
+there was typo..
+it was A containing $Y_T$.
+I have expanded it now}
+By restricting the dynamics to not spill beyond $A$, prevents large errors from occuring.  
+More details on the global dissipativity of $g$, as it is termed here, may be found in~\cite{Supp}.
 
@@
 To reach this point, one must first learn $\Gamma$.
-After first normalising the data to ensure that it has zero mean value and lies within the range $[-1,1]$ and required zero-padding \ednote{M: this zero padding is something that we need to tell the examiner}, we continue by driving the system with external input through the function $g$.
+After first normalising the data to ensure that it has zero mean value and lies within the range $[-1,1]$ and required zero-padding(see paragraph below for further discussion), we continue by driving the system with external input through the function $g$.
 
@@
 
-We make use of a Recurrent Neural Network (RNN) in our implementation to project the data on to a higher dimensional space, but we do not train the RNN in the conventional manner.
-\ednote{B: Should we say what the conventional manner is?} A few words about the RNNs: RNNs are widely used in the literature for temporal processing of data as a type of deep learning technique.
+We make use of a Recurrent Neural Network (RNN) in our implementation to project the data on to a higher dimensional space, but we do not train the RNN in the familiar ways in the literature but as in \cite{manjunath2021universal}.
+A few words about the RNNs: RNNs are widely used in the literature for temporal processing of data as a type of deep learning technique.
 They take the form of equations considered to be discrete-time version of continuous dynamical systems modelling a neural network of neurons \cite{jaeger2001echo}.
 
-We implement a feedforward network, (a network with no memory, unlike the RNN)\ednote{B: Added so that FFN is in more context} required to learn $\Gamma$ in the \emph{Python} programming language and making use of the \emph{Scipy} library, alongside the \emph{Keras} package in the \emph{Tensorflow} library \ednote{B: Is this something you want me to write in?
-M: Note!
-We do not implement the RNN, but the FN that learns Gamma through the Tensorflow.
-We called the RNN + FN set up a RCN in Adriaan's paper.
-FYI: FN's work only if there is a definite relationship between the $u_n$ and $u_{n+1}$ which is rare in temporal data.
-RNNs with ESP work when there is a definite relationship between $(\ldots,u_{n-1},u_n)$ and $u_{n+1}$ and is suited for temporal processing.}.
+We implement a feedforward network, (a network with no memory, unlike the RNN) required to learn $\Gamma$ in the \emph{Python} programming language and making use of the \emph{Scipy} library, alongside the \emph{Keras} package in the \emph{Tensorflow} library.
 
@@
 
-The two model equations for learning from data (\eqref{eqns_from_data}) allows us to determine what the state $x_{n+1}$ at time $n+1$ will be. 
-We specifically remind ourselves of \[x_{k+1}=\pi_2 \circ (\Gamma, \pi_2) \circ (\pi_2,g) (u_k,x_k)\]
+The two model equations for learning from data~\eqref{eqns_from_data} allows us to determine what the state $x_{n+1}$ at time $n+1$ will be.
+
+%We specifically remind ourselves of \[x_{k+1}=\pi_2 \circ (\Gamma, \pi_2) \circ (\pi_2,g) (u_k,x_k)\]
 
@@
 Then if we feed  the input values $u_n := \Phi_{\theta,2d}(\theta(w_{n}))$ to a driven system $g$ that is SI-invertible and has USP, the induced system $(Y_F,G_F)$ would be topologically conjugate to the inverse-limit space of
-$(\Phi_{\theta,2d}(W), F_\theta)$ as in Theorem~\ref{Thm_CET} \ednote{M: please see if this refers to the causal embedding theorem in the thesis}, and hence one could forecast $u_n,u_{n+1},\ldots$.
+$(\Phi_{\theta,2d}(W), F_\theta)$ as in Theorem~\ref{Thm_CET}, and hence one could forecast $u_n,u_{n+1},\ldots$.
 In other words, one could forecast  the values $\theta(w_n), \theta(w_{n+1}),\ldots$ by feeding these observations as delay coordinates.
@@
 We note that the twin equations together~\ref{eqns_from_data} generate a model from data.
-For a chaotic system, a small amount of noise due either to the input-noise or computational error would cause the solutions from two accurate models with the same initial conditions to  diverge as they are not identical.
-\ednote{B: edited this sentence} However, statistically, they would be the same since due to ergodic theory all typical orbits have the same visiting frequency to any region of the phase space of the dynamical system.
-\ednote{B: I'll need to recap this sentence with you before presenting as it's new to me.} To illustrate that we plot the densities of the trajectories of the actual and predicted solutions.
-
-Our implementation of the reservoir(A) and input(B) matrices in the RNN makes use of sparse matrices.
+For a chaotic system, a small amount of noise due either to the input-noise or computational error would cause the solutions from two accurate models with the same initial conditions to diverge.
+However, statistically, they would behave in the same way in the sense that the visiting frequency or the average time spent by an orbit in any region of the phase space of the dynamical system is identical for almost all initial conditions (almost all is specified as a measure), a result in ergodic theory (\url{https://en.wikipedia.org/wiki/Ergodic_theory} a topic beyond the scope of this thesis.
+\ednote{B: I'll need to recap this sentence with you before presenting as it's new to me.
+M: I rewrote it.} To illustrate that we plot the densities of the trajectories of the actual and predicted solutions.
+
+In our implementation of the input and reservoir  matrices $A$ and $B$ in the RNN are sparse matrices.
 We discuss three specifics relating to this concept.
@@
 \textbf{(Cite.)} By ensuring a condition number close to 1, we also ensure that our driven function $g$, depending on the matrices $A$ and $B$, will not be sensitive to adjustments in the input.
-In so doing we then also guarantee a measure of stability.
-\ednote{B: Not satisfactory.
-Edit again.}
+In so doing we then also guarantee another form stability.
+\ednote{B: Not complete enough.
+Edit further and cite M: Please feel free to cite a wikipedia page}
 
@@
 
-Note that to learn $\Gamma: (x_{n-1},x_{n}) \mapsto u_n$ throughout feedforward network, one could learn from the principal components of $(x_{n-1},x_{n})$ as well. 
-We employ this since empirically it gives slightly better results and not essential.
-The principal components are used only for an efficient state representation possibly to reduce the errors while learning and is not for a lossy approximation since all principal components are used.
-  To be explicit, we denote $X_{1:N}$ to be the matrix with the first $N$ states of the network data as row vectors.
-So, if $X_{1:N}=U\Sigma P^T$ represents the singular value decomposition of $X_{1:N}$, then the principal component matrix is denoted by $P$, and the principal components are given by
+Note that to learn $\Gamma: (x_{n-1},x_{n}) \mapsto u_n$ through our feedforward network, we could learn the principal components of $(x_{n-1},x_{n})$ as well. 
+We employ this since it provides slightly better results empirically but is not essential.
+
+Principal components are used only for a more efficient state representation to possibly reduce learning errors; it is not for a lossy approximation since we use all principal components \textbf{Cite.}.
+More explicitly, denote $X_{1:N}$ as the matrix with the first $N$ states of the network data represented by row vectors.
+If $X_{1:N}=U\Sigma P^T$ here represents the singular value decomposition ($U$ and $V$ are unitary matrices and $\Sigma$ is a diagonal matrix) of $X_{1:N}$ and we refer to ...\ednote{B: Todo.
+Refer or Cite M: Please add a Wiki citation in the bib}, we then denote the principal component matrix $P$, and the principal components are given by
  $Z_{1:N}=X_{1:N}P.$
-These principal components are then used to train a feedforward neural network. If we denote the row vectors of $Z_{1:N}$ by 
-$z_i^T, i=1,2,\cdots,N$ and the neural network by $NN$, then the network is trained to learn an approximation of the map
+These principal components are then used to train a feedforward neural network. Denote the row vectors of $Z_{1:N}$ by 
+$z_i^T, i=1,2,\cdots,N$ and the neural network by $NN$, the network is subsequently trained to learn an approximation of the map
 $NN: (z_{n-1},z_n) \mapsto  u_n.
@@
   \centering
-  \includegraphics[width=0.95\linewidth]{_dp_success_4coords_traj.eps} 
+  \includegraphics[width=0.95\linewidth]{Graphs/_dp_success_4coords_traj.pdf} 
   \captionof{figure}{Predicted(red) and actual(blue) trajectories of the $x$- and $y$-coordinates of each of the pendulum heads.
@@
  \end{figure}
+
+ Lastly, an run was done where only a single coordinate (the $y$-coordinate of the lower pendulum head) was fed into the system and we can illustrate through the obtained success that the coordinate contains all the information in its past states necessary to forecast its future evolution.
+ \textbf{Insert graphs here. Had to quickly change something here.}
 
@@
 
-\chapter{Conclusions} \label{ch6}
+\chapter{Conclusions}\label{ch6}
 
@@
 
-As technology has advanced over the past century, humanity’s ability to record and manipulate rich, complex data in order to produce forecasts has profoundly increased. 
-Collecting measurements from underlying artificial, natural, and physical systems through observables to realise a dynamical model with greater descriptive power is well-established in present-day science.
-
-Core to the success of these techniques is the fact that dynamics discerned from the observed space can often be characterised by a mapping of much lower functional complexity when compared to the map describing the unknown underlying system.
-Takens theorem and the numerous generalizations thereof is one such method establishing (under some generic conditions) the learnability of a system produced by creating a delay-coordinate vector from a sufficiently large number of previous observations of a dynamical system.
-This system is topologically conjugate to the system from which the observed time-series was obtained.
-
-A critical concern of this method is the requirement of full knowledge of the variables of the underlying state space that govern the complex dynamical system one is interested in.
-In real-world conditions this intelligence is seldom available and one must depend heavily on human expertise and insight.
+There are a myriad of techniques to generate accurate models for observed dynamical systems using data obtained from measurement functions.
+Core to the success of such methods is the fact that we can often characterise the dynamics from am observed space by a mapping of much lower functional complexity when compared to the true map describing the unknown underlying system.
+Takens theorem and the numerous generalizations thereof is one such method establishing (under some generic conditions) the learnability of a system produced by creating a delay-coordinate vector from a sufficiently large number of previous observations from a dynamical system.
+In this case the new system is topologically conjugate to the system from which the observed time-series was obtained.
+A critical concern of such methods is the requirement of full knowledge of the underlying state space variables governing the dynamical system one is interested in.
+In real-world conditions this information is rarely available and, when it is, is often contingent upon human expertise and insight.
 
 Each discrete-time state space model has at its core a driven dynamical system. 
-In this project, it has been shown that an input originating from a compact space entailing exactly one solution (i.e., if the driven system has the USP) is equivalent to representing information without distortion, a concept which was formalised by establishing a causal mapping’s existence.
-The established results are general, implying that any driven dynamical system with specific properties can give rise to a causal mapping, and in so doing  providing an answer as to the success of arbitrary choices of driven systems in information processing applications in \cite{jaeger2004harnessing, mass2002realtime}.
-
+In this project, it has been shown that an input originating from a compact space entailing exactly one solution (i.e., if the driven system has the USP) is equivalent to representing information without distortion, a concept which was formalised by establishing the existence of a causal mapping’s.
+The established results are general, implying that any driven dynamical system with specific properties, such SI-invertibility, can give rise to a causal mapping. 
 In particular, we have a topological conjugacy (semi-conjugacy) between the data comprising the single-delay lag dynamics of a driven system~\cite{manjunath2013echo} and the underlying homeomorphic dynamical system.
-Finite-length input data is sufficient for forecasting as driven system states are influenced only in an almost negligibly small manner by the left-infinite segment of the past of the input beyond some arbitrarily large but finite time (a result which holds due to the universal semi-conjugacy of a driven system’s continuity)
-
-Viewed through the lens reservoir computing approaches, we have also demonstrated that when single-delay dynamics are utilised, a learnable map in a reservoir computing network certainly exists.
-Moreover, such a driven system within a reservoir computing network enjoys certain proven qualities pertaining to the robustness w.r.t input- and parameter-perturbation.
-Accordingly, we are able to preserve the quality of a representation that would otherwise remain sensitive to various fluctuations, rendering the learning of the system’s dynamics wholly unpredictable.
-In addition, we obtain the robustness of the system against external noise, a phenomenon which is observed for the Fractal Dream Attractor.
-
-Finally, we consider some physical systems to observe the methodology in practice and verify that we can indeed predict the future evolution of chaotic systems such as the double pendulum.
-Empirical evidence shows that a linearity measure like a Pearson correlation coefficient points at reduced functional complexity while learning the dynamics in the state space of the driven system
-
-Considered with a philosophical-scientific lens, this project has aimed to establish a juncture between rationalism (the theory of nonautonomous dynamical systems) and empiricism (machine learning) to yield results profitable to the discipline of Applied Mathematics.
-This was done by showing that dynamics in an observed space can be established as topologically conjugate to the underlying system.
-As a result of the topological conjugacy, deep learning techniques are then subsequently applied to physically such equations from the data.
-In so doing, we acquire models of high fidelity that provide exceptional forecasting results on chaotic data that exhibit both statistical and topological consistency in the long run.
+Finite-length input data is shown to be sufficient for forecasting as the left-infinite segment of the past of the input beyond some arbitrarily large but finite time (a result which holds due to the universal semi-conjugacy of a driven system’s continuity) influences the driven system's states in an almost negligibly small manner.
+
+Viewed through the lens of reservoir computing, we have also demonstrated that when single-delay dynamics are utilised, a learnable map in a reservoir computing network certainly exists. 
+Moreover, within a reservoir computing network, such a driven system posssesses certain proven qualities pertaining to the robustness of the system to different forms of perturbation. 
+We are thus able to preserve the quality of a representation that would otherwise have remained sensitive to various fluctuations and rendering the learning of the system’s dynamics wholly unpredictable.
+
+Finally, we have considered some physical systems to observe the methodology in practice and verify that we can indeed predict the future evolution of chaotic systems such as the double pendulum.
+We demonstrated the robustness of our methodology to external noise by considering the cases of the  Dream Attractor and the Double Pendulum.
+
+\ednote{Loose paragraph commented out}
+% Considered from a philosophical-scientific  perspective, this project has aimed to establish a juncture between rationalism (the theory of nonautonomous dynamical systems) and empiricism (machine learning) to yield results profitable to the discipline of Applied Mathematics.
+% This was done by showing that dynamics in an observed space can be established as topologically conjugate to the underlying system. 
+% As a result of the topological conjugacy, deep learning techniques are then subsequently applied to physically such equations from the data. 
+% In so doing, we acquire models of high fidelity that provide exceptional forecasting results on chaotic data that exhibit both statistical and topological consistency in the long run.
 

### 2022-01-17 draft_16Jan/hiscurrent.tex
Changed abstract and conclusion

@@
 This is because differing orbits diverge quickly and to such an extent that it becomes seemingly impossible to retrace their steps back to the original conditions.
-Even the presence of usually' negligible' computational noise or measurement error renders long-term pointwise prediction infeasible.
-There are many difficulties that one encounters while modelling such systems:
+Even the presence of usually `negligible' computational noise or measurement error renders long-term pointwise prediction infeasible.
+We encounter many difficultiesin the process of modelling such systems:
 \vspace{-8mm}
@@
   \item Partially interpretable models capturing some modes of the dynamics, 
-  \item Non-Interpretable models, defined as such mainly due to the fact that they are defined on a different phase space that is usually high-dimensional. 
+  \item Non-interpretable models, defined as such mainly due to the fact that they are defined on a different phase space that is usually high-dimensional. 
 \end{enumerate}
@@
 
-Recent partially interpretable models available in the literature have been based on the Koopman operator (e.g.,[?, ?]) to employ observables mapping the data onto a higher-dimensional space.
+Recent partially interpretable models available in the literature have been based on the Koopman operator (see~\cite{koopman1932dynamical,budivsic2012applied}) to employ observables mapping the data onto a higher-dimensional space.
 This makes the dynamics in the higher-dimensional space more amenable for approximation by a linear transformation.
-Such methods do not guarantee exact reconstructions for nonlinear models, and in practice, provide poor long-term consistency for a large class of systems.
-(Citation)
-\textbf{(Citation)}
+Such methods do not guarantee exact reconstructions for nonlinear models, and in practice provide poor long-term consistency for a large class of chaotic dynamical systems \cite{Supp}.
 
 The non-interpretable models include the delay embedding and the machine learning algorithms.
-For example, one could learn a system conjugate to the underlying system by applying the Takens delay embedding~\cite{takens1981detecting} when one has "good" observations from the system; this learnt system could then be used to forecast the observed data.
-Takens delay embedding theorem ~\cite{takens1981detecting} and its various generalisations (e.g., \cite{sauer1991embedology, stark1999delay, gutman2018embedding}) establishes the learnability of a system constructed by concatenating sufficiently large previous time-series observations of a dynamical system into a vector (called delay coordinates).
+For example, one could learn a system conjugate to the underlying system by applying the Takens delay embedding method of delay-coordinates~\cite{takens1981detecting} when one has "good" observations from the system; this learnt system could then be used to forecast the observed data.
+Takens delay embedding theorem~\cite{takens1981detecting} and its various generalisations (see,~\cite{sauer1991embedology, stark1999delay, gutman2018embedding}) establishes the learnability of a system constructed by concatenating sufficiently large previous time-series observations of a dynamical system into a vector (called delay coordinates).
 This then confirms the existence of a map on the space of delay coordinates equivalent (or topologically conjugate) to the underlying map from which the observed time-series was first obtained.
-Although topological conjugacy guarantees an alternate representation of the underlying system, the quality still depends on numerous parameters, making the comprehension of the dynamics unreliable (e.g., [?]).
+Although topological conjugacy guarantees an alternate representation of the underlying system, the quality still depends on numerous parameters, making the comprehension of the dynamics unreliable (see, [?]).
 One reason for this fragility is that the embedded attractor in the reconstruction space is not always an attractor of the map learnt in the reconstruction space, despite unquestionably being an invariant set.
@@
 Practically, the application of Takens embedding involves learning a map through some technique, and consequently one wishes that these would have low functional complexity\cite{manjunath2021universal}, i.e., functions with fewer oscillatory graphs.
-On the other hand, pure machine learning methodology processes temporal information (like the echo state networks (Citations to echo state networks)) by mapping data onto a higher dimensional space for further processing.
-Although they perform  well on forecasting some dynamical systems, they fail completely on others (Citations here) as there is often no guarantee that the right function was learnt during training.
-
-This thesis involves implementing and analysing non-interpretable models that can guarantee exact reconstruction. 
+On the other hand, pure machine learning methodology processes temporal information (like the echo state networks~\cite{Manju_ESP, Manju_IEEE, grigoryeva2018echo}) by mapping data onto a higher dimensional space for further processing.
+Although they perform  well on forecasting some dynamical systems, they fail completely on others (\ednote{Ask for resources here}) as there is often no guarantee that the right function was learnt during training.
+
+This project involves implementing and analysing non-interpretable models that can guarantee exact reconstruction. 
 The project work concerns the study and implementation of a method (\cite{manjunath2021universal}) that incorporates learning a function by mapping the data on to a higher dimensional space using what is called a driven dynamical system (See Chapter~ref{ch.4}).
@@
 \newline Chapter~\ref{ch3} introduces the problem of forecasting dynamical systems, states the Takens delay embedding theorem, and discusses various issues faced while forecasting. 
-\newline In Chapter~\ref{ch4}, a driven dynamical system is defined and discuss the properties of a specific class of driven dynamical systems that we make use of in this thesis/report/paper. 
+\newline In Chapter~\ref{ch4}, a driven dynamical system is defined and discuss the properties of a specific class of driven dynamical systems that we make use of in this project.
 \newline Finally, in Chapter~\ref{ch5} we show the implementation of these forecasting methods, and conclusions are provided in Chapter~\ref{ch6}.
@@
 At its most elementary level, a dynamical system is just something that evolves deterministically through time.
-In the context of this thesis, deterministic refers to the fact that a system evolves according to specified rules rather than based on random events.
+In the context of this project, deterministic refers to the fact that a system evolves according to specified rules rather than based on random events.
 Dynamical systems arise in a variety of situations.
@@
 Once again, we may consider the field of ecology, where discrete dynamical systems are often directly derived or assumed.
-\textbf{Citation}
+\ednote{Ask for resource}
 There is a school of thought that advocates discrete-time dynamical systems to be more natural for modelling real-world observations than differential equations.
-We refer the interested reader to \cite{saber2010introduction}.
+We refer the interested reader to~\cite{saber2010introduction}.
 
@@
 For the example, $Tu =u^2$ on $[0,1]$, both $\{0\}$ and $[0,1]$ are attractors.
-It is known that an attractor for a dynamical system always exists in a compact space.
-\textbf{(Cite).}
+It is known that an attractor for a dynamical system always exists in a compact space~\cite{Milnor1985}.
 
@@
 For instance an invariant set could be just a single point or it could have an infinite set.
-If the invariant set is infinite, then complicated dynamics are possible.A particular well-studied phenomenon of such complexity gives rise to so-called  chaotic behaviour, studied in detail over the past fifty years.
+If the invariant set is infinite, then complicated dynamics are possible.
+A particular, well-studied phenomenon of such complexity gives rise to so-called chaotic behaviour, a subject studied in detail over the past fifty years.
 
@@
 
-There are several intricately different definitions of chaos available in the literature \textbf{(Cite)}, with each of them indicating some aspect of complexity.
-In practice, it is not possible to verify which definition or notion does the underlying dynamical system satisfies when only data is observed from it.
+Different authors propose a number of intricately different definitions of chaos in the literature~\cite{RasbandChaos, TaborChaos, WigginsChaos} and each of them indicate some aspect of complexity.
+
+In practice, when only data is observed from a system, it is not possible to verify which definition or notion is satisfied by the underlying dynamical system.
 We do, however, specifically recall the definition of chaos in the sense of Devaney \cite{devaney2018introduction,de2013elements} so as to understand some nuances behind the complexity.
@@
 \begin{Example} \rm
-  The logistic map(LM), a recursion relation of the form $x_{n+1}=rx_n(1-x_n)$ where $x_n\in[0,1]$ and $r\leq{4}$ is of particular interest to us as it is classically used to illustrate the influence of chaos on a system \textbf{(Cite)}.
+  The logistic map(LM), a recursion relation of the form $x_{n+1}=rx_n(1-x_n)$ where $x_n\in[0,1]$ and $r\leq{4}$ is a classic example used to illustrate the chaotic behaviour that a system may exhibit.
+When $r=4$, the one-dimensional recurrence relation $x_{n+1}=4x_n(1-x_n)$ can be used as the kindergarten's model to exhibit the presence of SDIC.
+\ednote{B: edited and removed unnecessary citation}
   In the graph below~\ref{fig:log_sdic} we plot the first 50 iterates of the LM for $r=4$ and two slightly different initial values $x_0$.
@@
 So please edit accordingly.}
-When $S$ is conjugate to $T$, the dynamics of the two systems are in some way ''dynamically equivalent''.
+When $S$ is conjugate to $T$, the dynamics of the two systems are in some way `dynamically equivalent'.
 Specifically, they are in one-to-one correspondence with one another.
@@
 When $S$ is semi-conjugate to $T$, it is also common to call $S$ a \emph{factor of $T$}, or conversely that $T$ is an \emph{extension of $S$}.
-In essence, an extension (e.g., \cite{de2013elements}) is a more extensive system capturing all of the essential dynamics of its factor.
+In essence, an extension (see, \cite{de2013elements}) is a more extensive system capturing all of the essential dynamics of its factor.
 
@@
 In this method, one considers a discrete-time dynamical system defined on a 'nice' space (a smooth manifold)\ednote{B: Should this be further explained?} that can be obtained as time-$K$ map of a continuous time-dynamical system, concepts formalised below.
-We make observations from such a system, i.e., we observe the evolution of $\theta(w_0)$ (where $w_0$ represents the initial state) i.e.\ednote{two i.e.'s}, we examine  a finite set of values of $u_n :=  \theta(w_n) = \theta(Tw_{n-1})$.
+We make observations from such a system, i.e., we observe the evolution of $\theta(w_0)$ (where $w_0$ represents the initial state) by examining  a finite set of values of $u_n :=  \theta(w_n) = \theta(Tw_{n-1})$.
 The sequence $\{u_n\}$ represents a scalar time-series and intuitively $\theta$ represents a probe inserted into a bigger system which is itself only measuring/extracting a small part of the greater system state at time $t=n$.
@@
 
-By generic we mean a residual set on a certain topology on appropriate spaces of functions (that we do not describe here) \textbf{(Cite)}. 
+By generic we mean a residual set on a certain topology on appropriate spaces of functions (that we do not describe here) \ednote{B: Where can I find a citation for this?}. 
 By $C^2$, we make reference to a twice-differentiable function  with a continuous second derivative.
@@
 This requirement is not overly restrictive, as a great number of chaotic systems do have a surjective map.
-(Cite) \ednote{B: I remember you saying this in one of our sessions, but don't have a source for it.}
+(Cite) \ednote{B: I remember you saying this in one of our sessions, but don't have a source for it.
+Do you have a specific reference?}
 
@@
 The reachable set itself can defined independently from $g$ possessing the ESP/USP.
-We note that $x\in{X_n(\overline{u})}$ if and only $g$ has a solution $\{x_k\}$ for $x_n=x$ and input $\overline{u}$.
-\textbf{(Cite)}.
+Note that $x\in{X_n(\overline{u})}$ if and only $g$ has a solution $\{x_k\}$ for $x_n=x$ and input $\overline{u}$, a result proved in nonautonomous dynamical systems literature~\cite{manjunath2014dynamics, manjunath2013echo}.
+\ednote{B: Adjusted sentence and cited}
 
@@
 This will lead to a result (for.
-e.g., \cite{manjunath2013echo}) that  \textit{$g$ being a topological contraction is equivalent to the existence of a exactly one entire solution.}
+see, \cite{manjunath2013echo}) that  \textit{$g$ being a topological contraction is equivalent to the existence of a exactly one entire solution.}
 
@@
 Fortunately, one does not need the entire left-infinite history of an input in practice thanks to the Uniform Attraction Property(UAP). 
-We use an alternate version of the definition of UAP -- since the notion of nonautonomous attractors in this thesis would take some time to establish and detracts from the principal thrust of this project, we refer to \cite{Manju_Nonlinearity}.
+We use an alternate version of the definition of UAP -- since the notion of nonautonomous attractors in this project would take some time to establish and detracts from the principal thrust of this project, we refer to \cite{Manju_Nonlinearity}.
 
@@
 \end{figure}
+\ednote{Ask M: Cite SIG-25May.}
 
@@
 
-To circumvent this: if a driven system with a saturation function\ednote{B: What is a saturation function?} is used, we can indirectly (via $g$) ensure the existence of an attractor $A$ of the map which implements $G_T$, and containing $Y_T^+$.
-\ednote{B: I'm not sure this sentence makes complete sense.}
+To circumvent this: if a driven system with a saturation function\ednote{B: What is a saturation function?} is used, we can ensure the existence of an attractor $A$ of the map which implements $G_T$, and containing $Y_T^+$ indirectly (via $g$).
+\ednote{B: A little vague.
+Are we saying that an attractor of a map implementing $G_T$ is also the one that contains $Y_T^+$?}
 By defining the dynamics to $A$, this prevents large errors from occuring.  
@@
 
-We implement a feedforward network, (a network with no memory, unlike the RNN)\ednote{B: Added so that FFN is in more context} required to learn $\Gamma$ in the \emph{Python} programming language and making use of the \emph{Scipy} library, alongside the \emph{Keras} package in the \emph{Tensorflow} library \ednote{B: Is this something you want me to write in?
-M: Note!
-We do not implement the RNN, but the FN that learns Gamma through the Tensorflow.
+We implement a feedforward network, (a network with no memory, unlike the RNN)\ednote{B: Expanded sentence s.t.
+FFN is in more context} required to learn $\Gamma$ in the \emph{Python} programming language and making use of the \emph{Scipy} library, alongside the \emph{Keras} package in the \emph{Tensorflow} library \ednote{B: Is this something you want me to write in?
+M: We do not implement the RNN, but the FN that learns Gamma through the Tensorflow.
 We called the RNN + FN set up a RCN in Adriaan's paper.
@@
 
-The two model equations for learning from data (\eqref{eqns_from_data}) allows us to determine what the state $x_{n+1}$ at time $n+1$ will be. 
+The two model equations for learning from data~\eqref{eqns_from_data} allows us to determine what the state $x_{n+1}$ at time $n+1$ will be. 
 We specifically remind ourselves of \[x_{k+1}=\pi_2 \circ (\Gamma, \pi_2) \circ (\pi_2,g) (u_k,x_k)\]
@@
 The condition number of a matrix $M$ (denoted $cond(M)$) is the product of the matrix norms of the matrix $M$ and its inverse $M^{-1}$ and is a measure of the sensitivity of output of a matrix multiplication when an input is changed.
-\textbf{(Cite.)} By ensuring a condition number close to 1, we also ensure that our driven function $g$, depending on the matrices $A$ and $B$, will not be sensitive to adjustments in the input.
-In so doing we then also guarantee a measure of stability.
-\ednote{B: Not satisfactory.
-Edit again.}
+\textbf{(Cite.)}\ednote{Ask M: Cite specific resource?} By ensuring a condition number close to 1, we also ensure that our driven function $g$, depending on the matrices $A$ and $B$, will not be sensitive to adjustments in the input.
+In so doing we then also guarantee another form stability.
+\ednote{B: Not complete enough.
+Edit further and cite}
 
 Lastly, we ensure the matrix $B$ has unit spectral radius.
-This is necessary due to the Causal Embedding Theorem~\ref{Thm_GT_Exists} which can then guarantee the existence of semi-conjugacy between the systems $(Y_T, G_T)$ and $(\widehat{U}_T, \widehat{T})$.
+This is necessary due to the Causal Embedding Theorem~\ref{Thm_CET} which can then guarantee the existence of semi-conjugacy between the systems $(Y_T, G_T)$ and $(\widehat{U}_T, \widehat{T})$.
 In fact, the theorem actually requires that $\alpha{B}$ have spectral radius $<1$, so we impose the restriction $\alpha<1$ and consider only the matrix $B$'s spectral radius from thereon.
 
-Note that to learn $\Gamma: (x_{n-1},x_{n}) \mapsto u_n$ throughout feedforward network, one could learn from the principal components of $(x_{n-1},x_{n})$ as well. 
-We employ this since empirically it gives slightly better results and not essential.
-The principal components are used only for an efficient state representation possibly to reduce the errors while learning and is not for a lossy approximation since all principal components are used.
-  To be explicit, we denote $X_{1:N}$ to be the matrix with the first $N$ states of the network data as row vectors.
-So, if $X_{1:N}=U\Sigma P^T$ represents the singular value decomposition of $X_{1:N}$, then the principal component matrix is denoted by $P$, and the principal components are given by
+Note that to learn $\Gamma: (x_{n-1},x_{n}) \mapsto u_n$ through our feedforward network, we could learn the principal components of $(x_{n-1},x_{n})$ as well. 
+We employ this since it provides slightly better results empirically but is not essential.
+
+Principal components are used only for a more efficient state representation to possibly reduce learning errors; it is not for a lossy approximation since we use all principal components \textbf{(Cite.)}\ednote{Ask M: Is there a specific resource to cite?}.
+More explicitly, denote $X_{1:N}$ as the matrix with the first $N$ states of the network data represented by row vectors.
+If $X_{1:N}=U\Sigma P^T$\ednote{Expand here} represents the singular value decomposition of $X_{1:N}$ \ednote{Ask: Is there a specific resource to cite?}, we then denote the principal component matrix $P$, and the principal components are given by
  $Z_{1:N}=X_{1:N}P.$
-These principal components are then used to train a feedforward neural network. If we denote the row vectors of $Z_{1:N}$ by 
-$z_i^T, i=1,2,\cdots,N$ and the neural network by $NN$, then the network is trained to learn an approximation of the map
+These principal components are then used to train a feedforward neural network. Denote the row vectors of $Z_{1:N}$ by 
+$z_i^T, i=1,2,\cdots,N$ and the neural network by $NN$, the network is subsequently trained to learn an approximation of the map
 $NN: (z_{n-1},z_n) \mapsto  u_n.
@@
 (See below)
-This system is a classical example of a system exhibiting chaos. \textbf{(Citation.)} Consider again~\ref{fig:dp_sdic}.
+This system is a classical example of a system exhibiting chaos and is often used as explanation of the very well-known Butterfly Effect.
+\ednote{B: Originally though a reference here was needed, but no longer think so.} Consider again~\ref{fig:dp_sdic}.
 The system has four variables: angle and angular velocity of each pendulum rod. An ideal system is assumed where the rods are massless and no friction in the pivots are considered.
@@
 We now consider an experiment where noise is added to the system and find it to be incredibly robust.
-We employ the same parameters as before and add noise from a normal-distribution with mean zero and standard deviation equal to 0.1.
+We employ the same parameters as before and add noise from a normal-distribution with mean zero and standard deviation equal to 0.1 which translates to roughly 39dB's of noise.
+(For the signal-to-noise ration we adopt the formula $\text{SNR}_{dB}=10\text{log}_{10}\big(\frac{P_\text{{signal}}}{P_{\text{noise}}}\big)$ where $P_\text{{signal}}$ refers to the variance observed in the normalised data and $P_{\text{noise}}$ represents the variance of the added noise.)
+
 Once again consider the actual and predicted trajectories for the noisy DP dataset in Figure~\ref{fig:noisydp_success_traj}.
@@
   \includegraphics[width=0.95\linewidth]{_dp_noise_FigCD_2030.eps} 
-  \captionof{figure}{Predicted(red) and actual(blue) trajectories of the $x$- and $y$-coordinates of each of the pendulum heads with noise added from a normal distribution with mean zero and standard deviation 0.1.
+  \captionof{figure}{Predicted(red) and actual(blue) trajectories of the $x$- and $y$-coordinates of each of the pendulum heads with noise added from a normal distribution with mean zero and standard deviation 0.1, equating to approximately 39dB's of noise.
   These graphs were constructed by predicting the DP 1000 timesteps into the future and in so doing illustrating the long-term consistency and accuracy of the learnt system.
@@
  \end{figure}
+
+ Lastly, a run was done where only a single coordinate (the $y$-coordinate of the lower pendulum head) was fed into the system and we can illustrate through the obtained success that the coordinate contains all the information in its past states necessary to forecast its future evolution.
+ \textbf{Insert graphs here. Had to quickly change something here.}
 
@@
 Two cases were considered: noise-free and the one perturbed by noise.
-Noise was generated from a normal distribution with mean 0 and standard deviation 0.05, which translates approximately to * dB \ednote{Please calculate dB as signal to noise ratio }
+Noise was generated from a normal distribution with mean 0 and standard deviation 0.05, which to translates approximately 33dB.
+($\text{SNR}_dB$ is calculated exactly as for the DP-data above)
+
 A sequence of 3000 observations in both the clear and noisy data-sets respectively were fed into the system with 300 entries discarded to simulate the network's loss of memory.
@@
 
-% Besides being able to learn the action of the Koopman operator exactly, our setup has other advantages over selecting or optimizing the choice of observables as in the Extended Dynamic Mode Decomposition (EDMD) algorithm (e.g., [21, 22]).
+% Besides being able to learn the action of the Koopman operator exactly, our setup has other advantages over selecting or optimizing the choice of observables as in the Extended Dynamic Mode Decomposition (EDMD) algorithm (see, [21, 22]).
 We can alter the functional complexity of G T by changing the number of observables or equivalently by altering the dimension of X – in the case of a RNN implementation of g, by simply changing the number of neurons in the network.
-Empirically, increasing the dimension of X is found to increase the linear relationship (or intuitively reduces the functional complexity of G T ) that is measured as a generalization of the Pearson correlation coefficient to random vectors (e.g., [44]) between (x_{n−1} , x_{n}) and G_T (x_{n−1} , x_n), and such numerical evidence is tabulated in [11, Table 1].
+Empirically, increasing the dimension of X is found to increase the linear relationship (or intuitively reduces the functional complexity of G T ) that is measured as a generalization of the Pearson correlation coefficient to random vectors (see, [44]) between (x_{n−1} , x_{n}) and G_T (x_{n−1} , x_n), and such numerical evidence is tabulated in [11, Table 1].
 In contrast, expanding the set of observables while using EDMD to learn a map with a lower functional complexity is a highly involved task, not least because so much that one often does not know how to guess a new observable, as well as the fact that adding observables does not necessarily retain the invariance of the span of the observables.
@@
 % Empirically (see Table ??), increasing the dimension of X is found to increase the linear relationship
-% (or intuitively reduce the functional complexity of G T ) that is measured as a generalization of the Pearson correlation coefficient to random vectors (e.g., [?]) between (x n−1 , x n ) and G T (x n−1 , x n ).
+% (or intuitively reduce the functional complexity of G T ) that is measured as a generalization of the Pearson correlation coefficient to random vectors (see, [?]) between (x n−1 , x n ) and G T (x n−1 , x n ).
 If Σ a and Σ b denotes the covariance matrices of the vectors (x n−1 , x n ) and G T (x n-1 , x n ) respectively, and if Σ ab denotes the covariance matrix between the vectors (x n-1 , x_n ) and G_T (x n-1 , x_n ), then the multidimensional 18correlation coefficient is computed by using the traces of these matrices:
@@
 Further, empirically it is found that $G_T$ is much more linear than $T$ as measured by what is called the multidimensional correlation coefficient $\rho$.
-This coefficient is a generalization of the Pearson correlation coefficient for real-valued random variables to random vectors (e.g., \cite{puccetti2019measuring}).
+This coefficient is a generalization of the Pearson correlation coefficient for real-valued random variables to random vectors (see, \cite{puccetti2019measuring}).
 We state teh details of computing this coefficient between
@@
 
-\chapter{Conclusions} \label{ch6}
-
-% The driven states $(x_{n-1},x_n)$ hold the information of the left-infinite history of the input.
-Furthermore, there is some evidence that the inverse-limit space (not system) seems to distinguish two dynamical systems that are not topologically conjugate for a class of systems in the sense their inverse-limit spaces are not homeomorphic to each other.
-So the relation $Y_T$, which is homeomorphic to the inverse-limit system, may be unique to a class of systems.
-(The reference is a paper on Ingram’s Conjecture by Marcy Barge, Henk Bruin and Sonja Stimac.) Further investigation is needed.
-\cite{budivsic2012applied, korda2020data}
-
-As technology has advanced over the past century, humanity’s ability to record and manipulate rich, complex data in order to produce forecasts has profoundly increased. 
-Collecting measurements from underlying artificial, natural, and physical systems through observables to realise a dynamical model with greater descriptive power is well-established in present-day science.
-
-Core to the success of these techniques is the fact that dynamics discerned from the observed space can often be characterised by a mapping of much lower functional complexity when compared to the map describing the unknown underlying system.
-Takens theorem and the numerous generalizations thereof is one such method establishing (under some generic conditions) the learnability of a system produced by creating a delay-coordinate vector from a sufficiently large number of previous observations of a dynamical system.
-This system is topologically conjugate to the system from which the observed time-series was obtained.
-
-A critical concern of this method is the requirement of full knowledge of the variables of the underlying state space that govern the complex dynamical system one is interested in.
-In real-world conditions this intelligence is seldom available and one must depend heavily on human expertise and insight.
-
-Each discrete-time state space model has at its core a driven dynamical system.
-In this project, it has been shown that an input originating from a compact space entailing exactly one solution (i.e., if the driven system has the USP) is equivalent to representing information without distortion, a concept which was formalised by establishing a causal mapping’s existence.
-The established results are general, implying that any driven dynamical system with specific properties can give rise to a causal mapping, and in so doing  providing an answer as to the success of arbitrary choices of driven systems in information processing applications in \cite{jaeger2004harnessing, mass2002realtime}.
-
-In particular, we have a topological conjugacy (semi-conjugacy) between the data comprising the single-delay lag dynamics of a driven system~\cite{manjunath2013echo} and the underlying homeomorphic dynamical system.
-Finite-length input data is sufficient for forecasting as driven system states are influenced only in an almost negligibly small manner by the left-infinite segment of the past of the input beyond some arbitrarily large but finite time (a result which holds due to the universal semi-conjugacy of a driven system’s continuity)
-
-Viewed through the lens reservoir computing approaches, we have also demonstrated that when single-delay dynamics are utilised, a learnable map in a reservoir computing network certainly exists.
-Moreover, such a driven system within a reservoir computing network enjoys certain proven qualities pertaining to the robustness w.r.t input- and parameter-perturbation.
-Accordingly, we are able to preserve the quality of a representation that would otherwise remain sensitive to various fluctuations, rendering the learning of the system’s dynamics wholly unpredictable.
-In addition, we obtain the robustness of the system against external noise, a phenomenon which is observed for the Fractal Dream Attractor.
-
-Finally, we consider some physical systems to observe the methodology in practice and verify that we can indeed predict the future evolution of chaotic systems such as the double pendulum.
-Empirical evidence shows that a linearity measure like a Pearson correlation coefficient points at reduced functional complexity while learning the dynamics in the state space of the driven system
-
-Considered with a philosophical-scientific lens, this project has aimed to establish a juncture between rationalism (the theory of nonautonomous dynamical systems) and empiricism (machine learning) to yield results profitable to the discipline of Applied Mathematics.
-This was done by showing that dynamics in an observed space can be established as topologically conjugate to the underlying system.
-As a result of the topological conjugacy, deep learning techniques are then subsequently applied to physically such equations from the data.
-In so doing, we acquire models of high fidelity that provide exceptional forecasting results on chaotic data that exhibit both statistical and topological consistency in the long run.
+\chapter{Conclusion}\label{ch6}
+\input{chapter6.tex}
 

### 2022-01-17 graphing.py

@@ -26,31 +26,58 @@
 # # plt.plot(X3, T(X3), '-k')
 # # plt.show()
 
+##################################### LOGISTIC MAP ############################
 
-def logistic(r, xn):
-    return r*xn*(1-xn)
+# def logistic(r, xn):
+#     return r*xn*(1-xn)
 
-x0=0.2
-x0_p=0.20001
-r0=4 #3.7
-N=50
+# x0=0.2
+# x0_p=0.20001
+# r0=4 #3.7
+# N=50
 
-Y1 = [x0]
-Y2= [x0_p]
-X = [i for i in range(N)]
-for i in range(1, N):
-    Y1.append(logistic(r0,Y1[i-1]))
-    Y2.append(logistic(r0,Y2[i-1]))
+# Y1 = [x0]
+# Y2= [x0_p]
+# X = [i for i in range(N)]
+# for i in range(1, N):
+#     Y1.append(logistic(r0,Y1[i-1]))
+#     Y2.append(logistic(r0,Y2[i-1]))
 
-import matplotlib.pyplot as plt
-plt.rcParams['figure.figsize'] = [10, 4]
-fig, ax = plt.figure(), plt.axes()
-plt.plot(X, Y1, 'darkblue', lw=0.75)
-plt.plot(X, Y2, 'orangered', lw=0.75)
+# import matplotlib.pyplot as plt
+# plt.rcParams['figure.figsize'] = [10, 4]
+# fig, ax = plt.figure(), plt.axes()
+# plt.plot(X, Y1, 'darkblue', lw=0.75)
+# plt.plot(X, Y2, 'orangered', lw=0.75)
 
-ax.axhspan(0.4-0.05, 0.4+0.05, alpha=0.6, color='lightgreen')
-ax.set_xlabel(r"n", fontsize=12)
-ax.set_ylabel(r"$x_n$", fontsize=12)
+# ax.axhspan(0.4-0.05, 0.4+0.05, alpha=0.6, color='lightgreen')
+# ax.set_xlabel(r"n", fontsize=12)
+# ax.set_ylabel(r"$x_n$", fontsize=12)
 
-# plt.savefig('_logistic_sdic.eps', format='eps', dpi=1000)
-plt.show()
+# # plt.savefig('_logistic_sdic.eps', format='eps', dpi=1000)
+# plt.show()
+
+
+######################## NOISE ADDED #######################################
+import numpy as np
+def calcDB(noise, actual):
+    import math
+    print("Noise in dB")
[…]

### 2022-01-17 projectReport/abstract.tex
Cleaner setup

--- 2022-01-17 draft_16Jan/draft7.tex
@@
-\documentclass[a4paper,12pt,twoside]{report}
-% \documentclass[a4paper,12pt,twoside]{book}
-\pagestyle{headings}
-\usepackage[utf8]{inputenc}
-\usepackage{filecontents}
-\usepackage[many]{tcolorbox}
-\usepackage{graphicx}
-\usepackage{harpoon}
-%\usepackage{apacite} 
-\usepackage{xcolor}
-\usepackage{latexsym}
-\usepackage{amssymb}
-\usepackage{amsmath}
-\usepackage{mdframed}
-%\usepackage[authoryear]{natbib}
-\usepackage{hyperref, url}
-\usepackage[show]{ed}
-\usepackage{subfigure}
-\usepackage[ntheorem]{empheq} 
-\usepackage{amsthm, amssymb, amsfonts, latexsym}
-% \usepackage{pst-all}
-
-\usepackage{enumitem}
-\usepackage{multirow}
-\usepackage{longtable}
-\usepackage[export]{adjustbox}
-% $f\arrowvert_A$ To restrict functions
-
-\setlength{\parskip}{1.7em}
-\setlength{\parindent}{0em}
-
-%\setlength{\baselineskip}{20pt}
-
-\setlength{\textwidth}{429.75499pt}
-\setlength{\textheight}{643.20255pt}
-\setlength{\oddsidemargin}{5 mm}
-\setlength{\evensidemargin}{5 mm}
-\setlength{\topmargin}{0 mm}
-\setlength{\headsep}{0 mm}
-\setlength{\headheight}{0 mm}
-
-\usepackage{pst-node}
-
-\newcommand{\citealtt}[1]{\citeauthor{#1},\citeyear{#1}}
-\newcommand{\myycite}[1]{\citep{#1}}
-
-\mathchardef\mhyphen="2D
-
-%\newcommand{\cev}[1]{\reflectbox{\ensuremath{\vec{\reflectbox{\ensuremath{#1}}}}}}
-
-\usepackage{pst-node,graphicx,pst-blur}
-%\uspackage{auto-pst-pdf}
-%\usepackage{tikz-cd}
-
-\makeatletter
-\DeclareRobustCommand{\cev}[1]{%
-  \mathpalette\do@cev{#1}%
-}
-\newcommand{\do@cev}[2]{%
-  \fix@cev{#1}{+}%
-  \reflectbox{$\m@th#1\vec{\reflectbox{$\fix@cev{#1}{-}\m@th#1#2\fix@cev{#1}{+}$}}$}%
-  \fix@cev{#1}{-}%
-}
-\newcommand{\fix@cev}[2]{%
-  \ifx#1\displaystyle
-    \mkern#20mu
-  \else
-    \ifx#1\textstyle
-      \mkern#20mu
-    \else
-      \ifx#1\scriptstyle
-        \mkern#26mu
-      \else
-        \mkern#26mu
-      \fi
-    \fi
-  \fi
-}
-
-\makeatother
-
-\newcommand{\xcm}{\epsfxsize=3.1cm}
-\newcommand{\fig}[1]{\epsfbox}
-% \newcommand{\bp}{\begin{minipage}{3.1cm}}
-% \newcommand{\ep}{\end{minipage}}
-
-\newtheorem{Definition}{Definition}[]
-%\newmdtheoremenv{Theorem}{Theorem}[]
-\newtheorem{Theorem}{Theorem}[]
-\newtheorem{Lemma}{Lemma}[]
-\newtheorem{Proposition}{Proposition}[]
-\newtheorem{Corollary}{Corollary}[]
-\newtheorem{Remark}{Remark}[]
-\newtheorem{Example}{Example}[]
-
-%Shortcut symbols
-\newcommand{\Ftheta}{\ensuremath{F_\theta}}
-
-\makeatletter 
-\renewcommand{\thefigure}{\@arabic\c@figure}
-\makeatother
-\usepackage{xr}
-
-%%%% PNAS GRAPHICS SETUP- FROM THE PNAS STYLE LATEX FILE
-\RequirePackage{graphicx,xcolor}
-\RequirePackage{colortbl}
-\RequirePackage{booktabs}
-\RequirePackage{algorithm}
-\RequirePackage[noend]{algpseudocode}
-\RequirePackage{changepage}
-\RequirePackage[twoside,%
-				letterpaper,includeheadfoot,%
-				layoutsize={8.125in,10.875in},%
-                layouthoffset=0.1875in,%
-                layoutvoffset=0.0625in,%
-                left=38.5pt,%
-                right=43pt,%
-                top=43pt,% 10pt provided by headsep
-                bottom=32pt,%
-                headheight=0pt,% No Header
-                headsep=30pt,%
-                footskip=25pt,
-                marginparwidth=38pt]{geometry}
-\RequirePackage[labelfont={bf,sf},%
-                labelsep=period,%
-                figurename=Fig. ]{caption}
-\setlength{\columnsep}{13.5pt} % Distance between the two columns of text
-\setlength{\parindent}{12pt} % Paragraph indent
-
-%% Figure caption style
-\DeclareCaptionFormat{pnasformat}{\normalfont\sffamily\fontsize{7}{9}\selectfont#1#2#3}
-\captionsetup*{format=pnasformat}
-
-%%% GREYBOX AROUND FIG
-\definecolor{lightgray}{gray}{0.95}
-\newcommand\greybox[1]{%
-  \vskip\baselineskip%
-  \par\noindent\colorbox{lightgray}{%
-    \begin{minipage}{\textwidth}#1\end{minipage}%
-  }%
-  \vskip\baselineskip%
-}
-
-\newtcolorbox{paperbox}[1][]{
-    enhanced,
-    colback=white,
-    boxrule=0pt,
-    boxsep=0pt,
-    arc=0mm,
-    width=0.8\linewidth,
-    fuzzy shadow={0mm}{-4pt}{-4pt}{1mm}{black!30!white},
-    #1
-}
-
-% \externaldocument{Supp_Koopman-0808}
-\begin{document}
-
-\title{Data-Driven Modelling \& \\ Forecasting of Chaotic Systems}
-
-\author{B.L. Nortier} 
-\date{January, 2022}
-\date{}
-\maketitle
-
 \begin{abstract}
@@
     \end{abstract}
-
-  \tableofcontents
-
-\chapter{Introduction}\label{ch1}
-
-Experimenting with biological, physical, and artificial systems to generate a more informative dynamical model is a well-established practice in modern science.
-Traditional methods for modelling physical systems are based on laws of physics that are based either on empirical relationships or intuition.
-For systems that evolve with time, physical laws yield mathematical equations that govern how the quantities evolve with time.
-However, our world is much more complex than that which can be distilled into elegant equations.
-We do not fully understand many of the more complex systems, nor do they provide us with a good physical intuition of the underlying principles governing system dynamics.
-To complicate this, the underlying systems we observe often display sensitive dependence on initial conditions despite having highly similar initial conditions.
-This is because differing orbits diverge quickly and to such an extent that it becomes seemingly impossible to retrace their steps back to the original conditions.
-Even the presence of usually' negligible' computational noise or measurement error renders long-term pointwise prediction infeasible.
-There are many difficulties that one encounters while modelling such systems:
-\vspace{-8mm}
-\begin{enumerate}[noitemsep, label=\roman*.]
-  \item One may not have access to the complete states of the systems
-  \item The system may be described by functions that behave wildly because their graphs have a wild oscillatory behaviour, i.e., a large functional complexity.
-\end{enumerate}
-
-Models derived primarily from data can be classified into three categories: 
-\vspace{-8mm}
-\begin{enumerate}[noitemsep, label=\roman*.]
-  \item  Interpretable models (i.e., they establish relationships between internal physical mechanisms), 
-  \item Partially interpretable models capturing some modes of the dynamics, 
-  \item Non-Interpretable models, defined as such mainly due to the fact that they are defined on a different phase space that is usually high-dimensional. 
-\end{enumerate}
-
-Examples in the literature attempting to forecast data from such systems have tried many different approaches with differing degrees of success. 
-When we have complete access to states of the system, an ordinary differential equation can be obtained from data (\cite{brunton2016discovering, champion2019data} and \cite{small2002modeling,xu2006modeling}) wherein one could approximate the vector field by a library of functions to obtain interpretable models.
-
-Recent partially interpretable models available in the literature have been based on the Koopman operator (e.g.,[?, ?]) to employ observables mapping the data onto a higher-dimensional space.
-This makes the dynamics in the higher-dimensional space more amenable for approximation by a linear transformation.
-Such methods do not guarantee exact reconstructions for nonlinear models, and in practice, provide poor long-term consistency for a large class of systems.
-(Citation)
-\textbf{(Citation)}
-
-The non-interpretable models include the delay embedding and the machine learning algorithms.
-For example, one could learn a system conjugate to the underlying system by applying the Takens delay embedding~\cite{takens1981detecting} when one has "good" observations from the system; this learnt system could then be used to forecast the observed data.
-Takens delay embedding theorem ~\cite{takens1981detecting} and its various generalisations (e.g., \cite{sauer1991embedology, stark1999delay, gutman2018embedding}) establishes the learnability of a system constructed by concatenating sufficiently large previous time-series observations of a dynamical system into a vector (called delay coordinates).
-This then confirms the existence of a map on the space of delay coordinates equivalent (or topologically conjugate) to the underlying map from which the observed time-series was first obtained.
-Although topological conjugacy guarantees an alternate representation of the underlying system, the quality still depends on numerous parameters, making the comprehension of the dynamics unreliable (e.g., [?]).
-One reason for this fragility is that the embedded attractor in the reconstruction space is not always an attractor of the map learnt in the reconstruction space, despite unquestionably being an invariant set.
-When the embedded object is not explicitly known to be an attractor (as explained in Chapter~ \ref{ch3}), it can cause predictions to fail.
-
-Practically, the application of Takens embedding involves learning a map through some technique, and consequently one wishes that these would have low functional complexity\cite{manjunath2021universal}, i.e., functions with fewer oscillatory graphs.
-On the other hand, pure machine learning methodology processes temporal information (like the echo state networks \cite{jaeger2001echo, jaeger2004harnessing} by mapping data onto a higher dimensional space for further processing.
-Although they perform  well on forecasting some dynamical systems, they fail completely on others (Citations here) as there is often no guarantee that the right function was learnt during training.
-
-This thesis involves implementing and analysing non-interpretable models that can guarantee exact reconstruction. 
-The project work concerns the study and implementation of a method (\cite{manjunath2021universal}) that incorporates learning a function by mapping the data on to a higher dimensional space using what is called a driven dynamical system (See Chapter~ref{ch.4}).
-With a clear understanding of how the data is mapped onto the higher dimensional space, the method then permits the learning of a dynamical system topologically conjugate to that of the underlying system.
-Instead of linear regression as in the training of echo state networks \cite{jaeger2001echo}, deep learning methods are employed to learn the correct function.
-With slight modifications to the implementation in the paper (with Adriaan), we show that one can construct accurate non-interpretable models with the ability to reconstruct attractors from more hard-to-forecast systems like the double pendulum.
-(The forecasting of the time-series from a double pendulum has not been reported before.) Moreover, we also demonstrate that long-term statistical consistency is preserved.
-
-\emph{By solidifying the mathematical underpinnings of our theory, we hope to guarantee the ability to construct models with predictive power ranging from molecular biology to neuroscience.
-in the near future.
-(We can modify this sentence when the report is completed.)}
-
-The more intricate details of proofs are referred to where relevant throughout. This report is organised as follows: 
-\newline In Chapter~\ref{ch2}, we recall the definition of a discrete-time dynamical system, how a discrete-system arises from a flow of an ODE and then proceed to define the inverse-limit space and topological conjugacy of autonomous systems.
-\newline Chapter~\ref{ch3} introduces the problem of forecasting dynamical systems, states the Takens delay embedding theorem, and discusses various issues faced while forecasting. 
-\newline In Chapter~\ref{ch4}, a driven dynamical system is defined and discuss the properties of a specific class of driven dynamical systems that we make use of in this thesis/report/paper. 
-\newline Finally, in Chapter~\ref{ch5} we show the implementation of these forecasting methods, and conclusions are provided in Chapter~\ref{ch6}.
-
-\chapter{Discrete-time Dynamical Systems}\label{ch2}
-
-This chapter briefly describes a discrete-time dynamical system and what it means to exhibit chaos. We refer to \cite{devaney2018introduction, de2013elements} for more details.
-
-At its most elementary level, a dynamical system is just something that evolves deterministically through time.
-In the context of this thesis, deterministic refers to the fact that a system evolves according to specified rules rather than based on random events.
-Dynamical systems arise in a variety of situations.
-A continuous-time dynamical system describes the states for all values of the time.
-Specifically, if the motion of a pendulum in which the quantities such as the angular position and angular momentum are known at all times, then it is a continuous-time dynamical system.
-The equations of the dynamical system can take the form of one or more ordinary differential equations that determine the relevant quantities at any future time if we know the initial location and momentum.
-In ecology, discrete-time dynamical systems are widely used to model population growth.
-The model in this case is a function calculating the following generation's population given the population of the previous generation.
-If we know the starting population, we may once again calculate the population at any time in the future.
-
-Formally, a function $T: U \to U$, where $U$ is some set is  a \emph{discrete-time dynamical system} and its iterates $\{u,Tu,T^2u,\ldots\}$, where $T^n$ denotes the $n$-fold composition of $T$ with itself, describe the evolution of an initial condition $u\in U$ (Note that we frequently drop the brackets and denote $T(u)$ by $Tu$ so as to simplify notations).
-Also, we assume that $U$ is endowed with a metric $d$.
-
-Continuous-time dynamical systems modelled using ordinary differential equations can give rise to discrete-time dynamical systems.
-To see this, consider a differential equation $\dot{x} = f(x)$, $f: \mathbb{R}^n \to \mathbb{R}^n$, $n\in\mathbb{N}$ given to have a unique solution passing through
-each point $x\in\mathbb{R}^{n}$, call it $x(t)$ where $x(0)=x_0$.
-
-\begin{Definition}
-  [\bf Flow of an Equation] \label{Dfn_Flow}\rm
-  The flow of the equation  $\dot{x} = f(x)$  is defined to be a mapping $\varphi: \mathbb{R}^n \times \mathbb{R} \to \mathbb{R}$ where $\varphi(x_0,t)= x(t)$ for the solution $x(t)$ with $x(0)=x_0$ 
-\end{Definition}
-
-By fixing $t=K \in (0,\infty)$, we can define the \emph{time-$K$ map} as  $T(x):= \varphi(x_0,K)$, and it is easily verified that $T\circ T(x_0) = \varphi(x_0,2K)$.
-In general the $m^{\mbox{th}}$ iterate of $x_0$ under $T$ would be the value of the solution of the ODE evaluated at time $mK$ with the initial condition $x_0$, i.e.
-$\varphi(x_0, mK)$.
-Thus ordinary differential equations give rise to a discrete-time dynamical system by sampling the value of the solution $x(t)$ at time intervals $K$ units apart.
-
-A numerical discretization of a differential equation can also give rise to a discrete-time dynamical system.
-For instance, Euler's method approximates $\dot{x}(t)$ by $(x(t+h)-x(t))/h$; if $h$ is fixed throughout, the solution of a differential equation $\dot{x}=f(x)$
-can be approximated at the time instant $t+(m+1)h$ by iterating the equation 
-\begin{equation}
-  x(t+(m+1)h) = x(t+mh) + h f(x(t+mh))
-\end{equation}
-
-Adopting more succinct notation by replacing $x(t+mh)$ with $u_m$, we rewrite the above equation as
-\begin{equation}
-u_{m+1} = u_m + hf(u_m)
-\end{equation}
-or in even simpler terms as the discrete-time dynamical system with map $T(u) = u + hf(u)$, where $T$, $u$ and $f$ are understood to be as above.
-
-Of course, discrete-time dynamical systems need not always arise through a differential equation.
-Once again, we may consider the field of ecology, where discrete dynamical systems are often directly derived or assumed.
-\textbf{Citation}
-There is a school of thought that advocates discrete-time dynamical systems to be more natural for modelling real-world observations than differential equations.
-We refer the interested reader to \cite{saber2010introduction}.
-
-\section{Invariant Sets}
-
-A core concept in the study of dynamical systems is that of invariance. 
-\begin{Definition}
-  [\bf Invariant Set]\label{Dfn_InvariantSet}\rm
-  Given a discrete-time dynamical system $T: U \to U$, a subset $A \subset U$ is said to be an \emph{invariant set} if $T(A) =A$. 
-\end{Definition}
-
-We also define the orbit of a function $T$.
-\begin{Definition}
-  [\bf Orbit of $T$]\label{Dfn_Orbit}\rm
-  The orbit of $T$ is to be defined the sequence $\bar{u} = \{u_n\}_{n\in \mathbb{Z}}$ obeying the update equation, $u_{n+1}=Tu_n$, $n \in \mathbb{Z}$. 
-\end{Definition}
-
-Two examples of invariant sets include a fixed point where $Tu=u$ for  $u\in U$, and a periodic orbit, i.e., a set of iterates $\{u,Tu, T^2u,\ldots,T^pu\}$, where $T^{p+1}u=u$ for some $p\in\mathbb{Z}$.
-The entire space $U$ could be also be invariant.
-Consider for example the space $U=[0,1]$, where $Tu=4u(1-u)$ and then $U$ is invariant (as every $u\in{U}$ can be written as $u = 4x(1-x)$ for some $x\in{U}$).
-
-We may learn a great deal about the iterates of a dynamical system by considering the types of invariant sets of a discrete-time dynamical system.
-For example, if $U=[0,1]$ has map $Tu= u/2$, then the only invariant set is $\{0\}$, and all orbits approach this invariant set as time flows in the forward direction.
-Indeed, if a some non-zero invariant set (call it $B$) exists, then there is  some $r\in(0,1]\cap{B}$.
-But $r\notin{T(B)}$ since any orbit with initial value $r$ will be a decreasing sequence.
-Moreover, every orbit of T will be decreasing and therefore approach the value $0$ as $n\rightarrow\infty$.
-
-One may ask if every orbit approaches an invariant set?
-In general the answer is no, since for the dynamical system $T: \mathbb{R} \to \mathbb{R}$ defined by $Tu=2u$, any orbit that does not intersect the invariant set $\{0\}$, will not approach any invariant set.
-
-However,  when the space $U$ is compact, all orbits approach an invariant set.
-This is since the set of limit points of the orbit can be shown to be invariant \cite{de2013elements}.
-So, when $U$ is compact,  the $\omega$-limit set $\omega(u;T)$ of a point $u$ defined to the collection of limit points of the sequence $\{x,Tu,T^2u,\ldots\}$ is nonempty, and $\omega(u;T)$ is invariant.
-
-\begin{Example}
-  For the map $Tu=u^2$ defined on $[0,1]$ all orbits lie in the invariant set $\{1\}$ or else would would approach the invariant set $\{0\}$. 
-\end{Example}
-
-Invariant sets have various properties. Invariant sets can be attracting or repelling depending on how orbits in their vicinity behave. 
-Recall the examples $Tu =u/2$ and $Tu=2u$ defined on $\mathbb{R}$ (where $\{0\}$ “attracts" orbits) or the example, $Tu=u^2$ defined on $[0,1]$ (where $\{1\}$ “repels" orbits). 
-We are interested in attractive invariant sets since they capture the long-term dynamics as time increases. 
-In particular, we are interested in those invariant sets named attractors.
-
-\begin{Definition}
-  [\bf Attractor]\label{Dfn_Attractor}\rm
-  Let $T: U \to U$, where $U$  is a metric space with metric $d$. A compact subset $A \subset U$ is said to be an attractor if it satisfies the three conditions: 
-  \vspace{-8mm}
-  \begin{enumerate}
-	\item $A$ is invariant. 
-	\item $A$ is asymptotically stable, i.e., for every $\epsilon > 0$ and for all $u$ so that $d(u,A) < \epsilon$, we have $d(T^nu,A) \to 0$ as $n\to \infty$. 
-	\item $A$ has Lyapunov stability, i.e., for every $\epsilon > 0$  there exists a $\delta(\epsilon) > 0$ so that $d(u,A) < \delta$ implies $d(T^nu,A) < \epsilon$ for all $n\ge 0$.  
-\end{enumerate}
-\end{Definition}
-
-Indeed in our previous examples, it can be verified that the system  $Tu=u/2$ defined on  $\mathbb{R}$ and $Tu=u^2$ defined on $[0,1]$  the singleton set $\{0\}$ is an attractor.
-For the system,  $Tu=1-|2u-1|$ on $[0,1]$, one may easily verify that the only attractor is the entire space $[0,1]$.
-This follows from the fact that between any two points $u< v$ in $[0,1]$, and for any $a,b$ so that $u\le a < b \le v$, we can find an $n$ so that $T^n(a,b)=[0,1]$ (as would be explained later in this chapter).
-
-A dynamical system can have several attractors and may also be contained in another attractor.
-For the example, $Tu =u^2$ on $[0,1]$, both $\{0\}$ and $[0,1]$ are attractors.
-It is known that an attractor for a dynamical system always exists in a compact space.
-\textbf{(Cite).}
-
-The dynamics restricted to an invariant set can be complicated.
-For instance an invariant set could be just a single point or it could have an infinite set.
-If the invariant set is infinite, then complicated dynamics are possible.A particular well-studied phenomenon of such complexity gives rise to so-called  chaotic behaviour, studied in detail over the past fifty years.
-
-% If the dynamics on the attractor are somewhat complicated in the sense that the attractor cannot be decomposed further, and if the attractor is infinite then there is a possibility of complicated dynamics, and a particular well-studied phenomenon of such complexity gives rise to what is called a chaotic behaviour.
-
-%If the dynamics on an attractor cannot be decomposed further and the attractor is itself also infinite, then there is a possibility of complicated dynamics, and a particular well-studied phenomenon of such complexity gives rise to what is called chaotic behaviour.
-
-\section{Chaos}
-
-In the 1960s, several mathematicians and mathematically interested scientists independently discovered chaos in the mathematical sense.
-The meteorologist Edward Lorenz may have been the first to explain this phenomenon in his 1963 paper \cite{lorenz1963deterministic}.
-The notions of invariance, attractivity, and chaos may also be described for continuous systems, and Lorenz's system comprised of a system of differential equations.
-The narrative of Lorenz's discovery of chaos and the history of other forerunners in this subject is fascinating.
-We highly recommend James Gleick's book Chaos: The Making of a New Science~\cite{gleick2008chaos} for those interested.
-It clearly illustrates these experiences, explains why chaos was such a startling and crucial mathematical and scientific discovery and describes the underlying mathematical notions for non-specialists.
-
-There are several intricately different definitions of chaos available in the literature \textbf{(Cite)}, with each of them indicating some aspect of complexity.
-In practice, it is not possible to verify which definition or notion does the underlying dynamical system satisfies when only data is observed from it.
-We do, however, specifically recall the definition of chaos in the sense of Devaney \cite{devaney2018introduction,de2013elements} so as to understand some nuances behind the complexity.
-Devaney's definition of chaos has three requirements, with the first being the notion  of sensitive dependence on initial conditions.
-
-\begin{Definition}\rm
-  [\bf {Sensitive Dependence on Initial Conditions}]\label{Dfn_SDIC}\rm
-A dynamical system $T: U \to U$ is said to have sensitive dependence on initial conditions (SDIC) if there exists a $\delta > 0$ such that for every $u \in U$ and in every neighborhood of $u \in U$ there exists a $v\in{U}$ and an integer $N:=N{(u,v)}\in\mathbb{Z}$ such that $d(T^Nu,T^Nv)>\delta$.
-\end{Definition}
-
-It is vital to acquire a sense of this definition as it can easily be misunderstood.
-It is a common misconception to interpret SDIC as two close points ($u$ and $v$) that eventually become separated by a distance $\delta$ under iteration by $T$.
-But this is not true.
-To fully comprehend all the subtleties of the concept, one needs to discuss it more thoroughly by considering each sentence with care and attention, whereafter one may examine how they fit together to convey the idea of sensitive dependence.
-
-To this end, we make 3 remarks:
-\vspace{-5mm}
-\begin{enumerate}
-  \item First, the $\delta>0$ in the definition of SDIC is independent of $u$. 
-  \item Second, in every neighborhood of $u$, we may not necessarily find all points $v$ in the neighborhood distinct from $u$ that would separate from the forward iterates of $u$. 
-  \item  Finally, $N$ depends upon $u$ and $v$ chosen, and their iterates may not separate forever (i.e., for all $n>N$) and we allow their iterates to get arbitrarily close in the future. 
-\end{enumerate}
-
-To illustrate the concept of SDIC, we present two examples - one from Mathematics and the other from the field of Physics.
-
-\begin{Example} \rm
-  The logistic map (LM), a recursion relation of the form $x_{n+1}=rx_n(1-x_n)$ where $x_n\in[0,1]$ and $0<r\leq{4}$.
-It is of particular interest to us as it is classically used to illustrate chaos \cite{devaney2018introduction} for certain parameters.
-  In the graph below~\ref{fig:log_sdic} we plot the first 50 iterates of the LM for $r=4$ and two slightly different initial values $x_0$.
-
-  \begin{figure}[ht]
-    \includegraphics[scale=0.74]{_logistic_sdic.eps}
-        \centering
-        \captionof{figure}{Figure generated for Logistic Map $x_{n+1}=rx_n(1-x_n)$ with $r=4$ to exhibit the presence of SDIC.
-Plotted are the values $x_n$ against time $n$ for the the first 50 step with initial values $x_0=0.2$ (blue) and $x_0=0.20001$ (red).
-Initially the two trajectories overlap, but they diverge completely at $n=12$ whereafter they follow distinct paths.}
-        \label{fig:log_sdic}
-      \end{figure}
-
-  % For an intuitive sense of the fact that the LM exhibits SDIC, consider again the definition and substitute specific values.
-Let $\delta=0.05$ and for purposes of illustration consider the value $u=0.4$.
-The green band in~\ref{fig:log_sdic} shades the region $[u-\delta, u+\delta]$ and we can see that regardless the value $v$  and in every neighborhood of $u \in U$ there exists a $v\in{U}$ and an integer $N:=N{(u,v)}\in\mathbb{Z}$ such that $d(T^Nu,T^Nv)>\delta$
-  % specific point, say $u=0.4$.
-Draw a green neighbourhood $[0.35,0.45]$ of diameters $\delta=0.02$ around the point.
-It is easy to see that every neighbourhood around $u=0.4$ will contain some $v_1, v_2\in[0,1]$ and natural number $N_1, N_2 \in\mathbb{N}$ respectively s.t.
-the distance $d(T^Nu_1,T^Nv)>\delta=0.05$ .
-\end{Example}
-
-\begin{Example} \rm
-  A second example pertains to a simple physical system called the double pendulum - a pendulum with another pendulum attached to its head given an initial position and angular velocity.
-
-  Below are two figures denoting trajectories of a double pendulum's second head after some elapsed time.
-Depicted are three pendulum heads with equal angular velocity but differing slightly in initial position.
-As can be seen below, the trajectories diverge very quickly to become completely different.
-We discuss the double pendulum in greater detail in a later chapter (\ref{ch5}), but we mention the example here to provide a numerical example of a practical system exhibiting SDIC.
-
-\begin{figure}[ht]
-  \centering
-\includegraphics[scale=0.8]{_dp_sdic.eps}
-    \captionof{figure}{Three double pendulum heads with equal angular velocity and initial angles differing by 0,025 radians}\label{fig:dp_sdic}
-  \end{figure}
-
-\end{Example}
-
-The second part in Devaney's definition of chaos concerns topological transitivity.
-
-\begin{Definition}
-  [\bf {Topologically Transitive}]\label{Dfn_TopolTrans}\rm
-	A dynamical system $F: U \to U$ is topologically transitive if for any pair of nonempty open sets $E_1$ and $E_2$ there exists a $n\in\mathbb{N}$ such that $T^n(E_1) \cap E_2 \not= \emptyset$. 
-\end{Definition}
-
-Topological transitivity implies that iterates of an open set of initial conditions get mixed up with other open sets.
-On a compact metric space, one may show that topological transitivity also implies the existence of a point whose forward iterates are dense \cite{de2013elements}; or in other words, the orbit going through this point will be dense in the compact metric space.
-In fact, it is topologically more likely that the choice of an arbitrary point will be one whose iterates are almost dense.
-
-\begin{Definition}
-  [\bf {Meager Set}]\label{Dfn_Meager Set}\rm
-A subset of a topological space $U$ is said to be a meager set if it can be written as a countable union of sets of with empty interior. The complement of the a meager set is set to be a residual set.
-\end{Definition}
-
-To be more precise (see \cite{de2013elements}), for a discrete-time dynamical system on a compact space, the set of points with dense iterates are residual, and they are typical or likely to be observed in practice.
-In this project, when data is observed from a topologically transitive system, we assume it arises from a dense orbit.
-
-We may now define the notion of Chaos as formulated by Devaney\cite{devaney2018introduction}.
-\begin{Definition}
-  [\bf {Devaney's Chaos}]\label{Dfn_ChaosDec}\rm
-	A dynamical system $T: U \to U$ is said to exhibit chaos in the sense of Devaney if it satisfies the three properties:
-	\vspace{-5mm}
-  \begin{enumerate}
-		\item $T$ has SDIC.
-		\item $T$ is topologically transitive.
-		\item The set of periodic points of $T$ are dense in $U$. 
-	\end{enumerate}
-\end{Definition}
-
-\begin{Example} \rm
-  The standard tent map $Tu=1-|2u-1|$ defined on $[0,1]$ is a well-known example of a dynamical system satisfying these three properties.
-
-  We now reason as to why this is true.
-The graph of the map $T$ is piecewise linear with two straight lines, one connecting the points $(0,0)$ and $(\frac{1}{2},1)$ and the other connecting $(\frac{1}{2},1)$ with $(1,0)$.
-They form a so-called tent with base centered at $u=1/2$.
-The graph of the map $T^2$ comprises two symmetric tents with their base centered at $1/4$ and $3/4$.
-See Figure \ref{fig:T2tentmap}
-
-  \begin{figure}[ht]
-    \includegraphics[scale=0.6]{_tentmap_2.eps}
-        \centering
-        \captionof{figure}{Graph of the $T^2$ map with bases at $x=0$, $x=\frac{1}{2}$ and $x=1$ respectively.}\label{fig:T2tentmap}
-  \end{figure}
-
-  In general, the graph of  $T^k$ contains $2^{k-1}$ tents.
-  Given any point $u$ and an open neighborhood $W \subset[0,1]$ of $u$, we can find $k\in\mathbb{Z}$ large enough so that $T^k(W) = [0,1]$ because we can accommodate a tent whose base is contained in $W$.
-% \ednote{B: This last phrase doesn't make sense to me.
-M: The number of tents grow with $k$ and hence this is obvious; I explained in one of our meetings.
-Does it make sense?}
-  This implies the existence of two points $w_1$ and $w_2$ in $W$ so that $d(T^kw_1,T^kw_2)= 1$, and by the triangle inequality  at least one of the inequalities $d(T^ku,T^kw_1)> \delta$ or $d(T^ku,T^kw_2)> \delta$ hold when $\delta\in (0,\frac{1}{2})$.
-So $T$ has sensitive dependence on initial conditions.
-
-  Next, let $W_1$ and $W_2$ be two nonempty open sets.
-Given any open set $W_1$, we can find a $k\in\mathbb{N}$ large enough so that $T^k(W_1) = [0,1]$ because we can accommodate a tent with base contained in $W$.
-So $T^k(W_1) \cap W_2 \not=\emptyset$, and thus $T$ has topological transitivity.
-
-  Finally, for every open interval $(a,b)$, the graph of $T^k$ intersects the graph of the identity map on $[0,1]$.
-This follows from the fact established above that there exists a $k\in\mathbb{N}$ such that $T^k(a,b) = [0,1]$.If the intersection point has coordinates of the form $(p,p)$, the point $p$ is then a fixed point of $T^k$ and therefore also a periodic point of $T$.
-Since we have found a periodic point in the interval $W$,  the set of periodic points are dense in $T$.
-  This follows because any fixed point of $T^k$ is also a periodic point, and we have found this in an arbitrary interval.
-
-\end{Example}
-
-\section{Conjugacy}
-
-We now turn our attention to the subject of conjugacy that describes when two dynamical systems are equivalent dynamically.
-
-To show topological similarity (or sameness) between two metric or topological spaces, one needs to establish a homeomorphism between the two spaces. 
-However, in the study of dynamical systems defined on two spaces, establishing a homeomorphism does not indicate that the systems are dynamically related in any way.
-For instance, the maps $Tu=u^2$ and $Tu=1-|2u-1|$ defined on $[0,1]$ have totally different behaviour. 
-Therefore, to find dynamically similar systems, one must establish a dynamical equivalence that we illustrate in a simple commutativity diagram below~\ref{eqn_conjugacy}.
-
-% Giving errors
-\begin{equation}  \label{eqn_conjugacy}
-%    \[ 
-    \psset{arrows=->, arrowinset=0.25, linewidth=0.6pt, nodesep=3pt, labelsep=2pt, rowsep=0.7cm, colsep = 1.1cm, shortput =tablr}
- \everypsbox{\scriptstyle}
- \begin{psmatrix}
-U & U\\%
-V & V.
- %%%
-%  \ncline{1,1}{1,2}^{T} 
-%  \ncline{1,1}{2,1} <{\phi}
-%   \ncline{2,1}{2,2}^{S}
-%  \ncline{1,2}{2,2} > {\phi}
- \end{psmatrix}
-% \]
-\end{equation}
-
-%Composite functions
-To understand the diagram, we note that if we travel "right and then down," the diagram instructs us to use $T$ (top arrow right) first, followed by $\phi$.
-(right arrow downwards).
-Consequently, the pathway amounts to finding $\phi(T(u))$.
-If we proceed "down and then right," the diagram instructs us to apply $\phi$ (left arrow downwards) first, and then apply $S$ (bottom arrow right).
-This then amounts to finding  $S(\phi(u))$.
-When the relationships in this diagram hold, we say $\phi(T(u))= S(\phi(u))$, and we formally denote it as $\phi \circ T=S\circ \phi$.
-
-\begin{Definition}\rm  
-[\bf {Conjugacy \& Semi-Conjugacy}]\label{Dfn_Conjugate}\rm
-   Consider the dynamical systems $T:U\to{U}$, $S:V\to{V}$ and suppose the relationship $\phi \circ T=S\circ \phi$ holds.
-If $\phi:U\to{V}$ is a homeomorphism, then $S$ is said to be conjugate to $T$ and $\phi$ is called a conjugacy with the underlying spaces.
-If we relax the criterion on $\phi$ and merely require $\phi$ to be continuous and surjective where $\phi:U\to{V}$, then $\phi$ is a semi-conjugacy between $T$ and $S$.
-\end{Definition}
-
-\ednote{B: Does defining explicitly the domain and codomain of $\phi$ make it clear enough?
-M: Saying $\phi$ is then a semi-conjugacy between $T$ and $S$ could be ambiguous as it does not specify which system is on the top in the commutativity.
-So please edit accordingly.
-M: Very often they are implicitly understood...
-so no worries, and I have edited.}
-When $S$ is conjugate to $T$, the dynamics of the two systems are in some way ''dynamically equivalent''.
-Specifically, they are in one-to-one correspondence with one another.
-However when $S$ is semi-conjugate to $T$ with $\phi:U\to{V}$ a many-to-one mapping, the dynamics on $V$ provide merely a coarse-grained description of the dynamics on $U$~\cite{de2013elements}.
-When $S$ is semi-conjugate to $T$, it is also common to call $S$ a \emph{factor of $T$}, or conversely that $T$ is an \emph{extension of $S$}.
-In essence, an extension (e.g., \cite{de2013elements}) is a more extensive system capturing all of the essential dynamics of its factor.
-
-It is a very hard or nearly an impossible task to establish the existence of a conjugacy or a semi-conjugacy $\phi$ between two systems \cite{devaney2018introduction}.
-However, one can verify that a function $\phi$ satisfies the commutativity diagram \ref{eqn_conjugacy}.
-
-\begin{Example}\rm
-  For example $\phi(x)=\sin\big(\frac{\pi}{2}x\big)^2$ is a conjugacy between two systems $Tu=1-|2u-1|$ and $Sv=4v(1-v)$  defined on $U=[0,1]$.
-
-  Indeed, $\phi:[0,1]\to[0,1]$ is a homeomorphism and by employing simple trigonometric identities, it is easily proved that $\phi$ satisfies~\ref{eqn_conjugacy}
-  \[
-    \begin{aligned}[t]
-      (\phi\circ{T})u 
-                &=\sin^{2}\Big(\frac{\pi}{2} - \frac{\pi}{2}|2u-1|\Big) = \sin^{2}(\pi{u})=\Big(2\sin(\frac{\pi}{2})\cos(\frac{\pi}{2})\Big)^{2}=4\sin^{2}(\frac{\pi}{2}u)\Big(1-\sin^{2}\frac{\pi}{2}u\Big) \\
-                                    &=({S}\circ\phi)u.
-    \end{aligned}
-\]
-And it follows that $\phi$ is a conjugacy between the two systems.
-\end{Example}
-
-By establishing that two systems are conjugate (semi-conjugate) to one another, we have also shown that one may choose to work with one system instead of the second and still be guaranteed to obtain information on the latter.
-It is instrumental during the process of forecasting the future evolution of dynamical systems.
-
-\chapter{The Learning Problem}\label{ch3}
-
-In this chapter, we acquaint ourselves with the question of forecasting dynamical systems with unknown underlying structures, state and discuss the shortcomings of the Takens delay embedding theorem and discuss various issues faced while forecasting.
-
-Consider a relatively simple learning problem: 
-Given the sequence $(u_0, u_1, \ldots, u_m)$ for $m\in\mathbb{Z}$, a finite segment of an orbit of the map $T$ where $T$ is defined by the update equation $u_{n+1} = Tu_n$, forecast the values $u_{m+1}, u_{m+1}$ where the map $T$ is unknown, given that $u_m$.
-
-A practical example of this would be the subsequent scenario: given the time-sequential coordinates of an object moving in space, predict the future positions of that object.
-We are very rarely, if at all, presented with a problem where the entire state information is available to us in the form $u_n$ at some timestep $n\in\mathbb{N}$.
-Consider then a more complicated learning problem:
-
-Suppose we only have the observations $\theta(u_0), \theta(u_1), \ldots, \theta(u_m)$ of the true system states $u$ in an unknown dynamical system $(U,T)$ and we wish to predict the values $\theta(u_{m+1}), \theta(u_{m+2})$ and so forth.
-
-First we establish the method of Takens delay embedding. 
-In this method, one considers a discrete-time dynamical system defined on a 'nice' space (a smooth manifold\footnote{A manifold is a special topological space that is locally homeomorphic at each point to an open set of $\mathbb{R}^d$} that can be obtained as time-$K$ map of a continuous time-dynamical system, concepts formalised below.
-We make observations from such a system, i.e., we observe the evolution of $\theta(w_0)$ (where $w_0$ represents the initial state) by examining  a finite set of values of $u_n :=  \theta(w_n) = \theta(Tw_{n-1})$.
-The sequence $\{u_n\}$ represents a scalar time-series and intuitively $\theta$ represents a probe inserted into a bigger system which is itself only measuring/extracting a small part of the greater system state at time $t=n$.
-Consider, for example, a thermometer erected to measure the ambient temperature in a local village.
-This measurement function, the thermometer, is capturing only a single aspect, the temperature, of a much grander dynamical system entailing the current weather of the surrounding area.
-Even more than that though- it is measuring a miniscule part of the global weather system.
-
-%Given the sequence $\{u_1, u_2, \ldots, u_n}$, we define the variable \newline $y_n:=[\theta_n, \theta_{n-\tau}, \ldots, \theta_{n-(2d-2)\tau}, \theta_{n-(2d-1)\tau}]^{T}\in\mathbb{R}_{d}\times{1}$ where $\tau$ represents the lag and $d$ the dimension of the attractor.
-(We shall return to these terms shortly and their discussion may be put on hold for the time being).
-
-Hence, when the actual state $u$ of a system is seldom, if ever, fully known.
-the most one can do is to insert probes into a system to obtain partial information through the measurements taken.
-Moreover, the process of taking a measurement itself introduces two additional aspects which complicate the problem:
-\vspace{-8mm}
-\begin{enumerate}[noitemsep, label=\roman*.]
-  \item A series of measurements over a specific time interval is inherently an orbit if discretisation  of the underlying continuous-time dynamical system.
-(Hence why we restricted our attention to discrete-time systems in the preceding section)
-  \item A probe will never be entirely accurate, and so the act of measurement introduces a certain measure of numerical noise/inaccuracy (and so we could consider some noise to be involved).
-\end{enumerate}
-
-Takens embedding theorem does not take into account the noise, and for describing it, we construct a multidimensional observable using the method of stacking previous observations, i.e., we create delay-coordinate map defined by
-$\Phi_{k,\theta}(w) := (\theta(T^{-k}w)\ldots,\theta(T^{-1}w),\theta(w))$.  
-The concept of an observable is understood in the sense of an observable as originally introduced by~\cite{takens1981detecting, genericObservableAeyels} to refer to a probe or measurement function inserted into the system.
-Takens' theorem, in essence, refers to the fact that for a sufficiently large $k$, we can define a dynamical system on the space $\mathbb{R}^{k+1}$ with states evolving as $\Phi_{k,\theta}(w)$, $\Phi_{k,\theta}(Tw)$, $\Phi_{k,\theta}(T^2w)$, etc.
-This dynamical system is topologically conjugate to the unknown underlying system $(W,T)$.
-We recall the Takens delay embedding theorem next.
-
-%We may then ask ourselves the question: Is it in any way possible to retain information about the system state $x(t)$ in this temporal data-series $\theta(x(t))$?
-The answer is easily yes if T is known.
-However, neither this, nor even the exact function $\theta$ is available.
-
-\section{Takens Embedding Theorem}\label{sect_Takens}
-
-We define first the concepts of homeomorphism and embedding:
-\begin{Definition}\rm
-  [\bf {Homeomorphism}]\label{Dfn_homeo}\rm
-  A homeomorphism is a function $f:Z\rightarrow Y$ between two topological spaces $Z$ and $Y$ that is continuous, bijective and has a continuous inverse. 
-\end{Definition}
-
-\begin{Definition}
-  [\bf {Embedding}]\label{Dfn_embed}\rm
-  Consider a homeomorphism $f:Z\rightarrow Y$ for $Y\subset X$. $Z$ is said to be embedded in $X$ by $f$.
-\end{Definition}
-
-Takens Theorem states a result establishing a relationship between the observed and underlying dynamical systems by showing that the concatenation of a sufficiently large number of previous observations into a vector will, under certain conditions, generate a map between the vectors from the respective systems.
-We formulate the theorem from \cite{takens1981detecting}.
-
-\begin{Theorem} 
-	[\bf Takens Embedding Theorem (adopted from \cite{takens1981detecting}] \label{Thm_Takens}
-         Let $W$ be a compact manifold of dimension $m$, and $d\ge m$ so that $2d$ is an integer. It is a 
-            generic property for the pair $(T, \theta)$,  where $T:W \to W$ is
-            a smooth diffeomorphism, and $\theta:W \to \mathbb{R}$ a smooth function, the map $\Phi_{2d,\theta}:W \to \mathbb{R}^{2d+1}$ defined on $W$ by 
-            $\Phi_{2d,\theta}(w) := (\theta(T^{-2d}w)\ldots,\theta(T^{-1}w),\theta(w))$
-            is a diffeomorphic embedding; by `smooth' we mean at least $C^2$.
-Consequently, there exists a map $F_\theta: \Phi_{2d,\theta}(W) \to \Phi_{2d,\theta}(W)$ defined by $$F_\theta: (\theta(T^{-2d}w),\ldots,\theta(T^{-1}w),\theta(w)) \mapsto
-            (\theta(T^{-2d+1}w),\ldots,\theta(w),\theta(Tw))$$
-           so that $(W,T)$ is topologically conjugate to 
-            $(\Phi_{2d,\theta}(W), F_\theta)$.    
-\end{Theorem}
-
-By generic we mean a residual set on a certain topology on appropriate spaces of functions (that we do not describe here) \textbf{(Cite)}. 
-By $C^2$, we make reference to a twice-differentiable function  with a continuous second derivative.
-In our scenario (and this remains important throughout) the input space $U$ is considered the attractor.~\label{attractor_U}
-
-Below we reproduce Figure 1 from \cite{Supp}.
-
-%Some explanation is beneficial.
-Takens establishes a delay-coordinate map $\Ftheta$ defined in \ref{eqn_takens} for $T$ the flow (\ref{defn_flow}) defined by the update equation  $Tu_n=u_{n+1}$ and $theta$ our measurement function.
-When confining a dynamical system to the manifold U (in which the underlying system is contained), Takens showed that if certain smoothness conditions are satisfied on $T$ and $\theta$, then the delay-coordinate map $\Ftheta$ embeds $U$ in the reconstruction space $R^{2d+1}$ \textbf{we haven't used the term reconstruction space yet, so this is falling out of the blue}, for almost every choice of measurement function $\theta$, the observable.
-By almost every choice here, we mean that we exclude some observations which do not give any information.
-To see a more precise discussion, see \textbf{xxx} for we do not concern ourselves with a further elaboration on this report.
-
-%Alternatively we may also make the statement: Takens' Embedding Theorem guarantees that almost all dynamical systems can be reconstructed from just one noiseless observation sequence \textbf{this isn't obvious from the theorem's formulation, is it?} i.e.
-for a great number of possible observation functions $\theta$, $\Ftheta$ preserves the topology of U.
-\textbf{'preserving the topology' is not something we've touched on before.
-Perhaps I should add to previous discussion so logic flows seamlessly here}.
-
-\begin{figure}[ht]
-  \includegraphics[scale=0.25]{_takensmap.eps}
-  \centering
-  \captionof{figure}{Schematic of Takens Embedding Theorem's usage and impact of iterates slipping from the attractor when the approximation $\tilde{F}_\theta$ of $F_\theta$ is learnt.}
-  \label{fig:takensmap}
-\end{figure}
-
-If we can learn the map $F_\theta$, from sufficient number of data points of $\{\phi_{2d,\theta}(w_n)\}$, then we also know how to forecast how $\theta(w_n)$.
-
-%Once again we pause and consider the graph to consolidate our understanding of the conjugacy (\textbf{are we being too simplistic?}).
-If we are in state $u$ and then evolve towards state $Tu$ before embedding into $\mathbb{R}^{2m-1}$, then this is equivalent to first embedding into $\mathbb{R}^{2m-1}$ and then evolving through the map $\Ftheta$, i.e.
-$\Phi_{2d,\theta}\circ{F}_{\theta} = T\circ\Phi_{2d,\theta}$.
-The map $F_theta$ is a homeomorphism as well.
-
-%Thus information about U can be retained in the time series' (or observation's) output.
-By preserving the topology on the manifold $U$ in the reconstruction space $X$ (\textbf{not discussed before}), we are also guaranteed the preservation of topological invariants of the manifold, of which dimensionality is one such invariant.
-(We note here that dimensionality will again be considered later in this project).
-
-\section{Practical Limitations to Takens' Theorem}
-Takens' Embedding Theorem is a powerful result, and it provides compelling reason to believe that one could conceivably reconstruct a system conjugate accurately to the underlying system. 
-Nonetheless, it does present some severe practical limitations:
-\vspace{-5mm}
-\begin{enumerate}
-\item Even supposing that we could indeed find $F_\theta$ , our \emph{approximation} of  $F_\theta$ is a map from a larger set $\mathbb{R}^{2d+1}$ containing the embedded attractor.
-There are, however, no theoretical guarantees that $F_\theta$ will retain  $V=\Phi_{2d,\theta}(W)$  as an attractor although $W$ could itself be an attractor.
-\item Takens theorem is stated only for noiseless observations.
-Due to noise $\epsilon_n$ the delay vector $\Phi_{2d,\theta}(w_n) + \epsilon_n$  may lie outside $V$.Furthermore, due to the chaotic nature of the underlying system (i.e.
-the fact that it has SIDC), the evolution of $\Phi_{2d,\theta}(w_n) + \epsilon_n$ under the map $F_\theta$ could move out of $V$ completely.
-This problem can be overcome  by using a driven dynamical system with some properties, and we discuss this in Chapter\ref{ch3}.
-\end{enumerate}
-
-% The Takens embedding theorem does not guarantee global dissipativity.
-% \ednote{B:Links with~\ref{subs_LearnGamma} in discussion of dissipativity of $\Gamma$.
-Sir, I'm not sure how to write this here - I think it a beneficial point to make so as to set the stage for the later discussion in ~\ref{subs_LearnGamma}, but I am unsure there as well.}
-
-Though not a uniquely related to the Takens embedding theorem, we do opt to make mention of another requirement for our work. 
-The methods described here are applicable for data originating from a surjective map. 
-This follows from the fact that when $T$ maps the space $U$ into a proper subset $B$ of$ $U, there exists some point in $U - A$ that does not have a preimage and we conclude that $T$ is not surjective.
-An example of this would be the case when a system displays energy-loss.
-Consider, for instance, consider the dying oscillations for a damped pendulum.
-(An account of greater depth  is provided in~\ref{ch5})
-
-This requirement is not overly restrictive, since if a map is topologically transitive on a subspace it has to be surjective on it, since if it were to be not surjective there would be an open set $V$ that does not intersect with any future iterates of an open set  $U$ disjoint from it.
-\ednote{B: I remember you saying this in one of our sessions, but don't have a source for it.
-Do you have a specific reference?
-M: Follows from the transitivity definition.
-It is enough}
-
-\chapter{Driven Dynamical Systems to Forecast Problems}\label{ch4}
-
-In this chapter, we discuss results on the mapping of temporal data obtained from a discrete-dynamical system onto a different space through the notion of a driven dynamical system. 
-We also consider the conditions a driven system should possess to avoid adding distortion to its state-space representation.
-We then describe the \emph{single-delay dynamics} (SDD) of the system and finally recall conditions on the driven system so that the SDD are conjugate (or at least semi-conjugate) to the underlying system.
-The SDD can then be used to forecast and reconstruct the underlying system.
-
-\section{Nonautonomous and Driven Dynamical Systems}
-
-\begin{Definition}
-  [\bf Nonautonomous Dynamical System (adopted from~\cite{Manju_ESP}))] \label{Dfn_NDS}\rm
-  A nonautonomous dynamical system (NDS) on a space $X$ is simply a dynamical system comprising a family of maps $\{g_n\}_{n \in \mathbb{Z}}$ where $g_n(\ldotp):=g(u_n,\ldotp):X\to{X}$ is continuous.
-Each $g_n$ arises as a consequence of an input $u_n$ from the input space $U$, a topological space.
-\end{Definition}
-
-We immediately recall the figure first defined in \eqref{eqn_conjugacy}  and remind ourselves of our ultimate goal to learn a system that is conjugate to an underlying or unknown dynamical system $(U,T)$ by making use of the measurements obtained from  the unknown system.
-
-%#####################################
-\begin{center}
-\psset{arrows=->, arrowinset=0.25, linewidth=0.6pt, nodesep=3pt, labelsep=2pt, rowsep=0.7cm, colsep = 1.1cm, shortput =tablr}
-\everypsbox{\scriptstyle}
-\begin{psmatrix}
-U & U\\%
-V & V.
-%%%
-%  \ncline{1,1}{1,2}^{T} 
-%  \ncline{1,1}{2,1} <{\phi}
-%   \ncline{2,1}{2,2}^{S}
-%  \ncline{1,2}{2,2} > {\phi}
-\end{psmatrix}
-%#####################################
-\end{center}
-
-To find a function $\phi$, the need now arises to consider a so-called driven dynamical system, a special case of a nonautonomous dynamical system.
-Two spaces are considered: input is taken from an input space $U$ and the state of a space $X$ is updated.
-(Previously, only the state was considered.)
-We do this to mimic the true conditions.
-By accounting for an extrogenous input sequence in $U$ which can also influence the  system-state at a specific time-step, it produces more general models than an autonomous system.
-
-\begin{Definition}
-  [\bf Driven, Compactly Driven Dynamical System] \label{Dfn_DDS} \rm
-A driven dynamical system comprises two topological metric spaces $U$, $X$ and a continuous function  $g:U\times{X}\to{X}$ where $g(u_n, x_n)=x_{n+1}$.
-If the input space $U$ is compact, we refer to the system as compactly driven. 
-\end{Definition}
-
-The update equation $x_{n+1} = g(u_n,x_n)$ for $n \in\mathbb{Z}$, input $u_n$ from $U$ and state $x_n$ belonging to $X$ (a compact space) generate the dynamics on X. 
-Abbreviated, we refer only to the \textit{driven system $g$}, where all other entities are understood implicitly.
-
-Notably, a nonautonomous dynamical system can be generated from $U$; any input $\overline{u}$, a bi-infinite sequence from $U$, gives rise to the sequence of self-maps $\{g(u_n, \cdot)\}_{n\in\mathbb{Z}}$ contained in $X$.
-Physically, one may think of this bi-infinite sequence as referring to a system that has been running for an extended period at the time of the first measurement taken from the system (alternatively, the first time a probe is inserted into the system to take an observation)
-
-\begin{Definition}
-  [\bf Entire Solution] \label{Dfn_Soln} \rm
-  A sequence $\{x_n\}_{n\in\mathbb{Z}}\subset X$ is called an entire solution (or simply a solution) of the driven system  $g$ with input $\overline{u}$ when it satisfies 
-  \[g(u_{n-1}, x_{n-1})=x_n\] for all $n\in\mathbb{Z}$
-\end{Definition}
-
-It is important to emphasise that a sequence $\{x_n\}$ satisfying the update equation above can only be a solution if $x_n\in{U}$ holds for all $n\in\mathbb{Z}$ and just for $n>0$.
-Consider the example below.
-
-\begin{Example} \rm \label{ex_halfux}
-  The only solution  $\{x_n\}_{n\in\mathbb{Z}}$ to the driven system  $g(u,x)=\frac{ux}{2}$, where $X=[0,1]$, $U=[0,1]$,  is the zero solution $x_n\equiv0$.
-  To see this, consider any $x_n=a\in[0,1]$ where $a\neq{0}$.  Let $\overline{u}\in{U}$ be an non-zero constant sequence , say $u_n=0.5$. 
-  The driven system may be rewritten as $x_{n-1}=\frac{2x_n}{u_{n-1}}=4x_n$ and the  iterates of $x_n$ in backward time will increase by a factor of 4 at each timestep. 
-  Thus for some $m\leq{n}$,  $1<x_m$ i.e. $x_m\notin{X}$. So ${\{x_n\}}_{n\in\mathbb{Z}}$ is not a solution and it follows that the only possible solution is the zero solution.
-\end{Example}
-% \ednote{Nicely written! Thank you}
-
-A system may also have multiple solutions, as evidenced in the example below.
-
-\begin{Example}\label{Ex_exp} \rm 
-  Consider the driven system $g(u,x)=x^2$ for $X=[0,1]$, $U=\mathbb{R}$.
-The system has an uncountable number of solutions, as there exists a solution for every $x\in{X}$ which also passes through the point $x$ and $lim_{n\to\infty}x_n=0$, $lim_{n\to{-}\infty}x_n=1$.
- The proof is deferred to immediately after the next paragraph (see~\ref{rem_proofEx}).
-\end{Example}
-
-As the solutions to a driven system are considered an important entity, we next identify a subspace $X_U$ of $X$ that contains all possible solutions.
-To realize such a subspace of a driven system $g$, the concept of a reachable set is defined.
-
-\begin{Definition}
-  [\bf Reachable Set]\label{Dfn_ReachableSet}\rm
-The reachable set of a driven system $g$ is exactly the union of all the elements of all the (entire) solutions, i.e., 
-\[X_U :=\Big \{x \in X:  x = x_k \mbox{ where  $\{x_n\}$  is a solution for some  $\bar{u}$} \Big \}.\]
-The set of all reachable states at a specific time $n$ for input $\overline{u}$ is denoted by $X_n(\overline{u})$
-\end{Definition}
-
-The reachable set itself can defined independently from $g$ possessing the ESP/USP.
-We note that $x\in{X_n(\overline{u})}$ if and only $g$ has a solution $\{x_k\}$ for $x_n=x$ and input $\overline{u}$.
-\textbf{(Cite)}.
-
-The set $X_n(\overline{u})$ is precisely the set of points $x$ through which some entire solution $\Psi$ passes at time $n$. We can give an alternate formulation of the set $X_n(\overline{u})$. 
-Now suppose, we for some fixed input 
-$\overline{u}$, we define $g_i = g(u_i,\cdot)$ for all $i\in \mathbb{Z}$, then the set of states
-
-\begin{equation} \label{eqn_association}
-X_{n,i}(\overline{u}) := g_{n-1} \circ \cdots g_{i+1} \circ g_i(X).
-\end{equation}
-
-The set $X_{n,i}(\overline{u})$ being the image of a finite composition of
-continuous maps is compact whenever $X$ is compact. Also $X_{n,i}(\overline{u})$ is
-nonempty.  Further, $X_{n,i}(\overline{u}) \supset X_{n,i-1}$. Hence $\bigcap_{i<n}
-X_{n,i}(\overline{u})$ is a nested intersection of closed nonempty subsets, and
-whenever $X$ is compact, the intersection is nonempty. We say $g$ is a topological contraction if $\bigcap_{i<n}X_{n,i}(\overline{u})$ is a singleton subset of $X$ for each $\overline{u}$.
-
-It turns out that $\bigcap_{i<n}X_{n,i}(\overline{u})$ is identically equal to $X_n((\overline{u})$ that we had defined earlier.
-This will lead to a result (for.
-e.g., \cite{manjunath2013echo}) that  \textit{$g$ being a topological contraction is equivalent to the existence of a exactly one entire solution.}
-
-% \begin{Definition}
-%   [\bf Topological Contraction]\label{Dfn_TopContr}\rm
-%   A function $g:U\times{X}\to{X}$ is a topological contraction if for all $n\in\mathbb{Z}$ and all $\overline{u}\subset{U}$, $X_n(\overline{u})$ is a singleton subset of $X$.  
-% \end{Definition}
-
-\begin{Remark}
-  [\bf Proof of Example~\ref{Ex_exp}] \label{rem_proofEx} \rm
-  For the driven system,  
-  $g(u,x)=x^2$ the evolution does not depend on the input at all.
-Hence $g(u,x)$ can be written as a single map $f(x)=x^2$ which is a homeomorphism on $[0,1]$.
-A left-infinite orbit is defined and it converges to $1$ while the right-infinite orbit converges to $0$.
-Or in other words the iterates of $f^{-1}$ converges to $1$ while the iterates of $f$ converge to $0$
-\end{Remark}
-
-Thus far, it has been demonstrated that a system may have one or more solutions; one may ask if a driven system always has a solution and, if so, whether it satisfies specific properties such as uniqueness.
-Should the driven system be compact, existence follows immediately, as shown in the following result.
-
-\begin{Theorem}\label{Thm_CompactExistence}
- Let $g$ be a driven system.  If $X$ is compact then for each input $\overline{u}$, there exists at least one solution to the driven system $g(\ldotp, x)$
-\end{Theorem}
-\begin{proof}
-  May be found in \cite{kloeden2011nonautonomous, manjunath2014dynamics, manjunath2013echo}
-\end{proof}
-
-% One may easily construct many systems with trivial solution-sets, such as $g(u,x)=x$ which has only the constant solution $x$ and so for $U=[-1,1]$, the system would have no solution if $|x|>1$.
-To refine the scenario, we consider only systems with unique solutions.
-
-\section{Unique Solution Property}
-
-\begin{Definition}
-  [\bf Unique Solution Property] \label{Dfn_usp}\rm
-  A driven system $g$ is said to have the Unique Solution Property (USP) if for each input $\overline{u}$ there exists exactly one solution. 
-  Alternatively we may formulate the USP as follows: $g$ has the Unique Solution Property if there exists a well-defined map $\Psi:{U}\to{X}$, with $\Psi({\overline{u}})$ denoting the unique solution.
-\end{Definition}
-
-The USP is equivalent called  the echo state property  \cite{Manju_ESP}, and systems that satisfy this property have other important and useful properties \cite{grigoryeva2018echo, grigoryeva2019differentiable, manjunath2020stability} for information processing.
-One of the first results obtained after defining the USP is that every solution will attract different initial conditions towards the component parts of the solution.
-If $g$ has the USP, then any solution to $g$ is also a uniform attractor in nonautonomous dynamical systems literature~\cite{Manju_Nonlinearity}.
-The discussion on nonautonomous attractors is beyond the scope of this project, and we refer the reader to \cite{Manju_ESP, esann2012ids} for further reading.
-
-% Paraphrased.
-%Additionally, $g$ having the USP is in our context also equivalent to $g$ being a topological contraction \cite{manjunath2021universal}, or alternatively also that $g$ exhibits the Echo State Property (ESP).
-
-In the majority of Reservoir Computing(RC) literature (a machine learning approach) where the input is mapped onto a different but higher dimensional through a system called the reservoir and a readout that measures the state of the reservoir is trained, a notion of forgetting the states of the reservoir asymptotically is often used.
-
-% \ednote{How should I explain/qualify RC better here?
-It's mentioned without really explaining.
-M: Have put it in one sentence; just to let you know people use optical devices to map signal into a ``reservoir".
-B: Thanks!},
-Concepts like the ESP,  fading memory \cite{boyd1985fading} are some of these notions.  
-If $g$ possesses the ESP (equivalent in our context to the USP), the we are guaranteed that the whole of an input's left-infinite history will exactly determine the current system state; i.e.
-there is only ever the possibility of a single reachable state at a given instance of time \cite{jaeger2001echo,Manju_2020}.
-% Defined formally, the ESP is intrinsically linked to an input (see \cite{Manju_ESP}) Unfortunately, definitions and results concerning the ESP do not have much practical benefit since a relationship between the network's response and the temporo-statistical properties of the input are not directly described.
-A more detailed presentation may be found in~\cite{jaeger2001echo}.
-
-% We do however consider it important to mention that the ESP is also often discussed in terms of a stability property that holds for all inputs of a system~\cite{manjunath2020stability} and, furthermore, plays a key role in the design and training of recurrent neural networks(RNN) within the field of RC \cite{Manju_ESP}.
-Beyond forgetting the past states the ESP is also often discussed in terms of a stability property that is nearby inputs create close by responses (solutions), a concept called input-related stability~\cite{manjunath2020stability} and, furthermore, plays a key role in the robustness of recurrent neural networks(RNN), a component of the implementation discussed in~\ref{ch5}.
-
-\section{Conjugacies}
-
-Having already defined the reachable set $X_U$ as the collection of all elements of all entire solutions, we pause briefly in order to fix additional notation.
-Defining $\cev{u}^{n}:=(\ldots,u_{n-2} ,u_{n-1})$ as the left-infinite subsequence of an input up until time $n$, $\overleftarrow{U}$ is then the notation for all these left-infinite sequences in $U$.
-Moreover, $\cev{u}^{n}v:=(\ldots,u_{n-2} ,u_{n-1}, v)$ will symbolise the input up to time $n$ with $v \in U$ being the specific input value at time $n$. 
-The introduction of a new input at time $n$ can be described by the mapping $\sigma_v:   \cev{u}^{n} \mapsto \cev{u}^{n}v$. 
-The right-shift map, $r\overleftarrow{U}\to\overleftarrow{U}$, of an input sequence is defined $r: (\cdots, u_{-2},u_{-1}) \mapsto(\cdots, u_{-3},u_{-2})$
-
-The question now assumes the form: Can we establish a semi-conjugacy for the driven system $g$ as presented below.
-
-\begin{equation}  \label{Scomm_h}
-  %    \[ 
-      \psset{arrows=->, arrowinset=0.25, linewidth=0.6pt, nodesep=3pt, labelsep=2pt, rowsep=0.7cm, colsep = 1.1cm, shortput =tablr}
-   \everypsbox{\scriptstyle}
-   \begin{psmatrix}
-   \overleftarrow{U} & \overleftarrow{U}\\%
-   X_U & X_U.
-   %%%
-  %  \ncline{1,1}{1,2}^{\sigma_v} \ncline{1,1}{2,1} <{h}
-  %  \ncline{1,2}{2,2} > {h}
-  %  \ncline{2,1}{2,2}^{g(v,\cdot)}
-   \end{psmatrix}
-  % \]
-  \end{equation}
-
-We proceed to consider a specific subclass of conjugacies.
-
-  \begin{Definition}
-    [\bf Universal Semi-Conjugacy]\label{Def_UnivSemiConj} \rm
-    Given a driven system $g$, we  call a continuous and surjective map $h : \overleftarrow{U} \to X_U$ a universal semi-conjugacy if  diagram \ref{Scomm_h} commutes for all $v \in U$.
-  \end{Definition}
-
-  If the universal semi-conjugacy $h$ exists (i.e. the diagram in \ref{Scomm_h} commutes) then the solution $\Psi(\bar{u})$ will intuitively have no more ``complexity' than the input $\bar{u}$.
-
-But does such a function $h$ for the above driven system $g$ above exist?
-Whenever $g$ has the USP and $\Psi(u)=\{x_n\}_{n\in\mathbb{Z}}$ it follows that $h$, defined by  $h(\cev{u}_n):=g(u_n,x_{n-1})=x_n$, will satisfy the semi-conjugacy in the graph above \ref{Scomm_h}.
-Regrettably, such a continuous mapping $h$ is not guaranteed to exist when $g$ does not have the USP \cite[Lemma 5]{Manju_Nonlinearity}.
-Note that even when $h$ does exist, we are not guaranteed its injectivity. Considering again example \ref{ex_halfux}, we see that even if $h$ were to exist, it could not be injective as $X_U=\{0\}$.
-
-Re-sketching the commutativity diagram \ref{Scomm_h} above by replacing $X_U$ by its left-infinite sequence space $\overleftarrow{X}_U$, we obtain the diagram below.
-In this case, the function $H:\overleftarrow{U}\to\overleftarrow{X}_U$, a map that is both continuous and surjective, is called a \emph{causal mapping} that is defined next.
-
-\begin{Definition}
-  [\bf Causal Mapping]\label{Def_CausMap}
-  A continuous, surjective map $H:\overleftarrow{U}\to\overleftarrow{X}_U$ such that \[H\circ\tilde{g}_v=\sigma_v \circ{H}\] holds for all $v \in U$ where $\tilde{g}_v$ maps $(\ldots, u_{-2}, u_{-1})$ to $(\ldots, u_{-2}, u_{-1}, g(v, u_{-1}))$ is called a causal mapping.
-\end{Definition}
-
-\begin{equation} \label{SCausal_H}
-    %    \[ 
-        \psset{arrows=->, arrowinset=0.25, linewidth=0.6pt, nodesep=3pt, labelsep=2pt, rowsep=0.7cm, colsep = 1.1cm, shortput =tablr}
-     \everypsbox{\scriptstyle}
-     \begin{psmatrix}
-     \overleftarrow{U} & \overleftarrow{U}\\%
-     \overleftarrow{X}_U & \overleftarrow{X}_U.
-     %%%
-    %  \ncline{1,1}{1,2}^{\sigma_v} \ncline{1,1}{2,1} <{h}
-    %  \ncline{1,2}{2,2} > {h}
-    %  \ncline{2,1}{2,2}^{\tilde{g}_v}
-     \end{psmatrix}
-    % \]
-  \end{equation}
-
- \begin{Theorem}
-  For a compactly driven system, a causal mapping $H$ exists if and only if $g$ has the USP. 
-\end{Theorem}
-{\bf Proof.}  See~\cite[Th.3]{manjunath2013echo}.
-
-The driven system $g$ can induce an embedding of $\overleftarrow{U}$ in $\overleftarrow{X}$ as follows: 
-If the causal mapping $H:\overleftarrow{U}{\to}{\overleftarrow{X}_U}$ is injective (in addition to being surjective), it becomes the embedding of $\overleftarrow{U}$ in $\overleftarrow{X}$ induced by the driven system $g$.
-The map $H^{-1}$ is continuous since $H$ is a continuous and surjective mapping of a compact space $\overleftarrow{U}$ to a Hausdorff space.
-and we refer to the map $H$ as a \emph{causal embedding}.
-
-When $g$ has the USP, the diagram below (adopted from~\cite{Manju_Nonlinearity}) illustrates the operation of the mappings $h$ and $H$.
-The mapping $h:\overleftarrow{U}\to{X_U}$ is also considered an observable as mentioned in the introduction of Chapter \ref{ch3}).
-
-\begin{figure}[ht]
-  \includegraphics[scale=0.4]{_actionofh_H.eps}
-  \centering
-  \captionof{figure}{Ladder-like behaviour of $h$ and $H$ to illustrate the causal mapping. (Figure reproduced from~\cite{Manju_Nonlinearity}) }
-  \label{fig:actionh_H}
-\end{figure}
-
-The function $H$ contains a countably infinite number of coordinate (or factor maps) since its codomain is $\overleftarrow{X}_U$.  
-As it is not practically feasible to consider a left-infinite sequence in $X_U$ (in other words an element of  $\overleftarrow{X}_U$) for learning, we wish to restrict the inputs we consider to a subspace of $\overleftarrow{U}$.
-Particularly, we restrict our attention to the  so-called inverse-limit space of a dynamical system and hope to embed it within some finite product of $X$.
-
-%We wish to explore the possibility of establishing a conjugacy rather than a mere semi-conjugacy and therefore are interested in the restriction of inputs to the left-infinite orbits of a dynamical system.
-
-The left-infinite orbit space is significant because a great amount of information about the original system can be gleaned from its topological structure~\cite{ingram2011inverse,Manju_IEEE}. 
-The formal topological structure for the inverse-limit space of a chaotic dynamical system is complicated, and further discussion is relegated to~\cite{kennedy2008inverse_limit, ingram2011inverse}.
-Broadly, we may conceive of the inverse-limit system of a dynamical system as a subspace of an infinite-dimensional space where each point in the inverse-limit space corresponds to a backward orbit of the map $T$.
-Following~\cite{manjunath2021universal, ingram2011inverse} we provide a definition:
-
-\begin{Definition}
-  [\bf Inverse-Limit Space, Inverse-Limit System]\label{Dfn_InverseL}\rm
- The subset $\widehat{U}_T\subset \overleftarrow{U}$ defined by \[\widehat{U}_T: = \{ (\ldots,u_{-2},u_{-1}): Tu_{i} = u_{i+1}\} \] is called the inverse-limit space of $(U,T)$.
- The self-map $\widehat{T}$  induced by $T$ on $\widehat{U}_T$  is defined by \[\widehat{T}: (\ldots,u_{-2},u_{-1}) \mapsto  (\ldots,u_{-2},u_{-1},T(u_{-1}))\] and the resulting dynamical system $(\widehat{U_T}, \widehat{T})$ is called the inverse-limit system of $(U,T)$.
-\end{Definition}
-
-The inverse-limit space is well-defined since $T : U \to{U}$ is surjective by assumption.
-We now get to the concept of embedding the inverse limit-space of the dynamical system in $X \times X$. 
-\begin{Definition} \rm
-	We say a driven system $g$ \emph{causally embeds} a dynamical system $(U,T)$ if it satisfies the two properties: (i) a universal semi-conjugacy exists and  (ii)  $H_2(\cev{u}) := (h(r\cev{u}),h(\cev{u}))$ embeds the inverse-limit space $\widehat{U}_T$ in $X \times X$.
-\end{Definition}
-
-  It is imperative to take note of a subtlety here: we refer to the \emph{map} $H$ as a (causal) embedding, whereas when the driven \emph{system} $g$ (causally) embeds another \emph{system} $(U,T)$ we are concerned with $H_2$.
-\ednote{B: inserted line}
-
-\section{Choosing the driven system $g$}
-
-So far it is not clear if the USP alone ensures the ability to causally embedding a dynamical system and we mention that a driving function $g$ cannot just be chosen in a frivolous manner.
-This could possibly complicate our work.
-%\ednote{B: Grammarly}
-
-%The driving function $g$ cannot be chosen in a frivolous manner as it could possibly complicate our work.
-Specifically, one must be careful to avoid a choice of $g$ which would  add complexity to the obtained solution.
-
-%When a causal embedding $H$ exists for the driven system $g$, one can map an arbitrary input ${u}$ onto the solution space $X$ without additional distortion or information-loss \textbf{(Cite.)}.
-
-%When an embedding is established, the question of possible additional complexity in the solution is removed by guaranteeing that, since the systems are conjugate (semi-conjugate, \textbf{(refer)}), $g$ does not add any (some) complexity to the system.
-
-%To achieve a causal embedding embedding not just a causal mapping, it is undesirable to choose a function $g$ that quenches the temporal structure in $u$ by contracting to such a degree that the ability to recover information from the original system is lost completely.
-In the above example(\ref{ex_halfux}), the input's temporal variation could not be related to the reachable set as $X_U$ only consisted of a single element; little, if not no, information is encoded.
-To obtain a suitably complex function $g$, it is thus desired that the reachable set of a driven system  be large enough to relate to the input. 
-This is true even for embedding  the inverse-limit space of a dynamical system \cite{manjunath2021universal}.
-
- To this end, we recall the notion of State-Input (SI) Invertibility.
-
-\begin{Definition}
-  [\bf SI-Invertibility]\label{Dfn_SIinv}\rm
-  A driven system $g$ is said to be SI-Invertible if $g(*,x): U \to X$ is invertible for all $x\in X$.
-Alternatively it may be said that if, given $x_n$ and $x_{n-1}$, $u_{n-1}$ can be uniquely determined from $x_n=g(u_n,x_{n-1})$, then $g$ is said to be SI-invertible.
-\end{Definition}
-
-SI-invertibility promises that 'enough' information is retained when $g$ is chosen without introducing one's choice of $g$ will still ensure 'enough' information is retained without introducing unwanted complexity.
-'Enough' here refers to the fact that we may always recover the previous input value $u_n$ (if we know successive states $x_{n-1}, x_n$).
-
-Subsequently we define the relation $Y_T$ induced by $(U,T)$ on $X_U\times{X_U}$ for a driven system $g$ possessing SI-invertibility.  
-To describe the  SDD formally, we consider a dynamical system $T: U \to U$ and define a relation on the reachable set $X_U$, i.e. a subset on $X_U \times X_U$  defined by 
-$$Y_T:=\{(x_{n-1},x_n): \{x_k\}_{k\in \mathbb{Z}} \mbox{ is a solution for some orbit of } T \mbox{ and } n \in \mathbb{Z}\}.$$
-
-The following theorem establishes the existence of a well-defined map $G_T$ describing the single-delay dynamics(SDD) of the system above.
-
-\begin{Theorem}\label{Thm_GT_Exists}
-  For an SI-invertible driven system $g$ and a dynamical system $(U,T)$, the map $G_T: Y_T \to Y_T$ defined by the relation $(x_{n-1},x_n) \mapsto (x_n,x_{n+1})$ is well-defined. 
-  (This results holds even in the absence of $g$ possessing over the USP) 
-  % Moreover, the mapping $(x_{n-1},x_n) \mapsto u_{n}$ is well-defined when $x_{n-1}$ and $x_n$ are successive points on a solution obtained for an input  $\{u_n\}$ that is an orbit of $T$.
-  \end{Theorem}
-  {\bf Proof.} See~\cite[Th.3]{Supp}
-
-%'Without additional complexity' is guaranteed by the invertibility of the function $g$ that guarantees a one-to-one mapping between $u_n, x_{n-1}$ and $x_n$.
-In obtaining an SI-invertible function, therefore, we are guaranteed a system which will not lose so much information about its previous states in the forward-flow of time that one may not make any credible claims as to the original system's behaviour and properties.
-Simultaneously, it does guarantee us that the important information encoded in previous inputs is indeed preserved.
-
-At this stage it is worth taking note of a specific driven system in the form of a discrete state-space model which has acquired some adherence in applications \cite{Manju_IEEE}- especially those pertaining to Echo State Networks and the ESP.
-The function
-\begin{equation}  \label{eqn_driving}
-  g(u,x) = (1-a)x + a\overline{\tanh}(Au + \alpha Bx)
-\end{equation} 
-is SI-invertible and, if $\alpha B$ has a spectral norm $<1$, also possesses the USP \cite[Th.2]{manjunath2013echo }. 
-It is easy to show that $g$ is SI-invertible by recovering $u_n$ in 
-\begin{equation} \label{eqn_SI_RNN}
-  u_{n-1} := A^{-1}\bigg(\overline{\tanh}^{^{-1}}\frac{1}{a}\Big(x_{n+1}-(1-a)x_n\Big) \bigg) - \alpha B x_n
-  \end{equation}
-  when $x_{n-1}$ and $x_n$ are known.
-
-This specific driving function $g$ is used in our implementation and is discussed more completely in chapter~\ref{ch5}
-
-Despite the ease that with which one manipulates a left-infinite history in the realm of theory, it is impossible to obtain or use such a sequence in any real-life application.  
-Fortunately, one does not need the entire left-infinite history of an input in practice thanks to the Uniform Attraction Property(UAP). 
-We use an alternate version of the definition of UAP -- since the notion of nonautonomous attractors in this thesis would take some time to establish and detracts from the principal thrust of this project, we refer to \cite{Manju_Nonlinearity}.
-
-\begin{Definition}
-  [\bf Uniform Attraction Property]\label{Dfn_UAP}\rm
-  A driven system $g$ has the Uniform Attraction Property (UAP) if we initialize the driven system
-with an arbitrary initial value $y_m \in X$, and the sequence $y_{m+1}, y_{m+2}, y_{m+3},...$ satisfying $y_{k+1}= g(u_k,y_k)$ for $k \geq m$ then approximates an actual solution $\{x_n\}$ uniformly in the sense that given $\epsilon>0$ (independent of $y_m$) there is an integer $n$ so that $d(x_{n+i}, y_{n+i})<\epsilon$ for all $i\ge 0$, where $\{x_m\}$ is a solution.
-\end{Definition}
-
-The UAP guarantees that all trajectories will converge to the same trajectory as time moves forward.
-An illustration of this is shown in Fig.~\ref{fig:memloss_conttime} where the trajectory in red locks on to the supposed actual solution when the system has the USP.
-
-\begin{figure}[ht]
-  \includegraphics[scale=0.35]{_Nonlinearity_Fig_response.eps}
-  \centering
-\captionof{figure}{A coordinate of a simulated solution $(x_0,x_1,\ldots,x_{5000})$ of a RNN  plotted in red (with parameter $\alpha=0.99$ (where $g$ has the USP) and blue ($a=1$ and $\alpha=1.05$ where $g$ does not have the USP) while the matrices $A$ and $B$ are randomly generated, and $B$ has unit spectral radius.
-The input is shown in black.
-Reproduced from \cite{Manju_Nonlinearity}. }
-\label{fig:memloss_conttime} 
-\end{figure} 
-\ednote{B: How can I explain the black line? I don't think this makes sense to the reader immediately. M: Oh it was the input in black that I have added.}
-
-One may now appreciate an even more astounding result: $g$ having the USP is equivalent to the UAP as is proved in~\cite{Manju_Nonlinearity}.
-
-%We restate the above theorem.
-
-%\begin{Theorem}\label{thm_usp-contr-uap}  The following statements are equivalent:  \vspace{-8mm}  \begin{enumerate}[noitemsep, label=\roman*.]    \item $g$ has the USP.
-\item $g$ is a topological contraction.
-\item $g$ has the UAP.
-\end{enumerate\end{Theorem}{\bf Proof.}  May be found in \cite[Th.1]{Manju_Nonlinearity}
-
-% One need therefore not establish any additional results to ensure that the sequence uniformly approaches the unique solution $\Psi$.
-This vastly simplifies the effort necessary to set up a problem in order to guarantee that the underlying system will be accurately represented by the conjugate system.
-
-%This also solves the problem of perturbation/noise introduced by the observable or measurement function.
-Recall from before that an observable is inherently a map that discretises the underlying continuous-time system.
-Moreover, measurement and device error introduce mistakes.
-Since we are working with a chaotic system displaying sensitive dependence on initial conditions, these small errors could potentially send the trajectory into a completely different attractor.
-%In the presence of the UAP, however, small measurement errors introduced into the system do not pose the same danger as before since the trajectories will eventually tend towards the actual solution.
-% \ednote{(B: Is this Complete enough?)}
-% \textbf{From the perspective of perturbation of an autonomous dynamical system, if the resulting nonautonomous system has the USP, then the temporal structure of noise relates to the dynamics on a nonautonomous uniform attractor through a topological semi-conjugacy or a conjugacy.}
-
-\section{The next step in Dynamics}
-
-% Recall that $g$ is only being given inputs from the orbits of $T$.
-
-%Incredibly, this permits one to initialise a driven system $g$ with an altogether arbitrary initial value $y_m\in{X}$ where $m\in\mathbb{Z}$ and the UAP then guarantees that the sequence $\{y_{m+1}, y_{m+2},\ldots\}$ which satisfies the relation $y_n=g(u_n, y_{n-1})$  for  $k\geq{m}$  will approach the elements $\{x_n\}$ of the actual solution.
-This vastly simplifies the effort necessary to set up a problem in order to guarantee that the underlying system will be accurately represented by the conjugate system.
-
-We are interested in determining whether $\widehat{T}$ and $G_T$ are related.
-For this, we introduce  the function $H_2{\overline{}} := (h(r\overline{u}, h(\overline{u})))$ mapping the entirety of a left-infinite sequence to some element in $X\times{X}$.
-Indeed, it may be shown that $g$ having SI-invertibility and the USP immediately guarantees the existence of at least a semi-conjugacy between the systems $(Y_T, G_T)$ and $(\widehat{U}_T, \widehat{T})$.
-This is formalised in the Causal Embedding Theorem, which we state next:
-
-\begin{Theorem}
-  [\bf Causal Embedding Theorem (adopted from \cite{Supp})]
-  % {\bf (Causal Embedding Theorem.)}
- \label{Thm_CET}
-	Let $g$ be a driven system with SI-invertibility and the USP.
-Let $h$ denote the universal semi-conjugacy and $H_2(\overleftarrow{u}) := (h(r\overleftarrow{u}),h(\overleftarrow{u}))$, where $r$ is the right-shift map.
- Let $(\widehat{U}_T, \widehat{T})$  be the inverse-limit system of a dynamical system $(U,T)$. 
- Then the restriction of $H_2$ to $\widehat{U}_T$ is a topological semi-conjugacy between the inverse-limit system $(\widehat{U}_T, \widehat{T})$ 
-and the induced dynamical system  $(Y_T,G_T)$, i.e., the following diagram commutes
-\begin{equation} \label{comm_H_CET}
-\psset{arrows=->, arrowinset=0.25, linewidth=0.6pt, nodesep=3pt, labelsep=2pt, rowsep=0.7cm, colsep = 1.1cm, shortput =tablr}
- \everypsbox{\scriptstyle}
- \begin{psmatrix}
- \widehat{U}_T & \widehat{U}_T\\%
- Y_T &  Y_T.
- %%%
-%  \ncline{1,1}{1,2}^{\widehat{T}} \ncline{1,1}{2,1} <{H_2}
-%  \ncline{1,2}{2,2} > {H_2}
-%  \ncline{2,1}{2,2}^{G_T}
- \end{psmatrix}
- \end{equation}
-or in other words, $(Y_T, G_T)$ is a factor of  $(\widehat{U}_T, \widehat{T})$. Further, if $T:U \to U$ is a homeomorphism, 
-then $H_2$ embeds $\widehat{U}_T$ in $X_U \times X_U$, and hence $(Y_T, G_T)$ is conjugate to $(\widehat{U}_T, \widehat{T})$.
-\end{Theorem}
-{\bf Proof.}  May be found as the proof of~\cite[Th.4]{Supp}
-
-Recall that $H_2$ maps an entire left-infinite solution sequence from $\Psi$ to an element in $X\times{X}$. 
-We are drawing nearer and nearer to our principal target and our results carry more and more weight: If we are able to learn $G_T$, we will also have learnt $\widehat{T}$; and since $\widehat{T}$ is an extension of $T$, we will have obtained $T$ as well.
-%\ednote{B: Grammarly check}
-
-%\begin{Theorem} Graph \ref{Scomm_h} is exactly the inverse-limit system $(\hat{U}, \hat{T})$.    \end{Theorem}
-
-%We now have the following (which may be compared with \ref{SCausal_H} above):
-
-%\begin{equation} \label{fig:inverse_limsystem}
-  %  \[       \psset{arrows=->, arrowinset=0.25, linewidth=0.6pt, nodesep=3pt, labelsep=2pt, rowsep=0.7cm, colsep = 1.1cm, shortput =tablr}      \everypsbox{\scriptstyle}
-%      \begin{psmatrix}      \widehat{U}_T  && \widehat{U}_T \\     Y_T && Y_T \\
-      %%%
-     %  \ncline{1,1}{1,2}^{\widehat{T}} \ncline{1,1}{2,1} <{h}
-     %  \ncline{1,2}{2,2} > {H_2}
-     %  \ncline{2,1}{2,2}^{G_T}
-%      \end{psmatrix}
-    %  \]
-%  \end{equation}
-
-%\begin{Theorem}
- %   $(Y_T, G_T)$ is semi-conjugate to $(\widehat{U}, \widehat{T})$.
-%\end{Theorem}
-% {\bf Proof.} See \cite[Th.3, Th.4]{manunath2021universal}
-
-\section*{Summarising the discussing thus far:}
-
-It is easy to lose the birds-eye view, so we take a moment to review our progress up until this point.
-
-\vspace{-8mm}
-\begin{enumerate}
-\item We are interested in a some dynamical system $(U,T)$ with unknown dynamics.
-\item To determine properties about this system $(U,T)$ and predict its future evolution, we determine the dynamics of the inverse-limit system $(\widehat{U}, \widehat{T})$.
-% Given certain assumptions, we can guarantee that $(\widehat{U}, \widehat{T})$ is at least semi-conjugate to $(U,T)$.
-\item If the driven system $g$ is SI-invertible (and $\{u_n\}\in{U}$  is an orbit of $T$), the map $G_T$ exists that describes the single delay dynamics. 
-\item If, furthermore, $g$ has the USP, $(Y_T, G_T)$ is semi-conjugate to $(\widehat{U}, \widehat{T})$.
-\item If we can assume that $T$ is a homeomorphism, $(Y_T, G_T)$ is topologically conjugate to $(\widehat{U}, \widehat{T})$, an extension space of $(U,T)$
-\end{enumerate}
-
-One can therefore in practice learn the SDD of the driven system states via $G_T$ with enough data thanks to the USP/UAP. This enables us to do at least two things: 
-\vspace{-8mm}
-\begin{itemize}
-\item Forecast  $x_{n+1},x_{n+2}, \ldots$ via iterates of $G_T$ (if $G_T$ can be learnt).
-\item Forecast future values of $u_n$ since $x_n$ and $x_{n+1}$ determine $u_n$ since $g$ is SI-invertible. 
-\end{itemize}
-
-Finally, we note that although $G_T$ exists with SI-invertibility, we need the USP as well(see \ref{fig_pictorialSummary} below for a visual illustration).
-If not, the driven states would have to be running for all time since they would not have forgotten their past states.
-
-\begin{figure}[ht]
-  \includegraphics[scale=0.3]{_summarypictorial.eps}
-  \centering
-  \captionof{figure}{Pictorial summary of results obtained up to now and how they overlap. Reproduced from \textbf{(Cite SIG-25May)} }
-  \label{fig:fig_pictorialSummary}
-\end{figure}
-
-\section{A discussion of $G_T$ }
-
-In the above sections, we established the map $G_T$ describing the SDD of a driven system.
-By establishing the existence of this map, we’ve essentially embedded the attractor $U$ (recall from \ref{attractor_U}) into the higher dimensional space $X\times{X}$.
-In layman’s terms, this ensures that there is more “dimensional room” for the underlying system’s underlying dynamics to “move”.
-Since the dynamics aren’t as “squashed”, we might therefore hope that the dynamics of $G_T$ are in some sense simpler than that of $T$.
-(Taken note of the fact that $G_T$ is a homeomorphism even when $T$ is just continuous)
-
-In \cite{manjunath2021universal} it is illustrated in an empirical fashion in that the map $G_T$ describes dynamics which are less functionally complex than that of $T$ or of the map $\Phi_{2d,\theta}$.
-This is done by implementing a Recurrent Neural Network (RNN), which is later mentiond in Chapter~\ref{ch5}.
-\ednote{B: Should I add in that we'll be using Pearson coefficient?
-Not necessary.
-Please check all the cross citation.
-You need to explicitly mention that it is a Chapter or else only a number would appear}
-
-We opt to learn $G_T$ in an indirect manner by defining a new map $\Gamma:(x_{n-1},x_n)\mapsto{u_n}$. It will follow immediately from $G_T$’s existence that $\Gamma$ also exists.
-
-The reasons for taking this roundabout approach remain to be discussed in Section~\ref{subs_LearnGamma}, but a pat answer may immediately be given: When $\Gamma$ has been learnt, the system can be driven autonomously and then $G_T$ is known anyway.
-We formalise this with a theorem.
-
-\begin{Theorem}
-  When $x_n$ and $x_{n-1}$ are successive points on a solution obtained for input-orbit of $T$ $\{u_n\}$, then the map $\Gamma: X\times{X}\to{U}$ defined by $(x_{n-1},x_n)\mapsto{u_n}$ exists whenever $G_T$ exists
-\end{Theorem}
-{\bf Proof.}  See \cite[Th. 3c]{manjunath2021universal}.
-
-Projection mappings $\pi_i$ are defined in the traditional meaning where a $k$-dimensional vector is projected to it's $i^{th}$ component such that $$\pi_i:(a_1,a_2, \ldotp, a_{k-1}, a_k)\to{a_i}$$
-
-The graph in equation \ref{fig:comm_H_CET} is then extended as below:
-
-\begin{equation}  \label{Scomm_imp}
-  %    \[ 
-      \psset{arrows=->, arrowinset=0.25, linewidth=0.6pt, nodesep=3pt, labelsep=2pt, rowsep=1.3cm, colsep = 1.1cm, shortput =tablr}
-   \everypsbox{\scriptstyle}
-   \begin{psmatrix}
-   \widehat{U}_T & \widehat{U}_T\\%
-   Y_T & Y_T\\
-   & \textcolor{red}{U \times X}
-   %%%
-  %  \ncline{1,1}{1,2}^{\widehat{T}} \ncline{1,1}{2,1} <{H_2}
-  %  \ncline{1,2}{2,2} > {H_2}
-  %  \ncline{2,1}{2,2}^{G_T}
-  %  \psset{linecolor=red}
-  %  \textcolor{red}{\ncline{2,1}{3,2}<{(\Gamma,\pi_2)}}
-  %  \ncline{3,2}{2,2}>{\textcolor{red}{(\pi_2,g)}}.
-   \end{psmatrix}
-  % \]
-  \end{equation}
-
-The problem finally simplifies to the issue of learning the map $\Gamma$ and combining this with the projection mapping $\pi_2$ and the function $g$, which will be known.
-A final set of equations is obtained - equations that have been entirely constructed from data.
-\begin{eqnarray}\label{eqns_from_data}
-	u_{k+1} &=& \pi_1 \circ (\Gamma, \pi_2) \circ (\pi_2,g) (u_k,x_k) \label{Seqn_u}\\
-	x_{k+1} &=& \pi_2 \circ (\Gamma, \pi_2) \circ (\pi_2,g) (u_k,x_k). \label{Seqn_x}
-\end{eqnarray}
-
-\section{Advantages of learning $\Gamma$} \label{subs_LearnGamma}
-
-One may immediately ask why we opt to take such a roundabout route; why not just learn the map $G_T$ from the get-go?
-On the surface, this seems to be an arbitrary decision path with no real reasoning, so we take a pause again and discuss the motivation for learning Gamma.
-There are a number of distinct advantages.
-
-In the first place, learning $\Gamma$ saves computational resources.
-This is due to the fact that the input $u_n$ lies in a lower-dimensional subspace in comparison to the high-dimensional space $X\times{X}$.
-
-%In practice, if the input is of a  lower dimension, one may easily embed it into the space $X\times{X}$ by padding the vector $u_n$ with zeroes.
-
-Secondly, the function $\Gamma$ is known to be stable.
-Learning $\Gamma$ makes use of equations \ref{eqns_from_data} which in turn employs the driven system $g$ possessing the USP, and offers distinct advantages with regards to stability in the presence of perturbation.
-It is desirable that inputs that differ only slightly would have only ‘slight’ effects on the output of the system.
-Moreover, one would hope to prevent numerical errors originating from input and measurement noise (see~\cite[Th.
-5]{manjunath2021universal}).
-As the data fed into the system is a function of both the input given and the parameters of the system itself (i.e.
-the value a, $\alpha$ in $\tanh$), we exploit both input-related and parameter-related stability.
-Both these notions are defined by way of the continuity of an encoding map of an input and design parameter respectively in \cite{manjunath2020stability}.
-
-Informally, we may consider input-related stability as concerned with the question of whether or not small variations in input would result in small responses.   
-If two inputs $\bar{u}$ and $\bar{u}_{noisy}$ are close in the product topology, then their tails could differ greatly due to the metric that generates the product topology has insensitivity to the differences in sequence's tails.
- However if the system has the USP then the function $h$ is continuous and so $h(\bar{u}), h(\bar{u}_{noisy})$ remain close-by for small levels of noise maintaining a measure of robustness to input and measurement noise.\cite{manjunath2021universal}.
-In a similar fashion, parameter-related stability is obtained thanks to a result \cite[Lemma 3.2]{manjunath2020stability} relating this form of stability with the ESP (which is here equivalent to the USP).
-
-% Secondly,  due to $g$ possessing the USP, both input- and parameter-related stability~\cite{manjunath2020stability} is obtained, which in turn prevents numerical errors that caused by input and measurement noise (see~\cite[Th.
-5]{manjunath2021universal}).
-
-Thirdly, learning a chaotic map on a set(here $Y_T$) with even a small measure of noise  means that the learnt system accepts arguments out of the set (Problems similar to Takens, cf.
-Section~\ref{sect_takenslimits}) \ednote{Please insert the right label}.
-Suppose iterates of  $(x_{n-1},x_{n})$ under the map $G_T$ is used to forecast two successive values of a solution of $g$ contained in $Y_T$, then these iterates would be contained in $Y_T$.
-However, in practice we are confronted with noise.
-Due to noise at some moment, an iterate could slip beyond $Y_T$ and into an
-expansion $Y_T^+$, for example a closed set that contains all the iterates of $(x_{n-1},x_{n})$ under the map $G_T^+$, where $G_T^+$ acts on $Y_T^+$.
-
-Now since $Y_T$ is not guaranteed to be an attractor of $G_T^+$ this could lead to errors. To see this, consider Fig.~\ref{fig:YtGtFailure}.
-Thus if one learns $G_T$ directly, we are in danger of replicating the problems arising for Takens.
-
-\begin{figure}[ht]\includegraphics[scale=0.25]{_YTerrors.eps}
-  \centering
-  \captionof{figure}{Diagram of potential errors developing when $G_T$ is learnt. (Figure from \cite{Supp}}
-  \label{fig:YtGtFailure}
-\end{figure}
-
-To circumvent this: if a driven system with a saturation function \footnote{a function $f(t)$ that is monotonic and approaches a constant as $|t|\to \infty$, for e.g., $\tanh(t)$ on $\mathbb{R}$} is used, we can ensure the existence of an attractor $A$ of the map that implements
-$(\pi_2,g) \circ (\Gamma,\pi_2)$ (printed in the red in ...\ednote{Please cite the commutativity diagram on the previous page}) on 
- which in practice turns out to be defined on $Y_T^+$ has an attractor $A$  containing $Y_T$.
-\ednote{B: A little vague.
-Are we saying that an attractor of a map implementing $G_T$ is also the one that contains $Y_T^+$?
-M: Please see now..
-there was typo..
-it was A containing $Y_T$.
-I have expanded it now}
-By restricting the dynamics to not spill beyond $A$, prevents large errors from occuring.  
-More details on the global dissipativity of $g$, as it is termed here, may be found in~\cite{Supp}.
-
-\begin{figure}[ht]
-\includegraphics[scale=0.25]{_autovsdriven.eps}
-  \centering
-  \captionof{figure}{Illustrating empirically the importance of global dissipativity.
-Plotted below are the three principal components of randomly initiated points(blue) of a learnt version of $G_T$ versus the trajectory $(x_{n-1}, x_n)$(red) of driven states obtained from the Fractal Dream Attractor's data~\ref{Thomas_Attractor}).
-}
-  \label{fig:learningFailure}
-\end{figure}
-
-In effect, global dissipativity prevents large numerical errors due to the noise in input data from arising (see Fig.~\ref{fig:learningFailure}). 
-If a system is not globally dissipative, neglibly small errors could lead to major inaccuracy which thus induce predictions to utterly fail.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\chapter{Implementation} \label{ch5}
-
-In this chapter we turn to the implementation of the function $G_T$ by discussing a method to learn the function $\Gamma$.
-We apply our methodology to some chaotic attractors.
-Most notably, we consider the reconstruction of the chaotic attractor of the double pendulum - an  attractor that has not been successfully reconstructed because of its intricate structure.
-In fact the previous data-driven approaches also have not been successful in building a model from the double pendulum.
-
-%We then indicate where this implementation differs from previous work done by (\cite{manjunath2021universal}) and empirically show that the SDD of $G_T$ are less functionally complex than $\Phi_{k, \theta}$ or $T$ by considering the Pearson correlation coefficient.
-Finally, we present and discuss some numerical results for a focus system of this project: the double pendulum.
-We also consider other chaotic attractors.
-
-\section{Implementing $G_T$}
-The map $G_T$ is implemented by applying the functions $\Gamma$, $g$ and the projection function $\pi_2$ in the relevant order as described by the equations~\ref{eqns_from_data}.
-To reach this point, one must first learn $\Gamma$.
-After first normalising the data to ensure that it has zero mean value and lies within the range $[-1,1]$ and required zero-padding(see paragraph below for further discussion), we continue by driving the system with external input through the function $g$.
-
-From \eqref{eqn_driving} we know that $g(u,x) = (1-a) x + a \tanh(Au + \alpha B x)$. Recall that $g$ is SI-invertible. We refer to $A$ as the input matrix and $B$ the reservoir matrix.
-
-We make use of a Recurrent Neural Network (RNN) in our implementation to project the data on to a higher dimensional space, but we do not train the RNN in the familiar ways in the literature but as in \cite{manjunath2021universal}.
-A few words about the RNNs: RNNs are widely used in the literature for temporal processing of data as a type of deep learning technique.
-They take the form of equations considered to be discrete-time version of continuous dynamical systems modelling a neural network of neurons \cite{jaeger2001echo}.
-
-We implement a feedforward network, (a network with no memory, unlike the RNN) required to learn $\Gamma$ in the \emph{Python} programming language and making use of the \emph{Scipy} library, alongside the \emph{Keras} package in the \emph{Tensorflow} library.
-
-If the RNN we will be using in the next step has $N$ neurons, then $A$ and $B$ are each $N\times{N}$ matrices.
-Inputs $u_n$ of dimension $K < N$ are padded by zeroes (as remarked earlier) to embed the input in a space of dimension $N$.
-As we do not need the entire left-infinite history of input, an arbitrary initialisation of the system with initial values followed by the driving of the system with the function $g$ means that the generated values will approach the actual solution elements in a uniform manner thanks to the UAP that we had discussed in the preceding chapter.
-
-Thirdly we prepare the network.
-Utilising a network $N$ neurons (the same value $N$ as in the dimensions of the matrices $A,B$).
-Learning of $\Gamma$ is done using the Adam Optimiser where the means square error (MSE) loss function is optimised.
-The system is trained several times (in our case 4) with incrementally smaller learning rates on each run (0.001 and thereafter divided by 10 at each run).
-Additional parameters such as the number of network-layers, neurons per layer and training-length are presented case-by-case below.
-
-We then reach the prediction step. Making use of $g$'s SI-invertibility and $g$ as defined in \eqref{eqn_driving}, we have the following expression for $u_n$ replicated from~\eqref{eqn_SI_RNN}.
-\begin{equation*}
-  u_n := A^{-1}\bigg(\overline{\tanh}^{^{-1}}\frac{1}{a}\Big(x_{n+1}-(1-a)x_n\Big) \bigg) - \alpha B x_n
-\end{equation*}
-
-The two model equations for learning from data~\eqref{eqns_from_data} allows us to determine what the state $x_{n+1}$ at time $n+1$ will be.
-
-%We specifically remind ourselves of \[x_{k+1}=\pi_2 \circ (\Gamma, \pi_2) \circ (\pi_2,g) (u_k,x_k)\]
-
-\section{Delay Coordinates}
-
-We also consider a more general problem of learning an attractor of dynamical system when only observations of an orbit are given.
-Explicitly, if $(W,T)$ is a dynamical system with dynamics generated by $w_{n+1}=Tw_n$, and if $\theta:W \to \mathbb{R}$ is an observable, then the task is to learn a dynamical system that is topologically conjugate to $(W,T)$ and predict $\theta(w_{m+1}),\theta(w_{m+2}),...$ using the data $\theta(w_{0}),\theta(w_{1}),\ldots,\theta(w_{m})$.
-So how do we do it?
-Suppose the input generated from the delay-coordinate map $\Phi_{\theta,2d}(\theta(w_{n})) := (\theta(w_{n-2d}),\ldots,\theta(w_{n-1}),\theta(w_{n}))$ is such that Takens delay embedding theorem (see Theorem~\ref{Thm_Takens}) holds, in which case, there exists a homeomorphism  $F_\theta: \Phi_{\theta,2d}(\theta(w_{n})) \mapsto \Phi_{\theta,2d}(\theta(w_{n+1}))$.
-Then if we feed  the input values $u_n := \Phi_{\theta,2d}(\theta(w_{n}))$ to a driven system $g$ that is SI-invertible and has USP, the induced system $(Y_F,G_F)$ would be topologically conjugate to the inverse-limit space of
-$(\Phi_{\theta,2d}(W), F_\theta)$ as in Theorem~\ref{Thm_CET}, and hence one could forecast $u_n,u_{n+1},\ldots$.
-In other words, one could forecast  the values $\theta(w_n), \theta(w_{n+1}),\ldots$ by feeding these observations as delay coordinates.
-The advantage of feeding delay-coordinates to a driven system is that the embedding is stable in the sense of global dissipativity that we have described in the previous chapter, i.e., $Y_F$ is made to contained in an attractor of a mapping induced on $Y_F$ that is induced by $g$.
-
-We note that the twin equations together~\ref{eqns_from_data} generate a model from data.
-For a chaotic system, a small amount of noise due either to the input-noise or computational error would cause the solutions from two accurate models with the same initial conditions to diverge.
-However, statistically, they would behave in the same way in the sense that the visiting frequency or the average time spent by an orbit in any region of the phase space of the dynamical system is identical for almost all initial conditions (almost all is specified as a measure), a result in ergodic theory (\url{https://en.wikipedia.org/wiki/Ergodic_theory} a topic beyond the scope of this thesis.
-\ednote{B: I'll need to recap this sentence with you before presenting as it's new to me.
-M: I rewrote it.} To illustrate that we plot the densities of the trajectories of the actual and predicted solutions.
-
-In our implementation of the input and reservoir  matrices $A$ and $B$ in the RNN are sparse matrices.
-We discuss three specifics relating to this concept.
-Firstly, the matrices are assigned a variable density in terms of the amount of non-zero entries and are scaled to have values within the range $[-1,1]$.
-For this, we make use of \textit{Python's SciPy} library and specifically the \textit{Sparse, LinAlg} modules.
-In our experiments, we generally opted for a density ratio of 10\%-25\%.
-Secondly, we opted to enforce the requirement that both matrices be well-conditioned, i.e.
-that they have a condition number of \~1.
-The condition number of a matrix $M$ (denoted $cond(M)$) is the product of the matrix norms of the matrix $M$ and its inverse $M^{-1}$ and is a measure of the sensitivity of output of a matrix multiplication when an input is changed.
-\textbf{(Cite.)} By ensuring a condition number close to 1, we also ensure that our driven function $g$, depending on the matrices $A$ and $B$, will not be sensitive to adjustments in the input.
-In so doing we then also guarantee another form stability.
-\ednote{B: Not complete enough.
-Edit further and cite M: Please feel free to cite a wikipedia page}
-
-Lastly, we ensure the matrix $B$ has unit spectral radius.
-This is necessary due to the Causal Embedding Theorem~\ref{Thm_GT_Exists} which can then guarantee the existence of semi-conjugacy between the systems $(Y_T, G_T)$ and $(\widehat{U}_T, \widehat{T})$.
-In fact, the theorem actually requires that $\alpha{B}$ have spectral radius $<1$, so we impose the restriction $\alpha<1$ and consider only the matrix $B$'s spectral radius from thereon.
-
-Note that to learn $\Gamma: (x_{n-1},x_{n}) \mapsto u_n$ through our feedforward network, we could learn the principal components of $(x_{n-1},x_{n})$ as well. 
-We employ this since it provides slightly better results empirically but is not essential.
-
-Principal components are used only for a more efficient state representation to possibly reduce learning errors; it is not for a lossy approximation since we use all principal components \textbf{Cite.}.
-More explicitly, denote $X_{1:N}$ as the matrix with the first $N$ states of the network data represented by row vectors.
-If $X_{1:N}=U\Sigma P^T$ here represents the singular value decomposition ($U$ and $V$ are unitary matrices and $\Sigma$ is a diagonal matrix) of $X_{1:N}$ and we refer to ...\ednote{B: Todo.
-Refer or Cite M: Please add a Wiki citation in the bib}, we then denote the principal component matrix $P$, and the principal components are given by
- $Z_{1:N}=X_{1:N}P.$
-These principal components are then used to train a feedforward neural network. Denote the row vectors of $Z_{1:N}$ by 
-$z_i^T, i=1,2,\cdots,N$ and the neural network by $NN$, the network is subsequently trained to learn an approximation of the map
-$NN: (z_{n-1},z_n) \mapsto  u_n.
-$
-The learnt approximation of $NN$ can be used to approximate $\Gamma$ since 
-$$
-NN\left( \begin{bmatrix} 
-P^Tx_{n-1} \\
-P^Tx_n
-\end{bmatrix}
-\right) = NN \circ 
-\begin{bmatrix}
-P^T & 0 \\
-0 & P^T 
-\end{bmatrix}\begin{bmatrix}
-x_{n-1}\\
-x_n
-\end{bmatrix} = u_n = \Gamma(x_{n-1},x_n).
-$$
-
-\section{Experimental Results}
-In this section we showcase numerical results pertaining to 3 attractors that were simulated using the methodology described.
-We first consider a simple physical system that exhibits complicated dynamics.
-\subsection{Double Pendulum}
-
-The Double Pendulum consists of two pendulums fastened to one another such that the system moves as a whole.
-The first (or higher pendulum) is attached to a fixed point while the second is attached to the endpoint of the first.
-(See below)
-This system is a classical example of a system exhibiting chaos. \textbf{(Citation.)} Consider again~\ref{fig:dp_sdic}.
-The system has four variables: angle and angular velocity of each pendulum rod. An ideal system is assumed where the rods are massless and no friction in the pivots are considered.
-\begin{figure}[ht]
-  \includegraphics[scale=0.35]{_dp_setup.eps}
-  \centering
-  \captionof{figure}{Pictorial representation of the setup of a DP. Image taken from~\cite{DPSetup}}
-  \label{fig:dp_setup}
-\end{figure}
-
-The scalar equations of motion are derived using Newtonian physics and are reproduced from~\cite{DPFormulas} in~\ref{eqns_dp}.
-Each pendulum component has an associated mass of the head, $m$, and length of the rod, $L_i$.
-Acceleration due to gravity is denoted by $g$ and initial velocities by $\dot{\theta}$
-\begin{eqnarray}\label{eqns_dp}
-  \ddot{\theta_{1}}  = \frac{-g(2m_1+m_2)\sin{\theta_1} - m_2g\sin(\theta_1-2\theta_2) - 2\sin(\theta_1-\theta_2)m_2({\dot{\theta_{2}}}^{2}L_2 + {\theta_{1}'}L_1\cos(\theta_1-\theta_2))} {L_1(2m_1 + m_2 -m_2\cos(2\theta_1 - 2\theta_2))}
-  \\
-  \ddot{\theta_{2}} = \frac{2\sin(\theta_1-\theta_2)(\theta_{1}'^{2}L_1(m_1+m_2) + g(m_1+m_2)\cos\theta_1 + \dot{\theta_{2}}^{2}L_2m_2\cos(\theta_1-\theta_2))}{L_2(2m_1 + m_2 -m_2\cos(2\theta_1 - 2\theta_2))}
-\end{eqnarray}
-
-The second-order system is converted to a first-order ODE system and solved using the Runge-Kutta numerical method.  
-Here we chose the values $m_1=2.0$, $m_2=2.0$, $L_1=1.5$ and $m_2=1.5$ with initial velocities $v_1=1.5, v_2=-2$ for the respective pendulum heads.
-Undamped DP data-sets were generated and used in this project due to the requirement that the map $T$ be surjective that we must not forget. 
-An example of this would be the case of a DP exhibiting energy-loss (i.e.
-where the map $T$ is not surjective).
-In such a case, the attractor is a single point in the phase space when the double pendulum comes to rest.
-It must be mentioned once again that if we use data outside the attractor, learning is not reliable. To see this, consder the graphs ~\ref{fig:damped_pendulum} and ~\ref{fig:dp_notsurjective}.
-
-\begin{figure}[ht]
-  \includegraphics[scale=0.4]{_dp_dying.eps}
-  \centering
-\caption{$x$- and $y$-coordinates of the two pendulum heads plotted for a dataset that exhibits dying oscillations.
-The map generating the (four-dimensional data) for a damped pendulum is not surjective.
-Data obtained from chaotic double pendulum dataset generated from videos takes of experimental pendulums~\cite{asseman2018learning}}
-\label{fig:damped_pendulum}
-\end{figure}
-
-\begin{figure}[ht]
-  \includegraphics[scale=0.3]{_dpfail_nonsurj.eps}
-  \centering
-\caption{Failure to forecast or learn a map which is not surjective: Evolution of the $y$-coordinate of the bottommost pendulum plotted here for 15000 timestep.
-Plotted here the predicted value(red) versus the actual value(blue).
-The plots overlap for the first 10000 steps, the duration of training to learn the map $\Gamma$.
-Afterwards, the trajectories start diverging and it the predicted value no longer accurately represents the complexity or energy-loss of the actual data.
-}
-\label{fig:dp_notsurjective}
-\end{figure}
-
-Having spent ample time to discuss the challenges, we now change tone so as to satisfy the optimist and present a successful run of the DP-data to predict the future movement of the pendulum.
-
-The system was driven for 15500 steps, of which the first 500 were immediately discarded to simulate the network 'forgetting'.
-The driving function $g$ had input and reservoir matrices were implemented as sparse matrices filled to 20\%.
-Parameter-values of $\alpha=0.99$ and $a=0.5$ were used.
-The remaining 15000 datapoints were used to train a network with 128 neurons. 
-The network was initialised to contain 16 hidden layers each with a layer dimension of 64, activated with the tanh activation function.
-Training was accomplished using 256 epochs with batch sizes 128 and a delay of 1.
-See~\ref{fig:dp_success_traj} for results pertaining to the trajectory taken by the DP's components and then~\ref{fig:dp_success_density} to witness associated densities of predicted vs.
-actual values.
-
-\begin{figure}[ht]
-  \centering
-  \includegraphics[width=0.95\linewidth]{Graphs/_dp_success_4coords_traj.pdf} 
-  \captionof{figure}{Predicted(red) and actual(blue) trajectories of the $x$- and $y$-coordinates of each of the pendulum heads.
-From top to bottom are as follows: $x$-coordinate of topmost pendulum head, $y$-coordinate of topmost pendulum head, $x$-coordinate of lower pendulum head, $y$-coordinate of lower pendulum head.
-  These graphs were constructed by predicting the DP 1000 timesteps into the future and in so doing illustrating the long-term consistency and accuracy of the learnt system.
-Here we lock onto the near-exact trajectory for about 300 timesteps and stay close up until about 600 seconds.
-}
-  \label{fig:dp_success_traj}
-\end{figure}
-\begin{figure}[ht]
-  \centering
-  \includegraphics[width=\linewidth]{_dp_success_4coords_hist.eps}
-  \captionof{figure}{Densities of predicted and actual values of $x$- and $y$-coordinate of pendulum heads.}
-  \label{fig:dp_success_density}
- \end{figure}
-
-We now consider an experiment where noise is added to the system and find it to be incredibly robust.
-We employ the same parameters as before and add noise from a normal-distribution with mean zero and standard deviation equal to 0.1.
-Once again consider the actual and predicted trajectories for the noisy DP dataset in Figure~\ref{fig:noisydp_success_traj}.
-Next we consider the densities of the predicted and actual values in Figure~\ref{fig:noisydp_success_density} and observe that they display highly similar distributions.
-
-\begin{figure}[ht]
-  \centering
-  \includegraphics[width=0.95\linewidth]{_dp_noise_FigCD_2030.eps} 
-  \captionof{figure}{Predicted(red) and actual(blue) trajectories of the $x$- and $y$-coordinates of each of the pendulum heads with noise added from a normal distribution with mean zero and standard deviation 0.1.
-  These graphs were constructed by predicting the DP 1000 timesteps into the future and in so doing illustrating the long-term consistency and accuracy of the learnt system.
-Here we lock onto the near-exact trajectory for about 250 timesteps and stay close up until about 400 seconds.
-}
-  \label{fig:noisydp_success_traj}
-\end{figure}
-\begin{figure}[ht]
-  \centering
-  \includegraphics[width=\linewidth]{_dp_noise_2B_2030.eps}
-  \captionof{figure}{Densities of predicted and actual values of x-coordinate of first pendulum head.}
-  \label{fig:noisydp_success_density}
- \end{figure}
-
- Lastly, an run was done where only a single coordinate (the $y$-coordinate of the lower pendulum head) was fed into the system and we can illustrate through the obtained success that the coordinate contains all the information in its past states necessary to forecast its future evolution.
- \textbf{Insert graphs here. Had to quickly change something here.}
-
-\subsection{Additional Attractors: Clifford \& Thomas}
-We follow the methodology presented in \cite{manjunath2021universal} but opt to consider attractors different from the Lorenz system and H\'enon map in this project.
-
-\subsubsection{Thomas' Cyclically Symmetric Strange Attractor}\label{Thomas_Attractor}
-Thomas' Cyclically Symmetric Strange Attractor, a 3D attractor proposed in 1999 by Ren\'e Thomas in \cite{ThomasAttractor}, is described by a set of three equations which is distinctly reminiscent of the Lorenz system, another deterministic three-equation model exhibiting chaotic behaviour.It has a single parameter $\beta$ and has been shown to transition to chaotic behavious when $\beta<0.208186$~\cite{Thomas_BetaParameter}.
-\begin{eqnarray}\label{eqns_thomas}
-  x_{n+1} = \sin(y_n) - \beta{x_n} \\
-  y_{n+1} = \sin(z_n) - \beta{y_n} \\
-  z_{n+1} = \sin(x_n) - \beta{z_n}
-\end{eqnarray}
-
-The behaviour of the system with a $\beta$-parameter value of 0.1056 was simulated by scaling it to fit inside the interval $[-1,1]$.
-Two cases were considered: noise-free and the one perturbed by noise.
-Noise was generated from a normal distribution with mean 0 and standard deviation 0.05, which translates approximately to * dB \ednote{Please calculate dB as signal to noise ratio }
- A sequence of 3000 observations in both the clear and noisy data-sets respectively were fed into the system with 300 entries discarded to simulate the network's loss of memory.
-The next 9000 steps were predicted and compared with true data.
-
-\begin{figure}[ht]
-  \centering
-  \minipage{0.5\linewidth}
-    \centering
-    \includegraphics[width=0.8\linewidth]{_thomas_noisy_0.1056.eps}
-  \endminipage\hfill
-  \minipage{0.5\linewidth}
-    \centering    \includegraphics[width=0.8\linewidth]{_thomas_clean_0.1056.eps}
-  \endminipage
-  \captionof{figure}{True(red) and predicted(blue) trajectories for the Thomas attractor with parameter value $\beta=0.1056$.
-To the left: Noise added from normal distribution with mean zero and standard deviation 0.05.
-Right: No noise added to dataset.}
-\end{figure}
-
-The RNN was initialised as follows: the dimension of the network is set equal to 500 neurons and 12 hidden layers of dimension 64 each were initialised.
-Training is realised through 150 epochs of batch sizes of 128 each; the map is learnt via the Adam Optimiser using finer and finer learning rates (0.001 and then divided by 10 at every iteration).
-
-\begin{figure}[ht]
-  \centering
-  \includegraphics[scale=0.35]{_Thomas_1.eps}\caption*{Predicted trajectories of the $x$- and $y$-coordinates for the Thomas attractor with parameter-value $\beta=0.1056$ demonstrate empirically the ability to predict the evolution of the trajectory for the next an estimated 100 timesteps into the future near-exactly.
-Here no noise was added.}
-  \includegraphics[scale=0.5]{_Thomas_3.eps}\caption*{Represented here the learnt (blue) and actual (red) densities of the first coordinate ($x$-coordinate) of the dataset for the Cyclically Symmetric Attractor with parameter-value $\beta=0.1056$.}
-\end{figure}
-
-\subsubsection{Fractal Dream Attractor}
-
-Another 2-equation discrete-time map named the Fractal Dream Attractor, or more commonly known as the Pickover Map (first discovered by Clifford A.
-Pickover and discussed in his fascinating book "Chaos in Wonderland"\cite{PickoverChaos}), was also considered.
-\begin{eqnarray}\label{eqns_clifford}
-  {x_{n+1}=\sin(ay_n) + c\cos(ax_n)} \\
-  {y_{n+1}=\sin(bx_n)+d\cos(by_n)}
-\end{eqnarray}
-
-\begin{figure}[ht]
-  \centering
-  \includegraphics[width=1.0\textwidth,left]{_Clifford_1_nonoise.eps}
-  \caption*{These graphs were constructed by predicting the Clifford system 1000 steps into the future and in so doing illustrating the long-term consistency and accuracy of the learnt system.
-As perceived here, we are able to lock on to the trajectory of the Clifford map almost exactly for the first 25 steps.}
-  \minipage{0.5\textwidth}
-      \centering
-      \includegraphics[width=\linewidth]{_Clifford_2_nonoise.eps}
-  \endminipage\hfill
-  \minipage{0.5\textwidth}
-    \centering
-    \includegraphics[width=\linewidth]{_Clifford_3_nonoise.eps}
-  \endminipage
-  \caption{The bottom row, left, shows a plot of the true attractor(red) and the predicted values(blue).
-Bottom row, right, shows the distribution of the first coordinate of the learnt system (blue) and the actual Clifford system (red).
-}
-  \label{fig:Clifford}
-\end{figure}
-
-The behaviour of the system with a parameter values of $x_0=0$, $y_0=0$, $a = -1.7$,  $b = 1.8$, $c = -1.9$ and $d = -0.4$ was examined.
-Data was scaled to fit inside the interval $[-1,1]$. Only the noise-free case was investigated. 
- A sequence of 25000 observationswas fed into the system with 15000 entries discarded to simulate the network's loss of memory. The next 1000 steps were predicted and compared with true data.
-The RNN was initialised exactly as for Thomas' Cyclically Symmetric Strange attractor: the dimension of the network is set equal to 500 neurons and 12 hidden layers of dimension 64 each were initialised.
-Training is realised through 150 epochs of batch sizes of 128 each; the map is learnt via the Adam Optimiser using finer and finer learning rates (0.001 and then divided by 10 at every iteration).
-
-\section{Functional Complexity Reduction in Noninvertible maps}
-
-% Besides being able to learn the action of the Koopman operator exactly, our setup has other advantages over selecting or optimizing the choice of observables as in the Extended Dynamic Mode Decomposition (EDMD) algorithm (e.g., [21, 22]).
-We can alter the functional complexity of G T by changing the number of observables or equivalently by altering the dimension of X – in the case of a RNN implementation of g, by simply changing the number of neurons in the network.
-Empirically, increasing the dimension of X is found to increase the linear relationship (or intuitively reduces the functional complexity of G T ) that is measured as a generalization of the Pearson correlation coefficient to random vectors (e.g., [44]) between (x_{n−1} , x_{n}) and G_T (x_{n−1} , x_n), and such numerical evidence is tabulated in [11, Table 1].
-In contrast, expanding the set of observables while using EDMD to learn a map with a lower functional complexity is a highly involved task, not least because so much that one often does not know how to guess a new observable, as well as the fact that adding observables does not necessarily retain the invariance of the span of the observables.
-
-% Empirical evidence shows that a linearity measure like a Pearson correlation coefficient points at reduced functional complexity while learning the dynamics in the state space of the driven system.
-
-% Empirically (see Table ??), increasing the dimension of X is found to increase the linear relationship
-% (or intuitively reduce the functional complexity of G T ) that is measured as a generalization of the Pearson correlation coefficient to random vectors (e.g., [?]) between (x n−1 , x n ) and G T (x n−1 , x n ).
-If Σ a and Σ b denotes the covariance matrices of the vectors (x n−1 , x n ) and G T (x n-1 , x n ) respectively, and if Σ ab denotes the covariance matrix between the vectors (x n-1 , x_n ) and G_T (x n-1 , x_n ), then the multidimensional 18correlation coefficient is computed by using the traces of these matrices:
-% \[\pho = \frac{tr(\sum_{ab})}{tr(\sqrt{\sum_{a}\sum_{b}})}\]
-
-% This multidimensional correlation coefficient satisfies most of the well-known properties of the one-dimensional Pearson coefficient.
-In particular, ρ = ±1 if and only Y = AX + b for some invertible matrix A and vector b.
-Hence ρ, and in practice, the estimator ρ̂, can be used to measure the linear relationship between two random vectors of the same dimension and thus serves as an indicator of functional complexity.
-
-% We remark that the estimation of such linear relationship in (see Table ??) obtained by sample correlations alone is justified whenever {u n } is a realization of an ergodic process since then when g has the USP, any solution {x n } is also a realization of an ergodic process [?].
-In the ergodic input case the generalized Pearson correlation coefficient (between (x n−1 , x n ) and G T (x n−1 , x n )) is independent of n.
-
-In the case of non-invertible maps, we note that the map $G_T$ is a homemorphism although $T$ is not.
-Further, empirically it is found that $G_T$ is much more linear than $T$ as measured by what is called the multidimensional correlation coefficient $\rho$.
-This coefficient is a generalization of the Pearson correlation coefficient for real-valued random variables to random vectors (e.g., \cite{puccetti2019measuring}).
-We state teh details of computing this coefficient between
-$(x_{n-1},x_n)$ and $G_T(x_{n-1},x_n)$. 
-If $\Sigma_{a}$ and $\Sigma_{b}$ denotes the covariance matrices of the vectors $(x_{n-1},x_n)$ and $G_T(x_{n-1},x_n)$ respectively, and if $\Sigma_{ab}$ denotes the covariance matrix between the vectors $(x_{n-1},x_n)$ and $G_T(x_{n-1},x_n)$, then the multidimensional correlation coefficient is computed by using the traces of these matrices:
-\[
-    \rho= \frac{\text{tr}({\Sigma_{ab}})}{\text{tr}({\sqrt{\Sigma_a\Sigma_b}})}.
-\]
-This multidimensional correlation coefficient satisfies one of the desired and well-known properties of the one-dimensional Pearson coefficient.
-That is, $\rho=\pm 1$ if and only if $Y\overset{d}{=}AX+b$ for some invertible matrix $A$ and vector $b$.
-Hence $\rho$, and in practice can be used  to measure the linear relationship between two random vectors of the same dimension and thus serves as an indicator of functional complexity.
-For the non-invertible system defined by the Clifford attractor, we indicate the difference by which $\rho$ changes for $G_T$ in Table~\ref{tbl_attractorsPearson}.
-Following \cite{manjunath2021universal}, we say that this multidimensional correlation coefficients $\rho$ to evidence a general reduction in the functional complexity of the map $G_T$ emerging due to a RNN.
-In the
-Table ~\ref{tbl_attractorsPearson}, the second column corresponds to the amount of artificial neurons (the dimension of $X$) in the implemented RNN.
-
-\begin{center}
-  \begin{table} 
-      \scalebox{0.75}
-      {\begin{tabular}{|c|c| c c |} 
-          \hline
-          Input & Dimension of $X$ & $u_n$ vs $u_{n+1}$ 
-          & $\begin{bmatrix}
-              x_{n-1}\\
-              x_n
-          \end{bmatrix}$ vs $\begin{bmatrix}
-              x_n\\
-              x_{n+1}
-          \end{bmatrix}$ \\
-          \hline      
-          % \hline
-          % $(w_n)$ is the Lorenz states, sampled every 0.1 timestamps. & & &\\
-          % \multirow{3}{*}{$u_n=\frac{1}{100}w_n$} & 10 &0.9311  & 0.9731\\
-          %     & 100 &  & 0.9934\\
-          %     & 1000 & & 0.9930\\              
-          % \multirow{3}{*}{$u_n=\frac{1}{10}(\sin(0.1w_{n,x})+\sin(0.1w_{n,y})+\sin(0.1w_{n,z}))$} 
-          %     & 10 &0.8401  & 0.9392\\
-          %     & 100 & & 0.9616\\
-          %     & 1000 & & 0.9737\\
-          % \hline
-
-               \hline  
-               $(w_n)$ comes from the Clifford Attractor \\ (for $a = -1.7; b = 1.8; c = -1.9; d = -0.4$): & & & \\
-        {$w_{n+1}= \begin{bmatrix} \sin(aw_{n,y}) + c\cos(aw_{n,x}) \\ 
-                                                                    \sin(bw_{n,x})+d\cos(bw_{n,y}) \end{bmatrix}$} & & & \\
-          \multirow{3}{*}{$u_n = w_n-\overline{w}$}
-              & 10 & -0.21713 & 0.78719 \\
-              & 100 & &  0.8334 \\
-              & 1000 & & 0.84854 \\
-          % \hline 
-          % $(w_n)$ comes from the Logistic map: & & & \\
-          % {$ w_{n+1}  = 4w_n(1-w_n)$} & & & \\
-          % \multirow{3}{*}{$u_n = w_n - 0.5$} 
-          %     & 10 &-0.0314  & 0.5577\\
-          %     & 100 &  & 0.7150\\
-          %     & 1000 & & 0.7826\\
-          % \hline
-          % \hline
-          % $(w_n)$ comes from the Svensson Attractor \\ (for $a = -1.7; b = 1.8; c = -1.9; d = -0.4$): & & & \\
-          % {$w_{n+1}= \begin{bmatrix} d\sin(aw_{n,x}) - \sin(bw_{n,y}) \\ 
-          %                                                             c\cos(aw_{n,x})+\cos(bw_{n,y}) \end{bmatrix}$} & & & \\
-          %   \multirow{3}{*}{$u_n = w_n-\overline{w}$}
-          %       & 10 & -0.2243 & 0.49698 \\
-          %       & 100 & &  0.63310 \\
-          %       & 1000 & & 0.71151 \\
-           \hline
-                \end{tabular}} %\ednote{http://paulbourke.net/fractals/peterdejong/}
- \label{Table_FC}
- \caption{Following \cite{manjunath2021universal}, we determine the multidimensional correlation coefficients $\rho$ to evidence a general reduction in the functional complexity of the map $G_T$ emerging due to a RNN.
-Rows correspond to distinct dynamical systems considered in the experiments
-      The second column corresponds to the amount of artificial neurons (the dimension of $X$) in the implemented RNN.
-      The last two columns comprise numerical estimates of $\rho$ for the relevant vectors. The linear association between $u_n$ vs $T(u_{n})$, as well as the linear relationship $\begin{bmatrix}
-              x_{n-1}\\
-              x_n
-          \end{bmatrix}$ vs $G_T \left( \begin{bmatrix}
-              x_{n-1}\\
-              x_n
-          \end{bmatrix}\right)$ may be contrasted by the reader. 
-          Finally, the RNN's dimension (from \eqref{Seq_RNN}) is provided to exhibiting that an increase in dimension of the driven system will 
-          typically result in a map $G_T$ with less functional complexity.}.
-      \end{table}\label{tbl_attractorsPearson}
-\end{center}
-
-\chapter{Conclusions}\label{ch6}
-
-% The driven states $(x_{n-1},x_n)$ hold the information of the left-infinite history of the input.
-Furthermore, there is some evidence that the inverse-limit space (not system) seems to distinguish two dynamical systems that are not topologically conjugate for a class of systems in the sense their inverse-limit spaces are not homeomorphic to each other.
-So the relation $Y_T$, which is homeomorphic to the inverse-limit system, may be unique to a class of systems.
-(The reference is a paper on Ingram’s Conjecture by Marcy Barge, Henk Bruin and Sonja Stimac.) Further investigation is needed.
-\cite{budivsic2012applied, korda2020data}
-
-There are a myriad of techniques to generate accurate models for observed dynamical systems using data obtained from measurement functions.
-Core to the success of such methods is the fact that we can often characterise the dynamics from am observed space by a mapping of much lower functional complexity when compared to the true map describing the unknown underlying system.
-Takens theorem and the numerous generalizations thereof is one such method establishing (under some generic conditions) the learnability of a system produced by creating a delay-coordinate vector from a sufficiently large number of previous observations from a dynamical system.
-In this case the new system is topologically conjugate to the system from which the observed time-series was obtained.
-A critical concern of such methods is the requirement of full knowledge of the underlying state space variables governing the dynamical system one is interested in.
-In real-world conditions this information is rarely available and, when it is, is often contingent upon human expertise and insight.
-
-Each discrete-time state space model has at its core a driven dynamical system. 
-In this project, it has been shown that an input originating from a compact space entailing exactly one solution (i.e., if the driven system has the USP) is equivalent to representing information without distortion, a concept which was formalised by establishing the existence of a causal mapping’s.
-The established results are general, implying that any driven dynamical system with specific properties, such SI-invertibility, can give rise to a causal mapping. 
-In particular, we have a topological conjugacy (semi-conjugacy) between the data comprising the single-delay lag dynamics of a driven system~\cite{manjunath2013echo} and the underlying homeomorphic dynamical system.
-Finite-length input data is shown to be sufficient for forecasting as the left-infinite segment of the past of the input beyond some arbitrarily large but finite time (a result which holds due to the universal semi-conjugacy of a driven system’s continuity) influences the driven system's states in an almost negligibly small manner.
-
-Viewed through the lens of reservoir computing, we have also demonstrated that when single-delay dynamics are utilised, a learnable map in a reservoir computing network certainly exists. 
-Moreover, within a reservoir computing network, such a driven system posssesses certain proven qualities pertaining to the robustness of the system to different forms of perturbation. 
-We are thus able to preserve the quality of a representation that would otherwise have remained sensitive to various fluctuations and rendering the learning of the system’s dynamics wholly unpredictable.
-
-Finally, we have considered some physical systems to observe the methodology in practice and verify that we can indeed predict the future evolution of chaotic systems such as the double pendulum.
-We demonstrated the robustness of our methodology to external noise by considering the cases of the  Dream Attractor and the Double Pendulum.
-
-\ednote{Loose paragraph commented out}
-% Considered from a philosophical-scientific  perspective, this project has aimed to establish a juncture between rationalism (the theory of nonautonomous dynamical systems) and empiricism (machine learning) to yield results profitable to the discipline of Applied Mathematics.
-% This was done by showing that dynamics in an observed space can be established as topologically conjugate to the underlying system. 
-% As a result of the topological conjugacy, deep learning techniques are then subsequently applied to physically such equations from the data. 
-% In so doing, we acquire models of high fidelity that provide exceptional forecasting results on chaotic data that exhibit both statistical and topological consistency in the long run.
-
-\bibliographystyle{pnas-new}
-\bibliography{pnas}
-
-\end{document}

### 2022-01-17 projectReport/chapter1.tex
Cleaner setup

--- 2022-01-17 draft_16Jan/hiscurrent.tex
@@
-\documentclass[a4paper,12pt,twoside]{report}
-% \documentclass[a4paper,12pt,twoside]{book}
-\pagestyle{headings}
-\usepackage[utf8]{inputenc}
-\usepackage{filecontents}
-\usepackage[many]{tcolorbox}
-\usepackage{graphicx}
-\usepackage{harpoon}
-%\usepackage{apacite} 
-\usepackage{xcolor}
-\usepackage{latexsym}
-\usepackage{amssymb}
-\usepackage{amsmath}
-\usepackage{mdframed}
-%\usepackage[authoryear]{natbib}
-\usepackage{hyperref, url}
-\usepackage[show]{ed}
-\usepackage{subfigure}
-\usepackage[ntheorem]{empheq} 
-\usepackage{amsthm, amssymb, amsfonts, latexsym}
-% \usepackage{pst-all}
-
-\usepackage{enumitem}
-\usepackage{multirow}
-\usepackage{longtable}
-\usepackage[export]{adjustbox}
-% $f\arrowvert_A$ To restrict functions
-
-\setlength{\parskip}{1.7em}
-\setlength{\parindent}{0em}
-
-%\setlength{\baselineskip}{20pt}
-
-\setlength{\textwidth}{429.75499pt}
-\setlength{\textheight}{643.20255pt}
-\setlength{\oddsidemargin}{5 mm}
-\setlength{\evensidemargin}{5 mm}
-\setlength{\topmargin}{0 mm}
-\setlength{\headsep}{0 mm}
-\setlength{\headheight}{0 mm}
-
-\usepackage{pst-node}
-
-\newcommand{\citealtt}[1]{\citeauthor{#1},\citeyear{#1}}
-\newcommand{\myycite}[1]{\citep{#1}}
-
-\mathchardef\mhyphen="2D
-
-%\newcommand{\cev}[1]{\reflectbox{\ensuremath{\vec{\reflectbox{\ensuremath{#1}}}}}}
-
-\usepackage{pst-node,graphicx,pst-blur}
-%\uspackage{auto-pst-pdf}
-%\usepackage{tikz-cd}
-
-\makeatletter
-\DeclareRobustCommand{\cev}[1]{%
-  \mathpalette\do@cev{#1}%
-}
-\newcommand{\do@cev}[2]{%
-  \fix@cev{#1}{+}%
-  \reflectbox{$\m@th#1\vec{\reflectbox{$\fix@cev{#1}{-}\m@th#1#2\fix@cev{#1}{+}$}}$}%
-  \fix@cev{#1}{-}%
-}
-\newcommand{\fix@cev}[2]{%
-  \ifx#1\displaystyle
-    \mkern#20mu
-  \else
-    \ifx#1\textstyle
-      \mkern#20mu
-    \else
-      \ifx#1\scriptstyle
-        \mkern#26mu
-      \else
-        \mkern#26mu
-      \fi
-    \fi
-  \fi
-}
-
-\makeatother
-
-\newcommand{\xcm}{\epsfxsize=3.1cm}
-\newcommand{\fig}[1]{\epsfbox}
-% \newcommand{\bp}{\begin{minipage}{3.1cm}}
-% \newcommand{\ep}{\end{minipage}}
-
-\newtheorem{Definition}{Definition}[]
-%\newmdtheoremenv{Theorem}{Theorem}[]
-\newtheorem{Theorem}{Theorem}[]
-\newtheorem{Lemma}{Lemma}[]
-\newtheorem{Proposition}{Proposition}[]
-\newtheorem{Corollary}{Corollary}[]
-\newtheorem{Remark}{Remark}[]
-\newtheorem{Example}{Example}[]
-
-%Shortcut symbols
-\newcommand{\Ftheta}{\ensuremath{F_\theta}}
-
-\makeatletter 
-\renewcommand{\thefigure}{\@arabic\c@figure}
-\makeatother
-\usepackage{xr}
-
-%%%% PNAS GRAPHICS SETUP- FROM THE PNAS STYLE LATEX FILE
-\RequirePackage{graphicx,xcolor}
-\RequirePackage{colortbl}
-\RequirePackage{booktabs}
-\RequirePackage{algorithm}
-\RequirePackage[noend]{algpseudocode}
-\RequirePackage{changepage}
-\RequirePackage[twoside,%
-				letterpaper,includeheadfoot,%
-				layoutsize={8.125in,10.875in},%
-                layouthoffset=0.1875in,%
-                layoutvoffset=0.0625in,%
-                left=38.5pt,%
-                right=43pt,%
-                top=43pt,% 10pt provided by headsep
-                bottom=32pt,%
-                headheight=0pt,% No Header
-                headsep=30pt,%
-                footskip=25pt,
-                marginparwidth=38pt]{geometry}
-\RequirePackage[labelfont={bf,sf},%
-                labelsep=period,%
-                figurename=Fig. ]{caption}
-\setlength{\columnsep}{13.5pt} % Distance between the two columns of text
-\setlength{\parindent}{12pt} % Paragraph indent
-
-%% Figure caption style
-\DeclareCaptionFormat{pnasformat}{\normalfont\sffamily\fontsize{7}{9}\selectfont#1#2#3}
-\captionsetup*{format=pnasformat}
-
-%%% GREYBOX AROUND FIG
-\definecolor{lightgray}{gray}{0.95}
-\newcommand\greybox[1]{%
-  \vskip\baselineskip%
-  \par\noindent\colorbox{lightgray}{%
-    \begin{minipage}{\textwidth}#1\end{minipage}%
-  }%
-  \vskip\baselineskip%
-}
-
-\newtcolorbox{paperbox}[1][]{
-    enhanced,
-    colback=white,
-    boxrule=0pt,
-    boxsep=0pt,
-    arc=0mm,
-    width=0.8\linewidth,
-    fuzzy shadow={0mm}{-4pt}{-4pt}{1mm}{black!30!white},
-    #1
-}
-
-% \externaldocument{Supp_Koopman-0808}
-\begin{document}
-
-\title{Data-Driven Modelling \& \\ Forecasting of Chaotic Systems}
-
-\author{B.L. Nortier} 
-\date{January, 2022}
-\date{}
-\maketitle
-
-\begin{abstract}
-  It is a well-established practice to collect measurements from underlying artificial, natural, and physical systems through observables in order to realise a dynamical model with greater descriptive power.
-  Many data-driven approaches to forecast the future evolution of systems exist.
-However, many of these methods rely on an initial understanding of the system and we often do not fully understand the many complexities of more intricate systems that we interact with on a daily basis.
-
-  In this project, the base assumption is that data is originates from  some discrete-time dynamical system, particularly one of a chaotic nature but no assumptions are made pertaining to a specific dynamical system.
-  A general learning problem is formulated and a well-known approach to the problem, alongside its practical limitations, is then considered.
-  The notion of driven dynamical systems is introduced, and we subsequently establish the theoretical existence of a dynamical system possessing certain properties which guarantees it to be topologically conjugate (`dynamically equivalent`) to the dynamics of the underlying system from which measurements were originally taken.
-  Our methodology is implemented practically by making use of recurrent neural networks, a form of deep learning, to learn a map topologically equivalent to the underlying system that forecasts future states.
-  In so doing we present the workability of an approach to obtain long-term topological and statistically consistent predictions for simple physical systems as well as a some chaotic attractors.
-  \end{abstract}
-
-  \tableofcontents
-
-\chapter{Introduction}\label{ch1}
-
 Experimenting with biological, physical, and artificial systems to generate a more informative dynamical model is a well-established practice in modern science.
@@
 \newline Finally, in Chapter~\ref{ch5} we show the implementation of these forecasting methods, and conclusions are provided in Chapter~\ref{ch6}.
-
-\chapter{Discrete-time Dynamical Systems}\label{ch2}
-
-This chapter briefly describes a discrete-time dynamical system and what it means to exhibit chaos. We refer to \cite{devaney2018introduction, de2013elements} for more details.
-
-At its most elementary level, a dynamical system is just something that evolves deterministically through time.
-In the context of this project, deterministic refers to the fact that a system evolves according to specified rules rather than based on random events.
-Dynamical systems arise in a variety of situations.
-A continuous-time dynamical system describes the states for all values of the time.
-Specifically, if the motion of a pendulum in which the quantities such as the angular position and angular momentum are known at all times, then it is a continuous-time dynamical system.
-The equations of the dynamical system can take the form of one or more ordinary differential equations that determine the relevant quantities at any future time if we know the initial location and momentum.
-In ecology, discrete-time dynamical systems are widely used to model population growth.
-The model in this case is a function calculating the following generation's population given the population of the previous generation.
-If we know the starting population, we may once again calculate the population at any time in the future.
-
-Formally, a function $T: U \to U$, where $U$ is some set is  a \emph{discrete-time dynamical system} and its iterates $\{u,Tu,T^2u,\ldots\}$, where $T^n$ denotes the $n$-fold composition of $T$ with itself, describe the evolution of an initial condition $u\in U$ (Note that we frequently drop the brackets and denote $T(u)$ by $Tu$ so as to simplify notations).
-
-Continuous-time dynamical systems modelled using ordinary differential equations can give rise to discrete-time dynamical systems.
-To see this, consider a differential equation $\dot{x} = f(x)$, $f: \mathbb{R}^n \to \mathbb{R}^n$, $n\in\mathbb{N}$ given to have a unique solution passing through
-each point $x\in\mathbb{R}^{n}$, call it $x(t)$ where $x(0)=x_0$.
-
-\begin{Definition}
-  [\bf Flow of an Equation] \label{Dfn_Flow}\rm
-  The flow of the equation  $\dot{x} = f(x)$  is defined to be a mapping $\varphi: \mathbb{R}^n \times \mathbb{R} \to \mathbb{R}$ where $\varphi(x_0,t)= x(t)$ for the solution $x(t)$ with $x(0)=x_0$ 
-\end{Definition}
-
-By fixing $t=K \in (0,\infty)$, we can define the \emph{time-$K$ map} as  $T(x):= \varphi(x_0,K)$, and it is easily verified that $T\circ T(x_0) = \varphi(x_0,2K)$.
-In general the $m^{\mbox{th}}$ iterate of $x_0$ under $T$ would be the value of the solution of the ODE evaluated at time $mK$ with the initial condition $x_0$, i.e.
-$\varphi(x_0, mK)$.
-Thus ordinary differential equations give rise to a discrete-time dynamical system by sampling the value of the solution $x(t)$ at time intervals $K$ units apart.
-
-A numerical discretization of a differential equation can also give rise to a discrete-time dynamical system.
-For instance, Euler's method approximates $\dot{x}(t)$ by $(x(t+h)-x(t))/h$; if $h$ is fixed throughout, the solution of a differential equation $\dot{x}=f(x)$
-can be approximated at the time instant $t+(m+1)h$ by iterating the equation 
-\begin{equation}
-  x(t+(m+1)h) = x(t+mh) + h f(x(t+mh))
-\end{equation}
-
-Adopting more succinct notation by replacing $x(t+mh)$ with $u_m$, we rewrite the above equation as
-\begin{equation}
-u_{m+1} = u_m + hf(u_m)
-\end{equation}
-or in even simpler terms as the discrete-time dynamical system with map $T(u) = u + hf(u)$, where $T$, $u$ and $f$ are understood to be as above.
-
-Of course, discrete-time dynamical systems need not always arise through a differential equation.
-Once again, we may consider the field of ecology, where discrete dynamical systems are often directly derived or assumed.
-\ednote{Ask for resource}
-There is a school of thought that advocates discrete-time dynamical systems to be more natural for modelling real-world observations than differential equations.
-We refer the interested reader to~\cite{saber2010introduction}.
-
-\section{Invariant Sets}
-
-A core concept in the study of dynamical systems is that of invariance. 
-\begin{Definition}
-  [\bf Invariant Set]\label{Dfn_InvariantSet}\rm
-  Given a discrete-time dynamical system $T: U \to U$, a subset $A \subset U$ is said to be an \emph{invariant set} if $T(A) =A$. 
-\end{Definition}
-
-We also define the orbit of a function $T$.
-\begin{Definition}
-  [\bf Orbit of $T$]\label{Dfn_Orbit}\rm
-  The orbit of $T$ is to be defined the sequence $\bar{u} = \{u_n\}_{n\in \mathbb{Z}}$ obeying the update equation, $u_{n+1}=Tu_n$, $n \in \mathbb{Z}$. 
-\end{Definition}
-
-Two examples of invariant sets include a fixed point where $Tu=u$ for  $u\in U$, and a periodic orbit, i.e., a set of iterates $\{u,Tu, T^2u,\ldots,T^pu\}$, where $T^{p+1}u=u$ for some $p\in\mathbb{Z}$.
-The entire space $U$ could be also be invariant.
-Consider for example the space $U=[0,1]$, where $Tu=4u(1-u)$ and then $U$ is invariant (as every $u\in{U}$ can be written as $u = 4x(1-x)$ for some $x\in{U}$).
-
-We may learn a great deal about the iterates of a dynamical system by considering the types of invariant sets of a discrete-time dynamical system.
-For example, if $U=[0,1]$ has map $Tu= u/2$, then the only invariant set is $\{0\}$, and all orbits approach this invariant set as time flows in the forward direction.
-Indeed, if a some non-zero invariant set (call it $B$) exists, then there is  some $r\in(0,1]\cap{B}$.
-But $r\notin{T(B)}$ since any orbit with initial value $r$ will be a decreasing sequence.
-Moreover, every orbit of T will be decreasing and therefore approach the value $0$ as $n\rightarrow\infty$.
-
-One may ask if every orbit approaches an invariant set?
-In general the answer is no, since for the dynamical system $T: \mathbb{R} \to \mathbb{R}$ defined by $Tu=2u$, any orbit that does not intersect the invariant set $\{0\}$, will not approach any invariant set.
-
-However,  when the space $U$ is compact, all orbits approach an invariant set.
-This is since the set of limit points of the orbit can be shown to be invariant \cite{de2013elements}.
-So, when $U$ is compact,  the $\omega$-limit set $\omega(u;T)$ of a point $u$ defined to the collection of limit points of the sequence $\{x,Tu,T^2u,\ldots\}$ is nonempty, and $\omega(u;T)$ is invariant.
-
-\begin{Example}
-  For the map $Tu=u^2$ defined on $[0,1]$ all orbits lie in the invariant set $\{1\}$ or else would would approach the invariant set $\{0\}$. 
-\end{Example}
-
-Invariant sets have various properties. Invariant sets can be attracting or repelling depending on how orbits in their vicinity behave. 
-Recall the examples $Tu =u/2$ and $Tu=2u$ defined on $\mathbb{R}$ (where $\{0\}$ “attracts" orbits) or the example, $Tu=u^2$ defined on $[0,1]$ (where $\{1\}$ “repels" orbits). 
-We are interested in attractive invariant sets since they capture the long-term dynamics as time increases. 
-In particular, we are interested in those invariant sets named attractors.
-
-\begin{Definition}
-  [\bf Attractor]\label{Dfn_Attractor}\rm
-  Let $T: U \to U$, where $U$  is a metric space with metric $d$. A compact subset $A \subset U$ is said to be an attractor if it satisfies the three conditions: 
-  \vspace{-8mm}
-  \begin{enumerate}
-	\item $A$ is invariant. 
-	\item $A$ is asymptotically stable, i.e., for every $\epsilon > 0$ and for all $u$ so that $d(u,A) < \epsilon$, we have $d(T^nu,A) \to 0$ as $n\to \infty$. 
-	\item $A$ has Lyapunov stability, i.e., for every $\epsilon > 0$  there exists a $\delta(\epsilon) > 0$ so that $d(u,A) < \delta$ implies $d(T^nu,A) < \epsilon$ for all $n\ge 0$.  
-\end{enumerate}
-\end{Definition}
-
-Indeed in our previous examples, it can be verified that the system  $Tu=u/2$ defined on  $\mathbb{R}$ and $Tu=u^2$ defined on $[0,1]$  the singleton set $\{0\}$ is an attractor.
-For the system,  $Tu=1-|2u-1|$ on $[0,1]$, one may easily verify that the only attractor is the entire space $[0,1]$.
-This follows from the fact that between any two points $u< v$ in $[0,1]$, and for any $a,b$ so that $u\le a < b \le v$, we can find an $n$ so that $T^n(a,b)=[0,1]$ (as would be explained later in this chapter).
-
-A dynamical system can have several attractors and may also be contained in another attractor.
-For the example, $Tu =u^2$ on $[0,1]$, both $\{0\}$ and $[0,1]$ are attractors.
-It is known that an attractor for a dynamical system always exists in a compact space~\cite{Milnor1985}.
-
-The dynamics restricted to an invariant set can be complicated.
-For instance an invariant set could be just a single point or it could have an infinite set.
-If the invariant set is infinite, then complicated dynamics are possible.
-A particular, well-studied phenomenon of such complexity gives rise to so-called chaotic behaviour, a subject studied in detail over the past fifty years.
-
-% If the dynamics on the attractor are somewhat complicated in the sense that the attractor cannot be decomposed further, and if the attractor is infinite then there is a possibility of complicated dynamics, and a particular well-studied phenomenon of such complexity gives rise to what is called a chaotic behaviour.
-
-%If the dynamics on an attractor cannot be decomposed further and the attractor is itself also infinite, then there is a possibility of complicated dynamics, and a particular well-studied phenomenon of such complexity gives rise to what is called chaotic behaviour.
-
-\section{Chaos}
-
-In the 1960s, several mathematicians and mathematically interested scientists independently discovered chaos in the mathematical sense.
-The meteorologist Edward Lorenz may have been the first to explain this phenomenon in his 1963 paper \cite{lorenz1963deterministic}.
-The notions of invariance, attractivity, and chaos may also be described for continuous systems, and Lorenz's system comprised of a system of differential equations.
-The narrative of Lorenz's discovery of chaos and the history of other forerunners in this subject is fascinating.
-We highly recommend James Gleick's book Chaos: The Making of a New Science~\cite{gleick2008chaos} for those interested.
-It clearly illustrates these experiences, explains why chaos was such a startling and crucial mathematical and scientific discovery and describes the underlying mathematical notions for non-specialists.
-
-Different authors propose a number of intricately different definitions of chaos in the literature~\cite{RasbandChaos, TaborChaos, WigginsChaos} and each of them indicate some aspect of complexity.
-
-In practice, when only data is observed from a system, it is not possible to verify which definition or notion is satisfied by the underlying dynamical system.
-We do, however, specifically recall the definition of chaos in the sense of Devaney \cite{devaney2018introduction,de2013elements} so as to understand some nuances behind the complexity.
-Devaney's definition of chaos has three requirements, with the first being the notion  of sensitive dependence on initial conditions.
-
-\begin{Definition}\rm
-  [\bf {Sensitive Dependence on Initial Conditions}]\label{Dfn_SDIC}\rm
-A dynamical system $T: U \to U$ is said to have sensitive dependence on initial conditions (SDIC) if there exists a $\delta > 0$ such that for every $u \in U$ and in every neighborhood of $u \in U$ there exists a $v\in{U}$ and an integer $N:=N{(u,v)}\in\mathbb{Z}$ such that $d(T^Nu,T^Nv)>\delta$.
-\end{Definition}
-
-It is vital to acquire a sense of this definition as it can easily be misunderstood.
-It is a common misconception to interpret SDIC as two close points ($u$ and $v$) that eventually become separated by a distance $\delta$ under iteration by $T$.
-But this is not true.
-To fully comprehend all the subtleties of the concept, one needs to discuss it more thoroughly by considering each sentence with care and attention, whereafter one may examine how they fit together to convey the idea of sensitive dependence.
-
-To this end, we make 3 remarks:
-\vspace{-5mm}
-\begin{enumerate}
-  \item First, the $\delta>0$ in the definition of SDIC is independent of $u$. 
-  \item Second, in every neighborhood of $u$, we may not necessarily find all points $v$ in the neighborhood distinct from $u$ that would separate from the forward iterates of $u$. 
-  \item  Finally, $N$ depends upon $u$ and $v$ chosen, and their iterates may not separate forever (i.e., for all $n>N$) and we allow their iterates to get arbitrarily close in the future. 
-\end{enumerate}
-
-To illustrate the concept of SDIC, we present two examples - one from Mathematics and the other from the field of Physics.
-
-\begin{Example} \rm
-  The logistic map(LM), a recursion relation of the form $x_{n+1}=rx_n(1-x_n)$ where $x_n\in[0,1]$ and $r\leq{4}$ is a classic example used to illustrate the chaotic behaviour that a system may exhibit.
-When $r=4$, the one-dimensional recurrence relation $x_{n+1}=4x_n(1-x_n)$ can be used as the kindergarten's model to exhibit the presence of SDIC.
-\ednote{B: edited and removed unnecessary citation}
-  In the graph below~\ref{fig:log_sdic} we plot the first 50 iterates of the LM for $r=4$ and two slightly different initial values $x_0$.
-
-  \begin{figure}[ht]
-    \includegraphics[scale=0.74]{_logistic_sdic.eps}
-        \centering
-        \captionof{figure}{Figure generated for Logistic Map $x_{n+1}=rx_n(1-x_n)$ with $r=4$ to exhibit the presence of SDIC.
-Plotted are the values $x_n$ against time $n$ for the the first 50 step with initial values $x_0=0.2$ (blue) and $x_0=0.20001$ (red).
-Initially the two trajectories overlap, but they diverge completely at $n=12$ whereafter they follow distinct paths.}
-        \label{fig:log_sdic}
-      \end{figure}
-  \ednote{B: Caption and figure adjusted}
-  % For an intuitive sense of the fact that the LM exhibits SDIC, consider again the definition and substitute specific values.
-Let $\delta=0.05$ and for purposes of illustration consider the value $u=0.4$.
-The green band in~\ref{fig:log_sdic} shades the region $[u-\delta, u+\delta]$ and we can see that regardless the value $v$  and in every neighborhood of $u \in U$ there exists a $v\in{U}$ and an integer $N:=N{(u,v)}\in\mathbb{Z}$ such that $d(T^Nu,T^Nv)>\delta$
-  % specific point, say $u=0.4$.
-Draw a green neighbourhood $[0.35,0.45]$ of diameters $\delta=0.02$ around the point.
-It is easy to see that every neighbourhood around $u=0.4$ will contain some $v_1, v_2\in[0,1]$ and natural number $N_1, N_2 \in\mathbb{N}$ respectively s.t.
-the distance $d(T^Nu_1,T^Nv)>\delta=0.05$ .
-\end{Example}
-
-\begin{Example} \rm
-  A second example pertains to a simple physical system called the double pendulum - a pendulum with another pendulum attached to its head given an initial position and angular velocity.
-
-  Below are two figures denoting trajectories of a double pendulum's second head after some elapsed time.
-Depicted are three pendulum heads with equal angular velocity but differing slightly in initial position.
-As can be seen below, the trajectories diverge very quickly to become completely different.
-We discuss the double pendulum in greater detail in a later chapter (\ref{ch5}), but we mention the example here to provide a numerical example of a practical system exhibiting SDIC.
-
-\begin{figure}[ht]
-  \centering
-\includegraphics[scale=0.8]{_dp_sdic.eps}
-    \captionof{figure}{Three double pendulum heads with equal angular velocity and initial angles differing by 0,025 radians}\label{fig:dp_sdic}
-  \end{figure}
-
-\end{Example}
-
-The second part in Devaney's definition of chaos concerns topological transitivity.
-
-\begin{Definition}
-  [\bf {Topologically Transitive}]\label{Dfn_TopolTrans}\rm
-	A dynamical system $F: U \to U$ is topologically transitive if for any pair of nonempty open sets $E_1$ and $E_2$ there exists a $n\in\mathbb{N}$ such that $T^n(E_1) \cap E_2 \not= \emptyset$. 
-\end{Definition}
-
-Topological transitivity implies that iterates of an open set of initial conditions get mixed up with other open sets.
-On a compact metric space, one may show that topological transitivity also implies the existence of a point whose forward iterates are dense \cite{de2013elements}; or in other words, the orbit going through this point will be dense in the compact metric space.
-In fact, it is topologically more likely that the choice of an arbitrary point will be one whose iterates are almost dense.
-
-\begin{Definition}
-  [\bf {Meager Set}]\label{Dfn_Meager Set}\rm
-A subset of a topological space $U$ is said to be a meager set if it can be written as a countable union of sets of with empty interior. The complement of the a meager set is set to be a residual set.
-\end{Definition}
-
-To be more precise (see \cite{de2013elements}), for a discrete-time dynamical system on a compact space, the set of points with dense iterates are residual, and they are typical or likely to be observed in practice.
-In this project, when data is observed from a topologically transitive system, we assume it arises from a dense orbit.
-
-We may now define the notion of Chaos as formulated by Devaney\cite{devaney2018introduction}.
-\begin{Definition}
-  [\bf {Devaney's Chaos}]\label{Dfn_ChaosDec}\rm
-	A dynamical system $T: U \to U$ is said to exhibit chaos in the sense of Devaney if it satisfies the three properties:
-	\vspace{-5mm}
-  \begin{enumerate}
-		\item $T$ has SDIC.
-		\item $T$ is topologically transitive.
-		\item The set of periodic points of $T$ are dense in $U$. 
-	\end{enumerate}
-\end{Definition}
-
-\begin{Example} \rm
-  The standard tent map $Tu=1-|2u-1|$ defined on $[0,1]$ is a well-known example of a dynamical system satisfying these three properties.
-
-  We now reason as to why this is true.
-The graph of the map $T$ is piecewise linear with two straight lines, one connecting the points $(0,0)$ and $(\frac{1}{2},1)$ and the other connecting $(\frac{1}{2},1)$ with $(1,0)$.
-They form a so-called tent with base centered at $u=1/2$.
-The graph of the map $T^2$ comprises two symmetric tents with their base centered at $1/4$ and $3/4$.
-See Figure \ref{fig:T2tentmap}
-
-  \begin{figure}[ht]
-    \includegraphics[scale=0.6]{_tentmap_2.eps}
-        \centering
-        \captionof{figure}{Graph of the $T^2$ map with bases at $x=0$, $x=\frac{1}{2}$ and $x=1$ respectively.}\label{fig:T2tentmap}
-  \end{figure}
-
-  In general, the graph of  $T^k$ contains $2^{k-1}$ tents.
-  Given any point $u$ and an open neighborhood $W \subset[0,1]$ of $u$, we can find $k\in\mathbb{Z}$ large enough so that $T^k(W) = [0,1]$ because we can accommodate a tent whose base is contained in $W$.
-% \ednote{B: This last phrase doesn't make sense to me.
-M: The number of tents grow with $k$ and hence this is obvious; I explained in one of our meetings.
-Does it make sense?}
-  This implies the existence of two points $w_1$ and $w_2$ in $W$ so that $d(T^kw_1,T^kw_2)= 1$, and by the triangle inequality  at least one of the inequalities $d(T^ku,T^kw_1)> \delta$ or $d(T^ku,T^kw_2)> \delta$ hold when $\delta\in (0,\frac{1}{2})$.
-So $T$ has sensitive dependence on initial conditions.
-
-  Next, let $W_1$ and $W_2$ be two nonempty open sets.
-Given any open set $W_1$, we can find a $k\in\mathbb{N}$ large enough so that $T^k(W_1) = [0,1]$ because we can accommodate a tent with base contained in $W$.
-So $T^k(W_1) \cap W_2 \not=\emptyset$, and thus $T$ has topological transitivity.
-
-  Finally, for every open interval $(a,b)$, the graph of $T^k$ intersects the graph of the identity map on $[0,1]$.
-This follows from the fact established above that there exists a $k\in\mathbb{N}$ such that $T^k(a,b) = [0,1]$.If the intersection point has coordinates of the form $(p,p)$, the point $p$ is then a fixed point of $T^k$ and therefore also a periodic point of $T$.
-Since we have found a periodic point in the interval $W$,  the set of periodic points are dense in $T$.
-  This follows because any fixed point of $T^k$ is also a periodic point, and we have found this in an arbitrary interval.
-
-\end{Example}
-
-\section{Conjugacy}
-
-We now turn our attention to the subject of conjugacy that describes when two dynamical systems are equivalent dynamically.
-
-To show topological similarity (or sameness) between two metric or topological spaces, one needs to establish a homeomorphism between the two spaces. 
-However, in the study of dynamical systems defined on two spaces, establishing a homeomorphism does not indicate that the systems are dynamically related in any way.
-For instance, the maps $Tu=u^2$ and $Tu=1-|2u-1|$ defined on $[0,1]$ have totally different behaviour. 
-Therefore, to find dynamically similar systems, one must establish a dynamical equivalence that we illustrate in a simple commutativity diagram below~\ref{eqn_conjugacy}.
-
-% Giving errors
-\begin{equation}  \label{eqn_conjugacy}
-%    \[ 
-    \psset{arrows=->, arrowinset=0.25, linewidth=0.6pt, nodesep=3pt, labelsep=2pt, rowsep=0.7cm, colsep = 1.1cm, shortput =tablr}
- \everypsbox{\scriptstyle}
- \begin{psmatrix}
-U & U\\%
-V & V.
- %%%
-%  \ncline{1,1}{1,2}^{T} 
-%  \ncline{1,1}{2,1} <{\phi}
-%   \ncline{2,1}{2,2}^{S}
-%  \ncline{1,2}{2,2} > {\phi}
- \end{psmatrix}
-% \]
-\end{equation}
-
-%Composite functions
-To understand the diagram, we note that if we travel "right and then down," the diagram instructs us to use $T$ (top arrow right) first, followed by $\phi$.
-(right arrow downwards).
-Consequently, the pathway amounts to finding $\phi(T(u))$.
-If we proceed "down and then right," the diagram instructs us to apply $\phi$ (left arrow downwards) first, and then apply $S$ (bottom arrow right).
-This then amounts to finding  $S(\phi(u))$.
-When the relationships in this diagram hold, we say $\phi(T(u))= S(\phi(u))$, and we formally denote it as $\phi \circ T=S\circ \phi$.
-
-\begin{Definition}\rm  
-[\bf {Conjugacy \& Semi-Conjugacy}]\label{Dfn_Conjugate}\rm
-   Consider the dynamical systems $T:U\to{U}$, $S:V\to{V}$ and suppose the relationship $\phi \circ T=S\circ \phi$ holds.
-If $\phi:U\to{V}$ is a homeomorphism, then $S$ is said to be conjugate to $T$ and $\phi$ is called a conjugacy.
-If we relax the criterion on $\phi$ and merely require $\phi$ to be continuous and surjective where $\phi:U\to{V}$, then $\phi$ is a semi-conjugacy between $T$ and $S$ where T has domain $U$ and S has domain $V$; S is said to be semi-conjugate to T.
-\end{Definition}
-
-\ednote{B: Does defining explicitly the domain and codomain of $\phi$ make it clear enough?
-M: Saying $\phi$ is then a semi-conjugacy between $T$ and $S$ could be ambiguous as it does not specify which system is on the top in the commutativity.
-So please edit accordingly.}
-When $S$ is conjugate to $T$, the dynamics of the two systems are in some way `dynamically equivalent'.
-Specifically, they are in one-to-one correspondence with one another.
-However when $S$ is semi-conjugate to $T$ with $\phi:U\to{V}$ a many-to-one mapping, the dynamics on $V$ provide merely a coarse-grained description of the dynamics on $U$~\cite{de2013elements}.
-When $S$ is semi-conjugate to $T$, it is also common to call $S$ a \emph{factor of $T$}, or conversely that $T$ is an \emph{extension of $S$}.
-In essence, an extension (see, \cite{de2013elements}) is a more extensive system capturing all of the essential dynamics of its factor.
-
-It is a very hard or nearly an impossible task to establish the existence of a conjugacy or a semi-conjugacy $\phi$ between two systems \cite{devaney2018introduction}.
-However, one can verify that a function $\phi$ satisfies the commutativity diagram \ref{eqn_conjugacy}.
-
-\begin{Example}\rm
-  For example $\phi(x)=\sin\big(\frac{\pi}{2}x\big)^2$ is a conjugacy between two systems $Tu=1-|2u-1|$ and $Sv=4v(1-v)$  defined on $U=[0,1]$.
-
-  Indeed, $\phi:[0,1]\to[0,1]$ is a homeomorphism and by employing simple trigonometric identities, it is easily proved that $\phi$ satisfies~\ref{eqn_conjugacy}
-  \[
-    \begin{aligned}[t]
-      (\phi\circ{T})u 
-                &=\sin^{2}\Big(\frac{\pi}{2} - \frac{\pi}{2}|2u-1|\Big) = \sin^{2}(\pi{u})=\Big(2\sin(\frac{\pi}{2})\cos(\frac{\pi}{2})\Big)^{2}=4\sin^{2}(\frac{\pi}{2}u)\Big(1-\sin^{2}\frac{\pi}{2}u\Big) \\
-                                    &=({S}\circ\phi)u.
-    \end{aligned}
-\]
-And it follows that $\phi$ is a conjugacy between the two systems.
-\end{Example}
-
-By establishing that two systems are conjugate (semi-conjugate) to one another, we have also shown that one may choose to work with one system instead of the second and still be guaranteed to obtain information on the latter.
-It is instrumental during the process of forecasting the future evolution of dynamical systems.
-
-\chapter{The Learning Problem}\label{ch3}
-
-In this chapter, we acquaint ourselves with the question of forecasting dynamical systems with unknown underlying structures, state and discuss the shortcomings of the Takens delay embedding theorem and discuss various issues faced while forecasting.
-
-Consider a relatively simple learning problem: 
-Given the sequence $(u_0, u_1, \ldots, u_m)$ for $m\in\mathbb{Z}$, a finite segment of an orbit of the map $T$ where $T$ is defined by the update equation $u_{n+1} = Tu_n$, forecast the values $u_{m+1}, u_{m+1}$ where the map $T$ is unknown, given that $u_m$.
-
-A practical example of this would be the subsequent scenario: given the time-sequential coordinates of an object moving in space, predict the future positions of that object.
-We are very rarely, if at all, presented with a problem where the entire state information is available to us in the form $u_n$ at some timestep $n\in\mathbb{N}$.
-Consider then a more complicated learning problem:
-
-Suppose we only have the observations $\theta(u_0), \theta(u_1), \ldots, \theta(u_m)$ of the true system states $u$ in an unknown dynamical system $(U,T)$ and we wish to predict the values $\theta(u_{m+1}), \theta(u_{m+2})$ and so forth.
-
-First we establish the method of Takens delay embedding. 
-In this method, one considers a discrete-time dynamical system defined on a 'nice' space (a smooth manifold)\ednote{B: Should this be further explained?} that can be obtained as time-$K$ map of a continuous time-dynamical system, concepts formalised below.
-We make observations from such a system, i.e., we observe the evolution of $\theta(w_0)$ (where $w_0$ represents the initial state) by examining  a finite set of values of $u_n :=  \theta(w_n) = \theta(Tw_{n-1})$.
-The sequence $\{u_n\}$ represents a scalar time-series and intuitively $\theta$ represents a probe inserted into a bigger system which is itself only measuring/extracting a small part of the greater system state at time $t=n$.
-Consider, for example, a thermometer erected to measure the ambient temperature in a local village.
-This measurement function, the thermometer, is capturing only a single aspect, the temperature, of a much grander dynamical system entailing the current weather of the surrounding area.
-Even more than that though- it is measuring a miniscule part of the global weather system.
-
-%Given the sequence $\{u_1, u_2, \ldots, u_n}$, we define the variable \newline $y_n:=[\theta_n, \theta_{n-\tau}, \ldots, \theta_{n-(2d-2)\tau}, \theta_{n-(2d-1)\tau}]^{T}\in\mathbb{R}_{d}\times{1}$ where $\tau$ represents the lag and $d$ the dimension of the attractor.
-(We shall return to these terms shortly and their discussion may be put on hold for the time being).
-
-However, the actual state $u$ of a system is seldom, if ever, fully known.
-the most one can do is to insert probes into a system to obtain partial information through the measurements taken.
-Moreover, the process of taking a measurement itself introduces two additional aspects which complicate the problem: \ednote{B: Perhaps this edit makes more sense?
-M: Maybe replace difficulty by saying that we need to consider two aspects.
-}
-\vspace{-8mm}
-\begin{enumerate}[noitemsep, label=\roman*.]
-  \item A series of measurements over a specific time interval is inherently a discretisation procedure of the underlying continuous-time dynamical system.
-(Hence why we restricted our attention to discrete-time systems in the preceding section)
-  \item A probe will never be entirely accurate, and so the act of measurement introduces a certain measure of numerical noise/inaccuracy.
-\end{enumerate}
-
-From here we construct a multidimensional observable using the method of stacking previous observations, i.e., we create delay-coordinate map defined by
-$\Phi_{k,\theta}(w) := (\theta(T^{-k}w)\ldots,\theta(T^{-1}w),\theta(w))$.  
-The concept of an observable is understood in the sense of an observable as originally introduced by~\cite{takens1981detecting, genericObservableAeyels} to refer to a probe or measurement function inserted into the system.
-Takens' theorem, in essence, refers to the fact that for a sufficiently large $k$, we can define a dynamical system on the space $\mathbb{R}^{k+1}$ with states $\Phi_{k,\theta}(w)$, $\Phi_{k,\theta}(Tw)$, $\Phi_{k,\theta}(T^2w)$, etc.
-This dynamical system is topologically conjugate to the unknown underlying system $(W,T)$.
-We recall the Takens delay embedding theorem next.
-
-%We may then ask ourselves the question: Is it in any way possible to retain information about the system state $x(t)$ in this temporal data-series $\theta(x(t))$?
-The answer is easily yes if T is known.
-However, neither this, nor even the exact function $\theta$ is available.
-
-\section{Takens Embedding Theorem}\label{sect_Takens}
-
-We define first the concepts of homeomorphism and embedding:
-\begin{Definition}\rm
-  [\bf {Homeomorphism}]\label{Dfn_homeo}\rm
-  A homeomorphism is a function $f:Z\rightarrow Y$ between two topological spaces $Z$ and $Y$ that is continuous, bijective and has a continuous inverse. 
-\end{Definition}
-
-\begin{Definition}
-  [\bf {Embedding}]\label{Dfn_embed}\rm
-  Consider a homeomorphism $f:Z\rightarrow Y$ for $Y\subset X$. $Z$ is said to be embedded in $X$ by $f$.
-\end{Definition}
-
-Takens Theorem states a result establishing a relationship between the observed and underlying dynamical systems by showing that the concatenation of a sufficiently large number of previous observations into a vector will, under certain conditions, generate a map between the vectors from the respective systems.
-We formulate the theorem from \cite{takens1981detecting}.
-
-\begin{Theorem} 
-	[\bf Takens Embedding Theorem (adopted from \cite{takens1981detecting}] \label{Thm_Takens}
-         Let $W$ be a compact manifold of dimension $m$, and $d\ge m$ so that $2d$ is an integer. It is a 
-            generic property for the pair $(T, \theta)$,  where $T:W \to W$ is
-            a smooth diffeomorphism, and $\theta:W \to \mathbb{R}$ a smooth function, the map $\Phi_{2d,\theta}:W \to \mathbb{R}^{2d+1}$ defined on $W$ by 
-            $\Phi_{2d,\theta}(w) := (\theta(T^{-2d}w)\ldots,\theta(T^{-1}w),\theta(w))$
-            is a diffeomorphic embedding; by `smooth' we mean at least $C^2$.
-Consequently, there exists a map $F_\theta: \Phi_{2d,\theta}(W) \to \Phi_{2d,\theta}(W)$ defined by $$F_\theta: (\theta(T^{-2d}w),\ldots,\theta(T^{-1}w),\theta(w)) \mapsto
-            (\theta(T^{-2d+1}w),\ldots,\theta(w),\theta(Tw))$$
-           so that $(W,T)$ is topologically conjugate to 
-            $(\Phi_{2d,\theta}(W), F_\theta)$.    
-\end{Theorem}
-
-By generic we mean a residual set on a certain topology on appropriate spaces of functions (that we do not describe here) \ednote{B: Where can I find a citation for this?}. 
-By $C^2$, we make reference to a twice-differentiable function  with a continuous second derivative.
-In our scenario (and this remains important throughout) the input space $U$ is considered the attractor.~\label{attractor_U}
-
-Below we reproduce Figure 1 from \cite{Supp}.
-
-%Some explanation is beneficial.
-Takens establishes a delay-coordinate map $\Ftheta$ defined in \ref{eqn_takens} for $T$ the flow (\ref{defn_flow}) defined by the update equation  $Tu_n=u_{n+1}$ and $theta$ our measurement function.
-When confining a dynamical system to the manifold U (in which the underlying system is contained), Takens showed that if certain smoothness conditions are satisfied on $T$ and $\theta$, then the delay-coordinate map $\Ftheta$ embeds $U$ in the reconstruction space $R^{2d+1}$ \textbf{we haven't used the term reconstruction space yet, so this is falling out of the blue}, for almost every choice of measurement function $\theta$, the observable.
-By almost every choice here, we mean that we exclude some observations which do not give any information.
-To see a more precise discussion, see \textbf{xxx} for we do not concern ourselves with a further elaboration on this report.
-
-%Alternatively we may also make the statement: Takens' Embedding Theorem guarantees that almost all dynamical systems can be reconstructed from just one noiseless observation sequence \textbf{this isn't obvious from the theorem's formulation, is it?} i.e.
-for a great number of possible observation functions $\theta$, $\Ftheta$ preserves the topology of U.
-\textbf{'preserving the topology' is not something we've touched on before.
-Perhaps I should add to previous discussion so logic flows seamlessly here}.
-
-\begin{figure}[ht]
-  \includegraphics[scale=0.25]{_takensmap.eps}
-  \centering
-  \captionof{figure}{Schematic of Takens Embedding Theorem's usage and impact of iterates slipping from the attractor when the approximation $\tilde{F}_\theta$ of $F_\theta$ is learnt.}
-  \label{fig:takensmap}
-\end{figure}
-
-If we can learn the map $F_\theta$, from sufficient number of data points of $\{\phi_{2d,\theta}(w_n)\}$, then we also know how to forecast how $\theta(w_n)$.
-
-%Once again we pause and consider the graph to consolidate our understanding of the conjugacy (\textbf{are we being too simplistic?}).
-If we are in state $u$ and then evolve towards state $Tu$ before embedding into $\mathbb{R}^{2m-1}$, then this is equivalent to first embedding into $\mathbb{R}^{2m-1}$ and then evolving through the map $\Ftheta$, i.e.
-$\Phi_{2d,\theta}\circ{F}_{\theta} = T\circ\Phi_{2d,\theta}$.
-The map $F_theta$ is a homeomorphism as well.
-
-%Thus information about U can be retained in the time series' (or observation's) output.
-By preserving the topology on the manifold $U$ in the reconstruction space $X$ (\textbf{not discussed before}), we are also guaranteed the preservation of topological invariants of the manifold, of which dimensionality is one such invariant.
-(We note here that dimensionality will again be considered later in this project).
-
-\section{Practical Limitations to Takens' Theorem}
-Takens' Embedding Theorem is a powerful result, and it provides compelling reason to believe that one could conceivably reconstruct a system conjugate accurately to the underlying system. 
-Nonetheless, it does present some severe practical limitations:
-\vspace{-5mm}
-\begin{enumerate}
-\item Even supposing that we could indeed find $F_\theta$ , our \emph{approximation} of  $F_\theta$ is a map from a larger set $\mathbb{R}^{2d+1}$ containing the embedded attractor.
-There are, however, no theoretical guarantees that $F_\theta$ will retain  $V=\Phi_{2d,\theta}(W)$  as an attractor although $W$ could itself be an attractor.
-\item Takens theorem is stated only for noiseless observations.
-Due to noise $\epsilon_n$ the delay vector $\Phi_{2d,\theta}(w_n) + \epsilon_n$  may lie outside $V$.Furthermore, due to the chaotic nature of the underlying system (i.e.
-the fact that it has SIDC), the evolution of $\Phi_{2d,\theta}(w_n) + \epsilon_n$ under the map $F_\theta$ could move out of $V$ completely.
-This problem can be overcome  by using a driven dynamical system with some properties, and we discuss this in Chapter\ref{ch3}.
-\end{enumerate}
-
-% The Takens embedding theorem does not guarantee global dissipativity.
-% \ednote{B:Links with~\ref{subs_LearnGamma} in discussion of dissipativity of $\Gamma$.
-Sir, I'm not sure how to write this here - I think it a beneficial point to make so as to set the stage for the later discussion in ~\ref{subs_LearnGamma}, but I am unsure there as well.}
-
-Though not a uniquely related to the Takens embedding theorem, we do opt to make mention of another requirement for our work. 
-The methods described here are applicable for data originating from a surjective map. 
-This follows from the fact that when $T$ maps the space $U$ into a proper subset $B$ of$ $U, there exists some point in $U - A$ that does not have a preimage and we conclude that $T$ is not surjective.
-An example of this would be the case when a system displays energy-loss.
-Consider, for instance, consider the dying oscillations for a damped pendulum.
-(An account of greater depth  is provided in~\ref{ch5})
-
-This requirement is not overly restrictive, as a great number of chaotic systems do have a surjective map.
-(Cite) \ednote{B: I remember you saying this in one of our sessions, but don't have a source for it.
-Do you have a specific reference?}
-
-\chapter{Driven Dynamical Systems to Forecast Problems}\label{ch4}
-
-In this chapter, we discuss results on the mapping of temporal data obtained from a discrete-dynamical system onto a different space through the notion of a driven dynamical system. 
-We also consider the conditions a driven system should possess to avoid adding distortion to its state-space representation.
-We then describe the \emph{single-delay dynamics} (SDD) of the system and finally recall conditions on the driven system so that the SDD are conjugate (or at least semi-conjugate) to the underlying system.
-The SDD can then be used to forecast and reconstruct the underlying system.
-
-\section{Nonautonomous and Driven Dynamical Systems}
-
-\begin{Definition}
-  [\bf Nonautonomous Dynamical System (adopted from~\cite{Manju_ESP}))] \label{Dfn_NDS}\rm
-  A nonautonomous dynamical system (NDS) on a space $X$ is simply a dynamical system comprising a family of maps $\{g_n\}_{n \in \mathbb{Z}}$ where $g_n(\ldotp):=g(u_n,\ldotp):X\to{X}$ is continuous.
-Each $g_n$ arises as a consequence of an input $u_n$ from the input space $U$, a topological space.
-\end{Definition}
-
-We immediately recall the figure first defined in \eqref{eqn_conjugacy}  and remind ourselves of our ultimate goal to learn a system that is conjugate to an underlying or unknown dynamical system $(U,T)$ by making use of the measurements obtained from  the unknown system.
-
-%#####################################
-\begin{center}
-\psset{arrows=->, arrowinset=0.25, linewidth=0.6pt, nodesep=3pt, labelsep=2pt, rowsep=0.7cm, colsep = 1.1cm, shortput =tablr}
-\everypsbox{\scriptstyle}
-\begin{psmatrix}
-U & U\\%
-V & V.
-%%%
-%  \ncline{1,1}{1,2}^{T} 
-%  \ncline{1,1}{2,1} <{\phi}
-%   \ncline{2,1}{2,2}^{S}
-%  \ncline{1,2}{2,2} > {\phi}
-\end{psmatrix}
-%#####################################
-\end{center}
-
-To find a function $\phi$, the need now arises to consider a so-called driven dynamical system, a special case of a nonautonomous dynamical system.
-Two spaces are considered: input is taken from an input space $U$ and the state of a space $X$ is updated.
-(Previously, only the state was considered.)
-We do this to mimic the true conditions.
-By accounting for an extrogenous input sequence in $U$ which can also influence the  system-state at a specific time-step, it produces more general models than an autonomous system.
-
-\begin{Definition}
-  [\bf Driven, Compactly Driven Dynamical System] \label{Dfn_DDS} \rm
-A driven dynamical system comprises two topological metric spaces $U$, $X$ and a continuous function  $g:U\times{X}\to{X}$ where $g(u_n, x_n)=x_{n+1}$.
-If the input space $U$ is compact, we refer to the system as compactly driven. 
-\end{Definition}
-
-The update equation $x_{n+1} = g(u_n,x_n)$ for $n \in\mathbb{Z}$, input $u_n$ from $U$ and state $x_n$ belonging to $X$ (a compact space) generate the dynamics on X. 
-Abbreviated, we refer only to the \textit{driven system $g$}, where all other entities are understood implicitly.
-
-Notably, a nonautonomous dynamical system can be generated from $U$; any input $\overline{u}$, a bi-infinite sequence from $U$, gives rise to the sequence of self-maps $\{g(u_n, \cdot)\}_{n\in\mathbb{Z}}$ contained in $X$.
-Physically, one may think of this bi-infinite sequence as referring to a system that has been running for an extended period at the time of the first measurement taken from the system (alternatively, the first time a probe is inserted into the system to take an observation)
-
-\begin{Definition}
-  [\bf Entire Solution] \label{Dfn_Soln} \rm
-  A sequence $\{x_n\}_{n\in\mathbb{Z}}\subset X$ is called an entire solution (or simply a solution) of the driven system  $g$ with input $\overline{u}$ when it satisfies 
-  \[g(u_{n-1}, x_{n-1})=x_n\] for all $n\in\mathbb{Z}$
-\end{Definition}
-
-It is important to emphasise that a sequence $\{x_n\}$ satisfying the update equation above can only be a solution if $x_n\in{U}$ holds for all $n\in\mathbb{Z}$ and just for $n>0$.
-Consider the example below.
-
-\begin{Example} \rm \label{ex_halfux}
-  The only solution  $\{x_n\}_{n\in\mathbb{Z}}$ to the driven system  $g(u,x)=\frac{ux}{2}$, where $X=[0,1]$, $U=[0,1]$,  is the zero solution $x_n\equiv0$.
-  To see this, consider any $x_n=a\in[0,1]$ where $a\neq{0}$.  Let $\overline{u}\in{U}$ be an non-zero constant sequence , say $u_n=0.5$. 
-  The driven system may be rewritten as $x_{n-1}=\frac{2x_n}{u_{n-1}}=4x_n$ and the  iterates of $x_n$ in backward time will increase by a factor of 4 at each timestep. 
-  Thus for some $m\leq{n}$,  $1<x_m$ i.e. $x_m\notin{X}$. So ${\{x_n\}}_{n\in\mathbb{Z}}$ is not a solution and it follows that the only possible solution is the zero solution.
-\end{Example}
-% \ednote{Nicely written! Thank you}
-
-A system may also have multiple solutions, as evidenced in the example below.
-
-\begin{Example}\label{Ex_exp} \rm 
-  Consider the driven system $g(u,x)=x^2$ for $X=[0,1]$, $U=\mathbb{R}$.
-The system has an uncountable number of solutions, as there exists a solution for every $x\in{X}$ which also passes through the point $x$ and $lim_{n\to\infty}x_n=0$, $lim_{n\to{-}\infty}x_n=1$.
- The proof is deferred to immediately after the next paragraph (see~\ref{rem_proofEx}).
-\end{Example}
-
-As the solutions to a driven system are considered an important entity, we next identify a subspace $X_U$ of $X$ that contains all possible solutions.
-To realize such a subspace of a driven system $g$, the concept of a reachable set is defined.
-
-\begin{Definition}
-  [\bf Reachable Set]\label{Dfn_ReachableSet}\rm
-The reachable set of a driven system $g$ is exactly the union of all the elements of all the (entire) solutions, i.e., 
-\[X_U :=\Big \{x \in X:  x = x_k \mbox{ where  $\{x_n\}$  is a solution for some  $\bar{u}$} \Big \}.\]
-The set of all reachable states at a specific time $n$ for input $\overline{u}$ is denoted by $X_n(\overline{u})$
-\end{Definition}
-
-The reachable set itself can defined independently from $g$ possessing the ESP/USP.
-Note that $x\in{X_n(\overline{u})}$ if and only $g$ has a solution $\{x_k\}$ for $x_n=x$ and input $\overline{u}$, a result proved in nonautonomous dynamical systems literature~\cite{manjunath2014dynamics, manjunath2013echo}.
-\ednote{B: Adjusted sentence and cited}
-
-The set $X_n(\overline{u})$ is precisely the set of points $x$ through which some entire solution $\Psi$ passes at time $n$. We can give an alternate formulation of the set $X_n(\overline{u})$. 
-Now suppose, we for some fixed input 
-$\overline{u}$, we define $g_i = g(u_i,\cdot)$ for all $i\in \mathbb{Z}$, then the set of states
-
-\begin{equation} \label{eqn_association}
-X_{n,i}(\overline{u}) := g_{n-1} \circ \cdots g_{i+1} \circ g_i(X).
-\end{equation}
-
-The set $X_{n,i}(\overline{u})$ being the image of a finite composition of
-continuous maps is compact whenever $X$ is compact. Also $X_{n,i}(\overline{u})$ is
-nonempty.  Further, $X_{n,i}(\overline{u}) \supset X_{n,i-1}$. Hence $\bigcap_{i<n}
-X_{n,i}(\overline{u})$ is a nested intersection of closed nonempty subsets, and
-whenever $X$ is compact, the intersection is nonempty. We say $g$ is a topological contraction if $\bigcap_{i<n}X_{n,i}(\overline{u})$ is a singleton subset of $X$ for each $\overline{u}$.
-
-It turns out that $\bigcap_{i<n}X_{n,i}(\overline{u})$ is identically equal to $X_n((\overline{u})$ that we had defined earlier.
-This will lead to a result (for.
-see, \cite{manjunath2013echo}) that  \textit{$g$ being a topological contraction is equivalent to the existence of a exactly one entire solution.}
-
-% \begin{Definition}
-%   [\bf Topological Contraction]\label{Dfn_TopContr}\rm
-%   A function $g:U\times{X}\to{X}$ is a topological contraction if for all $n\in\mathbb{Z}$ and all $\overline{u}\subset{U}$, $X_n(\overline{u})$ is a singleton subset of $X$.  
-% \end{Definition}
-
-\begin{Remark}
-  [\bf Proof of Example~\ref{Ex_exp}] \label{rem_proofEx} \rm
-  For the driven system,  
-  $g(u,x)=x^2$ the evolution does not depend on the input at all.
-Hence $g(u,x)$ can be written as a single map $f(x)=x^2$ which is a homeomorphism on $[0,1]$.
-A left-infinite orbit is defined and it converges to $1$ while the right-infinite orbit converges to $0$.
-Or in other words the iterates of $f^{-1}$ converges to $1$ while the iterates of $f$ converge to $0$
-\end{Remark}
-
-Thus far, it has been demonstrated that a system may have one or more solutions; one may ask if a driven system always has a solution and, if so, whether it satisfies specific properties such as uniqueness.
-Should the driven system be compact, existence follows immediately, as shown in the following result.
-
-\begin{Theorem}\label{Thm_CompactExistence}
- Let $g$ be a driven system.  If $X$ is compact then for each input $\overline{u}$, there exists at least one solution to the driven system $g(\ldotp, x)$
-\end{Theorem}
-\begin{proof}
-  May be found in \cite{kloeden2011nonautonomous, manjunath2014dynamics, manjunath2013echo}
-\end{proof}
-
-% One may easily construct many systems with trivial solution-sets, such as $g(u,x)=x$ which has only the constant solution $x$ and so for $U=[-1,1]$, the system would have no solution if $|x|>1$.
-To refine the scenario, we consider only systems with unique solutions.
-
-\section{Unique Solution Property}
-
-\begin{Definition}
-  [\bf Unique Solution Property] \label{Dfn_usp}\rm
-  A driven system $g$ is said to have the Unique Solution Property (USP) if for each input $\overline{u}$ there exists exactly one solution. 
-  Alternatively we may formulate the USP as follows: $g$ has the Unique Solution Property if there exists a well-defined map $\Psi:{U}\to{X}$, with $\Psi({\overline{u}})$ denoting the unique solution.
-\end{Definition}
-
-One of the first results obtained after defining the USP is that every solution will attract different initial conditions towards the component parts of the solution.
-If $g$ has the USP, then any solution to $g$ is also a uniform attractor in nonautonomous dynamical systems literature~\cite{Manju_Nonlinearity}.
-The discussion on nonautonomous attractors is beyond the scope of this project, and we refer the reader to \cite{Manju_ESP, esann2012ids} for further reading.
-
-% Paraphrased.
-%Additionally, $g$ having the USP is in our context also equivalent to $g$ being a topological contraction \cite{manjunath2021universal}, or alternatively also that $g$ exhibits the Echo State Property (ESP).
-
-In the majority of Reservoir Computing(RC) literature (a machine learning approach) where the input is mapped onto a different but higher dimensional through a system called the reservoir and a readout that measures the state of the reservoir is trained, a notion of forgetting the states of the reservoir asymptotically is often used.
-
-% \ednote{How should I explain/qualify RC better here?
-It's mentioned without really explaining.
-M: Have put it in one sentence; just to let you know people use optical devices to map signal into a ``reservoir".
-B: Thanks!},
-Concepts like the ESP,  fading memory \cite{boyd1985fading} are some of these notions.  
-If $g$ possesses the ESP (equivalent in our context to the USP), the we are guaranteed that the whole of an input's left-infinite history will exactly determine the current system state; i.e.
-there is only ever the possibility of a single reachable state at a given instance of time \cite{jaeger2001echo,Manju_2020}.
-% Defined formally, the ESP is intrinsically linked to an input (see \cite{Manju_ESP}) Unfortunately, definitions and results concerning the ESP do not have much practical benefit since a relationship between the network's response and the temporo-statistical properties of the input are not directly described.
-A more detailed presentation may be found in~\cite{jaeger2001echo}.
-
-% We do however consider it important to mention that the ESP is also often discussed in terms of a stability property that holds for all inputs of a system~\cite{manjunath2020stability} and, furthermore, plays a key role in the design and training of recurrent neural networks(RNN) within the field of RC \cite{Manju_ESP}.
-Beyond forgetting the past states the ESP is also often discussed in terms of a stability property that is nearby inputs create close by responses (solutions), a concept called input-related stability~\cite{manjunath2020stability} and, furthermore, plays a key role in the robustness of recurrent neural networks(RNN), a component of the implementation discussed in~\ref{ch5}.
-
-\section{Conjugacies}
-
-Having already defined the reachable set $X_U$ as the collection of all elements of all entire solutions, we pause briefly in order to fix additional notation.
-Defining $\cev{u}^{n}:=(\ldots,u_{n-2} ,u_{n-1})$ as the left-infinite subsequence of an input up until time $n$, $\overleftarrow{U}$ is then the notation for all these left-infinite sequences in $U$.
-Moreover, $\cev{u}^{n}v:=(\ldots,u_{n-2} ,u_{n-1}, v)$ will symbolise the input up to time $n$ with $v \in U$ being the specific input value at time $n$. 
-The introduction of a new input at time $n$ can be described by the mapping $\sigma_v:   \cev{u}^{n} \mapsto \cev{u}^{n}v$. 
-The right-shift map, $r\overleftarrow{U}\to\overleftarrow{U}$, of an input sequence is defined $r: (\cdots, u_{-2},u_{-1}) \mapsto(\cdots, u_{-3},u_{-2})$
-
-The question now assumes the form: Can we establish a semi-conjugacy for the driven system $g$ as presented below.
-
-\begin{equation}  \label{Scomm_h}
-  %    \[ 
-      \psset{arrows=->, arrowinset=0.25, linewidth=0.6pt, nodesep=3pt, labelsep=2pt, rowsep=0.7cm, colsep = 1.1cm, shortput =tablr}
-   \everypsbox{\scriptstyle}
-   \begin{psmatrix}
-   \overleftarrow{U} & \overleftarrow{U}\\%
-   X_U & X_U.
-   %%%
-  %  \ncline{1,1}{1,2}^{\sigma_v} \ncline{1,1}{2,1} <{h}
-  %  \ncline{1,2}{2,2} > {h}
-  %  \ncline{2,1}{2,2}^{g(v,\cdot)}
-   \end{psmatrix}
-  % \]
-  \end{equation}
-
-We proceed to consider a specific subclass of conjugacies.
-
-  \begin{Definition}
-    [\bf Universal Semi-Conjugacy]\label{Def_UnivSemiConj} \rm
-    Given a driven system $g$, we  call a continuous and surjective map $h : \overleftarrow{U} \to X_U$ a universal semi-conjugacy if  diagram \ref{Scomm_h} commutes for all $v \in U$.
-  \end{Definition}
-
-  If the universal semi-conjugacy $h$ exists (i.e. the diagram in \ref{Scomm_h} commutes) then the solution $\Psi(\bar{u})$ will intuitively have no more ``complexity' than the input $\bar{u}$.
-
-But does such a function $h$ for the above driven system $g$ above exist?
-Whenever $g$ has the USP and $\Psi(u)=\{x_n\}_{n\in\mathbb{Z}}$ it follows that $h$, defined by  $h(\cev{u}_n):=g(u_n,x_{n-1})=x_n$, will satisfy the semi-conjugacy in the graph above \ref{Scomm_h}.
-Regrettably, such a continuous mapping $h$ is not guaranteed to exist when $g$ does not have the USP \cite[Lemma 5]{Manju_Nonlinearity}.
-Note that even when $h$ does exist, we are not guaranteed its injectivity. Considering again example \ref{ex_halfux}, we see that even if $h$ were to exist, it could not be injective as $X_U=\{0\}$.
-
-Re-sketching the commutativity diagram \ref{Scomm_h} above by replacing $X_U$ by its left-infinite sequence space $\overleftarrow{X}_U$, we obtain the diagram below.
-In this case, the function $H:\overleftarrow{U}\to\overleftarrow{X}_U$, a map that is both continuous and surjective, is called a \emph{causal mapping} that is defined next.
-
-\begin{Definition}
-  [\bf Causal Mapping]\label{Def_CausMap}
-  A continuous, surjective map $H:\overleftarrow{U}\to\overleftarrow{X}_U$ such that \[H\circ\tilde{g}_v=\sigma\circ{H}\] holds for all $v \in U$ where $\tilde{g}_v$ maps $(\ldots, u_{-2}, u_{-1})$ to $(\ldots, u_{-2}, u_{-1}, g(v, u_{-1}))$ is called a causal mapping.
-\end{Definition}
-
-\begin{equation} \label{SCausal_H}
-    %    \[ 
-        \psset{arrows=->, arrowinset=0.25, linewidth=0.6pt, nodesep=3pt, labelsep=2pt, rowsep=0.7cm, colsep = 1.1cm, shortput =tablr}
-     \everypsbox{\scriptstyle}
-     \begin{psmatrix}
-     \overleftarrow{U} & \overleftarrow{U}\\%
-     \overleftarrow{X}_U & \overleftarrow{X}_U.
-     %%%
-    %  \ncline{1,1}{1,2}^{\sigma_v} \ncline{1,1}{2,1} <{h}
-    %  \ncline{1,2}{2,2} > {h}
-    %  \ncline{2,1}{2,2}^{\tilde{g}_v}
-     \end{psmatrix}
-    % \]
-  \end{equation}
-
- \begin{Theorem}
-  For a compactly driven system, a causal mapping $H$ exists if and only if $g$ has the USP. 
-\end{Theorem}
-{\bf Proof.}  See~\cite[Th.3]{manjunath2013echo}.
-
-The driven system $g$ can induce an embedding of $\overleftarrow{U}$ in $\overleftarrow{X}$ as follows: 
-If the causal mapping $H:\overleftarrow{U}{\to}{\overleftarrow{X}_U}$ is injective (in addition to being surjective), it becomes the embedding of $\overleftarrow{U}$ in $\overleftarrow{X}$ induced by the driven system $g$.
-The continuity of $H^{-1}$ follows from the fact that $H$ is itself a continuous and surjective mapping of a compact space $\overleftarrow{U}$ in a Hausdorff space.
-and we refer to the map $H$ as a \emph{causal embedding}. 
-\ednote{B: edited the sentence here}
-
-When $g$ has the USP, the diagram below (adopted from~\cite{Manju_Nonlinearity}) illustrates the operation of the mappings $h$ and $H$.
-The mapping $h:\overleftarrow{U}\to{X_U}$ is also considered an observable as mentioned in the introduction of Chapter \ref{ch3}).
-
-\begin{figure}[ht]
-  \includegraphics[scale=0.4]{_actionofh_H.eps}
-  \centering
-  \captionof{figure}{Ladder-like behaviour of $h$ and $H$ to illustrate the causal mapping. (Figure reproduced from~\cite{Manju_Nonlinearity}) }
-  \label{fig:actionh_H}
-\end{figure}
-
-The function $H$ contains a countably infinite number of coordinate (or factor maps) since its codomain is $\overleftarrow{X}_U$.  
-As it is not practically feasible to consider a left-infinite sequence in $X_U$ (in other words an element of  $\overleftarrow{X}_U$) for learning, we wish to restrict the inputs we consider to a subspace of $\overleftarrow{U}$.
-Particularly, we restrict our attention to the  so-called inverse-limit space of a dynamical system and hope to embed it within some finite product of $X$.
-
-%We wish to explore the possibility of establishing a conjugacy rather than a mere semi-conjugacy and therefore are interested in the restriction of inputs to the left-infinite orbits of a dynamical system.
-
-The left-infinite orbit space is significant because a great amount of information about the original system can be gleaned from its topological structure~\cite{ingram2011inverse,Manju_IEEE}. 
-The formal topological structure for the inverse-limit space of a chaotic dynamical system is complicated, and further discussion is relegated to~\cite{kennedy2008inverse_limit, ingram2011inverse}.
-Broadly, we may conceive of the inverse-limit system of a dynamical system as a subspace of an infinite-dimensional space where each point in the inverse-limit space corresponds to a backward orbit of the map $T$.
-Following~\cite{manjunath2021universal, ingram2011inverse} we provide a definition:
-
-\begin{Definition}
-  [\bf Inverse-Limit Space, Inverse-Limit System]\label{Dfn_InverseL}\rm
- The subset $\widehat{U}_T\subset \overleftarrow{U}$ defined by \[\widehat{U}_T: = \{ (\ldots,u_{-2},u_{-1}): Tu_{i} = u_{i+1}\} \] is called the inverse-limit space of $(U,T)$.
- The self-map $\widehat{T}$  induced by $T$ on $\widehat{U}_T$  is defined by \[\widehat{T}: (\ldots,u_{-2},u_{-1}) \mapsto  (\ldots,u_{-2},u_{-1},T(u_{-1}))\] and the resulting dynamical system $(\widehat{U_T}, \widehat{T})$ is called the inverse-limit system of $(U,T)$.
-\end{Definition}
-
-The inverse-limit space is well-defined since $T : U \to{U}$ is surjective by assumption.
-We now get to the concept of embedding the inverse limit-space of the dynamical system in $X \times X$. 
-\begin{Definition} \rm
-	We say a driven system $g$ \emph{causally embeds} a dynamical system $(U,T)$ if it satisfies the two properties: (i) a universal semi-conjugacy exists and  (ii)  $H_2(\cev{u}) := (h(r\cev{u}),h(\cev{u}))$ embeds the inverse-limit space $\widehat{U}_T$ in $X \times X$.
-\end{Definition}
-
-  It is imperative to take note of a subtlety here: we refer to the \emph{map} $H$ as a (causal) embedding, whereas the driven \emph{system} $g$ (causally) embeds another \emph{system} $(U,T)$.
-\ednote{B: inserted line}
-
-\section{Choosing the driven system $g$}
-
-So far it is not clear if the USP alone ensures the ability to causally embedding a dynamical system and we mention that a driving function $g$ cannot just be chosen in a frivolous manner.
-This could possibly complicate our work.
-%\ednote{B: Grammarly}
-
-%The driving function $g$ cannot be chosen in a frivolous manner as it could possibly complicate our work.
-Specifically, one must be careful to avoid a choice of $g$ which would  add complexity to the obtained solution.
-
-%When a causal embedding $H$ exists for the driven system $g$, one can map an arbitrary input ${u}$ onto the solution space $X$ without additional distortion or information-loss \textbf{(Cite.)}.
-
-%When an embedding is established, the question of possible additional complexity in the solution is removed by guaranteeing that, since the systems are conjugate (semi-conjugate, \textbf{(refer)}), $g$ does not add any (some) complexity to the system.
-
-%To achieve a causal embedding embedding not just a causal mapping, it is undesirable to choose a function $g$ that quenches the temporal structure in $u$ by contracting to such a degree that the ability to recover information from the original system is lost completely.
-In the above example(\ref{ex_halfux}), the input's temporal variation could not be related to the reachable set as $X_U$ only consisted of a single element; little, if not no, information is encoded.
-To obtain a suitably complex function $g$, it is thus desired that the reachable set of a driven system  be large enough to relate to the input. 
-This is true even for embedding  the inverse-limit space of a dynamical system. \ednote{B: Expand or refer?}
-
- To this end, we recall the notion of State-Input (SI) Invertibility.
-
-\begin{Definition}
-  [\bf SI-Invertibility]\label{Dfn_SIinv}\rm
-  A driven system $g$ is said to be SI-Invertible if $g(*,x): U \to X$ is invertible for all $x\in X$.
-Alternatively it may be said that if, given $x_n$ and $x_{n-1}$, $u_{n-1}$ can be uniquely determined from $x_n=g(u_n,x_{n-1})$, then $g$ is said to be SI-invertible.
-\end{Definition}
-
-SI-invertibility promises that 'enough' information is retained when $g$ is chosen without introducing one's choice of $g$ will still ensure 'enough' information is retained without introducing unwanted complexity.
-'Enough' here refers to the fact that we may always recover the previous input value $u_n$ (if we know successive states $x_{n-1}, x_n$).
-
-Subsequently we define the relation $Y_T$ induced by $(U,T)$ on $X_U\times{X_U}$ for a driven system $g$ possessing SI-invertibility.  
-To describe the  SDD formally, we consider a dynamical system $T: U \to U$ and define a relation on the reachable set $X_U$, i.e. a subset on $X_U \times X_U$  defined by 
-$$Y_T:=\{(x_{n-1},x_n): \{x_k\}_{k\in \mathbb{Z}} \mbox{ is a solution for some orbit of } T \mbox{ and } n \in \mathbb{Z}\}.$$
-
-The following theorem establishes the existence of a well-defined map $G_T$ describing the single-delay dynamics(SDD) of the system above.
-
-\begin{Theorem}\label{Thm_GT_Exists}
-  For an SI-invertible driven system $g$ and a dynamical system $(U,T)$, the map $G_T: Y_T \to Y_T$ defined by the relation $(x_{n-1},x_n) \mapsto (x_n,x_{n+1})$ is well-defined. 
-  (This results holds even in the absence of $g$ possessing over the USP) 
-  % Moreover, the mapping $(x_{n-1},x_n) \mapsto u_{n}$ is well-defined when $x_{n-1}$ and $x_n$ are successive points on a solution obtained for an input  $\{u_n\}$ that is an orbit of $T$.
-  \end{Theorem}
-  {\bf Proof.} See~\cite[Th.3]{Supp}
-
-%'Without additional complexity' is guaranteed by the invertibility of the function $g$ that guarantees a one-to-one mapping between $u_n, x_{n-1}$ and $x_n$.
-In obtaining an SI-invertible function, therefore, we are guaranteed a system which will not lose so much information about its previous states in the forward-flow of time that one may not make any credible claims as to the original system's behaviour and properties.
-Simultaneously, it does guarantee us that the important information encoded in previous inputs is indeed preserved.
-
-At this stage it is worth taking note of a specific driven system in the form of a discrete state-space model which has acquired some adherence in applications \cite{Manju_IEEE}- especially those pertaining to Echo State Networks and the ESP.
-The function
-\begin{equation}  \label{eqn_driving}
-  g(u,x) = (1-a)x + a\overline{\tanh}(Au + \alpha Bx)
-\end{equation} 
-is SI-invertible and, if $\alpha B$ has a spectral norm $<1$, also possesses the USP \cite[Th.2]{manjunath2013echo }. 
-It is easy to show that $g$ is SI-invertible by recovering $u_n$ in 
-\begin{equation} \label{eqn_SI_RNN}
-  u_{n-1} := A^{-1}\bigg(\overline{\tanh}^{^{-1}}\frac{1}{a}\Big(x_{n+1}-(1-a)x_n\Big) \bigg) - \alpha B x_n
-  \end{equation}
-  when $x_{n-1}$ and $x_n$ are known.
-
-This specific driving function $g$ is used in our implementation and is discussed more completely in chapter~\ref{ch5}
-
-Despite the ease that with which one manipulates a left-infinite history in the realm of theory, it is impossible to obtain or use such a sequence in any real-life application.  
-Fortunately, one does not need the entire left-infinite history of an input in practice thanks to the Uniform Attraction Property(UAP). 
-We use an alternate version of the definition of UAP -- since the notion of nonautonomous attractors in this project would take some time to establish and detracts from the principal thrust of this project, we refer to \cite{Manju_Nonlinearity}.
-
-\begin{Definition}
-  [\bf Uniform Attraction Property]\label{Dfn_UAP}\rm
-  A driven system $g$ has the Uniform Attraction Property (UAP) if we initialize the driven system
-with an arbitrary initial value $y_m \in X$, and the sequence $y_{m+1}, y_{m+2}, y_{m+3},...$ satisfying $y_{k+1}= g(u_k,y_k)$ for $k \geq m$ then approximates an actual solution $\{x_n\}$ uniformly in the sense that given $\epsilon>0$ (independent of $y_m$) there is an integer $n$ so that $d(x_{n+i}, y_{n+i})<\epsilon$ for all $i\ge 0$, where $\{x_m\}$ is a solution.
-\end{Definition}
-
-The UAP guarantees that all trajectories will converge to the same trajectory as time moves forward.
-An illustration of this is shown in Fig.~\ref{fig:memloss_conttime} where the trajectory in red locks on to the supposed actual solution when the system has the USP.
-
-\begin{figure}[ht]
-  \includegraphics[scale=0.35]{_Nonlinearity_Fig_response.eps}
-  \centering
-\captionof{figure}{A coordinate of a simulated solution $(x_0,x_1,\ldots,x_{5000})$ of a RNN  plotted in red (with parameter $\alpha=0.99$ (where $g$ has the USP) and blue ($a=1$ and $\alpha=1.05$ where $g$ does not have the USP) while the matrices $A$ and $B$ are randomly generated, and $B$ has unit spectral radius.
-Reproduced from \cite{Manju_Nonlinearity}. }
-\label{fig:memloss_conttime} 
-\end{figure} 
-\ednote{B: How can I explain the black line? I don't think this makes sense to the reader immediately.}
-
-One may now appreciate an even more astounding result: $g$ having the USP is equivalent to the UAP as is proved in~\cite{Manju_Nonlinearity}.
-
-%We restate the above theorem.
-
-%\begin{Theorem}\label{thm_usp-contr-uap}  The following statements are equivalent:  \vspace{-8mm}  \begin{enumerate}[noitemsep, label=\roman*.]    \item $g$ has the USP.
-\item $g$ is a topological contraction.
-\item $g$ has the UAP.
-\end{enumerate\end{Theorem}{\bf Proof.}  May be found in \cite[Th.1]{Manju_Nonlinearity}
-
-% One need therefore not establish any additional results to ensure that the sequence uniformly approaches the unique solution $\Psi$.
-This vastly simplifies the effort necessary to set up a problem in order to guarantee that the underlying system will be accurately represented by the conjugate system.
-
-%This also solves the problem of perturbation/noise introduced by the observable or measurement function.
-Recall from before that an observable is inherently a map that discretises the underlying continuous-time system.
-Moreover, measurement and device error introduce mistakes.
-Since we are working with a chaotic system displaying sensitive dependence on initial conditions, these small errors could potentially send the trajectory into a completely different attractor.
-%In the presence of the UAP, however, small measurement errors introduced into the system do not pose the same danger as before since the trajectories will eventually tend towards the actual solution.
-% \ednote{(B: Is this Complete enough?)}
-% \textbf{From the perspective of perturbation of an autonomous dynamical system, if the resulting nonautonomous system has the USP, then the temporal structure of noise relates to the dynamics on a nonautonomous uniform attractor through a topological semi-conjugacy or a conjugacy.}
-
-\section{The next step in Dynamics}
-
-% Recall that $g$ is only being given inputs from the orbits of $T$.
-
-%Incredibly, this permits one to initialise a driven system $g$ with an altogether arbitrary initial value $y_m\in{X}$ where $m\in\mathbb{Z}$ and the UAP then guarantees that the sequence $\{y_{m+1}, y_{m+2},\ldots\}$ which satisfies the relation $y_n=g(u_n, y_{n-1})$  for  $k\geq{m}$  will approach the elements $\{x_n\}$ of the actual solution.
-This vastly simplifies the effort necessary to set up a problem in order to guarantee that the underlying system will be accurately represented by the conjugate system.
-
-We are interested in determining whether $\widehat{T}$ and $G_T$ are related.
-For this, we introduce  the function $H_2{\overline{}} := (h(r\overline{u}, h(\overline{u})))$ mapping the entirety of a left-infinite sequence to some element in $X\times{X}$.
-Indeed, it may be shown that $g$ having SI-invertibility and the USP immediately guarantees the existence of at least a semi-conjugacy between the systems $(Y_T, G_T)$ and $(\widehat{U}_T, \widehat{T})$.
-This is formalised in the Causal Embedding Theorem, which we state next:
-
-\begin{Theorem}
-  [\bf Causal Embedding Theorem (adopted from \cite{Supp})]
-  % {\bf (Causal Embedding Theorem.)}
- \label{Thm_CET}
-	Let $g$ be a driven system with SI-invertibility and the USP.
-Let $h$ denote the universal semi-conjugacy and $H_2(\overleftarrow{u}) := (h(r\overleftarrow{u}),h(\overleftarrow{u}))$, where $r$ is the right-shift map.
- Let $(\widehat{U}_T, \widehat{T})$  be the inverse-limit system of a dynamical system $(U,T)$. 
- Then the restriction of $H_2$ to $\widehat{U}_T$ is a topological semi-conjugacy between the inverse-limit system $(\widehat{U}_T, \widehat{T})$ 
-and the induced dynamical system  $(Y_T,G_T)$, i.e., the following diagram commutes
-\begin{equation} \label{comm_H_CET}
-\psset{arrows=->, arrowinset=0.25, linewidth=0.6pt, nodesep=3pt, labelsep=2pt, rowsep=0.7cm, colsep = 1.1cm, shortput =tablr}
- \everypsbox{\scriptstyle}
- \begin{psmatrix}
- \widehat{U}_T & \widehat{U}_T\\%
- Y_T &  Y_T.
- %%%
-%  \ncline{1,1}{1,2}^{\widehat{T}} \ncline{1,1}{2,1} <{H_2}
-%  \ncline{1,2}{2,2} > {H_2}
-%  \ncline{2,1}{2,2}^{G_T}
- \end{psmatrix}
- \end{equation}
-or in other words, $(Y_T, G_T)$ is a factor of  $(\widehat{U}_T, \widehat{T})$. Further, if $T:U \to U$ is a homeomorphism, 
-then $H_2$ embeds $\widehat{U}_T$ in $X_U \times X_U$, and hence $(Y_T, G_T)$ is conjugate to $(\widehat{U}_T, \widehat{T})$.
-\end{Theorem}
-{\bf Proof.}  May be found as the proof of~\cite[Th.4]{Supp}
-
-Recall that $H_2$ maps an entire left-infinite solution sequence from $\Psi$ to an element in $X\times{X}$. 
-We are drawing nearer and nearer to our principal target and our results carry more and more weight: If we are able to learn $G_T$, we will also have learnt $\widehat{T}$; and since $\widehat{T}$ is an extension of $T$, we will have obtained $T$ as well.
-%\ednote{B: Grammarly check}
-
-%\begin{Theorem} Graph \ref{Scomm_h} is exactly the inverse-limit system $(\hat{U}, \hat{T})$.    \end{Theorem}
-
-%We now have the following (which may be compared with \ref{SCausal_H} above):
-
-%\begin{equation} \label{fig:inverse_limsystem}
-  %  \[       \psset{arrows=->, arrowinset=0.25, linewidth=0.6pt, nodesep=3pt, labelsep=2pt, rowsep=0.7cm, colsep = 1.1cm, shortput =tablr}      \everypsbox{\scriptstyle}
-%      \begin{psmatrix}      \widehat{U}_T  && \widehat{U}_T \\     Y_T && Y_T \\
-      %%%
-     %  \ncline{1,1}{1,2}^{\widehat{T}} \ncline{1,1}{2,1} <{h}
-     %  \ncline{1,2}{2,2} > {H_2}
-     %  \ncline{2,1}{2,2}^{G_T}
-%      \end{psmatrix}
-    %  \]
-%  \end{equation}
-
-%\begin{Theorem}
- %   $(Y_T, G_T)$ is semi-conjugate to $(\widehat{U}, \widehat{T})$.
-%\end{Theorem}
-% {\bf Proof.} See \cite[Th.3, Th.4]{manunath2021universal}
-
-\section*{Summarising the discussing thus far:}
-
-It is easy to lose the birds-eye view, so we take a moment to review our progress up until this point.
-
-\vspace{-8mm}
-\begin{enumerate}
-\item We are interested in a some dynamical system $(U,T)$ with unknown dynamics.
-\item To determine properties about this system $(U,T)$ and predict its future evolution, we determine the dynamics of the inverse-limit system $(\widehat{U}, \widehat{T})$.
-% Given certain assumptions, we can guarantee that $(\widehat{U}, \widehat{T})$ is at least semi-conjugate to $(U,T)$.
-\item If the driven system $g$ is SI-invertible (and $\{u_n\}\in{U}$  is an orbit of $T$), the map $G_T$ exists that describes the single delay dynamics. 
-\item If, furthermore, $g$ has the USP, $(Y_T, G_T)$ is semi-conjugate to $(\widehat{U}, \widehat{T})$.
-\item If we can assume that $T$ is a homeomorphism, $(Y_T, G_T)$ is topologically conjugate to $(\widehat{U}, \widehat{T})$, an extension space of $(U,T)$
-\end{enumerate}
-
-One can therefore in practice learn the SDD of the driven system states via $G_T$ with enough data thanks to the USP/UAP. This enables us to do at least two things: 
-\vspace{-8mm}
-\begin{itemize}
-\item Forecast  $x_{n+1},x_{n+2}, \ldots$ via iterates of $G_T$ (if $G_T$ can be learnt).
-\item Forecast future values of $u_n$ since $x_n$ and $x_{n+1}$ determine $u_n$ since $g$ is SI-invertible. 
-\end{itemize}
-
-Finally, we note that although $G_T$ exists with SI-invertibility, we need the USP as well(see \ref{fig_pictorialSummary} below for a visual illustration).
-If not, the driven states would have to be running for all time since they would not have forgotten their past states.
-
-\begin{figure}[ht]
-  \includegraphics[scale=0.3]{_summarypictorial.eps}
-  \centering
-  \captionof{figure}{Pictorial summary of results obtained up to now and how they overlap. Reproduced from \textbf{(Cite SIG-25May)} }
-  \label{fig:fig_pictorialSummary}
-\end{figure}
-\ednote{Ask M: Cite SIG-25May.}
-
-\section{A discussion of $G_T$ }
-
-In the above sections, we established the map $G_T$ describing the SDD of a driven system.
-By establishing the existence of this map, we’ve essentially embedded the attractor $U$ (recall from \ref{attractor_U}) into the higher dimensional space $X\times{X}$.
-In layman’s terms, this ensures that there is more “dimensional room” for the underlying system’s underlying dynamics to “move”.
-Since the dynamics aren’t as “squashed”, we might therefore hope that the dynamics of $G_T$ are in some sense simpler than that of $T$.
-(Taken note of the fact that $G_T$ is a homeomorphism even when $T$ is just continuous)
-
-In \cite{manjunath2021universal} it is illustrated in an empirical fashion in that the map $G_T$ describes dynamics which are less functionally complex than that of $T$ or of the map $\Phi_{2d,\theta}$.
-This is done by implementing a Recurrent Neural Network (RNN), which is discussed in~\ref{ch5}.
-\ednote{B: Should I add in that we'll be using Pearson coefficient?}
-
-We opt to learn $G_T$ in an indirect manner by defining a new map $\Gamma:(x_{n-1},x_n)\mapsto{u_n}$. It will follow immediately from $G_T$’s existence that $\Gamma$ also exists.
-
-The reasons for taking this roundabout approach remain to be discussed in \ref{subs_LearnGamma}, but a pat answer may immediately be given: When $\Gamma$ has been learnt, the system can be driven autonomously and then $G_T$ is known anyway.
-We formalise this with a theorem.
-
-\begin{Theorem}
-  When $x_n$ and $x_{n-1}$ are successive points on a solution obtained for input-orbit of $T$ $\{u_n\}$, then the map $\Gamma: X\times{X}\to{U}$ defined by $(x_{n-1},x_n)\mapsto{u_n}$ exists whenever $G_T$ exists
-\end{Theorem}
-{\bf Proof.}  See \cite[Th. 3c]{manjunath2021universal}.
-
-Projection mappings $\pi_i$ are defined in the traditional meaning where a $k$-dimensional vector is projected to it's $i^{th}$ component such that $$\pi_i:(a_1,a_2, \ldotp, a_{k-1}, a_k)\to{a_i}$$
-
-The graph in equation \ref{fig:comm_H_CET} is then extended as below:
-
-\begin{equation}  \label{Scomm_imp}
-  %    \[ 
-      \psset{arrows=->, arrowinset=0.25, linewidth=0.6pt, nodesep=3pt, labelsep=2pt, rowsep=1.3cm, colsep = 1.1cm, shortput =tablr}
-   \everypsbox{\scriptstyle}
-   \begin{psmatrix}
-   \widehat{U}_T & \widehat{U}_T\\%
-   Y_T & Y_T\\
-   & \textcolor{red}{U \times X}
-   %%%
-  %  \ncline{1,1}{1,2}^{\widehat{T}} \ncline{1,1}{2,1} <{H_2}
-  %  \ncline{1,2}{2,2} > {H_2}
-  %  \ncline{2,1}{2,2}^{G_T}
-  %  \psset{linecolor=red}
-  %  \textcolor{red}{\ncline{2,1}{3,2}<{(\Gamma,\pi_2)}}
-  %  \ncline{3,2}{2,2}>{\textcolor{red}{(\pi_2,g)}}.
-   \end{psmatrix}
-  % \]
-  \end{equation}
-
-The problem finally simplifies to the issue of learning the map $\Gamma$ and combining this with the projection mapping $\pi_2$ and the function $g$, which will be known.
-A final set of equations is obtained - equations that have been entirely constructed from data.
-\begin{eqnarray}\label{eqns_from_data}
-	u_{k+1} &=& \pi_1 \circ (\Gamma, \pi_2) \circ (\pi_2,g) (u_k,x_k) \label{Seqn_u}\\
-	x_{k+1} &=& \pi_2 \circ (\Gamma, \pi_2) \circ (\pi_2,g) (u_k,x_k). \label{Seqn_x}
-\end{eqnarray}
-
-\section{Advantages of learning $\Gamma$} \label{subs_LearnGamma}
-
-One may immediately ask why we opt to take such a roundabout route; why not just learn the map $G_T$ from the get-go?
-On the surface, this seems to be an arbitrary decision path with no real reasoning, so we take a pause again and discuss the motivation for learning Gamma.
-There are a number of distinct advantages.
-
-In the first place, learning $\Gamma$ saves computational resources.
-This is due to the fact that the input $u_n$ lies in a lower-dimensional subspace in comparison to the high-dimensional space $X\times{X}$.
-
-%In practice, if the input is of a  lower dimension, one may easily embed it into the space $X\times{X}$ by padding the vector $u_n$ with zeroes.
-
-Secondly, the function $\Gamma$ is known to be stable.
-Learning $\Gamma$ makes use of equations \ref{eqns_from_data} which in turn employs the driven system $g$ possessing the USP, and offers distinct advantages with regards to stability in the presence of perturbation.
-It is desirable that inputs that differ only slightly would have only ‘slight’ effects on the output of the system.
-Moreover, one would hope to prevent numerical errors originating from input and measurement noise (see~\cite[Th.
-5]{manjunath2021universal}).
-As the data fed into the system is a function of both the input given and the parameters of the system itself (i.e.
-the value a, $\alpha$ in $\tanh$), we exploit both input-related and parameter-related stability.
-Both these notions are defined by way of the continuity of an encoding map of an input and design parameter respectively in \cite{manjunath2020stability}.
-
-Informally, we may consider input-related stability as concerned with the question of whether or not small variations in input would result in small responses.   
-If two inputs $\bar{u}$ and $\bar{u}_{noisy}$ are close in the product topology, then their tails could differ greatly due to the metric that generates the product topology has insensitivity to the differences in sequence's tails.
- However if the system has the USP then the function $h$ is continuous and so $h(\bar{u}), h(\bar{u}_{noisy})$ remain close-by for small levels of noise maintaining a measure of robustness to input and measurement noise.\cite{manjunath2021universal}.
-In a similar fashion, parameter-related stability is obtained thanks to a result \cite[Lemma 3.2]{manjunath2020stability} relating this form of stability with the ESP (which is here equivalent to the USP).
-
-% Secondly,  due to $g$ possessing the USP, both input- and parameter-related stability~\cite{manjunath2020stability} is obtained, which in turn prevents numerical errors that caused by input and measurement noise (see~\cite[Th.
-5]{manjunath2021universal}).
-
-Thirdly, learning a chaotic map on a set(here $Y_T$) with even a small measure of noise  means that the learnt system accepts arguments out of the set (Problems similar to Takens, cf.~\ref{sect_takenslimits}).
-We thus have a map $G_T^+$ acting on $Y_T^+$ and since $Y_T$ is not guaranteed to be an attractor of $G_T$ this could lead to errors.
-If one learns $G_T$ directly, we are in danger of replicating the problems arising for Takens.  To see this, consider Fig.~\ref{fig:YtGtFailure}.
-
-\begin{figure}[ht]
-  \includegraphics[scale=0.25]{_YTerrors.eps}
-  \centering
-  \captionof{figure}{Diagram of potential errors developing when $G_T$ is learnt.}
-  \label{fig:YtGtFailure}
-\end{figure}
-
-To circumvent this: if a driven system with a saturation function\ednote{B: What is a saturation function?} is used, we can ensure the existence of an attractor $A$ of the map which implements $G_T$, and containing $Y_T^+$ indirectly (via $g$).
-\ednote{B: A little vague.
-Are we saying that an attractor of a map implementing $G_T$ is also the one that contains $Y_T^+$?}
-By defining the dynamics to $A$, this prevents large errors from occuring.  
-More details on the global dissipativity of $g$, as it is termed there, may be found in~\cite{Supp}.
-
-\begin{figure}[ht]
-  \includegraphics[scale=0.25]{_autovsdriven.eps}
-  \centering
-  \captionof{figure}{Illustrating empirically the importance of global dissipativity.
-Plotted below are the three principal components of randomly initiated points(blue) of a learnt version of $G_T$ versus the trajectory $(x_{n-1}, x_n)$(red) of driven states obtained from the Fractal Dream Attractor's data~\ref{Thomas_Attractor}).
-}
-  \label{fig:learningFailure}
-\end{figure}
-
-In effect, global dissipativity prevents large numerical errors due to the noise in input data from arising (see Fig.~\ref{fig:learningFailure}). 
-If a system is not globally dissipative, neglibly small errors could lead to major inaccuracy which thus induce predictions to utterly fail.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\chapter{Implementation} \label{ch5}
-
-In this chapter we turn to the implementation of the function $G_T$ by discussing a method to learn the function $\Gamma$.
-We apply our methodology to some chaotic attractors.
-Most notably, we consider the reconstruction of the chaotic attractor of the double pendulum - an  attractor that has not been successfully reconstructed because of its intricate structure.
-In fact the previous data-driven approaches also have not been successful in building a model from the double pendulum.
-
-%We then indicate where this implementation differs from previous work done by (\cite{manjunath2021universal}) and empirically show that the SDD of $G_T$ are less functionally complex than $\Phi_{k, \theta}$ or $T$ by considering the Pearson correlation coefficient.
-Finally, we present and discuss some numerical results for a focus system of this project: the double pendulum.
-We also consider other chaotic attractors.
-
-\section{Implementing $G_T$}
-The map $G_T$ is implemented by applying the functions $\Gamma$, $g$ and the projection function $\pi_2$ in the relevant order as described by the equations~\ref{eqns_from_data}.
-To reach this point, one must first learn $\Gamma$.
-After first normalising the data to ensure that it has zero mean value and lies within the range $[-1,1]$ and required zero-padding(see paragraph below for further discussion), we continue by driving the system with external input through the function $g$.
-
-From \eqref{eqn_driving} we know that $g(u,x) = (1-a) x + a \tanh(Au + \alpha B x)$. Recall that $g$ is SI-invertible. We refer to $A$ as the input matrix and $B$ the reservoir matrix.
-
-We make use of a Recurrent Neural Network (RNN) in our implementation to project the data on to a higher dimensional space, but we do not train the RNN in the conventional manner.
-\ednote{B: Should we say what the conventional manner is?} A few words about the RNNs: RNNs are widely used in the literature for temporal processing of data as a type of deep learning technique.
-They take the form of equations considered to be discrete-time version of continuous dynamical systems modelling a neural network of neurons \cite{jaeger2001echo}.
-
-We implement a feedforward network, (a network with no memory, unlike the RNN)\ednote{B: Expanded sentence s.t.
-FFN is in more context} required to learn $\Gamma$ in the \emph{Python} programming language and making use of the \emph{Scipy} library, alongside the \emph{Keras} package in the \emph{Tensorflow} library \ednote{B: Is this something you want me to write in?
-M: We do not implement the RNN, but the FN that learns Gamma through the Tensorflow.
-We called the RNN + FN set up a RCN in Adriaan's paper.
-FYI: FN's work only if there is a definite relationship between the $u_n$ and $u_{n+1}$ which is rare in temporal data.
-RNNs with ESP work when there is a definite relationship between $(\ldots,u_{n-1},u_n)$ and $u_{n+1}$ and is suited for temporal processing.}.
-
-If the RNN we will be using in the next step has $N$ neurons, then $A$ and $B$ are each $N\times{N}$ matrices.
-Inputs $u_n$ of dimension $K < N$ are padded by zeroes (as remarked earlier) to embed the input in a space of dimension $N$.
-As we do not need the entire left-infinite history of input, an arbitrary initialisation of the system with initial values followed by the driving of the system with the function $g$ means that the generated values will approach the actual solution elements in a uniform manner thanks to the UAP that we had discussed in the preceding chapter.
-
-Thirdly we prepare the network.
-Utilising a network $N$ neurons (the same value $N$ as in the dimensions of the matrices $A,B$).
-Learning of $\Gamma$ is done using the Adam Optimiser where the means square error (MSE) loss function is optimised.
-The system is trained several times (in our case 4) with incrementally smaller learning rates on each run (0.001 and thereafter divided by 10 at each run).
-Additional parameters such as the number of network-layers, neurons per layer and training-length are presented case-by-case below.
-
-We then reach the prediction step. Making use of $g$'s SI-invertibility and $g$ as defined in \eqref{eqn_driving}, we have the following expression for $u_n$ replicated from~\eqref{eqn_SI_RNN}.
-\begin{equation*}
-  u_n := A^{-1}\bigg(\overline{\tanh}^{^{-1}}\frac{1}{a}\Big(x_{n+1}-(1-a)x_n\Big) \bigg) - \alpha B x_n
-\end{equation*}
-
-The two model equations for learning from data~\eqref{eqns_from_data} allows us to determine what the state $x_{n+1}$ at time $n+1$ will be. 
-We specifically remind ourselves of \[x_{k+1}=\pi_2 \circ (\Gamma, \pi_2) \circ (\pi_2,g) (u_k,x_k)\]
-
-\section{Delay Coordinates}
-
-We also consider a more general problem of learning an attractor of dynamical system when only observations of an orbit are given.
-Explicitly, if $(W,T)$ is a dynamical system with dynamics generated by $w_{n+1}=Tw_n$, and if $\theta:W \to \mathbb{R}$ is an observable, then the task is to learn a dynamical system that is topologically conjugate to $(W,T)$ and predict $\theta(w_{m+1}),\theta(w_{m+2}),...$ using the data $\theta(w_{0}),\theta(w_{1}),\ldots,\theta(w_{m})$.
-So how do we do it?
-Suppose the input generated from the delay-coordinate map $\Phi_{\theta,2d}(\theta(w_{n})) := (\theta(w_{n-2d}),\ldots,\theta(w_{n-1}),\theta(w_{n}))$ is such that Takens delay embedding theorem (see Theorem~\ref{Thm_Takens}) holds, in which case, there exists a homeomorphism  $F_\theta: \Phi_{\theta,2d}(\theta(w_{n})) \mapsto \Phi_{\theta,2d}(\theta(w_{n+1}))$.
-Then if we feed  the input values $u_n := \Phi_{\theta,2d}(\theta(w_{n}))$ to a driven system $g$ that is SI-invertible and has USP, the induced system $(Y_F,G_F)$ would be topologically conjugate to the inverse-limit space of
-$(\Phi_{\theta,2d}(W), F_\theta)$ as in Theorem~\ref{Thm_CET}, and hence one could forecast $u_n,u_{n+1},\ldots$.
-In other words, one could forecast  the values $\theta(w_n), \theta(w_{n+1}),\ldots$ by feeding these observations as delay coordinates.
-The advantage of feeding delay-coordinates to a driven system is that the embedding is stable in the sense of global dissipativity that we have described in the previous chapter, i.e., $Y_F$ is made to contained in an attractor of a mapping induced on $Y_F$ that is induced by $g$.
-
-We note that the twin equations together~\ref{eqns_from_data} generate a model from data.
-For a chaotic system, a small amount of noise due either to the input-noise or computational error would cause the solutions from two accurate models with the same initial conditions to  diverge as they are not identical.
-\ednote{B: edited this sentence} However, statistically, they would be the same since due to ergodic theory all typical orbits have the same visiting frequency to any region of the phase space of the dynamical system.
-\ednote{B: I'll need to recap this sentence with you before presenting as it's new to me.} To illustrate that we plot the densities of the trajectories of the actual and predicted solutions.
-
-Our implementation of the reservoir(A) and input(B) matrices in the RNN makes use of sparse matrices.
-We discuss three specifics relating to this concept.
-Firstly, the matrices are assigned a variable density in terms of the amount of non-zero entries and are scaled to have values within the range $[-1,1]$.
-For this, we make use of \textit{Python's SciPy} library and specifically the \textit{Sparse, LinAlg} modules.
-In our experiments, we generally opted for a density ratio of 10\%-25\%.
-Secondly, we opted to enforce the requirement that both matrices be well-conditioned, i.e.
-that they have a condition number of \~1.
-The condition number of a matrix $M$ (denoted $cond(M)$) is the product of the matrix norms of the matrix $M$ and its inverse $M^{-1}$ and is a measure of the sensitivity of output of a matrix multiplication when an input is changed.
-\textbf{(Cite.)}\ednote{Ask M: Cite specific resource?} By ensuring a condition number close to 1, we also ensure that our driven function $g$, depending on the matrices $A$ and $B$, will not be sensitive to adjustments in the input.
-In so doing we then also guarantee another form stability.
-\ednote{B: Not complete enough.
-Edit further and cite}
-
-Lastly, we ensure the matrix $B$ has unit spectral radius.
-This is necessary due to the Causal Embedding Theorem~\ref{Thm_CET} which can then guarantee the existence of semi-conjugacy between the systems $(Y_T, G_T)$ and $(\widehat{U}_T, \widehat{T})$.
-In fact, the theorem actually requires that $\alpha{B}$ have spectral radius $<1$, so we impose the restriction $\alpha<1$ and consider only the matrix $B$'s spectral radius from thereon.
-
-Note that to learn $\Gamma: (x_{n-1},x_{n}) \mapsto u_n$ through our feedforward network, we could learn the principal components of $(x_{n-1},x_{n})$ as well. 
-We employ this since it provides slightly better results empirically but is not essential.
-
-Principal components are used only for a more efficient state representation to possibly reduce learning errors; it is not for a lossy approximation since we use all principal components \textbf{(Cite.)}\ednote{Ask M: Is there a specific resource to cite?}.
-More explicitly, denote $X_{1:N}$ as the matrix with the first $N$ states of the network data represented by row vectors.
-If $X_{1:N}=U\Sigma P^T$\ednote{Expand here} represents the singular value decomposition of $X_{1:N}$ \ednote{Ask: Is there a specific resource to cite?}, we then denote the principal component matrix $P$, and the principal components are given by
- $Z_{1:N}=X_{1:N}P.$
-These principal components are then used to train a feedforward neural network. Denote the row vectors of $Z_{1:N}$ by 
-$z_i^T, i=1,2,\cdots,N$ and the neural network by $NN$, the network is subsequently trained to learn an approximation of the map
-$NN: (z_{n-1},z_n) \mapsto  u_n.
-$
-The learnt approximation of $NN$ can be used to approximate $\Gamma$ since 
-$$
-NN\left( \begin{bmatrix} 
-P^Tx_{n-1} \\
-P^Tx_n
-\end{bmatrix}
-\right) = NN \circ 
-\begin{bmatrix}
-P^T & 0 \\
-0 & P^T 
-\end{bmatrix}\begin{bmatrix}
-x_{n-1}\\
-x_n
-\end{bmatrix} = u_n = \Gamma(x_{n-1},x_n).
-$$
-
-\section{Experimental Results}
-In this section we showcase numerical results pertaining to 3 attractors that were simulated using the methodology described.
-We first consider a simple physical system that exhibits complicated dynamics.
-\subsection{Double Pendulum}
-
-The Double Pendulum consists of two pendulums fastened to one another such that the system moves as a whole.
-The first (or higher pendulum) is attached to a fixed point while the second is attached to the endpoint of the first.
-(See below)
-This system is a classical example of a system exhibiting chaos and is often used as explanation of the very well-known Butterfly Effect.
-\ednote{B: Originally though a reference here was needed, but no longer think so.} Consider again~\ref{fig:dp_sdic}.
-The system has four variables: angle and angular velocity of each pendulum rod. An ideal system is assumed where the rods are massless and no friction in the pivots are considered.
-\begin{figure}[ht]
-  \includegraphics[scale=0.35]{_dp_setup.eps}
-  \centering
-  \captionof{figure}{Pictorial representation of the setup of a DP. Image taken from~\cite{DPSetup}}
-  \label{fig:dp_setup}
-\end{figure}
-
-The scalar equations of motion are derived using Newtonian physics and are reproduced from~\cite{DPFormulas} in~\ref{eqns_dp}.
-Each pendulum component has an associated mass of the head, $m$, and length of the rod, $L_i$.
-Acceleration due to gravity is denoted by $g$ and initial velocities by $\dot{\theta}$
-\begin{eqnarray}\label{eqns_dp}
-  \ddot{\theta_{1}}  = \frac{-g(2m_1+m_2)\sin{\theta_1} - m_2g\sin(\theta_1-2\theta_2) - 2\sin(\theta_1-\theta_2)m_2({\dot{\theta_{2}}}^{2}L_2 + {\theta_{1}'}L_1\cos(\theta_1-\theta_2))} {L_1(2m_1 + m_2 -m_2\cos(2\theta_1 - 2\theta_2))}
-  \\
-  \ddot{\theta_{2}} = \frac{2\sin(\theta_1-\theta_2)(\theta_{1}'^{2}L_1(m_1+m_2) + g(m_1+m_2)\cos\theta_1 + \dot{\theta_{2}}^{2}L_2m_2\cos(\theta_1-\theta_2))}{L_2(2m_1 + m_2 -m_2\cos(2\theta_1 - 2\theta_2))}
-\end{eqnarray}
-
-The second-order system is converted to a first-order ODE system and solved using the Runge-Kutta numerical method.  
-Here we chose the values $m_1=2.0$, $m_2=2.0$, $L_1=1.5$ and $m_2=1.5$ with initial velocities $v_1=1.5, v_2=-2$ for the respective pendulum heads.
-Undamped DP data-sets were generated and used in this project due to the requirement that the map $T$ be surjective that we must not forget. 
-An example of this would be the case of a DP exhibiting energy-loss (i.e.
-where the map $T$ is not surjective).
-In such a case, the attractor is a single point in the phase space when the double pendulum comes to rest.
-It must be mentioned once again that if we use data outside the attractor, learning is not reliable. To see this, consder the graphs ~\ref{fig:damped_pendulum} and ~\ref{fig:dp_notsurjective}.
-
-\begin{figure}[ht]
-  \includegraphics[scale=0.4]{_dp_dying.eps}
-  \centering
-\caption{$x$- and $y$-coordinates of the two pendulum heads plotted for a dataset that exhibits dying oscillations.
-The map generating the (four-dimensional data) for a damped pendulum is not surjective.
-Data obtained from chaotic double pendulum dataset generated from videos takes of experimental pendulums~\cite{asseman2018learning}}
-\label{fig:damped_pendulum}
-\end{figure}
-
-\begin{figure}[ht]
-  \includegraphics[scale=0.3]{_dpfail_nonsurj.eps}
-  \centering
-\caption{Failure to forecast or learn a map which is not surjective: Evolution of the $y$-coordinate of the bottommost pendulum plotted here for 15000 timestep.
-Plotted here the predicted value(red) versus the actual value(blue).
-The plots overlap for the first 10000 steps, the duration of training to learn the map $\Gamma$.
-Afterwards, the trajectories start diverging and it the predicted value no longer accurately represents the complexity or energy-loss of the actual data.
-}
-\label{fig:dp_notsurjective}
-\end{figure}
-
-Having spent ample time to discuss the challenges, we now change tone so as to satisfy the optimist and present a successful run of the DP-data to predict the future movement of the pendulum.
-
-The system was driven for 15500 steps, of which the first 500 were immediately discarded to simulate the network 'forgetting'.
-The driving function $g$ had input and reservoir matrices were implemented as sparse matrices filled to 20\%.
-Parameter-values of $\alpha=0.99$ and $a=0.5$ were used.
-The remaining 15000 datapoints were used to train a network with 128 neurons. 
-The network was initialised to contain 16 hidden layers each with a layer dimension of 64, activated with the tanh activation function.
-Training was accomplished using 256 epochs with batch sizes 128 and a delay of 1.
-See~\ref{fig:dp_success_traj} for results pertaining to the trajectory taken by the DP's components and then~\ref{fig:dp_success_density} to witness associated densities of predicted vs.
-actual values.
-
-\begin{figure}[ht]
-  \centering
-  \includegraphics[width=0.95\linewidth]{_dp_success_4coords_traj.eps} 
-  \captionof{figure}{Predicted(red) and actual(blue) trajectories of the $x$- and $y$-coordinates of each of the pendulum heads.
-From top to bottom are as follows: $x$-coordinate of topmost pendulum head, $y$-coordinate of topmost pendulum head, $x$-coordinate of lower pendulum head, $y$-coordinate of lower pendulum head.
-  These graphs were constructed by predicting the DP 1000 timesteps into the future and in so doing illustrating the long-term consistency and accuracy of the learnt system.
-Here we lock onto the near-exact trajectory for about 300 timesteps and stay close up until about 600 seconds.
-}
-  \label{fig:dp_success_traj}
-\end{figure}
-\begin{figure}[ht]
-  \centering
-  \includegraphics[width=\linewidth]{_dp_success_4coords_hist.eps}
-  \captionof{figure}{Densities of predicted and actual values of $x$- and $y$-coordinate of pendulum heads.}
-  \label{fig:dp_success_density}
- \end{figure}
-
-We now consider an experiment where noise is added to the system and find it to be incredibly robust.
-We employ the same parameters as before and add noise from a normal-distribution with mean zero and standard deviation equal to 0.1 which translates to roughly 39dB's of noise.
-(For the signal-to-noise ration we adopt the formula $\text{SNR}_{dB}=10\text{log}_{10}\big(\frac{P_\text{{signal}}}{P_{\text{noise}}}\big)$ where $P_\text{{signal}}$ refers to the variance observed in the normalised data and $P_{\text{noise}}$ represents the variance of the added noise.)
-
-Once again consider the actual and predicted trajectories for the noisy DP dataset in Figure~\ref{fig:noisydp_success_traj}.
-Next we consider the densities of the predicted and actual values in Figure~\ref{fig:noisydp_success_density} and observe that they display highly similar distributions.
-
-\begin{figure}[ht]
-  \centering
-  \includegraphics[width=0.95\linewidth]{_dp_noise_FigCD_2030.eps} 
-  \captionof{figure}{Predicted(red) and actual(blue) trajectories of the $x$- and $y$-coordinates of each of the pendulum heads with noise added from a normal distribution with mean zero and standard deviation 0.1, equating to approximately 39dB's of noise.
-  These graphs were constructed by predicting the DP 1000 timesteps into the future and in so doing illustrating the long-term consistency and accuracy of the learnt system.
-Here we lock onto the near-exact trajectory for about 250 timesteps and stay close up until about 400 seconds.
-}
-  \label{fig:noisydp_success_traj}
-\end{figure}
-\begin{figure}[ht]
-  \centering
-  \includegraphics[width=\linewidth]{_dp_noise_2B_2030.eps}
-  \captionof{figure}{Densities of predicted and actual values of x-coordinate of first pendulum head.}
-  \label{fig:noisydp_success_density}
- \end{figure}
-
- Lastly, a run was done where only a single coordinate (the $y$-coordinate of the lower pendulum head) was fed into the system and we can illustrate through the obtained success that the coordinate contains all the information in its past states necessary to forecast its future evolution.
- \textbf{Insert graphs here. Had to quickly change something here.}
-
-\subsection{Additional Attractors: Clifford \& Thomas}
-We follow the methodology presented in \cite{manjunath2021universal} but opt to consider attractors different from the Lorenz system and H\'enon map in this project.
-
-\subsubsection{Thomas' Cyclically Symmetric Strange Attractor}\label{Thomas_Attractor}
-Thomas' Cyclically Symmetric Strange Attractor, a 3D attractor proposed in 1999 by Ren\'e Thomas in \cite{ThomasAttractor}, is described by a set of three equations which is distinctly reminiscent of the Lorenz system, another deterministic three-equation model exhibiting chaotic behaviour.It has a single parameter $\beta$ and has been shown to transition to chaotic behavious when $\beta<0.208186$~\cite{Thomas_BetaParameter}.
-\begin{eqnarray}\label{eqns_thomas}
-  x_{n+1} = \sin(y_n) - \beta{x_n} \\
-  y_{n+1} = \sin(z_n) - \beta{y_n} \\
-  z_{n+1} = \sin(x_n) - \beta{z_n}
-\end{eqnarray}
-
-The behaviour of the system with a $\beta$-parameter value of 0.1056 was simulated by scaling it to fit inside the interval $[-1,1]$.
-Two cases were considered: noise-free and the one perturbed by noise.
-Noise was generated from a normal distribution with mean 0 and standard deviation 0.05, which to translates approximately 33dB.
-($\text{SNR}_dB$ is calculated exactly as for the DP-data above)
-
-A sequence of 3000 observations in both the clear and noisy data-sets respectively were fed into the system with 300 entries discarded to simulate the network's loss of memory.
-The next 9000 steps were predicted and compared with true data.
-
-\begin{figure}[ht]
-  \centering
-  \minipage{0.5\linewidth}
-    \centering
-    \includegraphics[width=0.8\linewidth]{_thomas_noisy_0.1056.eps}
-  \endminipage\hfill
-  \minipage{0.5\linewidth}
-    \centering    \includegraphics[width=0.8\linewidth]{_thomas_clean_0.1056.eps}
-  \endminipage
-  \captionof{figure}{True(red) and predicted(blue) trajectories for the Thomas attractor with parameter value $\beta=0.1056$.
-To the left: Noise added from normal distribution with mean zero and standard deviation 0.05.
-Right: No noise added to dataset.}
-\end{figure}
-
-The RNN was initialised as follows: the dimension of the network is set equal to 500 neurons and 12 hidden layers of dimension 64 each were initialised.
-Training is realised through 150 epochs of batch sizes of 128 each; the map is learnt via the Adam Optimiser using finer and finer learning rates (0.001 and then divided by 10 at every iteration).
-
-\begin{figure}[ht]
-  \centering
-  \includegraphics[scale=0.35]{_Thomas_1.eps}\caption*{Predicted trajectories of the $x$- and $y$-coordinates for the Thomas attractor with parameter-value $\beta=0.1056$ demonstrate empirically the ability to predict the evolution of the trajectory for the next an estimated 100 timesteps into the future near-exactly.
-Here no noise was added.}
-  \includegraphics[scale=0.5]{_Thomas_3.eps}\caption*{Represented here the learnt (blue) and actual (red) densities of the first coordinate ($x$-coordinate) of the dataset for the Cyclically Symmetric Attractor with parameter-value $\beta=0.1056$.}
-\end{figure}
-
-\subsubsection{Fractal Dream Attractor}
-
-Another 2-equation discrete-time map named the Fractal Dream Attractor, or more commonly known as the Pickover Map (first discovered by Clifford A.
-Pickover and discussed in his fascinating book "Chaos in Wonderland"\cite{PickoverChaos}), was also considered.
-\begin{eqnarray}\label{eqns_clifford}
-  {x_{n+1}=\sin(ay_n) + c\cos(ax_n)} \\
-  {y_{n+1}=\sin(bx_n)+d\cos(by_n)}
-\end{eqnarray}
-
-\begin{figure}[ht]
-  \centering
-  \includegraphics[width=1.0\textwidth,left]{_Clifford_1_nonoise.eps}
-  \caption*{These graphs were constructed by predicting the Clifford system 1000 steps into the future and in so doing illustrating the long-term consistency and accuracy of the learnt system.
-As perceived here, we are able to lock on to the trajectory of the Clifford map almost exactly for the first 25 steps.}
-  \minipage{0.5\textwidth}
-      \centering
-      \includegraphics[width=\linewidth]{_Clifford_2_nonoise.eps}
-  \endminipage\hfill
-  \minipage{0.5\textwidth}
-    \centering
-    \includegraphics[width=\linewidth]{_Clifford_3_nonoise.eps}
-  \endminipage
-  \caption{The bottom row, left, shows a plot of the true attractor(red) and the predicted values(blue).
-Bottom row, right, shows the distribution of the first coordinate of the learnt system (blue) and the actual Clifford system (red).
-}
-  \label{fig:Clifford}
-\end{figure}
-
-The behaviour of the system with a parameter values of $x_0=0$, $y_0=0$, $a = -1.7$,  $b = 1.8$, $c = -1.9$ and $d = -0.4$ was examined.
-Data was scaled to fit inside the interval $[-1,1]$. Only the noise-free case was investigated. 
- A sequence of 25000 observationswas fed into the system with 15000 entries discarded to simulate the network's loss of memory. The next 1000 steps were predicted and compared with true data.
-The RNN was initialised exactly as for Thomas' Cyclically Symmetric Strange attractor: the dimension of the network is set equal to 500 neurons and 12 hidden layers of dimension 64 each were initialised.
-Training is realised through 150 epochs of batch sizes of 128 each; the map is learnt via the Adam Optimiser using finer and finer learning rates (0.001 and then divided by 10 at every iteration).
-
-\section{Functional Complexity Reduction in Noninvertible maps}
-
-% Besides being able to learn the action of the Koopman operator exactly, our setup has other advantages over selecting or optimizing the choice of observables as in the Extended Dynamic Mode Decomposition (EDMD) algorithm (see, [21, 22]).
-We can alter the functional complexity of G T by changing the number of observables or equivalently by altering the dimension of X – in the case of a RNN implementation of g, by simply changing the number of neurons in the network.
-Empirically, increasing the dimension of X is found to increase the linear relationship (or intuitively reduces the functional complexity of G T ) that is measured as a generalization of the Pearson correlation coefficient to random vectors (see, [44]) between (x_{n−1} , x_{n}) and G_T (x_{n−1} , x_n), and such numerical evidence is tabulated in [11, Table 1].
-In contrast, expanding the set of observables while using EDMD to learn a map with a lower functional complexity is a highly involved task, not least because so much that one often does not know how to guess a new observable, as well as the fact that adding observables does not necessarily retain the invariance of the span of the observables.
-
-% Empirical evidence shows that a linearity measure like a Pearson correlation coefficient points at reduced functional complexity while learning the dynamics in the state space of the driven system.
-
-% Empirically (see Table ??), increasing the dimension of X is found to increase the linear relationship
-% (or intuitively reduce the functional complexity of G T ) that is measured as a generalization of the Pearson correlation coefficient to random vectors (see, [?]) between (x n−1 , x n ) and G T (x n−1 , x n ).
-If Σ a and Σ b denotes the covariance matrices of the vectors (x n−1 , x n ) and G T (x n-1 , x n ) respectively, and if Σ ab denotes the covariance matrix between the vectors (x n-1 , x_n ) and G_T (x n-1 , x_n ), then the multidimensional 18correlation coefficient is computed by using the traces of these matrices:
-% \[\pho = \frac{tr(\sum_{ab})}{tr(\sqrt{\sum_{a}\sum_{b}})}\]
-
-% This multidimensional correlation coefficient satisfies most of the well-known properties of the one-dimensional Pearson coefficient.
-In particular, ρ = ±1 if and only Y = AX + b for some invertible matrix A and vector b.
-Hence ρ, and in practice, the estimator ρ̂, can be used to measure the linear relationship between two random vectors of the same dimension and thus serves as an indicator of functional complexity.
-
-% We remark that the estimation of such linear relationship in (see Table ??) obtained by sample correlations alone is justified whenever {u n } is a realization of an ergodic process since then when g has the USP, any solution {x n } is also a realization of an ergodic process [?].
-In the ergodic input case the generalized Pearson correlation coefficient (between (x n−1 , x n ) and G T (x n−1 , x n )) is independent of n.
-
-In the case of non-invertible maps, we note that the map $G_T$ is a homemorphism although $T$ is not.
-Further, empirically it is found that $G_T$ is much more linear than $T$ as measured by what is called the multidimensional correlation coefficient $\rho$.
-This coefficient is a generalization of the Pearson correlation coefficient for real-valued random variables to random vectors (see, \cite{puccetti2019measuring}).
-We state teh details of computing this coefficient between
-$(x_{n-1},x_n)$ and $G_T(x_{n-1},x_n)$. 
-If $\Sigma_{a}$ and $\Sigma_{b}$ denotes the covariance matrices of the vectors $(x_{n-1},x_n)$ and $G_T(x_{n-1},x_n)$ respectively, and if $\Sigma_{ab}$ denotes the covariance matrix between the vectors $(x_{n-1},x_n)$ and $G_T(x_{n-1},x_n)$, then the multidimensional correlation coefficient is computed by using the traces of these matrices:
-\[
-    \rho= \frac{\text{tr}({\Sigma_{ab}})}{\text{tr}({\sqrt{\Sigma_a\Sigma_b}})}.
-\]
-This multidimensional correlation coefficient satisfies one of the desired and well-known properties of the one-dimensional Pearson coefficient.
-That is, $\rho=\pm 1$ if and only if $Y\overset{d}{=}AX+b$ for some invertible matrix $A$ and vector $b$.
-Hence $\rho$, and in practice can be used  to measure the linear relationship between two random vectors of the same dimension and thus serves as an indicator of functional complexity.
-For the non-invertible system defined by the Clifford attractor, we indicate the difference by which $\rho$ changes for $G_T$ in Table~\ref{tbl_attractorsPearson}.
-Following \cite{manjunath2021universal}, we say that this multidimensional correlation coefficients $\rho$ to evidence a general reduction in the functional complexity of the map $G_T$ emerging due to a RNN.
-In the
-Table ~\ref{tbl_attractorsPearson}, the second column corresponds to the amount of artificial neurons (the dimension of $X$) in the implemented RNN.
-
-\begin{center}
-  \begin{table} 
-      \scalebox{0.75}
-      {\begin{tabular}{|c|c| c c |} 
-          \hline
-          Input & Dimension of $X$ & $u_n$ vs $u_{n+1}$ 
-          & $\begin{bmatrix}
-              x_{n-1}\\
-              x_n
-          \end{bmatrix}$ vs $\begin{bmatrix}
-              x_n\\
-              x_{n+1}
-          \end{bmatrix}$ \\
-          \hline      
-          % \hline
-          % $(w_n)$ is the Lorenz states, sampled every 0.1 timestamps. & & &\\
-          % \multirow{3}{*}{$u_n=\frac{1}{100}w_n$} & 10 &0.9311  & 0.9731\\
-          %     & 100 &  & 0.9934\\
-          %     & 1000 & & 0.9930\\              
-          % \multirow{3}{*}{$u_n=\frac{1}{10}(\sin(0.1w_{n,x})+\sin(0.1w_{n,y})+\sin(0.1w_{n,z}))$} 
-          %     & 10 &0.8401  & 0.9392\\
-          %     & 100 & & 0.9616\\
-          %     & 1000 & & 0.9737\\
-          % \hline
-
-               \hline  
-               $(w_n)$ comes from the Clifford Attractor \\ (for $a = -1.7; b = 1.8; c = -1.9; d = -0.4$): & & & \\
-        {$w_{n+1}= \begin{bmatrix} \sin(aw_{n,y}) + c\cos(aw_{n,x}) \\ 
-                                                                    \sin(bw_{n,x})+d\cos(bw_{n,y}) \end{bmatrix}$} & & & \\
-          \multirow{3}{*}{$u_n = w_n-\overline{w}$}
-              & 10 & -0.21713 & 0.78719 \\
-              & 100 & &  0.8334 \\
-              & 1000 & & 0.84854 \\
-          % \hline 
-          % $(w_n)$ comes from the Logistic map: & & & \\
-          % {$ w_{n+1}  = 4w_n(1-w_n)$} & & & \\
-          % \multirow{3}{*}{$u_n = w_n - 0.5$} 
-          %     & 10 &-0.0314  & 0.5577\\
-          %     & 100 &  & 0.7150\\
-          %     & 1000 & & 0.7826\\
-          % \hline
-          % \hline
-          % $(w_n)$ comes from the Svensson Attractor \\ (for $a = -1.7; b = 1.8; c = -1.9; d = -0.4$): & & & \\
-          % {$w_{n+1}= \begin{bmatrix} d\sin(aw_{n,x}) - \sin(bw_{n,y}) \\ 
-          %                                                             c\cos(aw_{n,x})+\cos(bw_{n,y}) \end{bmatrix}$} & & & \\
-          %   \multirow{3}{*}{$u_n = w_n-\overline{w}$}
-          %       & 10 & -0.2243 & 0.49698 \\
-          %       & 100 & &  0.63310 \\
-          %       & 1000 & & 0.71151 \\
-           \hline
-                \end{tabular}} %\ednote{http://paulbourke.net/fractals/peterdejong/}
- \label{Table_FC}
- \caption{Following \cite{manjunath2021universal}, we determine the multidimensional correlation coefficients $\rho$ to evidence a general reduction in the functional complexity of the map $G_T$ emerging due to a RNN.
-Rows correspond to distinct dynamical systems considered in the experiments
-      The second column corresponds to the amount of artificial neurons (the dimension of $X$) in the implemented RNN.
-      The last two columns comprise numerical estimates of $\rho$ for the relevant vectors. The linear association between $u_n$ vs $T(u_{n})$, as well as the linear relationship $\begin{bmatrix}
-              x_{n-1}\\
-              x_n
-          \end{bmatrix}$ vs $G_T \left( \begin{bmatrix}
-              x_{n-1}\\
-              x_n
-          \end{bmatrix}\right)$ may be contrasted by the reader. 
-          Finally, the RNN's dimension (from \eqref{Seq_RNN}) is provided to exhibiting that an increase in dimension of the driven system will 
-          typically result in a map $G_T$ with less functional complexity.}.
-      \end{table}\label{tbl_attractorsPearson}
-\end{center}
-
-\chapter{Conclusion}\label{ch6}
-\input{chapter6.tex}
-
-\bibliographystyle{pnas-new}
-\bibliography{pnas}
-
-\end{document}

### 2022-01-17 projectReport/chapter2.tex
Cleaner setup

--- 2022-01-17 draft_16Jan/hiscurrent.tex
@@
-\documentclass[a4paper,12pt,twoside]{report}
-% \documentclass[a4paper,12pt,twoside]{book}
-\pagestyle{headings}
-\usepackage[utf8]{inputenc}
-\usepackage{filecontents}
-\usepackage[many]{tcolorbox}
-\usepackage{graphicx}
-\usepackage{harpoon}
-%\usepackage{apacite} 
-\usepackage{xcolor}
-\usepackage{latexsym}
-\usepackage{amssymb}
-\usepackage{amsmath}
-\usepackage{mdframed}
-%\usepackage[authoryear]{natbib}
-\usepackage{hyperref, url}
-\usepackage[show]{ed}
-\usepackage{subfigure}
-\usepackage[ntheorem]{empheq} 
-\usepackage{amsthm, amssymb, amsfonts, latexsym}
-% \usepackage{pst-all}
-
-\usepackage{enumitem}
-\usepackage{multirow}
-\usepackage{longtable}
-\usepackage[export]{adjustbox}
-% $f\arrowvert_A$ To restrict functions
-
-\setlength{\parskip}{1.7em}
-\setlength{\parindent}{0em}
-
-%\setlength{\baselineskip}{20pt}
-
-\setlength{\textwidth}{429.75499pt}
-\setlength{\textheight}{643.20255pt}
-\setlength{\oddsidemargin}{5 mm}
-\setlength{\evensidemargin}{5 mm}
-\setlength{\topmargin}{0 mm}
-\setlength{\headsep}{0 mm}
-\setlength{\headheight}{0 mm}
-
-\usepackage{pst-node}
-
-\newcommand{\citealtt}[1]{\citeauthor{#1},\citeyear{#1}}
-\newcommand{\myycite}[1]{\citep{#1}}
-
-\mathchardef\mhyphen="2D
-
-%\newcommand{\cev}[1]{\reflectbox{\ensuremath{\vec{\reflectbox{\ensuremath{#1}}}}}}
-
-\usepackage{pst-node,graphicx,pst-blur}
-%\uspackage{auto-pst-pdf}
-%\usepackage{tikz-cd}
-
-\makeatletter
-\DeclareRobustCommand{\cev}[1]{%
-  \mathpalette\do@cev{#1}%
-}
-\newcommand{\do@cev}[2]{%
-  \fix@cev{#1}{+}%
-  \reflectbox{$\m@th#1\vec{\reflectbox{$\fix@cev{#1}{-}\m@th#1#2\fix@cev{#1}{+}$}}$}%
-  \fix@cev{#1}{-}%
-}
-\newcommand{\fix@cev}[2]{%
-  \ifx#1\displaystyle
-    \mkern#20mu
-  \else
-    \ifx#1\textstyle
-      \mkern#20mu
-    \else
-      \ifx#1\scriptstyle
-        \mkern#26mu
-      \else
-        \mkern#26mu
-      \fi
-    \fi
-  \fi
-}
-
-\makeatother
-
-\newcommand{\xcm}{\epsfxsize=3.1cm}
-\newcommand{\fig}[1]{\epsfbox}
-% \newcommand{\bp}{\begin{minipage}{3.1cm}}
-% \newcommand{\ep}{\end{minipage}}
-
-\newtheorem{Definition}{Definition}[]
-%\newmdtheoremenv{Theorem}{Theorem}[]
-\newtheorem{Theorem}{Theorem}[]
-\newtheorem{Lemma}{Lemma}[]
-\newtheorem{Proposition}{Proposition}[]
-\newtheorem{Corollary}{Corollary}[]
-\newtheorem{Remark}{Remark}[]
-\newtheorem{Example}{Example}[]
-
-%Shortcut symbols
-\newcommand{\Ftheta}{\ensuremath{F_\theta}}
-
-\makeatletter 
-\renewcommand{\thefigure}{\@arabic\c@figure}
-\makeatother
-\usepackage{xr}
-
-%%%% PNAS GRAPHICS SETUP- FROM THE PNAS STYLE LATEX FILE
-\RequirePackage{graphicx,xcolor}
-\RequirePackage{colortbl}
-\RequirePackage{booktabs}
-\RequirePackage{algorithm}
-\RequirePackage[noend]{algpseudocode}
-\RequirePackage{changepage}
-\RequirePackage[twoside,%
-				letterpaper,includeheadfoot,%
-				layoutsize={8.125in,10.875in},%
-                layouthoffset=0.1875in,%
-                layoutvoffset=0.0625in,%
-                left=38.5pt,%
-                right=43pt,%
-                top=43pt,% 10pt provided by headsep
-                bottom=32pt,%
-                headheight=0pt,% No Header
-                headsep=30pt,%
-                footskip=25pt,
-                marginparwidth=38pt]{geometry}
-\RequirePackage[labelfont={bf,sf},%
-                labelsep=period,%
-                figurename=Fig. ]{caption}
-\setlength{\columnsep}{13.5pt} % Distance between the two columns of text
-\setlength{\parindent}{12pt} % Paragraph indent
-
-%% Figure caption style
-\DeclareCaptionFormat{pnasformat}{\normalfont\sffamily\fontsize{7}{9}\selectfont#1#2#3}
-\captionsetup*{format=pnasformat}
-
-%%% GREYBOX AROUND FIG
-\definecolor{lightgray}{gray}{0.95}
-\newcommand\greybox[1]{%
-  \vskip\baselineskip%
-  \par\noindent\colorbox{lightgray}{%
-    \begin{minipage}{\textwidth}#1\end{minipage}%
-  }%
-  \vskip\baselineskip%
-}
-
-\newtcolorbox{paperbox}[1][]{
-    enhanced,
-    colback=white,
-    boxrule=0pt,
-    boxsep=0pt,
-    arc=0mm,
-    width=0.8\linewidth,
-    fuzzy shadow={0mm}{-4pt}{-4pt}{1mm}{black!30!white},
-    #1
-}
-
-% \externaldocument{Supp_Koopman-0808}
-\begin{document}
-
-\title{Data-Driven Modelling \& \\ Forecasting of Chaotic Systems}
-
-\author{B.L. Nortier} 
-\date{January, 2022}
-\date{}
-\maketitle
-
-\begin{abstract}
-  It is a well-established practice to collect measurements from underlying artificial, natural, and physical systems through observables in order to realise a dynamical model with greater descriptive power.
-  Many data-driven approaches to forecast the future evolution of systems exist.
-However, many of these methods rely on an initial understanding of the system and we often do not fully understand the many complexities of more intricate systems that we interact with on a daily basis.
-
-  In this project, the base assumption is that data is originates from  some discrete-time dynamical system, particularly one of a chaotic nature but no assumptions are made pertaining to a specific dynamical system.
-  A general learning problem is formulated and a well-known approach to the problem, alongside its practical limitations, is then considered.
-  The notion of driven dynamical systems is introduced, and we subsequently establish the theoretical existence of a dynamical system possessing certain properties which guarantees it to be topologically conjugate (`dynamically equivalent`) to the dynamics of the underlying system from which measurements were originally taken.
-  Our methodology is implemented practically by making use of recurrent neural networks, a form of deep learning, to learn a map topologically equivalent to the underlying system that forecasts future states.
-  In so doing we present the workability of an approach to obtain long-term topological and statistically consistent predictions for simple physical systems as well as a some chaotic attractors.
-  \end{abstract}
-
-  \tableofcontents
-
-\chapter{Introduction}\label{ch1}
-
-Experimenting with biological, physical, and artificial systems to generate a more informative dynamical model is a well-established practice in modern science.
-Traditional methods for modelling physical systems are based on laws of physics that are based either on empirical relationships or intuition.
-For systems that evolve with time, physical laws yield mathematical equations that govern how the quantities evolve with time.
-However, our world is much more complex than that which can be distilled into elegant equations.
-We do not fully understand many of the more complex systems, nor do they provide us with a good physical intuition of the underlying principles governing system dynamics.
-To complicate this, the underlying systems we observe often display sensitive dependence on initial conditions despite having highly similar initial conditions.
-This is because differing orbits diverge quickly and to such an extent that it becomes seemingly impossible to retrace their steps back to the original conditions.
-Even the presence of usually `negligible' computational noise or measurement error renders long-term pointwise prediction infeasible.
-We encounter many difficultiesin the process of modelling such systems:
-\vspace{-8mm}
-\begin{enumerate}[noitemsep, label=\roman*.]
-  \item One may not have access to the complete states of the systems
-  \item The system may be described by functions that behave wildly because their graphs have a wild oscillatory behaviour, i.e., a large functional complexity.
-\end{enumerate}
-
-Models derived primarily from data can be classified into three categories: 
-\vspace{-8mm}
-\begin{enumerate}[noitemsep, label=\roman*.]
-  \item  Interpretable models (i.e., they establish relationships between internal physical mechanisms), 
-  \item Partially interpretable models capturing some modes of the dynamics, 
-  \item Non-interpretable models, defined as such mainly due to the fact that they are defined on a different phase space that is usually high-dimensional. 
-\end{enumerate}
-
-Examples in the literature attempting to forecast data from such systems have tried many different approaches with differing degrees of success. 
-When we have complete access to states of the system, an ordinary differential equation can be obtained from data (\cite{brunton2016discovering, champion2019data} and \cite{small2002modeling,xu2006modeling}) wherein one could approximate the vector field by a library of functions to obtain interpretable models.
-
-Recent partially interpretable models available in the literature have been based on the Koopman operator (see~\cite{koopman1932dynamical,budivsic2012applied}) to employ observables mapping the data onto a higher-dimensional space.
-This makes the dynamics in the higher-dimensional space more amenable for approximation by a linear transformation.
-Such methods do not guarantee exact reconstructions for nonlinear models, and in practice provide poor long-term consistency for a large class of chaotic dynamical systems \cite{Supp}.
-
-The non-interpretable models include the delay embedding and the machine learning algorithms.
-For example, one could learn a system conjugate to the underlying system by applying the Takens delay embedding method of delay-coordinates~\cite{takens1981detecting} when one has "good" observations from the system; this learnt system could then be used to forecast the observed data.
-Takens delay embedding theorem~\cite{takens1981detecting} and its various generalisations (see,~\cite{sauer1991embedology, stark1999delay, gutman2018embedding}) establishes the learnability of a system constructed by concatenating sufficiently large previous time-series observations of a dynamical system into a vector (called delay coordinates).
-This then confirms the existence of a map on the space of delay coordinates equivalent (or topologically conjugate) to the underlying map from which the observed time-series was first obtained.
-Although topological conjugacy guarantees an alternate representation of the underlying system, the quality still depends on numerous parameters, making the comprehension of the dynamics unreliable (see, [?]).
-One reason for this fragility is that the embedded attractor in the reconstruction space is not always an attractor of the map learnt in the reconstruction space, despite unquestionably being an invariant set.
-When the embedded object is not explicitly known to be an attractor (as explained in Chapter~ \ref{ch3}), it can cause predictions to fail.
-
-Practically, the application of Takens embedding involves learning a map through some technique, and consequently one wishes that these would have low functional complexity\cite{manjunath2021universal}, i.e., functions with fewer oscillatory graphs.
-On the other hand, pure machine learning methodology processes temporal information (like the echo state networks~\cite{Manju_ESP, Manju_IEEE, grigoryeva2018echo}) by mapping data onto a higher dimensional space for further processing.
-Although they perform  well on forecasting some dynamical systems, they fail completely on others (\ednote{Ask for resources here}) as there is often no guarantee that the right function was learnt during training.
-
-This project involves implementing and analysing non-interpretable models that can guarantee exact reconstruction. 
-The project work concerns the study and implementation of a method (\cite{manjunath2021universal}) that incorporates learning a function by mapping the data on to a higher dimensional space using what is called a driven dynamical system (See Chapter~ref{ch.4}).
-With a clear understanding of how the data is mapped onto the higher dimensional space, the method then permits the learning of a dynamical system topologically conjugate to that of the underlying system.
-Instead of linear regression as in the training of echo state networks, deep learning methods are employed to learn the correct function.
-With slight modifications to the implementation in the paper (\cite{manjunath2021universal}), we show that one can construct accurate non-interpretable models with the ability to reconstruct attractors from more hard-to-forecast systems like the double pendulum.
-(The forecasting of the time-series from a double pendulum has not been reported before.) Moreover, we also demonstrate that long-term statistical consistency is preserved.
-
-\emph{By solidifying the mathematical underpinnings of our theory, we hope to guarantee the ability to construct models with predictive power ranging from molecular biology to neuroscience.
-in the near future.
-(We can modify this sentence when the report is completed.)}
-
-The more intricate details of proofs are referred to where relevant throughout. This report is organised as follows: 
-\newline In Chapter~\ref{ch2}, we recall the definition of a discrete-time dynamical system, how a discrete-system arises from a flow of an ODE and then proceed to define the inverse-limit space and topological conjugacy of autonomous systems.
-\newline Chapter~\ref{ch3} introduces the problem of forecasting dynamical systems, states the Takens delay embedding theorem, and discusses various issues faced while forecasting. 
-\newline In Chapter~\ref{ch4}, a driven dynamical system is defined and discuss the properties of a specific class of driven dynamical systems that we make use of in this project.
-\newline Finally, in Chapter~\ref{ch5} we show the implementation of these forecasting methods, and conclusions are provided in Chapter~\ref{ch6}.
-
-\chapter{Discrete-time Dynamical Systems}\label{ch2}
-
 This chapter briefly describes a discrete-time dynamical system and what it means to exhibit chaos. We refer to \cite{devaney2018introduction, de2013elements} for more details.
@@
   \begin{figure}[ht]
-    \includegraphics[scale=0.74]{_logistic_sdic.eps}
+    \includegraphics[scale=0.74]{Graphs/_logistic_sdic.eps}
         \centering
@@
   \centering
-\includegraphics[scale=0.8]{_dp_sdic.eps}
+\includegraphics[scale=0.8]{Graphs/_dp_sdic.eps}
     \captionof{figure}{Three double pendulum heads with equal angular velocity and initial angles differing by 0,025 radians}\label{fig:dp_sdic}
@@
   \begin{figure}[ht]
-    \includegraphics[scale=0.6]{_tentmap_2.eps}
+    \includegraphics[scale=0.6]{Graphs/_tentmap_2.eps}
         \centering
@@
 It is instrumental during the process of forecasting the future evolution of dynamical systems.
-
-\chapter{The Learning Problem}\label{ch3}
-
-In this chapter, we acquaint ourselves with the question of forecasting dynamical systems with unknown underlying structures, state and discuss the shortcomings of the Takens delay embedding theorem and discuss various issues faced while forecasting.
-
-Consider a relatively simple learning problem: 
-Given the sequence $(u_0, u_1, \ldots, u_m)$ for $m\in\mathbb{Z}$, a finite segment of an orbit of the map $T$ where $T$ is defined by the update equation $u_{n+1} = Tu_n$, forecast the values $u_{m+1}, u_{m+1}$ where the map $T$ is unknown, given that $u_m$.
-
-A practical example of this would be the subsequent scenario: given the time-sequential coordinates of an object moving in space, predict the future positions of that object.
-We are very rarely, if at all, presented with a problem where the entire state information is available to us in the form $u_n$ at some timestep $n\in\mathbb{N}$.
-Consider then a more complicated learning problem:
-
-Suppose we only have the observations $\theta(u_0), \theta(u_1), \ldots, \theta(u_m)$ of the true system states $u$ in an unknown dynamical system $(U,T)$ and we wish to predict the values $\theta(u_{m+1}), \theta(u_{m+2})$ and so forth.
-
-First we establish the method of Takens delay embedding. 
-In this method, one considers a discrete-time dynamical system defined on a 'nice' space (a smooth manifold)\ednote{B: Should this be further explained?} that can be obtained as time-$K$ map of a continuous time-dynamical system, concepts formalised below.
-We make observations from such a system, i.e., we observe the evolution of $\theta(w_0)$ (where $w_0$ represents the initial state) by examining  a finite set of values of $u_n :=  \theta(w_n) = \theta(Tw_{n-1})$.
-The sequence $\{u_n\}$ represents a scalar time-series and intuitively $\theta$ represents a probe inserted into a bigger system which is itself only measuring/extracting a small part of the greater system state at time $t=n$.
-Consider, for example, a thermometer erected to measure the ambient temperature in a local village.
-This measurement function, the thermometer, is capturing only a single aspect, the temperature, of a much grander dynamical system entailing the current weather of the surrounding area.
-Even more than that though- it is measuring a miniscule part of the global weather system.
-
-%Given the sequence $\{u_1, u_2, \ldots, u_n}$, we define the variable \newline $y_n:=[\theta_n, \theta_{n-\tau}, \ldots, \theta_{n-(2d-2)\tau}, \theta_{n-(2d-1)\tau}]^{T}\in\mathbb{R}_{d}\times{1}$ where $\tau$ represents the lag and $d$ the dimension of the attractor.
-(We shall return to these terms shortly and their discussion may be put on hold for the time being).
-
-However, the actual state $u$ of a system is seldom, if ever, fully known.
-the most one can do is to insert probes into a system to obtain partial information through the measurements taken.
-Moreover, the process of taking a measurement itself introduces two additional aspects which complicate the problem: \ednote{B: Perhaps this edit makes more sense?
-M: Maybe replace difficulty by saying that we need to consider two aspects.
-}
-\vspace{-8mm}
-\begin{enumerate}[noitemsep, label=\roman*.]
-  \item A series of measurements over a specific time interval is inherently a discretisation procedure of the underlying continuous-time dynamical system.
-(Hence why we restricted our attention to discrete-time systems in the preceding section)
-  \item A probe will never be entirely accurate, and so the act of measurement introduces a certain measure of numerical noise/inaccuracy.
-\end{enumerate}
-
-From here we construct a multidimensional observable using the method of stacking previous observations, i.e., we create delay-coordinate map defined by
-$\Phi_{k,\theta}(w) := (\theta(T^{-k}w)\ldots,\theta(T^{-1}w),\theta(w))$.  
-The concept of an observable is understood in the sense of an observable as originally introduced by~\cite{takens1981detecting, genericObservableAeyels} to refer to a probe or measurement function inserted into the system.
-Takens' theorem, in essence, refers to the fact that for a sufficiently large $k$, we can define a dynamical system on the space $\mathbb{R}^{k+1}$ with states $\Phi_{k,\theta}(w)$, $\Phi_{k,\theta}(Tw)$, $\Phi_{k,\theta}(T^2w)$, etc.
-This dynamical system is topologically conjugate to the unknown underlying system $(W,T)$.
-We recall the Takens delay embedding theorem next.
-
-%We may then ask ourselves the question: Is it in any way possible to retain information about the system state $x(t)$ in this temporal data-series $\theta(x(t))$?
-The answer is easily yes if T is known.
-However, neither this, nor even the exact function $\theta$ is available.
-
-\section{Takens Embedding Theorem}\label{sect_Takens}
-
-We define first the concepts of homeomorphism and embedding:
-\begin{Definition}\rm
-  [\bf {Homeomorphism}]\label{Dfn_homeo}\rm
-  A homeomorphism is a function $f:Z\rightarrow Y$ between two topological spaces $Z$ and $Y$ that is continuous, bijective and has a continuous inverse. 
-\end{Definition}
-
-\begin{Definition}
-  [\bf {Embedding}]\label{Dfn_embed}\rm
-  Consider a homeomorphism $f:Z\rightarrow Y$ for $Y\subset X$. $Z$ is said to be embedded in $X$ by $f$.
-\end{Definition}
-
-Takens Theorem states a result establishing a relationship between the observed and underlying dynamical systems by showing that the concatenation of a sufficiently large number of previous observations into a vector will, under certain conditions, generate a map between the vectors from the respective systems.
-We formulate the theorem from \cite{takens1981detecting}.
-
-\begin{Theorem} 
-	[\bf Takens Embedding Theorem (adopted from \cite{takens1981detecting}] \label{Thm_Takens}
-         Let $W$ be a compact manifold of dimension $m$, and $d\ge m$ so that $2d$ is an integer. It is a 
-            generic property for the pair $(T, \theta)$,  where $T:W \to W$ is
-            a smooth diffeomorphism, and $\theta:W \to \mathbb{R}$ a smooth function, the map $\Phi_{2d,\theta}:W \to \mathbb{R}^{2d+1}$ defined on $W$ by 
-            $\Phi_{2d,\theta}(w) := (\theta(T^{-2d}w)\ldots,\theta(T^{-1}w),\theta(w))$
-            is a diffeomorphic embedding; by `smooth' we mean at least $C^2$.
-Consequently, there exists a map $F_\theta: \Phi_{2d,\theta}(W) \to \Phi_{2d,\theta}(W)$ defined by $$F_\theta: (\theta(T^{-2d}w),\ldots,\theta(T^{-1}w),\theta(w)) \mapsto
-            (\theta(T^{-2d+1}w),\ldots,\theta(w),\theta(Tw))$$
-           so that $(W,T)$ is topologically conjugate to 
-            $(\Phi_{2d,\theta}(W), F_\theta)$.    
-\end{Theorem}
-
-By generic we mean a residual set on a certain topology on appropriate spaces of functions (that we do not describe here) \ednote{B: Where can I find a citation for this?}. 
-By $C^2$, we make reference to a twice-differentiable function  with a continuous second derivative.
-In our scenario (and this remains important throughout) the input space $U$ is considered the attractor.~\label{attractor_U}
-
-Below we reproduce Figure 1 from \cite{Supp}.
-
-%Some explanation is beneficial.
-Takens establishes a delay-coordinate map $\Ftheta$ defined in \ref{eqn_takens} for $T$ the flow (\ref{defn_flow}) defined by the update equation  $Tu_n=u_{n+1}$ and $theta$ our measurement function.
-When confining a dynamical system to the manifold U (in which the underlying system is contained), Takens showed that if certain smoothness conditions are satisfied on $T$ and $\theta$, then the delay-coordinate map $\Ftheta$ embeds $U$ in the reconstruction space $R^{2d+1}$ \textbf{we haven't used the term reconstruction space yet, so this is falling out of the blue}, for almost every choice of measurement function $\theta$, the observable.
-By almost every choice here, we mean that we exclude some observations which do not give any information.
-To see a more precise discussion, see \textbf{xxx} for we do not concern ourselves with a further elaboration on this report.
-
-%Alternatively we may also make the statement: Takens' Embedding Theorem guarantees that almost all dynamical systems can be reconstructed from just one noiseless observation sequence \textbf{this isn't obvious from the theorem's formulation, is it?} i.e.
-for a great number of possible observation functions $\theta$, $\Ftheta$ preserves the topology of U.
-\textbf{'preserving the topology' is not something we've touched on before.
-Perhaps I should add to previous discussion so logic flows seamlessly here}.
-
-\begin{figure}[ht]
-  \includegraphics[scale=0.25]{_takensmap.eps}
-  \centering
-  \captionof{figure}{Schematic of Takens Embedding Theorem's usage and impact of iterates slipping from the attractor when the approximation $\tilde{F}_\theta$ of $F_\theta$ is learnt.}
-  \label{fig:takensmap}
-\end{figure}
-
-If we can learn the map $F_\theta$, from sufficient number of data points of $\{\phi_{2d,\theta}(w_n)\}$, then we also know how to forecast how $\theta(w_n)$.
-
-%Once again we pause and consider the graph to consolidate our understanding of the conjugacy (\textbf{are we being too simplistic?}).
-If we are in state $u$ and then evolve towards state $Tu$ before embedding into $\mathbb{R}^{2m-1}$, then this is equivalent to first embedding into $\mathbb{R}^{2m-1}$ and then evolving through the map $\Ftheta$, i.e.
-$\Phi_{2d,\theta}\circ{F}_{\theta} = T\circ\Phi_{2d,\theta}$.
-The map $F_theta$ is a homeomorphism as well.
-
-%Thus information about U can be retained in the time series' (or observation's) output.
-By preserving the topology on the manifold $U$ in the reconstruction space $X$ (\textbf{not discussed before}), we are also guaranteed the preservation of topological invariants of the manifold, of which dimensionality is one such invariant.
-(We note here that dimensionality will again be considered later in this project).
-
-\section{Practical Limitations to Takens' Theorem}
-Takens' Embedding Theorem is a powerful result, and it provides compelling reason to believe that one could conceivably reconstruct a system conjugate accurately to the underlying system. 
-Nonetheless, it does present some severe practical limitations:
-\vspace{-5mm}
-\begin{enumerate}
-\item Even supposing that we could indeed find $F_\theta$ , our \emph{approximation} of  $F_\theta$ is a map from a larger set $\mathbb{R}^{2d+1}$ containing the embedded attractor.
-There are, however, no theoretical guarantees that $F_\theta$ will retain  $V=\Phi_{2d,\theta}(W)$  as an attractor although $W$ could itself be an attractor.
-\item Takens theorem is stated only for noiseless observations.
-Due to noise $\epsilon_n$ the delay vector $\Phi_{2d,\theta}(w_n) + \epsilon_n$  may lie outside $V$.Furthermore, due to the chaotic nature of the underlying system (i.e.
-the fact that it has SIDC), the evolution of $\Phi_{2d,\theta}(w_n) + \epsilon_n$ under the map $F_\theta$ could move out of $V$ completely.
-This problem can be overcome  by using a driven dynamical system with some properties, and we discuss this in Chapter\ref{ch3}.
-\end{enumerate}
-
-% The Takens embedding theorem does not guarantee global dissipativity.
-% \ednote{B:Links with~\ref{subs_LearnGamma} in discussion of dissipativity of $\Gamma$.
-Sir, I'm not sure how to write this here - I think it a beneficial point to make so as to set the stage for the later discussion in ~\ref{subs_LearnGamma}, but I am unsure there as well.}
-
-Though not a uniquely related to the Takens embedding theorem, we do opt to make mention of another requirement for our work. 
-The methods described here are applicable for data originating from a surjective map. 
-This follows from the fact that when $T$ maps the space $U$ into a proper subset $B$ of$ $U, there exists some point in $U - A$ that does not have a preimage and we conclude that $T$ is not surjective.
-An example of this would be the case when a system displays energy-loss.
-Consider, for instance, consider the dying oscillations for a damped pendulum.
-(An account of greater depth  is provided in~\ref{ch5})
-
-This requirement is not overly restrictive, as a great number of chaotic systems do have a surjective map.
-(Cite) \ednote{B: I remember you saying this in one of our sessions, but don't have a source for it.
-Do you have a specific reference?}
-
-\chapter{Driven Dynamical Systems to Forecast Problems}\label{ch4}
-
-In this chapter, we discuss results on the mapping of temporal data obtained from a discrete-dynamical system onto a different space through the notion of a driven dynamical system. 
-We also consider the conditions a driven system should possess to avoid adding distortion to its state-space representation.
-We then describe the \emph{single-delay dynamics} (SDD) of the system and finally recall conditions on the driven system so that the SDD are conjugate (or at least semi-conjugate) to the underlying system.
-The SDD can then be used to forecast and reconstruct the underlying system.
-
-\section{Nonautonomous and Driven Dynamical Systems}
-
-\begin{Definition}
-  [\bf Nonautonomous Dynamical System (adopted from~\cite{Manju_ESP}))] \label{Dfn_NDS}\rm
-  A nonautonomous dynamical system (NDS) on a space $X$ is simply a dynamical system comprising a family of maps $\{g_n\}_{n \in \mathbb{Z}}$ where $g_n(\ldotp):=g(u_n,\ldotp):X\to{X}$ is continuous.
-Each $g_n$ arises as a consequence of an input $u_n$ from the input space $U$, a topological space.
-\end{Definition}
-
-We immediately recall the figure first defined in \eqref{eqn_conjugacy}  and remind ourselves of our ultimate goal to learn a system that is conjugate to an underlying or unknown dynamical system $(U,T)$ by making use of the measurements obtained from  the unknown system.
-
-%#####################################
-\begin{center}
-\psset{arrows=->, arrowinset=0.25, linewidth=0.6pt, nodesep=3pt, labelsep=2pt, rowsep=0.7cm, colsep = 1.1cm, shortput =tablr}
-\everypsbox{\scriptstyle}
-\begin{psmatrix}
-U & U\\%
-V & V.
-%%%
-%  \ncline{1,1}{1,2}^{T} 
-%  \ncline{1,1}{2,1} <{\phi}
-%   \ncline{2,1}{2,2}^{S}
-%  \ncline{1,2}{2,2} > {\phi}
-\end{psmatrix}
-%#####################################
-\end{center}
-
-To find a function $\phi$, the need now arises to consider a so-called driven dynamical system, a special case of a nonautonomous dynamical system.
-Two spaces are considered: input is taken from an input space $U$ and the state of a space $X$ is updated.
-(Previously, only the state was considered.)
-We do this to mimic the true conditions.
-By accounting for an extrogenous input sequence in $U$ which can also influence the  system-state at a specific time-step, it produces more general models than an autonomous system.
-
-\begin{Definition}
-  [\bf Driven, Compactly Driven Dynamical System] \label{Dfn_DDS} \rm
-A driven dynamical system comprises two topological metric spaces $U$, $X$ and a continuous function  $g:U\times{X}\to{X}$ where $g(u_n, x_n)=x_{n+1}$.
-If the input space $U$ is compact, we refer to the system as compactly driven. 
-\end{Definition}
-
-The update equation $x_{n+1} = g(u_n,x_n)$ for $n \in\mathbb{Z}$, input $u_n$ from $U$ and state $x_n$ belonging to $X$ (a compact space) generate the dynamics on X. 
-Abbreviated, we refer only to the \textit{driven system $g$}, where all other entities are understood implicitly.
-
-Notably, a nonautonomous dynamical system can be generated from $U$; any input $\overline{u}$, a bi-infinite sequence from $U$, gives rise to the sequence of self-maps $\{g(u_n, \cdot)\}_{n\in\mathbb{Z}}$ contained in $X$.
-Physically, one may think of this bi-infinite sequence as referring to a system that has been running for an extended period at the time of the first measurement taken from the system (alternatively, the first time a probe is inserted into the system to take an observation)
-
-\begin{Definition}
-  [\bf Entire Solution] \label{Dfn_Soln} \rm
-  A sequence $\{x_n\}_{n\in\mathbb{Z}}\subset X$ is called an entire solution (or simply a solution) of the driven system  $g$ with input $\overline{u}$ when it satisfies 
-  \[g(u_{n-1}, x_{n-1})=x_n\] for all $n\in\mathbb{Z}$
-\end{Definition}
-
-It is important to emphasise that a sequence $\{x_n\}$ satisfying the update equation above can only be a solution if $x_n\in{U}$ holds for all $n\in\mathbb{Z}$ and just for $n>0$.
-Consider the example below.
-
-\begin{Example} \rm \label{ex_halfux}
-  The only solution  $\{x_n\}_{n\in\mathbb{Z}}$ to the driven system  $g(u,x)=\frac{ux}{2}$, where $X=[0,1]$, $U=[0,1]$,  is the zero solution $x_n\equiv0$.
-  To see this, consider any $x_n=a\in[0,1]$ where $a\neq{0}$.  Let $\overline{u}\in{U}$ be an non-zero constant sequence , say $u_n=0.5$. 
-  The driven system may be rewritten as $x_{n-1}=\frac{2x_n}{u_{n-1}}=4x_n$ and the  iterates of $x_n$ in backward time will increase by a factor of 4 at each timestep. 
-  Thus for some $m\leq{n}$,  $1<x_m$ i.e. $x_m\notin{X}$. So ${\{x_n\}}_{n\in\mathbb{Z}}$ is not a solution and it follows that the only possible solution is the zero solution.
-\end{Example}
-% \ednote{Nicely written! Thank you}
-
-A system may also have multiple solutions, as evidenced in the example below.
-
-\begin{Example}\label{Ex_exp} \rm 
-  Consider the driven system $g(u,x)=x^2$ for $X=[0,1]$, $U=\mathbb{R}$.
-The system has an uncountable number of solutions, as there exists a solution for every $x\in{X}$ which also passes through the point $x$ and $lim_{n\to\infty}x_n=0$, $lim_{n\to{-}\infty}x_n=1$.
- The proof is deferred to immediately after the next paragraph (see~\ref{rem_proofEx}).
-\end{Example}
-
-As the solutions to a driven system are considered an important entity, we next identify a subspace $X_U$ of $X$ that contains all possible solutions.
-To realize such a subspace of a driven system $g$, the concept of a reachable set is defined.
-
-\begin{Definition}
-  [\bf Reachable Set]\label{Dfn_ReachableSet}\rm
-The reachable set of a driven system $g$ is exactly the union of all the elements of all the (entire) solutions, i.e., 
-\[X_U :=\Big \{x \in X:  x = x_k \mbox{ where  $\{x_n\}$  is a solution for some  $\bar{u}$} \Big \}.\]
-The set of all reachable states at a specific time $n$ for input $\overline{u}$ is denoted by $X_n(\overline{u})$
-\end{Definition}
-
-The reachable set itself can defined independently from $g$ possessing the ESP/USP.
-Note that $x\in{X_n(\overline{u})}$ if and only $g$ has a solution $\{x_k\}$ for $x_n=x$ and input $\overline{u}$, a result proved in nonautonomous dynamical systems literature~\cite{manjunath2014dynamics, manjunath2013echo}.
-\ednote{B: Adjusted sentence and cited}
-
-The set $X_n(\overline{u})$ is precisely the set of points $x$ through which some entire solution $\Psi$ passes at time $n$. We can give an alternate formulation of the set $X_n(\overline{u})$. 
-Now suppose, we for some fixed input 
-$\overline{u}$, we define $g_i = g(u_i,\cdot)$ for all $i\in \mathbb{Z}$, then the set of states
-
-\begin{equation} \label{eqn_association}
-X_{n,i}(\overline{u}) := g_{n-1} \circ \cdots g_{i+1} \circ g_i(X).
-\end{equation}
-
-The set $X_{n,i}(\overline{u})$ being the image of a finite composition of
-continuous maps is compact whenever $X$ is compact. Also $X_{n,i}(\overline{u})$ is
-nonempty.  Further, $X_{n,i}(\overline{u}) \supset X_{n,i-1}$. Hence $\bigcap_{i<n}
-X_{n,i}(\overline{u})$ is a nested intersection of closed nonempty subsets, and
-whenever $X$ is compact, the intersection is nonempty. We say $g$ is a topological contraction if $\bigcap_{i<n}X_{n,i}(\overline{u})$ is a singleton subset of $X$ for each $\overline{u}$.
-
-It turns out that $\bigcap_{i<n}X_{n,i}(\overline{u})$ is identically equal to $X_n((\overline{u})$ that we had defined earlier.
-This will lead to a result (for.
-see, \cite{manjunath2013echo}) that  \textit{$g$ being a topological contraction is equivalent to the existence of a exactly one entire solution.}
-
-% \begin{Definition}
-%   [\bf Topological Contraction]\label{Dfn_TopContr}\rm
-%   A function $g:U\times{X}\to{X}$ is a topological contraction if for all $n\in\mathbb{Z}$ and all $\overline{u}\subset{U}$, $X_n(\overline{u})$ is a singleton subset of $X$.  
-% \end{Definition}
-
-\begin{Remark}
-  [\bf Proof of Example~\ref{Ex_exp}] \label{rem_proofEx} \rm
-  For the driven system,  
-  $g(u,x)=x^2$ the evolution does not depend on the input at all.
-Hence $g(u,x)$ can be written as a single map $f(x)=x^2$ which is a homeomorphism on $[0,1]$.
-A left-infinite orbit is defined and it converges to $1$ while the right-infinite orbit converges to $0$.
-Or in other words the iterates of $f^{-1}$ converges to $1$ while the iterates of $f$ converge to $0$
-\end{Remark}
-
-Thus far, it has been demonstrated that a system may have one or more solutions; one may ask if a driven system always has a solution and, if so, whether it satisfies specific properties such as uniqueness.
-Should the driven system be compact, existence follows immediately, as shown in the following result.
-
-\begin{Theorem}\label{Thm_CompactExistence}
- Let $g$ be a driven system.  If $X$ is compact then for each input $\overline{u}$, there exists at least one solution to the driven system $g(\ldotp, x)$
-\end{Theorem}
-\begin{proof}
-  May be found in \cite{kloeden2011nonautonomous, manjunath2014dynamics, manjunath2013echo}
-\end{proof}
-
-% One may easily construct many systems with trivial solution-sets, such as $g(u,x)=x$ which has only the constant solution $x$ and so for $U=[-1,1]$, the system would have no solution if $|x|>1$.
-To refine the scenario, we consider only systems with unique solutions.
-
-\section{Unique Solution Property}
-
-\begin{Definition}
-  [\bf Unique Solution Property] \label{Dfn_usp}\rm
-  A driven system $g$ is said to have the Unique Solution Property (USP) if for each input $\overline{u}$ there exists exactly one solution. 
-  Alternatively we may formulate the USP as follows: $g$ has the Unique Solution Property if there exists a well-defined map $\Psi:{U}\to{X}$, with $\Psi({\overline{u}})$ denoting the unique solution.
-\end{Definition}
-
-One of the first results obtained after defining the USP is that every solution will attract different initial conditions towards the component parts of the solution.
-If $g$ has the USP, then any solution to $g$ is also a uniform attractor in nonautonomous dynamical systems literature~\cite{Manju_Nonlinearity}.
-The discussion on nonautonomous attractors is beyond the scope of this project, and we refer the reader to \cite{Manju_ESP, esann2012ids} for further reading.
-
-% Paraphrased.
-%Additionally, $g$ having the USP is in our context also equivalent to $g$ being a topological contraction \cite{manjunath2021universal}, or alternatively also that $g$ exhibits the Echo State Property (ESP).
-
-In the majority of Reservoir Computing(RC) literature (a machine learning approach) where the input is mapped onto a different but higher dimensional through a system called the reservoir and a readout that measures the state of the reservoir is trained, a notion of forgetting the states of the reservoir asymptotically is often used.
-
-% \ednote{How should I explain/qualify RC better here?
-It's mentioned without really explaining.
-M: Have put it in one sentence; just to let you know people use optical devices to map signal into a ``reservoir".
-B: Thanks!},
-Concepts like the ESP,  fading memory \cite{boyd1985fading} are some of these notions.  
-If $g$ possesses the ESP (equivalent in our context to the USP), the we are guaranteed that the whole of an input's left-infinite history will exactly determine the current system state; i.e.
-there is only ever the possibility of a single reachable state at a given instance of time \cite{jaeger2001echo,Manju_2020}.
-% Defined formally, the ESP is intrinsically linked to an input (see \cite{Manju_ESP}) Unfortunately, definitions and results concerning the ESP do not have much practical benefit since a relationship between the network's response and the temporo-statistical properties of the input are not directly described.
-A more detailed presentation may be found in~\cite{jaeger2001echo}.
-
-% We do however consider it important to mention that the ESP is also often discussed in terms of a stability property that holds for all inputs of a system~\cite{manjunath2020stability} and, furthermore, plays a key role in the design and training of recurrent neural networks(RNN) within the field of RC \cite{Manju_ESP}.
-Beyond forgetting the past states the ESP is also often discussed in terms of a stability property that is nearby inputs create close by responses (solutions), a concept called input-related stability~\cite{manjunath2020stability} and, furthermore, plays a key role in the robustness of recurrent neural networks(RNN), a component of the implementation discussed in~\ref{ch5}.
-
-\section{Conjugacies}
-
-Having already defined the reachable set $X_U$ as the collection of all elements of all entire solutions, we pause briefly in order to fix additional notation.
-Defining $\cev{u}^{n}:=(\ldots,u_{n-2} ,u_{n-1})$ as the left-infinite subsequence of an input up until time $n$, $\overleftarrow{U}$ is then the notation for all these left-infinite sequences in $U$.
-Moreover, $\cev{u}^{n}v:=(\ldots,u_{n-2} ,u_{n-1}, v)$ will symbolise the input up to time $n$ with $v \in U$ being the specific input value at time $n$. 
-The introduction of a new input at time $n$ can be described by the mapping $\sigma_v:   \cev{u}^{n} \mapsto \cev{u}^{n}v$. 
-The right-shift map, $r\overleftarrow{U}\to\overleftarrow{U}$, of an input sequence is defined $r: (\cdots, u_{-2},u_{-1}) \mapsto(\cdots, u_{-3},u_{-2})$
-
-The question now assumes the form: Can we establish a semi-conjugacy for the driven system $g$ as presented below.
-
-\begin{equation}  \label{Scomm_h}
-  %    \[ 
-      \psset{arrows=->, arrowinset=0.25, linewidth=0.6pt, nodesep=3pt, labelsep=2pt, rowsep=0.7cm, colsep = 1.1cm, shortput =tablr}
-   \everypsbox{\scriptstyle}
-   \begin{psmatrix}
-   \overleftarrow{U} & \overleftarrow{U}\\%
-   X_U & X_U.
-   %%%
-  %  \ncline{1,1}{1,2}^{\sigma_v} \ncline{1,1}{2,1} <{h}
-  %  \ncline{1,2}{2,2} > {h}
-  %  \ncline{2,1}{2,2}^{g(v,\cdot)}
-   \end{psmatrix}
-  % \]
-  \end{equation}
-
-We proceed to consider a specific subclass of conjugacies.
-
-  \begin{Definition}
-    [\bf Universal Semi-Conjugacy]\label{Def_UnivSemiConj} \rm
-    Given a driven system $g$, we  call a continuous and surjective map $h : \overleftarrow{U} \to X_U$ a universal semi-conjugacy if  diagram \ref{Scomm_h} commutes for all $v \in U$.
-  \end{Definition}
-
-  If the universal semi-conjugacy $h$ exists (i.e. the diagram in \ref{Scomm_h} commutes) then the solution $\Psi(\bar{u})$ will intuitively have no more ``complexity' than the input $\bar{u}$.
-
-But does such a function $h$ for the above driven system $g$ above exist?
-Whenever $g$ has the USP and $\Psi(u)=\{x_n\}_{n\in\mathbb{Z}}$ it follows that $h$, defined by  $h(\cev{u}_n):=g(u_n,x_{n-1})=x_n$, will satisfy the semi-conjugacy in the graph above \ref{Scomm_h}.
-Regrettably, such a continuous mapping $h$ is not guaranteed to exist when $g$ does not have the USP \cite[Lemma 5]{Manju_Nonlinearity}.
-Note that even when $h$ does exist, we are not guaranteed its injectivity. Considering again example \ref{ex_halfux}, we see that even if $h$ were to exist, it could not be injective as $X_U=\{0\}$.
-
-Re-sketching the commutativity diagram \ref{Scomm_h} above by replacing $X_U$ by its left-infinite sequence space $\overleftarrow{X}_U$, we obtain the diagram below.
-In this case, the function $H:\overleftarrow{U}\to\overleftarrow{X}_U$, a map that is both continuous and surjective, is called a \emph{causal mapping} that is defined next.
-
-\begin{Definition}
-  [\bf Causal Mapping]\label{Def_CausMap}
-  A continuous, surjective map $H:\overleftarrow{U}\to\overleftarrow{X}_U$ such that \[H\circ\tilde{g}_v=\sigma\circ{H}\] holds for all $v \in U$ where $\tilde{g}_v$ maps $(\ldots, u_{-2}, u_{-1})$ to $(\ldots, u_{-2}, u_{-1}, g(v, u_{-1}))$ is called a causal mapping.
-\end{Definition}
-
-\begin{equation} \label{SCausal_H}
-    %    \[ 
-        \psset{arrows=->, arrowinset=0.25, linewidth=0.6pt, nodesep=3pt, labelsep=2pt, rowsep=0.7cm, colsep = 1.1cm, shortput =tablr}
-     \everypsbox{\scriptstyle}
-     \begin{psmatrix}
-     \overleftarrow{U} & \overleftarrow{U}\\%
-     \overleftarrow{X}_U & \overleftarrow{X}_U.
-     %%%
-    %  \ncline{1,1}{1,2}^{\sigma_v} \ncline{1,1}{2,1} <{h}
-    %  \ncline{1,2}{2,2} > {h}
-    %  \ncline{2,1}{2,2}^{\tilde{g}_v}
-     \end{psmatrix}
-    % \]
-  \end{equation}
-
- \begin{Theorem}
-  For a compactly driven system, a causal mapping $H$ exists if and only if $g$ has the USP. 
-\end{Theorem}
-{\bf Proof.}  See~\cite[Th.3]{manjunath2013echo}.
-
-The driven system $g$ can induce an embedding of $\overleftarrow{U}$ in $\overleftarrow{X}$ as follows: 
-If the causal mapping $H:\overleftarrow{U}{\to}{\overleftarrow{X}_U}$ is injective (in addition to being surjective), it becomes the embedding of $\overleftarrow{U}$ in $\overleftarrow{X}$ induced by the driven system $g$.
-The continuity of $H^{-1}$ follows from the fact that $H$ is itself a continuous and surjective mapping of a compact space $\overleftarrow{U}$ in a Hausdorff space.
-and we refer to the map $H$ as a \emph{causal embedding}. 
-\ednote{B: edited the sentence here}
-
-When $g$ has the USP, the diagram below (adopted from~\cite{Manju_Nonlinearity}) illustrates the operation of the mappings $h$ and $H$.
-The mapping $h:\overleftarrow{U}\to{X_U}$ is also considered an observable as mentioned in the introduction of Chapter \ref{ch3}).
-
-\begin{figure}[ht]
-  \includegraphics[scale=0.4]{_actionofh_H.eps}
-  \centering
-  \captionof{figure}{Ladder-like behaviour of $h$ and $H$ to illustrate the causal mapping. (Figure reproduced from~\cite{Manju_Nonlinearity}) }
-  \label{fig:actionh_H}
-\end{figure}
-
-The function $H$ contains a countably infinite number of coordinate (or factor maps) since its codomain is $\overleftarrow{X}_U$.  
-As it is not practically feasible to consider a left-infinite sequence in $X_U$ (in other words an element of  $\overleftarrow{X}_U$) for learning, we wish to restrict the inputs we consider to a subspace of $\overleftarrow{U}$.
-Particularly, we restrict our attention to the  so-called inverse-limit space of a dynamical system and hope to embed it within some finite product of $X$.
-
-%We wish to explore the possibility of establishing a conjugacy rather than a mere semi-conjugacy and therefore are interested in the restriction of inputs to the left-infinite orbits of a dynamical system.
-
-The left-infinite orbit space is significant because a great amount of information about the original system can be gleaned from its topological structure~\cite{ingram2011inverse,Manju_IEEE}. 
-The formal topological structure for the inverse-limit space of a chaotic dynamical system is complicated, and further discussion is relegated to~\cite{kennedy2008inverse_limit, ingram2011inverse}.
-Broadly, we may conceive of the inverse-limit system of a dynamical system as a subspace of an infinite-dimensional space where each point in the inverse-limit space corresponds to a backward orbit of the map $T$.
-Following~\cite{manjunath2021universal, ingram2011inverse} we provide a definition:
-
-\begin{Definition}
-  [\bf Inverse-Limit Space, Inverse-Limit System]\label{Dfn_InverseL}\rm
- The subset $\widehat{U}_T\subset \overleftarrow{U}$ defined by \[\widehat{U}_T: = \{ (\ldots,u_{-2},u_{-1}): Tu_{i} = u_{i+1}\} \] is called the inverse-limit space of $(U,T)$.
- The self-map $\widehat{T}$  induced by $T$ on $\widehat{U}_T$  is defined by \[\widehat{T}: (\ldots,u_{-2},u_{-1}) \mapsto  (\ldots,u_{-2},u_{-1},T(u_{-1}))\] and the resulting dynamical system $(\widehat{U_T}, \widehat{T})$ is called the inverse-limit system of $(U,T)$.
-\end{Definition}
-
-The inverse-limit space is well-defined since $T : U \to{U}$ is surjective by assumption.
-We now get to the concept of embedding the inverse limit-space of the dynamical system in $X \times X$. 
-\begin{Definition} \rm
-	We say a driven system $g$ \emph{causally embeds} a dynamical system $(U,T)$ if it satisfies the two properties: (i) a universal semi-conjugacy exists and  (ii)  $H_2(\cev{u}) := (h(r\cev{u}),h(\cev{u}))$ embeds the inverse-limit space $\widehat{U}_T$ in $X \times X$.
-\end{Definition}
-
-  It is imperative to take note of a subtlety here: we refer to the \emph{map} $H$ as a (causal) embedding, whereas the driven \emph{system} $g$ (causally) embeds another \emph{system} $(U,T)$.
-\ednote{B: inserted line}
-
-\section{Choosing the driven system $g$}
-
-So far it is not clear if the USP alone ensures the ability to causally embedding a dynamical system and we mention that a driving function $g$ cannot just be chosen in a frivolous manner.
-This could possibly complicate our work.
-%\ednote{B: Grammarly}
-
-%The driving function $g$ cannot be chosen in a frivolous manner as it could possibly complicate our work.
-Specifically, one must be careful to avoid a choice of $g$ which would  add complexity to the obtained solution.
-
-%When a causal embedding $H$ exists for the driven system $g$, one can map an arbitrary input ${u}$ onto the solution space $X$ without additional distortion or information-loss \textbf{(Cite.)}.
-
-%When an embedding is established, the question of possible additional complexity in the solution is removed by guaranteeing that, since the systems are conjugate (semi-conjugate, \textbf{(refer)}), $g$ does not add any (some) complexity to the system.
-
-%To achieve a causal embedding embedding not just a causal mapping, it is undesirable to choose a function $g$ that quenches the temporal structure in $u$ by contracting to such a degree that the ability to recover information from the original system is lost completely.
-In the above example(\ref{ex_halfux}), the input's temporal variation could not be related to the reachable set as $X_U$ only consisted of a single element; little, if not no, information is encoded.
-To obtain a suitably complex function $g$, it is thus desired that the reachable set of a driven system  be large enough to relate to the input. 
-This is true even for embedding  the inverse-limit space of a dynamical system. \ednote{B: Expand or refer?}
-
- To this end, we recall the notion of State-Input (SI) Invertibility.
-
-\begin{Definition}
-  [\bf SI-Invertibility]\label{Dfn_SIinv}\rm
-  A driven system $g$ is said to be SI-Invertible if $g(*,x): U \to X$ is invertible for all $x\in X$.
-Alternatively it may be said that if, given $x_n$ and $x_{n-1}$, $u_{n-1}$ can be uniquely determined from $x_n=g(u_n,x_{n-1})$, then $g$ is said to be SI-invertible.
-\end{Definition}
-
-SI-invertibility promises that 'enough' information is retained when $g$ is chosen without introducing one's choice of $g$ will still ensure 'enough' information is retained without introducing unwanted complexity.
-'Enough' here refers to the fact that we may always recover the previous input value $u_n$ (if we know successive states $x_{n-1}, x_n$).
-
-Subsequently we define the relation $Y_T$ induced by $(U,T)$ on $X_U\times{X_U}$ for a driven system $g$ possessing SI-invertibility.  
-To describe the  SDD formally, we consider a dynamical system $T: U \to U$ and define a relation on the reachable set $X_U$, i.e. a subset on $X_U \times X_U$  defined by 
-$$Y_T:=\{(x_{n-1},x_n): \{x_k\}_{k\in \mathbb{Z}} \mbox{ is a solution for some orbit of } T \mbox{ and } n \in \mathbb{Z}\}.$$
-
-The following theorem establishes the existence of a well-defined map $G_T$ describing the single-delay dynamics(SDD) of the system above.
-
-\begin{Theorem}\label{Thm_GT_Exists}
-  For an SI-invertible driven system $g$ and a dynamical system $(U,T)$, the map $G_T: Y_T \to Y_T$ defined by the relation $(x_{n-1},x_n) \mapsto (x_n,x_{n+1})$ is well-defined. 
-  (This results holds even in the absence of $g$ possessing over the USP) 
-  % Moreover, the mapping $(x_{n-1},x_n) \mapsto u_{n}$ is well-defined when $x_{n-1}$ and $x_n$ are successive points on a solution obtained for an input  $\{u_n\}$ that is an orbit of $T$.
-  \end{Theorem}
-  {\bf Proof.} See~\cite[Th.3]{Supp}
-
-%'Without additional complexity' is guaranteed by the invertibility of the function $g$ that guarantees a one-to-one mapping between $u_n, x_{n-1}$ and $x_n$.
-In obtaining an SI-invertible function, therefore, we are guaranteed a system which will not lose so much information about its previous states in the forward-flow of time that one may not make any credible claims as to the original system's behaviour and properties.
-Simultaneously, it does guarantee us that the important information encoded in previous inputs is indeed preserved.
-
-At this stage it is worth taking note of a specific driven system in the form of a discrete state-space model which has acquired some adherence in applications \cite{Manju_IEEE}- especially those pertaining to Echo State Networks and the ESP.
-The function
-\begin{equation}  \label{eqn_driving}
-  g(u,x) = (1-a)x + a\overline{\tanh}(Au + \alpha Bx)
-\end{equation} 
-is SI-invertible and, if $\alpha B$ has a spectral norm $<1$, also possesses the USP \cite[Th.2]{manjunath2013echo }. 
-It is easy to show that $g$ is SI-invertible by recovering $u_n$ in 
-\begin{equation} \label{eqn_SI_RNN}
-  u_{n-1} := A^{-1}\bigg(\overline{\tanh}^{^{-1}}\frac{1}{a}\Big(x_{n+1}-(1-a)x_n\Big) \bigg) - \alpha B x_n
-  \end{equation}
-  when $x_{n-1}$ and $x_n$ are known.
-
-This specific driving function $g$ is used in our implementation and is discussed more completely in chapter~\ref{ch5}
-
-Despite the ease that with which one manipulates a left-infinite history in the realm of theory, it is impossible to obtain or use such a sequence in any real-life application.  
-Fortunately, one does not need the entire left-infinite history of an input in practice thanks to the Uniform Attraction Property(UAP). 
-We use an alternate version of the definition of UAP -- since the notion of nonautonomous attractors in this project would take some time to establish and detracts from the principal thrust of this project, we refer to \cite{Manju_Nonlinearity}.
-
-\begin{Definition}
-  [\bf Uniform Attraction Property]\label{Dfn_UAP}\rm
-  A driven system $g$ has the Uniform Attraction Property (UAP) if we initialize the driven system
-with an arbitrary initial value $y_m \in X$, and the sequence $y_{m+1}, y_{m+2}, y_{m+3},...$ satisfying $y_{k+1}= g(u_k,y_k)$ for $k \geq m$ then approximates an actual solution $\{x_n\}$ uniformly in the sense that given $\epsilon>0$ (independent of $y_m$) there is an integer $n$ so that $d(x_{n+i}, y_{n+i})<\epsilon$ for all $i\ge 0$, where $\{x_m\}$ is a solution.
-\end{Definition}
-
-The UAP guarantees that all trajectories will converge to the same trajectory as time moves forward.
-An illustration of this is shown in Fig.~\ref{fig:memloss_conttime} where the trajectory in red locks on to the supposed actual solution when the system has the USP.
-
-\begin{figure}[ht]
-  \includegraphics[scale=0.35]{_Nonlinearity_Fig_response.eps}
-  \centering
-\captionof{figure}{A coordinate of a simulated solution $(x_0,x_1,\ldots,x_{5000})$ of a RNN  plotted in red (with parameter $\alpha=0.99$ (where $g$ has the USP) and blue ($a=1$ and $\alpha=1.05$ where $g$ does not have the USP) while the matrices $A$ and $B$ are randomly generated, and $B$ has unit spectral radius.
-Reproduced from \cite{Manju_Nonlinearity}. }
-\label{fig:memloss_conttime} 
-\end{figure} 
-\ednote{B: How can I explain the black line? I don't think this makes sense to the reader immediately.}
-
-One may now appreciate an even more astounding result: $g$ having the USP is equivalent to the UAP as is proved in~\cite{Manju_Nonlinearity}.
-
-%We restate the above theorem.
-
-%\begin{Theorem}\label{thm_usp-contr-uap}  The following statements are equivalent:  \vspace{-8mm}  \begin{enumerate}[noitemsep, label=\roman*.]    \item $g$ has the USP.
-\item $g$ is a topological contraction.
-\item $g$ has the UAP.
-\end{enumerate\end{Theorem}{\bf Proof.}  May be found in \cite[Th.1]{Manju_Nonlinearity}
-
-% One need therefore not establish any additional results to ensure that the sequence uniformly approaches the unique solution $\Psi$.
-This vastly simplifies the effort necessary to set up a problem in order to guarantee that the underlying system will be accurately represented by the conjugate system.
-
-%This also solves the problem of perturbation/noise introduced by the observable or measurement function.
-Recall from before that an observable is inherently a map that discretises the underlying continuous-time system.
-Moreover, measurement and device error introduce mistakes.
-Since we are working with a chaotic system displaying sensitive dependence on initial conditions, these small errors could potentially send the trajectory into a completely different attractor.
-%In the presence of the UAP, however, small measurement errors introduced into the system do not pose the same danger as before since the trajectories will eventually tend towards the actual solution.
-% \ednote{(B: Is this Complete enough?)}
-% \textbf{From the perspective of perturbation of an autonomous dynamical system, if the resulting nonautonomous system has the USP, then the temporal structure of noise relates to the dynamics on a nonautonomous uniform attractor through a topological semi-conjugacy or a conjugacy.}
-
-\section{The next step in Dynamics}
-
-% Recall that $g$ is only being given inputs from the orbits of $T$.
-
-%Incredibly, this permits one to initialise a driven system $g$ with an altogether arbitrary initial value $y_m\in{X}$ where $m\in\mathbb{Z}$ and the UAP then guarantees that the sequence $\{y_{m+1}, y_{m+2},\ldots\}$ which satisfies the relation $y_n=g(u_n, y_{n-1})$  for  $k\geq{m}$  will approach the elements $\{x_n\}$ of the actual solution.
-This vastly simplifies the effort necessary to set up a problem in order to guarantee that the underlying system will be accurately represented by the conjugate system.
-
-We are interested in determining whether $\widehat{T}$ and $G_T$ are related.
-For this, we introduce  the function $H_2{\overline{}} := (h(r\overline{u}, h(\overline{u})))$ mapping the entirety of a left-infinite sequence to some element in $X\times{X}$.
-Indeed, it may be shown that $g$ having SI-invertibility and the USP immediately guarantees the existence of at least a semi-conjugacy between the systems $(Y_T, G_T)$ and $(\widehat{U}_T, \widehat{T})$.
-This is formalised in the Causal Embedding Theorem, which we state next:
-
-\begin{Theorem}
-  [\bf Causal Embedding Theorem (adopted from \cite{Supp})]
-  % {\bf (Causal Embedding Theorem.)}
- \label{Thm_CET}
-	Let $g$ be a driven system with SI-invertibility and the USP.
-Let $h$ denote the universal semi-conjugacy and $H_2(\overleftarrow{u}) := (h(r\overleftarrow{u}),h(\overleftarrow{u}))$, where $r$ is the right-shift map.
- Let $(\widehat{U}_T, \widehat{T})$  be the inverse-limit system of a dynamical system $(U,T)$. 
- Then the restriction of $H_2$ to $\widehat{U}_T$ is a topological semi-conjugacy between the inverse-limit system $(\widehat{U}_T, \widehat{T})$ 
-and the induced dynamical system  $(Y_T,G_T)$, i.e., the following diagram commutes
-\begin{equation} \label{comm_H_CET}
-\psset{arrows=->, arrowinset=0.25, linewidth=0.6pt, nodesep=3pt, labelsep=2pt, rowsep=0.7cm, colsep = 1.1cm, shortput =tablr}
- \everypsbox{\scriptstyle}
- \begin{psmatrix}
- \widehat{U}_T & \widehat{U}_T\\%
- Y_T &  Y_T.
- %%%
-%  \ncline{1,1}{1,2}^{\widehat{T}} \ncline{1,1}{2,1} <{H_2}
-%  \ncline{1,2}{2,2} > {H_2}
-%  \ncline{2,1}{2,2}^{G_T}
- \end{psmatrix}
- \end{equation}
-or in other words, $(Y_T, G_T)$ is a factor of  $(\widehat{U}_T, \widehat{T})$. Further, if $T:U \to U$ is a homeomorphism, 
-then $H_2$ embeds $\widehat{U}_T$ in $X_U \times X_U$, and hence $(Y_T, G_T)$ is conjugate to $(\widehat{U}_T, \widehat{T})$.
-\end{Theorem}
-{\bf Proof.}  May be found as the proof of~\cite[Th.4]{Supp}
-
-Recall that $H_2$ maps an entire left-infinite solution sequence from $\Psi$ to an element in $X\times{X}$. 
-We are drawing nearer and nearer to our principal target and our results carry more and more weight: If we are able to learn $G_T$, we will also have learnt $\widehat{T}$; and since $\widehat{T}$ is an extension of $T$, we will have obtained $T$ as well.
-%\ednote{B: Grammarly check}
-
-%\begin{Theorem} Graph \ref{Scomm_h} is exactly the inverse-limit system $(\hat{U}, \hat{T})$.    \end{Theorem}
-
-%We now have the following (which may be compared with \ref{SCausal_H} above):
-
-%\begin{equation} \label{fig:inverse_limsystem}
-  %  \[       \psset{arrows=->, arrowinset=0.25, linewidth=0.6pt, nodesep=3pt, labelsep=2pt, rowsep=0.7cm, colsep = 1.1cm, shortput =tablr}      \everypsbox{\scriptstyle}
-%      \begin{psmatrix}      \widehat{U}_T  && \widehat{U}_T \\     Y_T && Y_T \\
-      %%%
-     %  \ncline{1,1}{1,2}^{\widehat{T}} \ncline{1,1}{2,1} <{h}
-     %  \ncline{1,2}{2,2} > {H_2}
-     %  \ncline{2,1}{2,2}^{G_T}
-%      \end{psmatrix}
-    %  \]
-%  \end{equation}
-
-%\begin{Theorem}
- %   $(Y_T, G_T)$ is semi-conjugate to $(\widehat{U}, \widehat{T})$.
-%\end{Theorem}
-% {\bf Proof.} See \cite[Th.3, Th.4]{manunath2021universal}
-
-\section*{Summarising the discussing thus far:}
-
-It is easy to lose the birds-eye view, so we take a moment to review our progress up until this point.
-
-\vspace{-8mm}
-\begin{enumerate}
-\item We are interested in a some dynamical system $(U,T)$ with unknown dynamics.
-\item To determine properties about this system $(U,T)$ and predict its future evolution, we determine the dynamics of the inverse-limit system $(\widehat{U}, \widehat{T})$.
-% Given certain assumptions, we can guarantee that $(\widehat{U}, \widehat{T})$ is at least semi-conjugate to $(U,T)$.
-\item If the driven system $g$ is SI-invertible (and $\{u_n\}\in{U}$  is an orbit of $T$), the map $G_T$ exists that describes the single delay dynamics. 
-\item If, furthermore, $g$ has the USP, $(Y_T, G_T)$ is semi-conjugate to $(\widehat{U}, \widehat{T})$.
-\item If we can assume that $T$ is a homeomorphism, $(Y_T, G_T)$ is topologically conjugate to $(\widehat{U}, \widehat{T})$, an extension space of $(U,T)$
-\end{enumerate}
-
-One can therefore in practice learn the SDD of the driven system states via $G_T$ with enough data thanks to the USP/UAP. This enables us to do at least two things: 
-\vspace{-8mm}
-\begin{itemize}
-\item Forecast  $x_{n+1},x_{n+2}, \ldots$ via iterates of $G_T$ (if $G_T$ can be learnt).
-\item Forecast future values of $u_n$ since $x_n$ and $x_{n+1}$ determine $u_n$ since $g$ is SI-invertible. 
-\end{itemize}
-
-Finally, we note that although $G_T$ exists with SI-invertibility, we need the USP as well(see \ref{fig_pictorialSummary} below for a visual illustration).
-If not, the driven states would have to be running for all time since they would not have forgotten their past states.
-
-\begin{figure}[ht]
-  \includegraphics[scale=0.3]{_summarypictorial.eps}
-  \centering
-  \captionof{figure}{Pictorial summary of results obtained up to now and how they overlap. Reproduced from \textbf{(Cite SIG-25May)} }
-  \label{fig:fig_pictorialSummary}
-\end{figure}
-\ednote{Ask M: Cite SIG-25May.}
-
-\section{A discussion of $G_T$ }
-
-In the above sections, we established the map $G_T$ describing the SDD of a driven system.
-By establishing the existence of this map, we’ve essentially embedded the attractor $U$ (recall from \ref{attractor_U}) into the higher dimensional space $X\times{X}$.
-In layman’s terms, this ensures that there is more “dimensional room” for the underlying system’s underlying dynamics to “move”.
-Since the dynamics aren’t as “squashed”, we might therefore hope that the dynamics of $G_T$ are in some sense simpler than that of $T$.
-(Taken note of the fact that $G_T$ is a homeomorphism even when $T$ is just continuous)
-
-In \cite{manjunath2021universal} it is illustrated in an empirical fashion in that the map $G_T$ describes dynamics which are less functionally complex than that of $T$ or of the map $\Phi_{2d,\theta}$.
-This is done by implementing a Recurrent Neural Network (RNN), which is discussed in~\ref{ch5}.
-\ednote{B: Should I add in that we'll be using Pearson coefficient?}
-
-We opt to learn $G_T$ in an indirect manner by defining a new map $\Gamma:(x_{n-1},x_n)\mapsto{u_n}$. It will follow immediately from $G_T$’s existence that $\Gamma$ also exists.
-
-The reasons for taking this roundabout approach remain to be discussed in \ref{subs_LearnGamma}, but a pat answer may immediately be given: When $\Gamma$ has been learnt, the system can be driven autonomously and then $G_T$ is known anyway.
-We formalise this with a theorem.
-
-\begin{Theorem}
-  When $x_n$ and $x_{n-1}$ are successive points on a solution obtained for input-orbit of $T$ $\{u_n\}$, then the map $\Gamma: X\times{X}\to{U}$ defined by $(x_{n-1},x_n)\mapsto{u_n}$ exists whenever $G_T$ exists
-\end{Theorem}
-{\bf Proof.}  See \cite[Th. 3c]{manjunath2021universal}.
-
-Projection mappings $\pi_i$ are defined in the traditional meaning where a $k$-dimensional vector is projected to it's $i^{th}$ component such that $$\pi_i:(a_1,a_2, \ldotp, a_{k-1}, a_k)\to{a_i}$$
-
-The graph in equation \ref{fig:comm_H_CET} is then extended as below:
-
-\begin{equation}  \label{Scomm_imp}
-  %    \[ 
-      \psset{arrows=->, arrowinset=0.25, linewidth=0.6pt, nodesep=3pt, labelsep=2pt, rowsep=1.3cm, colsep = 1.1cm, shortput =tablr}
-   \everypsbox{\scriptstyle}
-   \begin{psmatrix}
-   \widehat{U}_T & \widehat{U}_T\\%
-   Y_T & Y_T\\
-   & \textcolor{red}{U \times X}
-   %%%
-  %  \ncline{1,1}{1,2}^{\widehat{T}} \ncline{1,1}{2,1} <{H_2}
-  %  \ncline{1,2}{2,2} > {H_2}
-  %  \ncline{2,1}{2,2}^{G_T}
-  %  \psset{linecolor=red}
-  %  \textcolor{red}{\ncline{2,1}{3,2}<{(\Gamma,\pi_2)}}
-  %  \ncline{3,2}{2,2}>{\textcolor{red}{(\pi_2,g)}}.
-   \end{psmatrix}
-  % \]
-  \end{equation}
-
-The problem finally simplifies to the issue of learning the map $\Gamma$ and combining this with the projection mapping $\pi_2$ and the function $g$, which will be known.
-A final set of equations is obtained - equations that have been entirely constructed from data.
-\begin{eqnarray}\label{eqns_from_data}
-	u_{k+1} &=& \pi_1 \circ (\Gamma, \pi_2) \circ (\pi_2,g) (u_k,x_k) \label{Seqn_u}\\
-	x_{k+1} &=& \pi_2 \circ (\Gamma, \pi_2) \circ (\pi_2,g) (u_k,x_k). \label{Seqn_x}
-\end{eqnarray}
-
-\section{Advantages of learning $\Gamma$} \label{subs_LearnGamma}
-
-One may immediately ask why we opt to take such a roundabout route; why not just learn the map $G_T$ from the get-go?
-On the surface, this seems to be an arbitrary decision path with no real reasoning, so we take a pause again and discuss the motivation for learning Gamma.
-There are a number of distinct advantages.
-
-In the first place, learning $\Gamma$ saves computational resources.
-This is due to the fact that the input $u_n$ lies in a lower-dimensional subspace in comparison to the high-dimensional space $X\times{X}$.
-
-%In practice, if the input is of a  lower dimension, one may easily embed it into the space $X\times{X}$ by padding the vector $u_n$ with zeroes.
-
-Secondly, the function $\Gamma$ is known to be stable.
-Learning $\Gamma$ makes use of equations \ref{eqns_from_data} which in turn employs the driven system $g$ possessing the USP, and offers distinct advantages with regards to stability in the presence of perturbation.
-It is desirable that inputs that differ only slightly would have only ‘slight’ effects on the output of the system.
-Moreover, one would hope to prevent numerical errors originating from input and measurement noise (see~\cite[Th.
-5]{manjunath2021universal}).
-As the data fed into the system is a function of both the input given and the parameters of the system itself (i.e.
-the value a, $\alpha$ in $\tanh$), we exploit both input-related and parameter-related stability.
-Both these notions are defined by way of the continuity of an encoding map of an input and design parameter respectively in \cite{manjunath2020stability}.
-
-Informally, we may consider input-related stability as concerned with the question of whether or not small variations in input would result in small responses.   
-If two inputs $\bar{u}$ and $\bar{u}_{noisy}$ are close in the product topology, then their tails could differ greatly due to the metric that generates the product topology has insensitivity to the differences in sequence's tails.
- However if the system has the USP then the function $h$ is continuous and so $h(\bar{u}), h(\bar{u}_{noisy})$ remain close-by for small levels of noise maintaining a measure of robustness to input and measurement noise.\cite{manjunath2021universal}.
-In a similar fashion, parameter-related stability is obtained thanks to a result \cite[Lemma 3.2]{manjunath2020stability} relating this form of stability with the ESP (which is here equivalent to the USP).
-
-% Secondly,  due to $g$ possessing the USP, both input- and parameter-related stability~\cite{manjunath2020stability} is obtained, which in turn prevents numerical errors that caused by input and measurement noise (see~\cite[Th.
-5]{manjunath2021universal}).
-
-Thirdly, learning a chaotic map on a set(here $Y_T$) with even a small measure of noise  means that the learnt system accepts arguments out of the set (Problems similar to Takens, cf.~\ref{sect_takenslimits}).
-We thus have a map $G_T^+$ acting on $Y_T^+$ and since $Y_T$ is not guaranteed to be an attractor of $G_T$ this could lead to errors.
-If one learns $G_T$ directly, we are in danger of replicating the problems arising for Takens.  To see this, consider Fig.~\ref{fig:YtGtFailure}.
-
-\begin{figure}[ht]
-  \includegraphics[scale=0.25]{_YTerrors.eps}
-  \centering
-  \captionof{figure}{Diagram of potential errors developing when $G_T$ is learnt.}
-  \label{fig:YtGtFailure}
-\end{figure}
-
-To circumvent this: if a driven system with a saturation function\ednote{B: What is a saturation function?} is used, we can ensure the existence of an attractor $A$ of the map which implements $G_T$, and containing $Y_T^+$ indirectly (via $g$).
-\ednote{B: A little vague.
-Are we saying that an attractor of a map implementing $G_T$ is also the one that contains $Y_T^+$?}
-By defining the dynamics to $A$, this prevents large errors from occuring.  
-More details on the global dissipativity of $g$, as it is termed there, may be found in~\cite{Supp}.
-
-\begin{figure}[ht]
-  \includegraphics[scale=0.25]{_autovsdriven.eps}
-  \centering
-  \captionof{figure}{Illustrating empirically the importance of global dissipativity.
-Plotted below are the three principal components of randomly initiated points(blue) of a learnt version of $G_T$ versus the trajectory $(x_{n-1}, x_n)$(red) of driven states obtained from the Fractal Dream Attractor's data~\ref{Thomas_Attractor}).
-}
-  \label{fig:learningFailure}
-\end{figure}
-
-In effect, global dissipativity prevents large numerical errors due to the noise in input data from arising (see Fig.~\ref{fig:learningFailure}). 
-If a system is not globally dissipative, neglibly small errors could lead to major inaccuracy which thus induce predictions to utterly fail.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\chapter{Implementation} \label{ch5}
-
-In this chapter we turn to the implementation of the function $G_T$ by discussing a method to learn the function $\Gamma$.
-We apply our methodology to some chaotic attractors.
-Most notably, we consider the reconstruction of the chaotic attractor of the double pendulum - an  attractor that has not been successfully reconstructed because of its intricate structure.
-In fact the previous data-driven approaches also have not been successful in building a model from the double pendulum.
-
-%We then indicate where this implementation differs from previous work done by (\cite{manjunath2021universal}) and empirically show that the SDD of $G_T$ are less functionally complex than $\Phi_{k, \theta}$ or $T$ by considering the Pearson correlation coefficient.
-Finally, we present and discuss some numerical results for a focus system of this project: the double pendulum.
-We also consider other chaotic attractors.
-
-\section{Implementing $G_T$}
-The map $G_T$ is implemented by applying the functions $\Gamma$, $g$ and the projection function $\pi_2$ in the relevant order as described by the equations~\ref{eqns_from_data}.
-To reach this point, one must first learn $\Gamma$.
-After first normalising the data to ensure that it has zero mean value and lies within the range $[-1,1]$ and required zero-padding(see paragraph below for further discussion), we continue by driving the system with external input through the function $g$.
-
-From \eqref{eqn_driving} we know that $g(u,x) = (1-a) x + a \tanh(Au + \alpha B x)$. Recall that $g$ is SI-invertible. We refer to $A$ as the input matrix and $B$ the reservoir matrix.
-
-We make use of a Recurrent Neural Network (RNN) in our implementation to project the data on to a higher dimensional space, but we do not train the RNN in the conventional manner.
-\ednote{B: Should we say what the conventional manner is?} A few words about the RNNs: RNNs are widely used in the literature for temporal processing of data as a type of deep learning technique.
-They take the form of equations considered to be discrete-time version of continuous dynamical systems modelling a neural network of neurons \cite{jaeger2001echo}.
-
-We implement a feedforward network, (a network with no memory, unlike the RNN)\ednote{B: Expanded sentence s.t.
-FFN is in more context} required to learn $\Gamma$ in the \emph{Python} programming language and making use of the \emph{Scipy} library, alongside the \emph{Keras} package in the \emph{Tensorflow} library \ednote{B: Is this something you want me to write in?
-M: We do not implement the RNN, but the FN that learns Gamma through the Tensorflow.
-We called the RNN + FN set up a RCN in Adriaan's paper.
-FYI: FN's work only if there is a definite relationship between the $u_n$ and $u_{n+1}$ which is rare in temporal data.
-RNNs with ESP work when there is a definite relationship between $(\ldots,u_{n-1},u_n)$ and $u_{n+1}$ and is suited for temporal processing.}.
-
-If the RNN we will be using in the next step has $N$ neurons, then $A$ and $B$ are each $N\times{N}$ matrices.
-Inputs $u_n$ of dimension $K < N$ are padded by zeroes (as remarked earlier) to embed the input in a space of dimension $N$.
-As we do not need the entire left-infinite history of input, an arbitrary initialisation of the system with initial values followed by the driving of the system with the function $g$ means that the generated values will approach the actual solution elements in a uniform manner thanks to the UAP that we had discussed in the preceding chapter.
-
-Thirdly we prepare the network.
-Utilising a network $N$ neurons (the same value $N$ as in the dimensions of the matrices $A,B$).
-Learning of $\Gamma$ is done using the Adam Optimiser where the means square error (MSE) loss function is optimised.
-The system is trained several times (in our case 4) with incrementally smaller learning rates on each run (0.001 and thereafter divided by 10 at each run).
-Additional parameters such as the number of network-layers, neurons per layer and training-length are presented case-by-case below.
-
-We then reach the prediction step. Making use of $g$'s SI-invertibility and $g$ as defined in \eqref{eqn_driving}, we have the following expression for $u_n$ replicated from~\eqref{eqn_SI_RNN}.
-\begin{equation*}
-  u_n := A^{-1}\bigg(\overline{\tanh}^{^{-1}}\frac{1}{a}\Big(x_{n+1}-(1-a)x_n\Big) \bigg) - \alpha B x_n
-\end{equation*}
-
-The two model equations for learning from data~\eqref{eqns_from_data} allows us to determine what the state $x_{n+1}$ at time $n+1$ will be. 
-We specifically remind ourselves of \[x_{k+1}=\pi_2 \circ (\Gamma, \pi_2) \circ (\pi_2,g) (u_k,x_k)\]
-
-\section{Delay Coordinates}
-
-We also consider a more general problem of learning an attractor of dynamical system when only observations of an orbit are given.
-Explicitly, if $(W,T)$ is a dynamical system with dynamics generated by $w_{n+1}=Tw_n$, and if $\theta:W \to \mathbb{R}$ is an observable, then the task is to learn a dynamical system that is topologically conjugate to $(W,T)$ and predict $\theta(w_{m+1}),\theta(w_{m+2}),...$ using the data $\theta(w_{0}),\theta(w_{1}),\ldots,\theta(w_{m})$.
-So how do we do it?
-Suppose the input generated from the delay-coordinate map $\Phi_{\theta,2d}(\theta(w_{n})) := (\theta(w_{n-2d}),\ldots,\theta(w_{n-1}),\theta(w_{n}))$ is such that Takens delay embedding theorem (see Theorem~\ref{Thm_Takens}) holds, in which case, there exists a homeomorphism  $F_\theta: \Phi_{\theta,2d}(\theta(w_{n})) \mapsto \Phi_{\theta,2d}(\theta(w_{n+1}))$.
-Then if we feed  the input values $u_n := \Phi_{\theta,2d}(\theta(w_{n}))$ to a driven system $g$ that is SI-invertible and has USP, the induced system $(Y_F,G_F)$ would be topologically conjugate to the inverse-limit space of
-$(\Phi_{\theta,2d}(W), F_\theta)$ as in Theorem~\ref{Thm_CET}, and hence one could forecast $u_n,u_{n+1},\ldots$.
-In other words, one could forecast  the values $\theta(w_n), \theta(w_{n+1}),\ldots$ by feeding these observations as delay coordinates.
-The advantage of feeding delay-coordinates to a driven system is that the embedding is stable in the sense of global dissipativity that we have described in the previous chapter, i.e., $Y_F$ is made to contained in an attractor of a mapping induced on $Y_F$ that is induced by $g$.
-
-We note that the twin equations together~\ref{eqns_from_data} generate a model from data.
-For a chaotic system, a small amount of noise due either to the input-noise or computational error would cause the solutions from two accurate models with the same initial conditions to  diverge as they are not identical.
-\ednote{B: edited this sentence} However, statistically, they would be the same since due to ergodic theory all typical orbits have the same visiting frequency to any region of the phase space of the dynamical system.
-\ednote{B: I'll need to recap this sentence with you before presenting as it's new to me.} To illustrate that we plot the densities of the trajectories of the actual and predicted solutions.
-
-Our implementation of the reservoir(A) and input(B) matrices in the RNN makes use of sparse matrices.
-We discuss three specifics relating to this concept.
-Firstly, the matrices are assigned a variable density in terms of the amount of non-zero entries and are scaled to have values within the range $[-1,1]$.
-For this, we make use of \textit{Python's SciPy} library and specifically the \textit{Sparse, LinAlg} modules.
-In our experiments, we generally opted for a density ratio of 10\%-25\%.
-Secondly, we opted to enforce the requirement that both matrices be well-conditioned, i.e.
-that they have a condition number of \~1.
-The condition number of a matrix $M$ (denoted $cond(M)$) is the product of the matrix norms of the matrix $M$ and its inverse $M^{-1}$ and is a measure of the sensitivity of output of a matrix multiplication when an input is changed.
-\textbf{(Cite.)}\ednote{Ask M: Cite specific resource?} By ensuring a condition number close to 1, we also ensure that our driven function $g$, depending on the matrices $A$ and $B$, will not be sensitive to adjustments in the input.
-In so doing we then also guarantee another form stability.
-\ednote{B: Not complete enough.
-Edit further and cite}
-
-Lastly, we ensure the matrix $B$ has unit spectral radius.
-This is necessary due to the Causal Embedding Theorem~\ref{Thm_CET} which can then guarantee the existence of semi-conjugacy between the systems $(Y_T, G_T)$ and $(\widehat{U}_T, \widehat{T})$.
-In fact, the theorem actually requires that $\alpha{B}$ have spectral radius $<1$, so we impose the restriction $\alpha<1$ and consider only the matrix $B$'s spectral radius from thereon.
-
-Note that to learn $\Gamma: (x_{n-1},x_{n}) \mapsto u_n$ through our feedforward network, we could learn the principal components of $(x_{n-1},x_{n})$ as well. 
-We employ this since it provides slightly better results empirically but is not essential.
-
-Principal components are used only for a more efficient state representation to possibly reduce learning errors; it is not for a lossy approximation since we use all principal components \textbf{(Cite.)}\ednote{Ask M: Is there a specific resource to cite?}.
-More explicitly, denote $X_{1:N}$ as the matrix with the first $N$ states of the network data represented by row vectors.
-If $X_{1:N}=U\Sigma P^T$\ednote{Expand here} represents the singular value decomposition of $X_{1:N}$ \ednote{Ask: Is there a specific resource to cite?}, we then denote the principal component matrix $P$, and the principal components are given by
- $Z_{1:N}=X_{1:N}P.$
-These principal components are then used to train a feedforward neural network. Denote the row vectors of $Z_{1:N}$ by 
-$z_i^T, i=1,2,\cdots,N$ and the neural network by $NN$, the network is subsequently trained to learn an approximation of the map
-$NN: (z_{n-1},z_n) \mapsto  u_n.
-$
-The learnt approximation of $NN$ can be used to approximate $\Gamma$ since 
-$$
-NN\left( \begin{bmatrix} 
-P^Tx_{n-1} \\
-P^Tx_n
-\end{bmatrix}
-\right) = NN \circ 
-\begin{bmatrix}
-P^T & 0 \\
-0 & P^T 
-\end{bmatrix}\begin{bmatrix}
-x_{n-1}\\
-x_n
-\end{bmatrix} = u_n = \Gamma(x_{n-1},x_n).
-$$
-
-\section{Experimental Results}
-In this section we showcase numerical results pertaining to 3 attractors that were simulated using the methodology described.
-We first consider a simple physical system that exhibits complicated dynamics.
-\subsection{Double Pendulum}
-
-The Double Pendulum consists of two pendulums fastened to one another such that the system moves as a whole.
-The first (or higher pendulum) is attached to a fixed point while the second is attached to the endpoint of the first.
-(See below)
-This system is a classical example of a system exhibiting chaos and is often used as explanation of the very well-known Butterfly Effect.
-\ednote{B: Originally though a reference here was needed, but no longer think so.} Consider again~\ref{fig:dp_sdic}.
-The system has four variables: angle and angular velocity of each pendulum rod. An ideal system is assumed where the rods are massless and no friction in the pivots are considered.
-\begin{figure}[ht]
-  \includegraphics[scale=0.35]{_dp_setup.eps}
-  \centering
-  \captionof{figure}{Pictorial representation of the setup of a DP. Image taken from~\cite{DPSetup}}
-  \label{fig:dp_setup}
-\end{figure}
-
-The scalar equations of motion are derived using Newtonian physics and are reproduced from~\cite{DPFormulas} in~\ref{eqns_dp}.
-Each pendulum component has an associated mass of the head, $m$, and length of the rod, $L_i$.
-Acceleration due to gravity is denoted by $g$ and initial velocities by $\dot{\theta}$
-\begin{eqnarray}\label{eqns_dp}
-  \ddot{\theta_{1}}  = \frac{-g(2m_1+m_2)\sin{\theta_1} - m_2g\sin(\theta_1-2\theta_2) - 2\sin(\theta_1-\theta_2)m_2({\dot{\theta_{2}}}^{2}L_2 + {\theta_{1}'}L_1\cos(\theta_1-\theta_2))} {L_1(2m_1 + m_2 -m_2\cos(2\theta_1 - 2\theta_2))}
-  \\
-  \ddot{\theta_{2}} = \frac{2\sin(\theta_1-\theta_2)(\theta_{1}'^{2}L_1(m_1+m_2) + g(m_1+m_2)\cos\theta_1 + \dot{\theta_{2}}^{2}L_2m_2\cos(\theta_1-\theta_2))}{L_2(2m_1 + m_2 -m_2\cos(2\theta_1 - 2\theta_2))}
-\end{eqnarray}
-
-The second-order system is converted to a first-order ODE system and solved using the Runge-Kutta numerical method.  
-Here we chose the values $m_1=2.0$, $m_2=2.0$, $L_1=1.5$ and $m_2=1.5$ with initial velocities $v_1=1.5, v_2=-2$ for the respective pendulum heads.
-Undamped DP data-sets were generated and used in this project due to the requirement that the map $T$ be surjective that we must not forget. 
-An example of this would be the case of a DP exhibiting energy-loss (i.e.
-where the map $T$ is not surjective).
-In such a case, the attractor is a single point in the phase space when the double pendulum comes to rest.
-It must be mentioned once again that if we use data outside the attractor, learning is not reliable. To see this, consder the graphs ~\ref{fig:damped_pendulum} and ~\ref{fig:dp_notsurjective}.
-
-\begin{figure}[ht]
-  \includegraphics[scale=0.4]{_dp_dying.eps}
-  \centering
-\caption{$x$- and $y$-coordinates of the two pendulum heads plotted for a dataset that exhibits dying oscillations.
-The map generating the (four-dimensional data) for a damped pendulum is not surjective.
-Data obtained from chaotic double pendulum dataset generated from videos takes of experimental pendulums~\cite{asseman2018learning}}
-\label{fig:damped_pendulum}
-\end{figure}
-
-\begin{figure}[ht]
-  \includegraphics[scale=0.3]{_dpfail_nonsurj.eps}
-  \centering
-\caption{Failure to forecast or learn a map which is not surjective: Evolution of the $y$-coordinate of the bottommost pendulum plotted here for 15000 timestep.
-Plotted here the predicted value(red) versus the actual value(blue).
-The plots overlap for the first 10000 steps, the duration of training to learn the map $\Gamma$.
-Afterwards, the trajectories start diverging and it the predicted value no longer accurately represents the complexity or energy-loss of the actual data.
-}
-\label{fig:dp_notsurjective}
-\end{figure}
-
-Having spent ample time to discuss the challenges, we now change tone so as to satisfy the optimist and present a successful run of the DP-data to predict the future movement of the pendulum.
-
-The system was driven for 15500 steps, of which the first 500 were immediately discarded to simulate the network 'forgetting'.
-The driving function $g$ had input and reservoir matrices were implemented as sparse matrices filled to 20\%.
-Parameter-values of $\alpha=0.99$ and $a=0.5$ were used.
-The remaining 15000 datapoints were used to train a network with 128 neurons. 
-The network was initialised to contain 16 hidden layers each with a layer dimension of 64, activated with the tanh activation function.
-Training was accomplished using 256 epochs with batch sizes 128 and a delay of 1.
-See~\ref{fig:dp_success_traj} for results pertaining to the trajectory taken by the DP's components and then~\ref{fig:dp_success_density} to witness associated densities of predicted vs.
-actual values.
-
-\begin{figure}[ht]
-  \centering
-  \includegraphics[width=0.95\linewidth]{_dp_success_4coords_traj.eps} 
-  \captionof{figure}{Predicted(red) and actual(blue) trajectories of the $x$- and $y$-coordinates of each of the pendulum heads.
-From top to bottom are as follows: $x$-coordinate of topmost pendulum head, $y$-coordinate of topmost pendulum head, $x$-coordinate of lower pendulum head, $y$-coordinate of lower pendulum head.
-  These graphs were constructed by predicting the DP 1000 timesteps into the future and in so doing illustrating the long-term consistency and accuracy of the learnt system.
-Here we lock onto the near-exact trajectory for about 300 timesteps and stay close up until about 600 seconds.
-}
-  \label{fig:dp_success_traj}
-\end{figure}
-\begin{figure}[ht]
-  \centering
-  \includegraphics[width=\linewidth]{_dp_success_4coords_hist.eps}
-  \captionof{figure}{Densities of predicted and actual values of $x$- and $y$-coordinate of pendulum heads.}
-  \label{fig:dp_success_density}
- \end{figure}
-
-We now consider an experiment where noise is added to the system and find it to be incredibly robust.
-We employ the same parameters as before and add noise from a normal-distribution with mean zero and standard deviation equal to 0.1 which translates to roughly 39dB's of noise.
-(For the signal-to-noise ration we adopt the formula $\text{SNR}_{dB}=10\text{log}_{10}\big(\frac{P_\text{{signal}}}{P_{\text{noise}}}\big)$ where $P_\text{{signal}}$ refers to the variance observed in the normalised data and $P_{\text{noise}}$ represents the variance of the added noise.)
-
-Once again consider the actual and predicted trajectories for the noisy DP dataset in Figure~\ref{fig:noisydp_success_traj}.
-Next we consider the densities of the predicted and actual values in Figure~\ref{fig:noisydp_success_density} and observe that they display highly similar distributions.
-
-\begin{figure}[ht]
-  \centering
-  \includegraphics[width=0.95\linewidth]{_dp_noise_FigCD_2030.eps} 
-  \captionof{figure}{Predicted(red) and actual(blue) trajectories of the $x$- and $y$-coordinates of each of the pendulum heads with noise added from a normal distribution with mean zero and standard deviation 0.1, equating to approximately 39dB's of noise.
-  These graphs were constructed by predicting the DP 1000 timesteps into the future and in so doing illustrating the long-term consistency and accuracy of the learnt system.
-Here we lock onto the near-exact trajectory for about 250 timesteps and stay close up until about 400 seconds.
-}
-  \label{fig:noisydp_success_traj}
-\end{figure}
-\begin{figure}[ht]
-  \centering
-  \includegraphics[width=\linewidth]{_dp_noise_2B_2030.eps}
-  \captionof{figure}{Densities of predicted and actual values of x-coordinate of first pendulum head.}
-  \label{fig:noisydp_success_density}
- \end{figure}
-
- Lastly, a run was done where only a single coordinate (the $y$-coordinate of the lower pendulum head) was fed into the system and we can illustrate through the obtained success that the coordinate contains all the information in its past states necessary to forecast its future evolution.
- \textbf{Insert graphs here. Had to quickly change something here.}
-
-\subsection{Additional Attractors: Clifford \& Thomas}
-We follow the methodology presented in \cite{manjunath2021universal} but opt to consider attractors different from the Lorenz system and H\'enon map in this project.
-
-\subsubsection{Thomas' Cyclically Symmetric Strange Attractor}\label{Thomas_Attractor}
-Thomas' Cyclically Symmetric Strange Attractor, a 3D attractor proposed in 1999 by Ren\'e Thomas in \cite{ThomasAttractor}, is described by a set of three equations which is distinctly reminiscent of the Lorenz system, another deterministic three-equation model exhibiting chaotic behaviour.It has a single parameter $\beta$ and has been shown to transition to chaotic behavious when $\beta<0.208186$~\cite{Thomas_BetaParameter}.
-\begin{eqnarray}\label{eqns_thomas}
-  x_{n+1} = \sin(y_n) - \beta{x_n} \\
-  y_{n+1} = \sin(z_n) - \beta{y_n} \\
-  z_{n+1} = \sin(x_n) - \beta{z_n}
-\end{eqnarray}
-
-The behaviour of the system with a $\beta$-parameter value of 0.1056 was simulated by scaling it to fit inside the interval $[-1,1]$.
-Two cases were considered: noise-free and the one perturbed by noise.
-Noise was generated from a normal distribution with mean 0 and standard deviation 0.05, which to translates approximately 33dB.
-($\text{SNR}_dB$ is calculated exactly as for the DP-data above)
-
-A sequence of 3000 observations in both the clear and noisy data-sets respectively were fed into the system with 300 entries discarded to simulate the network's loss of memory.
-The next 9000 steps were predicted and compared with true data.
-
-\begin{figure}[ht]
-  \centering
-  \minipage{0.5\linewidth}
-    \centering
-    \includegraphics[width=0.8\linewidth]{_thomas_noisy_0.1056.eps}
-  \endminipage\hfill
-  \minipage{0.5\linewidth}
-    \centering    \includegraphics[width=0.8\linewidth]{_thomas_clean_0.1056.eps}
-  \endminipage
-  \captionof{figure}{True(red) and predicted(blue) trajectories for the Thomas attractor with parameter value $\beta=0.1056$.
-To the left: Noise added from normal distribution with mean zero and standard deviation 0.05.
-Right: No noise added to dataset.}
-\end{figure}
-
-The RNN was initialised as follows: the dimension of the network is set equal to 500 neurons and 12 hidden layers of dimension 64 each were initialised.
-Training is realised through 150 epochs of batch sizes of 128 each; the map is learnt via the Adam Optimiser using finer and finer learning rates (0.001 and then divided by 10 at every iteration).
-
-\begin{figure}[ht]
-  \centering
-  \includegraphics[scale=0.35]{_Thomas_1.eps}\caption*{Predicted trajectories of the $x$- and $y$-coordinates for the Thomas attractor with parameter-value $\beta=0.1056$ demonstrate empirically the ability to predict the evolution of the trajectory for the next an estimated 100 timesteps into the future near-exactly.
-Here no noise was added.}
-  \includegraphics[scale=0.5]{_Thomas_3.eps}\caption*{Represented here the learnt (blue) and actual (red) densities of the first coordinate ($x$-coordinate) of the dataset for the Cyclically Symmetric Attractor with parameter-value $\beta=0.1056$.}
-\end{figure}
-
-\subsubsection{Fractal Dream Attractor}
-
-Another 2-equation discrete-time map named the Fractal Dream Attractor, or more commonly known as the Pickover Map (first discovered by Clifford A.
-Pickover and discussed in his fascinating book "Chaos in Wonderland"\cite{PickoverChaos}), was also considered.
-\begin{eqnarray}\label{eqns_clifford}
-  {x_{n+1}=\sin(ay_n) + c\cos(ax_n)} \\
-  {y_{n+1}=\sin(bx_n)+d\cos(by_n)}
-\end{eqnarray}
-
-\begin{figure}[ht]
-  \centering
-  \includegraphics[width=1.0\textwidth,left]{_Clifford_1_nonoise.eps}
-  \caption*{These graphs were constructed by predicting the Clifford system 1000 steps into the future and in so doing illustrating the long-term consistency and accuracy of the learnt system.
-As perceived here, we are able to lock on to the trajectory of the Clifford map almost exactly for the first 25 steps.}
-  \minipage{0.5\textwidth}
-      \centering
-      \includegraphics[width=\linewidth]{_Clifford_2_nonoise.eps}
-  \endminipage\hfill
-  \minipage{0.5\textwidth}
-    \centering
-    \includegraphics[width=\linewidth]{_Clifford_3_nonoise.eps}
-  \endminipage
-  \caption{The bottom row, left, shows a plot of the true attractor(red) and the predicted values(blue).
-Bottom row, right, shows the distribution of the first coordinate of the learnt system (blue) and the actual Clifford system (red).
-}
-  \label{fig:Clifford}
-\end{figure}
-
-The behaviour of the system with a parameter values of $x_0=0$, $y_0=0$, $a = -1.7$,  $b = 1.8$, $c = -1.9$ and $d = -0.4$ was examined.
-Data was scaled to fit inside the interval $[-1,1]$. Only the noise-free case was investigated. 
- A sequence of 25000 observationswas fed into the system with 15000 entries discarded to simulate the network's loss of memory. The next 1000 steps were predicted and compared with true data.
-The RNN was initialised exactly as for Thomas' Cyclically Symmetric Strange attractor: the dimension of the network is set equal to 500 neurons and 12 hidden layers of dimension 64 each were initialised.
-Training is realised through 150 epochs of batch sizes of 128 each; the map is learnt via the Adam Optimiser using finer and finer learning rates (0.001 and then divided by 10 at every iteration).
-
-\section{Functional Complexity Reduction in Noninvertible maps}
-
-% Besides being able to learn the action of the Koopman operator exactly, our setup has other advantages over selecting or optimizing the choice of observables as in the Extended Dynamic Mode Decomposition (EDMD) algorithm (see, [21, 22]).
-We can alter the functional complexity of G T by changing the number of observables or equivalently by altering the dimension of X – in the case of a RNN implementation of g, by simply changing the number of neurons in the network.
-Empirically, increasing the dimension of X is found to increase the linear relationship (or intuitively reduces the functional complexity of G T ) that is measured as a generalization of the Pearson correlation coefficient to random vectors (see, [44]) between (x_{n−1} , x_{n}) and G_T (x_{n−1} , x_n), and such numerical evidence is tabulated in [11, Table 1].
-In contrast, expanding the set of observables while using EDMD to learn a map with a lower functional complexity is a highly involved task, not least because so much that one often does not know how to guess a new observable, as well as the fact that adding observables does not necessarily retain the invariance of the span of the observables.
-
-% Empirical evidence shows that a linearity measure like a Pearson correlation coefficient points at reduced functional complexity while learning the dynamics in the state space of the driven system.
-
-% Empirically (see Table ??), increasing the dimension of X is found to increase the linear relationship
-% (or intuitively reduce the functional complexity of G T ) that is measured as a generalization of the Pearson correlation coefficient to random vectors (see, [?]) between (x n−1 , x n ) and G T (x n−1 , x n ).
-If Σ a and Σ b denotes the covariance matrices of the vectors (x n−1 , x n ) and G T (x n-1 , x n ) respectively, and if Σ ab denotes the covariance matrix between the vectors (x n-1 , x_n ) and G_T (x n-1 , x_n ), then the multidimensional 18correlation coefficient is computed by using the traces of these matrices:
-% \[\pho = \frac{tr(\sum_{ab})}{tr(\sqrt{\sum_{a}\sum_{b}})}\]
-
-% This multidimensional correlation coefficient satisfies most of the well-known properties of the one-dimensional Pearson coefficient.
-In particular, ρ = ±1 if and only Y = AX + b for some invertible matrix A and vector b.
-Hence ρ, and in practice, the estimator ρ̂, can be used to measure the linear relationship between two random vectors of the same dimension and thus serves as an indicator of functional complexity.
-
-% We remark that the estimation of such linear relationship in (see Table ??) obtained by sample correlations alone is justified whenever {u n } is a realization of an ergodic process since then when g has the USP, any solution {x n } is also a realization of an ergodic process [?].
-In the ergodic input case the generalized Pearson correlation coefficient (between (x n−1 , x n ) and G T (x n−1 , x n )) is independent of n.
-
-In the case of non-invertible maps, we note that the map $G_T$ is a homemorphism although $T$ is not.
-Further, empirically it is found that $G_T$ is much more linear than $T$ as measured by what is called the multidimensional correlation coefficient $\rho$.
-This coefficient is a generalization of the Pearson correlation coefficient for real-valued random variables to random vectors (see, \cite{puccetti2019measuring}).
-We state teh details of computing this coefficient between
-$(x_{n-1},x_n)$ and $G_T(x_{n-1},x_n)$. 
-If $\Sigma_{a}$ and $\Sigma_{b}$ denotes the covariance matrices of the vectors $(x_{n-1},x_n)$ and $G_T(x_{n-1},x_n)$ respectively, and if $\Sigma_{ab}$ denotes the covariance matrix between the vectors $(x_{n-1},x_n)$ and $G_T(x_{n-1},x_n)$, then the multidimensional correlation coefficient is computed by using the traces of these matrices:
-\[
-    \rho= \frac{\text{tr}({\Sigma_{ab}})}{\text{tr}({\sqrt{\Sigma_a\Sigma_b}})}.
-\]
-This multidimensional correlation coefficient satisfies one of the desired and well-known properties of the one-dimensional Pearson coefficient.
-That is, $\rho=\pm 1$ if and only if $Y\overset{d}{=}AX+b$ for some invertible matrix $A$ and vector $b$.
-Hence $\rho$, and in practice can be used  to measure the linear relationship between two random vectors of the same dimension and thus serves as an indicator of functional complexity.
-For the non-invertible system defined by the Clifford attractor, we indicate the difference by which $\rho$ changes for $G_T$ in Table~\ref{tbl_attractorsPearson}.
-Following \cite{manjunath2021universal}, we say that this multidimensional correlation coefficients $\rho$ to evidence a general reduction in the functional complexity of the map $G_T$ emerging due to a RNN.
-In the
-Table ~\ref{tbl_attractorsPearson}, the second column corresponds to the amount of artificial neurons (the dimension of $X$) in the implemented RNN.
-
-\begin{center}
-  \begin{table} 
-      \scalebox{0.75}
-      {\begin{tabular}{|c|c| c c |} 
-          \hline
-          Input & Dimension of $X$ & $u_n$ vs $u_{n+1}$ 
-          & $\begin{bmatrix}
-              x_{n-1}\\
-              x_n
-          \end{bmatrix}$ vs $\begin{bmatrix}
-              x_n\\
-              x_{n+1}
-          \end{bmatrix}$ \\
-          \hline      
-          % \hline
-          % $(w_n)$ is the Lorenz states, sampled every 0.1 timestamps. & & &\\
-          % \multirow{3}{*}{$u_n=\frac{1}{100}w_n$} & 10 &0.9311  & 0.9731\\
-          %     & 100 &  & 0.9934\\
-          %     & 1000 & & 0.9930\\              
-          % \multirow{3}{*}{$u_n=\frac{1}{10}(\sin(0.1w_{n,x})+\sin(0.1w_{n,y})+\sin(0.1w_{n,z}))$} 
-          %     & 10 &0.8401  & 0.9392\\
-          %     & 100 & & 0.9616\\
-          %     & 1000 & & 0.9737\\
-          % \hline
-
-               \hline  
-               $(w_n)$ comes from the Clifford Attractor \\ (for $a = -1.7; b = 1.8; c = -1.9; d = -0.4$): & & & \\
-        {$w_{n+1}= \begin{bmatrix} \sin(aw_{n,y}) + c\cos(aw_{n,x}) \\ 
-                                                                    \sin(bw_{n,x})+d\cos(bw_{n,y}) \end{bmatrix}$} & & & \\
-          \multirow{3}{*}{$u_n = w_n-\overline{w}$}
-              & 10 & -0.21713 & 0.78719 \\
-              & 100 & &  0.8334 \\
-              & 1000 & & 0.84854 \\
-          % \hline 
-          % $(w_n)$ comes from the Logistic map: & & & \\
-          % {$ w_{n+1}  = 4w_n(1-w_n)$} & & & \\
-          % \multirow{3}{*}{$u_n = w_n - 0.5$} 
-          %     & 10 &-0.0314  & 0.5577\\
-          %     & 100 &  & 0.7150\\
-          %     & 1000 & & 0.7826\\
-          % \hline
-          % \hline
-          % $(w_n)$ comes from the Svensson Attractor \\ (for $a = -1.7; b = 1.8; c = -1.9; d = -0.4$): & & & \\
-          % {$w_{n+1}= \begin{bmatrix} d\sin(aw_{n,x}) - \sin(bw_{n,y}) \\ 
-          %                                                             c\cos(aw_{n,x})+\cos(bw_{n,y}) \end{bmatrix}$} & & & \\
-          %   \multirow{3}{*}{$u_n = w_n-\overline{w}$}
-          %       & 10 & -0.2243 & 0.49698 \\
-          %       & 100 & &  0.63310 \\
-          %       & 1000 & & 0.71151 \\
-           \hline
-                \end{tabular}} %\ednote{http://paulbourke.net/fractals/peterdejong/}
- \label{Table_FC}
- \caption{Following \cite{manjunath2021universal}, we determine the multidimensional correlation coefficients $\rho$ to evidence a general reduction in the functional complexity of the map $G_T$ emerging due to a RNN.
-Rows correspond to distinct dynamical systems considered in the experiments
-      The second column corresponds to the amount of artificial neurons (the dimension of $X$) in the implemented RNN.
-      The last two columns comprise numerical estimates of $\rho$ for the relevant vectors. The linear association between $u_n$ vs $T(u_{n})$, as well as the linear relationship $\begin{bmatrix}
-              x_{n-1}\\
-              x_n
-          \end{bmatrix}$ vs $G_T \left( \begin{bmatrix}
-              x_{n-1}\\
-              x_n
-          \end{bmatrix}\right)$ may be contrasted by the reader. 
-          Finally, the RNN's dimension (from \eqref{Seq_RNN}) is provided to exhibiting that an increase in dimension of the driven system will 
-          typically result in a map $G_T$ with less functional complexity.}.
-      \end{table}\label{tbl_attractorsPearson}
-\end{center}
-
-\chapter{Conclusion}\label{ch6}
-\input{chapter6.tex}
-
-\bibliographystyle{pnas-new}
-\bibliography{pnas}
-
-\end{document}

### 2022-01-17 projectReport/chapter3.tex
Cleaner setup

--- 2022-01-17 draft_16Jan/hiscurrent.tex
@@
-\documentclass[a4paper,12pt,twoside]{report}
-% \documentclass[a4paper,12pt,twoside]{book}
-\pagestyle{headings}
-\usepackage[utf8]{inputenc}
-\usepackage{filecontents}
-\usepackage[many]{tcolorbox}
-\usepackage{graphicx}
-\usepackage{harpoon}
-%\usepackage{apacite} 
-\usepackage{xcolor}
-\usepackage{latexsym}
-\usepackage{amssymb}
-\usepackage{amsmath}
-\usepackage{mdframed}
-%\usepackage[authoryear]{natbib}
-\usepackage{hyperref, url}
-\usepackage[show]{ed}
-\usepackage{subfigure}
-\usepackage[ntheorem]{empheq} 
-\usepackage{amsthm, amssymb, amsfonts, latexsym}
-% \usepackage{pst-all}
-
-\usepackage{enumitem}
-\usepackage{multirow}
-\usepackage{longtable}
-\usepackage[export]{adjustbox}
-% $f\arrowvert_A$ To restrict functions
-
-\setlength{\parskip}{1.7em}
-\setlength{\parindent}{0em}
-
-%\setlength{\baselineskip}{20pt}
-
-\setlength{\textwidth}{429.75499pt}
-\setlength{\textheight}{643.20255pt}
-\setlength{\oddsidemargin}{5 mm}
-\setlength{\evensidemargin}{5 mm}
-\setlength{\topmargin}{0 mm}
-\setlength{\headsep}{0 mm}
-\setlength{\headheight}{0 mm}
-
-\usepackage{pst-node}
-
-\newcommand{\citealtt}[1]{\citeauthor{#1},\citeyear{#1}}
-\newcommand{\myycite}[1]{\citep{#1}}
-
-\mathchardef\mhyphen="2D
-
-%\newcommand{\cev}[1]{\reflectbox{\ensuremath{\vec{\reflectbox{\ensuremath{#1}}}}}}
-
-\usepackage{pst-node,graphicx,pst-blur}
-%\uspackage{auto-pst-pdf}
-%\usepackage{tikz-cd}
-
-\makeatletter
-\DeclareRobustCommand{\cev}[1]{%
-  \mathpalette\do@cev{#1}%
-}
-\newcommand{\do@cev}[2]{%
-  \fix@cev{#1}{+}%
-  \reflectbox{$\m@th#1\vec{\reflectbox{$\fix@cev{#1}{-}\m@th#1#2\fix@cev{#1}{+}$}}$}%
-  \fix@cev{#1}{-}%
-}
-\newcommand{\fix@cev}[2]{%
-  \ifx#1\displaystyle
-    \mkern#20mu
-  \else
-    \ifx#1\textstyle
-      \mkern#20mu
-    \else
-      \ifx#1\scriptstyle
-        \mkern#26mu
-      \else
-        \mkern#26mu
-      \fi
-    \fi
-  \fi
-}
-
-\makeatother
-
-\newcommand{\xcm}{\epsfxsize=3.1cm}
-\newcommand{\fig}[1]{\epsfbox}
-% \newcommand{\bp}{\begin{minipage}{3.1cm}}
-% \newcommand{\ep}{\end{minipage}}
-
-\newtheorem{Definition}{Definition}[]
-%\newmdtheoremenv{Theorem}{Theorem}[]
-\newtheorem{Theorem}{Theorem}[]
-\newtheorem{Lemma}{Lemma}[]
-\newtheorem{Proposition}{Proposition}[]
-\newtheorem{Corollary}{Corollary}[]
-\newtheorem{Remark}{Remark}[]
-\newtheorem{Example}{Example}[]
-
-%Shortcut symbols
-\newcommand{\Ftheta}{\ensuremath{F_\theta}}
-
-\makeatletter 
-\renewcommand{\thefigure}{\@arabic\c@figure}
-\makeatother
-\usepackage{xr}
-
-%%%% PNAS GRAPHICS SETUP- FROM THE PNAS STYLE LATEX FILE
-\RequirePackage{graphicx,xcolor}
-\RequirePackage{colortbl}
-\RequirePackage{booktabs}
-\RequirePackage{algorithm}
-\RequirePackage[noend]{algpseudocode}
-\RequirePackage{changepage}
-\RequirePackage[twoside,%
-				letterpaper,includeheadfoot,%
-				layoutsize={8.125in,10.875in},%
-                layouthoffset=0.1875in,%
-                layoutvoffset=0.0625in,%
-                left=38.5pt,%
-                right=43pt,%
-                top=43pt,% 10pt provided by headsep
-                bottom=32pt,%
-                headheight=0pt,% No Header
-                headsep=30pt,%
-                footskip=25pt,
-                marginparwidth=38pt]{geometry}
-\RequirePackage[labelfont={bf,sf},%
-                labelsep=period,%
-                figurename=Fig. ]{caption}
-\setlength{\columnsep}{13.5pt} % Distance between the two columns of text
-\setlength{\parindent}{12pt} % Paragraph indent
-
-%% Figure caption style
-\DeclareCaptionFormat{pnasformat}{\normalfont\sffamily\fontsize{7}{9}\selectfont#1#2#3}
-\captionsetup*{format=pnasformat}
-
-%%% GREYBOX AROUND FIG
-\definecolor{lightgray}{gray}{0.95}
-\newcommand\greybox[1]{%
-  \vskip\baselineskip%
-  \par\noindent\colorbox{lightgray}{%
-    \begin{minipage}{\textwidth}#1\end{minipage}%
-  }%
-  \vskip\baselineskip%
-}
-
-\newtcolorbox{paperbox}[1][]{
-    enhanced,
-    colback=white,
-    boxrule=0pt,
-    boxsep=0pt,
-    arc=0mm,
-    width=0.8\linewidth,
-    fuzzy shadow={0mm}{-4pt}{-4pt}{1mm}{black!30!white},
-    #1
-}
-
-% \externaldocument{Supp_Koopman-0808}
-\begin{document}
-
-\title{Data-Driven Modelling \& \\ Forecasting of Chaotic Systems}
-
-\author{B.L. Nortier} 
-\date{January, 2022}
-\date{}
-\maketitle
-
-\begin{abstract}
-  It is a well-established practice to collect measurements from underlying artificial, natural, and physical systems through observables in order to realise a dynamical model with greater descriptive power.
-  Many data-driven approaches to forecast the future evolution of systems exist.
-However, many of these methods rely on an initial understanding of the system and we often do not fully understand the many complexities of more intricate systems that we interact with on a daily basis.
-
-  In this project, the base assumption is that data is originates from  some discrete-time dynamical system, particularly one of a chaotic nature but no assumptions are made pertaining to a specific dynamical system.
-  A general learning problem is formulated and a well-known approach to the problem, alongside its practical limitations, is then considered.
-  The notion of driven dynamical systems is introduced, and we subsequently establish the theoretical existence of a dynamical system possessing certain properties which guarantees it to be topologically conjugate (`dynamically equivalent`) to the dynamics of the underlying system from which measurements were originally taken.
-  Our methodology is implemented practically by making use of recurrent neural networks, a form of deep learning, to learn a map topologically equivalent to the underlying system that forecasts future states.
-  In so doing we present the workability of an approach to obtain long-term topological and statistically consistent predictions for simple physical systems as well as a some chaotic attractors.
-  \end{abstract}
-
-  \tableofcontents
-
-\chapter{Introduction}\label{ch1}
-
-Experimenting with biological, physical, and artificial systems to generate a more informative dynamical model is a well-established practice in modern science.
-Traditional methods for modelling physical systems are based on laws of physics that are based either on empirical relationships or intuition.
-For systems that evolve with time, physical laws yield mathematical equations that govern how the quantities evolve with time.
-However, our world is much more complex than that which can be distilled into elegant equations.
-We do not fully understand many of the more complex systems, nor do they provide us with a good physical intuition of the underlying principles governing system dynamics.
-To complicate this, the underlying systems we observe often display sensitive dependence on initial conditions despite having highly similar initial conditions.
-This is because differing orbits diverge quickly and to such an extent that it becomes seemingly impossible to retrace their steps back to the original conditions.
-Even the presence of usually `negligible' computational noise or measurement error renders long-term pointwise prediction infeasible.
-We encounter many difficultiesin the process of modelling such systems:
-\vspace{-8mm}
-\begin{enumerate}[noitemsep, label=\roman*.]
-  \item One may not have access to the complete states of the systems
-  \item The system may be described by functions that behave wildly because their graphs have a wild oscillatory behaviour, i.e., a large functional complexity.
-\end{enumerate}
-
-Models derived primarily from data can be classified into three categories: 
-\vspace{-8mm}
-\begin{enumerate}[noitemsep, label=\roman*.]
-  \item  Interpretable models (i.e., they establish relationships between internal physical mechanisms), 
-  \item Partially interpretable models capturing some modes of the dynamics, 
-  \item Non-interpretable models, defined as such mainly due to the fact that they are defined on a different phase space that is usually high-dimensional. 
-\end{enumerate}
-
-Examples in the literature attempting to forecast data from such systems have tried many different approaches with differing degrees of success. 
-When we have complete access to states of the system, an ordinary differential equation can be obtained from data (\cite{brunton2016discovering, champion2019data} and \cite{small2002modeling,xu2006modeling}) wherein one could approximate the vector field by a library of functions to obtain interpretable models.
-
-Recent partially interpretable models available in the literature have been based on the Koopman operator (see~\cite{koopman1932dynamical,budivsic2012applied}) to employ observables mapping the data onto a higher-dimensional space.
-This makes the dynamics in the higher-dimensional space more amenable for approximation by a linear transformation.
-Such methods do not guarantee exact reconstructions for nonlinear models, and in practice provide poor long-term consistency for a large class of chaotic dynamical systems \cite{Supp}.
-
-The non-interpretable models include the delay embedding and the machine learning algorithms.
-For example, one could learn a system conjugate to the underlying system by applying the Takens delay embedding method of delay-coordinates~\cite{takens1981detecting} when one has "good" observations from the system; this learnt system could then be used to forecast the observed data.
-Takens delay embedding theorem~\cite{takens1981detecting} and its various generalisations (see,~\cite{sauer1991embedology, stark1999delay, gutman2018embedding}) establishes the learnability of a system constructed by concatenating sufficiently large previous time-series observations of a dynamical system into a vector (called delay coordinates).
-This then confirms the existence of a map on the space of delay coordinates equivalent (or topologically conjugate) to the underlying map from which the observed time-series was first obtained.
-Although topological conjugacy guarantees an alternate representation of the underlying system, the quality still depends on numerous parameters, making the comprehension of the dynamics unreliable (see, [?]).
-One reason for this fragility is that the embedded attractor in the reconstruction space is not always an attractor of the map learnt in the reconstruction space, despite unquestionably being an invariant set.
-When the embedded object is not explicitly known to be an attractor (as explained in Chapter~ \ref{ch3}), it can cause predictions to fail.
-
-Practically, the application of Takens embedding involves learning a map through some technique, and consequently one wishes that these would have low functional complexity\cite{manjunath2021universal}, i.e., functions with fewer oscillatory graphs.
-On the other hand, pure machine learning methodology processes temporal information (like the echo state networks~\cite{Manju_ESP, Manju_IEEE, grigoryeva2018echo}) by mapping data onto a higher dimensional space for further processing.
-Although they perform  well on forecasting some dynamical systems, they fail completely on others (\ednote{Ask for resources here}) as there is often no guarantee that the right function was learnt during training.
-
-This project involves implementing and analysing non-interpretable models that can guarantee exact reconstruction. 
-The project work concerns the study and implementation of a method (\cite{manjunath2021universal}) that incorporates learning a function by mapping the data on to a higher dimensional space using what is called a driven dynamical system (See Chapter~ref{ch.4}).
-With a clear understanding of how the data is mapped onto the higher dimensional space, the method then permits the learning of a dynamical system topologically conjugate to that of the underlying system.
-Instead of linear regression as in the training of echo state networks, deep learning methods are employed to learn the correct function.
-With slight modifications to the implementation in the paper (\cite{manjunath2021universal}), we show that one can construct accurate non-interpretable models with the ability to reconstruct attractors from more hard-to-forecast systems like the double pendulum.
-(The forecasting of the time-series from a double pendulum has not been reported before.) Moreover, we also demonstrate that long-term statistical consistency is preserved.
-
-\emph{By solidifying the mathematical underpinnings of our theory, we hope to guarantee the ability to construct models with predictive power ranging from molecular biology to neuroscience.
-in the near future.
-(We can modify this sentence when the report is completed.)}
-
-The more intricate details of proofs are referred to where relevant throughout. This report is organised as follows: 
-\newline In Chapter~\ref{ch2}, we recall the definition of a discrete-time dynamical system, how a discrete-system arises from a flow of an ODE and then proceed to define the inverse-limit space and topological conjugacy of autonomous systems.
-\newline Chapter~\ref{ch3} introduces the problem of forecasting dynamical systems, states the Takens delay embedding theorem, and discusses various issues faced while forecasting. 
-\newline In Chapter~\ref{ch4}, a driven dynamical system is defined and discuss the properties of a specific class of driven dynamical systems that we make use of in this project.
-\newline Finally, in Chapter~\ref{ch5} we show the implementation of these forecasting methods, and conclusions are provided in Chapter~\ref{ch6}.
-
-\chapter{Discrete-time Dynamical Systems}\label{ch2}
-
-This chapter briefly describes a discrete-time dynamical system and what it means to exhibit chaos. We refer to \cite{devaney2018introduction, de2013elements} for more details.
-
-At its most elementary level, a dynamical system is just something that evolves deterministically through time.
-In the context of this project, deterministic refers to the fact that a system evolves according to specified rules rather than based on random events.
-Dynamical systems arise in a variety of situations.
-A continuous-time dynamical system describes the states for all values of the time.
-Specifically, if the motion of a pendulum in which the quantities such as the angular position and angular momentum are known at all times, then it is a continuous-time dynamical system.
-The equations of the dynamical system can take the form of one or more ordinary differential equations that determine the relevant quantities at any future time if we know the initial location and momentum.
-In ecology, discrete-time dynamical systems are widely used to model population growth.
-The model in this case is a function calculating the following generation's population given the population of the previous generation.
-If we know the starting population, we may once again calculate the population at any time in the future.
-
-Formally, a function $T: U \to U$, where $U$ is some set is  a \emph{discrete-time dynamical system} and its iterates $\{u,Tu,T^2u,\ldots\}$, where $T^n$ denotes the $n$-fold composition of $T$ with itself, describe the evolution of an initial condition $u\in U$ (Note that we frequently drop the brackets and denote $T(u)$ by $Tu$ so as to simplify notations).
-
-Continuous-time dynamical systems modelled using ordinary differential equations can give rise to discrete-time dynamical systems.
-To see this, consider a differential equation $\dot{x} = f(x)$, $f: \mathbb{R}^n \to \mathbb{R}^n$, $n\in\mathbb{N}$ given to have a unique solution passing through
-each point $x\in\mathbb{R}^{n}$, call it $x(t)$ where $x(0)=x_0$.
-
-\begin{Definition}
-  [\bf Flow of an Equation] \label{Dfn_Flow}\rm
-  The flow of the equation  $\dot{x} = f(x)$  is defined to be a mapping $\varphi: \mathbb{R}^n \times \mathbb{R} \to \mathbb{R}$ where $\varphi(x_0,t)= x(t)$ for the solution $x(t)$ with $x(0)=x_0$ 
-\end{Definition}
-
-By fixing $t=K \in (0,\infty)$, we can define the \emph{time-$K$ map} as  $T(x):= \varphi(x_0,K)$, and it is easily verified that $T\circ T(x_0) = \varphi(x_0,2K)$.
-In general the $m^{\mbox{th}}$ iterate of $x_0$ under $T$ would be the value of the solution of the ODE evaluated at time $mK$ with the initial condition $x_0$, i.e.
-$\varphi(x_0, mK)$.
-Thus ordinary differential equations give rise to a discrete-time dynamical system by sampling the value of the solution $x(t)$ at time intervals $K$ units apart.
-
-A numerical discretization of a differential equation can also give rise to a discrete-time dynamical system.
-For instance, Euler's method approximates $\dot{x}(t)$ by $(x(t+h)-x(t))/h$; if $h$ is fixed throughout, the solution of a differential equation $\dot{x}=f(x)$
-can be approximated at the time instant $t+(m+1)h$ by iterating the equation 
-\begin{equation}
-  x(t+(m+1)h) = x(t+mh) + h f(x(t+mh))
-\end{equation}
-
-Adopting more succinct notation by replacing $x(t+mh)$ with $u_m$, we rewrite the above equation as
-\begin{equation}
-u_{m+1} = u_m + hf(u_m)
-\end{equation}
-or in even simpler terms as the discrete-time dynamical system with map $T(u) = u + hf(u)$, where $T$, $u$ and $f$ are understood to be as above.
-
-Of course, discrete-time dynamical systems need not always arise through a differential equation.
-Once again, we may consider the field of ecology, where discrete dynamical systems are often directly derived or assumed.
-\ednote{Ask for resource}
-There is a school of thought that advocates discrete-time dynamical systems to be more natural for modelling real-world observations than differential equations.
-We refer the interested reader to~\cite{saber2010introduction}.
-
-\section{Invariant Sets}
-
-A core concept in the study of dynamical systems is that of invariance. 
-\begin{Definition}
-  [\bf Invariant Set]\label{Dfn_InvariantSet}\rm
-  Given a discrete-time dynamical system $T: U \to U$, a subset $A \subset U$ is said to be an \emph{invariant set} if $T(A) =A$. 
-\end{Definition}
-
-We also define the orbit of a function $T$.
-\begin{Definition}
-  [\bf Orbit of $T$]\label{Dfn_Orbit}\rm
-  The orbit of $T$ is to be defined the sequence $\bar{u} = \{u_n\}_{n\in \mathbb{Z}}$ obeying the update equation, $u_{n+1}=Tu_n$, $n \in \mathbb{Z}$. 
-\end{Definition}
-
-Two examples of invariant sets include a fixed point where $Tu=u$ for  $u\in U$, and a periodic orbit, i.e., a set of iterates $\{u,Tu, T^2u,\ldots,T^pu\}$, where $T^{p+1}u=u$ for some $p\in\mathbb{Z}$.
-The entire space $U$ could be also be invariant.
-Consider for example the space $U=[0,1]$, where $Tu=4u(1-u)$ and then $U$ is invariant (as every $u\in{U}$ can be written as $u = 4x(1-x)$ for some $x\in{U}$).
-
-We may learn a great deal about the iterates of a dynamical system by considering the types of invariant sets of a discrete-time dynamical system.
-For example, if $U=[0,1]$ has map $Tu= u/2$, then the only invariant set is $\{0\}$, and all orbits approach this invariant set as time flows in the forward direction.
-Indeed, if a some non-zero invariant set (call it $B$) exists, then there is  some $r\in(0,1]\cap{B}$.
-But $r\notin{T(B)}$ since any orbit with initial value $r$ will be a decreasing sequence.
-Moreover, every orbit of T will be decreasing and therefore approach the value $0$ as $n\rightarrow\infty$.
-
-One may ask if every orbit approaches an invariant set?
-In general the answer is no, since for the dynamical system $T: \mathbb{R} \to \mathbb{R}$ defined by $Tu=2u$, any orbit that does not intersect the invariant set $\{0\}$, will not approach any invariant set.
-
-However,  when the space $U$ is compact, all orbits approach an invariant set.
-This is since the set of limit points of the orbit can be shown to be invariant \cite{de2013elements}.
-So, when $U$ is compact,  the $\omega$-limit set $\omega(u;T)$ of a point $u$ defined to the collection of limit points of the sequence $\{x,Tu,T^2u,\ldots\}$ is nonempty, and $\omega(u;T)$ is invariant.
-
-\begin{Example}
-  For the map $Tu=u^2$ defined on $[0,1]$ all orbits lie in the invariant set $\{1\}$ or else would would approach the invariant set $\{0\}$. 
-\end{Example}
-
-Invariant sets have various properties. Invariant sets can be attracting or repelling depending on how orbits in their vicinity behave. 
-Recall the examples $Tu =u/2$ and $Tu=2u$ defined on $\mathbb{R}$ (where $\{0\}$ “attracts" orbits) or the example, $Tu=u^2$ defined on $[0,1]$ (where $\{1\}$ “repels" orbits). 
-We are interested in attractive invariant sets since they capture the long-term dynamics as time increases. 
-In particular, we are interested in those invariant sets named attractors.
-
-\begin{Definition}
-  [\bf Attractor]\label{Dfn_Attractor}\rm
-  Let $T: U \to U$, where $U$  is a metric space with metric $d$. A compact subset $A \subset U$ is said to be an attractor if it satisfies the three conditions: 
-  \vspace{-8mm}
-  \begin{enumerate}
-	\item $A$ is invariant. 
-	\item $A$ is asymptotically stable, i.e., for every $\epsilon > 0$ and for all $u$ so that $d(u,A) < \epsilon$, we have $d(T^nu,A) \to 0$ as $n\to \infty$. 
-	\item $A$ has Lyapunov stability, i.e., for every $\epsilon > 0$  there exists a $\delta(\epsilon) > 0$ so that $d(u,A) < \delta$ implies $d(T^nu,A) < \epsilon$ for all $n\ge 0$.  
-\end{enumerate}
-\end{Definition}
-
-Indeed in our previous examples, it can be verified that the system  $Tu=u/2$ defined on  $\mathbb{R}$ and $Tu=u^2$ defined on $[0,1]$  the singleton set $\{0\}$ is an attractor.
-For the system,  $Tu=1-|2u-1|$ on $[0,1]$, one may easily verify that the only attractor is the entire space $[0,1]$.
-This follows from the fact that between any two points $u< v$ in $[0,1]$, and for any $a,b$ so that $u\le a < b \le v$, we can find an $n$ so that $T^n(a,b)=[0,1]$ (as would be explained later in this chapter).
-
-A dynamical system can have several attractors and may also be contained in another attractor.
-For the example, $Tu =u^2$ on $[0,1]$, both $\{0\}$ and $[0,1]$ are attractors.
-It is known that an attractor for a dynamical system always exists in a compact space~\cite{Milnor1985}.
-
-The dynamics restricted to an invariant set can be complicated.
-For instance an invariant set could be just a single point or it could have an infinite set.
-If the invariant set is infinite, then complicated dynamics are possible.
-A particular, well-studied phenomenon of such complexity gives rise to so-called chaotic behaviour, a subject studied in detail over the past fifty years.
-
-% If the dynamics on the attractor are somewhat complicated in the sense that the attractor cannot be decomposed further, and if the attractor is infinite then there is a possibility of complicated dynamics, and a particular well-studied phenomenon of such complexity gives rise to what is called a chaotic behaviour.
-
-%If the dynamics on an attractor cannot be decomposed further and the attractor is itself also infinite, then there is a possibility of complicated dynamics, and a particular well-studied phenomenon of such complexity gives rise to what is called chaotic behaviour.
-
-\section{Chaos}
-
-In the 1960s, several mathematicians and mathematically interested scientists independently discovered chaos in the mathematical sense.
-The meteorologist Edward Lorenz may have been the first to explain this phenomenon in his 1963 paper \cite{lorenz1963deterministic}.
-The notions of invariance, attractivity, and chaos may also be described for continuous systems, and Lorenz's system comprised of a system of differential equations.
-The narrative of Lorenz's discovery of chaos and the history of other forerunners in this subject is fascinating.
-We highly recommend James Gleick's book Chaos: The Making of a New Science~\cite{gleick2008chaos} for those interested.
-It clearly illustrates these experiences, explains why chaos was such a startling and crucial mathematical and scientific discovery and describes the underlying mathematical notions for non-specialists.
-
-Different authors propose a number of intricately different definitions of chaos in the literature~\cite{RasbandChaos, TaborChaos, WigginsChaos} and each of them indicate some aspect of complexity.
-
-In practice, when only data is observed from a system, it is not possible to verify which definition or notion is satisfied by the underlying dynamical system.
-We do, however, specifically recall the definition of chaos in the sense of Devaney \cite{devaney2018introduction,de2013elements} so as to understand some nuances behind the complexity.
-Devaney's definition of chaos has three requirements, with the first being the notion  of sensitive dependence on initial conditions.
-
-\begin{Definition}\rm
-  [\bf {Sensitive Dependence on Initial Conditions}]\label{Dfn_SDIC}\rm
-A dynamical system $T: U \to U$ is said to have sensitive dependence on initial conditions (SDIC) if there exists a $\delta > 0$ such that for every $u \in U$ and in every neighborhood of $u \in U$ there exists a $v\in{U}$ and an integer $N:=N{(u,v)}\in\mathbb{Z}$ such that $d(T^Nu,T^Nv)>\delta$.
-\end{Definition}
-
-It is vital to acquire a sense of this definition as it can easily be misunderstood.
-It is a common misconception to interpret SDIC as two close points ($u$ and $v$) that eventually become separated by a distance $\delta$ under iteration by $T$.
-But this is not true.
-To fully comprehend all the subtleties of the concept, one needs to discuss it more thoroughly by considering each sentence with care and attention, whereafter one may examine how they fit together to convey the idea of sensitive dependence.
-
-To this end, we make 3 remarks:
-\vspace{-5mm}
-\begin{enumerate}
-  \item First, the $\delta>0$ in the definition of SDIC is independent of $u$. 
-  \item Second, in every neighborhood of $u$, we may not necessarily find all points $v$ in the neighborhood distinct from $u$ that would separate from the forward iterates of $u$. 
-  \item  Finally, $N$ depends upon $u$ and $v$ chosen, and their iterates may not separate forever (i.e., for all $n>N$) and we allow their iterates to get arbitrarily close in the future. 
-\end{enumerate}
-
-To illustrate the concept of SDIC, we present two examples - one from Mathematics and the other from the field of Physics.
-
-\begin{Example} \rm
-  The logistic map(LM), a recursion relation of the form $x_{n+1}=rx_n(1-x_n)$ where $x_n\in[0,1]$ and $r\leq{4}$ is a classic example used to illustrate the chaotic behaviour that a system may exhibit.
-When $r=4$, the one-dimensional recurrence relation $x_{n+1}=4x_n(1-x_n)$ can be used as the kindergarten's model to exhibit the presence of SDIC.
-\ednote{B: edited and removed unnecessary citation}
-  In the graph below~\ref{fig:log_sdic} we plot the first 50 iterates of the LM for $r=4$ and two slightly different initial values $x_0$.
-
-  \begin{figure}[ht]
-    \includegraphics[scale=0.74]{_logistic_sdic.eps}
-        \centering
-        \captionof{figure}{Figure generated for Logistic Map $x_{n+1}=rx_n(1-x_n)$ with $r=4$ to exhibit the presence of SDIC.
-Plotted are the values $x_n$ against time $n$ for the the first 50 step with initial values $x_0=0.2$ (blue) and $x_0=0.20001$ (red).
-Initially the two trajectories overlap, but they diverge completely at $n=12$ whereafter they follow distinct paths.}
-        \label{fig:log_sdic}
-      \end{figure}
-  \ednote{B: Caption and figure adjusted}
-  % For an intuitive sense of the fact that the LM exhibits SDIC, consider again the definition and substitute specific values.
-Let $\delta=0.05$ and for purposes of illustration consider the value $u=0.4$.
-The green band in~\ref{fig:log_sdic} shades the region $[u-\delta, u+\delta]$ and we can see that regardless the value $v$  and in every neighborhood of $u \in U$ there exists a $v\in{U}$ and an integer $N:=N{(u,v)}\in\mathbb{Z}$ such that $d(T^Nu,T^Nv)>\delta$
-  % specific point, say $u=0.4$.
-Draw a green neighbourhood $[0.35,0.45]$ of diameters $\delta=0.02$ around the point.
-It is easy to see that every neighbourhood around $u=0.4$ will contain some $v_1, v_2\in[0,1]$ and natural number $N_1, N_2 \in\mathbb{N}$ respectively s.t.
-the distance $d(T^Nu_1,T^Nv)>\delta=0.05$ .
-\end{Example}
-
-\begin{Example} \rm
-  A second example pertains to a simple physical system called the double pendulum - a pendulum with another pendulum attached to its head given an initial position and angular velocity.
-
-  Below are two figures denoting trajectories of a double pendulum's second head after some elapsed time.
-Depicted are three pendulum heads with equal angular velocity but differing slightly in initial position.
-As can be seen below, the trajectories diverge very quickly to become completely different.
-We discuss the double pendulum in greater detail in a later chapter (\ref{ch5}), but we mention the example here to provide a numerical example of a practical system exhibiting SDIC.
-
-\begin{figure}[ht]
-  \centering
-\includegraphics[scale=0.8]{_dp_sdic.eps}
-    \captionof{figure}{Three double pendulum heads with equal angular velocity and initial angles differing by 0,025 radians}\label{fig:dp_sdic}
-  \end{figure}
-
-\end{Example}
-
-The second part in Devaney's definition of chaos concerns topological transitivity.
-
-\begin{Definition}
-  [\bf {Topologically Transitive}]\label{Dfn_TopolTrans}\rm
-	A dynamical system $F: U \to U$ is topologically transitive if for any pair of nonempty open sets $E_1$ and $E_2$ there exists a $n\in\mathbb{N}$ such that $T^n(E_1) \cap E_2 \not= \emptyset$. 
-\end{Definition}
-
-Topological transitivity implies that iterates of an open set of initial conditions get mixed up with other open sets.
-On a compact metric space, one may show that topological transitivity also implies the existence of a point whose forward iterates are dense \cite{de2013elements}; or in other words, the orbit going through this point will be dense in the compact metric space.
-In fact, it is topologically more likely that the choice of an arbitrary point will be one whose iterates are almost dense.
-
-\begin{Definition}
-  [\bf {Meager Set}]\label{Dfn_Meager Set}\rm
-A subset of a topological space $U$ is said to be a meager set if it can be written as a countable union of sets of with empty interior. The complement of the a meager set is set to be a residual set.
-\end{Definition}
-
-To be more precise (see \cite{de2013elements}), for a discrete-time dynamical system on a compact space, the set of points with dense iterates are residual, and they are typical or likely to be observed in practice.
-In this project, when data is observed from a topologically transitive system, we assume it arises from a dense orbit.
-
-We may now define the notion of Chaos as formulated by Devaney\cite{devaney2018introduction}.
-\begin{Definition}
-  [\bf {Devaney's Chaos}]\label{Dfn_ChaosDec}\rm
-	A dynamical system $T: U \to U$ is said to exhibit chaos in the sense of Devaney if it satisfies the three properties:
-	\vspace{-5mm}
-  \begin{enumerate}
-		\item $T$ has SDIC.
-		\item $T$ is topologically transitive.
-		\item The set of periodic points of $T$ are dense in $U$. 
-	\end{enumerate}
-\end{Definition}
-
-\begin{Example} \rm
-  The standard tent map $Tu=1-|2u-1|$ defined on $[0,1]$ is a well-known example of a dynamical system satisfying these three properties.
-
-  We now reason as to why this is true.
-The graph of the map $T$ is piecewise linear with two straight lines, one connecting the points $(0,0)$ and $(\frac{1}{2},1)$ and the other connecting $(\frac{1}{2},1)$ with $(1,0)$.
-They form a so-called tent with base centered at $u=1/2$.
-The graph of the map $T^2$ comprises two symmetric tents with their base centered at $1/4$ and $3/4$.
-See Figure \ref{fig:T2tentmap}
-
-  \begin{figure}[ht]
-    \includegraphics[scale=0.6]{_tentmap_2.eps}
-        \centering
-        \captionof{figure}{Graph of the $T^2$ map with bases at $x=0$, $x=\frac{1}{2}$ and $x=1$ respectively.}\label{fig:T2tentmap}
-  \end{figure}
-
-  In general, the graph of  $T^k$ contains $2^{k-1}$ tents.
-  Given any point $u$ and an open neighborhood $W \subset[0,1]$ of $u$, we can find $k\in\mathbb{Z}$ large enough so that $T^k(W) = [0,1]$ because we can accommodate a tent whose base is contained in $W$.
-% \ednote{B: This last phrase doesn't make sense to me.
-M: The number of tents grow with $k$ and hence this is obvious; I explained in one of our meetings.
-Does it make sense?}
-  This implies the existence of two points $w_1$ and $w_2$ in $W$ so that $d(T^kw_1,T^kw_2)= 1$, and by the triangle inequality  at least one of the inequalities $d(T^ku,T^kw_1)> \delta$ or $d(T^ku,T^kw_2)> \delta$ hold when $\delta\in (0,\frac{1}{2})$.
-So $T$ has sensitive dependence on initial conditions.
-
-  Next, let $W_1$ and $W_2$ be two nonempty open sets.
-Given any open set $W_1$, we can find a $k\in\mathbb{N}$ large enough so that $T^k(W_1) = [0,1]$ because we can accommodate a tent with base contained in $W$.
-So $T^k(W_1) \cap W_2 \not=\emptyset$, and thus $T$ has topological transitivity.
-
-  Finally, for every open interval $(a,b)$, the graph of $T^k$ intersects the graph of the identity map on $[0,1]$.
-This follows from the fact established above that there exists a $k\in\mathbb{N}$ such that $T^k(a,b) = [0,1]$.If the intersection point has coordinates of the form $(p,p)$, the point $p$ is then a fixed point of $T^k$ and therefore also a periodic point of $T$.
-Since we have found a periodic point in the interval $W$,  the set of periodic points are dense in $T$.
-  This follows because any fixed point of $T^k$ is also a periodic point, and we have found this in an arbitrary interval.
-
-\end{Example}
-
-\section{Conjugacy}
-
-We now turn our attention to the subject of conjugacy that describes when two dynamical systems are equivalent dynamically.
-
-To show topological similarity (or sameness) between two metric or topological spaces, one needs to establish a homeomorphism between the two spaces. 
-However, in the study of dynamical systems defined on two spaces, establishing a homeomorphism does not indicate that the systems are dynamically related in any way.
-For instance, the maps $Tu=u^2$ and $Tu=1-|2u-1|$ defined on $[0,1]$ have totally different behaviour. 
-Therefore, to find dynamically similar systems, one must establish a dynamical equivalence that we illustrate in a simple commutativity diagram below~\ref{eqn_conjugacy}.
-
-% Giving errors
-\begin{equation}  \label{eqn_conjugacy}
-%    \[ 
-    \psset{arrows=->, arrowinset=0.25, linewidth=0.6pt, nodesep=3pt, labelsep=2pt, rowsep=0.7cm, colsep = 1.1cm, shortput =tablr}
- \everypsbox{\scriptstyle}
- \begin{psmatrix}
-U & U\\%
-V & V.
- %%%
-%  \ncline{1,1}{1,2}^{T} 
-%  \ncline{1,1}{2,1} <{\phi}
-%   \ncline{2,1}{2,2}^{S}
-%  \ncline{1,2}{2,2} > {\phi}
- \end{psmatrix}
-% \]
-\end{equation}
-
-%Composite functions
-To understand the diagram, we note that if we travel "right and then down," the diagram instructs us to use $T$ (top arrow right) first, followed by $\phi$.
-(right arrow downwards).
-Consequently, the pathway amounts to finding $\phi(T(u))$.
-If we proceed "down and then right," the diagram instructs us to apply $\phi$ (left arrow downwards) first, and then apply $S$ (bottom arrow right).
-This then amounts to finding  $S(\phi(u))$.
-When the relationships in this diagram hold, we say $\phi(T(u))= S(\phi(u))$, and we formally denote it as $\phi \circ T=S\circ \phi$.
-
-\begin{Definition}\rm  
-[\bf {Conjugacy \& Semi-Conjugacy}]\label{Dfn_Conjugate}\rm
-   Consider the dynamical systems $T:U\to{U}$, $S:V\to{V}$ and suppose the relationship $\phi \circ T=S\circ \phi$ holds.
-If $\phi:U\to{V}$ is a homeomorphism, then $S$ is said to be conjugate to $T$ and $\phi$ is called a conjugacy.
-If we relax the criterion on $\phi$ and merely require $\phi$ to be continuous and surjective where $\phi:U\to{V}$, then $\phi$ is a semi-conjugacy between $T$ and $S$ where T has domain $U$ and S has domain $V$; S is said to be semi-conjugate to T.
-\end{Definition}
-
-\ednote{B: Does defining explicitly the domain and codomain of $\phi$ make it clear enough?
-M: Saying $\phi$ is then a semi-conjugacy between $T$ and $S$ could be ambiguous as it does not specify which system is on the top in the commutativity.
-So please edit accordingly.}
-When $S$ is conjugate to $T$, the dynamics of the two systems are in some way `dynamically equivalent'.
-Specifically, they are in one-to-one correspondence with one another.
-However when $S$ is semi-conjugate to $T$ with $\phi:U\to{V}$ a many-to-one mapping, the dynamics on $V$ provide merely a coarse-grained description of the dynamics on $U$~\cite{de2013elements}.
-When $S$ is semi-conjugate to $T$, it is also common to call $S$ a \emph{factor of $T$}, or conversely that $T$ is an \emph{extension of $S$}.
-In essence, an extension (see, \cite{de2013elements}) is a more extensive system capturing all of the essential dynamics of its factor.
-
-It is a very hard or nearly an impossible task to establish the existence of a conjugacy or a semi-conjugacy $\phi$ between two systems \cite{devaney2018introduction}.
-However, one can verify that a function $\phi$ satisfies the commutativity diagram \ref{eqn_conjugacy}.
-
-\begin{Example}\rm
-  For example $\phi(x)=\sin\big(\frac{\pi}{2}x\big)^2$ is a conjugacy between two systems $Tu=1-|2u-1|$ and $Sv=4v(1-v)$  defined on $U=[0,1]$.
-
-  Indeed, $\phi:[0,1]\to[0,1]$ is a homeomorphism and by employing simple trigonometric identities, it is easily proved that $\phi$ satisfies~\ref{eqn_conjugacy}
-  \[
-    \begin{aligned}[t]
-      (\phi\circ{T})u 
-                &=\sin^{2}\Big(\frac{\pi}{2} - \frac{\pi}{2}|2u-1|\Big) = \sin^{2}(\pi{u})=\Big(2\sin(\frac{\pi}{2})\cos(\frac{\pi}{2})\Big)^{2}=4\sin^{2}(\frac{\pi}{2}u)\Big(1-\sin^{2}\frac{\pi}{2}u\Big) \\
-                                    &=({S}\circ\phi)u.
-    \end{aligned}
-\]
-And it follows that $\phi$ is a conjugacy between the two systems.
-\end{Example}
-
-By establishing that two systems are conjugate (semi-conjugate) to one another, we have also shown that one may choose to work with one system instead of the second and still be guaranteed to obtain information on the latter.
-It is instrumental during the process of forecasting the future evolution of dynamical systems.
-
-\chapter{The Learning Problem}\label{ch3}
-
 In this chapter, we acquaint ourselves with the question of forecasting dynamical systems with unknown underlying structures, state and discuss the shortcomings of the Takens delay embedding theorem and discuss various issues faced while forecasting.
@@
 \begin{figure}[ht]
-  \includegraphics[scale=0.25]{_takensmap.eps}
+  \includegraphics[scale=0.25]{Graphs/_takensmap.eps}
   \centering
@@
 Do you have a specific reference?}
-
-\chapter{Driven Dynamical Systems to Forecast Problems}\label{ch4}
-
-In this chapter, we discuss results on the mapping of temporal data obtained from a discrete-dynamical system onto a different space through the notion of a driven dynamical system. 
-We also consider the conditions a driven system should possess to avoid adding distortion to its state-space representation.
-We then describe the \emph{single-delay dynamics} (SDD) of the system and finally recall conditions on the driven system so that the SDD are conjugate (or at least semi-conjugate) to the underlying system.
-The SDD can then be used to forecast and reconstruct the underlying system.
-
-\section{Nonautonomous and Driven Dynamical Systems}
-
-\begin{Definition}
-  [\bf Nonautonomous Dynamical System (adopted from~\cite{Manju_ESP}))] \label{Dfn_NDS}\rm
-  A nonautonomous dynamical system (NDS) on a space $X$ is simply a dynamical system comprising a family of maps $\{g_n\}_{n \in \mathbb{Z}}$ where $g_n(\ldotp):=g(u_n,\ldotp):X\to{X}$ is continuous.
-Each $g_n$ arises as a consequence of an input $u_n$ from the input space $U$, a topological space.
-\end{Definition}
-
-We immediately recall the figure first defined in \eqref{eqn_conjugacy}  and remind ourselves of our ultimate goal to learn a system that is conjugate to an underlying or unknown dynamical system $(U,T)$ by making use of the measurements obtained from  the unknown system.
-
-%#####################################
-\begin{center}
-\psset{arrows=->, arrowinset=0.25, linewidth=0.6pt, nodesep=3pt, labelsep=2pt, rowsep=0.7cm, colsep = 1.1cm, shortput =tablr}
-\everypsbox{\scriptstyle}
-\begin{psmatrix}
-U & U\\%
-V & V.
-%%%
-%  \ncline{1,1}{1,2}^{T} 
-%  \ncline{1,1}{2,1} <{\phi}
-%   \ncline{2,1}{2,2}^{S}
-%  \ncline{1,2}{2,2} > {\phi}
-\end{psmatrix}
-%#####################################
-\end{center}
-
-To find a function $\phi$, the need now arises to consider a so-called driven dynamical system, a special case of a nonautonomous dynamical system.
-Two spaces are considered: input is taken from an input space $U$ and the state of a space $X$ is updated.
-(Previously, only the state was considered.)
-We do this to mimic the true conditions.
-By accounting for an extrogenous input sequence in $U$ which can also influence the  system-state at a specific time-step, it produces more general models than an autonomous system.
-
-\begin{Definition}
-  [\bf Driven, Compactly Driven Dynamical System] \label{Dfn_DDS} \rm
-A driven dynamical system comprises two topological metric spaces $U$, $X$ and a continuous function  $g:U\times{X}\to{X}$ where $g(u_n, x_n)=x_{n+1}$.
-If the input space $U$ is compact, we refer to the system as compactly driven. 
-\end{Definition}
-
-The update equation $x_{n+1} = g(u_n,x_n)$ for $n \in\mathbb{Z}$, input $u_n$ from $U$ and state $x_n$ belonging to $X$ (a compact space) generate the dynamics on X. 
-Abbreviated, we refer only to the \textit{driven system $g$}, where all other entities are understood implicitly.
-
-Notably, a nonautonomous dynamical system can be generated from $U$; any input $\overline{u}$, a bi-infinite sequence from $U$, gives rise to the sequence of self-maps $\{g(u_n, \cdot)\}_{n\in\mathbb{Z}}$ contained in $X$.
-Physically, one may think of this bi-infinite sequence as referring to a system that has been running for an extended period at the time of the first measurement taken from the system (alternatively, the first time a probe is inserted into the system to take an observation)
-
-\begin{Definition}
-  [\bf Entire Solution] \label{Dfn_Soln} \rm
-  A sequence $\{x_n\}_{n\in\mathbb{Z}}\subset X$ is called an entire solution (or simply a solution) of the driven system  $g$ with input $\overline{u}$ when it satisfies 
-  \[g(u_{n-1}, x_{n-1})=x_n\] for all $n\in\mathbb{Z}$
-\end{Definition}
-
-It is important to emphasise that a sequence $\{x_n\}$ satisfying the update equation above can only be a solution if $x_n\in{U}$ holds for all $n\in\mathbb{Z}$ and just for $n>0$.
-Consider the example below.
-
-\begin{Example} \rm \label{ex_halfux}
-  The only solution  $\{x_n\}_{n\in\mathbb{Z}}$ to the driven system  $g(u,x)=\frac{ux}{2}$, where $X=[0,1]$, $U=[0,1]$,  is the zero solution $x_n\equiv0$.
-  To see this, consider any $x_n=a\in[0,1]$ where $a\neq{0}$.  Let $\overline{u}\in{U}$ be an non-zero constant sequence , say $u_n=0.5$. 
-  The driven system may be rewritten as $x_{n-1}=\frac{2x_n}{u_{n-1}}=4x_n$ and the  iterates of $x_n$ in backward time will increase by a factor of 4 at each timestep. 
-  Thus for some $m\leq{n}$,  $1<x_m$ i.e. $x_m\notin{X}$. So ${\{x_n\}}_{n\in\mathbb{Z}}$ is not a solution and it follows that the only possible solution is the zero solution.
-\end{Example}
-% \ednote{Nicely written! Thank you}
-
-A system may also have multiple solutions, as evidenced in the example below.
-
-\begin{Example}\label{Ex_exp} \rm 
-  Consider the driven system $g(u,x)=x^2$ for $X=[0,1]$, $U=\mathbb{R}$.
-The system has an uncountable number of solutions, as there exists a solution for every $x\in{X}$ which also passes through the point $x$ and $lim_{n\to\infty}x_n=0$, $lim_{n\to{-}\infty}x_n=1$.
- The proof is deferred to immediately after the next paragraph (see~\ref{rem_proofEx}).
-\end{Example}
-
-As the solutions to a driven system are considered an important entity, we next identify a subspace $X_U$ of $X$ that contains all possible solutions.
-To realize such a subspace of a driven system $g$, the concept of a reachable set is defined.
-
-\begin{Definition}
-  [\bf Reachable Set]\label{Dfn_ReachableSet}\rm
-The reachable set of a driven system $g$ is exactly the union of all the elements of all the (entire) solutions, i.e., 
-\[X_U :=\Big \{x \in X:  x = x_k \mbox{ where  $\{x_n\}$  is a solution for some  $\bar{u}$} \Big \}.\]
-The set of all reachable states at a specific time $n$ for input $\overline{u}$ is denoted by $X_n(\overline{u})$
-\end{Definition}
-
-The reachable set itself can defined independently from $g$ possessing the ESP/USP.
-Note that $x\in{X_n(\overline{u})}$ if and only $g$ has a solution $\{x_k\}$ for $x_n=x$ and input $\overline{u}$, a result proved in nonautonomous dynamical systems literature~\cite{manjunath2014dynamics, manjunath2013echo}.
-\ednote{B: Adjusted sentence and cited}
-
-The set $X_n(\overline{u})$ is precisely the set of points $x$ through which some entire solution $\Psi$ passes at time $n$. We can give an alternate formulation of the set $X_n(\overline{u})$. 
-Now suppose, we for some fixed input 
-$\overline{u}$, we define $g_i = g(u_i,\cdot)$ for all $i\in \mathbb{Z}$, then the set of states
-
-\begin{equation} \label{eqn_association}
-X_{n,i}(\overline{u}) := g_{n-1} \circ \cdots g_{i+1} \circ g_i(X).
-\end{equation}
-
-The set $X_{n,i}(\overline{u})$ being the image of a finite composition of
-continuous maps is compact whenever $X$ is compact. Also $X_{n,i}(\overline{u})$ is
-nonempty.  Further, $X_{n,i}(\overline{u}) \supset X_{n,i-1}$. Hence $\bigcap_{i<n}
-X_{n,i}(\overline{u})$ is a nested intersection of closed nonempty subsets, and
-whenever $X$ is compact, the intersection is nonempty. We say $g$ is a topological contraction if $\bigcap_{i<n}X_{n,i}(\overline{u})$ is a singleton subset of $X$ for each $\overline{u}$.
-
-It turns out that $\bigcap_{i<n}X_{n,i}(\overline{u})$ is identically equal to $X_n((\overline{u})$ that we had defined earlier.
-This will lead to a result (for.
-see, \cite{manjunath2013echo}) that  \textit{$g$ being a topological contraction is equivalent to the existence of a exactly one entire solution.}
-
-% \begin{Definition}
-%   [\bf Topological Contraction]\label{Dfn_TopContr}\rm
-%   A function $g:U\times{X}\to{X}$ is a topological contraction if for all $n\in\mathbb{Z}$ and all $\overline{u}\subset{U}$, $X_n(\overline{u})$ is a singleton subset of $X$.  
-% \end{Definition}
-
-\begin{Remark}
-  [\bf Proof of Example~\ref{Ex_exp}] \label{rem_proofEx} \rm
-  For the driven system,  
-  $g(u,x)=x^2$ the evolution does not depend on the input at all.
-Hence $g(u,x)$ can be written as a single map $f(x)=x^2$ which is a homeomorphism on $[0,1]$.
-A left-infinite orbit is defined and it converges to $1$ while the right-infinite orbit converges to $0$.
-Or in other words the iterates of $f^{-1}$ converges to $1$ while the iterates of $f$ converge to $0$
-\end{Remark}
-
-Thus far, it has been demonstrated that a system may have one or more solutions; one may ask if a driven system always has a solution and, if so, whether it satisfies specific properties such as uniqueness.
-Should the driven system be compact, existence follows immediately, as shown in the following result.
-
-\begin{Theorem}\label{Thm_CompactExistence}
- Let $g$ be a driven system.  If $X$ is compact then for each input $\overline{u}$, there exists at least one solution to the driven system $g(\ldotp, x)$
-\end{Theorem}
-\begin{proof}
-  May be found in \cite{kloeden2011nonautonomous, manjunath2014dynamics, manjunath2013echo}
-\end{proof}
-
-% One may easily construct many systems with trivial solution-sets, such as $g(u,x)=x$ which has only the constant solution $x$ and so for $U=[-1,1]$, the system would have no solution if $|x|>1$.
-To refine the scenario, we consider only systems with unique solutions.
-
-\section{Unique Solution Property}
-
-\begin{Definition}
-  [\bf Unique Solution Property] \label{Dfn_usp}\rm
-  A driven system $g$ is said to have the Unique Solution Property (USP) if for each input $\overline{u}$ there exists exactly one solution. 
-  Alternatively we may formulate the USP as follows: $g$ has the Unique Solution Property if there exists a well-defined map $\Psi:{U}\to{X}$, with $\Psi({\overline{u}})$ denoting the unique solution.
-\end{Definition}
-
-One of the first results obtained after defining the USP is that every solution will attract different initial conditions towards the component parts of the solution.
-If $g$ has the USP, then any solution to $g$ is also a uniform attractor in nonautonomous dynamical systems literature~\cite{Manju_Nonlinearity}.
-The discussion on nonautonomous attractors is beyond the scope of this project, and we refer the reader to \cite{Manju_ESP, esann2012ids} for further reading.
-
-% Paraphrased.
-%Additionally, $g$ having the USP is in our context also equivalent to $g$ being a topological contraction \cite{manjunath2021universal}, or alternatively also that $g$ exhibits the Echo State Property (ESP).
-
-In the majority of Reservoir Computing(RC) literature (a machine learning approach) where the input is mapped onto a different but higher dimensional through a system called the reservoir and a readout that measures the state of the reservoir is trained, a notion of forgetting the states of the reservoir asymptotically is often used.
-
-% \ednote{How should I explain/qualify RC better here?
-It's mentioned without really explaining.
-M: Have put it in one sentence; just to let you know people use optical devices to map signal into a ``reservoir".
-B: Thanks!},
-Concepts like the ESP,  fading memory \cite{boyd1985fading} are some of these notions.  
-If $g$ possesses the ESP (equivalent in our context to the USP), the we are guaranteed that the whole of an input's left-infinite history will exactly determine the current system state; i.e.
-there is only ever the possibility of a single reachable state at a given instance of time \cite{jaeger2001echo,Manju_2020}.
-% Defined formally, the ESP is intrinsically linked to an input (see \cite{Manju_ESP}) Unfortunately, definitions and results concerning the ESP do not have much practical benefit since a relationship between the network's response and the temporo-statistical properties of the input are not directly described.
-A more detailed presentation may be found in~\cite{jaeger2001echo}.
-
-% We do however consider it important to mention that the ESP is also often discussed in terms of a stability property that holds for all inputs of a system~\cite{manjunath2020stability} and, furthermore, plays a key role in the design and training of recurrent neural networks(RNN) within the field of RC \cite{Manju_ESP}.
-Beyond forgetting the past states the ESP is also often discussed in terms of a stability property that is nearby inputs create close by responses (solutions), a concept called input-related stability~\cite{manjunath2020stability} and, furthermore, plays a key role in the robustness of recurrent neural networks(RNN), a component of the implementation discussed in~\ref{ch5}.
-
-\section{Conjugacies}
-
-Having already defined the reachable set $X_U$ as the collection of all elements of all entire solutions, we pause briefly in order to fix additional notation.
-Defining $\cev{u}^{n}:=(\ldots,u_{n-2} ,u_{n-1})$ as the left-infinite subsequence of an input up until time $n$, $\overleftarrow{U}$ is then the notation for all these left-infinite sequences in $U$.
-Moreover, $\cev{u}^{n}v:=(\ldots,u_{n-2} ,u_{n-1}, v)$ will symbolise the input up to time $n$ with $v \in U$ being the specific input value at time $n$. 
-The introduction of a new input at time $n$ can be described by the mapping $\sigma_v:   \cev{u}^{n} \mapsto \cev{u}^{n}v$. 
-The right-shift map, $r\overleftarrow{U}\to\overleftarrow{U}$, of an input sequence is defined $r: (\cdots, u_{-2},u_{-1}) \mapsto(\cdots, u_{-3},u_{-2})$
-
-The question now assumes the form: Can we establish a semi-conjugacy for the driven system $g$ as presented below.
-
-\begin{equation}  \label{Scomm_h}
-  %    \[ 
-      \psset{arrows=->, arrowinset=0.25, linewidth=0.6pt, nodesep=3pt, labelsep=2pt, rowsep=0.7cm, colsep = 1.1cm, shortput =tablr}
-   \everypsbox{\scriptstyle}
-   \begin{psmatrix}
-   \overleftarrow{U} & \overleftarrow{U}\\%
-   X_U & X_U.
-   %%%
-  %  \ncline{1,1}{1,2}^{\sigma_v} \ncline{1,1}{2,1} <{h}
-  %  \ncline{1,2}{2,2} > {h}
-  %  \ncline{2,1}{2,2}^{g(v,\cdot)}
-   \end{psmatrix}
-  % \]
-  \end{equation}
-
-We proceed to consider a specific subclass of conjugacies.
-
-  \begin{Definition}
-    [\bf Universal Semi-Conjugacy]\label{Def_UnivSemiConj} \rm
-    Given a driven system $g$, we  call a continuous and surjective map $h : \overleftarrow{U} \to X_U$ a universal semi-conjugacy if  diagram \ref{Scomm_h} commutes for all $v \in U$.
-  \end{Definition}
-
-  If the universal semi-conjugacy $h$ exists (i.e. the diagram in \ref{Scomm_h} commutes) then the solution $\Psi(\bar{u})$ will intuitively have no more ``complexity' than the input $\bar{u}$.
-
-But does such a function $h$ for the above driven system $g$ above exist?
-Whenever $g$ has the USP and $\Psi(u)=\{x_n\}_{n\in\mathbb{Z}}$ it follows that $h$, defined by  $h(\cev{u}_n):=g(u_n,x_{n-1})=x_n$, will satisfy the semi-conjugacy in the graph above \ref{Scomm_h}.
-Regrettably, such a continuous mapping $h$ is not guaranteed to exist when $g$ does not have the USP \cite[Lemma 5]{Manju_Nonlinearity}.
-Note that even when $h$ does exist, we are not guaranteed its injectivity. Considering again example \ref{ex_halfux}, we see that even if $h$ were to exist, it could not be injective as $X_U=\{0\}$.
-
-Re-sketching the commutativity diagram \ref{Scomm_h} above by replacing $X_U$ by its left-infinite sequence space $\overleftarrow{X}_U$, we obtain the diagram below.
-In this case, the function $H:\overleftarrow{U}\to\overleftarrow{X}_U$, a map that is both continuous and surjective, is called a \emph{causal mapping} that is defined next.
-
-\begin{Definition}
-  [\bf Causal Mapping]\label{Def_CausMap}
-  A continuous, surjective map $H:\overleftarrow{U}\to\overleftarrow{X}_U$ such that \[H\circ\tilde{g}_v=\sigma\circ{H}\] holds for all $v \in U$ where $\tilde{g}_v$ maps $(\ldots, u_{-2}, u_{-1})$ to $(\ldots, u_{-2}, u_{-1}, g(v, u_{-1}))$ is called a causal mapping.
-\end{Definition}
-
-\begin{equation} \label{SCausal_H}
-    %    \[ 
-        \psset{arrows=->, arrowinset=0.25, linewidth=0.6pt, nodesep=3pt, labelsep=2pt, rowsep=0.7cm, colsep = 1.1cm, shortput =tablr}
-     \everypsbox{\scriptstyle}
-     \begin{psmatrix}
-     \overleftarrow{U} & \overleftarrow{U}\\%
-     \overleftarrow{X}_U & \overleftarrow{X}_U.
-     %%%
-    %  \ncline{1,1}{1,2}^{\sigma_v} \ncline{1,1}{2,1} <{h}
-    %  \ncline{1,2}{2,2} > {h}
-    %  \ncline{2,1}{2,2}^{\tilde{g}_v}
-     \end{psmatrix}
-    % \]
-  \end{equation}
-
- \begin{Theorem}
-  For a compactly driven system, a causal mapping $H$ exists if and only if $g$ has the USP. 
-\end{Theorem}
-{\bf Proof.}  See~\cite[Th.3]{manjunath2013echo}.
-
-The driven system $g$ can induce an embedding of $\overleftarrow{U}$ in $\overleftarrow{X}$ as follows: 
-If the causal mapping $H:\overleftarrow{U}{\to}{\overleftarrow{X}_U}$ is injective (in addition to being surjective), it becomes the embedding of $\overleftarrow{U}$ in $\overleftarrow{X}$ induced by the driven system $g$.
-The continuity of $H^{-1}$ follows from the fact that $H$ is itself a continuous and surjective mapping of a compact space $\overleftarrow{U}$ in a Hausdorff space.
-and we refer to the map $H$ as a \emph{causal embedding}. 
-\ednote{B: edited the sentence here}
-
-When $g$ has the USP, the diagram below (adopted from~\cite{Manju_Nonlinearity}) illustrates the operation of the mappings $h$ and $H$.
-The mapping $h:\overleftarrow{U}\to{X_U}$ is also considered an observable as mentioned in the introduction of Chapter \ref{ch3}).
-
-\begin{figure}[ht]
-  \includegraphics[scale=0.4]{_actionofh_H.eps}
-  \centering
-  \captionof{figure}{Ladder-like behaviour of $h$ and $H$ to illustrate the causal mapping. (Figure reproduced from~\cite{Manju_Nonlinearity}) }
-  \label{fig:actionh_H}
-\end{figure}
-
-The function $H$ contains a countably infinite number of coordinate (or factor maps) since its codomain is $\overleftarrow{X}_U$.  
-As it is not practically feasible to consider a left-infinite sequence in $X_U$ (in other words an element of  $\overleftarrow{X}_U$) for learning, we wish to restrict the inputs we consider to a subspace of $\overleftarrow{U}$.
-Particularly, we restrict our attention to the  so-called inverse-limit space of a dynamical system and hope to embed it within some finite product of $X$.
-
-%We wish to explore the possibility of establishing a conjugacy rather than a mere semi-conjugacy and therefore are interested in the restriction of inputs to the left-infinite orbits of a dynamical system.
-
-The left-infinite orbit space is significant because a great amount of information about the original system can be gleaned from its topological structure~\cite{ingram2011inverse,Manju_IEEE}. 
-The formal topological structure for the inverse-limit space of a chaotic dynamical system is complicated, and further discussion is relegated to~\cite{kennedy2008inverse_limit, ingram2011inverse}.
-Broadly, we may conceive of the inverse-limit system of a dynamical system as a subspace of an infinite-dimensional space where each point in the inverse-limit space corresponds to a backward orbit of the map $T$.
-Following~\cite{manjunath2021universal, ingram2011inverse} we provide a definition:
-
-\begin{Definition}
-  [\bf Inverse-Limit Space, Inverse-Limit System]\label{Dfn_InverseL}\rm
- The subset $\widehat{U}_T\subset \overleftarrow{U}$ defined by \[\widehat{U}_T: = \{ (\ldots,u_{-2},u_{-1}): Tu_{i} = u_{i+1}\} \] is called the inverse-limit space of $(U,T)$.
- The self-map $\widehat{T}$  induced by $T$ on $\widehat{U}_T$  is defined by \[\widehat{T}: (\ldots,u_{-2},u_{-1}) \mapsto  (\ldots,u_{-2},u_{-1},T(u_{-1}))\] and the resulting dynamical system $(\widehat{U_T}, \widehat{T})$ is called the inverse-limit system of $(U,T)$.
-\end{Definition}
-
-The inverse-limit space is well-defined since $T : U \to{U}$ is surjective by assumption.
-We now get to the concept of embedding the inverse limit-space of the dynamical system in $X \times X$. 
-\begin{Definition} \rm
-	We say a driven system $g$ \emph{causally embeds} a dynamical system $(U,T)$ if it satisfies the two properties: (i) a universal semi-conjugacy exists and  (ii)  $H_2(\cev{u}) := (h(r\cev{u}),h(\cev{u}))$ embeds the inverse-limit space $\widehat{U}_T$ in $X \times X$.
-\end{Definition}
-
-  It is imperative to take note of a subtlety here: we refer to the \emph{map} $H$ as a (causal) embedding, whereas the driven \emph{system} $g$ (causally) embeds another \emph{system} $(U,T)$.
-\ednote{B: inserted line}
-
-\section{Choosing the driven system $g$}
-
-So far it is not clear if the USP alone ensures the ability to causally embedding a dynamical system and we mention that a driving function $g$ cannot just be chosen in a frivolous manner.
-This could possibly complicate our work.
-%\ednote{B: Grammarly}
-
-%The driving function $g$ cannot be chosen in a frivolous manner as it could possibly complicate our work.
-Specifically, one must be careful to avoid a choice of $g$ which would  add complexity to the obtained solution.
-
-%When a causal embedding $H$ exists for the driven system $g$, one can map an arbitrary input ${u}$ onto the solution space $X$ without additional distortion or information-loss \textbf{(Cite.)}.
-
-%When an embedding is established, the question of possible additional complexity in the solution is removed by guaranteeing that, since the systems are conjugate (semi-conjugate, \textbf{(refer)}), $g$ does not add any (some) complexity to the system.
-
-%To achieve a causal embedding embedding not just a causal mapping, it is undesirable to choose a function $g$ that quenches the temporal structure in $u$ by contracting to such a degree that the ability to recover information from the original system is lost completely.
-In the above example(\ref{ex_halfux}), the input's temporal variation could not be related to the reachable set as $X_U$ only consisted of a single element; little, if not no, information is encoded.
-To obtain a suitably complex function $g$, it is thus desired that the reachable set of a driven system  be large enough to relate to the input. 
-This is true even for embedding  the inverse-limit space of a dynamical system. \ednote{B: Expand or refer?}
-
- To this end, we recall the notion of State-Input (SI) Invertibility.
-
-\begin{Definition}
-  [\bf SI-Invertibility]\label{Dfn_SIinv}\rm
-  A driven system $g$ is said to be SI-Invertible if $g(*,x): U \to X$ is invertible for all $x\in X$.
-Alternatively it may be said that if, given $x_n$ and $x_{n-1}$, $u_{n-1}$ can be uniquely determined from $x_n=g(u_n,x_{n-1})$, then $g$ is said to be SI-invertible.
-\end{Definition}
-
-SI-invertibility promises that 'enough' information is retained when $g$ is chosen without introducing one's choice of $g$ will still ensure 'enough' information is retained without introducing unwanted complexity.
-'Enough' here refers to the fact that we may always recover the previous input value $u_n$ (if we know successive states $x_{n-1}, x_n$).
-
-Subsequently we define the relation $Y_T$ induced by $(U,T)$ on $X_U\times{X_U}$ for a driven system $g$ possessing SI-invertibility.  
-To describe the  SDD formally, we consider a dynamical system $T: U \to U$ and define a relation on the reachable set $X_U$, i.e. a subset on $X_U \times X_U$  defined by 
-$$Y_T:=\{(x_{n-1},x_n): \{x_k\}_{k\in \mathbb{Z}} \mbox{ is a solution for some orbit of } T \mbox{ and } n \in \mathbb{Z}\}.$$
-
-The following theorem establishes the existence of a well-defined map $G_T$ describing the single-delay dynamics(SDD) of the system above.
-
-\begin{Theorem}\label{Thm_GT_Exists}
-  For an SI-invertible driven system $g$ and a dynamical system $(U,T)$, the map $G_T: Y_T \to Y_T$ defined by the relation $(x_{n-1},x_n) \mapsto (x_n,x_{n+1})$ is well-defined. 
-  (This results holds even in the absence of $g$ possessing over the USP) 
-  % Moreover, the mapping $(x_{n-1},x_n) \mapsto u_{n}$ is well-defined when $x_{n-1}$ and $x_n$ are successive points on a solution obtained for an input  $\{u_n\}$ that is an orbit of $T$.
-  \end{Theorem}
-  {\bf Proof.} See~\cite[Th.3]{Supp}
-
-%'Without additional complexity' is guaranteed by the invertibility of the function $g$ that guarantees a one-to-one mapping between $u_n, x_{n-1}$ and $x_n$.
-In obtaining an SI-invertible function, therefore, we are guaranteed a system which will not lose so much information about its previous states in the forward-flow of time that one may not make any credible claims as to the original system's behaviour and properties.
-Simultaneously, it does guarantee us that the important information encoded in previous inputs is indeed preserved.
-
-At this stage it is worth taking note of a specific driven system in the form of a discrete state-space model which has acquired some adherence in applications \cite{Manju_IEEE}- especially those pertaining to Echo State Networks and the ESP.
-The function
-\begin{equation}  \label{eqn_driving}
-  g(u,x) = (1-a)x + a\overline{\tanh}(Au + \alpha Bx)
-\end{equation} 
-is SI-invertible and, if $\alpha B$ has a spectral norm $<1$, also possesses the USP \cite[Th.2]{manjunath2013echo }. 
-It is easy to show that $g$ is SI-invertible by recovering $u_n$ in 
-\begin{equation} \label{eqn_SI_RNN}
-  u_{n-1} := A^{-1}\bigg(\overline{\tanh}^{^{-1}}\frac{1}{a}\Big(x_{n+1}-(1-a)x_n\Big) \bigg) - \alpha B x_n
-  \end{equation}
-  when $x_{n-1}$ and $x_n$ are known.
-
-This specific driving function $g$ is used in our implementation and is discussed more completely in chapter~\ref{ch5}
-
-Despite the ease that with which one manipulates a left-infinite history in the realm of theory, it is impossible to obtain or use such a sequence in any real-life application.  
-Fortunately, one does not need the entire left-infinite history of an input in practice thanks to the Uniform Attraction Property(UAP). 
-We use an alternate version of the definition of UAP -- since the notion of nonautonomous attractors in this project would take some time to establish and detracts from the principal thrust of this project, we refer to \cite{Manju_Nonlinearity}.
-
-\begin{Definition}
-  [\bf Uniform Attraction Property]\label{Dfn_UAP}\rm
-  A driven system $g$ has the Uniform Attraction Property (UAP) if we initialize the driven system
-with an arbitrary initial value $y_m \in X$, and the sequence $y_{m+1}, y_{m+2}, y_{m+3},...$ satisfying $y_{k+1}= g(u_k,y_k)$ for $k \geq m$ then approximates an actual solution $\{x_n\}$ uniformly in the sense that given $\epsilon>0$ (independent of $y_m$) there is an integer $n$ so that $d(x_{n+i}, y_{n+i})<\epsilon$ for all $i\ge 0$, where $\{x_m\}$ is a solution.
-\end{Definition}
-
-The UAP guarantees that all trajectories will converge to the same trajectory as time moves forward.
-An illustration of this is shown in Fig.~\ref{fig:memloss_conttime} where the trajectory in red locks on to the supposed actual solution when the system has the USP.
-
-\begin{figure}[ht]
-  \includegraphics[scale=0.35]{_Nonlinearity_Fig_response.eps}
-  \centering
-\captionof{figure}{A coordinate of a simulated solution $(x_0,x_1,\ldots,x_{5000})$ of a RNN  plotted in red (with parameter $\alpha=0.99$ (where $g$ has the USP) and blue ($a=1$ and $\alpha=1.05$ where $g$ does not have the USP) while the matrices $A$ and $B$ are randomly generated, and $B$ has unit spectral radius.
-Reproduced from \cite{Manju_Nonlinearity}. }
-\label{fig:memloss_conttime} 
-\end{figure} 
-\ednote{B: How can I explain the black line? I don't think this makes sense to the reader immediately.}
-
-One may now appreciate an even more astounding result: $g$ having the USP is equivalent to the UAP as is proved in~\cite{Manju_Nonlinearity}.
-
-%We restate the above theorem.
-
-%\begin{Theorem}\label{thm_usp-contr-uap}  The following statements are equivalent:  \vspace{-8mm}  \begin{enumerate}[noitemsep, label=\roman*.]    \item $g$ has the USP.
-\item $g$ is a topological contraction.
-\item $g$ has the UAP.
-\end{enumerate\end{Theorem}{\bf Proof.}  May be found in \cite[Th.1]{Manju_Nonlinearity}
-
-% One need therefore not establish any additional results to ensure that the sequence uniformly approaches the unique solution $\Psi$.
-This vastly simplifies the effort necessary to set up a problem in order to guarantee that the underlying system will be accurately represented by the conjugate system.
-
-%This also solves the problem of perturbation/noise introduced by the observable or measurement function.
-Recall from before that an observable is inherently a map that discretises the underlying continuous-time system.
-Moreover, measurement and device error introduce mistakes.
-Since we are working with a chaotic system displaying sensitive dependence on initial conditions, these small errors could potentially send the trajectory into a completely different attractor.
-%In the presence of the UAP, however, small measurement errors introduced into the system do not pose the same danger as before since the trajectories will eventually tend towards the actual solution.
-% \ednote{(B: Is this Complete enough?)}
-% \textbf{From the perspective of perturbation of an autonomous dynamical system, if the resulting nonautonomous system has the USP, then the temporal structure of noise relates to the dynamics on a nonautonomous uniform attractor through a topological semi-conjugacy or a conjugacy.}
-
-\section{The next step in Dynamics}
-
-% Recall that $g$ is only being given inputs from the orbits of $T$.
-
-%Incredibly, this permits one to initialise a driven system $g$ with an altogether arbitrary initial value $y_m\in{X}$ where $m\in\mathbb{Z}$ and the UAP then guarantees that the sequence $\{y_{m+1}, y_{m+2},\ldots\}$ which satisfies the relation $y_n=g(u_n, y_{n-1})$  for  $k\geq{m}$  will approach the elements $\{x_n\}$ of the actual solution.
-This vastly simplifies the effort necessary to set up a problem in order to guarantee that the underlying system will be accurately represented by the conjugate system.
-
-We are interested in determining whether $\widehat{T}$ and $G_T$ are related.
-For this, we introduce  the function $H_2{\overline{}} := (h(r\overline{u}, h(\overline{u})))$ mapping the entirety of a left-infinite sequence to some element in $X\times{X}$.
-Indeed, it may be shown that $g$ having SI-invertibility and the USP immediately guarantees the existence of at least a semi-conjugacy between the systems $(Y_T, G_T)$ and $(\widehat{U}_T, \widehat{T})$.
-This is formalised in the Causal Embedding Theorem, which we state next:
-
-\begin{Theorem}
-  [\bf Causal Embedding Theorem (adopted from \cite{Supp})]
-  % {\bf (Causal Embedding Theorem.)}
- \label{Thm_CET}
-	Let $g$ be a driven system with SI-invertibility and the USP.
-Let $h$ denote the universal semi-conjugacy and $H_2(\overleftarrow{u}) := (h(r\overleftarrow{u}),h(\overleftarrow{u}))$, where $r$ is the right-shift map.
- Let $(\widehat{U}_T, \widehat{T})$  be the inverse-limit system of a dynamical system $(U,T)$. 
- Then the restriction of $H_2$ to $\widehat{U}_T$ is a topological semi-conjugacy between the inverse-limit system $(\widehat{U}_T, \widehat{T})$ 
-and the induced dynamical system  $(Y_T,G_T)$, i.e., the following diagram commutes
-\begin{equation} \label{comm_H_CET}
-\psset{arrows=->, arrowinset=0.25, linewidth=0.6pt, nodesep=3pt, labelsep=2pt, rowsep=0.7cm, colsep = 1.1cm, shortput =tablr}
- \everypsbox{\scriptstyle}
- \begin{psmatrix}
- \widehat{U}_T & \widehat{U}_T\\%
- Y_T &  Y_T.
- %%%
-%  \ncline{1,1}{1,2}^{\widehat{T}} \ncline{1,1}{2,1} <{H_2}
-%  \ncline{1,2}{2,2} > {H_2}
-%  \ncline{2,1}{2,2}^{G_T}
- \end{psmatrix}
- \end{equation}
-or in other words, $(Y_T, G_T)$ is a factor of  $(\widehat{U}_T, \widehat{T})$. Further, if $T:U \to U$ is a homeomorphism, 
-then $H_2$ embeds $\widehat{U}_T$ in $X_U \times X_U$, and hence $(Y_T, G_T)$ is conjugate to $(\widehat{U}_T, \widehat{T})$.
-\end{Theorem}
-{\bf Proof.}  May be found as the proof of~\cite[Th.4]{Supp}
-
-Recall that $H_2$ maps an entire left-infinite solution sequence from $\Psi$ to an element in $X\times{X}$. 
-We are drawing nearer and nearer to our principal target and our results carry more and more weight: If we are able to learn $G_T$, we will also have learnt $\widehat{T}$; and since $\widehat{T}$ is an extension of $T$, we will have obtained $T$ as well.
-%\ednote{B: Grammarly check}
-
-%\begin{Theorem} Graph \ref{Scomm_h} is exactly the inverse-limit system $(\hat{U}, \hat{T})$.    \end{Theorem}
-
-%We now have the following (which may be compared with \ref{SCausal_H} above):
-
-%\begin{equation} \label{fig:inverse_limsystem}
-  %  \[       \psset{arrows=->, arrowinset=0.25, linewidth=0.6pt, nodesep=3pt, labelsep=2pt, rowsep=0.7cm, colsep = 1.1cm, shortput =tablr}      \everypsbox{\scriptstyle}
-%      \begin{psmatrix}      \widehat{U}_T  && \widehat{U}_T \\     Y_T && Y_T \\
-      %%%
-     %  \ncline{1,1}{1,2}^{\widehat{T}} \ncline{1,1}{2,1} <{h}
-     %  \ncline{1,2}{2,2} > {H_2}
-     %  \ncline{2,1}{2,2}^{G_T}
-%      \end{psmatrix}
-    %  \]
-%  \end{equation}
-
-%\begin{Theorem}
- %   $(Y_T, G_T)$ is semi-conjugate to $(\widehat{U}, \widehat{T})$.
-%\end{Theorem}
-% {\bf Proof.} See \cite[Th.3, Th.4]{manunath2021universal}
-
-\section*{Summarising the discussing thus far:}
-
-It is easy to lose the birds-eye view, so we take a moment to review our progress up until this point.
-
-\vspace{-8mm}
-\begin{enumerate}
-\item We are interested in a some dynamical system $(U,T)$ with unknown dynamics.
-\item To determine properties about this system $(U,T)$ and predict its future evolution, we determine the dynamics of the inverse-limit system $(\widehat{U}, \widehat{T})$.
-% Given certain assumptions, we can guarantee that $(\widehat{U}, \widehat{T})$ is at least semi-conjugate to $(U,T)$.
-\item If the driven system $g$ is SI-invertible (and $\{u_n\}\in{U}$  is an orbit of $T$), the map $G_T$ exists that describes the single delay dynamics. 
-\item If, furthermore, $g$ has the USP, $(Y_T, G_T)$ is semi-conjugate to $(\widehat{U}, \widehat{T})$.
-\item If we can assume that $T$ is a homeomorphism, $(Y_T, G_T)$ is topologically conjugate to $(\widehat{U}, \widehat{T})$, an extension space of $(U,T)$
-\end{enumerate}
-
-One can therefore in practice learn the SDD of the driven system states via $G_T$ with enough data thanks to the USP/UAP. This enables us to do at least two things: 
-\vspace{-8mm}
-\begin{itemize}
-\item Forecast  $x_{n+1},x_{n+2}, \ldots$ via iterates of $G_T$ (if $G_T$ can be learnt).
-\item Forecast future values of $u_n$ since $x_n$ and $x_{n+1}$ determine $u_n$ since $g$ is SI-invertible. 
-\end{itemize}
-
-Finally, we note that although $G_T$ exists with SI-invertibility, we need the USP as well(see \ref{fig_pictorialSummary} below for a visual illustration).
-If not, the driven states would have to be running for all time since they would not have forgotten their past states.
-
-\begin{figure}[ht]
-  \includegraphics[scale=0.3]{_summarypictorial.eps}
-  \centering
-  \captionof{figure}{Pictorial summary of results obtained up to now and how they overlap. Reproduced from \textbf{(Cite SIG-25May)} }
-  \label{fig:fig_pictorialSummary}
-\end{figure}
-\ednote{Ask M: Cite SIG-25May.}
-
-\section{A discussion of $G_T$ }
-
-In the above sections, we established the map $G_T$ describing the SDD of a driven system.
-By establishing the existence of this map, we’ve essentially embedded the attractor $U$ (recall from \ref{attractor_U}) into the higher dimensional space $X\times{X}$.
-In layman’s terms, this ensures that there is more “dimensional room” for the underlying system’s underlying dynamics to “move”.
-Since the dynamics aren’t as “squashed”, we might therefore hope that the dynamics of $G_T$ are in some sense simpler than that of $T$.
-(Taken note of the fact that $G_T$ is a homeomorphism even when $T$ is just continuous)
-
-In \cite{manjunath2021universal} it is illustrated in an empirical fashion in that the map $G_T$ describes dynamics which are less functionally complex than that of $T$ or of the map $\Phi_{2d,\theta}$.
-This is done by implementing a Recurrent Neural Network (RNN), which is discussed in~\ref{ch5}.
-\ednote{B: Should I add in that we'll be using Pearson coefficient?}
-
-We opt to learn $G_T$ in an indirect manner by defining a new map $\Gamma:(x_{n-1},x_n)\mapsto{u_n}$. It will follow immediately from $G_T$’s existence that $\Gamma$ also exists.
-
-The reasons for taking this roundabout approach remain to be discussed in \ref{subs_LearnGamma}, but a pat answer may immediately be given: When $\Gamma$ has been learnt, the system can be driven autonomously and then $G_T$ is known anyway.
-We formalise this with a theorem.
-
-\begin{Theorem}
-  When $x_n$ and $x_{n-1}$ are successive points on a solution obtained for input-orbit of $T$ $\{u_n\}$, then the map $\Gamma: X\times{X}\to{U}$ defined by $(x_{n-1},x_n)\mapsto{u_n}$ exists whenever $G_T$ exists
-\end{Theorem}
-{\bf Proof.}  See \cite[Th. 3c]{manjunath2021universal}.
-
-Projection mappings $\pi_i$ are defined in the traditional meaning where a $k$-dimensional vector is projected to it's $i^{th}$ component such that $$\pi_i:(a_1,a_2, \ldotp, a_{k-1}, a_k)\to{a_i}$$
-
-The graph in equation \ref{fig:comm_H_CET} is then extended as below:
-
-\begin{equation}  \label{Scomm_imp}
-  %    \[ 
-      \psset{arrows=->, arrowinset=0.25, linewidth=0.6pt, nodesep=3pt, labelsep=2pt, rowsep=1.3cm, colsep = 1.1cm, shortput =tablr}
-   \everypsbox{\scriptstyle}
-   \begin{psmatrix}
-   \widehat{U}_T & \widehat{U}_T\\%
-   Y_T & Y_T\\
-   & \textcolor{red}{U \times X}
-   %%%
-  %  \ncline{1,1}{1,2}^{\widehat{T}} \ncline{1,1}{2,1} <{H_2}
-  %  \ncline{1,2}{2,2} > {H_2}
-  %  \ncline{2,1}{2,2}^{G_T}
-  %  \psset{linecolor=red}
-  %  \textcolor{red}{\ncline{2,1}{3,2}<{(\Gamma,\pi_2)}}
-  %  \ncline{3,2}{2,2}>{\textcolor{red}{(\pi_2,g)}}.
-   \end{psmatrix}
-  % \]
-  \end{equation}
-
-The problem finally simplifies to the issue of learning the map $\Gamma$ and combining this with the projection mapping $\pi_2$ and the function $g$, which will be known.
-A final set of equations is obtained - equations that have been entirely constructed from data.
-\begin{eqnarray}\label{eqns_from_data}
-	u_{k+1} &=& \pi_1 \circ (\Gamma, \pi_2) \circ (\pi_2,g) (u_k,x_k) \label{Seqn_u}\\
-	x_{k+1} &=& \pi_2 \circ (\Gamma, \pi_2) \circ (\pi_2,g) (u_k,x_k). \label{Seqn_x}
-\end{eqnarray}
-
-\section{Advantages of learning $\Gamma$} \label{subs_LearnGamma}
-
-One may immediately ask why we opt to take such a roundabout route; why not just learn the map $G_T$ from the get-go?
-On the surface, this seems to be an arbitrary decision path with no real reasoning, so we take a pause again and discuss the motivation for learning Gamma.
-There are a number of distinct advantages.
-
-In the first place, learning $\Gamma$ saves computational resources.
-This is due to the fact that the input $u_n$ lies in a lower-dimensional subspace in comparison to the high-dimensional space $X\times{X}$.
-
-%In practice, if the input is of a  lower dimension, one may easily embed it into the space $X\times{X}$ by padding the vector $u_n$ with zeroes.
-
-Secondly, the function $\Gamma$ is known to be stable.
-Learning $\Gamma$ makes use of equations \ref{eqns_from_data} which in turn employs the driven system $g$ possessing the USP, and offers distinct advantages with regards to stability in the presence of perturbation.
-It is desirable that inputs that differ only slightly would have only ‘slight’ effects on the output of the system.
-Moreover, one would hope to prevent numerical errors originating from input and measurement noise (see~\cite[Th.
-5]{manjunath2021universal}).
-As the data fed into the system is a function of both the input given and the parameters of the system itself (i.e.
-the value a, $\alpha$ in $\tanh$), we exploit both input-related and parameter-related stability.
-Both these notions are defined by way of the continuity of an encoding map of an input and design parameter respectively in \cite{manjunath2020stability}.
-
-Informally, we may consider input-related stability as concerned with the question of whether or not small variations in input would result in small responses.   
-If two inputs $\bar{u}$ and $\bar{u}_{noisy}$ are close in the product topology, then their tails could differ greatly due to the metric that generates the product topology has insensitivity to the differences in sequence's tails.
- However if the system has the USP then the function $h$ is continuous and so $h(\bar{u}), h(\bar{u}_{noisy})$ remain close-by for small levels of noise maintaining a measure of robustness to input and measurement noise.\cite{manjunath2021universal}.
-In a similar fashion, parameter-related stability is obtained thanks to a result \cite[Lemma 3.2]{manjunath2020stability} relating this form of stability with the ESP (which is here equivalent to the USP).
-
-% Secondly,  due to $g$ possessing the USP, both input- and parameter-related stability~\cite{manjunath2020stability} is obtained, which in turn prevents numerical errors that caused by input and measurement noise (see~\cite[Th.
-5]{manjunath2021universal}).
-
-Thirdly, learning a chaotic map on a set(here $Y_T$) with even a small measure of noise  means that the learnt system accepts arguments out of the set (Problems similar to Takens, cf.~\ref{sect_takenslimits}).
-We thus have a map $G_T^+$ acting on $Y_T^+$ and since $Y_T$ is not guaranteed to be an attractor of $G_T$ this could lead to errors.
-If one learns $G_T$ directly, we are in danger of replicating the problems arising for Takens.  To see this, consider Fig.~\ref{fig:YtGtFailure}.
-
-\begin{figure}[ht]
-  \includegraphics[scale=0.25]{_YTerrors.eps}
-  \centering
-  \captionof{figure}{Diagram of potential errors developing when $G_T$ is learnt.}
-  \label{fig:YtGtFailure}
-\end{figure}
-
-To circumvent this: if a driven system with a saturation function\ednote{B: What is a saturation function?} is used, we can ensure the existence of an attractor $A$ of the map which implements $G_T$, and containing $Y_T^+$ indirectly (via $g$).
-\ednote{B: A little vague.
-Are we saying that an attractor of a map implementing $G_T$ is also the one that contains $Y_T^+$?}
-By defining the dynamics to $A$, this prevents large errors from occuring.  
-More details on the global dissipativity of $g$, as it is termed there, may be found in~\cite{Supp}.
-
-\begin{figure}[ht]
-  \includegraphics[scale=0.25]{_autovsdriven.eps}
-  \centering
-  \captionof{figure}{Illustrating empirically the importance of global dissipativity.
-Plotted below are the three principal components of randomly initiated points(blue) of a learnt version of $G_T$ versus the trajectory $(x_{n-1}, x_n)$(red) of driven states obtained from the Fractal Dream Attractor's data~\ref{Thomas_Attractor}).
-}
-  \label{fig:learningFailure}
-\end{figure}
-
-In effect, global dissipativity prevents large numerical errors due to the noise in input data from arising (see Fig.~\ref{fig:learningFailure}). 
-If a system is not globally dissipative, neglibly small errors could lead to major inaccuracy which thus induce predictions to utterly fail.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\chapter{Implementation} \label{ch5}
-
-In this chapter we turn to the implementation of the function $G_T$ by discussing a method to learn the function $\Gamma$.
-We apply our methodology to some chaotic attractors.
-Most notably, we consider the reconstruction of the chaotic attractor of the double pendulum - an  attractor that has not been successfully reconstructed because of its intricate structure.
-In fact the previous data-driven approaches also have not been successful in building a model from the double pendulum.
-
-%We then indicate where this implementation differs from previous work done by (\cite{manjunath2021universal}) and empirically show that the SDD of $G_T$ are less functionally complex than $\Phi_{k, \theta}$ or $T$ by considering the Pearson correlation coefficient.
-Finally, we present and discuss some numerical results for a focus system of this project: the double pendulum.
-We also consider other chaotic attractors.
-
-\section{Implementing $G_T$}
-The map $G_T$ is implemented by applying the functions $\Gamma$, $g$ and the projection function $\pi_2$ in the relevant order as described by the equations~\ref{eqns_from_data}.
-To reach this point, one must first learn $\Gamma$.
-After first normalising the data to ensure that it has zero mean value and lies within the range $[-1,1]$ and required zero-padding(see paragraph below for further discussion), we continue by driving the system with external input through the function $g$.
-
-From \eqref{eqn_driving} we know that $g(u,x) = (1-a) x + a \tanh(Au + \alpha B x)$. Recall that $g$ is SI-invertible. We refer to $A$ as the input matrix and $B$ the reservoir matrix.
-
-We make use of a Recurrent Neural Network (RNN) in our implementation to project the data on to a higher dimensional space, but we do not train the RNN in the conventional manner.
-\ednote{B: Should we say what the conventional manner is?} A few words about the RNNs: RNNs are widely used in the literature for temporal processing of data as a type of deep learning technique.
-They take the form of equations considered to be discrete-time version of continuous dynamical systems modelling a neural network of neurons \cite{jaeger2001echo}.
-
-We implement a feedforward network, (a network with no memory, unlike the RNN)\ednote{B: Expanded sentence s.t.
-FFN is in more context} required to learn $\Gamma$ in the \emph{Python} programming language and making use of the \emph{Scipy} library, alongside the \emph{Keras} package in the \emph{Tensorflow} library \ednote{B: Is this something you want me to write in?
-M: We do not implement the RNN, but the FN that learns Gamma through the Tensorflow.
-We called the RNN + FN set up a RCN in Adriaan's paper.
-FYI: FN's work only if there is a definite relationship between the $u_n$ and $u_{n+1}$ which is rare in temporal data.
-RNNs with ESP work when there is a definite relationship between $(\ldots,u_{n-1},u_n)$ and $u_{n+1}$ and is suited for temporal processing.}.
-
-If the RNN we will be using in the next step has $N$ neurons, then $A$ and $B$ are each $N\times{N}$ matrices.
-Inputs $u_n$ of dimension $K < N$ are padded by zeroes (as remarked earlier) to embed the input in a space of dimension $N$.
-As we do not need the entire left-infinite history of input, an arbitrary initialisation of the system with initial values followed by the driving of the system with the function $g$ means that the generated values will approach the actual solution elements in a uniform manner thanks to the UAP that we had discussed in the preceding chapter.
-
-Thirdly we prepare the network.
-Utilising a network $N$ neurons (the same value $N$ as in the dimensions of the matrices $A,B$).
-Learning of $\Gamma$ is done using the Adam Optimiser where the means square error (MSE) loss function is optimised.
-The system is trained several times (in our case 4) with incrementally smaller learning rates on each run (0.001 and thereafter divided by 10 at each run).
-Additional parameters such as the number of network-layers, neurons per layer and training-length are presented case-by-case below.
-
-We then reach the prediction step. Making use of $g$'s SI-invertibility and $g$ as defined in \eqref{eqn_driving}, we have the following expression for $u_n$ replicated from~\eqref{eqn_SI_RNN}.
-\begin{equation*}
-  u_n := A^{-1}\bigg(\overline{\tanh}^{^{-1}}\frac{1}{a}\Big(x_{n+1}-(1-a)x_n\Big) \bigg) - \alpha B x_n
-\end{equation*}
-
-The two model equations for learning from data~\eqref{eqns_from_data} allows us to determine what the state $x_{n+1}$ at time $n+1$ will be. 
-We specifically remind ourselves of \[x_{k+1}=\pi_2 \circ (\Gamma, \pi_2) \circ (\pi_2,g) (u_k,x_k)\]
-
-\section{Delay Coordinates}
-
-We also consider a more general problem of learning an attractor of dynamical system when only observations of an orbit are given.
-Explicitly, if $(W,T)$ is a dynamical system with dynamics generated by $w_{n+1}=Tw_n$, and if $\theta:W \to \mathbb{R}$ is an observable, then the task is to learn a dynamical system that is topologically conjugate to $(W,T)$ and predict $\theta(w_{m+1}),\theta(w_{m+2}),...$ using the data $\theta(w_{0}),\theta(w_{1}),\ldots,\theta(w_{m})$.
-So how do we do it?
-Suppose the input generated from the delay-coordinate map $\Phi_{\theta,2d}(\theta(w_{n})) := (\theta(w_{n-2d}),\ldots,\theta(w_{n-1}),\theta(w_{n}))$ is such that Takens delay embedding theorem (see Theorem~\ref{Thm_Takens}) holds, in which case, there exists a homeomorphism  $F_\theta: \Phi_{\theta,2d}(\theta(w_{n})) \mapsto \Phi_{\theta,2d}(\theta(w_{n+1}))$.
-Then if we feed  the input values $u_n := \Phi_{\theta,2d}(\theta(w_{n}))$ to a driven system $g$ that is SI-invertible and has USP, the induced system $(Y_F,G_F)$ would be topologically conjugate to the inverse-limit space of
-$(\Phi_{\theta,2d}(W), F_\theta)$ as in Theorem~\ref{Thm_CET}, and hence one could forecast $u_n,u_{n+1},\ldots$.
-In other words, one could forecast  the values $\theta(w_n), \theta(w_{n+1}),\ldots$ by feeding these observations as delay coordinates.
-The advantage of feeding delay-coordinates to a driven system is that the embedding is stable in the sense of global dissipativity that we have described in the previous chapter, i.e., $Y_F$ is made to contained in an attractor of a mapping induced on $Y_F$ that is induced by $g$.
-
-We note that the twin equations together~\ref{eqns_from_data} generate a model from data.
-For a chaotic system, a small amount of noise due either to the input-noise or computational error would cause the solutions from two accurate models with the same initial conditions to  diverge as they are not identical.
-\ednote{B: edited this sentence} However, statistically, they would be the same since due to ergodic theory all typical orbits have the same visiting frequency to any region of the phase space of the dynamical system.
-\ednote{B: I'll need to recap this sentence with you before presenting as it's new to me.} To illustrate that we plot the densities of the trajectories of the actual and predicted solutions.
-
-Our implementation of the reservoir(A) and input(B) matrices in the RNN makes use of sparse matrices.
-We discuss three specifics relating to this concept.
-Firstly, the matrices are assigned a variable density in terms of the amount of non-zero entries and are scaled to have values within the range $[-1,1]$.
-For this, we make use of \textit{Python's SciPy} library and specifically the \textit{Sparse, LinAlg} modules.
-In our experiments, we generally opted for a density ratio of 10\%-25\%.
-Secondly, we opted to enforce the requirement that both matrices be well-conditioned, i.e.
-that they have a condition number of \~1.
-The condition number of a matrix $M$ (denoted $cond(M)$) is the product of the matrix norms of the matrix $M$ and its inverse $M^{-1}$ and is a measure of the sensitivity of output of a matrix multiplication when an input is changed.
-\textbf{(Cite.)}\ednote{Ask M: Cite specific resource?} By ensuring a condition number close to 1, we also ensure that our driven function $g$, depending on the matrices $A$ and $B$, will not be sensitive to adjustments in the input.
-In so doing we then also guarantee another form stability.
-\ednote{B: Not complete enough.
-Edit further and cite}
-
-Lastly, we ensure the matrix $B$ has unit spectral radius.
-This is necessary due to the Causal Embedding Theorem~\ref{Thm_CET} which can then guarantee the existence of semi-conjugacy between the systems $(Y_T, G_T)$ and $(\widehat{U}_T, \widehat{T})$.
-In fact, the theorem actually requires that $\alpha{B}$ have spectral radius $<1$, so we impose the restriction $\alpha<1$ and consider only the matrix $B$'s spectral radius from thereon.
-
-Note that to learn $\Gamma: (x_{n-1},x_{n}) \mapsto u_n$ through our feedforward network, we could learn the principal components of $(x_{n-1},x_{n})$ as well. 
-We employ this since it provides slightly better results empirically but is not essential.
-
-Principal components are used only for a more efficient state representation to possibly reduce learning errors; it is not for a lossy approximation since we use all principal components \textbf{(Cite.)}\ednote{Ask M: Is there a specific resource to cite?}.
-More explicitly, denote $X_{1:N}$ as the matrix with the first $N$ states of the network data represented by row vectors.
-If $X_{1:N}=U\Sigma P^T$\ednote{Expand here} represents the singular value decomposition of $X_{1:N}$ \ednote{Ask: Is there a specific resource to cite?}, we then denote the principal component matrix $P$, and the principal components are given by
- $Z_{1:N}=X_{1:N}P.$
-These principal components are then used to train a feedforward neural network. Denote the row vectors of $Z_{1:N}$ by 
-$z_i^T, i=1,2,\cdots,N$ and the neural network by $NN$, the network is subsequently trained to learn an approximation of the map
-$NN: (z_{n-1},z_n) \mapsto  u_n.
-$
-The learnt approximation of $NN$ can be used to approximate $\Gamma$ since 
-$$
-NN\left( \begin{bmatrix} 
-P^Tx_{n-1} \\
-P^Tx_n
-\end{bmatrix}
-\right) = NN \circ 
-\begin{bmatrix}
-P^T & 0 \\
-0 & P^T 
-\end{bmatrix}\begin{bmatrix}
-x_{n-1}\\
-x_n
-\end{bmatrix} = u_n = \Gamma(x_{n-1},x_n).
-$$
-
-\section{Experimental Results}
-In this section we showcase numerical results pertaining to 3 attractors that were simulated using the methodology described.
-We first consider a simple physical system that exhibits complicated dynamics.
-\subsection{Double Pendulum}
-
-The Double Pendulum consists of two pendulums fastened to one another such that the system moves as a whole.
-The first (or higher pendulum) is attached to a fixed point while the second is attached to the endpoint of the first.
-(See below)
-This system is a classical example of a system exhibiting chaos and is often used as explanation of the very well-known Butterfly Effect.
-\ednote{B: Originally though a reference here was needed, but no longer think so.} Consider again~\ref{fig:dp_sdic}.
-The system has four variables: angle and angular velocity of each pendulum rod. An ideal system is assumed where the rods are massless and no friction in the pivots are considered.
-\begin{figure}[ht]
-  \includegraphics[scale=0.35]{_dp_setup.eps}
-  \centering
-  \captionof{figure}{Pictorial representation of the setup of a DP. Image taken from~\cite{DPSetup}}
-  \label{fig:dp_setup}
-\end{figure}
-
-The scalar equations of motion are derived using Newtonian physics and are reproduced from~\cite{DPFormulas} in~\ref{eqns_dp}.
-Each pendulum component has an associated mass of the head, $m$, and length of the rod, $L_i$.
-Acceleration due to gravity is denoted by $g$ and initial velocities by $\dot{\theta}$
-\begin{eqnarray}\label{eqns_dp}
-  \ddot{\theta_{1}}  = \frac{-g(2m_1+m_2)\sin{\theta_1} - m_2g\sin(\theta_1-2\theta_2) - 2\sin(\theta_1-\theta_2)m_2({\dot{\theta_{2}}}^{2}L_2 + {\theta_{1}'}L_1\cos(\theta_1-\theta_2))} {L_1(2m_1 + m_2 -m_2\cos(2\theta_1 - 2\theta_2))}
-  \\
-  \ddot{\theta_{2}} = \frac{2\sin(\theta_1-\theta_2)(\theta_{1}'^{2}L_1(m_1+m_2) + g(m_1+m_2)\cos\theta_1 + \dot{\theta_{2}}^{2}L_2m_2\cos(\theta_1-\theta_2))}{L_2(2m_1 + m_2 -m_2\cos(2\theta_1 - 2\theta_2))}
-\end{eqnarray}
-
-The second-order system is converted to a first-order ODE system and solved using the Runge-Kutta numerical method.  
-Here we chose the values $m_1=2.0$, $m_2=2.0$, $L_1=1.5$ and $m_2=1.5$ with initial velocities $v_1=1.5, v_2=-2$ for the respective pendulum heads.
-Undamped DP data-sets were generated and used in this project due to the requirement that the map $T$ be surjective that we must not forget. 
-An example of this would be the case of a DP exhibiting energy-loss (i.e.
-where the map $T$ is not surjective).
-In such a case, the attractor is a single point in the phase space when the double pendulum comes to rest.
-It must be mentioned once again that if we use data outside the attractor, learning is not reliable. To see this, consder the graphs ~\ref{fig:damped_pendulum} and ~\ref{fig:dp_notsurjective}.
-
-\begin{figure}[ht]
-  \includegraphics[scale=0.4]{_dp_dying.eps}
-  \centering
-\caption{$x$- and $y$-coordinates of the two pendulum heads plotted for a dataset that exhibits dying oscillations.
-The map generating the (four-dimensional data) for a damped pendulum is not surjective.
-Data obtained from chaotic double pendulum dataset generated from videos takes of experimental pendulums~\cite{asseman2018learning}}
-\label{fig:damped_pendulum}
-\end{figure}
-
-\begin{figure}[ht]
-  \includegraphics[scale=0.3]{_dpfail_nonsurj.eps}
-  \centering
-\caption{Failure to forecast or learn a map which is not surjective: Evolution of the $y$-coordinate of the bottommost pendulum plotted here for 15000 timestep.
-Plotted here the predicted value(red) versus the actual value(blue).
-The plots overlap for the first 10000 steps, the duration of training to learn the map $\Gamma$.
-Afterwards, the trajectories start diverging and it the predicted value no longer accurately represents the complexity or energy-loss of the actual data.
-}
-\label{fig:dp_notsurjective}
-\end{figure}
-
-Having spent ample time to discuss the challenges, we now change tone so as to satisfy the optimist and present a successful run of the DP-data to predict the future movement of the pendulum.
-
-The system was driven for 15500 steps, of which the first 500 were immediately discarded to simulate the network 'forgetting'.
-The driving function $g$ had input and reservoir matrices were implemented as sparse matrices filled to 20\%.
-Parameter-values of $\alpha=0.99$ and $a=0.5$ were used.
-The remaining 15000 datapoints were used to train a network with 128 neurons. 
-The network was initialised to contain 16 hidden layers each with a layer dimension of 64, activated with the tanh activation function.
-Training was accomplished using 256 epochs with batch sizes 128 and a delay of 1.
-See~\ref{fig:dp_success_traj} for results pertaining to the trajectory taken by the DP's components and then~\ref{fig:dp_success_density} to witness associated densities of predicted vs.
-actual values.
-
-\begin{figure}[ht]
-  \centering
-  \includegraphics[width=0.95\linewidth]{_dp_success_4coords_traj.eps} 
-  \captionof{figure}{Predicted(red) and actual(blue) trajectories of the $x$- and $y$-coordinates of each of the pendulum heads.
-From top to bottom are as follows: $x$-coordinate of topmost pendulum head, $y$-coordinate of topmost pendulum head, $x$-coordinate of lower pendulum head, $y$-coordinate of lower pendulum head.
-  These graphs were constructed by predicting the DP 1000 timesteps into the future and in so doing illustrating the long-term consistency and accuracy of the learnt system.
-Here we lock onto the near-exact trajectory for about 300 timesteps and stay close up until about 600 seconds.
-}
-  \label{fig:dp_success_traj}
-\end{figure}
-\begin{figure}[ht]
-  \centering
-  \includegraphics[width=\linewidth]{_dp_success_4coords_hist.eps}
-  \captionof{figure}{Densities of predicted and actual values of $x$- and $y$-coordinate of pendulum heads.}
-  \label{fig:dp_success_density}
- \end{figure}
-
-We now consider an experiment where noise is added to the system and find it to be incredibly robust.
-We employ the same parameters as before and add noise from a normal-distribution with mean zero and standard deviation equal to 0.1 which translates to roughly 39dB's of noise.
-(For the signal-to-noise ration we adopt the formula $\text{SNR}_{dB}=10\text{log}_{10}\big(\frac{P_\text{{signal}}}{P_{\text{noise}}}\big)$ where $P_\text{{signal}}$ refers to the variance observed in the normalised data and $P_{\text{noise}}$ represents the variance of the added noise.)
-
-Once again consider the actual and predicted trajectories for the noisy DP dataset in Figure~\ref{fig:noisydp_success_traj}.
-Next we consider the densities of the predicted and actual values in Figure~\ref{fig:noisydp_success_density} and observe that they display highly similar distributions.
-
-\begin{figure}[ht]
-  \centering
-  \includegraphics[width=0.95\linewidth]{_dp_noise_FigCD_2030.eps} 
-  \captionof{figure}{Predicted(red) and actual(blue) trajectories of the $x$- and $y$-coordinates of each of the pendulum heads with noise added from a normal distribution with mean zero and standard deviation 0.1, equating to approximately 39dB's of noise.
-  These graphs were constructed by predicting the DP 1000 timesteps into the future and in so doing illustrating the long-term consistency and accuracy of the learnt system.
-Here we lock onto the near-exact trajectory for about 250 timesteps and stay close up until about 400 seconds.
-}
-  \label{fig:noisydp_success_traj}
-\end{figure}
-\begin{figure}[ht]
-  \centering
-  \includegraphics[width=\linewidth]{_dp_noise_2B_2030.eps}
-  \captionof{figure}{Densities of predicted and actual values of x-coordinate of first pendulum head.}
-  \label{fig:noisydp_success_density}
- \end{figure}
-
- Lastly, a run was done where only a single coordinate (the $y$-coordinate of the lower pendulum head) was fed into the system and we can illustrate through the obtained success that the coordinate contains all the information in its past states necessary to forecast its future evolution.
- \textbf{Insert graphs here. Had to quickly change something here.}
-
-\subsection{Additional Attractors: Clifford \& Thomas}
-We follow the methodology presented in \cite{manjunath2021universal} but opt to consider attractors different from the Lorenz system and H\'enon map in this project.
-
-\subsubsection{Thomas' Cyclically Symmetric Strange Attractor}\label{Thomas_Attractor}
-Thomas' Cyclically Symmetric Strange Attractor, a 3D attractor proposed in 1999 by Ren\'e Thomas in \cite{ThomasAttractor}, is described by a set of three equations which is distinctly reminiscent of the Lorenz system, another deterministic three-equation model exhibiting chaotic behaviour.It has a single parameter $\beta$ and has been shown to transition to chaotic behavious when $\beta<0.208186$~\cite{Thomas_BetaParameter}.
-\begin{eqnarray}\label{eqns_thomas}
-  x_{n+1} = \sin(y_n) - \beta{x_n} \\
-  y_{n+1} = \sin(z_n) - \beta{y_n} \\
-  z_{n+1} = \sin(x_n) - \beta{z_n}
-\end{eqnarray}
-
-The behaviour of the system with a $\beta$-parameter value of 0.1056 was simulated by scaling it to fit inside the interval $[-1,1]$.
-Two cases were considered: noise-free and the one perturbed by noise.
-Noise was generated from a normal distribution with mean 0 and standard deviation 0.05, which to translates approximately 33dB.
-($\text{SNR}_dB$ is calculated exactly as for the DP-data above)
-
-A sequence of 3000 observations in both the clear and noisy data-sets respectively were fed into the system with 300 entries discarded to simulate the network's loss of memory.
-The next 9000 steps were predicted and compared with true data.
-
-\begin{figure}[ht]
-  \centering
-  \minipage{0.5\linewidth}
-    \centering
-    \includegraphics[width=0.8\linewidth]{_thomas_noisy_0.1056.eps}
-  \endminipage\hfill
-  \minipage{0.5\linewidth}
-    \centering    \includegraphics[width=0.8\linewidth]{_thomas_clean_0.1056.eps}
-  \endminipage
-  \captionof{figure}{True(red) and predicted(blue) trajectories for the Thomas attractor with parameter value $\beta=0.1056$.
-To the left: Noise added from normal distribution with mean zero and standard deviation 0.05.
-Right: No noise added to dataset.}
-\end{figure}
-
-The RNN was initialised as follows: the dimension of the network is set equal to 500 neurons and 12 hidden layers of dimension 64 each were initialised.
-Training is realised through 150 epochs of batch sizes of 128 each; the map is learnt via the Adam Optimiser using finer and finer learning rates (0.001 and then divided by 10 at every iteration).
-
-\begin{figure}[ht]
-  \centering
-  \includegraphics[scale=0.35]{_Thomas_1.eps}\caption*{Predicted trajectories of the $x$- and $y$-coordinates for the Thomas attractor with parameter-value $\beta=0.1056$ demonstrate empirically the ability to predict the evolution of the trajectory for the next an estimated 100 timesteps into the future near-exactly.
-Here no noise was added.}
-  \includegraphics[scale=0.5]{_Thomas_3.eps}\caption*{Represented here the learnt (blue) and actual (red) densities of the first coordinate ($x$-coordinate) of the dataset for the Cyclically Symmetric Attractor with parameter-value $\beta=0.1056$.}
-\end{figure}
-
-\subsubsection{Fractal Dream Attractor}
-
-Another 2-equation discrete-time map named the Fractal Dream Attractor, or more commonly known as the Pickover Map (first discovered by Clifford A.
-Pickover and discussed in his fascinating book "Chaos in Wonderland"\cite{PickoverChaos}), was also considered.
-\begin{eqnarray}\label{eqns_clifford}
-  {x_{n+1}=\sin(ay_n) + c\cos(ax_n)} \\
-  {y_{n+1}=\sin(bx_n)+d\cos(by_n)}
-\end{eqnarray}
-
-\begin{figure}[ht]
-  \centering
-  \includegraphics[width=1.0\textwidth,left]{_Clifford_1_nonoise.eps}
-  \caption*{These graphs were constructed by predicting the Clifford system 1000 steps into the future and in so doing illustrating the long-term consistency and accuracy of the learnt system.
-As perceived here, we are able to lock on to the trajectory of the Clifford map almost exactly for the first 25 steps.}
-  \minipage{0.5\textwidth}
-      \centering
-      \includegraphics[width=\linewidth]{_Clifford_2_nonoise.eps}
-  \endminipage\hfill
-  \minipage{0.5\textwidth}
-    \centering
-    \includegraphics[width=\linewidth]{_Clifford_3_nonoise.eps}
-  \endminipage
-  \caption{The bottom row, left, shows a plot of the true attractor(red) and the predicted values(blue).
-Bottom row, right, shows the distribution of the first coordinate of the learnt system (blue) and the actual Clifford system (red).
-}
-  \label{fig:Clifford}
-\end{figure}
-
-The behaviour of the system with a parameter values of $x_0=0$, $y_0=0$, $a = -1.7$,  $b = 1.8$, $c = -1.9$ and $d = -0.4$ was examined.
-Data was scaled to fit inside the interval $[-1,1]$. Only the noise-free case was investigated. 
- A sequence of 25000 observationswas fed into the system with 15000 entries discarded to simulate the network's loss of memory. The next 1000 steps were predicted and compared with true data.
-The RNN was initialised exactly as for Thomas' Cyclically Symmetric Strange attractor: the dimension of the network is set equal to 500 neurons and 12 hidden layers of dimension 64 each were initialised.
-Training is realised through 150 epochs of batch sizes of 128 each; the map is learnt via the Adam Optimiser using finer and finer learning rates (0.001 and then divided by 10 at every iteration).
-
-\section{Functional Complexity Reduction in Noninvertible maps}
-
-% Besides being able to learn the action of the Koopman operator exactly, our setup has other advantages over selecting or optimizing the choice of observables as in the Extended Dynamic Mode Decomposition (EDMD) algorithm (see, [21, 22]).
-We can alter the functional complexity of G T by changing the number of observables or equivalently by altering the dimension of X – in the case of a RNN implementation of g, by simply changing the number of neurons in the network.
-Empirically, increasing the dimension of X is found to increase the linear relationship (or intuitively reduces the functional complexity of G T ) that is measured as a generalization of the Pearson correlation coefficient to random vectors (see, [44]) between (x_{n−1} , x_{n}) and G_T (x_{n−1} , x_n), and such numerical evidence is tabulated in [11, Table 1].
-In contrast, expanding the set of observables while using EDMD to learn a map with a lower functional complexity is a highly involved task, not least because so much that one often does not know how to guess a new observable, as well as the fact that adding observables does not necessarily retain the invariance of the span of the observables.
-
-% Empirical evidence shows that a linearity measure like a Pearson correlation coefficient points at reduced functional complexity while learning the dynamics in the state space of the driven system.
-
-% Empirically (see Table ??), increasing the dimension of X is found to increase the linear relationship
-% (or intuitively reduce the functional complexity of G T ) that is measured as a generalization of the Pearson correlation coefficient to random vectors (see, [?]) between (x n−1 , x n ) and G T (x n−1 , x n ).
-If Σ a and Σ b denotes the covariance matrices of the vectors (x n−1 , x n ) and G T (x n-1 , x n ) respectively, and if Σ ab denotes the covariance matrix between the vectors (x n-1 , x_n ) and G_T (x n-1 , x_n ), then the multidimensional 18correlation coefficient is computed by using the traces of these matrices:
-% \[\pho = \frac{tr(\sum_{ab})}{tr(\sqrt{\sum_{a}\sum_{b}})}\]
-
-% This multidimensional correlation coefficient satisfies most of the well-known properties of the one-dimensional Pearson coefficient.
-In particular, ρ = ±1 if and only Y = AX + b for some invertible matrix A and vector b.
-Hence ρ, and in practice, the estimator ρ̂, can be used to measure the linear relationship between two random vectors of the same dimension and thus serves as an indicator of functional complexity.
-
-% We remark that the estimation of such linear relationship in (see Table ??) obtained by sample correlations alone is justified whenever {u n } is a realization of an ergodic process since then when g has the USP, any solution {x n } is also a realization of an ergodic process [?].
-In the ergodic input case the generalized Pearson correlation coefficient (between (x n−1 , x n ) and G T (x n−1 , x n )) is independent of n.
-
-In the case of non-invertible maps, we note that the map $G_T$ is a homemorphism although $T$ is not.
-Further, empirically it is found that $G_T$ is much more linear than $T$ as measured by what is called the multidimensional correlation coefficient $\rho$.
-This coefficient is a generalization of the Pearson correlation coefficient for real-valued random variables to random vectors (see, \cite{puccetti2019measuring}).
-We state teh details of computing this coefficient between
-$(x_{n-1},x_n)$ and $G_T(x_{n-1},x_n)$. 
-If $\Sigma_{a}$ and $\Sigma_{b}$ denotes the covariance matrices of the vectors $(x_{n-1},x_n)$ and $G_T(x_{n-1},x_n)$ respectively, and if $\Sigma_{ab}$ denotes the covariance matrix between the vectors $(x_{n-1},x_n)$ and $G_T(x_{n-1},x_n)$, then the multidimensional correlation coefficient is computed by using the traces of these matrices:
-\[
-    \rho= \frac{\text{tr}({\Sigma_{ab}})}{\text{tr}({\sqrt{\Sigma_a\Sigma_b}})}.
-\]
-This multidimensional correlation coefficient satisfies one of the desired and well-known properties of the one-dimensional Pearson coefficient.
-That is, $\rho=\pm 1$ if and only if $Y\overset{d}{=}AX+b$ for some invertible matrix $A$ and vector $b$.
-Hence $\rho$, and in practice can be used  to measure the linear relationship between two random vectors of the same dimension and thus serves as an indicator of functional complexity.
-For the non-invertible system defined by the Clifford attractor, we indicate the difference by which $\rho$ changes for $G_T$ in Table~\ref{tbl_attractorsPearson}.
-Following \cite{manjunath2021universal}, we say that this multidimensional correlation coefficients $\rho$ to evidence a general reduction in the functional complexity of the map $G_T$ emerging due to a RNN.
-In the
-Table ~\ref{tbl_attractorsPearson}, the second column corresponds to the amount of artificial neurons (the dimension of $X$) in the implemented RNN.
-
-\begin{center}
-  \begin{table} 
-      \scalebox{0.75}
-      {\begin{tabular}{|c|c| c c |} 
-          \hline
-          Input & Dimension of $X$ & $u_n$ vs $u_{n+1}$ 
-          & $\begin{bmatrix}
-              x_{n-1}\\
-              x_n
-          \end{bmatrix}$ vs $\begin{bmatrix}
-              x_n\\
-              x_{n+1}
-          \end{bmatrix}$ \\
-          \hline      
-          % \hline
-          % $(w_n)$ is the Lorenz states, sampled every 0.1 timestamps. & & &\\
-          % \multirow{3}{*}{$u_n=\frac{1}{100}w_n$} & 10 &0.9311  & 0.9731\\
-          %     & 100 &  & 0.9934\\
-          %     & 1000 & & 0.9930\\              
-          % \multirow{3}{*}{$u_n=\frac{1}{10}(\sin(0.1w_{n,x})+\sin(0.1w_{n,y})+\sin(0.1w_{n,z}))$} 
-          %     & 10 &0.8401  & 0.9392\\
-          %     & 100 & & 0.9616\\
-          %     & 1000 & & 0.9737\\
-          % \hline
-
-               \hline  
-               $(w_n)$ comes from the Clifford Attractor \\ (for $a = -1.7; b = 1.8; c = -1.9; d = -0.4$): & & & \\
-        {$w_{n+1}= \begin{bmatrix} \sin(aw_{n,y}) + c\cos(aw_{n,x}) \\ 
-                                                                    \sin(bw_{n,x})+d\cos(bw_{n,y}) \end{bmatrix}$} & & & \\
-          \multirow{3}{*}{$u_n = w_n-\overline{w}$}
-              & 10 & -0.21713 & 0.78719 \\
-              & 100 & &  0.8334 \\
-              & 1000 & & 0.84854 \\
-          % \hline 
-          % $(w_n)$ comes from the Logistic map: & & & \\
-          % {$ w_{n+1}  = 4w_n(1-w_n)$} & & & \\
-          % \multirow{3}{*}{$u_n = w_n - 0.5$} 
-          %     & 10 &-0.0314  & 0.5577\\
-          %     & 100 &  & 0.7150\\
-          %     & 1000 & & 0.7826\\
-          % \hline
-          % \hline
-          % $(w_n)$ comes from the Svensson Attractor \\ (for $a = -1.7; b = 1.8; c = -1.9; d = -0.4$): & & & \\
-          % {$w_{n+1}= \begin{bmatrix} d\sin(aw_{n,x}) - \sin(bw_{n,y}) \\ 
-          %                                                             c\cos(aw_{n,x})+\cos(bw_{n,y}) \end{bmatrix}$} & & & \\
-          %   \multirow{3}{*}{$u_n = w_n-\overline{w}$}
-          %       & 10 & -0.2243 & 0.49698 \\
-          %       & 100 & &  0.63310 \\
-          %       & 1000 & & 0.71151 \\
-           \hline
-                \end{tabular}} %\ednote{http://paulbourke.net/fractals/peterdejong/}
- \label{Table_FC}
- \caption{Following \cite{manjunath2021universal}, we determine the multidimensional correlation coefficients $\rho$ to evidence a general reduction in the functional complexity of the map $G_T$ emerging due to a RNN.
-Rows correspond to distinct dynamical systems considered in the experiments
-      The second column corresponds to the amount of artificial neurons (the dimension of $X$) in the implemented RNN.
-      The last two columns comprise numerical estimates of $\rho$ for the relevant vectors. The linear association between $u_n$ vs $T(u_{n})$, as well as the linear relationship $\begin{bmatrix}
-              x_{n-1}\\
-              x_n
-          \end{bmatrix}$ vs $G_T \left( \begin{bmatrix}
-              x_{n-1}\\
-              x_n
-          \end{bmatrix}\right)$ may be contrasted by the reader. 
-          Finally, the RNN's dimension (from \eqref{Seq_RNN}) is provided to exhibiting that an increase in dimension of the driven system will 
-          typically result in a map $G_T$ with less functional complexity.}.
-      \end{table}\label{tbl_attractorsPearson}
-\end{center}
-
-\chapter{Conclusion}\label{ch6}
-\input{chapter6.tex}
-
-\bibliographystyle{pnas-new}
-\bibliography{pnas}
-
-\end{document}

### 2022-01-17 projectReport/chapter4.tex
Cleaner setup

--- 2022-01-17 draft_16Jan/hiscurrent.tex
@@
-\documentclass[a4paper,12pt,twoside]{report}
-% \documentclass[a4paper,12pt,twoside]{book}
-\pagestyle{headings}
-\usepackage[utf8]{inputenc}
-\usepackage{filecontents}
-\usepackage[many]{tcolorbox}
-\usepackage{graphicx}
-\usepackage{harpoon}
-%\usepackage{apacite} 
-\usepackage{xcolor}
-\usepackage{latexsym}
-\usepackage{amssymb}
-\usepackage{amsmath}
-\usepackage{mdframed}
-%\usepackage[authoryear]{natbib}
-\usepackage{hyperref, url}
-\usepackage[show]{ed}
-\usepackage{subfigure}
-\usepackage[ntheorem]{empheq} 
-\usepackage{amsthm, amssymb, amsfonts, latexsym}
-% \usepackage{pst-all}
-
-\usepackage{enumitem}
-\usepackage{multirow}
-\usepackage{longtable}
-\usepackage[export]{adjustbox}
-% $f\arrowvert_A$ To restrict functions
-
-\setlength{\parskip}{1.7em}
-\setlength{\parindent}{0em}
-
-%\setlength{\baselineskip}{20pt}
-
-\setlength{\textwidth}{429.75499pt}
-\setlength{\textheight}{643.20255pt}
-\setlength{\oddsidemargin}{5 mm}
-\setlength{\evensidemargin}{5 mm}
-\setlength{\topmargin}{0 mm}
-\setlength{\headsep}{0 mm}
-\setlength{\headheight}{0 mm}
-
-\usepackage{pst-node}
-
-\newcommand{\citealtt}[1]{\citeauthor{#1},\citeyear{#1}}
-\newcommand{\myycite}[1]{\citep{#1}}
-
-\mathchardef\mhyphen="2D
-
-%\newcommand{\cev}[1]{\reflectbox{\ensuremath{\vec{\reflectbox{\ensuremath{#1}}}}}}
-
-\usepackage{pst-node,graphicx,pst-blur}
-%\uspackage{auto-pst-pdf}
-%\usepackage{tikz-cd}
-
-\makeatletter
-\DeclareRobustCommand{\cev}[1]{%
-  \mathpalette\do@cev{#1}%
-}
-\newcommand{\do@cev}[2]{%
-  \fix@cev{#1}{+}%
-  \reflectbox{$\m@th#1\vec{\reflectbox{$\fix@cev{#1}{-}\m@th#1#2\fix@cev{#1}{+}$}}$}%
-  \fix@cev{#1}{-}%
-}
-\newcommand{\fix@cev}[2]{%
-  \ifx#1\displaystyle
-    \mkern#20mu
-  \else
-    \ifx#1\textstyle
-      \mkern#20mu
-    \else
-      \ifx#1\scriptstyle
-        \mkern#26mu
-      \else
-        \mkern#26mu
-      \fi
-    \fi
-  \fi
-}
-
-\makeatother
-
-\newcommand{\xcm}{\epsfxsize=3.1cm}
-\newcommand{\fig}[1]{\epsfbox}
-% \newcommand{\bp}{\begin{minipage}{3.1cm}}
-% \newcommand{\ep}{\end{minipage}}
-
-\newtheorem{Definition}{Definition}[]
-%\newmdtheoremenv{Theorem}{Theorem}[]
-\newtheorem{Theorem}{Theorem}[]
-\newtheorem{Lemma}{Lemma}[]
-\newtheorem{Proposition}{Proposition}[]
-\newtheorem{Corollary}{Corollary}[]
-\newtheorem{Remark}{Remark}[]
-\newtheorem{Example}{Example}[]
-
-%Shortcut symbols
-\newcommand{\Ftheta}{\ensuremath{F_\theta}}
-
-\makeatletter 
-\renewcommand{\thefigure}{\@arabic\c@figure}
-\makeatother
-\usepackage{xr}
-
-%%%% PNAS GRAPHICS SETUP- FROM THE PNAS STYLE LATEX FILE
-\RequirePackage{graphicx,xcolor}
-\RequirePackage{colortbl}
-\RequirePackage{booktabs}
-\RequirePackage{algorithm}
-\RequirePackage[noend]{algpseudocode}
-\RequirePackage{changepage}
-\RequirePackage[twoside,%
-				letterpaper,includeheadfoot,%
-				layoutsize={8.125in,10.875in},%
-                layouthoffset=0.1875in,%
-                layoutvoffset=0.0625in,%
-                left=38.5pt,%
-                right=43pt,%
-                top=43pt,% 10pt provided by headsep
-                bottom=32pt,%
-                headheight=0pt,% No Header
-                headsep=30pt,%
-                footskip=25pt,
-                marginparwidth=38pt]{geometry}
-\RequirePackage[labelfont={bf,sf},%
-                labelsep=period,%
-                figurename=Fig. ]{caption}
-\setlength{\columnsep}{13.5pt} % Distance between the two columns of text
-\setlength{\parindent}{12pt} % Paragraph indent
-
-%% Figure caption style
-\DeclareCaptionFormat{pnasformat}{\normalfont\sffamily\fontsize{7}{9}\selectfont#1#2#3}
-\captionsetup*{format=pnasformat}
-
-%%% GREYBOX AROUND FIG
-\definecolor{lightgray}{gray}{0.95}
-\newcommand\greybox[1]{%
-  \vskip\baselineskip%
-  \par\noindent\colorbox{lightgray}{%
-    \begin{minipage}{\textwidth}#1\end{minipage}%
-  }%
-  \vskip\baselineskip%
-}
-
-\newtcolorbox{paperbox}[1][]{
-    enhanced,
-    colback=white,
-    boxrule=0pt,
-    boxsep=0pt,
-    arc=0mm,
-    width=0.8\linewidth,
-    fuzzy shadow={0mm}{-4pt}{-4pt}{1mm}{black!30!white},
-    #1
-}
-
-% \externaldocument{Supp_Koopman-0808}
-\begin{document}
-
-\title{Data-Driven Modelling \& \\ Forecasting of Chaotic Systems}
-
-\author{B.L. Nortier} 
-\date{January, 2022}
-\date{}
-\maketitle
-
-\begin{abstract}
-  It is a well-established practice to collect measurements from underlying artificial, natural, and physical systems through observables in order to realise a dynamical model with greater descriptive power.
-  Many data-driven approaches to forecast the future evolution of systems exist.
-However, many of these methods rely on an initial understanding of the system and we often do not fully understand the many complexities of more intricate systems that we interact with on a daily basis.
-
-  In this project, the base assumption is that data is originates from  some discrete-time dynamical system, particularly one of a chaotic nature but no assumptions are made pertaining to a specific dynamical system.
-  A general learning problem is formulated and a well-known approach to the problem, alongside its practical limitations, is then considered.
-  The notion of driven dynamical systems is introduced, and we subsequently establish the theoretical existence of a dynamical system possessing certain properties which guarantees it to be topologically conjugate (`dynamically equivalent`) to the dynamics of the underlying system from which measurements were originally taken.
-  Our methodology is implemented practically by making use of recurrent neural networks, a form of deep learning, to learn a map topologically equivalent to the underlying system that forecasts future states.
-  In so doing we present the workability of an approach to obtain long-term topological and statistically consistent predictions for simple physical systems as well as a some chaotic attractors.
-  \end{abstract}
-
-  \tableofcontents
-
-\chapter{Introduction}\label{ch1}
-
-Experimenting with biological, physical, and artificial systems to generate a more informative dynamical model is a well-established practice in modern science.
-Traditional methods for modelling physical systems are based on laws of physics that are based either on empirical relationships or intuition.
-For systems that evolve with time, physical laws yield mathematical equations that govern how the quantities evolve with time.
-However, our world is much more complex than that which can be distilled into elegant equations.
-We do not fully understand many of the more complex systems, nor do they provide us with a good physical intuition of the underlying principles governing system dynamics.
-To complicate this, the underlying systems we observe often display sensitive dependence on initial conditions despite having highly similar initial conditions.
-This is because differing orbits diverge quickly and to such an extent that it becomes seemingly impossible to retrace their steps back to the original conditions.
-Even the presence of usually `negligible' computational noise or measurement error renders long-term pointwise prediction infeasible.
-We encounter many difficultiesin the process of modelling such systems:
-\vspace{-8mm}
-\begin{enumerate}[noitemsep, label=\roman*.]
-  \item One may not have access to the complete states of the systems
-  \item The system may be described by functions that behave wildly because their graphs have a wild oscillatory behaviour, i.e., a large functional complexity.
-\end{enumerate}
-
-Models derived primarily from data can be classified into three categories: 
-\vspace{-8mm}
-\begin{enumerate}[noitemsep, label=\roman*.]
-  \item  Interpretable models (i.e., they establish relationships between internal physical mechanisms), 
-  \item Partially interpretable models capturing some modes of the dynamics, 
-  \item Non-interpretable models, defined as such mainly due to the fact that they are defined on a different phase space that is usually high-dimensional. 
-\end{enumerate}
-
-Examples in the literature attempting to forecast data from such systems have tried many different approaches with differing degrees of success. 
-When we have complete access to states of the system, an ordinary differential equation can be obtained from data (\cite{brunton2016discovering, champion2019data} and \cite{small2002modeling,xu2006modeling}) wherein one could approximate the vector field by a library of functions to obtain interpretable models.
-
-Recent partially interpretable models available in the literature have been based on the Koopman operator (see~\cite{koopman1932dynamical,budivsic2012applied}) to employ observables mapping the data onto a higher-dimensional space.
-This makes the dynamics in the higher-dimensional space more amenable for approximation by a linear transformation.
-Such methods do not guarantee exact reconstructions for nonlinear models, and in practice provide poor long-term consistency for a large class of chaotic dynamical systems \cite{Supp}.
-
-The non-interpretable models include the delay embedding and the machine learning algorithms.
-For example, one could learn a system conjugate to the underlying system by applying the Takens delay embedding method of delay-coordinates~\cite{takens1981detecting} when one has "good" observations from the system; this learnt system could then be used to forecast the observed data.
-Takens delay embedding theorem~\cite{takens1981detecting} and its various generalisations (see,~\cite{sauer1991embedology, stark1999delay, gutman2018embedding}) establishes the learnability of a system constructed by concatenating sufficiently large previous time-series observations of a dynamical system into a vector (called delay coordinates).
-This then confirms the existence of a map on the space of delay coordinates equivalent (or topologically conjugate) to the underlying map from which the observed time-series was first obtained.
-Although topological conjugacy guarantees an alternate representation of the underlying system, the quality still depends on numerous parameters, making the comprehension of the dynamics unreliable (see, [?]).
-One reason for this fragility is that the embedded attractor in the reconstruction space is not always an attractor of the map learnt in the reconstruction space, despite unquestionably being an invariant set.
-When the embedded object is not explicitly known to be an attractor (as explained in Chapter~ \ref{ch3}), it can cause predictions to fail.
-
-Practically, the application of Takens embedding involves learning a map through some technique, and consequently one wishes that these would have low functional complexity\cite{manjunath2021universal}, i.e., functions with fewer oscillatory graphs.
-On the other hand, pure machine learning methodology processes temporal information (like the echo state networks~\cite{Manju_ESP, Manju_IEEE, grigoryeva2018echo}) by mapping data onto a higher dimensional space for further processing.
-Although they perform  well on forecasting some dynamical systems, they fail completely on others (\ednote{Ask for resources here}) as there is often no guarantee that the right function was learnt during training.
-
-This project involves implementing and analysing non-interpretable models that can guarantee exact reconstruction. 
-The project work concerns the study and implementation of a method (\cite{manjunath2021universal}) that incorporates learning a function by mapping the data on to a higher dimensional space using what is called a driven dynamical system (See Chapter~ref{ch.4}).
-With a clear understanding of how the data is mapped onto the higher dimensional space, the method then permits the learning of a dynamical system topologically conjugate to that of the underlying system.
-Instead of linear regression as in the training of echo state networks, deep learning methods are employed to learn the correct function.
-With slight modifications to the implementation in the paper (\cite{manjunath2021universal}), we show that one can construct accurate non-interpretable models with the ability to reconstruct attractors from more hard-to-forecast systems like the double pendulum.
-(The forecasting of the time-series from a double pendulum has not been reported before.) Moreover, we also demonstrate that long-term statistical consistency is preserved.
-
-\emph{By solidifying the mathematical underpinnings of our theory, we hope to guarantee the ability to construct models with predictive power ranging from molecular biology to neuroscience.
-in the near future.
-(We can modify this sentence when the report is completed.)}
-
-The more intricate details of proofs are referred to where relevant throughout. This report is organised as follows: 
-\newline In Chapter~\ref{ch2}, we recall the definition of a discrete-time dynamical system, how a discrete-system arises from a flow of an ODE and then proceed to define the inverse-limit space and topological conjugacy of autonomous systems.
-\newline Chapter~\ref{ch3} introduces the problem of forecasting dynamical systems, states the Takens delay embedding theorem, and discusses various issues faced while forecasting. 
-\newline In Chapter~\ref{ch4}, a driven dynamical system is defined and discuss the properties of a specific class of driven dynamical systems that we make use of in this project.
-\newline Finally, in Chapter~\ref{ch5} we show the implementation of these forecasting methods, and conclusions are provided in Chapter~\ref{ch6}.
-
-\chapter{Discrete-time Dynamical Systems}\label{ch2}
-
-This chapter briefly describes a discrete-time dynamical system and what it means to exhibit chaos. We refer to \cite{devaney2018introduction, de2013elements} for more details.
-
-At its most elementary level, a dynamical system is just something that evolves deterministically through time.
-In the context of this project, deterministic refers to the fact that a system evolves according to specified rules rather than based on random events.
-Dynamical systems arise in a variety of situations.
-A continuous-time dynamical system describes the states for all values of the time.
-Specifically, if the motion of a pendulum in which the quantities such as the angular position and angular momentum are known at all times, then it is a continuous-time dynamical system.
-The equations of the dynamical system can take the form of one or more ordinary differential equations that determine the relevant quantities at any future time if we know the initial location and momentum.
-In ecology, discrete-time dynamical systems are widely used to model population growth.
-The model in this case is a function calculating the following generation's population given the population of the previous generation.
-If we know the starting population, we may once again calculate the population at any time in the future.
-
-Formally, a function $T: U \to U$, where $U$ is some set is  a \emph{discrete-time dynamical system} and its iterates $\{u,Tu,T^2u,\ldots\}$, where $T^n$ denotes the $n$-fold composition of $T$ with itself, describe the evolution of an initial condition $u\in U$ (Note that we frequently drop the brackets and denote $T(u)$ by $Tu$ so as to simplify notations).
-
-Continuous-time dynamical systems modelled using ordinary differential equations can give rise to discrete-time dynamical systems.
-To see this, consider a differential equation $\dot{x} = f(x)$, $f: \mathbb{R}^n \to \mathbb{R}^n$, $n\in\mathbb{N}$ given to have a unique solution passing through
-each point $x\in\mathbb{R}^{n}$, call it $x(t)$ where $x(0)=x_0$.
-
-\begin{Definition}
-  [\bf Flow of an Equation] \label{Dfn_Flow}\rm
-  The flow of the equation  $\dot{x} = f(x)$  is defined to be a mapping $\varphi: \mathbb{R}^n \times \mathbb{R} \to \mathbb{R}$ where $\varphi(x_0,t)= x(t)$ for the solution $x(t)$ with $x(0)=x_0$ 
-\end{Definition}
-
-By fixing $t=K \in (0,\infty)$, we can define the \emph{time-$K$ map} as  $T(x):= \varphi(x_0,K)$, and it is easily verified that $T\circ T(x_0) = \varphi(x_0,2K)$.
-In general the $m^{\mbox{th}}$ iterate of $x_0$ under $T$ would be the value of the solution of the ODE evaluated at time $mK$ with the initial condition $x_0$, i.e.
-$\varphi(x_0, mK)$.
-Thus ordinary differential equations give rise to a discrete-time dynamical system by sampling the value of the solution $x(t)$ at time intervals $K$ units apart.
-
-A numerical discretization of a differential equation can also give rise to a discrete-time dynamical system.
-For instance, Euler's method approximates $\dot{x}(t)$ by $(x(t+h)-x(t))/h$; if $h$ is fixed throughout, the solution of a differential equation $\dot{x}=f(x)$
-can be approximated at the time instant $t+(m+1)h$ by iterating the equation 
-\begin{equation}
-  x(t+(m+1)h) = x(t+mh) + h f(x(t+mh))
-\end{equation}
-
-Adopting more succinct notation by replacing $x(t+mh)$ with $u_m$, we rewrite the above equation as
-\begin{equation}
-u_{m+1} = u_m + hf(u_m)
-\end{equation}
-or in even simpler terms as the discrete-time dynamical system with map $T(u) = u + hf(u)$, where $T$, $u$ and $f$ are understood to be as above.
-
-Of course, discrete-time dynamical systems need not always arise through a differential equation.
-Once again, we may consider the field of ecology, where discrete dynamical systems are often directly derived or assumed.
-\ednote{Ask for resource}
-There is a school of thought that advocates discrete-time dynamical systems to be more natural for modelling real-world observations than differential equations.
-We refer the interested reader to~\cite{saber2010introduction}.
-
-\section{Invariant Sets}
-
-A core concept in the study of dynamical systems is that of invariance. 
-\begin{Definition}
-  [\bf Invariant Set]\label{Dfn_InvariantSet}\rm
-  Given a discrete-time dynamical system $T: U \to U$, a subset $A \subset U$ is said to be an \emph{invariant set} if $T(A) =A$. 
-\end{Definition}
-
-We also define the orbit of a function $T$.
-\begin{Definition}
-  [\bf Orbit of $T$]\label{Dfn_Orbit}\rm
-  The orbit of $T$ is to be defined the sequence $\bar{u} = \{u_n\}_{n\in \mathbb{Z}}$ obeying the update equation, $u_{n+1}=Tu_n$, $n \in \mathbb{Z}$. 
-\end{Definition}
-
-Two examples of invariant sets include a fixed point where $Tu=u$ for  $u\in U$, and a periodic orbit, i.e., a set of iterates $\{u,Tu, T^2u,\ldots,T^pu\}$, where $T^{p+1}u=u$ for some $p\in\mathbb{Z}$.
-The entire space $U$ could be also be invariant.
-Consider for example the space $U=[0,1]$, where $Tu=4u(1-u)$ and then $U$ is invariant (as every $u\in{U}$ can be written as $u = 4x(1-x)$ for some $x\in{U}$).
-
-We may learn a great deal about the iterates of a dynamical system by considering the types of invariant sets of a discrete-time dynamical system.
-For example, if $U=[0,1]$ has map $Tu= u/2$, then the only invariant set is $\{0\}$, and all orbits approach this invariant set as time flows in the forward direction.
-Indeed, if a some non-zero invariant set (call it $B$) exists, then there is  some $r\in(0,1]\cap{B}$.
-But $r\notin{T(B)}$ since any orbit with initial value $r$ will be a decreasing sequence.
-Moreover, every orbit of T will be decreasing and therefore approach the value $0$ as $n\rightarrow\infty$.
-
-One may ask if every orbit approaches an invariant set?
-In general the answer is no, since for the dynamical system $T: \mathbb{R} \to \mathbb{R}$ defined by $Tu=2u$, any orbit that does not intersect the invariant set $\{0\}$, will not approach any invariant set.
-
-However,  when the space $U$ is compact, all orbits approach an invariant set.
-This is since the set of limit points of the orbit can be shown to be invariant \cite{de2013elements}.
-So, when $U$ is compact,  the $\omega$-limit set $\omega(u;T)$ of a point $u$ defined to the collection of limit points of the sequence $\{x,Tu,T^2u,\ldots\}$ is nonempty, and $\omega(u;T)$ is invariant.
-
-\begin{Example}
-  For the map $Tu=u^2$ defined on $[0,1]$ all orbits lie in the invariant set $\{1\}$ or else would would approach the invariant set $\{0\}$. 
-\end{Example}
-
-Invariant sets have various properties. Invariant sets can be attracting or repelling depending on how orbits in their vicinity behave. 
-Recall the examples $Tu =u/2$ and $Tu=2u$ defined on $\mathbb{R}$ (where $\{0\}$ “attracts" orbits) or the example, $Tu=u^2$ defined on $[0,1]$ (where $\{1\}$ “repels" orbits). 
-We are interested in attractive invariant sets since they capture the long-term dynamics as time increases. 
-In particular, we are interested in those invariant sets named attractors.
-
-\begin{Definition}
-  [\bf Attractor]\label{Dfn_Attractor}\rm
-  Let $T: U \to U$, where $U$  is a metric space with metric $d$. A compact subset $A \subset U$ is said to be an attractor if it satisfies the three conditions: 
-  \vspace{-8mm}
-  \begin{enumerate}
-	\item $A$ is invariant. 
-	\item $A$ is asymptotically stable, i.e., for every $\epsilon > 0$ and for all $u$ so that $d(u,A) < \epsilon$, we have $d(T^nu,A) \to 0$ as $n\to \infty$. 
-	\item $A$ has Lyapunov stability, i.e., for every $\epsilon > 0$  there exists a $\delta(\epsilon) > 0$ so that $d(u,A) < \delta$ implies $d(T^nu,A) < \epsilon$ for all $n\ge 0$.  
-\end{enumerate}
-\end{Definition}
-
-Indeed in our previous examples, it can be verified that the system  $Tu=u/2$ defined on  $\mathbb{R}$ and $Tu=u^2$ defined on $[0,1]$  the singleton set $\{0\}$ is an attractor.
-For the system,  $Tu=1-|2u-1|$ on $[0,1]$, one may easily verify that the only attractor is the entire space $[0,1]$.
-This follows from the fact that between any two points $u< v$ in $[0,1]$, and for any $a,b$ so that $u\le a < b \le v$, we can find an $n$ so that $T^n(a,b)=[0,1]$ (as would be explained later in this chapter).
-
-A dynamical system can have several attractors and may also be contained in another attractor.
-For the example, $Tu =u^2$ on $[0,1]$, both $\{0\}$ and $[0,1]$ are attractors.
-It is known that an attractor for a dynamical system always exists in a compact space~\cite{Milnor1985}.
-
-The dynamics restricted to an invariant set can be complicated.
-For instance an invariant set could be just a single point or it could have an infinite set.
-If the invariant set is infinite, then complicated dynamics are possible.
-A particular, well-studied phenomenon of such complexity gives rise to so-called chaotic behaviour, a subject studied in detail over the past fifty years.
-
-% If the dynamics on the attractor are somewhat complicated in the sense that the attractor cannot be decomposed further, and if the attractor is infinite then there is a possibility of complicated dynamics, and a particular well-studied phenomenon of such complexity gives rise to what is called a chaotic behaviour.
-
-%If the dynamics on an attractor cannot be decomposed further and the attractor is itself also infinite, then there is a possibility of complicated dynamics, and a particular well-studied phenomenon of such complexity gives rise to what is called chaotic behaviour.
-
-\section{Chaos}
-
-In the 1960s, several mathematicians and mathematically interested scientists independently discovered chaos in the mathematical sense.
-The meteorologist Edward Lorenz may have been the first to explain this phenomenon in his 1963 paper \cite{lorenz1963deterministic}.
-The notions of invariance, attractivity, and chaos may also be described for continuous systems, and Lorenz's system comprised of a system of differential equations.
-The narrative of Lorenz's discovery of chaos and the history of other forerunners in this subject is fascinating.
-We highly recommend James Gleick's book Chaos: The Making of a New Science~\cite{gleick2008chaos} for those interested.
-It clearly illustrates these experiences, explains why chaos was such a startling and crucial mathematical and scientific discovery and describes the underlying mathematical notions for non-specialists.
-
-Different authors propose a number of intricately different definitions of chaos in the literature~\cite{RasbandChaos, TaborChaos, WigginsChaos} and each of them indicate some aspect of complexity.
-
-In practice, when only data is observed from a system, it is not possible to verify which definition or notion is satisfied by the underlying dynamical system.
-We do, however, specifically recall the definition of chaos in the sense of Devaney \cite{devaney2018introduction,de2013elements} so as to understand some nuances behind the complexity.
-Devaney's definition of chaos has three requirements, with the first being the notion  of sensitive dependence on initial conditions.
-
-\begin{Definition}\rm
-  [\bf {Sensitive Dependence on Initial Conditions}]\label{Dfn_SDIC}\rm
-A dynamical system $T: U \to U$ is said to have sensitive dependence on initial conditions (SDIC) if there exists a $\delta > 0$ such that for every $u \in U$ and in every neighborhood of $u \in U$ there exists a $v\in{U}$ and an integer $N:=N{(u,v)}\in\mathbb{Z}$ such that $d(T^Nu,T^Nv)>\delta$.
-\end{Definition}
-
-It is vital to acquire a sense of this definition as it can easily be misunderstood.
-It is a common misconception to interpret SDIC as two close points ($u$ and $v$) that eventually become separated by a distance $\delta$ under iteration by $T$.
-But this is not true.
-To fully comprehend all the subtleties of the concept, one needs to discuss it more thoroughly by considering each sentence with care and attention, whereafter one may examine how they fit together to convey the idea of sensitive dependence.
-
-To this end, we make 3 remarks:
-\vspace{-5mm}
-\begin{enumerate}
-  \item First, the $\delta>0$ in the definition of SDIC is independent of $u$. 
-  \item Second, in every neighborhood of $u$, we may not necessarily find all points $v$ in the neighborhood distinct from $u$ that would separate from the forward iterates of $u$. 
-  \item  Finally, $N$ depends upon $u$ and $v$ chosen, and their iterates may not separate forever (i.e., for all $n>N$) and we allow their iterates to get arbitrarily close in the future. 
-\end{enumerate}
-
-To illustrate the concept of SDIC, we present two examples - one from Mathematics and the other from the field of Physics.
-
-\begin{Example} \rm
-  The logistic map(LM), a recursion relation of the form $x_{n+1}=rx_n(1-x_n)$ where $x_n\in[0,1]$ and $r\leq{4}$ is a classic example used to illustrate the chaotic behaviour that a system may exhibit.
-When $r=4$, the one-dimensional recurrence relation $x_{n+1}=4x_n(1-x_n)$ can be used as the kindergarten's model to exhibit the presence of SDIC.
-\ednote{B: edited and removed unnecessary citation}
-  In the graph below~\ref{fig:log_sdic} we plot the first 50 iterates of the LM for $r=4$ and two slightly different initial values $x_0$.
-
-  \begin{figure}[ht]
-    \includegraphics[scale=0.74]{_logistic_sdic.eps}
-        \centering
-        \captionof{figure}{Figure generated for Logistic Map $x_{n+1}=rx_n(1-x_n)$ with $r=4$ to exhibit the presence of SDIC.
-Plotted are the values $x_n$ against time $n$ for the the first 50 step with initial values $x_0=0.2$ (blue) and $x_0=0.20001$ (red).
-Initially the two trajectories overlap, but they diverge completely at $n=12$ whereafter they follow distinct paths.}
-        \label{fig:log_sdic}
-      \end{figure}
-  \ednote{B: Caption and figure adjusted}
-  % For an intuitive sense of the fact that the LM exhibits SDIC, consider again the definition and substitute specific values.
-Let $\delta=0.05$ and for purposes of illustration consider the value $u=0.4$.
-The green band in~\ref{fig:log_sdic} shades the region $[u-\delta, u+\delta]$ and we can see that regardless the value $v$  and in every neighborhood of $u \in U$ there exists a $v\in{U}$ and an integer $N:=N{(u,v)}\in\mathbb{Z}$ such that $d(T^Nu,T^Nv)>\delta$
-  % specific point, say $u=0.4$.
-Draw a green neighbourhood $[0.35,0.45]$ of diameters $\delta=0.02$ around the point.
-It is easy to see that every neighbourhood around $u=0.4$ will contain some $v_1, v_2\in[0,1]$ and natural number $N_1, N_2 \in\mathbb{N}$ respectively s.t.
-the distance $d(T^Nu_1,T^Nv)>\delta=0.05$ .
-\end{Example}
-
-\begin{Example} \rm
-  A second example pertains to a simple physical system called the double pendulum - a pendulum with another pendulum attached to its head given an initial position and angular velocity.
-
-  Below are two figures denoting trajectories of a double pendulum's second head after some elapsed time.
-Depicted are three pendulum heads with equal angular velocity but differing slightly in initial position.
-As can be seen below, the trajectories diverge very quickly to become completely different.
-We discuss the double pendulum in greater detail in a later chapter (\ref{ch5}), but we mention the example here to provide a numerical example of a practical system exhibiting SDIC.
-
-\begin{figure}[ht]
-  \centering
-\includegraphics[scale=0.8]{_dp_sdic.eps}
-    \captionof{figure}{Three double pendulum heads with equal angular velocity and initial angles differing by 0,025 radians}\label{fig:dp_sdic}
-  \end{figure}
-
-\end{Example}
-
-The second part in Devaney's definition of chaos concerns topological transitivity.
-
-\begin{Definition}
-  [\bf {Topologically Transitive}]\label{Dfn_TopolTrans}\rm
-	A dynamical system $F: U \to U$ is topologically transitive if for any pair of nonempty open sets $E_1$ and $E_2$ there exists a $n\in\mathbb{N}$ such that $T^n(E_1) \cap E_2 \not= \emptyset$. 
-\end{Definition}
-
-Topological transitivity implies that iterates of an open set of initial conditions get mixed up with other open sets.
-On a compact metric space, one may show that topological transitivity also implies the existence of a point whose forward iterates are dense \cite{de2013elements}; or in other words, the orbit going through this point will be dense in the compact metric space.
-In fact, it is topologically more likely that the choice of an arbitrary point will be one whose iterates are almost dense.
-
-\begin{Definition}
-  [\bf {Meager Set}]\label{Dfn_Meager Set}\rm
-A subset of a topological space $U$ is said to be a meager set if it can be written as a countable union of sets of with empty interior. The complement of the a meager set is set to be a residual set.
-\end{Definition}
-
-To be more precise (see \cite{de2013elements}), for a discrete-time dynamical system on a compact space, the set of points with dense iterates are residual, and they are typical or likely to be observed in practice.
-In this project, when data is observed from a topologically transitive system, we assume it arises from a dense orbit.
-
-We may now define the notion of Chaos as formulated by Devaney\cite{devaney2018introduction}.
-\begin{Definition}
-  [\bf {Devaney's Chaos}]\label{Dfn_ChaosDec}\rm
-	A dynamical system $T: U \to U$ is said to exhibit chaos in the sense of Devaney if it satisfies the three properties:
-	\vspace{-5mm}
-  \begin{enumerate}
-		\item $T$ has SDIC.
-		\item $T$ is topologically transitive.
-		\item The set of periodic points of $T$ are dense in $U$. 
-	\end{enumerate}
-\end{Definition}
-
-\begin{Example} \rm
-  The standard tent map $Tu=1-|2u-1|$ defined on $[0,1]$ is a well-known example of a dynamical system satisfying these three properties.
-
-  We now reason as to why this is true.
-The graph of the map $T$ is piecewise linear with two straight lines, one connecting the points $(0,0)$ and $(\frac{1}{2},1)$ and the other connecting $(\frac{1}{2},1)$ with $(1,0)$.
-They form a so-called tent with base centered at $u=1/2$.
-The graph of the map $T^2$ comprises two symmetric tents with their base centered at $1/4$ and $3/4$.
-See Figure \ref{fig:T2tentmap}
-
-  \begin{figure}[ht]
-    \includegraphics[scale=0.6]{_tentmap_2.eps}
-        \centering
-        \captionof{figure}{Graph of the $T^2$ map with bases at $x=0$, $x=\frac{1}{2}$ and $x=1$ respectively.}\label{fig:T2tentmap}
-  \end{figure}
-
-  In general, the graph of  $T^k$ contains $2^{k-1}$ tents.
-  Given any point $u$ and an open neighborhood $W \subset[0,1]$ of $u$, we can find $k\in\mathbb{Z}$ large enough so that $T^k(W) = [0,1]$ because we can accommodate a tent whose base is contained in $W$.
-% \ednote{B: This last phrase doesn't make sense to me.
-M: The number of tents grow with $k$ and hence this is obvious; I explained in one of our meetings.
-Does it make sense?}
-  This implies the existence of two points $w_1$ and $w_2$ in $W$ so that $d(T^kw_1,T^kw_2)= 1$, and by the triangle inequality  at least one of the inequalities $d(T^ku,T^kw_1)> \delta$ or $d(T^ku,T^kw_2)> \delta$ hold when $\delta\in (0,\frac{1}{2})$.
-So $T$ has sensitive dependence on initial conditions.
-
-  Next, let $W_1$ and $W_2$ be two nonempty open sets.
-Given any open set $W_1$, we can find a $k\in\mathbb{N}$ large enough so that $T^k(W_1) = [0,1]$ because we can accommodate a tent with base contained in $W$.
-So $T^k(W_1) \cap W_2 \not=\emptyset$, and thus $T$ has topological transitivity.
-
-  Finally, for every open interval $(a,b)$, the graph of $T^k$ intersects the graph of the identity map on $[0,1]$.
-This follows from the fact established above that there exists a $k\in\mathbb{N}$ such that $T^k(a,b) = [0,1]$.If the intersection point has coordinates of the form $(p,p)$, the point $p$ is then a fixed point of $T^k$ and therefore also a periodic point of $T$.
-Since we have found a periodic point in the interval $W$,  the set of periodic points are dense in $T$.
-  This follows because any fixed point of $T^k$ is also a periodic point, and we have found this in an arbitrary interval.
-
-\end{Example}
-
-\section{Conjugacy}
-
-We now turn our attention to the subject of conjugacy that describes when two dynamical systems are equivalent dynamically.
-
-To show topological similarity (or sameness) between two metric or topological spaces, one needs to establish a homeomorphism between the two spaces. 
-However, in the study of dynamical systems defined on two spaces, establishing a homeomorphism does not indicate that the systems are dynamically related in any way.
-For instance, the maps $Tu=u^2$ and $Tu=1-|2u-1|$ defined on $[0,1]$ have totally different behaviour. 
-Therefore, to find dynamically similar systems, one must establish a dynamical equivalence that we illustrate in a simple commutativity diagram below~\ref{eqn_conjugacy}.
-
-% Giving errors
-\begin{equation}  \label{eqn_conjugacy}
-%    \[ 
-    \psset{arrows=->, arrowinset=0.25, linewidth=0.6pt, nodesep=3pt, labelsep=2pt, rowsep=0.7cm, colsep = 1.1cm, shortput =tablr}
- \everypsbox{\scriptstyle}
- \begin{psmatrix}
-U & U\\%
-V & V.
- %%%
-%  \ncline{1,1}{1,2}^{T} 
-%  \ncline{1,1}{2,1} <{\phi}
-%   \ncline{2,1}{2,2}^{S}
-%  \ncline{1,2}{2,2} > {\phi}
- \end{psmatrix}
-% \]
-\end{equation}
-
-%Composite functions
-To understand the diagram, we note that if we travel "right and then down," the diagram instructs us to use $T$ (top arrow right) first, followed by $\phi$.
-(right arrow downwards).
-Consequently, the pathway amounts to finding $\phi(T(u))$.
-If we proceed "down and then right," the diagram instructs us to apply $\phi$ (left arrow downwards) first, and then apply $S$ (bottom arrow right).
-This then amounts to finding  $S(\phi(u))$.
-When the relationships in this diagram hold, we say $\phi(T(u))= S(\phi(u))$, and we formally denote it as $\phi \circ T=S\circ \phi$.
-
-\begin{Definition}\rm  
-[\bf {Conjugacy \& Semi-Conjugacy}]\label{Dfn_Conjugate}\rm
-   Consider the dynamical systems $T:U\to{U}$, $S:V\to{V}$ and suppose the relationship $\phi \circ T=S\circ \phi$ holds.
-If $\phi:U\to{V}$ is a homeomorphism, then $S$ is said to be conjugate to $T$ and $\phi$ is called a conjugacy.
-If we relax the criterion on $\phi$ and merely require $\phi$ to be continuous and surjective where $\phi:U\to{V}$, then $\phi$ is a semi-conjugacy between $T$ and $S$ where T has domain $U$ and S has domain $V$; S is said to be semi-conjugate to T.
-\end{Definition}
-
-\ednote{B: Does defining explicitly the domain and codomain of $\phi$ make it clear enough?
-M: Saying $\phi$ is then a semi-conjugacy between $T$ and $S$ could be ambiguous as it does not specify which system is on the top in the commutativity.
-So please edit accordingly.}
-When $S$ is conjugate to $T$, the dynamics of the two systems are in some way `dynamically equivalent'.
-Specifically, they are in one-to-one correspondence with one another.
-However when $S$ is semi-conjugate to $T$ with $\phi:U\to{V}$ a many-to-one mapping, the dynamics on $V$ provide merely a coarse-grained description of the dynamics on $U$~\cite{de2013elements}.
-When $S$ is semi-conjugate to $T$, it is also common to call $S$ a \emph{factor of $T$}, or conversely that $T$ is an \emph{extension of $S$}.
-In essence, an extension (see, \cite{de2013elements}) is a more extensive system capturing all of the essential dynamics of its factor.
-
-It is a very hard or nearly an impossible task to establish the existence of a conjugacy or a semi-conjugacy $\phi$ between two systems \cite{devaney2018introduction}.
-However, one can verify that a function $\phi$ satisfies the commutativity diagram \ref{eqn_conjugacy}.
-
-\begin{Example}\rm
-  For example $\phi(x)=\sin\big(\frac{\pi}{2}x\big)^2$ is a conjugacy between two systems $Tu=1-|2u-1|$ and $Sv=4v(1-v)$  defined on $U=[0,1]$.
-
-  Indeed, $\phi:[0,1]\to[0,1]$ is a homeomorphism and by employing simple trigonometric identities, it is easily proved that $\phi$ satisfies~\ref{eqn_conjugacy}
-  \[
-    \begin{aligned}[t]
-      (\phi\circ{T})u 
-                &=\sin^{2}\Big(\frac{\pi}{2} - \frac{\pi}{2}|2u-1|\Big) = \sin^{2}(\pi{u})=\Big(2\sin(\frac{\pi}{2})\cos(\frac{\pi}{2})\Big)^{2}=4\sin^{2}(\frac{\pi}{2}u)\Big(1-\sin^{2}\frac{\pi}{2}u\Big) \\
-                                    &=({S}\circ\phi)u.
-    \end{aligned}
-\]
-And it follows that $\phi$ is a conjugacy between the two systems.
-\end{Example}
-
-By establishing that two systems are conjugate (semi-conjugate) to one another, we have also shown that one may choose to work with one system instead of the second and still be guaranteed to obtain information on the latter.
-It is instrumental during the process of forecasting the future evolution of dynamical systems.
-
-\chapter{The Learning Problem}\label{ch3}
-
-In this chapter, we acquaint ourselves with the question of forecasting dynamical systems with unknown underlying structures, state and discuss the shortcomings of the Takens delay embedding theorem and discuss various issues faced while forecasting.
-
-Consider a relatively simple learning problem: 
-Given the sequence $(u_0, u_1, \ldots, u_m)$ for $m\in\mathbb{Z}$, a finite segment of an orbit of the map $T$ where $T$ is defined by the update equation $u_{n+1} = Tu_n$, forecast the values $u_{m+1}, u_{m+1}$ where the map $T$ is unknown, given that $u_m$.
-
-A practical example of this would be the subsequent scenario: given the time-sequential coordinates of an object moving in space, predict the future positions of that object.
-We are very rarely, if at all, presented with a problem where the entire state information is available to us in the form $u_n$ at some timestep $n\in\mathbb{N}$.
-Consider then a more complicated learning problem:
-
-Suppose we only have the observations $\theta(u_0), \theta(u_1), \ldots, \theta(u_m)$ of the true system states $u$ in an unknown dynamical system $(U,T)$ and we wish to predict the values $\theta(u_{m+1}), \theta(u_{m+2})$ and so forth.
-
-First we establish the method of Takens delay embedding. 
-In this method, one considers a discrete-time dynamical system defined on a 'nice' space (a smooth manifold)\ednote{B: Should this be further explained?} that can be obtained as time-$K$ map of a continuous time-dynamical system, concepts formalised below.
-We make observations from such a system, i.e., we observe the evolution of $\theta(w_0)$ (where $w_0$ represents the initial state) by examining  a finite set of values of $u_n :=  \theta(w_n) = \theta(Tw_{n-1})$.
-The sequence $\{u_n\}$ represents a scalar time-series and intuitively $\theta$ represents a probe inserted into a bigger system which is itself only measuring/extracting a small part of the greater system state at time $t=n$.
-Consider, for example, a thermometer erected to measure the ambient temperature in a local village.
-This measurement function, the thermometer, is capturing only a single aspect, the temperature, of a much grander dynamical system entailing the current weather of the surrounding area.
-Even more than that though- it is measuring a miniscule part of the global weather system.
-
-%Given the sequence $\{u_1, u_2, \ldots, u_n}$, we define the variable \newline $y_n:=[\theta_n, \theta_{n-\tau}, \ldots, \theta_{n-(2d-2)\tau}, \theta_{n-(2d-1)\tau}]^{T}\in\mathbb{R}_{d}\times{1}$ where $\tau$ represents the lag and $d$ the dimension of the attractor.
-(We shall return to these terms shortly and their discussion may be put on hold for the time being).
-
-However, the actual state $u$ of a system is seldom, if ever, fully known.
-the most one can do is to insert probes into a system to obtain partial information through the measurements taken.
-Moreover, the process of taking a measurement itself introduces two additional aspects which complicate the problem: \ednote{B: Perhaps this edit makes more sense?
-M: Maybe replace difficulty by saying that we need to consider two aspects.
-}
-\vspace{-8mm}
-\begin{enumerate}[noitemsep, label=\roman*.]
-  \item A series of measurements over a specific time interval is inherently a discretisation procedure of the underlying continuous-time dynamical system.
-(Hence why we restricted our attention to discrete-time systems in the preceding section)
-  \item A probe will never be entirely accurate, and so the act of measurement introduces a certain measure of numerical noise/inaccuracy.
-\end{enumerate}
-
-From here we construct a multidimensional observable using the method of stacking previous observations, i.e., we create delay-coordinate map defined by
-$\Phi_{k,\theta}(w) := (\theta(T^{-k}w)\ldots,\theta(T^{-1}w),\theta(w))$.  
-The concept of an observable is understood in the sense of an observable as originally introduced by~\cite{takens1981detecting, genericObservableAeyels} to refer to a probe or measurement function inserted into the system.
-Takens' theorem, in essence, refers to the fact that for a sufficiently large $k$, we can define a dynamical system on the space $\mathbb{R}^{k+1}$ with states $\Phi_{k,\theta}(w)$, $\Phi_{k,\theta}(Tw)$, $\Phi_{k,\theta}(T^2w)$, etc.
-This dynamical system is topologically conjugate to the unknown underlying system $(W,T)$.
-We recall the Takens delay embedding theorem next.
-
-%We may then ask ourselves the question: Is it in any way possible to retain information about the system state $x(t)$ in this temporal data-series $\theta(x(t))$?
-The answer is easily yes if T is known.
-However, neither this, nor even the exact function $\theta$ is available.
-
-\section{Takens Embedding Theorem}\label{sect_Takens}
-
-We define first the concepts of homeomorphism and embedding:
-\begin{Definition}\rm
-  [\bf {Homeomorphism}]\label{Dfn_homeo}\rm
-  A homeomorphism is a function $f:Z\rightarrow Y$ between two topological spaces $Z$ and $Y$ that is continuous, bijective and has a continuous inverse. 
-\end{Definition}
-
-\begin{Definition}
-  [\bf {Embedding}]\label{Dfn_embed}\rm
-  Consider a homeomorphism $f:Z\rightarrow Y$ for $Y\subset X$. $Z$ is said to be embedded in $X$ by $f$.
-\end{Definition}
-
-Takens Theorem states a result establishing a relationship between the observed and underlying dynamical systems by showing that the concatenation of a sufficiently large number of previous observations into a vector will, under certain conditions, generate a map between the vectors from the respective systems.
-We formulate the theorem from \cite{takens1981detecting}.
-
-\begin{Theorem} 
-	[\bf Takens Embedding Theorem (adopted from \cite{takens1981detecting}] \label{Thm_Takens}
-         Let $W$ be a compact manifold of dimension $m$, and $d\ge m$ so that $2d$ is an integer. It is a 
-            generic property for the pair $(T, \theta)$,  where $T:W \to W$ is
-            a smooth diffeomorphism, and $\theta:W \to \mathbb{R}$ a smooth function, the map $\Phi_{2d,\theta}:W \to \mathbb{R}^{2d+1}$ defined on $W$ by 
-            $\Phi_{2d,\theta}(w) := (\theta(T^{-2d}w)\ldots,\theta(T^{-1}w),\theta(w))$
-            is a diffeomorphic embedding; by `smooth' we mean at least $C^2$.
-Consequently, there exists a map $F_\theta: \Phi_{2d,\theta}(W) \to \Phi_{2d,\theta}(W)$ defined by $$F_\theta: (\theta(T^{-2d}w),\ldots,\theta(T^{-1}w),\theta(w)) \mapsto
-            (\theta(T^{-2d+1}w),\ldots,\theta(w),\theta(Tw))$$
-           so that $(W,T)$ is topologically conjugate to 
-            $(\Phi_{2d,\theta}(W), F_\theta)$.    
-\end{Theorem}
-
-By generic we mean a residual set on a certain topology on appropriate spaces of functions (that we do not describe here) \ednote{B: Where can I find a citation for this?}. 
-By $C^2$, we make reference to a twice-differentiable function  with a continuous second derivative.
-In our scenario (and this remains important throughout) the input space $U$ is considered the attractor.~\label{attractor_U}
-
-Below we reproduce Figure 1 from \cite{Supp}.
-
-%Some explanation is beneficial.
-Takens establishes a delay-coordinate map $\Ftheta$ defined in \ref{eqn_takens} for $T$ the flow (\ref{defn_flow}) defined by the update equation  $Tu_n=u_{n+1}$ and $theta$ our measurement function.
-When confining a dynamical system to the manifold U (in which the underlying system is contained), Takens showed that if certain smoothness conditions are satisfied on $T$ and $\theta$, then the delay-coordinate map $\Ftheta$ embeds $U$ in the reconstruction space $R^{2d+1}$ \textbf{we haven't used the term reconstruction space yet, so this is falling out of the blue}, for almost every choice of measurement function $\theta$, the observable.
-By almost every choice here, we mean that we exclude some observations which do not give any information.
-To see a more precise discussion, see \textbf{xxx} for we do not concern ourselves with a further elaboration on this report.
-
-%Alternatively we may also make the statement: Takens' Embedding Theorem guarantees that almost all dynamical systems can be reconstructed from just one noiseless observation sequence \textbf{this isn't obvious from the theorem's formulation, is it?} i.e.
-for a great number of possible observation functions $\theta$, $\Ftheta$ preserves the topology of U.
-\textbf{'preserving the topology' is not something we've touched on before.
-Perhaps I should add to previous discussion so logic flows seamlessly here}.
-
-\begin{figure}[ht]
-  \includegraphics[scale=0.25]{_takensmap.eps}
-  \centering
-  \captionof{figure}{Schematic of Takens Embedding Theorem's usage and impact of iterates slipping from the attractor when the approximation $\tilde{F}_\theta$ of $F_\theta$ is learnt.}
-  \label{fig:takensmap}
-\end{figure}
-
-If we can learn the map $F_\theta$, from sufficient number of data points of $\{\phi_{2d,\theta}(w_n)\}$, then we also know how to forecast how $\theta(w_n)$.
-
-%Once again we pause and consider the graph to consolidate our understanding of the conjugacy (\textbf{are we being too simplistic?}).
-If we are in state $u$ and then evolve towards state $Tu$ before embedding into $\mathbb{R}^{2m-1}$, then this is equivalent to first embedding into $\mathbb{R}^{2m-1}$ and then evolving through the map $\Ftheta$, i.e.
-$\Phi_{2d,\theta}\circ{F}_{\theta} = T\circ\Phi_{2d,\theta}$.
-The map $F_theta$ is a homeomorphism as well.
-
-%Thus information about U can be retained in the time series' (or observation's) output.
-By preserving the topology on the manifold $U$ in the reconstruction space $X$ (\textbf{not discussed before}), we are also guaranteed the preservation of topological invariants of the manifold, of which dimensionality is one such invariant.
-(We note here that dimensionality will again be considered later in this project).
-
-\section{Practical Limitations to Takens' Theorem}
-Takens' Embedding Theorem is a powerful result, and it provides compelling reason to believe that one could conceivably reconstruct a system conjugate accurately to the underlying system. 
-Nonetheless, it does present some severe practical limitations:
-\vspace{-5mm}
-\begin{enumerate}
-\item Even supposing that we could indeed find $F_\theta$ , our \emph{approximation} of  $F_\theta$ is a map from a larger set $\mathbb{R}^{2d+1}$ containing the embedded attractor.
-There are, however, no theoretical guarantees that $F_\theta$ will retain  $V=\Phi_{2d,\theta}(W)$  as an attractor although $W$ could itself be an attractor.
-\item Takens theorem is stated only for noiseless observations.
-Due to noise $\epsilon_n$ the delay vector $\Phi_{2d,\theta}(w_n) + \epsilon_n$  may lie outside $V$.Furthermore, due to the chaotic nature of the underlying system (i.e.
-the fact that it has SIDC), the evolution of $\Phi_{2d,\theta}(w_n) + \epsilon_n$ under the map $F_\theta$ could move out of $V$ completely.
-This problem can be overcome  by using a driven dynamical system with some properties, and we discuss this in Chapter\ref{ch3}.
-\end{enumerate}
-
-% The Takens embedding theorem does not guarantee global dissipativity.
-% \ednote{B:Links with~\ref{subs_LearnGamma} in discussion of dissipativity of $\Gamma$.
-Sir, I'm not sure how to write this here - I think it a beneficial point to make so as to set the stage for the later discussion in ~\ref{subs_LearnGamma}, but I am unsure there as well.}
-
-Though not a uniquely related to the Takens embedding theorem, we do opt to make mention of another requirement for our work. 
-The methods described here are applicable for data originating from a surjective map. 
-This follows from the fact that when $T$ maps the space $U$ into a proper subset $B$ of$ $U, there exists some point in $U - A$ that does not have a preimage and we conclude that $T$ is not surjective.
-An example of this would be the case when a system displays energy-loss.
-Consider, for instance, consider the dying oscillations for a damped pendulum.
-(An account of greater depth  is provided in~\ref{ch5})
-
-This requirement is not overly restrictive, as a great number of chaotic systems do have a surjective map.
-(Cite) \ednote{B: I remember you saying this in one of our sessions, but don't have a source for it.
-Do you have a specific reference?}
-
-\chapter{Driven Dynamical Systems to Forecast Problems}\label{ch4}
-
 In this chapter, we discuss results on the mapping of temporal data obtained from a discrete-dynamical system onto a different space through the notion of a driven dynamical system. 
@@
 \begin{figure}[ht]
-  \includegraphics[scale=0.4]{_actionofh_H.eps}
+  \includegraphics[scale=0.4]{Graphs/_actionofh_H.eps}
   \centering
@@
 \begin{figure}[ht]
-  \includegraphics[scale=0.35]{_Nonlinearity_Fig_response.eps}
+  \includegraphics[scale=0.35]{Graphs/_Nonlinearity_Fig_response.eps}
   \centering
@@
 \begin{figure}[ht]
-  \includegraphics[scale=0.3]{_summarypictorial.eps}
+  \includegraphics[scale=0.3]{Graphs/_summarypictorial.eps}
   \centering
@@
 \begin{figure}[ht]
-  \includegraphics[scale=0.25]{_YTerrors.eps}
+  \includegraphics[scale=0.25]{Graphs/_YTerrors.eps}
   \centering
@@
 \begin{figure}[ht]
-  \includegraphics[scale=0.25]{_autovsdriven.eps}
+  \includegraphics[scale=0.25]{Graphs/_autovsdriven.eps}
   \centering
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\chapter{Implementation} \label{ch5}
-
-In this chapter we turn to the implementation of the function $G_T$ by discussing a method to learn the function $\Gamma$.
-We apply our methodology to some chaotic attractors.
-Most notably, we consider the reconstruction of the chaotic attractor of the double pendulum - an  attractor that has not been successfully reconstructed because of its intricate structure.
-In fact the previous data-driven approaches also have not been successful in building a model from the double pendulum.
-
-%We then indicate where this implementation differs from previous work done by (\cite{manjunath2021universal}) and empirically show that the SDD of $G_T$ are less functionally complex than $\Phi_{k, \theta}$ or $T$ by considering the Pearson correlation coefficient.
-Finally, we present and discuss some numerical results for a focus system of this project: the double pendulum.
-We also consider other chaotic attractors.
-
-\section{Implementing $G_T$}
-The map $G_T$ is implemented by applying the functions $\Gamma$, $g$ and the projection function $\pi_2$ in the relevant order as described by the equations~\ref{eqns_from_data}.
-To reach this point, one must first learn $\Gamma$.
-After first normalising the data to ensure that it has zero mean value and lies within the range $[-1,1]$ and required zero-padding(see paragraph below for further discussion), we continue by driving the system with external input through the function $g$.
-
-From \eqref{eqn_driving} we know that $g(u,x) = (1-a) x + a \tanh(Au + \alpha B x)$. Recall that $g$ is SI-invertible. We refer to $A$ as the input matrix and $B$ the reservoir matrix.
-
-We make use of a Recurrent Neural Network (RNN) in our implementation to project the data on to a higher dimensional space, but we do not train the RNN in the conventional manner.
-\ednote{B: Should we say what the conventional manner is?} A few words about the RNNs: RNNs are widely used in the literature for temporal processing of data as a type of deep learning technique.
-They take the form of equations considered to be discrete-time version of continuous dynamical systems modelling a neural network of neurons \cite{jaeger2001echo}.
-
-We implement a feedforward network, (a network with no memory, unlike the RNN)\ednote{B: Expanded sentence s.t.
-FFN is in more context} required to learn $\Gamma$ in the \emph{Python} programming language and making use of the \emph{Scipy} library, alongside the \emph{Keras} package in the \emph{Tensorflow} library \ednote{B: Is this something you want me to write in?
-M: We do not implement the RNN, but the FN that learns Gamma through the Tensorflow.
-We called the RNN + FN set up a RCN in Adriaan's paper.
-FYI: FN's work only if there is a definite relationship between the $u_n$ and $u_{n+1}$ which is rare in temporal data.
-RNNs with ESP work when there is a definite relationship between $(\ldots,u_{n-1},u_n)$ and $u_{n+1}$ and is suited for temporal processing.}.
-
-If the RNN we will be using in the next step has $N$ neurons, then $A$ and $B$ are each $N\times{N}$ matrices.
-Inputs $u_n$ of dimension $K < N$ are padded by zeroes (as remarked earlier) to embed the input in a space of dimension $N$.
-As we do not need the entire left-infinite history of input, an arbitrary initialisation of the system with initial values followed by the driving of the system with the function $g$ means that the generated values will approach the actual solution elements in a uniform manner thanks to the UAP that we had discussed in the preceding chapter.
-
-Thirdly we prepare the network.
-Utilising a network $N$ neurons (the same value $N$ as in the dimensions of the matrices $A,B$).
-Learning of $\Gamma$ is done using the Adam Optimiser where the means square error (MSE) loss function is optimised.
-The system is trained several times (in our case 4) with incrementally smaller learning rates on each run (0.001 and thereafter divided by 10 at each run).
-Additional parameters such as the number of network-layers, neurons per layer and training-length are presented case-by-case below.
-
-We then reach the prediction step. Making use of $g$'s SI-invertibility and $g$ as defined in \eqref{eqn_driving}, we have the following expression for $u_n$ replicated from~\eqref{eqn_SI_RNN}.
-\begin{equation*}
-  u_n := A^{-1}\bigg(\overline{\tanh}^{^{-1}}\frac{1}{a}\Big(x_{n+1}-(1-a)x_n\Big) \bigg) - \alpha B x_n
-\end{equation*}
-
-The two model equations for learning from data~\eqref{eqns_from_data} allows us to determine what the state $x_{n+1}$ at time $n+1$ will be. 
-We specifically remind ourselves of \[x_{k+1}=\pi_2 \circ (\Gamma, \pi_2) \circ (\pi_2,g) (u_k,x_k)\]
-
-\section{Delay Coordinates}
-
-We also consider a more general problem of learning an attractor of dynamical system when only observations of an orbit are given.
-Explicitly, if $(W,T)$ is a dynamical system with dynamics generated by $w_{n+1}=Tw_n$, and if $\theta:W \to \mathbb{R}$ is an observable, then the task is to learn a dynamical system that is topologically conjugate to $(W,T)$ and predict $\theta(w_{m+1}),\theta(w_{m+2}),...$ using the data $\theta(w_{0}),\theta(w_{1}),\ldots,\theta(w_{m})$.
-So how do we do it?
-Suppose the input generated from the delay-coordinate map $\Phi_{\theta,2d}(\theta(w_{n})) := (\theta(w_{n-2d}),\ldots,\theta(w_{n-1}),\theta(w_{n}))$ is such that Takens delay embedding theorem (see Theorem~\ref{Thm_Takens}) holds, in which case, there exists a homeomorphism  $F_\theta: \Phi_{\theta,2d}(\theta(w_{n})) \mapsto \Phi_{\theta,2d}(\theta(w_{n+1}))$.
-Then if we feed  the input values $u_n := \Phi_{\theta,2d}(\theta(w_{n}))$ to a driven system $g$ that is SI-invertible and has USP, the induced system $(Y_F,G_F)$ would be topologically conjugate to the inverse-limit space of
-$(\Phi_{\theta,2d}(W), F_\theta)$ as in Theorem~\ref{Thm_CET}, and hence one could forecast $u_n,u_{n+1},\ldots$.
-In other words, one could forecast  the values $\theta(w_n), \theta(w_{n+1}),\ldots$ by feeding these observations as delay coordinates.
-The advantage of feeding delay-coordinates to a driven system is that the embedding is stable in the sense of global dissipativity that we have described in the previous chapter, i.e., $Y_F$ is made to contained in an attractor of a mapping induced on $Y_F$ that is induced by $g$.
-
-We note that the twin equations together~\ref{eqns_from_data} generate a model from data.
-For a chaotic system, a small amount of noise due either to the input-noise or computational error would cause the solutions from two accurate models with the same initial conditions to  diverge as they are not identical.
-\ednote{B: edited this sentence} However, statistically, they would be the same since due to ergodic theory all typical orbits have the same visiting frequency to any region of the phase space of the dynamical system.
-\ednote{B: I'll need to recap this sentence with you before presenting as it's new to me.} To illustrate that we plot the densities of the trajectories of the actual and predicted solutions.
-
-Our implementation of the reservoir(A) and input(B) matrices in the RNN makes use of sparse matrices.
-We discuss three specifics relating to this concept.
-Firstly, the matrices are assigned a variable density in terms of the amount of non-zero entries and are scaled to have values within the range $[-1,1]$.
-For this, we make use of \textit{Python's SciPy} library and specifically the \textit{Sparse, LinAlg} modules.
-In our experiments, we generally opted for a density ratio of 10\%-25\%.
-Secondly, we opted to enforce the requirement that both matrices be well-conditioned, i.e.
-that they have a condition number of \~1.
-The condition number of a matrix $M$ (denoted $cond(M)$) is the product of the matrix norms of the matrix $M$ and its inverse $M^{-1}$ and is a measure of the sensitivity of output of a matrix multiplication when an input is changed.
-\textbf{(Cite.)}\ednote{Ask M: Cite specific resource?} By ensuring a condition number close to 1, we also ensure that our driven function $g$, depending on the matrices $A$ and $B$, will not be sensitive to adjustments in the input.
-In so doing we then also guarantee another form stability.
-\ednote{B: Not complete enough.
-Edit further and cite}
-
-Lastly, we ensure the matrix $B$ has unit spectral radius.
-This is necessary due to the Causal Embedding Theorem~\ref{Thm_CET} which can then guarantee the existence of semi-conjugacy between the systems $(Y_T, G_T)$ and $(\widehat{U}_T, \widehat{T})$.
-In fact, the theorem actually requires that $\alpha{B}$ have spectral radius $<1$, so we impose the restriction $\alpha<1$ and consider only the matrix $B$'s spectral radius from thereon.
-
-Note that to learn $\Gamma: (x_{n-1},x_{n}) \mapsto u_n$ through our feedforward network, we could learn the principal components of $(x_{n-1},x_{n})$ as well. 
-We employ this since it provides slightly better results empirically but is not essential.
-
-Principal components are used only for a more efficient state representation to possibly reduce learning errors; it is not for a lossy approximation since we use all principal components \textbf{(Cite.)}\ednote{Ask M: Is there a specific resource to cite?}.
-More explicitly, denote $X_{1:N}$ as the matrix with the first $N$ states of the network data represented by row vectors.
-If $X_{1:N}=U\Sigma P^T$\ednote{Expand here} represents the singular value decomposition of $X_{1:N}$ \ednote{Ask: Is there a specific resource to cite?}, we then denote the principal component matrix $P$, and the principal components are given by
- $Z_{1:N}=X_{1:N}P.$
-These principal components are then used to train a feedforward neural network. Denote the row vectors of $Z_{1:N}$ by 
-$z_i^T, i=1,2,\cdots,N$ and the neural network by $NN$, the network is subsequently trained to learn an approximation of the map
-$NN: (z_{n-1},z_n) \mapsto  u_n.
-$
-The learnt approximation of $NN$ can be used to approximate $\Gamma$ since 
-$$
-NN\left( \begin{bmatrix} 
-P^Tx_{n-1} \\
-P^Tx_n
-\end{bmatrix}
-\right) = NN \circ 
-\begin{bmatrix}
-P^T & 0 \\
-0 & P^T 
-\end{bmatrix}\begin{bmatrix}
-x_{n-1}\\
-x_n
-\end{bmatrix} = u_n = \Gamma(x_{n-1},x_n).
-$$
-
-\section{Experimental Results}
-In this section we showcase numerical results pertaining to 3 attractors that were simulated using the methodology described.
-We first consider a simple physical system that exhibits complicated dynamics.
-\subsection{Double Pendulum}
-
-The Double Pendulum consists of two pendulums fastened to one another such that the system moves as a whole.
-The first (or higher pendulum) is attached to a fixed point while the second is attached to the endpoint of the first.
-(See below)
-This system is a classical example of a system exhibiting chaos and is often used as explanation of the very well-known Butterfly Effect.
-\ednote{B: Originally though a reference here was needed, but no longer think so.} Consider again~\ref{fig:dp_sdic}.
-The system has four variables: angle and angular velocity of each pendulum rod. An ideal system is assumed where the rods are massless and no friction in the pivots are considered.
-\begin{figure}[ht]
-  \includegraphics[scale=0.35]{_dp_setup.eps}
-  \centering
-  \captionof{figure}{Pictorial representation of the setup of a DP. Image taken from~\cite{DPSetup}}
-  \label{fig:dp_setup}
-\end{figure}
-
-The scalar equations of motion are derived using Newtonian physics and are reproduced from~\cite{DPFormulas} in~\ref{eqns_dp}.
-Each pendulum component has an associated mass of the head, $m$, and length of the rod, $L_i$.
-Acceleration due to gravity is denoted by $g$ and initial velocities by $\dot{\theta}$
-\begin{eqnarray}\label{eqns_dp}
-  \ddot{\theta_{1}}  = \frac{-g(2m_1+m_2)\sin{\theta_1} - m_2g\sin(\theta_1-2\theta_2) - 2\sin(\theta_1-\theta_2)m_2({\dot{\theta_{2}}}^{2}L_2 + {\theta_{1}'}L_1\cos(\theta_1-\theta_2))} {L_1(2m_1 + m_2 -m_2\cos(2\theta_1 - 2\theta_2))}
-  \\
-  \ddot{\theta_{2}} = \frac{2\sin(\theta_1-\theta_2)(\theta_{1}'^{2}L_1(m_1+m_2) + g(m_1+m_2)\cos\theta_1 + \dot{\theta_{2}}^{2}L_2m_2\cos(\theta_1-\theta_2))}{L_2(2m_1 + m_2 -m_2\cos(2\theta_1 - 2\theta_2))}
-\end{eqnarray}
-
-The second-order system is converted to a first-order ODE system and solved using the Runge-Kutta numerical method.  
-Here we chose the values $m_1=2.0$, $m_2=2.0$, $L_1=1.5$ and $m_2=1.5$ with initial velocities $v_1=1.5, v_2=-2$ for the respective pendulum heads.
-Undamped DP data-sets were generated and used in this project due to the requirement that the map $T$ be surjective that we must not forget. 
-An example of this would be the case of a DP exhibiting energy-loss (i.e.
-where the map $T$ is not surjective).
-In such a case, the attractor is a single point in the phase space when the double pendulum comes to rest.
-It must be mentioned once again that if we use data outside the attractor, learning is not reliable. To see this, consder the graphs ~\ref{fig:damped_pendulum} and ~\ref{fig:dp_notsurjective}.
-
-\begin{figure}[ht]
-  \includegraphics[scale=0.4]{_dp_dying.eps}
-  \centering
-\caption{$x$- and $y$-coordinates of the two pendulum heads plotted for a dataset that exhibits dying oscillations.
-The map generating the (four-dimensional data) for a damped pendulum is not surjective.
-Data obtained from chaotic double pendulum dataset generated from videos takes of experimental pendulums~\cite{asseman2018learning}}
-\label{fig:damped_pendulum}
-\end{figure}
-
-\begin{figure}[ht]
-  \includegraphics[scale=0.3]{_dpfail_nonsurj.eps}
-  \centering
-\caption{Failure to forecast or learn a map which is not surjective: Evolution of the $y$-coordinate of the bottommost pendulum plotted here for 15000 timestep.
-Plotted here the predicted value(red) versus the actual value(blue).
-The plots overlap for the first 10000 steps, the duration of training to learn the map $\Gamma$.
-Afterwards, the trajectories start diverging and it the predicted value no longer accurately represents the complexity or energy-loss of the actual data.
-}
-\label{fig:dp_notsurjective}
-\end{figure}
-
-Having spent ample time to discuss the challenges, we now change tone so as to satisfy the optimist and present a successful run of the DP-data to predict the future movement of the pendulum.
-
-The system was driven for 15500 steps, of which the first 500 were immediately discarded to simulate the network 'forgetting'.
-The driving function $g$ had input and reservoir matrices were implemented as sparse matrices filled to 20\%.
-Parameter-values of $\alpha=0.99$ and $a=0.5$ were used.
-The remaining 15000 datapoints were used to train a network with 128 neurons. 
-The network was initialised to contain 16 hidden layers each with a layer dimension of 64, activated with the tanh activation function.
-Training was accomplished using 256 epochs with batch sizes 128 and a delay of 1.
-See~\ref{fig:dp_success_traj} for results pertaining to the trajectory taken by the DP's components and then~\ref{fig:dp_success_density} to witness associated densities of predicted vs.
-actual values.
-
-\begin{figure}[ht]
-  \centering
-  \includegraphics[width=0.95\linewidth]{_dp_success_4coords_traj.eps} 
-  \captionof{figure}{Predicted(red) and actual(blue) trajectories of the $x$- and $y$-coordinates of each of the pendulum heads.
-From top to bottom are as follows: $x$-coordinate of topmost pendulum head, $y$-coordinate of topmost pendulum head, $x$-coordinate of lower pendulum head, $y$-coordinate of lower pendulum head.
-  These graphs were constructed by predicting the DP 1000 timesteps into the future and in so doing illustrating the long-term consistency and accuracy of the learnt system.
-Here we lock onto the near-exact trajectory for about 300 timesteps and stay close up until about 600 seconds.
-}
-  \label{fig:dp_success_traj}
-\end{figure}
-\begin{figure}[ht]
-  \centering
-  \includegraphics[width=\linewidth]{_dp_success_4coords_hist.eps}
-  \captionof{figure}{Densities of predicted and actual values of $x$- and $y$-coordinate of pendulum heads.}
-  \label{fig:dp_success_density}
- \end{figure}
-
-We now consider an experiment where noise is added to the system and find it to be incredibly robust.
-We employ the same parameters as before and add noise from a normal-distribution with mean zero and standard deviation equal to 0.1 which translates to roughly 39dB's of noise.
-(For the signal-to-noise ration we adopt the formula $\text{SNR}_{dB}=10\text{log}_{10}\big(\frac{P_\text{{signal}}}{P_{\text{noise}}}\big)$ where $P_\text{{signal}}$ refers to the variance observed in the normalised data and $P_{\text{noise}}$ represents the variance of the added noise.)
-
-Once again consider the actual and predicted trajectories for the noisy DP dataset in Figure~\ref{fig:noisydp_success_traj}.
-Next we consider the densities of the predicted and actual values in Figure~\ref{fig:noisydp_success_density} and observe that they display highly similar distributions.
-
-\begin{figure}[ht]
-  \centering
-  \includegraphics[width=0.95\linewidth]{_dp_noise_FigCD_2030.eps} 
-  \captionof{figure}{Predicted(red) and actual(blue) trajectories of the $x$- and $y$-coordinates of each of the pendulum heads with noise added from a normal distribution with mean zero and standard deviation 0.1, equating to approximately 39dB's of noise.
-  These graphs were constructed by predicting the DP 1000 timesteps into the future and in so doing illustrating the long-term consistency and accuracy of the learnt system.
-Here we lock onto the near-exact trajectory for about 250 timesteps and stay close up until about 400 seconds.
-}
-  \label{fig:noisydp_success_traj}
-\end{figure}
-\begin{figure}[ht]
-  \centering
-  \includegraphics[width=\linewidth]{_dp_noise_2B_2030.eps}
-  \captionof{figure}{Densities of predicted and actual values of x-coordinate of first pendulum head.}
-  \label{fig:noisydp_success_density}
- \end{figure}
-
- Lastly, a run was done where only a single coordinate (the $y$-coordinate of the lower pendulum head) was fed into the system and we can illustrate through the obtained success that the coordinate contains all the information in its past states necessary to forecast its future evolution.
- \textbf{Insert graphs here. Had to quickly change something here.}
-
-\subsection{Additional Attractors: Clifford \& Thomas}
-We follow the methodology presented in \cite{manjunath2021universal} but opt to consider attractors different from the Lorenz system and H\'enon map in this project.
-
-\subsubsection{Thomas' Cyclically Symmetric Strange Attractor}\label{Thomas_Attractor}
-Thomas' Cyclically Symmetric Strange Attractor, a 3D attractor proposed in 1999 by Ren\'e Thomas in \cite{ThomasAttractor}, is described by a set of three equations which is distinctly reminiscent of the Lorenz system, another deterministic three-equation model exhibiting chaotic behaviour.It has a single parameter $\beta$ and has been shown to transition to chaotic behavious when $\beta<0.208186$~\cite{Thomas_BetaParameter}.
-\begin{eqnarray}\label{eqns_thomas}
-  x_{n+1} = \sin(y_n) - \beta{x_n} \\
-  y_{n+1} = \sin(z_n) - \beta{y_n} \\
-  z_{n+1} = \sin(x_n) - \beta{z_n}
-\end{eqnarray}
-
-The behaviour of the system with a $\beta$-parameter value of 0.1056 was simulated by scaling it to fit inside the interval $[-1,1]$.
-Two cases were considered: noise-free and the one perturbed by noise.
-Noise was generated from a normal distribution with mean 0 and standard deviation 0.05, which to translates approximately 33dB.
-($\text{SNR}_dB$ is calculated exactly as for the DP-data above)
-
-A sequence of 3000 observations in both the clear and noisy data-sets respectively were fed into the system with 300 entries discarded to simulate the network's loss of memory.
-The next 9000 steps were predicted and compared with true data.
-
-\begin{figure}[ht]
-  \centering
-  \minipage{0.5\linewidth}
-    \centering
-    \includegraphics[width=0.8\linewidth]{_thomas_noisy_0.1056.eps}
-  \endminipage\hfill
-  \minipage{0.5\linewidth}
-    \centering    \includegraphics[width=0.8\linewidth]{_thomas_clean_0.1056.eps}
-  \endminipage
-  \captionof{figure}{True(red) and predicted(blue) trajectories for the Thomas attractor with parameter value $\beta=0.1056$.
-To the left: Noise added from normal distribution with mean zero and standard deviation 0.05.
-Right: No noise added to dataset.}
-\end{figure}
-
-The RNN was initialised as follows: the dimension of the network is set equal to 500 neurons and 12 hidden layers of dimension 64 each were initialised.
-Training is realised through 150 epochs of batch sizes of 128 each; the map is learnt via the Adam Optimiser using finer and finer learning rates (0.001 and then divided by 10 at every iteration).
-
-\begin{figure}[ht]
-  \centering
-  \includegraphics[scale=0.35]{_Thomas_1.eps}\caption*{Predicted trajectories of the $x$- and $y$-coordinates for the Thomas attractor with parameter-value $\beta=0.1056$ demonstrate empirically the ability to predict the evolution of the trajectory for the next an estimated 100 timesteps into the future near-exactly.
-Here no noise was added.}
-  \includegraphics[scale=0.5]{_Thomas_3.eps}\caption*{Represented here the learnt (blue) and actual (red) densities of the first coordinate ($x$-coordinate) of the dataset for the Cyclically Symmetric Attractor with parameter-value $\beta=0.1056$.}
-\end{figure}
-
-\subsubsection{Fractal Dream Attractor}
-
-Another 2-equation discrete-time map named the Fractal Dream Attractor, or more commonly known as the Pickover Map (first discovered by Clifford A.
-Pickover and discussed in his fascinating book "Chaos in Wonderland"\cite{PickoverChaos}), was also considered.
-\begin{eqnarray}\label{eqns_clifford}
-  {x_{n+1}=\sin(ay_n) + c\cos(ax_n)} \\
-  {y_{n+1}=\sin(bx_n)+d\cos(by_n)}
-\end{eqnarray}
-
-\begin{figure}[ht]
-  \centering
-  \includegraphics[width=1.0\textwidth,left]{_Clifford_1_nonoise.eps}
-  \caption*{These graphs were constructed by predicting the Clifford system 1000 steps into the future and in so doing illustrating the long-term consistency and accuracy of the learnt system.
-As perceived here, we are able to lock on to the trajectory of the Clifford map almost exactly for the first 25 steps.}
-  \minipage{0.5\textwidth}
-      \centering
-      \includegraphics[width=\linewidth]{_Clifford_2_nonoise.eps}
-  \endminipage\hfill
-  \minipage{0.5\textwidth}
-    \centering
-    \includegraphics[width=\linewidth]{_Clifford_3_nonoise.eps}
-  \endminipage
-  \caption{The bottom row, left, shows a plot of the true attractor(red) and the predicted values(blue).
-Bottom row, right, shows the distribution of the first coordinate of the learnt system (blue) and the actual Clifford system (red).
-}
-  \label{fig:Clifford}
-\end{figure}
-
-The behaviour of the system with a parameter values of $x_0=0$, $y_0=0$, $a = -1.7$,  $b = 1.8$, $c = -1.9$ and $d = -0.4$ was examined.
-Data was scaled to fit inside the interval $[-1,1]$. Only the noise-free case was investigated. 
- A sequence of 25000 observationswas fed into the system with 15000 entries discarded to simulate the network's loss of memory. The next 1000 steps were predicted and compared with true data.
-The RNN was initialised exactly as for Thomas' Cyclically Symmetric Strange attractor: the dimension of the network is set equal to 500 neurons and 12 hidden layers of dimension 64 each were initialised.
-Training is realised through 150 epochs of batch sizes of 128 each; the map is learnt via the Adam Optimiser using finer and finer learning rates (0.001 and then divided by 10 at every iteration).
-
-\section{Functional Complexity Reduction in Noninvertible maps}
-
-% Besides being able to learn the action of the Koopman operator exactly, our setup has other advantages over selecting or optimizing the choice of observables as in the Extended Dynamic Mode Decomposition (EDMD) algorithm (see, [21, 22]).
-We can alter the functional complexity of G T by changing the number of observables or equivalently by altering the dimension of X – in the case of a RNN implementation of g, by simply changing the number of neurons in the network.
-Empirically, increasing the dimension of X is found to increase the linear relationship (or intuitively reduces the functional complexity of G T ) that is measured as a generalization of the Pearson correlation coefficient to random vectors (see, [44]) between (x_{n−1} , x_{n}) and G_T (x_{n−1} , x_n), and such numerical evidence is tabulated in [11, Table 1].
-In contrast, expanding the set of observables while using EDMD to learn a map with a lower functional complexity is a highly involved task, not least because so much that one often does not know how to guess a new observable, as well as the fact that adding observables does not necessarily retain the invariance of the span of the observables.
-
-% Empirical evidence shows that a linearity measure like a Pearson correlation coefficient points at reduced functional complexity while learning the dynamics in the state space of the driven system.
-
-% Empirically (see Table ??), increasing the dimension of X is found to increase the linear relationship
-% (or intuitively reduce the functional complexity of G T ) that is measured as a generalization of the Pearson correlation coefficient to random vectors (see, [?]) between (x n−1 , x n ) and G T (x n−1 , x n ).
-If Σ a and Σ b denotes the covariance matrices of the vectors (x n−1 , x n ) and G T (x n-1 , x n ) respectively, and if Σ ab denotes the covariance matrix between the vectors (x n-1 , x_n ) and G_T (x n-1 , x_n ), then the multidimensional 18correlation coefficient is computed by using the traces of these matrices:
-% \[\pho = \frac{tr(\sum_{ab})}{tr(\sqrt{\sum_{a}\sum_{b}})}\]
-
-% This multidimensional correlation coefficient satisfies most of the well-known properties of the one-dimensional Pearson coefficient.
-In particular, ρ = ±1 if and only Y = AX + b for some invertible matrix A and vector b.
-Hence ρ, and in practice, the estimator ρ̂, can be used to measure the linear relationship between two random vectors of the same dimension and thus serves as an indicator of functional complexity.
-
-% We remark that the estimation of such linear relationship in (see Table ??) obtained by sample correlations alone is justified whenever {u n } is a realization of an ergodic process since then when g has the USP, any solution {x n } is also a realization of an ergodic process [?].
-In the ergodic input case the generalized Pearson correlation coefficient (between (x n−1 , x n ) and G T (x n−1 , x n )) is independent of n.
-
-In the case of non-invertible maps, we note that the map $G_T$ is a homemorphism although $T$ is not.
-Further, empirically it is found that $G_T$ is much more linear than $T$ as measured by what is called the multidimensional correlation coefficient $\rho$.
-This coefficient is a generalization of the Pearson correlation coefficient for real-valued random variables to random vectors (see, \cite{puccetti2019measuring}).
-We state teh details of computing this coefficient between
-$(x_{n-1},x_n)$ and $G_T(x_{n-1},x_n)$. 
-If $\Sigma_{a}$ and $\Sigma_{b}$ denotes the covariance matrices of the vectors $(x_{n-1},x_n)$ and $G_T(x_{n-1},x_n)$ respectively, and if $\Sigma_{ab}$ denotes the covariance matrix between the vectors $(x_{n-1},x_n)$ and $G_T(x_{n-1},x_n)$, then the multidimensional correlation coefficient is computed by using the traces of these matrices:
-\[
-    \rho= \frac{\text{tr}({\Sigma_{ab}})}{\text{tr}({\sqrt{\Sigma_a\Sigma_b}})}.
-\]
-This multidimensional correlation coefficient satisfies one of the desired and well-known properties of the one-dimensional Pearson coefficient.
-That is, $\rho=\pm 1$ if and only if $Y\overset{d}{=}AX+b$ for some invertible matrix $A$ and vector $b$.
-Hence $\rho$, and in practice can be used  to measure the linear relationship between two random vectors of the same dimension and thus serves as an indicator of functional complexity.
-For the non-invertible system defined by the Clifford attractor, we indicate the difference by which $\rho$ changes for $G_T$ in Table~\ref{tbl_attractorsPearson}.
-Following \cite{manjunath2021universal}, we say that this multidimensional correlation coefficients $\rho$ to evidence a general reduction in the functional complexity of the map $G_T$ emerging due to a RNN.
-In the
-Table ~\ref{tbl_attractorsPearson}, the second column corresponds to the amount of artificial neurons (the dimension of $X$) in the implemented RNN.
-
-\begin{center}
-  \begin{table} 
-      \scalebox{0.75}
-      {\begin{tabular}{|c|c| c c |} 
-          \hline
-          Input & Dimension of $X$ & $u_n$ vs $u_{n+1}$ 
-          & $\begin{bmatrix}
-              x_{n-1}\\
-              x_n
-          \end{bmatrix}$ vs $\begin{bmatrix}
-              x_n\\
-              x_{n+1}
-          \end{bmatrix}$ \\
-          \hline      
-          % \hline
-          % $(w_n)$ is the Lorenz states, sampled every 0.1 timestamps. & & &\\
-          % \multirow{3}{*}{$u_n=\frac{1}{100}w_n$} & 10 &0.9311  & 0.9731\\
-          %     & 100 &  & 0.9934\\
-          %     & 1000 & & 0.9930\\              
-          % \multirow{3}{*}{$u_n=\frac{1}{10}(\sin(0.1w_{n,x})+\sin(0.1w_{n,y})+\sin(0.1w_{n,z}))$} 
-          %     & 10 &0.8401  & 0.9392\\
-          %     & 100 & & 0.9616\\
-          %     & 1000 & & 0.9737\\
-          % \hline
-
-               \hline  
-               $(w_n)$ comes from the Clifford Attractor \\ (for $a = -1.7; b = 1.8; c = -1.9; d = -0.4$): & & & \\
-        {$w_{n+1}= \begin{bmatrix} \sin(aw_{n,y}) + c\cos(aw_{n,x}) \\ 
-                                                                    \sin(bw_{n,x})+d\cos(bw_{n,y}) \end{bmatrix}$} & & & \\
-          \multirow{3}{*}{$u_n = w_n-\overline{w}$}
-              & 10 & -0.21713 & 0.78719 \\
-              & 100 & &  0.8334 \\
-              & 1000 & & 0.84854 \\
-          % \hline 
-          % $(w_n)$ comes from the Logistic map: & & & \\
-          % {$ w_{n+1}  = 4w_n(1-w_n)$} & & & \\
-          % \multirow{3}{*}{$u_n = w_n - 0.5$} 
-          %     & 10 &-0.0314  & 0.5577\\
-          %     & 100 &  & 0.7150\\
-          %     & 1000 & & 0.7826\\
-          % \hline
-          % \hline
-          % $(w_n)$ comes from the Svensson Attractor \\ (for $a = -1.7; b = 1.8; c = -1.9; d = -0.4$): & & & \\
-          % {$w_{n+1}= \begin{bmatrix} d\sin(aw_{n,x}) - \sin(bw_{n,y}) \\ 
-          %                                                             c\cos(aw_{n,x})+\cos(bw_{n,y}) \end{bmatrix}$} & & & \\
-          %   \multirow{3}{*}{$u_n = w_n-\overline{w}$}
-          %       & 10 & -0.2243 & 0.49698 \\
-          %       & 100 & &  0.63310 \\
-          %       & 1000 & & 0.71151 \\
-           \hline
-                \end{tabular}} %\ednote{http://paulbourke.net/fractals/peterdejong/}
- \label{Table_FC}
- \caption{Following \cite{manjunath2021universal}, we determine the multidimensional correlation coefficients $\rho$ to evidence a general reduction in the functional complexity of the map $G_T$ emerging due to a RNN.
-Rows correspond to distinct dynamical systems considered in the experiments
-      The second column corresponds to the amount of artificial neurons (the dimension of $X$) in the implemented RNN.
-      The last two columns comprise numerical estimates of $\rho$ for the relevant vectors. The linear association between $u_n$ vs $T(u_{n})$, as well as the linear relationship $\begin{bmatrix}
-              x_{n-1}\\
-              x_n
-          \end{bmatrix}$ vs $G_T \left( \begin{bmatrix}
-              x_{n-1}\\
-              x_n
-          \end{bmatrix}\right)$ may be contrasted by the reader. 
-          Finally, the RNN's dimension (from \eqref{Seq_RNN}) is provided to exhibiting that an increase in dimension of the driven system will 
-          typically result in a map $G_T$ with less functional complexity.}.
-      \end{table}\label{tbl_attractorsPearson}
-\end{center}
-
-\chapter{Conclusion}\label{ch6}
-\input{chapter6.tex}
-
-\bibliographystyle{pnas-new}
-\bibliography{pnas}
-
-\end{document}

### 2022-01-17 projectReport/chapter5.tex
Cleaner setup

--- 2022-01-17 draft_16Jan/hiscurrent.tex
@@
-\documentclass[a4paper,12pt,twoside]{report}
-% \documentclass[a4paper,12pt,twoside]{book}
-\pagestyle{headings}
-\usepackage[utf8]{inputenc}
-\usepackage{filecontents}
-\usepackage[many]{tcolorbox}
-\usepackage{graphicx}
-\usepackage{harpoon}
-%\usepackage{apacite} 
-\usepackage{xcolor}
-\usepackage{latexsym}
-\usepackage{amssymb}
-\usepackage{amsmath}
-\usepackage{mdframed}
-%\usepackage[authoryear]{natbib}
-\usepackage{hyperref, url}
-\usepackage[show]{ed}
-\usepackage{subfigure}
-\usepackage[ntheorem]{empheq} 
-\usepackage{amsthm, amssymb, amsfonts, latexsym}
-% \usepackage{pst-all}
-
-\usepackage{enumitem}
-\usepackage{multirow}
-\usepackage{longtable}
-\usepackage[export]{adjustbox}
-% $f\arrowvert_A$ To restrict functions
-
-\setlength{\parskip}{1.7em}
-\setlength{\parindent}{0em}
-
-%\setlength{\baselineskip}{20pt}
-
-\setlength{\textwidth}{429.75499pt}
-\setlength{\textheight}{643.20255pt}
-\setlength{\oddsidemargin}{5 mm}
-\setlength{\evensidemargin}{5 mm}
-\setlength{\topmargin}{0 mm}
-\setlength{\headsep}{0 mm}
-\setlength{\headheight}{0 mm}
-
-\usepackage{pst-node}
-
-\newcommand{\citealtt}[1]{\citeauthor{#1},\citeyear{#1}}
-\newcommand{\myycite}[1]{\citep{#1}}
-
-\mathchardef\mhyphen="2D
-
-%\newcommand{\cev}[1]{\reflectbox{\ensuremath{\vec{\reflectbox{\ensuremath{#1}}}}}}
-
-\usepackage{pst-node,graphicx,pst-blur}
-%\uspackage{auto-pst-pdf}
-%\usepackage{tikz-cd}
-
-\makeatletter
-\DeclareRobustCommand{\cev}[1]{%
-  \mathpalette\do@cev{#1}%
-}
-\newcommand{\do@cev}[2]{%
-  \fix@cev{#1}{+}%
-  \reflectbox{$\m@th#1\vec{\reflectbox{$\fix@cev{#1}{-}\m@th#1#2\fix@cev{#1}{+}$}}$}%
-  \fix@cev{#1}{-}%
-}
-\newcommand{\fix@cev}[2]{%
-  \ifx#1\displaystyle
-    \mkern#20mu
-  \else
-    \ifx#1\textstyle
-      \mkern#20mu
-    \else
-      \ifx#1\scriptstyle
-        \mkern#26mu
-      \else
-        \mkern#26mu
-      \fi
-    \fi
-  \fi
-}
-
-\makeatother
-
-\newcommand{\xcm}{\epsfxsize=3.1cm}
-\newcommand{\fig}[1]{\epsfbox}
-% \newcommand{\bp}{\begin{minipage}{3.1cm}}
-% \newcommand{\ep}{\end{minipage}}
-
-\newtheorem{Definition}{Definition}[]
-%\newmdtheoremenv{Theorem}{Theorem}[]
-\newtheorem{Theorem}{Theorem}[]
-\newtheorem{Lemma}{Lemma}[]
-\newtheorem{Proposition}{Proposition}[]
-\newtheorem{Corollary}{Corollary}[]
-\newtheorem{Remark}{Remark}[]
-\newtheorem{Example}{Example}[]
-
-%Shortcut symbols
-\newcommand{\Ftheta}{\ensuremath{F_\theta}}
-
-\makeatletter 
-\renewcommand{\thefigure}{\@arabic\c@figure}
-\makeatother
-\usepackage{xr}
-
-%%%% PNAS GRAPHICS SETUP- FROM THE PNAS STYLE LATEX FILE
-\RequirePackage{graphicx,xcolor}
-\RequirePackage{colortbl}
-\RequirePackage{booktabs}
-\RequirePackage{algorithm}
-\RequirePackage[noend]{algpseudocode}
-\RequirePackage{changepage}
-\RequirePackage[twoside,%
-				letterpaper,includeheadfoot,%
-				layoutsize={8.125in,10.875in},%
-                layouthoffset=0.1875in,%
-                layoutvoffset=0.0625in,%
-                left=38.5pt,%
-                right=43pt,%
-                top=43pt,% 10pt provided by headsep
-                bottom=32pt,%
-                headheight=0pt,% No Header
-                headsep=30pt,%
-                footskip=25pt,
-                marginparwidth=38pt]{geometry}
-\RequirePackage[labelfont={bf,sf},%
-                labelsep=period,%
-                figurename=Fig. ]{caption}
-\setlength{\columnsep}{13.5pt} % Distance between the two columns of text
-\setlength{\parindent}{12pt} % Paragraph indent
-
-%% Figure caption style
-\DeclareCaptionFormat{pnasformat}{\normalfont\sffamily\fontsize{7}{9}\selectfont#1#2#3}
-\captionsetup*{format=pnasformat}
-
-%%% GREYBOX AROUND FIG
-\definecolor{lightgray}{gray}{0.95}
-\newcommand\greybox[1]{%
-  \vskip\baselineskip%
-  \par\noindent\colorbox{lightgray}{%
-    \begin{minipage}{\textwidth}#1\end{minipage}%
-  }%
-  \vskip\baselineskip%
-}
-
-\newtcolorbox{paperbox}[1][]{
-    enhanced,
-    colback=white,
-    boxrule=0pt,
-    boxsep=0pt,
-    arc=0mm,
-    width=0.8\linewidth,
-    fuzzy shadow={0mm}{-4pt}{-4pt}{1mm}{black!30!white},
-    #1
-}
-
-% \externaldocument{Supp_Koopman-0808}
-\begin{document}
-
-\title{Data-Driven Modelling \& \\ Forecasting of Chaotic Systems}
-
-\author{B.L. Nortier} 
-\date{January, 2022}
-\date{}
-\maketitle
-
-\begin{abstract}
-  It is a well-established practice to collect measurements from underlying artificial, natural, and physical systems through observables in order to realise a dynamical model with greater descriptive power.
-  Many data-driven approaches to forecast the future evolution of systems exist.
-However, many of these methods rely on an initial understanding of the system and we often do not fully understand the many complexities of more intricate systems that we interact with on a daily basis.
-
-  In this project, the base assumption is that data is originates from  some discrete-time dynamical system, particularly one of a chaotic nature but no assumptions are made pertaining to a specific dynamical system.
-  A general learning problem is formulated and a well-known approach to the problem, alongside its practical limitations, is then considered.
-  The notion of driven dynamical systems is introduced, and we subsequently establish the theoretical existence of a dynamical system possessing certain properties which guarantees it to be topologically conjugate (`dynamically equivalent`) to the dynamics of the underlying system from which measurements were originally taken.
-  Our methodology is implemented practically by making use of recurrent neural networks, a form of deep learning, to learn a map topologically equivalent to the underlying system that forecasts future states.
-  In so doing we present the workability of an approach to obtain long-term topological and statistically consistent predictions for simple physical systems as well as a some chaotic attractors.
-  \end{abstract}
-
-  \tableofcontents
-
-\chapter{Introduction}\label{ch1}
-
-Experimenting with biological, physical, and artificial systems to generate a more informative dynamical model is a well-established practice in modern science.
-Traditional methods for modelling physical systems are based on laws of physics that are based either on empirical relationships or intuition.
-For systems that evolve with time, physical laws yield mathematical equations that govern how the quantities evolve with time.
-However, our world is much more complex than that which can be distilled into elegant equations.
-We do not fully understand many of the more complex systems, nor do they provide us with a good physical intuition of the underlying principles governing system dynamics.
-To complicate this, the underlying systems we observe often display sensitive dependence on initial conditions despite having highly similar initial conditions.
-This is because differing orbits diverge quickly and to such an extent that it becomes seemingly impossible to retrace their steps back to the original conditions.
-Even the presence of usually `negligible' computational noise or measurement error renders long-term pointwise prediction infeasible.
-We encounter many difficultiesin the process of modelling such systems:
-\vspace{-8mm}
-\begin{enumerate}[noitemsep, label=\roman*.]
-  \item One may not have access to the complete states of the systems
-  \item The system may be described by functions that behave wildly because their graphs have a wild oscillatory behaviour, i.e., a large functional complexity.
-\end{enumerate}
-
-Models derived primarily from data can be classified into three categories: 
-\vspace{-8mm}
-\begin{enumerate}[noitemsep, label=\roman*.]
-  \item  Interpretable models (i.e., they establish relationships between internal physical mechanisms), 
-  \item Partially interpretable models capturing some modes of the dynamics, 
-  \item Non-interpretable models, defined as such mainly due to the fact that they are defined on a different phase space that is usually high-dimensional. 
-\end{enumerate}
-
-Examples in the literature attempting to forecast data from such systems have tried many different approaches with differing degrees of success. 
-When we have complete access to states of the system, an ordinary differential equation can be obtained from data (\cite{brunton2016discovering, champion2019data} and \cite{small2002modeling,xu2006modeling}) wherein one could approximate the vector field by a library of functions to obtain interpretable models.
-
-Recent partially interpretable models available in the literature have been based on the Koopman operator (see~\cite{koopman1932dynamical,budivsic2012applied}) to employ observables mapping the data onto a higher-dimensional space.
-This makes the dynamics in the higher-dimensional space more amenable for approximation by a linear transformation.
-Such methods do not guarantee exact reconstructions for nonlinear models, and in practice provide poor long-term consistency for a large class of chaotic dynamical systems \cite{Supp}.
-
-The non-interpretable models include the delay embedding and the machine learning algorithms.
-For example, one could learn a system conjugate to the underlying system by applying the Takens delay embedding method of delay-coordinates~\cite{takens1981detecting} when one has "good" observations from the system; this learnt system could then be used to forecast the observed data.
-Takens delay embedding theorem~\cite{takens1981detecting} and its various generalisations (see,~\cite{sauer1991embedology, stark1999delay, gutman2018embedding}) establishes the learnability of a system constructed by concatenating sufficiently large previous time-series observations of a dynamical system into a vector (called delay coordinates).
-This then confirms the existence of a map on the space of delay coordinates equivalent (or topologically conjugate) to the underlying map from which the observed time-series was first obtained.
-Although topological conjugacy guarantees an alternate representation of the underlying system, the quality still depends on numerous parameters, making the comprehension of the dynamics unreliable (see, [?]).
-One reason for this fragility is that the embedded attractor in the reconstruction space is not always an attractor of the map learnt in the reconstruction space, despite unquestionably being an invariant set.
-When the embedded object is not explicitly known to be an attractor (as explained in Chapter~ \ref{ch3}), it can cause predictions to fail.
-
-Practically, the application of Takens embedding involves learning a map through some technique, and consequently one wishes that these would have low functional complexity\cite{manjunath2021universal}, i.e., functions with fewer oscillatory graphs.
-On the other hand, pure machine learning methodology processes temporal information (like the echo state networks~\cite{Manju_ESP, Manju_IEEE, grigoryeva2018echo}) by mapping data onto a higher dimensional space for further processing.
-Although they perform  well on forecasting some dynamical systems, they fail completely on others (\ednote{Ask for resources here}) as there is often no guarantee that the right function was learnt during training.
-
-This project involves implementing and analysing non-interpretable models that can guarantee exact reconstruction. 
-The project work concerns the study and implementation of a method (\cite{manjunath2021universal}) that incorporates learning a function by mapping the data on to a higher dimensional space using what is called a driven dynamical system (See Chapter~ref{ch.4}).
-With a clear understanding of how the data is mapped onto the higher dimensional space, the method then permits the learning of a dynamical system topologically conjugate to that of the underlying system.
-Instead of linear regression as in the training of echo state networks, deep learning methods are employed to learn the correct function.
-With slight modifications to the implementation in the paper (\cite{manjunath2021universal}), we show that one can construct accurate non-interpretable models with the ability to reconstruct attractors from more hard-to-forecast systems like the double pendulum.
-(The forecasting of the time-series from a double pendulum has not been reported before.) Moreover, we also demonstrate that long-term statistical consistency is preserved.
-
-\emph{By solidifying the mathematical underpinnings of our theory, we hope to guarantee the ability to construct models with predictive power ranging from molecular biology to neuroscience.
-in the near future.
-(We can modify this sentence when the report is completed.)}
-
-The more intricate details of proofs are referred to where relevant throughout. This report is organised as follows: 
-\newline In Chapter~\ref{ch2}, we recall the definition of a discrete-time dynamical system, how a discrete-system arises from a flow of an ODE and then proceed to define the inverse-limit space and topological conjugacy of autonomous systems.
-\newline Chapter~\ref{ch3} introduces the problem of forecasting dynamical systems, states the Takens delay embedding theorem, and discusses various issues faced while forecasting. 
-\newline In Chapter~\ref{ch4}, a driven dynamical system is defined and discuss the properties of a specific class of driven dynamical systems that we make use of in this project.
-\newline Finally, in Chapter~\ref{ch5} we show the implementation of these forecasting methods, and conclusions are provided in Chapter~\ref{ch6}.
-
-\chapter{Discrete-time Dynamical Systems}\label{ch2}
-
-This chapter briefly describes a discrete-time dynamical system and what it means to exhibit chaos. We refer to \cite{devaney2018introduction, de2013elements} for more details.
-
-At its most elementary level, a dynamical system is just something that evolves deterministically through time.
-In the context of this project, deterministic refers to the fact that a system evolves according to specified rules rather than based on random events.
-Dynamical systems arise in a variety of situations.
-A continuous-time dynamical system describes the states for all values of the time.
-Specifically, if the motion of a pendulum in which the quantities such as the angular position and angular momentum are known at all times, then it is a continuous-time dynamical system.
-The equations of the dynamical system can take the form of one or more ordinary differential equations that determine the relevant quantities at any future time if we know the initial location and momentum.
-In ecology, discrete-time dynamical systems are widely used to model population growth.
-The model in this case is a function calculating the following generation's population given the population of the previous generation.
-If we know the starting population, we may once again calculate the population at any time in the future.
-
-Formally, a function $T: U \to U$, where $U$ is some set is  a \emph{discrete-time dynamical system} and its iterates $\{u,Tu,T^2u,\ldots\}$, where $T^n$ denotes the $n$-fold composition of $T$ with itself, describe the evolution of an initial condition $u\in U$ (Note that we frequently drop the brackets and denote $T(u)$ by $Tu$ so as to simplify notations).
-
-Continuous-time dynamical systems modelled using ordinary differential equations can give rise to discrete-time dynamical systems.
-To see this, consider a differential equation $\dot{x} = f(x)$, $f: \mathbb{R}^n \to \mathbb{R}^n$, $n\in\mathbb{N}$ given to have a unique solution passing through
-each point $x\in\mathbb{R}^{n}$, call it $x(t)$ where $x(0)=x_0$.
-
-\begin{Definition}
-  [\bf Flow of an Equation] \label{Dfn_Flow}\rm
-  The flow of the equation  $\dot{x} = f(x)$  is defined to be a mapping $\varphi: \mathbb{R}^n \times \mathbb{R} \to \mathbb{R}$ where $\varphi(x_0,t)= x(t)$ for the solution $x(t)$ with $x(0)=x_0$ 
-\end{Definition}
-
-By fixing $t=K \in (0,\infty)$, we can define the \emph{time-$K$ map} as  $T(x):= \varphi(x_0,K)$, and it is easily verified that $T\circ T(x_0) = \varphi(x_0,2K)$.
-In general the $m^{\mbox{th}}$ iterate of $x_0$ under $T$ would be the value of the solution of the ODE evaluated at time $mK$ with the initial condition $x_0$, i.e.
-$\varphi(x_0, mK)$.
-Thus ordinary differential equations give rise to a discrete-time dynamical system by sampling the value of the solution $x(t)$ at time intervals $K$ units apart.
-
-A numerical discretization of a differential equation can also give rise to a discrete-time dynamical system.
-For instance, Euler's method approximates $\dot{x}(t)$ by $(x(t+h)-x(t))/h$; if $h$ is fixed throughout, the solution of a differential equation $\dot{x}=f(x)$
-can be approximated at the time instant $t+(m+1)h$ by iterating the equation 
-\begin{equation}
-  x(t+(m+1)h) = x(t+mh) + h f(x(t+mh))
-\end{equation}
-
-Adopting more succinct notation by replacing $x(t+mh)$ with $u_m$, we rewrite the above equation as
-\begin{equation}
-u_{m+1} = u_m + hf(u_m)
-\end{equation}
-or in even simpler terms as the discrete-time dynamical system with map $T(u) = u + hf(u)$, where $T$, $u$ and $f$ are understood to be as above.
-
-Of course, discrete-time dynamical systems need not always arise through a differential equation.
-Once again, we may consider the field of ecology, where discrete dynamical systems are often directly derived or assumed.
-\ednote{Ask for resource}
-There is a school of thought that advocates discrete-time dynamical systems to be more natural for modelling real-world observations than differential equations.
-We refer the interested reader to~\cite{saber2010introduction}.
-
-\section{Invariant Sets}
-
-A core concept in the study of dynamical systems is that of invariance. 
-\begin{Definition}
-  [\bf Invariant Set]\label{Dfn_InvariantSet}\rm
-  Given a discrete-time dynamical system $T: U \to U$, a subset $A \subset U$ is said to be an \emph{invariant set} if $T(A) =A$. 
-\end{Definition}
-
-We also define the orbit of a function $T$.
-\begin{Definition}
-  [\bf Orbit of $T$]\label{Dfn_Orbit}\rm
-  The orbit of $T$ is to be defined the sequence $\bar{u} = \{u_n\}_{n\in \mathbb{Z}}$ obeying the update equation, $u_{n+1}=Tu_n$, $n \in \mathbb{Z}$. 
-\end{Definition}
-
-Two examples of invariant sets include a fixed point where $Tu=u$ for  $u\in U$, and a periodic orbit, i.e., a set of iterates $\{u,Tu, T^2u,\ldots,T^pu\}$, where $T^{p+1}u=u$ for some $p\in\mathbb{Z}$.
-The entire space $U$ could be also be invariant.
-Consider for example the space $U=[0,1]$, where $Tu=4u(1-u)$ and then $U$ is invariant (as every $u\in{U}$ can be written as $u = 4x(1-x)$ for some $x\in{U}$).
-
-We may learn a great deal about the iterates of a dynamical system by considering the types of invariant sets of a discrete-time dynamical system.
-For example, if $U=[0,1]$ has map $Tu= u/2$, then the only invariant set is $\{0\}$, and all orbits approach this invariant set as time flows in the forward direction.
-Indeed, if a some non-zero invariant set (call it $B$) exists, then there is  some $r\in(0,1]\cap{B}$.
-But $r\notin{T(B)}$ since any orbit with initial value $r$ will be a decreasing sequence.
-Moreover, every orbit of T will be decreasing and therefore approach the value $0$ as $n\rightarrow\infty$.
-
-One may ask if every orbit approaches an invariant set?
-In general the answer is no, since for the dynamical system $T: \mathbb{R} \to \mathbb{R}$ defined by $Tu=2u$, any orbit that does not intersect the invariant set $\{0\}$, will not approach any invariant set.
-
-However,  when the space $U$ is compact, all orbits approach an invariant set.
-This is since the set of limit points of the orbit can be shown to be invariant \cite{de2013elements}.
-So, when $U$ is compact,  the $\omega$-limit set $\omega(u;T)$ of a point $u$ defined to the collection of limit points of the sequence $\{x,Tu,T^2u,\ldots\}$ is nonempty, and $\omega(u;T)$ is invariant.
-
-\begin{Example}
-  For the map $Tu=u^2$ defined on $[0,1]$ all orbits lie in the invariant set $\{1\}$ or else would would approach the invariant set $\{0\}$. 
-\end{Example}
-
-Invariant sets have various properties. Invariant sets can be attracting or repelling depending on how orbits in their vicinity behave. 
-Recall the examples $Tu =u/2$ and $Tu=2u$ defined on $\mathbb{R}$ (where $\{0\}$ “attracts" orbits) or the example, $Tu=u^2$ defined on $[0,1]$ (where $\{1\}$ “repels" orbits). 
-We are interested in attractive invariant sets since they capture the long-term dynamics as time increases. 
-In particular, we are interested in those invariant sets named attractors.
-
-\begin{Definition}
-  [\bf Attractor]\label{Dfn_Attractor}\rm
-  Let $T: U \to U$, where $U$  is a metric space with metric $d$. A compact subset $A \subset U$ is said to be an attractor if it satisfies the three conditions: 
-  \vspace{-8mm}
-  \begin{enumerate}
-	\item $A$ is invariant. 
-	\item $A$ is asymptotically stable, i.e., for every $\epsilon > 0$ and for all $u$ so that $d(u,A) < \epsilon$, we have $d(T^nu,A) \to 0$ as $n\to \infty$. 
-	\item $A$ has Lyapunov stability, i.e., for every $\epsilon > 0$  there exists a $\delta(\epsilon) > 0$ so that $d(u,A) < \delta$ implies $d(T^nu,A) < \epsilon$ for all $n\ge 0$.  
-\end{enumerate}
-\end{Definition}
-
-Indeed in our previous examples, it can be verified that the system  $Tu=u/2$ defined on  $\mathbb{R}$ and $Tu=u^2$ defined on $[0,1]$  the singleton set $\{0\}$ is an attractor.
-For the system,  $Tu=1-|2u-1|$ on $[0,1]$, one may easily verify that the only attractor is the entire space $[0,1]$.
-This follows from the fact that between any two points $u< v$ in $[0,1]$, and for any $a,b$ so that $u\le a < b \le v$, we can find an $n$ so that $T^n(a,b)=[0,1]$ (as would be explained later in this chapter).
-
-A dynamical system can have several attractors and may also be contained in another attractor.
-For the example, $Tu =u^2$ on $[0,1]$, both $\{0\}$ and $[0,1]$ are attractors.
-It is known that an attractor for a dynamical system always exists in a compact space~\cite{Milnor1985}.
-
-The dynamics restricted to an invariant set can be complicated.
-For instance an invariant set could be just a single point or it could have an infinite set.
-If the invariant set is infinite, then complicated dynamics are possible.
-A particular, well-studied phenomenon of such complexity gives rise to so-called chaotic behaviour, a subject studied in detail over the past fifty years.
-
-% If the dynamics on the attractor are somewhat complicated in the sense that the attractor cannot be decomposed further, and if the attractor is infinite then there is a possibility of complicated dynamics, and a particular well-studied phenomenon of such complexity gives rise to what is called a chaotic behaviour.
-
-%If the dynamics on an attractor cannot be decomposed further and the attractor is itself also infinite, then there is a possibility of complicated dynamics, and a particular well-studied phenomenon of such complexity gives rise to what is called chaotic behaviour.
-
-\section{Chaos}
-
-In the 1960s, several mathematicians and mathematically interested scientists independently discovered chaos in the mathematical sense.
-The meteorologist Edward Lorenz may have been the first to explain this phenomenon in his 1963 paper \cite{lorenz1963deterministic}.
-The notions of invariance, attractivity, and chaos may also be described for continuous systems, and Lorenz's system comprised of a system of differential equations.
-The narrative of Lorenz's discovery of chaos and the history of other forerunners in this subject is fascinating.
-We highly recommend James Gleick's book Chaos: The Making of a New Science~\cite{gleick2008chaos} for those interested.
-It clearly illustrates these experiences, explains why chaos was such a startling and crucial mathematical and scientific discovery and describes the underlying mathematical notions for non-specialists.
-
-Different authors propose a number of intricately different definitions of chaos in the literature~\cite{RasbandChaos, TaborChaos, WigginsChaos} and each of them indicate some aspect of complexity.
-
-In practice, when only data is observed from a system, it is not possible to verify which definition or notion is satisfied by the underlying dynamical system.
-We do, however, specifically recall the definition of chaos in the sense of Devaney \cite{devaney2018introduction,de2013elements} so as to understand some nuances behind the complexity.
-Devaney's definition of chaos has three requirements, with the first being the notion  of sensitive dependence on initial conditions.
-
-\begin{Definition}\rm
-  [\bf {Sensitive Dependence on Initial Conditions}]\label{Dfn_SDIC}\rm
-A dynamical system $T: U \to U$ is said to have sensitive dependence on initial conditions (SDIC) if there exists a $\delta > 0$ such that for every $u \in U$ and in every neighborhood of $u \in U$ there exists a $v\in{U}$ and an integer $N:=N{(u,v)}\in\mathbb{Z}$ such that $d(T^Nu,T^Nv)>\delta$.
-\end{Definition}
-
-It is vital to acquire a sense of this definition as it can easily be misunderstood.
-It is a common misconception to interpret SDIC as two close points ($u$ and $v$) that eventually become separated by a distance $\delta$ under iteration by $T$.
-But this is not true.
-To fully comprehend all the subtleties of the concept, one needs to discuss it more thoroughly by considering each sentence with care and attention, whereafter one may examine how they fit together to convey the idea of sensitive dependence.
-
-To this end, we make 3 remarks:
-\vspace{-5mm}
-\begin{enumerate}
-  \item First, the $\delta>0$ in the definition of SDIC is independent of $u$. 
-  \item Second, in every neighborhood of $u$, we may not necessarily find all points $v$ in the neighborhood distinct from $u$ that would separate from the forward iterates of $u$. 
-  \item  Finally, $N$ depends upon $u$ and $v$ chosen, and their iterates may not separate forever (i.e., for all $n>N$) and we allow their iterates to get arbitrarily close in the future. 
-\end{enumerate}
-
-To illustrate the concept of SDIC, we present two examples - one from Mathematics and the other from the field of Physics.
-
-\begin{Example} \rm
-  The logistic map(LM), a recursion relation of the form $x_{n+1}=rx_n(1-x_n)$ where $x_n\in[0,1]$ and $r\leq{4}$ is a classic example used to illustrate the chaotic behaviour that a system may exhibit.
-When $r=4$, the one-dimensional recurrence relation $x_{n+1}=4x_n(1-x_n)$ can be used as the kindergarten's model to exhibit the presence of SDIC.
-\ednote{B: edited and removed unnecessary citation}
-  In the graph below~\ref{fig:log_sdic} we plot the first 50 iterates of the LM for $r=4$ and two slightly different initial values $x_0$.
-
-  \begin{figure}[ht]
-    \includegraphics[scale=0.74]{_logistic_sdic.eps}
-        \centering
-        \captionof{figure}{Figure generated for Logistic Map $x_{n+1}=rx_n(1-x_n)$ with $r=4$ to exhibit the presence of SDIC.
-Plotted are the values $x_n$ against time $n$ for the the first 50 step with initial values $x_0=0.2$ (blue) and $x_0=0.20001$ (red).
-Initially the two trajectories overlap, but they diverge completely at $n=12$ whereafter they follow distinct paths.}
-        \label{fig:log_sdic}
-      \end{figure}
-  \ednote{B: Caption and figure adjusted}
-  % For an intuitive sense of the fact that the LM exhibits SDIC, consider again the definition and substitute specific values.
-Let $\delta=0.05$ and for purposes of illustration consider the value $u=0.4$.
-The green band in~\ref{fig:log_sdic} shades the region $[u-\delta, u+\delta]$ and we can see that regardless the value $v$  and in every neighborhood of $u \in U$ there exists a $v\in{U}$ and an integer $N:=N{(u,v)}\in\mathbb{Z}$ such that $d(T^Nu,T^Nv)>\delta$
-  % specific point, say $u=0.4$.
-Draw a green neighbourhood $[0.35,0.45]$ of diameters $\delta=0.02$ around the point.
-It is easy to see that every neighbourhood around $u=0.4$ will contain some $v_1, v_2\in[0,1]$ and natural number $N_1, N_2 \in\mathbb{N}$ respectively s.t.
-the distance $d(T^Nu_1,T^Nv)>\delta=0.05$ .
-\end{Example}
-
-\begin{Example} \rm
-  A second example pertains to a simple physical system called the double pendulum - a pendulum with another pendulum attached to its head given an initial position and angular velocity.
-
-  Below are two figures denoting trajectories of a double pendulum's second head after some elapsed time.
-Depicted are three pendulum heads with equal angular velocity but differing slightly in initial position.
-As can be seen below, the trajectories diverge very quickly to become completely different.
-We discuss the double pendulum in greater detail in a later chapter (\ref{ch5}), but we mention the example here to provide a numerical example of a practical system exhibiting SDIC.
-
-\begin{figure}[ht]
-  \centering
-\includegraphics[scale=0.8]{_dp_sdic.eps}
-    \captionof{figure}{Three double pendulum heads with equal angular velocity and initial angles differing by 0,025 radians}\label{fig:dp_sdic}
-  \end{figure}
-
-\end{Example}
-
-The second part in Devaney's definition of chaos concerns topological transitivity.
-
-\begin{Definition}
-  [\bf {Topologically Transitive}]\label{Dfn_TopolTrans}\rm
-	A dynamical system $F: U \to U$ is topologically transitive if for any pair of nonempty open sets $E_1$ and $E_2$ there exists a $n\in\mathbb{N}$ such that $T^n(E_1) \cap E_2 \not= \emptyset$. 
-\end{Definition}
-
-Topological transitivity implies that iterates of an open set of initial conditions get mixed up with other open sets.
-On a compact metric space, one may show that topological transitivity also implies the existence of a point whose forward iterates are dense \cite{de2013elements}; or in other words, the orbit going through this point will be dense in the compact metric space.
-In fact, it is topologically more likely that the choice of an arbitrary point will be one whose iterates are almost dense.
-
-\begin{Definition}
-  [\bf {Meager Set}]\label{Dfn_Meager Set}\rm
-A subset of a topological space $U$ is said to be a meager set if it can be written as a countable union of sets of with empty interior. The complement of the a meager set is set to be a residual set.
-\end{Definition}
-
-To be more precise (see \cite{de2013elements}), for a discrete-time dynamical system on a compact space, the set of points with dense iterates are residual, and they are typical or likely to be observed in practice.
-In this project, when data is observed from a topologically transitive system, we assume it arises from a dense orbit.
-
-We may now define the notion of Chaos as formulated by Devaney\cite{devaney2018introduction}.
-\begin{Definition}
-  [\bf {Devaney's Chaos}]\label{Dfn_ChaosDec}\rm
-	A dynamical system $T: U \to U$ is said to exhibit chaos in the sense of Devaney if it satisfies the three properties:
-	\vspace{-5mm}
-  \begin{enumerate}
-		\item $T$ has SDIC.
-		\item $T$ is topologically transitive.
-		\item The set of periodic points of $T$ are dense in $U$. 
-	\end{enumerate}
-\end{Definition}
-
-\begin{Example} \rm
-  The standard tent map $Tu=1-|2u-1|$ defined on $[0,1]$ is a well-known example of a dynamical system satisfying these three properties.
-
-  We now reason as to why this is true.
-The graph of the map $T$ is piecewise linear with two straight lines, one connecting the points $(0,0)$ and $(\frac{1}{2},1)$ and the other connecting $(\frac{1}{2},1)$ with $(1,0)$.
-They form a so-called tent with base centered at $u=1/2$.
-The graph of the map $T^2$ comprises two symmetric tents with their base centered at $1/4$ and $3/4$.
-See Figure \ref{fig:T2tentmap}
-
-  \begin{figure}[ht]
-    \includegraphics[scale=0.6]{_tentmap_2.eps}
-        \centering
-        \captionof{figure}{Graph of the $T^2$ map with bases at $x=0$, $x=\frac{1}{2}$ and $x=1$ respectively.}\label{fig:T2tentmap}
-  \end{figure}
-
-  In general, the graph of  $T^k$ contains $2^{k-1}$ tents.
-  Given any point $u$ and an open neighborhood $W \subset[0,1]$ of $u$, we can find $k\in\mathbb{Z}$ large enough so that $T^k(W) = [0,1]$ because we can accommodate a tent whose base is contained in $W$.
-% \ednote{B: This last phrase doesn't make sense to me.
-M: The number of tents grow with $k$ and hence this is obvious; I explained in one of our meetings.
-Does it make sense?}
-  This implies the existence of two points $w_1$ and $w_2$ in $W$ so that $d(T^kw_1,T^kw_2)= 1$, and by the triangle inequality  at least one of the inequalities $d(T^ku,T^kw_1)> \delta$ or $d(T^ku,T^kw_2)> \delta$ hold when $\delta\in (0,\frac{1}{2})$.
-So $T$ has sensitive dependence on initial conditions.
-
-  Next, let $W_1$ and $W_2$ be two nonempty open sets.
-Given any open set $W_1$, we can find a $k\in\mathbb{N}$ large enough so that $T^k(W_1) = [0,1]$ because we can accommodate a tent with base contained in $W$.
-So $T^k(W_1) \cap W_2 \not=\emptyset$, and thus $T$ has topological transitivity.
-
-  Finally, for every open interval $(a,b)$, the graph of $T^k$ intersects the graph of the identity map on $[0,1]$.
-This follows from the fact established above that there exists a $k\in\mathbb{N}$ such that $T^k(a,b) = [0,1]$.If the intersection point has coordinates of the form $(p,p)$, the point $p$ is then a fixed point of $T^k$ and therefore also a periodic point of $T$.
-Since we have found a periodic point in the interval $W$,  the set of periodic points are dense in $T$.
-  This follows because any fixed point of $T^k$ is also a periodic point, and we have found this in an arbitrary interval.
-
-\end{Example}
-
-\section{Conjugacy}
-
-We now turn our attention to the subject of conjugacy that describes when two dynamical systems are equivalent dynamically.
-
-To show topological similarity (or sameness) between two metric or topological spaces, one needs to establish a homeomorphism between the two spaces. 
-However, in the study of dynamical systems defined on two spaces, establishing a homeomorphism does not indicate that the systems are dynamically related in any way.
-For instance, the maps $Tu=u^2$ and $Tu=1-|2u-1|$ defined on $[0,1]$ have totally different behaviour. 
-Therefore, to find dynamically similar systems, one must establish a dynamical equivalence that we illustrate in a simple commutativity diagram below~\ref{eqn_conjugacy}.
-
-% Giving errors
-\begin{equation}  \label{eqn_conjugacy}
-%    \[ 
-    \psset{arrows=->, arrowinset=0.25, linewidth=0.6pt, nodesep=3pt, labelsep=2pt, rowsep=0.7cm, colsep = 1.1cm, shortput =tablr}
- \everypsbox{\scriptstyle}
- \begin{psmatrix}
-U & U\\%
-V & V.
- %%%
-%  \ncline{1,1}{1,2}^{T} 
-%  \ncline{1,1}{2,1} <{\phi}
-%   \ncline{2,1}{2,2}^{S}
-%  \ncline{1,2}{2,2} > {\phi}
- \end{psmatrix}
-% \]
-\end{equation}
-
-%Composite functions
-To understand the diagram, we note that if we travel "right and then down," the diagram instructs us to use $T$ (top arrow right) first, followed by $\phi$.
-(right arrow downwards).
-Consequently, the pathway amounts to finding $\phi(T(u))$.
-If we proceed "down and then right," the diagram instructs us to apply $\phi$ (left arrow downwards) first, and then apply $S$ (bottom arrow right).
-This then amounts to finding  $S(\phi(u))$.
-When the relationships in this diagram hold, we say $\phi(T(u))= S(\phi(u))$, and we formally denote it as $\phi \circ T=S\circ \phi$.
-
-\begin{Definition}\rm  
-[\bf {Conjugacy \& Semi-Conjugacy}]\label{Dfn_Conjugate}\rm
-   Consider the dynamical systems $T:U\to{U}$, $S:V\to{V}$ and suppose the relationship $\phi \circ T=S\circ \phi$ holds.
-If $\phi:U\to{V}$ is a homeomorphism, then $S$ is said to be conjugate to $T$ and $\phi$ is called a conjugacy.
-If we relax the criterion on $\phi$ and merely require $\phi$ to be continuous and surjective where $\phi:U\to{V}$, then $\phi$ is a semi-conjugacy between $T$ and $S$ where T has domain $U$ and S has domain $V$; S is said to be semi-conjugate to T.
-\end{Definition}
-
-\ednote{B: Does defining explicitly the domain and codomain of $\phi$ make it clear enough?
-M: Saying $\phi$ is then a semi-conjugacy between $T$ and $S$ could be ambiguous as it does not specify which system is on the top in the commutativity.
-So please edit accordingly.}
-When $S$ is conjugate to $T$, the dynamics of the two systems are in some way `dynamically equivalent'.
-Specifically, they are in one-to-one correspondence with one another.
-However when $S$ is semi-conjugate to $T$ with $\phi:U\to{V}$ a many-to-one mapping, the dynamics on $V$ provide merely a coarse-grained description of the dynamics on $U$~\cite{de2013elements}.
-When $S$ is semi-conjugate to $T$, it is also common to call $S$ a \emph{factor of $T$}, or conversely that $T$ is an \emph{extension of $S$}.
-In essence, an extension (see, \cite{de2013elements}) is a more extensive system capturing all of the essential dynamics of its factor.
-
-It is a very hard or nearly an impossible task to establish the existence of a conjugacy or a semi-conjugacy $\phi$ between two systems \cite{devaney2018introduction}.
-However, one can verify that a function $\phi$ satisfies the commutativity diagram \ref{eqn_conjugacy}.
-
-\begin{Example}\rm
-  For example $\phi(x)=\sin\big(\frac{\pi}{2}x\big)^2$ is a conjugacy between two systems $Tu=1-|2u-1|$ and $Sv=4v(1-v)$  defined on $U=[0,1]$.
-
-  Indeed, $\phi:[0,1]\to[0,1]$ is a homeomorphism and by employing simple trigonometric identities, it is easily proved that $\phi$ satisfies~\ref{eqn_conjugacy}
-  \[
-    \begin{aligned}[t]
-      (\phi\circ{T})u 
-                &=\sin^{2}\Big(\frac{\pi}{2} - \frac{\pi}{2}|2u-1|\Big) = \sin^{2}(\pi{u})=\Big(2\sin(\frac{\pi}{2})\cos(\frac{\pi}{2})\Big)^{2}=4\sin^{2}(\frac{\pi}{2}u)\Big(1-\sin^{2}\frac{\pi}{2}u\Big) \\
-                                    &=({S}\circ\phi)u.
-    \end{aligned}
-\]
-And it follows that $\phi$ is a conjugacy between the two systems.
-\end{Example}
-
-By establishing that two systems are conjugate (semi-conjugate) to one another, we have also shown that one may choose to work with one system instead of the second and still be guaranteed to obtain information on the latter.
-It is instrumental during the process of forecasting the future evolution of dynamical systems.
-
-\chapter{The Learning Problem}\label{ch3}
-
-In this chapter, we acquaint ourselves with the question of forecasting dynamical systems with unknown underlying structures, state and discuss the shortcomings of the Takens delay embedding theorem and discuss various issues faced while forecasting.
-
-Consider a relatively simple learning problem: 
-Given the sequence $(u_0, u_1, \ldots, u_m)$ for $m\in\mathbb{Z}$, a finite segment of an orbit of the map $T$ where $T$ is defined by the update equation $u_{n+1} = Tu_n$, forecast the values $u_{m+1}, u_{m+1}$ where the map $T$ is unknown, given that $u_m$.
-
-A practical example of this would be the subsequent scenario: given the time-sequential coordinates of an object moving in space, predict the future positions of that object.
-We are very rarely, if at all, presented with a problem where the entire state information is available to us in the form $u_n$ at some timestep $n\in\mathbb{N}$.
-Consider then a more complicated learning problem:
-
-Suppose we only have the observations $\theta(u_0), \theta(u_1), \ldots, \theta(u_m)$ of the true system states $u$ in an unknown dynamical system $(U,T)$ and we wish to predict the values $\theta(u_{m+1}), \theta(u_{m+2})$ and so forth.
-
-First we establish the method of Takens delay embedding. 
-In this method, one considers a discrete-time dynamical system defined on a 'nice' space (a smooth manifold)\ednote{B: Should this be further explained?} that can be obtained as time-$K$ map of a continuous time-dynamical system, concepts formalised below.
-We make observations from such a system, i.e., we observe the evolution of $\theta(w_0)$ (where $w_0$ represents the initial state) by examining  a finite set of values of $u_n :=  \theta(w_n) = \theta(Tw_{n-1})$.
-The sequence $\{u_n\}$ represents a scalar time-series and intuitively $\theta$ represents a probe inserted into a bigger system which is itself only measuring/extracting a small part of the greater system state at time $t=n$.
-Consider, for example, a thermometer erected to measure the ambient temperature in a local village.
-This measurement function, the thermometer, is capturing only a single aspect, the temperature, of a much grander dynamical system entailing the current weather of the surrounding area.
-Even more than that though- it is measuring a miniscule part of the global weather system.
-
-%Given the sequence $\{u_1, u_2, \ldots, u_n}$, we define the variable \newline $y_n:=[\theta_n, \theta_{n-\tau}, \ldots, \theta_{n-(2d-2)\tau}, \theta_{n-(2d-1)\tau}]^{T}\in\mathbb{R}_{d}\times{1}$ where $\tau$ represents the lag and $d$ the dimension of the attractor.
-(We shall return to these terms shortly and their discussion may be put on hold for the time being).
-
-However, the actual state $u$ of a system is seldom, if ever, fully known.
-the most one can do is to insert probes into a system to obtain partial information through the measurements taken.
-Moreover, the process of taking a measurement itself introduces two additional aspects which complicate the problem: \ednote{B: Perhaps this edit makes more sense?
-M: Maybe replace difficulty by saying that we need to consider two aspects.
-}
-\vspace{-8mm}
-\begin{enumerate}[noitemsep, label=\roman*.]
-  \item A series of measurements over a specific time interval is inherently a discretisation procedure of the underlying continuous-time dynamical system.
-(Hence why we restricted our attention to discrete-time systems in the preceding section)
-  \item A probe will never be entirely accurate, and so the act of measurement introduces a certain measure of numerical noise/inaccuracy.
-\end{enumerate}
-
-From here we construct a multidimensional observable using the method of stacking previous observations, i.e., we create delay-coordinate map defined by
-$\Phi_{k,\theta}(w) := (\theta(T^{-k}w)\ldots,\theta(T^{-1}w),\theta(w))$.  
-The concept of an observable is understood in the sense of an observable as originally introduced by~\cite{takens1981detecting, genericObservableAeyels} to refer to a probe or measurement function inserted into the system.
-Takens' theorem, in essence, refers to the fact that for a sufficiently large $k$, we can define a dynamical system on the space $\mathbb{R}^{k+1}$ with states $\Phi_{k,\theta}(w)$, $\Phi_{k,\theta}(Tw)$, $\Phi_{k,\theta}(T^2w)$, etc.
-This dynamical system is topologically conjugate to the unknown underlying system $(W,T)$.
-We recall the Takens delay embedding theorem next.
-
-%We may then ask ourselves the question: Is it in any way possible to retain information about the system state $x(t)$ in this temporal data-series $\theta(x(t))$?
-The answer is easily yes if T is known.
-However, neither this, nor even the exact function $\theta$ is available.
-
-\section{Takens Embedding Theorem}\label{sect_Takens}
-
-We define first the concepts of homeomorphism and embedding:
-\begin{Definition}\rm
-  [\bf {Homeomorphism}]\label{Dfn_homeo}\rm
-  A homeomorphism is a function $f:Z\rightarrow Y$ between two topological spaces $Z$ and $Y$ that is continuous, bijective and has a continuous inverse. 
-\end{Definition}
-
-\begin{Definition}
-  [\bf {Embedding}]\label{Dfn_embed}\rm
-  Consider a homeomorphism $f:Z\rightarrow Y$ for $Y\subset X$. $Z$ is said to be embedded in $X$ by $f$.
-\end{Definition}
-
-Takens Theorem states a result establishing a relationship between the observed and underlying dynamical systems by showing that the concatenation of a sufficiently large number of previous observations into a vector will, under certain conditions, generate a map between the vectors from the respective systems.
-We formulate the theorem from \cite{takens1981detecting}.
-
-\begin{Theorem} 
-	[\bf Takens Embedding Theorem (adopted from \cite{takens1981detecting}] \label{Thm_Takens}
-         Let $W$ be a compact manifold of dimension $m$, and $d\ge m$ so that $2d$ is an integer. It is a 
-            generic property for the pair $(T, \theta)$,  where $T:W \to W$ is
-            a smooth diffeomorphism, and $\theta:W \to \mathbb{R}$ a smooth function, the map $\Phi_{2d,\theta}:W \to \mathbb{R}^{2d+1}$ defined on $W$ by 
-            $\Phi_{2d,\theta}(w) := (\theta(T^{-2d}w)\ldots,\theta(T^{-1}w),\theta(w))$
-            is a diffeomorphic embedding; by `smooth' we mean at least $C^2$.
-Consequently, there exists a map $F_\theta: \Phi_{2d,\theta}(W) \to \Phi_{2d,\theta}(W)$ defined by $$F_\theta: (\theta(T^{-2d}w),\ldots,\theta(T^{-1}w),\theta(w)) \mapsto
-            (\theta(T^{-2d+1}w),\ldots,\theta(w),\theta(Tw))$$
-           so that $(W,T)$ is topologically conjugate to 
-            $(\Phi_{2d,\theta}(W), F_\theta)$.    
-\end{Theorem}
-
-By generic we mean a residual set on a certain topology on appropriate spaces of functions (that we do not describe here) \ednote{B: Where can I find a citation for this?}. 
-By $C^2$, we make reference to a twice-differentiable function  with a continuous second derivative.
-In our scenario (and this remains important throughout) the input space $U$ is considered the attractor.~\label{attractor_U}
-
-Below we reproduce Figure 1 from \cite{Supp}.
-
-%Some explanation is beneficial.
-Takens establishes a delay-coordinate map $\Ftheta$ defined in \ref{eqn_takens} for $T$ the flow (\ref{defn_flow}) defined by the update equation  $Tu_n=u_{n+1}$ and $theta$ our measurement function.
-When confining a dynamical system to the manifold U (in which the underlying system is contained), Takens showed that if certain smoothness conditions are satisfied on $T$ and $\theta$, then the delay-coordinate map $\Ftheta$ embeds $U$ in the reconstruction space $R^{2d+1}$ \textbf{we haven't used the term reconstruction space yet, so this is falling out of the blue}, for almost every choice of measurement function $\theta$, the observable.
-By almost every choice here, we mean that we exclude some observations which do not give any information.
-To see a more precise discussion, see \textbf{xxx} for we do not concern ourselves with a further elaboration on this report.
-
-%Alternatively we may also make the statement: Takens' Embedding Theorem guarantees that almost all dynamical systems can be reconstructed from just one noiseless observation sequence \textbf{this isn't obvious from the theorem's formulation, is it?} i.e.
-for a great number of possible observation functions $\theta$, $\Ftheta$ preserves the topology of U.
-\textbf{'preserving the topology' is not something we've touched on before.
-Perhaps I should add to previous discussion so logic flows seamlessly here}.
-
-\begin{figure}[ht]
-  \includegraphics[scale=0.25]{_takensmap.eps}
-  \centering
-  \captionof{figure}{Schematic of Takens Embedding Theorem's usage and impact of iterates slipping from the attractor when the approximation $\tilde{F}_\theta$ of $F_\theta$ is learnt.}
-  \label{fig:takensmap}
-\end{figure}
-
-If we can learn the map $F_\theta$, from sufficient number of data points of $\{\phi_{2d,\theta}(w_n)\}$, then we also know how to forecast how $\theta(w_n)$.
-
-%Once again we pause and consider the graph to consolidate our understanding of the conjugacy (\textbf{are we being too simplistic?}).
-If we are in state $u$ and then evolve towards state $Tu$ before embedding into $\mathbb{R}^{2m-1}$, then this is equivalent to first embedding into $\mathbb{R}^{2m-1}$ and then evolving through the map $\Ftheta$, i.e.
-$\Phi_{2d,\theta}\circ{F}_{\theta} = T\circ\Phi_{2d,\theta}$.
-The map $F_theta$ is a homeomorphism as well.
-
-%Thus information about U can be retained in the time series' (or observation's) output.
-By preserving the topology on the manifold $U$ in the reconstruction space $X$ (\textbf{not discussed before}), we are also guaranteed the preservation of topological invariants of the manifold, of which dimensionality is one such invariant.
-(We note here that dimensionality will again be considered later in this project).
-
-\section{Practical Limitations to Takens' Theorem}
-Takens' Embedding Theorem is a powerful result, and it provides compelling reason to believe that one could conceivably reconstruct a system conjugate accurately to the underlying system. 
-Nonetheless, it does present some severe practical limitations:
-\vspace{-5mm}
-\begin{enumerate}
-\item Even supposing that we could indeed find $F_\theta$ , our \emph{approximation} of  $F_\theta$ is a map from a larger set $\mathbb{R}^{2d+1}$ containing the embedded attractor.
-There are, however, no theoretical guarantees that $F_\theta$ will retain  $V=\Phi_{2d,\theta}(W)$  as an attractor although $W$ could itself be an attractor.
-\item Takens theorem is stated only for noiseless observations.
-Due to noise $\epsilon_n$ the delay vector $\Phi_{2d,\theta}(w_n) + \epsilon_n$  may lie outside $V$.Furthermore, due to the chaotic nature of the underlying system (i.e.
-the fact that it has SIDC), the evolution of $\Phi_{2d,\theta}(w_n) + \epsilon_n$ under the map $F_\theta$ could move out of $V$ completely.
-This problem can be overcome  by using a driven dynamical system with some properties, and we discuss this in Chapter\ref{ch3}.
-\end{enumerate}
-
-% The Takens embedding theorem does not guarantee global dissipativity.
-% \ednote{B:Links with~\ref{subs_LearnGamma} in discussion of dissipativity of $\Gamma$.
-Sir, I'm not sure how to write this here - I think it a beneficial point to make so as to set the stage for the later discussion in ~\ref{subs_LearnGamma}, but I am unsure there as well.}
-
-Though not a uniquely related to the Takens embedding theorem, we do opt to make mention of another requirement for our work. 
-The methods described here are applicable for data originating from a surjective map. 
-This follows from the fact that when $T$ maps the space $U$ into a proper subset $B$ of$ $U, there exists some point in $U - A$ that does not have a preimage and we conclude that $T$ is not surjective.
-An example of this would be the case when a system displays energy-loss.
-Consider, for instance, consider the dying oscillations for a damped pendulum.
-(An account of greater depth  is provided in~\ref{ch5})
-
-This requirement is not overly restrictive, as a great number of chaotic systems do have a surjective map.
-(Cite) \ednote{B: I remember you saying this in one of our sessions, but don't have a source for it.
-Do you have a specific reference?}
-
-\chapter{Driven Dynamical Systems to Forecast Problems}\label{ch4}
-
-In this chapter, we discuss results on the mapping of temporal data obtained from a discrete-dynamical system onto a different space through the notion of a driven dynamical system. 
-We also consider the conditions a driven system should possess to avoid adding distortion to its state-space representation.
-We then describe the \emph{single-delay dynamics} (SDD) of the system and finally recall conditions on the driven system so that the SDD are conjugate (or at least semi-conjugate) to the underlying system.
-The SDD can then be used to forecast and reconstruct the underlying system.
-
-\section{Nonautonomous and Driven Dynamical Systems}
-
-\begin{Definition}
-  [\bf Nonautonomous Dynamical System (adopted from~\cite{Manju_ESP}))] \label{Dfn_NDS}\rm
-  A nonautonomous dynamical system (NDS) on a space $X$ is simply a dynamical system comprising a family of maps $\{g_n\}_{n \in \mathbb{Z}}$ where $g_n(\ldotp):=g(u_n,\ldotp):X\to{X}$ is continuous.
-Each $g_n$ arises as a consequence of an input $u_n$ from the input space $U$, a topological space.
-\end{Definition}
-
-We immediately recall the figure first defined in \eqref{eqn_conjugacy}  and remind ourselves of our ultimate goal to learn a system that is conjugate to an underlying or unknown dynamical system $(U,T)$ by making use of the measurements obtained from  the unknown system.
-
-%#####################################
-\begin{center}
-\psset{arrows=->, arrowinset=0.25, linewidth=0.6pt, nodesep=3pt, labelsep=2pt, rowsep=0.7cm, colsep = 1.1cm, shortput =tablr}
-\everypsbox{\scriptstyle}
-\begin{psmatrix}
-U & U\\%
-V & V.
-%%%
-%  \ncline{1,1}{1,2}^{T} 
-%  \ncline{1,1}{2,1} <{\phi}
-%   \ncline{2,1}{2,2}^{S}
-%  \ncline{1,2}{2,2} > {\phi}
-\end{psmatrix}
-%#####################################
-\end{center}
-
-To find a function $\phi$, the need now arises to consider a so-called driven dynamical system, a special case of a nonautonomous dynamical system.
-Two spaces are considered: input is taken from an input space $U$ and the state of a space $X$ is updated.
-(Previously, only the state was considered.)
-We do this to mimic the true conditions.
-By accounting for an extrogenous input sequence in $U$ which can also influence the  system-state at a specific time-step, it produces more general models than an autonomous system.
-
-\begin{Definition}
-  [\bf Driven, Compactly Driven Dynamical System] \label{Dfn_DDS} \rm
-A driven dynamical system comprises two topological metric spaces $U$, $X$ and a continuous function  $g:U\times{X}\to{X}$ where $g(u_n, x_n)=x_{n+1}$.
-If the input space $U$ is compact, we refer to the system as compactly driven. 
-\end{Definition}
-
-The update equation $x_{n+1} = g(u_n,x_n)$ for $n \in\mathbb{Z}$, input $u_n$ from $U$ and state $x_n$ belonging to $X$ (a compact space) generate the dynamics on X. 
-Abbreviated, we refer only to the \textit{driven system $g$}, where all other entities are understood implicitly.
-
-Notably, a nonautonomous dynamical system can be generated from $U$; any input $\overline{u}$, a bi-infinite sequence from $U$, gives rise to the sequence of self-maps $\{g(u_n, \cdot)\}_{n\in\mathbb{Z}}$ contained in $X$.
-Physically, one may think of this bi-infinite sequence as referring to a system that has been running for an extended period at the time of the first measurement taken from the system (alternatively, the first time a probe is inserted into the system to take an observation)
-
-\begin{Definition}
-  [\bf Entire Solution] \label{Dfn_Soln} \rm
-  A sequence $\{x_n\}_{n\in\mathbb{Z}}\subset X$ is called an entire solution (or simply a solution) of the driven system  $g$ with input $\overline{u}$ when it satisfies 
-  \[g(u_{n-1}, x_{n-1})=x_n\] for all $n\in\mathbb{Z}$
-\end{Definition}
-
-It is important to emphasise that a sequence $\{x_n\}$ satisfying the update equation above can only be a solution if $x_n\in{U}$ holds for all $n\in\mathbb{Z}$ and just for $n>0$.
-Consider the example below.
-
-\begin{Example} \rm \label{ex_halfux}
-  The only solution  $\{x_n\}_{n\in\mathbb{Z}}$ to the driven system  $g(u,x)=\frac{ux}{2}$, where $X=[0,1]$, $U=[0,1]$,  is the zero solution $x_n\equiv0$.
-  To see this, consider any $x_n=a\in[0,1]$ where $a\neq{0}$.  Let $\overline{u}\in{U}$ be an non-zero constant sequence , say $u_n=0.5$. 
-  The driven system may be rewritten as $x_{n-1}=\frac{2x_n}{u_{n-1}}=4x_n$ and the  iterates of $x_n$ in backward time will increase by a factor of 4 at each timestep. 
-  Thus for some $m\leq{n}$,  $1<x_m$ i.e. $x_m\notin{X}$. So ${\{x_n\}}_{n\in\mathbb{Z}}$ is not a solution and it follows that the only possible solution is the zero solution.
-\end{Example}
-% \ednote{Nicely written! Thank you}
-
-A system may also have multiple solutions, as evidenced in the example below.
-
-\begin{Example}\label{Ex_exp} \rm 
-  Consider the driven system $g(u,x)=x^2$ for $X=[0,1]$, $U=\mathbb{R}$.
-The system has an uncountable number of solutions, as there exists a solution for every $x\in{X}$ which also passes through the point $x$ and $lim_{n\to\infty}x_n=0$, $lim_{n\to{-}\infty}x_n=1$.
- The proof is deferred to immediately after the next paragraph (see~\ref{rem_proofEx}).
-\end{Example}
-
-As the solutions to a driven system are considered an important entity, we next identify a subspace $X_U$ of $X$ that contains all possible solutions.
-To realize such a subspace of a driven system $g$, the concept of a reachable set is defined.
-
-\begin{Definition}
-  [\bf Reachable Set]\label{Dfn_ReachableSet}\rm
-The reachable set of a driven system $g$ is exactly the union of all the elements of all the (entire) solutions, i.e., 
-\[X_U :=\Big \{x \in X:  x = x_k \mbox{ where  $\{x_n\}$  is a solution for some  $\bar{u}$} \Big \}.\]
-The set of all reachable states at a specific time $n$ for input $\overline{u}$ is denoted by $X_n(\overline{u})$
-\end{Definition}
-
-The reachable set itself can defined independently from $g$ possessing the ESP/USP.
-Note that $x\in{X_n(\overline{u})}$ if and only $g$ has a solution $\{x_k\}$ for $x_n=x$ and input $\overline{u}$, a result proved in nonautonomous dynamical systems literature~\cite{manjunath2014dynamics, manjunath2013echo}.
-\ednote{B: Adjusted sentence and cited}
-
-The set $X_n(\overline{u})$ is precisely the set of points $x$ through which some entire solution $\Psi$ passes at time $n$. We can give an alternate formulation of the set $X_n(\overline{u})$. 
-Now suppose, we for some fixed input 
-$\overline{u}$, we define $g_i = g(u_i,\cdot)$ for all $i\in \mathbb{Z}$, then the set of states
-
-\begin{equation} \label{eqn_association}
-X_{n,i}(\overline{u}) := g_{n-1} \circ \cdots g_{i+1} \circ g_i(X).
-\end{equation}
-
-The set $X_{n,i}(\overline{u})$ being the image of a finite composition of
-continuous maps is compact whenever $X$ is compact. Also $X_{n,i}(\overline{u})$ is
-nonempty.  Further, $X_{n,i}(\overline{u}) \supset X_{n,i-1}$. Hence $\bigcap_{i<n}
-X_{n,i}(\overline{u})$ is a nested intersection of closed nonempty subsets, and
-whenever $X$ is compact, the intersection is nonempty. We say $g$ is a topological contraction if $\bigcap_{i<n}X_{n,i}(\overline{u})$ is a singleton subset of $X$ for each $\overline{u}$.
-
-It turns out that $\bigcap_{i<n}X_{n,i}(\overline{u})$ is identically equal to $X_n((\overline{u})$ that we had defined earlier.
-This will lead to a result (for.
-see, \cite{manjunath2013echo}) that  \textit{$g$ being a topological contraction is equivalent to the existence of a exactly one entire solution.}
-
-% \begin{Definition}
-%   [\bf Topological Contraction]\label{Dfn_TopContr}\rm
-%   A function $g:U\times{X}\to{X}$ is a topological contraction if for all $n\in\mathbb{Z}$ and all $\overline{u}\subset{U}$, $X_n(\overline{u})$ is a singleton subset of $X$.  
-% \end{Definition}
-
-\begin{Remark}
-  [\bf Proof of Example~\ref{Ex_exp}] \label{rem_proofEx} \rm
-  For the driven system,  
-  $g(u,x)=x^2$ the evolution does not depend on the input at all.
-Hence $g(u,x)$ can be written as a single map $f(x)=x^2$ which is a homeomorphism on $[0,1]$.
-A left-infinite orbit is defined and it converges to $1$ while the right-infinite orbit converges to $0$.
-Or in other words the iterates of $f^{-1}$ converges to $1$ while the iterates of $f$ converge to $0$
-\end{Remark}
-
-Thus far, it has been demonstrated that a system may have one or more solutions; one may ask if a driven system always has a solution and, if so, whether it satisfies specific properties such as uniqueness.
-Should the driven system be compact, existence follows immediately, as shown in the following result.
-
-\begin{Theorem}\label{Thm_CompactExistence}
- Let $g$ be a driven system.  If $X$ is compact then for each input $\overline{u}$, there exists at least one solution to the driven system $g(\ldotp, x)$
-\end{Theorem}
-\begin{proof}
-  May be found in \cite{kloeden2011nonautonomous, manjunath2014dynamics, manjunath2013echo}
-\end{proof}
-
-% One may easily construct many systems with trivial solution-sets, such as $g(u,x)=x$ which has only the constant solution $x$ and so for $U=[-1,1]$, the system would have no solution if $|x|>1$.
-To refine the scenario, we consider only systems with unique solutions.
-
-\section{Unique Solution Property}
-
-\begin{Definition}
-  [\bf Unique Solution Property] \label{Dfn_usp}\rm
-  A driven system $g$ is said to have the Unique Solution Property (USP) if for each input $\overline{u}$ there exists exactly one solution. 
-  Alternatively we may formulate the USP as follows: $g$ has the Unique Solution Property if there exists a well-defined map $\Psi:{U}\to{X}$, with $\Psi({\overline{u}})$ denoting the unique solution.
-\end{Definition}
-
-One of the first results obtained after defining the USP is that every solution will attract different initial conditions towards the component parts of the solution.
-If $g$ has the USP, then any solution to $g$ is also a uniform attractor in nonautonomous dynamical systems literature~\cite{Manju_Nonlinearity}.
-The discussion on nonautonomous attractors is beyond the scope of this project, and we refer the reader to \cite{Manju_ESP, esann2012ids} for further reading.
-
-% Paraphrased.
-%Additionally, $g$ having the USP is in our context also equivalent to $g$ being a topological contraction \cite{manjunath2021universal}, or alternatively also that $g$ exhibits the Echo State Property (ESP).
-
-In the majority of Reservoir Computing(RC) literature (a machine learning approach) where the input is mapped onto a different but higher dimensional through a system called the reservoir and a readout that measures the state of the reservoir is trained, a notion of forgetting the states of the reservoir asymptotically is often used.
-
-% \ednote{How should I explain/qualify RC better here?
-It's mentioned without really explaining.
-M: Have put it in one sentence; just to let you know people use optical devices to map signal into a ``reservoir".
-B: Thanks!},
-Concepts like the ESP,  fading memory \cite{boyd1985fading} are some of these notions.  
-If $g$ possesses the ESP (equivalent in our context to the USP), the we are guaranteed that the whole of an input's left-infinite history will exactly determine the current system state; i.e.
-there is only ever the possibility of a single reachable state at a given instance of time \cite{jaeger2001echo,Manju_2020}.
-% Defined formally, the ESP is intrinsically linked to an input (see \cite{Manju_ESP}) Unfortunately, definitions and results concerning the ESP do not have much practical benefit since a relationship between the network's response and the temporo-statistical properties of the input are not directly described.
-A more detailed presentation may be found in~\cite{jaeger2001echo}.
-
-% We do however consider it important to mention that the ESP is also often discussed in terms of a stability property that holds for all inputs of a system~\cite{manjunath2020stability} and, furthermore, plays a key role in the design and training of recurrent neural networks(RNN) within the field of RC \cite{Manju_ESP}.
-Beyond forgetting the past states the ESP is also often discussed in terms of a stability property that is nearby inputs create close by responses (solutions), a concept called input-related stability~\cite{manjunath2020stability} and, furthermore, plays a key role in the robustness of recurrent neural networks(RNN), a component of the implementation discussed in~\ref{ch5}.
-
-\section{Conjugacies}
-
-Having already defined the reachable set $X_U$ as the collection of all elements of all entire solutions, we pause briefly in order to fix additional notation.
-Defining $\cev{u}^{n}:=(\ldots,u_{n-2} ,u_{n-1})$ as the left-infinite subsequence of an input up until time $n$, $\overleftarrow{U}$ is then the notation for all these left-infinite sequences in $U$.
-Moreover, $\cev{u}^{n}v:=(\ldots,u_{n-2} ,u_{n-1}, v)$ will symbolise the input up to time $n$ with $v \in U$ being the specific input value at time $n$. 
-The introduction of a new input at time $n$ can be described by the mapping $\sigma_v:   \cev{u}^{n} \mapsto \cev{u}^{n}v$. 
-The right-shift map, $r\overleftarrow{U}\to\overleftarrow{U}$, of an input sequence is defined $r: (\cdots, u_{-2},u_{-1}) \mapsto(\cdots, u_{-3},u_{-2})$
-
-The question now assumes the form: Can we establish a semi-conjugacy for the driven system $g$ as presented below.
-
-\begin{equation}  \label{Scomm_h}
-  %    \[ 
-      \psset{arrows=->, arrowinset=0.25, linewidth=0.6pt, nodesep=3pt, labelsep=2pt, rowsep=0.7cm, colsep = 1.1cm, shortput =tablr}
-   \everypsbox{\scriptstyle}
-   \begin{psmatrix}
-   \overleftarrow{U} & \overleftarrow{U}\\%
-   X_U & X_U.
-   %%%
-  %  \ncline{1,1}{1,2}^{\sigma_v} \ncline{1,1}{2,1} <{h}
-  %  \ncline{1,2}{2,2} > {h}
-  %  \ncline{2,1}{2,2}^{g(v,\cdot)}
-   \end{psmatrix}
-  % \]
-  \end{equation}
-
-We proceed to consider a specific subclass of conjugacies.
-
-  \begin{Definition}
-    [\bf Universal Semi-Conjugacy]\label{Def_UnivSemiConj} \rm
-    Given a driven system $g$, we  call a continuous and surjective map $h : \overleftarrow{U} \to X_U$ a universal semi-conjugacy if  diagram \ref{Scomm_h} commutes for all $v \in U$.
-  \end{Definition}
-
-  If the universal semi-conjugacy $h$ exists (i.e. the diagram in \ref{Scomm_h} commutes) then the solution $\Psi(\bar{u})$ will intuitively have no more ``complexity' than the input $\bar{u}$.
-
-But does such a function $h$ for the above driven system $g$ above exist?
-Whenever $g$ has the USP and $\Psi(u)=\{x_n\}_{n\in\mathbb{Z}}$ it follows that $h$, defined by  $h(\cev{u}_n):=g(u_n,x_{n-1})=x_n$, will satisfy the semi-conjugacy in the graph above \ref{Scomm_h}.
-Regrettably, such a continuous mapping $h$ is not guaranteed to exist when $g$ does not have the USP \cite[Lemma 5]{Manju_Nonlinearity}.
-Note that even when $h$ does exist, we are not guaranteed its injectivity. Considering again example \ref{ex_halfux}, we see that even if $h$ were to exist, it could not be injective as $X_U=\{0\}$.
-
-Re-sketching the commutativity diagram \ref{Scomm_h} above by replacing $X_U$ by its left-infinite sequence space $\overleftarrow{X}_U$, we obtain the diagram below.
-In this case, the function $H:\overleftarrow{U}\to\overleftarrow{X}_U$, a map that is both continuous and surjective, is called a \emph{causal mapping} that is defined next.
-
-\begin{Definition}
-  [\bf Causal Mapping]\label{Def_CausMap}
-  A continuous, surjective map $H:\overleftarrow{U}\to\overleftarrow{X}_U$ such that \[H\circ\tilde{g}_v=\sigma\circ{H}\] holds for all $v \in U$ where $\tilde{g}_v$ maps $(\ldots, u_{-2}, u_{-1})$ to $(\ldots, u_{-2}, u_{-1}, g(v, u_{-1}))$ is called a causal mapping.
-\end{Definition}
-
-\begin{equation} \label{SCausal_H}
-    %    \[ 
-        \psset{arrows=->, arrowinset=0.25, linewidth=0.6pt, nodesep=3pt, labelsep=2pt, rowsep=0.7cm, colsep = 1.1cm, shortput =tablr}
-     \everypsbox{\scriptstyle}
-     \begin{psmatrix}
-     \overleftarrow{U} & \overleftarrow{U}\\%
-     \overleftarrow{X}_U & \overleftarrow{X}_U.
-     %%%
-    %  \ncline{1,1}{1,2}^{\sigma_v} \ncline{1,1}{2,1} <{h}
-    %  \ncline{1,2}{2,2} > {h}
-    %  \ncline{2,1}{2,2}^{\tilde{g}_v}
-     \end{psmatrix}
-    % \]
-  \end{equation}
-
- \begin{Theorem}
-  For a compactly driven system, a causal mapping $H$ exists if and only if $g$ has the USP. 
-\end{Theorem}
-{\bf Proof.}  See~\cite[Th.3]{manjunath2013echo}.
-
-The driven system $g$ can induce an embedding of $\overleftarrow{U}$ in $\overleftarrow{X}$ as follows: 
-If the causal mapping $H:\overleftarrow{U}{\to}{\overleftarrow{X}_U}$ is injective (in addition to being surjective), it becomes the embedding of $\overleftarrow{U}$ in $\overleftarrow{X}$ induced by the driven system $g$.
-The continuity of $H^{-1}$ follows from the fact that $H$ is itself a continuous and surjective mapping of a compact space $\overleftarrow{U}$ in a Hausdorff space.
-and we refer to the map $H$ as a \emph{causal embedding}. 
-\ednote{B: edited the sentence here}
-
-When $g$ has the USP, the diagram below (adopted from~\cite{Manju_Nonlinearity}) illustrates the operation of the mappings $h$ and $H$.
-The mapping $h:\overleftarrow{U}\to{X_U}$ is also considered an observable as mentioned in the introduction of Chapter \ref{ch3}).
-
-\begin{figure}[ht]
-  \includegraphics[scale=0.4]{_actionofh_H.eps}
-  \centering
-  \captionof{figure}{Ladder-like behaviour of $h$ and $H$ to illustrate the causal mapping. (Figure reproduced from~\cite{Manju_Nonlinearity}) }
-  \label{fig:actionh_H}
-\end{figure}
-
-The function $H$ contains a countably infinite number of coordinate (or factor maps) since its codomain is $\overleftarrow{X}_U$.  
-As it is not practically feasible to consider a left-infinite sequence in $X_U$ (in other words an element of  $\overleftarrow{X}_U$) for learning, we wish to restrict the inputs we consider to a subspace of $\overleftarrow{U}$.
-Particularly, we restrict our attention to the  so-called inverse-limit space of a dynamical system and hope to embed it within some finite product of $X$.
-
-%We wish to explore the possibility of establishing a conjugacy rather than a mere semi-conjugacy and therefore are interested in the restriction of inputs to the left-infinite orbits of a dynamical system.
-
-The left-infinite orbit space is significant because a great amount of information about the original system can be gleaned from its topological structure~\cite{ingram2011inverse,Manju_IEEE}. 
-The formal topological structure for the inverse-limit space of a chaotic dynamical system is complicated, and further discussion is relegated to~\cite{kennedy2008inverse_limit, ingram2011inverse}.
-Broadly, we may conceive of the inverse-limit system of a dynamical system as a subspace of an infinite-dimensional space where each point in the inverse-limit space corresponds to a backward orbit of the map $T$.
-Following~\cite{manjunath2021universal, ingram2011inverse} we provide a definition:
-
-\begin{Definition}
-  [\bf Inverse-Limit Space, Inverse-Limit System]\label{Dfn_InverseL}\rm
- The subset $\widehat{U}_T\subset \overleftarrow{U}$ defined by \[\widehat{U}_T: = \{ (\ldots,u_{-2},u_{-1}): Tu_{i} = u_{i+1}\} \] is called the inverse-limit space of $(U,T)$.
- The self-map $\widehat{T}$  induced by $T$ on $\widehat{U}_T$  is defined by \[\widehat{T}: (\ldots,u_{-2},u_{-1}) \mapsto  (\ldots,u_{-2},u_{-1},T(u_{-1}))\] and the resulting dynamical system $(\widehat{U_T}, \widehat{T})$ is called the inverse-limit system of $(U,T)$.
-\end{Definition}
-
-The inverse-limit space is well-defined since $T : U \to{U}$ is surjective by assumption.
-We now get to the concept of embedding the inverse limit-space of the dynamical system in $X \times X$. 
-\begin{Definition} \rm
-	We say a driven system $g$ \emph{causally embeds} a dynamical system $(U,T)$ if it satisfies the two properties: (i) a universal semi-conjugacy exists and  (ii)  $H_2(\cev{u}) := (h(r\cev{u}),h(\cev{u}))$ embeds the inverse-limit space $\widehat{U}_T$ in $X \times X$.
-\end{Definition}
-
-  It is imperative to take note of a subtlety here: we refer to the \emph{map} $H$ as a (causal) embedding, whereas the driven \emph{system} $g$ (causally) embeds another \emph{system} $(U,T)$.
-\ednote{B: inserted line}
-
-\section{Choosing the driven system $g$}
-
-So far it is not clear if the USP alone ensures the ability to causally embedding a dynamical system and we mention that a driving function $g$ cannot just be chosen in a frivolous manner.
-This could possibly complicate our work.
-%\ednote{B: Grammarly}
-
-%The driving function $g$ cannot be chosen in a frivolous manner as it could possibly complicate our work.
-Specifically, one must be careful to avoid a choice of $g$ which would  add complexity to the obtained solution.
-
-%When a causal embedding $H$ exists for the driven system $g$, one can map an arbitrary input ${u}$ onto the solution space $X$ without additional distortion or information-loss \textbf{(Cite.)}.
-
-%When an embedding is established, the question of possible additional complexity in the solution is removed by guaranteeing that, since the systems are conjugate (semi-conjugate, \textbf{(refer)}), $g$ does not add any (some) complexity to the system.
-
-%To achieve a causal embedding embedding not just a causal mapping, it is undesirable to choose a function $g$ that quenches the temporal structure in $u$ by contracting to such a degree that the ability to recover information from the original system is lost completely.
-In the above example(\ref{ex_halfux}), the input's temporal variation could not be related to the reachable set as $X_U$ only consisted of a single element; little, if not no, information is encoded.
-To obtain a suitably complex function $g$, it is thus desired that the reachable set of a driven system  be large enough to relate to the input. 
-This is true even for embedding  the inverse-limit space of a dynamical system. \ednote{B: Expand or refer?}
-
- To this end, we recall the notion of State-Input (SI) Invertibility.
-
-\begin{Definition}
-  [\bf SI-Invertibility]\label{Dfn_SIinv}\rm
-  A driven system $g$ is said to be SI-Invertible if $g(*,x): U \to X$ is invertible for all $x\in X$.
-Alternatively it may be said that if, given $x_n$ and $x_{n-1}$, $u_{n-1}$ can be uniquely determined from $x_n=g(u_n,x_{n-1})$, then $g$ is said to be SI-invertible.
-\end{Definition}
-
-SI-invertibility promises that 'enough' information is retained when $g$ is chosen without introducing one's choice of $g$ will still ensure 'enough' information is retained without introducing unwanted complexity.
-'Enough' here refers to the fact that we may always recover the previous input value $u_n$ (if we know successive states $x_{n-1}, x_n$).
-
-Subsequently we define the relation $Y_T$ induced by $(U,T)$ on $X_U\times{X_U}$ for a driven system $g$ possessing SI-invertibility.  
-To describe the  SDD formally, we consider a dynamical system $T: U \to U$ and define a relation on the reachable set $X_U$, i.e. a subset on $X_U \times X_U$  defined by 
-$$Y_T:=\{(x_{n-1},x_n): \{x_k\}_{k\in \mathbb{Z}} \mbox{ is a solution for some orbit of } T \mbox{ and } n \in \mathbb{Z}\}.$$
-
-The following theorem establishes the existence of a well-defined map $G_T$ describing the single-delay dynamics(SDD) of the system above.
-
-\begin{Theorem}\label{Thm_GT_Exists}
-  For an SI-invertible driven system $g$ and a dynamical system $(U,T)$, the map $G_T: Y_T \to Y_T$ defined by the relation $(x_{n-1},x_n) \mapsto (x_n,x_{n+1})$ is well-defined. 
-  (This results holds even in the absence of $g$ possessing over the USP) 
-  % Moreover, the mapping $(x_{n-1},x_n) \mapsto u_{n}$ is well-defined when $x_{n-1}$ and $x_n$ are successive points on a solution obtained for an input  $\{u_n\}$ that is an orbit of $T$.
-  \end{Theorem}
-  {\bf Proof.} See~\cite[Th.3]{Supp}
-
-%'Without additional complexity' is guaranteed by the invertibility of the function $g$ that guarantees a one-to-one mapping between $u_n, x_{n-1}$ and $x_n$.
-In obtaining an SI-invertible function, therefore, we are guaranteed a system which will not lose so much information about its previous states in the forward-flow of time that one may not make any credible claims as to the original system's behaviour and properties.
-Simultaneously, it does guarantee us that the important information encoded in previous inputs is indeed preserved.
-
-At this stage it is worth taking note of a specific driven system in the form of a discrete state-space model which has acquired some adherence in applications \cite{Manju_IEEE}- especially those pertaining to Echo State Networks and the ESP.
-The function
-\begin{equation}  \label{eqn_driving}
-  g(u,x) = (1-a)x + a\overline{\tanh}(Au + \alpha Bx)
-\end{equation} 
-is SI-invertible and, if $\alpha B$ has a spectral norm $<1$, also possesses the USP \cite[Th.2]{manjunath2013echo }. 
-It is easy to show that $g$ is SI-invertible by recovering $u_n$ in 
-\begin{equation} \label{eqn_SI_RNN}
-  u_{n-1} := A^{-1}\bigg(\overline{\tanh}^{^{-1}}\frac{1}{a}\Big(x_{n+1}-(1-a)x_n\Big) \bigg) - \alpha B x_n
-  \end{equation}
-  when $x_{n-1}$ and $x_n$ are known.
-
-This specific driving function $g$ is used in our implementation and is discussed more completely in chapter~\ref{ch5}
-
-Despite the ease that with which one manipulates a left-infinite history in the realm of theory, it is impossible to obtain or use such a sequence in any real-life application.  
-Fortunately, one does not need the entire left-infinite history of an input in practice thanks to the Uniform Attraction Property(UAP). 
-We use an alternate version of the definition of UAP -- since the notion of nonautonomous attractors in this project would take some time to establish and detracts from the principal thrust of this project, we refer to \cite{Manju_Nonlinearity}.
-
-\begin{Definition}
-  [\bf Uniform Attraction Property]\label{Dfn_UAP}\rm
-  A driven system $g$ has the Uniform Attraction Property (UAP) if we initialize the driven system
-with an arbitrary initial value $y_m \in X$, and the sequence $y_{m+1}, y_{m+2}, y_{m+3},...$ satisfying $y_{k+1}= g(u_k,y_k)$ for $k \geq m$ then approximates an actual solution $\{x_n\}$ uniformly in the sense that given $\epsilon>0$ (independent of $y_m$) there is an integer $n$ so that $d(x_{n+i}, y_{n+i})<\epsilon$ for all $i\ge 0$, where $\{x_m\}$ is a solution.
-\end{Definition}
-
-The UAP guarantees that all trajectories will converge to the same trajectory as time moves forward.
-An illustration of this is shown in Fig.~\ref{fig:memloss_conttime} where the trajectory in red locks on to the supposed actual solution when the system has the USP.
-
-\begin{figure}[ht]
-  \includegraphics[scale=0.35]{_Nonlinearity_Fig_response.eps}
-  \centering
-\captionof{figure}{A coordinate of a simulated solution $(x_0,x_1,\ldots,x_{5000})$ of a RNN  plotted in red (with parameter $\alpha=0.99$ (where $g$ has the USP) and blue ($a=1$ and $\alpha=1.05$ where $g$ does not have the USP) while the matrices $A$ and $B$ are randomly generated, and $B$ has unit spectral radius.
-Reproduced from \cite{Manju_Nonlinearity}. }
-\label{fig:memloss_conttime} 
-\end{figure} 
-\ednote{B: How can I explain the black line? I don't think this makes sense to the reader immediately.}
-
-One may now appreciate an even more astounding result: $g$ having the USP is equivalent to the UAP as is proved in~\cite{Manju_Nonlinearity}.
-
-%We restate the above theorem.
-
-%\begin{Theorem}\label{thm_usp-contr-uap}  The following statements are equivalent:  \vspace{-8mm}  \begin{enumerate}[noitemsep, label=\roman*.]    \item $g$ has the USP.
-\item $g$ is a topological contraction.
-\item $g$ has the UAP.
-\end{enumerate\end{Theorem}{\bf Proof.}  May be found in \cite[Th.1]{Manju_Nonlinearity}
-
-% One need therefore not establish any additional results to ensure that the sequence uniformly approaches the unique solution $\Psi$.
-This vastly simplifies the effort necessary to set up a problem in order to guarantee that the underlying system will be accurately represented by the conjugate system.
-
-%This also solves the problem of perturbation/noise introduced by the observable or measurement function.
-Recall from before that an observable is inherently a map that discretises the underlying continuous-time system.
-Moreover, measurement and device error introduce mistakes.
-Since we are working with a chaotic system displaying sensitive dependence on initial conditions, these small errors could potentially send the trajectory into a completely different attractor.
-%In the presence of the UAP, however, small measurement errors introduced into the system do not pose the same danger as before since the trajectories will eventually tend towards the actual solution.
-% \ednote{(B: Is this Complete enough?)}
-% \textbf{From the perspective of perturbation of an autonomous dynamical system, if the resulting nonautonomous system has the USP, then the temporal structure of noise relates to the dynamics on a nonautonomous uniform attractor through a topological semi-conjugacy or a conjugacy.}
-
-\section{The next step in Dynamics}
-
-% Recall that $g$ is only being given inputs from the orbits of $T$.
-
-%Incredibly, this permits one to initialise a driven system $g$ with an altogether arbitrary initial value $y_m\in{X}$ where $m\in\mathbb{Z}$ and the UAP then guarantees that the sequence $\{y_{m+1}, y_{m+2},\ldots\}$ which satisfies the relation $y_n=g(u_n, y_{n-1})$  for  $k\geq{m}$  will approach the elements $\{x_n\}$ of the actual solution.
-This vastly simplifies the effort necessary to set up a problem in order to guarantee that the underlying system will be accurately represented by the conjugate system.
-
-We are interested in determining whether $\widehat{T}$ and $G_T$ are related.
-For this, we introduce  the function $H_2{\overline{}} := (h(r\overline{u}, h(\overline{u})))$ mapping the entirety of a left-infinite sequence to some element in $X\times{X}$.
-Indeed, it may be shown that $g$ having SI-invertibility and the USP immediately guarantees the existence of at least a semi-conjugacy between the systems $(Y_T, G_T)$ and $(\widehat{U}_T, \widehat{T})$.
-This is formalised in the Causal Embedding Theorem, which we state next:
-
-\begin{Theorem}
-  [\bf Causal Embedding Theorem (adopted from \cite{Supp})]
-  % {\bf (Causal Embedding Theorem.)}
- \label{Thm_CET}
-	Let $g$ be a driven system with SI-invertibility and the USP.
-Let $h$ denote the universal semi-conjugacy and $H_2(\overleftarrow{u}) := (h(r\overleftarrow{u}),h(\overleftarrow{u}))$, where $r$ is the right-shift map.
- Let $(\widehat{U}_T, \widehat{T})$  be the inverse-limit system of a dynamical system $(U,T)$. 
- Then the restriction of $H_2$ to $\widehat{U}_T$ is a topological semi-conjugacy between the inverse-limit system $(\widehat{U}_T, \widehat{T})$ 
-and the induced dynamical system  $(Y_T,G_T)$, i.e., the following diagram commutes
-\begin{equation} \label{comm_H_CET}
-\psset{arrows=->, arrowinset=0.25, linewidth=0.6pt, nodesep=3pt, labelsep=2pt, rowsep=0.7cm, colsep = 1.1cm, shortput =tablr}
- \everypsbox{\scriptstyle}
- \begin{psmatrix}
- \widehat{U}_T & \widehat{U}_T\\%
- Y_T &  Y_T.
- %%%
-%  \ncline{1,1}{1,2}^{\widehat{T}} \ncline{1,1}{2,1} <{H_2}
-%  \ncline{1,2}{2,2} > {H_2}
-%  \ncline{2,1}{2,2}^{G_T}
- \end{psmatrix}
- \end{equation}
-or in other words, $(Y_T, G_T)$ is a factor of  $(\widehat{U}_T, \widehat{T})$. Further, if $T:U \to U$ is a homeomorphism, 
-then $H_2$ embeds $\widehat{U}_T$ in $X_U \times X_U$, and hence $(Y_T, G_T)$ is conjugate to $(\widehat{U}_T, \widehat{T})$.
-\end{Theorem}
-{\bf Proof.}  May be found as the proof of~\cite[Th.4]{Supp}
-
-Recall that $H_2$ maps an entire left-infinite solution sequence from $\Psi$ to an element in $X\times{X}$. 
-We are drawing nearer and nearer to our principal target and our results carry more and more weight: If we are able to learn $G_T$, we will also have learnt $\widehat{T}$; and since $\widehat{T}$ is an extension of $T$, we will have obtained $T$ as well.
-%\ednote{B: Grammarly check}
-
-%\begin{Theorem} Graph \ref{Scomm_h} is exactly the inverse-limit system $(\hat{U}, \hat{T})$.    \end{Theorem}
-
-%We now have the following (which may be compared with \ref{SCausal_H} above):
-
-%\begin{equation} \label{fig:inverse_limsystem}
-  %  \[       \psset{arrows=->, arrowinset=0.25, linewidth=0.6pt, nodesep=3pt, labelsep=2pt, rowsep=0.7cm, colsep = 1.1cm, shortput =tablr}      \everypsbox{\scriptstyle}
-%      \begin{psmatrix}      \widehat{U}_T  && \widehat{U}_T \\     Y_T && Y_T \\
-      %%%
-     %  \ncline{1,1}{1,2}^{\widehat{T}} \ncline{1,1}{2,1} <{h}
-     %  \ncline{1,2}{2,2} > {H_2}
-     %  \ncline{2,1}{2,2}^{G_T}
-%      \end{psmatrix}
-    %  \]
-%  \end{equation}
-
-%\begin{Theorem}
- %   $(Y_T, G_T)$ is semi-conjugate to $(\widehat{U}, \widehat{T})$.
-%\end{Theorem}
-% {\bf Proof.} See \cite[Th.3, Th.4]{manunath2021universal}
-
-\section*{Summarising the discussing thus far:}
-
-It is easy to lose the birds-eye view, so we take a moment to review our progress up until this point.
-
-\vspace{-8mm}
-\begin{enumerate}
-\item We are interested in a some dynamical system $(U,T)$ with unknown dynamics.
-\item To determine properties about this system $(U,T)$ and predict its future evolution, we determine the dynamics of the inverse-limit system $(\widehat{U}, \widehat{T})$.
-% Given certain assumptions, we can guarantee that $(\widehat{U}, \widehat{T})$ is at least semi-conjugate to $(U,T)$.
-\item If the driven system $g$ is SI-invertible (and $\{u_n\}\in{U}$  is an orbit of $T$), the map $G_T$ exists that describes the single delay dynamics. 
-\item If, furthermore, $g$ has the USP, $(Y_T, G_T)$ is semi-conjugate to $(\widehat{U}, \widehat{T})$.
-\item If we can assume that $T$ is a homeomorphism, $(Y_T, G_T)$ is topologically conjugate to $(\widehat{U}, \widehat{T})$, an extension space of $(U,T)$
-\end{enumerate}
-
-One can therefore in practice learn the SDD of the driven system states via $G_T$ with enough data thanks to the USP/UAP. This enables us to do at least two things: 
-\vspace{-8mm}
-\begin{itemize}
-\item Forecast  $x_{n+1},x_{n+2}, \ldots$ via iterates of $G_T$ (if $G_T$ can be learnt).
-\item Forecast future values of $u_n$ since $x_n$ and $x_{n+1}$ determine $u_n$ since $g$ is SI-invertible. 
-\end{itemize}
-
-Finally, we note that although $G_T$ exists with SI-invertibility, we need the USP as well(see \ref{fig_pictorialSummary} below for a visual illustration).
-If not, the driven states would have to be running for all time since they would not have forgotten their past states.
-
-\begin{figure}[ht]
-  \includegraphics[scale=0.3]{_summarypictorial.eps}
-  \centering
-  \captionof{figure}{Pictorial summary of results obtained up to now and how they overlap. Reproduced from \textbf{(Cite SIG-25May)} }
-  \label{fig:fig_pictorialSummary}
-\end{figure}
-\ednote{Ask M: Cite SIG-25May.}
-
-\section{A discussion of $G_T$ }
-
-In the above sections, we established the map $G_T$ describing the SDD of a driven system.
-By establishing the existence of this map, we’ve essentially embedded the attractor $U$ (recall from \ref{attractor_U}) into the higher dimensional space $X\times{X}$.
-In layman’s terms, this ensures that there is more “dimensional room” for the underlying system’s underlying dynamics to “move”.
-Since the dynamics aren’t as “squashed”, we might therefore hope that the dynamics of $G_T$ are in some sense simpler than that of $T$.
-(Taken note of the fact that $G_T$ is a homeomorphism even when $T$ is just continuous)
-
-In \cite{manjunath2021universal} it is illustrated in an empirical fashion in that the map $G_T$ describes dynamics which are less functionally complex than that of $T$ or of the map $\Phi_{2d,\theta}$.
-This is done by implementing a Recurrent Neural Network (RNN), which is discussed in~\ref{ch5}.
-\ednote{B: Should I add in that we'll be using Pearson coefficient?}
-
-We opt to learn $G_T$ in an indirect manner by defining a new map $\Gamma:(x_{n-1},x_n)\mapsto{u_n}$. It will follow immediately from $G_T$’s existence that $\Gamma$ also exists.
-
-The reasons for taking this roundabout approach remain to be discussed in \ref{subs_LearnGamma}, but a pat answer may immediately be given: When $\Gamma$ has been learnt, the system can be driven autonomously and then $G_T$ is known anyway.
-We formalise this with a theorem.
-
-\begin{Theorem}
-  When $x_n$ and $x_{n-1}$ are successive points on a solution obtained for input-orbit of $T$ $\{u_n\}$, then the map $\Gamma: X\times{X}\to{U}$ defined by $(x_{n-1},x_n)\mapsto{u_n}$ exists whenever $G_T$ exists
-\end{Theorem}
-{\bf Proof.}  See \cite[Th. 3c]{manjunath2021universal}.
-
-Projection mappings $\pi_i$ are defined in the traditional meaning where a $k$-dimensional vector is projected to it's $i^{th}$ component such that $$\pi_i:(a_1,a_2, \ldotp, a_{k-1}, a_k)\to{a_i}$$
-
-The graph in equation \ref{fig:comm_H_CET} is then extended as below:
-
-\begin{equation}  \label{Scomm_imp}
-  %    \[ 
-      \psset{arrows=->, arrowinset=0.25, linewidth=0.6pt, nodesep=3pt, labelsep=2pt, rowsep=1.3cm, colsep = 1.1cm, shortput =tablr}
-   \everypsbox{\scriptstyle}
-   \begin{psmatrix}
-   \widehat{U}_T & \widehat{U}_T\\%
-   Y_T & Y_T\\
-   & \textcolor{red}{U \times X}
-   %%%
-  %  \ncline{1,1}{1,2}^{\widehat{T}} \ncline{1,1}{2,1} <{H_2}
-  %  \ncline{1,2}{2,2} > {H_2}
-  %  \ncline{2,1}{2,2}^{G_T}
-  %  \psset{linecolor=red}
-  %  \textcolor{red}{\ncline{2,1}{3,2}<{(\Gamma,\pi_2)}}
-  %  \ncline{3,2}{2,2}>{\textcolor{red}{(\pi_2,g)}}.
-   \end{psmatrix}
-  % \]
-  \end{equation}
-
-The problem finally simplifies to the issue of learning the map $\Gamma$ and combining this with the projection mapping $\pi_2$ and the function $g$, which will be known.
-A final set of equations is obtained - equations that have been entirely constructed from data.
-\begin{eqnarray}\label{eqns_from_data}
-	u_{k+1} &=& \pi_1 \circ (\Gamma, \pi_2) \circ (\pi_2,g) (u_k,x_k) \label{Seqn_u}\\
-	x_{k+1} &=& \pi_2 \circ (\Gamma, \pi_2) \circ (\pi_2,g) (u_k,x_k). \label{Seqn_x}
-\end{eqnarray}
-
-\section{Advantages of learning $\Gamma$} \label{subs_LearnGamma}
-
-One may immediately ask why we opt to take such a roundabout route; why not just learn the map $G_T$ from the get-go?
-On the surface, this seems to be an arbitrary decision path with no real reasoning, so we take a pause again and discuss the motivation for learning Gamma.
-There are a number of distinct advantages.
-
-In the first place, learning $\Gamma$ saves computational resources.
-This is due to the fact that the input $u_n$ lies in a lower-dimensional subspace in comparison to the high-dimensional space $X\times{X}$.
-
-%In practice, if the input is of a  lower dimension, one may easily embed it into the space $X\times{X}$ by padding the vector $u_n$ with zeroes.
-
-Secondly, the function $\Gamma$ is known to be stable.
-Learning $\Gamma$ makes use of equations \ref{eqns_from_data} which in turn employs the driven system $g$ possessing the USP, and offers distinct advantages with regards to stability in the presence of perturbation.
-It is desirable that inputs that differ only slightly would have only ‘slight’ effects on the output of the system.
-Moreover, one would hope to prevent numerical errors originating from input and measurement noise (see~\cite[Th.
-5]{manjunath2021universal}).
-As the data fed into the system is a function of both the input given and the parameters of the system itself (i.e.
-the value a, $\alpha$ in $\tanh$), we exploit both input-related and parameter-related stability.
-Both these notions are defined by way of the continuity of an encoding map of an input and design parameter respectively in \cite{manjunath2020stability}.
-
-Informally, we may consider input-related stability as concerned with the question of whether or not small variations in input would result in small responses.   
-If two inputs $\bar{u}$ and $\bar{u}_{noisy}$ are close in the product topology, then their tails could differ greatly due to the metric that generates the product topology has insensitivity to the differences in sequence's tails.
- However if the system has the USP then the function $h$ is continuous and so $h(\bar{u}), h(\bar{u}_{noisy})$ remain close-by for small levels of noise maintaining a measure of robustness to input and measurement noise.\cite{manjunath2021universal}.
-In a similar fashion, parameter-related stability is obtained thanks to a result \cite[Lemma 3.2]{manjunath2020stability} relating this form of stability with the ESP (which is here equivalent to the USP).
-
-% Secondly,  due to $g$ possessing the USP, both input- and parameter-related stability~\cite{manjunath2020stability} is obtained, which in turn prevents numerical errors that caused by input and measurement noise (see~\cite[Th.
-5]{manjunath2021universal}).
-
-Thirdly, learning a chaotic map on a set(here $Y_T$) with even a small measure of noise  means that the learnt system accepts arguments out of the set (Problems similar to Takens, cf.~\ref{sect_takenslimits}).
-We thus have a map $G_T^+$ acting on $Y_T^+$ and since $Y_T$ is not guaranteed to be an attractor of $G_T$ this could lead to errors.
-If one learns $G_T$ directly, we are in danger of replicating the problems arising for Takens.  To see this, consider Fig.~\ref{fig:YtGtFailure}.
-
-\begin{figure}[ht]
-  \includegraphics[scale=0.25]{_YTerrors.eps}
-  \centering
-  \captionof{figure}{Diagram of potential errors developing when $G_T$ is learnt.}
-  \label{fig:YtGtFailure}
-\end{figure}
-
-To circumvent this: if a driven system with a saturation function\ednote{B: What is a saturation function?} is used, we can ensure the existence of an attractor $A$ of the map which implements $G_T$, and containing $Y_T^+$ indirectly (via $g$).
-\ednote{B: A little vague.
-Are we saying that an attractor of a map implementing $G_T$ is also the one that contains $Y_T^+$?}
-By defining the dynamics to $A$, this prevents large errors from occuring.  
-More details on the global dissipativity of $g$, as it is termed there, may be found in~\cite{Supp}.
-
-\begin{figure}[ht]
-  \includegraphics[scale=0.25]{_autovsdriven.eps}
-  \centering
-  \captionof{figure}{Illustrating empirically the importance of global dissipativity.
-Plotted below are the three principal components of randomly initiated points(blue) of a learnt version of $G_T$ versus the trajectory $(x_{n-1}, x_n)$(red) of driven states obtained from the Fractal Dream Attractor's data~\ref{Thomas_Attractor}).
-}
-  \label{fig:learningFailure}
-\end{figure}
-
-In effect, global dissipativity prevents large numerical errors due to the noise in input data from arising (see Fig.~\ref{fig:learningFailure}). 
-If a system is not globally dissipative, neglibly small errors could lead to major inaccuracy which thus induce predictions to utterly fail.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\chapter{Implementation} \label{ch5}
-
 In this chapter we turn to the implementation of the function $G_T$ by discussing a method to learn the function $\Gamma$.
@@
 \begin{figure}[ht]
-  \includegraphics[scale=0.35]{_dp_setup.eps}
+  \includegraphics[scale=0.35]{Graphs/_dp_setup.eps}
   \centering
@@
 \begin{figure}[ht]
-  \includegraphics[scale=0.4]{_dp_dying.eps}
+  \includegraphics[scale=0.4]{Graphs/_dp_dying.eps}
   \centering
@@
 \begin{figure}[ht]
-  \includegraphics[scale=0.3]{_dpfail_nonsurj.eps}
+  \includegraphics[scale=0.3]{Graphs/_dpfail_nonsurj.eps}
   \centering
@@
   \centering
-  \includegraphics[width=0.95\linewidth]{_dp_success_4coords_traj.eps} 
+  \includegraphics[width=0.95\linewidth]{Graphs/_dp_success_4coords_traj.eps} 
   \captionof{figure}{Predicted(red) and actual(blue) trajectories of the $x$- and $y$-coordinates of each of the pendulum heads.
@@
   \centering
-  \includegraphics[width=\linewidth]{_dp_success_4coords_hist.eps}
+  \includegraphics[width=\linewidth]{Graphs/_dp_success_4coords_hist.eps}
   \captionof{figure}{Densities of predicted and actual values of $x$- and $y$-coordinate of pendulum heads.}
@@
   \centering
-  \includegraphics[width=0.95\linewidth]{_dp_noise_FigCD_2030.eps} 
+  \includegraphics[width=0.95\linewidth]{Graphs/_dp_noise_FigCD_2030.eps} 
   \captionof{figure}{Predicted(red) and actual(blue) trajectories of the $x$- and $y$-coordinates of each of the pendulum heads with noise added from a normal distribution with mean zero and standard deviation 0.1, equating to approximately 39dB's of noise.
@@
   \centering
-  \includegraphics[width=\linewidth]{_dp_noise_2B_2030.eps}
+  \includegraphics[width=\linewidth]{Graphs/_dp_noise_2B_2030.eps}
   \captionof{figure}{Densities of predicted and actual values of x-coordinate of first pendulum head.}
@@
     \centering
-    \includegraphics[width=0.8\linewidth]{_thomas_noisy_0.1056.eps}
+    \includegraphics[width=0.8\linewidth]{Graphs/_thomas_noisy_0.1056.eps}
   \endminipage\hfill
   \minipage{0.5\linewidth}
-    \centering    \includegraphics[width=0.8\linewidth]{_thomas_clean_0.1056.eps}
+    \centering    \includegraphics[width=0.8\linewidth]{Graphs/_thomas_clean_0.1056.eps}
   \endminipage
@@
   \centering
-  \includegraphics[scale=0.35]{_Thomas_1.eps}\caption*{Predicted trajectories of the $x$- and $y$-coordinates for the Thomas attractor with parameter-value $\beta=0.1056$ demonstrate empirically the ability to predict the evolution of the trajectory for the next an estimated 100 timesteps into the future near-exactly.
+  \includegraphics[scale=0.35]{Graphs/_Thomas_1.eps}\caption*{Predicted trajectories of the $x$- and $y$-coordinates for the Thomas attractor with parameter-value $\beta=0.1056$ demonstrate empirically the ability to predict the evolution of the trajectory for the next an estimated 100 timesteps into the future near-exactly.
 Here no noise was added.}
-  \includegraphics[scale=0.5]{_Thomas_3.eps}\caption*{Represented here the learnt (blue) and actual (red) densities of the first coordinate ($x$-coordinate) of the dataset for the Cyclically Symmetric Attractor with parameter-value $\beta=0.1056$.}
+  \includegraphics[scale=0.5]{Graphs/_Thomas_3.eps}\caption*{Represented here the learnt (blue) and actual (red) densities of the first coordinate ($x$-coordinate) of the dataset for the Cyclically Symmetric Attractor with parameter-value $\beta=0.1056$.}
 \end{figure}
@@
   \centering
-  \includegraphics[width=1.0\textwidth,left]{_Clifford_1_nonoise.eps}
+  \includegraphics[width=1.0\textwidth,left]{Graphs/_Clifford_1_nonoise.eps}
   \caption*{These graphs were constructed by predicting the Clifford system 1000 steps into the future and in so doing illustrating the long-term consistency and accuracy of the learnt system.
@@
       \centering
-      \includegraphics[width=\linewidth]{_Clifford_2_nonoise.eps}
+      \includegraphics[width=\linewidth]{Graphs/_Clifford_2_nonoise.eps}
   \endminipage\hfill
@@
     \centering
-    \includegraphics[width=\linewidth]{_Clifford_3_nonoise.eps}
+    \includegraphics[width=\linewidth]{Graphs/_Clifford_3_nonoise.eps}
   \endminipage
@@
 \end{center}
-
-\chapter{Conclusion}\label{ch6}
-\input{chapter6.tex}
-
-\bibliographystyle{pnas-new}
-\bibliography{pnas}
-
-\end{document}

### 2022-01-17 projectReport/chapter6.tex
Cleaner setup

--- 2022-01-17 draft_16Jan/draft7.tex
@@
-\documentclass[a4paper,12pt,twoside]{report}
-% \documentclass[a4paper,12pt,twoside]{book}
-\pagestyle{headings}
-\usepackage[utf8]{inputenc}
-\usepackage{filecontents}
-\usepackage[many]{tcolorbox}
-\usepackage{graphicx}
-\usepackage{harpoon}
-%\usepackage{apacite} 
-\usepackage{xcolor}
-\usepackage{latexsym}
-\usepackage{amssymb}
-\usepackage{amsmath}
-\usepackage{mdframed}
-%\usepackage[authoryear]{natbib}
-\usepackage{hyperref, url}
-\usepackage[show]{ed}
-\usepackage{subfigure}
-\usepackage[ntheorem]{empheq} 
-\usepackage{amsthm, amssymb, amsfonts, latexsym}
-% \usepackage{pst-all}
-
-\usepackage{enumitem}
-\usepackage{multirow}
-\usepackage{longtable}
-\usepackage[export]{adjustbox}
-% $f\arrowvert_A$ To restrict functions
-
-\setlength{\parskip}{1.7em}
-\setlength{\parindent}{0em}
-
-%\setlength{\baselineskip}{20pt}
-
-\setlength{\textwidth}{429.75499pt}
-\setlength{\textheight}{643.20255pt}
-\setlength{\oddsidemargin}{5 mm}
-\setlength{\evensidemargin}{5 mm}
-\setlength{\topmargin}{0 mm}
-\setlength{\headsep}{0 mm}
-\setlength{\headheight}{0 mm}
-
-\usepackage{pst-node}
-
-\newcommand{\citealtt}[1]{\citeauthor{#1},\citeyear{#1}}
-\newcommand{\myycite}[1]{\citep{#1}}
-
-\mathchardef\mhyphen="2D
-
-%\newcommand{\cev}[1]{\reflectbox{\ensuremath{\vec{\reflectbox{\ensuremath{#1}}}}}}
-
-\usepackage{pst-node,graphicx,pst-blur}
-%\uspackage{auto-pst-pdf}
-%\usepackage{tikz-cd}
-
-\makeatletter
-\DeclareRobustCommand{\cev}[1]{%
-  \mathpalette\do@cev{#1}%
-}
-\newcommand{\do@cev}[2]{%
-  \fix@cev{#1}{+}%
-  \reflectbox{$\m@th#1\vec{\reflectbox{$\fix@cev{#1}{-}\m@th#1#2\fix@cev{#1}{+}$}}$}%
-  \fix@cev{#1}{-}%
-}
-\newcommand{\fix@cev}[2]{%
-  \ifx#1\displaystyle
-    \mkern#20mu
-  \else
-    \ifx#1\textstyle
-      \mkern#20mu
-    \else
-      \ifx#1\scriptstyle
-        \mkern#26mu
-      \else
-        \mkern#26mu
-      \fi
-    \fi
-  \fi
-}
-
-\makeatother
-
-\newcommand{\xcm}{\epsfxsize=3.1cm}
-\newcommand{\fig}[1]{\epsfbox}
-% \newcommand{\bp}{\begin{minipage}{3.1cm}}
-% \newcommand{\ep}{\end{minipage}}
-
-\newtheorem{Definition}{Definition}[]
-%\newmdtheoremenv{Theorem}{Theorem}[]
-\newtheorem{Theorem}{Theorem}[]
-\newtheorem{Lemma}{Lemma}[]
-\newtheorem{Proposition}{Proposition}[]
-\newtheorem{Corollary}{Corollary}[]
-\newtheorem{Remark}{Remark}[]
-\newtheorem{Example}{Example}[]
-
-%Shortcut symbols
-\newcommand{\Ftheta}{\ensuremath{F_\theta}}
-
-\makeatletter 
-\renewcommand{\thefigure}{\@arabic\c@figure}
-\makeatother
-\usepackage{xr}
-
-%%%% PNAS GRAPHICS SETUP- FROM THE PNAS STYLE LATEX FILE
-\RequirePackage{graphicx,xcolor}
-\RequirePackage{colortbl}
-\RequirePackage{booktabs}
-\RequirePackage{algorithm}
-\RequirePackage[noend]{algpseudocode}
-\RequirePackage{changepage}
-\RequirePackage[twoside,%
-				letterpaper,includeheadfoot,%
-				layoutsize={8.125in,10.875in},%
-                layouthoffset=0.1875in,%
-                layoutvoffset=0.0625in,%
-                left=38.5pt,%
-                right=43pt,%
-                top=43pt,% 10pt provided by headsep
-                bottom=32pt,%
-                headheight=0pt,% No Header
-                headsep=30pt,%
-                footskip=25pt,
-                marginparwidth=38pt]{geometry}
-\RequirePackage[labelfont={bf,sf},%
-                labelsep=period,%
-                figurename=Fig. ]{caption}
-\setlength{\columnsep}{13.5pt} % Distance between the two columns of text
-\setlength{\parindent}{12pt} % Paragraph indent
-
-%% Figure caption style
-\DeclareCaptionFormat{pnasformat}{\normalfont\sffamily\fontsize{7}{9}\selectfont#1#2#3}
-\captionsetup*{format=pnasformat}
-
-%%% GREYBOX AROUND FIG
-\definecolor{lightgray}{gray}{0.95}
-\newcommand\greybox[1]{%
-  \vskip\baselineskip%
-  \par\noindent\colorbox{lightgray}{%
-    \begin{minipage}{\textwidth}#1\end{minipage}%
-  }%
-  \vskip\baselineskip%
-}
-
-\newtcolorbox{paperbox}[1][]{
-    enhanced,
-    colback=white,
-    boxrule=0pt,
-    boxsep=0pt,
-    arc=0mm,
-    width=0.8\linewidth,
-    fuzzy shadow={0mm}{-4pt}{-4pt}{1mm}{black!30!white},
-    #1
-}
-
-% \externaldocument{Supp_Koopman-0808}
-\begin{document}
-
-\title{Data-Driven Modelling \& \\ Forecasting of Chaotic Systems}
-
-\author{B.L. Nortier} 
-\date{January, 2022}
-\date{}
-\maketitle
-
-\begin{abstract}
-  It is a well-established practice to collect measurements from underlying artificial, natural, and physical systems through observables in order to realise a dynamical model with greater descriptive power.
-  Many data-driven approaches to forecast the future evolution of systems exist.
-However, many of these methods rely on an initial understanding of the system and we often do not fully understand the many complexities of more intricate systems that we interact with on a daily basis.
-
-  In this project, the base assumption is that data is originates from  some discrete-time dynamical system, particularly one of a chaotic nature but no assumptions are made pertaining to a specific dynamical system.
-  A general learning problem is formulated and a well-known approach to the problem, alongside its practical limitations, is then considered.
-  The notion of driven dynamical systems is introduced, and we subsequently establish the theoretical existence of a dynamical system possessing certain properties which guarantees it to be topologically conjugate (`dynamically equivalent`) to the dynamics of the underlying system from which measurements were originally taken.
-  Our methodology is implemented practically by making use of recurrent neural networks, a form of deep learning, to learn a map topologically equivalent to the underlying system that forecasts future states.
-  In so doing we present the workability of an approach to obtain long-term topological and statistically consistent predictions for simple physical systems as well as a some chaotic attractors.
-  \end{abstract}
-
-  \tableofcontents
-
-\chapter{Introduction}\label{ch1}
-
-Experimenting with biological, physical, and artificial systems to generate a more informative dynamical model is a well-established practice in modern science.
-Traditional methods for modelling physical systems are based on laws of physics that are based either on empirical relationships or intuition.
-For systems that evolve with time, physical laws yield mathematical equations that govern how the quantities evolve with time.
-However, our world is much more complex than that which can be distilled into elegant equations.
-We do not fully understand many of the more complex systems, nor do they provide us with a good physical intuition of the underlying principles governing system dynamics.
-To complicate this, the underlying systems we observe often display sensitive dependence on initial conditions despite having highly similar initial conditions.
-This is because differing orbits diverge quickly and to such an extent that it becomes seemingly impossible to retrace their steps back to the original conditions.
-Even the presence of usually' negligible' computational noise or measurement error renders long-term pointwise prediction infeasible.
-There are many difficulties that one encounters while modelling such systems:
-\vspace{-8mm}
-\begin{enumerate}[noitemsep, label=\roman*.]
-  \item One may not have access to the complete states of the systems
-  \item The system may be described by functions that behave wildly because their graphs have a wild oscillatory behaviour, i.e., a large functional complexity.
-\end{enumerate}
-
-Models derived primarily from data can be classified into three categories: 
-\vspace{-8mm}
-\begin{enumerate}[noitemsep, label=\roman*.]
-  \item  Interpretable models (i.e., they establish relationships between internal physical mechanisms), 
-  \item Partially interpretable models capturing some modes of the dynamics, 
-  \item Non-Interpretable models, defined as such mainly due to the fact that they are defined on a different phase space that is usually high-dimensional. 
-\end{enumerate}
-
-Examples in the literature attempting to forecast data from such systems have tried many different approaches with differing degrees of success. 
-When we have complete access to states of the system, an ordinary differential equation can be obtained from data (\cite{brunton2016discovering, champion2019data} and \cite{small2002modeling,xu2006modeling}) wherein one could approximate the vector field by a library of functions to obtain interpretable models.
-
-Recent partially interpretable models available in the literature have been based on the Koopman operator (e.g.,[?, ?]) to employ observables mapping the data onto a higher-dimensional space.
-This makes the dynamics in the higher-dimensional space more amenable for approximation by a linear transformation.
-Such methods do not guarantee exact reconstructions for nonlinear models, and in practice, provide poor long-term consistency for a large class of systems.
-(Citation)
-\textbf{(Citation)}
-
-The non-interpretable models include the delay embedding and the machine learning algorithms.
-For example, one could learn a system conjugate to the underlying system by applying the Takens delay embedding~\cite{takens1981detecting} when one has "good" observations from the system; this learnt system could then be used to forecast the observed data.
-Takens delay embedding theorem ~\cite{takens1981detecting} and its various generalisations (e.g., \cite{sauer1991embedology, stark1999delay, gutman2018embedding}) establishes the learnability of a system constructed by concatenating sufficiently large previous time-series observations of a dynamical system into a vector (called delay coordinates).
-This then confirms the existence of a map on the space of delay coordinates equivalent (or topologically conjugate) to the underlying map from which the observed time-series was first obtained.
-Although topological conjugacy guarantees an alternate representation of the underlying system, the quality still depends on numerous parameters, making the comprehension of the dynamics unreliable (e.g., [?]).
-One reason for this fragility is that the embedded attractor in the reconstruction space is not always an attractor of the map learnt in the reconstruction space, despite unquestionably being an invariant set.
-When the embedded object is not explicitly known to be an attractor (as explained in Chapter~ \ref{ch3}), it can cause predictions to fail.
-
-Practically, the application of Takens embedding involves learning a map through some technique, and consequently one wishes that these would have low functional complexity\cite{manjunath2021universal}, i.e., functions with fewer oscillatory graphs.
-On the other hand, pure machine learning methodology processes temporal information (like the echo state networks \cite{jaeger2001echo, jaeger2004harnessing} by mapping data onto a higher dimensional space for further processing.
-Although they perform  well on forecasting some dynamical systems, they fail completely on others (Citations here) as there is often no guarantee that the right function was learnt during training.
-
-This thesis involves implementing and analysing non-interpretable models that can guarantee exact reconstruction. 
-The project work concerns the study and implementation of a method (\cite{manjunath2021universal}) that incorporates learning a function by mapping the data on to a higher dimensional space using what is called a driven dynamical system (See Chapter~ref{ch.4}).
-With a clear understanding of how the data is mapped onto the higher dimensional space, the method then permits the learning of a dynamical system topologically conjugate to that of the underlying system.
-Instead of linear regression as in the training of echo state networks \cite{jaeger2001echo}, deep learning methods are employed to learn the correct function.
-With slight modifications to the implementation in the paper (with Adriaan), we show that one can construct accurate non-interpretable models with the ability to reconstruct attractors from more hard-to-forecast systems like the double pendulum.
-(The forecasting of the time-series from a double pendulum has not been reported before.) Moreover, we also demonstrate that long-term statistical consistency is preserved.
-
-\emph{By solidifying the mathematical underpinnings of our theory, we hope to guarantee the ability to construct models with predictive power ranging from molecular biology to neuroscience.
-in the near future.
-(We can modify this sentence when the report is completed.)}
-
-The more intricate details of proofs are referred to where relevant throughout. This report is organised as follows: 
-\newline In Chapter~\ref{ch2}, we recall the definition of a discrete-time dynamical system, how a discrete-system arises from a flow of an ODE and then proceed to define the inverse-limit space and topological conjugacy of autonomous systems.
-\newline Chapter~\ref{ch3} introduces the problem of forecasting dynamical systems, states the Takens delay embedding theorem, and discusses various issues faced while forecasting. 
-\newline In Chapter~\ref{ch4}, a driven dynamical system is defined and discuss the properties of a specific class of driven dynamical systems that we make use of in this thesis/report/paper. 
-\newline Finally, in Chapter~\ref{ch5} we show the implementation of these forecasting methods, and conclusions are provided in Chapter~\ref{ch6}.
-
-\chapter{Discrete-time Dynamical Systems}\label{ch2}
-
-This chapter briefly describes a discrete-time dynamical system and what it means to exhibit chaos. We refer to \cite{devaney2018introduction, de2013elements} for more details.
-
-At its most elementary level, a dynamical system is just something that evolves deterministically through time.
-In the context of this thesis, deterministic refers to the fact that a system evolves according to specified rules rather than based on random events.
-Dynamical systems arise in a variety of situations.
-A continuous-time dynamical system describes the states for all values of the time.
-Specifically, if the motion of a pendulum in which the quantities such as the angular position and angular momentum are known at all times, then it is a continuous-time dynamical system.
-The equations of the dynamical system can take the form of one or more ordinary differential equations that determine the relevant quantities at any future time if we know the initial location and momentum.
-In ecology, discrete-time dynamical systems are widely used to model population growth.
-The model in this case is a function calculating the following generation's population given the population of the previous generation.
-If we know the starting population, we may once again calculate the population at any time in the future.
-
-Formally, a function $T: U \to U$, where $U$ is some set is  a \emph{discrete-time dynamical system} and its iterates $\{u,Tu,T^2u,\ldots\}$, where $T^n$ denotes the $n$-fold composition of $T$ with itself, describe the evolution of an initial condition $u\in U$ (Note that we frequently drop the brackets and denote $T(u)$ by $Tu$ so as to simplify notations).
-Also, we assume that $U$ is endowed with a metric $d$.
-
-Continuous-time dynamical systems modelled using ordinary differential equations can give rise to discrete-time dynamical systems.
-To see this, consider a differential equation $\dot{x} = f(x)$, $f: \mathbb{R}^n \to \mathbb{R}^n$, $n\in\mathbb{N}$ given to have a unique solution passing through
-each point $x\in\mathbb{R}^{n}$, call it $x(t)$ where $x(0)=x_0$.
-
-\begin{Definition}
-  [\bf Flow of an Equation] \label{Dfn_Flow}\rm
-  The flow of the equation  $\dot{x} = f(x)$  is defined to be a mapping $\varphi: \mathbb{R}^n \times \mathbb{R} \to \mathbb{R}$ where $\varphi(x_0,t)= x(t)$ for the solution $x(t)$ with $x(0)=x_0$ 
-\end{Definition}
-
-By fixing $t=K \in (0,\infty)$, we can define the \emph{time-$K$ map} as  $T(x):= \varphi(x_0,K)$, and it is easily verified that $T\circ T(x_0) = \varphi(x_0,2K)$.
-In general the $m^{\mbox{th}}$ iterate of $x_0$ under $T$ would be the value of the solution of the ODE evaluated at time $mK$ with the initial condition $x_0$, i.e.
-$\varphi(x_0, mK)$.
-Thus ordinary differential equations give rise to a discrete-time dynamical system by sampling the value of the solution $x(t)$ at time intervals $K$ units apart.
-
-A numerical discretization of a differential equation can also give rise to a discrete-time dynamical system.
-For instance, Euler's method approximates $\dot{x}(t)$ by $(x(t+h)-x(t))/h$; if $h$ is fixed throughout, the solution of a differential equation $\dot{x}=f(x)$
-can be approximated at the time instant $t+(m+1)h$ by iterating the equation 
-\begin{equation}
-  x(t+(m+1)h) = x(t+mh) + h f(x(t+mh))
-\end{equation}
-
-Adopting more succinct notation by replacing $x(t+mh)$ with $u_m$, we rewrite the above equation as
-\begin{equation}
-u_{m+1} = u_m + hf(u_m)
-\end{equation}
-or in even simpler terms as the discrete-time dynamical system with map $T(u) = u + hf(u)$, where $T$, $u$ and $f$ are understood to be as above.
-
-Of course, discrete-time dynamical systems need not always arise through a differential equation.
-Once again, we may consider the field of ecology, where discrete dynamical systems are often directly derived or assumed.
-\textbf{Citation}
-There is a school of thought that advocates discrete-time dynamical systems to be more natural for modelling real-world observations than differential equations.
-We refer the interested reader to \cite{saber2010introduction}.
-
-\section{Invariant Sets}
-
-A core concept in the study of dynamical systems is that of invariance. 
-\begin{Definition}
-  [\bf Invariant Set]\label{Dfn_InvariantSet}\rm
-  Given a discrete-time dynamical system $T: U \to U$, a subset $A \subset U$ is said to be an \emph{invariant set} if $T(A) =A$. 
-\end{Definition}
-
-We also define the orbit of a function $T$.
-\begin{Definition}
-  [\bf Orbit of $T$]\label{Dfn_Orbit}\rm
-  The orbit of $T$ is to be defined the sequence $\bar{u} = \{u_n\}_{n\in \mathbb{Z}}$ obeying the update equation, $u_{n+1}=Tu_n$, $n \in \mathbb{Z}$. 
-\end{Definition}
-
-Two examples of invariant sets include a fixed point where $Tu=u$ for  $u\in U$, and a periodic orbit, i.e., a set of iterates $\{u,Tu, T^2u,\ldots,T^pu\}$, where $T^{p+1}u=u$ for some $p\in\mathbb{Z}$.
-The entire space $U$ could be also be invariant.
-Consider for example the space $U=[0,1]$, where $Tu=4u(1-u)$ and then $U$ is invariant (as every $u\in{U}$ can be written as $u = 4x(1-x)$ for some $x\in{U}$).
-
-We may learn a great deal about the iterates of a dynamical system by considering the types of invariant sets of a discrete-time dynamical system.
-For example, if $U=[0,1]$ has map $Tu= u/2$, then the only invariant set is $\{0\}$, and all orbits approach this invariant set as time flows in the forward direction.
-Indeed, if a some non-zero invariant set (call it $B$) exists, then there is  some $r\in(0,1]\cap{B}$.
-But $r\notin{T(B)}$ since any orbit with initial value $r$ will be a decreasing sequence.
-Moreover, every orbit of T will be decreasing and therefore approach the value $0$ as $n\rightarrow\infty$.
-
-One may ask if every orbit approaches an invariant set?
-In general the answer is no, since for the dynamical system $T: \mathbb{R} \to \mathbb{R}$ defined by $Tu=2u$, any orbit that does not intersect the invariant set $\{0\}$, will not approach any invariant set.
-
-However,  when the space $U$ is compact, all orbits approach an invariant set.
-This is since the set of limit points of the orbit can be shown to be invariant \cite{de2013elements}.
-So, when $U$ is compact,  the $\omega$-limit set $\omega(u;T)$ of a point $u$ defined to the collection of limit points of the sequence $\{x,Tu,T^2u,\ldots\}$ is nonempty, and $\omega(u;T)$ is invariant.
-
-\begin{Example}
-  For the map $Tu=u^2$ defined on $[0,1]$ all orbits lie in the invariant set $\{1\}$ or else would would approach the invariant set $\{0\}$. 
-\end{Example}
-
-Invariant sets have various properties. Invariant sets can be attracting or repelling depending on how orbits in their vicinity behave. 
-Recall the examples $Tu =u/2$ and $Tu=2u$ defined on $\mathbb{R}$ (where $\{0\}$ “attracts" orbits) or the example, $Tu=u^2$ defined on $[0,1]$ (where $\{1\}$ “repels" orbits). 
-We are interested in attractive invariant sets since they capture the long-term dynamics as time increases. 
-In particular, we are interested in those invariant sets named attractors.
-
-\begin{Definition}
-  [\bf Attractor]\label{Dfn_Attractor}\rm
-  Let $T: U \to U$, where $U$  is a metric space with metric $d$. A compact subset $A \subset U$ is said to be an attractor if it satisfies the three conditions: 
-  \vspace{-8mm}
-  \begin{enumerate}
-	\item $A$ is invariant. 
-	\item $A$ is asymptotically stable, i.e., for every $\epsilon > 0$ and for all $u$ so that $d(u,A) < \epsilon$, we have $d(T^nu,A) \to 0$ as $n\to \infty$. 
-	\item $A$ has Lyapunov stability, i.e., for every $\epsilon > 0$  there exists a $\delta(\epsilon) > 0$ so that $d(u,A) < \delta$ implies $d(T^nu,A) < \epsilon$ for all $n\ge 0$.  
-\end{enumerate}
-\end{Definition}
-
-Indeed in our previous examples, it can be verified that the system  $Tu=u/2$ defined on  $\mathbb{R}$ and $Tu=u^2$ defined on $[0,1]$  the singleton set $\{0\}$ is an attractor.
-For the system,  $Tu=1-|2u-1|$ on $[0,1]$, one may easily verify that the only attractor is the entire space $[0,1]$.
-This follows from the fact that between any two points $u< v$ in $[0,1]$, and for any $a,b$ so that $u\le a < b \le v$, we can find an $n$ so that $T^n(a,b)=[0,1]$ (as would be explained later in this chapter).
-
-A dynamical system can have several attractors and may also be contained in another attractor.
-For the example, $Tu =u^2$ on $[0,1]$, both $\{0\}$ and $[0,1]$ are attractors.
-It is known that an attractor for a dynamical system always exists in a compact space.
-\textbf{(Cite).}
-
-The dynamics restricted to an invariant set can be complicated.
-For instance an invariant set could be just a single point or it could have an infinite set.
-If the invariant set is infinite, then complicated dynamics are possible.A particular well-studied phenomenon of such complexity gives rise to so-called  chaotic behaviour, studied in detail over the past fifty years.
-
-% If the dynamics on the attractor are somewhat complicated in the sense that the attractor cannot be decomposed further, and if the attractor is infinite then there is a possibility of complicated dynamics, and a particular well-studied phenomenon of such complexity gives rise to what is called a chaotic behaviour.
-
-%If the dynamics on an attractor cannot be decomposed further and the attractor is itself also infinite, then there is a possibility of complicated dynamics, and a particular well-studied phenomenon of such complexity gives rise to what is called chaotic behaviour.
-
-\section{Chaos}
-
-In the 1960s, several mathematicians and mathematically interested scientists independently discovered chaos in the mathematical sense.
-The meteorologist Edward Lorenz may have been the first to explain this phenomenon in his 1963 paper \cite{lorenz1963deterministic}.
-The notions of invariance, attractivity, and chaos may also be described for continuous systems, and Lorenz's system comprised of a system of differential equations.
-The narrative of Lorenz's discovery of chaos and the history of other forerunners in this subject is fascinating.
-We highly recommend James Gleick's book Chaos: The Making of a New Science~\cite{gleick2008chaos} for those interested.
-It clearly illustrates these experiences, explains why chaos was such a startling and crucial mathematical and scientific discovery and describes the underlying mathematical notions for non-specialists.
-
-There are several intricately different definitions of chaos available in the literature \textbf{(Cite)}, with each of them indicating some aspect of complexity.
-In practice, it is not possible to verify which definition or notion does the underlying dynamical system satisfies when only data is observed from it.
-We do, however, specifically recall the definition of chaos in the sense of Devaney \cite{devaney2018introduction,de2013elements} so as to understand some nuances behind the complexity.
-Devaney's definition of chaos has three requirements, with the first being the notion  of sensitive dependence on initial conditions.
-
-\begin{Definition}\rm
-  [\bf {Sensitive Dependence on Initial Conditions}]\label{Dfn_SDIC}\rm
-A dynamical system $T: U \to U$ is said to have sensitive dependence on initial conditions (SDIC) if there exists a $\delta > 0$ such that for every $u \in U$ and in every neighborhood of $u \in U$ there exists a $v\in{U}$ and an integer $N:=N{(u,v)}\in\mathbb{Z}$ such that $d(T^Nu,T^Nv)>\delta$.
-\end{Definition}
-
-It is vital to acquire a sense of this definition as it can easily be misunderstood.
-It is a common misconception to interpret SDIC as two close points ($u$ and $v$) that eventually become separated by a distance $\delta$ under iteration by $T$.
-But this is not true.
-To fully comprehend all the subtleties of the concept, one needs to discuss it more thoroughly by considering each sentence with care and attention, whereafter one may examine how they fit together to convey the idea of sensitive dependence.
-
-To this end, we make 3 remarks:
-\vspace{-5mm}
-\begin{enumerate}
-  \item First, the $\delta>0$ in the definition of SDIC is independent of $u$. 
-  \item Second, in every neighborhood of $u$, we may not necessarily find all points $v$ in the neighborhood distinct from $u$ that would separate from the forward iterates of $u$. 
-  \item  Finally, $N$ depends upon $u$ and $v$ chosen, and their iterates may not separate forever (i.e., for all $n>N$) and we allow their iterates to get arbitrarily close in the future. 
-\end{enumerate}
-
-To illustrate the concept of SDIC, we present two examples - one from Mathematics and the other from the field of Physics.
-
-\begin{Example} \rm
-  The logistic map (LM), a recursion relation of the form $x_{n+1}=rx_n(1-x_n)$ where $x_n\in[0,1]$ and $0<r\leq{4}$.
-It is of particular interest to us as it is classically used to illustrate chaos \cite{devaney2018introduction} for certain parameters.
-  In the graph below~\ref{fig:log_sdic} we plot the first 50 iterates of the LM for $r=4$ and two slightly different initial values $x_0$.
-
-  \begin{figure}[ht]
-    \includegraphics[scale=0.74]{_logistic_sdic.eps}
-        \centering
-        \captionof{figure}{Figure generated for Logistic Map $x_{n+1}=rx_n(1-x_n)$ with $r=4$ to exhibit the presence of SDIC.
-Plotted are the values $x_n$ against time $n$ for the the first 50 step with initial values $x_0=0.2$ (blue) and $x_0=0.20001$ (red).
-Initially the two trajectories overlap, but they diverge completely at $n=12$ whereafter they follow distinct paths.}
-        \label{fig:log_sdic}
-      \end{figure}
-
-  % For an intuitive sense of the fact that the LM exhibits SDIC, consider again the definition and substitute specific values.
-Let $\delta=0.05$ and for purposes of illustration consider the value $u=0.4$.
-The green band in~\ref{fig:log_sdic} shades the region $[u-\delta, u+\delta]$ and we can see that regardless the value $v$  and in every neighborhood of $u \in U$ there exists a $v\in{U}$ and an integer $N:=N{(u,v)}\in\mathbb{Z}$ such that $d(T^Nu,T^Nv)>\delta$
-  % specific point, say $u=0.4$.
-Draw a green neighbourhood $[0.35,0.45]$ of diameters $\delta=0.02$ around the point.
-It is easy to see that every neighbourhood around $u=0.4$ will contain some $v_1, v_2\in[0,1]$ and natural number $N_1, N_2 \in\mathbb{N}$ respectively s.t.
-the distance $d(T^Nu_1,T^Nv)>\delta=0.05$ .
-\end{Example}
-
-\begin{Example} \rm
-  A second example pertains to a simple physical system called the double pendulum - a pendulum with another pendulum attached to its head given an initial position and angular velocity.
-
-  Below are two figures denoting trajectories of a double pendulum's second head after some elapsed time.
-Depicted are three pendulum heads with equal angular velocity but differing slightly in initial position.
-As can be seen below, the trajectories diverge very quickly to become completely different.
-We discuss the double pendulum in greater detail in a later chapter (\ref{ch5}), but we mention the example here to provide a numerical example of a practical system exhibiting SDIC.
-
-\begin{figure}[ht]
-  \centering
-\includegraphics[scale=0.8]{_dp_sdic.eps}
-    \captionof{figure}{Three double pendulum heads with equal angular velocity and initial angles differing by 0,025 radians}\label{fig:dp_sdic}
-  \end{figure}
-
-\end{Example}
-
-The second part in Devaney's definition of chaos concerns topological transitivity.
-
-\begin{Definition}
-  [\bf {Topologically Transitive}]\label{Dfn_TopolTrans}\rm
-	A dynamical system $F: U \to U$ is topologically transitive if for any pair of nonempty open sets $E_1$ and $E_2$ there exists a $n\in\mathbb{N}$ such that $T^n(E_1) \cap E_2 \not= \emptyset$. 
-\end{Definition}
-
-Topological transitivity implies that iterates of an open set of initial conditions get mixed up with other open sets.
-On a compact metric space, one may show that topological transitivity also implies the existence of a point whose forward iterates are dense \cite{de2013elements}; or in other words, the orbit going through this point will be dense in the compact metric space.
-In fact, it is topologically more likely that the choice of an arbitrary point will be one whose iterates are almost dense.
-
-\begin{Definition}
-  [\bf {Meager Set}]\label{Dfn_Meager Set}\rm
-A subset of a topological space $U$ is said to be a meager set if it can be written as a countable union of sets of with empty interior. The complement of the a meager set is set to be a residual set.
-\end{Definition}
-
-To be more precise (see \cite{de2013elements}), for a discrete-time dynamical system on a compact space, the set of points with dense iterates are residual, and they are typical or likely to be observed in practice.
-In this project, when data is observed from a topologically transitive system, we assume it arises from a dense orbit.
-
-We may now define the notion of Chaos as formulated by Devaney\cite{devaney2018introduction}.
-\begin{Definition}
-  [\bf {Devaney's Chaos}]\label{Dfn_ChaosDec}\rm
-	A dynamical system $T: U \to U$ is said to exhibit chaos in the sense of Devaney if it satisfies the three properties:
-	\vspace{-5mm}
-  \begin{enumerate}
-		\item $T$ has SDIC.
-		\item $T$ is topologically transitive.
-		\item The set of periodic points of $T$ are dense in $U$. 
-	\end{enumerate}
-\end{Definition}
-
-\begin{Example} \rm
-  The standard tent map $Tu=1-|2u-1|$ defined on $[0,1]$ is a well-known example of a dynamical system satisfying these three properties.
-
-  We now reason as to why this is true.
-The graph of the map $T$ is piecewise linear with two straight lines, one connecting the points $(0,0)$ and $(\frac{1}{2},1)$ and the other connecting $(\frac{1}{2},1)$ with $(1,0)$.
-They form a so-called tent with base centered at $u=1/2$.
-The graph of the map $T^2$ comprises two symmetric tents with their base centered at $1/4$ and $3/4$.
-See Figure \ref{fig:T2tentmap}
-
-  \begin{figure}[ht]
-    \includegraphics[scale=0.6]{_tentmap_2.eps}
-        \centering
-        \captionof{figure}{Graph of the $T^2$ map with bases at $x=0$, $x=\frac{1}{2}$ and $x=1$ respectively.}\label{fig:T2tentmap}
-  \end{figure}
-
-  In general, the graph of  $T^k$ contains $2^{k-1}$ tents.
-  Given any point $u$ and an open neighborhood $W \subset[0,1]$ of $u$, we can find $k\in\mathbb{Z}$ large enough so that $T^k(W) = [0,1]$ because we can accommodate a tent whose base is contained in $W$.
-% \ednote{B: This last phrase doesn't make sense to me.
-M: The number of tents grow with $k$ and hence this is obvious; I explained in one of our meetings.
-Does it make sense?}
-  This implies the existence of two points $w_1$ and $w_2$ in $W$ so that $d(T^kw_1,T^kw_2)= 1$, and by the triangle inequality  at least one of the inequalities $d(T^ku,T^kw_1)> \delta$ or $d(T^ku,T^kw_2)> \delta$ hold when $\delta\in (0,\frac{1}{2})$.
-So $T$ has sensitive dependence on initial conditions.
-
-  Next, let $W_1$ and $W_2$ be two nonempty open sets.
-Given any open set $W_1$, we can find a $k\in\mathbb{N}$ large enough so that $T^k(W_1) = [0,1]$ because we can accommodate a tent with base contained in $W$.
-So $T^k(W_1) \cap W_2 \not=\emptyset$, and thus $T$ has topological transitivity.
-
-  Finally, for every open interval $(a,b)$, the graph of $T^k$ intersects the graph of the identity map on $[0,1]$.
-This follows from the fact established above that there exists a $k\in\mathbb{N}$ such that $T^k(a,b) = [0,1]$.If the intersection point has coordinates of the form $(p,p)$, the point $p$ is then a fixed point of $T^k$ and therefore also a periodic point of $T$.
-Since we have found a periodic point in the interval $W$,  the set of periodic points are dense in $T$.
-  This follows because any fixed point of $T^k$ is also a periodic point, and we have found this in an arbitrary interval.
-
-\end{Example}
-
-\section{Conjugacy}
-
-We now turn our attention to the subject of conjugacy that describes when two dynamical systems are equivalent dynamically.
-
-To show topological similarity (or sameness) between two metric or topological spaces, one needs to establish a homeomorphism between the two spaces. 
-However, in the study of dynamical systems defined on two spaces, establishing a homeomorphism does not indicate that the systems are dynamically related in any way.
-For instance, the maps $Tu=u^2$ and $Tu=1-|2u-1|$ defined on $[0,1]$ have totally different behaviour. 
-Therefore, to find dynamically similar systems, one must establish a dynamical equivalence that we illustrate in a simple commutativity diagram below~\ref{eqn_conjugacy}.
-
-% Giving errors
-\begin{equation}  \label{eqn_conjugacy}
-%    \[ 
-    \psset{arrows=->, arrowinset=0.25, linewidth=0.6pt, nodesep=3pt, labelsep=2pt, rowsep=0.7cm, colsep = 1.1cm, shortput =tablr}
- \everypsbox{\scriptstyle}
- \begin{psmatrix}
-U & U\\%
-V & V.
- %%%
-%  \ncline{1,1}{1,2}^{T} 
-%  \ncline{1,1}{2,1} <{\phi}
-%   \ncline{2,1}{2,2}^{S}
-%  \ncline{1,2}{2,2} > {\phi}
- \end{psmatrix}
-% \]
-\end{equation}
-
-%Composite functions
-To understand the diagram, we note that if we travel "right and then down," the diagram instructs us to use $T$ (top arrow right) first, followed by $\phi$.
-(right arrow downwards).
-Consequently, the pathway amounts to finding $\phi(T(u))$.
-If we proceed "down and then right," the diagram instructs us to apply $\phi$ (left arrow downwards) first, and then apply $S$ (bottom arrow right).
-This then amounts to finding  $S(\phi(u))$.
-When the relationships in this diagram hold, we say $\phi(T(u))= S(\phi(u))$, and we formally denote it as $\phi \circ T=S\circ \phi$.
-
-\begin{Definition}\rm  
-[\bf {Conjugacy \& Semi-Conjugacy}]\label{Dfn_Conjugate}\rm
-   Consider the dynamical systems $T:U\to{U}$, $S:V\to{V}$ and suppose the relationship $\phi \circ T=S\circ \phi$ holds.
-If $\phi:U\to{V}$ is a homeomorphism, then $S$ is said to be conjugate to $T$ and $\phi$ is called a conjugacy with the underlying spaces.
-If we relax the criterion on $\phi$ and merely require $\phi$ to be continuous and surjective where $\phi:U\to{V}$, then $\phi$ is a semi-conjugacy between $T$ and $S$.
-\end{Definition}
-
-\ednote{B: Does defining explicitly the domain and codomain of $\phi$ make it clear enough?
-M: Saying $\phi$ is then a semi-conjugacy between $T$ and $S$ could be ambiguous as it does not specify which system is on the top in the commutativity.
-So please edit accordingly.
-M: Very often they are implicitly understood...
-so no worries, and I have edited.}
-When $S$ is conjugate to $T$, the dynamics of the two systems are in some way ''dynamically equivalent''.
-Specifically, they are in one-to-one correspondence with one another.
-However when $S$ is semi-conjugate to $T$ with $\phi:U\to{V}$ a many-to-one mapping, the dynamics on $V$ provide merely a coarse-grained description of the dynamics on $U$~\cite{de2013elements}.
-When $S$ is semi-conjugate to $T$, it is also common to call $S$ a \emph{factor of $T$}, or conversely that $T$ is an \emph{extension of $S$}.
-In essence, an extension (e.g., \cite{de2013elements}) is a more extensive system capturing all of the essential dynamics of its factor.
-
-It is a very hard or nearly an impossible task to establish the existence of a conjugacy or a semi-conjugacy $\phi$ between two systems \cite{devaney2018introduction}.
-However, one can verify that a function $\phi$ satisfies the commutativity diagram \ref{eqn_conjugacy}.
-
-\begin{Example}\rm
-  For example $\phi(x)=\sin\big(\frac{\pi}{2}x\big)^2$ is a conjugacy between two systems $Tu=1-|2u-1|$ and $Sv=4v(1-v)$  defined on $U=[0,1]$.
-
-  Indeed, $\phi:[0,1]\to[0,1]$ is a homeomorphism and by employing simple trigonometric identities, it is easily proved that $\phi$ satisfies~\ref{eqn_conjugacy}
-  \[
-    \begin{aligned}[t]
-      (\phi\circ{T})u 
-                &=\sin^{2}\Big(\frac{\pi}{2} - \frac{\pi}{2}|2u-1|\Big) = \sin^{2}(\pi{u})=\Big(2\sin(\frac{\pi}{2})\cos(\frac{\pi}{2})\Big)^{2}=4\sin^{2}(\frac{\pi}{2}u)\Big(1-\sin^{2}\frac{\pi}{2}u\Big) \\
-                                    &=({S}\circ\phi)u.
-    \end{aligned}
-\]
-And it follows that $\phi$ is a conjugacy between the two systems.
-\end{Example}
-
-By establishing that two systems are conjugate (semi-conjugate) to one another, we have also shown that one may choose to work with one system instead of the second and still be guaranteed to obtain information on the latter.
-It is instrumental during the process of forecasting the future evolution of dynamical systems.
-
-\chapter{The Learning Problem}\label{ch3}
-
-In this chapter, we acquaint ourselves with the question of forecasting dynamical systems with unknown underlying structures, state and discuss the shortcomings of the Takens delay embedding theorem and discuss various issues faced while forecasting.
-
-Consider a relatively simple learning problem: 
-Given the sequence $(u_0, u_1, \ldots, u_m)$ for $m\in\mathbb{Z}$, a finite segment of an orbit of the map $T$ where $T$ is defined by the update equation $u_{n+1} = Tu_n$, forecast the values $u_{m+1}, u_{m+1}$ where the map $T$ is unknown, given that $u_m$.
-
-A practical example of this would be the subsequent scenario: given the time-sequential coordinates of an object moving in space, predict the future positions of that object.
-We are very rarely, if at all, presented with a problem where the entire state information is available to us in the form $u_n$ at some timestep $n\in\mathbb{N}$.
-Consider then a more complicated learning problem:
-
-Suppose we only have the observations $\theta(u_0), \theta(u_1), \ldots, \theta(u_m)$ of the true system states $u$ in an unknown dynamical system $(U,T)$ and we wish to predict the values $\theta(u_{m+1}), \theta(u_{m+2})$ and so forth.
-
-First we establish the method of Takens delay embedding. 
-In this method, one considers a discrete-time dynamical system defined on a 'nice' space (a smooth manifold\footnote{A manifold is a special topological space that is locally homeomorphic at each point to an open set of $\mathbb{R}^d$} that can be obtained as time-$K$ map of a continuous time-dynamical system, concepts formalised below.
-We make observations from such a system, i.e., we observe the evolution of $\theta(w_0)$ (where $w_0$ represents the initial state) by examining  a finite set of values of $u_n :=  \theta(w_n) = \theta(Tw_{n-1})$.
-The sequence $\{u_n\}$ represents a scalar time-series and intuitively $\theta$ represents a probe inserted into a bigger system which is itself only measuring/extracting a small part of the greater system state at time $t=n$.
-Consider, for example, a thermometer erected to measure the ambient temperature in a local village.
-This measurement function, the thermometer, is capturing only a single aspect, the temperature, of a much grander dynamical system entailing the current weather of the surrounding area.
-Even more than that though- it is measuring a miniscule part of the global weather system.
-
-%Given the sequence $\{u_1, u_2, \ldots, u_n}$, we define the variable \newline $y_n:=[\theta_n, \theta_{n-\tau}, \ldots, \theta_{n-(2d-2)\tau}, \theta_{n-(2d-1)\tau}]^{T}\in\mathbb{R}_{d}\times{1}$ where $\tau$ represents the lag and $d$ the dimension of the attractor.
-(We shall return to these terms shortly and their discussion may be put on hold for the time being).
-
-Hence, when the actual state $u$ of a system is seldom, if ever, fully known.
-the most one can do is to insert probes into a system to obtain partial information through the measurements taken.
-Moreover, the process of taking a measurement itself introduces two additional aspects which complicate the problem:
-\vspace{-8mm}
-\begin{enumerate}[noitemsep, label=\roman*.]
-  \item A series of measurements over a specific time interval is inherently an orbit if discretisation  of the underlying continuous-time dynamical system.
-(Hence why we restricted our attention to discrete-time systems in the preceding section)
-  \item A probe will never be entirely accurate, and so the act of measurement introduces a certain measure of numerical noise/inaccuracy (and so we could consider some noise to be involved).
-\end{enumerate}
-
-Takens embedding theorem does not take into account the noise, and for describing it, we construct a multidimensional observable using the method of stacking previous observations, i.e., we create delay-coordinate map defined by
-$\Phi_{k,\theta}(w) := (\theta(T^{-k}w)\ldots,\theta(T^{-1}w),\theta(w))$.  
-The concept of an observable is understood in the sense of an observable as originally introduced by~\cite{takens1981detecting, genericObservableAeyels} to refer to a probe or measurement function inserted into the system.
-Takens' theorem, in essence, refers to the fact that for a sufficiently large $k$, we can define a dynamical system on the space $\mathbb{R}^{k+1}$ with states evolving as $\Phi_{k,\theta}(w)$, $\Phi_{k,\theta}(Tw)$, $\Phi_{k,\theta}(T^2w)$, etc.
-This dynamical system is topologically conjugate to the unknown underlying system $(W,T)$.
-We recall the Takens delay embedding theorem next.
-
-%We may then ask ourselves the question: Is it in any way possible to retain information about the system state $x(t)$ in this temporal data-series $\theta(x(t))$?
-The answer is easily yes if T is known.
-However, neither this, nor even the exact function $\theta$ is available.
-
-\section{Takens Embedding Theorem}\label{sect_Takens}
-
-We define first the concepts of homeomorphism and embedding:
-\begin{Definition}\rm
-  [\bf {Homeomorphism}]\label{Dfn_homeo}\rm
-  A homeomorphism is a function $f:Z\rightarrow Y$ between two topological spaces $Z$ and $Y$ that is continuous, bijective and has a continuous inverse. 
-\end{Definition}
-
-\begin{Definition}
-  [\bf {Embedding}]\label{Dfn_embed}\rm
-  Consider a homeomorphism $f:Z\rightarrow Y$ for $Y\subset X$. $Z$ is said to be embedded in $X$ by $f$.
-\end{Definition}
-
-Takens Theorem states a result establishing a relationship between the observed and underlying dynamical systems by showing that the concatenation of a sufficiently large number of previous observations into a vector will, under certain conditions, generate a map between the vectors from the respective systems.
-We formulate the theorem from \cite{takens1981detecting}.
-
-\begin{Theorem} 
-	[\bf Takens Embedding Theorem (adopted from \cite{takens1981detecting}] \label{Thm_Takens}
-         Let $W$ be a compact manifold of dimension $m$, and $d\ge m$ so that $2d$ is an integer. It is a 
-            generic property for the pair $(T, \theta)$,  where $T:W \to W$ is
-            a smooth diffeomorphism, and $\theta:W \to \mathbb{R}$ a smooth function, the map $\Phi_{2d,\theta}:W \to \mathbb{R}^{2d+1}$ defined on $W$ by 
-            $\Phi_{2d,\theta}(w) := (\theta(T^{-2d}w)\ldots,\theta(T^{-1}w),\theta(w))$
-            is a diffeomorphic embedding; by `smooth' we mean at least $C^2$.
-Consequently, there exists a map $F_\theta: \Phi_{2d,\theta}(W) \to \Phi_{2d,\theta}(W)$ defined by $$F_\theta: (\theta(T^{-2d}w),\ldots,\theta(T^{-1}w),\theta(w)) \mapsto
-            (\theta(T^{-2d+1}w),\ldots,\theta(w),\theta(Tw))$$
-           so that $(W,T)$ is topologically conjugate to 
-            $(\Phi_{2d,\theta}(W), F_\theta)$.    
-\end{Theorem}
-
-By generic we mean a residual set on a certain topology on appropriate spaces of functions (that we do not describe here) \textbf{(Cite)}. 
-By $C^2$, we make reference to a twice-differentiable function  with a continuous second derivative.
-In our scenario (and this remains important throughout) the input space $U$ is considered the attractor.~\label{attractor_U}
-
-Below we reproduce Figure 1 from \cite{Supp}.
-
-%Some explanation is beneficial.
-Takens establishes a delay-coordinate map $\Ftheta$ defined in \ref{eqn_takens} for $T$ the flow (\ref{defn_flow}) defined by the update equation  $Tu_n=u_{n+1}$ and $theta$ our measurement function.
-When confining a dynamical system to the manifold U (in which the underlying system is contained), Takens showed that if certain smoothness conditions are satisfied on $T$ and $\theta$, then the delay-coordinate map $\Ftheta$ embeds $U$ in the reconstruction space $R^{2d+1}$ \textbf{we haven't used the term reconstruction space yet, so this is falling out of the blue}, for almost every choice of measurement function $\theta$, the observable.
-By almost every choice here, we mean that we exclude some observations which do not give any information.
-To see a more precise discussion, see \textbf{xxx} for we do not concern ourselves with a further elaboration on this report.
-
-%Alternatively we may also make the statement: Takens' Embedding Theorem guarantees that almost all dynamical systems can be reconstructed from just one noiseless observation sequence \textbf{this isn't obvious from the theorem's formulation, is it?} i.e.
-for a great number of possible observation functions $\theta$, $\Ftheta$ preserves the topology of U.
-\textbf{'preserving the topology' is not something we've touched on before.
-Perhaps I should add to previous discussion so logic flows seamlessly here}.
-
-\begin{figure}[ht]
-  \includegraphics[scale=0.25]{_takensmap.eps}
-  \centering
-  \captionof{figure}{Schematic of Takens Embedding Theorem's usage and impact of iterates slipping from the attractor when the approximation $\tilde{F}_\theta$ of $F_\theta$ is learnt.}
-  \label{fig:takensmap}
-\end{figure}
-
-If we can learn the map $F_\theta$, from sufficient number of data points of $\{\phi_{2d,\theta}(w_n)\}$, then we also know how to forecast how $\theta(w_n)$.
-
-%Once again we pause and consider the graph to consolidate our understanding of the conjugacy (\textbf{are we being too simplistic?}).
-If we are in state $u$ and then evolve towards state $Tu$ before embedding into $\mathbb{R}^{2m-1}$, then this is equivalent to first embedding into $\mathbb{R}^{2m-1}$ and then evolving through the map $\Ftheta$, i.e.
-$\Phi_{2d,\theta}\circ{F}_{\theta} = T\circ\Phi_{2d,\theta}$.
-The map $F_theta$ is a homeomorphism as well.
-
-%Thus information about U can be retained in the time series' (or observation's) output.
-By preserving the topology on the manifold $U$ in the reconstruction space $X$ (\textbf{not discussed before}), we are also guaranteed the preservation of topological invariants of the manifold, of which dimensionality is one such invariant.
-(We note here that dimensionality will again be considered later in this project).
-
-\section{Practical Limitations to Takens' Theorem}
-Takens' Embedding Theorem is a powerful result, and it provides compelling reason to believe that one could conceivably reconstruct a system conjugate accurately to the underlying system. 
-Nonetheless, it does present some severe practical limitations:
-\vspace{-5mm}
-\begin{enumerate}
-\item Even supposing that we could indeed find $F_\theta$ , our \emph{approximation} of  $F_\theta$ is a map from a larger set $\mathbb{R}^{2d+1}$ containing the embedded attractor.
-There are, however, no theoretical guarantees that $F_\theta$ will retain  $V=\Phi_{2d,\theta}(W)$  as an attractor although $W$ could itself be an attractor.
-\item Takens theorem is stated only for noiseless observations.
-Due to noise $\epsilon_n$ the delay vector $\Phi_{2d,\theta}(w_n) + \epsilon_n$  may lie outside $V$.Furthermore, due to the chaotic nature of the underlying system (i.e.
-the fact that it has SIDC), the evolution of $\Phi_{2d,\theta}(w_n) + \epsilon_n$ under the map $F_\theta$ could move out of $V$ completely.
-This problem can be overcome  by using a driven dynamical system with some properties, and we discuss this in Chapter\ref{ch3}.
-\end{enumerate}
-
-% The Takens embedding theorem does not guarantee global dissipativity.
-% \ednote{B:Links with~\ref{subs_LearnGamma} in discussion of dissipativity of $\Gamma$.
-Sir, I'm not sure how to write this here - I think it a beneficial point to make so as to set the stage for the later discussion in ~\ref{subs_LearnGamma}, but I am unsure there as well.}
-
-Though not a uniquely related to the Takens embedding theorem, we do opt to make mention of another requirement for our work. 
-The methods described here are applicable for data originating from a surjective map. 
-This follows from the fact that when $T$ maps the space $U$ into a proper subset $B$ of$ $U, there exists some point in $U - A$ that does not have a preimage and we conclude that $T$ is not surjective.
-An example of this would be the case when a system displays energy-loss.
-Consider, for instance, consider the dying oscillations for a damped pendulum.
-(An account of greater depth  is provided in~\ref{ch5})
-
-This requirement is not overly restrictive, since if a map is topologically transitive on a subspace it has to be surjective on it, since if it were to be not surjective there would be an open set $V$ that does not intersect with any future iterates of an open set  $U$ disjoint from it.
-\ednote{B: I remember you saying this in one of our sessions, but don't have a source for it.
-Do you have a specific reference?
-M: Follows from the transitivity definition.
-It is enough}
-
-\chapter{Driven Dynamical Systems to Forecast Problems}\label{ch4}
-
-In this chapter, we discuss results on the mapping of temporal data obtained from a discrete-dynamical system onto a different space through the notion of a driven dynamical system. 
-We also consider the conditions a driven system should possess to avoid adding distortion to its state-space representation.
-We then describe the \emph{single-delay dynamics} (SDD) of the system and finally recall conditions on the driven system so that the SDD are conjugate (or at least semi-conjugate) to the underlying system.
-The SDD can then be used to forecast and reconstruct the underlying system.
-
-\section{Nonautonomous and Driven Dynamical Systems}
-
-\begin{Definition}
-  [\bf Nonautonomous Dynamical System (adopted from~\cite{Manju_ESP}))] \label{Dfn_NDS}\rm
-  A nonautonomous dynamical system (NDS) on a space $X$ is simply a dynamical system comprising a family of maps $\{g_n\}_{n \in \mathbb{Z}}$ where $g_n(\ldotp):=g(u_n,\ldotp):X\to{X}$ is continuous.
-Each $g_n$ arises as a consequence of an input $u_n$ from the input space $U$, a topological space.
-\end{Definition}
-
-We immediately recall the figure first defined in \eqref{eqn_conjugacy}  and remind ourselves of our ultimate goal to learn a system that is conjugate to an underlying or unknown dynamical system $(U,T)$ by making use of the measurements obtained from  the unknown system.
-
-%#####################################
-\begin{center}
-\psset{arrows=->, arrowinset=0.25, linewidth=0.6pt, nodesep=3pt, labelsep=2pt, rowsep=0.7cm, colsep = 1.1cm, shortput =tablr}
-\everypsbox{\scriptstyle}
-\begin{psmatrix}
-U & U\\%
-V & V.
-%%%
-%  \ncline{1,1}{1,2}^{T} 
-%  \ncline{1,1}{2,1} <{\phi}
-%   \ncline{2,1}{2,2}^{S}
-%  \ncline{1,2}{2,2} > {\phi}
-\end{psmatrix}
-%#####################################
-\end{center}
-
-To find a function $\phi$, the need now arises to consider a so-called driven dynamical system, a special case of a nonautonomous dynamical system.
-Two spaces are considered: input is taken from an input space $U$ and the state of a space $X$ is updated.
-(Previously, only the state was considered.)
-We do this to mimic the true conditions.
-By accounting for an extrogenous input sequence in $U$ which can also influence the  system-state at a specific time-step, it produces more general models than an autonomous system.
-
-\begin{Definition}
-  [\bf Driven, Compactly Driven Dynamical System] \label{Dfn_DDS} \rm
-A driven dynamical system comprises two topological metric spaces $U$, $X$ and a continuous function  $g:U\times{X}\to{X}$ where $g(u_n, x_n)=x_{n+1}$.
-If the input space $U$ is compact, we refer to the system as compactly driven. 
-\end{Definition}
-
-The update equation $x_{n+1} = g(u_n,x_n)$ for $n \in\mathbb{Z}$, input $u_n$ from $U$ and state $x_n$ belonging to $X$ (a compact space) generate the dynamics on X. 
-Abbreviated, we refer only to the \textit{driven system $g$}, where all other entities are understood implicitly.
-
-Notably, a nonautonomous dynamical system can be generated from $U$; any input $\overline{u}$, a bi-infinite sequence from $U$, gives rise to the sequence of self-maps $\{g(u_n, \cdot)\}_{n\in\mathbb{Z}}$ contained in $X$.
-Physically, one may think of this bi-infinite sequence as referring to a system that has been running for an extended period at the time of the first measurement taken from the system (alternatively, the first time a probe is inserted into the system to take an observation)
-
-\begin{Definition}
-  [\bf Entire Solution] \label{Dfn_Soln} \rm
-  A sequence $\{x_n\}_{n\in\mathbb{Z}}\subset X$ is called an entire solution (or simply a solution) of the driven system  $g$ with input $\overline{u}$ when it satisfies 
-  \[g(u_{n-1}, x_{n-1})=x_n\] for all $n\in\mathbb{Z}$
-\end{Definition}
-
-It is important to emphasise that a sequence $\{x_n\}$ satisfying the update equation above can only be a solution if $x_n\in{U}$ holds for all $n\in\mathbb{Z}$ and just for $n>0$.
-Consider the example below.
-
-\begin{Example} \rm \label{ex_halfux}
-  The only solution  $\{x_n\}_{n\in\mathbb{Z}}$ to the driven system  $g(u,x)=\frac{ux}{2}$, where $X=[0,1]$, $U=[0,1]$,  is the zero solution $x_n\equiv0$.
-  To see this, consider any $x_n=a\in[0,1]$ where $a\neq{0}$.  Let $\overline{u}\in{U}$ be an non-zero constant sequence , say $u_n=0.5$. 
-  The driven system may be rewritten as $x_{n-1}=\frac{2x_n}{u_{n-1}}=4x_n$ and the  iterates of $x_n$ in backward time will increase by a factor of 4 at each timestep. 
-  Thus for some $m\leq{n}$,  $1<x_m$ i.e. $x_m\notin{X}$. So ${\{x_n\}}_{n\in\mathbb{Z}}$ is not a solution and it follows that the only possible solution is the zero solution.
-\end{Example}
-% \ednote{Nicely written! Thank you}
-
-A system may also have multiple solutions, as evidenced in the example below.
-
-\begin{Example}\label{Ex_exp} \rm 
-  Consider the driven system $g(u,x)=x^2$ for $X=[0,1]$, $U=\mathbb{R}$.
-The system has an uncountable number of solutions, as there exists a solution for every $x\in{X}$ which also passes through the point $x$ and $lim_{n\to\infty}x_n=0$, $lim_{n\to{-}\infty}x_n=1$.
- The proof is deferred to immediately after the next paragraph (see~\ref{rem_proofEx}).
-\end{Example}
-
-As the solutions to a driven system are considered an important entity, we next identify a subspace $X_U$ of $X$ that contains all possible solutions.
-To realize such a subspace of a driven system $g$, the concept of a reachable set is defined.
-
-\begin{Definition}
-  [\bf Reachable Set]\label{Dfn_ReachableSet}\rm
-The reachable set of a driven system $g$ is exactly the union of all the elements of all the (entire) solutions, i.e., 
-\[X_U :=\Big \{x \in X:  x = x_k \mbox{ where  $\{x_n\}$  is a solution for some  $\bar{u}$} \Big \}.\]
-The set of all reachable states at a specific time $n$ for input $\overline{u}$ is denoted by $X_n(\overline{u})$
-\end{Definition}
-
-The reachable set itself can defined independently from $g$ possessing the ESP/USP.
-We note that $x\in{X_n(\overline{u})}$ if and only $g$ has a solution $\{x_k\}$ for $x_n=x$ and input $\overline{u}$.
-\textbf{(Cite)}.
-
-The set $X_n(\overline{u})$ is precisely the set of points $x$ through which some entire solution $\Psi$ passes at time $n$. We can give an alternate formulation of the set $X_n(\overline{u})$. 
-Now suppose, we for some fixed input 
-$\overline{u}$, we define $g_i = g(u_i,\cdot)$ for all $i\in \mathbb{Z}$, then the set of states
-
-\begin{equation} \label{eqn_association}
-X_{n,i}(\overline{u}) := g_{n-1} \circ \cdots g_{i+1} \circ g_i(X).
-\end{equation}
-
-The set $X_{n,i}(\overline{u})$ being the image of a finite composition of
-continuous maps is compact whenever $X$ is compact. Also $X_{n,i}(\overline{u})$ is
-nonempty.  Further, $X_{n,i}(\overline{u}) \supset X_{n,i-1}$. Hence $\bigcap_{i<n}
-X_{n,i}(\overline{u})$ is a nested intersection of closed nonempty subsets, and
-whenever $X$ is compact, the intersection is nonempty. We say $g$ is a topological contraction if $\bigcap_{i<n}X_{n,i}(\overline{u})$ is a singleton subset of $X$ for each $\overline{u}$.
-
-It turns out that $\bigcap_{i<n}X_{n,i}(\overline{u})$ is identically equal to $X_n((\overline{u})$ that we had defined earlier.
-This will lead to a result (for.
-e.g., \cite{manjunath2013echo}) that  \textit{$g$ being a topological contraction is equivalent to the existence of a exactly one entire solution.}
-
-% \begin{Definition}
-%   [\bf Topological Contraction]\label{Dfn_TopContr}\rm
-%   A function $g:U\times{X}\to{X}$ is a topological contraction if for all $n\in\mathbb{Z}$ and all $\overline{u}\subset{U}$, $X_n(\overline{u})$ is a singleton subset of $X$.  
-% \end{Definition}
-
-\begin{Remark}
-  [\bf Proof of Example~\ref{Ex_exp}] \label{rem_proofEx} \rm
-  For the driven system,  
-  $g(u,x)=x^2$ the evolution does not depend on the input at all.
-Hence $g(u,x)$ can be written as a single map $f(x)=x^2$ which is a homeomorphism on $[0,1]$.
-A left-infinite orbit is defined and it converges to $1$ while the right-infinite orbit converges to $0$.
-Or in other words the iterates of $f^{-1}$ converges to $1$ while the iterates of $f$ converge to $0$
-\end{Remark}
-
-Thus far, it has been demonstrated that a system may have one or more solutions; one may ask if a driven system always has a solution and, if so, whether it satisfies specific properties such as uniqueness.
-Should the driven system be compact, existence follows immediately, as shown in the following result.
-
-\begin{Theorem}\label{Thm_CompactExistence}
- Let $g$ be a driven system.  If $X$ is compact then for each input $\overline{u}$, there exists at least one solution to the driven system $g(\ldotp, x)$
-\end{Theorem}
-\begin{proof}
-  May be found in \cite{kloeden2011nonautonomous, manjunath2014dynamics, manjunath2013echo}
-\end{proof}
-
-% One may easily construct many systems with trivial solution-sets, such as $g(u,x)=x$ which has only the constant solution $x$ and so for $U=[-1,1]$, the system would have no solution if $|x|>1$.
-To refine the scenario, we consider only systems with unique solutions.
-
-\section{Unique Solution Property}
-
-\begin{Definition}
-  [\bf Unique Solution Property] \label{Dfn_usp}\rm
-  A driven system $g$ is said to have the Unique Solution Property (USP) if for each input $\overline{u}$ there exists exactly one solution. 
-  Alternatively we may formulate the USP as follows: $g$ has the Unique Solution Property if there exists a well-defined map $\Psi:{U}\to{X}$, with $\Psi({\overline{u}})$ denoting the unique solution.
-\end{Definition}
-
-The USP is equivalent called  the echo state property  \cite{Manju_ESP}, and systems that satisfy this property have other important and useful properties \cite{grigoryeva2018echo, grigoryeva2019differentiable, manjunath2020stability} for information processing.
-One of the first results obtained after defining the USP is that every solution will attract different initial conditions towards the component parts of the solution.
-If $g$ has the USP, then any solution to $g$ is also a uniform attractor in nonautonomous dynamical systems literature~\cite{Manju_Nonlinearity}.
-The discussion on nonautonomous attractors is beyond the scope of this project, and we refer the reader to \cite{Manju_ESP, esann2012ids} for further reading.
-
-% Paraphrased.
-%Additionally, $g$ having the USP is in our context also equivalent to $g$ being a topological contraction \cite{manjunath2021universal}, or alternatively also that $g$ exhibits the Echo State Property (ESP).
-
-In the majority of Reservoir Computing(RC) literature (a machine learning approach) where the input is mapped onto a different but higher dimensional through a system called the reservoir and a readout that measures the state of the reservoir is trained, a notion of forgetting the states of the reservoir asymptotically is often used.
-
-% \ednote{How should I explain/qualify RC better here?
-It's mentioned without really explaining.
-M: Have put it in one sentence; just to let you know people use optical devices to map signal into a ``reservoir".
-B: Thanks!},
-Concepts like the ESP,  fading memory \cite{boyd1985fading} are some of these notions.  
-If $g$ possesses the ESP (equivalent in our context to the USP), the we are guaranteed that the whole of an input's left-infinite history will exactly determine the current system state; i.e.
-there is only ever the possibility of a single reachable state at a given instance of time \cite{jaeger2001echo,Manju_2020}.
-% Defined formally, the ESP is intrinsically linked to an input (see \cite{Manju_ESP}) Unfortunately, definitions and results concerning the ESP do not have much practical benefit since a relationship between the network's response and the temporo-statistical properties of the input are not directly described.
-A more detailed presentation may be found in~\cite{jaeger2001echo}.
-
-% We do however consider it important to mention that the ESP is also often discussed in terms of a stability property that holds for all inputs of a system~\cite{manjunath2020stability} and, furthermore, plays a key role in the design and training of recurrent neural networks(RNN) within the field of RC \cite{Manju_ESP}.
-Beyond forgetting the past states the ESP is also often discussed in terms of a stability property that is nearby inputs create close by responses (solutions), a concept called input-related stability~\cite{manjunath2020stability} and, furthermore, plays a key role in the robustness of recurrent neural networks(RNN), a component of the implementation discussed in~\ref{ch5}.
-
-\section{Conjugacies}
-
-Having already defined the reachable set $X_U$ as the collection of all elements of all entire solutions, we pause briefly in order to fix additional notation.
-Defining $\cev{u}^{n}:=(\ldots,u_{n-2} ,u_{n-1})$ as the left-infinite subsequence of an input up until time $n$, $\overleftarrow{U}$ is then the notation for all these left-infinite sequences in $U$.
-Moreover, $\cev{u}^{n}v:=(\ldots,u_{n-2} ,u_{n-1}, v)$ will symbolise the input up to time $n$ with $v \in U$ being the specific input value at time $n$. 
-The introduction of a new input at time $n$ can be described by the mapping $\sigma_v:   \cev{u}^{n} \mapsto \cev{u}^{n}v$. 
-The right-shift map, $r\overleftarrow{U}\to\overleftarrow{U}$, of an input sequence is defined $r: (\cdots, u_{-2},u_{-1}) \mapsto(\cdots, u_{-3},u_{-2})$
-
-The question now assumes the form: Can we establish a semi-conjugacy for the driven system $g$ as presented below.
-
-\begin{equation}  \label{Scomm_h}
-  %    \[ 
-      \psset{arrows=->, arrowinset=0.25, linewidth=0.6pt, nodesep=3pt, labelsep=2pt, rowsep=0.7cm, colsep = 1.1cm, shortput =tablr}
-   \everypsbox{\scriptstyle}
-   \begin{psmatrix}
-   \overleftarrow{U} & \overleftarrow{U}\\%
-   X_U & X_U.
-   %%%
-  %  \ncline{1,1}{1,2}^{\sigma_v} \ncline{1,1}{2,1} <{h}
-  %  \ncline{1,2}{2,2} > {h}
-  %  \ncline{2,1}{2,2}^{g(v,\cdot)}
-   \end{psmatrix}
-  % \]
-  \end{equation}
-
-We proceed to consider a specific subclass of conjugacies.
-
-  \begin{Definition}
-    [\bf Universal Semi-Conjugacy]\label{Def_UnivSemiConj} \rm
-    Given a driven system $g$, we  call a continuous and surjective map $h : \overleftarrow{U} \to X_U$ a universal semi-conjugacy if  diagram \ref{Scomm_h} commutes for all $v \in U$.
-  \end{Definition}
-
-  If the universal semi-conjugacy $h$ exists (i.e. the diagram in \ref{Scomm_h} commutes) then the solution $\Psi(\bar{u})$ will intuitively have no more ``complexity' than the input $\bar{u}$.
-
-But does such a function $h$ for the above driven system $g$ above exist?
-Whenever $g$ has the USP and $\Psi(u)=\{x_n\}_{n\in\mathbb{Z}}$ it follows that $h$, defined by  $h(\cev{u}_n):=g(u_n,x_{n-1})=x_n$, will satisfy the semi-conjugacy in the graph above \ref{Scomm_h}.
-Regrettably, such a continuous mapping $h$ is not guaranteed to exist when $g$ does not have the USP \cite[Lemma 5]{Manju_Nonlinearity}.
-Note that even when $h$ does exist, we are not guaranteed its injectivity. Considering again example \ref{ex_halfux}, we see that even if $h$ were to exist, it could not be injective as $X_U=\{0\}$.
-
-Re-sketching the commutativity diagram \ref{Scomm_h} above by replacing $X_U$ by its left-infinite sequence space $\overleftarrow{X}_U$, we obtain the diagram below.
-In this case, the function $H:\overleftarrow{U}\to\overleftarrow{X}_U$, a map that is both continuous and surjective, is called a \emph{causal mapping} that is defined next.
-
-\begin{Definition}
-  [\bf Causal Mapping]\label{Def_CausMap}
-  A continuous, surjective map $H:\overleftarrow{U}\to\overleftarrow{X}_U$ such that \[H\circ\tilde{g}_v=\sigma_v \circ{H}\] holds for all $v \in U$ where $\tilde{g}_v$ maps $(\ldots, u_{-2}, u_{-1})$ to $(\ldots, u_{-2}, u_{-1}, g(v, u_{-1}))$ is called a causal mapping.
-\end{Definition}
-
-\begin{equation} \label{SCausal_H}
-    %    \[ 
-        \psset{arrows=->, arrowinset=0.25, linewidth=0.6pt, nodesep=3pt, labelsep=2pt, rowsep=0.7cm, colsep = 1.1cm, shortput =tablr}
-     \everypsbox{\scriptstyle}
-     \begin{psmatrix}
-     \overleftarrow{U} & \overleftarrow{U}\\%
-     \overleftarrow{X}_U & \overleftarrow{X}_U.
-     %%%
-    %  \ncline{1,1}{1,2}^{\sigma_v} \ncline{1,1}{2,1} <{h}
-    %  \ncline{1,2}{2,2} > {h}
-    %  \ncline{2,1}{2,2}^{\tilde{g}_v}
-     \end{psmatrix}
-    % \]
-  \end{equation}
-
- \begin{Theorem}
-  For a compactly driven system, a causal mapping $H$ exists if and only if $g$ has the USP. 
-\end{Theorem}
-{\bf Proof.}  See~\cite[Th.3]{manjunath2013echo}.
-
-The driven system $g$ can induce an embedding of $\overleftarrow{U}$ in $\overleftarrow{X}$ as follows: 
-If the causal mapping $H:\overleftarrow{U}{\to}{\overleftarrow{X}_U}$ is injective (in addition to being surjective), it becomes the embedding of $\overleftarrow{U}$ in $\overleftarrow{X}$ induced by the driven system $g$.
-The map $H^{-1}$ is continuous since $H$ is a continuous and surjective mapping of a compact space $\overleftarrow{U}$ to a Hausdorff space.
-and we refer to the map $H$ as a \emph{causal embedding}.
-
-When $g$ has the USP, the diagram below (adopted from~\cite{Manju_Nonlinearity}) illustrates the operation of the mappings $h$ and $H$.
-The mapping $h:\overleftarrow{U}\to{X_U}$ is also considered an observable as mentioned in the introduction of Chapter \ref{ch3}).
-
-\begin{figure}[ht]
-  \includegraphics[scale=0.4]{_actionofh_H.eps}
-  \centering
-  \captionof{figure}{Ladder-like behaviour of $h$ and $H$ to illustrate the causal mapping. (Figure reproduced from~\cite{Manju_Nonlinearity}) }
-  \label{fig:actionh_H}
-\end{figure}
-
-The function $H$ contains a countably infinite number of coordinate (or factor maps) since its codomain is $\overleftarrow{X}_U$.  
-As it is not practically feasible to consider a left-infinite sequence in $X_U$ (in other words an element of  $\overleftarrow{X}_U$) for learning, we wish to restrict the inputs we consider to a subspace of $\overleftarrow{U}$.
-Particularly, we restrict our attention to the  so-called inverse-limit space of a dynamical system and hope to embed it within some finite product of $X$.
-
-%We wish to explore the possibility of establishing a conjugacy rather than a mere semi-conjugacy and therefore are interested in the restriction of inputs to the left-infinite orbits of a dynamical system.
-
-The left-infinite orbit space is significant because a great amount of information about the original system can be gleaned from its topological structure~\cite{ingram2011inverse,Manju_IEEE}. 
-The formal topological structure for the inverse-limit space of a chaotic dynamical system is complicated, and further discussion is relegated to~\cite{kennedy2008inverse_limit, ingram2011inverse}.
-Broadly, we may conceive of the inverse-limit system of a dynamical system as a subspace of an infinite-dimensional space where each point in the inverse-limit space corresponds to a backward orbit of the map $T$.
-Following~\cite{manjunath2021universal, ingram2011inverse} we provide a definition:
-
-\begin{Definition}
-  [\bf Inverse-Limit Space, Inverse-Limit System]\label{Dfn_InverseL}\rm
- The subset $\widehat{U}_T\subset \overleftarrow{U}$ defined by \[\widehat{U}_T: = \{ (\ldots,u_{-2},u_{-1}): Tu_{i} = u_{i+1}\} \] is called the inverse-limit space of $(U,T)$.
- The self-map $\widehat{T}$  induced by $T$ on $\widehat{U}_T$  is defined by \[\widehat{T}: (\ldots,u_{-2},u_{-1}) \mapsto  (\ldots,u_{-2},u_{-1},T(u_{-1}))\] and the resulting dynamical system $(\widehat{U_T}, \widehat{T})$ is called the inverse-limit system of $(U,T)$.
-\end{Definition}
-
-The inverse-limit space is well-defined since $T : U \to{U}$ is surjective by assumption.
-We now get to the concept of embedding the inverse limit-space of the dynamical system in $X \times X$. 
-\begin{Definition} \rm
-	We say a driven system $g$ \emph{causally embeds} a dynamical system $(U,T)$ if it satisfies the two properties: (i) a universal semi-conjugacy exists and  (ii)  $H_2(\cev{u}) := (h(r\cev{u}),h(\cev{u}))$ embeds the inverse-limit space $\widehat{U}_T$ in $X \times X$.
-\end{Definition}
-
-  It is imperative to take note of a subtlety here: we refer to the \emph{map} $H$ as a (causal) embedding, whereas when the driven \emph{system} $g$ (causally) embeds another \emph{system} $(U,T)$ we are concerned with $H_2$.
-\ednote{B: inserted line}
-
-\section{Choosing the driven system $g$}
-
-So far it is not clear if the USP alone ensures the ability to causally embedding a dynamical system and we mention that a driving function $g$ cannot just be chosen in a frivolous manner.
-This could possibly complicate our work.
-%\ednote{B: Grammarly}
-
-%The driving function $g$ cannot be chosen in a frivolous manner as it could possibly complicate our work.
-Specifically, one must be careful to avoid a choice of $g$ which would  add complexity to the obtained solution.
-
-%When a causal embedding $H$ exists for the driven system $g$, one can map an arbitrary input ${u}$ onto the solution space $X$ without additional distortion or information-loss \textbf{(Cite.)}.
-
-%When an embedding is established, the question of possible additional complexity in the solution is removed by guaranteeing that, since the systems are conjugate (semi-conjugate, \textbf{(refer)}), $g$ does not add any (some) complexity to the system.
-
-%To achieve a causal embedding embedding not just a causal mapping, it is undesirable to choose a function $g$ that quenches the temporal structure in $u$ by contracting to such a degree that the ability to recover information from the original system is lost completely.
-In the above example(\ref{ex_halfux}), the input's temporal variation could not be related to the reachable set as $X_U$ only consisted of a single element; little, if not no, information is encoded.
-To obtain a suitably complex function $g$, it is thus desired that the reachable set of a driven system  be large enough to relate to the input. 
-This is true even for embedding  the inverse-limit space of a dynamical system \cite{manjunath2021universal}.
-
- To this end, we recall the notion of State-Input (SI) Invertibility.
-
-\begin{Definition}
-  [\bf SI-Invertibility]\label{Dfn_SIinv}\rm
-  A driven system $g$ is said to be SI-Invertible if $g(*,x): U \to X$ is invertible for all $x\in X$.
-Alternatively it may be said that if, given $x_n$ and $x_{n-1}$, $u_{n-1}$ can be uniquely determined from $x_n=g(u_n,x_{n-1})$, then $g$ is said to be SI-invertible.
-\end{Definition}
-
-SI-invertibility promises that 'enough' information is retained when $g$ is chosen without introducing one's choice of $g$ will still ensure 'enough' information is retained without introducing unwanted complexity.
-'Enough' here refers to the fact that we may always recover the previous input value $u_n$ (if we know successive states $x_{n-1}, x_n$).
-
-Subsequently we define the relation $Y_T$ induced by $(U,T)$ on $X_U\times{X_U}$ for a driven system $g$ possessing SI-invertibility.  
-To describe the  SDD formally, we consider a dynamical system $T: U \to U$ and define a relation on the reachable set $X_U$, i.e. a subset on $X_U \times X_U$  defined by 
-$$Y_T:=\{(x_{n-1},x_n): \{x_k\}_{k\in \mathbb{Z}} \mbox{ is a solution for some orbit of } T \mbox{ and } n \in \mathbb{Z}\}.$$
-
-The following theorem establishes the existence of a well-defined map $G_T$ describing the single-delay dynamics(SDD) of the system above.
-
-\begin{Theorem}\label{Thm_GT_Exists}
-  For an SI-invertible driven system $g$ and a dynamical system $(U,T)$, the map $G_T: Y_T \to Y_T$ defined by the relation $(x_{n-1},x_n) \mapsto (x_n,x_{n+1})$ is well-defined. 
-  (This results holds even in the absence of $g$ possessing over the USP) 
-  % Moreover, the mapping $(x_{n-1},x_n) \mapsto u_{n}$ is well-defined when $x_{n-1}$ and $x_n$ are successive points on a solution obtained for an input  $\{u_n\}$ that is an orbit of $T$.
-  \end{Theorem}
-  {\bf Proof.} See~\cite[Th.3]{Supp}
-
-%'Without additional complexity' is guaranteed by the invertibility of the function $g$ that guarantees a one-to-one mapping between $u_n, x_{n-1}$ and $x_n$.
-In obtaining an SI-invertible function, therefore, we are guaranteed a system which will not lose so much information about its previous states in the forward-flow of time that one may not make any credible claims as to the original system's behaviour and properties.
-Simultaneously, it does guarantee us that the important information encoded in previous inputs is indeed preserved.
-
-At this stage it is worth taking note of a specific driven system in the form of a discrete state-space model which has acquired some adherence in applications \cite{Manju_IEEE}- especially those pertaining to Echo State Networks and the ESP.
-The function
-\begin{equation}  \label{eqn_driving}
-  g(u,x) = (1-a)x + a\overline{\tanh}(Au + \alpha Bx)
-\end{equation} 
-is SI-invertible and, if $\alpha B$ has a spectral norm $<1$, also possesses the USP \cite[Th.2]{manjunath2013echo }. 
-It is easy to show that $g$ is SI-invertible by recovering $u_n$ in 
-\begin{equation} \label{eqn_SI_RNN}
-  u_{n-1} := A^{-1}\bigg(\overline{\tanh}^{^{-1}}\frac{1}{a}\Big(x_{n+1}-(1-a)x_n\Big) \bigg) - \alpha B x_n
-  \end{equation}
-  when $x_{n-1}$ and $x_n$ are known.
-
-This specific driving function $g$ is used in our implementation and is discussed more completely in chapter~\ref{ch5}
-
-Despite the ease that with which one manipulates a left-infinite history in the realm of theory, it is impossible to obtain or use such a sequence in any real-life application.  
-Fortunately, one does not need the entire left-infinite history of an input in practice thanks to the Uniform Attraction Property(UAP). 
-We use an alternate version of the definition of UAP -- since the notion of nonautonomous attractors in this thesis would take some time to establish and detracts from the principal thrust of this project, we refer to \cite{Manju_Nonlinearity}.
-
-\begin{Definition}
-  [\bf Uniform Attraction Property]\label{Dfn_UAP}\rm
-  A driven system $g$ has the Uniform Attraction Property (UAP) if we initialize the driven system
-with an arbitrary initial value $y_m \in X$, and the sequence $y_{m+1}, y_{m+2}, y_{m+3},...$ satisfying $y_{k+1}= g(u_k,y_k)$ for $k \geq m$ then approximates an actual solution $\{x_n\}$ uniformly in the sense that given $\epsilon>0$ (independent of $y_m$) there is an integer $n$ so that $d(x_{n+i}, y_{n+i})<\epsilon$ for all $i\ge 0$, where $\{x_m\}$ is a solution.
-\end{Definition}
-
-The UAP guarantees that all trajectories will converge to the same trajectory as time moves forward.
-An illustration of this is shown in Fig.~\ref{fig:memloss_conttime} where the trajectory in red locks on to the supposed actual solution when the system has the USP.
-
-\begin{figure}[ht]
-  \includegraphics[scale=0.35]{_Nonlinearity_Fig_response.eps}
-  \centering
-\captionof{figure}{A coordinate of a simulated solution $(x_0,x_1,\ldots,x_{5000})$ of a RNN  plotted in red (with parameter $\alpha=0.99$ (where $g$ has the USP) and blue ($a=1$ and $\alpha=1.05$ where $g$ does not have the USP) while the matrices $A$ and $B$ are randomly generated, and $B$ has unit spectral radius.
-The input is shown in black.
-Reproduced from \cite{Manju_Nonlinearity}. }
-\label{fig:memloss_conttime} 
-\end{figure} 
-\ednote{B: How can I explain the black line? I don't think this makes sense to the reader immediately. M: Oh it was the input in black that I have added.}
-
-One may now appreciate an even more astounding result: $g$ having the USP is equivalent to the UAP as is proved in~\cite{Manju_Nonlinearity}.
-
-%We restate the above theorem.
-
-%\begin{Theorem}\label{thm_usp-contr-uap}  The following statements are equivalent:  \vspace{-8mm}  \begin{enumerate}[noitemsep, label=\roman*.]    \item $g$ has the USP.
-\item $g$ is a topological contraction.
-\item $g$ has the UAP.
-\end{enumerate\end{Theorem}{\bf Proof.}  May be found in \cite[Th.1]{Manju_Nonlinearity}
-
-% One need therefore not establish any additional results to ensure that the sequence uniformly approaches the unique solution $\Psi$.
-This vastly simplifies the effort necessary to set up a problem in order to guarantee that the underlying system will be accurately represented by the conjugate system.
-
-%This also solves the problem of perturbation/noise introduced by the observable or measurement function.
-Recall from before that an observable is inherently a map that discretises the underlying continuous-time system.
-Moreover, measurement and device error introduce mistakes.
-Since we are working with a chaotic system displaying sensitive dependence on initial conditions, these small errors could potentially send the trajectory into a completely different attractor.
-%In the presence of the UAP, however, small measurement errors introduced into the system do not pose the same danger as before since the trajectories will eventually tend towards the actual solution.
-% \ednote{(B: Is this Complete enough?)}
-% \textbf{From the perspective of perturbation of an autonomous dynamical system, if the resulting nonautonomous system has the USP, then the temporal structure of noise relates to the dynamics on a nonautonomous uniform attractor through a topological semi-conjugacy or a conjugacy.}
-
-\section{The next step in Dynamics}
-
-% Recall that $g$ is only being given inputs from the orbits of $T$.
-
-%Incredibly, this permits one to initialise a driven system $g$ with an altogether arbitrary initial value $y_m\in{X}$ where $m\in\mathbb{Z}$ and the UAP then guarantees that the sequence $\{y_{m+1}, y_{m+2},\ldots\}$ which satisfies the relation $y_n=g(u_n, y_{n-1})$  for  $k\geq{m}$  will approach the elements $\{x_n\}$ of the actual solution.
-This vastly simplifies the effort necessary to set up a problem in order to guarantee that the underlying system will be accurately represented by the conjugate system.
-
-We are interested in determining whether $\widehat{T}$ and $G_T$ are related.
-For this, we introduce  the function $H_2{\overline{}} := (h(r\overline{u}, h(\overline{u})))$ mapping the entirety of a left-infinite sequence to some element in $X\times{X}$.
-Indeed, it may be shown that $g$ having SI-invertibility and the USP immediately guarantees the existence of at least a semi-conjugacy between the systems $(Y_T, G_T)$ and $(\widehat{U}_T, \widehat{T})$.
-This is formalised in the Causal Embedding Theorem, which we state next:
-
-\begin{Theorem}
-  [\bf Causal Embedding Theorem (adopted from \cite{Supp})]
-  % {\bf (Causal Embedding Theorem.)}
- \label{Thm_CET}
-	Let $g$ be a driven system with SI-invertibility and the USP.
-Let $h$ denote the universal semi-conjugacy and $H_2(\overleftarrow{u}) := (h(r\overleftarrow{u}),h(\overleftarrow{u}))$, where $r$ is the right-shift map.
- Let $(\widehat{U}_T, \widehat{T})$  be the inverse-limit system of a dynamical system $(U,T)$. 
- Then the restriction of $H_2$ to $\widehat{U}_T$ is a topological semi-conjugacy between the inverse-limit system $(\widehat{U}_T, \widehat{T})$ 
-and the induced dynamical system  $(Y_T,G_T)$, i.e., the following diagram commutes
-\begin{equation} \label{comm_H_CET}
-\psset{arrows=->, arrowinset=0.25, linewidth=0.6pt, nodesep=3pt, labelsep=2pt, rowsep=0.7cm, colsep = 1.1cm, shortput =tablr}
- \everypsbox{\scriptstyle}
- \begin{psmatrix}
- \widehat{U}_T & \widehat{U}_T\\%
- Y_T &  Y_T.
- %%%
-%  \ncline{1,1}{1,2}^{\widehat{T}} \ncline{1,1}{2,1} <{H_2}
-%  \ncline{1,2}{2,2} > {H_2}
-%  \ncline{2,1}{2,2}^{G_T}
- \end{psmatrix}
- \end{equation}
-or in other words, $(Y_T, G_T)$ is a factor of  $(\widehat{U}_T, \widehat{T})$. Further, if $T:U \to U$ is a homeomorphism, 
-then $H_2$ embeds $\widehat{U}_T$ in $X_U \times X_U$, and hence $(Y_T, G_T)$ is conjugate to $(\widehat{U}_T, \widehat{T})$.
-\end{Theorem}
-{\bf Proof.}  May be found as the proof of~\cite[Th.4]{Supp}
-
-Recall that $H_2$ maps an entire left-infinite solution sequence from $\Psi$ to an element in $X\times{X}$. 
-We are drawing nearer and nearer to our principal target and our results carry more and more weight: If we are able to learn $G_T$, we will also have learnt $\widehat{T}$; and since $\widehat{T}$ is an extension of $T$, we will have obtained $T$ as well.
-%\ednote{B: Grammarly check}
-
-%\begin{Theorem} Graph \ref{Scomm_h} is exactly the inverse-limit system $(\hat{U}, \hat{T})$.    \end{Theorem}
-
-%We now have the following (which may be compared with \ref{SCausal_H} above):
-
-%\begin{equation} \label{fig:inverse_limsystem}
-  %  \[       \psset{arrows=->, arrowinset=0.25, linewidth=0.6pt, nodesep=3pt, labelsep=2pt, rowsep=0.7cm, colsep = 1.1cm, shortput =tablr}      \everypsbox{\scriptstyle}
-%      \begin{psmatrix}      \widehat{U}_T  && \widehat{U}_T \\     Y_T && Y_T \\
-      %%%
-     %  \ncline{1,1}{1,2}^{\widehat{T}} \ncline{1,1}{2,1} <{h}
-     %  \ncline{1,2}{2,2} > {H_2}
-     %  \ncline{2,1}{2,2}^{G_T}
-%      \end{psmatrix}
-    %  \]
-%  \end{equation}
-
-%\begin{Theorem}
- %   $(Y_T, G_T)$ is semi-conjugate to $(\widehat{U}, \widehat{T})$.
-%\end{Theorem}
-% {\bf Proof.} See \cite[Th.3, Th.4]{manunath2021universal}
-
-\section*{Summarising the discussing thus far:}
-
-It is easy to lose the birds-eye view, so we take a moment to review our progress up until this point.
-
-\vspace{-8mm}
-\begin{enumerate}
-\item We are interested in a some dynamical system $(U,T)$ with unknown dynamics.
-\item To determine properties about this system $(U,T)$ and predict its future evolution, we determine the dynamics of the inverse-limit system $(\widehat{U}, \widehat{T})$.
-% Given certain assumptions, we can guarantee that $(\widehat{U}, \widehat{T})$ is at least semi-conjugate to $(U,T)$.
-\item If the driven system $g$ is SI-invertible (and $\{u_n\}\in{U}$  is an orbit of $T$), the map $G_T$ exists that describes the single delay dynamics. 
-\item If, furthermore, $g$ has the USP, $(Y_T, G_T)$ is semi-conjugate to $(\widehat{U}, \widehat{T})$.
-\item If we can assume that $T$ is a homeomorphism, $(Y_T, G_T)$ is topologically conjugate to $(\widehat{U}, \widehat{T})$, an extension space of $(U,T)$
-\end{enumerate}
-
-One can therefore in practice learn the SDD of the driven system states via $G_T$ with enough data thanks to the USP/UAP. This enables us to do at least two things: 
-\vspace{-8mm}
-\begin{itemize}
-\item Forecast  $x_{n+1},x_{n+2}, \ldots$ via iterates of $G_T$ (if $G_T$ can be learnt).
-\item Forecast future values of $u_n$ since $x_n$ and $x_{n+1}$ determine $u_n$ since $g$ is SI-invertible. 
-\end{itemize}
-
-Finally, we note that although $G_T$ exists with SI-invertibility, we need the USP as well(see \ref{fig_pictorialSummary} below for a visual illustration).
-If not, the driven states would have to be running for all time since they would not have forgotten their past states.
-
-\begin{figure}[ht]
-  \includegraphics[scale=0.3]{_summarypictorial.eps}
-  \centering
-  \captionof{figure}{Pictorial summary of results obtained up to now and how they overlap. Reproduced from \textbf{(Cite SIG-25May)} }
-  \label{fig:fig_pictorialSummary}
-\end{figure}
-
-\section{A discussion of $G_T$ }
-
-In the above sections, we established the map $G_T$ describing the SDD of a driven system.
-By establishing the existence of this map, we’ve essentially embedded the attractor $U$ (recall from \ref{attractor_U}) into the higher dimensional space $X\times{X}$.
-In layman’s terms, this ensures that there is more “dimensional room” for the underlying system’s underlying dynamics to “move”.
-Since the dynamics aren’t as “squashed”, we might therefore hope that the dynamics of $G_T$ are in some sense simpler than that of $T$.
-(Taken note of the fact that $G_T$ is a homeomorphism even when $T$ is just continuous)
-
-In \cite{manjunath2021universal} it is illustrated in an empirical fashion in that the map $G_T$ describes dynamics which are less functionally complex than that of $T$ or of the map $\Phi_{2d,\theta}$.
-This is done by implementing a Recurrent Neural Network (RNN), which is later mentiond in Chapter~\ref{ch5}.
-\ednote{B: Should I add in that we'll be using Pearson coefficient?
-Not necessary.
-Please check all the cross citation.
-You need to explicitly mention that it is a Chapter or else only a number would appear}
-
-We opt to learn $G_T$ in an indirect manner by defining a new map $\Gamma:(x_{n-1},x_n)\mapsto{u_n}$. It will follow immediately from $G_T$’s existence that $\Gamma$ also exists.
-
-The reasons for taking this roundabout approach remain to be discussed in Section~\ref{subs_LearnGamma}, but a pat answer may immediately be given: When $\Gamma$ has been learnt, the system can be driven autonomously and then $G_T$ is known anyway.
-We formalise this with a theorem.
-
-\begin{Theorem}
-  When $x_n$ and $x_{n-1}$ are successive points on a solution obtained for input-orbit of $T$ $\{u_n\}$, then the map $\Gamma: X\times{X}\to{U}$ defined by $(x_{n-1},x_n)\mapsto{u_n}$ exists whenever $G_T$ exists
-\end{Theorem}
-{\bf Proof.}  See \cite[Th. 3c]{manjunath2021universal}.
-
-Projection mappings $\pi_i$ are defined in the traditional meaning where a $k$-dimensional vector is projected to it's $i^{th}$ component such that $$\pi_i:(a_1,a_2, \ldotp, a_{k-1}, a_k)\to{a_i}$$
-
-The graph in equation \ref{fig:comm_H_CET} is then extended as below:
-
-\begin{equation}  \label{Scomm_imp}
-  %    \[ 
-      \psset{arrows=->, arrowinset=0.25, linewidth=0.6pt, nodesep=3pt, labelsep=2pt, rowsep=1.3cm, colsep = 1.1cm, shortput =tablr}
-   \everypsbox{\scriptstyle}
-   \begin{psmatrix}
-   \widehat{U}_T & \widehat{U}_T\\%
-   Y_T & Y_T\\
-   & \textcolor{red}{U \times X}
-   %%%
-  %  \ncline{1,1}{1,2}^{\widehat{T}} \ncline{1,1}{2,1} <{H_2}
-  %  \ncline{1,2}{2,2} > {H_2}
-  %  \ncline{2,1}{2,2}^{G_T}
-  %  \psset{linecolor=red}
-  %  \textcolor{red}{\ncline{2,1}{3,2}<{(\Gamma,\pi_2)}}
-  %  \ncline{3,2}{2,2}>{\textcolor{red}{(\pi_2,g)}}.
-   \end{psmatrix}
-  % \]
-  \end{equation}
-
-The problem finally simplifies to the issue of learning the map $\Gamma$ and combining this with the projection mapping $\pi_2$ and the function $g$, which will be known.
-A final set of equations is obtained - equations that have been entirely constructed from data.
-\begin{eqnarray}\label{eqns_from_data}
-	u_{k+1} &=& \pi_1 \circ (\Gamma, \pi_2) \circ (\pi_2,g) (u_k,x_k) \label{Seqn_u}\\
-	x_{k+1} &=& \pi_2 \circ (\Gamma, \pi_2) \circ (\pi_2,g) (u_k,x_k). \label{Seqn_x}
-\end{eqnarray}
-
-\section{Advantages of learning $\Gamma$} \label{subs_LearnGamma}
-
-One may immediately ask why we opt to take such a roundabout route; why not just learn the map $G_T$ from the get-go?
-On the surface, this seems to be an arbitrary decision path with no real reasoning, so we take a pause again and discuss the motivation for learning Gamma.
-There are a number of distinct advantages.
-
-In the first place, learning $\Gamma$ saves computational resources.
-This is due to the fact that the input $u_n$ lies in a lower-dimensional subspace in comparison to the high-dimensional space $X\times{X}$.
-
-%In practice, if the input is of a  lower dimension, one may easily embed it into the space $X\times{X}$ by padding the vector $u_n$ with zeroes.
-
-Secondly, the function $\Gamma$ is known to be stable.
-Learning $\Gamma$ makes use of equations \ref{eqns_from_data} which in turn employs the driven system $g$ possessing the USP, and offers distinct advantages with regards to stability in the presence of perturbation.
-It is desirable that inputs that differ only slightly would have only ‘slight’ effects on the output of the system.
-Moreover, one would hope to prevent numerical errors originating from input and measurement noise (see~\cite[Th.
-5]{manjunath2021universal}).
-As the data fed into the system is a function of both the input given and the parameters of the system itself (i.e.
-the value a, $\alpha$ in $\tanh$), we exploit both input-related and parameter-related stability.
-Both these notions are defined by way of the continuity of an encoding map of an input and design parameter respectively in \cite{manjunath2020stability}.
-
-Informally, we may consider input-related stability as concerned with the question of whether or not small variations in input would result in small responses.   
-If two inputs $\bar{u}$ and $\bar{u}_{noisy}$ are close in the product topology, then their tails could differ greatly due to the metric that generates the product topology has insensitivity to the differences in sequence's tails.
- However if the system has the USP then the function $h$ is continuous and so $h(\bar{u}), h(\bar{u}_{noisy})$ remain close-by for small levels of noise maintaining a measure of robustness to input and measurement noise.\cite{manjunath2021universal}.
-In a similar fashion, parameter-related stability is obtained thanks to a result \cite[Lemma 3.2]{manjunath2020stability} relating this form of stability with the ESP (which is here equivalent to the USP).
-
-% Secondly,  due to $g$ possessing the USP, both input- and parameter-related stability~\cite{manjunath2020stability} is obtained, which in turn prevents numerical errors that caused by input and measurement noise (see~\cite[Th.
-5]{manjunath2021universal}).
-
-Thirdly, learning a chaotic map on a set(here $Y_T$) with even a small measure of noise  means that the learnt system accepts arguments out of the set (Problems similar to Takens, cf.
-Section~\ref{sect_takenslimits}) \ednote{Please insert the right label}.
-Suppose iterates of  $(x_{n-1},x_{n})$ under the map $G_T$ is used to forecast two successive values of a solution of $g$ contained in $Y_T$, then these iterates would be contained in $Y_T$.
-However, in practice we are confronted with noise.
-Due to noise at some moment, an iterate could slip beyond $Y_T$ and into an
-expansion $Y_T^+$, for example a closed set that contains all the iterates of $(x_{n-1},x_{n})$ under the map $G_T^+$, where $G_T^+$ acts on $Y_T^+$.
-
-Now since $Y_T$ is not guaranteed to be an attractor of $G_T^+$ this could lead to errors. To see this, consider Fig.~\ref{fig:YtGtFailure}.
-Thus if one learns $G_T$ directly, we are in danger of replicating the problems arising for Takens.
-
-\begin{figure}[ht]\includegraphics[scale=0.25]{_YTerrors.eps}
-  \centering
-  \captionof{figure}{Diagram of potential errors developing when $G_T$ is learnt. (Figure from \cite{Supp}}
-  \label{fig:YtGtFailure}
-\end{figure}
-
-To circumvent this: if a driven system with a saturation function \footnote{a function $f(t)$ that is monotonic and approaches a constant as $|t|\to \infty$, for e.g., $\tanh(t)$ on $\mathbb{R}$} is used, we can ensure the existence of an attractor $A$ of the map that implements
-$(\pi_2,g) \circ (\Gamma,\pi_2)$ (printed in the red in ...\ednote{Please cite the commutativity diagram on the previous page}) on 
- which in practice turns out to be defined on $Y_T^+$ has an attractor $A$  containing $Y_T$.
-\ednote{B: A little vague.
-Are we saying that an attractor of a map implementing $G_T$ is also the one that contains $Y_T^+$?
-M: Please see now..
-there was typo..
-it was A containing $Y_T$.
-I have expanded it now}
-By restricting the dynamics to not spill beyond $A$, prevents large errors from occuring.  
-More details on the global dissipativity of $g$, as it is termed here, may be found in~\cite{Supp}.
-
-\begin{figure}[ht]
-\includegraphics[scale=0.25]{_autovsdriven.eps}
-  \centering
-  \captionof{figure}{Illustrating empirically the importance of global dissipativity.
-Plotted below are the three principal components of randomly initiated points(blue) of a learnt version of $G_T$ versus the trajectory $(x_{n-1}, x_n)$(red) of driven states obtained from the Fractal Dream Attractor's data~\ref{Thomas_Attractor}).
-}
-  \label{fig:learningFailure}
-\end{figure}
-
-In effect, global dissipativity prevents large numerical errors due to the noise in input data from arising (see Fig.~\ref{fig:learningFailure}). 
-If a system is not globally dissipative, neglibly small errors could lead to major inaccuracy which thus induce predictions to utterly fail.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\chapter{Implementation} \label{ch5}
-
-In this chapter we turn to the implementation of the function $G_T$ by discussing a method to learn the function $\Gamma$.
-We apply our methodology to some chaotic attractors.
-Most notably, we consider the reconstruction of the chaotic attractor of the double pendulum - an  attractor that has not been successfully reconstructed because of its intricate structure.
-In fact the previous data-driven approaches also have not been successful in building a model from the double pendulum.
-
-%We then indicate where this implementation differs from previous work done by (\cite{manjunath2021universal}) and empirically show that the SDD of $G_T$ are less functionally complex than $\Phi_{k, \theta}$ or $T$ by considering the Pearson correlation coefficient.
-Finally, we present and discuss some numerical results for a focus system of this project: the double pendulum.
-We also consider other chaotic attractors.
-
-\section{Implementing $G_T$}
-The map $G_T$ is implemented by applying the functions $\Gamma$, $g$ and the projection function $\pi_2$ in the relevant order as described by the equations~\ref{eqns_from_data}.
-To reach this point, one must first learn $\Gamma$.
-After first normalising the data to ensure that it has zero mean value and lies within the range $[-1,1]$ and required zero-padding(see paragraph below for further discussion), we continue by driving the system with external input through the function $g$.
-
-From \eqref{eqn_driving} we know that $g(u,x) = (1-a) x + a \tanh(Au + \alpha B x)$. Recall that $g$ is SI-invertible. We refer to $A$ as the input matrix and $B$ the reservoir matrix.
-
-We make use of a Recurrent Neural Network (RNN) in our implementation to project the data on to a higher dimensional space, but we do not train the RNN in the familiar ways in the literature but as in \cite{manjunath2021universal}.
-A few words about the RNNs: RNNs are widely used in the literature for temporal processing of data as a type of deep learning technique.
-They take the form of equations considered to be discrete-time version of continuous dynamical systems modelling a neural network of neurons \cite{jaeger2001echo}.
-
-We implement a feedforward network, (a network with no memory, unlike the RNN) required to learn $\Gamma$ in the \emph{Python} programming language and making use of the \emph{Scipy} library, alongside the \emph{Keras} package in the \emph{Tensorflow} library.
-
-If the RNN we will be using in the next step has $N$ neurons, then $A$ and $B$ are each $N\times{N}$ matrices.
-Inputs $u_n$ of dimension $K < N$ are padded by zeroes (as remarked earlier) to embed the input in a space of dimension $N$.
-As we do not need the entire left-infinite history of input, an arbitrary initialisation of the system with initial values followed by the driving of the system with the function $g$ means that the generated values will approach the actual solution elements in a uniform manner thanks to the UAP that we had discussed in the preceding chapter.
-
-Thirdly we prepare the network.
-Utilising a network $N$ neurons (the same value $N$ as in the dimensions of the matrices $A,B$).
-Learning of $\Gamma$ is done using the Adam Optimiser where the means square error (MSE) loss function is optimised.
-The system is trained several times (in our case 4) with incrementally smaller learning rates on each run (0.001 and thereafter divided by 10 at each run).
-Additional parameters such as the number of network-layers, neurons per layer and training-length are presented case-by-case below.
-
-We then reach the prediction step. Making use of $g$'s SI-invertibility and $g$ as defined in \eqref{eqn_driving}, we have the following expression for $u_n$ replicated from~\eqref{eqn_SI_RNN}.
-\begin{equation*}
-  u_n := A^{-1}\bigg(\overline{\tanh}^{^{-1}}\frac{1}{a}\Big(x_{n+1}-(1-a)x_n\Big) \bigg) - \alpha B x_n
-\end{equation*}
-
-The two model equations for learning from data~\eqref{eqns_from_data} allows us to determine what the state $x_{n+1}$ at time $n+1$ will be.
-
-%We specifically remind ourselves of \[x_{k+1}=\pi_2 \circ (\Gamma, \pi_2) \circ (\pi_2,g) (u_k,x_k)\]
-
-\section{Delay Coordinates}
-
-We also consider a more general problem of learning an attractor of dynamical system when only observations of an orbit are given.
-Explicitly, if $(W,T)$ is a dynamical system with dynamics generated by $w_{n+1}=Tw_n$, and if $\theta:W \to \mathbb{R}$ is an observable, then the task is to learn a dynamical system that is topologically conjugate to $(W,T)$ and predict $\theta(w_{m+1}),\theta(w_{m+2}),...$ using the data $\theta(w_{0}),\theta(w_{1}),\ldots,\theta(w_{m})$.
-So how do we do it?
-Suppose the input generated from the delay-coordinate map $\Phi_{\theta,2d}(\theta(w_{n})) := (\theta(w_{n-2d}),\ldots,\theta(w_{n-1}),\theta(w_{n}))$ is such that Takens delay embedding theorem (see Theorem~\ref{Thm_Takens}) holds, in which case, there exists a homeomorphism  $F_\theta: \Phi_{\theta,2d}(\theta(w_{n})) \mapsto \Phi_{\theta,2d}(\theta(w_{n+1}))$.
-Then if we feed  the input values $u_n := \Phi_{\theta,2d}(\theta(w_{n}))$ to a driven system $g$ that is SI-invertible and has USP, the induced system $(Y_F,G_F)$ would be topologically conjugate to the inverse-limit space of
-$(\Phi_{\theta,2d}(W), F_\theta)$ as in Theorem~\ref{Thm_CET}, and hence one could forecast $u_n,u_{n+1},\ldots$.
-In other words, one could forecast  the values $\theta(w_n), \theta(w_{n+1}),\ldots$ by feeding these observations as delay coordinates.
-The advantage of feeding delay-coordinates to a driven system is that the embedding is stable in the sense of global dissipativity that we have described in the previous chapter, i.e., $Y_F$ is made to contained in an attractor of a mapping induced on $Y_F$ that is induced by $g$.
-
-We note that the twin equations together~\ref{eqns_from_data} generate a model from data.
-For a chaotic system, a small amount of noise due either to the input-noise or computational error would cause the solutions from two accurate models with the same initial conditions to diverge.
-However, statistically, they would behave in the same way in the sense that the visiting frequency or the average time spent by an orbit in any region of the phase space of the dynamical system is identical for almost all initial conditions (almost all is specified as a measure), a result in ergodic theory (\url{https://en.wikipedia.org/wiki/Ergodic_theory} a topic beyond the scope of this thesis.
-\ednote{B: I'll need to recap this sentence with you before presenting as it's new to me.
-M: I rewrote it.} To illustrate that we plot the densities of the trajectories of the actual and predicted solutions.
-
-In our implementation of the input and reservoir  matrices $A$ and $B$ in the RNN are sparse matrices.
-We discuss three specifics relating to this concept.
-Firstly, the matrices are assigned a variable density in terms of the amount of non-zero entries and are scaled to have values within the range $[-1,1]$.
-For this, we make use of \textit{Python's SciPy} library and specifically the \textit{Sparse, LinAlg} modules.
-In our experiments, we generally opted for a density ratio of 10\%-25\%.
-Secondly, we opted to enforce the requirement that both matrices be well-conditioned, i.e.
-that they have a condition number of \~1.
-The condition number of a matrix $M$ (denoted $cond(M)$) is the product of the matrix norms of the matrix $M$ and its inverse $M^{-1}$ and is a measure of the sensitivity of output of a matrix multiplication when an input is changed.
-\textbf{(Cite.)} By ensuring a condition number close to 1, we also ensure that our driven function $g$, depending on the matrices $A$ and $B$, will not be sensitive to adjustments in the input.
-In so doing we then also guarantee another form stability.
-\ednote{B: Not complete enough.
-Edit further and cite M: Please feel free to cite a wikipedia page}
-
-Lastly, we ensure the matrix $B$ has unit spectral radius.
-This is necessary due to the Causal Embedding Theorem~\ref{Thm_GT_Exists} which can then guarantee the existence of semi-conjugacy between the systems $(Y_T, G_T)$ and $(\widehat{U}_T, \widehat{T})$.
-In fact, the theorem actually requires that $\alpha{B}$ have spectral radius $<1$, so we impose the restriction $\alpha<1$ and consider only the matrix $B$'s spectral radius from thereon.
-
-Note that to learn $\Gamma: (x_{n-1},x_{n}) \mapsto u_n$ through our feedforward network, we could learn the principal components of $(x_{n-1},x_{n})$ as well. 
-We employ this since it provides slightly better results empirically but is not essential.
-
-Principal components are used only for a more efficient state representation to possibly reduce learning errors; it is not for a lossy approximation since we use all principal components \textbf{Cite.}.
-More explicitly, denote $X_{1:N}$ as the matrix with the first $N$ states of the network data represented by row vectors.
-If $X_{1:N}=U\Sigma P^T$ here represents the singular value decomposition ($U$ and $V$ are unitary matrices and $\Sigma$ is a diagonal matrix) of $X_{1:N}$ and we refer to ...\ednote{B: Todo.
-Refer or Cite M: Please add a Wiki citation in the bib}, we then denote the principal component matrix $P$, and the principal components are given by
- $Z_{1:N}=X_{1:N}P.$
-These principal components are then used to train a feedforward neural network. Denote the row vectors of $Z_{1:N}$ by 
-$z_i^T, i=1,2,\cdots,N$ and the neural network by $NN$, the network is subsequently trained to learn an approximation of the map
-$NN: (z_{n-1},z_n) \mapsto  u_n.
-$
-The learnt approximation of $NN$ can be used to approximate $\Gamma$ since 
-$$
-NN\left( \begin{bmatrix} 
-P^Tx_{n-1} \\
-P^Tx_n
-\end{bmatrix}
-\right) = NN \circ 
-\begin{bmatrix}
-P^T & 0 \\
-0 & P^T 
-\end{bmatrix}\begin{bmatrix}
-x_{n-1}\\
-x_n
-\end{bmatrix} = u_n = \Gamma(x_{n-1},x_n).
-$$
-
-\section{Experimental Results}
-In this section we showcase numerical results pertaining to 3 attractors that were simulated using the methodology described.
-We first consider a simple physical system that exhibits complicated dynamics.
-\subsection{Double Pendulum}
-
-The Double Pendulum consists of two pendulums fastened to one another such that the system moves as a whole.
-The first (or higher pendulum) is attached to a fixed point while the second is attached to the endpoint of the first.
-(See below)
-This system is a classical example of a system exhibiting chaos. \textbf{(Citation.)} Consider again~\ref{fig:dp_sdic}.
-The system has four variables: angle and angular velocity of each pendulum rod. An ideal system is assumed where the rods are massless and no friction in the pivots are considered.
-\begin{figure}[ht]
-  \includegraphics[scale=0.35]{_dp_setup.eps}
-  \centering
-  \captionof{figure}{Pictorial representation of the setup of a DP. Image taken from~\cite{DPSetup}}
-  \label{fig:dp_setup}
-\end{figure}
-
-The scalar equations of motion are derived using Newtonian physics and are reproduced from~\cite{DPFormulas} in~\ref{eqns_dp}.
-Each pendulum component has an associated mass of the head, $m$, and length of the rod, $L_i$.
-Acceleration due to gravity is denoted by $g$ and initial velocities by $\dot{\theta}$
-\begin{eqnarray}\label{eqns_dp}
-  \ddot{\theta_{1}}  = \frac{-g(2m_1+m_2)\sin{\theta_1} - m_2g\sin(\theta_1-2\theta_2) - 2\sin(\theta_1-\theta_2)m_2({\dot{\theta_{2}}}^{2}L_2 + {\theta_{1}'}L_1\cos(\theta_1-\theta_2))} {L_1(2m_1 + m_2 -m_2\cos(2\theta_1 - 2\theta_2))}
-  \\
-  \ddot{\theta_{2}} = \frac{2\sin(\theta_1-\theta_2)(\theta_{1}'^{2}L_1(m_1+m_2) + g(m_1+m_2)\cos\theta_1 + \dot{\theta_{2}}^{2}L_2m_2\cos(\theta_1-\theta_2))}{L_2(2m_1 + m_2 -m_2\cos(2\theta_1 - 2\theta_2))}
-\end{eqnarray}
-
-The second-order system is converted to a first-order ODE system and solved using the Runge-Kutta numerical method.  
-Here we chose the values $m_1=2.0$, $m_2=2.0$, $L_1=1.5$ and $m_2=1.5$ with initial velocities $v_1=1.5, v_2=-2$ for the respective pendulum heads.
-Undamped DP data-sets were generated and used in this project due to the requirement that the map $T$ be surjective that we must not forget. 
-An example of this would be the case of a DP exhibiting energy-loss (i.e.
-where the map $T$ is not surjective).
-In such a case, the attractor is a single point in the phase space when the double pendulum comes to rest.
-It must be mentioned once again that if we use data outside the attractor, learning is not reliable. To see this, consder the graphs ~\ref{fig:damped_pendulum} and ~\ref{fig:dp_notsurjective}.
-
-\begin{figure}[ht]
-  \includegraphics[scale=0.4]{_dp_dying.eps}
-  \centering
-\caption{$x$- and $y$-coordinates of the two pendulum heads plotted for a dataset that exhibits dying oscillations.
-The map generating the (four-dimensional data) for a damped pendulum is not surjective.
-Data obtained from chaotic double pendulum dataset generated from videos takes of experimental pendulums~\cite{asseman2018learning}}
-\label{fig:damped_pendulum}
-\end{figure}
-
-\begin{figure}[ht]
-  \includegraphics[scale=0.3]{_dpfail_nonsurj.eps}
-  \centering
-\caption{Failure to forecast or learn a map which is not surjective: Evolution of the $y$-coordinate of the bottommost pendulum plotted here for 15000 timestep.
-Plotted here the predicted value(red) versus the actual value(blue).
-The plots overlap for the first 10000 steps, the duration of training to learn the map $\Gamma$.
-Afterwards, the trajectories start diverging and it the predicted value no longer accurately represents the complexity or energy-loss of the actual data.
-}
-\label{fig:dp_notsurjective}
-\end{figure}
-
-Having spent ample time to discuss the challenges, we now change tone so as to satisfy the optimist and present a successful run of the DP-data to predict the future movement of the pendulum.
-
-The system was driven for 15500 steps, of which the first 500 were immediately discarded to simulate the network 'forgetting'.
-The driving function $g$ had input and reservoir matrices were implemented as sparse matrices filled to 20\%.
-Parameter-values of $\alpha=0.99$ and $a=0.5$ were used.
-The remaining 15000 datapoints were used to train a network with 128 neurons. 
-The network was initialised to contain 16 hidden layers each with a layer dimension of 64, activated with the tanh activation function.
-Training was accomplished using 256 epochs with batch sizes 128 and a delay of 1.
-See~\ref{fig:dp_success_traj} for results pertaining to the trajectory taken by the DP's components and then~\ref{fig:dp_success_density} to witness associated densities of predicted vs.
-actual values.
-
-\begin{figure}[ht]
-  \centering
-  \includegraphics[width=0.95\linewidth]{Graphs/_dp_success_4coords_traj.pdf} 
-  \captionof{figure}{Predicted(red) and actual(blue) trajectories of the $x$- and $y$-coordinates of each of the pendulum heads.
-From top to bottom are as follows: $x$-coordinate of topmost pendulum head, $y$-coordinate of topmost pendulum head, $x$-coordinate of lower pendulum head, $y$-coordinate of lower pendulum head.
-  These graphs were constructed by predicting the DP 1000 timesteps into the future and in so doing illustrating the long-term consistency and accuracy of the learnt system.
-Here we lock onto the near-exact trajectory for about 300 timesteps and stay close up until about 600 seconds.
-}
-  \label{fig:dp_success_traj}
-\end{figure}
-\begin{figure}[ht]
-  \centering
-  \includegraphics[width=\linewidth]{_dp_success_4coords_hist.eps}
-  \captionof{figure}{Densities of predicted and actual values of $x$- and $y$-coordinate of pendulum heads.}
-  \label{fig:dp_success_density}
- \end{figure}
-
-We now consider an experiment where noise is added to the system and find it to be incredibly robust.
-We employ the same parameters as before and add noise from a normal-distribution with mean zero and standard deviation equal to 0.1.
-Once again consider the actual and predicted trajectories for the noisy DP dataset in Figure~\ref{fig:noisydp_success_traj}.
-Next we consider the densities of the predicted and actual values in Figure~\ref{fig:noisydp_success_density} and observe that they display highly similar distributions.
-
-\begin{figure}[ht]
-  \centering
-  \includegraphics[width=0.95\linewidth]{_dp_noise_FigCD_2030.eps} 
-  \captionof{figure}{Predicted(red) and actual(blue) trajectories of the $x$- and $y$-coordinates of each of the pendulum heads with noise added from a normal distribution with mean zero and standard deviation 0.1.
-  These graphs were constructed by predicting the DP 1000 timesteps into the future and in so doing illustrating the long-term consistency and accuracy of the learnt system.
-Here we lock onto the near-exact trajectory for about 250 timesteps and stay close up until about 400 seconds.
-}
-  \label{fig:noisydp_success_traj}
-\end{figure}
-\begin{figure}[ht]
-  \centering
-  \includegraphics[width=\linewidth]{_dp_noise_2B_2030.eps}
-  \captionof{figure}{Densities of predicted and actual values of x-coordinate of first pendulum head.}
-  \label{fig:noisydp_success_density}
- \end{figure}
-
- Lastly, an run was done where only a single coordinate (the $y$-coordinate of the lower pendulum head) was fed into the system and we can illustrate through the obtained success that the coordinate contains all the information in its past states necessary to forecast its future evolution.
- \textbf{Insert graphs here. Had to quickly change something here.}
-
-\subsection{Additional Attractors: Clifford \& Thomas}
-We follow the methodology presented in \cite{manjunath2021universal} but opt to consider attractors different from the Lorenz system and H\'enon map in this project.
-
-\subsubsection{Thomas' Cyclically Symmetric Strange Attractor}\label{Thomas_Attractor}
-Thomas' Cyclically Symmetric Strange Attractor, a 3D attractor proposed in 1999 by Ren\'e Thomas in \cite{ThomasAttractor}, is described by a set of three equations which is distinctly reminiscent of the Lorenz system, another deterministic three-equation model exhibiting chaotic behaviour.It has a single parameter $\beta$ and has been shown to transition to chaotic behavious when $\beta<0.208186$~\cite{Thomas_BetaParameter}.
-\begin{eqnarray}\label{eqns_thomas}
-  x_{n+1} = \sin(y_n) - \beta{x_n} \\
-  y_{n+1} = \sin(z_n) - \beta{y_n} \\
-  z_{n+1} = \sin(x_n) - \beta{z_n}
-\end{eqnarray}
-
-The behaviour of the system with a $\beta$-parameter value of 0.1056 was simulated by scaling it to fit inside the interval $[-1,1]$.
-Two cases were considered: noise-free and the one perturbed by noise.
-Noise was generated from a normal distribution with mean 0 and standard deviation 0.05, which translates approximately to * dB \ednote{Please calculate dB as signal to noise ratio }
- A sequence of 3000 observations in both the clear and noisy data-sets respectively were fed into the system with 300 entries discarded to simulate the network's loss of memory.
-The next 9000 steps were predicted and compared with true data.
-
-\begin{figure}[ht]
-  \centering
-  \minipage{0.5\linewidth}
-    \centering
-    \includegraphics[width=0.8\linewidth]{_thomas_noisy_0.1056.eps}
-  \endminipage\hfill
-  \minipage{0.5\linewidth}
-    \centering    \includegraphics[width=0.8\linewidth]{_thomas_clean_0.1056.eps}
-  \endminipage
-  \captionof{figure}{True(red) and predicted(blue) trajectories for the Thomas attractor with parameter value $\beta=0.1056$.
-To the left: Noise added from normal distribution with mean zero and standard deviation 0.05.
-Right: No noise added to dataset.}
-\end{figure}
-
-The RNN was initialised as follows: the dimension of the network is set equal to 500 neurons and 12 hidden layers of dimension 64 each were initialised.
-Training is realised through 150 epochs of batch sizes of 128 each; the map is learnt via the Adam Optimiser using finer and finer learning rates (0.001 and then divided by 10 at every iteration).
-
-\begin{figure}[ht]
-  \centering
-  \includegraphics[scale=0.35]{_Thomas_1.eps}\caption*{Predicted trajectories of the $x$- and $y$-coordinates for the Thomas attractor with parameter-value $\beta=0.1056$ demonstrate empirically the ability to predict the evolution of the trajectory for the next an estimated 100 timesteps into the future near-exactly.
-Here no noise was added.}
-  \includegraphics[scale=0.5]{_Thomas_3.eps}\caption*{Represented here the learnt (blue) and actual (red) densities of the first coordinate ($x$-coordinate) of the dataset for the Cyclically Symmetric Attractor with parameter-value $\beta=0.1056$.}
-\end{figure}
-
-\subsubsection{Fractal Dream Attractor}
-
-Another 2-equation discrete-time map named the Fractal Dream Attractor, or more commonly known as the Pickover Map (first discovered by Clifford A.
-Pickover and discussed in his fascinating book "Chaos in Wonderland"\cite{PickoverChaos}), was also considered.
-\begin{eqnarray}\label{eqns_clifford}
-  {x_{n+1}=\sin(ay_n) + c\cos(ax_n)} \\
-  {y_{n+1}=\sin(bx_n)+d\cos(by_n)}
-\end{eqnarray}
-
-\begin{figure}[ht]
-  \centering
-  \includegraphics[width=1.0\textwidth,left]{_Clifford_1_nonoise.eps}
-  \caption*{These graphs were constructed by predicting the Clifford system 1000 steps into the future and in so doing illustrating the long-term consistency and accuracy of the learnt system.
-As perceived here, we are able to lock on to the trajectory of the Clifford map almost exactly for the first 25 steps.}
-  \minipage{0.5\textwidth}
-      \centering
-      \includegraphics[width=\linewidth]{_Clifford_2_nonoise.eps}
-  \endminipage\hfill
-  \minipage{0.5\textwidth}
-    \centering
-    \includegraphics[width=\linewidth]{_Clifford_3_nonoise.eps}
-  \endminipage
-  \caption{The bottom row, left, shows a plot of the true attractor(red) and the predicted values(blue).
-Bottom row, right, shows the distribution of the first coordinate of the learnt system (blue) and the actual Clifford system (red).
-}
-  \label{fig:Clifford}
-\end{figure}
-
-The behaviour of the system with a parameter values of $x_0=0$, $y_0=0$, $a = -1.7$,  $b = 1.8$, $c = -1.9$ and $d = -0.4$ was examined.
-Data was scaled to fit inside the interval $[-1,1]$. Only the noise-free case was investigated. 
- A sequence of 25000 observationswas fed into the system with 15000 entries discarded to simulate the network's loss of memory. The next 1000 steps were predicted and compared with true data.
-The RNN was initialised exactly as for Thomas' Cyclically Symmetric Strange attractor: the dimension of the network is set equal to 500 neurons and 12 hidden layers of dimension 64 each were initialised.
-Training is realised through 150 epochs of batch sizes of 128 each; the map is learnt via the Adam Optimiser using finer and finer learning rates (0.001 and then divided by 10 at every iteration).
-
-\section{Functional Complexity Reduction in Noninvertible maps}
-
-% Besides being able to learn the action of the Koopman operator exactly, our setup has other advantages over selecting or optimizing the choice of observables as in the Extended Dynamic Mode Decomposition (EDMD) algorithm (e.g., [21, 22]).
-We can alter the functional complexity of G T by changing the number of observables or equivalently by altering the dimension of X – in the case of a RNN implementation of g, by simply changing the number of neurons in the network.
-Empirically, increasing the dimension of X is found to increase the linear relationship (or intuitively reduces the functional complexity of G T ) that is measured as a generalization of the Pearson correlation coefficient to random vectors (e.g., [44]) between (x_{n−1} , x_{n}) and G_T (x_{n−1} , x_n), and such numerical evidence is tabulated in [11, Table 1].
-In contrast, expanding the set of observables while using EDMD to learn a map with a lower functional complexity is a highly involved task, not least because so much that one often does not know how to guess a new observable, as well as the fact that adding observables does not necessarily retain the invariance of the span of the observables.
-
-% Empirical evidence shows that a linearity measure like a Pearson correlation coefficient points at reduced functional complexity while learning the dynamics in the state space of the driven system.
-
-% Empirically (see Table ??), increasing the dimension of X is found to increase the linear relationship
-% (or intuitively reduce the functional complexity of G T ) that is measured as a generalization of the Pearson correlation coefficient to random vectors (e.g., [?]) between (x n−1 , x n ) and G T (x n−1 , x n ).
-If Σ a and Σ b denotes the covariance matrices of the vectors (x n−1 , x n ) and G T (x n-1 , x n ) respectively, and if Σ ab denotes the covariance matrix between the vectors (x n-1 , x_n ) and G_T (x n-1 , x_n ), then the multidimensional 18correlation coefficient is computed by using the traces of these matrices:
-% \[\pho = \frac{tr(\sum_{ab})}{tr(\sqrt{\sum_{a}\sum_{b}})}\]
-
-% This multidimensional correlation coefficient satisfies most of the well-known properties of the one-dimensional Pearson coefficient.
-In particular, ρ = ±1 if and only Y = AX + b for some invertible matrix A and vector b.
-Hence ρ, and in practice, the estimator ρ̂, can be used to measure the linear relationship between two random vectors of the same dimension and thus serves as an indicator of functional complexity.
-
-% We remark that the estimation of such linear relationship in (see Table ??) obtained by sample correlations alone is justified whenever {u n } is a realization of an ergodic process since then when g has the USP, any solution {x n } is also a realization of an ergodic process [?].
-In the ergodic input case the generalized Pearson correlation coefficient (between (x n−1 , x n ) and G T (x n−1 , x n )) is independent of n.
-
-In the case of non-invertible maps, we note that the map $G_T$ is a homemorphism although $T$ is not.
-Further, empirically it is found that $G_T$ is much more linear than $T$ as measured by what is called the multidimensional correlation coefficient $\rho$.
-This coefficient is a generalization of the Pearson correlation coefficient for real-valued random variables to random vectors (e.g., \cite{puccetti2019measuring}).
-We state teh details of computing this coefficient between
-$(x_{n-1},x_n)$ and $G_T(x_{n-1},x_n)$. 
-If $\Sigma_{a}$ and $\Sigma_{b}$ denotes the covariance matrices of the vectors $(x_{n-1},x_n)$ and $G_T(x_{n-1},x_n)$ respectively, and if $\Sigma_{ab}$ denotes the covariance matrix between the vectors $(x_{n-1},x_n)$ and $G_T(x_{n-1},x_n)$, then the multidimensional correlation coefficient is computed by using the traces of these matrices:
-\[
-    \rho= \frac{\text{tr}({\Sigma_{ab}})}{\text{tr}({\sqrt{\Sigma_a\Sigma_b}})}.
-\]
-This multidimensional correlation coefficient satisfies one of the desired and well-known properties of the one-dimensional Pearson coefficient.
-That is, $\rho=\pm 1$ if and only if $Y\overset{d}{=}AX+b$ for some invertible matrix $A$ and vector $b$.
-Hence $\rho$, and in practice can be used  to measure the linear relationship between two random vectors of the same dimension and thus serves as an indicator of functional complexity.
-For the non-invertible system defined by the Clifford attractor, we indicate the difference by which $\rho$ changes for $G_T$ in Table~\ref{tbl_attractorsPearson}.
-Following \cite{manjunath2021universal}, we say that this multidimensional correlation coefficients $\rho$ to evidence a general reduction in the functional complexity of the map $G_T$ emerging due to a RNN.
-In the
-Table ~\ref{tbl_attractorsPearson}, the second column corresponds to the amount of artificial neurons (the dimension of $X$) in the implemented RNN.
-
-\begin{center}
-  \begin{table} 
-      \scalebox{0.75}
-      {\begin{tabular}{|c|c| c c |} 
-          \hline
-          Input & Dimension of $X$ & $u_n$ vs $u_{n+1}$ 
-          & $\begin{bmatrix}
-              x_{n-1}\\
-              x_n
-          \end{bmatrix}$ vs $\begin{bmatrix}
-              x_n\\
-              x_{n+1}
-          \end{bmatrix}$ \\
-          \hline      
-          % \hline
-          % $(w_n)$ is the Lorenz states, sampled every 0.1 timestamps. & & &\\
-          % \multirow{3}{*}{$u_n=\frac{1}{100}w_n$} & 10 &0.9311  & 0.9731\\
-          %     & 100 &  & 0.9934\\
-          %     & 1000 & & 0.9930\\              
-          % \multirow{3}{*}{$u_n=\frac{1}{10}(\sin(0.1w_{n,x})+\sin(0.1w_{n,y})+\sin(0.1w_{n,z}))$} 
-          %     & 10 &0.8401  & 0.9392\\
-          %     & 100 & & 0.9616\\
-          %     & 1000 & & 0.9737\\
-          % \hline
-
-               \hline  
-               $(w_n)$ comes from the Clifford Attractor \\ (for $a = -1.7; b = 1.8; c = -1.9; d = -0.4$): & & & \\
-        {$w_{n+1}= \begin{bmatrix} \sin(aw_{n,y}) + c\cos(aw_{n,x}) \\ 
-                                                                    \sin(bw_{n,x})+d\cos(bw_{n,y}) \end{bmatrix}$} & & & \\
-          \multirow{3}{*}{$u_n = w_n-\overline{w}$}
-              & 10 & -0.21713 & 0.78719 \\
-              & 100 & &  0.8334 \\
-              & 1000 & & 0.84854 \\
-          % \hline 
-          % $(w_n)$ comes from the Logistic map: & & & \\
-          % {$ w_{n+1}  = 4w_n(1-w_n)$} & & & \\
-          % \multirow{3}{*}{$u_n = w_n - 0.5$} 
-          %     & 10 &-0.0314  & 0.5577\\
-          %     & 100 &  & 0.7150\\
-          %     & 1000 & & 0.7826\\
-          % \hline
-          % \hline
-          % $(w_n)$ comes from the Svensson Attractor \\ (for $a = -1.7; b = 1.8; c = -1.9; d = -0.4$): & & & \\
-          % {$w_{n+1}= \begin{bmatrix} d\sin(aw_{n,x}) - \sin(bw_{n,y}) \\ 
-          %                                                             c\cos(aw_{n,x})+\cos(bw_{n,y}) \end{bmatrix}$} & & & \\
-          %   \multirow{3}{*}{$u_n = w_n-\overline{w}$}
-          %       & 10 & -0.2243 & 0.49698 \\
-          %       & 100 & &  0.63310 \\
-          %       & 1000 & & 0.71151 \\
-           \hline
-                \end{tabular}} %\ednote{http://paulbourke.net/fractals/peterdejong/}
- \label{Table_FC}
- \caption{Following \cite{manjunath2021universal}, we determine the multidimensional correlation coefficients $\rho$ to evidence a general reduction in the functional complexity of the map $G_T$ emerging due to a RNN.
-Rows correspond to distinct dynamical systems considered in the experiments
-      The second column corresponds to the amount of artificial neurons (the dimension of $X$) in the implemented RNN.
-      The last two columns comprise numerical estimates of $\rho$ for the relevant vectors. The linear association between $u_n$ vs $T(u_{n})$, as well as the linear relationship $\begin{bmatrix}
-              x_{n-1}\\
-              x_n
-          \end{bmatrix}$ vs $G_T \left( \begin{bmatrix}
-              x_{n-1}\\
-              x_n
-          \end{bmatrix}\right)$ may be contrasted by the reader. 
-          Finally, the RNN's dimension (from \eqref{Seq_RNN}) is provided to exhibiting that an increase in dimension of the driven system will 
-          typically result in a map $G_T$ with less functional complexity.}.
-      \end{table}\label{tbl_attractorsPearson}
-\end{center}
-
-\chapter{Conclusions}\label{ch6}
+\chapter{Conclusions}~\label{ch6}
 
@@
 % In so doing, we acquire models of high fidelity that provide exceptional forecasting results on chaotic data that exhibit both statistical and topological consistency in the long run.
-
-\bibliographystyle{pnas-new}
-\bibliography{pnas}
-
-\end{document}

### 2022-01-18 projectReport/chapter1.tex
Worked through all tex errors

@@
 We encounter many difficultiesin the process of modelling such systems:
+
 \vspace{-8mm}
@@
 Examples in the literature attempting to forecast data from such systems have tried many different approaches with differing degrees of success. 
-When we have complete access to states of the system, an ordinary differential equation can be obtained from data (\cite{brunton2016discovering, champion2019data} and \cite{small2002modeling,xu2006modeling}) wherein one could approximate the vector field by a library of functions to obtain interpretable models.
+When we have complete access to states of the system, an ordinary differential equation can be obtained from data (\cite{brunton2016discovering, champion2019data} and~\cite{small2002modeling,xu2006modeling}) wherein one could approximate the vector field by a library of functions to obtain interpretable models.
 
@@
 This makes the dynamics in the higher-dimensional space more amenable for approximation by a linear transformation.
-Such methods do not guarantee exact reconstructions for nonlinear models, and in practice provide poor long-term consistency for a large class of chaotic dynamical systems \cite{Supp}.
+Such methods do not guarantee exact reconstructions for nonlinear models, and in practice provide poor long-term consistency for a large class of chaotic dynamical systems~\cite{Supp}.
 
 The non-interpretable models include the delay embedding and the machine learning algorithms.
-For example, one could learn a system conjugate to the underlying system by applying the Takens delay embedding method of delay-coordinates~\cite{takens1981detecting} when one has "good" observations from the system; this learnt system could then be used to forecast the observed data.
+For example, one could learn a system conjugate to the underlying system by applying the Takens delay embedding method of delay-coordinates~\cite{takens1981detecting} when one has `good' observations from the system; this learnt system could then be used to forecast the observed data.
 Takens delay embedding theorem~\cite{takens1981detecting} and its various generalisations (see,~\cite{sauer1991embedology, stark1999delay, gutman2018embedding}) establishes the learnability of a system constructed by concatenating sufficiently large previous time-series observations of a dynamical system into a vector (called delay coordinates).
@@
 One reason for this fragility is that the embedded attractor in the reconstruction space is not always an attractor of the map learnt in the reconstruction space, despite unquestionably being an invariant set.
-When the embedded object is not explicitly known to be an attractor (as explained in Chapter~ \ref{ch3}), it can cause predictions to fail.
+When the embedded object is not explicitly known to be an attractor (as explained in Chapter~\ref{ch3}), it can cause predictions to fail.
 
@@
 
-\emph{By solidifying the mathematical underpinnings of our theory, we hope to guarantee the ability to construct models with predictive power ranging from molecular biology to neuroscience.
-in the near future.
+\emph{By solidifying the mathematical underpinnings of our theory, we hope to guarantee the ability to construct models with predictive power ranging from molecular biology to neuroscience in the near future.
 (We can modify this sentence when the report is completed.)}

### 2022-01-18 projectReport/chapter2.tex
Worked through all tex errors

@@
-This chapter briefly describes a discrete-time dynamical system and what it means to exhibit chaos. We refer to \cite{devaney2018introduction, de2013elements} for more details.
+This chapter briefly describes a discrete-time dynamical system and what it means to exhibit chaos. We refer to~\cite{devaney2018introduction, de2013elements} for more details.
 
@@
 \begin{Definition}
-  [\bf Flow of an Equation] \label{Dfn_Flow}\rm
+  [\bf Flow of an Equation]\label{Dfn_Flow}\rm
   The flow of the equation  $\dot{x} = f(x)$  is defined to be a mapping $\varphi: \mathbb{R}^n \times \mathbb{R} \to \mathbb{R}$ where $\varphi(x_0,t)= x(t)$ for the solution $x(t)$ with $x(0)=x_0$ 
@@
 \end{Definition}
-
 We also define the orbit of a function $T$.
@@
   [\bf Orbit of $T$]\label{Dfn_Orbit}\rm
-  The orbit of $T$ is to be defined the sequence $\bar{u} = \{u_n\}_{n\in \mathbb{Z}}$ obeying the update equation, $u_{n+1}=Tu_n$, $n \in \mathbb{Z}$. 
+  The orbit of $T$ is to be defined the sequence $\bar{u} = {\{u_n\}}_{n\in \mathbb{Z}}$ obeying the update equation, $u_{n+1}=Tu_n$, $n \in \mathbb{Z}$. 
 \end{Definition}
 
-Two examples of invariant sets include a fixed point where $Tu=u$ for  $u\in U$, and a periodic orbit, i.e., a set of iterates $\{u,Tu, T^2u,\ldots,T^pu\}$, where $T^{p+1}u=u$ for some $p\in\mathbb{Z}$.
+Two examples of invariant sets include a fixed point where $Tu=u$ for  $u\in U$, and a periodic orbit, i.e., a set of iterates $\{u,Tu, T^2u,\ldots,T^{pu}\}$, where $T^{p+1}u=u$ for some $p\in\mathbb{Z}$.
 The entire space $U$ could be also be invariant.
@@
 However,  when the space $U$ is compact, all orbits approach an invariant set.
-This is since the set of limit points of the orbit can be shown to be invariant \cite{de2013elements}.
+This is since the set of limit points of the orbit can be shown to be invariant~\cite{de2013elements}.
 So, when $U$ is compact,  the $\omega$-limit set $\omega(u;T)$ of a point $u$ defined to the collection of limit points of the sequence $\{x,Tu,T^2u,\ldots\}$ is nonempty, and $\omega(u;T)$ is invariant.
 
-\begin{Example}
+\begin{Example}\rm
   For the map $Tu=u^2$ defined on $[0,1]$ all orbits lie in the invariant set $\{1\}$ or else would would approach the invariant set $\{0\}$. 
@@
 Invariant sets have various properties. Invariant sets can be attracting or repelling depending on how orbits in their vicinity behave. 
-Recall the examples $Tu =u/2$ and $Tu=2u$ defined on $\mathbb{R}$ (where $\{0\}$ “attracts" orbits) or the example, $Tu=u^2$ defined on $[0,1]$ (where $\{1\}$ “repels" orbits). 
+Recall the examples $Tu =u/2$ and $Tu=2u$ defined on $\mathbb{R}$ (where $\{0\}$ `attracts' orbits) or the example, $Tu=u^2$ defined on $[0,1]$ (where $\{1\}$ `repels' orbits). 
 We are interested in attractive invariant sets since they capture the long-term dynamics as time increases. 
@@
 	\item $A$ is invariant. 
-	\item $A$ is asymptotically stable, i.e., for every $\epsilon > 0$ and for all $u$ so that $d(u,A) < \epsilon$, we have $d(T^nu,A) \to 0$ as $n\to \infty$. 
+	\item $A$ is asymptotically stable, i.e., for every $\epsilon > 0$ and for all $u$ so that $d(u,A) < \epsilon$, we have $d(T^nu,A){\to}{0}$ as $n\to \infty$. 
 	\item $A$ has Lyapunov stability, i.e., for every $\epsilon > 0$  there exists a $\delta(\epsilon) > 0$ so that $d(u,A) < \delta$ implies $d(T^nu,A) < \epsilon$ for all $n\ge 0$.  
@@
 In the 1960s, several mathematicians and mathematically interested scientists independently discovered chaos in the mathematical sense.
-The meteorologist Edward Lorenz may have been the first to explain this phenomenon in his 1963 paper \cite{lorenz1963deterministic}.
+The meteorologist Edward Lorenz may have been the first to explain this phenomenon in his 1963 paper~\cite{lorenz1963deterministic}.
 The notions of invariance, attractivity, and chaos may also be described for continuous systems, and Lorenz's system comprised of a system of differential equations.
@@
 In practice, when only data is observed from a system, it is not possible to verify which definition or notion is satisfied by the underlying dynamical system.
-We do, however, specifically recall the definition of chaos in the sense of Devaney \cite{devaney2018introduction,de2013elements} so as to understand some nuances behind the complexity.
+We do, however, specifically recall the definition of chaos in the sense of Devaney~\cite{devaney2018introduction,de2013elements} so as to understand some nuances behind the complexity.
 Devaney's definition of chaos has three requirements, with the first being the notion  of sensitive dependence on initial conditions.
@@
 
-To illustrate the concept of SDIC, we present two examples - one from Mathematics and the other from the field of Physics.
+To illustrate the concept of SDIC, we present two examples -- one from Mathematics and the other from the field of Physics.
 
 \begin{Example} \rm
-  The logistic map(LM), a recursion relation of the form $x_{n+1}=rx_n(1-x_n)$ where $x_n\in[0,1]$ and $r\leq{4}$ is a classic example used to illustrate the chaotic behaviour that a system may exhibit.
+  The logistic map (LM), a recursion relation of the form $x_{n+1}=rx_n(1-x_n)$ where $x_n\in[0,1]$ and $r\leq{4}$ is a classic example used to illustrate the chaotic behaviour that a system may exhibit.
 When $r=4$, the one-dimensional recurrence relation $x_{n+1}=4x_n(1-x_n)$ can be used as the kindergarten's model to exhibit the presence of SDIC.
-\ednote{B: edited and removed unnecessary citation}
-  In the graph below~\ref{fig:log_sdic} we plot the first 50 iterates of the LM for $r=4$ and two slightly different initial values $x_0$.
+    \ednote{B:edited and removed unnecessary citation}
+  In Graph~\ref{fig:log_sdic} below, we plot the first 50 iterates of the LM for $r=4$ and two slightly different initial values $x_0$.
 
@@
 \begin{Example} \rm
-  A second example pertains to a simple physical system called the double pendulum - a pendulum with another pendulum attached to its head given an initial position and angular velocity.
+  A further example pertains to the simple physical system of the double pendulum -- a pendulum with another pendulum attached to its head given an initial position and angular velocity (see the setup of the pendulum in Figure~\ref{fig:dp_setup}).
 
@@
 Topological transitivity implies that iterates of an open set of initial conditions get mixed up with other open sets.
-On a compact metric space, one may show that topological transitivity also implies the existence of a point whose forward iterates are dense \cite{de2013elements}; or in other words, the orbit going through this point will be dense in the compact metric space.
+On a compact metric space, one may show that topological transitivity also implies the existence of a point whose forward iterates are dense~\cite{de2013elements}; or in other words, the orbit going through this point will be dense in the compact metric space.
 In fact, it is topologically more likely that the choice of an arbitrary point will be one whose iterates are almost dense.
@@
 
-To be more precise (see \cite{de2013elements}), for a discrete-time dynamical system on a compact space, the set of points with dense iterates are residual, and they are typical or likely to be observed in practice.
+To be more precise (see~\cite{de2013elements}), for a discrete-time dynamical system on a compact space, the set of points with dense iterates are residual, and they are typical or likely to be observed in practice.
 In this project, when data is observed from a topologically transitive system, we assume it arises from a dense orbit.
@@
 The graph of the map $T^2$ comprises two symmetric tents with their base centered at $1/4$ and $3/4$.
-See Figure \ref{fig:T2tentmap}
+See Figure~\ref{fig:T2tentmap}
 
@@
 %Composite functions
-To understand the diagram, we note that if we travel "right and then down," the diagram instructs us to use $T$ (top arrow right) first, followed by $\phi$.
+To understand the diagram, we note that if we travel `right and then down', the diagram instructs us to use $T$ (top arrow right) first, followed by $\phi$.
 (right arrow downwards).
 Consequently, the pathway amounts to finding $\phi(T(u))$.
-If we proceed "down and then right," the diagram instructs us to apply $\phi$ (left arrow downwards) first, and then apply $S$ (bottom arrow right).
+If we proceed `down and then right', the diagram instructs us to apply $\phi$ (left arrow downwards) first, and then apply $S$ (bottom arrow right).
 This then amounts to finding  $S(\phi(u))$.
@@
 When $S$ is semi-conjugate to $T$, it is also common to call $S$ a \emph{factor of $T$}, or conversely that $T$ is an \emph{extension of $S$}.
-In essence, an extension (see, \cite{de2013elements}) is a more extensive system capturing all of the essential dynamics of its factor.
+In essence, an extension (see,~\cite{de2013elements}) is a more extensive system capturing all of the essential dynamics of its factor.
 
-It is a very hard or nearly an impossible task to establish the existence of a conjugacy or a semi-conjugacy $\phi$ between two systems \cite{devaney2018introduction}.
-However, one can verify that a function $\phi$ satisfies the commutativity diagram \ref{eqn_conjugacy}.
+It is a very hard or nearly an impossible task to establish the existence of a conjugacy or a semi-conjugacy $\phi$ between two systems~\cite{devaney2018introduction}.
+However, one can verify that a function $\phi$ satisfies the commutativity diagram~\ref{eqn_conjugacy}.
 
 \begin{Example}\rm
-  For example $\phi(x)=\sin\big(\frac{\pi}{2}x\big)^2$ is a conjugacy between two systems $Tu=1-|2u-1|$ and $Sv=4v(1-v)$  defined on $U=[0,1]$.
+  For example $\phi(x)=\sin{\big(\frac{\pi}{2}x\big)}^2$ is a conjugacy between two systems $Tu=1-|2u-1|$ and $Sv=4v(1-v)$  defined on $U=[0,1]$.
 
@@
       (\phi\circ{T})u 
-                &=\sin^{2}\Big(\frac{\pi}{2} - \frac{\pi}{2}|2u-1|\Big) = \sin^{2}(\pi{u})=\Big(2\sin(\frac{\pi}{2})\cos(\frac{\pi}{2})\Big)^{2}=4\sin^{2}(\frac{\pi}{2}u)\Big(1-\sin^{2}\frac{\pi}{2}u\Big) \\
+                &=\sin^{2}\Big(\frac{\pi}{2} - \frac{\pi}{2}|2u-1|\Big) = \sin^{2}(\pi{u})={\Big(2\sin(\frac{\pi}{2})\cos(\frac{\pi}{2})\Big)}^{2}=4\sin^{2}(\frac{\pi}{2}u)\Big(1-\sin^{2}\frac{\pi}{2}u\Big) \\
                                     &=({S}\circ\phi)u.

### 2022-01-18 projectReport/chapter3.tex
Worked through all tex errors

@@
 First we establish the method of Takens delay embedding. 
-In this method, one considers a discrete-time dynamical system defined on a 'nice' space (a smooth manifold)\ednote{B: Should this be further explained?} that can be obtained as time-$K$ map of a continuous time-dynamical system, concepts formalised below.
+In this method, one considers a discrete-time dynamical system defined on a `nice' space (a smooth manifold)\ednote{B: Should this be further explained?} that can be obtained as time-$K$ map of a continuous time-dynamical system, concepts formalised below.
 We make observations from such a system, i.e., we observe the evolution of $\theta(w_0)$ (where $w_0$ represents the initial state) by examining  a finite set of values of $u_n :=  \theta(w_n) = \theta(Tw_{n-1})$.
@@
 However, the actual state $u$ of a system is seldom, if ever, fully known.
-the most one can do is to insert probes into a system to obtain partial information through the measurements taken.
+The most one can do is to insert probes into a system to obtain partial information through the measurements taken.
 Moreover, the process of taking a measurement itself introduces two additional aspects which complicate the problem: \ednote{B: Perhaps this edit makes more sense?
@@
 \begin{Definition}\rm
-  [\bf {Homeomorphism}]\label{Dfn_homeo}\rm
+  {\bf(Homeomorphism)}\label{Dfn_homeo}\rm
   A homeomorphism is a function $f:Z\rightarrow Y$ between two topological spaces $Z$ and $Y$ that is continuous, bijective and has a continuous inverse. 
@@
 \begin{Definition}
-  [\bf {Embedding}]\label{Dfn_embed}\rm
+  [\bf{Embedding}]\label{Dfn_embed}\rm
   Consider a homeomorphism $f:Z\rightarrow Y$ for $Y\subset X$. $Z$ is said to be embedded in $X$ by $f$.
@@
 Takens Theorem states a result establishing a relationship between the observed and underlying dynamical systems by showing that the concatenation of a sufficiently large number of previous observations into a vector will, under certain conditions, generate a map between the vectors from the respective systems.
-We formulate the theorem from \cite{takens1981detecting}.
+We formulate the theorem from~\cite{takens1981detecting}.
 
 \begin{Theorem} 
-	[\bf Takens Embedding Theorem (adopted from \cite{takens1981detecting}] \label{Thm_Takens}
+	[\bf Takens Embedding Theorem (adopted from~\cite{takens1981detecting})]\label{Thm_Takens}
          Let $W$ be a compact manifold of dimension $m$, and $d\ge m$ so that $2d$ is an integer. It is a 
@@
             is a diffeomorphic embedding; by `smooth' we mean at least $C^2$.
-Consequently, there exists a map $F_\theta: \Phi_{2d,\theta}(W) \to \Phi_{2d,\theta}(W)$ defined by $$F_\theta: (\theta(T^{-2d}w),\ldots,\theta(T^{-1}w),\theta(w)) \mapsto
-            (\theta(T^{-2d+1}w),\ldots,\theta(w),\theta(Tw))$$
+Consequently, there exists a map $F_\theta: \Phi_{2d,\theta}(W) \to \Phi_{2d,\theta}(W)$ defined by \[F_\theta: (\theta(T^{-2d}w),\ldots,\theta(T^{-1}w),\theta(w)) \mapsto
+            (\theta(T^{-2d+1}w),\ldots,\theta(w),\theta(Tw)) \]
            so that $(W,T)$ is topologically conjugate to 
@@
 
-Below we reproduce Figure 1 from \cite{Supp}.
+Below we reproduce Figure 1 from~\cite{Supp}.
 
@@
 \begin{enumerate}
-\item Even supposing that we could indeed find $F_\theta$ , our \emph{approximation} of  $F_\theta$ is a map from a larger set $\mathbb{R}^{2d+1}$ containing the embedded attractor.
+\item Even supposing that we could indeed find $F_\theta$, our \emph{approximation} of  $F_\theta$ is a map from a larger set $\mathbb{R}^{2d+1}$ containing the embedded attractor.
 There are, however, no theoretical guarantees that $F_\theta$ will retain  $V=\Phi_{2d,\theta}(W)$  as an attractor although $W$ could itself be an attractor.
@@
 This requirement is not overly restrictive, as a great number of chaotic systems do have a surjective map.
-(Cite) \ednote{B: I remember you saying this in one of our sessions, but don't have a source for it.
+(Cite) \ednote{B:I remember you saying this in one of our sessions, but don't have a source for it.
 Do you have a specific reference?}

### 2022-01-18 projectReport/chapter4.tex
Worked through all tex errors

@@
 \begin{Definition}
-  [\bf Nonautonomous Dynamical System (adopted from~\cite{Manju_ESP}))] \label{Dfn_NDS}\rm
-  A nonautonomous dynamical system (NDS) on a space $X$ is simply a dynamical system comprising a family of maps $\{g_n\}_{n \in \mathbb{Z}}$ where $g_n(\ldotp):=g(u_n,\ldotp):X\to{X}$ is continuous.
+  [\bf Nonautonomous Dynamical System (adopted from~\cite{Manju_ESP})]\label{Dfn_NDS}\rm
+  A nonautonomous dynamical system (NDS) on a space $X$ is simply a dynamical system comprising a family of maps ${\{g_n\}}_{n \in \mathbb{Z}}$ where $g_n(\ldotp):=g(u_n,\ldotp):X\to{X}$ is continuous.
 Each $g_n$ arises as a consequence of an input $u_n$ from the input space $U$, a topological space.
@@
 
-We immediately recall the figure first defined in \eqref{eqn_conjugacy}  and remind ourselves of our ultimate goal to learn a system that is conjugate to an underlying or unknown dynamical system $(U,T)$ by making use of the measurements obtained from  the unknown system.
+We immediately recall the figure first defined in\eqref{eqn_conjugacy}  and remind ourselves of our ultimate goal to learn a system that is conjugate to an underlying or unknown dynamical system $(U,T)$ by making use of the measurements obtained from  the unknown system.
+See this sketched in the diagram~\ref{UV_conjugacy}
 
 %#####################################
-\begin{center}
+\begin{equation}
+  % \begin{center}
+    \centering
     \psset{arrows=->, arrowinset=0.25, linewidth=0.6pt, nodesep=3pt, labelsep=2pt, rowsep=0.7cm, colsep = 1.1cm, shortput =tablr}
     \everypsbox{\scriptstyle}
-\begin{psmatrix}
+    \begin{psmatrix}\label{UV_conjugacy}
     U & U\\%
@@
     %#####################################
-\end{center}
+    % \end{center}
+\end{equation}
 
@@
 \begin{Definition}
-  [\bf Driven, Compactly Driven Dynamical System] \label{Dfn_DDS} \rm
+  [\bf Driven, Compactly Driven Dynamical System]\label{Dfn_DDS} \rm
 A driven dynamical system comprises two topological metric spaces $U$, $X$ and a continuous function  $g:U\times{X}\to{X}$ where $g(u_n, x_n)=x_{n+1}$.
@@
 
-Notably, a nonautonomous dynamical system can be generated from $U$; any input $\overline{u}$, a bi-infinite sequence from $U$, gives rise to the sequence of self-maps $\{g(u_n, \cdot)\}_{n\in\mathbb{Z}}$ contained in $X$.
+Notably, a nonautonomous dynamical system can be generated from $U$; any input $\overline{u}$, a bi-infinite sequence from $U$, gives rise to the sequence of self-maps ${\{g(u_n, \cdot)\}}_{n\in\mathbb{Z}}$ contained in $X$.
 Physically, one may think of this bi-infinite sequence as referring to a system that has been running for an extended period at the time of the first measurement taken from the system (alternatively, the first time a probe is inserted into the system to take an observation)
@@
 \begin{Definition}
-  [\bf Entire Solution] \label{Dfn_Soln} \rm
-  A sequence $\{x_n\}_{n\in\mathbb{Z}}\subset X$ is called an entire solution (or simply a solution) of the driven system  $g$ with input $\overline{u}$ when it satisfies 
-  \[g(u_{n-1}, x_{n-1})=x_n\] for all $n\in\mathbb{Z}$
+  [\bf Entire Solution]\label{Dfn_Soln} \rm
+  A sequence ${\{x_n\}}_{n\in\mathbb{Z}}\subset X$ is called an entire solution (or simply a solution) of the driven system  $g$ with input $\overline{u}$ when it satisfies 
+  \begin{equation}
+    g(u_{n-1}, x_{n-1})=x_n \text{ for all }n\in\mathbb{Z}
+  \end{equation}
+
 \end{Definition}
@@
 
-\begin{Example} \rm \label{ex_halfux}
-  The only solution  $\{x_n\}_{n\in\mathbb{Z}}$ to the driven system  $g(u,x)=\frac{ux}{2}$, where $X=[0,1]$, $U=[0,1]$,  is the zero solution $x_n\equiv0$.
-  To see this, consider any $x_n=a\in[0,1]$ where $a\neq{0}$.  Let $\overline{u}\in{U}$ be an non-zero constant sequence , say $u_n=0.5$. 
+\begin{Example}\label{ex_halfux} \rm
+  The only solution  ${\{x_n\}}_{n\in\mathbb{Z}}$ to the driven system  $g(u,x)=\frac{ux}{2}$, where $X=[0,1]$, $U=[0,1]$,  is the zero solution $x_n\equiv0$.
+  To see this, consider any $x_n=a\in[0,1]$ where $a\neq{0}$.  Let $\overline{u}\in{U}$ be an non-zero constant sequence, say $u_n=0.5$. 
   The driven system may be rewritten as $x_{n-1}=\frac{2x_n}{u_{n-1}}=4x_n$ and the  iterates of $x_n$ in backward time will increase by a factor of 4 at each timestep. 
@@
   Consider the driven system $g(u,x)=x^2$ for $X=[0,1]$, $U=\mathbb{R}$.
-The system has an uncountable number of solutions, as there exists a solution for every $x\in{X}$ which also passes through the point $x$ and $lim_{n\to\infty}x_n=0$, $lim_{n\to{-}\infty}x_n=1$.
+The system has an uncountable number of solutions, as there exists a solution for every $x\in{X}$ which also passes through the point $x$ and $\lim_{n\to\infty}x_n=0$, $\lim_{n\to{-}\infty}x_n=1$.
  The proof is deferred to immediately after the next paragraph (see~\ref{rem_proofEx}).
@@
 The reachable set of a driven system $g$ is exactly the union of all the elements of all the (entire) solutions, i.e., 
-\[X_U :=\Big \{x \in X:  x = x_k \mbox{ where  $\{x_n\}$  is a solution for some  $\bar{u}$} \Big \}.\]
+\begin{equation}
+  X_U :=\Big \{x \in X:  x = x_k \mbox{ where  $\{x_n\}$  is a solution for some  $\bar{u}$} \Big \}.  
+\end{equation}
+
 The set of all reachable states at a specific time $n$ for input $\overline{u}$ is denoted by $X_n(\overline{u})$
@@
 $\overline{u}$, we define $g_i = g(u_i,\cdot)$ for all $i\in \mathbb{Z}$, then the set of states
-
 \begin{equation} \label{eqn_association}
@@
 It turns out that $\bigcap_{i<n}X_{n,i}(\overline{u})$ is identically equal to $X_n((\overline{u})$ that we had defined earlier.
-This will lead to a result (for.
-see, \cite{manjunath2013echo}) that  \textit{$g$ being a topological contraction is equivalent to the existence of a exactly one entire solution.}
+This will lead to a result (see~\cite{manjunath2013echo}) that \textit{$g$ being a topological contraction is equivalent to the existence of a exactly one entire solution.}
 
@@
 \begin{Remark}
-  [\bf Proof of Example~\ref{Ex_exp}] \label{rem_proofEx} \rm
+  [\bf Proof of Example~\ref{Ex_exp}]\label{rem_proofEx} \rm
   For the driven system,  
@@
 \end{Theorem}
-\begin{proof}
-  May be found in \cite{kloeden2011nonautonomous, manjunath2014dynamics, manjunath2013echo}
-\end{proof}
-
+{\bf Proof.} May be found in~\cite{kloeden2011nonautonomous, manjunath2014dynamics, manjunath2013echo}
 % One may easily construct many systems with trivial solution-sets, such as $g(u,x)=x$ which has only the constant solution $x$ and so for $U=[-1,1]$, the system would have no solution if $|x|>1$.
 To refine the scenario, we consider only systems with unique solutions.
-
 \section{Unique Solution Property}
@@
 \begin{Definition}
-  [\bf Unique Solution Property] \label{Dfn_usp}\rm
+  [\bf Unique Solution Property]\label{Dfn_usp}\rm
   A driven system $g$ is said to have the Unique Solution Property (USP) if for each input $\overline{u}$ there exists exactly one solution. 
@@
 If $g$ has the USP, then any solution to $g$ is also a uniform attractor in nonautonomous dynamical systems literature~\cite{Manju_Nonlinearity}.
-The discussion on nonautonomous attractors is beyond the scope of this project, and we refer the reader to \cite{Manju_ESP, esann2012ids} for further reading.
+The discussion on nonautonomous attractors is beyond the scope of this project, and we refer the reader to~\cite{Manju_ESP, esann2012ids} for further reading.
 
@@
 
-In the majority of Reservoir Computing(RC) literature (a machine learning approach) where the input is mapped onto a different but higher dimensional through a system called the reservoir and a readout that measures the state of the reservoir is trained, a notion of forgetting the states of the reservoir asymptotically is often used.
+In the majority of Reservoir Computing (RC) literature- a machine learning approach- where the input is mapped onto a different but higher dimensional through a system called the reservoir and a readout that measures the state of the reservoir is trained, a notion of forgetting the states of the reservoir asymptotically is often used.
 
@@
 B: Thanks!},
-Concepts like the ESP,  fading memory \cite{boyd1985fading} are some of these notions.  
+Concepts like the ESP,  fading memory~\cite{boyd1985fading} are some of these notions.  
 If $g$ possesses the ESP (equivalent in our context to the USP), the we are guaranteed that the whole of an input's left-infinite history will exactly determine the current system state; i.e.
-there is only ever the possibility of a single reachable state at a given instance of time \cite{jaeger2001echo,Manju_2020}.
+there is only ever the possibility of a single reachable state at a given instance of time~\cite{jaeger2001echo,Manju_2020}.
 % Defined formally, the ESP is intrinsically linked to an input (see \cite{Manju_ESP}) Unfortunately, definitions and results concerning the ESP do not have much practical benefit since a relationship between the network's response and the temporo-statistical properties of the input are not directly described.
@@
 % We do however consider it important to mention that the ESP is also often discussed in terms of a stability property that holds for all inputs of a system~\cite{manjunath2020stability} and, furthermore, plays a key role in the design and training of recurrent neural networks(RNN) within the field of RC \cite{Manju_ESP}.
-Beyond forgetting the past states the ESP is also often discussed in terms of a stability property that is nearby inputs create close by responses (solutions), a concept called input-related stability~\cite{manjunath2020stability} and, furthermore, plays a key role in the robustness of recurrent neural networks(RNN), a component of the implementation discussed in~\ref{ch5}.
+Beyond forgetting the past states the ESP is also often discussed in terms of a stability property that is nearby inputs create close by responses (solutions), a concept called input-related stability~\cite{manjunath2020stability} and, furthermore, plays a key role in the robustness of recurrent neural networks (RNN), a component of the implementation discussed in~\ref{ch5}.
 
@@
     [\bf Universal Semi-Conjugacy]\label{Def_UnivSemiConj} \rm
-    Given a driven system $g$, we  call a continuous and surjective map $h : \overleftarrow{U} \to X_U$ a universal semi-conjugacy if  diagram \ref{Scomm_h} commutes for all $v \in U$.
+    Given a driven system $g$, we  call a continuous and surjective map $h : \overleftarrow{U} \to X_U$ a universal semi-conjugacy if  diagram~\ref{Scomm_h} commutes for all $v \in U$.
   \end{Definition}
 
-  If the universal semi-conjugacy $h$ exists (i.e. the diagram in \ref{Scomm_h} commutes) then the solution $\Psi(\bar{u})$ will intuitively have no more ``complexity' than the input $\bar{u}$.
+  If the universal semi-conjugacy $h$ exists (i.e. the diagram in~\ref{Scomm_h} commutes) then the solution $\Psi(\bar{u})$ will intuitively have no more ``complexity' than the input $\bar{u}$.
 
 But does such a function $h$ for the above driven system $g$ above exist?
-Whenever $g$ has the USP and $\Psi(u)=\{x_n\}_{n\in\mathbb{Z}}$ it follows that $h$, defined by  $h(\cev{u}_n):=g(u_n,x_{n-1})=x_n$, will satisfy the semi-conjugacy in the graph above \ref{Scomm_h}.
-Regrettably, such a continuous mapping $h$ is not guaranteed to exist when $g$ does not have the USP \cite[Lemma 5]{Manju_Nonlinearity}.
-Note that even when $h$ does exist, we are not guaranteed its injectivity. Considering again example \ref{ex_halfux}, we see that even if $h$ were to exist, it could not be injective as $X_U=\{0\}$.
+Whenever $g$ has the USP and $\Psi(u)={\{x_n\}}_{n\in\mathbb{Z}}$ it follows that $h$, defined by  $h(\cev{u}_n):=g(u_n,x_{n-1})=x_n$, will satisfy the semi-conjugacy in the graph above~\ref{Scomm_h}.
+Regrettably, such a continuous mapping $h$ is not guaranteed to exist when $g$ does not have the USP~\cite[Lemma 5]{Manju_Nonlinearity}.
+Note that even when $h$ does exist, we are not guaranteed its injectivity. Considering again example~\ref{ex_halfux}, we see that even if $h$ were to exist, it could not be injective as $X_U=\{0\}$.
 
-Re-sketching the commutativity diagram \ref{Scomm_h} above by replacing $X_U$ by its left-infinite sequence space $\overleftarrow{X}_U$, we obtain the diagram below.
+Re-sketching the commutativity diagram~\ref{Scomm_h} above by replacing $X_U$ by its left-infinite sequence space $\overleftarrow{X}_U$, we obtain the diagram below.
 In this case, the function $H:\overleftarrow{U}\to\overleftarrow{X}_U$, a map that is both continuous and surjective, is called a \emph{causal mapping} that is defined next.
@@
 When $g$ has the USP, the diagram below (adopted from~\cite{Manju_Nonlinearity}) illustrates the operation of the mappings $h$ and $H$.
-The mapping $h:\overleftarrow{U}\to{X_U}$ is also considered an observable as mentioned in the introduction of Chapter \ref{ch3}).
+The mapping $h:\overleftarrow{U}\to{X_U}$ is also considered an observable as mentioned in the introduction of Chapter~\ref{ch3}).
 
@@
   [\bf Inverse-Limit Space, Inverse-Limit System]\label{Dfn_InverseL}\rm
- The subset $\widehat{U}_T\subset \overleftarrow{U}$ defined by \[\widehat{U}_T: = \{ (\ldots,u_{-2},u_{-1}): Tu_{i} = u_{i+1}\} \] is called the inverse-limit space of $(U,T)$.
- The self-map $\widehat{T}$  induced by $T$ on $\widehat{U}_T$  is defined by \[\widehat{T}: (\ldots,u_{-2},u_{-1}) \mapsto  (\ldots,u_{-2},u_{-1},T(u_{-1}))\] and the resulting dynamical system $(\widehat{U_T}, \widehat{T})$ is called the inverse-limit system of $(U,T)$.
+ The subset $\widehat{U}_T\subset \overleftarrow{U}$ defined by 
+ \begin{equation}
+  \widehat{U}_T: = \{ (\ldots,u_{-2},u_{-1}): Tu_{i} = u_{i+1}\}  
+ \end{equation}
+ is called the inverse-limit space of $(U,T)$.
+ The self-map $\widehat{T}$  induced by $T$ on $\widehat{U}_T$  is defined by 
+ \begin{equation}
+  \widehat{T}: (\ldots,u_{-2},u_{-1}) \mapsto  (\ldots,u_{-2},u_{-1},T(u_{-1}))  
+ \end{equation}
+ and the resulting dynamical system $(\widehat{U_T}, \widehat{T})$ is called the inverse-limit system of $(U,T)$.
 \end{Definition}
@@
 %To achieve a causal embedding embedding not just a causal mapping, it is undesirable to choose a function $g$ that quenches the temporal structure in $u$ by contracting to such a degree that the ability to recover information from the original system is lost completely.
-In the above example(\ref{ex_halfux}), the input's temporal variation could not be related to the reachable set as $X_U$ only consisted of a single element; little, if not no, information is encoded.
+In the above example (\ref{ex_halfux}), the input's temporal variation could not be related to the reachable set as $X_U$ only consisted of a single element; little, if not no, information is encoded.
 To obtain a suitably complex function $g$, it is thus desired that the reachable set of a driven system  be large enough to relate to the input. 
@@
 
-SI-invertibility promises that 'enough' information is retained when $g$ is chosen without introducing one's choice of $g$ will still ensure 'enough' information is retained without introducing unwanted complexity.
-'Enough' here refers to the fact that we may always recover the previous input value $u_n$ (if we know successive states $x_{n-1}, x_n$).
+SI-invertibility promises that `enough' information is retained when $g$ is chosen without introducing one's choice of $g$ will still ensure `enough' information is retained without introducing unwanted complexity.
+`Enough' here refers to the fact that we may always recover the previous input value $u_n$ (if we know successive states $x_{n-1}, x_n$).
 
@@
 To describe the  SDD formally, we consider a dynamical system $T: U \to U$ and define a relation on the reachable set $X_U$, i.e. a subset on $X_U \times X_U$  defined by 
-$$Y_T:=\{(x_{n-1},x_n): \{x_k\}_{k\in \mathbb{Z}} \mbox{ is a solution for some orbit of } T \mbox{ and } n \in \mathbb{Z}\}.$$
+\begin{equation}
+  Y_T:=\{(x_{n-1},x_n): {\{x_k\}}_{k\in \mathbb{Z}} \mbox{ is a solution for some orbit of } T \mbox{ and } n \in \mathbb{Z}\}.  
+\end{equation}
 
-The following theorem establishes the existence of a well-defined map $G_T$ describing the single-delay dynamics(SDD) of the system above.
+The following theorem establishes the existence of a well-defined map $G_T$ describing the single-delay dynamics (SDD) of the system above.
 
@@
 
-At this stage it is worth taking note of a specific driven system in the form of a discrete state-space model which has acquired some adherence in applications \cite{Manju_IEEE}- especially those pertaining to Echo State Networks and the ESP.
-The function
+At this stage it is worth taking note of a specific driven system in the form of a discrete state-space model which has acquired some adherence in applications~\cite{Manju_IEEE}- especially those pertaining to Echo State Networks and the ESP.\@The function
 \begin{equation}  \label{eqn_driving}
@@
 \end{equation} 
-is SI-invertible and, if $\alpha B$ has a spectral norm $<1$, also possesses the USP \cite[Th.2]{manjunath2013echo }. 
+is SI-invertible and, if $\alpha B$ has a spectral norm $<1$, also possesses the USP~\cite[Th.2]{manjunath2013echo }. 
 It is easy to show that $g$ is SI-invertible by recovering $u_n$ in 
@@
 
-This specific driving function $g$ is used in our implementation and is discussed more completely in chapter~\ref{ch5}
+This specific driving function $g$ is used in our implementation and is discussed more completely in Chapter~\ref{ch5}
 
 Despite the ease that with which one manipulates a left-infinite history in the realm of theory, it is impossible to obtain or use such a sequence in any real-life application.  
-Fortunately, one does not need the entire left-infinite history of an input in practice thanks to the Uniform Attraction Property(UAP). 
-We use an alternate version of the definition of UAP -- since the notion of nonautonomous attractors in this project would take some time to establish and detracts from the principal thrust of this project, we refer to \cite{Manju_Nonlinearity}.
+Fortunately, one does not need the entire left-infinite history of an input in practice thanks to the Uniform Attraction Property (UAP). 
+We use an alternate version of the definition of UAP -- since the notion of nonautonomous attractors in this project would take some time to establish and detracts from the principal thrust of this project, we refer to~\cite{Manju_Nonlinearity}.
 
@@
   A driven system $g$ has the Uniform Attraction Property (UAP) if we initialize the driven system
-with an arbitrary initial value $y_m \in X$, and the sequence $y_{m+1}, y_{m+2}, y_{m+3},...$ satisfying $y_{k+1}= g(u_k,y_k)$ for $k \geq m$ then approximates an actual solution $\{x_n\}$ uniformly in the sense that given $\epsilon>0$ (independent of $y_m$) there is an integer $n$ so that $d(x_{n+i}, y_{n+i})<\epsilon$ for all $i\ge 0$, where $\{x_m\}$ is a solution.
+with an arbitrary initial value $y_m \in X$, and the sequence $y_{m+1}, y_{m+2}, y_{m+3},\ldots$ satisfying $y_{k+1}= g(u_k,y_k)$ for $k \geq m$ then approximates an actual solution $\{x_n\}$ uniformly in the sense that given $\epsilon>0$ (independent of $y_m$) there is an integer $n$ so that $d(x_{n+i}, y_{n+i})<\epsilon$ for all $i\ge 0$, where $\{x_m\}$ is a solution.
 \end{Definition}
@@
   \centering
-\captionof{figure}{A coordinate of a simulated solution $(x_0,x_1,\ldots,x_{5000})$ of a RNN  plotted in red (with parameter $\alpha=0.99$ (where $g$ has the USP) and blue ($a=1$ and $\alpha=1.05$ where $g$ does not have the USP) while the matrices $A$ and $B$ are randomly generated, and $B$ has unit spectral radius.
-Reproduced from \cite{Manju_Nonlinearity}. }
+\captionof{figure}{A coordinate of a simulated solution $(x_0,x_1,\ldots,x_{5000})$ of a RNN  plotted in red (with parameter $\alpha=0.99$ where $g$ has the USP) and blue ($a=1$, $\alpha=1.05$ where $g$ does not have the USP) while the matrices $A$ and $B$ are randomly generated, and $B$ has unit spectral radius.
+Reproduced from~\cite{Manju_Nonlinearity}.
+}
 \label{fig:memloss_conttime} 
@@
 \begin{Theorem}
-  [\bf Causal Embedding Theorem (adopted from \cite{Supp})]
+  [\bf Causal Embedding Theorem (adopted from~\cite{Supp})]
   % {\bf (Causal Embedding Theorem.)}
@@
 
-Finally, we note that although $G_T$ exists with SI-invertibility, we need the USP as well(see \ref{fig_pictorialSummary} below for a visual illustration).
+Finally, we note that although $G_T$ exists with SI-invertibility, we need the USP as well (see~\ref{fig:pictorialSummary} below for a visual illustration). 
 If not, the driven states would have to be running for all time since they would not have forgotten their past states.
@@
   \captionof{figure}{Pictorial summary of results obtained up to now and how they overlap. Reproduced from \textbf{(Cite SIG-25May)} }
-  \label{fig:fig_pictorialSummary}
+  \label{fig:pictorialSummary}
 \end{figure}
@@
 In the above sections, we established the map $G_T$ describing the SDD of a driven system.
-By establishing the existence of this map, we’ve essentially embedded the attractor $U$ (recall from \ref{attractor_U}) into the higher dimensional space $X\times{X}$.
+By establishing the existence of this map, we’ve essentially embedded the attractor $U$ (recall from~\ref{attractor_U}) into the higher dimensional space $X\times{X}$.
 In layman’s terms, this ensures that there is more “dimensional room” for the underlying system’s underlying dynamics to “move”.
@@
 
-In \cite{manjunath2021universal} it is illustrated in an empirical fashion in that the map $G_T$ describes dynamics which are less functionally complex than that of $T$ or of the map $\Phi_{2d,\theta}$.
-This is done by implementing a Recurrent Neural Network (RNN), which is discussed in~\ref{ch5}.
+In~\cite{manjunath2021universal} it is illustrated in an empirical fashion in that the map $G_T$ describes dynamics which are less functionally complex than that of $T$ or of the map $\Phi_{2d,\theta}$.
+This is done by implementing a Recurrent Neural Network (RNN), which is discussed in Chapter~\ref{ch5}.
 \ednote{B: Should I add in that we'll be using Pearson coefficient?}
@@
 
-The reasons for taking this roundabout approach remain to be discussed in \ref{subs_LearnGamma}, but a pat answer may immediately be given: When $\Gamma$ has been learnt, the system can be driven autonomously and then $G_T$ is known anyway.
+The reasons for taking this roundabout approach remain to be discussed in~\ref{subs_LearnGamma}, but a pat answer may immediately be given: When $\Gamma$ has been learnt, the system can be driven autonomously and then $G_T$ is known anyway.
 We formalise this with a theorem.
@@
 \end{Theorem}
-{\bf Proof.}  See \cite[Th. 3c]{manjunath2021universal}.
+{\bf Proof.}  See~\cite[Th. 3c]{manjunath2021universal}.
 
-Projection mappings $\pi_i$ are defined in the traditional meaning where a $k$-dimensional vector is projected to it's $i^{th}$ component such that $$\pi_i:(a_1,a_2, \ldotp, a_{k-1}, a_k)\to{a_i}$$
+Projection mappings $\pi_i$ are defined in the traditional meaning where a $k$-dimensional vector is projected to it's $i^{th}$ component such that \[\pi_i:(a_1,a_2, \ldotp, a_{k-1}, a_k)\to{a_i}\]
 
-The graph in equation \ref{fig:comm_H_CET} is then extended as below:
+The graph in equation \ref{fig:comm_H_CET} is then extended to the diagram~\ref{Scomm_imp} below:
 
@@
 
-\section{Advantages of learning $\Gamma$} \label{subs_LearnGamma}
+\section{Advantages of learning $\Gamma$}\label{subs_LearnGamma}
 
@@
 Secondly, the function $\Gamma$ is known to be stable.
-Learning $\Gamma$ makes use of equations \ref{eqns_from_data} which in turn employs the driven system $g$ possessing the USP, and offers distinct advantages with regards to stability in the presence of perturbation.
+Learning $\Gamma$ makes use of equations~\ref{eqns_from_data} which in turn employs the driven system $g$ possessing the USP, and offers distinct advantages with regards to stability in the presence of perturbation.
 It is desirable that inputs that differ only slightly would have only ‘slight’ effects on the output of the system.
@@
 the value a, $\alpha$ in $\tanh$), we exploit both input-related and parameter-related stability.
-Both these notions are defined by way of the continuity of an encoding map of an input and design parameter respectively in \cite{manjunath2020stability}.
+Both these notions are defined by way of the continuity of an encoding map of an input and design parameter respectively in~\cite{manjunath2020stability}.
 
@@
  However if the system has the USP then the function $h$ is continuous and so $h(\bar{u}), h(\bar{u}_{noisy})$ remain close-by for small levels of noise maintaining a measure of robustness to input and measurement noise.\cite{manjunath2021universal}.
-In a similar fashion, parameter-related stability is obtained thanks to a result \cite[Lemma 3.2]{manjunath2020stability} relating this form of stability with the ESP (which is here equivalent to the USP).
+In a similar fashion, parameter-related stability is obtained thanks to a result~\cite[Lemma 3.2]{manjunath2020stability} relating this form of stability with the ESP (which is here equivalent to the USP).
 
@@
 
-Thirdly, learning a chaotic map on a set(here $Y_T$) with even a small measure of noise  means that the learnt system accepts arguments out of the set (Problems similar to Takens, cf.~\ref{sect_takenslimits}).
+Thirdly, learning a chaotic map on a set (here $Y_T$) with even a small measure of noise  means that the learnt system accepts arguments out of the set (Problems similar to Takens, cf.~\ref{sect_takenslimits}).
 We thus have a map $G_T^+$ acting on $Y_T^+$ and since $Y_T$ is not guaranteed to be an attractor of $G_T$ this could lead to errors.
@@
   \captionof{figure}{Illustrating empirically the importance of global dissipativity.
-Plotted below are the three principal components of randomly initiated points(blue) of a learnt version of $G_T$ versus the trajectory $(x_{n-1}, x_n)$(red) of driven states obtained from the Fractal Dream Attractor's data~\ref{Thomas_Attractor}).
+Plotted below are the three principal components of randomly initiated points (blue) of a learnt version of $G_T$ versus the trajectory $(x_{n-1}, x_n)$ (red) of driven states obtained from the Fractal Dream Attractor's data~\ref{Thomas_Attractor}.
 }

### 2022-01-18 projectReport/chapter5.tex
Worked through all tex errors

@@
 We apply our methodology to some chaotic attractors.
-Most notably, we consider the reconstruction of the chaotic attractor of the double pendulum - an  attractor that has not been successfully reconstructed because of its intricate structure.
+Most notably, we consider the reconstruction of the chaotic attractor of the double pendulum -- an  attractor that has not been successfully reconstructed because of its intricate structure.
 In fact the previous data-driven approaches also have not been successful in building a model from the double pendulum.
@@
 To reach this point, one must first learn $\Gamma$.
-After first normalising the data to ensure that it has zero mean value and lies within the range $[-1,1]$ and required zero-padding(see paragraph below for further discussion), we continue by driving the system with external input through the function $g$.
+After first normalising the data to ensure that it has zero mean value and lies within the range $[-1,1]$ and required zero-padding (see paragraph below for further discussion), we continue by driving the system with external input through the function $g$.
 
-From \eqref{eqn_driving} we know that $g(u,x) = (1-a) x + a \tanh(Au + \alpha B x)$. Recall that $g$ is SI-invertible. We refer to $A$ as the input matrix and $B$ the reservoir matrix.
+From~\eqref{eqn_driving} we know that $g(u,x) = (1-a) x + a \tanh(Au + \alpha B x)$. Recall that $g$ is SI-invertible. We refer to $A$ as the input matrix and $B$ the reservoir matrix.
 
@@
 \ednote{B: Should we say what the conventional manner is?} A few words about the RNNs: RNNs are widely used in the literature for temporal processing of data as a type of deep learning technique.
-They take the form of equations considered to be discrete-time version of continuous dynamical systems modelling a neural network of neurons \cite{jaeger2001echo}.
+They take the form of equations considered to be discrete-time version of continuous dynamical systems modelling a neural network of neurons~\cite{jaeger2001echo}.
 
@@
 
-We then reach the prediction step. Making use of $g$'s SI-invertibility and $g$ as defined in \eqref{eqn_driving}, we have the following expression for $u_n$ replicated from~\eqref{eqn_SI_RNN}.
+We then reach the prediction step. Making use of $g$'s SI-invertibility and $g$ as defined in~\eqref{eqn_driving}, we have the following expression for $u_n$ replicated from~\eqref{eqn_SI_RNN}.
 \begin{equation*}
@@
 We also consider a more general problem of learning an attractor of dynamical system when only observations of an orbit are given.
-Explicitly, if $(W,T)$ is a dynamical system with dynamics generated by $w_{n+1}=Tw_n$, and if $\theta:W \to \mathbb{R}$ is an observable, then the task is to learn a dynamical system that is topologically conjugate to $(W,T)$ and predict $\theta(w_{m+1}),\theta(w_{m+2}),...$ using the data $\theta(w_{0}),\theta(w_{1}),\ldots,\theta(w_{m})$.
+Explicitly, if $(W,T)$ is a dynamical system with dynamics generated by $w_{n+1}=Tw_n$, and if $\theta:W \to \mathbb{R}$ is an observable, then the task is to learn a dynamical system that is topologically conjugate to $(W,T)$ and predict $\theta(w_{m+1}),\theta(w_{m+2}),\ldots$ using the data $\theta(w_{0}),\theta(w_{1}),\ldots,\theta(w_{m})$.
 So how do we do it?
@@
 
-Our implementation of the reservoir(A) and input(B) matrices in the RNN makes use of sparse matrices.
+Our implementation of the reservoir (A) and input (B) matrices in the RNN makes use of sparse matrices.
 We discuss three specifics relating to this concept.
@@
 If $X_{1:N}=U\Sigma P^T$\ednote{Expand here} represents the singular value decomposition of $X_{1:N}$ \ednote{Ask: Is there a specific resource to cite?}, we then denote the principal component matrix $P$, and the principal components are given by
- $Z_{1:N}=X_{1:N}P.$
+ $Z_{1:N}=X_{1:N}P$.
 These principal components are then used to train a feedforward neural network. Denote the row vectors of $Z_{1:N}$ by 
-$z_i^T, i=1,2,\cdots,N$ and the neural network by $NN$, the network is subsequently trained to learn an approximation of the map
+$z_i^T, i=1,2,\ldots,N$ and the neural network by $NN$, the network is subsequently trained to learn an approximation of the map
 $NN: (z_{n-1},z_n) \mapsto  u_n.
@@
 The learnt approximation of $NN$ can be used to approximate $\Gamma$ since 
-$$
+\[
 NN\left( \begin{bmatrix} 
@@
 \end{bmatrix} = u_n = \Gamma(x_{n-1},x_n).
-$$
+\]
 
@@
 The first (or higher pendulum) is attached to a fixed point while the second is attached to the endpoint of the first.
-(See below)
+(See below the Figure~\ref{fig:dp_setup})
 This system is a classical example of a system exhibiting chaos and is often used as explanation of the very well-known Butterfly Effect.
-\ednote{B: Originally though a reference here was needed, but no longer think so.} Consider again~\ref{fig:dp_sdic}.
+\ednote{B: Originally though a reference here was needed, but no longer think so.} Consider again the Figure~\ref{fig:dp_sdic}.
 The system has four variables: angle and angular velocity of each pendulum rod. An ideal system is assumed where the rods are massless and no friction in the pivots are considered.
@@
 In such a case, the attractor is a single point in the phase space when the double pendulum comes to rest.
-It must be mentioned once again that if we use data outside the attractor, learning is not reliable. To see this, consder the graphs ~\ref{fig:damped_pendulum} and ~\ref{fig:dp_notsurjective}.
+It must be mentioned once again that if we use data outside the attractor, learning is not reliable. 
+To see this, consider Figure~\ref{fig:damped_pendulum} and~\ref{fig:dp_notsurjective}.
 
@@
 \caption{$x$- and $y$-coordinates of the two pendulum heads plotted for a dataset that exhibits dying oscillations.
-The map generating the (four-dimensional data) for a damped pendulum is not surjective.
+The map generating the (four-dimensional data) for a damped pendulum is not surjective and so learning $\Gamma$ for this dataset will fail.
+Consider below the Figure~\ref{fig:dp_notsurjective}.
 Data obtained from chaotic double pendulum dataset generated from videos takes of experimental pendulums~\cite{asseman2018learning}}
@@
   \centering
-\caption{Failure to forecast or learn a map which is not surjective: Evolution of the $y$-coordinate of the bottommost pendulum plotted here for 15000 timestep.
-Plotted here the predicted value(red) versus the actual value(blue).
+\caption{Failure to forecast or learn a map which is not surjective: Evolution of the $y$-coordinate of the bottommost pendulum plotted here for 15000 timesteps.
+Plotted here the predicted value (red) versus the actual value (blue).
 The plots overlap for the first 10000 steps, the duration of training to learn the map $\Gamma$.
+Parameter-values are exactly as in the successful case discussed in the paragraph directly afterwards.
 Afterwards, the trajectories start diverging and it the predicted value no longer accurately represents the complexity or energy-loss of the actual data.
@@
 \end{figure}
+\ednote{Can I say in the caption \emph{Parameter-values are exactly as in the successful case discussed in the paragraph directly afterwards.}}
 
@@
 
-The system was driven for 15500 steps, of which the first 500 were immediately discarded to simulate the network 'forgetting'.
+The system was driven for 15500 steps, of which the first 500 were immediately discarded to simulate the network `forgetting'.
 The driving function $g$ had input and reservoir matrices were implemented as sparse matrices filled to 20\%.
@@
 Training was accomplished using 256 epochs with batch sizes 128 and a delay of 1.
-See~\ref{fig:dp_success_traj} for results pertaining to the trajectory taken by the DP's components and then~\ref{fig:dp_success_density} to witness associated densities of predicted vs.
-actual values.
+See~\ref{fig:dp_success_traj} for results pertaining to the trajectory taken by the DP's components and then~\ref{fig:dp_success_density} to witness associated densities of predicted versus actual values.
 
@@
   \includegraphics[width=0.95\linewidth]{Graphs/_dp_success_4coords_traj.eps} 
-  \captionof{figure}{Predicted(red) and actual(blue) trajectories of the $x$- and $y$-coordinates of each of the pendulum heads.
+  \captionof{figure}{Predicted (red) and actual (blue) trajectories of the $x$- and $y$-coordinates of each of the pendulum heads.
 From top to bottom are as follows: $x$-coordinate of topmost pendulum head, $y$-coordinate of topmost pendulum head, $x$-coordinate of lower pendulum head, $y$-coordinate of lower pendulum head.
@@
 We now consider an experiment where noise is added to the system and find it to be incredibly robust.
-We employ the same parameters as before and add noise from a normal-distribution with mean zero and standard deviation equal to 0.1 which translates to roughly 39dB's of noise.
+We employ the same parameters as before and add noise from a normal-distribution with mean zero and standard deviation equal to 0.1, translating to roughly 39dB's of noise.
 (For the signal-to-noise ration we adopt the formula $\text{SNR}_{dB}=10\text{log}_{10}\big(\frac{P_\text{{signal}}}{P_{\text{noise}}}\big)$ where $P_\text{{signal}}$ refers to the variance observed in the normalised data and $P_{\text{noise}}$ represents the variance of the added noise.)
@@
   \includegraphics[width=0.95\linewidth]{Graphs/_dp_noise_FigCD_2030.eps} 
-  \captionof{figure}{Predicted(red) and actual(blue) trajectories of the $x$- and $y$-coordinates of each of the pendulum heads with noise added from a normal distribution with mean zero and standard deviation 0.1, equating to approximately 39dB's of noise.
+  \captionof{figure}{Predicted (red) and actual (blue) trajectories of the $x$- and $y$-coordinates of each of the pendulum heads with noise added from a normal distribution with mean zero and standard deviation 0.1, equating to approximately 39dB's of noise.
   These graphs were constructed by predicting the DP 1000 timesteps into the future and in so doing illustrating the long-term consistency and accuracy of the learnt system.
@@
 \subsection{Additional Attractors: Clifford \& Thomas}
-We follow the methodology presented in \cite{manjunath2021universal} but opt to consider attractors different from the Lorenz system and H\'enon map in this project.
+We follow the methodology presented in~\cite{manjunath2021universal} but opt to consider attractors different from the Lorenz system and H\'enon map in this project.
 
 \subsubsection{Thomas' Cyclically Symmetric Strange Attractor}\label{Thomas_Attractor}
-Thomas' Cyclically Symmetric Strange Attractor, a 3D attractor proposed in 1999 by Ren\'e Thomas in \cite{ThomasAttractor}, is described by a set of three equations which is distinctly reminiscent of the Lorenz system, another deterministic three-equation model exhibiting chaotic behaviour.It has a single parameter $\beta$ and has been shown to transition to chaotic behavious when $\beta<0.208186$~\cite{Thomas_BetaParameter}.
+Thomas' Cyclically Symmetric Strange Attractor, a 3D attractor proposed in 1999 by Ren\'e Thomas in~\cite{ThomasAttractor}, is described by a set of three equations which is distinctly reminiscent of the Lorenz system, another deterministic three-equation model exhibiting chaotic behaviour.
+It has a single parameter $\beta$ and has been shown to transition to chaotic behavious when $\beta<0.208186$~\cite{Thomas_BetaParameter}.
 \begin{eqnarray}\label{eqns_thomas}
@@
 
-The behaviour of the system with a $\beta$-parameter value of 0.1056 was simulated by scaling it to fit inside the interval $[-1,1]$.
-Two cases were considered: noise-free and the one perturbed by noise.
-Noise was generated from a normal distribution with mean 0 and standard deviation 0.05, which to translates approximately 33dB.
-($\text{SNR}_dB$ is calculated exactly as for the DP-data above)
+The behaviour of the system with a $\beta$-parameter value of 0.1056 was simulated by scaling it to fit inside the interval $[-1,1]$. Two cases were considered: noise-free and perturbed by noise. 
+Noise was generated  by adding data points taken from a normal distribution with mean 0 and standard deviation 0.05, which to translates approximately 33dB. 
+($\text{SNR}_{dB}$ is calculated exactly as for the DP-data above)
 
 A sequence of 3000 observations in both the clear and noisy data-sets respectively were fed into the system with 300 entries discarded to simulate the network's loss of memory. 
-The next 9000 steps were predicted and compared with true data.
+The next 9000 steps were predicted and compared with true data. See Figure~\ref{fig:Thomas_trajectories} for the actual and predicted trajectories of the noise-free and perturbed systems respectively.
 
@@
   \endminipage
-  \captionof{figure}{True(red) and predicted(blue) trajectories for the Thomas attractor with parameter value $\beta=0.1056$.
+ \captionof{figure}{True (red) and predicted (blue) trajectories for the Thomas attractor with parameter value $\beta=0.1056$.
 To the left: Noise added from normal distribution with mean zero and standard deviation 0.05.
 Right: No noise added to dataset.}
+\label{fig:Thomas_trajectories}
 \end{figure}
@@
 Another 2-equation discrete-time map named the Fractal Dream Attractor, or more commonly known as the Pickover Map (first discovered by Clifford A.
-Pickover and discussed in his fascinating book "Chaos in Wonderland"\cite{PickoverChaos}), was also considered.
+Pickover and discussed in his fascinating book `Chaos in Wonderland'\cite{PickoverChaos}), was also considered.
 \begin{eqnarray}\label{eqns_clifford}
@@
   \endminipage
-  \caption{The bottom row, left, shows a plot of the true attractor(red) and the predicted values(blue).
+  \caption{The bottom row, left, shows a plot of the true attractor (red) and the predicted values (blue).
 Bottom row, right, shows the distribution of the first coordinate of the learnt system (blue) and the actual Clifford system (red).
@@
 The behaviour of the system with a parameter values of $x_0=0$, $y_0=0$, $a = -1.7$,  $b = 1.8$, $c = -1.9$ and $d = -0.4$ was examined. 
-Data was scaled to fit inside the interval $[-1,1]$. Only the noise-free case was investigated. 
- A sequence of 25000 observationswas fed into the system with 15000 entries discarded to simulate the network's loss of memory. The next 1000 steps were predicted and compared with true data.
+Data was scaled to fit inside the interval $[-1,1]$. Only the noise-free case was investigated. Depictions of the empirical results available in Figure~\ref{fig:Clifford}.
+ A sequence of 25000 observations was fed into the system with 15000 entries discarded to simulate the network's loss of memory. The next 1000 steps were predicted and compared with true data.
 The RNN was initialised exactly as for Thomas' Cyclically Symmetric Strange attractor: the dimension of the network is set equal to 500 neurons and 12 hidden layers of dimension 64 each were initialised.
@@
 
-\section{Functional Complexity Reduction in Noninvertible maps}
-
+\section{Reducing Functional Complexity in Noninvertible Maps}
+\ednote{section header is overfull}
 % Besides being able to learn the action of the Koopman operator exactly, our setup has other advantages over selecting or optimizing the choice of observables as in the Extended Dynamic Mode Decomposition (EDMD) algorithm (see, [21, 22]).
@@
 Further, empirically it is found that $G_T$ is much more linear than $T$ as measured by what is called the multidimensional correlation coefficient $\rho$.
-This coefficient is a generalization of the Pearson correlation coefficient for real-valued random variables to random vectors (see, \cite{puccetti2019measuring}).
+This coefficient is a generalization of the Pearson correlation coefficient for real-valued random variables to random vectors (see,~\cite{puccetti2019measuring}).
 We state teh details of computing this coefficient between
@@
 For the non-invertible system defined by the Clifford attractor, we indicate the difference by which $\rho$ changes for $G_T$ in Table~\ref{tbl_attractorsPearson}.
-Following \cite{manjunath2021universal}, we say that this multidimensional correlation coefficients $\rho$ to evidence a general reduction in the functional complexity of the map $G_T$ emerging due to a RNN.
+Following~\cite{manjunath2021universal}, we say that this multidimensional correlation coefficients $\rho$ to evidence a general reduction in the functional complexity of the map $G_T$ emerging due to a RNN.
 In the
-Table ~\ref{tbl_attractorsPearson}, the second column corresponds to the amount of artificial neurons (the dimension of $X$) in the implemented RNN.
+Table~\ref{tbl_attractorsPearson}, the second column corresponds to the amount of artificial neurons (the dimension of $X$) in the implemented RNN.
 
@@
       {\begin{tabular}{|c|c| c c |} 
-          \hline
+          \toprule
           Input & Dimension of $X$ & $u_n$ vs $u_{n+1}$ 
@@
           \end{bmatrix}$ \\
-          \hline      
+          \midrule      
           % \hline
@@
 
-               \hline  
+               \midrule  
                $(w_n)$ comes from the Clifford Attractor \\ (for $a = -1.7; b = 1.8; c = -1.9; d = -0.4$): & & & \\
@@
           %       & 1000 & & 0.71151 \\
-           \hline
+           \bottomrule
                 \end{tabular}} %\ednote{http://paulbourke.net/fractals/peterdejong/}
- \label{Table_FC}
- \caption{Following \cite{manjunath2021universal}, we determine the multidimensional correlation coefficients $\rho$ to evidence a general reduction in the functional complexity of the map $G_T$ emerging due to a RNN.
+ \caption{Following~\cite{manjunath2021universal}, we determine the multidimensional correlation coefficients $\rho$ to evidence a general reduction in the functional complexity of the map $G_T$ emerging due to a RNN.
 Rows correspond to distinct dynamical systems considered in the experiments
@@
           \end{bmatrix}\right)$ may be contrasted by the reader. 
-          Finally, the RNN's dimension (from \eqref{Seq_RNN}) is provided to exhibiting that an increase in dimension of the driven system will 
+          Finally, the RNN's dimension (from~\eqref{Seq_RNN}) is provided to exhibiting that an increase in dimension of the driven system will 
           typically result in a map $G_T$ with less functional complexity.}.

### 2022-01-18 projectReport/chapter3.tex
checkpoint

@@
 
-\section{Practical Limitations to Takens' Theorem}
+\section{Practical Limitations to Takens' Theorem}\label{sect_takenslimits}
 Takens' Embedding Theorem is a powerful result, and it provides compelling reason to believe that one could conceivably reconstruct a system conjugate accurately to the underlying system. 

### 2022-01-18 projectReport/chapter5.tex
checkpoint

@@
 The learnt approximation of $NN$ can be used to approximate $\Gamma$ since 
-\[
+\begin{equation}\label{Seq_RNN}
   NN\left( \begin{bmatrix} 
@@
     \end{bmatrix} = u_n = \Gamma(x_{n-1},x_n).
-\]
+\end{equation}
 
@@
       \scalebox{0.75}
+      \centering
       {\begin{tabular}{|c|c| c c |} 
@@
           \end{bmatrix}\right)$ may be contrasted by the reader. 
-          Finally, the RNN's dimension (from~\eqref{Seq_RNN}) is provided to exhibiting that an increase in dimension of the driven system will 
+          Finally, the RNN's dimension (from~\eqref{Seq_RNN}\ednote{Is this reference correct?}) is provided to exhibiting that an increase in dimension of the driven system will 
 typically result in a map $G_T$ with less functional complexity.}.

### 2022-01-18 projectReport/abstract.tex
M final edits to Ch.5 and Michelle edits up to Ch.4

@@
     It is a well-established practice to collect measurements from underlying artificial, natural, and physical systems through observables in order to realise a dynamical model with greater descriptive power.
-    Many data-driven approaches to forecast the future evolution of systems exist.
-However, many of these methods rely on an initial understanding of the system and we often do not fully understand the many complexities of more intricate systems that we interact with on a daily basis.
+    Many approaches to forecasting exist but they necessitate an understanding of systems we often don't know much about.
 
-    In this project, the base assumption is that data is originates from  some discrete-time dynamical system, particularly one of a chaotic nature but no assumptions are made pertaining to a specific dynamical system.
+    In this project, the base assumption is that data is originates from  some discrete-time dynamical system, particularly one of a chaotic nature, but no assumptions are made pertaining to a specific dynamical system.
     A general learning problem is formulated and a well-known approach to the problem, alongside its practical limitations, is then considered.
-    The notion of driven dynamical systems is introduced, and we subsequently establish the theoretical existence of a dynamical system possessing certain properties which guarantees it to be topologically conjugate (`dynamically equivalent`) to the dynamics of the underlying system from which measurements were originally taken.
-    Our methodology is implemented practically by making use of recurrent neural networks, a form of deep learning, to learn a map topologically equivalent to the underlying system that forecasts future states.
+    The notion of driven dynamical systems is introduced, and we subsequently establish the theoretical existence of a dynamical system possessing certain properties which guarantees it to be topologically conjugate (`dynamically equivalent') to the dynamics of the underlying system from which measurements were originally taken.
+    Our methodology is implemented practically by making use of recurrent neural networks to learn a map topologically equivalent to the underlying system that forecasts future states.
     In so doing we present the workability of an approach to obtain long-term topological and statistically consistent predictions for simple physical systems as well as a some chaotic attractors.

### 2022-01-18 projectReport/chapter1.tex
M final edits to Ch.5 and Michelle edits up to Ch.4

@@
 This then confirms the existence of a map on the space of delay coordinates equivalent (or topologically conjugate) to the underlying map from which the observed time-series was first obtained.
-Although topological conjugacy guarantees an alternate representation of the underlying system, the quality still depends on numerous parameters, making the comprehension of the dynamics unreliable (see, [?]).
+Although topological conjugacy guarantees an alternate representation of the underlying system, the quality still depends on numerous parameters, making the understanding of the dynamics unreliable (see \cite{principe1992prediction}).
 One reason for this fragility is that the embedded attractor in the reconstruction space is not always an attractor of the map learnt in the reconstruction space, despite unquestionably being an invariant set.
@@
 On the other hand, pure machine learning methodology processes temporal information (like the echo state networks~\cite{Manju_ESP, Manju_IEEE, grigoryeva2018echo}) by mapping data onto a higher dimensional space for further processing.
-Although they perform  well on forecasting some dynamical systems, they fail completely on others (\ednote{Ask for resources here}) as there is often no guarantee that the right function was learnt during training.
+Although they perform  well on forecasting some dynamical systems, they fail completely on others~\cite{Supp} as there is often no guarantee that the right function was learnt during training.
 
 This project involves implementing and analysing non-interpretable models that can guarantee exact reconstruction. 
-The project work concerns the study and implementation of a method (\cite{manjunath2021universal}) that incorporates learning a function by mapping the data on to a higher dimensional space using what is called a driven dynamical system (See Chapter~ref{ch.4}).
+The project work concerns the study and implementation of a method (\cite{manjunath2021universal}) that incorporates learning a function by mapping the data on to a higher dimensional space using what is called a driven dynamical system (See Chapter~\ref{ch4}).
 With a clear understanding of how the data is mapped onto the higher dimensional space, the method then permits the learning of a dynamical system topologically conjugate to that of the underlying system.
@@
 With slight modifications to the implementation in the paper (\cite{manjunath2021universal}), we show that one can construct accurate non-interpretable models with the ability to reconstruct attractors from more hard-to-forecast systems like the double pendulum.
-(The forecasting of the time-series from a double pendulum has not been reported before.) Moreover, we also demonstrate that long-term statistical consistency is preserved.
+The forecasting of the time-series from a double pendulum has not been reported before.
+Moreover, we also demonstrate that long-term statistical consistency is preserved.
 
-\emph{By solidifying the mathematical underpinnings of our theory, we hope to guarantee the ability to construct models with predictive power ranging from molecular biology to neuroscience in the near future.
+% \emph{By solidifying the mathematical underpinnings of our theory, we hope to guarantee the ability to construct models with predictive power ranging from molecular biology to neuroscience in the near future.
 (We can modify this sentence when the report is completed.)}

### 2022-01-18 projectReport/chapter2.tex
M final edits to Ch.5 and Michelle edits up to Ch.4

@@
 
-Formally, a function $T: U \to U$, where $U$ is some set is  a \emph{discrete-time dynamical system} and its iterates $\{u,Tu,T^2u,\ldots\}$, where $T^n$ denotes the $n$-fold composition of $T$ with itself, describe the evolution of an initial condition $u\in U$ (Note that we frequently drop the brackets and denote $T(u)$ by $Tu$ so as to simplify notations).
+Formally, a function $T: U \to U$, where $U$ is some set is  a \emph{discrete-time dynamical system} and its iterates $\{u,Tu,T^2u,\ldots\}$, where $T^n$ denotes the $n$-fold composition of $T$ with itself, describe the evolution of an initial condition $u\in U$.
+Note that we frequently drop the brackets and denote $T(u)$ by $Tu$ so as to simplify notations.
 
-Continuous-time dynamical systems modelled using ordinary differential equations can give rise to discrete-time dynamical systems.
-To see this, consider a differential equation $\dot{x} = f(x)$, $f: \mathbb{R}^n \to \mathbb{R}^n$, $n\in\mathbb{N}$ given to have a unique solution passing through
+Continuous-time dynamical systems modelled using ordinary differential equations can give rise to discrete-time dynamical systems. To see this, consider a differential equation 
+\[\dot{x} = f(x), f: \mathbb{R}^n \to \mathbb{R}^n, n\in\mathbb{N}\]
+  given to have a unique solution passing through 
 each point $x\in\mathbb{R}^{n}$, call it $x(t)$ where $x(0)=x_0$.
@@
   [\bf Flow of an Equation]\label{Dfn_Flow}\rm
-  The flow of the equation  $\dot{x} = f(x)$  is defined to be a mapping $\varphi: \mathbb{R}^n \times \mathbb{R} \to \mathbb{R}$ where $\varphi(x_0,t)= x(t)$ for the solution $x(t)$ with $x(0)=x_0$ 
+  The flow of the equation  $\dot{x} = f(x)$  is defined to be a mapping $\varphi: \mathbb{R}^n \times \mathbb{R} \to \mathbb{R}$ where $\varphi(x_0,t)= x(t)$ for the solution $x(t)$ with $x(0)=x_0$.
 \end{Definition}
@@
 \begin{equation}
-  x(t+(m+1)h) = x(t+mh) + h f(x(t+mh))
+  x(t+(m+1)h) = x(t+mh) + h f(x(t+mh)).
 \end{equation}
@@
 Once again, we may consider the field of ecology, where discrete dynamical systems are often directly derived or assumed.
-\ednote{Ask for resource}
+\ednote{Could you refer me to a resource here?}
 There is a school of thought that advocates discrete-time dynamical systems to be more natural for modelling real-world observations than differential equations.
@@
   [\bf Invariant Set]\label{Dfn_InvariantSet}\rm
-  Given a discrete-time dynamical system $T: U \to U$, a subset $A \subset U$ is said to be an \emph{invariant set} if $T(A) =A$. 
+  Given a discrete-time dynamical system $T: U \to U$, a non-empty subset $A \subset U$ is said to be an \emph{invariant set} if $T(A) =A$. 
 \end{Definition}
@@
   [\bf Orbit of $T$]\label{Dfn_Orbit}\rm
-  The orbit of $T$ is to be defined the sequence $\bar{u} = {\{u_n\}}_{n\in \mathbb{Z}}$ obeying the update equation, $u_{n+1}=Tu_n$, $n \in \mathbb{Z}$. 
+  The orbit of $T$ is defined  is to the sequence $\bar{u} = {\{u_n\}}_{n\in \mathbb{Z}}$ obeying the update equation, $u_{n+1}=Tu_n$, $n \in \mathbb{Z}$. 
 \end{Definition}
@@
 The entire space $U$ could be also be invariant.
-Consider for example the space $U=[0,1]$, where $Tu=4u(1-u)$ and then $U$ is invariant (as every $u\in{U}$ can be written as $u = 4x(1-x)$ for some $x\in{U}$).
+Consider for example the space $U=[0,1]$, where $Tu=4u(1-u)$ and then $U$ is invariant as every $u\in{U}$ can be written as $u = 4x(1-x)$ for some $x\in{U}$.
 
 We may learn a great deal about the iterates of a dynamical system by considering the types of invariant sets of a discrete-time dynamical system.
-For example, if $U=[0,1]$ has map $Tu= u/2$, then the only invariant set is $\{0\}$, and all orbits approach this invariant set as time flows in the forward direction.
+
+\begin{Example}\label{ex_halfu} \rm
+If $U=[0,1]$ has map $Tu= u/2$, then the only invariant set is $\{0\}$, and all orbits approach this invariant set as time flows in the forward direction. 
 Indeed, if a some non-zero invariant set (call it $B$) exists, then there is  some $r\in(0,1]\cap{B}$.
@@
 Moreover, every orbit of T will be decreasing and therefore approach the value $0$ as $n\rightarrow\infty$.
+\end{Example}
 
@@
 
-\begin{Example}\rm
+\begin{Example}\label{ex_Tu2}\rm
   For the map $Tu=u^2$ defined on $[0,1]$ all orbits lie in the invariant set $\{1\}$ or else would would approach the invariant set $\{0\}$. 
@@
 
-Indeed in our previous examples, it can be verified that the system  $Tu=u/2$ defined on  $\mathbb{R}$ and $Tu=u^2$ defined on $[0,1]$  the singleton set $\{0\}$ is an attractor.
+Indeed in our previous example~\ref{ex_halfu, ex_Tu2}, it can be verified that the system  $Tu=u/2$ defined on  $\mathbb{R}$ and $Tu=u^2$ defined on $[0,1]$  the singleton set $\{0\}$ is an attractor.
 For the system,  $Tu=1-|2u-1|$ on $[0,1]$, one may easily verify that the only attractor is the entire space $[0,1]$.
-This follows from the fact that between any two points $u< v$ in $[0,1]$, and for any $a,b$ so that $u\le a < b \le v$, we can find an $n$ so that $T^n(a,b)=[0,1]$ (as would be explained later in this chapter).
+This follows from the fact that between any two points $u$ and $v$ s.t.
+$u< v$ in $[0,1]$, and for any $a,b$ so that $u\le a < b \le v$, we can find an $n$ so that $T^n(a,b)=[0,1]$  as is explained in Ex.~\ref{ex_standardTentMap}.
 
@@
 
-\begin{Definition}\rm
-  [\bf {Sensitive Dependence on Initial Conditions}]\label{Dfn_SDIC}\rm
-A dynamical system $T: U \to U$ is said to have sensitive dependence on initial conditions (SDIC) if there exists a $\delta > 0$ such that for every $u \in U$ and in every neighborhood of $u \in U$ there exists a $v\in{U}$ and an integer $N:=N{(u,v)}\in\mathbb{Z}$ such that $d(T^Nu,T^Nv)>\delta$.
+\begin{Definition}
+  [\bf {(Sensitive Dependence on Initial Conditions)}]\label{Dfn_SDIC}\rm
+A dynamical system $T: U \to U$ is said to have \textit{sensitive dependence on initial conditions (SDIC)} if there exists a $\delta > 0$ such that for every $u \in U$ and in every neighborhood of $u \in U$ there exists a $v\in{U}$ and an integer $N:=N{(u,v)}\in\mathbb{Z}$ such that $d(T^Nu,T^Nv)>\delta$.
 \end{Definition}
@@
 \begin{enumerate}
-  \item First, the $\delta>0$ in the definition of SDIC is independent of $u$. 
-  \item Second, in every neighborhood of $u$, we may not necessarily find all points $v$ in the neighborhood distinct from $u$ that would separate from the forward iterates of $u$. 
-  \item  Finally, $N$ depends upon $u$ and $v$ chosen, and their iterates may not separate forever (i.e., for all $n>N$) and we allow their iterates to get arbitrarily close in the future. 
+  \item The $\delta>0$ in the definition of SDIC is independent of $u$. 
+  \item In every neighborhood of $u$, we may not necessarily find all points $v$ in the neighborhood distinct from $u$ that would separate from the forward iterates of $u$. 
+  \item  $N$ depends upon $u$ and $v$ chosen, and their iterates may not separate forever (i.e., for all $n>N$) and we allow their iterates to get arbitrarily close in the future. 
 \end{enumerate}
@@
 
-\begin{Example} \rm
-  The logistic map (LM), a recursion relation of the form $x_{n+1}=rx_n(1-x_n)$ where $x_n\in[0,1]$ and $r\leq{4}$ is a classic example used to illustrate the chaotic behaviour that a system may exhibit.
+\begin{Example}\label{ex_LM} \rm
+  The logistic map (LM), a recursion relation of the form $x_{n+1}=rx_n(1-x_n)$ where $x_n\in[0,1]$ and $r\leq{4}$, is a classic example used to illustrate the chaotic behaviour that a system may exhibit.
 When $r=4$, the one-dimensional recurrence relation $x_{n+1}=4x_n(1-x_n)$ can be used as the kindergarten's model to exhibit the presence of SDIC.
-    \ednote{B:edited and removed unnecessary citation}
+%\ednote{B:edited and removed unnecessary citation}
   In Graph~\ref{fig:log_sdic} below, we plot the first 50 iterates of the LM for $r=4$ and two slightly different initial values $x_0$. 
+  The values of $x_n$ are plotted agains time $n$  where they have initial values $x_0=0.2$ (blue) and $x_0=0.20001$ (red).
+Initially the two trajectories overlap, but they diverge completely at $n=12$ whereafter they follow distinct paths.
 
@@
         \centering
-        \captionof{figure}{Figure generated for Logistic Map $x_{n+1}=rx_n(1-x_n)$ with $r=4$ to exhibit the presence of SDIC.
-Plotted are the values $x_n$ against time $n$ for the the first 50 step with initial values $x_0=0.2$ (blue) and $x_0=0.20001$ (red).
-Initially the two trajectories overlap, but they diverge completely at $n=12$ whereafter they follow distinct paths.}
+        \captionof{figure}{Figure generated for Logistic Map $x_{n+1}=rx_n(1-x_n)$ with $r=4$ to exhibit the presence of SDIC. }
    \label{fig:log_sdic}
       \end{figure}
-  \ednote{B: Caption and figure adjusted}
-  % For an intuitive sense of the fact that the LM exhibits SDIC, consider again the definition and substitute specific values.
-Let $\delta=0.05$ and for purposes of illustration consider the value $u=0.4$.
-The green band in~\ref{fig:log_sdic} shades the region $[u-\delta, u+\delta]$ and we can see that regardless the value $v$  and in every neighborhood of $u \in U$ there exists a $v\in{U}$ and an integer $N:=N{(u,v)}\in\mathbb{Z}$ such that $d(T^Nu,T^Nv)>\delta$
-  % specific point, say $u=0.4$.
-Draw a green neighbourhood $[0.35,0.45]$ of diameters $\delta=0.02$ around the point.
-It is easy to see that every neighbourhood around $u=0.4$ will contain some $v_1, v_2\in[0,1]$ and natural number $N_1, N_2 \in\mathbb{N}$ respectively s.t.
-the distance $d(T^Nu_1,T^Nv)>\delta=0.05$ .
 \end{Example}
 
-\begin{Example} \rm
+\begin{Example}\label{ex_DP} \rm
   A further example pertains to the simple physical system of the double pendulum -- a pendulum with another pendulum attached to its head given an initial position and angular velocity (see the setup of the pendulum in Figure~\ref{fig:dp_setup}).
@@
   [\bf {Topologically Transitive}]\label{Dfn_TopolTrans}\rm
-	A dynamical system $F: U \to U$ is topologically transitive if for any pair of nonempty open sets $E_1$ and $E_2$ there exists a $n\in\mathbb{N}$ such that $T^n(E_1) \cap E_2 \not= \emptyset$. 
+	A dynamical system $F: U \to U$ is topologically transitive if for any pair of nonempty open sets $E_1$ and $E_2$ there exists a $n\in\mathbb{N}$ such that $$T^n(E_1) \cap E_2 \not= \emptyset$$. 
 \end{Definition}
@@
   [\bf {Meager Set}]\label{Dfn_Meager Set}\rm
-A subset of a topological space $U$ is said to be a meager set if it can be written as a countable union of sets of with empty interior. The complement of the a meager set is set to be a residual set.
+A subset of a topological space $U$ is said to be a meager set if it can be written as a countable union of sets of with empty interior.
+The complement of the a meager set is said to be a residual set.
 \end{Definition}
@@
 
-\begin{Example} \rm
-  The standard tent map $Tu=1-|2u-1|$ defined on $[0,1]$ is a well-known example of a dynamical system satisfying these three properties.
+\begin{Example}\label{ex_standardTentMap}\rm
+  The standard tent map $Tu=1-|2u-1|$ defined on $[0,1]$ is a well-known example of a dynamical system satisfying these three properties.  We now reason as to why this is true.
 
-  We now reason as to why this is true.
+  \begin{proof}
     The graph of the map $T$ is piecewise linear with two straight lines, one connecting the points $(0,0)$ and $(\frac{1}{2},1)$ and the other connecting $(\frac{1}{2},1)$ with $(1,0)$.
@@
     This follows because any fixed point of $T^k$ is also a periodic point, and we have found this in an arbitrary interval.
-
+  \end{proof}
 \end{Example}
@@
 %Composite functions
-To understand the diagram, we note that if we travel `right and then down', the diagram instructs us to use $T$ (top arrow right) first, followed by $\phi$.
-(right arrow downwards).
+To understand the diagram, we note that if we travel `right and then down', the diagram instructs us to use $T$ (top arrow right) first, followed by $\phi$ (right arrow downwards).
 Consequently, the pathway amounts to finding $\phi(T(u))$.
@@
 
-\begin{Definition}\rm  
-[\bf {Conjugacy \& Semi-Conjugacy}]\label{Dfn_Conjugate}\rm
+\begin{Definition}
+  [\bf {Conjugacy and Semi-Conjugacy}]\label{Dfn_Conjugate}\rm
    Consider the dynamical systems $T:U\to{U}$, $S:V\to{V}$ and suppose the relationship $\phi \circ T=S\circ \phi$ holds.

### 2022-01-18 projectReport/chapter3.tex
M final edits to Ch.5 and Michelle edits up to Ch.4

@@
 First we establish the method of Takens delay embedding. 
-In this method, one considers a discrete-time dynamical system defined on a `nice' space (a smooth manifold)\ednote{B: Should this be further explained?} that can be obtained as time-$K$ map of a continuous time-dynamical system, concepts formalised below.
+In this method, one considers a discrete-time dynamical system defined on a `nice' space (a smooth manifold)\ednote{B: Where can I find a citation for this?} that can be obtained as time-$K$ map of a continuous time-dynamical system, concepts which are formalised below.
 We make observations from such a system, i.e., we observe the evolution of $\theta(w_0)$ (where $w_0$ represents the initial state) by examining  a finite set of values of $u_n :=  \theta(w_n) = \theta(Tw_{n-1})$.
@@
 We define first the concepts of homeomorphism and embedding:
-\begin{Definition}\rm
+\begin{Definition}
   {\bf(Homeomorphism)}\label{Dfn_homeo}\rm
@@
 Takens Theorem states a result establishing a relationship between the observed and underlying dynamical systems by showing that the concatenation of a sufficiently large number of previous observations into a vector will, under certain conditions, generate a map between the vectors from the respective systems.
-We formulate the theorem from~\cite{takens1981detecting}.
+We formulate the theorem from~\cite{takens1981detecting} after establishing preliminary definitions.
+
+\ednote{It was suggested that I define these two concepts before stating the theorem.}
+\begin{Definition}
+  [\bf{Manifold, Compact Manifold}]\label{Dfn_compactManif}\rm
+  A topological space is called a manifold if every point in the space has a neighbourhood homeomorphic to $\mathbb{R}^n$. Succinctly, a manifold is a locally Euclidean topological space.
+  If a manifold is compact as a space topologically, then it is termed a compact manifold
+\end{Definition}
+
+\begin{Definition}
+  [\bf{Diffeomorphism}]\label{Dfn_diffeo}\rm
+  A smooth, differential map $f:X\to{Y}$ with a differentiable inverse is said to be a diffeomorphism.
+By `smooth' we mean at least $C^2$, where $C^2$ refers to any twice-differentiable function  with a continuous second derivative.
+\end{Definition}
 
@@
             $\Phi_{2d,\theta}(w) := (\theta(T^{-2d}w)\ldots,\theta(T^{-1}w),\theta(w))$
-            is a diffeomorphic embedding; by `smooth' we mean at least $C^2$.
+            is a diffeomorphic embedding.
 Consequently, there exists a map $F_\theta: \Phi_{2d,\theta}(W) \to \Phi_{2d,\theta}(W)$ defined by \[F_\theta: (\theta(T^{-2d}w),\ldots,\theta(T^{-1}w),\theta(w)) \mapsto
@@
 
-By generic we mean a residual set on a certain topology on appropriate spaces of functions (that we do not describe here) \ednote{B: Where can I find a citation for this?}. 
-By $C^2$, we make reference to a twice-differentiable function  with a continuous second derivative.
+By generic, we mean a residual set on a certain topology on appropriate spaces of functions (that we do not describe here) \ednote{B: Where can I find a citation for this?}. 
 In our scenario (and this remains important throughout) the input space $U$ is considered the attractor.~\label{attractor_U}
@@
 This requirement is not overly restrictive, as a great number of chaotic systems do have a surjective map.
-(Cite) \ednote{B:I remember you saying this in one of our sessions, but don't have a source for it.
-Do you have a specific reference?}
+\textbf{(Cite)} \ednote{B: I remember you saying this in one of our sessions, but don't have a source.
+Could you point me to a resource?}

### 2022-01-18 projectReport/chapter4.tex
checkpoint
M final edits to Ch.5 and Michelle edits up to Ch.4

@@
-In this chapter, we discuss results on the mapping of temporal data obtained from a discrete-dynamical system onto a different space through the notion of a driven dynamical system. 
+In this chapter, we discuss mathematical results relating to the mapping of temporal data obtained from a discrete-dynamical system onto a different space through the notion of a driven dynamical system.
 We also consider the conditions a driven system should possess to avoid adding distortion to its state-space representation.
-We then describe the \emph{single-delay dynamics} (SDD) of the system and finally recall conditions on the driven system so that the SDD are conjugate (or at least semi-conjugate) to the underlying system.
+We then describe the \emph{single-delay dynamics} (SDD) of the system and finally state conditions on the driven system so that the SDD are conjugate (or at least semi-conjugate) to the underlying system.
 The SDD can then be used to forecast and reconstruct the underlying system.
@@
 \begin{Definition}
-  [\bf Nonautonomous Dynamical System (adopted from~\cite{Manju_ESP})]\label{Dfn_NDS}\rm
+  [\bf Nonautonomous Dynamical System (as defined in~\cite{Manju_ESP})]\label{Dfn_NDS}\rm
   A nonautonomous dynamical system (NDS) on a space $X$ is simply a dynamical system comprising a family of maps ${\{g_n\}}_{n \in \mathbb{Z}}$ where $g_n(\ldotp):=g(u_n,\ldotp):X\to{X}$ is continuous.
@@
 Note that $x\in{X_n(\overline{u})}$ if and only $g$ has a solution $\{x_k\}$ for $x_n=x$ and input $\overline{u}$, a result proved in nonautonomous dynamical systems literature~\cite{manjunath2014dynamics, manjunath2013echo}.
-\ednote{B: Adjusted sentence and cited}
 
@@
 \end{Theorem}
+\vspace{-6mm}
 {\bf Proof.} May be found in~\cite{kloeden2011nonautonomous, manjunath2014dynamics, manjunath2013echo}
@@
 \end{Theorem}
+\vspace{-6mm}
 {\bf Proof.}  See~\cite[Th.3]{manjunath2013echo}.
@@
 and we refer to the map $H$ as a \emph{causal embedding}. 
-\ednote{B: edited the sentence here}
+%\ednote{B: edited the sentence here}
 
-When $g$ has the USP, the diagram below (adopted from~\cite{Manju_Nonlinearity}) illustrates the operation of the mappings $h$ and $H$.
+When $g$ has the USP, the diagram below (taken from~\cite{Manju_Nonlinearity}) illustrates the operation of the mappings $h$ and $H$.
 The mapping $h:\overleftarrow{U}\to{X_U}$ is also considered an observable as mentioned in the introduction of Chapter~\ref{ch3}).
@@
 \begin{Definition} \rm
-	We say a driven system $g$ \emph{causally embeds} a dynamical system $(U,T)$ if it satisfies the two properties: (i) a universal semi-conjugacy exists and  (ii)  $H_2(\cev{u}) := (h(r\cev{u}),h(\cev{u}))$ embeds the inverse-limit space $\widehat{U}_T$ in $X \times X$.
+	We say a driven system $g$ \emph{causally embeds} a dynamical system $(U,T)$ if it satisfies the two properties: 
+  \vspace{-8mm}
+\begin{enumerate}[noitemsep, label=\roman*.]
+  \item  a universal semi-conjugacy exists and  
+  \item  $H_2(\cev{u}) := (h(r\cev{u}),h(\cev{u}))$ embeds the inverse-limit space $\widehat{U}_T$ in $X \times X$. 
+\end{enumerate}
 \end{Definition}
@@
 It is imperative to take note of a subtlety here: we refer to the \emph{map} $H$ as a (causal) embedding, whereas the driven \emph{system} $g$ (causally) embeds another \emph{system} $(U,T)$.
-\ednote{B: inserted line}
+\ednote{Check this sentence again.}
 
@@
 
-So far it is not clear if the USP alone ensures the ability to causally embedding a dynamical system and we mention that a driving function $g$ cannot just be chosen in a frivolous manner.
+So far it is not clear whether the USP alone ensures the ability to causally embedding a dynamical system and we mention that a driving function $g$ cannot just be chosen in a frivolous manner.
 This could possibly complicate our work.
@@
 To obtain a suitably complex function $g$, it is thus desired that the reachable set of a driven system  be large enough to relate to the input. 
-This is true even for embedding  the inverse-limit space of a dynamical system. \ednote{B: Expand or refer?}
+This is even true for embedding  the inverse-limit space of a dynamical system. \ednote{B: Expand or refer?}
 
@@
 \end{Definition}
+
+As mentioned, $g$ can also introduce additional complexity to a solution.
+Consider the Figure~\ref{fig:memloss_conttime} where a RNN's solution was numerically simulated for randomly generated matrices $A$ and $B$ where $B$ has unit spectral radius.
+An input sequence (black) $(x_0,x_1,\ldots,x_{5000})$ was fed to the network and we plot it as well as the the coordinates of two simulated solutions in red (with parameter $a=1$, $\alpha=0.99$ where $g$ has the USP) and blue ($a=1$, $\alpha=1.05$ where $g$ does not have the USP).
+The red solution mimics the input sequence well in its variation, whereas the blue solution is disordered and changes a lot.
+In this we may say that the blue solution has introduced additional complexity to the system.
+\ednote{B: Added this paragraph}
 
@@
   \end{Theorem}
+  \vspace{-6mm}
   {\bf Proof.} See~\cite[Th.3]{Supp}
@@
 The UAP guarantees that all trajectories will converge to the same trajectory as time moves forward. 
-An illustration of this is shown in Fig.~\ref{fig:memloss_conttime} where the trajectory in red locks on to the supposed actual solution when the system has the USP.
+An illustration of this is shown in Fig.~\ref{fig:memloss_conttime} where the trajectory in red seems to lock on to the supposed actual solution in black when the system has the USP.
 
@@
   \centering
-\captionof{figure}{A coordinate of a simulated solution $(x_0,x_1,\ldots,x_{5000})$ of a RNN  plotted in red (with parameter $\alpha=0.99$ where $g$ has the USP) and blue ($a=1$, $\alpha=1.05$ where $g$ does not have the USP) while the matrices $A$ and $B$ are randomly generated, and $B$ has unit spectral radius.
+\captionof{figure}{Input sequence (black) and a coordinate of a simulated solution $(x_0,x_1,\ldots,x_{5000})$ of a RNN  plotted in red (with parameter $a=1$, $\alpha=0.99$ where $g$ has the USP) and blue ($a=1$, $\alpha=1.05$ where $g$ does not have the USP) for randomly generated matrices $A$ and $B$.
+$B$ has unit spectral radius.
 Reproduced from~\cite{Manju_Nonlinearity}.
@@
 \end{figure}
-\ednote{B: How can I explain the black line? I don't think this makes sense to the reader immediately.}
 
@@
 \end{Theorem}
+\vspace{-6mm}
 {\bf Proof.}  May be found as the proof of~\cite[Th.4]{Supp}
@@
   \centering
-  \captionof{figure}{Pictorial summary of results obtained up to now and how they overlap. Reproduced from \textbf{(Cite SIG-25May)} }
+  \captionof{figure}{Pictorial summary of results obtained up to now and how they overlap. Reproduced from~\cite{SIG25May} }
   \label{fig:pictorialSummary}
 \end{figure}
-\ednote{Ask M: Cite SIG-25May.}
+\ednote{Cited SIG25May here, but unsure if bibtex reference is correct}
 
@@
 \end{Theorem}
+\vspace{-6mm}
 {\bf Proof.}  See~\cite[Th. 3c]{manjunath2021universal}.
@@
 
-The graph in equation \ref{fig:comm_H_CET} is then extended to the diagram~\ref{Scomm_imp} below:
+The graph in equation~\ref{comm_H_CET} is then extended to the diagram~\ref{Scomm_imp} below:
 

### 2022-01-18 projectReport/chapter5.tex
M final edits to Ch.5 and Michelle edits up to Ch.4

@@
 
-From~\eqref{eqn_driving} we know that $g(u,x) = (1-a) x + a \tanh(Au + \alpha B x)$. Recall that $g$ is SI-invertible. We refer to $A$ as the input matrix and $B$ the reservoir matrix.
+From~\eqref{eqn_driving} we recall the equation of the RNN: $g(u,x) = (1-a) x + a \tanh(Au + \alpha B x)$.
+Recall that $g$ is SI-invertible when $A$ is invertible and has same dimension as $B$.
+We refer to $A$ as the input matrix and $B$ the reservoir matrix.
 
@@
 
-We implement a feedforward network, (a network with no memory, unlike the RNN)\ednote{B: Expanded sentence s.t.
-FFN is in more context} required to learn $\Gamma$ in the \emph{Python} programming language and making use of the \emph{Scipy} library, alongside the \emph{Keras} package in the \emph{Tensorflow} library \ednote{B: Is this something you want me to write in?
-M: We do not implement the RNN, but the FN that learns Gamma through the Tensorflow.
+We implement a feedforward network, (a network with no memory, unlike the RNN) required to learn $\Gamma$ in the \emph{Python} programming language and making use of the \emph{Scipy} library, alongside the \emph{Keras} package in the \emph{Tensorflow} library
+% \ednote{B: Is this something you want me to write in?
+M: We do not implement the RNN, but the FFN that learns Gamma through Tensorflow.
 We called the RNN + FN set up a RCN in Adriaan's paper.
@@
 
-Thirdly we prepare the network.
-Utilising a network $N$ neurons (the same value $N$ as in the dimensions of the matrices $A,B$).
+Thirdly, we prepare the network by setting up a network with $N$ neurons (the same value $N$ as in the dimensions of the matrices $A,B$).
 Learning of $\Gamma$ is done using the Adam Optimiser where the means square error (MSE) loss function is optimised.
@@
 
-We then reach the prediction step. Making use of $g$'s SI-invertibility and $g$ as defined in~\eqref{eqn_driving}, we have the following expression for $u_n$ replicated from~\eqref{eqn_SI_RNN}.
-\begin{equation*}
-  u_n := A^{-1}\bigg(\overline{\tanh}^{^{-1}}\frac{1}{a}\Big(x_{n+1}-(1-a)x_n\Big) \bigg) - \alpha B x_n
-\end{equation*}
+Once $\Gamma$ is learnt then reach the prediction step, where we forecast $u_n$ from $(x_{n-1},x_{n})$.
+Note that  due to causal embedding $(x_{n-1},x_{n})$ is in one-to-one correspondence with the left-infinite orbit $(\ldots,u_{n-1},u_{n-1})$.
+Once $u_n$ is computed using $\Gamma$, we feed that to the driven system $g$ along with  $x_n$ to obtain $x_{n+1}$.
+Now we apply $\Gamma$ on $(x_n,x_{n+1})$ to forecast $u_{n+1}$ and the whole process repeats.
+Note that this corresponds to implementing the two model equations for learning from data~\eqref{eqns_from_data}.
+% We then reach the prediction step. Making use of $g$'s SI-invertibility and $g$ as defined in~\eqref{eqn_driving}, we have the following expression for $u_n$ replicated from~\eqref{eqn_SI_RNN}.
+% \begin{equation*}
+%   u_n := A^{-1}\bigg(\overline{\tanh}^{^{-1}}\frac{1}{a}\Big(x_{n+1}-(1-a)x_n\Big) \bigg) - \alpha B x_n
+% \end{equation*}
 
-The two model equations for learning from data~\eqref{eqns_from_data} allows us to determine what the state $x_{n+1}$ at time $n+1$ will be. 
-We specifically remind ourselves of \[x_{k+1}=\pi_2 \circ (\Gamma, \pi_2) \circ (\pi_2,g) (u_k,x_k)\]
+% The two model equations for learning from data~\eqref{eqns_from_data} allows us to determine what the state $x_{n+1}$ at time $n+1$ will be. 
+% We specifically remind ourselves of \[x_{k+1}=\pi_2 \circ (\Gamma, \pi_2) \circ (\pi_2,g) (u_k,x_k)\]
 
@@
 Then if we feed  the input values $u_n := \Phi_{\theta,2d}(\theta(w_{n}))$ to a driven system $g$ that is SI-invertible and has USP, the induced system $(Y_F,G_F)$ would be topologically conjugate to the inverse-limit space of
-$(\Phi_{\theta,2d}(W), F_\theta)$ as in Theorem~\ref{Thm_CET}, and hence one could forecast $u_n,u_{n+1},\ldots$.
-In other words, one could forecast  the values $\theta(w_n), \theta(w_{n+1}),\ldots$ by feeding these observations as delay coordinates.
-The advantage of feeding delay-coordinates to a driven system is that the embedding is stable in the sense of global dissipativity that we have described in the previous chapter, i.e., $Y_F$ is made to contained in an attractor of a mapping induced on $Y_F$ that is induced by $g$.
+$(\Phi_{\theta,2d}(W), F_\theta)$ due to Theorem~\ref{Thm_CET}, and hence one could forecast $u_n,u_{n+1},\ldots$.
+In summary, one could forecast  the values $\theta(w_n), \theta(w_{n+1}),\ldots$ by feeding these observations as delay coordinates since the input is again an orbit from a dynamical system.
+The advantage of feeding delay-coordinates to a driven system  over learning a map that describes the evolution of the delay coordinates is that the embedding is stable in the sense of global dissipativity that we have described in the previous chapter, i.e., there is an attractor $A$ of a map that implements $G_T$ on an expansion $Y_F^+$ providing some robustness to noise.
 
-We note that the twin equations together~\ref{eqns_from_data} generate a model from data.
-For a chaotic system, a small amount of noise due either to the input-noise or computational error would cause the solutions from two accurate models with the same initial conditions to  diverge as they are not identical.
-\ednote{B: edited this sentence} However, statistically, they would be the same since due to ergodic theory all typical orbits have the same visiting frequency to any region of the phase space of the dynamical system.
-\ednote{B: I'll need to recap this sentence with you before presenting as it's new to me.} To illustrate that we plot the densities of the trajectories of the actual and predicted solutions.
+Now discuss on another practical issue with forecasting chaotic dynamical systems.  
+We note that the twin equations~\ref{eqns_from_data} generate a model from data.
+For a chaotic system, a small amount of noise due either to the input-noise or computational error would cause the solutions from even an accurate model with the same initial conditions to diverge.
+However, statistically, a chaotic system would behave well in the sense that the visiting frequency or the average time spent by an orbit in any region of the phase space of the dynamical system is identical for almost all initial conditions (almost all is specified as a measure), a result in ergodic theory (\cite{wikiErgodicTheory}) a topic beyond the scope of this thesis.
+To illustrate the accuracy of the twin  equations~\ref{eqns_from_data} in modeling the data we plot the densities of the trajectories of the actual and predicted solutions in~\ref{noisydp_success_density, fig:dp_success_density}.
 
-Our implementation of the reservoir (A) and input (B) matrices in the RNN makes use of sparse matrices.
-We discuss three specifics relating to this concept.
+In our implementation, the input and reservoir  matrices $A$ and $B$ in the RNN are sparse matrices.
+This sparsity is not essential but seems to perform slightly empirically better than choosing the matrices arbitrarily.
+We also discuss three specifics relating to this concept.
 Firstly, the matrices are assigned a variable density in terms of the amount of non-zero entries and are scaled to have values within the range $[-1,1]$.
@@
 In our experiments, we generally opted for a density ratio of 10\%-25\%.
-Secondly, we opted to enforce the requirement that both matrices be well-conditioned, i.e.
+Secondly, we opt to enforce the requirement that both matrices be well-conditioned, i.e.
 that they have a condition number of \~1.
 The condition number of a matrix $M$ (denoted $cond(M)$) is the product of the matrix norms of the matrix $M$ and its inverse $M^{-1}$ and is a measure of the sensitivity of output of a matrix multiplication when an input is changed.
-\textbf{(Cite.)}\ednote{Ask M: Cite specific resource?} By ensuring a condition number close to 1, we also ensure that our driven function $g$, depending on the matrices $A$ and $B$, will not be sensitive to adjustments in the input.
+We do not have any theoretical justifications for this.
+
+%By ensuring a condition number close to 1, we also ensure that our driven function $g$, depending on the matrices $A$ and $B$, will not be sensitive to adjustments in the input.
 In so doing we then also guarantee another form stability.
 \ednote{B: Not complete enough.
-Edit further and cite}
+Edit further and cite M: Please feel free to cite a wikipedia page}
 
-Lastly, we ensure the matrix $B$ has unit spectral radius.
+% Lastly, we ensure the matrix $B$ has unit spectral radius.
 This is necessary due to the Causal Embedding Theorem~\ref{Thm_CET} which can then guarantee the existence of semi-conjugacy between the systems $(Y_T, G_T)$ and $(\widehat{U}_T, \widehat{T})$.
 In fact, the theorem actually requires that $\alpha{B}$ have spectral radius $<1$, so we impose the restriction $\alpha<1$ and consider only the matrix $B$'s spectral radius from thereon.
+Lastly, we ensure the matrix $B$ has unit spectral radius so that the variable $\alpha$ in the RNN equation \eqref{eqn_driving} can be used to know if the USP is satisfied.
+It is a  feature that a matrix with a spectral radius $<1$ generally has the USP \cite{yildiz2012re}.
 
-Note that to learn $\Gamma: (x_{n-1},x_{n}) \mapsto u_n$ through our feedforward network, we could learn the principal components of $(x_{n-1},x_{n})$ as well. 
-We employ this since it provides slightly better results empirically but is not essential.
-
-Principal components are used only for a more efficient state representation to possibly reduce learning errors; it is not for a lossy approximation since we use all principal components \textbf{(Cite.)}\ednote{Ask M: Is there a specific resource to cite?}.
+We remark on one more aspect that helps produce slightly better empirical results.
+To learn $\Gamma: (x_{n-1},x_{n}) \mapsto u_n$ through our feedforward network, we could feed the principal components.
+Principal components are used only for a more efficient state representation to possibly reduce learning errors; it is not for a lossy approximation since we use all principal components.
 More explicitly, denote $X_{1:N}$ as the matrix with the first $N$ states of the network data represented by row vectors.
-If $X_{1:N}=U\Sigma P^T$\ednote{Expand here} represents the singular value decomposition of $X_{1:N}$ \ednote{Ask: Is there a specific resource to cite?}, we then denote the principal component matrix $P$, and the principal components are given by
+If $X_{1:N}=U\Sigma P^T$ here represents the singular value decomposition ($U$ and $V$ are unitary matrices and $\Sigma$ is a diagonal matrix) of $X_{1:N}$ and we refer to \cite{wikiSVdecomp}.
+We then denote the principal component matrix $P$, and the principal components are given by
  $Z_{1:N}=X_{1:N}P$.
@@
 \section{Experimental Results}
-In this section we showcase numerical results pertaining to 3 attractors that were simulated using the methodology described.
+In this section we showcase numerical results pertaining to 3 attractors that were simulated using the theory and methodology described.
 We first consider a simple physical system that exhibits complicated dynamics.
@@
 (See below the Figure~\ref{fig:dp_setup})
-This system is a classical example of a system exhibiting chaos and is often used as explanation of the very well-known Butterfly Effect.
-\ednote{B: Originally though a reference here was needed, but no longer think so.} Consider again the Figure~\ref{fig:dp_sdic}.
+This system shows chaos strikingly. Consider again Figure~\ref{fig:dp_sdic}.
 The system has four variables: angle and angular velocity of each pendulum rod. An ideal system is assumed where the rods are massless and no friction in the pivots are considered.
@@
 
-The scalar equations of motion are derived using Newtonian physics and are reproduced from~\cite{DPFormulas} in~\ref{eqns_dp}.
+The scalar equations of motion are derived using Newtonian physics and are reproduced from~\cite{DPFormulas} in Equations~\ref{eqns_dp}.
 Each pendulum component has an associated mass of the head, $m$, and length of the rod, $L_i$.
@@
 
-The second-order system is converted to a first-order ODE system and solved using the Runge-Kutta numerical method.  
+% The second-order system is converted to a first-order ODE system and solved using the Runge-Kutta numerical method.  
 Here we chose the values $m_1=2.0$, $m_2=2.0$, $L_1=1.5$ and $m_2=1.5$ with initial velocities $v_1=1.5, v_2=-2$ for the respective pendulum heads.
-Undamped DP data-sets were generated and used in this project due to the requirement that the map $T$ be surjective that we must not forget. 
-An example of this would be the case of a DP exhibiting energy-loss (i.e.
+With a change of variables the second-order system can be 
+written as a system of first order equations with four variables, where two variables devoted to each pendulum denote the $x$ and $y$ coordinates of the mass (the lengthy equations are not shown here).
+The ODE system is then solved using the fourth-order Runge-Kutta numerical method.
+
+% Undamped DP data-sets were generated and used in this project due to the requirement that the map $T$ be surjective that we must not forget. 
+% An example of this would be the case of a DP exhibiting energy-loss (i.e.
 where the map $T$ is not surjective).
 In such a case, the attractor is a single point in the phase space when the double pendulum comes to rest.
-It must be mentioned once again that if we use data outside the attractor, learning is not reliable. 
-To see this, consider Figure~\ref{fig:damped_pendulum} and~\ref{fig:dp_notsurjective}.
+% It must be mentioned once again that if we use data outside the attractor, learning is not reliable. 
+% To see this, consider Figure~\ref{fig:damped_pendulum} and~\ref{fig:dp_notsurjective}.
+
+We need the assumption the system is not damped so that the map $T$ resulting from the discretization of the system  is surjective.
+If we assume damping surjectivity is lost on any phase space with more than one point.
+An example of this would be the case of a DP exhibiting energy-loss. In such a case, the attractor is a single point in the phase space when the double pendulum comes to rest. 
+For such a system, it may be noted that if we use data outside the attractor, learning is not reliable. To see this, consider the plots ~\ref{fig:damped_pendulum} and ~\ref{fig:dp_notsurjective}.
 
@@
 \end{figure}
-\ednote{Can I say in the caption \emph{Parameter-values are exactly as in the successful case discussed in the paragraph directly afterwards.}}
+% \ednote{Can I say in the caption \emph{Parameter-values are exactly as in the successful case discussed in the paragraph directly afterwards.}}
 
@@
 The remaining 15000 datapoints were used to train a network with 128 neurons. 
-The network was initialised to contain 16 hidden layers each with a layer dimension of 64, activated with the tanh activation function.
+The network was initialised to contain 16 hidden layers each with a layer dimension of 64, activated with the $\tanh$ activation function.
 Training was accomplished using 256 epochs with batch sizes 128 and a delay of 1.
-See~\ref{fig:dp_success_traj} for results pertaining to the trajectory taken by the DP's components and then~\ref{fig:dp_success_density} to witness associated densities of predicted versus actual values.
+See Fig.~\ref{fig:dp_success_traj} for results pertaining to the trajectory taken by the DP's components and then Fig.~\ref{fig:dp_success_density} to witness associated densities of predicted versus actual values.
 
@@
 We now consider an experiment where noise is added to the system and find it to be incredibly robust.
-We employ the same parameters as before and add noise from a normal-distribution with mean zero and standard deviation equal to 0.1, translating to roughly 39dB's of noise.
-(For the signal-to-noise ration we adopt the formula $\text{SNR}_{dB}=10\text{log}_{10}\big(\frac{P_\text{{signal}}}{P_{\text{noise}}}\big)$ where $P_\text{{signal}}$ refers to the variance observed in the normalised data and $P_{\text{noise}}$ represents the variance of the added noise.)
+We employ the same parameters as in the preceding experiment and add noise from a normal-distribution with mean zero and standard deviation equal to 0.1 which gives a signal-to-noise ratio of 39dB.
 
+For the signal-to-noise ration we adopt the formula $\text{SNR}_{dB}=10\text{log}_{10}\big(\frac{P_\text{{signal}}}{P_{\text{noise}}}\big)$ where $P_\text{{signal}}$ refers to the variance observed in the normalised data and $P_{\text{noise}}$ represents the variance of the added noise.
+\ednote{B:Added this sentence.}
 Once again consider the actual and predicted trajectories for the noisy DP dataset in Figure~\ref{fig:noisydp_success_traj}.
@@
 
- Lastly, a run was done where only a single coordinate (the $y$-coordinate of the lower pendulum head) was fed into the system and we can illustrate through the obtained success that the coordinate contains all the information in its past states necessary to forecast its future evolution.
- \textbf{Insert graphs here. Had to quickly change something here.}
+%  Lastly, a run was done where only a single coordinate (the $y$-coordinate of the lower pendulum head) was fed into the system and we can illustrate through the obtained success that the coordinate contains all the information in its past states necessary to forecast its future evolution.
+%  \textbf{Insert graphs here. Had to quickly change something here.}
 
@@
 
-The behaviour of the system with a $\beta$-parameter value of 0.1056 was simulated by scaling it to fit inside the interval $[-1,1]$. Two cases were considered: noise-free and perturbed by noise. 
-Noise was generated  by adding data points taken from a normal distribution with mean 0 and standard deviation 0.05, which to translates approximately 33dB. 
+The behaviour of the system with a $\beta$-parameter value of 0.1056 was simulated by scaling the data from it to fit inside the interval $[-1,1]$.
+Two cases were considered: one noise-free and another perturbed by noise.
+Noise was generated  by adding data points taken from a normal distribution with mean 0 and standard deviation 0.05, which translates approximately to a signal-to-noise ration of 33dB. 
 ($\text{SNR}_{dB}$ is calculated exactly as for the DP-data above)
@@
 
-Another 2-equation discrete-time map named the Fractal Dream Attractor, or more commonly known as the Pickover Map (first discovered by Clifford A.
-Pickover and discussed in his fascinating book `Chaos in Wonderland'\cite{PickoverChaos}), was also considered.
+Another two-equation discrete-time map named the Fractal Dream Attractor, or more commonly as the Pickover or Clifford Map (first discovered by Clifford A.
+Pickover and discussed in his fascinating book``Chaos in Wonderland" \cite{PickoverChaos}), was also considered.
 \begin{eqnarray}\label{eqns_clifford}
@@
 The behaviour of the system with a parameter values of $x_0=0$, $y_0=0$, $a = -1.7$,  $b = 1.8$, $c = -1.9$ and $d = -0.4$ was examined. 
-Data was scaled to fit inside the interval $[-1,1]$. Only the noise-free case was investigated. Depictions of the empirical results available in Figure~\ref{fig:Clifford}.
- A sequence of 25000 observations was fed into the system with 15000 entries discarded to simulate the network's loss of memory. The next 1000 steps were predicted and compared with true data.
+Data was scaled to fit inside the interval $[-1,1]$. Only the noise-free case was investigated. 
+Depictions of the empirical results available in Figure~\ref{fig:Clifford}.
+ A sequence of 25000 observations was fed into the system with 15000 entries discarded \ednote{M: Why did you have to discard 15000?} to simulate the network's loss of memory.
+The next 1000 steps were predicted and compared with true data.
+
  The RNN was initialised exactly as for Thomas' Cyclically Symmetric Strange attractor: the dimension of the network is set equal to 500 neurons and 12 hidden layers of dimension 64 each were initialised.
@@
 
-\section{Reducing Functional Complexity in Noninvertible Maps}
-\ednote{section header is overfull}
+\section{Functional Complexity Reduction in Noninvertible Maps}
 % Besides being able to learn the action of the Koopman operator exactly, our setup has other advantages over selecting or optimizing the choice of observables as in the Extended Dynamic Mode Decomposition (EDMD) algorithm (see, [21, 22]).
@@
 
-In the case of non-invertible maps, we note that the map $G_T$ is a homemorphism although $T$ is not.
+% In the case of non-invertible maps, we note that the map $G_T$ is a homemorphism although $T$ is not.
 Further, empirically it is found that $G_T$ is much more linear than $T$ as measured by what is called the multidimensional correlation coefficient $\rho$.
@@
 We state teh details of computing this coefficient between
+The map $G_T : Y_T \to Y_T$ is a homeomorphism since it is conjugate to the homeomorphism $\widehat{T}$ even though $T$ may be a non-invertible map.
+We remark that in the case of a non-invertible map, empirically it is found that $G_T$ is much more linear than $T$ as measured by what is called the multidimensional correlation coefficient $\rho$.
+This amplification in linearity  or the increased value of the multidimensional correlation coefficient $\rho$  is also true for invertible maps and we refer to empirical results in \cite{Supp}.
+Here we consider only one non-invertible map pertaining to the Clifford attractor to add to the results in \cite{Supp}.
+This coefficient is a generalization of the Pearson correlation coefficient for real-valued random variables to random vectors (e.g., \cite{puccetti2019measuring}).
+We state the details of computing this coefficient between
 $(x_{n-1},x_n)$ and $G_T(x_{n-1},x_n)$. 
@@
 
-\begin{center}
+% \begin{center}
   \begin{table} 
-      \scalebox{0.75}
+      \scalebox{0.65}
       \centering
@@
           % \hline         
-
                \midrule  
@@
            \bottomrule
-                \end{tabular}} %\ednote{http://paulbourke.net/fractals/peterdejong/}
- \caption{Following~\cite{manjunath2021universal}, we determine the multidimensional correlation coefficients $\rho$ to evidence a general reduction in the functional complexity of the map $G_T$ emerging due to a RNN.
+                \end{tabular}}\label{Table_FC}%\ednote{http://paulbourke.net/fractals/peterdejong/}
+     \caption{
+              Following~\cite{manjunath2021universal}, we determine the multidimensional correlation coefficients $\rho$ to evidence a general reduction in the functional complexity of the map $G_T$ emerging due to a RNN.
 Rows correspond to distinct dynamical systems considered in the experiments
               The second column corresponds to the amount of artificial neurons (the dimension of $X$) in the implemented RNN.
-      The last two columns comprise numerical estimates of $\rho$ for the relevant vectors. The linear association between $u_n$ vs $T(u_{n})$, as well as the linear relationship $\begin{bmatrix}
+              The last two columns comprise numerical estimates of $\rho$ for the relevant vectors. The linear association between $u_n$ vs $T(u_{n})$, as well as the linear relationship 
+              $\begin{bmatrix}
                 x_{n-1}\\
                 x_n
-          \end{bmatrix}$ vs $G_T \left( \begin{bmatrix}
+            \end{bmatrix}$ vs 
+            $G_T 
+            \left({\begin{bmatrix}
                 x_{n-1}\\
                 x_n
-          \end{bmatrix}\right)$ may be contrasted by the reader. 
-          Finally, the RNN's dimension (from~\eqref{Seq_RNN}\ednote{Is this reference correct?}) is provided to exhibiting that an increase in dimension of the driven system will 
-typically result in a map $G_T$ with less functional complexity.}.
+            \end{bmatrix}}\right)$ 
+              may be contrasted by the reader. 
+              Finally, the RNN's dimension (from \eqref{Seq_RNN})
+              is provided to exhibiting that an increase in dimension of the driven system will 
+              typically result in a map $G_T$ with less functional complexity.}
       \end{table}\label{tbl_attractorsPearson}
-\end{center}
+% \end{center}
+
+In summary, the numerical results not only corroborate the theory of employing the notion of causally embedding a dynamical system for data-driven modeling by showing both long-term topological consistency (the attractor reconstruction) and statistical consistency (the densities of the actual and predicted match).

### 2022-01-19 projectReport/chapter2.tex
Rereading the project. On 5.4

@@
   [\bf Attractor]\label{Dfn_Attractor}\rm
-  Let $T: U \to U$, where $U$  is a metric space with metric $d$. A compact subset $A \subset U$ is said to be an attractor if it satisfies the three conditions: 
+  Let $T: U \to U$, where $U$  is a metric space with metric $d$. A compact subset $A \subset U$ is said to be an attractor if it satisfies the three conditions.
   \vspace{-8mm}
@@
 
-To this end, we make 3 remarks:
+To this end, we make 3 remarks.
 \vspace{-5mm}
@@
 When $r=4$, the one-dimensional recurrence relation $x_{n+1}=4x_n(1-x_n)$ can be used as the kindergarten's model to exhibit the presence of SDIC.
-%\ednote{B:edited and removed unnecessary citation}
   In Graph~\ref{fig:log_sdic} below, we plot the first 50 iterates of the LM for $r=4$ and two slightly different initial values $x_0$. 
@@
   [\bf {Devaney's Chaos}]\label{Dfn_ChaosDec}\rm
-	A dynamical system $T: U \to U$ is said to exhibit chaos in the sense of Devaney if it satisfies the three properties:
+	A dynamical system $T: U \to U$ is said to exhibit chaos in the sense of Devaney if it satisfies the three properties.
 	\vspace{-5mm}
@@
     Given any point $u$ and an open neighborhood $W \subset[0,1]$ of $u$, we can find $k\in\mathbb{Z}$ large enough so that $T^k(W) = [0,1]$ because we can accommodate a tent whose base is contained in $W$.
-% \ednote{B: This last phrase doesn't make sense to me.
-M: The number of tents grow with $k$ and hence this is obvious; I explained in one of our meetings.
-Does it make sense?}
     This implies the existence of two points $w_1$ and $w_2$ in $W$ so that $d(T^kw_1,T^kw_2)= 1$, and by the triangle inequality  at least one of the inequalities $d(T^ku,T^kw_1)> \delta$ or $d(T^ku,T^kw_2)> \delta$ hold when $\delta\in (0,\frac{1}{2})$.
@@
 
-We now turn our attention to the subject of conjugacy that describes when two dynamical systems are equivalent dynamically.
+We now turn our attention to the subject of conjugacy to describe when two dynamical systems are dynamically equivalent.
 
-To show topological similarity (or sameness) between two metric or topological spaces, one needs to establish a homeomorphism between the two spaces. 
-However, in the study of dynamical systems defined on two spaces, establishing a homeomorphism does not indicate that the systems are dynamically related in any way.
+To show topological similarity (or sameness) between two metric or topological spaces, one needs to establish a homeomorphism between these two spaces. 
+However, in the study of dynamical systems, establishing a homeomorphism does not indicate that the systems are dynamically related in any way.
 For instance, the maps $Tu=u^2$ and $Tu=1-|2u-1|$ defined on $[0,1]$ have totally different behaviour. 
-Therefore, to find dynamically similar systems, one must establish a dynamical equivalence that we illustrate in a simple commutativity diagram below~\ref{eqn_conjugacy}.
+Therefore, to find dynamically similar systems, one must establish a dynamical equivalence as illustrated in the simple commutativity diagram below~\ref{eqn_conjugacy}.
 
-% Giving errors
-\begin{equation}  \label{eqn_conjugacy}
-%    \[ 
-    \psset{arrows=->, arrowinset=0.25, linewidth=0.6pt, nodesep=3pt, labelsep=2pt, rowsep=0.7cm, colsep = 1.1cm, shortput =tablr}
- \everypsbox{\scriptstyle}
- \begin{psmatrix}
-U & U\\%
-V & V.
- %%%
-%  \ncline{1,1}{1,2}^{T} 
-%  \ncline{1,1}{2,1} <{\phi}
-%   \ncline{2,1}{2,2}^{S}
-%  \ncline{1,2}{2,2} > {\phi}
- \end{psmatrix}
-% \]
+\begin{equation}\label{eqn_conjugacy}
+  \begin{tikzcd}
+    U && U \\
+    V && V \\
+    \arrow["T", from=1-1, to=1-3]
+    \arrow["\phi", from=1-3, to=2-3]
+    \arrow["\phi"', from=1-1, to=2-1]
+    \arrow["S"', from=2-1, to=2-3]
+  \end{tikzcd}
   \end{equation}
@@
 %Composite functions
-To understand the diagram, we note that if we travel `right and then down', the diagram instructs us to use $T$ (top arrow right) first, followed by $\phi$ (right arrow downwards).
+To understand this diagram, we note that if we travel `right and then down', the diagram instructs us to apply $T$ (top arrow right) first, followed by $\phi$ (right arrow downwards).
 Consequently, the pathway amounts to finding $\phi(T(u))$.
@@
 If $\phi:U\to{V}$ is a homeomorphism, then $S$ is said to be conjugate to $T$ and $\phi$ is called a conjugacy.
-If we relax the criterion on $\phi$ and merely require $\phi$ to be continuous and surjective where $\phi:U\to{V}$, then $\phi$ is a semi-conjugacy between $T$ and $S$ where T has domain $U$ and S has domain $V$; S is said to be semi-conjugate to T.
+If we relax the criterion on $\phi$ and merely require $\phi$ to be continuous and surjective where $\phi:U\to{V}$, then $\phi$ is a semi-conjugacy between $T$ and $S$, where T has domain $U$ and S has domain $V$; S is said to be semi-conjugate to T.
 \end{Definition}
 
-\ednote{B: Does defining explicitly the domain and codomain of $\phi$ make it clear enough?
+% \ednote{B: Does defining explicitly the domain and codomain of $\phi$ make it clear enough?
 M: Saying $\phi$ is then a semi-conjugacy between $T$ and $S$ could be ambiguous as it does not specify which system is on the top in the commutativity.
@@
 However when $S$ is semi-conjugate to $T$ with $\phi:U\to{V}$ a many-to-one mapping, the dynamics on $V$ provide merely a coarse-grained description of the dynamics on $U$~\cite{de2013elements}.
-When $S$ is semi-conjugate to $T$, it is also common to call $S$ a \emph{factor of $T$}, or conversely that $T$ is an \emph{extension of $S$}.
+When $S$ is semi-conjugate to $T$, it is common to call $S$ a \emph{factor of $T$}, or conversely that $T$ is an \emph{extension of $S$}.
 In essence, an extension (see,~\cite{de2013elements}) is a more extensive system capturing all of the essential dynamics of its factor.
@@
 It is a very hard or nearly an impossible task to establish the existence of a conjugacy or a semi-conjugacy $\phi$ between two systems~\cite{devaney2018introduction}.
-However, one can verify that a function $\phi$ satisfies the commutativity diagram~\ref{eqn_conjugacy}.
+However, one can more easily verify that a function $\phi$ satisfies the commutativity diagram~\ref{eqn_conjugacy}.
 
 \begin{Example}\rm
-  For example $\phi(x)=\sin{\big(\frac{\pi}{2}x\big)}^2$ is a conjugacy between two systems $Tu=1-|2u-1|$ and $Sv=4v(1-v)$  defined on $U=[0,1]$.
+  $\phi(x)=\sin{\big(\frac{\pi}{2}x\big)}^2$ is a conjugacy between two systems $Tu=1-|2u-1|$ and $Sv=4v(1-v)$  defined on $U=[0,1]$.
 
@@
 \]
-And it follows that $\phi$ is a conjugacy between the two systems.
+So $\phi$ is a conjugacy between the two systems.
 \end{Example}
 
-By establishing that two systems are conjugate (semi-conjugate) to one another, we have also shown that one may choose to work with one system instead of the second and still be guaranteed to obtain information on the latter.
-It is instrumental during the process of forecasting the future evolution of dynamical systems.
+By establishing that two systems are conjugate (semi-conjugate) to one another, we have simultaneously shown that one may choose to work with one system instead of the second and still be guaranteed information pertaining to the latter.
+This is instrumental during the process of forecasting the future evolution of dynamical systems.

### 2022-01-19 projectReport/chapter3.tex
Rereading the project. On 5.4

@@
-In this chapter, we acquaint ourselves with the question of forecasting dynamical systems with unknown underlying structures, state and discuss the shortcomings of the Takens delay embedding theorem and discuss various issues faced while forecasting.
+In this chapter, we acquaint ourselves with the question of forecasting dynamical systems that have unknown underlying structures, state and discuss the shortcomings of the Takens delay embedding theorem, and discuss various issues faced while forecasting.
 
 Consider a relatively simple learning problem:
-Given the sequence $(u_0, u_1, \ldots, u_m)$ for $m\in\mathbb{Z}$, a finite segment of an orbit of the map $T$ where $T$ is defined by the update equation $u_{n+1} = Tu_n$, forecast the values $u_{m+1}, u_{m+1}$ where the map $T$ is unknown, given that $u_m$.
+Given the sequence $(u_0, u_1, \ldots, u_m)$ for $m\in\mathbb{Z}$, a finite segment of an orbit of the map $T$ where $T$ is defined by the update equation $u_{n+1} = Tu_n$, forecast the values $u_{m+1}, u_{m+2}, \ldots$ where the map $T$ is unknown.
 
-A practical example of this would be the subsequent scenario: given the time-sequential coordinates of an object moving in space, predict the future positions of that object.
-We are very rarely, if at all, presented with a problem where the entire state information is available to us in the form $u_n$ at some timestep $n\in\mathbb{N}$.
-Consider then a more complicated learning problem:
+A practical example of this would be to predict the future positions of an object moving in space given the time-sequential coordinates of that object. 
+We are rarely, if ever, presented with a problem where the entire state information is available in the form $u_n$ at some timestep $n\in\mathbb{N}$.
+Hence, consider a more complicated learning problem.
 
-Suppose we only have the observations $\theta(u_0), \theta(u_1), \ldots, \theta(u_m)$ of the true system states $u$ in an unknown dynamical system $(U,T)$ and we wish to predict the values $\theta(u_{m+1}), \theta(u_{m+2})$ and so forth.
+Suppose only have the observations $\theta(u_0), \theta(u_1), \ldots, \theta(u_m)$ of the true system states $u$ in an unknown dynamical system $(U,T)$ are available and, we wish to predict the values $\theta(u_{m+1})$, $\theta(u_{m+2})$ and so forth.
 
-First we establish the method of Takens delay embedding. 
-In this method, one considers a discrete-time dynamical system defined on a `nice' space (a smooth manifold)\ednote{B: Where can I find a citation for this?} that can be obtained as time-$K$ map of a continuous time-dynamical system, concepts which are formalised below.
+We start by establishing the method of Takens delay embedding. 
+In this method, one considers a discrete-time dynamical system defined on a `nice' space 
+(a smooth manifold~\footnote{A manifold is understood to be a special topological space that is locally homeomorphic at each point to an open set of $\mathbb{R}^d$})
+that can be obtained as a time-$K$ map of a continuous-time dynamical system, concepts which are formalised below.
 We make observations from such a system, i.e., we observe the evolution of $\theta(w_0)$ (where $w_0$ represents the initial state) by examining  a finite set of values of $u_n :=  \theta(w_n) = \theta(Tw_{n-1})$.
-The sequence $\{u_n\}$ represents a scalar time-series and intuitively $\theta$ represents a probe inserted into a bigger system which is itself only measuring/extracting a small part of the greater system state at time $t=n$.
+The sequence $\{u_n\}$ represents a scalar time-series.
+Intuitively $\theta$ represents a probe inserted into the larger system which is itself only measuring/extracting a small part of the greater system state at time $t=n$.
 Consider, for example, a thermometer erected to measure the ambient temperature in a local village.
-This measurement function, the thermometer, is capturing only a single aspect, the temperature, of a much grander dynamical system entailing the current weather of the surrounding area.
-Even more than that though- it is measuring a miniscule part of the global weather system.
+This measurement function- the thermometer- captures only a single aspect- the temperature- of a much grander dynamical system encompassing the current weather of the surrounding area.
+Even more than that- this probe is measuring a miniscule part of the global weather system.
 
@@
 
-However, the actual state $u$ of a system is seldom, if ever, fully known.
-The most one can do is to insert probes into a system to obtain partial information through the measurements taken.
-Moreover, the process of taking a measurement itself introduces two additional aspects which complicate the problem: \ednote{B: Perhaps this edit makes more sense?
-M: Maybe replace difficulty by saying that we need to consider two aspects.
-}
-\vspace{-8mm}
+As already mentioned, the actual state $u$ of a system is nearly never fully known.
+The most one can do is to place probes into a system to obtain partial information through the measurements taken.
+Moreover, the process of taking a measurement itself introduces two additional aspects which complicate the problem.
+\vspace{-6mm}
 \begin{enumerate}[noitemsep, label=\roman*.]
@@
 (Hence why we restricted our attention to discrete-time systems in the preceding section)
-  \item A probe will never be entirely accurate, and so the act of measurement introduces a certain measure of numerical noise/inaccuracy.
+  \item A probe will never be entirely accurate; the act of measurement introduces a certain measure of numerical noise/inaccuracy.
 \end{enumerate}
 
-From here we construct a multidimensional observable using the method of stacking previous observations, i.e., we create delay-coordinate map defined by
+Takens embedding theorem does not take into account the noise, and to describe it we construct a multidimensional observable using the method of stacking previous observations.
+The concept of an observable is understood in the sense of an observable as originally introduced by~\cite{takens1981detecting, genericObservableAeyels} to refer to a probe or measurement function inserted into the system.
+When speaking of stacking, we mean that a delay-coordinate map is defined by
 $\Phi_{k,\theta}(w) := (\theta(T^{-k}w)\ldots,\theta(T^{-1}w),\theta(w))$.  
-The concept of an observable is understood in the sense of an observable as originally introduced by~\cite{takens1981detecting, genericObservableAeyels} to refer to a probe or measurement function inserted into the system.
-Takens' theorem, in essence, refers to the fact that for a sufficiently large $k$, we can define a dynamical system on the space $\mathbb{R}^{k+1}$ with states $\Phi_{k,\theta}(w)$, $\Phi_{k,\theta}(Tw)$, $\Phi_{k,\theta}(T^2w)$, etc.
+In essence, Takens theorem refers to the fact that for a sufficiently large $k$, we can define a dynamical system on the space $\mathbb{R}^{k+1}$ with states evolving as $\Phi_{k,\theta}(w)$, $\Phi_{k,\theta}(Tw)$, $\Phi_{k,\theta}(T^2w)$, etc.
 This dynamical system is topologically conjugate to the unknown underlying system $(W,T)$.
@@
 
-We define first the concepts of homeomorphism and embedding:
-\begin{Definition}
-  {\bf(Homeomorphism)}\label{Dfn_homeo}\rm
+We define first the concepts of homeomorphism and embedding.
+\begin{Definition} \rm
+  [\bf {Homeomorphism}]\label{Dfn_homeo}\rm
   A homeomorphism is a function $f:Z\rightarrow Y$ between two topological spaces $Z$ and $Y$ that is continuous, bijective and has a continuous inverse. 
@@
 
-Takens Theorem states a result establishing a relationship between the observed and underlying dynamical systems by showing that the concatenation of a sufficiently large number of previous observations into a vector will, under certain conditions, generate a map between the vectors from the respective systems.
-We formulate the theorem from~\cite{takens1981detecting} after establishing preliminary definitions.
-
-\ednote{It was suggested that I define these two concepts before stating the theorem.}
-\begin{Definition}
-  [\bf{Manifold, Compact Manifold}]\label{Dfn_compactManif}\rm
-  A topological space is called a manifold if every point in the space has a neighbourhood homeomorphic to $\mathbb{R}^n$. Succinctly, a manifold is a locally Euclidean topological space.
-  If a manifold is compact as a space topologically, then it is termed a compact manifold
-\end{Definition}
-
 \begin{Definition}
@@
 
+Takens Theorem states a result establishing a relationship between the observed and underlying dynamical systems by showing that the concatenation of a sufficiently large number of previous observations into a vector will, under certain conditions, generate a map between the vectors from the respective systems.
+We formulate the theorem from~\cite{takens1981detecting}.
+
 \begin{Theorem} 
-	[\bf Takens Embedding Theorem (adopted from~\cite{takens1981detecting})]\label{Thm_Takens}
+	[\bf Takens Embedding Theorem (taken from~\cite{takens1981detecting})]\label{Thm_Takens}
          Let $W$ be a compact manifold of dimension $m$, and $d\ge m$ so that $2d$ is an integer. It is a 
@@
 By generic, we mean a residual set on a certain topology on appropriate spaces of functions (that we do not describe here) \ednote{B: Where can I find a citation for this?}. 
-In our scenario (and this remains important throughout) the input space $U$ is considered the attractor.~\label{attractor_U}
+In our scenario -- and this remains important throughout -- the input space $U$ is considered the attractor.~\label{attractor_U}
 
@@
 \begin{figure}[ht]
-  \includegraphics[scale=0.25]{Graphs/_takensmap.eps}
+  \includegraphics[width=0.7\linewidth]{Graphs/_takensmap.eps}
   \centering
@@
 \section{Practical Limitations to Takens' Theorem}\label{sect_takenslimits}
-Takens' Embedding Theorem is a powerful result, and it provides compelling reason to believe that one could conceivably reconstruct a system conjugate accurately to the underlying system. 
-Nonetheless, it does present some severe practical limitations:
+The theorem above is a powerful result, and it provides compelling reason to believe that one could conceivably accurately reconstruct a system conjugate to the underlying system. 
+Nonetheless, it does present some severe practical limitations.
 \vspace{-5mm}
@@
 \item Even supposing that we could indeed find $F_\theta$, our \emph{approximation} of  $F_\theta$ is a map from a larger set $\mathbb{R}^{2d+1}$ containing the embedded attractor.
-There are, however, no theoretical guarantees that $F_\theta$ will retain  $V=\Phi_{2d,\theta}(W)$  as an attractor although $W$ could itself be an attractor.
+There are, however, no theoretical guarantees that $F_\theta$ will retain  $V:=\Phi_{2d,\theta}(W)$  as an attractor although $W$ could itself be an attractor.
 \item Takens theorem is stated only for noiseless observations.
-Due to noise $\epsilon_n$ the delay vector $\Phi_{2d,\theta}(w_n) + \epsilon_n$  may lie outside $V$.Furthermore, due to the chaotic nature of the underlying system (i.e.
+Due to noise $\epsilon_n$, the delay vector $\Phi_{2d,\theta}(w_n) + \epsilon_n$  may lie outside $V$.
+Furthermore, due to the chaotic nature of the underlying system (i.e.
 the fact that it has SIDC), the evolution of $\Phi_{2d,\theta}(w_n) + \epsilon_n$ under the map $F_\theta$ could move out of $V$ completely.
-This problem can be overcome  by using a driven dynamical system with some properties, and we discuss this in Chapter\ref{ch3}.
+This problem can be overcome  by using a driven dynamical system with specific properties, and we discuss this in Chapter~\ref{ch3}.
 \end{enumerate}
@@
 
-Though not a uniquely related to the Takens embedding theorem, we do opt to make mention of another requirement for our work. 
-The methods described here are applicable for data originating from a surjective map. 
-This follows from the fact that when $T$ maps the space $U$ into a proper subset $B$ of$ $U, there exists some point in $U - A$ that does not have a preimage and we conclude that $T$ is not surjective.
+Though not uniquely related to the Takens embedding theorem, we opt to make mention of another requirement to our work here. 
+The methods described in this project are applicable for data originating from a surjective map. 
+This follows from the fact that when $T$ maps the space $U$ into a proper subset $B$ of $U$, there exists some point in $U - A$ that does not have a preimage such that $T$ cannot be not surjective. 
 An example of this would be the case when a system displays energy-loss.
-Consider, for instance, consider the dying oscillations for a damped pendulum.
-(An account of greater depth  is provided in~\ref{ch5})
-
-This requirement is not overly restrictive, as a great number of chaotic systems do have a surjective map.
-\textbf{(Cite)} \ednote{B: I remember you saying this in one of our sessions, but don't have a source.
-Could you point me to a resource?}
+Consider, for instance, consider the dying oscillations for a damped pendulum (see Figure~\ref{fig:damped_pendulum}).
+The double pendulum is discussed in greater depth  is provided in Chapter~\ref{ch5}.
+The requirement is not overly restrictive as map that is topologically transitive on a subspace must also be surjective on it; if it were to be not surjective there would be an open set $V$ that does not intersect with any future iterates of an open set  $U$ disjoint from it.

### 2022-01-19 projectReport/chapter4.tex
Rereading the project. On 5.4

@@
 We also consider the conditions a driven system should possess to avoid adding distortion to its state-space representation. 
-We then describe the \emph{single-delay dynamics} (SDD) of the system and finally state conditions on the driven system so that the SDD are conjugate (or at least semi-conjugate) to the underlying system.
+We then describe the \emph{single-delay dynamics} (SDD) of the system, and finally state conditions on the driven system for the SDD to be conjugate (or at least semi-conjugate) to the underlying system.
 The SDD can then be used to forecast and reconstruct the underlying system.
@@
 
-We immediately recall the figure first defined in\eqref{eqn_conjugacy}  and remind ourselves of our ultimate goal to learn a system that is conjugate to an underlying or unknown dynamical system $(U,T)$ by making use of the measurements obtained from  the unknown system.
-See this sketched in the diagram~\ref{UV_conjugacy}
-
-%#####################################
-\begin{equation}
-  % \begin{center}
-    \centering
-    \psset{arrows=->, arrowinset=0.25, linewidth=0.6pt, nodesep=3pt, labelsep=2pt, rowsep=0.7cm, colsep = 1.1cm, shortput =tablr}
-    \everypsbox{\scriptstyle}
-    \begin{psmatrix}\label{UV_conjugacy}
-    U & U\\%
-    V & V.
-    %%%
-    %  \ncline{1,1}{1,2}^{T} 
-    %  \ncline{1,1}{2,1} <{\phi}
-    %   \ncline{2,1}{2,2}^{S}
-    %  \ncline{1,2}{2,2} > {\phi}
-    \end{psmatrix}
-    %#####################################
-    % \end{center}
-\end{equation}
+We immediately recall the figure first defined in\eqref{eqn_conjugacy}  and remind ourselves of our ultimate goal to learn a system that is conjugate to an underlying/unknown dynamical system $(U,T)$ by making use of measurements obtained from  this unknown system.
+We again sketch the commutativity diagram~\ref{eqn_conjugacy}.
+\begin{equation*}\label{UV_conjugacy}
+  \begin{tikzcd}
+    U && U \\
+    V && V \\
+    \arrow["T", from=1-1, to=1-3]
+    \arrow["\phi", from=1-3, to=2-3]
+    \arrow["\phi"', from=1-1, to=2-1]
+    \arrow["S"', from=2-1, to=2-3]
+  \end{tikzcd}
+  \end{equation*}
 
 To find a function $\phi$, the need now arises to consider a so-called driven dynamical system, a special case of a nonautonomous dynamical system. 
-Two spaces are considered: input is taken from an input space $U$ and the state of a space $X$ is updated.
-(Previously, only the state was considered.)
+Two spaces are considered: input is taken from an input space $U$, and the state of a space $X$ is updated. (Previously, only the state was considered.)
 We do this to mimic the true conditions.
-By accounting for an extrogenous input sequence in $U$ which can also influence the  system-state at a specific time-step, it produces more general models than an autonomous system.
+By accounting for an extrogenous input sequence in $U$ which can also influence the  system-state at a specific time-step, models more general than an autonomous system are produced.
 
 \begin{Definition}
-  [\bf Driven, Compactly Driven Dynamical System]\label{Dfn_DDS} \rm
-A driven dynamical system comprises two topological metric spaces $U$, $X$ and a continuous function  $g:U\times{X}\to{X}$ where $g(u_n, x_n)=x_{n+1}$.
+  [\bf Driven and Compactly Driven Dynamical System]\label{Dfn_DDS} \rm
+A driven dynamical system comprises two topological metric spaces $U$ and $X$, as well as  a continuous function  ${g:U\times{X}\to{X}}$ where ${g(u_n, x_n)=x_{n+1}}$.
 If the input space $U$ is compact, we refer to the system as compactly driven. 
@@
 
-The update equation $x_{n+1} = g(u_n,x_n)$ for $n \in\mathbb{Z}$, input $u_n$ from $U$ and state $x_n$ belonging to $X$ (a compact space) generate the dynamics on X. 
+The update equation $x_{n+1} = g(u_n,x_n)$ for $n \in\mathbb{Z}$, input $u_n$ from $U$ and state $x_n$ belonging to $X$,a compact space, generates the dynamics on $X$. 
 Abbreviated, we refer only to the \textit{driven system $g$}, where all other entities are understood implicitly.
 
-Notably, a nonautonomous dynamical system can be generated from $U$; any input $\overline{u}$, a bi-infinite sequence from $U$, gives rise to the sequence of self-maps ${\{g(u_n, \cdot)\}}_{n\in\mathbb{Z}}$ contained in $X$.
-Physically, one may think of this bi-infinite sequence as referring to a system that has been running for an extended period at the time of the first measurement taken from the system (alternatively, the first time a probe is inserted into the system to take an observation)
+Notably, a nonautonomous dynamical system may be generated from $U$.
+Any input $\overline{u}$, a bi-infinite sequence from $U$, gives rise to the sequence of self-maps ${\{g(u_n, \cdot)\}}_{n\in\mathbb{Z}}$ contained in $X$.
+Physically, one may think of this bi-infinite sequence as referring to a system that has been running for an extended period at the time of the first measurement taken from the system.
+Alternatively, one may conceive of the scenario where this is the first time a probe is inserted into a system to take an observation.
 
@@
 
-\begin{Example}\label{ex_halfux} \rm
+\begin{Example}\label{ex_halfux}\rm
   The only solution  ${\{x_n\}}_{n\in\mathbb{Z}}$ to the driven system  $g(u,x)=\frac{ux}{2}$, where $X=[0,1]$, $U=[0,1]$,  is the zero solution $x_n\equiv0$.
@@
 
-A system may also have multiple solutions, as evidenced in the example below.
+A system may also have multiple solutions, as evidenced in Example~\ref{Ex_exp}.
 
-\begin{Example}\label{Ex_exp} \rm 
+\begin{Example}\label{Ex_exp}\rm 
   Consider the driven system $g(u,x)=x^2$ for $X=[0,1]$, $U=\mathbb{R}$.
-The system has an uncountable number of solutions, as there exists a solution for every $x\in{X}$ which also passes through the point $x$ and $\lim_{n\to\infty}x_n=0$, $\lim_{n\to{-}\infty}x_n=1$.
+The system has an uncountable number of solutions as there exists a solution for every $x\in{X}$ which also passes through the point $x$ and $\lim_{n\to\infty}x_n=0$, $\lim_{n\to{-}\infty}x_n=1$.
  The proof is deferred to immediately after the next paragraph (see~\ref{rem_proofEx}).
@@
 
-As the solutions to a driven system are considered an important entity, we next identify a subspace $X_U$ of $X$ that contains all possible solutions.
-To realize such a subspace of a driven system $g$, the concept of a reachable set is defined.
+Solutions to a driven system are considered an important entity, so we next identify a subspace $X_U$ of $X$ that contains all possible entire solutions. 
+% To realize such a subspace of a driven system $g$, the concept of a reachable set is defined.
 
@@
 
-The reachable set itself can defined independently from $g$ possessing the ESP/USP.
+\ednote{Moved to after defn of USP/ESP: The reachable set itself can defined independently from $g$ possessing the ESP/USP. } 
 Note that $x\in{X_n(\overline{u})}$ if and only $g$ has a solution $\{x_k\}$ for $x_n=x$ and input $\overline{u}$, a result proved in nonautonomous dynamical systems literature~\cite{manjunath2014dynamics, manjunath2013echo}.
 
-The set $X_n(\overline{u})$ is precisely the set of points $x$ through which some entire solution $\Psi$ passes at time $n$. We can give an alternate formulation of the set $X_n(\overline{u})$. 
-Now suppose, we for some fixed input 
-$\overline{u}$, we define $g_i = g(u_i,\cdot)$ for all $i\in \mathbb{Z}$, then the set of states
+The set $X_n(\overline{u})$ is precisely the set of points $x$ through which some entire solution $\Psi$ passes at time $n$. We can give an alternate formulation of the set $X_n(\overline{u})$: 
+Suppose for some fixed input 
+$\overline{u}$, we define $g_i = g(u_i,\cdot)$ where $i\in \mathbb{Z}$, then the set of states
 \begin{equation} \label{eqn_association}
@@
 X_{n,i}(\overline{u})$ is a nested intersection of closed nonempty subsets, and
-whenever $X$ is compact, the intersection is nonempty. We say $g$ is a topological contraction if $\bigcap_{i<n}X_{n,i}(\overline{u})$ is a singleton subset of $X$ for each $\overline{u}$.
+whenever $X$ is compact, the intersection is nonempty.
 
-It turns out that $\bigcap_{i<n}X_{n,i}(\overline{u})$ is identically equal to $X_n((\overline{u})$ that we had defined earlier.
-This will lead to a result (see~\cite{manjunath2013echo}) that \textit{$g$ being a topological contraction is equivalent to the existence of a exactly one entire solution.}
+\begin{Definition}
+  [\bf Topological Contraction]\label{Dfn_topContr}\rm
+  The driven system $g$ is a topological contraction if $\bigcap_{i<n}X_{n,i}(\overline{u})$ is a singleton subset of $X$ for each $\overline{u}$. 
+\end{Definition}
 
-% \begin{Definition}
-%   [\bf Topological Contraction]\label{Dfn_TopContr}\rm
-%   A function $g:U\times{X}\to{X}$ is a topological contraction if for all $n\in\mathbb{Z}$ and all $\overline{u}\subset{U}$, $X_n(\overline{u})$ is a singleton subset of $X$.  
-% \end{Definition}
+It turns out that $\bigcap_{i<n}X_{n,i}(\overline{u})$ is identically equal to $X_n((\overline{u})$ that we had defined earlier. This will lead to a result (see~\cite{manjunath2013echo}) that 
+\textit{$g$ being a topological contraction is equivalent to the existence of exactly one entire solution.}
 
@@
   For the driven system,  
-  $g(u,x)=x^2$ the evolution does not depend on the input at all.
-Hence $g(u,x)$ can be written as a single map $f(x)=x^2$ which is a homeomorphism on $[0,1]$.
+  $g(u,x)=x^2$ the evolution does not depend on the input at all. Hence $g(u,x)$ can be written as a single map $f(x)=x^2$ which is a homeomorphism on $[0,1]$. 
   A left-infinite orbit is defined and it converges to $1$ while the right-infinite orbit converges to $0$. 
-Or in other words the iterates of $f^{-1}$ converges to $1$ while the iterates of $f$ converge to $0$
+  In other words the iterates of $f^{-1}$ converges to $1$ while the iterates of $f$ converge to $0$ 
 \end{Remark}
 
-Thus far, it has been demonstrated that a system may have one or more solutions; one may ask if a driven system always has a solution and, if so, whether it satisfies specific properties such as uniqueness.
+Thus far, it has been demonstrated that a system can have a single or multiple solutions.
+One may therefore reasonably ask if a driven system always has a solution and, if so, whether it satisfies specific properties such as uniqueness.
 Should the driven system be compact, existence follows immediately, as shown in the following result.
@@
 To refine the scenario, we consider only systems with unique solutions.
+
 \section{Unique Solution Property}
@@
   A driven system $g$ is said to have the Unique Solution Property (USP) if for each input $\overline{u}$ there exists exactly one solution. 
-  Alternatively we may formulate the USP as follows: $g$ has the Unique Solution Property if there exists a well-defined map $\Psi:{U}\to{X}$, with $\Psi({\overline{u}})$ denoting the unique solution.
+  Alternatively we may formulate the USP as follows: $g$ has the Unique Solution Property if there exists a well-defined map $\Psi:{U}\to{X}$, where $\Psi({\overline{u}})$ denotes the unique solution.
 \end{Definition}
 
-One of the first results obtained after defining the USP is that every solution will attract different initial conditions towards the component parts of the solution.
+The USP is equivalently called  the echo state property (ESP)~\cite{Manju_ESP}, and systems that possess the ESP have other important and useful properties for information processing~\cite{grigoryeva2018echo, grigoryeva2019differentiable, manjunath2020stability} .
+
+One of the first results obtained after defining the USP refers to the fact that every solution will attract different initial conditions towards the component parts of the solution.
 If $g$ has the USP, then any solution to $g$ is also a uniform attractor in nonautonomous dynamical systems literature~\cite{Manju_Nonlinearity}.
@@
 
+Secondly, it is worth noting that the reachable set~\ref{Dfn_ReachableSet} can be defined independently from $g$ possessing the USP. \ednote{hmm. does it fit here}
 % Paraphrased.
@@
 
-In the majority of Reservoir Computing (RC) literature- a machine learning approach- where the input is mapped onto a different but higher dimensional through a system called the reservoir and a readout that measures the state of the reservoir is trained, a notion of forgetting the states of the reservoir asymptotically is often used.
+In the majority of Reservoir Computing (RC) literature -- a machine learning approach -- where the input is also mapped onto a different, but higher dimensional, space through a system called the reservoir.
+A readout that measures the state of the reservoir is trained, a notion of forgetting the states of the reservoir asymptotically is often used.
+\ednote{this sentence falls out of the blue}
 
@@
 B: Thanks!},
-Concepts like the ESP,  fading memory~\cite{boyd1985fading} are some of these notions.  
-If $g$ possesses the ESP (equivalent in our context to the USP), the we are guaranteed that the whole of an input's left-infinite history will exactly determine the current system state; i.e.
+Concepts like the echo state property (ESP) and fading memory~\cite{boyd1985fading} are some of these notions.  
+If $g$ possesses the ESP (equivalent in our context to the USP), then we are guaranteed that the whole of an input's left-infinite history will determine exactly the current system state; i.e.
 there is only ever the possibility of a single reachable state at a given instance of time~\cite{jaeger2001echo,Manju_2020}.
@@
 % We do however consider it important to mention that the ESP is also often discussed in terms of a stability property that holds for all inputs of a system~\cite{manjunath2020stability} and, furthermore, plays a key role in the design and training of recurrent neural networks(RNN) within the field of RC \cite{Manju_ESP}.
-Beyond forgetting the past states the ESP is also often discussed in terms of a stability property that is nearby inputs create close by responses (solutions), a concept called input-related stability~\cite{manjunath2020stability} and, furthermore, plays a key role in the robustness of recurrent neural networks (RNN), a component of the implementation discussed in~\ref{ch5}.
+Beyond forgetting the past states, the ESP is also often discussed in terms of a stability property termed input-related stability where nearby inputs create nearby responses (solutions)~\cite{manjunath2020stability}.
+Furthermore, it plays a key role in the robustness of recurrent neural networks (RNN), a component of the implementation discussed in~\ref{ch5}.
 
@@
 Having already defined the reachable set $X_U$ as the collection of all elements of all entire solutions, we pause briefly in order to fix additional notation.
-Defining $\cev{u}^{n}:=(\ldots,u_{n-2} ,u_{n-1})$ as the left-infinite subsequence of an input up until time $n$, $\overleftarrow{U}$ is then the notation for all these left-infinite sequences in $U$.
-Moreover, $\cev{u}^{n}v:=(\ldots,u_{n-2} ,u_{n-1}, v)$ will symbolise the input up to time $n$ with $v \in U$ being the specific input value at time $n$. 
+Defining $\cev{u}^{n}:=(\ldots,u_{n-2} ,u_{n-1})$ as the left-infinite subsequence of an input up until time $n$, $\overleftarrow{U}$ then denotes all such left-infinite sequences in $U$. 
+Moreover, $\cev{u}^{n}v:=(\ldots,u_{n-2} ,u_{n-1}, v)$ will symbolise the input up to time $n$, with $v \in U$ being the specific input value at time $n$. 
 The introduction of a new input at time $n$ can be described by the mapping $\sigma_v:   \cev{u}^{n} \mapsto \cev{u}^{n}v$. 
@@
 
-The question now assumes the form: Can we establish a semi-conjugacy for the driven system $g$ as presented below.
+The question now assumes the form: Can we establish a semi-conjugacy for the driven system $g$ as presented below?
 
-\begin{equation}  \label{Scomm_h}
-  %    \[ 
-      \psset{arrows=->, arrowinset=0.25, linewidth=0.6pt, nodesep=3pt, labelsep=2pt, rowsep=0.7cm, colsep = 1.1cm, shortput =tablr}
-   \everypsbox{\scriptstyle}
-   \begin{psmatrix}
-   \overleftarrow{U} & \overleftarrow{U}\\%
-   X_U & X_U.
-   %%%
-  %  \ncline{1,1}{1,2}^{\sigma_v} \ncline{1,1}{2,1} <{h}
-  %  \ncline{1,2}{2,2} > {h}
-  %  \ncline{2,1}{2,2}^{g(v,\cdot)}
-   \end{psmatrix}
-  % \]
+  \begin{equation}\label{Scomm_h}
+    \begin{tikzcd}
+	{\overleftarrow{U}} && {\overleftarrow{U}} \\
+	{X_U} && {X_U} \\
+	\arrow["{\sigma_v}", from=1-1, to=1-3]
+	\arrow["h"', from=1-3, to=2-3]
+	\arrow["h", from=1-1, to=2-1]
+	\arrow["{g(v,\cdot)}"', from=2-1, to=2-3]
+\end{tikzcd}
 \end{equation}
@@
 But does such a function $h$ for the above driven system $g$ above exist?
-Whenever $g$ has the USP and $\Psi(u)={\{x_n\}}_{n\in\mathbb{Z}}$ it follows that $h$, defined by  $h(\cev{u}_n):=g(u_n,x_{n-1})=x_n$, will satisfy the semi-conjugacy in the graph above~\ref{Scomm_h}.
-Regrettably, such a continuous mapping $h$ is not guaranteed to exist when $g$ does not have the USP~\cite[Lemma 5]{Manju_Nonlinearity}.
-Note that even when $h$ does exist, we are not guaranteed its injectivity. Considering again example~\ref{ex_halfux}, we see that even if $h$ were to exist, it could not be injective as $X_U=\{0\}$.
+Whenever $g$ has the USP and $\Psi(u)={\{x_n\}}_{n\in\mathbb{Z}}$, it follows that $h$ defined by  $h(\cev{u}_n):=g(u_n,x_{n-1})=x_n$ will satisfy the semi-conjugacy in the graph above~\ref{Scomm_h}.
+Regrettably, such a continuous mapping $h$ is not guaranteed to exist when $g$ does not have the USP.
 
-Re-sketching the commutativity diagram~\ref{Scomm_h} above by replacing $X_U$ by its left-infinite sequence space $\overleftarrow{X}_U$, we obtain the diagram below.
-In this case, the function $H:\overleftarrow{U}\to\overleftarrow{X}_U$, a map that is both continuous and surjective, is called a \emph{causal mapping} that is defined next.
+\begin{Theorem}
+  A compactly driven system $g$ induces a universal semi-conjugacy $h:\overleftarrow{U}\to{X_U}$ if and only if $g$ has the USP. 
+\end{Theorem}
+\vspace{-6mm}
+{\bf Proof.} See the proof of Lemma 5 provided in~\cite{Manju_Nonlinearity}
+
+Note that even when $h$ does exist, we are not guaranteed its injectivity. In example~\ref{ex_halfux} we observe that even if $h$ were to exist, it could not be injective as $X_U=\{0\}$.
+
+Re-sketching the commutativity diagram~\ref{Scomm_h} above by replacing $X_U$ with its left-infinite sequence space $\overleftarrow{X}_U$, we obtain the diagram~\ref{SCausal_H}. 
+In this case, the function $H:\overleftarrow{U}\to\overleftarrow{X}_U$, a map that is both continuous and surjective, is called a \emph{causal mapping}.
 
@@
   [\bf Causal Mapping]\label{Def_CausMap}
-  A continuous, surjective map $H:\overleftarrow{U}\to\overleftarrow{X}_U$ such that \[H\circ\tilde{g}_v=\sigma\circ{H}\] holds for all $v \in U$ where $\tilde{g}_v$ maps $(\ldots, u_{-2}, u_{-1})$ to $(\ldots, u_{-2}, u_{-1}, g(v, u_{-1}))$ is called a causal mapping.
+  A continuous, surjective map $H:\overleftarrow{U}\to\overleftarrow{X}_U$ such that \[H\circ\tilde{g}_v=\sigma_v \circ{H}\] holds for all $v \in U$ where $\tilde{g}_v$ maps $(\ldots, u_{-2}, u_{-1})$ to $(\ldots, u_{-2}, u_{-1}, g(v, u_{-1}))$ is called a causal mapping.
 \end{Definition}
 
-\begin{equation} \label{SCausal_H}
-    %    \[ 
-        \psset{arrows=->, arrowinset=0.25, linewidth=0.6pt, nodesep=3pt, labelsep=2pt, rowsep=0.7cm, colsep = 1.1cm, shortput =tablr}
-     \everypsbox{\scriptstyle}
-     \begin{psmatrix}
-     \overleftarrow{U} & \overleftarrow{U}\\%
-     \overleftarrow{X}_U & \overleftarrow{X}_U.
-     %%%
-    %  \ncline{1,1}{1,2}^{\sigma_v} \ncline{1,1}{2,1} <{h}
-    %  \ncline{1,2}{2,2} > {h}
-    %  \ncline{2,1}{2,2}^{\tilde{g}_v}
-     \end{psmatrix}
-    % \]
+  \begin{equation}\label{SCausal_H}
+    \begin{tikzcd}
+    \overleftarrow{U} && \overleftarrow{U}\\
+     \overleftarrow{X}_U && \overleftarrow{X}_U.
+     \\
+    \arrow["{\sigma_v}", from=1-1, to=1-3]
+	\arrow["h"', from=1-3, to=2-3]
+	\arrow["h", from=1-1, to=2-1]
+	\arrow["{\tilde{g_v}}"', from=2-1, to=2-3]
+\end{tikzcd}
 \end{equation}
@@
 \vspace{-6mm}
-{\bf Proof.}  See~\cite[Th.3]{manjunath2013echo}.
+{\bf Proof.}  See~\cite[Theorem.3]{manjunath2013echo}.
 
 The driven system $g$ can induce an embedding of $\overleftarrow{U}$ in $\overleftarrow{X}$ as follows: 
-If the causal mapping $H:\overleftarrow{U}{\to}{\overleftarrow{X}_U}$ is injective (in addition to being surjective), it becomes the embedding of $\overleftarrow{U}$ in $\overleftarrow{X}$ induced by the driven system $g$.
+If the causal mapping $H:\overleftarrow{U}{\to}{\overleftarrow{X}_U}$ is injective (in addition to its surjectivity), it becomes the embedding of $\overleftarrow{U}$ in $\overleftarrow{X}$ induced by the driven system $g$.
 The continuity of $H^{-1}$ follows from the fact that $H$ is itself a continuous and surjective mapping of a compact space $\overleftarrow{U}$ in a Hausdorff space.
-and we refer to the map $H$ as a \emph{causal embedding}. 
+and we refer to the map $H$ as a \emph{causal embedding}. (Compare this with Definition~\ref{defn_causallyEmbed}.) 
 %\ednote{B: edited the sentence here}
@@
 When $g$ has the USP, the diagram below (taken from~\cite{Manju_Nonlinearity}) illustrates the operation of the mappings $h$ and $H$. 
-The mapping $h:\overleftarrow{U}\to{X_U}$ is also considered an observable as mentioned in the introduction of Chapter~\ref{ch3}).
+The mapping $h:\overleftarrow{U}\to{X_U}$ is an observable as mentioned in the introduction of Chapter~\ref{ch3}).
 
 \begin{figure}[ht]
-  \includegraphics[scale=0.4]{Graphs/_actionofh_H.eps}
+  \includegraphics[width=0.6\linewidth]{Graphs/_actionofh_H.eps}
   \centering
@@
 
-The function $H$ contains a countably infinite number of coordinate (or factor maps) since its codomain is $\overleftarrow{X}_U$.  
-As it is not practically feasible to consider a left-infinite sequence in $X_U$ (in other words an element of  $\overleftarrow{X}_U$) for learning, we wish to restrict the inputs we consider to a subspace of $\overleftarrow{U}$.
-Particularly, we restrict our attention to the  so-called inverse-limit space of a dynamical system and hope to embed it within some finite product of $X$.
+The function $H$ contains a countably infinite number of coordinate (or \textit{factor maps}) since its codomain is $\overleftarrow{X}_U$.  
+It is not practically feasible to consider a left-infinite sequence in $X_U$ (in other words an element of  $\overleftarrow{X}_U$) for learning.
+We wish to restrict the inputs to a subspace of $\overleftarrow{U}$.
+Particularly, we restrict our attention to the  so-called inverse-limit space (see Definition~\ref{Dfn_InverseL}) of a dynamical system and hope to embed it within a finite product of $X$.
 
@@
 Broadly, we may conceive of the inverse-limit system of a dynamical system as a subspace of an infinite-dimensional space where each point in the inverse-limit space corresponds to a backward orbit of the map $T$.
-Following~\cite{manjunath2021universal, ingram2011inverse} we provide a definition:
-
+Following~\cite{manjunath2021universal, ingram2011inverse} we provide a definition.
 \begin{Definition}
@@
 
-The inverse-limit space is well-defined since $T : U \to{U}$ is surjective by assumption.
+The inverse-limit space is well-defined as $T : U \to{U}$ is surjective by assumption.
 We now get to the concept of embedding the inverse limit-space of the dynamical system in $X \times X$. 
-\begin{Definition} \rm
+\begin{Definition}\label{defn_causallyEmbed} \rm
 	We say a driven system $g$ \emph{causally embeds} a dynamical system $(U,T)$ if it satisfies the two properties: 
@@
 
-It is imperative to take note of a subtlety here: we refer to the \emph{map} $H$ as a (causal) embedding, whereas the driven \emph{system} $g$ (causally) embeds another \emph{system} $(U,T)$.
-\ednote{Check this sentence again.}
+It is imperative to take note of a subtlety here: we refer to the map $H$ as a causal \textit{embedding}, whereas the driven \textit{system} $g$ causally embeds another \emph{system} $(U,T)$.
 
@@
 So far it is not clear whether the USP alone ensures the ability to causally embedding a dynamical system and we mention that a driving function $g$ cannot just be chosen in a frivolous manner.
-This could possibly complicate our work.
+This could possibly complicate our task.
 %\ednote{B: Grammarly}
@@
 %To achieve a causal embedding embedding not just a causal mapping, it is undesirable to choose a function $g$ that quenches the temporal structure in $u$ by contracting to such a degree that the ability to recover information from the original system is lost completely.
-In the above example (\ref{ex_halfux}), the input's temporal variation could not be related to the reachable set as $X_U$ only consisted of a single element; little, if not no, information is encoded.
-To obtain a suitably complex function $g$, it is thus desired that the reachable set of a driven system  be large enough to relate to the input. 
-This is even true for embedding  the inverse-limit space of a dynamical system. \ednote{B: Expand or refer?}
+In the above example~\ref{ex_halfux}, the input's temporal variation could not be related to the reachable set as $X_U$ only consisted of a single element and little to no information was encoded.
+To obtain a suitably complex function $g$, we wish to have a reachable set for a driven system  that is large enough to relate to the input. 
+This is even true for embedding  the inverse-limit space of a dynamical system~\cite{manjunath2021universal}.
 
- To this end, we recall the notion of State-Input (SI) Invertibility.
+ To this end, we define the notion of State-Input (SI) Invertibility.
 
@@
   [\bf SI-Invertibility]\label{Dfn_SIinv}\rm
-  A driven system $g$ is said to be SI-Invertible if $g(*,x): U \to X$ is invertible for all $x\in X$.
-Alternatively it may be said that if, given $x_n$ and $x_{n-1}$, $u_{n-1}$ can be uniquely determined from $x_n=g(u_n,x_{n-1})$, then $g$ is said to be SI-invertible.
+  A driven system $g$ is said to be SI-Invertible if ${g(*,x): U \to X}$ is invertible for all $x\in X$.
+Alternatively, if for  $x_n$ and $x_{n-1}$ given, one can uniquely determine $u_{n-1}$ from $x_n=g(u_n,x_{n-1})$, $g$ is said to be SI-invertible.
 \end{Definition}
 
-As mentioned, $g$ can also introduce additional complexity to a solution.
+As mentioned, $g$ can introduce additional complexity to a solution.
 Consider the Figure~\ref{fig:memloss_conttime} where a RNN's solution was numerically simulated for randomly generated matrices $A$ and $B$ where $B$ has unit spectral radius.
-An input sequence (black) $(x_0,x_1,\ldots,x_{5000})$ was fed to the network and we plot it as well as the the coordinates of two simulated solutions in red (with parameter $a=1$, $\alpha=0.99$ where $g$ has the USP) and blue ($a=1$, $\alpha=1.05$ where $g$ does not have the USP).
-The red solution mimics the input sequence well in its variation, whereas the blue solution is disordered and changes a lot.
-In this we may say that the blue solution has introduced additional complexity to the system.
-\ednote{B: Added this paragraph}
+An input sequence in black $(x_0,x_1,\ldots,x_{5000})$ was fed to the network and it, as well as the the coordinates of two simulated solutions in red (with parameter $a=1$, $\alpha=0.99$ where $g$ has the USP) and blue ($a=1$, $\alpha=1.05$ where $g$ does not have the USP), were plotted.
+The red solution mimicked the input sequence's temporal variation quite well, whereas the blue solution is observed to be disordered and oft-changing.
+Through this we may state that the blue solution has introduced additional complexity to the system.
 
-SI-invertibility promises that `enough' information is retained when $g$ is chosen without introducing one's choice of $g$ will still ensure `enough' information is retained without introducing unwanted complexity.
-`Enough' here refers to the fact that we may always recover the previous input value $u_n$ (if we know successive states $x_{n-1}, x_n$).
+SI-invertibility promises that `enough' information is retained when $g$ is chosen without introducing without introducing unwanted complexity. 
+`Enough' here refers to the fact that one can always recover the previous input value $u_n$ if the successive states $x_{n-1}, x_n$ are known.
 
-Subsequently we define the relation $Y_T$ induced by $(U,T)$ on $X_U\times{X_U}$ for a driven system $g$ possessing SI-invertibility.  
-To describe the  SDD formally, we consider a dynamical system $T: U \to U$ and define a relation on the reachable set $X_U$, i.e. a subset on $X_U \times X_U$  defined by 
+Subsequently, we define the relation $Y_T$ induced by $(U,T)$ on $X_U\times{X_U}$ for a driven system $g$ possessing SI-invertibility.  
+To describe the  SDD formally, consider a dynamical system $T: U \to U$ and define a relation on the reachable set $X_U$. The subset on $X_U \times X_U$  is defined by 
 \begin{equation}
@@
   For an SI-invertible driven system $g$ and a dynamical system $(U,T)$, the map $G_T: Y_T \to Y_T$ defined by the relation $(x_{n-1},x_n) \mapsto (x_n,x_{n+1})$ is well-defined. 
-  (This results holds even in the absence of $g$ possessing over the USP) 
+  (This results holds even in the absence of $g$ possessing the USP). 
   % Moreover, the mapping $(x_{n-1},x_n) \mapsto u_{n}$ is well-defined when $x_{n-1}$ and $x_n$ are successive points on a solution obtained for an input  $\{u_n\}$ that is an orbit of $T$.
@@
   \vspace{-6mm}
-  {\bf Proof.} See~\cite[Th.3]{Supp}
+  {\bf Proof.} See~\cite[Theorem.3]{Supp}
 
@@
 
-At this stage it is worth taking note of a specific driven system in the form of a discrete state-space model which has acquired some adherence in applications~\cite{Manju_IEEE}- especially those pertaining to Echo State Networks and the ESP.\@The function
+At this stage it is worth taking note of a specific driven system in the form of a discrete-time state-space model which has acquired some adherence in applications~\cite{Manju_IEEE} -- especially those pertaining to echo state betworks and the ESP.
+The function
 \begin{equation}  \label{eqn_driving}
@@
 \end{equation} 
-is SI-invertible and, if $\alpha B$ has a spectral norm $<1$, also possesses the USP~\cite[Th.2]{manjunath2013echo }. 
+is SI-invertible and, if $\alpha B$ has a spectral norm $<1$, also possesses the USP~\cite[Theorem.2]{manjunath2013echo }. 
 It is easy to show that $g$ is SI-invertible by recovering $u_n$ in 
@@
   when $x_{n-1}$ and $x_n$ are known.
+This specific driving function $g$ is used in our implementation and is discussed more completely in Chapter~\ref{ch5}.
 
-This specific driving function $g$ is used in our implementation and is discussed more completely in Chapter~\ref{ch5}
-
-Despite the ease that with which one manipulates a left-infinite history in the realm of theory, it is impossible to obtain or use such a sequence in any real-life application.  
-Fortunately, one does not need the entire left-infinite history of an input in practice thanks to the Uniform Attraction Property (UAP). 
-We use an alternate version of the definition of UAP -- since the notion of nonautonomous attractors in this project would take some time to establish and detracts from the principal thrust of this project, we refer to~\cite{Manju_Nonlinearity}.
+Despite the ease with which one manipulates a left-infinite history in the realm of theory, it is impossible to either obtain or use such a sequence in any real-life application.
+Fortunately, one does not need an input's entire left-infinite history in practice in practice thanks to the Uniform Attraction Property (UAP). 
+We use an alternate version of the definition of UAP since the notion of nonautonomous attractors in this project would take some time to establish and detracts from the principal thrust of this project.
+For more specific results, we refer to~\cite{Manju_Nonlinearity}.
 
@@
 
-The UAP guarantees that all trajectories will converge to the same trajectory as time moves forward. 
+The UAP guarantees that all trajectories will converge to the same trajectory in forward time. 
 An illustration of this is shown in Fig.~\ref{fig:memloss_conttime} where the trajectory in red seems to lock on to the supposed actual solution in black when the system has the USP.
@@
 
-One may now appreciate an even more astounding result: $g$ having the USP is equivalent to the UAP as is proved in~\cite{Manju_Nonlinearity}.
+One may now appreciate an even more astounding result.
+
+\begin{Theorem}
+  For a driven system $g$, the USP is equivalent to the UAP. 
+\end{Theorem}
+\vspace{-6mm}
+{\bf Proof} See~\cite{Manju_Nonlinearity}.
 
@@
 \item $g$ has the UAP.
-\end{enumerate\end{Theorem}{\bf Proof.}  May be found in \cite[Th.1]{Manju_Nonlinearity}
+\end{enumerate\end{Theorem}{\bf Proof.}  May be found in \cite[Theorem.1]{Manju_Nonlinearity}
 
@@
 We are interested in determining whether $\widehat{T}$ and $G_T$ are related.
-For this, we introduce  the function $H_2{\overline{}} := (h(r\overline{u}, h(\overline{u})))$ mapping the entirety of a left-infinite sequence to some element in $X\times{X}$.
+For this, we introduce  the function ${H_2{\overline{}} := (h(r\overline{u}, h(\overline{u}))}$ which maps the entirety of a left-infinite sequence in $\Psi$ to some element in $X\times{X}$.
 Indeed, it may be shown that $g$ having SI-invertibility and the USP immediately guarantees the existence of at least a semi-conjugacy between the systems $(Y_T, G_T)$ and $(\widehat{U}_T, \widehat{T})$.
-This is formalised in the Causal Embedding Theorem, which we state next:
+This is formalised in the Causal Embedding Theorem, which we state next.
 
@@
 \begin{equation} \label{comm_H_CET}
-\psset{arrows=->, arrowinset=0.25, linewidth=0.6pt, nodesep=3pt, labelsep=2pt, rowsep=0.7cm, colsep = 1.1cm, shortput =tablr}
- \everypsbox{\scriptstyle}
- \begin{psmatrix}
- \widehat{U}_T & \widehat{U}_T\\%
- Y_T &  Y_T.
- %%%
-%  \ncline{1,1}{1,2}^{\widehat{T}} \ncline{1,1}{2,1} <{H_2}
-%  \ncline{1,2}{2,2} > {H_2}
-%  \ncline{2,1}{2,2}^{G_T}
- \end{psmatrix}
+	\begin{tikzcd}
+	{\widehat{U}_T} && {\widehat{U}_T} \\
+		{Y_T} && {Y_T} \\
+	\arrow["{\widehat{T}}", from=1-1, to=1-3]
+			\arrow["{H_2}"', from=1-3, to=2-3]
+			\arrow["H_2"', from=1-1, to=2-1]
+		\arrow["G_T"', from=2-1, to=2-3]
+	\end{tikzcd}
 	\end{equation}
@@
 \vspace{-6mm}
-{\bf Proof.}  May be found as the proof of~\cite[Th.4]{Supp}
+{\bf Proof.}  May be found as the proof of~\cite[Theorem.4]{Supp}
 
-Recall that $H_2$ maps an entire left-infinite solution sequence from $\Psi$ to an element in $X\times{X}$. 
-We are drawing nearer and nearer to our principal target and our results carry more and more weight: If we are able to learn $G_T$, we will also have learnt $\widehat{T}$; and since $\widehat{T}$ is an extension of $T$, we will have obtained $T$ as well.
+% Recall that $H_2$ maps an entire left-infinite solution sequence from $\Psi$ to a single element in $X\times{X}$. 
+We are drawing nearer and nearer to our principal target and our results carry more and more weight.
+If we are able to learn $G_T$, we will also have learnt $\widehat{T}$; and since $\widehat{T}$ is an extension of $T$, we will have obtained $T$ as well.
 %\ednote{B: Grammarly check}
@@
 %\end{Theorem}
-% {\bf Proof.} See \cite[Th.3, Th.4]{manunath2021universal}
+% {\bf Proof.} See \cite[Theorem.3, Theorem.4]{manunath2021universal}
 
-\section*{Summarising the discussing thus far:}
+\section*{Summarising the discussing thus far}
 
@@
 \item We are interested in a some dynamical system $(U,T)$ with unknown dynamics.
-\item To determine properties about this system $(U,T)$ and predict its future evolution, we determine the dynamics of the inverse-limit system $(\widehat{U}, \widehat{T})$.
+\item To determine information about this system $(U,T)$ and predict its future evolution, we determine the dynamics of the inverse-limit system $(\widehat{U}, \widehat{T})$.
 % Given certain assumptions, we can guarantee that $(\widehat{U}, \widehat{T})$ is at least semi-conjugate to $(U,T)$.
@@
 
-One can therefore in practice learn the SDD of the driven system states via $G_T$ with enough data thanks to the USP/UAP. This enables us to do at least two things: 
+One can therefore in practice learn the SDD of the driven system states via $G_T$ with enough data thanks to the USP/UAP. This enables us to do at least two things. 
 \vspace{-8mm}
@@
 
-Finally, we note that although $G_T$ exists with SI-invertibility, we need the USP as well (see~\ref{fig:pictorialSummary} below for a visual illustration). 
+Finally, we note that although $G_T$ exists with SI-invertibility, we need the USP as well (see Figure~\ref{fig:pictorialSummary} below for a visual illustration). 
 If not, the driven states would have to be running for all time since they would not have forgotten their past states.
@@
 \end{figure}
-\ednote{Cited SIG25May here, but unsure if bibtex reference is correct}
 
@@
 
-In the above sections, we established the map $G_T$ describing the SDD of a driven system.
-By establishing the existence of this map, we’ve essentially embedded the attractor $U$ (recall from~\ref{attractor_U}) into the higher dimensional space $X\times{X}$.
-In layman’s terms, this ensures that there is more “dimensional room” for the underlying system’s underlying dynamics to “move”.
-Since the dynamics aren’t as “squashed”, we might therefore hope that the dynamics of $G_T$ are in some sense simpler than that of $T$.
-(Taken note of the fact that $G_T$ is a homeomorphism even when $T$ is just continuous)
+In the above sections, we established the existence of a $G_T$ describing the SDD of a driven system.
+By establishing its existence, we’ve essentially embedded the attractor $U$ (recall from~\ref{attractor_U}) into the higher dimensional space $X\times{X}$.
+In layman’s terms, this ensures that there is more `dimensional room' for the underlying system’s underlying dynamics to `move'.
+Since the dynamics aren’t as `squashed', we might therefore hope that the dynamics of $G_T$ are in some sense simpler than that of $T$.
+In~\cite{manjunath2021universal} it is illustrated in an empirical fashion in that the map $G_T$ describes dynamics which are less functionally complex than that of $T$ or of the map $\Phi_{2d,\theta}$.
+This is done by implementing a Recurrent Neural Network (RNN), which is later discussed in Chapter~\ref{ch5}.
+Note that $G_T$ is a homeomorphism even when $T$ is merely continuous.
 
-In~\cite{manjunath2021universal} it is illustrated in an empirical fashion in that the map $G_T$ describes dynamics which are less functionally complex than that of $T$ or of the map $\Phi_{2d,\theta}$.
-This is done by implementing a Recurrent Neural Network (RNN), which is discussed in Chapter~\ref{ch5}.
-\ednote{B: Should I add in that we'll be using Pearson coefficient?}
+We opt to learn $G_T$ in an indirect manner by defining a new map $\Gamma:(x_{n-1},x_n)\mapsto{u_n}$. It follows immediately from $G_T$’s existence that $\Gamma$ also exists.
 
-We opt to learn $G_T$ in an indirect manner by defining a new map $\Gamma:(x_{n-1},x_n)\mapsto{u_n}$. It will follow immediately from $G_T$’s existence that $\Gamma$ also exists.
-
-The reasons for taking this roundabout approach remain to be discussed in~\ref{subs_LearnGamma}, but a pat answer may immediately be given: When $\Gamma$ has been learnt, the system can be driven autonomously and then $G_T$ is known anyway.
+The reasons for taking this roundabout approach remain to be discussed in Section~\ref{subs_LearnGamma}, but a pat answer may immediately be given.
+When $\Gamma$ has been learnt, the system can be driven autonomously and then $G_T$ is known anyway.
 We formalise this with a theorem.
@@
 \vspace{-6mm}
-{\bf Proof.}  See~\cite[Th. 3c]{manjunath2021universal}.
+{\bf Proof.}  See~\cite[Theorem. 3c]{manjunath2021universal}.
 
+The graph in equation~\ref{comm_H_CET} is then extended to the diagram~\ref{Scomm_imp} below.
 Projection mappings $\pi_i$ are defined in the traditional meaning where a $k$-dimensional vector is projected to it's $i^{th}$ component such that \[\pi_i:(a_1,a_2, \ldotp, a_{k-1}, a_k)\to{a_i}\]
 
-The graph in equation~\ref{comm_H_CET} is then extended to the diagram~\ref{Scomm_imp} below:
-
 \begin{equation} \label{Scomm_imp}
-  %    \[ 
-      \psset{arrows=->, arrowinset=0.25, linewidth=0.6pt, nodesep=3pt, labelsep=2pt, rowsep=1.3cm, colsep = 1.1cm, shortput =tablr}
-   \everypsbox{\scriptstyle}
-   \begin{psmatrix}
-   \widehat{U}_T & \widehat{U}_T\\%
-   Y_T & Y_T\\
-   & \textcolor{red}{U \times X}
-   %%%
-  %  \ncline{1,1}{1,2}^{\widehat{T}} \ncline{1,1}{2,1} <{H_2}
-  %  \ncline{1,2}{2,2} > {H_2}
-  %  \ncline{2,1}{2,2}^{G_T}
-  %  \psset{linecolor=red}
-  %  \textcolor{red}{\ncline{2,1}{3,2}<{(\Gamma,\pi_2)}}
-  %  \ncline{3,2}{2,2}>{\textcolor{red}{(\pi_2,g)}}.
-   \end{psmatrix}
-  % \]
+\begin{tikzcd}
+	{\widehat{U}_T} && {\widehat{U}_T} \\
+	\\
+	{Y_T} && {Y_T} \\
+	\\
+	\\
+	&& {U\times X}
+	\arrow["{\widehat{T}}"', from=1-1, to=1-3]
+	\arrow["{H_2}", from=1-1, to=3-1]
+	\arrow["{H_2}"', from=1-3, to=3-3]
+	\arrow["{G_T}", from=3-1, to=3-3]
+	\arrow["{(\Gamma, \pi_2)}"{description}, color={rgb,255:red,214;green,92;blue,92}, from=3-1, to=6-3]
+	\arrow["{(\pi_2, g)}"{description}, color={rgb,255:red,214;green,92;blue,92}, from=6-3, to=3-3]
+\end{tikzcd}
 \end{equation}
 
-The problem finally simplifies to the issue of learning the map $\Gamma$ and combining this with the projection mapping $\pi_2$ and the function $g$, which will be known.
-A final set of equations is obtained - equations that have been entirely constructed from data.
+Our task now finally simplifies to the question of learning the map $\Gamma$ and then combining it with the projection mapping $\pi_2$ and the function $g$, which will both be known.
+A final set of equations is obtained -- equations that have been \textit{entirely} constructed from data.
 \begin{eqnarray}\label{eqns_from_data}
@@
 
-One may immediately ask why we opt to take such a roundabout route; why not just learn the map $G_T$ from the get-go?
-On the surface, this seems to be an arbitrary decision path with no real reasoning, so we take a pause again and discuss the motivation for learning Gamma.
+It is a completely valid to wonder why we opt to take such a roundabout route; why not just learn the map $G_T$ from the get-go?
+The decision does seem arbitrary on the surface, so we take a pause again and discuss the motivation for learning Gamma.
 There are a number of distinct advantages.
 
-In the first place, learning $\Gamma$ saves computational resources.
-This is due to the fact that the input $u_n$ lies in a lower-dimensional subspace in comparison to the high-dimensional space $X\times{X}$.
+Firstly, learning $\Gamma$ saves computational resources. This is due to the fact that the input $u_n$ lies in a lower-dimensional subspace in comparison to the high-dimensional space $X\times{X}$.
 
@@
 Secondly, the function $\Gamma$ is known to be stable.
-Learning $\Gamma$ makes use of equations~\ref{eqns_from_data} which in turn employs the driven system $g$ possessing the USP, and offers distinct advantages with regards to stability in the presence of perturbation.
-It is desirable that inputs that differ only slightly would have only ‘slight’ effects on the output of the system.
-Moreover, one would hope to prevent numerical errors originating from input and measurement noise (see~\cite[Th.
+Learning $\Gamma$ makes use of equations~\ref{eqns_from_data}\ednote{how can I reference both 4.13 and 4.14?} which in turn employ the driven system $g$ possessing the USP.
+This offers distinct advantages with regards to stability in the presence of perturbation.
+It is desirable that inputs differing slightly would have only ‘slight’ effects on the system's output.
+Moreover, one would hope to prevent numerical errors originating from input and measurement noise (see~\cite[Theorem.
 5]{manjunath2021universal}).
-As the data fed into the system is a function of both the input given and the parameters of the system itself (i.e.
+As the data fed into the system is a function of both the input and the parameters of the system itself (i.e.
 the value a, $\alpha$ in $\tanh$), we exploit both input-related and parameter-related stability.
-Both these notions are defined by way of the continuity of an encoding map of an input and design parameter respectively in~\cite{manjunath2020stability}.
+Both these notions are defined by way of the continuity of an encoding map of the respective input and design parameters~\cite{manjunath2020stability}.
 
-Informally, we may consider input-related stability as concerned with the question of whether or not small variations in input would result in small responses.   
-If two inputs $\bar{u}$ and $\bar{u}_{noisy}$ are close in the product topology, then their tails could differ greatly due to the metric that generates the product topology has insensitivity to the differences in sequence's tails.
- However if the system has the USP then the function $h$ is continuous and so $h(\bar{u}), h(\bar{u}_{noisy})$ remain close-by for small levels of noise maintaining a measure of robustness to input and measurement noise.\cite{manjunath2021universal}.
+Informally, we may consider input-related stability relating to the question of whether or not small variations in input would result in small responses.   
+If two inputs $\bar{u}$ and $\bar{u}_{noisy}$ are close in the product topology, their tails could differ greatly due to the metric that generates the product topology insensitive to the differences in sequence's tails.
+ However, if the system has the USP, then the function $h$ is continuous and so $h(\bar{u})$, $h(\bar{u}_{noisy})$ remain close-by for low levels of noise; this maintains a measure of robustness to input and measurement noise.\cite{manjunath2021universal}.
 In a similar fashion, parameter-related stability is obtained thanks to a result~\cite[Lemma 3.2]{manjunath2020stability} relating this form of stability with the ESP (which is here equivalent to the USP).
 
-% Secondly,  due to $g$ possessing the USP, both input- and parameter-related stability~\cite{manjunath2020stability} is obtained, which in turn prevents numerical errors that caused by input and measurement noise (see~\cite[Th.
+% Secondly,  due to $g$ possessing the USP, both input- and parameter-related stability~\cite{manjunath2020stability} is obtained, which in turn prevents numerical errors that caused by input and measurement noise (see~\cite[Theorem.
 5]{manjunath2021universal}).
 
-Thirdly, learning a chaotic map on a set (here $Y_T$) with even a small measure of noise  means that the learnt system accepts arguments out of the set (Problems similar to Takens, cf.~\ref{sect_takenslimits}).
+Thirdly, learning a chaotic map on a set (here $Y_T$) where even a small amount of noise is present would lead to a learnt system accepting arguments out of the set (Problems similar to Takens, cf.~\ref{sect_takenslimits}).
 We thus have a map $G_T^+$ acting on $Y_T^+$ and since $Y_T$ is not guaranteed to be an attractor of $G_T$ this could lead to errors.
@@
 \begin{figure}[ht]
-  \includegraphics[scale=0.25]{Graphs/_YTerrors.eps}
+  \includegraphics[width=0.6\linewidth]{Graphs/_YTerrors.eps}
   \centering
@@
 
-To circumvent this: if a driven system with a saturation function\ednote{B: What is a saturation function?} is used, we can ensure the existence of an attractor $A$ of the map which implements $G_T$, and containing $Y_T^+$ indirectly (via $g$).
-\ednote{B: A little vague.
-Are we saying that an attractor of a map implementing $G_T$ is also the one that contains $Y_T^+$?}
-By defining the dynamics to $A$, this prevents large errors from occuring.  
-More details on the global dissipativity of $g$, as it is termed there, may be found in~\cite{Supp}.
-
+To circumvent this: if a driven system with a saturation function \footnote{A function $f(t)$ that is monotonic and approaches a constant as $|t|\to \infty$, for e.g., $\tanh(t)$ on $\mathbb{R}$} is used, we can ensure the existence of an attractor $A$ of the map that implements
+$(\pi_2,g) \circ (\Gamma,\pi_2)$ (printed in the red on Figure~\ref{Scomm_imp})
+ which in practice turns out to be defined on $Y_T^+$ has an attractor $A$  containing $Y_T$.
+By restricting the dynamics to not spill beyond $A$, prevents large errors from occuring.  
+More details on the global dissipativity of $g$, as it is termed here, may be found in~\cite{Supp}.
 \begin{figure}[ht]

### 2022-01-19 projectReport/chapter5.tex
Rereading the project. On 5.4

@@
-In this chapter we turn to the implementation of the function $G_T$ by discussing a method to learn the function $\Gamma$.
-We apply our methodology to some chaotic attractors.
-Most notably, we consider the reconstruction of the chaotic attractor of the double pendulum -- an  attractor that has not been successfully reconstructed because of its intricate structure.
+In this chapter we turn to the implementation of the function $G_T$ by discussing a method to learn the function $\Gamma$. We apply our methodology to some chaotic attractors. 
+Most notably, we consider the reconstruction of the chaotic attractor of the double pendulum -- an  attractor that has not previously been successfully reconstructed because of its intricate structure.
 In fact the previous data-driven approaches also have not been successful in building a model from the double pendulum.
@@
 \section{Implementing $G_T$}
-The map $G_T$ is implemented by applying the functions $\Gamma$, $g$ and the projection function $\pi_2$ in the relevant order as described by the equations~\ref{eqns_from_data}.
-To reach this point, one must first learn $\Gamma$.
-After first normalising the data to ensure that it has zero mean value and lies within the range $[-1,1]$ and required zero-padding (see paragraph below for further discussion), we continue by driving the system with external input through the function $g$.
+The map $G_T$ is implemented by applying the functions $\Gamma$, $g$ and the projection function $\pi_2$ in the relevant order as described by the equations~\ref{eqns_from_data}\ednote{How do I get both?}.
+To reach this point though, one must first learn $\Gamma$.
+After first normalising the data to ensure that it has zero mean value, lies within the range $[-1,1]$, and has been zero-padded as required (see paragraph below for further discussion), we drive the system with external input through the function $g$.
 
-From~\eqref{eqn_driving} we recall the equation of the RNN: $g(u,x) = (1-a) x + a \tanh(Au + \alpha B x)$.
-Recall that $g$ is SI-invertible when $A$ is invertible and has same dimension as $B$.
-We refer to $A$ as the input matrix and $B$ the reservoir matrix.
+From~\eqref{eqn_driving} we recall the equation of the RNN: \[g(u,x) = (1-a) x + a \tanh(Au + \alpha B x).\]
+Recall that $g$ is SI-invertible when the matrix $A$ has the same dimension as $B$ and is invertible. We refer to $A$ as the \textit{input} matrix and $B$ the \textit{reservoir} matrix.
 
-We make use of a Recurrent Neural Network (RNN) in our implementation to project the data on to a higher dimensional space, but we do not train the RNN in the conventional manner.
-\ednote{B: Should we say what the conventional manner is?} A few words about the RNNs: RNNs are widely used in the literature for temporal processing of data as a type of deep learning technique.
-They take the form of equations considered to be discrete-time version of continuous dynamical systems modelling a neural network of neurons~\cite{jaeger2001echo}.
+We make use of a Recurrent Neural Network (RNN) in our implementation to project the data on to a higher dimensional space.
+We do not train the RNN in the ways familiar to literature and opt to follow the methodology of~\cite{manjunath2021universal}.
+A few words about the RNN's: RNN's are widely used in the literature for temporal processing of data as a type of deep learning technique. 
+They take the form of equations considered to be discrete-time version of continuous dynamical systems that model a neural network of neurons~\cite{jaeger2001echo}.
 
-We implement a feedforward network, (a network with no memory, unlike the RNN) required to learn $\Gamma$ in the \emph{Python} programming language and making use of the \emph{Scipy} library, alongside the \emph{Keras} package in the \emph{Tensorflow} library
-% \ednote{B: Is this something you want me to write in?
-M: We do not implement the RNN, but the FFN that learns Gamma through Tensorflow.
-We called the RNN + FN set up a RCN in Adriaan's paper.
-FYI: FN's work only if there is a definite relationship between the $u_n$ and $u_{n+1}$ which is rare in temporal data.
-RNNs with ESP work when there is a definite relationship between $(\ldots,u_{n-1},u_n)$ and $u_{n+1}$ and is suited for temporal processing.}.
+We implement a feedforward network (a network with no memory, unlike the RNN) to learn $\Gamma$ in the \emph{Python} programming language and making use of the \emph{Scipy} library, alongside the \emph{Keras} package in the \emph{Tensorflow} library .
 
-If the RNN we will be using in the next step has $N$ neurons, then $A$ and $B$ are each $N\times{N}$ matrices.
+If the RNN we will be using has $N$ neurons, then $A$ and $B$ are each $N\times{N}$ matrices.
 Inputs $u_n$ of dimension $K < N$ are padded by zeroes (as remarked earlier) to embed the input in a space of dimension $N$.
-As we do not need the entire left-infinite history of input, an arbitrary initialisation of the system with initial values followed by the driving of the system with the function $g$ means that the generated values will approach the actual solution elements in a uniform manner thanks to the UAP that we had discussed in the preceding chapter.
+As we do not need the entire left-infinite history of an input, an arbitrary initialisation of the system with initial values, followed by the driving of the system with the function $g$ will mean that the generated values approach the actual solution elements in a uniform manner.
+This follows from the UAP that discussed in the preceding chapter (see~\ref{Dfn_UAP} and subsequent discussion).
 
-Thirdly, we prepare the network by setting up a network with $N$ neurons (the same value $N$ as in the dimensions of the matrices $A,B$).
-Learning of $\Gamma$ is done using the Adam Optimiser where the means square error (MSE) loss function is optimised.
+Thirdly, we prepare the network by setting up a network with $N$ neurons  to obtain successive state values as samples in the space $Y_T$. 
+The actual learning of $\Gamma$ is done using the Adam Optimiser where the means square error (MSE) loss function is optimised. 
 The system is trained several times (in our case 4) with incrementally smaller learning rates on each run (0.001 and thereafter divided by 10 at each run). 
@@
 
-Once $\Gamma$ is learnt then reach the prediction step, where we forecast $u_n$ from $(x_{n-1},x_{n})$.
+Once $\Gamma$ is learnt, we progress to the the prediction step where $u_n$ is forecasted from $(x_{n-1},x_{n})$.
 Note that  due to causal embedding $(x_{n-1},x_{n})$ is in one-to-one correspondence with the left-infinite orbit $(\ldots,u_{n-1},u_{n-1})$.
-Once $u_n$ is computed using $\Gamma$, we feed that to the driven system $g$ along with  $x_n$ to obtain $x_{n+1}$.
+When $u_n$ is computed using $\Gamma$, it is fed to the driven system $g$ along with  $x_n$ to obtain $x_{n+1}$.
 Now we apply $\Gamma$ on $(x_n,x_{n+1})$ to forecast $u_{n+1}$ and the whole process repeats.
 Note that this corresponds to implementing the two model equations for learning from data~\eqref{eqns_from_data}.
-% We then reach the prediction step. Making use of $g$'s SI-invertibility and $g$ as defined in~\eqref{eqn_driving}, we have the following expression for $u_n$ replicated from~\eqref{eqn_SI_RNN}.
-% \begin{equation*}
-%   u_n := A^{-1}\bigg(\overline{\tanh}^{^{-1}}\frac{1}{a}\Big(x_{n+1}-(1-a)x_n\Big) \bigg) - \alpha B x_n
-% \end{equation*}
-
-% The two model equations for learning from data~\eqref{eqns_from_data} allows us to determine what the state $x_{n+1}$ at time $n+1$ will be. 
-% We specifically remind ourselves of \[x_{k+1}=\pi_2 \circ (\Gamma, \pi_2) \circ (\pi_2,g) (u_k,x_k)\]
 
@@
 Explicitly, if $(W,T)$ is a dynamical system with dynamics generated by $w_{n+1}=Tw_n$, and if $\theta:W \to \mathbb{R}$ is an observable, then the task is to learn a dynamical system that is topologically conjugate to $(W,T)$ and predict $\theta(w_{m+1}),\theta(w_{m+2}),\ldots$ using the data $\theta(w_{0}),\theta(w_{1}),\ldots,\theta(w_{m})$.
+
 So how do we do it?
-Suppose the input generated from the delay-coordinate map $\Phi_{\theta,2d}(\theta(w_{n})) := (\theta(w_{n-2d}),\ldots,\theta(w_{n-1}),\theta(w_{n}))$ is such that Takens delay embedding theorem (see Theorem~\ref{Thm_Takens}) holds, in which case, there exists a homeomorphism  $F_\theta: \Phi_{\theta,2d}(\theta(w_{n})) \mapsto \Phi_{\theta,2d}(\theta(w_{n+1}))$.
-Then if we feed  the input values $u_n := \Phi_{\theta,2d}(\theta(w_{n}))$ to a driven system $g$ that is SI-invertible and has USP, the induced system $(Y_F,G_F)$ would be topologically conjugate to the inverse-limit space of
-$(\Phi_{\theta,2d}(W), F_\theta)$ due to Theorem~\ref{Thm_CET}, and hence one could forecast $u_n,u_{n+1},\ldots$.
+Suppose the input generated from the delay-coordinate map $\Phi_{\theta,2d}(\theta(w_{n})) := (\theta(w_{n-2d}),\ldots,\theta(w_{n-1}),\theta(w_{n}))$ is such that Takens delay embedding theorem (see Theorem~\ref{Thm_Takens}) holds.
+In that case, there exists a homeomorphism  $F_\theta: \Phi_{\theta,2d}(\theta(w_{n})) \mapsto \Phi_{\theta,2d}(\theta(w_{n+1}))$.
+If we then feed the input values $u_n := \Phi_{\theta,2d}(\theta(w_{n}))$ to a driven system $g$ having SI-invertibility and the USP, the induced system $(Y_F,G_F)$ would be topologically conjugate to the inverse-limit space of
+$(\Phi_{\theta,2d}(W), F_\theta)$ due to Theorem~\ref{Thm_CET}, and hence one could forecast the values ($u_n,u_{n+1},\ldots$).
+
 In summary, one could forecast  the values $\theta(w_n), \theta(w_{n+1}),\ldots$ by feeding these observations as delay coordinates since the input is again an orbit from a dynamical system. 
-The advantage of feeding delay-coordinates to a driven system  over learning a map that describes the evolution of the delay coordinates is that the embedding is stable in the sense of global dissipativity that we have described in the previous chapter, i.e., there is an attractor $A$ of a map that implements $G_T$ on an expansion $Y_F^+$ providing some robustness to noise.
+The advantage of feeding delay-coordinates to a driven system over learning a map that describes the evolution of the delay coordinates is that the embedding is stable in the sense of global dissipativity that we have described in a previous section~\ref{subs_LearnGamma}.
+I.e., there is an attractor $A$ of a map that implements $G_T$ on an expansion $Y_F^+$ providing some robustness to noise.
 
-Now discuss on another practical issue with forecasting chaotic dynamical systems.  
-We note that the twin equations~\ref{eqns_from_data} generate a model from data.
-For a chaotic system, a small amount of noise due either to the input-noise or computational error would cause the solutions from even an accurate model with the same initial conditions to diverge.
-However, statistically, a chaotic system would behave well in the sense that the visiting frequency or the average time spent by an orbit in any region of the phase space of the dynamical system is identical for almost all initial conditions (almost all is specified as a measure), a result in ergodic theory (\cite{wikiErgodicTheory}) a topic beyond the scope of this thesis.
-To illustrate the accuracy of the twin  equations~\ref{eqns_from_data} in modeling the data we plot the densities of the trajectories of the actual and predicted solutions in~\ref{noisydp_success_density, fig:dp_success_density}.
+We now turn our attention to another practical issue related to the forecasting of chaotic dynamical systems.  
+Recall that the twin equations~\ref{eqns_from_data} generate a model from data. 
+For a chaotic system, a small amount of noise due either to input-noise or computational error would cause solutions from even a very accurate model with the same initial conditions to diverge.
+%\ednote{Grammarly}
+Statistically, however,  a chaotic system will behave well in the sense that the visiting frequency (the average time spent by an orbit in any region of the phase space of the dynamical system) is identical for almost all initial conditions.
+The phrase `almost all' here is specified as a measure and refers to a result in ergodic theory (\cite{wikiErgodicTheory}) which is a topic beyond the scope of this project.
+To illustrate the accuracy of the twin  equations~\ref{eqns_from_data} in modelling the data, we plot the densities of the trajectories of the actual and predicted solutions in ~\ref{fig:dp_success_density},~\ref{fig:noisydp_success_density} and ~\ref{fig:Thomas_TrajDensity}.
 
@@
 Secondly, we opt to enforce the requirement that both matrices be well-conditioned, i.e.
-that they have a condition number of \~1.
-The condition number of a matrix $M$ (denoted $cond(M)$) is the product of the matrix norms of the matrix $M$ and its inverse $M^{-1}$ and is a measure of the sensitivity of output of a matrix multiplication when an input is changed.
+that they have a condition number of 1.
+The condition number of a matrix $M$ is the product of the matrix norms of the matrix $M$ and its inverse $M^{-1}$ and is a measure of the sensitivity of output of a matrix multiplication when an input is changed.
 We do not have any theoretical justifications for this.
 
-%By ensuring a condition number close to 1, we also ensure that our driven function $g$, depending on the matrices $A$ and $B$, will not be sensitive to adjustments in the input.
-In so doing we then also guarantee another form stability.
-\ednote{B: Not complete enough.
-Edit further and cite M: Please feel free to cite a wikipedia page}
-
-% Lastly, we ensure the matrix $B$ has unit spectral radius.
-This is necessary due to the Causal Embedding Theorem~\ref{Thm_CET} which can then guarantee the existence of semi-conjugacy between the systems $(Y_T, G_T)$ and $(\widehat{U}_T, \widehat{T})$.
-In fact, the theorem actually requires that $\alpha{B}$ have spectral radius $<1$, so we impose the restriction $\alpha<1$ and consider only the matrix $B$'s spectral radius from thereon.
 Lastly, we ensure the matrix $B$ has unit spectral radius so that the variable $\alpha$ in the RNN equation \eqref{eqn_driving} can be used to know if the USP is satisfied. 
-It is a  feature that a matrix with a spectral radius $<1$ generally has the USP \cite{yildiz2012re}.
+It is a  feature that a matrix with a spectral radius less than 1 generally has the USP~\cite{yildiz2012re}.
 
@@
 To learn $\Gamma: (x_{n-1},x_{n}) \mapsto u_n$ through our feedforward network, we could feed the principal components.
-Principal components are used only for a more efficient state representation to possibly reduce learning errors; it is not for a lossy approximation since we use all principal components.
+Principal components are used only for a more efficient state representation to possibly reduce learning errors; it is not a lossy approximation since we use all principal components.
+
 More explicitly, denote $X_{1:N}$ as the matrix with the first $N$ states of the network data represented by row vectors. 
-If $X_{1:N}=U\Sigma P^T$ here represents the singular value decomposition ($U$ and $V$ are unitary matrices and $\Sigma$ is a diagonal matrix) of $X_{1:N}$ and we refer to \cite{wikiSVdecomp}.
-We then denote the principal component matrix $P$, and the principal components are given by
+$X_{1:N}=U\Sigma P^T$ here represents the singular value decomposition of $X_{1:N}$ ($U$ and $V$ are unitary matrices, and $\Sigma$ is a diagonal matrix) and we refer to \cite{wikiSVdecomp}. 
+We then denote the principal component matrix by the symbol $P$, and the principal components are given by 
  $Z_{1:N}=X_{1:N}P$.
-These principal components are then used to train a feedforward neural network. Denote the row vectors of $Z_{1:N}$ by 
-$z_i^T, i=1,2,\ldots,N$ and the neural network by $NN$, the network is subsequently trained to learn an approximation of the map
+These principal components are used to train a feedforward neural network.
+
+Denoting the row vectors of $Z_{1:N}$ by  $z_i^T, i=1,2,\ldots,N$ and the neural network by $NN$, the network is trained to learn an approximation of the map
 $NN: (z_{n-1},z_n) \mapsto  u_n.
 $
-The learnt approximation of $NN$ can be used to approximate $\Gamma$ since 
+The learnt approximation of $NN$ can be used to approximate $\Gamma$ because 
 \begin{equation}\label{Seq_RNN}
@@
 \section{Experimental Results}
-In this section we showcase numerical results pertaining to 3 attractors that were simulated using the theory and methodology described.
+In this section we showcase numerical results pertaining to three attractors that were simulated using the theory and described methodology.
 We first consider a simple physical system that exhibits complicated dynamics.
@@
 The first (or higher pendulum) is attached to a fixed point while the second is attached to the endpoint of the first.
-(See below the Figure~\ref{fig:dp_setup})
-This system shows chaos strikingly. Consider again Figure~\ref{fig:dp_sdic}.
-The system has four variables: angle and angular velocity of each pendulum rod. An ideal system is assumed where the rods are massless and no friction in the pivots are considered.
+(See Figure~\ref{fig:dp_setup} below.)
+This system shows chaos in a striking manner -- consider once more Figure~\ref{fig:dp_sdic}.
+The system has four variables: for each pendulum rod an associated angle and angular velocity. An ideal system is assumed where the rods are massless and pivots are frictionless.
 \begin{figure}[ht]
-  \includegraphics[scale=0.35]{Graphs/_dp_setup.eps}
+  \includegraphics[width=0.8\linewidth]{Graphs/_dp_setup.eps}
   \centering
-  \captionof{figure}{Pictorial representation of the setup of a DP. Image taken from~\cite{DPSetup}}
+  \captionof{figure}{Pictorial representation of the setup of a DP where $\theta, m, l$ represent the angle, rod-length, and pendulum head mass respectively.}
 \label{fig:dp_setup}
@@
 The scalar equations of motion are derived using Newtonian physics and are reproduced from~\cite{DPFormulas} in Equations~\ref{eqns_dp}. 
-Each pendulum component has an associated mass of the head, $m$, and length of the rod, $L_i$.
-Acceleration due to gravity is denoted by $g$ and initial velocities by $\dot{\theta}$
+Each pendulum component has an associated mass of the head, $m$, and length of the rod, $L_i$. Acceleration due to gravity is denoted by the variable $g$ and initial velocities by $\dot{\theta}$
 \begin{eqnarray}\label{eqns_dp}
@@
 Here we chose the values $m_1=2.0$, $m_2=2.0$, $L_1=1.5$ and $m_2=1.5$ with initial velocities $v_1=1.5, v_2=-2$ for the respective pendulum heads.
-With a change of variables the second-order system can be 
-written as a system of first order equations with four variables, where two variables devoted to each pendulum denote the $x$ and $y$ coordinates of the mass (the lengthy equations are not shown here).
+With a change of variables the second-order system can be written as a system of first order equations with four variables, where each pendulum has an associated $x$ and $y$ coordinates of the mass (the lengthy equations are not shown here).
 The ODE system is then solved using the fourth-order Runge-Kutta numerical method.
@@
 
-We need the assumption the system is not damped so that the map $T$ resulting from the discretization of the system  is surjective.
-If we assume damping surjectivity is lost on any phase space with more than one point.
-An example of this would be the case of a DP exhibiting energy-loss. In such a case, the attractor is a single point in the phase space when the double pendulum comes to rest. 
-For such a system, it may be noted that if we use data outside the attractor, learning is not reliable. To see this, consider the plots ~\ref{fig:damped_pendulum} and ~\ref{fig:dp_notsurjective}.
+We need the assumption of an undamped system so that the map $T$ resulting from the discretization of the system  is surjective.
+If we assume, damping surjectivity is lost on any phase space with more than one point.
+An example of this would be the case of a DP exhibiting energy-loss.
+In such a case, the attractor is a single point in the phase space representing the instant when the double pendulum comes to rest.
+If we use data outside the attractor in this scenario, learning is not reliable. To see this, consider the plots~\ref{fig:damped_pendulum} and~\ref{fig:dp_notsurjective}.
 
@@
 The system was driven for 15500 steps, of which the first 500 were immediately discarded to simulate the network `forgetting'. 
-The driving function $g$ had input and reservoir matrices were implemented as sparse matrices filled to 20\%.
-Parameter-values of $\alpha=0.99$ and $a=0.5$ were used.
+Input and reservoir matrices of the driven system $g$ were implemented as sparse matrices filled to 20\%. Parameter-values of $\alpha=0.99$ and $a=0.5$ were used.
 The remaining 15000 datapoints were used to train a network with 128 neurons. 
-The network was initialised to contain 16 hidden layers each with a layer dimension of 64, activated with the $\tanh$ activation function.
-Training was accomplished using 256 epochs with batch sizes 128 and a delay of 1.
-See Fig.~\ref{fig:dp_success_traj} for results pertaining to the trajectory taken by the DP's components and then Fig.~\ref{fig:dp_success_density} to witness associated densities of predicted versus actual values.
+The network was initialised to contain 16 hidden layers each with layer dimension 64 and activated with the $\tanh$ activation function.
+Training was accomplished using 256 epochs with batch sizes 128.
+See Fig.~\ref{fig:dp_success_traj} for results pertaining to the trajectory taken component-wise from the DP's and then Fig.~\ref{fig:dp_success_density} to witness associated densities of predicted versus actual values.
 
@@
 We now consider an experiment where noise is added to the system and find it to be incredibly robust.
-We employ the same parameters as in the preceding experiment and add noise from a normal-distribution with mean zero and standard deviation equal to 0.1 which gives a signal-to-noise ratio of 39dB.
+Employing the same parameters as in the preceding experiment, we include noise from a normal-distribution with zero mean value and standard deviation equal to 0.1 which equates to a signal-to-noise ratio of 39dB.
 
 For the signal-to-noise ration we adopt the formula $\text{SNR}_{dB}=10\text{log}_{10}\big(\frac{P_\text{{signal}}}{P_{\text{noise}}}\big)$ where $P_\text{{signal}}$ refers to the variance observed in the normalised data and $P_{\text{noise}}$ represents the variance of the added noise.
-\ednote{B:Added this sentence.}
-Once again consider the actual and predicted trajectories for the noisy DP dataset in Figure~\ref{fig:noisydp_success_traj}.
+Once again we display the actual and predicted trajectories for the noisy DP data in Figure~\ref{fig:noisydp_success_traj}. 
 Next we consider the densities of the predicted and actual values in Figure~\ref{fig:noisydp_success_density} and observe that they display highly similar distributions.
@@
 \subsection{Additional Attractors: Clifford \& Thomas}
-We follow the methodology presented in~\cite{manjunath2021universal} but opt to consider attractors different from the Lorenz system and H\'enon map in this project.
+We follow the methodology presented in~\cite{manjunath2021universal} but opt to consider attractors different from the Lorenz system and H\'enon map for this project.
 
 \subsubsection{Thomas' Cyclically Symmetric Strange Attractor}\label{Thomas_Attractor}
-Thomas' Cyclically Symmetric Strange Attractor, a 3D attractor proposed in 1999 by Ren\'e Thomas in~\cite{ThomasAttractor}, is described by a set of three equations which is distinctly reminiscent of the Lorenz system, another deterministic three-equation model exhibiting chaotic behaviour.
+Thomas' Cyclically Symmetric Strange Attractor, a 3D attractor proposed in 1999 by Ren\'e Thomas in~\cite{ThomasAttractor}, is described by a set of three equations which is distinctly reminiscent of another deterministic three-equation model exhibiting chaotic behaviour proposed by Lorenz.
 It has a single parameter $\beta$ and has been shown to transition to chaotic behavious when $\beta<0.208186$~\cite{Thomas_BetaParameter}.
@@
   y_{n+1} = \sin(z_n) - \beta{y_n} \\
-  z_{n+1} = \sin(x_n) - \beta{z_n}
+  z_{n+1} = \sin(x_n) - \beta{z_n}.
 \end{eqnarray}
 
-The behaviour of the system with a $\beta$-parameter value of 0.1056 was simulated by scaling the data from it to fit inside the interval $[-1,1]$.
+The behaviour of the system with a $\beta$-parameter value of 0.1056 was simulated by scaling the data generated from the equations to fall inside the interval $[-1,1]$.
 Two cases were considered: one noise-free and another perturbed by noise.
-Noise was generated  by adding data points taken from a normal distribution with mean 0 and standard deviation 0.05, which translates approximately to a signal-to-noise ration of 33dB. 
-($\text{SNR}_{dB}$ is calculated exactly as for the DP-data above)
+Noise was generated  by adding data points taken from a normal distribution with a mean of zero and standard deviation 0.05, which translates approximately to a signal-to-noise ration of 33dB. 
+$\text{SNR}_{dB}$ is calculated exactly as for the DP-data.
 
-A sequence of 3000 observations in both the clear and noisy data-sets respectively were fed into the system with 300 entries discarded to simulate the network's loss of memory. 
+A sequence of 3000 observations in both the clear and noisy data-sets respectively was fed into the system with 300 entries discarded to simulate the network's loss of memory. 
 The next 9000 steps were predicted and compared with true data. See Figure~\ref{fig:Thomas_trajectories} for the actual and predicted trajectories of the noise-free and perturbed systems respectively.
@@
   \endminipage
- \captionof{figure}{True (red) and predicted (blue) trajectories for the Thomas attractor with parameter value $\beta=0.1056$.
-To the left: Noise added from normal distribution with mean zero and standard deviation 0.05.
+ \captionof{figure}{True (red) and predicted (blue) trajectories for the Thomas attractor with parameter value $\beta=0.1056$.Left: Noise added from normal distribution with mean zero and standard deviation 0.05.
 Right: No noise added to dataset.}
@@
   \includegraphics[scale=0.5]{Graphs/_Thomas_3.eps}\caption*{Represented here the learnt (blue) and actual (red) densities of the first coordinate ($x$-coordinate) of the dataset for the Cyclically Symmetric Attractor with parameter-value $\beta=0.1056$.}
-\end{figure}
+\end{figure} \label{fig:Thomas_TrajDensity}
 
@@
 
-Another two-equation discrete-time map named the Fractal Dream Attractor, or more commonly as the Pickover or Clifford Map (first discovered by Clifford A.
-Pickover and discussed in his fascinating book``Chaos in Wonderland" \cite{PickoverChaos}), was also considered.
+A finaldiscrete-time map named the Fractal Dream Attractor -- more commonly as the Pickover or Clifford Map (first discovered by Clifford A.
+Pickover and discussed in his fascinating book ``Chaos in Wonderland" \cite{PickoverChaos}) -- was also considered.
 \begin{eqnarray}\label{eqns_clifford}
   {x_{n+1}=\sin(ay_n) + c\cos(ax_n)} \\
-  {y_{n+1}=\sin(bx_n)+d\cos(by_n)}
+  {y_{n+1}=\sin(bx_n)+d\cos(by_n)}.
 \end{eqnarray}
+
+The behaviour of the system with a parameter values of $x_0=0$, $y_0=0$, $a = -1.7$,  $b = 1.8$, $c = -1.9$ and $d = -0.4$ was examined. 
+Data was scaled to fit inside the interval $[-1,1]$. Only the noise-free case was investigated. 
+Depictions of the empirical results are available in Figure~\ref{fig:Clifford}.
+ A sequence of 25000 observations was fed into the system with 15000 entries discarded. 
+ The reason for such a great number of discarded entries is the following: the system seems to have an incredibly long transient;driving it for less than 5000 entries only gives an accurate prediction to about 5 seconds and it starts improving when the discard is in the range $[10000,15000]$ to simulate the network's loss of memory.
+The next 1000 steps were predicted and compared with true data.
 
@@
 
-The behaviour of the system with a parameter values of $x_0=0$, $y_0=0$, $a = -1.7$,  $b = 1.8$, $c = -1.9$ and $d = -0.4$ was examined. 
-Data was scaled to fit inside the interval $[-1,1]$. Only the noise-free case was investigated. 
-Depictions of the empirical results available in Figure~\ref{fig:Clifford}.
- A sequence of 25000 observations was fed into the system with 15000 entries discarded \ednote{M: Why did you have to discard 15000?} to simulate the network's loss of memory.
-The next 1000 steps were predicted and compared with true data.
-
  The RNN was initialised exactly as for Thomas' Cyclically Symmetric Strange attractor: the dimension of the network is set equal to 500 neurons and 12 hidden layers of dimension 64 each were initialised.
@@
 \section{Functional Complexity Reduction in Noninvertible Maps}
-% Besides being able to learn the action of the Koopman operator exactly, our setup has other advantages over selecting or optimizing the choice of observables as in the Extended Dynamic Mode Decomposition (EDMD) algorithm (see, [21, 22]).
-We can alter the functional complexity of G T by changing the number of observables or equivalently by altering the dimension of X – in the case of a RNN implementation of g, by simply changing the number of neurons in the network.
-Empirically, increasing the dimension of X is found to increase the linear relationship (or intuitively reduces the functional complexity of G T ) that is measured as a generalization of the Pearson correlation coefficient to random vectors (see, [44]) between (x_{n−1} , x_{n}) and G_T (x_{n−1} , x_n), and such numerical evidence is tabulated in [11, Table 1].
-In contrast, expanding the set of observables while using EDMD to learn a map with a lower functional complexity is a highly involved task, not least because so much that one often does not know how to guess a new observable, as well as the fact that adding observables does not necessarily retain the invariance of the span of the observables.
-
-% Empirical evidence shows that a linearity measure like a Pearson correlation coefficient points at reduced functional complexity while learning the dynamics in the state space of the driven system.
-
-% Empirically (see Table ??), increasing the dimension of X is found to increase the linear relationship
-% (or intuitively reduce the functional complexity of G T ) that is measured as a generalization of the Pearson correlation coefficient to random vectors (see, [?]) between (x n−1 , x n ) and G T (x n−1 , x n ).
-If Σ a and Σ b denotes the covariance matrices of the vectors (x n−1 , x n ) and G T (x n-1 , x n ) respectively, and if Σ ab denotes the covariance matrix between the vectors (x n-1 , x_n ) and G_T (x n-1 , x_n ), then the multidimensional 18correlation coefficient is computed by using the traces of these matrices:
-% \[\pho = \frac{tr(\sum_{ab})}{tr(\sqrt{\sum_{a}\sum_{b}})}\]
-
-% This multidimensional correlation coefficient satisfies most of the well-known properties of the one-dimensional Pearson coefficient.
-In particular, ρ = ±1 if and only Y = AX + b for some invertible matrix A and vector b.
-Hence ρ, and in practice, the estimator ρ̂, can be used to measure the linear relationship between two random vectors of the same dimension and thus serves as an indicator of functional complexity.
-
-% We remark that the estimation of such linear relationship in (see Table ??) obtained by sample correlations alone is justified whenever {u n } is a realization of an ergodic process since then when g has the USP, any solution {x n } is also a realization of an ergodic process [?].
-In the ergodic input case the generalized Pearson correlation coefficient (between (x n−1 , x n ) and G T (x n−1 , x n )) is independent of n.
-
-% In the case of non-invertible maps, we note that the map $G_T$ is a homemorphism although $T$ is not.
-Further, empirically it is found that $G_T$ is much more linear than $T$ as measured by what is called the multidimensional correlation coefficient $\rho$.
-This coefficient is a generalization of the Pearson correlation coefficient for real-valued random variables to random vectors (see,~\cite{puccetti2019measuring}).
-We state teh details of computing this coefficient between
 The map $G_T : Y_T \to Y_T$ is a homeomorphism since it is conjugate to the homeomorphism $\widehat{T}$ even though $T$ may be a non-invertible map.
@@
 $(x_{n-1},x_n)$ and $G_T(x_{n-1},x_n)$. 
-If $\Sigma_{a}$ and $\Sigma_{b}$ denotes the covariance matrices of the vectors $(x_{n-1},x_n)$ and $G_T(x_{n-1},x_n)$ respectively, and if $\Sigma_{ab}$ denotes the covariance matrix between the vectors $(x_{n-1},x_n)$ and $G_T(x_{n-1},x_n)$, then the multidimensional correlation coefficient is computed by using the traces of these matrices:
+If $\Sigma_{a}$ and $\Sigma_{b}$ denotes the covariance matrices of the vectors $(x_{n-1},x_n)$ and $G_T(x_{n-1},x_n)$ respectively, and if $\Sigma_{ab}$ denotes the covariance matrix between the vectors $(x_{n-1},x_n)$ and $G_T(x_{n-1},x_n)$, then the multidimensional correlation coefficient is computed by using the traces of these matrices.
 \[

### 2022-01-19 projectReport/chapter5.tex
Rereading done til end Ch.5

@@
 \section{Functional Complexity Reduction in Noninvertible Maps}
-The map $G_T : Y_T \to Y_T$ is a homeomorphism since it is conjugate to the homeomorphism $\widehat{T}$ even though $T$ may be a non-invertible map.
-We remark that in the case of a non-invertible map, empirically it is found that $G_T$ is much more linear than $T$ as measured by what is called the multidimensional correlation coefficient $\rho$.
-This amplification in linearity  or the increased value of the multidimensional correlation coefficient $\rho$  is also true for invertible maps and we refer to empirical results in \cite{Supp}.
-Here we consider only one non-invertible map pertaining to the Clifford attractor to add to the results in \cite{Supp}.
-This coefficient is a generalization of the Pearson correlation coefficient for real-valued random variables to random vectors (e.g., \cite{puccetti2019measuring}).
-We state the details of computing this coefficient between
-$(x_{n-1},x_n)$ and $G_T(x_{n-1},x_n)$. 
-If $\Sigma_{a}$ and $\Sigma_{b}$ denotes the covariance matrices of the vectors $(x_{n-1},x_n)$ and $G_T(x_{n-1},x_n)$ respectively, and if $\Sigma_{ab}$ denotes the covariance matrix between the vectors $(x_{n-1},x_n)$ and $G_T(x_{n-1},x_n)$, then the multidimensional correlation coefficient is computed by using the traces of these matrices.
+The map $G_T : Y_T \to Y_T$ is a homeomorphism since it is conjugate to the homeomorphism $\widehat{T}$ even though $T$ may itself be a non-invertible map.
+In the case of a non-invertible map, it is empirically found that $G_T$ is much more linear than $T$ when measured by the multidimensional correlation coefficient $\rho$.
+This amplification in linearity visible through the increased value of the multidimensional correlation coefficient $\rho$  also holds for invertible maps and we refer to empirical results in \cite{Supp}.
+
+Here we consider only the non-invertible map of the Clifford attractor to supplement the results in \cite{Supp}. 
+This statistic is a generalisation of the Pearson correlation coefficient for real-valued random variables to random vectors (e.g., \cite{puccetti2019measuring}) and we present the details of computing this coefficient between $(x_{n-1},x_n)$ and $G_T(x_{n-1},x_n)$.
+If $\Sigma_{a}$ and $\Sigma_{b}$ denote the covariance matrices of the vectors $(x_{n-1},x_n)$ and $G_T(x_{n-1},x_n)$ respectively, and $\Sigma_{ab}$ denotes the covariance matrix between the vectors $(x_{n-1},x_n)$ and $G_T(x_{n-1},x_n)$, then the multidimensional correlation coefficient is computed by using the traces of these matrices.
 \[
@@
 \]
-This multidimensional correlation coefficient satisfies one of the desired and well-known properties of the one-dimensional Pearson coefficient.
-That is, $\rho=\pm 1$ if and only if $Y\overset{d}{=}AX+b$ for some invertible matrix $A$ and vector $b$.
-Hence $\rho$, and in practice can be used  to measure the linear relationship between two random vectors of the same dimension and thus serves as an indicator of functional complexity.
+The value $\rho$ also satisfies one of the desired and well-known properties of the one-dimensional Pearson coefficient; that is, $\rho=\pm 1$ if and only if $Y\overset{d}{=}AX+b$ for some invertible matrix $A$ and vector $b$.
+In practice then, $\rho$, can be used to measure the linear relationship between two random vectors of the same dimension and thus serves as an indicator of functional complexity. 
 For the non-invertible system defined by the Clifford attractor, we indicate the difference by which $\rho$ changes for $G_T$ in Table~\ref{tbl_attractorsPearson}.
-Following~\cite{manjunath2021universal}, we say that this multidimensional correlation coefficients $\rho$ to evidence a general reduction in the functional complexity of the map $G_T$ emerging due to a RNN.
-In the
-Table~\ref{tbl_attractorsPearson}, the second column corresponds to the amount of artificial neurons (the dimension of $X$) in the implemented RNN.
+Following~\cite{manjunath2021universal}, we state that the multidimensional correlation coefficient $\rho$ evidences a general reduction in the functional complexity of the map $G_T$ emerging due to a RNN.
+In Table~\ref{tbl_attractorsPearson}, the second column corresponds to the amount of artificial neurons (the dimension of $X$) in the implemented RNN.
 
@@
           \end{bmatrix}$ \\
-          \midrule      
-          % \hline
-          % $(w_n)$ is the Lorenz states, sampled every 0.1 timestamps. & & &\\
-          % \multirow{3}{*}{$u_n=\frac{1}{100}w_n$} & 10 &0.9311  & 0.9731\\
-          %     & 100 &  & 0.9934\\
-          %     & 1000 & & 0.9930\\              
-          % \multirow{3}{*}{$u_n=\frac{1}{10}(\sin(0.1w_{n,x})+\sin(0.1w_{n,y})+\sin(0.1w_{n,z}))$} 
-          %     & 10 &0.8401  & 0.9392\\
-          %     & 100 & & 0.9616\\
-          %     & 1000 & & 0.9737\\
-          % \hline         
+
                \midrule  
@@
               & 1000 & & 0.84854 \\
-          % \hline 
-          % $(w_n)$ comes from the Logistic map: & & & \\
-          % {$ w_{n+1}  = 4w_n(1-w_n)$} & & & \\
-          % \multirow{3}{*}{$u_n = w_n - 0.5$} 
-          %     & 10 &-0.0314  & 0.5577\\
-          %     & 100 &  & 0.7150\\
-          %     & 1000 & & 0.7826\\
-          % \hline
-          % \hline
-          % $(w_n)$ comes from the Svensson Attractor \\ (for $a = -1.7; b = 1.8; c = -1.9; d = -0.4$): & & & \\
-          % {$w_{n+1}= \begin{bmatrix} d\sin(aw_{n,x}) - \sin(bw_{n,y}) \\ 
-          %                                                             c\cos(aw_{n,x})+\cos(bw_{n,y}) \end{bmatrix}$} & & & \\
-          %   \multirow{3}{*}{$u_n = w_n-\overline{w}$}
-          %       & 10 & -0.2243 & 0.49698 \\
-          %       & 100 & &  0.63310 \\
-          %       & 1000 & & 0.71151 \\
+
            \bottomrule
@@
               is provided to exhibiting that an increase in dimension of the driven system will 
-              typically result in a map $G_T$ with less functional complexity.}
-      \end{table}\label{tbl_attractorsPearson}
+              typically result in a map $G_T$ with less functional complexity.}\label{tbl_attractorsPearson}
+      \end{table}
 % \end{center}
 
-In summary, the numerical results not only corroborate the theory of employing the notion of causally embedding a dynamical system for data-driven modeling by showing both long-term topological consistency (the attractor reconstruction) and statistical consistency (the densities of the actual and predicted match).
+In summary, the numerical results presented here corroborate the theory that causally embedding a dynamical system for data-driven modelling show long-term topological consistency (the attractor reconstruction) and statistical consistency (the densities of the actual and predicted match).

### 2022-01-19 projectReport/abstract.tex
All G edits done

@@
 
-    In this project, the base assumption is that data is originates from  some discrete-time dynamical system, particularly one of a chaotic nature, but no assumptions are made pertaining to a specific dynamical system.
+    In this project, the base assumption is that data originates from  some discrete-time dynamical system, particularly one of a chaotic nature, 
+    but no assumptions are made pertaining to a specific dynamical system.
     A general learning problem is formulated and a well-known approach to the problem, alongside its practical limitations, is then considered.
-    The notion of driven dynamical systems is introduced, and we subsequently establish the theoretical existence of a dynamical system possessing certain properties which guarantees it to be topologically conjugate (`dynamically equivalent') to the dynamics of the underlying system from which measurements were originally taken.
-    Our methodology is implemented practically by making use of recurrent neural networks to learn a map topologically equivalent to the underlying system that forecasts future states.
-    In so doing we present the workability of an approach to obtain long-term topological and statistically consistent predictions for simple physical systems as well as a some chaotic attractors.
+    The notion of driven dynamical systems is discussed, and we establish the theoretical conditions that guarantee its single-delay dynamics to be topologically conjugate (`dynamically equivalent') to the dynamics of the underlying system from which measurements were originally taken.
+    The methodology is implemented practically by making use of recurrent neural networks and deep learning methods to learn a map that in turn helps forecast the future observations of the underlying system.
+    In so doing we present the workability of an approach to obtain long-term topological and statistically consistent predictions for some chaotic attractors, and notably the double pendulum -- a system difficult to model from data.
     \end{abstract}

### 2022-01-19 projectReport/chapter2.tex
All G edits done

@@
 Of course, discrete-time dynamical systems need not always arise through a differential equation.
-Once again, we may consider the field of ecology, where discrete dynamical systems are often directly derived or assumed.
-\ednote{Could you refer me to a resource here?}
+Once again, we may consider the field of ecology, where discrete dynamical systems are often directly derived or assumed~\cite{saber2010introduction}.
 There is a school of thought that advocates discrete-time dynamical systems to be more natural for modelling real-world observations than differential equations.
@@
 
-To this end, we make 3 remarks.
+To this end, we make 3 remarks:
 \vspace{-5mm}
@@
 When $r=4$, the one-dimensional recurrence relation $x_{n+1}=4x_n(1-x_n)$ can be used as the kindergarten's model to exhibit the presence of SDIC.
-  In Graph~\ref{fig:log_sdic} below, we plot the first 50 iterates of the LM for $r=4$ and two slightly different initial values $x_0$. 
+  In the plot Fig.~\ref{fig:log_sdic} below, we plot the first 50 iterates of the LM for $r=4$ and two slightly different initial values $x_0$. 
   The values of $x_n$ are plotted agains time $n$  where they have initial values $x_0=0.2$ (blue) and $x_0=0.20001$ (red).
@@
   [\bf {Devaney's Chaos}]\label{Dfn_ChaosDec}\rm
-	A dynamical system $T: U \to U$ is said to exhibit chaos in the sense of Devaney if it satisfies the three properties.
+	A dynamical system $T: U \to U$ is said to exhibit chaos in the sense of Devaney if it satisfies the three properties:
 	\vspace{-5mm}
@@
     \begin{figure}[ht]
-      \includegraphics[scale=0.6]{Graphs/_tentmap_2.eps}
+      \includegraphics[width=0.5\linewidth]{Graphs/_tentmap_2.eps}
           \centering
@@
 To show topological similarity (or sameness) between two metric or topological spaces, one needs to establish a homeomorphism between these two spaces. 
-However, in the study of dynamical systems, establishing a homeomorphism does not indicate that the systems are dynamically related in any way.
+However, in the study of dynamical systems defined on two spaces, establishing a homeomorphism does not indicate that the systems are dynamically related in any way.
 For instance, the maps $Tu=u^2$ and $Tu=1-|2u-1|$ defined on $[0,1]$ have totally different behaviour. 
@@
 If $\phi:U\to{V}$ is a homeomorphism, then $S$ is said to be conjugate to $T$ and $\phi$ is called a conjugacy.
-If we relax the criterion on $\phi$ and merely require $\phi$ to be continuous and surjective where $\phi:U\to{V}$, then $\phi$ is a semi-conjugacy between $T$ and $S$, where T has domain $U$ and S has domain $V$; S is said to be semi-conjugate to T.
+If we relax the criterion on $\phi$ and merely require $\phi$ to be continuous and surjective where $\phi:U\to{V}$, then $\phi$ is a semi-conjugacy between $T$ and $S$ where T has domain $U$ and S has domain $V$; S is said to be semi-conjugate to T.
 \end{Definition}
 
-% \ednote{B: Does defining explicitly the domain and codomain of $\phi$ make it clear enough?
-M: Saying $\phi$ is then a semi-conjugacy between $T$ and $S$ could be ambiguous as it does not specify which system is on the top in the commutativity.
-So please edit accordingly.}
 When $S$ is conjugate to $T$, the dynamics of the two systems are in some way `dynamically equivalent'.

### 2022-01-19 projectReport/chapter3.tex
All G edits done

@@
-In this chapter, we acquaint ourselves with the question of forecasting dynamical systems that have unknown underlying structures, state and discuss the shortcomings of the Takens delay embedding theorem, and discuss various issues faced while forecasting.
+In this chapter, we acquaint ourselves with the question of forecasting dynamical systems with unknown underlying structures, state and discuss the shortcomings of the Takens delay embedding theorem and discuss various issues faced while forecasting.
 
@@
 The most one can do is to place probes into a system to obtain partial information through the measurements taken.
-Moreover, the process of taking a measurement itself introduces two additional aspects which complicate the problem.
+Moreover, the process of taking a measurement itself introduces two additional aspects which complicate the problem:
 \vspace{-6mm}
@@
 
+Takens Theorem states a result establishing a relationship between the observed and underlying dynamical systems by showing that the concatenation of a sufficiently large number of previous observations into a vector will, under certain conditions, generate a map between the vectors from the respective systems.
+We formulate the theorem from \cite{takens1981detecting} after establishing preliminary definitions.
+
+\begin{Definition}
+  [\bf{Manifold, Compact Manifold}]\label{Dfn_compactManif}\rm
+  A topological space is called a manifold if every point in the space has a neighbourhood homeomorphic to $\mathbb{R}^n$. Succinctly, a manifold is a locally Euclidean topological space.
+  If a manifold is compact as a space topologically, then it is termed a compact manifold.
+\end{Definition}
+
 \begin{Definition}
@@
 
-By generic, we mean a residual set on a certain topology on appropriate spaces of functions (that we do not describe here) \ednote{B: Where can I find a citation for this?}. 
-In our scenario -- and this remains important throughout -- the input space $U$ is considered the attractor.~\label{attractor_U}
+By generic we mean a residual set on a certain topology on appropriate spaces of functions~\cite{takens1981detecting}.
+% In our scenario -- and this remains important throughout -- the input space $U$ is considered the attractor.~\label{attractor_U}
 
@@
 The double pendulum is discussed in greater depth  is provided in Chapter~\ref{ch5}.
+
 The requirement is not overly restrictive as map that is topologically transitive on a subspace must also be surjective on it; if it were to be not surjective there would be an open set $V$ that does not intersect with any future iterates of an open set  $U$ disjoint from it.

### 2022-01-19 projectReport/chapter4.tex
All G edits done

@@
 
-\ednote{Moved to after defn of USP/ESP: The reachable set itself can defined independently from $g$ possessing the ESP/USP. } 
 Note that $x\in{X_n(\overline{u})}$ if and only $g$ has a solution $\{x_k\}$ for $x_n=x$ and input $\overline{u}$, a result proved in nonautonomous dynamical systems literature~\cite{manjunath2014dynamics, manjunath2013echo}.
@@
 {\bf Proof.} May be found in~\cite{kloeden2011nonautonomous, manjunath2014dynamics, manjunath2013echo}
-% One may easily construct many systems with trivial solution-sets, such as $g(u,x)=x$ which has only the constant solution $x$ and so for $U=[-1,1]$, the system would have no solution if $|x|>1$.
-To refine the scenario, we consider only systems with unique solutions.
 
@@
 
-Secondly, it is worth noting that the reachable set~\ref{Dfn_ReachableSet} can be defined independently from $g$ possessing the USP. \ednote{hmm. does it fit here}
+% Secondly, it is worth noting that the reachable set~\ref{Dfn_ReachableSet} can be defined independently from $g$ possessing the USP. \ednote{hmm. does it fit here}
 % Paraphrased.
@@
 and we refer to the map $H$ as a \emph{causal embedding}. (Compare this with Definition~\ref{defn_causallyEmbed}.)
-%\ednote{B: edited the sentence here}
 
@@
 
-So far it is not clear whether the USP alone ensures the ability to causally embedding a dynamical system and we mention that a driving function $g$ cannot just be chosen in a frivolous manner.
-This could possibly complicate our task.
-%\ednote{B: Grammarly}
+So far it is not clear whether the USP alone ensures the ability to causally embed a dynamical system.
+We also mention that a driving function $g$ cannot just be chosen in a frivolous manner. 
+This could possibly complicate our task. %\ednote{B: Grammarly}
 
@@
 \begin{figure}[ht]
-  \includegraphics[scale=0.35]{Graphs/_Nonlinearity_Fig_response.eps}
+  \includegraphics[width=0.5\linewidth]{Graphs/_Nonlinearity_Fig_response.eps}
   \centering
@@
 \begin{figure}[ht]
-  \includegraphics[scale=0.3]{Graphs/_summarypictorial.eps}
+  \includegraphics[width=0.7\linewidth]{Graphs/_summarypictorial.eps}
   \centering
@@
 It is desirable that inputs differing slightly would have only ‘slight’ effects on the system's output. 
-Moreover, one would hope to prevent numerical errors originating from input and measurement noise (see~\cite[Theorem.
-5]{manjunath2021universal}).
+Moreover, one would hope to prevent numerical errors originating from input and measurement noise (see~\cite[Theorem. 5]{manjunath2021universal}). 
 As the data fed into the system is a function of both the input and the parameters of the system itself (i.e.
@@
 To circumvent this: if a driven system with a saturation function \footnote{A function $f(t)$ that is monotonic and approaches a constant as $|t|\to \infty$, for e.g., $\tanh(t)$ on $\mathbb{R}$} is used, we can ensure the existence of an attractor $A$ of the map that implements
-$(\pi_2,g) \circ (\Gamma,\pi_2)$ (printed in the red on Figure~\ref{Scomm_imp})
+$(\pi_2,g) \circ (\Gamma,\pi_2)$ (printed in the red on Equation~\ref{Scomm_imp})
  which in practice turns out to be defined on $Y_T^+$ has an attractor $A$  containing $Y_T$.

### 2022-01-19 projectReport/chapter5.tex
All G edits done

@@
 \section{Implementing $G_T$}
-The map $G_T$ is implemented by applying the functions $\Gamma$, $g$ and the projection function $\pi_2$ in the relevant order as described by the equations~\ref{eqns_from_data}\ednote{How do I get both?}.
+The map $G_T$ is implemented by applying the functions $\Gamma$, $g$ and the projection function $\pi_2$ in the relevant order as described by the equations~\eqref{eqns_from_data}\ednote{How do I get both?}.
 To reach this point though, one must first learn $\Gamma$.
@@
 Recall that the twin equations~\ref{eqns_from_data} generate a model from data. 
-For a chaotic system, a small amount of noise due either to input-noise or computational error would cause solutions from even a very accurate model with the same initial conditions to diverge.
-%\ednote{Grammarly}
+For a chaotic system, a small amount of noise from even a very accurate model would cause solutions with the same initial conditions to diverge.
+This noise could be due either to input-noise or computational error.
 Statistically, however,  a chaotic system will behave well in the sense that the visiting frequency (the average time spent by an orbit in any region of the phase space of the dynamical system) is identical for almost all initial conditions.
 The phrase `almost all' here is specified as a measure and refers to a result in ergodic theory (\cite{wikiErgodicTheory}) which is a topic beyond the scope of this project.
-To illustrate the accuracy of the twin  equations~\ref{eqns_from_data} in modelling the data, we plot the densities of the trajectories of the actual and predicted solutions in ~\ref{fig:dp_success_density},~\ref{fig:noisydp_success_density} and ~\ref{fig:Thomas_TrajDensity}.
+To illustrate the accuracy of the twin  equations~\ref{eqns_from_data} in modelling the data, we plot the densities of the trajectories of the actual and predicted solutions in ~\ref{fig:dp_success_density},~\ref{fig:noisydp_success_density} and~\ref{fig:Thomas_TrajDensity}.
 
@@
   \centering
-  \includegraphics[width=0.95\linewidth]{Graphs/_dp_success_4coords_traj.eps} 
+  \includegraphics[width=\linewidth]{Graphs/_dp_success_4coords_traj.eps} 
   \captionof{figure}{Predicted (red) and actual (blue) trajectories of the $x$- and $y$-coordinates of each of the pendulum heads.
@@
   \includegraphics[scale=0.5]{Graphs/_Thomas_3.eps}\caption*{Represented here the learnt (blue) and actual (red) densities of the first coordinate ($x$-coordinate) of the dataset for the Cyclically Symmetric Attractor with parameter-value $\beta=0.1056$.}
-\end{figure} \label{fig:Thomas_TrajDensity}
+\end{figure}\label{fig:Thomas_TrajDensity}
 
@@
   \centering
-  \includegraphics[width=1.0\textwidth,left]{Graphs/_Clifford_1_nonoise.eps}
+  \includegraphics[width=\textwidth,left]{Graphs/_Clifford_1_nonoise.eps}
   \caption*{These graphs were constructed by predicting the Clifford system 1000 steps into the future and in so doing illustrating the long-term consistency and accuracy of the learnt system.

### 2022-01-19 projectReport/chapter6.tex
Worked through all tex errors
All G edits done

@@
-\chapter{Conclusions}~\label{ch6}
-
 % The driven states $(x_{n-1},x_n)$ hold the information of the left-infinite history of the input.
@@
 
-There are a myriad of techniques to generate accurate models for observed dynamical systems using data obtained from measurement functions.
-Core to the success of such methods is the fact that we can often characterise the dynamics from am observed space by a mapping of much lower functional complexity when compared to the true map describing the unknown underlying system.
+There are few established techniques to generate data-driven models where data arises from a dynamical system.
+
 Takens theorem and the numerous generalizations thereof is one such method establishing (under some generic conditions) the learnability of a system produced by creating a delay-coordinate vector from a sufficiently large number of previous observations from a dynamical system.
 In this case the new system is topologically conjugate to the system from which the observed time-series was obtained.
-A critical concern of such methods is the requirement of full knowledge of the underlying state space variables governing the dynamical system one is interested in.
+
+Core to the success of data-driven models is the fact that we can often characterise the dynamics from am observed space by a mapping of much lower functional complexity when compared to the true map describing the unknown underlying system.
+
+A critical concern of Takens delay embedding is the fact that the functional complexity of the underlying system is unknown.
+The less theoretically accurate methods that approximate a differential equation to fit the data requires full knowledge of the underlying state space variables governing the dynamical system one is interested in.
 In real-world conditions this information is rarely available and, when it is, is often contingent upon human expertise and insight.
@@
 Each discrete-time state space model has at its core a driven dynamical system. 
-In this project, it has been shown that an input originating from a compact space entailing exactly one solution (i.e., if the driven system has the USP) is equivalent to representing information without distortion, a concept which was formalised by establishing the existence of a causal mapping’s.
-The established results are general, implying that any driven dynamical system with specific properties, such SI-invertibility, can give rise to a causal mapping. 
-In particular, we have a topological conjugacy (semi-conjugacy) between the data comprising the single-delay lag dynamics of a driven system~\cite{manjunath2013echo} and the underlying homeomorphic dynamical system.
-Finite-length input data is shown to be sufficient for forecasting as the left-infinite segment of the past of the input beyond some arbitrarily large but finite time (a result which holds due to the universal semi-conjugacy of a driven system’s continuity) influences the driven system's states in an almost negligibly small manner.
+In this project, it has been observed that an input originating from a compact space entailing exactly one solution (i.e., if the driven system has the USP) is equivalent to representing information without distortion, a concept which was formalised by establishing the existence of a causal mapping.
+The established results are general, implying that any driven dynamical system with specific properties, such as SI-invertibility, can give rise to a causal embedding. 
+In particular, we have a topological conjugacy between the inverse-limit space of the underlying dynamical system and the single-delay lag dynamics of a driven system~\cite{manjunath2013echo}.
+Finite-length input data is shown to be sufficient for forecasting as the left-infinite segment of the past of the input beyond some arbitrarily large but finite time (a result which holds due to the continuity of the universal semi-conjugacy) influences the driven system's states in an almost negligibly small manner.
 
 Viewed through the lens of reservoir computing, we have also demonstrated that when single-delay dynamics are utilised, a learnable map in a reservoir computing network certainly exists. 
-Moreover, within a reservoir computing network, such a driven system posssesses certain proven qualities pertaining to the robustness of the system to different forms of perturbation. 
-We are thus able to preserve the quality of a representation that would otherwise have remained sensitive to various fluctuations and rendering the learning of the system’s dynamics wholly unpredictable.
+Moreover, within a reservoir computing network, such a driven system possesses certain proven qualities pertaining to the robustness of the system to different forms of perturbation. 
+We are thus able to preserve the quality of a representation that would otherwise have remained sensitive to various fluctuations and rendering the learning of the system's dynamics wholly unpredictable.
 
 Finally, we have considered some physical systems to observe the methodology in practice and verify that we can indeed predict the future evolution of chaotic systems such as the double pendulum.
-We demonstrated the robustness of our methodology to external noise by considering the cases of the  Dream Attractor and the Double Pendulum.
-
-\ednote{Loose paragraph commented out}
+We demonstrated the robustness of our methodology to external noise by considering the cases of the  Dream Attractor and the Thomas' attractor.
+\ednote{A sufficient ending?}
+%\ednote{Loose paragraph commented out}
 % Considered from a philosophical-scientific  perspective, this project has aimed to establish a juncture between rationalism (the theory of nonautonomous dynamical systems) and empiricism (machine learning) to yield results profitable to the discipline of Applied Mathematics.

### 2022-01-26 projectReport/abstract.tex
Found too many small errors

@@
 \begin{abstract}
-    It is a well-established practice to collect measurements from underlying artificial, natural, and physical systems through observables in order to realise a dynamical model with greater descriptive power.
+    It is a well-established practice to collect measurements from underlying artificial, natural, and physical systems through observables to realise a dynamical model with greater descriptive power.
     Many approaches to forecasting exist but they necessitate an understanding of systems we often don't know much about.

### 2022-01-26 projectReport/chapter1.tex
Rereading the project. On 5.4
All G edits done
Found too many small errors

@@
 Even the presence of usually `negligible' computational noise or measurement error renders long-term pointwise prediction infeasible.
-We encounter many difficultiesin the process of modelling such systems:
+We encounter many difficulties in the process of modelling such systems:
 
@@
 
-Models derived primarily from data can be classified into three categories: 
+Models derived primarily from data can be classified into three categories.
 \vspace{-8mm}
@@
   \item Partially interpretable models capturing some modes of the dynamics, 
-  \item Non-interpretable models, defined as such mainly due to the fact that they are defined on a different phase space that is usually high-dimensional. 
+  \item Non-interpretable models, defined as such mainly because they are defined on a different phase space that is usually high-dimensional. 
 \end{enumerate}
@@
 This then confirms the existence of a map on the space of delay coordinates equivalent (or topologically conjugate) to the underlying map from which the observed time-series was first obtained.
-Although topological conjugacy guarantees an alternate representation of the underlying system, the quality still depends on numerous parameters, making the understanding of the dynamics unreliable (see \cite{principe1992prediction}).
+Although topological conjugacy guarantees an alternate representation of the underlying system, the quality still depends on numerous parameters, making the understanding of the dynamics unreliable (see~\cite{principe1992prediction}).
 One reason for this fragility is that the embedded attractor in the reconstruction space is not always an attractor of the map learnt in the reconstruction space, despite unquestionably being an invariant set.

### 2022-01-26 projectReport/chapter2.tex
Found too many small errors

@@
 A continuous-time dynamical system describes the states for all values of the time.
-Specifically, if the motion of a pendulum in which the quantities such as the angular position and angular momentum are known at all times, then it is a continuous-time dynamical system.
+Specifically, if the motion of a pendulum in which the quantities such as the angular position and angular momentum are always known, then it is a continuous-time dynamical system.
 The equations of the dynamical system can take the form of one or more ordinary differential equations that determine the relevant quantities at any future time if we know the initial location and momentum.
@@
   [\bf Orbit of $T$]\label{Dfn_Orbit}\rm
-  The orbit of $T$ is defined  is to the sequence $\bar{u} = {\{u_n\}}_{n\in \mathbb{Z}}$ obeying the update equation, $u_{n+1}=Tu_n$, $n \in \mathbb{Z}$. 
+  The orbit of $T$ is defined to be the sequence $\bar{u} = {\{u_n\}}_{n\in \mathbb{Z}}$ obeying the update equation, $u_{n+1}=Tu_n$, $n \in \mathbb{Z}$. 
 \end{Definition}
@@
 Two examples of invariant sets include a fixed point where $Tu=u$ for  $u\in U$, and a periodic orbit, i.e., a set of iterates $\{u,Tu, T^2u,\ldots,T^{pu}\}$, where $T^{p+1}u=u$ for some $p\in\mathbb{Z}$.
-The entire space $U$ could be also be invariant.
+The entire space $U$ could also be invariant.
 Consider for example the space $U=[0,1]$, where $Tu=4u(1-u)$ and then $U$ is invariant as every $u\in{U}$ can be written as $u = 4x(1-x)$ for some $x\in{U}$.
@@
 If $U=[0,1]$ has map $Tu= u/2$, then the only invariant set is $\{0\}$, and all orbits approach this invariant set as time flows in the forward direction. 
-Indeed, if a some non-zero invariant set (call it $B$) exists, then there is  some $r\in(0,1]\cap{B}$.
+Indeed, if some non-zero invariant set (call it $B$) exists, then there is  some $r\in(0,1]\cap{B}$.
 But $r\notin{T(B)}$ since any orbit with initial value $r$ will be a decreasing sequence.
@@
 
-One may ask if every orbit approaches an invariant set?
+One may ask if every orbit approaches an invariant set.
 In general the answer is no, since for the dynamical system $T: \mathbb{R} \to \mathbb{R}$ defined by $Tu=2u$, any orbit that does not intersect the invariant set $\{0\}$, will not approach any invariant set.
@@
 \begin{Example}\label{ex_Tu2}\rm
-  For the map $Tu=u^2$ defined on $[0,1]$ all orbits lie in the invariant set $\{1\}$ or else would would approach the invariant set $\{0\}$. 
+  For the map $Tu=u^2$ defined on $[0,1]$ all orbits lie in the invariant set $\{1\}$ or else would approach the invariant set $\{0\}$. 
 \end{Example}
@@
 We are interested in attractive invariant sets since they capture the long-term dynamics as time increases. 
-In particular, we are interested in those invariant sets named attractors.
+We are interested in those invariant sets named attractors.
 
@@
 
-Indeed in our previous example~\ref{ex_halfu, ex_Tu2}, it can be verified that the system  $Tu=u/2$ defined on  $\mathbb{R}$ and $Tu=u^2$ defined on $[0,1]$  the singleton set $\{0\}$ is an attractor.
+Indeed, in our previous example~\ref{ex_halfu},~\ref{ex_Tu2}, it can be verified that the system  $Tu=u/2$ defined on  $\mathbb{R}$ and $Tu=u^2$ defined on $[0,1]$  the singleton set $\{0\}$ is an attractor.
 For the system,  $Tu=1-|2u-1|$ on $[0,1]$, one may easily verify that the only attractor is the entire space $[0,1]$.
@@
 The dynamics restricted to an invariant set can be complicated.
-For instance an invariant set could be just a single point or it could have an infinite set.
+For instance, an invariant set could be just a single point or it could have an infinite set.
 If the invariant set is infinite, then complicated dynamics are possible.
-A particular, well-studied phenomenon of such complexity gives rise to so-called chaotic behaviour, a subject studied in detail over the past fifty years.
+A well-studied phenomenon of such complexity gives rise to so-called chaotic behaviour, a subject studied in detail over the past fifty years.
 
@@
   [\bf {(Sensitive Dependence on Initial Conditions)}]\label{Dfn_SDIC}\rm
-A dynamical system $T: U \to U$ is said to have \textit{sensitive dependence on initial conditions (SDIC)} if there exists a $\delta > 0$ such that for every $u \in U$ and in every neighborhood of $u \in U$ there exists a $v\in{U}$ and an integer $N:=N{(u,v)}\in\mathbb{Z}$ such that $d(T^Nu,T^Nv)>\delta$.
+A dynamical system $T: U \to U$ is said to have \textit{sensitive dependence on initial conditions (SDIC)} if there exists a $\delta > 0$ such that for every $u \in U$ and in every neighbourhood of $u \in U$ there exists a $v\in{U}$ and an integer $N:=N{(u,v)}\in\mathbb{Z}$ such that $d(T^Nu,T^Nv)>\delta$.
 \end{Definition}
@@
   \item The $\delta>0$ in the definition of SDIC is independent of $u$. 
-  \item In every neighborhood of $u$, we may not necessarily find all points $v$ in the neighborhood distinct from $u$ that would separate from the forward iterates of $u$. 
+  \item In every neighbourhood of $u$, we may not necessarily find all points $v$ in the neighbourhood distinct from $u$ that would separate from the forward iterates of $u$. 
   \item  $N$ depends upon $u$ and $v$ chosen, and their iterates may not separate forever (i.e., for all $n>N$) and we allow their iterates to get arbitrarily close in the future. 
@@
   In the plot Fig.~\ref{fig:log_sdic} below, we plot the first 50 iterates of the LM for $r=4$ and two slightly different initial values $x_0$. 
-  The values of $x_n$ are plotted agains time $n$  where they have initial values $x_0=0.2$ (blue) and $x_0=0.20001$ (red).
+  The values of $x_n$ are plotted against time $n$  where they have initial values $x_0=0.2$ (blue) and $x_0=0.20001$ (red).
 Initially the two trajectories overlap, but they diverge completely at $n=12$ whereafter they follow distinct paths.
@@
   [\bf {Topologically Transitive}]\label{Dfn_TopolTrans}\rm
-	A dynamical system $F: U \to U$ is topologically transitive if for any pair of nonempty open sets $E_1$ and $E_2$ there exists a $n\in\mathbb{N}$ such that $$T^n(E_1) \cap E_2 \not= \emptyset$$. 
+	A dynamical system $F: U \to U$ is topologically transitive if for any pair of nonempty open sets $E_1$ and $E_2$ there exists a $n\in\mathbb{N}$ such that \[T^n(E_1) \cap E_2 \not= \emptyset.\]
 \end{Definition}
@@
 \begin{Definition}
-  [\bf {Meager Set}]\label{Dfn_Meager Set}\rm
-A subset of a topological space $U$ is said to be a meager set if it can be written as a countable union of sets of with empty interior.
-The complement of the a meager set is said to be a residual set.
+  [\bf {Meagre Set}]\label{Dfn_Meagre Set}\rm
+A subset of a topological space $U$ is said to be a meagre set if it can be written as a countable union of sets of with empty interior. The complement of the meagre set is said to be a residual set.
 \end{Definition}
@@
     The graph of the map $T$ is piecewise linear with two straight lines, one connecting the points $(0,0)$ and $(\frac{1}{2},1)$ and the other connecting $(\frac{1}{2},1)$ with $(1,0)$.
-They form a so-called tent with base centered at $u=1/2$.
-The graph of the map $T^2$ comprises two symmetric tents with their base centered at $1/4$ and $3/4$.
+They form a so-called tent with base centred at $u=1/2$.
+The graph of the map $T^2$ comprises two symmetric tents with their base centred at $1/4$ and $3/4$.
 See Figure~\ref{fig:T2tentmap}
@@
     In general, the graph of  $T^k$ contains $2^{k-1}$ tents.
-    Given any point $u$ and an open neighborhood $W \subset[0,1]$ of $u$, we can find $k\in\mathbb{Z}$ large enough so that $T^k(W) = [0,1]$ because we can accommodate a tent whose base is contained in $W$.
+    Given any point $u$ and an open neighbourhood $W \subset[0,1]$ of $u$, we can find $k\in\mathbb{Z}$ large enough so that $T^k(W) = [0,1]$ because we can accommodate a tent whose base is contained in $W$.
     This implies the existence of two points $w_1$ and $w_2$ in $W$ so that $d(T^kw_1,T^kw_2)= 1$, and by the triangle inequality  at least one of the inequalities $d(T^ku,T^kw_1)> \delta$ or $d(T^ku,T^kw_2)> \delta$ hold when $\delta\in (0,\frac{1}{2})$.
@@
 When $S$ is semi-conjugate to $T$, it is common to call $S$ a \emph{factor of $T$}, or conversely that $T$ is an \emph{extension of $S$}.
-In essence, an extension (see,~\cite{de2013elements}) is a more extensive system capturing all of the essential dynamics of its factor.
+In essence, an extension (see,~\cite{de2013elements}) is a more extensive system capturing all the essential dynamics of its factor.
 

### 2022-01-26 projectReport/chapter3.tex
Found too many small errors

@@
 
-Takens embedding theorem does not take into account the noise, and to describe it we construct a multidimensional observable using the method of stacking previous observations.
+Takens embedding theorem does not consider the noise, and to describe it we construct a multidimensional observable using the method of stacking previous observations.
 The concept of an observable is understood in the sense of an observable as originally introduced by~\cite{takens1981detecting, genericObservableAeyels} to refer to a probe or measurement function inserted into the system.
@@
 Takens Theorem states a result establishing a relationship between the observed and underlying dynamical systems by showing that the concatenation of a sufficiently large number of previous observations into a vector will, under certain conditions, generate a map between the vectors from the respective systems.
-We formulate the theorem from \cite{takens1981detecting} after establishing preliminary definitions.
+We formulate the theorem from~\cite{takens1981detecting} after establishing preliminary definitions.
 
@@
 \end{Definition}
-
-Takens Theorem states a result establishing a relationship between the observed and underlying dynamical systems by showing that the concatenation of a sufficiently large number of previous observations into a vector will, under certain conditions, generate a map between the vectors from the respective systems.
-We formulate the theorem from~\cite{takens1981detecting}.
 
@@
 % The Takens embedding theorem does not guarantee global dissipativity.
-% \ednote{B:Links with~\ref{subs_LearnGamma} in discussion of dissipativity of $\Gamma$.
-Sir, I'm not sure how to write this here - I think it a beneficial point to make so as to set the stage for the later discussion in ~\ref{subs_LearnGamma}, but I am unsure there as well.}
 

### 2022-01-26 projectReport/chapter4.tex
Found too many small errors

@@
   A nonautonomous dynamical system (NDS) on a space $X$ is simply a dynamical system comprising a family of maps ${\{g_n\}}_{n \in \mathbb{Z}}$ where $g_n(\ldotp):=g(u_n,\ldotp):X\to{X}$ is continuous.
-Each $g_n$ arises as a consequence of an input $u_n$ from the input space $U$, a topological space.
+Each $g_n$ arises because of an input $u_n$ from the input space $U$, a topological space.
 \end{Definition}
@@
   [\bf Driven and Compactly Driven Dynamical System]\label{Dfn_DDS} \rm
-A driven dynamical system comprises two topological metric spaces $U$ and $X$, as well as  a continuous function  ${g:U\times{X}\to{X}}$ where ${g(u_n, x_n)=x_{n+1}}$.
+A driven dynamical system comprises two topological metric spaces $U$ and $X$, as well as  a continuous function  ${g:U\times{X}\to{X}}$ where ${g(u_{n-1}, x_{n-1})=x_{n}}$.
 If the input space $U$ is compact, we refer to the system as compactly driven. 
@@
   The driven system may be rewritten as $x_{n-1}=\frac{2x_n}{u_{n-1}}=4x_n$ and the  iterates of $x_n$ in backward time will increase by a factor of 4 at each timestep. 
-  Thus for some $m\leq{n}$,  $1<x_m$ i.e. $x_m\notin{X}$. So ${\{x_n\}}_{n\in\mathbb{Z}}$ is not a solution and it follows that the only possible solution is the zero solution.
+  Thus for some $m\leq{n}$,  $1<x_m$ i.e., $x_m\notin{X}$. So ${\{x_n\}}_{n\in\mathbb{Z}}$ is not a solution, and it follows that the only possible solution is the zero solution.
 \end{Example}
-% \ednote{Nicely written! Thank you}
 
@@
   $g(u,x)=x^2$ the evolution does not depend on the input at all. Hence $g(u,x)$ can be written as a single map $f(x)=x^2$ which is a homeomorphism on $[0,1]$. 
-  A left-infinite orbit is defined and it converges to $1$ while the right-infinite orbit converges to $0$. 
-  In other words the iterates of $f^{-1}$ converges to $1$ while the iterates of $f$ converge to $0$ 
+  A left-infinite orbit is defined, and it converges to $1$ while the right-infinite orbit converges to $0$. 
+  In other words, the iterates of $f^{-1}$ converges to $1$ while the iterates of $f$ converge to $0$ 
 \end{Remark}
@@
 
-The USP is equivalently called  the echo state property (ESP)~\cite{Manju_ESP}, and systems that possess the ESP have other important and useful properties for information processing~\cite{grigoryeva2018echo, grigoryeva2019differentiable, manjunath2020stability} .
+The USP is equivalently called  the echo state property (ESP)~\cite{Manju_ESP}, and systems that possess the ESP have other important and useful properties for information processing~\cite{grigoryeva2018echo, grigoryeva2019differentiable, manjunath2020stability}.
 
@@
 
-In the majority of Reservoir Computing (RC) literature -- a machine learning approach -- where the input is also mapped onto a different, but higher dimensional, space through a system called the reservoir.
+In most of the Reservoir Computing (RC) literature -- a machine learning approach -- where the input is also mapped onto a different, but higher dimensional, space through a system called the reservoir.
 A readout that measures the state of the reservoir is trained, a notion of forgetting the states of the reservoir asymptotically is often used.
-\ednote{this sentence falls out of the blue}
+%\ednote{this sentence falls out of the blue}
 
@@
 If $g$ possesses the ESP (equivalent in our context to the USP), then we are guaranteed that the whole of an input's left-infinite history will determine exactly the current system state; i.e.
-there is only ever the possibility of a single reachable state at a given instance of time~\cite{jaeger2001echo,Manju_2020}.
+there is only ever the possibility of a single reachable state at a given instance of time~\cite{jaeger2001echo,manjunath2020stability}.
 % Defined formally, the ESP is intrinsically linked to an input (see \cite{Manju_ESP}) Unfortunately, definitions and results concerning the ESP do not have much practical benefit since a relationship between the network's response and the temporo-statistical properties of the input are not directly described.
@@
 
-Having already defined the reachable set $X_U$ as the collection of all elements of all entire solutions, we pause briefly in order to fix additional notation.
+Having already defined the reachable set $X_U$ as the collection of all elements of all entire solutions, we pause briefly to fix additional notation.
 Defining $\cev{u}^{n}:=(\ldots,u_{n-2} ,u_{n-1})$ as the left-infinite subsequence of an input up until time $n$, $\overleftarrow{U}$ then denotes all such left-infinite sequences in $U$. 
@@
 The introduction of a new input at time $n$ can be described by the mapping $\sigma_v:   \cev{u}^{n} \mapsto \cev{u}^{n}v$. 
-The right-shift map, $r\overleftarrow{U}\to\overleftarrow{U}$, of an input sequence is defined $r: (\cdots, u_{-2},u_{-1}) \mapsto(\cdots, u_{-3},u_{-2})$
+The right-shift map, $r:\overleftarrow{U}\to\overleftarrow{U}$, of an input sequence is defined $r: (\cdots, u_{-2},u_{-1}) \mapsto(\cdots, u_{-3},u_{-2})$
 
@@
 
-  If the universal semi-conjugacy $h$ exists (i.e. the diagram in~\ref{Scomm_h} commutes) then the solution $\Psi(\bar{u})$ will intuitively have no more ``complexity' than the input $\bar{u}$.
+  If the universal semi-conjugacy $h$ exists (i.e., the diagram in~\ref{Scomm_h} commutes) then the solution $\Psi(\bar{u})$ will intuitively have no more ``complexity' than the input $\bar{u}$.
 
@@
 The continuity of $H^{-1}$ follows from the fact that $H$ is itself a continuous and surjective mapping of a compact space $\overleftarrow{U}$ in a Hausdorff space.
-and we refer to the map $H$ as a \emph{causal embedding}. (Compare this with Definition~\ref{defn_causallyEmbed}.)
+We refer to the map $H$ as a \emph{causal embedding}. (Compare this with Definition~\ref{defn_causallyEmbed}.)
 
@@
 
-The function $H$ contains a countably infinite number of coordinate (or \textit{factor maps}) since its codomain is $\overleftarrow{X}_U$.  
+The function $H$ contains a countably infinite number of coordinate (or \textit{factor}) maps since its codomain is $\overleftarrow{X}_U$.  
 It is not practically feasible to consider a left-infinite sequence in $X_U$ (in other words an element of  $\overleftarrow{X}_U$) for learning.
@@
 The left-infinite orbit space is significant because a great amount of information about the original system can be gleaned from its topological structure~\cite{ingram2011inverse,Manju_IEEE}. 
-The formal topological structure for the inverse-limit space of a chaotic dynamical system is complicated, and further discussion is relegated to~\cite{kennedy2008inverse_limit, ingram2011inverse}.
+The formal topological structure for the inverse-limit space of a chaotic dynamical system is complicated, and further discussion is relegated to~\cite{kennedy2008inverse, ingram2011inverse}.
 Broadly, we may conceive of the inverse-limit system of a dynamical system as a subspace of an infinite-dimensional space where each point in the inverse-limit space corresponds to a backward orbit of the map $T$.
@@
 Consider the Figure~\ref{fig:memloss_conttime} where a RNN's solution was numerically simulated for randomly generated matrices $A$ and $B$ where $B$ has unit spectral radius.
-An input sequence in black $(x_0,x_1,\ldots,x_{5000})$ was fed to the network and it, as well as the the coordinates of two simulated solutions in red (with parameter $a=1$, $\alpha=0.99$ where $g$ has the USP) and blue ($a=1$, $\alpha=1.05$ where $g$ does not have the USP), were plotted.
-The red solution mimicked the input sequence's temporal variation quite well, whereas the blue solution is observed to be disordered and oft-changing.
+An input sequence in black $(x_0,x_1,\ldots,x_{5000})$ was fed to the network and it, as well as the coordinates of two simulated solutions in red (with parameter $a=1$, $\alpha=0.99$ where $g$ has the USP) and blue ($a=1$, $\alpha=1.05$ where $g$ does not have the USP), were plotted.
+The red solution mimicked the input sequence's temporal variation quite well, whereas the blue solution is observed to be disordered and oft changing.
 Through this we may state that the blue solution has introduced additional complexity to the system.
@@
   For an SI-invertible driven system $g$ and a dynamical system $(U,T)$, the map $G_T: Y_T \to Y_T$ defined by the relation $(x_{n-1},x_n) \mapsto (x_n,x_{n+1})$ is well-defined. 
-  (This results holds even in the absence of $g$ possessing the USP). 
+  (This result holds even in the absence of $g$ possessing the USP). 
   % Moreover, the mapping $(x_{n-1},x_n) \mapsto u_{n}$ is well-defined when $x_{n-1}$ and $x_n$ are successive points on a solution obtained for an input  $\{u_n\}$ that is an orbit of $T$.
@@
 
-At this stage it is worth taking note of a specific driven system in the form of a discrete-time state-space model which has acquired some adherence in applications~\cite{Manju_IEEE} -- especially those pertaining to echo state betworks and the ESP.
+At this stage it is worth taking note of a specific driven system in the form of a discrete-time state-space model which has acquired some adherence in applications~\cite{Manju_IEEE} -- especially those pertaining to echo state networks and the ESP.
 The function
@@
 \end{equation} 
-is SI-invertible and, if $\alpha B$ has a spectral norm $<1$, also possesses the USP~\cite[Theorem.2]{manjunath2013echo }. 
+is SI-invertible and, if $\alpha B$ has a spectral norm $<1$, also possesses the USP~\cite[Theorem.2]{manjunath2013echo}. 
 It is easy to show that $g$ is SI-invertible by recovering $u_n$ in 
@@
 % One need therefore not establish any additional results to ensure that the sequence uniformly approaches the unique solution $\Psi$.
-This vastly simplifies the effort necessary to set up a problem in order to guarantee that the underlying system will be accurately represented by the conjugate system.
+This vastly simplifies the effort necessary to set up a problem to guarantee that the underlying system will be accurately represented by the conjugate system.
 
@@
 %Incredibly, this permits one to initialise a driven system $g$ with an altogether arbitrary initial value $y_m\in{X}$ where $m\in\mathbb{Z}$ and the UAP then guarantees that the sequence $\{y_{m+1}, y_{m+2},\ldots\}$ which satisfies the relation $y_n=g(u_n, y_{n-1})$  for  $k\geq{m}$  will approach the elements $\{x_n\}$ of the actual solution.
-This vastly simplifies the effort necessary to set up a problem in order to guarantee that the underlying system will be accurately represented by the conjugate system.
+This vastly simplifies the effort necessary to set up a problem to guarantee that the underlying system will be accurately represented by the conjugate system.
 
 We are interested in determining whether $\widehat{T}$ and $G_T$ are related.
-For this, we introduce  the function ${H_2{\overline{}} := (h(r\overline{u}, h(\overline{u}))}$ which maps the entirety of a left-infinite sequence in $\Psi$ to some element in $X\times{X}$.
+For this, we introduce  the function ${H_2{\overline{}} := (h(r\overline{u}, h(\overline{u})))}$ which maps the entirety of a left-infinite sequence in $\Psi$ to some element in $X\times{X}$.
 Indeed, it may be shown that $g$ having SI-invertibility and the USP immediately guarantees the existence of at least a semi-conjugacy between the systems $(Y_T, G_T)$ and $(\widehat{U}_T, \widehat{T})$.
@@
 %\begin{Theorem}
- %   $(Y_T, G_T)$ is semi-conjugate to $(\widehat{U}, \widehat{T})$.
+ %   $(Y_T, G_T)$ is semi-conjugate to $(\widehat{U}_T, \widehat{T})$.
 %\end{Theorem}
@@
 \item We are interested in a some dynamical system $(U,T)$ with unknown dynamics.
-\item To determine information about this system $(U,T)$ and predict its future evolution, we determine the dynamics of the inverse-limit system $(\widehat{U}, \widehat{T})$.
-% Given certain assumptions, we can guarantee that $(\widehat{U}, \widehat{T})$ is at least semi-conjugate to $(U,T)$.
+\item To determine information about this system $(U,T)$ and predict its future evolution, we determine the dynamics of the inverse-limit system $(\widehat{U}_T, \widehat{T})$.
+% Given certain assumptions, we can guarantee that $(\widehat{U}_T, \widehat{T})$ is at least semi-conjugate to $(U,T)$.
 \item If the driven system $g$ is SI-invertible (and $\{u_n\}\in{U}$  is an orbit of $T$), the map $G_T$ exists that describes the single delay dynamics. 
-\item If, furthermore, $g$ has the USP, $(Y_T, G_T)$ is semi-conjugate to $(\widehat{U}, \widehat{T})$.
-\item If we can assume that $T$ is a homeomorphism, $(Y_T, G_T)$ is topologically conjugate to $(\widehat{U}, \widehat{T})$, an extension space of $(U,T)$
+\item If, furthermore, $g$ has the USP, $(Y_T, G_T)$ is semi-conjugate to $(\widehat{U}_T, \widehat{T})$.
+\item If we can assume that $T$ is a homeomorphism, $(Y_T, G_T)$ is topologically conjugate to $(\widehat{U}_T, \widehat{T})$, an extension space of $(U,T)$
 \end{enumerate}
@@
 In the above sections, we established the existence of a $G_T$ describing the SDD of a driven system.
-By establishing its existence, we’ve essentially embedded the attractor $U$ (recall from~\ref{attractor_U}) into the higher dimensional space $X\times{X}$.
+By establishing its existence, we’ve essentially embedded the attractor $U$ into the higher dimensional space $X\times{X}$.
 In layman’s terms, this ensures that there is more `dimensional room' for the underlying system’s underlying dynamics to `move'.
@@
 	u_{k+1} &=& \pi_1 \circ (\Gamma, \pi_2) \circ (\pi_2,g) (u_k,x_k) \label{Seqn_u}\\
-	x_{k+1} &=& \pi_2 \circ (\Gamma, \pi_2) \circ (\pi_2,g) (u_k,x_k). \label{Seqn_x}
+	x_{k+1} &=& \pi_2 \circ (\Gamma, \pi_2) \circ (\pi_2,g) (u_k,x_k). \nonumber
+	\label{Seqn_x}
 \end{eqnarray}
@@
 The decision does seem arbitrary on the surface, so we take a pause again and discuss the motivation for learning Gamma.
-There are a number of distinct advantages.
+There are several distinct advantages.
 
-Firstly, learning $\Gamma$ saves computational resources. This is due to the fact that the input $u_n$ lies in a lower-dimensional subspace in comparison to the high-dimensional space $X\times{X}$.
+Firstly, learning $\Gamma$ saves computational resources. This is since the input $u_n$ lies in a lower-dimensional subspace in comparison to the high-dimensional space $X\times{X}$.
 
@@
 Secondly, the function $\Gamma$ is known to be stable.
-Learning $\Gamma$ makes use of equations~\ref{eqns_from_data}\ednote{how can I reference both 4.13 and 4.14?} which in turn employ the driven system $g$ possessing the USP.
+Learning $\Gamma$ makes use of equations~\ref{eqns_from_data} which in turn employ the driven system $g$ possessing the USP.
 This offers distinct advantages with regards to stability in the presence of perturbation.
@@
 Moreover, one would hope to prevent numerical errors originating from input and measurement noise (see~\cite[Theorem. 5]{manjunath2021universal}). 
-As the data fed into the system is a function of both the input and the parameters of the system itself (i.e.
-the value a, $\alpha$ in $\tanh$), we exploit both input-related and parameter-related stability.
+As the data fed into the system is a function of both the input and the parameters of the system itself (i.e., the value a, $\alpha$ in $\tanh$), we exploit both input-related and parameter-related stability.
 Both these notions are defined by way of the continuity of an encoding map of the respective input and design parameters~\cite{manjunath2020stability}.
 
-Informally, we may consider input-related stability relating to the question of whether or not small variations in input would result in small responses.   
+Informally, we may consider input-related stability relating to the question of whether small variations in input would result in small responses.   
 If two inputs $\bar{u}$ and $\bar{u}_{noisy}$ are close in the product topology, their tails could differ greatly due to the metric that generates the product topology insensitive to the differences in sequence's tails.
- However, if the system has the USP, then the function $h$ is continuous and so $h(\bar{u})$, $h(\bar{u}_{noisy})$ remain close-by for low levels of noise; this maintains a measure of robustness to input and measurement noise.\cite{manjunath2021universal}.
+ However, if the system has the USP, then the function $h$ is continuous and so $h(\bar{u})$, $h(\bar{u}_{noisy})$ remain close-by for low levels of noise; this maintains a measure of robustness to input and measurement noise~\cite{manjunath2021universal}.
 In a similar fashion, parameter-related stability is obtained thanks to a result~\cite[Lemma 3.2]{manjunath2020stability} relating this form of stability with the ESP (which is here equivalent to the USP).
@@
  which in practice turns out to be defined on $Y_T^+$ has an attractor $A$  containing $Y_T$.
-By restricting the dynamics to not spill beyond $A$, prevents large errors from occuring.  
+By restricting the dynamics to not spill beyond $A$, prevents large errors from occurring.  
 More details on the global dissipativity of $g$, as it is termed here, may be found in~\cite{Supp}.

### 2022-01-26 projectReport/chapter5.tex
Found too many small errors

@@
 Most notably, we consider the reconstruction of the chaotic attractor of the double pendulum -- an  attractor that has not previously been successfully reconstructed because of its intricate structure.
-In fact the previous data-driven approaches also have not been successful in building a model from the double pendulum.
+In fact, the previous data-driven approaches also have not been successful in building a model from the double pendulum.
 
@@
 \section{Implementing $G_T$}
-The map $G_T$ is implemented by applying the functions $\Gamma$, $g$ and the projection function $\pi_2$ in the relevant order as described by the equations~\eqref{eqns_from_data}\ednote{How do I get both?}.
+The map $G_T$ is implemented by applying the functions $\Gamma$, $g$ and the projection function $\pi_2$ in the relevant order as described by the equations~\eqref{eqns_from_data}.
 To reach this point though, one must first learn $\Gamma$.
@@
 
-We implement a feedforward network (a network with no memory, unlike the RNN) to learn $\Gamma$ in the \emph{Python} programming language and making use of the \emph{Scipy} library, alongside the \emph{Keras} package in the \emph{Tensorflow} library .
+We implement a feedforward network (a network with no memory, unlike the RNN) to learn $\Gamma$ in the \emph{Python} programming language and making use of the \emph{Scipy} library, alongside the \emph{Keras} package in the \emph{Tensorflow} library.
 
@@
 
-Once $\Gamma$ is learnt, we progress to the the prediction step where $u_n$ is forecasted from $(x_{n-1},x_{n})$.
+Once $\Gamma$ is learnt, we progress to the prediction step where $u_n$ is forecasted from $(x_{n-1},x_{n})$.
 Note that  due to causal embedding $(x_{n-1},x_{n})$ is in one-to-one correspondence with the left-infinite orbit $(\ldots,u_{n-1},u_{n-1})$.
@@
 Statistically, however,  a chaotic system will behave well in the sense that the visiting frequency (the average time spent by an orbit in any region of the phase space of the dynamical system) is identical for almost all initial conditions.
-The phrase `almost all' here is specified as a measure and refers to a result in ergodic theory (\cite{wikiErgodicTheory}) which is a topic beyond the scope of this project.
-To illustrate the accuracy of the twin  equations~\ref{eqns_from_data} in modelling the data, we plot the densities of the trajectories of the actual and predicted solutions in ~\ref{fig:dp_success_density},~\ref{fig:noisydp_success_density} and~\ref{fig:Thomas_TrajDensity}.
+The phrase `almost all' here is specified as a measure and refers to a result in ergodic theory (\cite{walters2000introduction}) which is a topic beyond the scope of this project.
+To illustrate the accuracy of the twin  equations~\ref{eqns_from_data} in modelling the data, we plot the densities of the trajectories of the actual and predicted solutions in~\ref{fig:dp_success_density},~\ref{fig:noisydp_success_density} and~\ref{fig:Thomas_TrajDensity}.
 
@@
 In our experiments, we generally opted for a density ratio of 10\%-25\%.
-Secondly, we opt to enforce the requirement that both matrices be well-conditioned, i.e.
-that they have a condition number of 1.
+Secondly, we opt to enforce the requirement that both matrices be well-conditioned, i.e.,that they have a condition number of 1.
 The condition number of a matrix $M$ is the product of the matrix norms of the matrix $M$ and its inverse $M^{-1}$ and is a measure of the sensitivity of output of a matrix multiplication when an input is changed.
@@
 
-Lastly, we ensure the matrix $B$ has unit spectral radius so that the variable $\alpha$ in the RNN equation \eqref{eqn_driving} can be used to know if the USP is satisfied. 
+Lastly, we ensure the matrix $B$ has unit spectral radius so that the variable $\alpha$ in the RNN equation~\eqref{eqn_driving} can be used to know if the USP is satisfied. 
 It is a  feature that a matrix with a spectral radius less than 1 generally has the USP~\cite{yildiz2012re}.
@@
 More explicitly, denote $X_{1:N}$ as the matrix with the first $N$ states of the network data represented by row vectors. 
-$X_{1:N}=U\Sigma P^T$ here represents the singular value decomposition of $X_{1:N}$ ($U$ and $V$ are unitary matrices, and $\Sigma$ is a diagonal matrix) and we refer to \cite{wikiSVdecomp}. 
+$X_{1:N}=U\Sigma P^T$ here represents the singular value decomposition of $X_{1:N}$ ($U$ and $V$ are unitary matrices, and $\Sigma$ is a diagonal matrix) and we refer to~\cite{walters2000introduction}.
 We then denote the principal component matrix by the symbol $P$, and the principal components are given by 
@@
 This system shows chaos in a striking manner -- consider once more Figure~\ref{fig:dp_sdic}.
-The system has four variables: for each pendulum rod an associated angle and angular velocity. An ideal system is assumed where the rods are massless and pivots are frictionless.
+The system has four variables: for each pendulum rod an associated angle and angular velocity. An ideal system is assumed where the rods are massless, and pivots are frictionless.
 \begin{figure}[ht]
-  \includegraphics[width=0.8\linewidth]{Graphs/_dp_setup.eps}
+  \includegraphics[width=0.4\linewidth]{Graphs/_dp_setup.eps}
   \centering
@@
 Here we chose the values $m_1=2.0$, $m_2=2.0$, $L_1=1.5$ and $m_2=1.5$ with initial velocities $v_1=1.5, v_2=-2$ for the respective pendulum heads.
-With a change of variables the second-order system can be written as a system of first order equations with four variables, where each pendulum has an associated $x$ and $y$ coordinates of the mass (the lengthy equations are not shown here).
+With a change of variables, the second-order system can be written as a system of first order equations with four variables, where each pendulum has an associated $x$ and $y$ coordinates of the mass (the lengthy equations are not shown here).
 The ODE system is then solved using the fourth-order Runge-Kutta numerical method.
 
-% Undamped DP data-sets were generated and used in this project due to the requirement that the map $T$ be surjective that we must not forget. 
-% An example of this would be the case of a DP exhibiting energy-loss (i.e.
-where the map $T$ is not surjective).
+% Undamped DP datasets were generated and used in this project due to the requirement that the map $T$ be surjective that we must not forget. 
+% An example of this would be the case of a DP exhibiting energy-loss (i.e.,where the map $T$ is not surjective).
 In such a case, the attractor is a single point in the phase space when the double pendulum comes to rest.
@@
 
-Having spent ample time to discuss the challenges, we now change tone so as to satisfy the optimist and present a successful run of the DP-data to predict the future movement of the pendulum.
+Having spent ample time to discuss the challenges, we now change tone to satisfy the optimist and present a successful run of the DP-data to predict the future movement of the pendulum.
 
@@
   These graphs were constructed by predicting the DP 1000 timesteps into the future and in so doing illustrating the long-term consistency and accuracy of the learnt system.
-Here we lock onto the near-exact trajectory for about 300 timesteps and stay close up until about 600 seconds.
+Here we lock onto the near-exact trajectory for about 300 timesteps and stay close until about 600 seconds.
 }
@@
 For the signal-to-noise ration we adopt the formula $\text{SNR}_{dB}=10\text{log}_{10}\big(\frac{P_\text{{signal}}}{P_{\text{noise}}}\big)$ where $P_\text{{signal}}$ refers to the variance observed in the normalised data and $P_{\text{noise}}$ represents the variance of the added noise.
-Once again we display the actual and predicted trajectories for the noisy DP data in Figure~\ref{fig:noisydp_success_traj}. 
-Next we consider the densities of the predicted and actual values in Figure~\ref{fig:noisydp_success_density} and observe that they display highly similar distributions.
+Once again, we display the actual and predicted trajectories for the noisy DP data in Figure~\ref{fig:noisydp_success_traj}. 
+Next, we consider the densities of the predicted and actual values in Figure~\ref{fig:noisydp_success_density} and observe that they display highly similar distributions.
 
@@
   These graphs were constructed by predicting the DP 1000 timesteps into the future and in so doing illustrating the long-term consistency and accuracy of the learnt system.
-Here we lock onto the near-exact trajectory for about 250 timesteps and stay close up until about 400 seconds.
+Here we lock onto the near-exact trajectory for about 250 timesteps and stay close until about 400 seconds.
 }
@@
 Thomas' Cyclically Symmetric Strange Attractor, a 3D attractor proposed in 1999 by Ren\'e Thomas in~\cite{ThomasAttractor}, is described by a set of three equations which is distinctly reminiscent of another deterministic three-equation model exhibiting chaotic behaviour proposed by Lorenz.
-It has a single parameter $\beta$ and has been shown to transition to chaotic behavious when $\beta<0.208186$~\cite{Thomas_BetaParameter}.
+It has a single parameter $\beta$ and has been shown to transition to chaotic behaviour when $\beta<0.208186$~\cite{Thomas_BetaParameter}.
 \begin{eqnarray}\label{eqns_thomas}
@@
 
-A sequence of 3000 observations in both the clear and noisy data-sets respectively was fed into the system with 300 entries discarded to simulate the network's loss of memory. 
+A sequence of 3000 observations in both the clear and noisy datasets respectively was fed into the system with 300 entries discarded to simulate the network's loss of memory. 
 The next 9000 steps were predicted and compared with true data. See Figure~\ref{fig:Thomas_trajectories} for the actual and predicted trajectories of the noise-free and perturbed systems respectively.
@@
   \endminipage
- \captionof{figure}{True (red) and predicted (blue) trajectories for the Thomas attractor with parameter value $\beta=0.1056$.Left: Noise added from normal distribution with mean zero and standard deviation 0.05.
+ \captionof{figure}{True (red) and predicted (blue) trajectories for the Thomas attractor with parameter value $\beta=0.1056$.
+Left: Noise added from normal distribution with mean zero and standard deviation 0.05.
 Right: No noise added to dataset.}
@@
 A finaldiscrete-time map named the Fractal Dream Attractor -- more commonly as the Pickover or Clifford Map (first discovered by Clifford A.
-Pickover and discussed in his fascinating book ``Chaos in Wonderland" \cite{PickoverChaos}) -- was also considered.
+Pickover and discussed in his fascinating book ``Chaos in Wonderland"~\cite{PickoverChaos}) -- was also considered.
 \begin{eqnarray}\label{eqns_clifford}
@@
 
-The behaviour of the system with a parameter values of $x_0=0$, $y_0=0$, $a = -1.7$,  $b = 1.8$, $c = -1.9$ and $d = -0.4$ was examined. 
+The behaviour of the system with parameter values of $x_0=0$, $y_0=0$, $a = -1.7$,  $b = 1.8$, $c = -1.9$ and $d = -0.4$ was examined. 
 Data was scaled to fit inside the interval $[-1,1]$. Only the noise-free case was investigated. 
@@
  A sequence of 25000 observations was fed into the system with 15000 entries discarded. 
- The reason for such a great number of discarded entries is the following: the system seems to have an incredibly long transient;driving it for less than 5000 entries only gives an accurate prediction to about 5 seconds and it starts improving when the discard is in the range $[10000,15000]$ to simulate the network's loss of memory.
+ The reason for such a great number of discarded entries is the following: the system seems to have an incredibly long transient; driving it for less than 5000 entries only gives an accurate prediction to about 5 seconds and it starts improving when the discard is in the range $[10000,15000]$ to simulate the network's loss of memory.
 The next 1000 steps were predicted and compared with true data.
@@
   \caption*{These graphs were constructed by predicting the Clifford system 1000 steps into the future and in so doing illustrating the long-term consistency and accuracy of the learnt system.
-As perceived here, we are able to lock on to the trajectory of the Clifford map almost exactly for the first 25 steps.}
+As perceived here, we can lock on to the trajectory of the Clifford map almost exactly for the first 25 steps.}
   \minipage{0.5\textwidth}
@@
 
-Here we consider only the non-invertible map of the Clifford attractor to supplement the results in \cite{Supp}. 
-This statistic is a generalisation of the Pearson correlation coefficient for real-valued random variables to random vectors (e.g., \cite{puccetti2019measuring}) and we present the details of computing this coefficient between $(x_{n-1},x_n)$ and $G_T(x_{n-1},x_n)$.
+Here we consider only the non-invertible map of the Clifford attractor to supplement the results in~\cite{Supp}. 
+This statistic is a generalisation of the Pearson correlation coefficient for real-valued random variables to random vectors (e.g.,~\cite{puccetti2019measuring}) and we present the details of computing this coefficient between $(x_{n-1},x_n)$ and $G_T(x_{n-1},x_n)$.
 If $\Sigma_{a}$ and $\Sigma_{b}$ denote the covariance matrices of the vectors $(x_{n-1},x_n)$ and $G_T(x_{n-1},x_n)$ respectively, and $\Sigma_{ab}$ denotes the covariance matrix between the vectors $(x_{n-1},x_n)$ and $G_T(x_{n-1},x_n)$, then the multidimensional correlation coefficient is computed by using the traces of these matrices.
@@
 \]
-The value $\rho$ also satisfies one of the desired and well-known properties of the one-dimensional Pearson coefficient; that is, $\rho=\pm 1$ if and only if $Y\overset{d}{=}AX+b$ for some invertible matrix $A$ and vector $b$.
+The value $\rho$ also satisfies one of the desired and well-known properties of the one-dimensional Pearson coefficient: that is, $\rho=\pm 1$ if and only if $Y\overset{d}{=}AX+b$ for some invertible matrix $A$ and vector $b$.
 In practice then, $\rho$, can be used to measure the linear relationship between two random vectors of the same dimension and thus serves as an indicator of functional complexity. 
@@
 Following~\cite{manjunath2021universal}, we state that the multidimensional correlation coefficient $\rho$ evidences a general reduction in the functional complexity of the map $G_T$ emerging due to a RNN.
-In Table~\ref{tbl_attractorsPearson}, the second column corresponds to the amount of artificial neurons (the dimension of $X$) in the implemented RNN.
+In Table~\ref{tbl_attractorsPearson}, the second column corresponds number of artificial neurons (the dimension of $X$) in the implemented RNN.
 
@@
 Rows correspond to distinct dynamical systems considered in the experiments
-              The second column corresponds to the amount of artificial neurons (the dimension of $X$) in the implemented RNN.
+              The second column corresponds number of artificial neurons (the dimension of $X$) in the implemented RNN.
               The last two columns comprise numerical estimates of $\rho$ for the relevant vectors. The linear association between $u_n$ vs $T(u_{n})$, as well as the linear relationship 
@@
               may be contrasted by the reader. 
-              Finally, the RNN's dimension (from \eqref{Seq_RNN})
+              Finally, the RNN's dimension (from~\eqref{Seq_RNN})
               is provided to exhibiting that an increase in dimension of the driven system will 

### 2022-01-26 projectReport/chapter6.tex
Found too many small errors

@@
 The established results are general, implying that any driven dynamical system with specific properties, such as SI-invertibility, can give rise to a causal embedding. 
-In particular, we have a topological conjugacy between the inverse-limit space of the underlying dynamical system and the single-delay lag dynamics of a driven system~\cite{manjunath2013echo}.
+We have a topological conjugacy between the inverse-limit space of the underlying dynamical system and the single-delay lag dynamics of a driven system~\cite{manjunath2013echo}.
 Finite-length input data is shown to be sufficient for forecasting as the left-infinite segment of the past of the input beyond some arbitrarily large but finite time (a result which holds due to the continuity of the universal semi-conjugacy) influences the driven system's states in an almost negligibly small manner.
@@
 Finally, we have considered some physical systems to observe the methodology in practice and verify that we can indeed predict the future evolution of chaotic systems such as the double pendulum.
-We demonstrated the robustness of our methodology to external noise by considering the cases of the  Dream Attractor and the Thomas' attractor.
-\ednote{A sufficient ending?}
+We demonstrated the robustness of our methodology to external noise by considering the cases of the  Fractal Dream Attractor and the Thomas' attractor.
+
 %\ednote{Loose paragraph commented out}
